### 2012-06-19 Notes/ContTime.tex
added notes and poster

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
% Preamble:
\usepackage{a4wide}
\usepackage[centertags]{amsmath}
%\usepackage{ams} for finding documentation only
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage[sort&compress,numbers]{natbib}
%\usepackage{citeB}
\usepackage{ifpdf}
%\usepackage{graphicx}
%\usepackage{graphics} for finding documentation only
%\usepackage{xcolor}
%\usepackage{pgf}
\ifpdf
\usepackage[pdftex,bookmarks,bookmarksopen,pdfstartview=FitH]{hyperref}
\else
\usepackage[hypertex]{hyperref}
%\DeclareGraphicsRule{.png}{eps}{.bb}{}
\fi
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
\vfuzz2pt % Don't report over-full v-boxes if over-edge is small
\hfuzz2pt % Don't report over-full h-boxes if over-edge is small
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{mydefs.tex}
\input{newsymb.tex}
%logic
\newcommand{\means}{\Longleftrightarrow}
\newcommand{\requires}{\Longleftarrow}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\eq}{\mathbf{p}^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\DeclareMathOperator{\SNR}{SNR}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
% ----------------------------------------------------------------
\usepackage{amsthm}
\newtheoremstyle{sldefinition}%
  {3pt}%space above
  {3pt}%space below
  {}%body font
  {}%indent amount
  {\bfseries}%theorem head font
  {}%theorem head punctuation
  {\newline}%space after head
  {\thmname{#1}\thmnumber{ #2}:{\bfseries\thmnote{ #3}}}%head spec
\newtheoremstyle{slplain}%
  {3pt}%space above
  {3pt}%space below
  {}%body font
  {}%indent amount
  {\bfseries}%theorem head font
  {}%theorem head punctuation
  {\newline}%space after head
  {\thmname{#1}\thmnumber{ #2}{\bfseries\thmnote{ (#3)}}: }%head spec
%
\theoremstyle{slplain}
\newtheorem{thm}{Theorem}
\newtheorem{cor}[thm]{Corollary}
%
\theoremstyle{sldefinition}
\newtheorem{defn}{Definition}
%
\theoremstyle{remark}
\newtheorem*{notn}{Notation}
\newtheorem*{rem}{Remark}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Area maximisation in continuous time}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
  We try to find the continuous time Markov process that has the maximal area under the signal-to-noise curve.
\end{abstract}

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beginning of Article:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Continuous time Markov processes}\label{sec:ContMarkov}

In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time.
It is a straightforward generalisation of material that can be found in \cite{kemeny1960finite} with some ideas from \cite{hunter2000survey}.

\subsection{Notation}\label{sec:not}

For any matrix $\mathbf{A}$, we define matrices $\mathbf{A}\dg$ and $\overline{\mathbf{A}}$ as
%
\begin{equation}\label{eq:dgdef}
  \mathbf{A}\dg_{ij} \equiv \delta_{ij}\mathbf{A}_{ij},
  \qquad
  \overline{\mathbf{A}} \equiv \mathbf{A}-\mathbf{A}\dg.
\end{equation}
%
We let $\onev$ denote a column-vector of ones and $\onem=\onev\onev\trans$ denote a matrix of ones.

A Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$.
The probabilities of being in each state at time $t$, the row-vector $\mathbf{p}(t)$, evolve according to
%
\begin{equation}\label{eq:rowsum}
  \diff{\mathbf{p}(t)}{t} = \mathbf{p}(t) \MM,
  \qquad
  \mathbf{p}(t)\onev=1,
  \qquad
  \MM\onev=0.
\end{equation}
%
In Surya's notes, $\MM=r\W^F$.

The equilibrium probabilities, $\eq$, satisfy
%
\begin{equation}\label{eq:equilibrium}
  \eq\MM=0,
  \qquad
  \eq\onev=1.
\end{equation}
%
As we assume an ergodic process, this eigenvalue is non-degenerate.
If all other eigenvalues have strictly negative real parts, the process is regular (aperiodic).

We define additional matrices
%
\begin{equation}\label{eq:defDLP}
  \Lb \equiv (-\MM\dg)\inv,
  \qquad
  \Pb \equiv \I + \Lb\MM.
\end{equation}
%
It can be shown that $\Lb_{ii}$ is the mean time it takes to leave state $i$ and $\Pb_{ij}$ is the probability the the next transition from state $i$ goes to state $j$:
%
\begin{equation}\label{eq:LamdaPcmpt}
  \Lb_{ii} = \frac{1}{\sum_{j \neq j} \MM_{ij}},
  \qquad
  \Pb_{ij} = %\left\{
%  \begin{aligned}
%     &0                                         && \text{if }i=j, \\
%     &\frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &&\text{otherwise}.
%  \end{aligned}
%  \right.
  \begin{cases}
     0                                         &\text{if }i=j, \\
     \frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &\text{otherwise}.
  \end{cases}
\end{equation}
%
Furthermore, we also define
%
\begin{equation}\label{eq:pdotD}
  \D \equiv \operatorname{diag}(\eq)\inv,
  \qquad\implies\qquad
  \eq\D=\onev\trans.
\end{equation}
%

\subsection{Fundamental matrix}\label{sec:fund}

\begin{defn}[Fundamental matrix]
  %
  \begin{equation}\label{eq:funddef}
    \fund \equiv (-\MM + \onev\pib)\inv,
  \end{equation}
  %
  where $\pib$ is any row-vector with $\pib\onev=1/\tau\neq0$.
\end{defn}
Note that the canonical choice for the discrete time version, $\pib=\eq$, is not available here due to problems with units.
It will be helpful to choose $\pib$ to be independent of $\MM$, \eg $\pib=\onev\trans/(n\tau)$.
All quantities that we calculate using $\fund$ below will be independent of this choice.

\begin{thm}
  The definition of $\fund$ is valid, \ie $(-\MM + \onev\pib)$ is invertible.
\end{thm}
\begin{proof}
  Assume there exists an $\mathbf{x}$ such that
  %
  \begin{equation}\label{eq:fundinvkern}
    (-\MM + \onev\pib)\mathbf{x}=0.
  \end{equation}
  %
  Multiplying from the left with $\eq$ gives
  %
  \begin{equation}\label{eq:fundinvpix}
    \pib \mathbf{x} = 0.
  \end{equation}
  %
  Substituting back into \eqref{eq:fundinvkern} gives
  %
  \begin{equation*}
    \MM\mathbf{x}=0.
  \end{equation*}
  %
  As we assume an ergodic process, the zero eigenvalue is non-degenerate.
  Therefore, $\mathbf{x}=\lambda\onev$.
  Substituting this into \eqref{eq:fundinvpix} gives
  %
  \begin{equation*}
    \lambda\pib\onev = \frac{\lambda}{\tau} = 0.
  \end{equation*}
  %
  As we defined $\pib$ such that $1/\tau\neq0$, this means $\lambda=0 \implies \mathbf{x}=0$.
\end{proof}

\begin{cor}
  %
  \begin{align}
    \pib\fund &= \eq, \label{eq:fundprob}\\
    \fund\onev &= \tau\onev,\label{eq:fundrowsum}\\
    \I+\MM\fund &= \onev\eq, \label{eq:fundQZ}\\
    \I+\fund\MM &= \tau\onev\pib. \label{eq:fundZQ}
  \end{align}
  %
\end{cor}
\begin{proof}
  We can deduce \eqref{eq:fundprob} and \eqref{eq:fundrowsum} be pre/post-multiplying the following equations by $\fund$:
  %
  \begin{equation*}
    \begin{aligned}
      \eq(-\MM + \onev\pib) &= \pib, \\
      (-\MM + \onev\pib)\onev &= \frac{\onev}{\tau}.
    \end{aligned}
  \end{equation*}
  %
  We can then deduce \eqref{eq:fundQZ} and \eqref{eq:fundZQ} by substituting these into
  %
  \begin{equation*}
    (-\MM + \onev\pib)\fund = \fund(-\MM + \onev\pib) = \I.
  \end{equation*}
  %
\end{proof}

%We will not need to use the concept of a g-inverse anywhere, so you can skip ahead to \sref{sec:fpt} now. For the sake of completeness, we can define $\MM^\#$ as
%%
%\begin{equation}\label{eq:ginvdef}
%  \MM^\# \equiv \tau\onev\eq - \fund
%\end{equation}
%%
%\begin{thm}
%  The matrix $\MM^\#$ satisfies the following defining properties of a g-inverse of $\MM$:
%  %
%  \begin{align}
%    \MM^\# \MM \MM^\# &= \MM^\#,     \label{eq:ginvpropo} \\
%    \MM \MM^\# \MM    &= \MM,        \label{eq:ginvpropi} \\
%    \MM \MM^\#        &= \MM^\# \MM, \label{eq:ginvcomm}
%  \end{align}
%  %
%  with \eqref{eq:ginvcomm} holding iff $\eq$ and $\pib$ are proportional to each other.
%\end{thm}
%\begin{proof}
%  Using \eqref{eq:rowsum}, \eqref{eq:equilibrium}, \eqref{eq:fundprob} and \eqref{eq:fundrowsum}, one can verify
%  %
%  \begin{alignat*}{2}
%    \MM \MM^\# &= -\MM \fund,
%    &\qquad
%    \pib \MM^\# &= 0,
%    \\
%    \MM^\# \MM &= -\fund \MM,
%    &
%    \MM^\# \onev &= 0.
%  \end{alignat*}
%  %
%  Substituting into \eqref{eq:fundQZ} and \eqref{eq:fundZQ} leads to
%  %
%  \begin{equation*}
%    \I - \MM \MM^\# = \onev \eq,
%    \qquad
%    \I - \MM^\# \MM = \tau \onev \pib.
%  \end{equation*}
%  %
%  Pre/post-multiplying these by $\MM$ or $\MM^\#$ produces \eqref{eq:ginvpropo} and \eqref{eq:ginvpropi}. Subtracting produces
%  %
%  \begin{equation*}
%    [\MM,\MM^\#] = \onev(\tau\pib-\eq),
%  \end{equation*}
%  %
%  which vanishes iff $\eq$ and $\pib$ are proportional to each other (due to $\pib\onev=1/\tau$ and $\eq\onev=1$, the constant of proportionality could only be $\tau$).
%\end{proof}

\subsection{First passage times}\label{sec:fpt}

\begin{defn}[First passage time matrix]
 We define $\fptb_{ij}$ as the mean time it takes the process to reach state $j$ for the first time, starting from state $i$.
 We also define $\fpt\dg_{ii}$ as the mean time it takes the process to return to state $i$.
 As usual, $\fpt=\fptb+\fpt\dg$.
\end{defn}

Consider the first time the process leaves state $i$.
On average, this will take time $\Lb_{ii}$.
With probability $\Pb_{ij}$, it will go directly to $j$, so the conditional mean time would be $\Lb_{ii}$.
On the other hand, if it goes to some other state, $k$, with probability $\Pb_{ik}$, the conditional mean time would be $\fptb_{kj}+\Lb_{ii}$.
Combining these, we get the recursion relation
%
\begin{equation}\label{eq:fptrecursion}
  \begin{aligned}
  \fpt_{ij} &= \sum_{k\neq j} \Pb_{ik}(\fptb_{kj}+\Lb_{ii}) + \Pb_{ij}\Lb_{ii} \\
    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \sum_{k} \Pb_{ik}\Lb_{ii}\\
    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \Lb_{ii},\\
   \fpt &= \Pb \fptb + \Lb\onem,
  \end{aligned}
\end{equation}
%
or, using \eqref{eq:defDLP},
%
\begin{equation}\label{eq:fptrecdg}
  \fpt\dg = \Lb\MM\fptb+\Lb\onem.
\end{equation}
%

\begin{thm}
  The recurrence times are given by
  %
  \begin{equation}\label{eq:recurtime}
    \fpt\dg = \Lb\D.
  \end{equation}
  %
\end{thm}
\begin{proof}
  Pre-multiply \eqref{eq:fptrecdg} by $\eq\Lb\inv$:
  %
  \begin{equation*}
    \eq\Lb\inv\fpt\dg = \onev\trans,
  \end{equation*}
  %
  or in component form
  %
  \begin{equation*}
    \eq_i \Lb_{ii}\inv \fpt\dg_{ii} = 1.
  \end{equation*}
  %
  This is the same as \eqref{eq:recurtime} in component form: $\fpt\dg_{ii} = \Lb_{ii}/\eq_i$.
\end{proof}
I think the extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.

\begin{thm}
  The solution of \eqref{eq:fptrecdg} is unique.
\end{thm}
\begin{proof}
  We have already seen that the diagonal part is determined by \eqref{eq:fptrecdg}.
  Suppose we have two solutions for the off-diagonal part, $\fptb$ and $\fptb'$.
  Subtracting \eqref{eq:fptrecdg} for these two solutions gives
  %
  \begin{equation*}
    \Lb\MM(\fptb-\fptb')=0.
  \end{equation*}
  %
  As we are assuming an ergodic process, the zero eigenvalue is non-degenerate.
  Therefore
  %
  \begin{equation*}
    \fptb-\fptb' = \onev\mathbf{y},
  \end{equation*}
  %
  for some row-vector $\mathbf{y}$. Looking at the diagonal components
  %
  \begin{equation*}
    \mathbf{y}_{i} = \fptb_{ii}-\fptb'_{ii} = 0,
  \end{equation*}
  %
  and therefore $\fptb-\fptb'=0$.
\end{proof}

\begin{thm}
  The off-diagonal mean first passage times are given by
  %
  \begin{equation}\label{eq:fptfund}
    \fptb = (\onem\fund\dg - \fund)\D.
  \end{equation}
  %
\end{thm}
\begin{proof}
  We just need to show that this satisfies \eqref{eq:fptrecdg} with \eqref{eq:recurtime}, and that the diagonal elements vanish.
  Premultiply \eqref{eq:fptfund} by $\Lb\MM$ and subtract from \eqref{eq:fptrecdg}:
  %
  \begin{equation*}
  \begin{aligned}
    \fpt\dg - \Lb\MM\fptb &= \Lb(\I+\MM\fund)\D\\
     &= \Lb\onev\eq\D \\
     &= \Lb\onem,
  \end{aligned}
  \end{equation*}
  %
  where we used \eqref{eq:fundQZ} and \eqref{eq:pdotD}.
  The vanishing of the diagonal elements follows trivially from the component form of \eqref{eq:fptfund}:
%
\begin{equation}\label{eq:fptfundcmpt}
  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
\end{equation}
%

\end{proof}

\subsection{Mixing time (Kemeny's constant)}\label{sec:mixtime}

\begin{thm}\label{th:kemenyconst}
  The quantity
  %
  \begin{equation}\label{eq:mixdef}
    \eta \equiv \sum_j \fptb_{ij}\eq_j
  \end{equation}
  %
  is independent of $i$.
\end{thm}
\begin{proof}
  We use \eqref{eq:fptfund}, \eqref{eq:fundrowsum} and the transpose of \eqref{eq:pdotD}:
  %
  \begin{equation*}
    \begin{aligned}
      \fptb (\eq)\trans &= (\onem\fund\dg - \fund) \D (\eq)\trans \\
        &= (\onev\onev\trans\fund\dg - \fund) \onev \\
        &= (\onev\trans\fund\dg\onev)\onev - \fund\onev \\
        &= (\tr\fund - \tau)\onev.
    \end{aligned}
  \end{equation*}
  %
  which proves \eqref{eq:mixdef} with $\eta = \tr\fund - \tau$.
\end{proof}

Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version \cite{hunter2006mixing} where this would only shift $\eta$ by 1.

\subsection{Sensitivity of equilibrium distribution}\label{sec:sensitivity}

Suppose that the Markov process, defined by $\MM$, depends on some parameter $\alpha$.
Differentiating \eqref{eq:funddef} gives
%
\begin{equation}\label{eq:diffZ}
  \diff{\fund}{\alpha} = \fund \diff{\MM}{\alpha} \fund.
\end{equation}
%
We can substitute this into the derivative of \eqref{eq:fundprob}:
%
\begin{equation}\label{eq:diffp}
  \diff{\eq}{\alpha} = \pib \fund \diff{\MM}{\alpha} \fund = \eq \diff{\MM}{\alpha} \fund.
\end{equation}
%
We can rewrite this in component form and use the fact that $\MM_{ii} = - \sum_{i\neq j} \MM_{ij}$:
%
\begin{equation}\label{eq:diffpT}
\begin{aligned}
  \diff{\eq_k}{\alpha} &= \sum_{i,j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} \\
    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} + \sum_i \eq_i \diff{\MM_{ii}}{\alpha} \fund_{ik} \\
    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} (\fund_{jk} - \fund_{ik}) \\
    &= \sum_{i\neq j} \diff{\MM_{ij}}{\alpha} \eq_i \eq_k (\fptb_{ik} - \fptb_{jk}),
\end{aligned}
\end{equation}
%
which is the result of \cite{cho2000markov} that we need.
Note that the summand vanishes for $i=j$, so we can drop the restriction $i\neq j$ from the range of the sum.

\subsection{Subsets and flux}\label{sec:subsets}

Let us denote the set of states by $\CS$.
Consider a subset $\CA\subset\CS$.
We can define a projection operator onto this subset:
%
\begin{equation}\label{eq:proj}
  \prn{\I^\CA}_{ij} =
    \begin{cases}
      1 &\text{if $i=j\in\CA$,}\\
      0 &\text{otherwise.}
    \end{cases}
\end{equation}
%
We will use superscripts/subscripts to denote projection onto/summation over a subset:
%
\begin{equation}\label{eq:projsum}
  \begin{aligned}
    \pib^\CA &=\pib\I^\CA, &
    \M^{\cdot\CA}&=\M\I^\CA, &
    \M^{\CA\cdot}&=\I^\CA\M, &
    \mathbf{x}^\CA &= \I^\CA\mathbf{x},
    \\
    \pib_\CA &=\pib\onev^\CA, &
    \M_{\cdot\CA}&=\M\onev^\CA, &
    \M_{\CA\cdot}&=\prn{\onev^\CA}\trans\M, &
    \mathbf{x}_\CA &= \prn{\onev^\CA}\trans\mathbf{x},
  \end{aligned}
\end{equation}
%
where $\pib$ is a row vector, $\CM$ is a matrix and $\mathbf{x}$ is a column vector.

We can define a flux matrix, a.k.a.\ ergodic flow:
%
\begin{equation}\label{eq:flux}
  \F = \D\inv\MM,
  \qquad
  \F_{ij} = \eq_i \MM_{ij}.
\end{equation}
%
This measures the flow of probability between states in the equilibrium distribution.
Detailed balance, a.k.a.\ reversibility, is equivalent to $\F=\F\trans$.

The flux between two subsets is a particularly useful quantity:
%
\begin{equation}\label{eq:subflux}
  \F_{\CA\CB} = {\eq}^\CA\MM\onev^\CB.
\end{equation}
%
One can show that
%
\begin{equation}\label{eq:compflux}
  \F_{\CA\CA\comp} = \F_{\CA\comp\CA} = -\F_{\CA\CA} = -\F_{\CA\comp\CA\comp}
\end{equation}
%
using $\prn{{\eq}^\CA + {\eq}^{\CA\comp}}\MM=0$ and $\MM\prn{{\onev}^\CA + {\onev}^{\CA\comp}}=0$.

\subsection{Lumpability}\label{sec:lump}

Suppose we have partitioned the states into disjoint subsets, $\brc{\CA_\alpha}$:
%
\begin{equation}\label{eq:partition}
  \bigcup_\alpha \CA_\alpha = \CS,
  \qquad
  \CA_\alpha \cap \CA_\beta = \delta_{\alpha\beta}\CA_\alpha.
\end{equation}
%
We will use $\alpha$ instead of $\CA_\alpha$ in superscripts and subscripts in the following.
The fact that these subsets are disjoint and exhaustive allows us to define the function
%
\begin{equation}\label{eq:whichsey}
  \sigma(i)=\alpha
  \qquad\means\qquad
  i\in\CA_\alpha.
\end{equation}
%

We can use this partition to define a new stochastic process associated with the original Markov chain.
At time $t$, if the state of the original process is $i$, the state of the new process is $\sigma(i)$.

One may ask if this new process is a Markov chain.
The answer is yes, if the original Markov chain has a property called lumpability \wrt the partition \cite[\S6.3]{kemeny1960finite}:
%
\begin{equation}\label{eq:lump}
  \sigma(i)=\sigma(j)
  \quad\implies \quad
  \MM_{i\alpha}=\MM_{j\alpha} \equiv \sum_{k\in\CA_\alpha} \MM_{jk},
\end{equation}
%
\ie the total transition rate from some state to some subset is the same for all starting states within the same subset.
This common value is the transition rate for the new lumped Markov chain.

This can be rewritten with the aid of two matrices
%
\begin{equation}\label{eq:lumpmats}
  U_{\alpha i} = \frac{\delta_{\alpha\sigma(i)}}{\abs{\CA_\alpha}},
  \qquad
  V_{i\alpha} = \delta_{\sigma(i)\alpha}.
\end{equation}
%
Left multiplication by $U$ averages over subsets, right multiplication by $V$ sums over subsets.
For $U$, we chose the uniform measure in each subset. Any measure would work equally well, \eg one proportional to the equilibrium distribution:
%
\begin{equation}\label{eq:altlumpmats}
  U_{\alpha i} = \frac{{\eq_i}^\alpha}{\eq_\alpha}.
\end{equation}
%

One can show that $(UV)_{\alpha\beta}=\delta_{\alpha\beta}$.
The matrix $VU$ is also interesting.
It has a block diagonal structure, with each block corresponding to a subset.
Each block is a discrete-time ergodic Markov matrix (it is an independent trials process with probabilities given by the measure chosen for $U$).
This means that the right eigenvectors with eigenvalue 1 will be those that are constant in each subset:
%
\begin{equation}\label{eq:setconst}
  VU\mathbf{x}=\mathbf{x}
  \qquad\means\qquad
  \mathbf{x} = \sum_\alpha x_\alpha \onev^\alpha.
\end{equation}
%

This allows us to write the lumpability condition \eqref{eq:lump}, and the transition matrix for the lumped process compactly:
%
\begin{equation}\label{eq:lumpcompact}
  VU\MM V = \MM V,
  \qquad
  \widehat{\MM} = U\MM V.
\end{equation}
%
By induction, one can show that similar relations hold for all powers:
%
\begin{equation}\label{eq:lumppower}
  VU\MM^nV = \MM^nV,
  \qquad
  \widehat{\MM}^n = U\MM^n V,
\end{equation}
%
and, via the Taylor series, for the exponential as well:
%
\begin{equation}\label{eq:lumpexp}
  VU \e^{t\MM} V = \e^{t\MM}V,
  \qquad
  \e^{t\widehat{\MM}} = U\e^{t\MM} V.
\end{equation}
%
The equilibrium distribution of the lumped process is given by
%
\begin{equation}\label{eq:lumpeq}
  \widehat{\mathbf{p}}^\infty = \eq V.
\end{equation}
%

\section{Signal-to-Noise ratio (SNR)}\label{sec:SNR}

In this section we will look at the signal-to-noise curve, and put an upper bound on its initial value.
We will only consider ergodic Markov chains.
Transient states would be unoccupied in equilibrium and would not be accessed by the signal creation process, therefore they could be removed from the analysis.
Absorbing chains are degenerate cases: they have zero initial signal but infinite decay times, so they can only be approached as the limit of a sequence of ergodic chains.

\subsection{Framework}\label{sec:framework}

The individual potentiation/depression events will be described by \emph{discrete}-time Markov chains:
%
\begin{equation}\label{eq:MWdef}
  \M^\pm \equiv \I + \W^\pm,
  \qquad
  \M^\pm\onev = \onev,
  \qquad
  \M^\pm_{ij} \in [0,1].
\end{equation}
%
The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf^\pm$, where $f^++f^-=1$ are the fraction of plasticity events that are potentiating/depressing respectively.
This means that the ``forgetting'' process will be described by the \emph{continuous}-time Markov chain:
%
\begin{equation}\label{eq:forgetting}
  \MM = r \W^F \equiv r\prn{f^+\W^+ + f^-\W^-}.
\end{equation}
%
We only require that this Markov chain is ergodic. The Markov chains described by $\M^\pm$ need not be.

We assume that the probability distribution starts in the equilibrium distribution \eqref{eq:equilibrium}.
During the initial signal creation, a fraction $f^+$ will change to $\eq\M^+$ and a fraction $f^-$ will change to $\eq\M^-$.
After this, probabilities will evolve according to \eqref{eq:rowsum}.

\subsection{SNR curve}\label{sec:SNRcurve}

As we ignore correlations between different synapses, the only quantity of interest will be the number of synapses with each strength.
The signal can be defined as the increase in synaptic strength due to potentiation plus the decrease in synaptic strength due to depression.
If there are $N$ synapses and the strength of the synapse in each state is given by the column vector, $\w$, this is given by
%
\begin{equation}\label{eq:signal}
  S(t) = N f^+ [\mathbf{p}(t|+)-\eq]\w - N f^+ [\mathbf{p}(t|-)-\eq]\w,
\end{equation}
%
where $\mathbf{p}(t|\pm)$ is the probability of being in each state conditional on potentiation/depression at $t=0$:
%
\begin{equation}\label{eq:initprob}
  \mathbf{p}(t|\pm) = \eq\M^\pm \e^{rt\W^F}.
\end{equation}
%
We can rewrite the signal as
%
\begin{equation}\label{eq:signalr}
\begin{aligned}
  S(t) &= N \eq (f^+\W^+ - f^-\W^-)\e^{rt\W^F}\w\\
    & N (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\W^F}\w.
\end{aligned}
\end{equation}
%
For the noise, we can use the standard deviation of the synaptic strength:
%
\begin{equation}\label{eq:noise}
  \text{Noise} = \sqrt{N \prn{\sum_i \eq_i\w_i^2 - (\eq\w)^2}}.
\end{equation}
%
From now on we shall assume that there are only two possible synaptic strengths, strong and weak, which we will describe by $\w_i=\pm1$. This splits the set of states into two subsets:
%
\begin{equation}\label{eq:plusminusstates}
  \boldsymbol{+} = \set{k}{\w_k=+1},
  \qquad
  \boldsymbol{-} = \set{k}{\w_k=-1}.
\end{equation}
%
This simplifies the noise:
%
\begin{equation}\label{eq:noisepm}
  \text{Noise} = \sqrt{N \prn{4\eq_+\eq_-}}.
\end{equation}
%
The factors of $\eq_+\eq_-$ will have little effect on anything that follows, so we will drop them.\footnote{They will not affect the optimal transition rates for most quantities, and the optima will correspond to the symmetric case when $\eq_+=\eq_-=\half$.} This leaves the signal-to-noise ratio as
%
\begin{equation}\label{eq:SNRcurve}
  \SNR(t) = \sqrt{N} (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\W^F}\w.
\end{equation}
%

We can express this in terms of the one parameter family of transition matrices:
%
\begin{equation}\label{eq:Walpha}
  \begin{aligned}
  \W(\alpha) &= \alpha\W^++(1-\alpha)\W^-,
  &\quad&\implies&
    \W^F&=\W(f^+),\\&&&&
    \W^+ - \W^- &= \diff{\W}{\alpha},\\&&&&
    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \W^F.
  \end{aligned}
\end{equation}
%
Then \eqref{eq:SNRcurve} becomes
%
\begin{equation}\label{eq:SNRalpha}
  \SNR(t) = \sqrt{N} (2f^+f^-) \diff{\eq}{\alpha}(-\W^F)\,\e^{rt\W^F}\w.
\end{equation}
%

\subsection{Degrees of freedom and constraints}\label{sec:constraints}

We would like to use the components of $\W^\pm$ as the independent degrees of freedom when we try to maximise quantities associated with the SNR curve \eqref{eq:SNRcurve}.
However, \eqref{eq:MWdef} imposes some restrictions on the space of allowed values.
We can solve some of these by setting
%
\begin{equation}\label{eq:Wdiag}
  \W^\pm_{ii} = -\sum_{j\neq i} \W^\pm_{ij}.
\end{equation}
%
Then we treat the off-diagonal components as the independent degrees of freedom.
These must satisfy the following inequalities:
%
\begin{align}\label{eq:Wmin}
  \W^\pm_{ij} &\geq 0 \qquad \text{for $i \neq j$,}\\
\label{eq:Wmax}
  \sum_{j \neq i} \W^\pm_{ij} &\leq 1.
\end{align}
%
The inequalities $\W^\pm_{ij} \leq 1$, $-1\leq\W^\pm_{ii}\leq0$ follow automatically from these.

In \sref{sec:rescale}, we will see that the area under the SNR curve is invariant under the scaling $\W^\pm\to\lambda\W^\pm$.
This means that we can ignore the constraint \eqref{eq:Wmax} when maximising the area:
after the area has been maximised, we can use this degree of freedom to enforce \eqref{eq:Wmax} without changing the area.
However, this scaling does change the SNR curve, in particular its initial value, so we can't ignore \eqref{eq:Wmax} when maximising that.

\subsection{Lumpability}\label{sec:SNRlump}

Suppose that we have a partition such that $\W^+$ and $\W^-$ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \sref{sec:lump}):
%
\begin{equation}\label{eq:lumpablesynapse}
  VU\W^\pm V = \W^\pm V,
  \qquad
  VU\w=\w.
\end{equation}
%
We can define a new synapse with
%
\begin{equation}\label{eq:lumpedsynapse}
  \widehat{\W}^\pm = U\W^\pm V,
  \qquad
  \widehat{\w} = U \w,
  \qquad
  \widehat{\mathbf{p}}^\infty = \eq V.
\end{equation}
%
This synapse has an SNR curve:
%
\begin{equation}\label{eq:lumpedSNR}
  \begin{aligned}
    \frac{\SNR(t)}{\sqrt{N} (2f^+f^-)} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}^+ - \widehat{\W}^-)\e^{rt\widehat{\W}^F}\widehat{\w}. \\
      &= \eq VU (\W^+ - \W^-) VU \e^{rt\W^F} VU \w. \\
      &= \eq (\W^+ - \W^-) VU \e^{rt\W^F} VU \w. \\
      &= \eq (\W^+ - \W^-)\e^{rt\W^F} VU \w. \\
      &= \eq (\W^+ - \W^-)\e^{rt\W^F}\w. \\
  \end{aligned}
\end{equation}
%
\ie the lumped process has exactly the same SNR as the original one.

\subsection{Initial SNR and flux}\label{sec:initflux}

Using the first line of \eqref{eq:signalr}, we can write the initial SNR as
%
\begin{equation}\label{eq:init}
  \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f^+\W^+-f^-\W^-)(\onev^+-\onev^-).
\end{equation}
%
Using $\W^\pm(\onev^++\onev^-)=0$ and \eqref{eq:compflux}:
%
\begin{equation*}
  r{\eq}^-(f^+\W^++f^-\W^-)\onev^+ = \F_{-+} = \F_{+-} = r{\eq}^+(f^+\W^++f^-\W^-)\onev^-,
\end{equation*}
%
we can rewrite \eqref{eq:init} as
%
\begin{equation}\label{eq:initflux}
  I = \frac{4\F_{-+}}{r} - 4{\eq}^+f^+\W^+\onev^- - 4{\eq}^-f^-\W^-\onev^+.
\end{equation}
%
The last two terms are guaranteed to be negative, as the diagonal parts of $\W^\pm$ cannot contribute.
Therefore
%
\begin{equation}\label{eq:initfluxineq}
  \SNR(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}.
\end{equation}
%
This inequality is saturated if potentiation never takes it from a $+$ state to a $-$ state and depression never takes it from a $-$ state to a $+$ state.

\subsection{Maximising initial SNR}\label{sec:initmax}

The flux will be maximised if potentiation is guaranteed to take it to a $+$ state and depression is guaranteed to take it to a $-$ state:
%
\begin{equation}\label{eq:guarantee}
  \sum_{j\in+} \W^+_{ij} = 1 \quad \forall\, i\in-,
  \qquad
  \sum_{j\in-} \W^-_{ij} = 1 \quad \forall\, i\in+.
\end{equation}
%
This implies that the process is lumpable to a two state system (see \sref{sec:lump} and \sref{sec:SNRlump}):
%
\begin{multline}\label{eq:binarylump}
%\begin{aligned}
  \W^+ = \begin{pmatrix}
           -1 & 1 \\
           0 & 0 \\
         \end{pmatrix}
  ,\quad
  \W^- = \begin{pmatrix}
           0 & 0 \\
           1 & -1 \\
         \end{pmatrix}
  ,\quad
  \w = \begin{pmatrix}
           -1 \\
           1 \\
         \end{pmatrix}
  ,\\ \implies\quad
  \W^F = \begin{pmatrix}
           -f^+ & f^+ \\
           f^- & -f^- \\
         \end{pmatrix}
  ,\quad
  \eq = \begin{pmatrix}
           f^- & f^+ \\
         \end{pmatrix}
  ,%&
%  \e^{rt\W^F} &= \I+(1-\e^{-rt})\W^F
%         \begin{pmatrix}
%           f^-+f^+\e^{-rt} & f^+(1-\e^{-rt}) \\
%           f^-(1-\e^{-rt}) & f^++f^-\e^{-rt} \\
%         \end{pmatrix}
%  ,
%  \end{aligned}
\end{multline}
%
which leads to the initial SNR %curve
\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to
%$\SNR(t) = \sqrt{N (4f^+f^-)} \,\e^{-rt}$,
$\SNR(0) \leq \sqrt{N (4f^+f^-)}$,
which would have no effect on what follows.}
%
\begin{equation}\label{eq:binarySNR}
 % \SNR(t) = \sqrt{N} (4f^+f^-) \,\e^{-rt}.
  \SNR(0) \leq \sqrt{N} (4f^+f^-) .
\end{equation}
%
%The initial SNR is maximised at $f^+=f^-=\half$:
This is maximised at $f^+=f^-=\half$:
%
\begin{equation}\label{eq:initmax}
  \SNR(0) \leq \sqrt{N}.
\end{equation}
%

\section{Area maximisation}\label{sec:areamax}

In this section we will find an upper bound on the area under the signal-to-noise curve.
As in \sref{sec:SNR}, we will only consider ergodic Markov chains.
We will see in \sref{sec:multistate} that the optimal chain is absorbing, so it lies on the boundary of the (open) set of ergodic chains, but it still puts an upper bound on the area.

\subsection{Area under signal-to-noise curve}\label{sec:area}

The signal-to-noise curve is given by \eqref{eq:SNRalpha}.
%%
%\begin{equation}\label{eq:SNR}
%  \SNR(t) = \sqrt{N}(2f^+f^-) \diff{\eq}{\alpha} (-\W^F) \e^{rt\W^F} \w.
%\end{equation}
%%
The area is computed by integrating this
%
\begin{equation}\label{eq:area}
\begin{aligned}
  A &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\W^F}}_0^\infty \w \\
    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \w.
\end{aligned}
\end{equation}
%

We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f^+f^-)\hat{A}$ and $q_{ij} \equiv \diff{\W^F_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
%
\begin{equation}\label{eq:areaT}
  \hat{A} = \sum_{i,j,k} \eq_i q_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
\end{equation}
%

\begin{defn}[Partial mixing times]
We define the $\pm$ mixing times as
%
\begin{equation}\label{eq:mixingpm}
\begin{aligned}
  \eta^\pm_i &\equiv \sum_k \fptb_{ik} \eq_k \prn{\frac{1 \pm \w_k}{2}}
%    &&= \sum_{\set{k}{\w_k=\pm1}} \fptb_{ik} \eq_k \\
    &&= \sum_{k\in\pm} \fptb_{ik} \eq_k \\
    &= \sum_k \prn{\fund_{kk}-\fund_{ik}} \prn{\frac{1 \pm \w_k}{2}}
    &&= \sum_{k\in\pm} \prn{\fund_{kk}-\fund_{ik}} .
\end{aligned}
\end{equation}
%
We can think of $\eta^+_i$ as a measure of the ``distance'' to the $\w_k=+1$ states and $\eta^-_i$ as the ``distance'' to the $\w_k=-1$ states.
\end{defn}
Using \eqref{eq:mixdef}, we can write:
%
\begin{equation}\label{eq:mixingrels}
\begin{aligned}
  \eta^+_i + \eta^-_i &= \eta,\\
  2(\eta^+_i - \eta^+_j) &= \sum_k (\fptb_{ik}-\fptb_{jk}) \eq_k \w_k
    = \sum_k (\fund_{jk}-\fund_{ik}) \w_k.
\end{aligned}
\end{equation}
%
We could arrange the states in order of decreasing $\eta^+$, which is the same as the order of increasing $\eta^-$.

We can rewrite \eqref{eq:areaT} as
%
\begin{equation}\label{eq:areaEta}
\begin{aligned}
  \hat{A} &= 2\sum_{i,j} q_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}) &
    &= -2\sum_{i,j} q_{ij} \eq_i \eta^+_{j} \\
    &= 2\sum_{i,j} q_{ij} \eq_i (\eta^-_{j} - \eta^-_{i}) &
    &= 2\sum_{i,j} q_{ij} \eq_i \eta^-_{j}.
\end{aligned}
\end{equation}
%
We can also express it in terms of the fundamental matrix \eqref{eq:funddef} as
%
\begin{equation}\label{eq:areaZ}
  \hat{A} = \sum_{i,j,k,l} q_{ij} \pib_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
    = \pib \fund q \fund \w.
\end{equation}
%

It is also helpful to define the following quantities:
%
\begin{equation}\label{eq:areacoeffs}
  \begin{aligned}
    c_k &= \diff{\ln\eq_k}{\alpha}
      = \sum_{ij} \eq_i q_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
      = - \prn{\eq q \fptb}_k
      = \frac{(\eq q \fund)_k}{\eq_k}, \\
    a_i &= \sum_j q_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
    \implies
    \hat{A} &= \sum_k c_k \eq_k \w_k
      = 2\sum_i a_i.
  \end{aligned}
\end{equation}
%
Note that the optimal choice of $\w$ is $\w_k = \sgn(c_k)$.

\subsection{Derivatives \wrt $\W^\pm$}\label{sec:deriv}

As discussed in \sref{sec:constraints},
we will regard the off-diagonal elements of $\W^\pm_{ij}$ to be the independent variables,
with $\W^\pm_{ii}=-\sum_{j \neq i} \W^\pm_{ij}$ imposed by hand.
Thus,
%
\begin{equation}\label{eq:basicderivs}
  \pdiff{\W^F_{ij}}{\W^\pm_{gh}} = f^\pm \delta_{gi}(\delta_{hj}-\delta_{ij}),
  \qquad
  \pdiff{q_{ij}}{\W^\pm_{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
\end{equation}
%
The implicit $g \neq h$ that comes with all derivatives is unnecessary, as the derivatives above vanish when $g=h$.

In particular, differentiating \eqref{eq:funddef},
%
\begin{equation}\label{eq:derivZ}
  \pdiff{\fund_{ij}}{\W^\pm_{gh}} = rf^\pm \fund_{ig} (\fund_{hj}-\fund_{gj}).
\end{equation}
%
We can then differentiate expression \eqref{eq:areaZ} to get
%
\begin{equation}\label{eq:derivA}
  \begin{aligned}
    \pdiff{\hat{A}}{\W^\pm_{gh}} =&\,
      rf^\pm \sum_{ijkl} q_{ij} \w_k \pib_l \brk{
         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
    &\pm \sum_{kl} \w_k \pib_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
    =&\,
      2rf^\pm \sum_{ij} q_{ij} \eq_i \eq_g \brk{
         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
         (\fptb_{ig}-\fptb_{jg})(\eta^+_{g}-\eta^+_{h}) } \\
      &\pm 2 \eq_g (\eta^+_g - \eta^+_h) \\
    =&\,
      2rf^\pm \sum_{ij} q_{ij} \eq_i \eq_g
         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
      \pm 2 \eq_g (\eta^+_g - \eta^+_h)\brk{1+\pdiff{\ln(\eq_g)}{\,\ln(f^\pm)}} \\
    =&\,
      2rf^\pm \eq_g \brk{\sum_i a_i (\fptb_{gi}-\fptb_{hi}) + c_g (\eta^+_{g}-\eta^+_{h})}
      \pm 2 \eq_g (\eta^+_g - \eta^+_h). \\
  \end{aligned}
\end{equation}
%
where $a_i$ and $c_k$ were defined in \eqref{eq:areacoeffs}.

It is sometimes useful to consider the following derivatives:
%
\begin{equation}\label{eq:derivqw}
  \begin{aligned}
    \pdiff{}{\W^F_{gh}} &\equiv     \pdiff{}{\W^+_{gh}} +     \pdiff{}{\W^-_{gh}} ,&\qquad
    \pdiff{}{q_{gh}}    &\equiv f^- \pdiff{}{\W^+_{gh}} - f^+ \pdiff{}{\W^-_{gh}} .
  \end{aligned}
\end{equation}
%
Each of these derivatives behaves as their names suggest:
%
\begin{equation}\label{eq:derivqweff}
  \pdiff{\W^F_{ij}}{\W^F_{gh}} = \pdiff{q_{ij}}{q_{gh}}
  =  \delta_{gi}(\delta_{hj}-\delta_{ij}),
  \qquad
  \pdiff{q_{ij}}{\W^F_{gh}} = \pdiff{\W^F_{ij}}{q_{gh}}  = 0.
\end{equation}
%
This is because we could treat $\W^F$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W^\pm$.

\subsubsection{Scaling mode}\label{sec:rescale}

Consider the following differential operator:
%
\begin{equation}\label{eq:scaleop}
  \Delta \equiv \sum_{g,h} \W^+_{gh}\pdiff{}{\W^+_{gh}} + \W^-_{gh}\pdiff{}{\W^-_{gh}}.
\end{equation}
%
This corresponds to the scaling, $\W^\pm \to (1+\epsilon)\W^\pm$.
Intuitively, this has two effects: it scales up the initial potentiation/depression and it scales down all timescales.
This intuition is confirmed by the following results:
%
\begin{equation}\label{eq:scaleeffects}
  \begin{aligned}
    \Delta \fund &= \tau\onev\eq - \fund ,&
    \Delta \eq  &= 0 ,&
    \Delta \fpt  &= -\fpt ,
    \\
    \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
    \Delta q_{ij} &= q_{ij} ,&
    \Delta \hat{A}  &= 0 ,&
  \end{aligned}
\end{equation}
%
The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W^\pm$.

As the area is invariant under this scaling, we can consider the $\W^\pm$ to be projective coordinates.
Therefore we don't need to enforce the constraint \eqref{eq:Wmax} while looking for the maximum area, as we can use this null-mode to enforce it afterwards without changing the area.
We also don't have to worry about the boundary at infinity,
as sending some of them to infinity is equivalent to sending the rest to zero.

\subsection{Kuhn-Tucker conditions}\label{sec:kuhntucker}

Consider the Lagrangian
%
\begin{equation}\label{eq:lagrangian}
  \CL = \hat{A} + \sum_{\pm}\sum_{i\neq j} \mu^\pm_{ij} \W^\pm_{ij} +\lambda\eq\w.
\end{equation}
%
The last term ensures that we hold $\eq_\pm$ fixed during this extremisation, so that we can safely ignore the factors that we dropped from the noise in \eqref{eq:noisepm}.
Necessary conditions for an extremum are
%
\begin{equation}\label{eq:extremum}
  \pdiff{\CL}{\W^\pm_{gh}} = 0,
  \qquad
    \mu^\pm_{gh} \geq 0,\quad
    \W^\pm_{gh} \geq 0,\quad
    \mu^\pm_{gh}\W^\pm_{gh} = 0.
\end{equation}
%
with $g \neq h$. This enforces the constraints \eqref{eq:Wmin}, but not \eqref{eq:Wmax}. As discussed in \sref{sec:constraints} and \sref{sec:rescale}, that can be enforced after finding the maximum with the null scaling degree of freedom.

\subsubsection{Triangularity}\label{sec:triangular}

Consider
%
\begin{equation}\label{eq:shiftqderiv}
  \pdiff{\CL}{q_{gh}} =
  f^- \pdiff{\CL}{\W^+_{gh}} - f^+ \pdiff{\CL}{\W^-_{gh}}
   = (f^- \mu^+_{gh} - f^+ \mu^-_{gh}) + 2\eq_g (\eta^+_g - \eta^+_h)
   = 0.
\end{equation}
%
This corresponds to the shift
%
\begin{equation}\label{eq:shiftq}
  \W^+_{ij} \to \W^+_{ij} + f^-\epsilon_{ij},
  \qquad
  \W^-_{ij} \to \W^-_{ij} - f^+\epsilon_{ij},
  \qquad
  \sum_j \epsilon_{ij} = 0,
\end{equation}
%
which leaves $\W^F$ unchanged, and therefore $\eq$, $\fpt$ and $\eta^\pm$ as well.

Assume $\eta^+_g > \eta^+_h$. Then
%
\begin{equation}\label{eq:lowertriangular}
 f^- \mu^+_{gh} - f^+ \mu^-_{gh} <0
 \qquad\implies\qquad
 \mu^-_{gh} >0
 \qquad\implies\qquad
 \W^-_{gh}=0.
\end{equation}
%
Similarly, if $\eta^+_g < \eta^+_h$, then
%
\begin{equation}\label{eq:uppertriangular}
 f^- \mu^+_{gh} - f^+ \mu^-_{gh} > 0
 \qquad\implies\qquad
 \mu^+_{gh} >0
 \qquad\implies\qquad
 \W^+_{gh}=0.
\end{equation}
%
Thus, if we arrange the states in order of decreasing $\eta^+$, $\W^+$ is upper-triangular and $\W^-$ is lower triangular.

We have ignored the possibility that $\eq_g=0$, as this would imply that $\fpt_{ig}=\infty$, which would in turn imply that the Markov process is not ergodic.

%Note that if we multiply \eqref{eq:shiftqderiv} by $q_{gh}$ and sum over $(g,h)$, we get
%%
%\begin{equation}\label{eq:stupid}
%  \hat{A} = \sum_\pm\sum_{g\neq h} \mu^\pm_{gh} \W^F_{gh},
%\end{equation}
%%
%which is probably useless

\subsubsection{Increasing $c_k$}\label{sec:areacoeffincr}

Consider the following combinations of derivatives:
%
\begin{align}
%  \begin{aligned}
\label{eq:areacoeffincrderiv}
    \Delta_{gh} &\equiv
      \frac{1}{\eq_{g}} \pdiffc{}{\W^F_{gh}}
      + \frac{1}{\eq_h} \pdiffc{}{\W^F_{hg}}, \\
%  \end{aligned}
\end{align}
%
Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
%
\begin{equation}\label{eq:sareacoeffincrprob}
  \Delta_{gh} \eq_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:areacoeffincrfinite}
  \W^\pm \to \W^\pm + \D\boldsymbol{\epsilon},
  \qquad
  \begin{aligned}
  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
  \\
  \boldsymbol{\epsilon} \onev &= 0.
  \end{aligned}
\end{equation}
%
But more interestingly:
%
\begin{align}
%  \begin{aligned}
\label{eq:areacoeffincr}
    \Delta_{gh}\CL &=
      \frac{\mu^+_{gh}+\mu^-_{gh}}{\eq_g} + \frac{\mu^+_{hg}+\mu^-_{hg}}{\eq_h}
      + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
%  \end{aligned}
\end{align}
%
where $c_k$ were defined in \eqref{eq:areacoeffs}.

Using the non-negativity of the Kuhn-Tucker multipliers, $\mu^\pm_{ij}$, \eqref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).

Note that, according to \sref{sec:triangular}, either $\W^+_{gh}$ or $\W^-_{gh}$ will be zero at the maximum, therefore we can expect one of $\mu^+_{gh}+\mu^-_{gh}$ to be non-zero.
This would rule out degeneracy of the $c_k$ or $\eta^+_k$.
Looking at \eqref{eq:shiftqderiv} closely, the only way $\mu^+_{gh}+\mu^-_{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.

\subsubsection{Shortcuts}\label{sec:shortcuts}

Now consider the following combinations of derivatives for $m>1$:
%
\begin{align}
%  \begin{aligned}
\label{eq:shortcutderiv}
    \widetilde{\Delta}^\pm_{g,m} &\equiv
      \brk{ \sum_{k=0}^{m-1} \frac{1}{\eq_{g\pm k}} \pdiffc{}{\W^\pm_{g\pm k,g\pm(k+1)}} }
      - \frac{1}{\eq_g} \pdiffc{}{\W^\pm_{g,g\pm m}}.
%  \end{aligned}
\end{align}
%
Once again, they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).

One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
%
\begin{equation}\label{eq:shortcutprob}
  \widetilde{\Delta}^\pm_{g,m} \eq_i = 0,
\end{equation}
%
which makes it possible to integrate the perturbation:
%
\begin{equation}\label{eq:shortcutfinite}
  \W^\pm \to \W^\pm + \D\boldsymbol{\epsilon}^{\pm(g,m)},
  \qquad
  \begin{aligned}
    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g,g\pm m}
      &\!\!\!\!\!\!&= -\epsilon,\\
    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm(k+1)}
      &\!\!\!\!\!\!&= \epsilon
        &\;
        &\forall\, k \in [0,m-1],\\
    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm k}
      &\!\!\!\!\!\!&= -\epsilon
        &
        &\forall\, k \in [1,m-1].
  \end{aligned}
\end{equation}
%
But more interestingly for our purposes:
%
\begin{align}
%  \begin{aligned}
\label{eq:shortcutarea}
    \widetilde{\Delta}^\pm_{g,m}\CL &=
      \brk{ \sum_{k=0}^{m-1} \frac{\mu^\pm_{g\pm k,g\pm(k+1)}}{\eq_{g\pm k}}
      - \frac{\mu^\pm_{g,g\pm m}}{\eq_g}}
      + 2rf^\pm \sum_{k=0}^{m-1} \prn{\eta^+_{g\pm k} - \eta^+_{g\pm(k+1)}} \prn{c_{g\pm k} - c_{g}},
%  \end{aligned}
\end{align}
%

If we put the states in order of decreasing $\eta^+_k$, the results of the \sref{sec:areacoeffincr} tell us that the $c_k$ are non-decreasing.
This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu^\pm_{g,g\pm m}>0$, which in turn implies that $\W^\pm_{g,g\pm m}=0$.
This would tell us that the process with the maximal area has to have a multi-state topology.

%From the expression for the area in \eqref{eq:areacoeffs}, it will clearly also be optimal to have the $\w_k$ non-decreasing, \ie all the positive $\w_k$ should lie ahead of all the negative $\w_k$.

\subsubsection{Summary}\label{sec:KTsummary}

Using the Kuhn-Tucker formalism, we have shown that,
with the states arranged in order of non-increasing $\eta^+_i$:
%
\begin{itemize}
  \item There can be no ergodic maximum for which $\W^+$ contains backwards transitions or $\W^-$ contains forwards transitions.
  \item There can be no ergodic maximum with the $c_k$ decreasing.
  \item The $c_k$ may only be degenerate at an ergodic maximum if the corresponding $\eta^+_k$ are also degenerate.
  \item If the $c_k$ increase and the $\eta^+_i$ decrease, there can be no ergodic maximum with shortcuts.
\end{itemize}
%
These were shown by finding allowed perturbations that increase the area.

%Suppose we start with some arbitrary ergodic process. If $\W^\pm$ contain backwards/forwards transitions or the $c_k$ decrease, then we can increase the area by applying the perturbations \eqref{eq:shiftqderiv} and \eqref{eq:areacoeffincrderiv}.
%The second perturbation will generate the unwanted backwards/forwards transitions, requiring subsequent application of first perturbation.
%The first perturbation will change the $c_k$, possibly changing their order, requiring subsequent application of second perturbation.
%
%This procedure can end in several ways:
%%
%\begin{itemize}
%  \item Applying \eqref{eq:areacoeffincrderiv} an infinite amount sends two transition rates to infinity, which due to the scaling mode (\sref{sec:rescale}), is equivalent to sending all other transitions to zero.
%      Subsequent application of \eqref{eq:shiftqderiv} just removes the backwards transition from $\W^+$ and the forward transition from $\W^-$.
%  \item Applying \eqref{eq:areacoeffincrderiv} ends in a process with degenerate $c_k$ or $\eta^+_k$, and the subsequent application of \eqref{eq:shiftqderiv} doesn't change this degeneracy.
%  \item As above, but the subsequent application of \eqref{eq:shiftqderiv} lifts the degeneracy, resulting in a chain with increasing $c_k$.
%\end{itemize}
%%
%In the third case, we can then apply the perturbation \eqref{eq:shortcutderiv} to remove all the shortcuts.
%Either this, or the first case, results in a multi-state topology, possibly disconnected or without the full set of states.

This leaves two possibilities for the maximum area Markov chain.
Either there is no degeneracy and no shortcuts, which implies the Multi-state topology that we'll discuss in \sref{sec:multistate}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \eqref{eq:shortcutarea}).

Degeneracy tends to be very delicate. It is usually hard to arrange without some symmetry relating degenerate state. Such a symmetry would imply lumpability (see \sref{sec:lump}). The lumped chain would not have any shortcuts, as an entire degenerate set cannot be bypassed. As this lumped chain has the same area (see \sref{sec:SNRlump}), we would need only consider the multi-state topology.

\subsection{Multi-state topology}\label{sec:multistate}

The multi-state topology is defined by:
%
\begin{equation}\label{eq:multistatedef}
  \W^+_{ij} = q^+_i \delta_{i+1,j},
  \qquad
  \W^-_{ij} = q^-_j \delta_{i,j+1}.
\end{equation}
%
It saturates various inequalities:
%
\begin{equation}\label{eq:multistateineq}
  \begin{aligned}
    \fptb_{ik} - \fptb_{jk} &=
      \begin{cases}
        \fptb_{ij},  &\text{if}\quad i \leq j \leq k \quad\text{or}\quad i \geq j \geq k,\\
        -\fptb_{ji}, &\text{if}\quad j \leq i \leq k \quad\text{or}\quad j \geq i \geq k,\\
      \end{cases} \\
    r\eq_i \W^F_{ij} \prn{\fptb_{ij}+\fptb_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
  \end{aligned}
\end{equation}
%
and it satisfies detailed balance (a.k.a.\ reversibility a.k.a. $\CL^2_{\eq}$ self-adjointness):
%
\begin{equation}\label{eq:multistateprob}
  f^+ q^+_i \eq_i = f^- q^-_i \eq_{i+1},
\end{equation}
%
which means we can always choose the transition rates, $q^\pm_i$, to give any desired equilibrium probabilities, $\eq_i$.

This allows us to calculate the $c_k$'s:
%
\begin{equation}\label{eq:areacoeffchain}
\begin{aligned}
  c_k =&\, \sum_{i<k} \fpt_{i,i+1} \prn{\eq_{i}q^+_i+\eq_{i+1}q^-_i}
    %\\&
    - \sum_{i\geq k} \fpt_{i+1,i} \prn{\eq_{i}q^+_i+\eq_{i+1}q^-_i}
  ,\\
  c_{k+1} - c_k =&\, \prn{\fpt_{k,k+1}+\fpt_{k+1,k}} \prn{\frac{\eq_{k}\W^F_{k,k+1}}{f^+}+\frac{\eq_{k+1}\W^F_{k+1,k}}{f^-}}
    =%\\=&\,
    \frac{1}{rf^+f^-},\\
  \sum_k c_k \eq_k =&\, \sum_{ij} \eq_i q_{ij} (\eta-\eta) = 0,\\
%  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f^+f^-},
  \implies c_k =&\, \frac{k - \sum_j j\eq_j}{rf^+f^-},
\end{aligned}
\end{equation}
%
where we used \eqref{eq:multistateineq} to derive the first two equations respectively and Th.\ref{th:kemenyconst} to derive the third. This allows us to write the area as
%
\begin{equation}\label{eq:multistatearea}
 % A = 2\sqrt{N} \sum_k \brk{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \eq_k \w_k.
  A = \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k
    = \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
\end{equation}
%
where we used $\w_k=\sgn(c_k)$, as discussed after \eqref{eq:areacoeffs}.
%It will also help to define $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
%Note that
%%
%\begin{equation}\label{eq:allpplusorminus}
%  \begin{aligned}
%    \eq_+ &=1
%     \quad&\implies\quad
%     \eq_k \w_k &= \eq_k
%     \quad&\implies\quad
%     A=0,\\
%    \eq_- &=1
%     \quad&\implies\quad
%     \eq_k \w_k &= -\eq_k
%     \quad&\implies\quad
%     A=0,
%  \end{aligned}
%\end{equation}
%%
%neither of which are optimal.
%
%Consider the Lagrangian
%%
%\begin{equation}\label{eq:multistatelagrangian}
%  \CL = \frac{A}{2\sqrt{N}} + \lambda\prn{1-\sum_i \eq_i} + \sum_i \mu_i \eq_i,
%\end{equation}
%%
%where $\lambda$ is a Lagrange multiplier and $\mu_i$ are Kuhn-Tucker multipliers (see \eqref{eq:extremum}). Extremising \wrt $\eq_i$:
%%
%\begin{equation}\label{eq:multistateext}
%%  \pdiff{\CL}{\eq_i} = \brk{\prn{i-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \w_i
%%    - \prn{i-\frac{n+1}{2}}\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
%  \pdiff{\CL}{\eq_i} = \brk{i - \sum_j j\eq_j} \w_i
%    - i\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
%\end{equation}
%%
%Consider an $i$ such that $\w_i=\w_{i+1}$:
%%
%\begin{equation}\label{eq:multistatemudiff}
%  \mu_{i+1} - \mu_i = \prn{\eq_+-\eq_-} - \w_i.
%\end{equation}
%%
%As $\abs{\eq_+-\eq_-}<1$ (unless we want $A=0$, \eqref{eq:allpplusorminus}), this means the $\mu_i$ increase amongst the negative $\w_i$ and decrease amongst the positive $\w_i$.
%Therefore the only non-zero equilibrium probabilities at optimum are $\eq_1$ and $\eq_n$.

First let us maximise \eqref{eq:multistatearea} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
Clearly this will happen when we put all of the probability at the ends: $\eq_1=\eq_-$ and $\eq_n=\eq_+$ are the only non-zero $\eq_k$.
This gives an area of
\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_1\eq_n)}(n-1)/r$, which would have no effect on what follows.}

%
\begin{equation}\label{eq:multistateextarea}
  A \leq \frac{\sqrt{N}}{r}(n-1)\prn{4\eq_1\eq_n}.
\end{equation}
%
This is maximised at $\eq_1=\eq_n=\half$:
%
\begin{equation}\label{eq:maxarea}
  A \leq \frac{\sqrt{N}}{r}(n-1).
\end{equation}
%
Note that this is not an ergodic chain, the two states at each end are absorbing.

\appendix
\section{Other quantities}

Here we collect useless facts.

\subsection{Differences in $c_k$'s}\label{sec:areacoeffdiff}

In the general case, making full use of \cite[Cor.6.2.7]{kemeny1960finite}, we can show that:
%
\begin{equation}\label{eq:areacoeffdiff}
  \begin{aligned}
    c_{k+1} - c_k &= \frac{N^{k+1}_{kk}}{\eq_k} \brk{
      \sum_{i<j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{H^{k+1}_{i,k} + H^{k}_{j,k+1} - 1}
      +\sum_{i>j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{H^{k+1}_{j,k} + H^{k}_{i,k+1} - 1}
    } \\
    &= \frac{N^{k+1}_{kk}}{\eq_k} \brk{
      \sum_{i<j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{1 - H^{k+1}_{j,k} + H^{k}_{i,k+1}}
      +\sum_{i>j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{1 - H^{k+1}_{i,k} + H^{k}_{j,k+1}}
    }, \\
    \eta^+_{k} - \eta^+_{k+1} &=
      \fpt_{k,k+1} \eq_+ - \sum_{\set{i}{\w_i>0}} N^{k+1}_{ii} \prn{1 - H^{i}_{k,k+1}} \\&=
      \sum_{\set{i}{\w_i>0}} N^{k}_{ii} \prn{1 - H^{i}_{k+1,k}} - \fpt_{k+1,k} \eq_+ \\
    = \eta^-_{k+1} - \eta^-_{k} &=
      \fpt_{k+1,k} \eq_- - \sum_{\set{i}{\w_i<0}} N^{k}_{ii} \prn{1 - H^{i}_{k+1,k}} \\&=
      \sum_{\set{i}{\w_i<0}} N^{k+1}_{ii} \prn{1 - H^{i}_{k,k+1}} - \fpt_{k,k+1} \eq_- .
  \end{aligned}
\end{equation}
%
We need to show that the positivity of one ($\forall k$) implies that the positivity of the other.

\subsection{Derivatives \wrt $\fptb$}\label{sec:derivT}

The Markov chain, $\W^F$, is completely determined by the off-diagonal mean first passage times, $\fptb$, \cite[Th.4.4.12]{kemeny1960finite}:
%
\begin{equation}\label{eq:fptdet}
  \begin{aligned}
    \eq &= \eta \prn{\fptb\inv\onev}\trans, \\
    r\W^F= \MM &= (\D-\onem)\fptb\inv.
  \end{aligned}
\end{equation}
%
where $\eta$ can be determined by normalising $\eq$ and $\D$ is determined by $\eq$, \eqref{eq:pdotD}. Clearly $q$ will be undetermined.

We can then define derivatives \wrt $\fptb$:
%
\begin{equation}\label{eq:derivTdef}
  \pdiff{}{\fptb_{gh}} = \sum_{ij} \pdiffc{\W^F_{ij}}{\fptb_{gh}} \pdiff{}{\W^F_{ij}},
\end{equation}
%
where
%
\begin{equation}\label{eq:derivTWp}
  \pdiff{\W^F_{ij}}{\fptb_{gh}} = r \eq_h \W^F_{ig} \prn{\W^F_{ij}-\W^F_{hj}} ,
  \qquad
  \pdiff{\eq_{i}}{\fptb_{gh}} = -r \eq_h \eq_i \W^F_{ig} .
\end{equation}
%
These derivatives of $q$ will vanish.

This means that the derivatives of the area are:
%
\begin{equation}\label{derivTarea}
  \pdiff{\hat{A}}{\fptb_{gh}} = -\eq_h\brk{
    \sum_i \eq_i q_{ig} \w_h
    + r\sum_{ijk} \prn{\W^F_{ig}+\W^F_{kg}} \eq_i q_{ij} \prn{\fptb_{ik}-\fptb_{jk}} \eq_k \w_k
    }.
\end{equation}
%

\subsection{Initial SNR}\label{sec:initSNR}

Let us write $\SNR(0) = \sqrt{N} (2f^+f^-) \hat{I}$. Using \eqref{eq:SNRcurve}, we can write
%
\begin{equation}\label{eq:initSNR}
  \begin{aligned}
    \hat{I} &= \diff{\eq}{\alpha}(-\W^F)\w &
      &= - \pib \fund \diff{\MM}{\alpha} \fund \W^F \w &
      &= - \pib \fund \diff{\W^F}{\alpha} \fund \MM \w \\
      &= \pib \fund \diff{\W^F}{\alpha} (\I - \tau\onev\pib) \w &
      &= \pib \fund \diff{\W^F}{\alpha} \w &
      &= \eq \diff{\W^F}{\alpha} \w .
  \end{aligned}
\end{equation}
%
We can also write this in component form
%
\begin{equation}\label{eq:initSNRcmpt}
  \begin{aligned}
     \hat{I} &= \sum_{i,j} \eq_i q_{ij} \w_j \\
      &= \sum_{i \neq j} \eq_i q_{ij} \w_j + \sum_i \eq_i q_{ii} \w_i \\
      &= \sum_{i \neq j} q_{ij} \eq_i (\w_j - \w_i).
  \end{aligned}
\end{equation}
%
We can differentiate the expression in terms of $\fund$ to get
%
\begin{equation}\label{eq:initSNRdiff}
 \begin{aligned}
  \pdiff{\hat{I}}{\W^\pm_{gh}}
    &= rf^\pm \sum_{ijk} \pib_k \fund_{kg} (\fund_{hi}-\fund_{gi}) q_{ij} \w_j
    \pm \sum_k \pib_k \fund_{kg} (\w_h - \w_g) \\
    &= rf^\pm \sum_{ij} \eq_g q_{ij} (\fptb_{gi}-\fptb_{hi}) \eq_i \w_j
    \pm \eq_g (\w_h - \w_g).
 \end{aligned}
\end{equation}
%
If we wished to maximise the area with fixed initial SNR, $\hat{I}=I_0$, we would add $\beta(\hat{I}-I_0)$ to the Lagrangian \eqref{eq:lagrangian} and extremise wrt $\beta$.

\subsection{Fixing mixing time (Kemeny's  constant)}\label{sec:fixmix}

The mixing time, $\eta$ \eqref{eq:mixdef}, can be thought of as the mean time to pass between two states drawn from the equilibrium distribution \cite{hunter2006mixing,levene2002kemeny}. It should be a measure of the time it takes to reach equilibrium, and therefore related to the time it takes the SNR to decay. It could be interesting to hold it fixed while maximising the area, as this could help prevent degenerate solutions with very low initial SNR but long decay time. This can be done by adding $\gamma(\eta-\eta_0)$ to the Lagrangian \eqref{eq:lagrangian} and extremising wrt $\gamma$.

It would help to calculate its derivatives.
%
\begin{equation}\label{eq:mixingdiff}
  \begin{aligned}
    \pdiff{\eta}{\W^\pm_{gh}} &= \pdiff{}{\W^\pm_{gh}}\brk{\sum_i \fund_{ii} - \tau} \\
      &= rf^\pm \sum_i \fund_{ig}(\fund_{hi}-\fund_{gi}) \\
      &= rf^\pm (\fund^2_{hg} - \fund^2_{gg}).
  \end{aligned}
\end{equation}
%

\subsection{Scaling mode}\label{sec:rescaleconstr}

Consider the following differential operator:
%
\begin{equation}\label{eq:scaleopapp}
  \Delta \equiv \sum_{g,h} \W^+_{gh}\pdiff{}{\W^+_{gh}} + \W^-_{gh}\pdiff{}{\W^-_{gh}}.
\end{equation}
%
This corresponds to the scaling, $\W^\pm \to (1+\epsilon)\W^\pm$. Intuitively, this has two effects: it scales up the initial potentiation/depression and it scales down all timescales. This intuition is confirmed by the following results:
%
\begin{equation}\label{eq:scaleeffectsapp}
  \begin{aligned}
    \Delta \fund &= \tau\onev\eq - \fund ,&
    \Delta \eq  &= 0 ,&
    \Delta \fpt  &= -\fpt ,&
    \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
    \Delta q_{ij} &= q_{ij} ,
    \\&&
    \Delta \hat{A}  &= 0 ,&
    \Delta \hat{I} &= \hat{I} ,&
    \Delta \eta  &= -\eta .
  \end{aligned}
\end{equation}
%
The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W^\pm$.

Suppose we don't hold anything fixed when maximising the area. Then we'd have
%
\begin{equation*}
  \Delta \CL = 0,
\end{equation*}
%
i.e. this scaling is a null direction. Area maximisation can only determine the ratios of transition rates, not their absolute values.

Now suppose we only fix $\hat{I}$ when maximising the area. Then we'd have
%
\begin{equation*}
  \Delta \CL = \beta \hat{I} = \beta I_0 = 0.
\end{equation*}
%
As the Lagrange multiplier is zero, the constraint doesn't change the derivatives of the Lagrangian wrt $\W^\pm$. The sole effect of the constraint would be to fix the scaling mode. It wouldn't change the ratios of transition rates, topology, etc.

Exactly the same conclusions would follow if we only fixed $\eta$ when maximising the area. This would also only fix the scaling mode. However, if we held them both fixed, it would have more of an effect:
%
\begin{equation}\label{eq:scaleconstr}
  \Delta \CL = \beta \hat{I} - \gamma \eta = \beta I_0 - \gamma \eta_0 = 0.
\end{equation}
%
The Lagrange multipliers are non-zero, so this could affect the topology of the optimal Markov chain.

%\subsection*{Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection*{Appendices}
%\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{utcaps_sl}
\bibliography{maths}

\end{document}

### 2012-06-20 Code/CascMSFig2.m
rearranged panels

@@ -25,4 +25,4 @@ legend({'Cascade','Casc MS','Alt MS'},'Location','Best')
 embiggen
 set(gcf, 'PaperPositionMode', 'auto');
 %%
-clear Wp Wm w x beta n
+clear Wp Wm w x beta n q qm

### 2012-06-26 Notes/ContTime.tex
all figs done
lumpability refs

@@
 \input{newsymb.tex}
-%logic
-\newcommand{\means}{\Longleftrightarrow}
-\newcommand{\requires}{\Longleftarrow}
 %matrices
@@
 One may ask if this new process is a Markov chain.
-The answer is yes, if the original Markov chain has a property called lumpability \wrt the partition \cite[\S6.3]{kemeny1960finite}:
+The answer is yes, if the original Markov chain has a property called lumpability \wrt the partition 
+(see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time):
 %

### 2012-11-12 Cosyne/Abstract.tex
started cosyne abstract

% -*- TeX -*- -*- US -*-
\documentclass[12pt]{article}
\input{sl_preamble.tex}
%\input{sl_graphics_preamble.tex}
\usepackage{abstract}
\renewcommand{\abstractname}{Summary}
\renewcommand{\abstractnamefont}{\normalfont\large\bfseries}
\renewcommand{\abstracttextfont}{\normalfont}
\renewcommand{\absnamepos}{flushleft}
\setlength{\absleftindent}{0in}
\setlength{\absrightindent}{0in}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
%\input{sl_symbols.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Learning and memory with complex synapses}
%
% Author List:
%
\author{Subhaneil Lahiri\thanks{\emaillink{sulahiri@stanford.edu}}~ and Surya Ganguli\thanks{\emaillink{sganguli@stanford.edu}}\\
%
\small{
Department of Applied Physics, Stanford University, Stanford CA
}
}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
  We consider the storage of long term memories through synaptic modifications in existing networks. Recent experimental work suggests that single synapses are digital, in the sense that, from the perspective of extracellular physiology, they can only take on a finite number of discrete values for their strength. This imposes catastrophic limits on the memory capacity of classical models of memory that have relied on a continuum of analog synaptic strengths. However, synapses have many internal molecular states, suggesting we should model synapses themselves as complex molecular networks, rather than by a single scalar value, or strength. We develop new theorems bounding the memory capacity of such complex synaptic models and describe the structural organization of internal molecular networks necessary for achieving these limits. 
\end{abstract}

\subsection*{Additional detail}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliographystyle{utcaps_sl}
%\bibliography{maths,neuro}

\end{document}

### 2012-11-12 Notes/ContTime.tex
started cosyne abstract

@@
 
+We can substitute this into \eqref{eq:fptrecdg} to get a recursion relation for the off diagonal part:
+%
+\begin{equation}\label{eq:fptbrec}
+  \MM\fptb = \D-\onem.
+\end{equation}
+%
 \begin{thm}
-  The solution of \eqref{eq:fptrecdg} is unique.
+  The off-diagonal solution of \eqref{eq:fptbrec} is unique.
 \end{thm}
 \begin{proof}
-  We have already seen that the diagonal part is determined by \eqref{eq:fptrecdg}.
   Suppose we have two solutions for the off-diagonal part, $\fptb$ and $\fptb'$.
-  Subtracting \eqref{eq:fptrecdg} for these two solutions gives
+  Subtracting \eqref{eq:fptbrec} for these two solutions gives
   %
   \begin{equation*}
-    \Lb\MM(\fptb-\fptb')=0.
+    \MM(\fptb-\fptb')=0.
   \end{equation*}
@@
 \begin{proof}
-  We just need to show that this satisfies \eqref{eq:fptrecdg} with \eqref{eq:recurtime}, and that the diagonal elements vanish.
-  Premultiply \eqref{eq:fptfund} by $\Lb\MM$ and subtract from \eqref{eq:fptrecdg}:
+  We just need to show that this satisfies \eqref{eq:fptbrec} and that the diagonal elements vanish.
+  Premultiply \eqref{eq:fptfund} by $\MM$:
   %
@@
   \begin{aligned}
-    \fpt\dg - \Lb\MM\fptb &= \Lb(\I+\MM\fund)\D\\
-     &= \Lb\onev\eq\D \\
-     &= \Lb\onem,
+    \MM\fptb &= -\MM\fund\D\\
+     &= (\I-\onev\eq)\D \\
+     &= \D-\onem,
   \end{aligned}
@@
 Degeneracy tends to be very delicate.
-It is usually hard to arrange without some symmetry relating degenerate state.
+It is usually hard to arrange without some symmetry relating degenerate states.
 Such a symmetry would imply lumpability (see \sref{sec:lump}).
@@
 
-The multi-state topology is defined by:
+The multi-state topology is defined by (see \cite{amit1994learning,Fusi2007multistate}):
 %
@@
 This gives an area of
-\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_1\eq_n)}(n-1)/r$, which would have no effect on what follows.}
+\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_+\eq_-)}(n-1)/r$, which would have no effect on what follows.}
 
@@
 \begin{equation}\label{eq:multistateextarea}
-  A \leq \frac{\sqrt{N}}{r}(n-1)\prn{4\eq_1\eq_n}.
-\end{equation}
-%
-This is maximised at $\eq_1=\eq_n=\half$:
+  A \leq \frac{\sqrt{N}}{r}(n-1)\prn{4\eq_+\eq_-}.
+\end{equation}
+%
+This is maximised at $\eq_+=\eq_-=\half$:
 %
@@
 %
-Note that this is not an ergodic chain, the two states at each end are absorbing.
+Note that the chain that achieves this is not ergodic, the two states at each end are absorbing.
+This is similar to the results found numerically in \cite{Barrett2008discrete} in a slightly different situation.
 
@@
 \bibliographystyle{utcaps_sl}
-\bibliography{maths}
+\bibliography{maths,neuro}
 

### 2012-11-12 Code/Calc/Kemeney.m

@@ -1,0 +1,15 @@
+function [ eta ] = Kemeney( W )
+%eta=KEMENEY(W) calculate Kemeney's constant
+%   eta = sum_j T_ij p_j = Tr Z - tau 
+
+error(CheckSize(W,@ismat));%matrix
+error(CheckSize(W,@issquare));%square
+
+
+qa=eig(-W);
+qa=sort(qa);
+
+eta = sum(1./qa(2:end));
+
+end
+

### 2012-11-12 Code/Tests/PlotClique.m

@@ -4,7 +4,7 @@
 
 %allvars={'crqm','crq','c2q','init','cqm','ev','L2','AI','Eqbnd','Tmax'};
 % varnames={'crq','crqm','c2q'};
-varnames={'crq','Eqbnd','init','ev'};
+varnames={'crq','qPi','init','A','etaPi'};
 numpts=100;
 n=zeros(1,numpts);
 output=zeros(length(varnames),numpts);

### 2012-11-12 Code/Tests/TestClique.m

@@ -8,8 +8,8 @@ assert(mod(n,2)==0);
 % [wp,wm,w]=CliqueChain(n);
 
 q=ones(1,n-1);
-% q(n/2+1)=1/n;
-q(1)=1/n;
+q(n/2+1)=1/n;
+% q(1)=1/n;
 [wp,wm,w]=MakeSMS(q);
 
 % q=ones(1,n-1);
@@ -96,6 +96,11 @@ if tf
     output(loc)=(c.*q)'*recqaqb*(c.*q);
 end
 %
+[tf,loc]=ismember('A',varnames);
+if tf
+    output(loc)=real(sum(c));
+end
+%
 [tf,loc]=ismember('AI',varnames);
 if tf
     output(loc)=real(sum(c)*c'*q);
@@ -111,6 +116,17 @@ end
 if tf
     T=FPT((wp+wm)/2);
     output(loc)=max(max(T));
+end
+%
+[tf,loc]=ismember('qPi',varnames);
+if tf
+    output(loc)=-w'*(wp+wm)*w/(2*n);
+end
+%
+[tf,loc]=ismember('etaPi',varnames);
+if tf
+    Zinv=ones(size(wp))-(wp+wm)/2;
+    output(loc)=w'*(Zinv\w)/(2*n);
 end
 % allvars={'crqm','crq','c2q','init','cqm','ev','L2','AI','Eqbnd','Tmax'};
 end

### 2012-11-12 Cosyne/Abstract.tex
bit more of cosyne abstract

@@
 %\input{sl_graphics_preamble.tex}
-\usepackage{abstract}
-\renewcommand{\abstractname}{Summary}
-\renewcommand{\abstractnamefont}{\normalfont\large\bfseries}
-\renewcommand{\abstracttextfont}{\normalfont}
-\renewcommand{\absnamepos}{flushleft}
-\setlength{\absleftindent}{0in}
-\setlength{\absrightindent}{0in}
+%\usepackage{abstract}
+%\renewcommand{\abstractname}{Summary}
+%\renewcommand{\abstractnamefont}{\normalfont\large\bfseries}
+%\renewcommand{\abstracttextfont}{\normalfont}
+%\renewcommand{\absnamepos}{flushleft}
+%\setlength{\absleftindent}{0in}
+%\setlength{\absrightindent}{0in}
 % ----------------------------------------------------------------
@@
 
-\begin{abstract}
+\subsection*{Summary}
+
 We consider the storage of long term memories through synaptic modifications in existing networks. 
@@
 We develop new theorems bounding the memory capacity of such complex synaptic models and describe the structural organization of internal molecular networks necessary for achieving these limits.
-\end{abstract}
 
@@
 
+Much of the previous work on complex synapses has focused on the study of specific models \cite{amit1994learning,Fusi2005cascade,Fusi2007multistate}.
+If we wish to understand the structure of the molecular networks responsible for synaptic plasticity (see \cite{Coba2009phosphorylation}), it will be vital to have explored the space of all possible molecular networks so that we can develop a correspondence between features of these networks and desirable/undersirable properties of the synapse.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%\bibliographystyle{utcaps_sl}
-%\bibliography{maths,neuro}
+\bibliographystyle{utcaps_sl}
+\bibliography{maths,neuro}
 

### 2012-11-13 Cosyne/Abstract.tex
more cosyne abstract, incl figures

@@
 \input{sl_preamble.tex}
-%\input{sl_graphics_preamble.tex}
+\input{sl_graphics_preamble.tex}
+\graphicspath{{Figures/}}
 %\usepackage{abstract}
@@
 %
-\author{Subhaneil Lahiri\thanks{\emaillink{sulahiri@stanford.edu}}~ and Surya Ganguli\thanks{\emaillink{sganguli@stanford.edu}}\\
-%
-\small{
-Department of Applied Physics, Stanford University, Stanford CA
-}
-}
+\author{Subhaneil Lahiri and Surya Ganguli}
+%\author{Subhaneil Lahiri\thanks{\emaillink{sulahiri@stanford.edu}}~ and Surya Ganguli\thanks{\emaillink{sganguli@stanford.edu}}\\
+%%
+%\small{
+%Department of Applied Physics, Stanford University, Stanford CA
+%}
+%}
+\date{}
 
@@
 
+%\begin{abstract}
 \subsection*{Summary}
@@
 Recent experimental work suggests that single synapses are digital, in the sense that, from the perspective of extracellular physiology, they can only take on a finite number of discrete values for their strength.
-This imposes catastrophic limits on the memory capacity of classical models of memory that have relied on a continuum of analog synaptic strengths. 
-However, synapses have many internal molecular states, suggesting we should model synapses themselves as complex molecular networks, rather than by a single scalar value, or strength. 
+This imposes catastrophic limits on the memory capacity of classical models of memory that have relied on a continuum of analog synaptic strengths [Amit and Fusi, Network: Computation in Neural Systems (1992), Neural Computation (1994)]\nocite{amit1992constraints,amit1994learning}.
+
+However, synapses have many internal molecular states [Coba et al., Sci Signal (2009)]\nocite{Coba2009phosphorylation}, suggesting we should model synapses themselves as complex molecular networks, rather than by a single scalar value, or strength.
 We develop new theorems bounding the memory capacity of such complex synaptic models and describe the structural organization of internal molecular networks necessary for achieving these limits.
+%\end{abstract}
 
@@
 Much of the previous work on complex synapses has focused on the study of specific models \cite{amit1994learning,Fusi2005cascade,Fusi2007multistate}.
-If we wish to understand the structure of the molecular networks responsible for synaptic plasticity (see \cite{Coba2009phosphorylation}), it will be vital to have explored the space of all possible molecular networks so that we can develop a correspondence between features of these networks and desirable/undersirable properties of the synapse.
+If we wish to understand the structure of the molecular networks responsible for synaptic plasticit, it will be vital to have explored the space of all possible molecular networks so that we can develop a correspondence between features of these networks and desirable/undersirable properties of the synapse.
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+   \item\aligntop{\includegraphics[width=4cm]{synapse.svg}}
+   \hspace{0.25cm}
+   \item\aligntop{\includegraphics[width=2cm]{pot.svg}}
+   \hspace{0.25cm}
+   \item\aligntop{\includegraphics[width=2cm]{dep.svg}}
+ \end{myenuma}
+ \end{center}
+ \caption{Model of a complex synapse.}\label{fig:compsynapse}
+\end{figure}
+
+We model synaptic plasticity with two Markov processes between the internal states, one for potentiation and one for depression (see \fref{fig:compsynapse}).
+These are responsible for the initial creation of a memory and the subsequent forgetting due to ongoing plasticity.
+The performance of the synapse is quantified by the signal-to-noise ratio (SNR) as a function of time.
+
+We prove upper limits on the SNR curve.
+For example we show that the area under the curve has an upper bound proportional to the number of internal states.
+This upper limit is saturated by a transition network with a linear chain topology,
+We also find an upper limit on the instantaneous value of the SNR at any time.
+This bounds the length of time that any memory can be recovered, and hence the memory capacity of such synapses.
 

### 2012-11-14 Cosyne/Abstract.tex
cosyne abstract: first draft

@@
 \graphicspath{{Figures/}}
+\usepackage{mciteplus}
+\usepackage{titling}
+\setlength{\droptitle}{-1.5in}
+\renewcommand{\maketitlehookd}{\vspace*{-0.5in}}
 %\usepackage{abstract}
@@
 Recent experimental work suggests that single synapses are digital, in the sense that, from the perspective of extracellular physiology, they can only take on a finite number of discrete values for their strength.
-This imposes catastrophic limits on the memory capacity of classical models of memory that have relied on a continuum of analog synaptic strengths [Amit and Fusi, Network: Computation in Neural Systems (1992), Neural Computation (1994)]\nocite{amit1992constraints,amit1994learning}.
+This imposes catastrophic limits on the memory capacity of classical models of memory that have relied on a continuum of analog synaptic strengths [Amit and Fusi, Network: Computation in Neural Systems (1992)]\nocite{amit1992constraints}.
 
@@
 We develop new theorems bounding the memory capacity of such complex synaptic models and describe the structural organization of internal molecular networks necessary for achieving these limits.
+
+In particular, we prove an upper bound on the area under the signal-to-noise ratio (SNR) curves for \emph{all} models of this type, as well as the SNR at any fixed time and the memory capacity of these synapses.
+
+This type of exploration of the space of possible synaptic models should prove to be essential for understanding the internal molecular structure of real synapses.
 %\end{abstract}
@@
 
-Much of the previous work on complex synapses has focused on the study of specific models \cite{amit1994learning,Fusi2005cascade,Fusi2007multistate}.
-If we wish to understand the structure of the molecular networks responsible for synaptic plasticit, it will be vital to have explored the space of all possible molecular networks so that we can develop a correspondence between features of these networks and desirable/undersirable properties of the synapse.
+Much of the previous work on complex synapses has focused on the study of specific models \cite{amit1994learning,*Fusi2005cascade,*Fusi2007multistate}.
+If we wish to understand the structure of the molecular networks responsible for synaptic plasticity, it will be vital to have explored the space of all possible molecular networks so that we can develop a correspondence between features of these networks and desirable/undersirable properties of the synapse.
 
@@
  \begin{myenuma}
-   \item\aligntop{\includegraphics[width=4cm]{synapse.svg}}
+   \item\aligntop{\includegraphics[width=4cm]{synapse.svg}}\label{fig:synapse}
    \hspace{0.25cm}
-   \item\aligntop{\includegraphics[width=2cm]{pot.svg}}
+   \item\aligntop{\includegraphics[width=2cm]{pot.svg}}\label{fig:pot}
    \hspace{0.25cm}
-   \item\aligntop{\includegraphics[width=2cm]{dep.svg}}
+   \item\aligntop{\includegraphics[width=2cm]{dep.svg}}\label{fig:dep}
  \end{myenuma}
  \end{center}
- \caption{Model of a complex synapse.}\label{fig:compsynapse}
+ \caption{Models of complex synapses.
+(\ref{fig:synapse}) The complex synapse has a number of internal states, some of which correspond to a strong synaptic strength and some to weak.
+(\ref{fig:pot},\ref{fig:dep}) Potentiation and depression induce stochastic transitions between these states.}\label{fig:compsynapse}
 \end{figure}
@@
 
-We prove upper limits on the SNR curve.
-For example we show that the area under the curve has an upper bound proportional to the number of internal states.
-This upper limit is saturated by a transition network with a linear chain topology,
+We can use the transition probabilities to optimize features of the SNR, thereby proving upper limits on the performance of the synapse.
+For example we show that the area under the curve has an upper bound that grow linearly with the number of internal states.
+This upper limit is saturated by a particular transition network with a linear chain topology,
 We also find an upper limit on the instantaneous value of the SNR at any time.
-This bounds the length of time that any memory can be recovered, and hence the memory capacity of such synapses.
+This allows us to bound the length of time that any memory can be recovered, and hence the memory capacity of such synapses.
+
+The models that maximize the SNR at one time are shown to be dominated by a single time-scale.
+When this is extended to maximizing the SNR at several, well separated times, we are led to models with multiple time-scales.
 
@@
 
-\bibliographystyle{utcaps_sl}
-\bibliography{maths,neuro}
+\bibliographystyle{utcaps_m}
+\bibliography{Abstract-minimal}
+%\bibliography{maths,neuro}
 

### 2012-11-14 Cosyne/Abstract.tex
cosyne abstract: refs

@@
 We consider the storage of long term memories through synaptic modifications in existing networks.
-Recent experimental work suggests that single synapses are digital, in the sense that, from the perspective of extracellular physiology, they can only take on a finite number of discrete values for their strength.
+Recent experimental work suggests that single synapses are digital, in the sense that, from the perspective of extracellular physiology, they can only take on a finite number of discrete values for their strength [Bliss and Collingridge, Nature (1993), Bredt and Nicoll, Neuron (2003)]\nocite{Bliss1993,*Bredt2003361}.
 This imposes catastrophic limits on the memory capacity of classical models of memory that have relied on a continuum of analog synaptic strengths [Amit and Fusi, Network: Computation in Neural Systems (1992)]\nocite{amit1992constraints}.
@@
 We can use the transition probabilities to optimize features of the SNR, thereby proving upper limits on the performance of the synapse.
-For example we show that the area under the curve has an upper bound that grow linearly with the number of internal states.
+For example we show that the area under the curve has an upper bound that grows linearly with the number of internal states.
 This upper limit is saturated by a particular transition network with a linear chain topology,

### 2012-12-04 Cosyne/Abstract.tex
two-time envelope

@@
 We consider the storage of long term memories through synaptic modifications in existing networks.
-Recent experimental work suggests that single synapses are digital, in the sense that, from the perspective of extracellular physiology, they can only take on a finite number of discrete values for their strength [Bliss and Collingridge, Nature (1993), Bredt and Nicoll, Neuron (2003)]\nocite{Bliss1993,*Bredt2003361}.
+Recent experimental work suggests that single synapses are digital, in the sense that, from the perspective of extracellular physiology, they can only take on a finite number of discrete values for their strength [Bliss and Collingridge, Nature (1993)]\nocite{Bliss1993}.
 This imposes catastrophic limits on the memory capacity of classical models of memory that have relied on a continuum of analog synaptic strengths [Amit and Fusi, Network: Computation in Neural Systems (1992)]\nocite{amit1992constraints}.
@@
 
-Much of the previous work on complex synapses has focused on the study of specific models \cite{amit1994learning,*Fusi2005cascade,*Fusi2007multistate}.
+Much of the previous work on complex synapses has focused on the study of specific models \cite{amit1994learning,*Fusi2005cascade}.
 If we wish to understand the structure of the molecular networks responsible for synaptic plasticity, it will be vital to have explored the space of all possible molecular networks so that we can develop a correspondence between features of these networks and desirable/undersirable properties of the synapse.

### 2012-12-04 Code/DoubleEnv/Case1Exp1.m

@@ -1,0 +1,44 @@
+function [ S1 ] = Case1Exp1( t1,t2,S2,n,varargin )
+%S1=CASE1EXP1(t1,t2,S2,n,...) Two-time envelope in case 1 with 1 exponential
+%   t1 = time of maximisation
+%   t2 = time of fixed SNR
+%   S2 = SNR(t2)
+%   n = # states
+%
+%   Case# = sum_i Theta(mu_i) 2^i-1
+%   Constraints:
+%       1: sum_a c_a q_a < 1
+%       2: sum_a c_a     < n-1
+%       3: c_a sqrt(q_a) < gamma
+%   Parameter/Value pairs
+%       t = time to evaluate SNR envelope (default=t1)
+%       Constraint3 = Use constraint 3? (default=false)
+
+error(CheckSize(t1,@isscalar))
+error(CheckValue(t1,@(x)x>0))
+error(CheckSize(t2,@isscalar))
+error(CheckValue(t2,@(x)x>0))
+error(CheckSize(S2,@isscalar))
+error(CheckValue(S2,@(x)x>0))
+error(CheckSize(n,@isscalar))
+error(CheckValue(n,@(x)x>0))
+
+gammasq=128/pi^4;
+
+t=t1;
+Constraint3=false;
+varargin=assignApplicable(varargin);
+
+
+S1=S2^(t/t2);
+
+valid = t1<=t2;
+valid = valid && t2 <= - (n-1)*log(S2);
+if Constraint3
+    valid = valid && t2 <= - gammasq*log(S2);
+end%if
+
+S1=S1*valid;
+
+end%function
+

### 2012-12-04 Code/DoubleEnv/Case1Exp2.m

@@ -1,0 +1,30 @@
+function [ S1 ] = Case1Exp2( t1,t2,S2,n,varargin )
+%S1=CASE1EXP2(t1,t2,S2,n,...) Two-time envelope in case 1 with 2
+%exponentials
+%   t1 = time of maximisation
+%   t2 = time of fixed SNR
+%   S2 = SNR(t2)
+%   n = # states
+%
+%   Case# = sum_i Theta(mu_i) 2^i-1
+%   Constraints:
+%       1: sum_a c_a q_a < 1
+%       2: sum_a c_a     < n-1
+%       3: c_a sqrt(q_a) < gamma
+%   Parameter/Value pairs
+%       t = time to evaluate SNR envelope (default=t1)
+%       Constraint3 = Use constraint 3? (default=false)
+
+error(CheckSize(t1,@isscalar))
+error(CheckValue(t1,@(x)x>0))
+error(CheckSize(t2,@isscalar))
+error(CheckValue(t2,@(x)x>0))
+error(CheckSize(S2,@isscalar))
+error(CheckValue(S2,@(x)x>0))
+error(CheckSize(n,@isscalar))
+error(CheckValue(n,@(x)x>0))
+
+S1=0;
+
+end%function
+

### 2012-12-04 Code/DoubleEnv/Case2Exp1.m

@@ -1,0 +1,46 @@
+function [ S1 ] = Case2Exp1( t1,t2,S2,n,varargin )
+%S1=CASE2EXP1(t1,t2,S2,n,...) Two-time envelope in case 2 with 1 exponential
+%   t1 = time of maximisation
+%   t2 = time of fixed SNR
+%   S2 = SNR(t2)
+%   n = # states
+%
+%   Case# = sum_i Theta(mu_i) 2^i-1
+%   Constraints:
+%       1: sum_a c_a q_a < 1
+%       2: sum_a c_a     < n-1
+%       3: c_a sqrt(q_a) < gamma
+%   Parameter/Value pairs
+%       t = time to evaluate SNR envelope (default=t1)
+%       Constraint3 = Use constraint 3? (default=false)
+
+error(CheckSize(t1,@isscalar))
+error(CheckValue(t1,@(x)x>0))
+error(CheckSize(t2,@isscalar))
+error(CheckValue(t2,@(x)x>0))
+error(CheckSize(S2,@isscalar))
+error(CheckValue(S2,@(x)x>0))
+error(CheckSize(n,@isscalar))
+error(CheckValue(n,@(x)x>0))
+
+gammasq=128/pi^4;
+
+t=t1;
+Constraint3=false;
+varargin=assignApplicable(varargin);
+
+c=n-1;
+q=-lambertw(-S2*t2/(n-1))/t2;
+
+S1=c*q*exp(-q*t);
+
+valid = c*q<=1;
+valid = valid && (t1-t2)/(1-q*t2)>=0;
+if Constraint3
+    valid = valid && c^2*q <= gammasq;
+end%if
+
+S1=S1*valid;
+
+end%function
+

### 2012-12-04 Code/DoubleEnv/Case2Exp2.m

@@ -1,0 +1,82 @@
+function [ S1 ] = Case2Exp2( t1,t2,S2,n,varargin )
+%S1=CASE2EXP2(t1,t2,S2,n,...) Two-time envelope in case 1 with 2
+%exponentials
+%   t1 = time of maximisation
+%   t2 = time of fixed SNR
+%   S2 = SNR(t2)
+%   n = # states
+%
+%   Case# = sum_i Theta(mu_i) 2^i-1
+%   Constraints:
+%       1: sum_a c_a q_a < 1
+%       2: sum_a c_a     < n-1
+%       3: c_a sqrt(q_a) < gamma
+%   Parameter/Value pairs
+%       t = time to evaluate SNR envelope (default=t1)
+%       Constraint3 = Use constraint 3? (default=false)
+
+error(CheckSize(t1,@isscalar))
+error(CheckValue(t1,@(x)x>0))
+error(CheckSize(t2,@isscalar))
+error(CheckValue(t2,@(x)x>0))
+error(CheckSize(S2,@isscalar))
+error(CheckValue(S2,@(x)x>0))
+error(CheckSize(n,@isscalar))
+error(CheckValue(n,@(x)x>0))
+
+gammasq=128/pi^4;
+
+t=t1;
+Constraint3=false;
+varargin=assignApplicable(varargin);
+
+
+q1guess=(1+exp(1)*t1/t2)/t2;
+q2guess=1/t1;
+
+qc=fsolve(@eqs,[q1guess,q2guess]);
+
+q1=qc(1);
+q2=qc(2);
+
+
+c1=(S2-(n-1)*q2*exp(-q2*t2))/(q1*exp(-q1*t2)-q2*exp(-q2*t2));
+c2=n-1-c1;
+
+S1=c1*q1*exp(-q1*t)+c2*q2*exp(-q2*t);
+
+valid = c1*q1+c2*q2<=1;
+valid = valid && mu2(q1)>=0;
+if Constraint3
+    valid = valid && c1^2*q1 <= gammasq;
+    valid = valid && c2^2*q2 <= gammasq;
+end%if
+
+S1=S1*valid;
+
+
+
+    function lam=lambda(q)
+        lam=-(1-q*t1)/(1-q*t2)*exp(q*(t2-t1));
+    end
+
+    function mu=mu2(q)
+        mu=q^2*(t1-t2)/(1-q*t2)*exp(-q*t1);
+    end
+
+    function val=eq1(q1,q2)
[…]

### 2013-01-11 Notes/ContTime.tex
used global preamble
notes for model of VOR expt

@@
 \documentclass[12pt]{article}
-% Preamble:
+%Preamble:
 \usepackage{a4wide}
-\usepackage[centertags]{amsmath}
-%\usepackage{ams} for finding documentation only
-\usepackage{amssymb}
-%\usepackage{amsthm}
-\usepackage[sort&compress,numbers]{natbib}
-%\usepackage{citeB}
-\usepackage{ifpdf}
-%\usepackage{graphicx}
-%\usepackage{graphics} for finding documentation only
-%\usepackage{xcolor}
-%\usepackage{pgf}
-\ifpdf
-\usepackage[pdftex,bookmarks,bookmarksopen,pdfstartview=FitH]{hyperref}
-\else
-\usepackage[hypertex]{hyperref}
-%\DeclareGraphicsRule{.png}{eps}{.bb}{}
-\fi
+\input{sl_preamble.tex}
 %
@@
 % ----------------------------------------------------------------
-\vfuzz2pt % Don't report over-full v-boxes if over-edge is small
-\hfuzz2pt % Don't report over-full h-boxes if over-edge is small
 %\numberwithin{equation}{section}
@@
 % New commands etc.
-\input{mydefs.tex}
-\input{newsymb.tex}
+\input{sl_definitions.tex}
+\input{sl_symbols.tex}
 %matrices
@@
 % ----------------------------------------------------------------
-\usepackage{amsthm}
-\newtheoremstyle{sldefinition}%
-  {3pt}%space above
-  {3pt}%space below
-  {}%body font
-  {}%indent amount
-  {\bfseries}%theorem head font
-  {}%theorem head punctuation
-  {\newline}%space after head
-  {\thmname{#1}\thmnumber{ #2}:{\bfseries\thmnote{ #3}}}%head spec
-\newtheoremstyle{slplain}%
-  {3pt}%space above
-  {3pt}%space below
-  {}%body font
-  {}%indent amount
-  {\bfseries}%theorem head font
-  {}%theorem head punctuation
-  {\newline}%space after head
-  {\thmname{#1}\thmnumber{ #2}{\bfseries\thmnote{ (#3)}}: }%head spec
-%
-\theoremstyle{slplain}
-\newtheorem{thm}{Theorem}
-\newtheorem{cor}[thm]{Corollary}
-%
-\theoremstyle{sldefinition}
-\newtheorem{defn}{Definition}
-%
-\theoremstyle{remark}
-\newtheorem*{notn}{Notation}
-\newtheorem*{rem}{Remark}
+\input{sl_theorems_preamble.tex}
 % ----------------------------------------------------------------
@@
 
-\subsection{Derivatives \wrt $\W^\pm$}\label{sec:deriv}
+\subsection{Derivatives \wrt \texorpdfstring{$\W^\pm$}{W+/-}}\label{sec:deriv}
 
@@
 
-\subsubsection{Increasing $c_k$}\label{sec:areacoeffincr}
+\subsubsection{Increasing \texorpdfstring{$c_k$}{c(k)}}\label{sec:areacoeffincr}
 
@@
 
-\subsection{Differences in $c_k$'s}\label{sec:areacoeffdiff}
+\subsection{Differences in \texorpdfstring{$c_k$}{c(k)}'s}\label{sec:areacoeffdiff}
 
@@
 
-\subsection{Derivatives \wrt $\fptb$}\label{sec:derivT}
+\subsection{Derivatives \wrt \texorpdfstring{$\fptb$}{T}}\label{sec:derivT}
 

### 2013-01-11 Notes/VORmodel.tex
notes for model of VOR expt

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\DeclareMathOperator{\SNR}{SNR}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\kn}{_{\text{D$^b$K$^b$}}}
\newcommand{\tpre}{t_{\text{pre}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Models of VOR learning in MHC knockout mice}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
  We see if we can model VOR gain increase and decrease learning in mice with a knockout in MHC as well as wild type.
\end{abstract}

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beginning of Article:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The setup}\label{sec:setup}

\subsection{Models of synapses}\label{sec:synapse}

We make the following assumptions:
\begin{itemize}
  \item There are $N$ identical synapses with $M$ internal functional states.
  \item The states of different synapses are independent of each other.
  \item The synapses that are eligible for plasticity are chosen randomly.
  \item The potentiating/depressing plasticity event timings are distributes as Poisson processes with rates $rf\potdep$, where $f\pot+f\dep=1$.
  \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
  \item The synaptic weights of the internal states are given by the column vector $\w$. This can only take two values that we can call $\pm1$.
\end{itemize}

The independence and identicalness of synapses means that the state of the system can be completely described by the probability distribution of the internal states, the row vector $\pr(t)$.

The evolution of this probability is described by a forgetting matrix, $\frg$:
%
\begin{equation}\label{eq:evolve}
  \diff{\pr(t)}{t} = r\pr(t)\frg,
  \qquad
  \frg = f\pot\M\pot+f\dep\M\dep-\I.
\end{equation}
%
Eventually, this will settle into the equilibrium distribution:
%
\begin{equation}\label{eq:eqprob}
  \eq\frg=0.
\end{equation}
%

With only two possible synaptic weights, the distribution of synaptic weights is completely described by the mean, $\pr(t)\w$.

\subsection{Model of VOR learning experiment}\label{sec:learning}

Training the animal will not change the internal dynamics of a synapse under potentiation or depression.
It will change the environment, will lead to a change in how often potentiation and depression occur.
It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
We will model this by changing $f\potdep$, leaving $r$ and $\M\potdep$ unchanged.

The untrained case will be described by $f\pot=f\pot_0$.
Gain increase training  will be described by $f\pot=f\pot\inc<f\pot_0$, and
gain decrease training  will be described by $f\pot=f\pot\inc<f\pot_0$. Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depends on $f\pot$, which we will indicate with subscripts.

Before training, the synaptic distribution will in equilibrium with $f\pot_0$.
During gain increase training, it will evolve according to \eqref{eq:evolve} with $f\pot\inc$:
%
\begin{equation}\label{eq:nopre}
  \pr(t) = \eq_0 \exp\prn{rt\frg\inc}.
\end{equation}
%
On the other hand, if the gain increase training follows gain decrease pre-training for some time period, $\tpre$:
%
\begin{equation}\label{eq:withpre}
  \pr(t) = \eq_0 \exp\prn{r\tpre\frg\dec} \exp\prn{r(t-\tpre)\frg\inc}.
\end{equation}
%

We will describe the effect of training by the decrease in mean synaptic weight:
%
\begin{equation}\label{eq:learning}
  L(t) = (\pr(0)-\pr(t))\w.
\end{equation}
%
The behavioural output (VOR gain) will be some non-linear function of the synaptic weights, so the best we can hope for is to reproduce qualitative features of the experiment.

The MHC knockout has a lower threshold for depression.
We can model this by changing $\W\dep\wt$ to $\W\dep\kn$, which should have larger matrix elements.

QUESTION: Should $f\pot\wt=f\pot\kn$?

If they are equal, this would change the mean synaptic weight in equilibrium.
This seems like it would affect the ability of the network to perform its function, and one might expect adaptation to the environment to produce an equilibrium state that has the same performance.
In any case, if the synaptic weights are different, the electrical activity will be different, and there will be no reason to expect the same statistics for potentiation or depression.

On the other hand, one could imagine adjusting $f\pot\wt$ and $f\pot\kn$ so that $\eq\wt\w = \eq\kn\w$.
But, now that the synaptic weights are the same, the electrical activity will be the same, and there will be no reason to expect different statistics for potentiation or depression.

%\subsection*{Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection*{Appendices}
%\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{utcaps_sl}
\bibliography{maths,neuro}

\end{document}

### 2013-01-12 Notes/VORmodel.tex
wrote up model of expt

@@
 \newcommand{\potdep}{^{\text{pot/dep}}}
+\newcommand{\norm}{_0}
 \newcommand{\inc}{_{\text{inc}}}
@@
   \item The synapses that are eligible for plasticity are chosen randomly.
-  \item The potentiating/depressing plasticity event timings are distributes as Poisson processes with rates $rf\potdep$, where $f\pot+f\dep=1$.
+  \item The potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $rf\potdep$, where $f\pot+f\dep=1$.
   \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
@@
 
-The untrained case will be described by $f\pot=f\pot_0$.
-Gain increase training  will be described by $f\pot=f\pot\inc<f\pot_0$, and
-gain decrease training  will be described by $f\pot=f\pot\inc<f\pot_0$.
+The untrained case will be described by $f\pot=f\pot\norm$.
+Gain increase training  will be described by $f\pot=f\pot\inc<f\pot\norm$, and
+gain decrease training  will be described by $f\pot=f\pot\inc<f\pot\norm$.
 Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depends on $f\pot$, which we will indicate with subscripts.
 
-Before training, the synaptic distribution will in equilibrium with $f\pot_0$.
+Before training, the synaptic distribution will in equilibrium with $f\pot\norm$.
 During gain increase training, it will evolve according to \eqref{eq:evolve} with $f\pot\inc$:
@@
 \begin{equation}\label{eq:nopre}
-  \pr(t) = \eq_0 \exp\prn{rt\frg\inc}.
+  \pr(t) = \eq\norm \exp\prn{rt\frg\inc}.
 \end{equation}
@@
 \begin{equation}\label{eq:withpre}
-  \pr(t) = \eq_0 \exp\prn{r\tpre\frg\dec} \exp\prn{r(t-\tpre)\frg\inc}.
+  \pr(t) = \eq\norm \exp\prn{r\tpre\frg\dec} \exp\prn{r(t-\tpre)\frg\inc}.
 \end{equation}
@@
 \begin{equation}\label{eq:learning}
-  L(t) = (\pr(0)-\pr(t))\w.
+  L(t) = \prn{\pr(0)-\pr(t)}\w.
 \end{equation}
@@
 
-\bibliographystyle{utcaps_sl}
-\bibliography{maths,neuro}
+%\bibliographystyle{utcaps_sl}
+%\bibliography{maths,neuro}
 

### 2013-01-15 Notes/VORmodel.tex
models of VOR

@@
 \input{sl_preamble.tex}
+\input{sl_graphics_preamble.tex}
+\graphicspath{{"Figures/"}}
 %
@@
 \newcommand{\wt}{_{\text{WT}}}
-\newcommand{\kn}{_{\text{D$^b$K$^b$}}}
+\newcommand{\kn}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$}}}
+\newcommand{\Kn}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
 \newcommand{\tpre}{t_{\text{pre}}}
@@
 Gain increase training  will be described by $f\pot=f\pot\inc<f\pot\norm$, and
-gain decrease training  will be described by $f\pot=f\pot\inc<f\pot\norm$.
+gain decrease training  will be described by $f\pot=f\pot\dec<f\pot\norm$.
 Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depends on $f\pot$, which we will indicate with subscripts.
@@
 
+\section{Simulations}\label{sec:sims}
+
+\subsection{Models and parameters}
+
+We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade}), the multistate model (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary}).
+
+For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \Kn\ mutant.
+We will use a larger value for $x$ for depression in the \Kn\ mutant.
+
+For the multistate and two state models, we will use the same value for the transition probabilities, $q$ for potentiation and depression in the wild-type as well as potentiation in the \Kn\ mutant.
+We will use a larger value for $q$ for depression in the \Kn\ mutant.
+
+In each case, we set $f\pot\norm=\half$, $f\pot\inc=f\pot\norm+\Delta f$ and $f\pot\dec=f\pot\norm-\Delta f$. The values of all these parameters are listed in \autoref{tab:params}.
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade}\hspace{0.5cm}
+  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate}\hspace{0.5cm}
+  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary}
+ \end{myenuma}
+ \end{center}
+  \caption{Transition probabilities for different models.
+(\ref{fig:cascade}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
+(\ref{fig:multistate}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
+(\ref{fig:binary}) The two-state model is parameterised by the two transition probailities.}\label{fig:models}
+\end{figure}
+
+\begin{table}
+ \begin{center}
+  \begin{tabular}{|l|c|c|c|c|}
+    \hline
+    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+    Model & \# states & pot,WT dep & \Kn\ dep & $\Delta f$\\
+    \hline
+    Cascade & 10 & $x=0.25$ & $x=0.3$ & -0.1 \\
+    Multistate & 10 & $q=0.6$ & $q=0.8$ & -0.1 \\
+    Two-state & 2 & $q=0.6$ & $q=0.8$ & -0.1 \\
+    \hline
+  \end{tabular}
+ \end{center}
+  \caption{Parameters used in simulations.}\label{tab:params}
+\end{table}
+
 %\subsection*{Acknowledgements}
@@
 
-%\bibliographystyle{utcaps_sl}
-%\bibliography{maths,neuro}
+\bibliographystyle{utcaps_sl}
+\bibliography{maths,neuro}
 

### 2013-01-15 Code/VORcomparison.m

@@ -6,11 +6,25 @@ function VORcomparison( Wp,Wm,w,t,df,tchange,WpM,WmM,varargin )
 %   DF = Change in fraction of potentiation transitions
 %   w = Weights of states (+/-1)
 
-phWT=VORexpt(Wp,Wm,w,t,df,tchange,varargin{:});
-phKn=VORexpt(WpM,WmM,w,t,df,tchange,'LinSpec','g',varargin{:});
-legend([phWT(1);phKn(1)],{'WT';'Knockout'},'Location','Best')
+phWT=VORexpt(Wp,Wm,w,t,df,tchange,'LinSpec','k',varargin{:});
+phKn=VORexpt(WpM,WmM,w,t,df,tchange,'LinSpec','r',varargin{:});
 xlabel('Training time')
-ylabel('Learning (mean synaptic weight)')
+ylabel('Learning (change in mean synaptic weight)')
+yl=ylim;
+xl=xlim;
+[x,y]=dsxy2figxy([xl(1) xl(2)], 0.95*yl(2)*[1 1]);
+annotation('doublearrow',x,y);
+pos=dsxy2figxy([0.6*xl(1)+0.4*xl(2) 0.9*yl(2) 0.2*(xl(2)-xl(1)) 0.05*yl(2)]);
+annotation('textbox',pos,'String','training','VerticalAlignment','top','LineStyle','none','HorizontalAlignment','center');
+[x,y]=dsxy2figxy([xl(1) tchange], 0.95*yl(1)*[1 1]);
+annotation('doublearrow',x,y);
+pos=dsxy2figxy([0.6*xl(1)+0.4*tchange 0.95*yl(1) 0.2*(tchange-xl(1)) abs(0.05*yl(1))]);
+annotation('textbox',pos,'String','pre-training','VerticalAlignment','bottom','LineStyle','none','HorizontalAlignment','center');
+[x,y]=dsxy2figxy([tchange xl(2)], 0.95*yl(1)*[1 1]);
+annotation('doublearrow',x,y);
+pos=dsxy2figxy([0.6*tchange+0.4*xl(2) 0.95*yl(1) 0.2*(xl(2)-tchange) abs(0.05*yl(1))]);
+annotation('textbox',pos,'String','training','VerticalAlignment','bottom','LineStyle','none','HorizontalAlignment','center');
+legend([phWT(1);phKn(1)],{'WT';'D^bK^b-/-'},'Location','Best')
 embiggen
 
 end

### 2013-01-19 Notes/VORmodel.tex
comparison ov VOR training schemes
VOR: analytic discussion of multistate and two-state models

@@
 \newcommand{\wt}{_{\text{WT}}}
-\newcommand{\kn}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$}}}
+\newcommand{\kn}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
 \newcommand{\Kn}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
@@
 The untrained case will be described by $f\pot=f\pot\norm$.
-Gain increase training  will be described by $f\pot=f\pot\inc<f\pot\norm$, and
-gain decrease training  will be described by $f\pot=f\pot\dec<f\pot\norm$.
+Gain-increase training  will be described by $f\pot=f\pot\inc<f\pot\norm$, and
+gain-decrease training  will be described by $f\pot=f\pot\dec<f\pot\norm$.
 Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depends on $f\pot$, which we will indicate with subscripts.
@@
 Before training, the synaptic distribution will in equilibrium with $f\pot\norm$.
-During gain increase training, it will evolve according to \eqref{eq:evolve} with $f\pot\inc$:
+During gain-increase training, it will evolve according to \eqref{eq:evolve} with $f\pot\inc$:
 %
@@
 %
-On the other hand, if the gain increase training follows gain decrease pre-training for some time period, $\tpre$:
+On the other hand, if the gain-increase training follows gain-decrease pre-training for some time period, $\tpre$:
 %
@@
     \hline
-    Cascade & 10 & $x=0.25$ & $x=0.3$ & -0.1 \\
+    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 \\
     Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 \\
@@
 \end{table}
+
+\subsection{Results}\label{sec:results}
+
+The features of the actual experiments that we'd like to capture are:
+\begin{itemize}
+  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the mutant.
+  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.
+  \item After pre-training, gain-increase learning is significantly faster in the mutant than in the wild-type.
+  \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
+  \item For the mutant, gain-increase learning is significantly faster with pre-training than without it.
+\end{itemize}
+
+We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
+This is proportional to the net-flux between the $\w=+1$ states and the $\w=-1$ states, measuring using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
+
+\subsubsection{Cascade}\label{sec:cascade}
+
+\subsubsection{Multistate}\label{sec:multistate}
+
+Consider the general uniform multistate model. 
+Then the equilibrium distribution is given by
+%
+\begin{equation}\label{eq:mutltieq}
+  \eq_i = \frac{1-\alpha}{1-\alpha^M}\,\alpha^{i-1},
+  \qquad \text{where} \quad
+  \alpha=\frac{f\pot q\pot}{f\dep q\dep}.
+\end{equation}
+%
+If we take the limit $\alpha\rightarrow1$, this becomes $\frac{1}{M}$.
+
+The net-flux from the $\w=+1$ states to the $\w=-1$ states is:
+%
+\begin{equation}\label{eq:multiflux}
+  \Phi = \eq_{M/2+1}f'{}\dep q\dep - \eq_{M/2}f'{}\pot q\pot = \frac{1-\alpha}{1-\alpha^M}\,\alpha^{M/2-1}\prn{\alpha-\alpha'}f'{}\dep q\dep,
+\end{equation}
+%
+where primed values correspond to the new value of $f\pot$.
+
+First, consider the wild-type, for which $q\pot=q\dep=q$.
+Without pre-training:
+%
+\begin{equation}\label{eq:multiWTnopre}
+  \Phi = -\frac{2\Delta fq}{M},
+\end{equation}
+%
+where it's worth remembering that $\Delta f<0$.
+With pre-training:
+%
+\begin{equation}\label{eq:multiWTpre}
+\begin{aligned}
+  \Phi &= 16(\Delta f)^2q \, \frac{(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1}}
+          {(1-2\Delta f)^M - (1+2\Delta f)^M} \\
+       &= -\frac{4\Delta fq}{M} + \CO(\Delta f)^2.
+\end{aligned}
+\end{equation}
+%
+So, we see that pre-training will speed up learning when $\Delta f$ is small.
+On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot.
+
+Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
+Without pre-training:
+%
+\begin{equation}\label{eq:multiKNnopre}
+  \Phi = -\frac{2(1-\beta)\beta^{M/2-1}\Delta fq}{1-\beta^M}.
+\end{equation}
+%
+With pre-training:
+%
+\begin{equation}\label{eq:multiKnpre}
+\begin{aligned}
+  \Phi &= -4\Delta f q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
+          {(1-2\Delta f)^M - \beta^M(1+2\Delta f)^M}   \,
+          \beta^{M/2-1}(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1}.
+\end{aligned}
+\end{equation}
+%
+
+\subsubsection{Two-state}\label{sec:binary}
+
+This model can be solved exactly:
+%
+\begin{multline}\label{eq:binarysol}
+  \eq = \frac{(f\dep q\dep, f\pot q\pot)}{\lambda},
+  \qquad
+  \pr(t) = \eq + (\pr(t)-\eq)\e^{-\lambda rt},\\
+  \qquad \text{where} \quad
+  \lambda = f\pot q\pot + f\dep q\dep.
+\end{multline}
+%
+But it is easier to just substitute $M=2$ into the formulae in \sref{sec:multistate}.
+In this case, the initial rate of change encapsulates the whole solution, as there is only a single exponential decay.
+
+First, consider the wild-type, for which $q\pot=q\dep=q$.
+Without pre-training:
+%
+\begin{equation}\label{eq:binWTnopre}
+  \Phi = -\Delta f q,
+\end{equation}
+%
+where it's worth remembering that $\Delta f<0$.
+With pre-training:
+%
+\begin{equation}\label{eq:binWTpre}
+\begin{aligned}
+  \Phi &= -2\Delta f q.
+\end{aligned}
+\end{equation}
+%
+So, we see that pre-training will always speed up learning.
+
+Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
+Without pre-training:
+%
+\begin{equation}\label{eq:binKNnopre}
+  \Phi = -\frac{2\Delta f q}{1+\beta},
+\end{equation}
+%
+which is always larger that \eqref{eq:binWTnopre}.
+With pre-training:
+%
+\begin{equation}\label{eq:binKnpre}
+\begin{aligned}
+  \Phi &= -4\Delta f q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
+          {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
+\end{aligned}
+\end{equation}
+%
 

### 2013-01-23 Notes/VORmodel.tex
added VOR graphs

@@
 \newcommand{\wt}{_{\text{WT}}}
-\newcommand{\kn}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
-\newcommand{\Kn}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
+\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
+\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
 \newcommand{\tpre}{t_{\text{pre}}}
@@
 The MHC knockout has a lower threshold for depression.
-We can model this by changing $\W\dep\wt$ to $\W\dep\kn$, which should have larger matrix elements.
+We can model this by changing $\W\dep\wt$ to $\W\dep\ko$, which should have larger matrix elements.
 
-QUESTION: Should $f\pot\wt=f\pot\kn$?
+QUESTION: Should $f\pot\wt=f\pot\ko$?
 
@@
 
-On the other hand, one could imagine adjusting $f\pot\wt$ and $f\pot\kn$ so that $\eq\wt\w = \eq\kn\w$.
+On the other hand, one could imagine adjusting $f\pot\wt$ and $f\pot\ko$ so that $\eq\wt\w = \eq\ko\w$.
 But, now that the synaptic weights are the same, the electrical activity will be the same, and there will be no reason to expect different statistics for potentiation or depression.
+If we did this for the multistate or two-state models, we'd remove all differences between the wild-type and \KO\ knockout.
 
@@
 
-For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \Kn\ mutant.
-We will use a larger value for $x$ for depression in the \Kn\ mutant.
+For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
+We will use a larger value for $x$ for depression in the \KO\ mutant.
 
-For the multistate and two state models, we will use the same value for the transition probabilities, $q$ for potentiation and depression in the wild-type as well as potentiation in the \Kn\ mutant. 
-We will use a larger value for $q$ for depression in the \Kn\ mutant.
+For the multistate and two state models, we will use the same value for the transition probabilities, $q$ for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
+We will use a larger value for $q$ for depression in the \KO\ mutant.
 
@@
   (\ref{fig:multistate}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
-  (\ref{fig:binary}) The two-state model is parameterised by the two transition probailities.}\label{fig:models}
+  (\ref{fig:binary}) The two-state model is parameterised by the two transition probailities.} \label{fig:models}
 \end{figure}
@@
  \begin{center}
-  \begin{tabular}{|l|c|c|c|c|}
+  \begin{tabular}{|l|c|c|c|c|c|c|}
     \hline
     % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-    Model & \# states & pot,WT dep & \Kn\ dep & $\Delta f$\\
+    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $rt_\text{train}$ & $rt_\text{pre}$ \\
     \hline
-    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 \\
-    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 \\
-    Two-state  & 2  & $q=0.6$  & $q=0.8$  & -0.1 \\
+    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50 \\
+    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150 \\
+    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50 \\
+    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20 \\
+    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5 & 5 \\
     \hline
@@
  \end{center}
-  \caption{Parameters used in simulations.}\label{tab:params}
+  \caption{Parameters used in simulations.} \label{tab:params}
 \end{table}
@@
 
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{cascade_short_learn.eps}}\label{fig:cascade_short_learn}
+  \item\aligntop{\includegraphics[width=7cm]{cascade_short_learnS.eps}}\label{fig:cascade_short_learnS}
+  \item\label{fig:cascade_short_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{cascade_short_eq_WT.eps}}\label{fig:cascade_short_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{cascade_short_eq_KO.eps}}\label{fig:cascade_short_eq_KO}
+                  \end{myenumi}
+  \item\label{fig:cascade_short_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_WT_nopre.eps}}\label{fig:cascade_short_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_WT_pre.eps}}\label{fig:cascade_short_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_KO_nopre.eps}}\label{fig:cascade_short_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_KO_pre.eps}}\label{fig:cascade_short_pr_KO_pre}
+                  \end{myenumi}
+ \end{myenuma}
+ \end{center}
+  \caption{Simulation results for cascade model with short pre-training ($rt_\text{pre}=50$).
+  (\ref{fig:cascade_short_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  (\ref{fig:cascade_short_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:cascade_short_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_short_eq_WT}) wild-type and (\ref{fig:cascade_short_eq_KO}) \KO\ knockout.
+  (\ref{fig:cascade_short_pr}) Evolution of probability distributions for (\ref{fig:cascade_short_pr_WT_nopre}\ref{fig:cascade_short_pr_WT_pre}) wild-type and  (\ref{fig:cascade_short_pr_KO_nopre}\ref{fig:cascade_short_pr_KO_pre}) \KO\ knockout without (\ref{fig:cascade_short_pr_WT_nopre},\ref{fig:cascade_short_pr_KO_nopre}) and with (\ref{fig:cascade_short_pr_WT_pre},\ref{fig:cascade_short_pr_KO_pre}) pre-training.
+} \label{fig:cascade_short}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{cascade_long_learn.eps}}\label{fig:cascade_long_learn}
+  \item\aligntop{\includegraphics[width=7cm]{cascade_long_learnS.eps}}\label{fig:cascade_long_learnS}
+  \item\label{fig:cascade_long_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{cascade_long_eq_WT.eps}}\label{fig:cascade_long_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{cascade_long_eq_KO.eps}}\label{fig:cascade_long_eq_KO}
+                  \end{myenumi}
+  \item\label{fig:cascade_long_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_WT_nopre.eps}}\label{fig:cascade_long_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_WT_pre.eps}}\label{fig:cascade_long_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_KO_nopre.eps}}\label{fig:cascade_long_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_KO_pre.eps}}\label{fig:cascade_long_pr_KO_pre}
+                  \end{myenumi}
+ \end{myenuma}
+ \end{center}
+  \caption{Simulation results for cascade model with long pre-training ($rt_\text{pre}=150$).
+  (\ref{fig:cascade_long_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  (\ref{fig:cascade_long_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:cascade_long_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_long_eq_WT}) wild-type and (\ref{fig:cascade_long_eq_KO}) \KO\ knockout.
+  (\ref{fig:cascade_long_pr}) Evolution of probability distributions for (\ref{fig:cascade_long_pr_WT_nopre}\ref{fig:cascade_long_pr_WT_pre}) wild-type and  (\ref{fig:cascade_long_pr_KO_nopre}\ref{fig:cascade_long_pr_KO_pre}) \KO\ knockout without (\ref{fig:cascade_long_pr_WT_nopre},\ref{fig:cascade_long_pr_KO_nopre}) and with (\ref{fig:cascade_long_pr_WT_pre},\ref{fig:cascade_long_pr_KO_pre}) pre-training.
+} \label{fig:cascade_long}
+\end{figure}
+
 \subsubsection{Multistate}\label{sec:multistate}
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{multistate_weak_learn.eps}}\label{fig:multistate_weak_learn}
+  \item\aligntop{\includegraphics[width=7cm]{multistate_weak_learnS.eps}}\label{fig:multistate_weak_learnS}
+  \item\label{fig:multistate_weak_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{multistate_weak_eq_WT.eps}}\label{fig:multistate_weak_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{multistate_weak_eq_KO.eps}}\label{fig:multistate_weak_eq_KO}
+                  \end{myenumi}
+  \item\label{fig:multistate_weak_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_WT_nopre.eps}}\label{fig:multistate_weak_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_WT_pre.eps}}\label{fig:multistate_weak_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_KO_nopre.eps}}\label{fig:multistate_weak_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_KO_pre.eps}}\label{fig:multistate_weak_pr_KO_pre}
+                  \end{myenumi}
+ \end{myenuma}
+ \end{center}
+  \caption{Simulation results for multistate model with weak training ($\Delta f=-0.1$).
+  (\ref{fig:multistate_weak_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  (\ref{fig:multistate_weak_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:multistate_weak_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:multistate_weak_eq_WT}) wild-type and (\ref{fig:multistate_weak_eq_KO}) \KO\ knockout.
+  (\ref{fig:multistate_weak_pr}) Evolution of probability distributions for (\ref{fig:multistate_weak_pr_WT_nopre}\ref{fig:multistate_weak_pr_WT_pre}) wild-type and  (\ref{fig:multistate_weak_pr_KO_nopre}\ref{fig:multistate_weak_pr_KO_pre}) \KO\ knockout without (\ref{fig:multistate_weak_pr_WT_nopre},\ref{fig:multistate_weak_pr_KO_nopre}) and with (\ref{fig:multistate_weak_pr_WT_pre},\ref{fig:multistate_weak_pr_KO_pre}) pre-training.
+} \label{fig:multistate_weak}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{multistate_strong_learn.eps}}\label{fig:multistate_strong_learn}
+  \item\aligntop{\includegraphics[width=7cm]{multistate_strong_learnS.eps}}\label{fig:multistate_strong_learnS}
+  \item\label{fig:multistate_strong_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{multistate_strong_eq_WT.eps}}\label{fig:multistate_strong_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{multistate_strong_eq_KO.eps}}\label{fig:multistate_strong_eq_KO}
+                  \end{myenumi}
+  \item\label{fig:multistate_strong_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_WT_nopre.eps}}\label{fig:multistate_strong_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_WT_pre.eps}}\label{fig:multistate_strong_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_KO_nopre.eps}}\label{fig:multistate_strong_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_KO_pre.eps}}\label{fig:multistate_strong_pr_KO_pre}
+                  \end{myenumi}
+ \end{myenuma}
+ \end{center}
+  \caption{Simulation results for multistate model with strong training ($\Delta f=-0.4$).
+  (\ref{fig:multistate_strong_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  (\ref{fig:multistate_strong_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:multistate_strong_eq}) Equilibrium distributions with/without gain-increase/decrease training for (\ref{fig:multistate_strong_eq_WT}) wild-type and (\ref{fig:multistate_strong_eq_KO}) \KO\ knockout.
+  (\ref{fig:multistate_strong_pr}) Evolution of probability distributions for (\ref{fig:multistate_strong_pr_WT_nopre}\ref{fig:multistate_strong_pr_WT_pre}) wild-type and  (\ref{fig:multistate_strong_pr_KO_nopre}\ref{fig:multistate_strong_pr_KO_pre}) \KO\ knockout without (\ref{fig:multistate_strong_pr_WT_nopre},\ref{fig:multistate_strong_pr_KO_nopre}) and with (\ref{fig:multistate_strong_pr_WT_pre},\ref{fig:multistate_strong_pr_KO_pre}) pre-training.
+} \label{fig:multistate_strong}
+\end{figure}
 
@@
 %
-\begin{equation}\label{eq:multiKNnopre}
+\begin{equation}\label{eq:multiKOnopre}
   \Phi = -\frac{2(1-\beta)\beta^{M/2-1}\Delta fq}{1-\beta^M}.
@@
 %
-\begin{equation}\label{eq:multiKnpre}
+\begin{equation}\label{eq:multiKOpre}
 \begin{aligned}
@@
 \subsubsection{Two-state}\label{sec:binary}
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{binary_learn.eps}}\label{fig:binary_learn}
+  \item\aligntop{\includegraphics[width=7cm]{binary_learnS.eps}}\label{fig:binary_learnS}
+  \item\label{fig:binary_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{binary_eq_WT.eps}}\label{fig:binary_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{binary_eq_KO.eps}}\label{fig:binary_eq_KO}
+                  \end{myenumi}
+  \item\label{fig:binary_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_WT_nopre.eps}}\label{fig:binary_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_WT_pre.eps}}\label{fig:binary_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_KO_nopre.eps}}\label{fig:binary_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_KO_pre.eps}}\label{fig:binary_pr_KO_pre}
+                  \end{myenumi}
+ \end{myenuma}
+ \end{center}
+  \caption{Simulation results for two-state model.
+  (\ref{fig:binary_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  (\ref{fig:binary_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:binary_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:binary_eq_WT}) wild-type and (\ref{fig:binary_eq_KO}) \KO\ knockout.
+  (\ref{fig:binary_pr}) Evolution of probability distributions for (\ref{fig:binary_pr_WT_nopre}\ref{fig:binary_pr_WT_pre}) wild-type and  (\ref{fig:binary_pr_KO_nopre}\ref{fig:binary_pr_KO_pre}) \KO\ knockout without (\ref{fig:binary_pr_WT_nopre},\ref{fig:binary_pr_KO_nopre}) and with (\ref{fig:binary_pr_WT_pre},\ref{fig:binary_pr_KO_pre}) pre-training.
+} \label{fig:binary_res}
+\end{figure}
 
@@
 %
-\begin{equation}\label{eq:binKNnopre}
+\begin{equation}\label{eq:binKOnopre}
   \Phi = -\frac{2\Delta f q}{1+\beta},
@@
 %
-\begin{equation}\label{eq:binKnpre}
+\begin{equation}\label{eq:binKOpre}
 \begin{aligned}

### 2013-01-23 Code/VOR/ProbEvol.m

@@ -1,0 +1,19 @@
+function ProbEvol( Pt,t,titletext,varargin )
+%PROBEVOL(Pt,t,titletext,...) Summary of this function goes here
+%   Detailed explanation goes here
+
+Parent=gca;
+FontSize=10;
+varargin=assignApplicable(varargin);
+cla(Parent);
+
+imagesc(t,1:size(Pt,2),Pt','Parent',Parent,varargin{:});
+ylabel(Parent,'State');
+xlabel(Parent,'Training time');
+title(Parent,titletext);
+h=colorbar('peer',Parent);
+% set(get(h,'YLabel'),'String','Probability','Rotation',270,'VerticalAlignment','bottom','FontSize',FontSize);
+colorbarlabel(h,'Probability','FontSize',FontSize);
+        
+end
+

### 2013-01-23 Code/VOR/VORexpt.m

@@ -17,7 +17,7 @@ error(CheckSize(w,@iscol));
 fp=0.5;
 LinSpec='b';
 Parent=gca;
-altax=[];
+% altax=[];
 
 varargin=assignApplicable(varargin);
 
@@ -27,13 +27,13 @@ hold(Parent,'on');
 [S,P_pre]=LearningCurve(fp*Wp+(1-fp)*Wm,w,t,(fp-df)*Wp+(1-fp+df)*Wm,tchange,(fp+df)*Wp+(1-fp-df)*Wm);
 ph(2)=plot(t,S(1)-S,LinSpec,'Parent',Parent,varargin{:});
 
-if ~isempty(altax)
-    plot(altax,[EqProb(fp*Wp+(1-fp)*Wm); EqProb((fp+df)*Wp+(1-fp-df)*Wm); EqProb((fp-df)*Wp+(1-fp+df)*Wm)]');
-%     plot(altax,pT');
-    xlabel(altax,'State');
-    ylabel(altax,'Equilibrium prob.');
-    legend(altax,{'Untrained','Gain increase','Gain decrease'},'Location','Best');
-end
-
+% if ~isempty(altax)
+%     plot(altax,[EqProb(fp*Wp+(1-fp)*Wm); EqProb((fp+df)*Wp+(1-fp-df)*Wm); EqProb((fp-df)*Wp+(1-fp+df)*Wm)]');
+% %     plot(altax,pT');
+%     xlabel(altax,'State');
+%     ylabel(altax,'Equilibrium prob.');
+%     legend(altax,{'Untrained','Gain increase','Gain decrease'},'Location','Best');
+% end
+% 
 end
 

### 2013-01-23 Notes/VORmodel.tex
written interpretations of graphs and equations

@@
 
+\section{Introduction and summary}\label{sec:intro}
+
 \section{The setup}\label{sec:setup}
@@
 
-The MHC knockout has a lower threshold for depression.
+The MHC mutant has a lower threshold for depression.
 We can model this by changing $\W\dep\wt$ to $\W\dep\ko$, which should have larger matrix elements.
@@
 But, now that the synaptic weights are the same, the electrical activity will be the same, and there will be no reason to expect different statistics for potentiation or depression.
-If we did this for the multistate or two-state models, we'd remove all differences between the wild-type and \KO\ knockout.
+If we did this for the multistate or two-state models, we'd remove all differences between the wild-type and \KO\ mutant.
 
@@
   \caption{Simulation results for cascade model with short pre-training ($rt_\text{pre}=50$).
-  (\ref{fig:cascade_short_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  Other parameters can be found in \autoref{tab:params}.
+  (\ref{fig:cascade_short_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
   (\ref{fig:cascade_short_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:cascade_short_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_short_eq_WT}) wild-type and (\ref{fig:cascade_short_eq_KO}) \KO\ knockout.
-  (\ref{fig:cascade_short_pr}) Evolution of probability distributions for (\ref{fig:cascade_short_pr_WT_nopre}\ref{fig:cascade_short_pr_WT_pre}) wild-type and  (\ref{fig:cascade_short_pr_KO_nopre}\ref{fig:cascade_short_pr_KO_pre}) \KO\ knockout without (\ref{fig:cascade_short_pr_WT_nopre},\ref{fig:cascade_short_pr_KO_nopre}) and with (\ref{fig:cascade_short_pr_WT_pre},\ref{fig:cascade_short_pr_KO_pre}) pre-training.
+  (\ref{fig:cascade_short_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_short_eq_WT}) wild-type and (\ref{fig:cascade_short_eq_KO}) \KO\ mutant.
+  (\ref{fig:cascade_short_pr}) Evolution of probability distributions for (\ref{fig:cascade_short_pr_WT_nopre}\ref{fig:cascade_short_pr_WT_pre}) wild-type and  (\ref{fig:cascade_short_pr_KO_nopre}\ref{fig:cascade_short_pr_KO_pre}) \KO\ mutant without (\ref{fig:cascade_short_pr_WT_nopre},\ref{fig:cascade_short_pr_KO_nopre}) and with (\ref{fig:cascade_short_pr_WT_pre},\ref{fig:cascade_short_pr_KO_pre}) pre-training.
 } \label{fig:cascade_short}
@@
   \caption{Simulation results for cascade model with long pre-training ($rt_\text{pre}=150$).
-  (\ref{fig:cascade_long_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  Other parameters can be found in \autoref{tab:params}.
+  (\ref{fig:cascade_long_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
   (\ref{fig:cascade_long_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:cascade_long_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_long_eq_WT}) wild-type and (\ref{fig:cascade_long_eq_KO}) \KO\ knockout.
-  (\ref{fig:cascade_long_pr}) Evolution of probability distributions for (\ref{fig:cascade_long_pr_WT_nopre}\ref{fig:cascade_long_pr_WT_pre}) wild-type and  (\ref{fig:cascade_long_pr_KO_nopre}\ref{fig:cascade_long_pr_KO_pre}) \KO\ knockout without (\ref{fig:cascade_long_pr_WT_nopre},\ref{fig:cascade_long_pr_KO_nopre}) and with (\ref{fig:cascade_long_pr_WT_pre},\ref{fig:cascade_long_pr_KO_pre}) pre-training.
+  (\ref{fig:cascade_long_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_long_eq_WT}) wild-type and (\ref{fig:cascade_long_eq_KO}) \KO\ mutant.
+  (\ref{fig:cascade_long_pr}) Evolution of probability distributions for (\ref{fig:cascade_long_pr_WT_nopre}\ref{fig:cascade_long_pr_WT_pre}) wild-type and  (\ref{fig:cascade_long_pr_KO_nopre}\ref{fig:cascade_long_pr_KO_pre}) \KO\ mutant without (\ref{fig:cascade_long_pr_WT_nopre},\ref{fig:cascade_long_pr_KO_nopre}) and with (\ref{fig:cascade_long_pr_WT_pre},\ref{fig:cascade_long_pr_KO_pre}) pre-training.
 } \label{fig:cascade_long}
 \end{figure}
+
+The results of simulations of the cascade model can be seen in \autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
+
+In both of these, we see that the mutant is slower than wild-type without pre-training but faster with it.
+This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
+With pre-training, some of them will now be potentiated (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_pre}), and the enhanced depression can speed up learning.
+
+With shorter pre-training, we see that it speeds up learning in both wild-type and mutant (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the mutant.
+This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning.
+
+We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_learnS} that longer pre-training pushes the synapses further down the cascade, slow in down learning for the wild-type.
+This effect is weaker for the mutant, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
 
@@
   \caption{Simulation results for multistate model with weak training ($\Delta f=-0.1$).
-  (\ref{fig:multistate_weak_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  Other parameters can be found in \autoref{tab:params}.
+  (\ref{fig:multistate_weak_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
   (\ref{fig:multistate_weak_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:multistate_weak_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:multistate_weak_eq_WT}) wild-type and (\ref{fig:multistate_weak_eq_KO}) \KO\ knockout.
-  (\ref{fig:multistate_weak_pr}) Evolution of probability distributions for (\ref{fig:multistate_weak_pr_WT_nopre}\ref{fig:multistate_weak_pr_WT_pre}) wild-type and  (\ref{fig:multistate_weak_pr_KO_nopre}\ref{fig:multistate_weak_pr_KO_pre}) \KO\ knockout without (\ref{fig:multistate_weak_pr_WT_nopre},\ref{fig:multistate_weak_pr_KO_nopre}) and with (\ref{fig:multistate_weak_pr_WT_pre},\ref{fig:multistate_weak_pr_KO_pre}) pre-training.
+  (\ref{fig:multistate_weak_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:multistate_weak_eq_WT}) wild-type and (\ref{fig:multistate_weak_eq_KO}) \KO\ mutant.
+  (\ref{fig:multistate_weak_pr}) Evolution of probability distributions for (\ref{fig:multistate_weak_pr_WT_nopre}\ref{fig:multistate_weak_pr_WT_pre}) wild-type and  (\ref{fig:multistate_weak_pr_KO_nopre}\ref{fig:multistate_weak_pr_KO_pre}) \KO\ mutant without (\ref{fig:multistate_weak_pr_WT_nopre},\ref{fig:multistate_weak_pr_KO_nopre}) and with (\ref{fig:multistate_weak_pr_WT_pre},\ref{fig:multistate_weak_pr_KO_pre}) pre-training.
 } \label{fig:multistate_weak}
@@
   \caption{Simulation results for multistate model with strong training ($\Delta f=-0.4$). 
-  (\ref{fig:multistate_strong_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  Other parameters can be found in \autoref{tab:params}.
+  (\ref{fig:multistate_strong_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
   (\ref{fig:multistate_strong_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:multistate_strong_eq}) Equilibrium distributions with/without gain-increase/decrease training for (\ref{fig:multistate_strong_eq_WT}) wild-type and (\ref{fig:multistate_strong_eq_KO}) \KO\ knockout.
-  (\ref{fig:multistate_strong_pr}) Evolution of probability distributions for (\ref{fig:multistate_strong_pr_WT_nopre}\ref{fig:multistate_strong_pr_WT_pre}) wild-type and  (\ref{fig:multistate_strong_pr_KO_nopre}\ref{fig:multistate_strong_pr_KO_pre}) \KO\ knockout without (\ref{fig:multistate_strong_pr_WT_nopre},\ref{fig:multistate_strong_pr_KO_nopre}) and with (\ref{fig:multistate_strong_pr_WT_pre},\ref{fig:multistate_strong_pr_KO_pre}) pre-training.
+  (\ref{fig:multistate_strong_eq}) Equilibrium distributions with/without gain-increase/decrease training for (\ref{fig:multistate_strong_eq_WT}) wild-type and (\ref{fig:multistate_strong_eq_KO}) \KO\ mutant.
+  (\ref{fig:multistate_strong_pr}) Evolution of probability distributions for (\ref{fig:multistate_strong_pr_WT_nopre}\ref{fig:multistate_strong_pr_WT_pre}) wild-type and  (\ref{fig:multistate_strong_pr_KO_nopre}\ref{fig:multistate_strong_pr_KO_pre}) \KO\ mutant without (\ref{fig:multistate_strong_pr_WT_nopre},\ref{fig:multistate_strong_pr_KO_nopre}) and with (\ref{fig:multistate_strong_pr_WT_pre},\ref{fig:multistate_strong_pr_KO_pre}) pre-training.
 } \label{fig:multistate_strong}
 \end{figure}
+
+The results of simulations of the multistate model can be seen in \autoref{fig:multistate_weak} and \autoref{fig:multistate_strong}.
+However, we can get some insight into this model analytically.
 
@@
 %
-So, we see that pre-training will speed up learning when $\Delta f$ is small.
-On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot.
+So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS}.
+On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.
+
+Intuitively, the flux depends on the slope of the distribution at the centre of the chain (with an offset due to the difference between $f\pot$ and $f\dep$).
+Pre-training has two effects: it produces a slope in the right direction (see \autoref{fig:multistate_weak}\ref{fig:multistate_weak_eq}), but it also reduces distribution at the centre (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}).
+For small $\Delta f$, the first effect is stronger and learning speeds up.
+For larger $\Delta f$, the second effect wins and learning slows down.
 
@@
 \begin{equation}\label{eq:multiKOnopre}
-  \Phi = -\frac{2(1-\beta)\beta^{M/2-1}\Delta fq}{1-\beta^M}.
+  \Phi = -2\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
 \end{equation}
 %
+This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$. 
+This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.
+
+There are two effects here as well.
+Smaller $\beta$ will increase the probability of crossing the centre of the chain, speeding up learning, but it will also concentrate probability at the ends of the chain, depleting the centre and slowing down learning.
+The first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ and will be more significant for smaller $\beta$ or in a longer chain.
+
 With pre-training:
@@
           {(1-2\Delta f)^M - \beta^M(1+2\Delta f)^M}   \,
-          \beta^{M/2-1}(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1}.
+          \beta^{M/2-1}(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1} \\
+       &= -4\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
 \end{aligned}
@@
 %
+Once again, we see that pre-training will speed up learning when $\Delta f$ is small, whereas, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning.
+
+Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
+As we would like pre-training to slow down learning in the wild-type but speed it up in the mutant, it would seem that we require $\Delta f^*(\beta,M) < \Delta f < \Delta f^*(1,M)$ (remembering once more that $\Delta f<0$ and that the wild-type corresponds to $\beta=1$).
+However, as seen in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, which would make this impossible.
+
+But, all is not lost.
+This is only the initial, instantaneous rate of change.
+Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f < \Delta f^*(1,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
+We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}) drifting to the centre.
+This happens sooner for the mutant than the wild-type, due to the stronger depressing transitions.
+This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the mutant.
+
+In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f < \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the mutant learns slower than wild-type without pre-training, and that pre-training speeds up learning in the mutant but slows it down in the wild-type.
+However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}), which is \emph{not} seen in the experiment.
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
+  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
+ \end{myenuma}
+ \end{center}
+  \caption{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$ and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$.}\label{fig:multistate_star}
+\end{figure}
 
@@
   \caption{Simulation results for two-state model.
-  (\ref{fig:binary_learn}) Learning curves for wild-type and \KO\ knockout with and without pre-training.
+  Parameters can be found in \autoref{tab:params}.
+  (\ref{fig:binary_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
   (\ref{fig:binary_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:binary_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:binary_eq_WT}) wild-type and (\ref{fig:binary_eq_KO}) \KO\ knockout.
-  (\ref{fig:binary_pr}) Evolution of probability distributions for (\ref{fig:binary_pr_WT_nopre}\ref{fig:binary_pr_WT_pre}) wild-type and  (\ref{fig:binary_pr_KO_nopre}\ref{fig:binary_pr_KO_pre}) \KO\ knockout without (\ref{fig:binary_pr_WT_nopre},\ref{fig:binary_pr_KO_nopre}) and with (\ref{fig:binary_pr_WT_pre},\ref{fig:binary_pr_KO_pre}) pre-training.
+  (\ref{fig:binary_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:binary_eq_WT}) wild-type and (\ref{fig:binary_eq_KO}) \KO\ mutant.
+  (\ref{fig:binary_pr}) Evolution of probability distributions for (\ref{fig:binary_pr_WT_nopre}\ref{fig:binary_pr_WT_pre}) wild-type and  (\ref{fig:binary_pr_KO_nopre}\ref{fig:binary_pr_KO_pre}) \KO\ mutant without (\ref{fig:binary_pr_WT_nopre},\ref{fig:binary_pr_KO_nopre}) and with (\ref{fig:binary_pr_WT_pre},\ref{fig:binary_pr_KO_pre}) pre-training.
 } \label{fig:binary_res}
@@
 
-This model can be solved exactly:
+The results of simulations of the multistate model can be seen in \autoref{fig:binary_res}.
+However, this model can be solved exactly:
 %
@@
 %
-where it's worth remembering that $\Delta f<0$.
+where, again, it's worth remembering that $\Delta f<0$.
 With pre-training:
@@
 %
-So, we see that pre-training will always speed up learning.
+So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
+This can be seen in \autoref{fig:binary_res}\ref{fig:binary_learn},\ref{fig:binary_learnS}.
 
@@
 %
-which is always larger that \eqref{eq:binWTnopre}.
+which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
+This is also seen in \autoref{fig:binary_res}\ref{fig:binary_learn},\ref{fig:binary_learnS}.
+As discussed below \eqref{eq:multiKOnopre}, the smaller values of has two effects.
+Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
+The second effect is much smaller for the very short chain considered here.
+
 With pre-training:
@@
 %
+As we've already ruled out this model, we won't waste any time analysing this formula.
 

### 2013-01-25 Notes/VORmodel.tex
added pooled resource model

@@
 \newcommand{\tpre}{t_{\text{pre}}}
+\newcommand{\ttrain}{t_{\text{train}}}
+\newcommand{\lmax}{_{\text{max}}}
+\newcommand{\lmin}{_{\text{min}}}
 % ----------------------------------------------------------------
@@
 \begin{abstract}
-  We see if we can model VOR gain increase and decrease learning in mice with a knockout in MHC as well as wild type.
+  We see if we can model VOR gain increase and decrease learning in mice with a knockout in MHC as well as wild-type.
 \end{abstract}
@@
   \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
-  \item The synaptic weights of the internal states are given by the column vector $\w$. This can only take two values that we can call $\pm1$.
+  \item The synaptic weights of the internal states are given by the column vector $\w$.
+This can only take two values that we can call $\pm1$, except for the pooled resource model that we will discuss in \sref{sec:pooledmodel} and \sref{sec:pooled}.
 \end{itemize}
@@
     % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $rt_\text{train}$ & $rt_\text{pre}$ \\
+    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
     \hline
@@
 \end{table}
+
+\subsubsection{Pooled resource model}\label{sec:pooledmodel}
+
+Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed.
+We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single pseudo-synapse.
+
+We will model the individual synapses with the two-state model. 
+Let $i$ be the number of them that are potentiated.
+We will model the effect of resource depletion linearly with the potentiation/depression probabilities for the individual synapses:
+%
+\begin{equation}\label{eq:depletion}
+  \begin{aligned}
+    q\pot &= \frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}, \qquad
+    q\dep &= \frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}.
+  \end{aligned}
+\end{equation}
+%
+At each plasticity event for the pseudo-synapse, one of the individual synapses will be chosen randomly for update.
+This effectively reduces the transition probabilities by $1/P$.
+
+This pseudo-synapse would seem to have $2^P$ internal states.
+However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
+The transition network topology will then have a multistate toplogy (see \autoref{fig:models}\ref{fig:multistate}) but the transition probabilities will no longer be uniform and the weight of the pseudo-synapse is the mean of its constituent synapses:
+%
+\begin{equation}\label{eq:pooledweight}
+  \w_i = \frac{2i}{P}-1.
+\end{equation}
+%
+
+The Markov process is lumpable \wrt this partition of states (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
+The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$.
+The result must be the same for all states in lump $i$.
+
+For any state in lump $i$, there are $P-i$ synapses that can be potentiated to go to lump $i+1$.
+Each of these transition probabilities is $q\pot/P$.
+Similarly, there are $i$ synapses that can be depressed to go to lump $i-1$.
+Each of these transition probabilities is $q\dep/P$.
+Thus:
+%
+\begin{equation}\label{eq:pooledpotdep}
+  \begin{aligned}
+    \M\pot_{ii+1} &=  \brk{\frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}} \frac{P-i}{P}, \\
+    \M\dep_{ii-1} &=  \brk{\frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}} \frac{i}{P},
+  \end{aligned}
+\end{equation}
+%
+with all other off-diagonal elements equal to zero.
+The diagonal elements are chosen so that the rows sum to one.
 
@@
 
-\subsubsection{Cascade}\label{sec:cascade}
+\subsubsection{Cascade model}\label{sec:cascade}
 
@@
 
-\subsubsection{Multistate}\label{sec:multistate}
+\subsubsection{Multistate model}\label{sec:multistate}
 
@@
 
-\subsubsection{Two-state}\label{sec:binary}
+\subsubsection{Two-state model}\label{sec:binary}
 
@@
 
+\subsubsection{Pooled resource model}\label{sec:pooled}
+
 %\subsection*{Acknowledgements}

### 2013-01-25 Code/VOR/MakeVORcomparison.m

@@ -30,6 +30,12 @@ if strcmpi(modelname,'multistate')
 elseif strcmpi(modelname,'cascade')
     [Wp,Wm,w]=CascadeOriginal(paramWT,paramWT,n);
     [~,WmKn]=CascadeOriginal(paramWT,paramKn,n);
+elseif strcmpi(modelname,'pooled')
+    w=(-1:2/(n-1):1)';
+    q=(paramWT(1):diff(paramWT)/(n-2):paramWT(2)).*(n-1:-1:1)/(n-1);
+    [Wp,Wm]=MakeSMS(q);
+    q=(paramKn(1):diff(paramKn)/(n-2):paramKn(2)).*(n-1:-1:1)/(n-1);
+    [~,WmKn]=MakeSMS(q);
 else
     error(['invalid model name: ' modelname]);
 end

### 2013-01-26 Notes/VORmodel.tex
pooled model figs

@@
   \item The potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $rf\potdep$, where $f\pot+f\dep=1$.
+      In other words, $r$ is the total rate of plasticity events per synapse and $f\potdep$ are the fraction of these events that are potentiating/depressing.
   \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
   \item The synaptic weights of the internal states are given by the column vector $\w$.
-This can only take two values that we can call $\pm1$, except for the pooled resource model that we will discuss in \sref{sec:pooledmodel} and \sref{sec:pooled}.
+This can only take two values that we can call $\pm1$, except for the pooled resource model that we will discuss in \autoref{sec:pooledmodel} and \autoref{sec:pooled}.
 \end{itemize}
@@
 gain-decrease training  will be described by $f\pot=f\pot\dec<f\pot\norm$.
-Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depends on $f\pot$, which we will indicate with subscripts.
+Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depend on $f\pot$, which we will indicate with subscripts.
 
@@
 If they are equal, this would change the mean synaptic weight in equilibrium.
-This seems like it would affect the ability of the network to perform its function, and one might expect adaptation to the environment to produce an equilibrium state that has the same performance.
+This seems like it could affect the ability of the network to perform its function, whereas, in reality, the wild-type and mutant have the same baseline performance.
 In any case, if the synaptic weights are different, the electrical activity will be different, and there will be no reason to expect the same statistics for potentiation or depression.
+In setting $f\pot\wt=f\pot\ko$, we are assuming that the rest of the brain compensates for the difference in synaptic weights somehow.
 
@@
 But, now that the synaptic weights are the same, the electrical activity will be the same, and there will be no reason to expect different statistics for potentiation or depression.
-If we did this for the multistate or two-state models, we'd remove all differences between the wild-type and \KO\ mutant.
+If we did this for the multistate or two-state models, we'd remove all differences between the wild-type and \KO\ mutant, so we won't.
 
@@
 We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade}), the multistate model (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary}).
+We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.
 
@@
 
-For the multistate and two state models, we will use the same value for the transition probabilities, $q$ for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
+For the multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
 We will use a larger value for $q$ for depression in the \KO\ mutant.
 
-In each case, we set $f\pot\norm=\half$, $f\pot\inc=f\pot\norm+\Delta f$ and $f\pot\dec=f\pot\norm-\Delta f$. The values of all these parameters are listed in \autoref{tab:params}.
+In each case, we set $f\pot\norm=\half$, $f\pot\inc=f\pot\norm+\Delta f$ and $f\pot\dec=f\pot\norm-\Delta f$. 
+We use the same values for wild-type and mutant and treat gain-increase and decrease symmetrically.
+The values of all these parameters are listed in \autoref{tab:params}.
 
@@
     Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5   \\
+    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
+                                          & -0.1 & 70 & 70 \\
+    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
+                                          & -0.1 & 70 & 70 \\
     \hline
@@
 The diagonal elements are chosen so that the rows sum to one.
+
+This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
+We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
+We will use larger values for depression in the \KO\ mutant.
 
@@
 
-The results of simulations of the multistate model can be seen in \autoref{fig:binary_res}.
+The results of simulations of the two-state model can be seen in \autoref{fig:binary_res}.
 However, this model can be solved exactly:
@@
 %
-But it is easier to just substitute $M=2$ into the formulae in \sref{sec:multistate}.
+But it is easier to just substitute $M=2$ into the formulae in \autoref{sec:multistate}.
 In this case, the initial rate of change encapsulates the whole solution, as there is only a single exponential decay.
@@
 So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
-This can be seen in \autoref{fig:binary_res}\ref{fig:binary_learn},\ref{fig:binary_learnS}.
+This can be seen in \autoref{fig:binary_res}\ref{fig:binary_learnS}.
 
@@
 
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_learn.eps}}\label{fig:pooled_plenty_learn}
+  \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_learnS.eps}}\label{fig:pooled_plenty_learnS}
+  \item\label{fig:pooled_plenty_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_eq_WT.eps}}\label{fig:pooled_plenty_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_eq_KO.eps}}\label{fig:pooled_plenty_eq_KO}
+                  \end{myenumi}
+  \item\label{fig:pooled_plenty_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_WT_nopre.eps}}\label{fig:pooled_plenty_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_WT_pre.eps}}\label{fig:pooled_plenty_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_KO_nopre.eps}}\label{fig:pooled_plenty_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_KO_pre.eps}}\label{fig:pooled_plenty_pr_KO_pre}
+                  \end{myenumi}
+ \end{myenuma}
+ \end{center}
+  \caption{Simulation results for pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$).
+  Parameters can be found in \autoref{tab:params}.
+  (\ref{fig:pooled_plenty_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
+  (\ref{fig:pooled_plenty_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:pooled_plenty_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:pooled_plenty_eq_WT}) wild-type and (\ref{fig:pooled_plenty_eq_KO}) \KO\ mutant.
+  (\ref{fig:pooled_plenty_pr}) Evolution of probability distributions for (\ref{fig:pooled_plenty_pr_WT_nopre}\ref{fig:pooled_plenty_pr_WT_pre}) wild-type and  (\ref{fig:pooled_plenty_pr_KO_nopre}\ref{fig:pooled_plenty_pr_KO_pre}) \KO\ mutant without (\ref{fig:pooled_plenty_pr_WT_nopre},\ref{fig:pooled_plenty_pr_KO_nopre}) and with (\ref{fig:pooled_plenty_pr_WT_pre},\ref{fig:pooled_plenty_pr_KO_pre}) pre-training.
+} \label{fig:pooled_plenty}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{pooled_scarce_learn.eps}}\label{fig:pooled_scarce_learn}
+  \item\aligntop{\includegraphics[width=7cm]{pooled_scarce_learnS.eps}}\label{fig:pooled_scarce_learnS}
+  \item\label{fig:pooled_scarce_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{pooled_scarce_eq_WT.eps}}\label{fig:pooled_scarce_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{pooled_scarce_eq_KO.eps}}\label{fig:pooled_scarce_eq_KO}
+                  \end{myenumi}
+  \item\label{fig:pooled_scarce_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{pooled_scarce_pr_WT_nopre.eps}}\label{fig:pooled_scarce_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{pooled_scarce_pr_WT_pre.eps}}\label{fig:pooled_scarce_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{pooled_scarce_pr_KO_nopre.eps}}\label{fig:pooled_scarce_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{pooled_scarce_pr_KO_pre.eps}}\label{fig:pooled_scarce_pr_KO_pre}
+                  \end{myenumi}
+ \end{myenuma}
+ \end{center}
+  \caption{Simulation results for pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$).
+  Parameters can be found in \autoref{tab:params}.
+  (\ref{fig:pooled_scarce_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
+  (\ref{fig:pooled_scarce_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:pooled_scarce_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:pooled_scarce_eq_WT}) wild-type and (\ref{fig:pooled_scarce_eq_KO}) \KO\ mutant.
+  (\ref{fig:pooled_scarce_pr}) Evolution of probability distributions for (\ref{fig:pooled_scarce_pr_WT_nopre}\ref{fig:pooled_scarce_pr_WT_pre}) wild-type and  (\ref{fig:pooled_scarce_pr_KO_nopre}\ref{fig:pooled_scarce_pr_KO_pre}) \KO\ mutant without (\ref{fig:pooled_scarce_pr_WT_nopre},\ref{fig:pooled_scarce_pr_KO_nopre}) and with (\ref{fig:pooled_scarce_pr_WT_pre},\ref{fig:pooled_scarce_pr_KO_pre}) pre-training.
+} \label{fig:pooled_scarce}
+\end{figure}
+
+The results of simulations of the pooled resource model can be seen in \autoref{fig:pooled_plenty} and \autoref{fig:pooled_scarce}.
+
+If we compare gain-increase learning in the mutant to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_eq}: \ref{fig:pooled_plenty_eq_WT} and \ref{fig:pooled_plenty_eq_KO} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_eq}: \ref{fig:pooled_scarce_eq_WT} and \ref{fig:pooled_scarce_eq_KO}), where there are fewer synapses available for depression and resources are depleted.
+When resource depletion is less severe, the first effect dominates and the mutant learns faster than wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learn},\ref{fig:pooled_plenty_learnS}).
+However, when resource depletion is more severe, the second effect dominates and the mutant learns slower than wild-type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learn},\ref{fig:pooled_scarce_learnS}), which matches what is seen in the experiment.
+
+Gain-decrease pre-training lessens the second effect, and results in the mutant learning faster than wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learn},\ref{fig:pooled_plenty_learnS} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learn},\ref{fig:pooled_scarce_learnS}), as seen in the experiment.
+
+Gain-decrease pre-training will shift the distribution to the potentiated site, where there are more synapses available for depression and resources are more plentiful, for both the mutant and wild type.
+This means that the pre-trained animal will learn faster then the untrained one for both mutant and wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learnS} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learnS}).
+This differs from what is seen experimentally, where the pre-trained wild-type learns slower than the untrained one.
+
 %\subsection*{Acknowledgements}

### 2013-01-26 Code/VOR/MakeVORfigs.m

@@ -13,15 +13,20 @@ DoPrint= strcmpi(PrintView,'Print');
 numpts=100;
 close('all');
 
-fig=figure('WindowStyle','docked','PaperPositionMode','auto');
-figs=figure('WindowStyle','docked','PaperPositionMode','auto');
-altfig(1)=figure('WindowStyle','docked','PaperPositionMode','auto');
-fig3(1)=figure('WindowStyle','docked','PaperPositionMode','auto');
 if DoPrint
-    altfig(2)=figure('WindowStyle','docked','PaperPositionMode','auto');
-    fig3(2)=figure('WindowStyle','docked','PaperPositionMode','auto');
-    fig3(3)=figure('WindowStyle','docked','PaperPositionMode','auto');
-    fig3(4)=figure('WindowStyle','docked','PaperPositionMode','auto');
+    fig=figure('PaperPositionMode','auto','Position',[60 60 1000 1000]);
+    figs=figure('PaperPositionMode','auto','Position',[60 60 1000 1000]);
+    altfig(1)=figure('PaperPositionMode','auto','Position',[60 60 1000 500]);
+    altfig(2)=figure('PaperPositionMode','auto','Position',[60 60 1000 500]);
+    fig3(1)=figure('PaperPositionMode','auto','Position',[60 60 600 500]);
+    fig3(2)=figure('PaperPositionMode','auto','Position',[60 60 600 500]);
+    fig3(3)=figure('PaperPositionMode','auto','Position',[60 60 600 500]);
+    fig3(4)=figure('PaperPositionMode','auto','Position',[60 60 600 500]);
+else
+    fig=figure('WindowStyle','docked','PaperPositionMode','auto');
+    figs=figure('WindowStyle','docked','PaperPositionMode','auto');
+    altfig=figure('WindowStyle','docked','PaperPositionMode','auto');
+    fig3=figure('WindowStyle','docked','PaperPositionMode','auto');
 end
 
 if strncmpi(prefix,'cascade',3)
@@ -63,6 +68,7 @@ if DoPrint
     print(fig3(2),[prefix '_pr_WT_pre.eps'],'-depsc');
     print(fig3(3),[prefix '_pr_KO_nopre.eps'],'-depsc');
     print(fig3(4),[prefix '_pr_KO_pre.eps'],'-depsc');
+    close('all');
 end
 
 

### 2013-01-26 Notes/Figures/MakeFigs.m

@@ -1,0 +1,7 @@
+MakeVORfigs('multistate_strong',0.6,0.8,-0.4,20,20,'print')
+MakeVORfigs('multistate_weak',0.6,0.8,-0.1,50,50,'print')
+MakeVORfigs('cascade_short',0.25,0.33,-0.3,50,50,'print')
+MakeVORfigs('cascade_long',0.25,0.33,-0.3,50,150,'print')
+MakeVORfigs('multistate_weak',0.6,0.8,-0.1,50,50,'print')
+MakeVORfigs('multistate_strong',0.6,0.8,-0.4,20,20,'print')
+MakeVORfigs('binary',0.4,0.8,-0.1,5,5,'print')

### 2013-01-26 Notes/VORmodel.tex
introductory blather

@@
 \section{Introduction and summary}\label{sec:intro}
+
+We will see if we can capture the essential features of VOR gain-increase learning in wild type and MHC \KO\ mutants and wild-type, with and without gain-decrease pre-training.
+We will try to do this using models of complex synapses, such as the cascade model \cite{Fusi2005cascade} or the multistate model \cite{amit1994learning,Fusi2007multistate}.
+We will also construct a new model in \autoref{sec:pooledmodel} that incorporates the idea that biasing the synapses in one direction could deplete some resource, thereby impairing further plasticity in the same direction in nearby synapses.
+
+These models all assume independent, statistically identical synapses.
+This has the benefit of freeing us from modelling the neurons and their networks, allowing us to focus on the synapses themselves.
+However, we will lose certain aspects of real synapses.
+For example, during gain-increase learning, certain synapses will be potentiated more often and others will be depressed more often.
+In these models, however, synaptic identity is not kept track of, so training can only be described by changing the overall levels of potentiation and depression.
+
+As the MHC \KO\ mutation lowers the threshold for LTD, we will model these mice by increasing the probabilities for internal synaptic transitions associated with depression, whilst leaving potentiation unchanged.
+
+We will be parsimonious in our modelling.
+We will assume that the wild-type has a symmetry between potentiation and depression and we will treat gain-increase and decrease training in a symmetric way.
+We will also assume that the rates of potentiation/depression in the mutant and wild-type are the same.
+None of these assumptions are necessary or justified experimentally (although \cite{Fusi2007multistate} notes that imbalance between potentiation and depression can drastically impair memory).
+They do reduce the number of adjustable parameters we have, simplifying searches for good models.
+Presumably, relaxing these assumptions and enlarging the parameter space would allow for a better fit to data, but the number of parameters would start to look large compared to the number of features we're trying to match.
+
+If we imagine something like an LN model for the Purkinje cell, the output would depend on some linear combination of the synaptic weights (weighted by the activities of the corresponding parallel fibres).
+If we're not keeping track of synaptic identity, the most natural linear combination to use would be an equal sum of them all.
+Thus we will quantify learning by the change in the mean synaptic weight.
+
+The output of the Purkinje cell, and the behavioural output of the animal, will depend on this quantity in some complicated nonlinear way.
+Modelling this is far beyond the scope of this work.
+Instead we will look at more qualitative features, asking whether learning is slower or faster in some situation compared to another.
+
+Within the parameter range explored, we will see that synaptic complexity is essential.
+The simple two-state model always shows enhanced learning in the mutant (see \autoref{fig:binary_res}), in contradiction with experiment.
+It seems that the enhanced depressing transition more than compensates for saturation effects.
+
+On the other hand, the cascade, multistate and pooled resource models all have parameter ranges where the mutant has impaired learning without pre-training (see \autoref{fig:cascade_short}, \autoref{fig:cascade_long}, \autoref{fig:multistate_weak}, \autoref{fig:multistate_strong} and \autoref{fig:pooled_scarce}).
+We have found parameter ranges where the reverse happens for the multistate (see discussion below \eqref{eq:multiKOnopre}) and the pooled resource model (see \autoref{fig:pooled_plenty}).
+These rely on properties of the mutation, so it would be difficult to explore with further experiments.
+
+All of these models show pre-training enhancing learning for the mutant, as expected.
+It is harder to find regimes where pre-training impairs learning for the wild-type.
+In the cascade model, this required a much longer duration for pe-training (see \autoref{fig:cascade_long}).
+In the multistate model, this required stronger training, which resulted in gain-decrease learning being faster for the mutant then wild-type (see \autoref{fig:multistate_strong}), unlike what is seen experimentally.
+It is possible that these features could be improved by lessening the symmetry between gain-increase and decrease or potentiation and depression.
+We haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.
 
@@
 Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed.
-We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single pseudo-synapse.
+We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single compound synapse.
 
@@
 %
-At each plasticity event for the pseudo-synapse, one of the individual synapses will be chosen randomly for update.
+At each plasticity event for the compound synapse, one of the individual synapses will be chosen randomly for update.
 This effectively reduces the transition probabilities by $1/P$.
 
-This pseudo-synapse would seem to have $2^P$ internal states.
+This compound synapse would seem to have $2^P$ internal states.
 However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
-The transition network topology will then have a multistate toplogy (see \autoref{fig:models}\ref{fig:multistate}) but the transition probabilities will no longer be uniform and the weight of the pseudo-synapse is the mean of its constituent synapses:
+The transition network topology will then have a multistate toplogy (see \autoref{fig:models}\ref{fig:multistate}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
 %

### 2013-01-29 Notes/VORmodel.tex
row refs

@@
 \graphicspath{{"Figures/"}}
+%
+\usepackage{array}
+\newcounter{tablerow}
+\newcommand{\tablerowautorefname}{row}
+\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
+%\usepackage{etoolbox}
+%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
 %
@@
  \begin{center}
-  \begin{tabular}{|l|c|c|c|c|c|c|}
-    \hline
+  \begin{tabularn}{|l|c|c|c|c|c|c|}
+%  \begin{tabular}{|l|c|c|c|c|c|c|>{\refstepcounter{tablerow}}l}
+    \cline{1-7}
     % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
     Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
-    \hline
-    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50  \\
-    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150 \\
-    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50  \\
-    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20  \\
-    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5   \\
+    \cline{1-7}
+    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50  &\label{tr:cascade_short} \\
+    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150 &\label{tr:cascade_long} \\
+    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50  &\label{tr:multistate_weak} \\
+    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20  & \label{tr:multistate_strong} \\
+    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5   &\label{tr:binary} \\
     Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-                                          & -0.1 & 70 & 70 \\
+                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
     Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
-                                          & -0.1 & 70 & 70 \\
-    \hline
-  \end{tabular}
+                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
+    \cline{1-7}
+  \end{tabularn}
  \end{center}
@@
   \caption{Simulation results for cascade model with short pre-training ($rt_\text{pre}=50$).
-  Other parameters can be found in \autoref{tab:params}.
+  Other parameters can be found in \autoref{tr:cascade_short} of \autoref{tab:params}.
   (\ref{fig:cascade_short_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
@@
   \caption{Simulation results for cascade model with long pre-training ($rt_\text{pre}=150$).
-  Other parameters can be found in \autoref{tab:params}.
+  Other parameters can be found in \autoref{tr:cascade_long} of \autoref{tab:params}.
   (\ref{fig:cascade_long_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
@@
   \caption{Simulation results for multistate model with weak training ($\Delta f=-0.1$).
-  Other parameters can be found in \autoref{tab:params}.
+  Other parameters can be found in \autoref{tr:multistate_weak} of \autoref{tab:params}.
   (\ref{fig:multistate_weak_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
@@
   \caption{Simulation results for multistate model with strong training ($\Delta f=-0.4$).
-  Other parameters can be found in \autoref{tab:params}.
+  Other parameters can be found in \autoref{tr:multistate_strong} of \autoref{tab:params}.
   (\ref{fig:multistate_strong_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
@@
   \caption{Simulation results for two-state model.
-  Parameters can be found in \autoref{tab:params}.
+  Parameters can be found in \autoref{tr:binary} of \autoref{tab:params}.
   (\ref{fig:binary_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
@@
   \caption{Simulation results for pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$).
-  Parameters can be found in \autoref{tab:params}.
+  Parameters can be found in \autoref{tr:pooled_plenty} of \autoref{tab:params}.
   (\ref{fig:pooled_plenty_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
@@
   \caption{Simulation results for pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$).
-  Parameters can be found in \autoref{tab:params}.
+  Parameters can be found in \autoref{tr:pooled_scarce} of \autoref{tab:params}.
   (\ref{fig:pooled_scarce_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.

### 2013-01-29 Notes/VORmodel.tex
typo and blah

@@
 It is possible that these features could be improved by lessening the symmetry between gain-increase and decrease or potentiation and depression.
-We haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.
+For the pooled resource model, we haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.
+
+There seems to be two take home messages.
+First, impaired learning without pre-training in the mutant requires synaptic complexity, as it is not seen in the simple two-state model.
+Second, impaired learning with pre-training in the wild-type requires a form of metaplasticity where repeated potentiation makes subsequent depression less likely.
+This is a feature of the cascade model and our version of the multistate model, but not the pooled resource model.
 
@@
 
-We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade}), the multistate model (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary}).
+We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade}), the multistate model\footnote{The version of the multistate model that we'll study is slightly different than the original definition.
+We will only allow the synaptic weight to take two values, whereas it originally varied linearly across the states.
+} (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary}).
 We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.

### 2013-02-04 Notes/VORmodel.tex
simplified figures

@@
 We will assume that the wild-type has a symmetry between potentiation and depression and we will treat gain-increase and decrease training in a symmetric way.
-We will also assume that the rates of potentiation/depression in the mutant and wild-type are the same.
+We will also assume that the relative rates of potentiation/depression in the mutant and wild-type are the same and that the overall rate of plasticity events is the same in all situations.
 None of these assumptions are necessary or justified experimentally (although \cite{Fusi2007multistate} notes that imbalance between potentiation and depression can drastically impair memory).
@@
 Within the parameter range explored, we will see that synaptic complexity is essential.
-The simple two-state model always shows enhanced learning in the mutant (see \autoref{fig:binary_res}), in contradiction with experiment.
+The simple two-state model always shows enhanced learning in the mutant (see \autoref{fig:binary}), in contradiction with experiment.
 It seems that the enhanced depressing transition more than compensates for saturation effects.
@@
 In the multistate model, this required stronger training, which resulted in gain-decrease learning being faster for the mutant then wild-type (see \autoref{fig:multistate_strong}), unlike what is seen experimentally.
-It is possible that these features could be improved by lessening the symmetry between gain-increase and decrease or potentiation and depression.
+It is possible that these features could be improved by lessening the symmetry between gain-increase and decrease or potentiation and depression, or by allowing the rate of plasticity events to vary.
 For the pooled resource model, we haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.
@@
 
+We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equlibrium weights mentioned above.
+
 \section{Simulations}\label{sec:sims}
@@
 
-We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade}), the multistate model\footnote{The version of the multistate model that we'll study is slightly different than the original definition.
+We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the multistate model\footnote{The version of the multistate model that we'll study is slightly different than the original definition.
 We will only allow the synaptic weight to take two values, whereas it originally varied linearly across the states.
-} (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary}).
+} (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary_model}).
 We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.
@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary}
+  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}\hspace{0.5cm}
+  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
+  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}
  \end{myenuma}
@@
   \caption{Transition probabilities for different models.
-  (\ref{fig:cascade}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
-  (\ref{fig:multistate}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
-  (\ref{fig:binary}) The two-state model is parameterised by the two transition probailities.} \label{fig:models}
+  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
+  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
+  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probailities.} \label{fig:models}
 \end{figure}
@@
   \begin{tabularn}{|l|c|c|c|c|c|c|}
-%  \begin{tabular}{|l|c|c|c|c|c|c|>{\refstepcounter{tablerow}}l}
     \cline{1-7}
@@
     \cline{1-7}
+%   \hline
+%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
+%    \cline{1-7}
+%    \label{tr:cascade_short}&
+%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
+%    \label{tr:cascade_long}&
+%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
+%    \label{tr:multistate_weak}&
+%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
+%    \label{tr:multistate_strong}&
+%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
+%    \label{tr:binary}&
+%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
+%    \label{tr:pooled_plenty}&
+%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
+%                                          & -0.1 & 70 & 70 \\
+%    \label{tr:pooled_scarce}&
+%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
+%                                          & -0.1 & 70 & 70 \\
+%    \cline{1-7}
   \end{tabularn}
@@
 However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
-The transition network topology will then have a multistate toplogy (see \autoref{fig:models}\ref{fig:multistate}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
+The transition network topology will then have a multistate toplogy (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
 %
@@
 
-\subsubsection{Cascade model}\label{sec:cascade}
-
-\begin{figure}
+\newcommand{\resultsfig}[2]{\begin{figure}
  \begin{center}
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{cascade_short_learn.eps}}\label{fig:cascade_short_learn}
-  \item\aligntop{\includegraphics[width=7cm]{cascade_short_learnS.eps}}\label{fig:cascade_short_learnS}
-  \item\label{fig:cascade_short_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{cascade_short_eq_WT.eps}}\label{fig:cascade_short_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{cascade_short_eq_KO.eps}}\label{fig:cascade_short_eq_KO}
+  \item\aligntop{\includegraphics[width=7cm]{#1_learn.eps}}\label{fig:#1_learn}
+  \item\aligntop{\includegraphics[width=7cm]{#1_learnS.eps}}\label{fig:#1_learnS}
+  \item\label{fig:#1_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_WT.eps}}\label{fig:#1_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_KO.eps}}\label{fig:#1_eq_KO}
                   \end{myenumi}
-  \item\label{fig:cascade_short_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_WT_nopre.eps}}\label{fig:cascade_short_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_WT_pre.eps}}\label{fig:cascade_short_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_KO_nopre.eps}}\label{fig:cascade_short_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{cascade_short_pr_KO_pre.eps}}\label{fig:cascade_short_pr_KO_pre}
+  \item\label{fig:#1_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_nopre.eps}}\label{fig:#1_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_pre.eps}}\label{fig:#1_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_nopre.eps}}\label{fig:#1_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_pre.eps}}\label{fig:#1_pr_KO_pre}
                   \end{myenumi}
@@
  \end{center}
-  \caption{Simulation results for cascade model with short pre-training ($rt_\text{pre}=50$).
-  Other parameters can be found in \autoref{tr:cascade_short} of \autoref{tab:params}.
-  (\ref{fig:cascade_short_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:cascade_short_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:cascade_short_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_short_eq_WT}) wild-type and (\ref{fig:cascade_short_eq_KO}) \KO\ mutant.
-  (\ref{fig:cascade_short_pr}) Evolution of probability distributions for (\ref{fig:cascade_short_pr_WT_nopre}\ref{fig:cascade_short_pr_WT_pre}) wild-type and  (\ref{fig:cascade_short_pr_KO_nopre}\ref{fig:cascade_short_pr_KO_pre}) \KO\ mutant without (\ref{fig:cascade_short_pr_WT_nopre},\ref{fig:cascade_short_pr_KO_nopre}) and with (\ref{fig:cascade_short_pr_WT_pre},\ref{fig:cascade_short_pr_KO_pre}) pre-training.
-} \label{fig:cascade_short}
-\end{figure}
+  \caption[#2]{#2.
+  Other parameters can be found in \autoref{tr:#1} of \autoref{tab:params}.
+  Time is measured in units of $1/r$, the mean time between plasticity events.
+  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
+  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant.
+  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre}\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre}\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training.
+} \label{fig:#1}
+\end{figure}}
 
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{cascade_long_learn.eps}}\label{fig:cascade_long_learn}
-  \item\aligntop{\includegraphics[width=7cm]{cascade_long_learnS.eps}}\label{fig:cascade_long_learnS}
-  \item\label{fig:cascade_long_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{cascade_long_eq_WT.eps}}\label{fig:cascade_long_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{cascade_long_eq_KO.eps}}\label{fig:cascade_long_eq_KO}
-                  \end{myenumi}
-  \item\label{fig:cascade_long_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_WT_nopre.eps}}\label{fig:cascade_long_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_WT_pre.eps}}\label{fig:cascade_long_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_KO_nopre.eps}}\label{fig:cascade_long_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{cascade_long_pr_KO_pre.eps}}\label{fig:cascade_long_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}
-  \caption{Simulation results for cascade model with long pre-training ($rt_\text{pre}=150$).
-  Other parameters can be found in \autoref{tr:cascade_long} of \autoref{tab:params}.
-  (\ref{fig:cascade_long_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:cascade_long_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:cascade_long_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:cascade_long_eq_WT}) wild-type and (\ref{fig:cascade_long_eq_KO}) \KO\ mutant.
-  (\ref{fig:cascade_long_pr}) Evolution of probability distributions for (\ref{fig:cascade_long_pr_WT_nopre}\ref{fig:cascade_long_pr_WT_pre}) wild-type and  (\ref{fig:cascade_long_pr_KO_nopre}\ref{fig:cascade_long_pr_KO_pre}) \KO\ mutant without (\ref{fig:cascade_long_pr_WT_nopre},\ref{fig:cascade_long_pr_KO_nopre}) and with (\ref{fig:cascade_long_pr_WT_pre},\ref{fig:cascade_long_pr_KO_pre}) pre-training.
-} \label{fig:cascade_long}
-\end{figure}
+\subsubsection{Cascade model}\label{sec:cascade}
+
+\resultsfig{cascade_short}{Simulation results for cascade model with short pre-training ($rt_\text{pre}=50$)}
+
+\resultsfig{cascade_long}{Simulation results for cascade model with long pre-training ($rt_\text{pre}=150$)}
 
@@
 
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_weak_learn.eps}}\label{fig:multistate_weak_learn}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_weak_learnS.eps}}\label{fig:multistate_weak_learnS}
-  \item\label{fig:multistate_weak_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{multistate_weak_eq_WT.eps}}\label{fig:multistate_weak_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{multistate_weak_eq_KO.eps}}\label{fig:multistate_weak_eq_KO}
-                  \end{myenumi}
-  \item\label{fig:multistate_weak_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_WT_nopre.eps}}\label{fig:multistate_weak_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_WT_pre.eps}}\label{fig:multistate_weak_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_KO_nopre.eps}}\label{fig:multistate_weak_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{multistate_weak_pr_KO_pre.eps}}\label{fig:multistate_weak_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}
-  \caption{Simulation results for multistate model with weak training ($\Delta f=-0.1$).
-  Other parameters can be found in \autoref{tr:multistate_weak} of \autoref{tab:params}.
-  (\ref{fig:multistate_weak_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:multistate_weak_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:multistate_weak_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:multistate_weak_eq_WT}) wild-type and (\ref{fig:multistate_weak_eq_KO}) \KO\ mutant.
-  (\ref{fig:multistate_weak_pr}) Evolution of probability distributions for (\ref{fig:multistate_weak_pr_WT_nopre}\ref{fig:multistate_weak_pr_WT_pre}) wild-type and  (\ref{fig:multistate_weak_pr_KO_nopre}\ref{fig:multistate_weak_pr_KO_pre}) \KO\ mutant without (\ref{fig:multistate_weak_pr_WT_nopre},\ref{fig:multistate_weak_pr_KO_nopre}) and with (\ref{fig:multistate_weak_pr_WT_pre},\ref{fig:multistate_weak_pr_KO_pre}) pre-training.
-} \label{fig:multistate_weak}
-\end{figure}
+\resultsfig{multistate_weak}{Simulation results for multistate model with weak training ($\Delta f=-0.1$)}
 
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_strong_learn.eps}}\label{fig:multistate_strong_learn}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_strong_learnS.eps}}\label{fig:multistate_strong_learnS}
-  \item\label{fig:multistate_strong_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{multistate_strong_eq_WT.eps}}\label{fig:multistate_strong_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{multistate_strong_eq_KO.eps}}\label{fig:multistate_strong_eq_KO}
-                  \end{myenumi}
-  \item\label{fig:multistate_strong_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_WT_nopre.eps}}\label{fig:multistate_strong_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_WT_pre.eps}}\label{fig:multistate_strong_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_KO_nopre.eps}}\label{fig:multistate_strong_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{multistate_strong_pr_KO_pre.eps}}\label{fig:multistate_strong_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}
-  \caption{Simulation results for multistate model with strong training ($\Delta f=-0.4$).
-  Other parameters can be found in \autoref{tr:multistate_strong} of \autoref{tab:params}.
-  (\ref{fig:multistate_strong_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:multistate_strong_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:multistate_strong_eq}) Equilibrium distributions with/without gain-increase/decrease training for (\ref{fig:multistate_strong_eq_WT}) wild-type and (\ref{fig:multistate_strong_eq_KO}) \KO\ mutant.
-  (\ref{fig:multistate_strong_pr}) Evolution of probability distributions for (\ref{fig:multistate_strong_pr_WT_nopre}\ref{fig:multistate_strong_pr_WT_pre}) wild-type and  (\ref{fig:multistate_strong_pr_KO_nopre}\ref{fig:multistate_strong_pr_KO_pre}) \KO\ mutant without (\ref{fig:multistate_strong_pr_WT_nopre},\ref{fig:multistate_strong_pr_KO_nopre}) and with (\ref{fig:multistate_strong_pr_WT_pre},\ref{fig:multistate_strong_pr_KO_pre}) pre-training.
-} \label{fig:multistate_strong}
-\end{figure}
+\resultsfig{multistate_strong}{Simulation results for multistate model with strong training ($\Delta f=-0.4$)}
 
@@
 
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{binary_learn.eps}}\label{fig:binary_learn}
-  \item\aligntop{\includegraphics[width=7cm]{binary_learnS.eps}}\label{fig:binary_learnS}
-  \item\label{fig:binary_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{binary_eq_WT.eps}}\label{fig:binary_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{binary_eq_KO.eps}}\label{fig:binary_eq_KO}
-                  \end{myenumi}
-  \item\label{fig:binary_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_WT_nopre.eps}}\label{fig:binary_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_WT_pre.eps}}\label{fig:binary_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_KO_nopre.eps}}\label{fig:binary_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{binary_pr_KO_pre.eps}}\label{fig:binary_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}
-  \caption{Simulation results for two-state model.
-  Parameters can be found in \autoref{tr:binary} of \autoref{tab:params}.
-  (\ref{fig:binary_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:binary_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:binary_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:binary_eq_WT}) wild-type and (\ref{fig:binary_eq_KO}) \KO\ mutant.
-  (\ref{fig:binary_pr}) Evolution of probability distributions for (\ref{fig:binary_pr_WT_nopre}\ref{fig:binary_pr_WT_pre}) wild-type and  (\ref{fig:binary_pr_KO_nopre}\ref{fig:binary_pr_KO_pre}) \KO\ mutant without (\ref{fig:binary_pr_WT_nopre},\ref{fig:binary_pr_KO_nopre}) and with (\ref{fig:binary_pr_WT_pre},\ref{fig:binary_pr_KO_pre}) pre-training.
-} \label{fig:binary_res}
-\end{figure}
+\resultsfig{binary}{Simulation results for two-state model}
 
-The results of simulations of the two-state model can be seen in \autoref{fig:binary_res}.
+The results of simulations of the two-state model can be seen in \autoref{fig:binary}.
 However, this model can be solved exactly:
@@
 So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
-This can be seen in \autoref{fig:binary_res}\ref{fig:binary_learnS}.
+This can be seen in \autoref{fig:binary}\ref{fig:binary_learnS}.
 
@@
 which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
-This is also seen in \autoref{fig:binary_res}\ref{fig:binary_learn},\ref{fig:binary_learnS}.
+This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS}.
 As discussed below \eqref{eq:multiKOnopre}, the smaller values of has two effects.
@@
 
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_learn.eps}}\label{fig:pooled_plenty_learn}
-  \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_learnS.eps}}\label{fig:pooled_plenty_learnS}
-  \item\label{fig:pooled_plenty_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_eq_WT.eps}}\label{fig:pooled_plenty_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{pooled_plenty_eq_KO.eps}}\label{fig:pooled_plenty_eq_KO}
-                  \end{myenumi}
-  \item\label{fig:pooled_plenty_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_WT_nopre.eps}}\label{fig:pooled_plenty_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_WT_pre.eps}}\label{fig:pooled_plenty_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_KO_nopre.eps}}\label{fig:pooled_plenty_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{pooled_plenty_pr_KO_pre.eps}}\label{fig:pooled_plenty_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}
-  \caption{Simulation results for pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$).
-  Parameters can be found in \autoref{tr:pooled_plenty} of \autoref{tab:params}.
-  (\ref{fig:pooled_plenty_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:pooled_plenty_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:pooled_plenty_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:pooled_plenty_eq_WT}) wild-type and (\ref{fig:pooled_plenty_eq_KO}) \KO\ mutant.
-  (\ref{fig:pooled_plenty_pr}) Evolution of probability distributions for (\ref{fig:pooled_plenty_pr_WT_nopre}\ref{fig:pooled_plenty_pr_WT_pre}) wild-type and  (\ref{fig:pooled_plenty_pr_KO_nopre}\ref{fig:pooled_plenty_pr_KO_pre}) \KO\ mutant without (\ref{fig:pooled_plenty_pr_WT_nopre},\ref{fig:pooled_plenty_pr_KO_nopre}) and with (\ref{fig:pooled_plenty_pr_WT_pre},\ref{fig:pooled_plenty_pr_KO_pre}) pre-training.
-} \label{fig:pooled_plenty}
-\end{figure}
+\resultsfig{pooled_plenty}{Simulation results for pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$)}
 
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{pooled_scarce_learn.eps}}\label{fig:pooled_scarce_learn}
-  \item\aligntop{\includegraphics[width=7cm]{pooled_scarce_learnS.eps}}\label{fig:pooled_scarce_learnS}
-  \item\label{fig:pooled_scarce_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{pooled_scarce_eq_WT.eps}}\label{fig:pooled_scarce_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{pooled_scarce_eq_KO.eps}}\label{fig:pooled_scarce_eq_KO}
-                  \end{myenumi}
-  \item\label{fig:pooled_scarce_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{pooled_scarce_pr_WT_nopre.eps}}\label{fig:pooled_scarce_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{pooled_scarce_pr_WT_pre.eps}}\label{fig:pooled_scarce_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{pooled_scarce_pr_KO_nopre.eps}}\label{fig:pooled_scarce_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{pooled_scarce_pr_KO_pre.eps}}\label{fig:pooled_scarce_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}
-  \caption{Simulation results for pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$).
-  Parameters can be found in \autoref{tr:pooled_scarce} of \autoref{tab:params}.
-  (\ref{fig:pooled_scarce_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:pooled_scarce_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:pooled_scarce_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:pooled_scarce_eq_WT}) wild-type and (\ref{fig:pooled_scarce_eq_KO}) \KO\ mutant.
-  (\ref{fig:pooled_scarce_pr}) Evolution of probability distributions for (\ref{fig:pooled_scarce_pr_WT_nopre}\ref{fig:pooled_scarce_pr_WT_pre}) wild-type and  (\ref{fig:pooled_scarce_pr_KO_nopre}\ref{fig:pooled_scarce_pr_KO_pre}) \KO\ mutant without (\ref{fig:pooled_scarce_pr_WT_nopre},\ref{fig:pooled_scarce_pr_KO_nopre}) and with (\ref{fig:pooled_scarce_pr_WT_pre},\ref{fig:pooled_scarce_pr_KO_pre}) pre-training.
-} \label{fig:pooled_scarce}
-\end{figure}
+\resultsfig{pooled_scarce}{Simulation results for pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$)}
 
@@
 When resource depletion is less severe, the first effect dominates and the mutant learns faster than wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learn},\ref{fig:pooled_plenty_learnS}).
+This is reasonable, as when resource depletion is removed it will reduce to the two-state model, which showed this behaviour.
 However, when resource depletion is more severe, the second effect dominates and the mutant learns slower than wild-type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learn},\ref{fig:pooled_scarce_learnS}), which matches what is seen in the experiment.

### 2013-02-06 Notes/VORmodel.tex
typo

@@
   \qquad
-  \frg = f\pot\M\pot+f\dep\M\dep-\I.
+  \frg = f\pot\M\pot+f\dep\M\dep-\I,
 \end{equation}
+where $\I$ is the identity matrix.
 %
@@
 Gain-increase training  will be described by $f\pot=f\pot\inc<f\pot\norm$, and
-gain-decrease training  will be described by $f\pot=f\pot\dec<f\pot\norm$.
+gain-decrease training  will be described by $f\pot=f\pot\dec>f\pot\norm$.
 Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depend on $f\pot$, which we will indicate with subscripts.

### 2013-02-08 Notes/VORmodel.tex
talked about less symmetry between gain up & down

@@
 Instead we will look at more qualitative features, asking whether learning is slower or faster in some situation compared to another.
+Provided that the nonlinearity is monotonic, these types of questions will not be affected by it.
 
@@
 \begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
-          {(1-2\Delta f)^M - \beta^M(1+2\Delta f)^M}   \,
+  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f) - \beta(1-2\Delta f)}
+          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M}   \,
           \beta^{M/2-1}(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1} \\
@@
 However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}), which is \emph{not} seen in the experiment.
+
+Now consider what would happen if we did not treat gain-increase and gain-decrease symmetrically.
+Let us set $f\pot\dec = f\pot\norm - \overline{\Delta f}$.
+Then, \eqref{eq:multiKOpre} becomes
+%
+\begin{equation}\label{eq:multiKOdiffpre}
+\begin{aligned}
+  \Phi &= -2\prn{\Delta f + \overline{\Delta f}} q \, \frac{(1+2\overline{\Delta f}) - \beta(1-2\overline{\Delta f})}
+          {(1+2\overline{\Delta f})^M - \beta^M(1-2\overline{\Delta f})^M}   \,
+          \beta^{M/2-1}(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} \\
+       &= -2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
+\end{aligned}
+\end{equation}
+%
+We can then look at the wild-type by setting $\beta=1$:
+%
+\begin{equation}\label{eq:multiWTdiffpre}
+\begin{aligned}
+  \Phi &= -8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \, 
+          \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
+          {(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}   \\
+       &= -\frac{2\prn{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
+\end{aligned}
+\end{equation}
+%
+
+We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the mutant, the important thing is to have \emph{strong pre-training}, not strong training.
+Thus, removing the symmetry between gain-increase and gain-decrease will not help with the unrealistic feature of this model -- the fact that pre-training proceeds much faster in the mutant than wild-type.
 

### 2013-02-10 Notes/VORmodel.tex
started OG multistate section

@@
   (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
-  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probailities.} \label{fig:models}
+  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.} \label{fig:models}
 \end{figure}
@@
 
-\resultsfig{cascade_short}{Simulation results for cascade model with short pre-training ($rt_\text{pre}=50$)}
+\resultsfig{cascade_short}{Simulation results for the cascade model with short pre-training ($rt_\text{pre}=50$)}
 
-\resultsfig{cascade_long}{Simulation results for cascade model with long pre-training ($rt_\text{pre}=150$)}
+\resultsfig{cascade_long}{Simulation results for the cascade model with long pre-training ($rt_\text{pre}=150$)}
 
@@
 
-\resultsfig{multistate_weak}{Simulation results for multistate model with weak training ($\Delta f=-0.1$)}
+\resultsfig{multistate_weak}{Simulation results for the multistate model with weak training ($\Delta f=-0.1$)}
 
-\resultsfig{multistate_strong}{Simulation results for multistate model with strong training ($\Delta f=-0.4$)}
+\resultsfig{multistate_strong}{Simulation results for the multistate model with strong training ($\Delta f=-0.4$)}
 
@@
 
+\subsubsection{OG multistate model}\label{sec:multistate_lin}
+
+In this section, we will consider the multistate model as originally defined in \cite{amit1994learning}, \ie with linearly varying synaptic weight:
+%
+\begin{equation}\label{eq:multistateLinWeight}
+  \w_i = \frac{2i-M-1}{M-1}.
+\end{equation}
+%
+The equilibrium distribution, \eqref{eq:mutltieq}, still applies.
+However, now the rate of change of our learning metric, \eqref{eq:learning}, will be proportional to the sum of the net fluxes between adjacent states:
+%
+\begin{equation}\label{eq:multiLinFlux}
+  \begin{aligned}
+    \Phi &= \sum_{i=1}^{M-1} \eq_{i+1} f'{}\dep q\dep - \eq_i f'{}\pot q\pot 
+         &= \sum_{i=1}^{M-1} \eq_i (\alpha-\alpha') f'{}\dep q\dep \\
+         &= \frac{1-\alpha^{M-1}}{1-\alpha^M} \, (\alpha-\alpha') f'{}\dep q\dep.
+  \end{aligned}
+\end{equation}
+%
+
+First, consider the wild-type, for which $q\pot=q\dep=q$.
+Without pre-training:
+%
+\begin{equation}\label{eq:multiLinWTnopre}
+  \Phi = -2\Delta fq\,\frac{M-1}{M},
+\end{equation}
+%
+where it's worth remembering that $\Delta f<0$.
+With pre-training:
+%
+\begin{equation}\label{eq:multiLinWTpre}
+\begin{aligned}
+  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
+          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
+       &= -{4\Delta f q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
+\end{aligned}
+\end{equation}
+%
+
+Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
+Without pre-training:
+%
+\begin{equation}\label{eq:multiLinKOnopre}
+  \Phi = -2\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
+\end{equation}
+%
+With pre-training:
+%
+\begin{equation}\label{eq:multiLinKOpre}
+\begin{aligned}
+  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
+          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M} \\
+       &= -4\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
+\end{aligned}
+\end{equation}
+%
+
 \subsubsection{Two-state model}\label{sec:binary}
@@
 
+We made a number of assumptions in our analysis so far.
+Before we declare this model to be falsified, we should think about what would happen if we relaxed these assumptions and explored a larger parameter space.
+
+First, we assumed that the wild type had symmetric potentiation and depression.
+Relaxing this assumption would correspond modelling the wild-type like the mutant, but with a different value for $\beta$.
+We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the mutant.
+
+Secondly, we assumed that there would be equal rates of potentiation and depression in the untrained situation.
+However, relaxing this assumption merely changes \eqref{eq:binKOnopre} to
+%
+\begin{equation}\label{eq:binKOnoprediff}
+  \Phi = -\frac{\Delta f q}{f\dep+\beta f\pot}.
+\end{equation}
+%
+This will not change our conclusions either.
+
+As we didn't need the pre-training paradigm to rule out this model, we don't need to look at removing the symmetry between gain-increase and decrease.
+
 \subsubsection{Pooled resource model}\label{sec:pooled}
 
-\resultsfig{pooled_plenty}{Simulation results for pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$)}
+\resultsfig{pooled_plenty}{Simulation results for the pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$)}
 
-\resultsfig{pooled_scarce}{Simulation results for pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$)}
+\resultsfig{pooled_scarce}{Simulation results for the pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$)}
 

### 2013-02-12 Notes/VORmodel.tex
Fixed mistake in multistate analysis
Included multistate figs with intermediate training strength.

@@
     Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
-    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20  & \label{tr:multistate_strong} \\
+    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
+    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
     Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
@@
 
+\resultsfig{multistate_med}{Simulation results for the multistate model with intermediate training ($\Delta f=-0.17$)}
+
 \resultsfig{multistate_strong}{Simulation results for the multistate model with strong training ($\Delta f=-0.4$)}
 
-The results of simulations of the multistate model can be seen in \autoref{fig:multistate_weak} and \autoref{fig:multistate_strong}.
+The results of simulations of the multistate model can be seen in \autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
 However, we can get some insight into this model analytically.
@@
 Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
+Values of $\Delta f$ that are smaller than this (stronger training) will correspond to slower learning after pre-training.
 As we would like pre-training to slow down learning in the wild-type but speed it up in the mutant, it would seem that we require $\Delta f^*(\beta,M) < \Delta f < \Delta f^*(1,M)$ (remembering once more that $\Delta f<0$ and that the wild-type corresponds to $\beta=1$).
-However, as seen in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, which would make this impossible.
+As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) < \Delta f^*(1,M)$, so this range always exists.
+We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.
 
-But, all is not lost.
-This is only the initial, instantaneous rate of change.
+However, this is only the initial, instantaneous rate of change.
 Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f < \Delta f^*(1,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
@@
 This happens sooner for the mutant than the wild-type, due to the stronger depressing transitions.
-This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the mutant.
+This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the mutant, even outside the desired range of values for $\Delta f$.
+As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}, even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.
 
@@
  \end{center}
-  \caption{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$ and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$.}\label{fig:multistate_star}
+  \caption{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
 \end{figure}

### 2013-02-12 Notes/Figures/MakeFigs.m

@@ -1,3 +1,4 @@
+MakeVORfigs('multistate_med',0.6,0.8,-0.17,10,30,'print')
 MakeVORfigs('multistate_strong',0.6,0.8,-0.4,20,20,'print')
 MakeVORfigs('multistate_weak',0.6,0.8,-0.1,50,50,'print')
 MakeVORfigs('cascade_short',0.25,0.33,-0.3,50,50,'print')

### 2013-02-12 Notes/VORmodel.tex
Talk about crossovers

@@
 %
-where it's worth remembering that $\Delta f<0$.
+where, once more, we should note that $\Delta f<0$.
 With pre-training:
@@
 But it is easier to just substitute $M=2$ into the formulae in \autoref{sec:multistate}.
-In this case, the initial rate of change encapsulates the whole solution, as there is only a single exponential decay.
+In this case, the initial rate of change and the total change encapsulate the whole solution, as there is only a single exponential decay.
 
@@
 %
-where, again, it's worth remembering that $\Delta f<0$.
+where, once again, it's worth remembering that $\Delta f<0$.
 With pre-training:
@@
 
+The only thing we have to worry about is when the wild-type catches up with the mutant.
+As the total change will be larger for the wild-type than mutant, it must overtake eventually.
+If this happens early enough, the initial period, where the mutant learns faster than wild-type, would not be seen in the experiment.
+However, the timescale for this crossover would be similar to the timescale of the exponentials.
+In the data, it looks like that timescale is longer than the gaps between successive measurements, so we needn't worry about this.
+
 \subsubsection{Pooled resource model}\label{sec:pooled}

### 2013-02-12 Notes/VORmodel.tex
Discussion of OG multistate

@@
 \tableofcontents
+\listoffigures
 
@@
  \end{center}
-  \caption{Transition probabilities for different models.
+  \caption[Transition probabilities for different models]{Transition probabilities for different models.
   (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
@@
     Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
+    OG Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.3  & 20 & 20  &\label{tr:multistate_lin} \\
     Two-state     & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
@@
 We will model the individual synapses with the two-state model.
-Let $i$ be the number of them that are potentiated.
+Let $i=0\ldots P$ be the number of them that are potentiated.
 We will model the effect of resource depletion linearly with the potentiation/depression probabilities for the individual synapses:
@@
   \begin{aligned}
-    q\pot &= \frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}, \qquad
-    q\dep &= \frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}.
+    q\pot &= \frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}, \quad& i &= 0 \ldots P-1,\\
+    q\dep &= \frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}, & i &= 1 \ldots P.
   \end{aligned}
@@
 We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
-This is proportional to the net-flux between the $\w=+1$ states and the $\w=-1$ states, measuring using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
+This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measuring using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
+This should be multiplied by the difference in synaptic weights 
+But, as this quantity is constant within a model and doesn't change when comparing wild-type and mutant or with/without pre-training, it can be ignored.
 
@@
   (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant.
-  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre}\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre}\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training.
-} \label{fig:#1}
+  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant,
+  with states of weak synaptic strength to the left and strong states to the right.
+  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre}\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre}\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
+  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#1}
 \end{figure}}
@@
 The results of simulations of the cascade model can be seen in \autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
+It is difficult to study this model analytically.
+However, the numerical results can helps us understand it qualitatively.
 
@@
 So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS}.
-On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.
+On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot due to the factor of $(1+2\Delta f)^{M/2-1}$, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.
 
@@
           {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M}   \,
-          \beta^{M/2-1}(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1} \\
+          \brk{\beta (1-2\Delta f) (1+2\Delta f)}^{M/2-1} \\
        &= -4\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
@@
           {(1+2\overline{\Delta f})^M - \beta^M(1-2\overline{\Delta f})^M}   \,
-          \beta^{M/2-1}(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} \\
+          \brk{\beta (1-2\overline{\Delta f}) (1+2\overline{\Delta f})}^{M/2-1} \\
        &= -2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
@@
 We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the mutant, the important thing is to have \emph{strong pre-training}, not strong training.
-Thus, removing the symmetry between gain-increase and gain-decrease will not help with the unrealistic feature of this model -- the fact that pre-training proceeds much faster in the mutant than wild-type.
+Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the mutant than wild-type.
 
@@
  \end{center}
-  \caption{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
+  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
 \end{figure}
@@
 \subsubsection{OG multistate model}\label{sec:multistate_lin}
+
+\resultsfig{multistate_lin}{Simulation results for multistate model with linear weights}%\label{fig:multistate_lin}
 
@@
 %
+The numerical results can be seen in \autoref{fig:multistate_lin}, but we can get some analytic insight into this model as well.
+
 The equilibrium distribution, \eqref{eq:mutltieq}, still applies.
@@
 %
+This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
+This means that the mutant will initially learn faster than the wild-type.
+However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}
+
 With pre-training:
@@
 %
+The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the lowest end of the interval $\Delta f \in \brk{-\half,0}$, where it takes the value
+%
+\begin{equation}\label{eq:multiLinprenopre}
+  \frac{\Phi_{\text{with pre}}}{\Phi_{\text{without pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
+   \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,
+\end{equation}
+%
+which is always greater than 1.
+Therefore, pre-training will always enhance learning, for both mutant and wild-type.
 
@@
 
-\resultsfig{pooled_plenty}{Simulation results for the pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$)}
+\resultsfig{pooled_plenty}{Simulation results for the pooled resource model with light depletion ($q\lmin/q\lmax=0.75$)}
 
-\resultsfig{pooled_scarce}{Simulation results for the pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$)}
+\resultsfig{pooled_scarce}{Simulation results for the pooled resource model with heavy depletion ($q\lmin/q\lmax=0.25$)}
 
 The results of simulations of the pooled resource model can be seen in \autoref{fig:pooled_plenty} and \autoref{fig:pooled_scarce}.
+It is difficult to study this model analytically.
+However, the numerical results can helps us understand it qualitatively.
 

### 2013-02-12 Notes/Figures/MakeFigs.m

@@ -6,3 +6,4 @@ MakeVORfigs('cascade_long',0.25,0.33,-0.3,50,150,'print')
 MakeVORfigs('multistate_weak',0.6,0.8,-0.1,50,50,'print')
 MakeVORfigs('multistate_strong',0.6,0.8,-0.4,20,20,'print')
 MakeVORfigs('binary',0.4,0.8,-0.1,5,5,'print')
+MakeVORfigs('multistate_lin',0.6,0.8,-0.3,20,20,'print','LinWeights',true)

### 2013-02-13 Notes/VORmodel.tex
typos etc

@@
 
-We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equlibrium weights mentioned above.
+We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
 
@@
 However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
-The transition network topology will then have a multistate toplogy (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
+The transition network topology will then have a multistate topology (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
 %
@@
 
-\resultsfig{multistate_med}{Simulation results for the multistate model with intermediate training ($\Delta f=-0.17$)}
+\resultsfig{multistate_med}{Simulation results for the multistate model with moderate training ($\Delta f=-0.17$)}
 

### 2013-02-14 Notes/VORmodel.tex
redesigned figure captions

@@
 
-\newcommand{\resultsfig}[2]{\begin{figure}
+\newcommand{\resultsfig}[3]{\begin{figure}
  \begin{center}
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{#1_learn.eps}}\label{fig:#1_learn}
-  \item\aligntop{\includegraphics[width=7cm]{#1_learnS.eps}}\label{fig:#1_learnS}
-  \item\label{fig:#1_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_WT.eps}}\label{fig:#1_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_KO.eps}}\label{fig:#1_eq_KO}
+  \item\aligntop{\includegraphics[width=7cm]{#3_learn.eps}}\label{fig:#3_learn}
+  \item\aligntop{\includegraphics[width=7cm]{#3_learnS.eps}}\label{fig:#3_learnS}
+  \item\label{fig:#3_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_WT.eps}}\label{fig:#3_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_KO.eps}}\label{fig:#3_eq_KO}
                   \end{myenumi}
-  \item\label{fig:#1_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_nopre.eps}}\label{fig:#1_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_pre.eps}}\label{fig:#1_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_nopre.eps}}\label{fig:#1_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_pre.eps}}\label{fig:#1_pr_KO_pre}
+  \item\label{fig:#3_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_nopre.eps}}\label{fig:#3_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_pre.eps}}\label{fig:#3_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_nopre.eps}}\label{fig:#3_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_pre.eps}}\label{fig:#3_pr_KO_pre}
                   \end{myenumi}
@@
  \end{center}
-  \caption[#2]{#2.
-  Other parameters can be found in \autoref{tr:#1} of \autoref{tab:params}.
+  \caption[#1]{#1 #2.
+  Other parameters can be found in \autoref{tr:#3} of \autoref{tab:params}.
   Time is measured in units of $1/r$, the mean time between plasticity events.
-  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant,
+  (\ref{fig:#3_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
+  (\ref{fig:#3_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:#3_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#3_eq_WT}) wild-type and (\ref{fig:#3_eq_KO}) \KO\ mutant,
   with states of weak synaptic strength to the left and strong states to the right.
-  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre}\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre}\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
-  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#1}
+  (\ref{fig:#3_pr}) Evolution of probability distributions for (\ref{fig:#3_pr_WT_nopre}\ref{fig:#3_pr_WT_pre}) wild-type and  (\ref{fig:#3_pr_KO_nopre}\ref{fig:#3_pr_KO_pre}) \KO\ mutant without (\ref{fig:#3_pr_WT_nopre},\ref{fig:#3_pr_KO_nopre}) and with (\ref{fig:#3_pr_WT_pre},\ref{fig:#3_pr_KO_pre}) pre-training,
+  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#3}
+\end{figure}}
+
+\newcommand{\resultsfign}[2]{\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{#2_learn.eps}}\label{fig:#2_learn}
+  \item\aligntop{\includegraphics[width=7cm]{#2_learnS.eps}}\label{fig:#2_learnS}
+  \item\label{fig:#2_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_WT.eps}}\label{fig:#2_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_KO.eps}}\label{fig:#2_eq_KO}
+                  \end{myenumi}
+  \item\label{fig:#2_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_nopre.eps}}\label{fig:#2_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_pre.eps}}\label{fig:#2_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_nopre.eps}}\label{fig:#2_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_pre.eps}}\label{fig:#2_pr_KO_pre}
+                  \end{myenumi}
+ \end{myenuma}
+ \end{center}
+  \caption[#1]{#1.
+  Parameters can be found in \autoref{tr:#2} of \autoref{tab:params}.
+  Time is measured in units of $1/r$, the mean time between plasticity events.
+  (\ref{fig:#2_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
+  (\ref{fig:#2_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:#2_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#2_eq_WT}) wild-type and (\ref{fig:#2_eq_KO}) \KO\ mutant,
+  with states of weak synaptic strength to the left and strong states to the right.
+  (\ref{fig:#2_pr}) Evolution of probability distributions for (\ref{fig:#2_pr_WT_nopre}\ref{fig:#2_pr_WT_pre}) wild-type and  (\ref{fig:#2_pr_KO_nopre}\ref{fig:#2_pr_KO_pre}) \KO\ mutant without (\ref{fig:#2_pr_WT_nopre},\ref{fig:#2_pr_KO_nopre}) and with (\ref{fig:#2_pr_WT_pre},\ref{fig:#2_pr_KO_pre}) pre-training,
+  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#2}
 \end{figure}}
@@
 
-\resultsfig{cascade_short}{Simulation results for the cascade model with short pre-training ($rt_\text{pre}=50$)}
+\resultsfig{Simulation results for the cascade model with short pre-training}{($rt_\text{pre}=50$)}{cascade_short}
 
-\resultsfig{cascade_long}{Simulation results for the cascade model with long pre-training ($rt_\text{pre}=150$)}
+\resultsfig{Simulation results for the cascade model with long pre-training}{($rt_\text{pre}=150$)}{cascade_long}
 
@@
 
-\resultsfig{multistate_weak}{Simulation results for the multistate model with weak training ($\Delta f=-0.1$)}
+\resultsfig{Simulation results for the multistate model with weak training}{($\Delta f=-0.1$)}{multistate_weak}
 
-\resultsfig{multistate_med}{Simulation results for the multistate model with moderate training ($\Delta f=-0.17$)}
+\resultsfig{Simulation results for the multistate model with moderate training}{($\Delta f=-0.17$)}{multistate_med}
 
-\resultsfig{multistate_strong}{Simulation results for the multistate model with strong training ($\Delta f=-0.4$)}
+\resultsfig{Simulation results for the multistate model with strong training}{($\Delta f=-0.4$)}{multistate_strong}
 
@@
 
-\resultsfig{multistate_lin}{Simulation results for multistate model with linear weights}%\label{fig:multistate_lin}
+\resultsfign{Simulation results for multistate model with linear weights}{multistate_lin}
 
@@
 
-\resultsfig{binary}{Simulation results for two-state model}
+\resultsfign{Simulation results for two-state model}{binary}
 
@@
 
-\resultsfig{pooled_plenty}{Simulation results for the pooled resource model with light depletion ($q\lmin/q\lmax=0.75$)}
+\resultsfig{Simulation results for the pooled resource model with light depletion}{($q\lmin/q\lmax=0.75$)}{pooled_plenty}
 
-\resultsfig{pooled_scarce}{Simulation results for the pooled resource model with heavy depletion ($q\lmin/q\lmax=0.25$)}
+\resultsfig{Simulation results for the pooled resource model with heavy depletion}{($q\lmin/q\lmax=0.25$)}{pooled_scarce}
 

### 2013-02-15 Notes/VORmodel.tex
renamed linear and step multistate models

@@
 We will see if we can capture the essential features of VOR gain-increase learning in wild type and MHC \KO\ mutants and wild-type, with and without gain-decrease pre-training.
-We will try to do this using models of complex synapses, such as the cascade model \cite{Fusi2005cascade} or the multistate model \cite{amit1994learning,Fusi2007multistate}.
+We will try to do this using models of complex synapses, such as the cascade model \cite{Fusi2005cascade} or the multistate model \cite{amit1994learning,Fusi2007multistate}.\footnote{The original definition of the multistate model had a synaptic weight that varied linearly along the chain.
+We will refer to this as the linear multistate model.
+We will also look at the step-multistate model, for which the synaptic weight to take two values.}
 We will also construct a new model in \autoref{sec:pooledmodel} that incorporates the idea that biasing the synapses in one direction could deplete some resource, thereby impairing further plasticity in the same direction in nearby synapses.
@@
 
-On the other hand, the cascade, multistate and pooled resource models all have parameter ranges where the mutant has impaired learning without pre-training (see \autoref{fig:cascade_short}, \autoref{fig:cascade_long}, \autoref{fig:multistate_weak}, \autoref{fig:multistate_strong} and \autoref{fig:pooled_scarce}).
+A similar thing happens for the linear multistate model (see \autoref{fig:multistate_lin}), but the wild-type catches up and overtakes the mutant very quickly.
+
+On the other hand, the cascade, step-multistate and pooled resource models all have parameter ranges where the mutant has impaired learning without pre-training (see \autoref{fig:cascade_short}, \autoref{fig:cascade_long}, \autoref{fig:multistate_weak}, \autoref{fig:multistate_strong} and \autoref{fig:pooled_scarce}).
 We have found parameter ranges where the reverse happens for the multistate (see discussion below \eqref{eq:multiKOnopre}) and the pooled resource model (see \autoref{fig:pooled_plenty}).
@@
 In the cascade model, this required a much longer duration for pe-training (see \autoref{fig:cascade_long}).
-In the multistate model, this required stronger training, which resulted in gain-decrease learning being faster for the mutant then wild-type (see \autoref{fig:multistate_strong}), unlike what is seen experimentally.
+In the step-multistate model, this required stronger training, which resulted in gain-decrease learning being faster for the mutant then wild-type (see \autoref{fig:multistate_strong}), unlike what is seen experimentally.
 It is possible that these features could be improved by lessening the symmetry between gain-increase and decrease or potentiation and depression, or by allowing the rate of plasticity events to vary.
-For the pooled resource model, we haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.
+For the linear multistate and pooled resource models, we haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.
 
@@
 Second, impaired learning with pre-training in the wild-type requires a form of metaplasticity where repeated potentiation makes subsequent depression less likely.
-This is a feature of the cascade model and our version of the multistate model, but not the pooled resource model.
+This is a feature of the cascade model and our version of the multistate model, but not the pooled resource model or linear multistate.
 
@@
 
-We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the multistate model\footnote{The version of the multistate model that we'll study is slightly different than the original definition.
-We will only allow the synaptic weight to take two values, whereas it originally varied linearly across the states.
-} (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary_model}).
+We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the multistate model (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary_model}).
 We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.
@@
     Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:cascade_long} \\
-    Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
-    Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
-    Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
-    OG Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.3  & 20 & 20  &\label{tr:multistate_lin} \\
+    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
+    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
+    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
+    Linear Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.3  & 20 & 20  &\label{tr:multistate_lin} \\
     Two-state     & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
@@
 
-\subsubsection{Multistate model}\label{sec:multistate}
+\subsubsection{Step-multistate model}\label{sec:multistate}
 
-\resultsfig{Simulation results for the multistate model with weak training}{($\Delta f=-0.1$)}{multistate_weak}
+\resultsfig{Simulation results for the step-multistate model with weak training}{($\Delta f=-0.1$)}{multistate_weak}
 
-\resultsfig{Simulation results for the multistate model with moderate training}{($\Delta f=-0.17$)}{multistate_med}
+\resultsfig{Simulation results for the step-multistate model with moderate training}{($\Delta f=-0.17$)}{multistate_med}
 
-\resultsfig{Simulation results for the multistate model with strong training}{($\Delta f=-0.4$)}{multistate_strong}
+\resultsfig{Simulation results for the step-multistate model with strong training}{($\Delta f=-0.4$)}{multistate_strong}
 
-The results of simulations of the multistate model can be seen in \autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
+The results of simulations of the step-multistate model can be seen in \autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
 However, we can get some insight into this model analytically.
@@
 
-\subsubsection{OG multistate model}\label{sec:multistate_lin}
+\subsubsection{Linear multistate model}\label{sec:multistate_lin}
 

### 2013-02-15 Notes/VORmodel.tex
more blather

@@
 
-QUESTION: Should $f\pot\wt=f\pot\ko$?
+%QUESTION: Should $f\pot\wt=f\pot\ko$?
+%
+%If they are equal, this would change the mean synaptic weight in equilibrium.
+%This seems like it could affect the ability of the network to perform its function, whereas, in reality, the wild-type and mutant have the same baseline performance.
+%In any case, if the synaptic weights are different, the electrical activity will be different, and there will be no reason to expect the same statistics for potentiation or depression.
+%In setting $f\pot\wt=f\pot\ko$, we are assuming that the rest of the brain compensates for the difference in synaptic weights somehow.
+%
+%On the other hand, one could imagine adjusting $f\pot\wt$ and $f\pot\ko$ so that $\eq\wt\w = \eq\ko\w$.
+%But, now that the synaptic weights are the same, the electrical activity will be the same, and there will be no reason to expect different statistics for potentiation or depression.
+%If we did this for the multistate or two-state models, we'd remove all differences between the wild-type and \KO\ mutant, so we won't.
 
-If they are equal, this would change the mean synaptic weight in equilibrium.
-This seems like it could affect the ability of the network to perform its function, whereas, in reality, the wild-type and mutant have the same baseline performance.
-In any case, if the synaptic weights are different, the electrical activity will be different, and there will be no reason to expect the same statistics for potentiation or depression.
-In setting $f\pot\wt=f\pot\ko$, we are assuming that the rest of the brain compensates for the difference in synaptic weights somehow.
-
-On the other hand, one could imagine adjusting $f\pot\wt$ and $f\pot\ko$ so that $\eq\wt\w = \eq\ko\w$.
-But, now that the synaptic weights are the same, the electrical activity will be the same, and there will be no reason to expect different statistics for potentiation or depression.
-If we did this for the multistate or two-state models, we'd remove all differences between the wild-type and \KO\ mutant, so we won't.
+We will assume $f\pot\wt=f\pot\ko$.
+This is because the mutation is well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
+Therefore the rates of potentiation and depression should not change very much either.
+This implies that the mean synaptic weight in equilibrium differs between the mutant and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
+As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
+We will model this compensation as a simple linear offset, as could be produced by another population of neurons/synapses whose effect cancels with these neurons/synapses.
 
@@
 The features of the actual experiments that we'd like to capture are:
-\begin{itemize}
+\begin{enumerate}
   \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the mutant.
-  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.
+  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
   \item After pre-training, gain-increase learning is significantly faster in the mutant than in the wild-type.
@@
   \item For the mutant, gain-increase learning is significantly faster with pre-training than without it.
-\end{itemize}
+\end{enumerate}
+These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
+We will not worry too much about \autoref{it:gaindown}, as gain-decrease learning uses a different mechanism to gain-increase.
+We are only modelling the effect of gain-decrease training on \emph{these} synapses.
+
+We will not worry too much about the curvature of the learning curves, as this can be changes by the nonlinearity.
+However, both the wild-type and mutant should be affected in the same way, so the difference in the curvature of the mutant and wild-type curves after pre-training would be nice to capture.
 

### 2013-02-22 Notes/MHCmodelFig.tex
fig for MHC paper

\documentclass[12pt,landscape,letterpaper]{article}
%Preamble:
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
\usepackage[margin=1in]{geometry}
\usepackage{array}
\setlength{\extrarowheight}{0.5cm}

\begin{document}
\pagestyle{empty}

\begin{flushleft}
\begin{tabular}{>{\sffamily\bfseries}rc!{\hspace{0.5cm}}c!{\hspace{0.5cm}}c}
  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
  a & \aligntop{\includegraphics[height=2cm]{binary.svg}} & \aligntop{\includegraphics[height=2.4cm]{pooled.svg}} & \aligntop{\includegraphics[height=2cm]{multistate.svg}} \\
  b & \aligntop{\includegraphics[height=6cm]{binary_learnS.eps}} & \aligntop{\includegraphics[height=6cm]{pooled_scarce_learnS.eps}} & \aligntop{\includegraphics[height=6cm]{multistate_med_learnS.eps}} \\
\end{tabular}
\end{flushleft}

\end{document}

### 2013-02-25 Notes/MHCsupp.tex
MHC supplemental

--- 2013-02-15 Notes/VORmodel.tex
@@
-% -*- TeX -*- -*- UK -*-
-% ----------------------------------------------------------------
-% arXiv Paper ************************************************
-%
-% Subhaneil Lahiri's template
-%
-% Before submitting:
-%    Comment out hyperref
-%    Comment out showkeys
-%    Replace \input{mydefs.tex} with its contents
-%       or include mydefs.tex in zip/tar file
-%    Replace \input{newsymb.tex} with its contents
-%       or include newsymb.tex in zip/tar file
-%    Put this file, the .bbl file, any picture or
-%       other additional files and natbib.sty
-%       file in a zip/tar file
-%
-% **** -----------------------------------------------------------
 \documentclass[12pt]{article}
@@
 %
-% >> Only for drafts! <<
-\usepackage[notref,notcite]{showkeys}
-% ----------------------------------------------------------------
-%\numberwithin{equation}{section}
-%\renewcommand{\baselinestretch}{1.5}
 % ----------------------------------------------------------------
@@
 %matrices
-\newcommand{\inv}{^{-1}}
-\newcommand{\dg}{^\mathrm{dg}}
-\newcommand{\trans}{^\mathrm{T}}
 \newcommand{\I}{\mathbf{I}}
 %vec of ones
-\newcommand{\onev}{\mathbf{e}}
-%mat of ones
-\newcommand{\onem}{\mathbf{E}}
-%Markov matrix
-\newcommand{\MM}{\mathbf{Q}}
 %equilibrium distribution
@@
 \newcommand{\eq}{\pr^\infty}
-%first passage times
-\newcommand{\fpt}{\mathbf{T}}
-%off-diag first passage times
-\newcommand{\fptb}{\overline{\fpt}}
-%fundamental matrix
-\newcommand{\fund}{\mathbf{Z}}
 %other symbols
-\newcommand{\Pb}{\mathbf{P}}
-\newcommand{\D}{\mathbf{D}}
-\newcommand{\pib}{\boldsymbol{\pi}}
-\newcommand{\Lb}{\boldsymbol{\Lambda}}
 \newcommand{\w}{\mathbf{w}}
@@
 \newcommand{\F}{\boldsymbol{\Phi}}
-\DeclareMathOperator{\SNR}{SNR}
-\newcommand{\CS}{\mathcal{S}}
-\newcommand{\CA}{\mathcal{A}}
-\newcommand{\CB}{\mathcal{B}}
-\newcommand{\comp}{^\mathrm{c}}
 \newcommand{\pot}{^{\text{pot}}}
@@
 % Title info:
-\title{Models of VOR learning in MHC knockout mice}
+\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
 %
-% Author List:
-%
-\author{Subhaneil Lahiri
-%
-}
+%% Author List:
+%%
+%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
+%%
+%\small{\emph{$^a$ Address 1}}\\
+%\small{\emph{$^b$ Address 2}}\\
+%\small{\emph{$^c$ Address 3}}\\
+%}
 
@@
 
-\begin{abstract}
-  We see if we can model VOR gain increase and decrease learning in mice with a knockout in MHC as well as wild-type.
-\end{abstract}
+%\begin{abstract}
+%  We see if we can model VOR gain increase and decrease learning in mice with a knockout in MHC as well as wild-type.
+%\end{abstract}
 
@@
 
-\section{Introduction and summary}\label{sec:intro}
-
-We will see if we can capture the essential features of VOR gain-increase learning in wild type and MHC \KO\ mutants and wild-type, with and without gain-decrease pre-training.
-We will try to do this using models of complex synapses, such as the cascade model \cite{Fusi2005cascade} or the multistate model \cite{amit1994learning,Fusi2007multistate}.\footnote{The original definition of the multistate model had a synaptic weight that varied linearly along the chain.
-We will refer to this as the linear multistate model.
-We will also look at the step-multistate model, for which the synaptic weight to take two values.}
-We will also construct a new model in \autoref{sec:pooledmodel} that incorporates the idea that biasing the synapses in one direction could deplete some resource, thereby impairing further plasticity in the same direction in nearby synapses.
-
-These models all assume independent, statistically identical synapses.
-This has the benefit of freeing us from modelling the neurons and their networks, allowing us to focus on the synapses themselves.
-However, we will lose certain aspects of real synapses.
-For example, during gain-increase learning, certain synapses will be potentiated more often and others will be depressed more often.
-In these models, however, synaptic identity is not kept track of, so training can only be described by changing the overall levels of potentiation and depression.
-
-As the MHC \KO\ mutation lowers the threshold for LTD, we will model these mice by increasing the probabilities for internal synaptic transitions associated with depression, whilst leaving potentiation unchanged.
-
-We will be parsimonious in our modelling.
-We will assume that the wild-type has a symmetry between potentiation and depression and we will treat gain-increase and decrease training in a symmetric way.
-We will also assume that the relative rates of potentiation/depression in the mutant and wild-type are the same and that the overall rate of plasticity events is the same in all situations.
-None of these assumptions are necessary or justified experimentally (although \cite{Fusi2007multistate} notes that imbalance between potentiation and depression can drastically impair memory).
-They do reduce the number of adjustable parameters we have, simplifying searches for good models.
-Presumably, relaxing these assumptions and enlarging the parameter space would allow for a better fit to data, but the number of parameters would start to look large compared to the number of features we're trying to match.
-
-If we imagine something like an LN model for the Purkinje cell, the output would depend on some linear combination of the synaptic weights (weighted by the activities of the corresponding parallel fibres).
-If we're not keeping track of synaptic identity, the most natural linear combination to use would be an equal sum of them all.
-Thus we will quantify learning by the change in the mean synaptic weight.
-
-The output of the Purkinje cell, and the behavioural output of the animal, will depend on this quantity in some complicated nonlinear way.
-Modelling this is far beyond the scope of this work.
-Instead we will look at more qualitative features, asking whether learning is slower or faster in some situation compared to another.
-Provided that the nonlinearity is monotonic, these types of questions will not be affected by it.
-
-Within the parameter range explored, we will see that synaptic complexity is essential.
-The simple two-state model always shows enhanced learning in the mutant (see \autoref{fig:binary}), in contradiction with experiment.
-It seems that the enhanced depressing transition more than compensates for saturation effects.
-
-A similar thing happens for the linear multistate model (see \autoref{fig:multistate_lin}), but the wild-type catches up and overtakes the mutant very quickly.
-
-On the other hand, the cascade, step-multistate and pooled resource models all have parameter ranges where the mutant has impaired learning without pre-training (see \autoref{fig:cascade_short}, \autoref{fig:cascade_long}, \autoref{fig:multistate_weak}, \autoref{fig:multistate_strong} and \autoref{fig:pooled_scarce}).
-We have found parameter ranges where the reverse happens for the multistate (see discussion below \eqref{eq:multiKOnopre}) and the pooled resource model (see \autoref{fig:pooled_plenty}).
-These rely on properties of the mutation, so it would be difficult to explore with further experiments.
-
-All of these models show pre-training enhancing learning for the mutant, as expected.
-It is harder to find regimes where pre-training impairs learning for the wild-type.
-In the cascade model, this required a much longer duration for pe-training (see \autoref{fig:cascade_long}).
-In the step-multistate model, this required stronger training, which resulted in gain-decrease learning being faster for the mutant then wild-type (see \autoref{fig:multistate_strong}), unlike what is seen experimentally.
-It is possible that these features could be improved by lessening the symmetry between gain-increase and decrease or potentiation and depression, or by allowing the rate of plasticity events to vary.
-For the linear multistate and pooled resource models, we haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.
-
-There seems to be two take home messages.
-First, impaired learning without pre-training in the mutant requires synaptic complexity, as it is not seen in the simple two-state model.
-Second, impaired learning with pre-training in the wild-type requires a form of metaplasticity where repeated potentiation makes subsequent depression less likely.
-This is a feature of the cascade model and our version of the multistate model, but not the pooled resource model or linear multistate.
-
-\section{The setup}\label{sec:setup}
+\section{Mathematical formalism}\label{sec:setup}
 
@@
   \item The synaptic weights of the internal states are given by the column vector $\w$.
-This can only take two values that we can call $\pm1$, except for the pooled resource model that we will discuss in \autoref{sec:pooledmodel} and \autoref{sec:pooled}.
+This can only take values in a finite range that we can shift to $\pm1$.
+Most of the models will only use the two extreme values.
 \end{itemize}
@@
 
-With only two possible synaptic weights, the distribution of synaptic weights is completely described by the mean, $\pr(t)\w$.
-
-\subsection{Model of VOR learning experiment}\label{sec:learning}
-
-Training the animal will not change the internal dynamics of a synapse under potentiation or depression.
-It will change the environment, will lead to a change in how often potentiation and depression occur.
-It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
-We will model this by changing $f\potdep$, leaving $r$ and $\M\potdep$ unchanged.
-
-The untrained case will be described by $f\pot=f\pot\norm$.
-Gain-increase training  will be described by $f\pot=f\pot\inc<f\pot\norm$, and
-gain-decrease training  will be described by $f\pot=f\pot\dec>f\pot\norm$.
-Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depend on $f\pot$, which we will indicate with subscripts.
-
-Before training, the synaptic distribution will in equilibrium with $f\pot\norm$.
-During gain-increase training, it will evolve according to \eqref{eq:evolve} with $f\pot\inc$:
-%
-\begin{equation}\label{eq:nopre}
-  \pr(t) = \eq\norm \exp\prn{rt\frg\inc}.
-\end{equation}
-%
-On the other hand, if the gain-increase training follows gain-decrease pre-training for some time period, $\tpre$:
-%
-\begin{equation}\label{eq:withpre}
-  \pr(t) = \eq\norm \exp\prn{r\tpre\frg\dec} \exp\prn{r(t-\tpre)\frg\inc}.
-\end{equation}
-%
-
-We will describe the effect of training by the decrease in mean synaptic weight:
-%
-\begin{equation}\label{eq:learning}
-  L(t) = \prn{\pr(0)-\pr(t)}\w.
-\end{equation}
-%
-The behavioural output (VOR gain) will be some non-linear function of the synaptic weights, so the best we can hope for is to reproduce qualitative features of the experiment.
+For models with only two possible synaptic weights, the distribution of synaptic weights is completely described by the mean, $\pr(t)\w$.
 
 The MHC mutant has a lower threshold for depression.
-We can model this by changing $\W\dep\wt$ to $\W\dep\ko$, which should have larger matrix elements.
+We can model this by changing $\M\dep\wt$ to $\M\dep\ko$, which should have larger off-diagonal matrix elements.
 
-%QUESTION: Should $f\pot\wt=f\pot\ko$?
-%
-%If they are equal, this would change the mean synaptic weight in equilibrium.
-%This seems like it could affect the ability of the network to perform its function, whereas, in reality, the wild-type and mutant have the same baseline performance.
-%In any case, if the synaptic weights are different, the electrical activity will be different, and there will be no reason to expect the same statistics for potentiation or depression.
-%In setting $f\pot\wt=f\pot\ko$, we are assuming that the rest of the brain compensates for the difference in synaptic weights somehow.
-%
-%On the other hand, one could imagine adjusting $f\pot\wt$ and $f\pot\ko$ so that $\eq\wt\w = \eq\ko\w$.
-%But, now that the synaptic weights are the same, the electrical activity will be the same, and there will be no reason to expect different statistics for potentiation or depression.
-%If we did this for the multistate or two-state models, we'd remove all differences between the wild-type and \KO\ mutant, so we won't.
+We will look at several different models,
+the multistate model\footnote{The original definition of the multistate model had a synaptic weight that varied linearly along the chain.
+We will refer to this as the linear multistate model.
+We will also look at the step-multistate model, for which the synaptic weight takes only two values.} (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}),
+and the two-state model (which can be thought of as a special case of the multistate model, see \autoref{fig:models}\ref{fig:binary_model}),
+and the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}).
+We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.
 
-We will assume $f\pot\wt=f\pot\ko$.
-This is because the mutation is well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
-Therefore the rates of potentiation and depression should not change very much either.
-This implies that the mean synaptic weight in equilibrium differs between the mutant and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
-As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
-We will model this compensation as a simple linear offset, as could be produced by another population of neurons/synapses whose effect cancels with these neurons/synapses.
-
-We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
-
-\section{Simulations}\label{sec:sims}
-
-\subsection{Models and parameters}
-
-We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the multistate model (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary_model}).
-We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.
+For the multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
+We will use a larger value for $q$ for depression in the \KO\ mutant.
 
@@
 
-For the multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
-We will use a larger value for $q$ for depression in the \KO\ mutant.
-
-In each case, we set $f\pot\norm=\half$, $f\pot\inc=f\pot\norm+\Delta f$ and $f\pot\dec=f\pot\norm-\Delta f$.
-We use the same values for wild-type and mutant and treat gain-increase and decrease symmetrically.
 The values of all these parameters are listed in \autoref{tab:params}.
@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}\hspace{0.5cm}
   \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}
+  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\hspace{0.5cm}
+  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}
  \end{myenuma}
@@
   \caption[Transition probabilities for different models]{Transition probabilities for different models.
-  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
+  Potentiation induces transitions indicated by solid red arrows, depression indicated by dashed blue arrows.
+  States of strong/weak synaptic weight indicated by red/blue circles.
+  States of strong/weak synaptic weight
   (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
-  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.} \label{fig:models}
+  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.
+  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}) and synaptic weight takes only two values.
+  } \label{fig:models}
 \end{figure}
-
-\begin{table}
- \begin{center}
-  \begin{tabularn}{|l|c|c|c|c|c|c|}
-    \cline{1-7}
-    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
-    \cline{1-7}
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 50  &\label{tr:cascade_short} \\
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:cascade_long} \\
-    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
-    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
-    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
-    Linear Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.3  & 20 & 20  &\label{tr:multistate_lin} \\
-    Two-state     & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
-    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
-    Pooled        & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
-                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
-    \cline{1-7}
-%   \hline
-%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
-%    \cline{1-7}
-%    \label{tr:cascade_short}&
-%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
-%    \label{tr:cascade_long}&
-%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
-%    \label{tr:multistate_weak}&
-%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
-%    \label{tr:multistate_strong}&
-%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
-%    \label{tr:binary}&
-%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
-%    \label{tr:pooled_plenty}&
-%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-%                                          & -0.1 & 70 & 70 \\
-%    \label{tr:pooled_scarce}&
-%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
-%                                          & -0.1 & 70 & 70 \\
-%    \cline{1-7}
-  \end{tabularn}
- \end{center}
-  \caption{Parameters used in simulations.} \label{tab:params}
-\end{table}
 
@@
 We will use larger values for depression in the \KO\ mutant.
+The values of these parameters are listed in \autoref{tab:params}.
 
-\subsection{Results}\label{sec:results}
+\begin{figure}
+ \begin{center}
+  \aligntop{\includegraphics[height=1.7cm]{pooled.svg}}
+ \end{center}
+  \caption[The pooled resource model]{The pooled resource model.
+  Several two-state synapses share resources that are required for potentiation and depression.
+  These resources are depleted as more synapses are potentiated or depressed.
+  This pool of synapses can be modelled as one compound synapse.} \label{fig:pooled_model}
+\end{figure}
+
+\subsection{Model of VOR learning experiment}\label{sec:learning}
+
+Training the animal will not change the internal dynamics of a synapse under potentiation or depression.
+It will change the environment, will lead to a change in how often potentiation and depression occur.
+It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
+We will model this by changing $f\potdep$, leaving $r$ and $\M\potdep$ unchanged.
+
+The untrained case will be described by $f\pot=f\pot\norm$.
+Gain-increase training  will be described by $f\pot=f\pot\inc<f\pot\norm$, and
+gain-decrease training  will be described by $f\pot=f\pot\dec>f\pot\norm$.
+Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depend on $f\pot$, which we will indicate with subscripts.
+
+Before training, the synaptic distribution will in equilibrium with $f\pot\norm$.
+During gain-increase training, it will evolve according to \eqref{eq:evolve} with $f\pot\inc$:
+%
+\begin{equation}\label{eq:nopre}
+  \pr(t) = \eq\norm \exp\prn{rt\frg\inc}.
+\end{equation}
+%
+On the other hand, if the gain-increase training follows gain-decrease pre-training for some time period, $\tpre$:
+%
+\begin{equation}\label{eq:withpre}
+  \pr(t) = \eq\norm \exp\prn{r\tpre\frg\dec} \exp\prn{r(t-\tpre)\frg\inc}.
+\end{equation}
+%
+
+We will describe the effect of training by the decrease in mean synaptic weight:
+%
+\begin{equation}\label{eq:learning}
+  L(t) = \prn{\pr(0)-\pr(t)}\w.
+\end{equation}
+%
+If we imagine something like an LN model for the Purkinje cell, the output would depend on some linear combination of the synaptic weights (weighted by the activities of the corresponding parallel fibres).
+If we're not keeping track of synaptic identity, the most natural linear combination to use would be an equal sum of them all.
+The behavioural output (VOR gain) will be some unknown, non-linear function of the synaptic weights, so the best we can hope for is to reproduce qualitative features of the experiment, such as whether learning is enhanced or impaired by the mutation or pre-training.
+
+We will assume $f\pot\wt=f\pot\ko$.
+This is because the mutation is well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
+Therefore the rates of potentiation and depression should not change very much either.
+This implies that the mean synaptic weight in equilibrium differs between the mutant and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
+As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
+We will model this compensation as a simple linear offset, as could be produced by another population of neurons/synapses whose effect cancels with these neurons/synapses.
+
+For the most part, we set $f\pot\norm=\half$, $f\pot\inc=f\pot\norm+\Delta f$ and $f\pot\dec=f\pot\norm-\Delta f$, with $\Delta f<0$.
+We use the same values for wild-type and mutant and treat gain-increase and decrease symmetrically.
+The values of these parameters are listed in \autoref{tab:params}.
+
+We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
+This ensures that the qualitative questions (enhancement or impairment of learning) will not be affected.
+
+\section{Simulation and analysis of models}\label{sec:results}
 
@@
 But, as this quantity is constant within a model and doesn't change when comparing wild-type and mutant or with/without pre-training, it can be ignored.
+
+The values for all parameters we will use can be found in \autoref{tab:params}.
+
+\begin{table}
+ \begin{center}
+  \begin{tabularn}{|l|c|c|c|c|c|c|}
+    \cline{1-7}
+    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
+    \cline{1-7}
+    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 50  &\label{tr:cascade_short} \\
+    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:cascade_long} \\
+    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
+    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
+    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
+    Linear Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.3  & 20 & 20  &\label{tr:multistate_lin} \\
+    Two-state     & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
+    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
+                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
+    Pooled        & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
+                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
+    \cline{1-7}
+%   \hline
+%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
+%    \cline{1-7}
+%    \label{tr:cascade_short}&
+%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
+%    \label{tr:cascade_long}&
+%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
+%    \label{tr:multistate_weak}&
+%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
+%    \label{tr:multistate_strong}&
+%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
+%    \label{tr:binary}&
+%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
+%    \label{tr:pooled_plenty}&
+%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
+%                                          & -0.1 & 70 & 70 \\
+%    \label{tr:pooled_scarce}&
+%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
+%                                          & -0.1 & 70 & 70 \\
+%    \cline{1-7}
+  \end{tabularn}
+ \end{center}
+  \caption{Parameters used in simulations.} \label{tab:params}
+\end{table}
 
@@
 
-\subsubsection{Cascade model}\label{sec:cascade}
-
-\resultsfig{Simulation results for the cascade model with short pre-training}{($rt_\text{pre}=50$)}{cascade_short}
-
-\resultsfig{Simulation results for the cascade model with long pre-training}{($rt_\text{pre}=150$)}{cascade_long}
-
-The results of simulations of the cascade model can be seen in \autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
-It is difficult to study this model analytically.
-However, the numerical results can helps us understand it qualitatively.
-
-In both of these, we see that the mutant is slower than wild-type without pre-training but faster with it.
-This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
-With pre-training, some of them will now be potentiated (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_pre}), and the enhanced depression can speed up learning.
-
-With shorter pre-training, we see that it speeds up learning in both wild-type and mutant (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the mutant.
-This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning.
-
-We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_learnS} that longer pre-training pushes the synapses further down the cascade, slow in down learning for the wild-type.
-This effect is weaker for the mutant, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
-
-\subsubsection{Step-multistate model}\label{sec:multistate}
+\subsection{Step-multistate model}\label{sec:multistate}
 
@@
 
-\subsubsection{Linear multistate model}\label{sec:multistate_lin}
-
-\resultsfign{Simulation results for multistate model with linear weights}{multistate_lin}
-
-In this section, we will consider the multistate model as originally defined in \cite{amit1994learning}, \ie with linearly varying synaptic weight:
-%
-\begin{equation}\label{eq:multistateLinWeight}
-  \w_i = \frac{2i-M-1}{M-1}.
-\end{equation}
-%
-The numerical results can be seen in \autoref{fig:multistate_lin}, but we can get some analytic insight into this model as well.
-
-The equilibrium distribution, \eqref{eq:mutltieq}, still applies.
-However, now the rate of change of our learning metric, \eqref{eq:learning}, will be proportional to the sum of the net fluxes between adjacent states:
-%
-\begin{equation}\label{eq:multiLinFlux}
-  \begin{aligned}
-    \Phi &= \sum_{i=1}^{M-1} \eq_{i+1} f'{}\dep q\dep - \eq_i f'{}\pot q\pot
-         &= \sum_{i=1}^{M-1} \eq_i (\alpha-\alpha') f'{}\dep q\dep \\
-         &= \frac{1-\alpha^{M-1}}{1-\alpha^M} \, (\alpha-\alpha') f'{}\dep q\dep.
-  \end{aligned}
-\end{equation}
-%
-
-First, consider the wild-type, for which $q\pot=q\dep=q$.
-Without pre-training:
-%
-\begin{equation}\label{eq:multiLinWTnopre}
-  \Phi = -2\Delta fq\,\frac{M-1}{M},
-\end{equation}
-%
-where, once more, we should note that $\Delta f<0$.
-With pre-training:
-%
-\begin{equation}\label{eq:multiLinWTpre}
-\begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
-          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
-       &= -{4\Delta f q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
-\end{aligned}
-\end{equation}
-%
-
-Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
-Without pre-training:
-%
-\begin{equation}\label{eq:multiLinKOnopre}
-  \Phi = -2\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
-\end{equation}
-%
-This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
-This means that the mutant will initially learn faster than the wild-type.
-However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}
-
-With pre-training:
-%
-\begin{equation}\label{eq:multiLinKOpre}
-\begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
-          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M} \\
-       &= -4\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
-\end{aligned}
-\end{equation}
-%
-The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the lowest end of the interval $\Delta f \in \brk{-\half,0}$, where it takes the value
-%
-\begin{equation}\label{eq:multiLinprenopre}
-  \frac{\Phi_{\text{with pre}}}{\Phi_{\text{without pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
-   \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,
-\end{equation}
-%
-which is always greater than 1.
-Therefore, pre-training will always enhance learning, for both mutant and wild-type.
-
-\subsubsection{Two-state model}\label{sec:binary}
+\subsection{Two-state model}\label{sec:binary}
 
@@
 
-\subsubsection{Pooled resource model}\label{sec:pooled}
+\subsection{Linear multistate model}\label{sec:multistate_lin}
+
+\resultsfign{Simulation results for multistate model with linear weights}{multistate_lin}
+
+In this section, we will consider the multistate model as originally defined in \cite{amit1994learning}, \ie with linearly varying synaptic weight:
+%
+\begin{equation}\label{eq:multistateLinWeight}
+  \w_i = \frac{2i-M-1}{M-1}.
+\end{equation}
+%
+The numerical results can be seen in \autoref{fig:multistate_lin}, but we can get some analytic insight into this model as well.
+In essence, this model is like a series of two-state models attached to each other, in contrast to the step-multistate model for which the synaptic weight only changes between one of the pairs of states.
+
+The equilibrium distribution, \eqref{eq:mutltieq}, still applies.
+However, now the rate of change of our learning metric, \eqref{eq:learning}, will be proportional to the sum of the net fluxes between adjacent states:
+%
+\begin{equation}\label{eq:multiLinFlux}
+  \begin{aligned}
+    \Phi &= \sum_{i=1}^{M-1} \eq_{i+1} f'{}\dep q\dep - \eq_i f'{}\pot q\pot
+         &= \sum_{i=1}^{M-1} \eq_i (\alpha-\alpha') f'{}\dep q\dep \\
+         &= \frac{1-\alpha^{M-1}}{1-\alpha^M} \, (\alpha-\alpha') f'{}\dep q\dep.
+  \end{aligned}
+\end{equation}
+%
+
+First, consider the wild-type, for which $q\pot=q\dep=q$.
+Without pre-training:
+%
+\begin{equation}\label{eq:multiLinWTnopre}
+  \Phi = -2\Delta fq\,\frac{M-1}{M},
+\end{equation}
+%
+where, once more, we should note that $\Delta f<0$.
+With pre-training:
+%
+\begin{equation}\label{eq:multiLinWTpre}
+\begin{aligned}
+  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
+          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
+       &= -{4\Delta f q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
+\end{aligned}
+\end{equation}
+%
+
+Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
+Without pre-training:
+%
+\begin{equation}\label{eq:multiLinKOnopre}
+  \Phi = -2\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
+\end{equation}
+%
+This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
+This means that the mutant will initially learn faster than the wild-type.
+However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}
+
+With pre-training:
+%
+\begin{equation}\label{eq:multiLinKOpre}
+\begin{aligned}
+  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
+          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M} \\
+       &= -4\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
+\end{aligned}
+\end{equation}
+%
+The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the lowest end of the interval $\Delta f \in \brk{-\half,0}$, where it takes the value
+%
+\begin{equation}\label{eq:multiLinprenopre}
+  \frac{\Phi_{\text{with pre}}}{\Phi_{\text{without pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
+   \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,
+\end{equation}
+%
+which is always greater than 1.
+Therefore, pre-training will always enhance learning, for both mutant and wild-type.
+
+\subsection{Pooled resource model}\label{sec:pooled}
 
@@
 
-%\subsection*{Acknowledgements}
+\subsection{Cascade model}\label{sec:cascade}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection*{Appendices}
-%\appendix
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\resultsfig{Simulation results for the cascade model with short pre-training}{($rt_\text{pre}=50$)}{cascade_short}
+
+\resultsfig{Simulation results for the cascade model with long pre-training}{($rt_\text{pre}=150$)}{cascade_long}
+
+The results of simulations of the cascade model can be seen in \autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
+It is difficult to study this model analytically.
+However, the numerical results can helps us understand it qualitatively.
+
+In both of these, we see that the mutant is slower than wild-type without pre-training but faster with it.
+This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
+With pre-training, some of them will now be potentiated (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_pre}), and the enhanced depression can speed up learning.
+
+With shorter pre-training, we see that it speeds up learning in both wild-type and mutant (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the mutant.
+This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning.
+
+We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_learnS} that longer pre-training pushes the synapses further down the cascade, slow in down learning for the wild-type.
+This effect is weaker for the mutant, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
 

### 2013-02-26 Notes/MHCsupp.tex
typos

@@
+% -*- TeX -*- -*- UK -*-
+% ----------------------------------------------------------------
 \documentclass[12pt]{article}
@@
 \input{sl_preamble.tex}
+\hypersetup{colorlinks,linkcolor=blue,citecolor=violet}
 \input{sl_graphics_preamble.tex}
@@
 We will also look at the step-multistate model, for which the synaptic weight takes only two values.} (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}),
-and the two-state model (which can be thought of as a special case of the multistate model, see \autoref{fig:models}\ref{fig:binary_model}),
+the two-state model (which can be thought of as a special case of the multistate model, see \autoref{fig:models}\ref{fig:binary_model}),
 and the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}).
@@
 
-The values of all these parameters are listed in \autoref{tab:params}.
+The values used for all these parameters in simulations are listed in \autoref{tab:params}.
 
@@
 However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
-The transition network topology will then have a multistate topology (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
+The transition network will then have a multistate topology (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
 %
@@
   \begin{aligned}
-    \M\pot_{ii+1} &=  \brk{\frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}} \frac{P-i}{P}, \\
-    \M\dep_{ii-1} &=  \brk{\frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}} \frac{i}{P},
+    \M\pot_{ii+1} &=  \brk{\frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}} \frac{P-i}{P} ,  
+      \quad& i &= 0 \ldots P-1,\\
+    \M\dep_{ii-1} &=  \brk{\frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}} \frac{i}{P} ,   
+           & i &= 1 \ldots P,
   \end{aligned}
@@
 Training the animal will not change the internal dynamics of a synapse under potentiation or depression.
-It will change the environment, will lead to a change in how often potentiation and depression occur.
+It will change the environment, which will lead to a change in how often potentiation and depression occur.
 It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
@@
 
-Before training, the synaptic distribution will in equilibrium with $f\pot\norm$.
+Before training, the synaptic distribution will be in the equilibrium distribution corresponding to $f\pot\norm$.
 During gain-increase training, it will evolve according to \eqref{eq:evolve} with $f\pot\inc$:
@@
 We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
-This ensures that the qualitative questions (enhancement or impairment of learning) will not be affected.
+This ensures that the qualitative questions mentioned above (enhancement or impairment of learning) will not be affected.
 
@@
 
-We will not worry too much about the curvature of the learning curves, as this can be changes by the nonlinearity.
+We will not worry too much about the curvature of the learning curves, as this can be changed by the nonlinear relation between synaptic weight and VOR gain.
 However, both the wild-type and mutant should be affected in the same way, so the difference in the curvature of the mutant and wild-type curves after pre-training would be nice to capture.
@@
 This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measuring using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
-This should be multiplied by the difference in synaptic weights
+This should be multiplied by the difference in synaptic weights.
 But, as this quantity is constant within a model and doesn't change when comparing wild-type and mutant or with/without pre-training, it can be ignored.
@@
 Intuitively, the flux depends on the slope of the distribution at the centre of the chain (with an offset due to the difference between $f\pot$ and $f\dep$).
-Pre-training has two effects: it produces a slope in the right direction (see \autoref{fig:multistate_weak}\ref{fig:multistate_weak_eq}), but it also reduces distribution at the centre (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}).
+Pre-training has two effects: it produces a slope in the right direction (see \autoref{fig:multistate_weak}\ref{fig:multistate_weak_eq}), but it also reduces the distribution at the centre (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}).
 For small $\Delta f$, the first effect is stronger and learning speeds up.
@@
 In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f < \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the mutant learns slower than wild-type without pre-training, and that pre-training speeds up learning in the mutant but slows it down in the wild-type.
-However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}), which is \emph{not} seen in the experiment.
+However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}), which is \emph{not} seen in the experiment.
 

### 2013-02-26 Notes/MHCsupp.tex
improved references to figs

@@
 %
-So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS}.
+So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS} (black curves).
 On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot due to the factor of $(1+2\Delta f)^{M/2-1}$, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.
@@
 Intuitively, the flux depends on the slope of the distribution at the centre of the chain (with an offset due to the difference between $f\pot$ and $f\dep$).
-Pre-training has two effects: it produces a slope in the right direction (see \autoref{fig:multistate_weak}\ref{fig:multistate_weak_eq}), but it also reduces the distribution at the centre (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}).
+Pre-training has two effects: it produces a slope in the right direction (see \autoref{fig:multistate_weak}\ref{fig:multistate_weak_eq}, red trace), but it also reduces the distribution at the centre (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace).
 For small $\Delta f$, the first effect is stronger and learning speeds up.
@@
 There are two effects here as well.
-Smaller $\beta$ will increase the probability of crossing the centre of the chain, speeding up learning, but it will also concentrate probability at the ends of the chain, depleting the centre and slowing down learning.
+Smaller $\beta$ will increase the probability of crossing the centre of the chain, speeding up learning, but it will also concentrate probability at the ends of the chain, depleting the centre and slowing down learning (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}\ref{fig:multistate_strong_eq_KO}, blue trace).
 The first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ and will be more significant for smaller $\beta$ or in a longer chain.
@@
 Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f < \Delta f^*(1,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
-We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}) drifting to the centre.
+We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
 This happens sooner for the mutant than the wild-type, due to the stronger depressing transitions.
 This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the mutant, even outside the desired range of values for $\Delta f$.
-As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}, even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.
+As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} (black curves), even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
+  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
+ \end{myenuma}
+ \end{center}
+  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
+\end{figure}
 
 In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f < \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the mutant learns slower than wild-type without pre-training, and that pre-training speeds up learning in the mutant but slows it down in the wild-type.
-However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}), which is \emph{not} seen in the experiment.
+However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.
 
@@
 
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
- \end{myenuma}
- \end{center}
-  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
-\end{figure}
+Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the mutant (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
+This is due to the mass of probability at the right end of the chain (see \autoref{fig:multistate_med}\ref{fig:multistate_med_eq} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) that is unable to contribute to change in synaptic weight until it has drifted to the centre.
+This will last for a shorter time in the mutant, due to the enhanced transitions.
+This will not be a feature of any of the other models that we study here.
+Now, the curvature can be changed by the nonlinear relation between synaptic weight and VOR gain, but both curves would be affected in the same way.
+The difference in these two curvatures is therefore a good feature of this model.
 
@@
 So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
-This can be seen in \autoref{fig:binary}\ref{fig:binary_learnS}.
+This can be seen in \autoref{fig:binary}\ref{fig:binary_learnS} (black curves).
 
@@
 which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
-This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS}.
+This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS} (solid curves).
 As discussed below \eqref{eq:multiKOnopre}, the smaller values of has two effects.
@@
 This means that the mutant will initially learn faster than the wild-type.
-However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}
+However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS} (solid curves).
 
@@
 which is always greater than 1.
-Therefore, pre-training will always enhance learning, for both mutant and wild-type.
+Therefore, pre-training will always enhance learning, for both mutant and wild-type, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
 
@@
 
-If we compare gain-increase learning in the mutant to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_eq}: \ref{fig:pooled_plenty_eq_WT} and \ref{fig:pooled_plenty_eq_KO} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_eq}: \ref{fig:pooled_scarce_eq_WT} and \ref{fig:pooled_scarce_eq_KO}), where there are fewer synapses available for depression and resources are depleted.
-When resource depletion is less severe, the first effect dominates and the mutant learns faster than wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learn},\ref{fig:pooled_plenty_learnS}).
+If we compare gain-increase learning in the mutant to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_eq}: \ref{fig:pooled_plenty_eq_WT} and \ref{fig:pooled_plenty_eq_KO} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_eq}: \ref{fig:pooled_scarce_eq_WT} and \ref{fig:pooled_scarce_eq_KO}, blue traces), where there are fewer synapses available for depression and resources are depleted.
+When resource depletion is less severe, the first effect dominates and the mutant learns faster than wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learn},\ref{fig:pooled_plenty_learnS}, solid curves).
 This is reasonable, as when resource depletion is removed it will reduce to the two-state model, which showed this behaviour.
-However, when resource depletion is more severe, the second effect dominates and the mutant learns slower than wild-type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learn},\ref{fig:pooled_scarce_learnS}), which matches what is seen in the experiment.
+However, when resource depletion is more severe, the second effect dominates and the mutant learns slower than wild-type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learn},\ref{fig:pooled_scarce_learnS}, solid curves), which matches what is seen in the experiment.
 
@@
 
-Gain-decrease pre-training will shift the distribution to the potentiated site, where there are more synapses available for depression and resources are more plentiful, for both the mutant and wild type.
+Gain-decrease pre-training will shift the distribution to the potentiated site, where there are more synapses available for depression and resources are more plentiful, for both the mutant and wild type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_eq}, red traces).
 This means that the pre-trained animal will learn faster then the untrained one for both mutant and wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learnS} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learnS}).
@@
 With shorter pre-training, we see that it speeds up learning in both wild-type and mutant (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the mutant.
-This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning.
+This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning (see \autoref{fig:cascade_short}\ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_WT_pre}).
 
-We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_learnS} that longer pre-training pushes the synapses further down the cascade, slow in down learning for the wild-type.
+We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_pr}\ref{fig:cascade_long_pr_WT_pre} that longer pre-training pushes the synapses further down the cascade, slowing down learning for the wild-type.
 This effect is weaker for the mutant, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).

### 2013-04-03 MHCpaper/MHCmodelFig.tex
rearranged MHC folder

\documentclass[12pt,landscape,letterpaper]{article}
%Preamble:
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
\usepackage[margin=1in]{geometry}
\usepackage{array}
\setlength{\extrarowheight}{0.5cm}

\begin{document}
\pagestyle{empty}

\begin{flushleft}
\begin{tabular}{>{\sffamily\bfseries}rc!{\hspace{0.5cm}}c!{\hspace{0.5cm}}c}
  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
  a & \aligntop{\includegraphics[height=2cm]{binary.svg}} & \aligntop{\includegraphics[height=2.4cm]{pooled.svg}} & \aligntop{\includegraphics[height=2cm]{multistate.svg}} \\
  b & \aligntop{\includegraphics[height=6cm]{binary_learnS.eps}} & \aligntop{\includegraphics[height=6cm]{pooled_scarce_learnS.eps}} & \aligntop{\includegraphics[height=6cm]{multistate_med_learnS.eps}} \\
  c & \aligntop{\includegraphics[height=3cm]{binary_eq_WT.eps}} & \aligntop{\includegraphics[height=3cm]{pooled_scarce_eq_WT.eps}} & \aligntop{\includegraphics[height=3cm]{multistate_med_eq_WT.eps}} \\
   & \aligntop{\includegraphics[height=3cm]{binary_eq_KO.eps}} & \aligntop{\includegraphics[height=3cm]{pooled_scarce_eq_KO.eps}} & \aligntop{\includegraphics[height=3cm]{multistate_med_eq_KO.eps}} \\
\end{tabular}
\end{flushleft}

\end{document}

### 2013-04-04 MHCpaper/MHCmodelFig.tex
put figs together

@@
-\documentclass[12pt,landscape,letterpaper]{article}
+\documentclass[12pt,letterpaper]{article}
 %Preamble:
@@
 \graphicspath{{"Figures/"}}
+\newenvironment{figenuma}{\begin{inparaenum}[\sffamily a]}{\end{inparaenum}}
+\newcommand{\hp}{\hspace{0.5cm}}
+\newcommand{\vp}{\vspace{1cm}}
 \usepackage[margin=1in]{geometry}
-\usepackage{array}
-\setlength{\extrarowheight}{0.5cm}
-
+%\usepackage{array}
+%\setlength{\extrarowheight}{0.5cm}
+%
 \begin{document}
 \pagestyle{empty}
+%
+%\begin{flushleft}
+%\begin{tabular}{>{\sffamily\bfseries}rc!{\hspace{0.5cm}}c!{\hspace{0.5cm}}c}
+%  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+%  a & \aligntop{\includegraphics[height=2cm]{binary.svg}} & \aligntop{\includegraphics[height=2.4cm]{pooled.svg}} & \aligntop{\includegraphics[height=2cm]{multistate.svg}} \\
+%  b & \aligntop{\includegraphics[height=6cm]{binary_learnS.eps}} & \aligntop{\includegraphics[height=6cm]{pooled_scarce_learnS.eps}} & \aligntop{\includegraphics[height=6cm]{multistate_med_learnS.eps}} \\
+%  c & \aligntop{\includegraphics[height=3cm]{binary_eq_WT.eps}} & \aligntop{\includegraphics[height=3cm]{pooled_scarce_eq_WT.eps}} & \aligntop{\includegraphics[height=3cm]{multistate_med_eq_WT.eps}} \\
+%   & \aligntop{\includegraphics[height=3cm]{binary_eq_KO.eps}} & \aligntop{\includegraphics[height=3cm]{pooled_scarce_eq_KO.eps}} & \aligntop{\includegraphics[height=3cm]{multistate_med_eq_KO.eps}} \\
+%\end{tabular}
+%\end{flushleft}
 
+%
+\begin{center}
+  \Large{\textsf{Figure 5\_Raymond}}
+\end{center}
+%
+\vp
 \begin{flushleft}
-\begin{tabular}{>{\sffamily\bfseries}rc!{\hspace{0.5cm}}c!{\hspace{0.5cm}}c}
-  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-  a & \aligntop{\includegraphics[height=2cm]{binary.svg}} & \aligntop{\includegraphics[height=2.4cm]{pooled.svg}} & \aligntop{\includegraphics[height=2cm]{multistate.svg}} \\
-  b & \aligntop{\includegraphics[height=6cm]{binary_learnS.eps}} & \aligntop{\includegraphics[height=6cm]{pooled_scarce_learnS.eps}} & \aligntop{\includegraphics[height=6cm]{multistate_med_learnS.eps}} \\
-  c & \aligntop{\includegraphics[height=3cm]{binary_eq_WT.eps}} & \aligntop{\includegraphics[height=3cm]{pooled_scarce_eq_WT.eps}} & \aligntop{\includegraphics[height=3cm]{multistate_med_eq_WT.eps}} \\
-   & \aligntop{\includegraphics[height=3cm]{binary_eq_KO.eps}} & \aligntop{\includegraphics[height=3cm]{pooled_scarce_eq_KO.eps}} & \aligntop{\includegraphics[height=3cm]{multistate_med_eq_KO.eps}} \\
-\end{tabular}
+\begin{figenuma}
+  \item \aligntop{\includegraphics[height=5cm]{data_shifted.svg}}
+  \hp
+  \item \aligntop{\includegraphics[height=5cm]{binary_learnS.eps}}
+  \\ \vp
+  \item \hp\aligntop{\includegraphics[height=5cm]{binary_bar_wt_wo.svg}}
+  \hp
+  \item \hp\aligntop{\includegraphics[height=5cm]{binary_bar_ko_wo.svg}}
+  \hp
+  \item \hp\aligntop{\includegraphics[height=5cm]{binary_bar_wt_w.svg}}
+  \hp
+  \item \hp\aligntop{\includegraphics[height=5cm]{binary_bar_ko_w.svg}}
+  \\ \vp
+  \item \aligntop{
+  \includegraphics[height=4.5cm]{multistate_lin.svg}}
+  \hp
+  \item \aligntop{\includegraphics[height=5cm]{multistate_lin_learnS.eps}}
+\end{figenuma}
+\end{flushleft}
+
+\newpage
+
+%
+\begin{center}
+  \Large{\textsf{Figure 6\_Raymond}}
+\end{center}
+%
+\vp
+\begin{flushleft}
+\begin{figenuma}
+  \item \aligntop{\includegraphics[height=4cm]{pooled_main.svg}}
+  \hp
+  \item \aligntop{\includegraphics[height=5cm]{pooled_scarce_learnS.eps}}
+\end{figenuma}
+\end{flushleft}
+
+\newpage
+
+%
+\begin{center}
+  \Large{\textsf{Figure 7\_Raymond}}
+\end{center}
+%
+\vp
+\begin{flushleft}
+\begin{figenuma}
+  \item \hp\aligntop{\includegraphics[height=3cm]{multistate_bar_wt_wo.svg}}
+  \hp
+  \item \hp\aligntop{\includegraphics[height=3cm]{multistate_bar_ko_wo.svg}}
+  \hp
+  \item \hp\aligntop{\includegraphics[height=3cm]{multistate_bar_wt_w.svg}}
+  \hp
+  \item \hp\aligntop{\includegraphics[height=3cm]{multistate_bar_ko_w.svg}}
+  \\ \vp
+  \item \aligntop{\includegraphics[height=5cm]{multistate_med_learnS.eps}}
+  \item \aligntop{\includegraphics[height=4cm]{cascade_main.svg}}
+  \item \aligntop{\includegraphics[height=5cm]{cascade_long_learnS.eps}}
+\end{figenuma}
 \end{flushleft}

### 2013-04-08 MHCpaper/MHCsupp.tex
rearranged MHC folder
adjusted spacing
match main text better

--- 2013-02-26 Notes/MHCsupp.tex
@@
 \input{sl_preamble.tex}
-\hypersetup{colorlinks,linkcolor=blue,citecolor=violet}
+\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
 \input{sl_graphics_preamble.tex}
@@
   \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}
+  \item\aligntop{\includegraphics[height=2.2cm]{cascade.svg}}\label{fig:cascade_model}
  \end{myenuma}
@@
  \begin{center}
-  \aligntop{\includegraphics[height=1.7cm]{pooled.svg}}
+  \aligntop{\includegraphics[height=4cm]{pooled.svg}}
  \end{center}
@@
   \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the mutant.
-  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
-  \item After pre-training, gain-increase learning is significantly faster in the mutant than in the wild-type.
   \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
   \item For the mutant, gain-increase learning is significantly faster with pre-training than without it.
+  \item After pre-training, gain-increase learning is significantly faster in the mutant than in the wild-type.
+  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
 \end{enumerate}
@@
 We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
-This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measuring using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
+This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measured using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
 This should be multiplied by the difference in synaptic weights.
@@
   \item\aligntop{\includegraphics[width=7cm]{#3_learnS.eps}}\label{fig:#3_learnS}
+  \\ \vp\vp
   \item\label{fig:#3_eq}\begin{myenumi}
@@
                   \end{myenumi}
+  \\ \vp \vp
   \item\label{fig:#3_pr}\begin{myenumi}
@@
   \item\aligntop{\includegraphics[width=7cm]{#2_learnS.eps}}\label{fig:#2_learnS}
+  \\ \vp \vp
   \item\label{fig:#2_eq}\begin{myenumi}
@@
                   \end{myenumi}
+  \\ \vp \vp
   \item\label{fig:#2_pr}\begin{myenumi}
@@
 \begin{equation}\label{eq:multiWTnopre}
-  \Phi = -\frac{2\Delta fq}{M},
+  \Phi = \frac{2\abs{\Delta f}q}{M},
 \end{equation}
@@
           {(1-2\Delta f)^M - (1+2\Delta f)^M} \\
-       &= -\frac{4\Delta fq}{M} + \CO(\Delta f)^2.
+       &= \frac{4\abs{\Delta f}q}{M} + \CO(\Delta f)^2.
 \end{aligned}
@@
 \begin{equation}\label{eq:multiKOnopre}
-  \Phi = -2\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
+  \Phi = 2\abs{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
 \end{equation}
@@
 \begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f) - \beta(1-2\Delta f)}
+  \Phi &= 4\abs{\Delta f} q \, \frac{(1+2\Delta f) - \beta(1-2\Delta f)}
           {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M}   \,
           \brk{\beta (1-2\Delta f) (1+2\Delta f)}^{M/2-1} \\
-       &= -4\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
+       &= 4\abs{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
 \end{aligned}
@@
 \begin{aligned}
-  \Phi &= -2\prn{\Delta f + \overline{\Delta f}} q \, \frac{(1+2\overline{\Delta f}) - \beta(1-2\overline{\Delta f})}
+  \Phi &= 2\abs{\Delta f + \overline{\Delta f}} q \, \frac{(1+2\overline{\Delta f}) - \beta(1-2\overline{\Delta f})}
           {(1+2\overline{\Delta f})^M - \beta^M(1-2\overline{\Delta f})^M}   \,
           \brk{\beta (1-2\overline{\Delta f}) (1+2\overline{\Delta f})}^{M/2-1} \\
-       &= -2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
+       &= 2\abs{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
 \end{aligned}
@@
 \begin{aligned}
-  \Phi &= -8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \,
+  \Phi &= 8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \,
           \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
-          {(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}   \\
-       &= -\frac{2\prn{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
+          {\abs{(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}}   \\
+       &= \frac{2\abs{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
 \end{aligned}
@@
 \begin{equation}\label{eq:binWTnopre}
-  \Phi = -\Delta f q,
+  \Phi = \abs{\Delta f} q,
 \end{equation}
@@
 \begin{aligned}
-  \Phi &= -2\Delta f q.
+  \Phi &= 2\abs{\Delta f} q.
 \end{aligned}
@@
 \begin{equation}\label{eq:binKOnopre}
-  \Phi = -\frac{2\Delta f q}{1+\beta},
+  \Phi = \frac{2\abs{\Delta f} q}{1+\beta},
 \end{equation}
@@
 This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS} (solid curves).
-As discussed below \eqref{eq:multiKOnopre}, the smaller values of has two effects.
+As discussed above (below \eqref{eq:multiKOnopre}), the larger value of $q\dep=q/\beta$ has two effects.
 Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
-The second effect is much smaller for the very short chain considered here.
+For the step-multistate model, we argued that the first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ to leading order.
+Here $M=2$, so the leading part of the two effects will cancel.
+The subleading effects are responsible for the faster learning in the mutant.
 
@@
 \begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
+  \Phi &= 4\abs{\Delta f} q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
           {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
@@
 \begin{equation}\label{eq:binKOnoprediff}
-  \Phi = -\frac{\Delta f q}{f\dep+\beta f\pot}.
+  \Phi = \frac{\abs{\Delta f} q}{f\dep+\beta f\pot}.
 \end{equation}
@@
 \begin{equation}\label{eq:multiLinWTnopre}
-  \Phi = -2\Delta fq\,\frac{M-1}{M},
+  \Phi = 2\abs{\Delta f}q\,\frac{M-1}{M},
 \end{equation}
@@
 \begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
+  \Phi &= 4\abs{\Delta f} q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
           {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
-       &= -{4\Delta f q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
+       &= {4\abs{\Delta f} q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
 \end{aligned}
@@
 \begin{equation}\label{eq:multiLinKOnopre}
-  \Phi = -2\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
+  \Phi = 2\abs{\Delta f} q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
 \end{equation}
@@
 \begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
+  \Phi &= 4\abs{\Delta f} q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
           {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M} \\
-       &= -4\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
+       &= 4\abs{\Delta f} q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
 \end{aligned}
@@
 \begin{equation}\label{eq:multiLinprenopre}
-  \frac{\Phi_{\text{with pre}}}{\Phi_{\text{without pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
+  \frac{\Phi_{\text{w/ pre}}}{\Phi_{\text{w/o pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
    \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,

### 2013-04-10 MHCpaper/MHCsupp.tex
minor
minor update

@@
 \end{equation}
+%
 where $\I$ is the identity matrix.
-%
 Eventually, this will settle into the equilibrium distribution:
@@
 
-Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed.
+Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed and replenished when this is reversed.
 We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single compound synapse.

### 2013-04-11 Code/Calc/OneTimeMax.m
envelope slides

@@ -1,0 +1,22 @@
+function [ s ] = OneTimeMax( t,t1,n )
+%s=ONETIMEMAX(t,t1,n) curve that maximises SNR at t1, with 3 constraints
+%s(t) maximises s(t1).
+%   t=full range of times
+%   t1=time of max
+%   n=#states
+gammasq=(128)/pi^4;
+
+if t1<gammasq/2
+    s=exp(-t/gammasq);
+elseif t1<(n-1)^2/(2*gammasq)
+    s=sqrt(gammasq/(2*t1))*exp(-t/(2*t1));
+elseif t1<(n-1)^2/(gammasq)
+    s=gammasq/(n-1)*exp(-t*gammasq/(n-1)^2);
+else
+    s=(n-1)/t1*exp(-t/t1);
+end
+
+
+
+end
+

### 2013-04-11 Code/Calc/OneTimeMax2.m

@@ -1,0 +1,15 @@
+function [ s ] = OneTimeMax2( t,t1,n )
+%s=ONETIMEMAX2(t,t1,n) curve that maximises SNR at t1, with 2 constraints
+%s(t) maximises s(t1).
+%   t=full range of times
+%   t1=time of max
+%   n=#states
+
+if t1<n-1
+    s=exp(-t/n-1);
+else
+    s=(n-1)/t1*exp(-t/t1);
+end
+
+end
+

### 2013-04-11 Code/Figs/EnvSlideFigs.m

@@ -1,0 +1,78 @@
+fig=figure('PaperPositionMode','auto','Position',[60 60 800 400]);
+ax=axes('Parent',fig,'FontSize',16);
+hold(ax,'on')
+
+t=0.5:0.5:4000;
+n=40;
+
+[env,bnds]=SNRenvelope(t,n);
+[env2,bnds2]=SNRenvelope2(t,n);
+yl=[env(end)/5 env2(1)];
+
+
+t1=200;
+s1=OneTimeMax2(t,t1,n);
+tn1=find(t==t1);
+
+lh1=line([1;1]*t1,yl','Color','r','LineWidth',2,'LineStyle','--','Parent',ax);
+ph1=plot(t,s1,'b','LineWidth',2,'Parent',ax);
+mh1=plot(t1,s1(tn1),'dg','Parent',ax,'MarkerFaceColor','g');
+
+
+
+set(ax,'XScale','log','YScale','log')
+xlim(ax,[t(1) t(end)])
+ylim(ax,yl)
+xlabel(ax,'Time','FontSize',30)
+ylabel(ax,'SNR','FontSize',30);
+
+pause;
+print(fig,'env2_t1.eps','-depsc');
+
+delete([lh1,ph1]);
+clear('t1','s1','tn1','lh1','ph1');
+
+t2=20;
+s2=OneTimeMax2(t,t2,n);
+tn2=find(t==t2);
+
+lh2=line([1;1]*t2,yl','Color','r','LineWidth',2,'LineStyle','--','Parent',ax);
+ph2=plot(t,s2,'b','LineWidth',2,'Parent',ax);
+mh2=plot(t2,s2(tn2),'dg','Parent',ax,'MarkerFaceColor','g');
+
+pause;
+print(fig,'env2_t2.eps','-depsc');
+
+
+delete([mh1,mh2,lh2,ph2]);
+clear('mh1','mh2','t2','s2','tn2','lh2','ph2');
+
+plot(t,env2,'g','LineWidth',3,'Parent',ax);
+lh2=line([1;1]*bnds2',yl'*ones(1,length(bnds2)),'Color','k','LineWidth',1.5,'LineStyle','--','Parent',ax);
+
+pause;
+print(fig,'env2_all.eps','-depsc');
+
+nvals=[4 10 16 28 40];
+epsvals=0.1.^(1:0.5:3);
+for nn=nvals
+s=UniSMScurve(t,nn,1);
+plot(t,s,'b','LineWidth',1.5)
+end
[…]

### 2013-04-15 MHCpaper/MHCsupp.tex
rearranged figures

@@
   States of strong/weak synaptic weight indicated by red/blue circles.
-  States of strong/weak synaptic weight
   (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
@@
 
-\newcommand{\resultsfig}[3]{\begin{figure}
- \begin{center}
+\newcommand{\arrangefig}[1]{\begin{center}
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{#3_learn.eps}}\label{fig:#3_learn}
-  \item\aligntop{\includegraphics[width=7cm]{#3_learnS.eps}}\label{fig:#3_learnS}
+  \item\aligntop{\includegraphics[width=7cm]{#1_learn.eps}}\label{fig:#1_learn}
+  \item\aligntop{\includegraphics[width=7cm]{#1_learnS.eps}}\label{fig:#1_learnS}
   \\ \vp\vp
-  \item\label{fig:#3_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_WT.eps}}\label{fig:#3_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_KO.eps}}\label{fig:#3_eq_KO}
+  \item\label{fig:#1_eq}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_WT.eps}}\label{fig:#1_eq_WT}
+                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_KO.eps}}\label{fig:#1_eq_KO}
                   \end{myenumi}
   \\ \vp \vp
-  \item\label{fig:#3_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_nopre.eps}}\label{fig:#3_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_pre.eps}}\label{fig:#3_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_nopre.eps}}\label{fig:#3_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_pre.eps}}\label{fig:#3_pr_KO_pre}
+  \item\label{fig:#1_pr}\begin{myenumi}
+                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_nopre.eps}}\label{fig:#1_pr_WT_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_pre.eps}}\label{fig:#1_pr_WT_pre}
+                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_nopre.eps}}\label{fig:#1_pr_KO_nopre}
+                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_pre.eps}}\label{fig:#1_pr_KO_pre}
                   \end{myenumi}
  \end{myenuma}
- \end{center}
-  \caption[#1]{#1 #2.
-  Other parameters can be found in \autoref{tr:#3} of \autoref{tab:params}.
+ \end{center}}
+
+\newcommand{\captiontext}[1]{ can be found in \autoref{tr:#1} of \autoref{tab:params}.
   Time is measured in units of $1/r$, the mean time between plasticity events.
-  (\ref{fig:#3_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:#3_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:#3_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#3_eq_WT}) wild-type and (\ref{fig:#3_eq_KO}) \KO\ mutant,
+  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
+  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
+  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant,
   with states of weak synaptic strength to the left and strong states to the right.
-  (\ref{fig:#3_pr}) Evolution of probability distributions for (\ref{fig:#3_pr_WT_nopre}\ref{fig:#3_pr_WT_pre}) wild-type and  (\ref{fig:#3_pr_KO_nopre}\ref{fig:#3_pr_KO_pre}) \KO\ mutant without (\ref{fig:#3_pr_WT_nopre},\ref{fig:#3_pr_KO_nopre}) and with (\ref{fig:#3_pr_WT_pre},\ref{fig:#3_pr_KO_pre}) pre-training,
-  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#3}
+  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre}\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre}\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
+  with states of weak synaptic strength at the top and strong states underneath. }
+
+\newcommand{\captionstart}[2]{Simulation results for the \textbf{#1} model with #2}
+\newcommand{\captionstarts}[2]{Simulation results for the #1 model with #2}
+\newcommand{\captionstartn}[1]{Simulation results for the \textbf{#1} model}
+\newcommand{\captionstartsn}[1]{Simulation results for the #1 model}
+
+\newcommand{\resultsfig}[4]{\begin{figure}
+  \arrangefig{#4}
+  \caption[\captionstarts{#1}{#2}]{\captionstart{#1}{#2 #3}.
+  Other parameters \captiontext{#4} } \label{fig:#4}
 \end{figure}}
@@
 \newcommand{\resultsfign}[2]{\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{#2_learn.eps}}\label{fig:#2_learn}
-  \item\aligntop{\includegraphics[width=7cm]{#2_learnS.eps}}\label{fig:#2_learnS}
-  \\ \vp \vp
-  \item\label{fig:#2_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_WT.eps}}\label{fig:#2_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_KO.eps}}\label{fig:#2_eq_KO}
-                  \end{myenumi}
-  \\ \vp \vp
-  \item\label{fig:#2_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_nopre.eps}}\label{fig:#2_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_pre.eps}}\label{fig:#2_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_nopre.eps}}\label{fig:#2_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_pre.eps}}\label{fig:#2_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}
-  \caption[#1]{#1.
-  Parameters can be found in \autoref{tr:#2} of \autoref{tab:params}.
-  Time is measured in units of $1/r$, the mean time between plasticity events.
-  (\ref{fig:#2_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:#2_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:#2_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#2_eq_WT}) wild-type and (\ref{fig:#2_eq_KO}) \KO\ mutant,
-  with states of weak synaptic strength to the left and strong states to the right.
-  (\ref{fig:#2_pr}) Evolution of probability distributions for (\ref{fig:#2_pr_WT_nopre}\ref{fig:#2_pr_WT_pre}) wild-type and  (\ref{fig:#2_pr_KO_nopre}\ref{fig:#2_pr_KO_pre}) \KO\ mutant without (\ref{fig:#2_pr_WT_nopre},\ref{fig:#2_pr_KO_nopre}) and with (\ref{fig:#2_pr_WT_pre},\ref{fig:#2_pr_KO_pre}) pre-training,
-  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#2}
+  \arrangefig{#2}
+  \caption[\captionstartsn{#1}]{\captionstartn{#1}.
+  Parameters \captiontext{#2} } \label{fig:#2}
 \end{figure}}
@@
 
-\resultsfig{Simulation results for the step-multistate model with weak training}{($\Delta f=-0.1$)}{multistate_weak}
+\resultsfig{step-multistate}{weak training}{($\Delta f=-0.1$)}{multistate_weak}
 
-\resultsfig{Simulation results for the step-multistate model with moderate training}{($\Delta f=-0.17$)}{multistate_med}
+\resultsfig{step-multistate}{moderate training}{($\Delta f=-0.17$)}{multistate_med}
 
-\resultsfig{Simulation results for the step-multistate model with strong training}{($\Delta f=-0.4$)}{multistate_strong}
+\resultsfig{step-multistate}{strong training}{($\Delta f=-0.4$)}{multistate_strong}
 
@@
 
-\resultsfign{Simulation results for two-state model}{binary}
+\resultsfign{two-state model}{binary}
 
@@
 
-\resultsfign{Simulation results for multistate model with linear weights}{multistate_lin}
+\resultsfign{linear multistate model}{multistate_lin}
 
@@
 
-\resultsfig{Simulation results for the pooled resource model with light depletion}{($q\lmin/q\lmax=0.75$)}{pooled_plenty}
+\resultsfig{pooled resource}{light depletion}{($q\lmin/q\lmax=0.75$)}{pooled_plenty}
 
-\resultsfig{Simulation results for the pooled resource model with heavy depletion}{($q\lmin/q\lmax=0.25$)}{pooled_scarce}
+\resultsfig{pooled resource}{heavy depletion}{($q\lmin/q\lmax=0.25$)}{pooled_scarce}
 
@@
 
-\resultsfig{Simulation results for the cascade model with short pre-training}{($rt_\text{pre}=50$)}{cascade_short}
+\resultsfig{cascade}{short pre-training}{($rt_\text{pre}=50$)}{cascade_short}
 
-\resultsfig{Simulation results for the cascade model with long pre-training}{($rt_\text{pre}=150$)}{cascade_long}
+\resultsfig{cascade}{long pre-training}{($rt_\text{pre}=150$)}{cascade_long}
 

### 2013-04-17 MHCpaper/MHCsupp.tex
updated plot parameters

@@
     \cline{1-7}
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 50  &\label{tr:cascade_short} \\
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:cascade_long} \\
-    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
-    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
-    Step-multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
-    Linear Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.3  & 20 & 20  &\label{tr:multistate_lin} \\
-    Two-state     & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
+    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & -0.1  & 5 & 20  &\label{tr:multistate_weak} \\
+    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & -0.3  & 5 & 20  &\label{tr:multistate_med} \\
+    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & -0.45 & 5 & 30  &\label{tr:multistate_strong} \\
+    Two-state          & 2  & $q=0.1$  & $q=0.2$  & -0.1  & 5 & 5   &\label{tr:binary} \\
+    Linear Multistate  & 10 & $q=0.3$  & $q=0.4$  & -0.3  & 5 & 5   &\label{tr:multistate_lin} \\
     Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
-    Pooled        & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
-                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
+                                          & -0.1 & 5 & 20 &\label{tr:pooled_plenty}\\
+    Pooled        & 10 & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
+                                          & -0.1 & 5 & 20 &\label{tr:pooled_scarce}\\
+    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 5 & 20  &\label{tr:cascade_short} \\
+    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 5 & 100 &\label{tr:cascade_long} \\
     \cline{1-7}
@@
 
-\resultsfig{step-multistate}{moderate training}{($\Delta f=-0.17$)}{multistate_med}
+\resultsfig{step-multistate}{moderate training}{($\Delta f=-0.3$)}{multistate_med}
 
-\resultsfig{step-multistate}{strong training}{($\Delta f=-0.4$)}{multistate_strong}
+\resultsfig{step-multistate}{strong training}{($\Delta f=-0.45$)}{multistate_strong}
 
@@
 
-\resultsfig{pooled resource}{heavy depletion}{($q\lmin/q\lmax=0.25$)}{pooled_scarce}
+\resultsfig{pooled resource}{heavy depletion}{($q\lmin/q\lmax=0.125$)}{pooled_scarce}
 
@@
 
-\resultsfig{cascade}{short pre-training}{($rt_\text{pre}=50$)}{cascade_short}
+\resultsfig{cascade}{short pre-training}{($rt_\text{pre}=20$)}{cascade_short}
 
-\resultsfig{cascade}{long pre-training}{($rt_\text{pre}=150$)}{cascade_long}
+\resultsfig{cascade}{long pre-training}{($rt_\text{pre}=100$)}{cascade_long}
 

### 2013-04-17 MHCpaper/Figures/MakeFigs.m

@@ -1,11 +1,9 @@
-MakeVORfigs('multistate_med',0.6,0.8,-0.17,10,30,'print')
-MakeVORfigs('multistate_strong',0.6,0.8,-0.4,20,20,'print')
-MakeVORfigs('multistate_weak',0.6,0.8,-0.1,50,50,'print')
-MakeVORfigs('pooled_plenty',[0.3 0.4],[0.6 0.8],-0.1,70,70,'print','Pooled',true)
-MakeVORfigs('pooled_scarce',[0.1 0.4],[0.2 0.8],-0.1,70,70,'print','Pooled',true)
-MakeVORfigs('cascade_short',0.25,0.33,-0.3,50,50,'print')
-MakeVORfigs('cascade_long',0.25,0.33,-0.3,50,150,'print')
-MakeVORfigs('multistate_weak',0.6,0.8,-0.1,50,50,'print')
-MakeVORfigs('multistate_strong',0.6,0.8,-0.4,20,20,'print')
-MakeVORfigs('binary',0.4,0.8,-0.1,5,5,'print')
-MakeVORfigs('multistate_lin',0.6,0.8,-0.3,20,20,'print','LinWeights',true)
+MakeVORfigs('multistate_med',0.3,0.4,-0.3,5,20,'print')
+MakeVORfigs('multistate_strong',0.3,0.4,-0.45,5,30,'print')
+MakeVORfigs('multistate_weak',0.3,0.4,-0.1,5,20,'print')
+MakeVORfigs('binary',0.1,0.2,-0.1,5,5,'print')
+MakeVORfigs('multistate_lin',0.3,0.4,-0.3,5,5,'print','LinWeights',true)
+MakeVORfigs('pooled_plenty',[0.3 0.4],[0.6 0.8],-0.1,5,20,'print','Pooled',true)
+MakeVORfigs('pooled_scarce',[0.05 0.4],[0.1 0.8],-0.1,5,20,'print','Pooled',true)
+MakeVORfigs('cascade_short',0.25,0.33,-0.3,5,20,'print')
+MakeVORfigs('cascade_long',0.25,0.33,-0.3,5,100,'print')

### 2013-04-25 Notes/ContTime.tex
wrote up some finite time max stuff

@@
 \newcommand{\w}{\mathbf{w}}
+\newcommand{\uv}{\mathbf{u}}
+\newcommand{\vv}{\mathbf{v}}
 \newcommand{\W}{\mathbf{W}}
 \newcommand{\M}{\mathbf{M}}
+\newcommand{\enc}{\mathbf{q}}
+\newcommand{\frg}{\W^{\mathrm{F}}}
 \newcommand{\F}{\boldsymbol{\Phi}}
@@
 %
-In Surya's notes, $\MM=r\W^F$.
+In Surya's notes, $\MM=r\frg$.
 
@@
 \begin{equation}\label{eq:forgetting}
-  \MM = r \W^F \equiv r\prn{f^+\W^+ + f^-\W^-}.
+  \MM = r \frg \equiv r\prn{f^+\W^+ + f^-\W^-}.
 \end{equation}
@@
 \begin{equation}\label{eq:initprob}
-  \mathbf{p}(t|\pm) = \eq\M^\pm \e^{rt\W^F}.
+  \mathbf{p}(t|\pm) = \eq\M^\pm \e^{rt\frg}.
 \end{equation}
@@
 \begin{aligned}
-  S(t) &= N \eq (f^+\W^+ - f^-\W^-)\e^{rt\W^F}\w\\
-    & N (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\W^F}\w.
+  S(t) &= N \eq (f^+\W^+ - f^-\W^-)\e^{rt\frg}\w\\
+    & N (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\frg}\w.
 \end{aligned}
@@
 \begin{equation}\label{eq:SNRcurve}
-  \SNR(t) = \sqrt{N} (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\W^F}\w.
+  \SNR(t) = \sqrt{N} (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\frg}\w.
 \end{equation}
@@
   &\quad&\implies&
-    \W^F&=\W(f^+),\\&&&&
+    \frg&=\W(f^+),\\&&&&
     \W^+ - \W^- &= \diff{\W}{\alpha},\\&&&&
-    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \W^F.
+    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \frg.
   \end{aligned}
@@
 \begin{equation}\label{eq:SNRalpha}
-  \SNR(t) = \sqrt{N} (2f^+f^-) \diff{\eq}{\alpha}(-\W^F)\,\e^{rt\W^F}\w.
+  \SNR(t) = \sqrt{N} (2f^+f^-) \diff{\eq}{\alpha}(-\frg)\,\e^{rt\frg}\w.
 \end{equation}
@@
     \frac{\SNR(t)}{\sqrt{N} (2f^+f^-)} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}^+ - \widehat{\W}^-)\e^{rt\widehat{\W}^F}\widehat{\w}. \\
-      &= \eq VU (\W^+ - \W^-) VU \e^{rt\W^F} VU \w. \\
-      &= \eq (\W^+ - \W^-) VU \e^{rt\W^F} VU \w. \\
-      &= \eq (\W^+ - \W^-)\e^{rt\W^F} VU \w. \\
-      &= \eq (\W^+ - \W^-)\e^{rt\W^F}\w. \\
+      &= \eq VU (\W^+ - \W^-) VU \e^{rt\frg} VU \w. \\
+      &= \eq (\W^+ - \W^-) VU \e^{rt\frg} VU \w. \\
+      &= \eq (\W^+ - \W^-)\e^{rt\frg} VU \w. \\
+      &= \eq (\W^+ - \W^-)\e^{rt\frg}\w. \\
   \end{aligned}
@@
   ,\\ \implies\quad
-  \W^F = \begin{pmatrix}
+  \frg = \begin{pmatrix}
            -f^+ & f^+ \\
@@
   ,%&
-%  \e^{rt\W^F} &= \I+(1-\e^{-rt})\W^F
+%  \e^{rt\frg} &= \I+(1-\e^{-rt})\frg
 %         \begin{pmatrix}
@@
 %\begin{equation}\label{eq:SNR}
-%  \SNR(t) = \sqrt{N}(2f^+f^-) \diff{\eq}{\alpha} (-\W^F) \e^{rt\W^F} \w.
+%  \SNR(t) = \sqrt{N}(2f^+f^-) \diff{\eq}{\alpha} (-\frg) \e^{rt\frg} \w.
 %\end{equation}
@@
 \begin{aligned}
-  A &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\W^F}}_0^\infty \w \\
+  A &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\frg}}_0^\infty \w \\
     &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
@@
 
-We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f^+f^-)\hat{A}$ and $q_{ij} \equiv \diff{\W^F_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
+We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f^+f^-)\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
 %
 \begin{equation}\label{eq:areaT}
-  \hat{A} = \sum_{i,j,k} \eq_i q_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
+  \hat{A} = \sum_{i,j,k} \eq_i \enc_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
 \end{equation}
@@
 \begin{aligned}
-  \hat{A} &= 2\sum_{i,j} q_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}) &
-    &= -2\sum_{i,j} q_{ij} \eq_i \eta^+_{j} \\
-    &= 2\sum_{i,j} q_{ij} \eq_i (\eta^-_{j} - \eta^-_{i}) &
-    &= 2\sum_{i,j} q_{ij} \eq_i \eta^-_{j}.
+  \hat{A} &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}) &
+    &= -2\sum_{i,j} \enc_{ij} \eq_i \eta^+_{j} \\
+    &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^-_{j} - \eta^-_{i}) &
+    &= 2\sum_{i,j} \enc_{ij} \eq_i \eta^-_{j}.
 \end{aligned}
@@
 \begin{equation}\label{eq:areaZ}
-  \hat{A} = \sum_{i,j,k,l} q_{ij} \pib_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
+  \hat{A} = \sum_{i,j,k,l} \enc_{ij} \pib_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
     = \pib \fund q \fund \w.
@@
     c_k &= \diff{\ln\eq_k}{\alpha}
-      = \sum_{ij} \eq_i q_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
+      = \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
       = - \prn{\eq q \fptb}_k
       = \frac{(\eq q \fund)_k}{\eq_k}, \\
-    a_i &= \sum_j q_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
+    a_i &= \sum_j \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
     \implies
@@
 \begin{equation}\label{eq:basicderivs}
-  \pdiff{\W^F_{ij}}{\W^\pm_{gh}} = f^\pm \delta_{gi}(\delta_{hj}-\delta_{ij}),
-  \qquad
-  \pdiff{q_{ij}}{\W^\pm_{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
+  \pdiff{\frg_{ij}}{\W^\pm_{gh}} = f^\pm \delta_{gi}(\delta_{hj}-\delta_{ij}),
+  \qquad
+  \pdiff{\enc_{ij}}{\W^\pm_{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
 \end{equation}
@@
     \pdiff{\hat{A}}{\W^\pm_{gh}} =&\,
-      rf^\pm \sum_{ijkl} q_{ij} \w_k \pib_l \brk{
+      rf^\pm \sum_{ijkl} \enc_{ij} \w_k \pib_l \brk{
          \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
@@
     =&\,
-      2rf^\pm \sum_{ij} q_{ij} \eq_i \eq_g \brk{
+      2rf^\pm \sum_{ij} \enc_{ij} \eq_i \eq_g \brk{
          (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
@@
     =&\,
-      2rf^\pm \sum_{ij} q_{ij} \eq_i \eq_g
+      2rf^\pm \sum_{ij} \enc_{ij} \eq_i \eq_g
          (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
@@
   \begin{aligned}
-    \pdiff{}{\W^F_{gh}} &\equiv     \pdiff{}{\W^+_{gh}} +     \pdiff{}{\W^-_{gh}} ,&\qquad
-    \pdiff{}{q_{gh}}    &\equiv f^- \pdiff{}{\W^+_{gh}} - f^+ \pdiff{}{\W^-_{gh}} .
+    \pdiff{}{\frg_{gh}} &\equiv     \pdiff{}{\W^+_{gh}} +     \pdiff{}{\W^-_{gh}} ,&\qquad
+    \pdiff{}{\enc_{gh}}    &\equiv f^- \pdiff{}{\W^+_{gh}} - f^+ \pdiff{}{\W^-_{gh}} .
   \end{aligned}
@@
 \begin{equation}\label{eq:derivqweff}
-  \pdiff{\W^F_{ij}}{\W^F_{gh}} = \pdiff{q_{ij}}{q_{gh}}
+  \pdiff{\frg_{ij}}{\frg_{gh}} = \pdiff{\enc_{ij}}{\enc_{gh}}
   =  \delta_{gi}(\delta_{hj}-\delta_{ij}),
   \qquad
-  \pdiff{q_{ij}}{\W^F_{gh}} = \pdiff{\W^F_{ij}}{q_{gh}}  = 0.
-\end{equation}
-%
-This is because we could treat $\W^F$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W^\pm$.
+  \pdiff{\enc_{ij}}{\frg_{gh}} = \pdiff{\frg_{ij}}{q_{gh}}  = 0.
+\end{equation}
+%
+This is because we could treat $\frg$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W^\pm$.
 
@@
     \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
-    \Delta q_{ij} &= q_{ij} ,&
+    \Delta \enc_{ij} &= \enc_{ij} ,&
     \Delta \hat{A}  &= 0 ,&
@@
 \begin{equation}\label{eq:shiftqderiv}
-  \pdiff{\CL}{q_{gh}} =
+  \pdiff{\CL}{\enc_{gh}} =
   f^- \pdiff{\CL}{\W^+_{gh}} - f^+ \pdiff{\CL}{\W^-_{gh}}
@@
 %
-which leaves $\W^F$ unchanged, and therefore $\eq$, $\fpt$ and $\eta^\pm$ as well.
+which leaves $\frg$ unchanged, and therefore $\eq$, $\fpt$ and $\eta^\pm$ as well.
 
@@
 
-%Note that if we multiply \eqref{eq:shiftqderiv} by $q_{gh}$ and sum over $(g,h)$, we get
+%Note that if we multiply \eqref{eq:shiftqderiv} by $\enc_{gh}$ and sum over $(g,h)$, we get
 %%
 %\begin{equation}\label{eq:stupid}
-%  \hat{A} = \sum_\pm\sum_{g\neq h} \mu^\pm_{gh} \W^F_{gh},
+%  \hat{A} = \sum_\pm\sum_{g\neq h} \mu^\pm_{gh} \frg_{gh},
 %\end{equation}
@@
     \Delta_{gh} &\equiv
-      \frac{1}{\eq_{g}} \pdiffc{}{\W^F_{gh}}
-      + \frac{1}{\eq_h} \pdiffc{}{\W^F_{hg}}, \\
+      \frac{1}{\eq_{g}} \pdiffc{}{\frg_{gh}}
+      + \frac{1}{\eq_h} \pdiffc{}{\frg_{hg}}, \\
 %  \end{aligned}
@@
       \end{cases} \\
-    r\eq_i \W^F_{ij} \prn{\fptb_{ij}+\fptb_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
+    r\eq_i \frg_{ij} \prn{\fptb_{ij}+\fptb_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
   \end{aligned}
@@
   ,\\
-  c_{k+1} - c_k =&\, \prn{\fpt_{k,k+1}+\fpt_{k+1,k}} \prn{\frac{\eq_{k}\W^F_{k,k+1}}{f^+}+\frac{\eq_{k+1}\W^F_{k+1,k}}{f^-}}
+  c_{k+1} - c_k =&\, \prn{\fpt_{k,k+1}+\fpt_{k+1,k}} \prn{\frac{\eq_{k}\frg_{k,k+1}}{f^+}+\frac{\eq_{k+1}\frg_{k+1,k}}{f^-}}
     =%\\=&\,
     \frac{1}{rf^+f^-},\\
-  \sum_k c_k \eq_k =&\, \sum_{ij} \eq_i q_{ij} (\eta-\eta) = 0,\\
+  \sum_k c_k \eq_k =&\, \sum_{ij} \eq_i \enc_{ij} (\eta-\eta) = 0,\\
 %  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f^+f^-},
@@
 
+\section{Finite time SNR}\label{sec:finitetime}
+
+The SNR at time $t$ was given in \eqref{eq:SNRcurve} as
+%
+\begin{equation*}
+  \SNR(t) = \sqrt{N} (2f^+f^-) \eq \enc \, \e^{rt\frg}\w.
+\end{equation*}
+%
+Let us try to maximise this at time $t$ only.
+We will consider the independent parameters to be the off-diagonal elements of $\M^\pm$, with the diagonal elements determined via $\M^\pm_{ii} = \sum_{j\neq i} \M^\pm_{ij}$.
+Consider the Lagrangian
+%
+\begin{equation}\label{eq:finlagrangian}
+  \CL = \eq \enc\e^{rt\frg}\w + \sum_\pm \sum_{ij} \mu^\pm_{ij} \M^\pm_{ij}.
+\end{equation}
+%
+Necessary conditions for an extremum are
+%
+\begin{equation*}
+  \pdiff{\CL}{\M^\pm_{gh}} = 0,
+  \qquad
+    \mu^\pm_{gh} \geq 0,\quad
+    \M^\pm_{gh} \geq 0,\quad
+    \mu^\pm_{gh}\M^\pm_{gh} = 0.
+\end{equation*}
+%
+We can use the results of \cite{Ortiz2001diffmatexp} to write down the shift of this Lagrangian under an infinitesimal perturbation
+%
+\begin{equation}\label{eq:finpert}
+\begin{aligned}
+  \delta\CL =&\, \eq \delta\frg \fund\, \enc \, \e^{rt\frg} \w 
+    + \eq \delta\enc \, \e^{rt\frg} \w %\\ &
+   + r\int_{0}^{t}\!\!\dr\tau\, \eq \enc\, \e^{r\tau\frg} \delta\frg \e^{r(t-\tau)\frg} \w\\
+   &+ \sum_\pm \sum_{ij} \mu^\pm_{ij} \delta\M^\pm_{ij}
+\end{aligned}
+\end{equation}
+%
+This can be rewritten using the eigenvector decomposition of $\frg$.
+%
+\begin{equation}\label{eq:eigendecomp}
+  \frg = \sum_a -q_a \uv^a \vv^a,
+  \quad
+  \vv^a \uv^b = \delta_{ab},
+  \quad
+  \frg \uv^a = -q_a \uv^a,
+  \quad
+  \vv^a \frg = -q_a \vv^a.
+\end{equation}
+%
+People often write $\vv$ as $\vv\trans$, presumably because they find the concept of row vectors too difficult to handle. 
+We're better than that.
+We can choose $q_1=0$, $\uv^1=\onev$ and $\vv=\eq$.
+
+Then, defining
+%
+\begin{equation}\label{eq:eigdiff}
+  F_{ab} = \left\{
+           \begin{aligned}
+             &rt\,\e^{-q_art}, &\quad&\text{if } q_a=q_b,\\
+             &\frac{\e^{-q_art}-\e^{-q_brt}}{q_b-q_a}, &&\text{otherwise,}
+           \end{aligned}
+           \right.
+\end{equation}
+%
+we get
+%
+\begin{equation}\label{findiff}
+\begin{aligned}
+  \pdiff{\CL}{\M^\pm_{gh}} =&\,
+    f^\pm \eq_g \prn{\brk{\fund \, \enc \, \e^{rt\frg} \w}_h - \brk{\fund \, \enc \, \e^{rt\frg} \w}_g} \\
+    &\pm \eq_g \prn{\brk{\e^{rt\frg} \w}_h - \brk{\e^{rt\frg} \w}_g} \\
+    &+ \sum_{ab} F_{ab} (\eq\enc\uv^a) \vv^{ag} (\uv^b_h-\uv^b_g) (\vv^a\w) \\
+    &+ \mu^\pm_{gh} - \mu^\pm_{gg}.
+\end{aligned}
+\end{equation}
+%
+
+Consider
+%
+\begin{equation}\label{eq:finshiftq}
+\begin{aligned}
+  f^- \pdiff{\CL}{\M^+_{gh}} - f^+ \pdiff{\CL}{\M^-_{gh}}
+   =&\, \eq_g \prn{\brk{\e^{rt\frg} \w}_h - \brk{\e^{rt\frg} \w}_g}\\
+    &+ f^- \mu^+_{gh} + f^+\mu^-_{gg} - f^+ \mu^-_{gh} - f^- \mu^+_{gg}\\
+   =&\, 0,
+\end{aligned}
+\end{equation}
+%
+which corresponds to the shift considered in \S\ref{sec:triangular}. We can see that
+%
+\begin{equation}\label{eq:etalambdaineq}
+  \begin{aligned}
+    \brk{\e^{rt\frg} \w}_g &< \brk{\e^{rt\frg} \w}_h
+%    &\quad & \implies \quad&
+%    f^+ \mu^-_{gh} + f^- \mu^+_{gg} &> 0
+    &\quad & \implies \quad&
+    \mu^-_{gh} &> 0
+    &\quad & \text{or} \quad&
+    \mu^+_{gg} &> 0
+    \\ &
+    &\quad & \implies \quad&
+    \M^-_{gh} &= 0
+    &\quad & \text{or} \quad&
+    \M^+_{gg} &= 0,
+    \\
+    \brk{\e^{rt\frg} \w}_g &> \brk{\e^{rt\frg} \w}_h
+%    &\quad & \implies \quad&
+%    f^- \mu^+_{gh} + f^+\mu^-_{gg} &> 0
+    &\quad & \implies \quad&
+    \mu^+_{gh} &> 0
+    &\quad & \text{or} \quad&
+    \mu^-_{gg} &> 0
+    \\ &
+    &\quad & \implies \quad&
+    \M^+_{gh} &= 0
+    &\quad & \text{or} \quad&
+    \M^-_{gg} &= 0.
+  \end{aligned}
+\end{equation}
+%
+
 \appendix
@@
     c_{k+1} - c_k &= \frac{N^{k+1}_{kk}}{\eq_k} \brk{
-      \sum_{i<j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{H^{k+1}_{i,k} + H^{k}_{j,k+1} - 1}
-      +\sum_{i>j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{H^{k+1}_{j,k} + H^{k}_{i,k+1} - 1}
+      \sum_{i<j} \eq_i \frac{\frg_{ij}}{f^+} \prn{H^{k+1}_{i,k} + H^{k}_{j,k+1} - 1}
+\right.\\&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\left.
+      +\sum_{i>j} \eq_i \frac{\frg_{ij}}{f^+} \prn{H^{k+1}_{j,k} + H^{k}_{i,k+1} - 1}
     } \\
     &= \frac{N^{k+1}_{kk}}{\eq_k} \brk{
-      \sum_{i<j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{1 - H^{k+1}_{j,k} + H^{k}_{i,k+1}}
-      +\sum_{i>j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{1 - H^{k+1}_{i,k} + H^{k}_{j,k+1}}
+      \sum_{i<j} \eq_i \frac{\frg_{ij}}{f^+} \prn{1 - H^{k+1}_{j,k} + H^{k}_{i,k+1}}
+\right.\\&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\left.
+      +\sum_{i>j} \eq_i \frac{\frg_{ij}}{f^+} \prn{1 - H^{k+1}_{i,k} + H^{k}_{j,k+1}}
     }, \\
@@
 
-The Markov chain, $\W^F$, is completely determined by the off-diagonal mean first passage times, $\fptb$, \cite[Th.4.4.12]{kemeny1960finite}:
+The Markov chain, $\frg$, is completely determined by the off-diagonal mean first passage times, $\fptb$, \cite[Th.4.4.12]{kemeny1960finite}:
 %
@@
     \eq &= \eta \prn{\fptb\inv\onev}\trans, \\
-    r\W^F= \MM &= (\D-\onem)\fptb\inv.
+    r\frg= \MM &= (\D-\onem)\fptb\inv.
   \end{aligned}
@@
 \begin{equation}\label{eq:derivTdef}
-  \pdiff{}{\fptb_{gh}} = \sum_{ij} \pdiffc{\W^F_{ij}}{\fptb_{gh}} \pdiff{}{\W^F_{ij}},
+  \pdiff{}{\fptb_{gh}} = \sum_{ij} \pdiffc{\frg_{ij}}{\fptb_{gh}} \pdiff{}{\frg_{ij}},
 \end{equation}
@@
 \begin{equation}\label{eq:derivTWp}
-  \pdiff{\W^F_{ij}}{\fptb_{gh}} = r \eq_h \W^F_{ig} \prn{\W^F_{ij}-\W^F_{hj}} ,
-  \qquad
-  \pdiff{\eq_{i}}{\fptb_{gh}} = -r \eq_h \eq_i \W^F_{ig} .
+  \pdiff{\frg_{ij}}{\fptb_{gh}} = r \eq_h \frg_{ig} \prn{\frg_{ij}-\frg_{hj}} ,
+  \qquad
+  \pdiff{\eq_{i}}{\fptb_{gh}} = -r \eq_h \eq_i \frg_{ig} .
 \end{equation}
@@
   \pdiff{\hat{A}}{\fptb_{gh}} = -\eq_h\brk{
-    \sum_i \eq_i q_{ig} \w_h
-    + r\sum_{ijk} \prn{\W^F_{ig}+\W^F_{kg}} \eq_i q_{ij} \prn{\fptb_{ik}-\fptb_{jk}} \eq_k \w_k
+    \sum_i \eq_i \enc_{ig} \w_h
+    + r\sum_{ijk} \prn{\frg_{ig}+\frg_{kg}} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}} \eq_k \w_k
     }.
@@
   \begin{aligned}
-    \hat{I} &= \diff{\eq}{\alpha}(-\W^F)\w &
-      &= - \pib \fund \diff{\MM}{\alpha} \fund \W^F \w &
-      &= - \pib \fund \diff{\W^F}{\alpha} \fund \MM \w \\
-      &= \pib \fund \diff{\W^F}{\alpha} (\I - \tau\onev\pib) \w &
-      &= \pib \fund \diff{\W^F}{\alpha} \w &
-      &= \eq \diff{\W^F}{\alpha} \w .
+    \hat{I} &= \diff{\eq}{\alpha}(-\frg)\w &
+      &= - \pib \fund \diff{\MM}{\alpha} \fund \frg \w &
+      &= - \pib \fund \diff{\frg}{\alpha} \fund \MM \w \\
+      &= \pib \fund \diff{\frg}{\alpha} (\I - \tau\onev\pib) \w &
+      &= \pib \fund \diff{\frg}{\alpha} \w &
+      &= \eq \diff{\frg}{\alpha} \w .
   \end{aligned}
@@
   \begin{aligned}
-     \hat{I} &= \sum_{i,j} \eq_i q_{ij} \w_j \\
-      &= \sum_{i \neq j} \eq_i q_{ij} \w_j + \sum_i \eq_i q_{ii} \w_i \\
-      &= \sum_{i \neq j} q_{ij} \eq_i (\w_j - \w_i).
+     \hat{I} &= \sum_{i,j} \eq_i \enc_{ij} \w_j \\
+      &= \sum_{i \neq j} \eq_i \enc_{ij} \w_j + \sum_i \eq_i \enc_{ii} \w_i \\
+      &= \sum_{i \neq j} \enc_{ij} \eq_i (\w_j - \w_i).
   \end{aligned}
@@
   \pdiff{\hat{I}}{\W^\pm_{gh}}
-    &= rf^\pm \sum_{ijk} \pib_k \fund_{kg} (\fund_{hi}-\fund_{gi}) q_{ij} \w_j
+    &= rf^\pm \sum_{ijk} \pib_k \fund_{kg} (\fund_{hi}-\fund_{gi}) \enc_{ij} \w_j
     \pm \sum_k \pib_k \fund_{kg} (\w_h - \w_g) \\
-    &= rf^\pm \sum_{ij} \eq_g q_{ij} (\fptb_{gi}-\fptb_{hi}) \eq_i \w_j
+    &= rf^\pm \sum_{ij} \eq_g \enc_{ij} (\fptb_{gi}-\fptb_{hi}) \eq_i \w_j
     \pm \eq_g (\w_h - \w_g).
@@
     \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
-    \Delta q_{ij} &= q_{ij} ,
+    \Delta \enc_{ij} &= \enc_{ij} ,
     \\&&

### 2013-04-27 Notes/ContTime.tex
typo

@@
     &\pm \eq_g \prn{\brk{\e^{rt\frg} \w}_h - \brk{\e^{rt\frg} \w}_g} \\
-    &+ \sum_{ab} F_{ab} (\eq\enc\uv^a) \vv^{ag} (\uv^b_h-\uv^b_g) (\vv^a\w) \\
+    &+ \sum_{ab} F_{ab} (\eq\enc\uv^a) \vv^a_g (\uv^b_h-\uv^b_g) (\vv^a\w) \\
     &+ \mu^\pm_{gh} - \mu^\pm_{gg}.

### 2013-04-30 MHCpaper/MHCsupp.tex
changed sign of df

@@
   \item The synapses that are eligible for plasticity are chosen randomly.
-  \item The potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $rf\potdep$, where $f\pot+f\dep=1$.
-      In other words, $r$ is the total rate of plasticity events per synapse and $f\potdep$ are the fraction of these events that are potentiating/depressing.
+  \item Candidate potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $r\potdep$.
   \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
@@
 \end{itemize}
+
+In practice, the overall rate of candidate plasticity events will only set the timescale, and only the relative rates will matter.
+With this in mind, we define
+%
+\begin{equation}\label{eq:fpotdep}
+  r = r\pot + r\dep,
+  \qquad
+  f\pot = \frac{r\pot}{r},
+  \qquad
+  f\dep = \frac{r\dep}{r}.
+\end{equation}
+%
+The quantities $f\potdep$ measure the fraction of candidate plasticity events that are potentiating/depressing.
 
@@
 It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
-We will model this by changing $f\potdep$, leaving $r$ and $\M\potdep$ unchanged.
+We will model this by changing $f\potdep$, leaving $\M\potdep$ unchanged.
 
-The untrained case will be described by $f\pot=f\pot\norm$.
-Gain-increase training  will be described by $f\pot=f\pot\inc<f\pot\norm$, and
-gain-decrease training  will be described by $f\pot=f\pot\dec>f\pot\norm$.
-Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depend on $f\pot$, which we will indicate with subscripts.
+The untrained case will be described by $f\dep=f\dep\norm$.
+Gain-increase training  will be described by $f\dep=f\dep\inc>f\dep\norm$, and
+gain-decrease training  will be described by $f\dep=f\dep\dec<f\dep\norm$. 
+Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depend on $f\dep$, which we will indicate with subscripts.
 
-Before training, the synaptic distribution will be in the equilibrium distribution corresponding to $f\pot\norm$.
+Before training, the synaptic distribution will be in the equilibrium distribution corresponding to $f\dep\norm$.
 During gain-increase training, it will evolve according to \eqref{eq:evolve} with $f\pot\inc$:
@@
 
-We will assume $f\pot\wt=f\pot\ko$.
-This is because the mutation is well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
+We will assume $f\dep\wt=f\dep\ko$.
+This is because the effects of the mutation is well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
 Therefore the rates of potentiation and depression should not change very much either.
@@
 
-For the most part, we set $f\pot\norm=\half$, $f\pot\inc=f\pot\norm+\Delta f$ and $f\pot\dec=f\pot\norm-\Delta f$, with $\Delta f<0$.
-We use the same values for wild-type and mutant and treat gain-increase and decrease symmetrically.
+For the most part, we set $f\dep\norm=\half$, $f\dep\inc=f\dep\norm+\Delta f$ and $f\dep\dec=f\dep\norm-\Delta f$, with $\Delta f>0$.
+We use the same values for wild-type and mutant for the reasons discussed above.
+We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
+We could adjust $r$ to keep $r\pot$ unchanged, if so desired, but this would only change the overall timescale and would not affect any of the qualitative comparisons that we are concerned with here. 
 The values of these parameters are listed in \autoref{tab:params}.
@@
     \cline{1-7}
-    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & -0.1  & 5 & 20  &\label{tr:multistate_weak} \\
-    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & -0.3  & 5 & 20  &\label{tr:multistate_med} \\
-    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & -0.45 & 5 & 30  &\label{tr:multistate_strong} \\
-    Two-state          & 2  & $q=0.1$  & $q=0.2$  & -0.1  & 5 & 5   &\label{tr:binary} \\
-    Linear Multistate  & 10 & $q=0.3$  & $q=0.4$  & -0.3  & 5 & 5   &\label{tr:multistate_lin} \\
+    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & 0.1  & 5 & 20  &\label{tr:multistate_weak} \\
+    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 20  &\label{tr:multistate_med} \\
+    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & 0.45 & 5 & 30  &\label{tr:multistate_strong} \\
+    Two-state          & 2  & $q=0.1$  & $q=0.2$  & 0.1  & 5 & 5   &\label{tr:binary} \\
+    Linear Multistate  & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 5   &\label{tr:multistate_lin} \\
     Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-                                          & -0.1 & 5 & 20 &\label{tr:pooled_plenty}\\
+                                          & 0.1 & 5 & 20 &\label{tr:pooled_plenty}\\
     Pooled        & 10 & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
-                                          & -0.1 & 5 & 20 &\label{tr:pooled_scarce}\\
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 5 & 20  &\label{tr:cascade_short} \\
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 5 & 100 &\label{tr:cascade_long} \\
+                                          & 0.1 & 5 & 20 &\label{tr:pooled_scarce}\\
+    Cascade       & 10 & $x=0.25$ & $x=0.33$ & 0.3  & 5 & 20  &\label{tr:cascade_short} \\
+    Cascade       & 10 & $x=0.25$ & $x=0.33$ & 0.3  & 5 & 100 &\label{tr:cascade_long} \\
     \cline{1-7}
@@
 \begin{equation}\label{eq:multiWTnopre}
-  \Phi = \frac{2\abs{\Delta f}q}{M},
+  \Phi = \frac{2{\Delta f}q}{M}.
 \end{equation}
 %
-where it's worth remembering that $\Delta f<0$.
+%where it's worth remembering that $\Delta f<0$.
 With pre-training:
@@
 \begin{aligned}
-  \Phi &= 16(\Delta f)^2q \, \frac{(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1}}
-          {(1-2\Delta f)^M - (1+2\Delta f)^M} \\
-       &= \frac{4\abs{\Delta f}q}{M} + \CO(\Delta f)^2.
+  \Phi &= 16(\Delta f)^2q \, \frac{(1+2\Delta f)^{M/2-1} (1-2\Delta f)^{M/2-1}}
+          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
+       &= \frac{4{\Delta f}\,q}{M} + \CO(\Delta f)^2.
 \end{aligned}
@@
 So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS} (black curves).
-On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot due to the factor of $(1+2\Delta f)^{M/2-1}$, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.
+On the other hand, if $\Delta f$ is close to $\half$, pre-training will initially slow down learning a lot due to the factor of $(1-2\Delta f)^{M/2-1}$, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.
 
@@
 
-Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
+Now, consider the mutant, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
 Without pre-training:
@@
 \begin{equation}\label{eq:multiKOnopre}
-  \Phi = 2\abs{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
+  \Phi = 2{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
 \end{equation}
 %
-This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$.
+This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$, where $\beta^*(M)$ is defined as the value at which they are equal.
 This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.
@@
 \begin{aligned}
-  \Phi &= 4\abs{\Delta f} q \, \frac{(1+2\Delta f) - \beta(1-2\Delta f)}
-          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M}   \,
+  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
+          {(1-2\Delta f)^M - \beta^M(1+2\Delta f)^M}   \,
           \brk{\beta (1-2\Delta f) (1+2\Delta f)}^{M/2-1} \\
-       &= 4\abs{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
+       &= 4{\Delta f}\, q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
 \end{aligned}
@@
 Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
-Values of $\Delta f$ that are smaller than this (stronger training) will correspond to slower learning after pre-training.
-As we would like pre-training to slow down learning in the wild-type but speed it up in the mutant, it would seem that we require $\Delta f^*(\beta,M) < \Delta f < \Delta f^*(1,M)$ (remembering once more that $\Delta f<0$ and that the wild-type corresponds to $\beta=1$).
-As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) < \Delta f^*(1,M)$, so this range always exists.
+Values of $\Delta f$ that are larger than this (stronger training) will correspond to slower learning after pre-training.
+As we would like pre-training to slow down learning in the wild-type but speed it up in the mutant, it would seem that we require $\Delta f^*(1,M) < \Delta f < \Delta f^*(\beta,M)$ (remembering that the wild-type corresponds to $\beta=1$).
+As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, so this range always exists.
 We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.
@@
 However, this is only the initial, instantaneous rate of change.
-Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f < \Delta f^*(1,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
+Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
 We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
@@
 \begin{aligned}
-  \Phi &= 2\abs{\Delta f + \overline{\Delta f}} q \, \frac{(1+2\overline{\Delta f}) - \beta(1-2\overline{\Delta f})}
-          {(1+2\overline{\Delta f})^M - \beta^M(1-2\overline{\Delta f})^M}   \,
+  \Phi &= 2\prn{\Delta f + \overline{\Delta f}} q \, \frac{(1-2\overline{\Delta f}) - \beta(1+2\overline{\Delta f})}
+          {(1-2\overline{\Delta f})^M - \beta^M(1+2\overline{\Delta f})^M}   \,
           \brk{\beta (1-2\overline{\Delta f}) (1+2\overline{\Delta f})}^{M/2-1} \\
-       &= 2\abs{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
+       &= 2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
 \end{aligned}
@@
           \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
-          {\abs{(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}}   \\
-       &= \frac{2\abs{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
+          {{(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}}   \\
+       &= \frac{2\prn{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
 \end{aligned}
@@
 \begin{equation}\label{eq:binWTnopre}
-  \Phi = \abs{\Delta f} q,
+  \Phi = {\Delta f}\, q.
 \end{equation}
 %
-where, once again, it's worth remembering that $\Delta f<0$.
 With pre-training:
@@
 \begin{aligned}
-  \Phi &= 2\abs{\Delta f} q.
+  \Phi &= 2{\Delta f}\, q.
 \end{aligned}
@@
 \begin{equation}\label{eq:binKOnopre}
-  \Phi = \frac{2\abs{\Delta f} q}{1+\beta},
+  \Phi = \frac{2{\Delta f}\, q}{1+\beta},
 \end{equation}
@@
 \begin{aligned}
-  \Phi &= 4\abs{\Delta f} q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
+  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
           {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
@@
 \begin{equation}\label{eq:binKOnoprediff}
-  \Phi = \frac{\abs{\Delta f} q}{f\dep+\beta f\pot}.
+  \Phi = \frac{{\Delta f}\, q}{f\dep+\beta f\pot} = \frac{{\Delta f}\, q\pot q\dep}{f\dep q\dep+ f\pot q\pot}.
 \end{equation}
@@
 \begin{equation}\label{eq:multiLinWTnopre}
-  \Phi = 2\abs{\Delta f}q\,\frac{M-1}{M},
+  \Phi = 2{\Delta f}\,q\,\frac{M-1}{M},
 \end{equation}
@@
 \begin{aligned}
-  \Phi &= 4\abs{\Delta f} q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
+  \Phi &= 4{\Delta f}\, q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
           {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
-       &= {4\abs{\Delta f} q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
+       &= {4{\Delta f}\, q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
 \end{aligned}
@@
 \begin{equation}\label{eq:multiLinKOnopre}
-  \Phi = 2\abs{\Delta f} q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
+  \Phi = 2{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
 \end{equation}
@@
 \begin{aligned}
-  \Phi &= 4\abs{\Delta f} q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
-          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M} \\
-       &= 4\abs{\Delta f} q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
+  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f)^{M-1} - \beta^{M-1}(1+2\Delta f)^{M-1}}
+          {(1-2\Delta f)^M - \beta^M(1+2\Delta f)^M} \\
+       &= 4{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
 \end{aligned}
@@
 %
-The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the lowest end of the interval $\Delta f \in \brk{-\half,0}$, where it takes the value
+The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the upper end of the interval $\Delta f \in \brk{0,\half}$, where it takes the value
 %

### 2013-04-30 MHCpaper/MHCsupp.tex
clean up

@@
 We will assume $f\dep\wt=f\dep\ko$.
-This is because the effects of the mutation is well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
+This is because the effects of the mutation are well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
 Therefore the rates of potentiation and depression should not change very much either.
@@
 Here $M=2$, so the leading part of the two effects will cancel.
-The subleading effects are responsible for the faster learning in the mutant.
+The subleading effects (the normalisation of the probabilities) are responsible for the faster learning in the mutant.
 

### 2013-05-10 NIPS/SynapseNIPS.tex
started NIPS paper

--- 2013-02-14 Notes/VORmodel.tex
@@
-% -*- TeX -*- -*- UK -*-
-% ----------------------------------------------------------------
-% arXiv Paper ************************************************
-%
-% Subhaneil Lahiri's template
-%
-% Before submitting:
-%    Comment out hyperref
-%    Comment out showkeys
-%    Replace \input{mydefs.tex} with its contents
-%       or include mydefs.tex in zip/tar file
-%    Replace \input{newsymb.tex} with its contents
-%       or include newsymb.tex in zip/tar file
-%    Put this file, the .bbl file, any picture or
-%       other additional files and natbib.sty
-%       file in a zip/tar file
-%
-% **** -----------------------------------------------------------
-\documentclass[12pt]{article}
-%Preamble:
-\usepackage{a4wide}
+\documentclass{article} % For LaTeX2e
+\usepackage{nips13submit_e,times}
 \input{sl_preamble.tex}
 \input{sl_graphics_preamble.tex}
-\graphicspath{{"Figures/"}}
-%
-\usepackage{array}
-\newcounter{tablerow}
-\newcommand{\tablerowautorefname}{row}
-\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
-%\usepackage{etoolbox}
-%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
-%
+\graphicspath{{"Figs/"}}
 % >> Only for drafts! <<
@@
 \input{sl_symbols.tex}
+%
+%additional symbols:
+%
+\DeclareMathOperator{\SNR}{SNR}
 %matrices
@@
 \newcommand{\MM}{\mathbf{Q}}
-%equilibrium distribution
+%prob distributions
 \newcommand{\pr}{\mathbf{p}}
@@
 \newcommand{\fund}{\mathbf{Z}}
-%other symbols
+%other symbols for matrices
 \newcommand{\Pb}{\mathbf{P}}
@@
 \newcommand{\w}{\mathbf{w}}
+\newcommand{\uv}{\mathbf{u}}
+\newcommand{\vv}{\mathbf{v}}
 \newcommand{\W}{\mathbf{W}}
-\newcommand{\frg}{\W^\mathrm{F}}
 \newcommand{\M}{\mathbf{M}}
+\newcommand{\enc}{\mathbf{q}}
+\newcommand{\frg}{\W^{\mathrm{F}}}
 \newcommand{\F}{\boldsymbol{\Phi}}
-\DeclareMathOperator{\SNR}{SNR}
+%superscripts
+\newcommand{\pot}{^{\text{pot}}}
+\newcommand{\dep}{^{\text{dep}}}
+\newcommand{\potdep}{^{\text{pot/dep}}}
+%sets
 \newcommand{\CS}{\mathcal{S}}
@@
 \newcommand{\comp}{^\mathrm{c}}
-\newcommand{\pot}{^{\text{pot}}}
-\newcommand{\dep}{^{\text{dep}}}
-\newcommand{\potdep}{^{\text{pot/dep}}}
-\newcommand{\norm}{_0}
-\newcommand{\inc}{_{\text{inc}}}
-\newcommand{\dec}{_{\text{dec}}}
-\newcommand{\incdec}{_{\text{inc/dec}}}
-\newcommand{\wt}{_{\text{WT}}}
-\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
-\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
-\newcommand{\tpre}{t_{\text{pre}}}
-\newcommand{\ttrain}{t_{\text{train}}}
-\newcommand{\lmax}{_{\text{max}}}
-\newcommand{\lmin}{_{\text{min}}}
-% ----------------------------------------------------------------
+
 %
@@
 % Title info:
-\title{Models of VOR learning in MHC knockout mice}
+\title{A memory frontier for complex synapses}
 %
@@
 %
-\author{Subhaneil Lahiri
+\author{Subhaneil Lahiri and Surya Ganguli\\
+Applied Physics Department, Stanford University, Stanford CA\\
+\emaillink{sulahiri@stanford.edu}, \emaillink{sulahiri@stanford.edu}
 %
 }
+
+%\nipsfinalcopy % Uncomment for camera-ready version
 
@@
 \begin{abstract}
-  We see if we can model VOR gain increase and decrease learning in mice with a knockout in MHC as well as wild-type.
+  Blah blah blah.
 \end{abstract}
 
-\tableofcontents
-\listoffigures
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Beginning of Article:
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Introduction and summary}\label{sec:intro}
+\section{Introduction}\label{sec:intro}
 
-We will see if we can capture the essential features of VOR gain-increase learning in wild type and MHC \KO\ mutants and wild-type, with and without gain-decrease pre-training.
-We will try to do this using models of complex synapses, such as the cascade model \cite{Fusi2005cascade} or the multistate model \cite{amit1994learning,Fusi2007multistate}.
-We will also construct a new model in \autoref{sec:pooledmodel} that incorporates the idea that biasing the synapses in one direction could deplete some resource, thereby impairing further plasticity in the same direction in nearby synapses.
+\section{Mathematical setup}\label{sec:setup}
 
-These models all assume independent, statistically identical synapses.
-This has the benefit of freeing us from modelling the neurons and their networks, allowing us to focus on the synapses themselves.
-However, we will lose certain aspects of real synapses.
-For example, during gain-increase learning, certain synapses will be potentiated more often and others will be depressed more often.
-In these models, however, synaptic identity is not kept track of, so training can only be described by changing the overall levels of potentiation and depression.
-
-As the MHC \KO\ mutation lowers the threshold for LTD, we will model these mice by increasing the probabilities for internal synaptic transitions associated with depression, whilst leaving potentiation unchanged.
-
-We will be parsimonious in our modelling.
-We will assume that the wild-type has a symmetry between potentiation and depression and we will treat gain-increase and decrease training in a symmetric way.
-We will also assume that the relative rates of potentiation/depression in the mutant and wild-type are the same and that the overall rate of plasticity events is the same in all situations.
-None of these assumptions are necessary or justified experimentally (although \cite{Fusi2007multistate} notes that imbalance between potentiation and depression can drastically impair memory).
-They do reduce the number of adjustable parameters we have, simplifying searches for good models.
-Presumably, relaxing these assumptions and enlarging the parameter space would allow for a better fit to data, but the number of parameters would start to look large compared to the number of features we're trying to match.
-
-If we imagine something like an LN model for the Purkinje cell, the output would depend on some linear combination of the synaptic weights (weighted by the activities of the corresponding parallel fibres).
-If we're not keeping track of synaptic identity, the most natural linear combination to use would be an equal sum of them all.
-Thus we will quantify learning by the change in the mean synaptic weight.
-
-The output of the Purkinje cell, and the behavioural output of the animal, will depend on this quantity in some complicated nonlinear way.
-Modelling this is far beyond the scope of this work.
-Instead we will look at more qualitative features, asking whether learning is slower or faster in some situation compared to another.
-Provided that the nonlinearity is monotonic, these types of questions will not be affected by it.
-
-Within the parameter range explored, we will see that synaptic complexity is essential.
-The simple two-state model always shows enhanced learning in the mutant (see \autoref{fig:binary}), in contradiction with experiment.
-It seems that the enhanced depressing transition more than compensates for saturation effects.
-
-On the other hand, the cascade, multistate and pooled resource models all have parameter ranges where the mutant has impaired learning without pre-training (see \autoref{fig:cascade_short}, \autoref{fig:cascade_long}, \autoref{fig:multistate_weak}, \autoref{fig:multistate_strong} and \autoref{fig:pooled_scarce}).
-We have found parameter ranges where the reverse happens for the multistate (see discussion below \eqref{eq:multiKOnopre}) and the pooled resource model (see \autoref{fig:pooled_plenty}).
-These rely on properties of the mutation, so it would be difficult to explore with further experiments.
-
-All of these models show pre-training enhancing learning for the mutant, as expected.
-It is harder to find regimes where pre-training impairs learning for the wild-type.
-In the cascade model, this required a much longer duration for pe-training (see \autoref{fig:cascade_long}).
-In the multistate model, this required stronger training, which resulted in gain-decrease learning being faster for the mutant then wild-type (see \autoref{fig:multistate_strong}), unlike what is seen experimentally.
-It is possible that these features could be improved by lessening the symmetry between gain-increase and decrease or potentiation and depression, or by allowing the rate of plasticity events to vary.
-For the pooled resource model, we haven't found any parameter regime where pre-training impairs learning for the wild-type, and we discuss why it seems unlikely in \autoref{sec:pooled}.
-
-There seems to be two take home messages.
-First, impaired learning without pre-training in the mutant requires synaptic complexity, as it is not seen in the simple two-state model.
-Second, impaired learning with pre-training in the wild-type requires a form of metaplasticity where repeated potentiation makes subsequent depression less likely.
-This is a feature of the cascade model and our version of the multistate model, but not the pooled resource model.
-
-\section{The setup}\label{sec:setup}
-
-\subsection{Models of synapses}\label{sec:synapse}
+We use a well established formalism for the study of learning and memory with complex synapses (see \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}).
+In this approach, potentiating and depressing plasticity events occur at random times, with all information about the neural activity and learning rules responsible for them absorbed into their rates.
 
@@
 \begin{itemize}
-  \item There are $N$ identical synapses with $M$ internal functional states.
-  \item The states of different synapses are independent of each other.
-  \item The synapses that are eligible for plasticity are chosen randomly.
-  \item The potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $rf\potdep$, where $f\pot+f\dep=1$.
-      In other words, $r$ is the total rate of plasticity events per synapse and $f\potdep$ are the fraction of these events that are potentiating/depressing.
-  \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
-  \item The synaptic weights of the internal states are given by the column vector $\w$.
-This can only take two values that we can call $\pm1$, except for the pooled resource model that we will discuss in \autoref{sec:pooledmodel} and \autoref{sec:pooled}.
+  \item There are $N$ identical synapses with $M$ internal functional states each.
+  \item There are no spatial or temporal correlations in the pattern of potentiating and depressing event.
+  \item There are no correlations in the states of different synapse.
+  \item Plasticity events occur at random (Poisson distributed) times at rate $r$.
+  \item The fraction of these that are potentiating or depressing are given by $f\pot$ and $f\dep$ respectively.
 \end{itemize}
 
-The independence and identicalness of synapses means that the state of the system can be completely described by the probability distribution of the internal states, the row vector $\pr(t)$.
+\section{Upper bounds}\label{sec:bounds}
 
-The evolution of this probability is described by a forgetting matrix, $\frg$:
-%
-\begin{equation}\label{eq:evolve}
-  \diff{\pr(t)}{t} = r\pr(t)\frg,
-  \qquad
-  \frg = f\pot\M\pot+f\dep\M\dep-\I,
-\end{equation}
-where $\I$ is the identity matrix.
-%
-Eventually, this will settle into the equilibrium distribution:
-%
-\begin{equation}\label{eq:eqprob}
-  \eq\frg=0.
-\end{equation}
-%
+\section{Memory curve envelope}\label{sec:env}
 
-With only two possible synaptic weights, the distribution of synaptic weights is completely described by the mean, $\pr(t)\w$.
+\section{Discussion}\label{sec:disc}
 
-\subsection{Model of VOR learning experiment}\label{sec:learning}
-
-Training the animal will not change the internal dynamics of a synapse under potentiation or depression.
-It will change the environment, will lead to a change in how often potentiation and depression occur.
-It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
-We will model this by changing $f\potdep$, leaving $r$ and $\M\potdep$ unchanged.
-
-The untrained case will be described by $f\pot=f\pot\norm$.
-Gain-increase training  will be described by $f\pot=f\pot\inc<f\pot\norm$, and
-gain-decrease training  will be described by $f\pot=f\pot\dec>f\pot\norm$.
-Note that the forgetting matrix \eqref{eq:evolve} and the equilibrium distribution \eqref{eq:eqprob} depend on $f\pot$, which we will indicate with subscripts.
-
-Before training, the synaptic distribution will in equilibrium with $f\pot\norm$.
-During gain-increase training, it will evolve according to \eqref{eq:evolve} with $f\pot\inc$:
-%
-\begin{equation}\label{eq:nopre}
-  \pr(t) = \eq\norm \exp\prn{rt\frg\inc}.
-\end{equation}
-%
-On the other hand, if the gain-increase training follows gain-decrease pre-training for some time period, $\tpre$:
-%
-\begin{equation}\label{eq:withpre}
-  \pr(t) = \eq\norm \exp\prn{r\tpre\frg\dec} \exp\prn{r(t-\tpre)\frg\inc}.
-\end{equation}
-%
-
-We will describe the effect of training by the decrease in mean synaptic weight:
-%
-\begin{equation}\label{eq:learning}
-  L(t) = \prn{\pr(0)-\pr(t)}\w.
-\end{equation}
-%
-The behavioural output (VOR gain) will be some non-linear function of the synaptic weights, so the best we can hope for is to reproduce qualitative features of the experiment.
-
-The MHC mutant has a lower threshold for depression.
-We can model this by changing $\W\dep\wt$ to $\W\dep\ko$, which should have larger matrix elements.
-
-QUESTION: Should $f\pot\wt=f\pot\ko$?
-
-If they are equal, this would change the mean synaptic weight in equilibrium.
-This seems like it could affect the ability of the network to perform its function, whereas, in reality, the wild-type and mutant have the same baseline performance.
-In any case, if the synaptic weights are different, the electrical activity will be different, and there will be no reason to expect the same statistics for potentiation or depression.
-In setting $f\pot\wt=f\pot\ko$, we are assuming that the rest of the brain compensates for the difference in synaptic weights somehow.
-
-On the other hand, one could imagine adjusting $f\pot\wt$ and $f\pot\ko$ so that $\eq\wt\w = \eq\ko\w$.
-But, now that the synaptic weights are the same, the electrical activity will be the same, and there will be no reason to expect different statistics for potentiation or depression.
-If we did this for the multistate or two-state models, we'd remove all differences between the wild-type and \KO\ mutant, so we won't.
-
-We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
-
-\section{Simulations}\label{sec:sims}
-
-\subsection{Models and parameters}
-
-We will look at three different models, the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the multistate model\footnote{The version of the multistate model that we'll study is slightly different than the original definition.
-We will only allow the synaptic weight to take two values, whereas it originally varied linearly across the states.
-} (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}) and the two-state model (which can be thought of as a special case of the previous two, see \autoref{fig:models}\ref{fig:binary_model}).
-We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.
-
-For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
-We will use a larger value for $x$ for depression in the \KO\ mutant.
-
-For the multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
-We will use a larger value for $q$ for depression in the \KO\ mutant.
-
-In each case, we set $f\pot\norm=\half$, $f\pot\inc=f\pot\norm+\Delta f$ and $f\pot\dec=f\pot\norm-\Delta f$.
-We use the same values for wild-type and mutant and treat gain-increase and decrease symmetrically.
-The values of all these parameters are listed in \autoref{tab:params}.
-
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}
- \end{myenuma}
- \end{center}
-  \caption[Transition probabilities for different models]{Transition probabilities for different models.
-  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
-  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
-  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.} \label{fig:models}
-\end{figure}
-
-\begin{table}
- \begin{center}
-  \begin{tabularn}{|l|c|c|c|c|c|c|}
-    \cline{1-7}
-    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
-    \cline{1-7}
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 50  &\label{tr:cascade_short} \\
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:cascade_long} \\
-    Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
-    Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
-    Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
-    OG Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.3  & 20 & 20  &\label{tr:multistate_lin} \\
-    Two-state     & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
-    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
-    Pooled        & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
-                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
-    \cline{1-7}
-%   \hline
-%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
-%    \cline{1-7}
-%    \label{tr:cascade_short}&
-%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
-%    \label{tr:cascade_long}&
-%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
-%    \label{tr:multistate_weak}&
-%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
-%    \label{tr:multistate_strong}&
-%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
-%    \label{tr:binary}&
-%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
-%    \label{tr:pooled_plenty}&
-%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-%                                          & -0.1 & 70 & 70 \\
-%    \label{tr:pooled_scarce}&
-%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
-%                                          & -0.1 & 70 & 70 \\
-%    \cline{1-7}
-  \end{tabularn}
- \end{center}
-  \caption{Parameters used in simulations.} \label{tab:params}
-\end{table}
-
-\subsubsection{Pooled resource model}\label{sec:pooledmodel}
-
-Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed.
-We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single compound synapse.
-
-We will model the individual synapses with the two-state model.
-Let $i=0\ldots P$ be the number of them that are potentiated.
-We will model the effect of resource depletion linearly with the potentiation/depression probabilities for the individual synapses:
-%
-\begin{equation}\label{eq:depletion}
-  \begin{aligned}
-    q\pot &= \frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}, \quad& i &= 0 \ldots P-1,\\
-    q\dep &= \frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}, & i &= 1 \ldots P.
-  \end{aligned}
-\end{equation}
-%
-At each plasticity event for the compound synapse, one of the individual synapses will be chosen randomly for update.
-This effectively reduces the transition probabilities by $1/P$.
-
-This compound synapse would seem to have $2^P$ internal states.
-However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
-The transition network topology will then have a multistate topology (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
-%
-\begin{equation}\label{eq:pooledweight}
-  \w_i = \frac{2i}{P}-1.
-\end{equation}
-%
-
-The Markov process is lumpable \wrt this partition of states (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$.
-The result must be the same for all states in lump $i$.
-
-For any state in lump $i$, there are $P-i$ synapses that can be potentiated to go to lump $i+1$.
-Each of these transition probabilities is $q\pot/P$.
-Similarly, there are $i$ synapses that can be depressed to go to lump $i-1$.
-Each of these transition probabilities is $q\dep/P$.
-Thus:
-%
-\begin{equation}\label{eq:pooledpotdep}
-  \begin{aligned}
-    \M\pot_{ii+1} &=  \brk{\frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}} \frac{P-i}{P}, \\
-    \M\dep_{ii-1} &=  \brk{\frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}} \frac{i}{P},
-  \end{aligned}
-\end{equation}
-%
-with all other off-diagonal elements equal to zero.
-The diagonal elements are chosen so that the rows sum to one.
-
-This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
-We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
-We will use larger values for depression in the \KO\ mutant.
-
-\subsection{Results}\label{sec:results}
-
-The features of the actual experiments that we'd like to capture are:
-\begin{itemize}
-  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the mutant.
-  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.
-  \item After pre-training, gain-increase learning is significantly faster in the mutant than in the wild-type.
-  \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
-  \item For the mutant, gain-increase learning is significantly faster with pre-training than without it.
-\end{itemize}
-
-We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
-This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measuring using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
-This should be multiplied by the difference in synaptic weights
-But, as this quantity is constant within a model and doesn't change when comparing wild-type and mutant or with/without pre-training, it can be ignored.
-
-\newcommand{\resultsfig}[3]{\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{#3_learn.eps}}\label{fig:#3_learn}
-  \item\aligntop{\includegraphics[width=7cm]{#3_learnS.eps}}\label{fig:#3_learnS}
-  \item\label{fig:#3_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_WT.eps}}\label{fig:#3_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_KO.eps}}\label{fig:#3_eq_KO}
-                  \end{myenumi}
-  \item\label{fig:#3_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_nopre.eps}}\label{fig:#3_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_pre.eps}}\label{fig:#3_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_nopre.eps}}\label{fig:#3_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_pre.eps}}\label{fig:#3_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}
-  \caption[#1]{#1 #2.
-  Other parameters can be found in \autoref{tr:#3} of \autoref{tab:params}.
-  Time is measured in units of $1/r$, the mean time between plasticity events.
-  (\ref{fig:#3_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:#3_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:#3_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#3_eq_WT}) wild-type and (\ref{fig:#3_eq_KO}) \KO\ mutant,
-  with states of weak synaptic strength to the left and strong states to the right.
-  (\ref{fig:#3_pr}) Evolution of probability distributions for (\ref{fig:#3_pr_WT_nopre}\ref{fig:#3_pr_WT_pre}) wild-type and  (\ref{fig:#3_pr_KO_nopre}\ref{fig:#3_pr_KO_pre}) \KO\ mutant without (\ref{fig:#3_pr_WT_nopre},\ref{fig:#3_pr_KO_nopre}) and with (\ref{fig:#3_pr_WT_pre},\ref{fig:#3_pr_KO_pre}) pre-training,
-  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#3}
-\end{figure}}
-
-\newcommand{\resultsfign}[2]{\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{#2_learn.eps}}\label{fig:#2_learn}
-  \item\aligntop{\includegraphics[width=7cm]{#2_learnS.eps}}\label{fig:#2_learnS}
-  \item\label{fig:#2_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_WT.eps}}\label{fig:#2_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_KO.eps}}\label{fig:#2_eq_KO}
-                  \end{myenumi}
-  \item\label{fig:#2_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_nopre.eps}}\label{fig:#2_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_pre.eps}}\label{fig:#2_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_nopre.eps}}\label{fig:#2_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_pre.eps}}\label{fig:#2_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}
-  \caption[#1]{#1.
-  Parameters can be found in \autoref{tr:#2} of \autoref{tab:params}.
-  Time is measured in units of $1/r$, the mean time between plasticity events.
-  (\ref{fig:#2_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
-  (\ref{fig:#2_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:#2_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#2_eq_WT}) wild-type and (\ref{fig:#2_eq_KO}) \KO\ mutant,
-  with states of weak synaptic strength to the left and strong states to the right.
-  (\ref{fig:#2_pr}) Evolution of probability distributions for (\ref{fig:#2_pr_WT_nopre}\ref{fig:#2_pr_WT_pre}) wild-type and  (\ref{fig:#2_pr_KO_nopre}\ref{fig:#2_pr_KO_pre}) \KO\ mutant without (\ref{fig:#2_pr_WT_nopre},\ref{fig:#2_pr_KO_nopre}) and with (\ref{fig:#2_pr_WT_pre},\ref{fig:#2_pr_KO_pre}) pre-training,
-  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#2}
-\end{figure}}
-
-\subsubsection{Cascade model}\label{sec:cascade}
-
-\resultsfig{Simulation results for the cascade model with short pre-training}{($rt_\text{pre}=50$)}{cascade_short}
-
-\resultsfig{Simulation results for the cascade model with long pre-training}{($rt_\text{pre}=150$)}{cascade_long}
-
-The results of simulations of the cascade model can be seen in \autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
-It is difficult to study this model analytically.
-However, the numerical results can helps us understand it qualitatively.
-
-In both of these, we see that the mutant is slower than wild-type without pre-training but faster with it.
-This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
-With pre-training, some of them will now be potentiated (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_pre}), and the enhanced depression can speed up learning.
-
-With shorter pre-training, we see that it speeds up learning in both wild-type and mutant (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the mutant.
-This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning.
-
-We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_learnS} that longer pre-training pushes the synapses further down the cascade, slow in down learning for the wild-type.
-This effect is weaker for the mutant, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
-
-\subsubsection{Multistate model}\label{sec:multistate}
-
-\resultsfig{Simulation results for the multistate model with weak training}{($\Delta f=-0.1$)}{multistate_weak}
-
-\resultsfig{Simulation results for the multistate model with moderate training}{($\Delta f=-0.17$)}{multistate_med}
-
-\resultsfig{Simulation results for the multistate model with strong training}{($\Delta f=-0.4$)}{multistate_strong}
-
-The results of simulations of the multistate model can be seen in \autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
-However, we can get some insight into this model analytically.
-
-Consider the general uniform multistate model.
-Then the equilibrium distribution is given by
-%
-\begin{equation}\label{eq:mutltieq}
-  \eq_i = \frac{1-\alpha}{1-\alpha^M}\,\alpha^{i-1},
-  \qquad \text{where} \quad
-  \alpha=\frac{f\pot q\pot}{f\dep q\dep}.
-\end{equation}
-%
-If we take the limit $\alpha\rightarrow1$, this becomes $\frac{1}{M}$.
-
-The net-flux from the $\w=+1$ states to the $\w=-1$ states is:
-%
-\begin{equation}\label{eq:multiflux}
-  \Phi = \eq_{M/2+1}f'{}\dep q\dep - \eq_{M/2}f'{}\pot q\pot = \frac{1-\alpha}{1-\alpha^M}\,\alpha^{M/2-1}\prn{\alpha-\alpha'}f'{}\dep q\dep,
-\end{equation}
-%
-where primed values correspond to the new value of $f\pot$.
-
-First, consider the wild-type, for which $q\pot=q\dep=q$.
-Without pre-training:
-%
-\begin{equation}\label{eq:multiWTnopre}
-  \Phi = -\frac{2\Delta fq}{M},
-\end{equation}
-%
-where it's worth remembering that $\Delta f<0$.
-With pre-training:
-%
-\begin{equation}\label{eq:multiWTpre}
-\begin{aligned}
-  \Phi &= 16(\Delta f)^2q \, \frac{(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1}}
-          {(1-2\Delta f)^M - (1+2\Delta f)^M} \\
-       &= -\frac{4\Delta fq}{M} + \CO(\Delta f)^2.
-\end{aligned}
-\end{equation}
-%
-So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS}.
-On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot due to the factor of $(1+2\Delta f)^{M/2-1}$, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.
-
-Intuitively, the flux depends on the slope of the distribution at the centre of the chain (with an offset due to the difference between $f\pot$ and $f\dep$).
-Pre-training has two effects: it produces a slope in the right direction (see \autoref{fig:multistate_weak}\ref{fig:multistate_weak_eq}), but it also reduces distribution at the centre (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}).
-For small $\Delta f$, the first effect is stronger and learning speeds up.
-For larger $\Delta f$, the second effect wins and learning slows down.
-
-Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
-Without pre-training:
-%
-\begin{equation}\label{eq:multiKOnopre}
-  \Phi = -2\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
-\end{equation}
-%
-This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$.
-This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.
-
-There are two effects here as well.
-Smaller $\beta$ will increase the probability of crossing the centre of the chain, speeding up learning, but it will also concentrate probability at the ends of the chain, depleting the centre and slowing down learning.
-The first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ and will be more significant for smaller $\beta$ or in a longer chain.
-
-With pre-training:
-%
-\begin{equation}\label{eq:multiKOpre}
-\begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f) - \beta(1-2\Delta f)}
-          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M}   \,
-          \brk{\beta (1-2\Delta f) (1+2\Delta f)}^{M/2-1} \\
-       &= -4\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
-\end{aligned}
-\end{equation}
-%
-Once again, we see that pre-training will speed up learning when $\Delta f$ is small, whereas, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning.
-
-Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
-Values of $\Delta f$ that are smaller than this (stronger training) will correspond to slower learning after pre-training.
-As we would like pre-training to slow down learning in the wild-type but speed it up in the mutant, it would seem that we require $\Delta f^*(\beta,M) < \Delta f < \Delta f^*(1,M)$ (remembering once more that $\Delta f<0$ and that the wild-type corresponds to $\beta=1$).
-As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) < \Delta f^*(1,M)$, so this range always exists.
-We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.
-
-However, this is only the initial, instantaneous rate of change.
-Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f < \Delta f^*(1,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
-We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}) drifting to the centre.
-This happens sooner for the mutant than the wild-type, due to the stronger depressing transitions.
-This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the mutant, even outside the desired range of values for $\Delta f$.
-As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}, even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.
-
-In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f < \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the mutant learns slower than wild-type without pre-training, and that pre-training speeds up learning in the mutant but slows it down in the wild-type.
-However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}), which is \emph{not} seen in the experiment.
-
-Now consider what would happen if we did not treat gain-increase and gain-decrease symmetrically.
-Let us set $f\pot\dec = f\pot\norm - \overline{\Delta f}$.
-Then, \eqref{eq:multiKOpre} becomes
-%
-\begin{equation}\label{eq:multiKOdiffpre}
-\begin{aligned}
-  \Phi &= -2\prn{\Delta f + \overline{\Delta f}} q \, \frac{(1+2\overline{\Delta f}) - \beta(1-2\overline{\Delta f})}
-          {(1+2\overline{\Delta f})^M - \beta^M(1-2\overline{\Delta f})^M}   \,
-          \brk{\beta (1-2\overline{\Delta f}) (1+2\overline{\Delta f})}^{M/2-1} \\
-       &= -2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
-\end{aligned}
-\end{equation}
-%
-We can then look at the wild-type by setting $\beta=1$:
-%
-\begin{equation}\label{eq:multiWTdiffpre}
-\begin{aligned}
-  \Phi &= -8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \,
-          \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
-          {(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}   \\
-       &= -\frac{2\prn{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
-\end{aligned}
-\end{equation}
-%
-
-We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the mutant, the important thing is to have \emph{strong pre-training}, not strong training.
-Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the mutant than wild-type.
-
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
- \end{myenuma}
- \end{center}
-  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
-\end{figure}
-
-\subsubsection{OG multistate model}\label{sec:multistate_lin}
-
-\resultsfign{Simulation results for multistate model with linear weights}{multistate_lin}
-
-In this section, we will consider the multistate model as originally defined in \cite{amit1994learning}, \ie with linearly varying synaptic weight:
-%
-\begin{equation}\label{eq:multistateLinWeight}
-  \w_i = \frac{2i-M-1}{M-1}.
-\end{equation}
-%
-The numerical results can be seen in \autoref{fig:multistate_lin}, but we can get some analytic insight into this model as well.
-
-The equilibrium distribution, \eqref{eq:mutltieq}, still applies.
-However, now the rate of change of our learning metric, \eqref{eq:learning}, will be proportional to the sum of the net fluxes between adjacent states:
-%
-\begin{equation}\label{eq:multiLinFlux}
-  \begin{aligned}
-    \Phi &= \sum_{i=1}^{M-1} \eq_{i+1} f'{}\dep q\dep - \eq_i f'{}\pot q\pot
-         &= \sum_{i=1}^{M-1} \eq_i (\alpha-\alpha') f'{}\dep q\dep \\
-         &= \frac{1-\alpha^{M-1}}{1-\alpha^M} \, (\alpha-\alpha') f'{}\dep q\dep.
-  \end{aligned}
-\end{equation}
-%
-
-First, consider the wild-type, for which $q\pot=q\dep=q$.
-Without pre-training:
-%
-\begin{equation}\label{eq:multiLinWTnopre}
-  \Phi = -2\Delta fq\,\frac{M-1}{M},
-\end{equation}
-%
-where, once more, we should note that $\Delta f<0$.
-With pre-training:
-%
-\begin{equation}\label{eq:multiLinWTpre}
-\begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
-          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
-       &= -{4\Delta f q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
-\end{aligned}
-\end{equation}
-%
-
-Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
-Without pre-training:
-%
-\begin{equation}\label{eq:multiLinKOnopre}
-  \Phi = -2\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
-\end{equation}
-%
-This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
-This means that the mutant will initially learn faster than the wild-type.
-However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}
-
-With pre-training:
-%
-\begin{equation}\label{eq:multiLinKOpre}
-\begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
-          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M} \\
-       &= -4\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
-\end{aligned}
-\end{equation}
-%
-The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the lowest end of the interval $\Delta f \in \brk{-\half,0}$, where it takes the value
-%
-\begin{equation}\label{eq:multiLinprenopre}
-  \frac{\Phi_{\text{with pre}}}{\Phi_{\text{without pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
-   \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,
-\end{equation}
-%
-which is always greater than 1.
-Therefore, pre-training will always enhance learning, for both mutant and wild-type.
-
-\subsubsection{Two-state model}\label{sec:binary}
-
-\resultsfign{Simulation results for two-state model}{binary}
-
-The results of simulations of the two-state model can be seen in \autoref{fig:binary}.
-However, this model can be solved exactly:
-%
-\begin{multline}\label{eq:binarysol}
-  \eq = \frac{(f\dep q\dep, f\pot q\pot)}{\lambda},
-  \qquad
-  \pr(t) = \eq + (\pr(t)-\eq)\e^{-\lambda rt},\\
-  \qquad \text{where} \quad
-  \lambda = f\pot q\pot + f\dep q\dep.
-\end{multline}
-%
-But it is easier to just substitute $M=2$ into the formulae in \autoref{sec:multistate}.
-In this case, the initial rate of change and the total change encapsulate the whole solution, as there is only a single exponential decay.
-
-First, consider the wild-type, for which $q\pot=q\dep=q$.
-Without pre-training:
-%
-\begin{equation}\label{eq:binWTnopre}
-  \Phi = -\Delta f q,
-\end{equation}
-%
-where, once again, it's worth remembering that $\Delta f<0$.
-With pre-training:
-%
-\begin{equation}\label{eq:binWTpre}
-\begin{aligned}
-  \Phi &= -2\Delta f q.
-\end{aligned}
-\end{equation}
-%
-So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
-This can be seen in \autoref{fig:binary}\ref{fig:binary_learnS}.
-
-Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
-Without pre-training:
-%
-\begin{equation}\label{eq:binKOnopre}
-  \Phi = -\frac{2\Delta f q}{1+\beta},
-\end{equation}
-%
-which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
-This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS}.
-As discussed below \eqref{eq:multiKOnopre}, the smaller values of has two effects.
-Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
-The second effect is much smaller for the very short chain considered here.
-
-With pre-training:
-%
-\begin{equation}\label{eq:binKOpre}
-\begin{aligned}
-  \Phi &= -4\Delta f q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
-          {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
-\end{aligned}
-\end{equation}
-%
-As we've already ruled out this model, we won't waste any time analysing this formula.
-
-We made a number of assumptions in our analysis so far.
-Before we declare this model to be falsified, we should think about what would happen if we relaxed these assumptions and explored a larger parameter space.
-
-First, we assumed that the wild type had symmetric potentiation and depression.
-Relaxing this assumption would correspond modelling the wild-type like the mutant, but with a different value for $\beta$.
-We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the mutant.
-
-Secondly, we assumed that there would be equal rates of potentiation and depression in the untrained situation.
-However, relaxing this assumption merely changes \eqref{eq:binKOnopre} to
-%
-\begin{equation}\label{eq:binKOnoprediff}
-  \Phi = -\frac{\Delta f q}{f\dep+\beta f\pot}.
-\end{equation}
-%
-This will not change our conclusions either.
-
-As we didn't need the pre-training paradigm to rule out this model, we don't need to look at removing the symmetry between gain-increase and decrease.
-
-The only thing we have to worry about is when the wild-type catches up with the mutant.
-As the total change will be larger for the wild-type than mutant, it must overtake eventually.
-If this happens early enough, the initial period, where the mutant learns faster than wild-type, would not be seen in the experiment.
-However, the timescale for this crossover would be similar to the timescale of the exponentials.
-In the data, it looks like that timescale is longer than the gaps between successive measurements, so we needn't worry about this.
-
-\subsubsection{Pooled resource model}\label{sec:pooled}
-
-\resultsfig{Simulation results for the pooled resource model with light depletion}{($q\lmin/q\lmax=0.75$)}{pooled_plenty}
-
-\resultsfig{Simulation results for the pooled resource model with heavy depletion}{($q\lmin/q\lmax=0.25$)}{pooled_scarce}
-
-The results of simulations of the pooled resource model can be seen in \autoref{fig:pooled_plenty} and \autoref{fig:pooled_scarce}.
-It is difficult to study this model analytically.
-However, the numerical results can helps us understand it qualitatively.
-
-If we compare gain-increase learning in the mutant to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_eq}: \ref{fig:pooled_plenty_eq_WT} and \ref{fig:pooled_plenty_eq_KO} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_eq}: \ref{fig:pooled_scarce_eq_WT} and \ref{fig:pooled_scarce_eq_KO}), where there are fewer synapses available for depression and resources are depleted.
-When resource depletion is less severe, the first effect dominates and the mutant learns faster than wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learn},\ref{fig:pooled_plenty_learnS}).
-This is reasonable, as when resource depletion is removed it will reduce to the two-state model, which showed this behaviour.
-However, when resource depletion is more severe, the second effect dominates and the mutant learns slower than wild-type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learn},\ref{fig:pooled_scarce_learnS}), which matches what is seen in the experiment.
-
-Gain-decrease pre-training lessens the second effect, and results in the mutant learning faster than wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learn},\ref{fig:pooled_plenty_learnS} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learn},\ref{fig:pooled_scarce_learnS}), as seen in the experiment.
-
-Gain-decrease pre-training will shift the distribution to the potentiated site, where there are more synapses available for depression and resources are more plentiful, for both the mutant and wild type.
-This means that the pre-trained animal will learn faster then the untrained one for both mutant and wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learnS} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learnS}).
-This differs from what is seen experimentally, where the pre-trained wild-type learns slower than the untrained one.
-
-%\subsection*{Acknowledgements}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection*{Appendices}
-%\appendix
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection*{Acknowledgements}
 

### 2013-05-14 NIPS/SynapseNIPS.tex
wrote first pass of setup section

@@
 \DeclareMathOperator{\SNR}{SNR}
+\DeclareMathOperator{\snr}{SNR}
+\newcommand{\wv}{\vec{w}}
+\newcommand{\wvi}{\vec{w}_\text{ideal}}
 %matrices
@@
 \newcommand{\w}{\mathbf{w}}
-\newcommand{\uv}{\mathbf{u}}
-\newcommand{\vv}{\mathbf{v}}
 \newcommand{\W}{\mathbf{W}}
@@
 \newcommand{\comp}{^\mathrm{c}}
+%eigenmodes
+\newcommand{\uv}{\mathbf{u}}
+\newcommand{\vv}{\mathbf{v}}
+\newcommand{\CI}{\mathcal{I}}
 
@@
 In this approach, potentiating and depressing plasticity events occur at random times, with all information about the neural activity and learning rules responsible for them absorbed into their rates.
+We assume that there are no spatial or temporal correlations in the pattern of potentiating and depressing events, and these events cause Markovian transitions between the internal states of the synapses.
+As a result of these assumptions, the states of different synapses will be independent, and the system can be fully described by the probability distribution across these states, which we will indicate with the row-vector $\pr(t)$.
 
-We make the following assumptions:
-\begin{itemize}
-  \item There are $N$ identical synapses with $M$ internal functional states each.
-  \item There are no spatial or temporal correlations in the pattern of potentiating and depressing event.
-  \item There are no correlations in the states of different synapse.
-  \item Plasticity events occur at random (Poisson distributed) times at rate $r$.
-  \item The fraction of these that are potentiating or depressing are given by $f\pot$ and $f\dep$ respectively.
-\end{itemize}
+We also employ an ``ideal observer'' approach to the memory readout, where the synaptic weights are read directly.
+This provides an upper bound to the quality of any readout using neural activity.
+These synaptic weights will be restricted to two values, which we can shift and scale to $\pm1$.
+
+For any single memory, there will be an ideal pattern of synaptic weights, the $N$-element vector $\wvi$, that is $+1$ at all synapses that were potentiated and $-1$ at all synapses that were depressed.
+The actual pattern of synaptic weights at some later time, $t$, will be $\wv(t)$.
+We can use the overlap between these, $\wvi\cdt\wv(t)$, as a measure of the quality of the memory.
+As $t\to\infty$, the system will return to its steady state distribution which will be uncorrelated with the memory in question.
+The probability distribution of the quantity $\wvi\cdt\wv(\infty)$ can be used as a ``null model'' for comparison.
+
+The extent to which the memory has been stored is described by a signal-to-noise ratio (SNR):
+%
+\begin{equation}\label{eq:SNRdef}
+  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t) - \wv_\text{ideal}\cdt\wv(\infty)}}
+     {\sqrt{\var(\wv_\text{ideal}\cdt\wv(\infty))}}.
+\end{equation}
+%
+The noise is essentially $\sqrt{N}$. 
+There is a correction when potentiation and depression are imbalanced, but this will not affect the upper bounds that we will discuss below and will be ignored in the subsequent formulae.
+
+All of the preceding plasticity events will put the system in its steady-state distribution, $\eq$.
+The memory we are tracking at $t=0$ will change this to $\eq\M\pot$ in those synapses that are potentiated and $\eq\M\dep$ in those synapses that are depressed, where $\M\potdep$ are $M\times M$ matrices of transition probabilities for the Markov processes describing potentiation and depression.
+As the potentiating/depressing nature of the subsequent memories is independent of $\wvi$, we can average over all sequences, resulting in the evolution of the probability distribution:
+%
+\begin{equation}\label{eq:evol}
+  \diff{\pr(t)}{t} = r\pr(t)\frg,
+  \qquad \text{where} \quad
+  \frg = f\pot\M\pot + f\dep\M\dep - \I.
+\end{equation}
+%
+Here, $r$ is the rate of the Poisson process describing the timing of plasticity events,
+$f\potdep$ is the fraction of events that are potentiating/depressing and $\I$ is the identity matrix.
+
+This results in the following SNR
+%
+\begin{equation}\label{eq:SNRcalc}
+  \snr(t) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \e^{rt\frg} \w.
+\end{equation}
+%
+We will frequently refer to this function as the memory curve.
+
+The parameters must satisfy the following constraints:
+%
+\begin{equation}\label{eq:constr}
+\begin{aligned}
+  \M\potdep_{ij} &\in [0,1], &\quad
+  f\potdep &\in [0,1], &\quad
+  \eq\frg &= 0, &\quad
+  \w_i &= \pm 1, \\
+  \sum_j \M\potdep_{ij} &= 1, &
+  f\pot + f\dep &= 1, &
+  \sum_i \eq_i &= 1.
+\end{aligned}
+\end{equation}
+%
+The upper bounds on $\M\potdep_{ij}$ and $f\potdep$ follow automatically from the other constraints.
+
+The question is: what do these constraints imply for the memory curve above?
+In practice, to make any statements about finite times, we need to got to the eigenmode description:
+%
+\begin{equation}\label{eq:eigendecomp}
+  \frg = \sum_a -q_a \uv^a \vv^a,
+  \quad
+  \vv^a \uv^b = \delta_{ab},
+  \quad
+  \frg \uv^a = -q_a \uv^a,
+  \quad
+  \vv^a \frg = -q_a \vv^a.
+\end{equation}
+%
+Where $q_a$ are the negative of the eigenvalues, $\uv^a$ are the right (column) eigenvectors and $\vv^a$ are the left (row) eigenvectors.
+This allow us to write
+%
+\begin{equation}\label{eq:SNReigen}
+\begin{aligned}
+  \snr(t) &= \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a},
+  &\quad \text{where}&\;&
+  \CI_a &= \prn{2 f\pot f\dep} \eq (\M\pot-\M\dep) \uv^a  \vv^a \w,\\&
+  & \text{and}&&
+  \tau_a &= \frac{1}{q_a}.
+\end{aligned}
+\end{equation}
+%
+We can ask then ask the question: what are the constraints on these quantities implied by the constraints \eqref{eq:constr}?
+We will find some of these constraints in the next section.
 

### 2013-05-14 NIPS/SynapseNIPS.tex
wrote initial SNR section

@@
 
-\section{Mathematical setup}\label{sec:setup}
+\section{Synaptic models and their memory curves}\label{sec:setup}
+
+In this section, we will describe the class of models of synaptic plasticity that we are studying and how we quantify their performance in memory storage.
+In the subsequent sections, we will find upper bounds on this performance.
 
@@
 \begin{equation}\label{eq:SNRcalc}
-  \snr(t) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \e^{rt\frg} \w.
+  \snr(t) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \e^{rt\frg} \w,
 \end{equation}
 %
+where the weight of the synapse when it is in its $i$'th state is given by the corresponding element of the column vector $\w$.
 We will frequently refer to this function as the memory curve.
@@
 
+In this section, we will find some upper bounds on certain properties of the memory curve discussed in the previous section, namely its initial value and the area under it.
+We will use these bounds to determine an upper bound on the SNR at finite times in the next section.
+
+\subsection{Initial SNR}\label{sec:initial}
+
+Now we will discuss the SNR at $t=0$:
+%
+\begin{equation}\label{eq:init}
+  \snr(0) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \w.
+\end{equation}
+%
+We will find an upper bound on this quantity for \emph{all} possible models and also find the model that saturates this bound.
+
+A useful quantity is the equilibrium probability flux between two disjoint sets of states, $\CA$ and $\CB$:
+%
+\begin{equation}\label{eq:flux}
+  \F_{\CA\CB} = \sum_{i\in\CA} \sum_{j\in\CB} r \eq_i \frg_{ij}.
+\end{equation}
+%
+The initial SNR is closely related to the flux between the states with $\w_i=-1$ and those with $\w_j=+1$:
+%
+\begin{equation}\label{eq:initflux}
+  \snr(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}\,.
+\end{equation}
+%
+This inequality becomes an equality if potentiation never decreases the synaptic weight and depression never increases it, which should be a property of any sensible model.
+
+To maximise this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
+AN example of such a model is shown in \autoref{fig:maxinit}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
+These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
+They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \autoref{fig:maxinit}(\ref{fig:binary_det}).
+
+This two state model has the equilibrium distribution $\eq=(f\dep,f\pot)$ and its flux is given by $\F_{-+} = rf\pot f\dep$. 
+This is maximized when $f\pot=f\dep=\half$, leading to the upper bound:
+%
+\begin{equation}\label{eq:maxinit}
+  \snr(0) \leq \sqrt{N}.
+\end{equation}
+%
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\hp\aligntop{\includegraphics[width=0.25\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
+  \item\aligntop{\includegraphics[width=0.25\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
+  \item\aligntop{\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
+  \end{myenuma}
+ \end{center}
+  \caption{Synaptic models that maximize initial SNR. 
+  (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
+  (\ref{fig:max_init_pot}) Depression is similar to potentiation, but with strong and weak interchanged. 
+  (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
+  }\label{fig:maxinit}
+\end{figure}
+
+\subsection{Area}\label{sec:area}
+
+Now consider the area under the memory curve:
+%
+\begin{equation}\label{eq:area}
+  A = \int_0^\infty\!\!\dr t \, \snr(t).
+\end{equation}
+%
+We will find an upper bound on this quantity as well as the model that saturates it.
+
 \section{Memory curve envelope}\label{sec:env}

### 2013-05-15 NIPS/SynapseNIPS.tex
wrote area section

@@
+% -*- TeX -*- -*- US -*-
 \documentclass{article} % For LaTeX2e
@@
 We will find an upper bound on this quantity for \emph{all} possible models and also find the model that saturates this bound.
+As maximizing initial SNR does not place any value on later parts of the memory curve, this will produce models with fast decay that may not be good for memory storage.
+Nevertheless, the upper bound will be useful in \autoref{sec:env}.
 
@@
 %
+Whilst this model has high initial SNR, it also has very fast decay -- with a timescale $\tau\sim\frac{1}{r}$.
+As the synapse is very plastic, the initial memory is encoded very easily, but the subsequent memories also overwrite it easily.
+This is not a good design for a synapse, but the resulting bound on the initial SNR will prove useful in \autoref{sec:env}.
 
@@
  \begin{myenuma}
-  \item\hp\aligntop{\includegraphics[width=0.25\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
-  \item\aligntop{\includegraphics[width=0.25\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
-  \item\aligntop{\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
+  \item\hp\aligntop{\includegraphics[width=0.15\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
+  \item\aligntop{\includegraphics[width=0.15\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
+  \item\aligntop{\includegraphics[width=0.05\linewidth]{binary_det.svg}}\label{fig:binary_det}
   \end{myenuma}
@@
 We will find an upper bound on this quantity as well as the model that saturates it.
+The area is useful as it also provides an upper bound on the lifetime of a memory, $\tau(\epsilon)$, defined as the time at which $\snr(\tau)=\epsilon$:
+%
+\begin{equation}\label{eq:arealife}
+  A = \int_0^\infty\!\!\dr t \, \snr(t) 
+    > \int_0^{\tau(\epsilon)}\!\!\dr t \, \snr(t)
+    > \int_0^{\tau(\epsilon)}\!\!\dr t \, \epsilon
+    = \epsilon \tau(\epsilon).
+\end{equation}
+%
+However, the area still receives a contribution from the tail of the memory curve, after the SNR has decayed below any useful value.
+This means that maximizing SNR will favor models with long tails that will not necessarily be good for storing memories.
+
+One important property of the area is that it has a null ``scaling'' degree of freedom.
+If all off-diagonal elements of $\M\potdep$ are multiplied by a constant, with diagonal elements adjusted to keep row-sums fixed, this will have two effects.
+First, it will scale up the initial signal creation.
+Second, it will scale down time.
+Combining these two effects, it will leave the area invariant.
+
+This feature allows us to ignore the lower bound on the diagonal elements, $\M\potdep_{ii}>0$.
+If we maximize the area ignoring this constraint, the ``scaling'' mode can be used to satisfy it without changing the area.
+
+We also develop an important organizing principle using the theory of first passage times.
+The mean first passage time matrix, $\fptb_{ij}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $j$.
+The diagonal elements are defined to be zero.
+Then the quantity
+%
+\begin{equation}\label{eq:kemenyconst}
+  \eta \equiv \sum_j \fptb_{ij}\eq_j
+\end{equation}
+%
+is independent of the starting state $i$.
+It is known as Kemeny's constant (see \cite[\S4.4]{kemeny1960finite}).
+We can define analogous quantities
+%
+\begin{equation}\label{eq:kemenypm}
+   \eta^+_i = \sum_{j\in+} \fptb_{ij} \eq_j,
+   \qquad
+   \eta^-_i = \sum_{j\in-} \fptb_{ij} \eq_j.
+\end{equation}
+%
+These can be thought of as the average time it takes to reach the strong/weak states respectively.
+We can put the states in order of decreasing $\eta^+_i$ or increasing $\eta^-_i$.
+Because $\eta^+_i+\eta^-_i=\eta$ is independent of $i$, the two orderings are the same.
+
+With the states in this order, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
+Some of these perturbations are shown in \autoref{fig:perts}.
+The only thing that can prevent these perturbations from increasing the area is when they require the decrease of a matrix element that has already been set to 0.
+This determines the topology (non-zero transition probabilities) of the model with maximal area.
+It is of the form shown in \autoref{fig:max_area}, with potentiation moving one step to the right and depression moving one step to the left.
+Any other topology would allow some perturbation that could increase the area.
+
+As these perturbations do not change the equilibrium distribution, this means that the area of \emph{any} model is bounded by that of a linear chain with the same equilibrium distribution.
+This area is given by 
+%
+\begin{equation}\label{eq:multistatearea}
+  A \leq \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k
+    = \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
+\end{equation}
+%
+where we chose $\w_k=\sgn\brk{k - \sum_j j\eq_j}$.
+We can then maximize this by pushing all of the equilibrium distribution symmetrically to the two end states.
+This can be done by reducing the transition probabilities out of these states, as in \autoref{fig:max_area}.
+This makes it very difficult to get out of the end states once they have been entered.
+The resulting area is
+%
+\begin{equation}\label{eq:max_area}
+  A \leq \frac{\sqrt{N}(M-1)}{r}.
+\end{equation}
+%
+The ``sticky'' end states result in very slow decay, but they also make it difficult to encode the memory in the first place.
+The memory curve for this model starts off very low, but hovers above zero for a very long time.
+This is also not a good design for a synapse, but the resulting bound on the area will also prove useful in \autoref{sec:env}.
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
+  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
+  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut.svg}}\label{fig:shortcut}
+ \end{myenuma}
+ \end{center}
+  \caption{Perturbations that increase the area.
+  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$.
+It can no longer be used when $\M\dep$ is lower triangular.
+  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$.
+It can no longer be used when $\M\pot$ is upper triangular.
+  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions.
+It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
+  }\label{fig:perts}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+  \includegraphics[width=0.4\linewidth]{multistate_sticky.svg}
+ \end{center}
+  \caption{Model that maximizes the area. 
+  Unlabeled transitions have probability 1. 
+  Labeled transitions have probability $\epsilon\to0$.
+  }\label{fig:max_area}
+\end{figure}
 

### 2013-05-15 NIPS/SynapseNIPS.tex
typos

@@
 
-To maximise this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
-AN example of such a model is shown in \autoref{fig:maxinit}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
+To maximize this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
+An example of such a model is shown in \autoref{fig:maxinit}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
 These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
@@
   \item\aligntop{\includegraphics[width=0.15\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
-  \item\aligntop{\includegraphics[width=0.05\linewidth]{binary_det.svg}}\label{fig:binary_det}
+  \item\hp\aligntop{\includegraphics[width=0.05\linewidth]{binary_det.svg}}\label{fig:binary_det}
   \end{myenuma}
@@
 It is of the form shown in \autoref{fig:max_area}, with potentiation moving one step to the right and depression moving one step to the left.
-Any other topology would allow some perturbation that could increase the area.
+Any other topology would allow some perturbation to further increase the area.
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
+  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
+  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut.svg}}\label{fig:shortcut}
+ \end{myenuma}
+ \end{center}
+  \caption{Perturbations that increase the area.
+  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$.
+It can no longer be used when $\M\dep$ is lower triangular.
+  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$.
+It can no longer be used when $\M\pot$ is upper triangular.
+  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions.
+It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
+  }\label{fig:perts}
+\end{figure}
 
@@
  \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
-  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
-  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut.svg}}\label{fig:shortcut}
- \end{myenuma}
- \end{center}
-  \caption{Perturbations that increase the area.
-  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$.
-It can no longer be used when $\M\dep$ is lower triangular.
-  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$.
-It can no longer be used when $\M\pot$ is upper triangular.
-  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions.
-It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
-  }\label{fig:perts}
-\end{figure}
-
-\begin{figure}
- \begin{center}
   \includegraphics[width=0.4\linewidth]{multistate_sticky.svg}

### 2013-05-15 NIPS/SynapseNIPS.tex
wrote envelope section

@@
 
-In this section, we will describe the class of models of synaptic plasticity that we are studying and how we quantify their performance in memory storage.
+In this section, we will describe the class of models of synaptic plasticity that we are studying and how we quantify their performance as memory storage devices.
 In the subsequent sections, we will find upper bounds on this performance.
@@
 To maximize this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
-An example of such a model is shown in \autoref{fig:maxinit}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
+An example of such a model is shown in \autoref{fig:max_init}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
 These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \autoref{fig:maxinit}(\ref{fig:binary_det}).
+They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \autoref{fig:max_init}(\ref{fig:binary_det}).
 
@@
 %
-\begin{equation}\label{eq:maxinit}
+\begin{equation}\label{eq:max_init}
   \snr(0) \leq \sqrt{N}.
@@
   (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
-  }\label{fig:maxinit}
+  }\label{fig:max_init}
 \end{figure}
@@
   Unlabeled transitions have probability 1.
-  Labeled transitions have probability $\epsilon\to0$.
+  Labeled transitions have probability $\varepsilon\to0$.
   }\label{fig:max_area}
@@
 \section{Memory curve envelope}\label{sec:env}
+
+Now we will look at the implications of the upper bounds found in the previous section for the SNR at finite times.
+As argued in \eqref{eq:SNReigen}, the memory curve can be written in the form
+%
+\begin{equation}\label{eq:SNRmodes}
+  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a}.
+\end{equation}
+%
+The upper bounds on the initial SNR, \eqref{eq:max_init}, and the area, \eqref{eq:max_area}, imply the following constraints on the parameters $\brc{\CI_a,\tau_a}$:
+%
+\begin{equation}\label{eq:modeconstr}
+  \sum_a \CI_a \leq 1,
+  \qquad\qquad
+  \sum_a \CI_a \tau_a \leq M-1.
+\end{equation}
+%
+We are not claiming that these are a complete set of constraints: not every set $\brc{\CI_a,\tau_a}$ that satisfies these inequalities will actually be achievable by a synaptic model.
+However, any set that violates either inequality will definitely not be achievable.
+
+Now we can pick some fixed time, $t_0$, and maximize the SNR at that time \wrt the parameters $\brc{\CI_a,\tau_a}$, subject to the constraints above.
+This always results in a single nonzero $\CI_a$:
+%
+\begin{equation}\label{eq:env}
+\begin{aligned}
+  t_0 &\leq \frac{M-1}{r} &
+   \implies 
+  \snr(t) &= \sqrt{N}\e^{-rt/(M-1)} &
+   \implies 
+  \snr(t_0) &= \sqrt{N}\e^{-rt_0/(M-1)},\\
+  t_0 &\geq \frac{M-1}{r} &
+   \implies 
+  \snr(t) &= \frac{\sqrt{N}(M-1)\e^{-t/t_0}}{rt_0} &
+   \implies 
+  \snr(t_0) &= \frac{\sqrt{N}(M-1)}{\e rt_0}.
+\end{aligned}
+\end{equation}
+%
+Both the initial SNR bound and the area bound are saturated at early times.
+At late times, only the area bound is saturated.
+The function $\snr(t_0)$, the green curve in \autoref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope.
+No synaptic model can have an SNR that is greater than this at any time.
+
+This leaves the question of whether this bound is achievable.
+At any time, can we find a model whose memory curve touched the envelope?
+The red curves in \autoref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization with random initialization and by hand designed models.
+
+We see that here is a gap between the upper bound that we can prove and what we can achieve with actual models.
+There may be other models we haven't found that could beat the ones we have, but it is probably too much to expect them to reach this envelope.
+That would require saturating the area bound, and we saw that the only model that does that has very low initial SNR.
+It seems that we need to find additional constraints to add to those we have \eqref{eq:modeconstr}, thereby bringing the envelope down.
+We do not necessarily need to find \emph{every} constraint.
+We have already seen that not every constraints is saturated for certain time ranges, so it may be that a small subset of them will be sufficient to find an envelope that is tight.
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.4\linewidth]{env.eps}}\label{fig:envplot}\hp
+  \item\aligntop{\includegraphics[width=0.3\linewidth]{diffjump.svg}}\label{fig:diffjump}
+ \end{myenuma}
+ \end{center}
+  \caption{The memory curve envelope for $N=100$, $M=12$.
+  (\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
+  The red dashed curve shows the result of numerical optimization of synaptic models with random initialization.
+  The solid red curve shows the highest SNR we have found with hand designed models.
+  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states.
+  At late times they are of the form shown in \autoref{fig:max_area} with different values of $\varepsilon$.
+  }\label{fig:env}
+\end{figure}
 

### 2013-05-16 NIPS/SynapseNIPS.tex
first stab at intro section

@@
 \section{Introduction}\label{sec:intro}
+
+In the field of theoretical neuroscience there are several classic results regarding the storage capacity of synaptic memory, \eg the Hopfield model's capacity is proportional to the number of neurons, or the Perceptron's capacity is proportional to the number of synapses.
+However, these rely on one unrealistic feature of these models: the synaptic weights are allowed to take a large number of different values.
+
+Measurements of real synapses have shown that the synaptic efficacy can only take a small number of discrete values, possibly as few as two \cite{Bliss1993LTP,Petersen1998allornone,O'Connor2005switch} (see \cite{Enoki2009graded} for dissenting results).
+Restricting synaptic weights to a small set of values has a catastrophic effect on the storage capacity of these theoretical models, growing only logarithmically with the size of the system if all else is held fixed \cite{amit1992constraints,amit1994learning}.
+Due to newer memories overwriting older ones, there is a trade-off between learning and remembering: if synapses are too plastic older memories will be wiped out too easily, if they are too rigid it becomes difficult to learn new memories in the first place.
+
+This can be improved when we take into account the fact that there is a complex dynamical system underlying synaptic plasticity (\eg see \cite{Bliss1993LTP,Bredt2003AMPA,Coba2009phosphorylation}).
+Allowing the synapse to have hidden functional states, with potentiation and depression causing transitions between these states, can dramatically improve their performance.
+
+Much of the previous work on these types of synaptic models has focussed on specific models, to understand their ability to store memories \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial}.
+Two examples of these models are shown in \autoref{fig:models}.
+we see that they have different memory properties.
+The serial model has performs very well at one particular timescale, but it performs poorly at other times.
+The cascade model does not perform quite as well at that time, but it maintains it performance over a wider range of timescales.
+
+\begin{figure}[tbp]
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
+  \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
+  \item\aligntop{\includegraphics[width=0.35\linewidth]{cascms.eps}}\label{fig:cascms}
+ \end{myenuma}
+ \end{center}
+  \caption{Models of complex synapse.
+  (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
+  (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
+  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined in \autoref{sec:setup}) as subsequent memories are stored.
+  }\label{fig:models}
+\end{figure}
+
+We are going to take a different approach.
+We are going to consider the \emph{entire} space of these models to find upper bounds on the capability of \emph{any} of them.
+As the performance of a synaptic model is quantified by a memory curve (describing the decay of a memory trace as more and more subsequent memories are stored) rather than a single number, there is no good notion of optimality.
+What type of memory curve is best will depend on what the synapses are required to do and over what timescales.
+
+We begin in \autoref{sec:setup} by describing the class of models that we will consider and define the memory curve that we use to quantify their performance.
+In \autoref{sec:bounds} we will prove upper bounds on two quantities, the initial value of the memory curve and the area under it, as well as finding the models that saturate them.
+We will use these bounds in \autoref{sec:env} to find an envelope curve -- one that cannot be crossed by the memory curve of any model at any time.
 
@@
 
-\begin{figure}
+\begin{figure}[tbp]
  \begin{center}
@@
 We will find an upper bound on this quantity as well as the model that saturates it.
-The area is useful as it also provides an upper bound on the lifetime of a memory, $\tau(\epsilon)$, defined as the time at which $\snr(\tau)=\epsilon$:
+The area is useful as it also provides an upper bound on the lifetime of a memory, $\tau(\epsilon)$, defined as the time at which $\snr(\tau)=\epsilon$ (we often take $\epsilon=1$):
 %
@@
 %
+This result is similar to that found numerically by Barrett and van Rossum \cite{Barrett2008discrete} in a slightly different situation.
+
 The ``sticky'' end states result in very slow decay, but they also make it difficult to encode the memory in the first place.
@@
 No synaptic model can have an SNR that is greater than this at any time.
+We can use this to find an upper bound on the memory lifetime, $\tau(\epsilon)$, by finding the point at which the envelope crosses $\epsilon$:
+%
+\begin{equation}\label{eq:lifebound}
+  \tau(\epsilon) \leq \frac{\sqrt{N}(M-1)}{\epsilon\e r},
+\end{equation}
+%
+where we assume $N>(\epsilon\e)^2$.
 

### 2013-05-16 Code/Figs/CascMSFig.m

@@ -1,4 +1,4 @@
-t=0.1:2000;
+t=10.^(-1:0.2:3);
 x=0.5;
 q=(x.^abs((1:19)-10))/(1-x);
 [Wp,Wm,w]=CascadeMSinterp(q,q,0.5,1);
@@ -14,8 +14,8 @@ xlim([t(1) t(end)])
 ylim([casc(end) casc(1)])
 xlabel('Time')
 ylabel('SNR')
-legend({'Cascade','Multistate'},'Location','Best')
-embiggen
+legend({'Cascade','Serial'},'Location','Best')
+embiggen(gca,20)
 set(gcf, 'PaperPositionMode', 'auto');
 %%
 clear x q Wp Wm w t m1 casc

### 2013-05-16 Code/MakeMat/CascadeMSinterp.m

@@ -38,8 +38,8 @@ Wm(nn+1,nn)=qm(nn)-fp/(1-fp)*qp(nn+1)*lambda;
 Wp(1,nn+1)=qp(1)*lambda;
 Wm(n,nn)=qm(n-1)*lambda;
 
-Wp=Stochastify(Wp);
-Wm=Stochastify(Wm);
+Wp=StochastifyC(Wp);
+Wm=StochastifyC(Wm);
 
 w=[-ones(nn,1);ones(nn,1)];
 

### 2013-05-29 NIPS/SynapseNIPS.tex
Surya's changes

--- 2013-05-15 NIPS/SynapseNIPS.tex
@@
 
+% references to add:
+%
+% Hopfield: original paper
+% Somp: memory capacity of Hopfield
+% Gardner: memory capacity of perceptron
+% Madison/Montgomery:  State-dependent heterogeneity in synaptic depression between pyramidal cell pairs. Neuron.
+% Abraham/Bear: Metaplasticity: the plasticity of synaptic plasticity.
+% Grant:  Evolution of synapse complexity and diversity (Ann Rev)
+
 \section{Introduction}\label{sec:intro}
 
-\section{Synaptic models and their memory curves}\label{sec:setup}
+It is widely thought that our very ability to remember the past over long time scales depends crucially on our ability to modify synapses in our brain in an experience dependent manner.
+Classical models of synaptic plasticity model synaptic efficacy as an analog scalar value, denoting the size of a post-synaptic potential injected into one neuron from another.
+Theoretical work has shown that such models have a reasonable, extensive memory capacity, in which the number of long term associations that can be stored by a neuron is proportional its number of afferent synapses \cite{Hopfield1982model,Amit1985hopfield,Gardner1988perceptron}.
+However, recent experimental work has shown that many synapses are more digital than analog; they cannot robustly assume an infinite continuum of analog values, but rather can only take on a finite number of distinguishable strengths, a number than can be as small as two \cite{Bliss1993LTP,Petersen1998allornone,O'Connor2005switch} (though see \cite{Enoki2009graded}).
+This one simple modification leads to a catastrophe in memory capacity:  classical models with digital synapses, when operating in a palimpset mode in which the ongoing storage of new memories can overwrite previous memories, have a memory capacity proportional to the logarithm of the number of synapses \cite{amit1992constraints,amit1994learning}.
+Intuitively, when synapses are digital, the storage of a new memory can flip a population of synaptic switches, thereby rapidly erasing previous memories.
+This result indicates that the dominant theoretical basis for the storage of long term memories in modifiable synaptic switches is flawed.
 
-In this section, we will describe the class of models of synaptic plasticity that we are studying and how we quantify their performance as memory storage devices.
+Recent work \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial} has suggested that a way out of this logarithmic catastrophe is to expand our theoretical conception of a synapse from a single scalar value to an entire stochastic dynamical system in its own right.
+This conceptual expansion is further necessitated by the experimental reality that synapses contain within them immensely complex molecular signaling pathways, with many internal molecular functional states (\eg see \cite{Bliss1993LTP,Bredt2003AMPA,Coba2009phosphorylation}).
+While externally, synaptic efficacy could be digital, candidate patterns of electrical activity leading to potentiation or depression could yield transitions between these internal molecular states without necessarily inducing an associated change in synaptic efficacy.
+This form of synaptic change, known as metaplasticity \cite{Abraham1996metaplasticity,Montgomery2002765}, can allow the probability of synaptic potentiation or depression to acquire a rich dependence on the history of prior changes in efficacy, thereby potentially improving memory capacity.
+
+Theoretical studies of complex, metaplastic synapses, have focused on analyzing the memory performance of a limited number of very specific, molecular dynamical systems, characterized by a number of internal states in which potentiation and depression each induce a specific set of allowable transitions between states (\eg see \autoref{fig:models} below).
+While these models can vastly outperform simple binary synaptic switches, these analyses leave open several deep and important questions.
+For example, how does the structure of a synaptic dynamical system determine its memory performance?
+What are the fundamental limits of memory performance over the space of all possible synaptic dynamical systems?
+What is the structural organization of synaptic dynamical systems that achieve these limits?
+Moreover, from an experimental perspective, it is unlikely to be the case that there is a single canonical synaptic model;  just like the case of neurons, there is an incredible diversity of molecular networks underlying synapses both across species and across brain regions within a single organism \cite{Emes2012synapserev}.
+In order to elucidate the functional contribution of this diverse molecular complexity to learning and memory, it is essential to move beyond the analysis of specific models and instead develop a general theory of learning and memory for complex synapses.
+Moreover, such a general theory of complex synapses could aid in development of novel artificial memory storage devices.  % Sutskever, memory for learning?
+
+Here we initiate such a general theory by proving upper bounds on the memory curve associated with any synaptic dynamical system, within the well established ideal
+observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}.
+Along the way we develop principles based on first passage time theory to order the structure of synaptic dynamical systems and relate this structure to memory performance.
+We summarize our main results in the \hyperref[sec:disc]{discussion section}.
+
+\section{Overall framework: synaptic models and their memory curves}\label{sec:setup}
+
+In this section, we will describe the class of models of synaptic plasticity that we are studying and how we quantify their memory performance.
 In the subsequent sections, we will find upper bounds on this performance.
@@
 We use a well established formalism for the study of learning and memory with complex synapses (see \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}).
-In this approach, potentiating and depressing plasticity events occur at random times, with all information about the neural activity and learning rules responsible for them absorbed into their rates.
-We assume that there are no spatial or temporal correlations in the pattern of potentiating and depressing events, and these events cause Markovian transitions between the internal states of the synapses.
-As a result of these assumptions, the states of different synapses will be independent, and the system can be fully described by the probability distribution across these states, which we will indicate with the row-vector $\pr(t)$.
+In this approach, electrical patterns of activity corresponding to candidate potentiating and depressing plasticity events occur randomly and independently at all synapses at a Poisson rate $r$.
+These events reflect possible synaptic changes due to either spontaneous network activity, or the storage of new memories.
+We let  $f\pot$ and  $f\dep$ denote the fraction of these events that are candidate potentiating or depressing events respectively.
+Furthermore, we assume our synaptic model has $M$ internal molecular functional states, and that a candidate potentiating (depotentiating) event induces a stochastic transition in the internal state described by an $M \times M$ discrete time Markov transition matrix  $\M\pot$ ($\M\dep$).
+In this framework, the states of different synapses will be independent, and the entire synaptic population can be fully described by the probability distribution across these states, which we will indicate with the row-vector $\pr(t)$.
+Thus the $i$'th component of $\pr(t)$ denotes the fraction of the synaptic population in state $i$.
+Furthermore, each state $i$ has its own synaptic weight, which we take, in the worst case scenario, to be restricted to two values.
+After shifting and scaling these two values,  we can assume they are $\pm1$, without loss of generality.
 
 We also employ an ``ideal observer'' approach to the memory readout, where the synaptic weights are read directly.
-This provides an upper bound to the quality of any readout using neural activity.
-These synaptic weights will be restricted to two values, which we can shift and scale to $\pm1$.
+This provides an upper bound on the quality of any readout using neural activity.
 
-For any single memory, there will be an ideal pattern of synaptic weights, the $N$-element vector $\wvi$, that is $+1$ at all synapses that were potentiated and $-1$ at all synapses that were depressed.
-The actual pattern of synaptic weights at some later time, $t$, will be $\wv(t)$.
+For any single memory, stored at time $t=0$, we assume there will be an ideal pattern of synaptic weights across a population of $N$ synapses, the $N$-element vector $\wvi$, that is $+1$ at all synapses that experience a candidate potentiation event, and $-1$ at all synapses that experienced a candidate depression event at the time of memory storage.
+The actual pattern of synaptic weights at some later time, $t$, will change to $\wv(t)$ due to further modifications from subsequent memories.
 We can use the overlap between these, $\wvi\cdt\wv(t)$, as a measure of the quality of the memory.
-As $t\to\infty$, the system will return to its steady state distribution which will be uncorrelated with the memory in question.
+As $t\to\infty$, the system will return to its steady state distribution which will be uncorrelated with the memory stored at $t=0$.
 The probability distribution of the quantity $\wvi\cdt\wv(\infty)$ can be used as a ``null model'' for comparison.
 
-The extent to which the memory has been stored is described by a signal-to-noise ratio (SNR):
+The extent to which the memory has been stored is described by a signal-to-noise ratio (SNR) \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}:
 %
@@
 %
-The noise is essentially $\sqrt{N}$.
+The noise in the denominator is essentially $\sqrt{N}$.
 There is a correction when potentiation and depression are imbalanced, but this will not affect the upper bounds that we will discuss below and will be ignored in the subsequent formulae.
 
-All of the preceding plasticity events will put the system in its steady-state distribution, $\eq$.
-The memory we are tracking at $t=0$ will change this to $\eq\M\pot$ in those synapses that are potentiated and $\eq\M\dep$ in those synapses that are depressed, where $\M\potdep$ are $M\times M$ matrices of transition probabilities for the Markov processes describing potentiation and depression.
+A simple mean memory curve can be derived as follows.
+All of the preceding plasticity events, prior to $t=0$, will put the population of synapses in its steady-state distribution, $\eq$.
+The memory we are tracking at $t=0$ will change the internal state distribution to $\eq\M\pot$ in those synapses that are potentiated and $\eq\M\dep$ in those synapses that are depressed.
 As the potentiating/depressing nature of the subsequent memories is independent of $\wvi$, we can average over all sequences, resulting in the evolution of the probability distribution:
@@
 %
-Here, $r$ is the rate of the Poisson process describing the timing of plasticity events,
-$f\potdep$ is the fraction of events that are potentiating/depressing and $\I$ is the identity matrix.
+Here $\frg$ is a continuous time transition matrix that models the process of forgetting the memory stored at time $t=0$ due to random candidate potentiation/depression events occurring at each synapse due to the storage of subsequent memories.
 
@@
 We will frequently refer to this function as the memory curve.
+It can be thought of as the fraction of synapses that maintain the ideal synaptic strength at time $t$, dictated by the stored memory at time $t=0$.
 
-The parameters must satisfy the following constraints:
+Much of the previous work on these types of complex synaptic models has focused on understanding the memory curves of specific models, or choices of $\M\potdep$.
+Two examples of these models are shown in \autoref{fig:models}. We see that they have different memory properties.
+The serial model performs relatively well at one particular timescale, but it performs poorly at other times.
+The cascade model does not perform quite as well at that time, but it maintains it performance over a wider range of timescales.
+
+\begin{figure}[tbp]
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.25\linewidth]{cascade-generated.pdf}}\label{fig:cascade}
+  \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate-generated.pdf}}\label{fig:multistate}
+  \item\aligntop{\includegraphics[width=0.35\linewidth]{cascms.eps}}\label{fig:cascms}
+ \end{myenuma}
+ \end{center}
+  \caption{Models of complex synapses.
+  (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
+  (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
+  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined in \autoref{sec:setup}) as subsequent memories are stored.
+  }\label{fig:models}
+\end{figure}
+
+In this work, rather than analyzing specific models,  we take a different approach, in order to obtain a more general theory.
+We consider the \emph{entire} space of these models and find upper bounds on the memory capacity of \emph{any} of them.
+The space of models with a fixed number of internal states $M$ is parameterized by the pair of $M \times M$ discrete time stochastic
+transition matrices $\M\pot$ and $\M\dep$, in addition to $f\potdep$.  The parameters must satisfy the following constraints:
 %
@@
 
-The question is: what do these constraints imply for the memory curve above?
-In practice, to make any statements about finite times, we need to got to the eigenmode description:
+The critical question is: what do these constraints imply about the space of achievable memory curves in \eqref{eq:SNRcalc}?
+To answer this question, especially for limits on achievable memory at finite times, it will be useful to employ the eigenmode description:
 %
@@
 %
-Where $q_a$ are the negative of the eigenvalues, $\uv^a$ are the right (column) eigenvectors and $\vv^a$ are the left (row) eigenvectors.
-This allow us to write
+Here $q_a$ are the negative of the eigenvalues of the forgetting process $\frg$, $\uv^a$ are the right (column) eigenvectors and $\vv^a$ are the left (row) eigenvectors.
+This decomposition allows us to write the memory curve as a sum of exponentials,
 %
@@
 %
-We can ask then ask the question: what are the constraints on these quantities implied by the constraints \eqref{eq:constr}?
-We will find some of these constraints in the next section.
+We can then ask the question: what are the constraints on these quantities, namely initial SNR's $\CI_a$ and time constants $\tau_a$ for the modes,  implied by the constraints in \eqref{eq:constr}?
+We will derive some of these constraints in the next section.
 
-\section{Upper bounds}\label{sec:bounds}
+\section{Upper bounds on achievable memory capacity}\label{sec:bounds}
 
-In this section, we will find some upper bounds on certain properties of the memory curve discussed in the previous section, namely its initial value and the area under it.
-We will use these bounds to determine an upper bound on the SNR at finite times in the next section.
+In the previous section, in \eqref{eq:SNRcalc} we have described an analytic expression for a memory curve as a function of the structure of a synaptic dynamical system, described by the pair of stochastic transition matrices  $\M\potdep$.
+Since the performance measure for memory is an entire memory curve, and not just a single number, there is no universal scalar notion of optimal memory in the space of synaptic dynamical systems.
+Instead there are tradeoffs between storing proximal and distal memories;  often attempts to increase memory at late (early) times by changing $\M\potdep$, incurs a performance loss in memory at early (late) times in specific models considered so far \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial}.
+Thus our goal, achieved in \autoref{sec:env}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the entire memory curve of {\it any} model.
+In order to achieve this goal, in this section, we must first derive upper bounds, over the space of all possible synaptic models, on two different scalar functions of the memory curve: its initial SNR, and the area under the memory curve.
+As a corollary of these results, we also derive an upper-bound on the memory lifetime, or the time at which the SNR drops to a given detection threshold.
 
@@
 
-\begin{figure}
+\begin{figure}[tbp]
  \begin{center}
  \begin{myenuma}
-  \item\hp\aligntop{\includegraphics[width=0.15\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
-  \item\aligntop{\includegraphics[width=0.15\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
-  \item\hp\aligntop{\includegraphics[width=0.05\linewidth]{binary_det.svg}}\label{fig:binary_det}
+  \item\hp\aligntop{\includegraphics[width=0.15\linewidth]{max_init_pot-generated.pdf}}\label{fig:max_init_pot}\hp
+  \item\aligntop{\includegraphics[width=0.15\linewidth]{max_init_dep-generated.pdf}}\label{fig:max_init_dep}\hp
+  \item\hp\aligntop{\includegraphics[width=0.05\linewidth]{binary_det-generated.pdf}}\label{fig:binary_det}
   \end{myenuma}
@@
 
-\begin{figure}
+\begin{figure}[tbp]
  \begin{center}
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
-  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
-  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut.svg}}\label{fig:shortcut}
+  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right-generated.pdf}}\label{fig:tri_right}\hp
+  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left-generated.pdf}}\label{fig:tri_left}\hp
+  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut-generated.pdf}}\label{fig:shortcut}
  \end{myenuma}
@@
 
-\begin{figure}
+\begin{figure}[tbp]
  \begin{center}
-  \includegraphics[width=0.4\linewidth]{multistate_sticky.svg}
+  \includegraphics[width=0.4\linewidth]{multistate_sticky-generated.pdf}
  \end{center}
@@
   \item\aligntop{\includegraphics[width=0.4\linewidth]{env.eps}}\label{fig:envplot}\hp
-  \item\aligntop{\includegraphics[width=0.3\linewidth]{diffjump.svg}}\label{fig:diffjump}
+  \item\aligntop{\includegraphics[width=0.3\linewidth]{diffjump-generated.pdf}}\label{fig:diffjump}
  \end{myenuma}

### 2013-05-30 NIPS/SynapseNIPS.tex
added supp

@@
 
+In terms of these quantities, the area can be written as
+%
+\begin{equation}\label{eq:areaEta}
+\begin{aligned}
+  A &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \prn{\M\pot_{ij}-\M\dep_{ij}} \prn{\eta^+_i-\eta^+_j}\\
+    &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \prn{\M\pot_{ij}-\M\dep_{ij}} \prn{\eta^-_j-\eta^-_i}.
+\end{aligned}
+\end{equation}
+%
+
 With the states in this order, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
@@
 
+We have developed a general theory of learning and memory with complex synapses, allowing for an exploration of the entire space of models.
+If we are to understand the wealth of data that is arriving regarding the molecular mechanisms of synaptic plasticity \cite{Bliss1993LTP,Bredt2003AMPA,Coba2009phosphorylation}, we will need more of this exploration so that we can build an understanding of the relation between structure and function.
+
+We have begun this exploration by finding upper bounds on the memory storage of these synapses.
+
 %\subsubsection*{Acknowledgements}

### 2013-05-30 NIPS/SynapseNIPSsupp.tex
added supp

--- 2013-01-11 Notes/ContTime.tex
@@
-% -*- TeX -*- -*- UK -*-
-% ----------------------------------------------------------------
-% arXiv Paper ************************************************
-%
-% Subhaneil Lahiri's template
-%
-% Before submitting:
-%    Comment out hyperref
-%    Comment out showkeys
-%    Replace \input{mydefs.tex} with its contents
-%       or include mydefs.tex in zip/tar file
-%    Replace \input{newsymb.tex} with its contents
-%       or include newsymb.tex in zip/tar file
-%    Put this file, the .bbl file, any picture or
-%       other additional files and natbib.sty
-%       file in a zip/tar file
-%
-% **** -----------------------------------------------------------
-\documentclass[12pt]{article}
-%Preamble:
-\usepackage{a4wide}
+% -*- TeX -*- -*- US -*-
+\documentclass{article} % For LaTeX2e
+\usepackage{nips13submit_e,times}
 \input{sl_preamble.tex}
-%
+\input{sl_graphics_preamble.tex}
+\graphicspath{{"Figs/"}}
 % >> Only for drafts! <<
@@
 \input{sl_symbols.tex}
+%
+%additional symbols:
+%
+\DeclareMathOperator{\SNR}{SNR}
+\DeclareMathOperator{\snr}{SNR}
+\newcommand{\wv}{\vec{w}}
+\newcommand{\wvi}{\vec{w}_\text{ideal}}
 %matrices
@@
 \newcommand{\MM}{\mathbf{Q}}
-%equilibrium distribution
-\newcommand{\eq}{\mathbf{p}^\infty}
+%prob distributions
+\newcommand{\pr}{\mathbf{p}}
+\newcommand{\eq}{\pr^\infty}
 %first passage times
@@
 \newcommand{\M}{\mathbf{M}}
+\newcommand{\enc}{\mathbf{q}}
+\newcommand{\frg}{\W^{\mathrm{F}}}
 \newcommand{\F}{\boldsymbol{\Phi}}
-\DeclareMathOperator{\SNR}{SNR}
+%superscripts
+\newcommand{\pot}{^{\text{pot}}}
+\newcommand{\dep}{^{\text{dep}}}
+\newcommand{\potdep}{^{\text{pot/dep}}}
+%sets
 \newcommand{\CS}{\mathcal{S}}
@@
 \newcommand{\comp}{^\mathrm{c}}
+%eigenmodes
+\newcommand{\uv}{\mathbf{u}}
+\newcommand{\vv}{\mathbf{v}}
+\newcommand{\CI}{\mathcal{I}}
 % ----------------------------------------------------------------
@@
 % ----------------------------------------------------------------
+
 %
@@
 % Title info:
-\title{Area maximisation in continuous time}
+\title{A memory frontier for complex synapses: Supplementary material}
 %
@@
 %
-\author{Subhaneil Lahiri
+\author{Subhaneil Lahiri and Surya Ganguli\\
+Applied Physics Department, Stanford University, Stanford CA\\
+\emaillink{sulahiri@stanford.edu}, \emaillink{sulahiri@stanford.edu}
 %
@@
 
+%\nipsfinalcopy % Uncomment for camera-ready version
+
 \begin{document}
@@
 
-\begin{abstract}
-  We try to find the continuous time Markov process that has the maximal area under the signal-to-noise curve.
-\end{abstract}
-
-\tableofcontents
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Beginning of Article:
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Continuous time Markov processes}\label{sec:ContMarkov}
 
-In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time.
+In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time to define notation and .
 It is a straightforward generalisation of material that can be found in \cite{kemeny1960finite} with some ideas from \cite{hunter2000survey}.
@@
 %
-
 \section{Signal-to-Noise ratio (SNR)}\label{sec:SNR}
@@
 \begin{equation}\label{eq:MWdef}
-  \M^\pm \equiv \I + \W^\pm,
-  \qquad
-  \M^\pm\onev = \onev,
-  \qquad
-  \M^\pm_{ij} \in [0,1].
-\end{equation}
-%
-The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf^\pm$, where $f^++f^-=1$ are the fraction of plasticity events that are potentiating/depressing respectively.
+  \M\potdep \equiv \I + \W\potdep,
+  \qquad
+  \M\potdep\onev = \onev,
+  \qquad
+  \M\potdep_{ij} \in [0,1].
+\end{equation}
+%
+The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf\potdep$, where $f\pot +f\dep =1$ are the fraction of plasticity events that are potentiating/depressing respectively.
 This means that the ``forgetting'' process will be described by the \emph{continuous}-time Markov chain:
@@
 \begin{equation}\label{eq:forgetting}
-  \MM = r \W^F \equiv r\prn{f^+\W^+ + f^-\W^-}.
-\end{equation}
-%
-We only require that this Markov chain is ergodic. The Markov chains described by $\M^\pm$ need not be.
+  \MM = r \frg \equiv r\prn{f\pot\W\pot + f\dep\W\dep}.
+\end{equation}
+%
+We only require that this Markov chain is ergodic. The Markov chains described by $\M\potdep $ need not be.
 
 We assume that the probability distribution starts in the equilibrium distribution \eqref{eq:equilibrium}.
-During the initial signal creation, a fraction $f^+$ will change to $\eq\M^+$ and a fraction $f^-$ will change to $\eq\M^-$.
+During the initial signal creation, a fraction $f\pot$ will change to $\eq\M\pot$ and a fraction $f\dep$ will change to $\eq\M\dep$.
 After this, probabilities will evolve according to \eqref{eq:rowsum}.
@@
 
-As we ignore correlations between different synapses, the only quantity of interest will be the number of synapses with each strength.
-The signal can be defined as the increase in synaptic strength due to potentiation plus the decrease in synaptic strength due to depression.
-If there are $N$ synapses and the strength of the synapse in each state is given by the column vector, $\w$, this is given by
-%
-\begin{equation}\label{eq:signal}
-  S(t) = N f^+ [\mathbf{p}(t|+)-\eq]\w - N f^+ [\mathbf{p}(t|-)-\eq]\w,
-\end{equation}
-%
-where $\mathbf{p}(t|\pm)$ is the probability of being in each state conditional on potentiation/depression at $t=0$:
-%
-\begin{equation}\label{eq:initprob}
-  \mathbf{p}(t|\pm) = \eq\M^\pm \e^{rt\W^F}.
-\end{equation}
-%
-We can rewrite the signal as
-%
-\begin{equation}\label{eq:signalr}
-\begin{aligned}
-  S(t) &= N \eq (f^+\W^+ - f^-\W^-)\e^{rt\W^F}\w\\
-    & N (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\W^F}\w.
-\end{aligned}
-\end{equation}
-%
-For the noise, we can use the standard deviation of the synaptic strength:
-%
-\begin{equation}\label{eq:noise}
-  \text{Noise} = \sqrt{N \prn{\sum_i \eq_i\w_i^2 - (\eq\w)^2}}.
-\end{equation}
-%
-From now on we shall assume that there are only two possible synaptic strengths, strong and weak, which we will describe by $\w_i=\pm1$. This splits the set of states into two subsets:
-%
-\begin{equation}\label{eq:plusminusstates}
-  \boldsymbol{+} = \set{k}{\w_k=+1},
-  \qquad
-  \boldsymbol{-} = \set{k}{\w_k=-1}.
-\end{equation}
-%
-This simplifies the noise:
-%
-\begin{equation}\label{eq:noisepm}
-  \text{Noise} = \sqrt{N \prn{4\eq_+\eq_-}}.
-\end{equation}
-%
-The factors of $\eq_+\eq_-$ will have little effect on anything that follows, so we will drop them.\footnote{They will not affect the optimal transition rates for most quantities, and the optima will correspond to the symmetric case when $\eq_+=\eq_-=\half$.} This leaves the signal-to-noise ratio as
+As discussed in the main text, the signal-to-noise ratio is given by
 %
 \begin{equation}\label{eq:SNRcurve}
-  \SNR(t) = \sqrt{N} (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\W^F}\w.
+  \SNR(t) = \sqrt{N} (2f\pot f\dep ) \eq (\W\pot  - \W\dep )\e^{rt\frg}\w.
 \end{equation}
@@
   \begin{aligned}
-  \W(\alpha) &= \alpha\W^++(1-\alpha)\W^-,
+  \W(\alpha) &= \alpha\W\pot +(1-\alpha)\W\dep ,
   &\quad&\implies&
-    \W^F&=\W(f^+),\\&&&&
-    \W^+ - \W^- &= \diff{\W}{\alpha},\\&&&&
-    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \W^F.
+    \frg&=\W(f\pot ),\\&&&&
+    \W\pot  - \W\dep  &= \diff{\W}{\alpha},\\&&&&
+    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \frg.
   \end{aligned}
@@
 \begin{equation}\label{eq:SNRalpha}
-  \SNR(t) = \sqrt{N} (2f^+f^-) \diff{\eq}{\alpha}(-\W^F)\,\e^{rt\W^F}\w.
-\end{equation}
-%
-
-\subsection{Degrees of freedom and constraints}\label{sec:constraints}
-
-We would like to use the components of $\W^\pm$ as the independent degrees of freedom when we try to maximise quantities associated with the SNR curve \eqref{eq:SNRcurve}.
-However, \eqref{eq:MWdef} imposes some restrictions on the space of allowed values.
-We can solve some of these by setting
-%
-\begin{equation}\label{eq:Wdiag}
-  \W^\pm_{ii} = -\sum_{j\neq i} \W^\pm_{ij}.
-\end{equation}
-%
-Then we treat the off-diagonal components as the independent degrees of freedom.
-These must satisfy the following inequalities:
-%
-\begin{align}\label{eq:Wmin}
-  \W^\pm_{ij} &\geq 0 \qquad \text{for $i \neq j$,}\\
-\label{eq:Wmax}
-  \sum_{j \neq i} \W^\pm_{ij} &\leq 1.
-\end{align}
-%
-The inequalities $\W^\pm_{ij} \leq 1$, $-1\leq\W^\pm_{ii}\leq0$ follow automatically from these.
-
-In \sref{sec:rescale}, we will see that the area under the SNR curve is invariant under the scaling $\W^\pm\to\lambda\W^\pm$.
-This means that we can ignore the constraint \eqref{eq:Wmax} when maximising the area:
-after the area has been maximised, we can use this degree of freedom to enforce \eqref{eq:Wmax} without changing the area.
-However, this scaling does change the SNR curve, in particular its initial value, so we can't ignore \eqref{eq:Wmax} when maximising that.
+  \SNR(t) = \sqrt{N} (2f\pot f\dep ) \diff{\eq}{\alpha}(-\frg)\,\e^{rt\frg}\w.
+\end{equation}
+%
 
@@
 
-Suppose that we have a partition such that $\W^+$ and $\W^-$ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \sref{sec:lump}):
+Suppose that we have a partition such that $\W\pot $ and $\W\dep $ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \sref{sec:lump}):
 %
 \begin{equation}\label{eq:lumpablesynapse}
-  VU\W^\pm V = \W^\pm V,
+  VU\W\potdep  V = \W\potdep  V,
   \qquad
@@
 \begin{equation}\label{eq:lumpedsynapse}
-  \widehat{\W}^\pm = U\W^\pm V,
+  \widehat{\W}\potdep  = U\W\potdep  V,
   \qquad
@@
   \begin{aligned}
-    \frac{\SNR(t)}{\sqrt{N} (2f^+f^-)} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}^+ - \widehat{\W}^-)\e^{rt\widehat{\W}^F}\widehat{\w}. \\
-      &= \eq VU (\W^+ - \W^-) VU \e^{rt\W^F} VU \w. \\
-      &= \eq (\W^+ - \W^-) VU \e^{rt\W^F} VU \w. \\
-      &= \eq (\W^+ - \W^-)\e^{rt\W^F} VU \w. \\
-      &= \eq (\W^+ - \W^-)\e^{rt\W^F}\w. \\
+    \frac{\SNR(t)}{\sqrt{N} (2f\pot f\dep )} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}\pot  - \widehat{\W}\dep )\e^{rt\widehat{\W}^F}\widehat{\w}. \\
+      &= \eq VU (\W\pot  - \W\dep ) VU \e^{rt\frg} VU \w. \\
+      &= \eq (\W\pot  - \W\dep ) VU \e^{rt\frg} VU \w. \\
+      &= \eq (\W\pot  - \W\dep )\e^{rt\frg} VU \w. \\
+      &= \eq (\W\pot  - \W\dep )\e^{rt\frg}\w. \\
   \end{aligned}
@@
 
-Using the first line of \eqref{eq:signalr}, we can write the initial SNR as
+Using $\eq\frg=0$, we can write the initial SNR as
 %
 \begin{equation}\label{eq:init}
-  \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f^+\W^+-f^-\W^-)(\onev^+-\onev^-).
-\end{equation}
-%
-Using $\W^\pm(\onev^++\onev^-)=0$ and \eqref{eq:compflux}:
+  \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f\pot \W\pot -f\dep \W\dep )(\onev^+-\onev^-).
+\end{equation}
+%
+Using $\W\potdep (\onev^++\onev^-)=0$ and \eqref{eq:compflux}:
 %
 \begin{equation*}
-  r{\eq}^-(f^+\W^++f^-\W^-)\onev^+ = \F_{-+} = \F_{+-} = r{\eq}^+(f^+\W^++f^-\W^-)\onev^-,
+  r{\eq}^-(f\pot \W\pot +f\dep \W\dep )\onev^+ = \F_{-+} = \F_{+-} = r{\eq}^+(f\pot \W\pot +f\dep \W\dep )\onev^-,
 \end{equation*}
@@
 \begin{equation}\label{eq:initflux}
-  I = \frac{4\F_{-+}}{r} - 4{\eq}^+f^+\W^+\onev^- - 4{\eq}^-f^-\W^-\onev^+.
-\end{equation}
-%
-The last two terms are guaranteed to be negative, as the diagonal parts of $\W^\pm$ cannot contribute.
+  I = \frac{4\F_{-+}}{r} - 4{\eq}^+f\pot \W\pot \onev^- - 4{\eq}^-f\dep \W\dep \onev^+.
+\end{equation}
+%
+The last two terms are guaranteed to be negative, as the diagonal parts of $\W\potdep $ cannot contribute.
 Therefore
@@
 This inequality is saturated if potentiation never takes it from a $+$ state to a $-$ state and depression never takes it from a $-$ state to a $+$ state.
-
-\subsection{Maximising initial SNR}\label{sec:initmax}
-
-The flux will be maximised if potentiation is guaranteed to take it to a $+$ state and depression is guaranteed to take it to a $-$ state:
-%
-\begin{equation}\label{eq:guarantee}
-  \sum_{j\in+} \W^+_{ij} = 1 \quad \forall\, i\in-,
-  \qquad
-  \sum_{j\in-} \W^-_{ij} = 1 \quad \forall\, i\in+.
-\end{equation}
-%
-This implies that the process is lumpable to a two state system (see \sref{sec:lump} and \sref{sec:SNRlump}):
-%
-\begin{multline}\label{eq:binarylump}
-%\begin{aligned}
-  \W^+ = \begin{pmatrix}
-           -1 & 1 \\
-           0 & 0 \\
-         \end{pmatrix}
-  ,\quad
-  \W^- = \begin{pmatrix}
-           0 & 0 \\
-           1 & -1 \\
-         \end{pmatrix}
-  ,\quad
-  \w = \begin{pmatrix}
-           -1 \\
-           1 \\
-         \end{pmatrix}
-  ,\\ \implies\quad
-  \W^F = \begin{pmatrix}
-           -f^+ & f^+ \\
-           f^- & -f^- \\
-         \end{pmatrix}
-  ,\quad
-  \eq = \begin{pmatrix}
-           f^- & f^+ \\
-         \end{pmatrix}
-  ,%&
-%  \e^{rt\W^F} &= \I+(1-\e^{-rt})\W^F
-%         \begin{pmatrix}
-%           f^-+f^+\e^{-rt} & f^+(1-\e^{-rt}) \\
-%           f^-(1-\e^{-rt}) & f^++f^-\e^{-rt} \\
-%         \end{pmatrix}
-%  ,
-%  \end{aligned}
-\end{multline}
-%
-which leads to the initial SNR %curve
-\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to
-%$\SNR(t) = \sqrt{N (4f^+f^-)} \,\e^{-rt}$,
-$\SNR(0) \leq \sqrt{N (4f^+f^-)}$,
-which would have no effect on what follows.}
-%
-\begin{equation}\label{eq:binarySNR}
- % \SNR(t) = \sqrt{N} (4f^+f^-) \,\e^{-rt}.
-  \SNR(0) \leq \sqrt{N} (4f^+f^-) .
-\end{equation}
-%
-%The initial SNR is maximised at $f^+=f^-=\half$:
-This is maximised at $f^+=f^-=\half$:
-%
-\begin{equation}\label{eq:initmax}
-  \SNR(0) \leq \sqrt{N}.
-\end{equation}
-%
 
@@
 %\begin{equation}\label{eq:SNR}
-%  \SNR(t) = \sqrt{N}(2f^+f^-) \diff{\eq}{\alpha} (-\W^F) \e^{rt\W^F} \w.
+%  \SNR(t) = \sqrt{N}(2f^+f\dep ) \diff{\eq}{\alpha} (-\frg) \e^{rt\frg} \w.
 %\end{equation}
@@
 \begin{aligned}
-  A &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\W^F}}_0^\infty \w \\
-    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
-    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \w.
-\end{aligned}
-\end{equation}
-%
-
-We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f^+f^-)\hat{A}$ and $q_{ij} \equiv \diff{\W^F_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
+  A &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\frg}}_0^\infty \w \\
+    &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
+    &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} \w.
+\end{aligned}
+\end{equation}
+%
+
+We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f\pot f\dep )\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W\pot _{ij}-\W\dep _{ij}$
 %
 \begin{equation}\label{eq:areaT}
-  \hat{A} = \sum_{i,j,k} \eq_i q_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
+  \hat{A} = \sum_{i,j,k} \eq_i \enc_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
 \end{equation}
@@
 
-\subsection{Derivatives \wrt \texorpdfstring{$\W^\pm$}{W+/-}}\label{sec:deriv}
-
-As discussed in \sref{sec:constraints},
-we will regard the off-diagonal elements of $\W^\pm_{ij}$ to be the independent variables,
-with $\W^\pm_{ii}=-\sum_{j \neq i} \W^\pm_{ij}$ imposed by hand.
+\subsection{Derivatives \wrt \texorpdfstring{$\W\potdep $}{W+/-}}\label{sec:deriv}
+
+As discussed in the main text,
+we will regard the off-diagonal elements of $\W\potdep _{ij}$ to be the independent variables,
+with $\W\potdep _{ii}=-\sum_{j \neq i} \W\potdep _{ij}$ imposed by hand.
 Thus,
@@
 \begin{equation}\label{eq:basicderivs}
-  \pdiff{\W^F_{ij}}{\W^\pm_{gh}} = f^\pm \delta_{gi}(\delta_{hj}-\delta_{ij}),
-  \qquad
-  \pdiff{q_{ij}}{\W^\pm_{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
+  \pdiff{\frg_{ij}}{\W\potdep _{gh}} = f\potdep  \delta_{gi}(\delta_{hj}-\delta_{ij}),
+  \qquad
+  \pdiff{\enc_{ij}}{\W\potdep _{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
 \end{equation}
@@
 \begin{equation}\label{eq:derivZ}
-  \pdiff{\fund_{ij}}{\W^\pm_{gh}} = rf^\pm \fund_{ig} (\fund_{hj}-\fund_{gj}).
+  \pdiff{\fund_{ij}}{\W\potdep _{gh}} = rf\potdep  \fund_{ig} (\fund_{hj}-\fund_{gj}).
 \end{equation}
@@
   \begin{aligned}
-    \pdiff{\hat{A}}{\W^\pm_{gh}} =&\,
-      rf^\pm \sum_{ijkl} q_{ij} \w_k \pib_l \brk{
-         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
-    &\pm \sum_{kl} \w_k \pib_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
-    =&\,
-      2rf^\pm \sum_{ij} q_{ij} \eq_i \eq_g \brk{
-         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
-         (\fptb_{ig}-\fptb_{jg})(\eta^+_{g}-\eta^+_{h}) } \\
-      &\pm 2 \eq_g (\eta^+_g - \eta^+_h) \\
-    =&\,
-      2rf^\pm \sum_{ij} q_{ij} \eq_i \eq_g
-         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
-      \pm 2 \eq_g (\eta^+_g - \eta^+_h)\brk{1+\pdiff{\ln(\eq_g)}{\,\ln(f^\pm)}} \\
-    =&\,
-      2rf^\pm \eq_g \brk{\sum_i a_i (\fptb_{gi}-\fptb_{hi}) + c_g (\eta^+_{g}-\eta^+_{h})}
+    \pdiff{\hat{A}}{\W\potdep _{gh}} =&\,
+%      rf\potdep  \sum_{ijkl} \enc_{ij} \w_k \pib_l \brk{
+%         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
+%    &\pm \sum_{kl} \w_k \pib_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
+%    =&\,
+%      2rf\potdep  \sum_{ij} \enc_{ij} \eq_i \eq_g \brk{
+%         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
+%         (\fptb_{ig}-\fptb_{jg})(\eta^+_{g}-\eta^+_{h}) } \\
+%      &\pm 2 \eq_g (\eta^+_g - \eta^+_h) \\
+%    =&\,
+%      2rf\potdep  \sum_{ij} \enc_{ij} \eq_i \eq_g
+%         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
+%      \pm 2 \eq_g (\eta^+_g - \eta^+_h)\brk{1+\pdiff{\ln(\eq_g)}{\,\ln(f\potdep )}} \\
+%    =&\,
+      2rf\potdep  \eq_g \brk{\sum_i a_i (\fptb_{gi}-\fptb_{hi}) + c_g (\eta^+_{g}-\eta^+_{h})}
       \pm 2 \eq_g (\eta^+_g - \eta^+_h). \\
@@
   \begin{aligned}
-    \pdiff{}{\W^F_{gh}} &\equiv     \pdiff{}{\W^+_{gh}} +     \pdiff{}{\W^-_{gh}} ,&\qquad
-    \pdiff{}{q_{gh}}    &\equiv f^- \pdiff{}{\W^+_{gh}} - f^+ \pdiff{}{\W^-_{gh}} .
+    \pdiff{}{\frg_{gh}} &\equiv     \pdiff{}{\W\pot _{gh}} +     \pdiff{}{\W\dep _{gh}} ,&\qquad
+    \pdiff{}{\enc_{gh}}    &\equiv f\dep  \pdiff{}{\W\pot _{gh}} - f\pot  \pdiff{}{\W\dep _{gh}} .
   \end{aligned}
@@
 \begin{equation}\label{eq:derivqweff}
-  \pdiff{\W^F_{ij}}{\W^F_{gh}} = \pdiff{q_{ij}}{q_{gh}}
+  \pdiff{\frg_{ij}}{\frg_{gh}} = \pdiff{\enc_{ij}}{\enc_{gh}}
   =  \delta_{gi}(\delta_{hj}-\delta_{ij}),
   \qquad
-  \pdiff{q_{ij}}{\W^F_{gh}} = \pdiff{\W^F_{ij}}{q_{gh}}  = 0.
-\end{equation}
-%
-This is because we could treat $\W^F$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W^\pm$.
+  \pdiff{\enc_{ij}}{\frg_{gh}} = \pdiff{\frg_{ij}}{q_{gh}}  = 0.
+\end{equation}
+%
+This is because we could treat $\frg$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W\potdep $.
 
@@
 \begin{equation}\label{eq:scaleop}
-  \Delta \equiv \sum_{g,h} \W^+_{gh}\pdiff{}{\W^+_{gh}} + \W^-_{gh}\pdiff{}{\W^-_{gh}}.
-\end{equation}
-%
-This corresponds to the scaling, $\W^\pm \to (1+\epsilon)\W^\pm$.
+  \Delta \equiv \sum_{g,h} \W\pot _{gh}\pdiff{}{\W\pot _{gh}} + \W\dep _{gh}\pdiff{}{\W\dep _{gh}}.
+\end{equation}
+%
+This corresponds to the scaling, $\W\potdep  \to (1+\epsilon)\W\potdep $.
 Intuitively, this has two effects: it scales up the initial potentiation/depression and it scales down all timescales.
@@
 %
-The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W^\pm$.
-
-As the area is invariant under this scaling, we can consider the $\W^\pm$ to be projective coordinates.
-Therefore we don't need to enforce the constraint \eqref{eq:Wmax} while looking for the maximum area, as we can use this null-mode to enforce it afterwards without changing the area.
-We also don't have to worry about the boundary at infinity,
-as sending some of them to infinity is equivalent to sending the rest to zero.
+The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W\potdep $.
+
+As the area is invariant under this scaling, we can consider the $\W\potdep $ to be projective coordinates.
+Therefore we don't need to enforce the lower bound on the diagonal matrix elements while looking for the maximum area, as we can use this null-mode to enforce it afterwards without changing the area.
+%We also don't have to worry about the boundary at infinity, as sending some of them to infinity is equivalent to sending the rest to zero.
 
@@
 \begin{equation}\label{eq:lagrangian}
-  \CL = \hat{A} + \sum_{\pm}\sum_{i\neq j} \mu^\pm_{ij} \W^\pm_{ij} +\lambda\eq\w.
-\end{equation}
-%
-The last term ensures that we hold $\eq_\pm$ fixed during this extremisation, so that we can safely ignore the factors that we dropped from the noise in \eqref{eq:noisepm}.
+  \CL = \hat{A} + \sum_{\text{pot/dep}}\sum_{i\neq j} \mu\potdep _{ij} \W\potdep _{ij} .%  +\lambda\eq\w.
+\end{equation}
+%
+%The last term ensures that we hold $\eq_\pm$ fixed during this extremisation, so that we can safely ignore the factors that we dropped from the noise in \eqref{eq:noisepm}.
 Necessary conditions for an extremum are
@@
 \begin{equation}\label{eq:extremum}
-  \pdiff{\CL}{\W^\pm_{gh}} = 0,
-  \qquad
-    \mu^\pm_{gh} \geq 0,\quad
-    \W^\pm_{gh} \geq 0,\quad
-    \mu^\pm_{gh}\W^\pm_{gh} = 0.
+  \pdiff{\CL}{\W\potdep _{gh}} = 0,
+  \qquad
+    \mu\potdep _{gh} \geq 0,\quad
+    \W\potdep _{gh} \geq 0,\quad
+    \mu\potdep _{gh}\W\potdep _{gh} = 0.
 \end{equation}
@@
 with $g \neq h$.
-This enforces the constraints \eqref{eq:Wmin}, but not \eqref{eq:Wmax}.
-As discussed in \sref{sec:constraints} and \sref{sec:rescale}, that can be enforced after finding the maximum with the null scaling degree of freedom.
+This enforces the positivity constraints on the off-diagonal elements, but not the diagonals.
+As discussed in \sref{sec:rescale}, that can be enforced after finding the maximum using the null scaling degree of freedom.
 
@@
 \begin{equation}\label{eq:shiftqderiv}
-  \pdiff{\CL}{q_{gh}} =
-  f^- \pdiff{\CL}{\W^+_{gh}} - f^+ \pdiff{\CL}{\W^-_{gh}}
-   = (f^- \mu^+_{gh} - f^+ \mu^-_{gh}) + 2\eq_g (\eta^+_g - \eta^+_h)
+  \pdiff{\CL}{\enc_{gh}} =
+  f\dep  \pdiff{\CL}{\W\pot _{gh}} - f\pot  \pdiff{\CL}{\W\dep _{gh}}
+   = (f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh}) + 2\eq_g (\eta^+_g - \eta^+_h)
    = 0.
@@
 \begin{equation}\label{eq:shiftq}
-  \W^+_{ij} \to \W^+_{ij} + f^-\epsilon_{ij},
-  \qquad
-  \W^-_{ij} \to \W^-_{ij} - f^+\epsilon_{ij},
+  \W\pot _{ij} \to \W\pot _{ij} + f\dep \epsilon_{ij},
+  \qquad
+  \W\dep _{ij} \to \W\dep _{ij} - f\pot \epsilon_{ij},
   \qquad
@@
 \begin{equation}\label{eq:lowertriangular}
- f^- \mu^+_{gh} - f^+ \mu^-_{gh} <0
+ f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh} <0
  \qquad\implies\qquad
- \mu^-_{gh} >0
+ \mu\dep _{gh} >0
  \qquad\implies\qquad
- \W^-_{gh}=0.
+ \W\dep _{gh}=0.
 \end{equation}
@@
 \begin{equation}\label{eq:uppertriangular}
- f^- \mu^+_{gh} - f^+ \mu^-_{gh} > 0
+ f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh} > 0
  \qquad\implies\qquad
- \mu^+_{gh} >0
+ \mu\pot _{gh} >0
  \qquad\implies\qquad
- \W^+_{gh}=0.
-\end{equation}
-%
-Thus, if we arrange the states in order of decreasing $\eta^+$, $\W^+$ is upper-triangular and $\W^-$ is lower triangular.
+ \W\pot _{gh}=0.
+\end{equation}
+%
+Thus, if we arrange the states in order of decreasing $\eta^+$, $\W\pot $ is upper-triangular and $\W\dep $ is lower triangular.
 
@@
 
-%Note that if we multiply \eqref{eq:shiftqderiv} by $q_{gh}$ and sum over $(g,h)$, we get
+%Note that if we multiply \eqref{eq:shiftqderiv} by $\enc_{gh}$ and sum over $(g,h)$, we get
 %%
 %\begin{equation}\label{eq:stupid}
-%  \hat{A} = \sum_\pm\sum_{g\neq h} \mu^\pm_{gh} \W^F_{gh},
+%  \hat{A} = \sum_\pm\sum_{g\neq h} \mu\potdep _{gh} \frg_{gh},
 %\end{equation}
@@
 \begin{equation}\label{eq:areacoeffincrfinite}
-  \W^\pm \to \W^\pm + \D\boldsymbol{\epsilon},
+  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon},
   \qquad
@@
     \Delta_{gh}\CL &=
-      \frac{\mu^+_{gh}+\mu^-_{gh}}{\eq_g} + \frac{\mu^+_{hg}+\mu^-_{hg}}{\eq_h}
+      \frac{\mu\pot _{gh}+\mu\dep _{gh}}{\eq_g} + \frac{\mu\pot _{hg}+\mu\dep _{hg}}{\eq_h}
       + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
@@
 
-Using the non-negativity of the Kuhn-Tucker multipliers, $\mu^\pm_{ij}$, \eqref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
-
-Note that, according to \sref{sec:triangular}, either $\W^+_{gh}$ or $\W^-_{gh}$ will be zero at the maximum, therefore we can expect one of $\mu^+_{gh}+\mu^-_{gh}$ to be non-zero.
+Using the non-negativity of the Kuhn-Tucker multipliers, $\mu\potdep _{ij}$, \eqref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
+
+Note that, according to \sref{sec:triangular}, either $\W\pot _{gh}$ or $\W\dep _{gh}$ will be zero at the maximum, therefore we can expect one of $\mu\pot _{gh}+\mu\dep _{gh}$ to be non-zero.
 This would rule out degeneracy of the $c_k$ or $\eta^+_k$.
-Looking at \eqref{eq:shiftqderiv} closely, the only way $\mu^+_{gh}+\mu^-_{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
+Looking at \eqref{eq:shiftqderiv} closely, the only way $\mu\pot _{gh}+\mu\dep _{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
 
@@
 \label{eq:shortcutderiv}
-    \widetilde{\Delta}^\pm_{g,m} &\equiv
-      \brk{ \sum_{k=0}^{m-1} \frac{1}{\eq_{g\pm k}} \pdiffc{}{\W^\pm_{g\pm k,g\pm(k+1)}} }
-      - \frac{1}{\eq_g} \pdiffc{}{\W^\pm_{g,g\pm m}}.
+    \widetilde{\Delta}\potdep _{g,m} &\equiv
+      \brk{ \sum_{k=0}^{m-1} \frac{1}{\eq_{g\pm k}} \pdiffc{}{\W\potdep _{g\pm k,g\pm(k+1)}} }
+      - \frac{1}{\eq_g} \pdiffc{}{\W\potdep _{g,g\pm m}}.
 %  \end{aligned}
@@
 \begin{equation}\label{eq:shortcutprob}
-  \widetilde{\Delta}^\pm_{g,m} \eq_i = 0,
+  \widetilde{\Delta}\potdep _{g,m} \eq_i = 0,
 \end{equation}
@@
 \begin{equation}\label{eq:shortcutfinite}
-  \W^\pm \to \W^\pm + \D\boldsymbol{\epsilon}^{\pm(g,m)},
+  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon}^{\pm(g,m)},
   \qquad
@@
 \label{eq:shortcutarea}
-    \widetilde{\Delta}^\pm_{g,m}\CL &=
-      \brk{ \sum_{k=0}^{m-1} \frac{\mu^\pm_{g\pm k,g\pm(k+1)}}{\eq_{g\pm k}}
-      - \frac{\mu^\pm_{g,g\pm m}}{\eq_g}}
-      + 2rf^\pm \sum_{k=0}^{m-1} \prn{\eta^+_{g\pm k} - \eta^+_{g\pm(k+1)}} \prn{c_{g\pm k} - c_{g}},
+    \widetilde{\Delta}\potdep _{g,m}\CL &=
+      \brk{ \sum_{k=0}^{m-1} \frac{\mu\potdep _{g\pm k,g\pm(k+1)}}{\eq_{g\pm k}}
+      - \frac{\mu\potdep _{g,g\pm m}}{\eq_g}}
+      + 2rf\potdep  \sum_{k=0}^{m-1} \prn{\eta^+_{g\pm k} - \eta^+_{g\pm(k+1)}} \prn{c_{g\pm k} - c_{g}},
 %  \end{aligned}
@@
 This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
-If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu^\pm_{g,g\pm m}>0$, which in turn implies that $\W^\pm_{g,g\pm m}=0$.
+If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu\potdep _{g,g\pm m}>0$, which in turn implies that $\W\potdep _{g,g\pm m}=0$.
 This would tell us that the process with the maximal area has to have a multi-state topology.
@@
 \begin{itemize}
-  \item There can be no ergodic maximum for which $\W^+$ contains backwards transitions or $\W^-$ contains forwards transitions.
+  \item There can be no ergodic maximum for which $\W\pot $ contains backwards transitions or $\W\dep $ contains forwards transitions.
   \item There can be no ergodic maximum with the $c_k$ decreasing.
@@
 %Suppose we start with some arbitrary ergodic process.
-If $\W^\pm$ contain backwards/forwards transitions or the $c_k$ decrease, then we can increase the area by applying the perturbations \eqref{eq:shiftqderiv} and \eqref{eq:areacoeffincrderiv}.
+If $\W\potdep $ contain backwards/forwards transitions or the $c_k$ decrease, then we can increase the area by applying the perturbations \eqref{eq:shiftqderiv} and \eqref{eq:areacoeffincrderiv}.
 %The second perturbation will generate the unwanted backwards/forwards transitions, requiring subsequent application of first perturbation.
@@
 %  \item Applying \eqref{eq:areacoeffincrderiv} an infinite amount sends two transition rates to infinity, which due to the scaling mode (\sref{sec:rescale}), is equivalent to sending all other transitions to zero.
-%      Subsequent application of \eqref{eq:shiftqderiv} just removes the backwards transition from $\W^+$ and the forward transition from $\W^-$.
+%      Subsequent application of \eqref{eq:shiftqderiv} just removes the backwards transition from $\W\pot $ and the forward transition from $\W\dep $.
 %  \item Applying \eqref{eq:areacoeffincrderiv} ends in a process with degenerate $c_k$ or $\eta^+_k$, and the subsequent application of \eqref{eq:shiftqderiv} doesn't change this degeneracy.
@@
 This leaves two possibilities for the maximum area Markov chain.
-Either there is no degeneracy and no shortcuts, which implies the Multi-state topology that we'll discuss in \sref{sec:multistate}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \eqref{eq:shortcutarea}).
+Either there is no degeneracy and no shortcuts, which implies the Multi-state/serial topology that we'll discuss in \sref{sec:multistate}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \eqref{eq:shortcutarea}).
 
@@
 
-\subsection{Multi-state topology}\label{sec:multistate}
-
-The multi-state topology is defined by (see \cite{amit1994learning,Fusi2007multistate}):
+\subsection{Multi-state/Serial topology}\label{sec:multistate}
+
+The multi-state/serial topology is defined by (see \cite{amit1994learning,Fusi2007multistate,Leibold2008serial}):
 %
 \begin{equation}\label{eq:multistatedef}
-  \W^+_{ij} = q^+_i \delta_{i+1,j},
-  \qquad
-  \W^-_{ij} = q^-_j \delta_{i,j+1}.
+  \W\pot _{ij} = q\pot _i \delta_{i+1,j},
+  \qquad
+  \W\dep _{ij} = q\dep _j \delta_{i,j+1}.
 \end{equation}
@@
 \begin{equation}\label{eq:multistateprob}
-  f^+ q^+_i \eq_i = f^- q^-_i \eq_{i+1},
-\end{equation}
-%
-which means we can always choose the transition rates, $q^\pm_i$, to give any desired equilibrium probabilities, $\eq_i$.
+  f\pot  q\pot _i \eq_i = f\dep  q\dep _i \eq_{i+1},
+\end{equation}
+%
+which means we can always choose the transition rates, $q\potdep _i$, to give any desired equilibrium probabilities, $\eq_i$.
 
@@
 \begin{aligned}
-  c_k =&\, \sum_{i<k} \fpt_{i,i+1} \prn{\eq_{i}q^+_i+\eq_{i+1}q^-_i}
+  c_k =&\, \sum_{i<k} \fpt_{i,i+1} \prn{\eq_{i}q\pot _i+\eq_{i+1}q\dep _i}
     %\\&
-    - \sum_{i\geq k} \fpt_{i+1,i} \prn{\eq_{i}q^+_i+\eq_{i+1}q^-_i}
+    - \sum_{i\geq k} \fpt_{i+1,i} \prn{\eq_{i}q\pot _i+\eq_{i+1}q\dep _i}
   ,\\
-  c_{k+1} - c_k =&\, \prn{\fpt_{k,k+1}+\fpt_{k+1,k}} \prn{\frac{\eq_{k}\W^F_{k,k+1}}{f^+}+\frac{\eq_{k+1}\W^F_{k+1,k}}{f^-}}
+  c_{k+1} - c_k =&\, \prn{\fpt_{k,k+1}+\fpt_{k+1,k}} \prn{\frac{\eq_{k}\frg_{k,k+1}}{f\pot }+\frac{\eq_{k+1}\frg_{k+1,k}}{f\dep }}
     =%\\=&\,
-    \frac{1}{rf^+f^-},\\
-  \sum_k c_k \eq_k =&\, \sum_{ij} \eq_i q_{ij} (\eta-\eta) = 0,\\
-%  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f^+f^-},
-  \implies c_k =&\, \frac{k - \sum_j j\eq_j}{rf^+f^-},
+    \frac{1}{rf\pot f\dep },\\
+  \sum_k c_k \eq_k =&\, \sum_{ij} \eq_i \enc_{ij} (\eta-\eta) = 0,\\
+%  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f\pot f\dep },
+  \implies c_k =&\, \frac{k - \sum_j j\eq_j}{rf\pot f\dep },
 \end{aligned}
@@
 This gives an area of
-\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_+\eq_-)}(n-1)/r$, which would have no effect on what follows.}
+%\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_+\eq_-)}(n-1)/r$, which would have no effect on what follows.}
 
@@
 \begin{equation}\label{eq:multistateextarea}
-  A \leq \frac{\sqrt{N}}{r}(n-1)\prn{4\eq_+\eq_-}.
+  A \leq \frac{\sqrt{N}}{r}(M-1)\prn{4\eq_+\eq_-}.
 \end{equation}
@@
 \begin{equation}\label{eq:maxarea}
-  A \leq \frac{\sqrt{N}}{r}(n-1).
+  A \leq \frac{\sqrt{N}}{r}(M-1).
 \end{equation}
@@
 
-\appendix
-\section{Other quantities}
-
-Here we collect useless facts.
-
-\subsection{Differences in \texorpdfstring{$c_k$}{c(k)}'s}\label{sec:areacoeffdiff}
-
-In the general case, making full use of \cite[Cor.6.2.7]{kemeny1960finite}, we can show that:
-%
-\begin{equation}\label{eq:areacoeffdiff}
-  \begin{aligned}
-    c_{k+1} - c_k &= \frac{N^{k+1}_{kk}}{\eq_k} \brk{
-      \sum_{i<j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{H^{k+1}_{i,k} + H^{k}_{j,k+1} - 1}
-      +\sum_{i>j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{H^{k+1}_{j,k} + H^{k}_{i,k+1} - 1}
-    } \\
-    &= \frac{N^{k+1}_{kk}}{\eq_k} \brk{
-      \sum_{i<j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{1 - H^{k+1}_{j,k} + H^{k}_{i,k+1}}
-      +\sum_{i>j} \eq_i \frac{\W^F_{ij}}{f^+} \prn{1 - H^{k+1}_{i,k} + H^{k}_{j,k+1}}
-    }, \\
-    \eta^+_{k} - \eta^+_{k+1} &=
-      \fpt_{k,k+1} \eq_+ - \sum_{\set{i}{\w_i>0}} N^{k+1}_{ii} \prn{1 - H^{i}_{k,k+1}} \\&=
-      \sum_{\set{i}{\w_i>0}} N^{k}_{ii} \prn{1 - H^{i}_{k+1,k}} - \fpt_{k+1,k} \eq_+ \\
-    = \eta^-_{k+1} - \eta^-_{k} &=
-      \fpt_{k+1,k} \eq_- - \sum_{\set{i}{\w_i<0}} N^{k}_{ii} \prn{1 - H^{i}_{k+1,k}} \\&=
-      \sum_{\set{i}{\w_i<0}} N^{k+1}_{ii} \prn{1 - H^{i}_{k,k+1}} - \fpt_{k,k+1} \eq_- .
-  \end{aligned}
-\end{equation}
-%
-We need to show that the positivity of one ($\forall k$) implies that the positivity of the other.
-
-\subsection{Derivatives \wrt \texorpdfstring{$\fptb$}{T}}\label{sec:derivT}
-
-The Markov chain, $\W^F$, is completely determined by the off-diagonal mean first passage times, $\fptb$, \cite[Th.4.4.12]{kemeny1960finite}:
-%
-\begin{equation}\label{eq:fptdet}
-  \begin{aligned}
-    \eq &= \eta \prn{\fptb\inv\onev}\trans, \\
-    r\W^F= \MM &= (\D-\onem)\fptb\inv.
-  \end{aligned}
-\end{equation}
-%
-where $\eta$ can be determined by normalising $\eq$ and $\D$ is determined by $\eq$, \eqref{eq:pdotD}. Clearly $q$ will be undetermined.
-
-We can then define derivatives \wrt $\fptb$:
-%
-\begin{equation}\label{eq:derivTdef}
-  \pdiff{}{\fptb_{gh}} = \sum_{ij} \pdiffc{\W^F_{ij}}{\fptb_{gh}} \pdiff{}{\W^F_{ij}},
-\end{equation}
-%
-where
-%
-\begin{equation}\label{eq:derivTWp}
-  \pdiff{\W^F_{ij}}{\fptb_{gh}} = r \eq_h \W^F_{ig} \prn{\W^F_{ij}-\W^F_{hj}} ,
-  \qquad
-  \pdiff{\eq_{i}}{\fptb_{gh}} = -r \eq_h \eq_i \W^F_{ig} .
-\end{equation}
-%
-These derivatives of $q$ will vanish.
-
-This means that the derivatives of the area are:
-%
-\begin{equation}\label{derivTarea}
-  \pdiff{\hat{A}}{\fptb_{gh}} = -\eq_h\brk{
-    \sum_i \eq_i q_{ig} \w_h
-    + r\sum_{ijk} \prn{\W^F_{ig}+\W^F_{kg}} \eq_i q_{ij} \prn{\fptb_{ik}-\fptb_{jk}} \eq_k \w_k
-    }.
-\end{equation}
-%
-
-\subsection{Initial SNR}\label{sec:initSNR}
-
-Let us write $\SNR(0) = \sqrt{N} (2f^+f^-) \hat{I}$. Using \eqref{eq:SNRcurve}, we can write
-%
-\begin{equation}\label{eq:initSNR}
-  \begin{aligned}
-    \hat{I} &= \diff{\eq}{\alpha}(-\W^F)\w &
-      &= - \pib \fund \diff{\MM}{\alpha} \fund \W^F \w &
-      &= - \pib \fund \diff{\W^F}{\alpha} \fund \MM \w \\
-      &= \pib \fund \diff{\W^F}{\alpha} (\I - \tau\onev\pib) \w &
-      &= \pib \fund \diff{\W^F}{\alpha} \w &
-      &= \eq \diff{\W^F}{\alpha} \w .
-  \end{aligned}
-\end{equation}
-%
-We can also write this in component form
-%
-\begin{equation}\label{eq:initSNRcmpt}
-  \begin{aligned}
-     \hat{I} &= \sum_{i,j} \eq_i q_{ij} \w_j \\
-      &= \sum_{i \neq j} \eq_i q_{ij} \w_j + \sum_i \eq_i q_{ii} \w_i \\
-      &= \sum_{i \neq j} q_{ij} \eq_i (\w_j - \w_i).
-  \end{aligned}
-\end{equation}
-%
-We can differentiate the expression in terms of $\fund$ to get
-%
-\begin{equation}\label{eq:initSNRdiff}
- \begin{aligned}
-  \pdiff{\hat{I}}{\W^\pm_{gh}}
-    &= rf^\pm \sum_{ijk} \pib_k \fund_{kg} (\fund_{hi}-\fund_{gi}) q_{ij} \w_j
-    \pm \sum_k \pib_k \fund_{kg} (\w_h - \w_g) \\
-    &= rf^\pm \sum_{ij} \eq_g q_{ij} (\fptb_{gi}-\fptb_{hi}) \eq_i \w_j
-    \pm \eq_g (\w_h - \w_g).
- \end{aligned}
-\end{equation}
-%
-If we wished to maximise the area with fixed initial SNR, $\hat{I}=I_0$, we would add $\beta(\hat{I}-I_0)$ to the Lagrangian \eqref{eq:lagrangian} and extremise wrt $\beta$.
-
-\subsection{Fixing mixing time (Kemeny's  constant)}\label{sec:fixmix}
-
-The mixing time, $\eta$ \eqref{eq:mixdef}, can be thought of as the mean time to pass between two states drawn from the equilibrium distribution \cite{hunter2006mixing,levene2002kemeny}.
-It should be a measure of the time it takes to reach equilibrium, and therefore related to the time it takes the SNR to decay.
-It could be interesting to hold it fixed while maximising the area, as this could help prevent degenerate solutions with very low initial SNR but long decay time.
-This can be done by adding $\gamma(\eta-\eta_0)$ to the Lagrangian \eqref{eq:lagrangian} and extremising wrt $\gamma$.
-
-It would help to calculate its derivatives.
-%
-\begin{equation}\label{eq:mixingdiff}
-  \begin{aligned}
-    \pdiff{\eta}{\W^\pm_{gh}} &= \pdiff{}{\W^\pm_{gh}}\brk{\sum_i \fund_{ii} - \tau} \\
-      &= rf^\pm \sum_i \fund_{ig}(\fund_{hi}-\fund_{gi}) \\
-      &= rf^\pm (\fund^2_{hg} - \fund^2_{gg}).
-  \end{aligned}
-\end{equation}
-%
-
-\subsection{Scaling mode}\label{sec:rescaleconstr}
-
-Consider the following differential operator:
-%
-\begin{equation}\label{eq:scaleopapp}
-  \Delta \equiv \sum_{g,h} \W^+_{gh}\pdiff{}{\W^+_{gh}} + \W^-_{gh}\pdiff{}{\W^-_{gh}}.
-\end{equation}
-%
-This corresponds to the scaling, $\W^\pm \to (1+\epsilon)\W^\pm$.
-Intuitively, this has two effects: it scales up the initial potentiation/depression and it scales down all timescales.
-This intuition is confirmed by the following results:
-%
-\begin{equation}\label{eq:scaleeffectsapp}
-  \begin{aligned}
-    \Delta \fund &= \tau\onev\eq - \fund ,&
-    \Delta \eq  &= 0 ,&
-    \Delta \fpt  &= -\fpt ,&
-    \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
-    \Delta q_{ij} &= q_{ij} ,
-    \\&&
-    \Delta \hat{A}  &= 0 ,&
-    \Delta \hat{I} &= \hat{I} ,&
-    \Delta \eta  &= -\eta .
-  \end{aligned}
-\end{equation}
-%
-The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W^\pm$.
-
-Suppose we don't hold anything fixed when maximising the area. Then we'd have
-%
-\begin{equation*}
-  \Delta \CL = 0,
-\end{equation*}
-%
-i.e. this scaling is a null direction. Area maximisation can only determine the ratios of transition rates, not their absolute values.
-
-Now suppose we only fix $\hat{I}$ when maximising the area. Then we'd have
-%
-\begin{equation*}
-  \Delta \CL = \beta \hat{I} = \beta I_0 = 0.
-\end{equation*}
-%
-As the Lagrange multiplier is zero, the constraint doesn't change the derivatives of the Lagrangian wrt $\W^\pm$.
-The sole effect of the constraint would be to fix the scaling mode.
-It wouldn't change the ratios of transition rates, topology, etc.
-
-Exactly the same conclusions would follow if we only fixed $\eta$ when maximising the area.
-This would also only fix the scaling mode.
-However, if we held them both fixed, it would have more of an effect:
-%
-\begin{equation}\label{eq:scaleconstr}
-  \Delta \CL = \beta \hat{I} - \gamma \eta = \beta I_0 - \gamma \eta_0 = 0.
-\end{equation}
-%
-The Lagrange multipliers are non-zero, so this could affect the topology of the optimal Markov chain.
-
-%\subsection*{Acknowledgements}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection*{Appendices}
-%\appendix
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-05-30 NIPS/SynapseNIPS.tex
nothing
almost final version

--- 2013-05-29 NIPS/SynapseNIPS.tex
@@
 % >> Only for drafts! <<
-\usepackage[notref,notcite]{showkeys}
+%\usepackage[notref,notcite]{showkeys}
 % ----------------------------------------------------------------
@@
 Applied Physics Department, Stanford University, Stanford CA\\
-\emaillink{sulahiri@stanford.edu}, \emaillink{sulahiri@stanford.edu}
+\emaillink{sulahiri@stanford.edu}, \emaillink{sganguli@stanford.edu}
 %
@@
 \begin{abstract}
-  Blah blah blah.
+  An incredible gulf separates theoretical models of synapses, often described solely by a single scalar value denoting the size of a postsynaptic potential, from the immense complexity of molecular signaling pathways underlying real synapses.
+  To understand the functional contribution of such molecular complexity to learning and memory, it is essential to expand our theoretical conception of a synapse from a single scalar to an entire dynamical system with many internal molecular functional states.
+  Moreover, theoretical considerations alone demand such an expansion; network models with scalar synapses assuming finite numbers of distinguishable synaptic strengths have strikingly limited memory capacity.
+  This raises the fundamental question, how does synaptic complexity give rise to memory?
+  To address this, we develop new mathematical theorems elucidating the relationship between the structural organization and memory properties of complex synapses that are themselves molecular networks.
+  Moreover, in proving such theorems, we uncover a framework, based on first passage time theory, to impose an order on the internal states of complex synaptic models, thereby simplifying the relationship between synaptic structure and function.
+
+%Overall, we uncover general design principles governing the functional organization of complex molecular networks, and suggest new experimental observables in synaptic physiology, based on first %passage time theory, that connect molecular complexity to memory.
 \end{abstract}
@@
 This one simple modification leads to a catastrophe in memory capacity:  classical models with digital synapses, when operating in a palimpset mode in which the ongoing storage of new memories can overwrite previous memories, have a memory capacity proportional to the logarithm of the number of synapses \cite{amit1992constraints,amit1994learning}.
-Intuitively, when synapses are digital, the storage of a new memory can flip a population of synaptic switches, thereby rapidly erasing previous memories.
+Intuitively, when synapses are digital, the storage of a new memory can flip a population of synaptic switches, thereby rapidly erasing previous memories stored in the same synaptic population.
 This result indicates that the dominant theoretical basis for the storage of long term memories in modifiable synaptic switches is flawed.
@@
 What is the structural organization of synaptic dynamical systems that achieve these limits?
-Moreover, from an experimental perspective, it is unlikely to be the case that there is a single canonical synaptic model;  just like the case of neurons, there is an incredible diversity of molecular networks underlying synapses both across species and across brain regions within a single organism \cite{Emes2012synapserev}.
+Moreover, from an experimental perspective, it is unlikely that all synapses can be described by a single canonical synaptic model;  just like the case of neurons, there is an incredible diversity of molecular networks underlying synapses both across species and across brain regions within a single organism \cite{Emes2012synapserev}.
 In order to elucidate the functional contribution of this diverse molecular complexity to learning and memory, it is essential to move beyond the analysis of specific models and instead develop a general theory of learning and memory for complex synapses.
@@
 Here $\frg$ is a continuous time transition matrix that models the process of forgetting the memory stored at time $t=0$ due to random candidate potentiation/depression events occurring at each synapse due to the storage of subsequent memories.
+Its stationary distribution is $\eq$.
 
@@
 We will frequently refer to this function as the memory curve.
-It can be thought of as the fraction of synapses that maintain the ideal synaptic strength at time $t$, dictated by the stored memory at time $t=0$.
+It can be thought of as the excess fraction of synapses (relative to equilibrium) that maintain their ideal synaptic strength at time $t$, as dictated by the stored memory at time $t=0$.
 
@@
 The critical question is: what do these constraints imply about the space of achievable memory curves in \eqref{eq:SNRcalc}?
-To answer this question, especially for limits on achievable memory at finite times, it will be useful to employ the eigenmode description:
+To answer this question, especially for limits on achievable memory at finite times, it will be useful to employ the eigenmode decomposition:
 %
@@
 Instead there are tradeoffs between storing proximal and distal memories;  often attempts to increase memory at late (early) times by changing $\M\potdep$, incurs a performance loss in memory at early (late) times in specific models considered so far \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial}.
-Thus our goal, achieved in \autoref{sec:env}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the entire memory curve of {\it any} model.
+Thus our end goal, achieved in \autoref{sec:env}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the {\it entire} memory curve of {\it any} model.
 In order to achieve this goal, in this section, we must first derive upper bounds, over the space of all possible synaptic models, on two different scalar functions of the memory curve: its initial SNR, and the area under the memory curve.
-As a corollary of these results, we also derive an upper-bound on the memory lifetime, or the time at which the SNR drops to a given detection threshold.
+In the process of upper-bounding the area, we will develop an essential framework to organize the structure of synaptic dynamical systems based on first passage time theory.
 
-\subsection{Initial SNR}\label{sec:initial}
+\subsection{Bounding initial SNR}\label{sec:initial}
 
-Now we will discuss the SNR at $t=0$:
+We now give an upper bound on the initial SNR,
 %
 \begin{equation}\label{eq:init}
-  \snr(0) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \w.
+  \snr(0) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \w,
 \end{equation}
 %
-We will find an upper bound on this quantity for \emph{all} possible models and also find the model that saturates this bound.
-As maximizing initial SNR does not place any value on later parts of the memory curve, this will produce models with fast decay that may not be good for memory storage.
-Nevertheless, the upper bound will be useful in \autoref{sec:env}.
-
+over \emph{all} possible models and also find the model that saturates this bound.
 A useful quantity is the equilibrium probability flux between two disjoint sets of states, $\CA$ and $\CB$:
@@
 %
-The initial SNR is closely related to the flux between the states with $\w_i=-1$ and those with $\w_j=+1$:
+The initial SNR is closely related to the flux from the states with $\w_i=-1$ to those with $\w_j=+1$:
 %
@@
 %
-Whilst this model has high initial SNR, it also has very fast decay -- with a timescale $\tau\sim\frac{1}{r}$.
-As the synapse is very plastic, the initial memory is encoded very easily, but the subsequent memories also overwrite it easily.
-This is not a good design for a synapse, but the resulting bound on the initial SNR will prove useful in \autoref{sec:env}.
-
+We note that while this model has high initial SNR, it also has very fast memory decay -- with a timescale $\tau\sim\frac{1}{r}$.
+As the synapse is very plastic, the initial memory is encoded very easily, but the subsequent memories also overwrite it rapidly.
+This is one example of the tradeoff between optimizing memory at early versus late times.
 \begin{figure}[tbp]
@@
 
-\subsection{Area}\label{sec:area}
+\subsection{Imposing order on internal states through first passage times}\label{sec:order}
+
+Our goal of understanding the relationship between structure and function in the space of all possible synaptic models is complicated by the fact that this space contains many different possible network topologies, encoded in the nonzero matrix elements of $\M\potdep$.
+To systematically analyze this entire space, we develop an important organizing principle using the theory of first passage times in the stochastic process of forgetting, described by $\frg$.
+The mean first passage time matrix, $\fptb_{ij}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $j$.
+The diagonal elements are defined to be zero.
+
+A remarkable theorem we will exploit is that the quantity
+%
+\begin{equation}\label{eq:kemenyconst}
+  \eta \equiv \sum_j \fptb_{ij}\eq_j
+\end{equation}
+%
+is independent of the starting state $i$.
+It is known as Kemeny's constant (see \cite[\S4.4]{kemeny1960finite}).
+Intuitively, \eqref{eq:kemenyconst} states that the average time it takes to reach any state, weighted by its equilibrium probability, is independent of the starting state, implying a hidden constancy inherent in any stochastic process.
+
+In the context of complex synapses, we can define the partial sums
+%
+\begin{equation}\label{eq:kemenypm}
+   \eta^+_i = \sum_{j\in+} \fptb_{ij} \eq_j,
+   \qquad
+   \eta^-_i = \sum_{j\in-} \fptb_{ij} \eq_j.
+\end{equation}
+%
+These can be thought of as the average time it takes to reach the strong/weak states respectively.
+Using these definitions, we can then impose an order on the states by arranging them in order of decreasing $\eta^+_i$ or increasing $\eta^-_i$.
+Because $\eta^+_i+\eta^-_i=\eta$ is independent of $i$, the two orderings are the same.
+In this order, which depends sensitively on the structure of $\M\potdep$, states later (to the right in figures below) can be considered to be more potentiated than states earlier (to the left in figures below), despite the fact that they have the same synaptic efficacy.
+In essence, in this order, a state is considered to be more potentiated if the average time it takes to reach all the strong efficacy states is shorter. 
+We will see that synaptic models that optimize various measures of memory have an exceedingly simple structure when, and only when, their states are arranged in this order.
+
+\subsection{Bounding area}\label{sec:area}
 
@@
 We will find an upper bound on this quantity as well as the model that saturates it.
-The area is useful as it also provides an upper bound on the lifetime of a memory, $\tau(\epsilon)$, defined as the time at which $\snr(\tau)=\epsilon$ (we often take $\epsilon=1$):
+
+First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \eqref{eq:kemenypm}:
 %
-\begin{equation}\label{eq:arealife}
-  A = \int_0^\infty\!\!\dr t \, \snr(t)
-    > \int_0^{\tau(\epsilon)}\!\!\dr t \, \snr(t)
-    > \int_0^{\tau(\epsilon)}\!\!\dr t \, \epsilon
-    = \epsilon \tau(\epsilon).
+\begin{equation}\label{eq:areaEta}
+\begin{aligned}
+  A &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \prn{\M\pot_{ij}-\M\dep_{ij}} \prn{\eta^+_i-\eta^+_j}\\
+    &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \prn{\M\pot_{ij}-\M\dep_{ij}} \prn{\eta^-_j-\eta^-_i}.
+\end{aligned}
 \end{equation}
 %
-However, the area still receives a contribution from the tail of the memory curve, after the SNR has decayed below any useful value.
-This means that maximizing SNR will favor models with long tails that will not necessarily be good for storing memories.
 
-One important property of the area is that it has a null ``scaling'' degree of freedom.
-If all off-diagonal elements of $\M\potdep$ are multiplied by a constant, with diagonal elements adjusted to keep row-sums fixed, this will have two effects.
-First, it will scale up the initial signal creation.
-Second, it will scale down time.
-Combining these two effects, it will leave the area invariant.
-
-This feature allows us to ignore the lower bound on the diagonal elements, $\M\potdep_{ii}>0$.
-If we maximize the area ignoring this constraint, the ``scaling'' mode can be used to satisfy it without changing the area.
-
-We also develop an important organizing principle using the theory of first passage times.
-The mean first passage time matrix, $\fptb_{ij}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $j$.
-The diagonal elements are defined to be zero.
-Then the quantity
-%
-\begin{equation}\label{eq:kemenyconst}
-  \eta \equiv \sum_j \fptb_{ij}\eq_j
-\end{equation}
-%
-is independent of the starting state $i$.
-It is known as Kemeny's constant (see \cite[\S4.4]{kemeny1960finite}).
-We can define analogous quantities
-%
-\begin{equation}\label{eq:kemenypm}
-   \eta^+_i = \sum_{j\in+} \fptb_{ij} \eq_j,
-   \qquad
-   \eta^-_i = \sum_{j\in-} \fptb_{ij} \eq_j.
-\end{equation}
-%
-These can be thought of as the average time it takes to reach the strong/weak states respectively.
-We can put the states in order of decreasing $\eta^+_i$ or increasing $\eta^-_i$.
-Because $\eta^+_i+\eta^-_i=\eta$ is independent of $i$, the two orderings are the same.
-
-With the states in this order, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
+With the states in the order described above, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
 Some of these perturbations are shown in \autoref{fig:perts}.
+For example, in \autoref{fig:perts}\ref{fig:tri_right}, for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
+$\eta^+_i > \eta^+_j$), we have proven that increasing $\M\pot_{ij}$ and decreasing $\M\dep_{ij}$ leads to an increase in area.
 The only thing that can prevent these perturbations from increasing the area is when they require the decrease of a matrix element that has already been set to 0.
@@
 As these perturbations do not change the equilibrium distribution, this means that the area of \emph{any} model is bounded by that of a linear chain with the same equilibrium distribution.
-This area is given by
+The area of a linear chain model can be expressed directly in terms of its equilibrium state distribution, $\eq$, yielding the following upper bound on the area of any model with the same $\eq$:
 %
@@
 %
-This result is similar to that found numerically by Barrett and van Rossum \cite{Barrett2008discrete} in a slightly different situation.
+This analytical result is similar to a numerical result found in \cite{Barrett2008discrete} under a slightly different information theoretic measure of memory performance.
 
-The ``sticky'' end states result in very slow decay, but they also make it difficult to encode the memory in the first place.
-The memory curve for this model starts off very low, but hovers above zero for a very long time.
-This is also not a good design for a synapse, but the resulting bound on the area will also prove useful in \autoref{sec:env}.
+The ``sticky'' end states result in very slow decay of memory, but they also make it difficult to encode the memory in the first place, since a small fraction of synapses are able to change synaptic efficacy during the storage of a new memory.
+Thus models that optimize area optimize memory at late times, at the expense of early times.
 
@@
  \begin{center}
-  \includegraphics[width=0.4\linewidth]{multistate_sticky-generated.pdf}
+  \includegraphics[width=0.35\linewidth]{multistate_sticky-generated.pdf}
  \end{center}
@@
 
-Now we will look at the implications of the upper bounds found in the previous section for the SNR at finite times.
+Now we will look at the implications of the upper bounds found in the \hyperref[sec:bounds]{previous section} for the SNR at finite times.
 As argued in \eqref{eq:SNReigen}, the memory curve can be written in the form
@@
 Now we can pick some fixed time, $t_0$, and maximize the SNR at that time \wrt the parameters $\brc{\CI_a,\tau_a}$, subject to the constraints above.
-This always results in a single nonzero $\CI_a$:
+This always results in a single nonzero $\CI_a$; in essence, optimizing memory at a single time requires a single exponential.
+The resulting optimal memory curve, along with the achieved memory at the chosen time, depends on $t_0$ as follows:
 %
@@
 At late times, only the area bound is saturated.
-The function $\snr(t_0)$, the green curve in \autoref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope.
+The function $\snr(t_0)$, the green curve in \autoref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim t_0\inv$.
 No synaptic model can have an SNR that is greater than this at any time.
@@
 where we assume $N>(\epsilon\e)^2$.
+Intriguingly, both the lifetime and memory envelope expand linearly with the number of internal states $M$, and increase as the square root of the number of synapses $N$.
 
@@
 At any time, can we find a model whose memory curve touched the envelope?
-The red curves in \autoref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization with random initialization and by hand designed models.
+The red curves in \autoref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
 
-We see that here is a gap between the upper bound that we can prove and what we can achieve with actual models.
-There may be other models we haven't found that could beat the ones we have, but it is probably too much to expect them to reach this envelope.
-That would require saturating the area bound, and we saw that the only model that does that has very low initial SNR.
-It seems that we need to find additional constraints to add to those we have \eqref{eq:modeconstr}, thereby bringing the envelope down.
-We do not necessarily need to find \emph{every} constraint.
-We have already seen that not every constraints is saturated for certain time ranges, so it may be that a small subset of them will be sufficient to find an envelope that is tight.
+We see that at early, but not late times, there is a gap between the upper bound that we can prove and what we can achieve with actual models.
+There may be other models we haven't found that could beat the ones we have, and come closer to our proven envelope.
+However, we suspect that the area constraint is not the bottleneck for optimizing memory at times less than  $\CO(\frac{M}{r})$.
+We believe there is some other constraint that prevents models from approaching the envelope, and currently are exploring several mathematical conjectures for the precise form of this constraint, in order to obtain a potentially tighter envelope.
+Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \autoref{fig:max_area} can come close to the envelope.
 
@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.4\linewidth]{env.eps}}\label{fig:envplot}\hp
+  \item\aligntop{\includegraphics[width=0.38\linewidth]{env.eps}}\label{fig:envplot}\hp
   \item\aligntop{\includegraphics[width=0.3\linewidth]{diffjump-generated.pdf}}\label{fig:diffjump}
@@
   The solid red curve shows the highest SNR we have found with hand designed models.
-  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states.
+  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states, and all transition probabilities equal to $1$.
   At late times they are of the form shown in \autoref{fig:max_area} with different values of $\varepsilon$.
@@
 
+We have initiated the development of  a general theory of learning and memory with complex synapses, allowing for an exploration of the entire space of complex synaptic models, rather than analyzing individual models one at a time.
+In doing so, we have obtained several new mathematical results delineating the functional limits of memory achievable by synaptic complexity, and the structural characterization of synaptic dynamical systems that achieve these limits.
+In particular, operating within the ideal observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}, we have shown that for a population of $N$ synapses with $M$ internal states, (a) the initial SNR of any synaptic model cannot exceed $\sqrt{N}$, and any model that achieves this bound is equivalent to a binary synapse, (b) the area under the memory curve of any model cannot exceed that of a linear chain model with the same equilibrium distribution, (c) both the area and memory lifetime of any model cannot exceed $\CO(\sqrt{N}M)$, and the model that achieves this limit has a linear chain topology, (d) we have derived an envelope memory curve in the SNR-time plane that cannot be exceeded by the memory curve of any model, and models that approach this envelope for times greater than $\CO(\frac{M}{r})$ are linear chain models, and (e) this late-time envelope is a power-law proportional to $O({\sqrt{N} M}/{rt})$, indicating that synaptic complexity can strongly enhance the limits of achievable memory.
+
+This theoretical study opens up several avenues for further inquiry.
+In particular, the tightness of our envelope for early times, less than $\CO(\frac{M}{r})$, remains an open question, and we are currently pursuing several conjectures.
+We have also derived memory constrained envelopes, by asking in the space of models that achieve a given SNR at a given time, what is the maximal SNR achievable at other times.
+If these two times are beyond a threshold separation, optimal constrained models require two exponentials.
+It would be interesting to systematically analyze the space of models that achieve good memory at multiple times, and understand their structural organization, and how they give rise to multiple exponentials, leading to power law memory decays.
+Finally, it would be interesting to design physiological experiments in order to perform optimal systems identification of potential Markovian dynamical systems hiding within biological synapses.
+
+In summary, we hope that a deeper theoretical understanding of the functional role of synaptic complexity, initiated here, will help advance our understanding of the neurobiology of learning and memory, aid in the design of engineered memory circuits, and lead to new mathematical theorems about stochastic processes.
+
 %\subsubsection*{Acknowledgements}
+%We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
 

### 2013-05-30 NIPS/SynapseNIPSsupp.tex
almost final version

@@
 
+Although the theory and derivations outlined in the main text are relatively self-contained, here we provide more details for the aid of the reviewer.
+
 \section{Continuous time Markov processes}\label{sec:ContMarkov}
@@
 
-Consider the first time the process leaves state $i$.
-On average, this will take time $\Lb_{ii}$.
-With probability $\Pb_{ij}$, it will go directly to $j$, so the conditional mean time would be $\Lb_{ii}$.
-On the other hand, if it goes to some other state, $k$, with probability $\Pb_{ik}$, the conditional mean time would be $\fptb_{kj}+\Lb_{ii}$.
-Combining these, we get the recursion relation
-%
-\begin{equation}\label{eq:fptrecursion}
-  \begin{aligned}
-  \fpt_{ij} &= \sum_{k\neq j} \Pb_{ik}(\fptb_{kj}+\Lb_{ii}) + \Pb_{ij}\Lb_{ii} \\
-    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \sum_{k} \Pb_{ik}\Lb_{ii}\\
-    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \Lb_{ii},\\
-   \fpt &= \Pb \fptb + \Lb\onem,
-  \end{aligned}
-\end{equation}
-%
-or, using \eqref{eq:defDLP},
-%
-\begin{equation}\label{eq:fptrecdg}
-  \fpt\dg = \Lb\MM\fptb+\Lb\onem.
-\end{equation}
-%
-
-\begin{thm}
+This matrix is given by
+%
+\begin{equation}\label{eq:fpt}
+  \fpt=(\onem\fund\dg - \fund + \Lb)\D,
+\end{equation}
+%
+see \cite{Yao1985fpt} for a proof.
+We can separate this into its diagonal and off-diagonal pieces.
+
 The recurrence times are given by
@@
 \end{equation}
-  %
-\end{thm}
-\begin{proof}
-  Pre-multiply \eqref{eq:fptrecdg} by $\eq\Lb\inv$:
-  %
-  \begin{equation*}
-    \eq\Lb\inv\fpt\dg = \onev\trans,
-  \end{equation*}
 %
@@
 %
-  This is the same as \eqref{eq:recurtime} in component form: $\fpt\dg_{ii} = \Lb_{ii}/\eq_i$.
-\end{proof}
-I think the extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.
-
-We can substitute this into \eqref{eq:fptrecdg} to get a recursion relation for the off diagonal part:
-%
-\begin{equation}\label{eq:fptbrec}
-  \MM\fptb = \D-\onem.
-\end{equation}
-%
-\begin{thm}
-  The off-diagonal solution of \eqref{eq:fptbrec} is unique.
-\end{thm}
-\begin{proof}
-  Suppose we have two solutions for the off-diagonal part, $\fptb$ and $\fptb'$.
-  Subtracting \eqref{eq:fptbrec} for these two solutions gives
-  %
-  \begin{equation*}
-    \MM(\fptb-\fptb')=0.
-  \end{equation*}
-  %
-  As we are assuming an ergodic process, the zero eigenvalue is non-degenerate.
-  Therefore
-  %
-  \begin{equation*}
-    \fptb-\fptb' = \onev\mathbf{y},
-  \end{equation*}
-  %
-  for some row-vector $\mathbf{y}$. Looking at the diagonal components
-  %
-  \begin{equation*}
-    \mathbf{y}_{i} = \fptb_{ii}-\fptb'_{ii} = 0,
-  \end{equation*}
-  %
-  and therefore $\fptb-\fptb'=0$.
-\end{proof}
-
-\begin{thm}
+The extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.
+
 The off-diagonal mean first passage times are given by
@@
 %
-\end{thm}
-\begin{proof}
-  We just need to show that this satisfies \eqref{eq:fptbrec} and that the diagonal elements vanish.
-  Premultiply \eqref{eq:fptfund} by $\MM$:
-  %
-  \begin{equation*}
-  \begin{aligned}
-    \MM\fptb &= -\MM\fund\D\\
-     &= (\I-\onev\eq)\D \\
-     &= \D-\onem,
-  \end{aligned}
-  \end{equation*}
-  %
-  where we used \eqref{eq:fundQZ} and \eqref{eq:pdotD}.
-  The vanishing of the diagonal elements follows trivially from the component form of \eqref{eq:fptfund}:
+or in component form of \eqref{eq:fptfund}:
 %
@@
 %
-
-\end{proof}
 
@@
 
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators using \eqref{eq:diffpT}:
 %
@@
 
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators \eqref{eq:diffpT}:
 %

### 2013-05-31 NIPS/SynapseNIPS.tex
fixed typo, back to svg

@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.25\linewidth]{cascade-generated.pdf}}\label{fig:cascade}
-  \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate-generated.pdf}}\label{fig:multistate}
+  \item\aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
+  \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
   \item\aligntop{\includegraphics[width=0.35\linewidth]{cascms.eps}}\label{fig:cascms}
@@
  \begin{myenuma}
-  \item\hp\aligntop{\includegraphics[width=0.15\linewidth]{max_init_pot-generated.pdf}}\label{fig:max_init_pot}\hp
-  \item\aligntop{\includegraphics[width=0.15\linewidth]{max_init_dep-generated.pdf}}\label{fig:max_init_dep}\hp
-  \item\hp\aligntop{\includegraphics[width=0.05\linewidth]{binary_det-generated.pdf}}\label{fig:binary_det}
+  \item\hp\aligntop{\includegraphics[width=0.15\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
+  \item\aligntop{\includegraphics[width=0.15\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
+  \item\hp\aligntop{\includegraphics[width=0.05\linewidth]{binary_det.svg}}\label{fig:binary_det}
   \end{myenuma}
@@
 To systematically analyze this entire space, we develop an important organizing principle using the theory of first passage times in the stochastic process of forgetting, described by $\frg$.
-The mean first passage time matrix, $\fptb_{ij}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $j$.
+The mean first passage time matrix, $\fptb_{ij}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $i$.
 The diagonal elements are defined to be zero.
@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right-generated.pdf}}\label{fig:tri_right}\hp
-  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left-generated.pdf}}\label{fig:tri_left}\hp
-  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut-generated.pdf}}\label{fig:shortcut}
+  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
+  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
+  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut.svg}}\label{fig:shortcut}
  \end{myenuma}
@@
  \begin{center}
-  \includegraphics[width=0.35\linewidth]{multistate_sticky-generated.pdf}
+  \includegraphics[width=0.35\linewidth]{multistate_sticky.svg}
  \end{center}
@@
   \item\aligntop{\includegraphics[width=0.38\linewidth]{env.eps}}\label{fig:envplot}\hp
-  \item\aligntop{\includegraphics[width=0.3\linewidth]{diffjump-generated.pdf}}\label{fig:diffjump}
+  \item\aligntop{\includegraphics[width=0.3\linewidth]{diffjump.svg}}\label{fig:diffjump}
  \end{myenuma}
@@
 %We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
+%We thank larry Abbott, Marcus Benna, Stefano Fusi, Jascha Sohl-Dickstein and David Sussillo for useful discussions.
 

### 2013-07-09 MHCslides/MHCslides.tex
started mhc slides

% -*- TeX -*- -*- UK -*- -*- BMR -*-
% ----------------------------------------------------------------
% Beamer presentation ************************************************
% **** -----------------------------------------------------------
\documentclass{beamer}%[hyperref={backref=slide}]

\input{sl_slide_preamble.tex}
%\input{sl_slide_preamble_nonotes.tex}
\input{sl_slide_graphics_preamble.tex}
\input{sl_definitions.tex}
\input{sl_slide_symbols.tex}
%
\graphicspath{{"Figures/"}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\I}{\mathbf{I}}
%prob vector
\newcommand{\pr}{\mathbf{p}}
%equilibrium distribution
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\wv}{\vec{w}}
%super/subscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
%quantities
\newcommand{\initial}{\mathcal{I}}
\newcommand{\area}{\mathcal{A}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\comp}{^\mathrm{c}}
\renewcommand{\e}{\mathsf{e}}
%---------Title-----------------------------------------------------------

\title[Impaired learning with enhanced plasticity]{A saturation model for impaired learning with enhanced plasticity}
%
\subtitle{\small{based on work in preparation by: T.D. Barbara Nguyen-Vu, Grace Q. Zhao, Han-Mi Lee, SL, Surya Ganguli, Carla J. Shatz, Jennifer L. Raymond
}}
%
\author{Subhaneil Lahiri%\inst{1}
}
%
\institute[Stanford]{%
%\inst{1}
Stanford University, Applied Physics
}
%
%\slideCaption{}

%---------Beginning--------------------------------------------------------

\begin{document}

%-------------Slide--------------------------------------------------------

\begin{frame}
%
 \titlepage
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Outline}
%
 \tableofcontents[hideallsubsections]
 %
 \note[item]{review terrible properties of simple synapses.}
 \note[item]{mathematical formalism of model, quantify performance (memory decay over time)}
 \note[item]{upper bounds on single numbers that depend on whole memory curve (decay over time)}
 \note[item]{upper bounds at finite times}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}[allowframebreaks]{References}
%

 {\small
 \bibliographystyle{unsrt_slides}
 \bibliography{maths,neuro}
 }
%
\end{frame}

%-----End----------------------------------------------------------------

\end{document}

### 2013-07-10 MHCslides/MHCslides.tex
results in slides

@@
 %
-\graphicspath{{"Figures/"}}
+\graphicspath{{"Figs/"}}
 \DeclareMathOperator{\SNR}{SNR}
@@
  %
- \note[item]{review terrible properties of simple synapses.}
- \note[item]{mathematical formalism of model, quantify performance (memory decay over time)}
- \note[item]{upper bounds on single numbers that depend on whole memory curve (decay over time)}
- \note[item]{upper bounds at finite times}
+%
+\end{frame}
+
+%-------------Section--------------------------------------------------------
+
+\section{VOR learning and the cerebellum}
+
+%-------------Section--------------------------------------------------------
+
+\section{The effects of enhanced plasticity and saturation}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Questions}
+%
+ \begin{itemize}
+   \item Can we find a purely synaptic explanation of these results?
+   \item Can the saturation effect overcome the enhanced plasticity?
+   \item How can a little reverse bias help, but too much hurt?
+ \end{itemize}
+%
+\end{frame}
+
+%-------------Section--------------------------------------------------------
+
+\section{Modelling approach}
+
+%-------------Section--------------------------------------------------------
+
+\section{Modelling results}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Binary synapse}
+%
+ %
+ \begin{center}
+   \includegraphics[width=0.15\linewidth]{binary.svg} \\[1cm]
+   \aligntop{\includegraphics[width=0.35\linewidth]{binary_learnS.eps}}
+   \hspace{2cm}
+   \aligntop{\includegraphics[width=0.35\linewidth]{comparisons_binary.svg}}
+ \end{center}
+ %
+ \note[item]{understand why next slide}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Binary synapse: initial distributions}
+%
+ %
+ \begin{center}
+ \begin{tabular}{ll}
+   \aligntop{\includegraphics[height=0.25\linewidth]{binary_bar_wt_wo.svg}}
+   \hspace{2cm} &
+   \aligntop{\includegraphics[height=0.25\linewidth]{binary_bar_ko_wo.svg}}
+   \\[3cm]
+   \aligntop{\includegraphics[height=0.25\linewidth]{binary_bar_wt_w.svg}}
+   &
+   \aligntop{\includegraphics[height=0.25\linewidth]{binary_bar_ko_w.svg}}
+ \end{tabular}
+ \end{center}
+ %
+ \note[item]{understand why next slide}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Serial synapse}
+%
+ %
+ \begin{center}
+   \includegraphics[width=0.5\linewidth]{serial.svg} \\[1cm]
+   \aligntop{\includegraphics[width=0.35\linewidth]{serial_learnS.eps}}
+   \hspace{2cm}
+   \aligntop{\includegraphics[width=0.35\linewidth]{comparisons_serial.svg}}
+ \end{center}
+ %
+ \note[item]{understand why next slide}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Serial synapse: initial distributions}
+%
+ %
+ \begin{center}
+ \begin{tabular}{ll}
+   \aligntop{\includegraphics[height=0.2\linewidth]{serial_bar_wt_wo.svg}}
+   \hspace{1cm} &
+   \aligntop{\includegraphics[height=0.2\linewidth]{serial_bar_ko_wo.svg}}
+   \\[3cm]
+   \aligntop{\includegraphics[height=0.2\linewidth]{serial_bar_wt_w.svg}}
+   &
+   \aligntop{\includegraphics[height=0.2\linewidth]{serial_bar_ko_w.svg}}
+ \end{tabular}
+ \end{center}
+ %
+ \note[item]{understand why next slide}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Mathematics}
+%
+ Serial synapse: $\eq_i \sim \CN \prn{\frac{q\pot}{q\dep}}^i$.
+
+ \vp Learning rate $\sim \eq_{M/2}\prn{\frac{q\dep}{q\pot}} = \CN \prn{\frac{q\pot}{q\dep}}^{\frac{M}{2}-1}$.
+
+ \vp For $M>2$, larger $q\dep \implies$ slower learning.
+
+ \vp For $M=2$, larger $q\dep \implies$ larger $\CN \implies$ faster learning.
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Essential features}
+%
+ The success od the serial model relies on two features:
+ \begin{itemize}
+   \item Enhancing the effect of saturation,
+   \note[item]{due to exponential decay}
+   \item Metaplasticity -- repeated potentiation makes subsequent depression harder.
+   \note[item]{push away from boundary where signal generated}
+ \end{itemize}
+ %
+ \begin{center}
+   \includegraphics[width=0.4\linewidth]{serial.svg}
+ \end{center}
+ %
+ \note[item]{borne out by other models that fail/succeed}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Other models that fail}
+%
+ \begin{tabular}{ll}
+   % & for next tab, \\ for new line...
+    \aligntop{\includegraphics[width=0.4\linewidth]{multistate.svg}}\hspace{1.5cm} &
+    \aligntop{\includegraphics[width=0.28\linewidth]{multistate_learnS.eps}} \\[2cm]
+    \aligntop{\includegraphics[width=0.4\linewidth]{pooled.svg}} &
+    \aligntop{\includegraphics[width=0.28\linewidth]{pooled_scarce_learnS.eps}}
+ \end{tabular}
+
+ \citerr{amit1994learning}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Other models that work}
+%
+ \begin{tabular}{ll}
+   % & for next tab, \\ for new line...
+    \aligntop{\includegraphics[width=0.4\linewidth]{multistate_nonuni.svg}}\hspace{1.5cm} &
+    \aligntop{\includegraphics[width=0.28\linewidth]{nonuni_learnS.eps}} \\[2cm]
+    \aligntop{\includegraphics[width=0.4\linewidth]{cascade.svg}} &
+    \aligntop{\includegraphics[width=0.28\linewidth]{cascade_long_learnS.eps}}
+ \end{tabular}
+
+ \citerr{Fusi2005cascade}
 %

### 2013-07-10 Notes/VORmodel.tex
results in slides

@@
                                           & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
+    Non-uniform MS       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:nonuni} \\
     \cline{1-7}
@@
 
+\subsubsection{Non-uniform linear multistate model}\label{sec:nonuni}
+
+\resultsfign{Simulation results for a non-uniform linear multistate model}{nonuni}
+
 %\subsection*{Acknowledgements}

### 2013-07-10 MHCslides/MHCslides.tex
mhc slides except for expt section

@@
 
+\usepackage{array}
 \input{sl_slide_preamble.tex}
@@
 
-\title[Impaired learning with enhanced plasticity]{A saturation model for impaired learning with enhanced plasticity}
+\title[Saturation by enh. plasticity impairs learning]{A saturation model for impaired learning with enhanced plasticity}
 %
@@
 
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Vestibular Occular Reflex}
+%
+ \parbox[t]{0.4\linewidth}{\aligntop{\includegraphics[width=0.99\linewidth]{VOR.svg}}}
+ \parbox[t]{0.59\linewidth}{%
+ Eye movements compensate for head movements to maintain fixation.
+
+ \vp Requires control of $\text{VOR gain} = \frac{\text{eye velocity}}{\text{head velocity}}$.
+
+ \vp Needs to be adjusted as eye muscles age, \etc
+ }
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{VOR training}
+%
+ \parbox[t]{0.2\linewidth}{\aligntop{\includegraphics[width=0.99\linewidth]{VORinc.svg}}\\[1cm]
+   \note[item]{trick brain into thinking VOR gain needs adjusting my moving visual stimulus}
+   \note[item]{anti-phase \lto increase gain}
+ \aligntop{\includegraphics[width=0.99\linewidth]{VORdec.svg}}}
+ \note[item]{in phase \lto decrease gain}
+ \hspace{0.15\linewidth}
+ \parbox[t]{0.59\linewidth}{\aligntop{\includegraphics[width=0.99\linewidth]{VORcircuit.svg}}}
+ \note[item]{Gain change involves cerebellum}
+ \note[item]{Marr-Albus-Ito: Pf-Pk synapses}
+ \note[item]{Lisberger-Miles: Vestibular input-VN synapses}
+ \note[item]{Different mechs for different freq, head angle, gain up/down.}
+ \note[item]{Different Pk cells have different tunings.}
+ \note[item]{Gain up in case of interest: LTD in Pf-Pk in flucculus}
+ \note[item]{Gain down: uses different mech for behaviour, but does reverse LTD in Pf-Pk in flucculus}
+%
+\end{frame}
+
 %-------------Section--------------------------------------------------------
@@
  \begin{itemize}
+   \item Can the saturation effect overcome the enhanced plasticity?
+   \note[item]{in competition}
+   \item How can a little reverse bias help, but too much hurt?
+   \note[item]{first makes sense, but second?}
    \item Can we find a purely synaptic explanation of these results?
-   \item Can the saturation effect overcome the enhanced plasticity?
-   \item How can a little reverse bias help, but too much hurt?
+   \note[item]{This is a question about synaptic populations after all.}
  \end{itemize}
@@
 \section{Modelling approach}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Model of circuit}
+%
+ %
+ \begin{center}
+   \includegraphics[width=0.9\linewidth]{VORmodel.svg}
+ \end{center}
+ %
+ \note[item]{Contralateral baseline shift compensates for Our baseline shift}
+ \note[item]{Gain increase due to LTD at lightning}
+ \note[item]{Gain decease due to plasticity elsewhere, but also reverses LTD at lightning.}
+ \note[item]{Nonlinearity here won't affect our questions, as long as it doesn't change}
+ \note[item]{Nonlinearity before compensation could change things}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Simplifying assumptions}
+%
+\begin{itemize}
+  \item No spatial/temporal correlations in plasticity events.
+      %
+      \note[item]{allows us to concentrate on synapse, not neuron/network}
+      %
+%  \item Which synapses eligible for plasticity chosen randomly.
+      %
+      \note[item]{No filing system}
+      %
+  \item Potentiating/depressing plasticity events $\sim$ Poisson processes.% with rates $rf\potdep$, where $f\pot+f\dep=1$.
+      %
+      \note[item]{don't care if STDP...}
+%      %
+%      \note[item]{$r=$ total rate of plasticity events per synapse, $f\potdep=$ fraction of events that are potentiating/depressing.}
+      %
+  \item Potentiation and depression are described by Markov processes.% with transition probabilities $\M\potdep$.
+%      %
+%      \note[item]{matrix elements: transition prob from $i\to j$, given pot/dep}
+%  \item Ideal observer: read weights directly.
+%      \note[item]{not electrical activity: don't model neurons/network}
+%      \note[item]{upper bound on electrical activity readout}
+%      %
+%  \item Synaptic weights can only take values $\pm1$.
+      %
+  \citerr{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}
+      %
+      \note[item]{looks like binary synapse from outside. Inside...}
+\end{itemize}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Dynamics}
+%
+  There are $N$ identical synapses with $M$ internal functional states.
+  %
+  \begin{center}
+    \includegraphics[width=0.7\linewidth]{model.svg}
+  \end{center}
+  %
+  \note[item]{stoch process has steady state.}
+  \note[item]{Prior activity puts it in this state. row vec.}
+  \note[item]{plasticity events at rate r}
+  \note[item]{fraction pot/dep}
+  \note[item]{probs changed by Markov matrices, prob $i\to j$}
+  \note[item]{Readout: synaptic weight vec when in each state.}
+  \vp
+% At $t=0$, the memory is created by $\M\potdep$ with probability $f\potdep$.
+% \note[item]{for this one, we keep track of pot/dep, look for inc/dec of $\w$}
+%
+% \vp Forgetting caused by subsequent memories, evolving as
+%
+  \begin{equation*}
+  \begin{aligned}
+    \diff{\pr(t)}{t} &= r\pr(t)\frg,
+    &\qquad
+    \frg &= f\pot\M\pot+f\dep\M\dep-\I,\\&&
+    \eq\frg &=0.
+  \end{aligned}
+  \end{equation*}
+%
+  \note[item]{Memory at $t=0$, keep track of pot/dep}
+  \note[item]{subsequent: average over pot/dep}
+% \note[item]{$\frg$ is forgetting matrix, $\I=$identity, don't keep track of pot/dep}
+% Eventually, this will settle into the equilibrium distribution:
+% %
+% %
+% \note[item]{In equilibrium prior to memory creation}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Modelling VOR learning}
+%
+\begin{tabular}{>{\bfseries}ll>{\impl}l}
+    Mutation: &
+    Changes mechanism of LTD  &
+    change $\M\dep$.
+    \note[item]{lower threshold \lto increase off-diagonal elements.}
+ \\[1cm]
+    Training: &
+    Changes statistics of LTP/LTD  &
+    change $r,f\pot,f\dep$.
+    \note[item]{Only parameters we have. Don't care about $r$.}
+ \\[1cm]
+    Learning: &
+    Change in VOR gain &
+    decrease in $\av{\w}$.
+    \note[item]{Only output we have. Don't keep track of synaptic identity.}
+\end{tabular}
+%
+\end{frame}
 
@@
  %
+ \note[item]{Compare solid curves}
+ \note[item]{Compare black curves}
  \note[item]{understand why next slide}
@@
  %
- \begin{center}
- \begin{tabular}{ll}
-   \aligntop{\includegraphics[height=0.25\linewidth]{binary_bar_wt_wo.svg}}
-   \hspace{2cm} &
-   \aligntop{\includegraphics[height=0.25\linewidth]{binary_bar_ko_wo.svg}}
-   \\[3cm]
-   \aligntop{\includegraphics[height=0.25\linewidth]{binary_bar_wt_w.svg}}
-   &
-   \aligntop{\includegraphics[height=0.25\linewidth]{binary_bar_ko_w.svg}}
+\adjustbox{valign=m}{\parbox[t]{0.4\linewidth}{
+  \begin{raggedleft}
+  Learning rate: \phantom{rate of events} \\
+  \hp\hp\parbox{0.75\linewidth}{%
+  $\sim$ rate of events\\
+  $\times$ prob.\ of transition\\
+  $\times$ prob.\ ready for $\Delta w$\\
+  $\times$ (-$\Delta w$)}
+  \end{raggedleft}
+}}
+% \begin{center}
+\alignmid{\parbox[t]{0.5\linewidth}{
+ \begin{tabular}{ccc}
+   \alignmid{\includegraphics[height=0.5\linewidth]{binary_bar_wt_wo.svg}}
+   & \alignmid{\includegraphics[height=0.12\linewidth]{left_bad.svg}} &
+   \alignmid{\includegraphics[height=0.5\linewidth]{binary_bar_ko_wo.svg}}
+   \\[1.5cm] \alignmid{\includegraphics[height=0.12\linewidth]{up_bad.svg}}
+   && \alignmid{\includegraphics[height=0.12\linewidth]{up_good.svg}}\\[0.5cm]
+   \alignmid{\includegraphics[height=0.5\linewidth]{binary_bar_wt_w.svg}}
+   & \alignmid{\includegraphics[height=0.12\linewidth]{left_good.svg}} &
+   \alignmid{\includegraphics[height=0.5\linewidth]{binary_bar_ko_w.svg}}
  \end{tabular}
- \end{center}
+}}
+% \end{center}
  %
- \note[item]{understand why next slide}
+ \note[item]{WT: start with everything equal -- just for illustration, not essential}
+ \note[item]{WT: during training, increase $f\dep$ (green arrow) $\to$ weakening.}
+ \note[item]{KO: inc $q\pot\to$ bias}
+ \note[item]{KO: competition between inc prob trans \& dec prob ready}
+ \note[item]{KO: first one wins. see why after next model}
+ \note[item]{pre: reduces/reverses bias. always helps.}
 %
@@
    \includegraphics[width=0.5\linewidth]{serial.svg} \\[1cm]
+ \note[item]{Still looks binary from outside. Hidden states (not essential).}
+ \note[item]{Only see $\Delta w$ at boundary.}
    \aligntop{\includegraphics[width=0.35\linewidth]{serial_learnS.eps}}
@@
  %
+
  \note[item]{understand why next slide}
@@
  \begin{center}
- \begin{tabular}{ll}
-   \aligntop{\includegraphics[height=0.2\linewidth]{serial_bar_wt_wo.svg}}
-   \hspace{1cm} &
-   \aligntop{\includegraphics[height=0.2\linewidth]{serial_bar_ko_wo.svg}}
-   \\[3cm]
-   \aligntop{\includegraphics[height=0.2\linewidth]{serial_bar_wt_w.svg}}
-   &
-   \aligntop{\includegraphics[height=0.2\linewidth]{serial_bar_ko_w.svg}}
+ \begin{tabular}{ccc}
+   \alignmid{\includegraphics[height=0.19\linewidth]{serial_bar_wt_wo.svg}}
+   & \alignmid{\includegraphics[height=0.05\linewidth]{right_good.svg}} &
+   \alignmid{\includegraphics[height=0.19\linewidth]{serial_bar_ko_wo.svg}}
+   \\[1.5cm] \alignmid{\includegraphics[height=0.05\linewidth]{down_good.svg}}
+   && \alignmid{\includegraphics[height=0.05\linewidth]{up_good.svg}}\\[0.5cm]
+   \alignmid{\includegraphics[height=0.19\linewidth]{serial_bar_wt_w.svg}}
+   & \alignmid{\includegraphics[height=0.05\linewidth]{left_good.svg}} &
+   \alignmid{\includegraphics[height=0.19\linewidth]{serial_bar_ko_w.svg}}
  \end{tabular}
@@
  %
- \note[item]{understand why next slide}
+ \note[item]{WT: start with everything equal -- just for illustration, not essential}
+ \note[item]{WT: during training, increase $f\dep$ (green arrow) $\to$ weakening.}
+ \note[item]{KO: inc $q\pot\to$ bias, now exponential}
+ \note[item]{KO: comp. between inc prob trans \& dec prob ready, now only at bndry}
+ \note[item]{KO: second one wins, now exponential}
+ \note[item]{pre: reduces/reverses bias.}
+ \note[item]{pre: little reverse bias repopulates bndry, helps.}
+ \note[item]{pre: too much reverse bias moves away from bndry, hurts.}
+ \note[item]{maths next slide}
 %
@@
 
-\begin{frame}{Mathematics}
+\begin{frame}{Mathematical explanation}
 %
@@
  \vp Learning rate $\sim \eq_{M/2}\prn{\frac{q\dep}{q\pot}} = \CN \prn{\frac{q\pot}{q\dep}}^{\frac{M}{2}-1}$.
+ \note[item]{Detailed balance. Exponential decay.}
 
- \vp For $M>2$, larger $q\dep \implies$ slower learning.
+ \vp For $M>2$: larger $q\dep \implies$ slower learning.
+ \note[item]{for large enough $M,q\pot$, overcome $\CN$}
 
- \vp For $M=2$, larger $q\dep \implies$ larger $\CN \implies$ faster learning.
+ \vp For $M=2$: larger $q\dep \implies$ larger $\CN \implies$ faster learning.
+ \note[item]{Other factor in $\eq$ smaller $\implies \CN$ larger.}
 %
@@
 %
- The success od the serial model relies on two features:
+ The success of the serial model relies on two features:
  \begin{itemize}
@@
     \aligntop{\includegraphics[width=0.4\linewidth]{multistate.svg}}\hspace{1.5cm} &
-    \aligntop{\includegraphics[width=0.28\linewidth]{multistate_learnS.eps}} \\[2cm]
+    \aligntop{\includegraphics[width=0.28\linewidth]{multistate_learnS.eps}}
+    \note[item]{MS: linear weights, unlike serial.}
+    \note[item]{like bunch of binary synapses in series.}
+    \note[item]{solid curves: fails early on , but catches up quickly}
+    \note[item]{black curves: fails badly}
+    \note[item]{No real enhancement of saturation, no metaplasticity.}
+    \note[item]{All transitions contribute: pushing to end has little effect.}
+    \\[2cm]
     \aligntop{\includegraphics[width=0.4\linewidth]{pooled.svg}} &
     \aligntop{\includegraphics[width=0.28\linewidth]{pooled_scarce_learnS.eps}}
+    \note[item]{Pooled: resource depleted by pot/dep. replenished by reverse.}
+    \note[item]{solid curves succeed: enhanced saturation}
+    \note[item]{black curves fail: opposite metaplasticity, pot makes dep easier}
  \end{tabular}
@@
  \end{tabular}
+ \note[item]{Both models, trans probs decay exponentially from centre.}
+ \note[item]{Nonuni: linear weights. Cascade: binary weights.}
+ \note[item]{Enhanced saturation and metaplasticity}
+ \note[item]{Pushing to end makes pot and dep harder}
+ \note[item]{Note: hidden states not necessary}
 
  \citerr{Fusi2005cascade}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Conclusions and further questions}
+%
+ \begin{itemize}
+   \item The saturation effect overcome the enhanced plasticity, if it is enhanced.
+   \alert{Requires complexity}
+   \note[item]{\eg exponential deacy, resource depletion,\ldots}
+   \item A little reverse bias can help, but too much hurts, if repeated potentiation makes depression harder.
+   \alert{Requires metaplasticity}
+   \note[item]{\eg moving away from weight boundary, or weaker transitions.}
+   \item We can find a purely synaptic explanation of VOR behaviour, iff the synapses have these features. 
+   \note[item]{Other explanations? Non-linearity in PK cell?}
+   \item We used behaviour to constrain molecular structure of synapses!
+   \item Can we constrain it further with more experiments?
+ \end{itemize}
 %

### 2013-07-12 MHCslides/MHCslides.tex
expt results section

@@
 \usepackage{array}
+\usepackage{cancel}
 \input{sl_slide_preamble.tex}
+%\usecolortheme[named=darkblue]{structure}
+\usecolortheme[rgb={0.6372549 0.27843137 0.27058824}]{structure}
+%\setbeamercolor{alerted text}{fg=blue!80!black}
 %\input{sl_slide_preamble_nonotes.tex}
@@
  \titlepage
+ \note[item]{Acknowledge Barbara and Grace}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Introduction}
+%
+ One might expect enhanced plasticity to enhance learning.
+ \note[item]{It does help in some cases}
+ But often it actually causes impairment.
+
+ \vp We argue that this impaired learning is due to basal activity biasing the synapses prior to learning, leaving fewer synapses available for further plasticity.
+
+ \vp By analysing several models of complex synapses, we see that the behaviour of enhanced LTD mice in a motor learning task can constrain the synaptic structure.
 %
@@
  }
+
+ \citerr{duLac1995review,Boyden2004review}
 %
@@
 %
- \parbox[t]{0.2\linewidth}{\aligntop{\includegraphics[width=0.99\linewidth]{VORinc.svg}}\\[1cm]
+ \parbox[t]{0.2\linewidth}{%
+   \aligntop{\includegraphics[width=0.99\linewidth]{VORinc.svg}}\\[1cm]
      \note[item]{trick brain into thinking VOR gain needs adjusting my moving visual stimulus}
      \note[item]{anti-phase \lto increase gain}
- \aligntop{\includegraphics[width=0.99\linewidth]{VORdec.svg}}}
+   \aligntop{\includegraphics[width=0.99\linewidth]{VORdec.svg}}
      \note[item]{in phase \lto decrease gain}
+   }
  \hspace{0.15\linewidth}
- \parbox[t]{0.59\linewidth}{\aligntop{\includegraphics[width=0.99\linewidth]{VORcircuit.svg}}}
+ \parbox[t]{0.59\linewidth}{%
+   \aligntop{\includegraphics[width=0.99\linewidth]{VORcircuit.svg}}
+
+   \vp PF-Pk: \parbox[t]{0.5\linewidth}{PF+CF \lto LTD, \\
+              PF+\cancel{CF} \lto LTP.}
+ }
    \note[item]{Gain change involves cerebellum}
@@
    \note[item]{Different Pk cells have different tunings.}
- \note[item]{Gain up in case of interest: LTD in Pf-Pk in flucculus}
- \note[item]{Gain down: uses different mech for behaviour, but does reverse LTD in Pf-Pk in flucculus}
+   \note[item]{Gain up in case of interest: LTD in Pf-Pk in flocculus}
+   \note[item]{Gain down: uses different mech for behaviour, but does reverse LTD in Pf-Pk in flocculus}
+
+   \citerr{Marr1969cerebellum,Albus1971cerebellum,ito1972cerebellum,Ito1982VOR,Miles1981VOR}
 %
@@
 \section{The effects of enhanced plasticity and saturation}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{MHC-I D$^\mathrm{b}$K$^\mathrm{b} -/-$ knockout}
+%
+ Knockout of molecule lowers threshold for LTD in PF-Pk synapses.\\
+ \citerr{McConnell2009MHCcerebellum}
+ \note[item]{Easier LTD \lto expect better learning}
+
+ \vp \begin{center}
+ \begin{tabular}{ccc}
+   Baseline & \hspace{0.1\linewidth} & Gain increase at $1.0\,\mathrm{Hz}$ \\
+   \alignmid{\includegraphics[width=0.4\linewidth]{baseline.svg}} && 
+   \alignmid{\includegraphics[width=0.3\linewidth]{gain_inc.svg}} 
+ \end{tabular}
+ \end{center}
+ \note[item]{No difference at baseline}
+ \note[item]{Impairment of learning}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Saturation hypothesis}
+%
+ \begin{center}
+ \begin{tabular}{ccc}
+   Error model & \hspace{0.1\linewidth} & Saturation model \\[1cm]
+   \alignmid{\includegraphics[width=0.3\linewidth]{err_model.svg}} &&
+   \alignmid{\includegraphics[width=0.3\linewidth]{sat_model.svg}}
+ \end{tabular}
+ \end{center}
+ \note[item]{Older explanation: error model}
+ \note[item]{Our model: baseline activity \lto saturation \lto less depression possible}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Evidence: level of depression}
+%
+ \aligntop{\includegraphics[width=0.3\linewidth]{pSer880.svg}}
+ \hp
+ \parbox[t]{0.6\linewidth}{%
+ Basal level of GluR2 phosphorylation at serine 880 in AMPA receptor.
+
+ \vp Biochemical signature of PF-Pk LTD.
+
+ \vp Shows that \# depressed synapses in flocculus is larger in KO than WT.
+ \note[item]{Predicted by saturation hypothesis}
+ }
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Evidence: saturation by CF stimulation}
+%
+ Use Channelrhodopsin to stimulate CF \parbox[t]{0.45\linewidth}{%
+ \lto increase LTD in PF-Pk synapses \\
+ \lto simulate saturation in WT.
+ }
+ \note[item]{should result in similar behaviour to KO}
+
+ \vp\begin{center}
+   \alignmid{\includegraphics[width=0.4\linewidth]{CFstim.svg}}
+   \hspace{0.1\linewidth}
+   \alignmid{\includegraphics[width=0.4\linewidth]{CFstim_results.svg}}
+ \end{center}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Evidence: desaturation by gain-decrease pretraining}
+%
+ %
+ \begin{center}
+   \includegraphics[width=0.7\linewidth]{desat.svg}
+
+   \vp \includegraphics[width=0.7\linewidth]{gain_dec.svg}
+ \end{center}
+ %
+ \note[item]{gain dec reverses LTD}
+ \note[item]{but behaviour from elsewhere}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
+\begin{frame}{Summary of training results}
+%
+ \begin{center}
+   \alignmid{\includegraphics[width=0.6\linewidth]{data_shifted.svg}}
+   \hspace{0.05\linewidth}
+   \alignmid{\includegraphics[width=0.3\linewidth]{comparisons.svg}}
+ \end{center}
+ \note[item]{Restricted to gain inc for comparison}
+ \note[item]{Black: WT. Red: KO}
+ \note[item]{Solid: no pre. Dashed: with pre}
+ \note[item]{Horz and vert comparisons: conceptual}
+ \note[item]{Diagonal comparisons: paramter fitting. Depend on size of KO vs.\ pretraining}
+ \note[item]{KO hurts w/o, but helps w/}
+ \note[item]{pre helps KO but hurts WT}
+ \note[item]{top and left most restrictive}
+%
+\end{frame}
 
@@
 
+\begin{frame}{Complex synapses}
+%
+  \aligntop{\includegraphics[width=5cm]{synapse.svg}}\hp
+  %
+  \note[item]{functional states, not molecules}
+  \note[item]{synaptic weight depends on state}
+  \note[item]{many states can have same weight}
+  %
+  \hp\aligntop{\includegraphics[width=2cm]{pot.svg}}
+  \hp\aligntop{\includegraphics[width=2cm]{dep.svg}}
+  %
+  \note[item]{stochastic transitions}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
 \begin{frame}{Simplifying assumptions}
@@
 %  \item Synaptic weights can only take values $\pm1$.
+%      %
+%      \note[item]{looks like binary synapse from outside. Inside...}
       %
   \citerr{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}
-      %
-      \note[item]{looks like binary synapse from outside. Inside...}
 \end{itemize}
@@
 
-\begin{frame}{Dynamics}
+\begin{frame}{Synaptic dynamics}
 %
@@
     \note[item]{Only parameters we have. Don't care about $r$.}
+    \note[item]{Same PF+CF input \lto same $r,f\pot,f\dep$ in each case.}
  \\[1cm]
@@
     Change in VOR gain &
-    decrease in $\av{\w}$.
+    fn.\ of decrease in $\av{\w}$.
     \note[item]{Only output we have. Don't keep track of synaptic identity.}
+    \note[item]{Input to Pk, some linear combination of $\w$'s. }
 \end{tabular}
@@
  \begin{center}
-   \includegraphics[width=0.5\linewidth]{serial.svg} \\[1cm]
+   \includegraphics[width=0.5\linewidth]{serial.svg}\\
+   \citerr{Leibold2008serial} \\[1cm]
  \note[item]{Still looks binary from outside. Hidden states (not essential).}
@@
 
+%%-------------Slide--------------------------------------------------------
+%
+%\begin{frame}{Acknowledgements}
+%%
+% 
+%%
+%\end{frame}
+%
 %-------------Slide--------------------------------------------------------

### 2013-07-12 MHCslides/MHCslides.tex
typos

@@
 
-\begin{frame}{Vestibular Occular Reflex}
+\begin{frame}{Vestibulo-Occular Reflex}
 %
@@
 %
- Knockout of molecule lowers threshold for LTD in PF-Pk synapses.\\
+ Knockout of molecules lowers threshold for LTD in PF-Pk synapses.\\
  \citerr{McConnell2009MHCcerebellum}
@@
 
+\begin{frame}{Viral rescue removes defect}
+%
+ Local injection of Pk specific virus used to restore the knocked-out molecules.
+
+ \vp\begin{center}
+   \alignmid{\includegraphics[width=0.3\linewidth]{rescue.svg}}
+   \hspace{0.1\linewidth}
+   \alignmid{\includegraphics[width=0.5\linewidth]{rescue_results.svg}}
+ \end{center}
+ \note[item]{removes defect}
+ \note[item]{tells us that it is these PF-Pk synapses that are important}
+ \note[item]{But what mechanism?}
+%
+\end{frame}
+
+%-------------Slide--------------------------------------------------------
+
 \begin{frame}{Saturation hypothesis}
@@
  \note[item]{Our model: baseline activity \lto saturation \lto less depression possible}
+ \note[item]{Saturation has to compete with enhanced plsaticity. Which will win?}
 %
@@
  \note[item]{top and left most restrictive}
+ \note[item]{Pay attention to solid: black above red}
+ \note[item]{Pay attention to black: solid above dashed}
+ \note[item]{Concentrate on initial slope}
 %

### 2013-07-13 MHCslides/MHCslides.tex
fixed some errors

@@
       %
-  \item Potentiating/depressing plasticity events $\sim$ Poisson processes.% with rates $rf\potdep$, where $f\pot+f\dep=1$.
+  \vp\item Potentiating/depressing plasticity events $\sim$ Poisson processes.% with rates $rf\potdep$, where $f\pot+f\dep=1$.
       %
@@
       %
-  \item Potentiation and depression are described by Markov processes.% with transition probabilities $\M\potdep$.
+  \vp\item Potentiation and depression are described by Markov processes.% with transition probabilities $\M\potdep$.
 %      %
@@
       %
-  \citerr{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}
+  \vp\citerr{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}
 \end{itemize}
@@
  %
- \note[item]{WT: start with everything equal -- just for illustration, not essential}
- \note[item]{WT: during training, increase $f\dep$ (green arrow) $\to$ weakening.}
+ \note[item]{Now only get signal from crossing boundary}
  \note[item]{KO: inc $q\pot\to$ bias, now exponential}
- \note[item]{KO: comp. between inc prob trans \& dec prob ready, now only at bndry}
- \note[item]{KO: second one wins, now exponential}
+ \note[item]{KO: prob ready wins over prob trans, now exponential}
  \note[item]{pre: reduces/reverses bias.}
@@
 %
- Serial synapse: $\eq_i \sim \CN \prn{\frac{q\pot}{q\dep}}^i$.
+ %
+ \begin{center}
+   \includegraphics[width=0.4\linewidth]{serial_label.svg}
+ \end{center}
+ %
+ \vp Serial synapse: $\eq_i \sim \CN \prn{\frac{q\pot}{q\dep}}^i$.
 
@@
 %
- The success of the serial model relies on two features:
+ %
+ \begin{center}
+   \includegraphics[width=0.4\linewidth]{serial.svg}
+ \end{center}
+ %
+ \vp The success of the serial model relies on two features:
  \begin{itemize}
@@
  \end{itemize}
- %
- \begin{center}
-   \includegraphics[width=0.4\linewidth]{serial.svg}
- \end{center}
- %
  \note[item]{borne out by other models that fail/succeed}

### 2013-07-15 Notes/Figures/MakeFigs.m
started 15min talk

@@ -7,4 +7,4 @@ MakeVORfigs('pooled_plenty',[0.3 0.4],[0.6 0.8],-0.1,5,20,'print','Pooled',true)
 MakeVORfigs('pooled_scarce',[0.05 0.4],[0.1 0.8],-0.1,5,20,'print','Pooled',true)
 MakeVORfigs('cascade_short',0.25,0.33,-0.3,5,20,'print')
 MakeVORfigs('cascade_long',0.25,0.33,-0.3,5,100,'print')
-MakeVORfigs('nonuni',0.25,0.33,-0.3,5,100,'print')
+MakeVORfigs('nonuni',0.25,0.33,-0.3,5,100,'print','LinWeights',true)

### 2013-07-15 Code/VOR/VORcomparison.m
changes for 15 min talk

@@ -15,7 +15,7 @@ FontSize=20;
 EqFontSize=20;
 txFontSize=20;
 Pooled=false;
-mutant='D^bK^b-/-';
+mutant='KO';
 red=[192 0 0]/255;
 varargin=assignApplicable(varargin);
 
@@ -84,7 +84,7 @@ else
     annotation('doublearrow',x,y);
     pos=dsxy2figxy([0.6*tchange+0.4*xl(2) (0.95*yl(1)+0.05*yl(2)) 0.2*(xl(2)-tchange) 0.05*(yl(2)-yl(1))]);
     annotation('textbox',pos,'String','training','VerticalAlignment','bottom','LineStyle','none','HorizontalAlignment','center','FontSize',txFontSize);
-    legend([phWT(1);phKn(1)],{'WT';'D^bK^b-/-'},'Location','Best')
+    legend([phWT(1);phKn(1)],{'WT'; mutant},'Location','Best')
     drawnow;
     
     if isscalar(fig3)

### 2013-07-22 MHCpaper/MHCsupp.tex
MHC supp: added nonuni and changed colours

@@
 We will look at several different models,
-the multistate model\footnote{The original definition of the multistate model had a synaptic weight that varied linearly along the chain.
-We will refer to this as the linear multistate model.
-We will also look at the step-multistate model, for which the synaptic weight takes only two values.} (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}),
-the two-state model (which can be thought of as a special case of the multistate model, see \autoref{fig:models}\ref{fig:binary_model}),
+the serial model (see \cite{Leibold2008serial} and \autoref{fig:models}\ref{fig:serial_model}) which has only two values for the synaptic weight,
+the two-state model (which can be thought of as a special case of the serial model, see \autoref{fig:models}\ref{fig:binary_model}),
+the multistate model (see \cite{amit1994learning,Fusi2007multistate} and \autoref{fig:models}\ref{fig:multistate_model}) which has a linearly varying synaptic weight,
 and the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}).
-We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel}.
+We will also look at a new, pooled resource model that we will define below in \autoref{sec:pooledmodel} and a non-uniform version of the multistate model that we will describe in \autoref{sec:nonunimodel}.
 
-For the multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
+For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
 We will use a larger value for $q$ for depression in the \KO\ mutant.
@@
  \begin{center}
+ \parbox{0.8\linewidth}{%
  \begin{myenuma}
+  \item\aligntop{\includegraphics[height=1.7cm]{serial.svg}}\label{fig:serial_model}\hspace{0.5cm}
+  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\hspace{0.5cm}\\[1cm]
   \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=2.2cm]{cascade.svg}}\label{fig:cascade_model}
+  \item\aligntop{\includegraphics[height=2.5cm]{cascade.svg}}\label{fig:cascade_model}
  \end{myenuma}
+ }
  \end{center}
   \caption[Transition probabilities for different models]{Transition probabilities for different models.
-  Potentiation induces transitions indicated by solid red arrows, depression indicated by dashed blue arrows.
-  States of strong/weak synaptic weight indicated by red/blue circles.
-  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
+  Potentiation induces transitions indicated by orange arrows, depression indicated by blue arrows.
+  States of strong/weak synaptic weight indicated by orange/blue circles.
+  (\ref{fig:serial_model}) In the serial model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
+  The synaptic weight takes only two values, $\pm1$.
   (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.
+  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised X
+  by these two values.
+  The synaptic weight varies linearly in the interval $[-1,1]$.
   (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}) and synaptic weight takes only two values.
@@
   This pool of synapses can be modelled as one compound synapse.} \label{fig:pooled_model}
+\end{figure}
+
+\subsubsection{Non-uniform multistate model}\label{sec:nonunimodel}
+
+This model is similar to the multistate model (see \autoref{fig:models}\ref{fig:multistate_model})), as it only has transitions between adjacent states and a linearly varying synaptic weight.
+However, like the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the transition probabilities decay exponentially away from the central transition.
+More precisely:
+%
+\begin{equation}\label{eq:nonunidef}
+    \M\pot_{ii+1} = \M\dep_{i+1i} =  x^{\abs{\frac{M+1}{2}-i}+\half},
+      \quad i = 1 \ldots M-1,
+\end{equation}
+%
+with all other off-diagonal elements equal to zero.
+The diagonal elements are chosen so that the rows sum to one.
+
+This model is parameterised by the values of $x$ chosen for potentiation and depression.
+We will use a larger value for depression in the \KO\ mutant.
+The values used are listed in \autoref{tab:params}.
+
+\begin{figure}
+ \begin{center}
+  \aligntop{\includegraphics[height=3cm]{multistate_nonuni.svg}}
+ \end{center}
+  \caption[The non-uniform multistate model]{The non-uniform multistate model.
+  Similar to the multistate model, \autoref{fig:models}\ref{fig:multistate_model}), the synaptic weight varies linearly in the interval $[-1,1]$, but the transition probabilities between adjacent states decays exponentially away from the central transition for both potentiation and depression.} \label{fig:nonuni_model}
 \end{figure}
@@
     \cline{1-7}
-    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & 0.1  & 5 & 20  &\label{tr:multistate_weak} \\
-    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 20  &\label{tr:multistate_med} \\
-    Step-multistate    & 10 & $q=0.3$  & $q=0.4$  & 0.45 & 5 & 30  &\label{tr:multistate_strong} \\
+    Serial        & 10 & $q=0.3$  & $q=0.4$  & 0.1  & 5 & 20  &\label{tr:multistate_weak} \\
+    Serial        & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 20  &\label{tr:multistate_med} \\
+    Serial        & 10 & $q=0.3$  & $q=0.4$  & 0.45 & 5 & 30  &\label{tr:multistate_strong} \\
     Two-state     & 2  & $q=0.1$  & $q=0.2$  & 0.1  & 5 & 5   &\label{tr:binary} \\
-    Linear Multistate  & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 5   &\label{tr:multistate_lin} \\
+    Multistate    & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 5   &\label{tr:multistate_lin} \\
     Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
@@
     Cascade       & 10 & $x=0.25$ & $x=0.33$ & 0.3  & 5 & 100 &\label{tr:cascade_long} \\
+    Non-uniform MS       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:nonuni} \\
     \cline{1-7}
@@
 
-\subsection{Step-multistate model}\label{sec:multistate}
+\subsection{Serial model}\label{sec:multistate}
 
-\resultsfig{step-multistate}{weak training}{($\Delta f=-0.1$)}{multistate_weak}
+\resultsfig{serial}{weak training}{($\Delta f=-0.1$)}{multistate_weak}
 
-\resultsfig{step-multistate}{moderate training}{($\Delta f=-0.3$)}{multistate_med}
+\resultsfig{serial}{moderate training}{($\Delta f=-0.3$)}{multistate_med}
 
-\resultsfig{step-multistate}{strong training}{($\Delta f=-0.45$)}{multistate_strong}
+\resultsfig{serial}{strong training}{($\Delta f=-0.45$)}{multistate_strong}
 
-The results of simulations of the step-multistate model can be seen in \autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
+The results of simulations of the serial model can be seen in \autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
 However, we can get some insight into this model analytically.
@@
 
-In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f < \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the mutant learns slower than wild-type without pre-training, and that pre-training speeds up learning in the mutant but slows it down in the wild-type.
+In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the mutant learns slower than wild-type without pre-training, and that pre-training speeds up learning in the mutant but slows it down in the wild-type.
 However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.
@@
 Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
-For the step-multistate model, we argued that the first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ to leading order.
+For the serial model, we argued that the first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ to leading order.
 Here $M=2$, so the leading part of the two effects will cancel.
@@
 
-\subsection{Linear multistate model}\label{sec:multistate_lin}
+\subsection{Multistate model}\label{sec:multistate_lin}
 
-\resultsfign{linear multistate model}{multistate_lin}
+\resultsfign{multistate model}{multistate_lin}
 
@@
 The numerical results can be seen in \autoref{fig:multistate_lin}, but we can get some analytic insight into this model as well.
-In essence, this model is like a series of two-state models attached to each other, in contrast to the step-multistate model for which the synaptic weight only changes between one of the pairs of states.
+In essence, this model is like a series of two-state models attached to each other, in contrast to the serial model for which the synaptic weight only changes between one of the pairs of states.
 
@@
 
-Gain-decrease pre-training will shift the distribution to the potentiated site, where there are more synapses available for depression and resources are more plentiful, for both the mutant and wild type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_eq}, red traces).
+Gain-decrease pre-training will shift the distribution to the potentiated side, where there are more synapses available for depression and resources are more plentiful, for both the mutant and wild type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_eq}, red traces).
 This means that the pre-trained animal will learn faster then the untrained one for both mutant and wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learnS} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learnS}).
@@
 
+\subsection{Non-uniform multistate model}\label{sec:nonuni}
+
+\resultsfign{non-uniform multistate}{nonuni}
+
+The results of simulations of the non-uniform multistate model can be seen in \autoref{fig:nonuni}.
+It is difficult to study this model analytically.
+However, the numerical results can helps us understand it qualitatively.
+
+First, we see that the mutant is slower than wild-type without pre-training, as seen experimentally.
+This seems to be because the enhanced depression results in a greater fraction of synapses begin in states of weaker synaptic weight (see \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_KO}) where the depressing transitions have lower probability (see \autoref{fig:nonuni_model}).
+
+However, gain-decrease pre-training will push synapses towards the states of stronger synaptic weight.
+The wild-type will be pushed further that the mutant, as the latter started further to the weaker side as explained above (compare \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_WT} and \ref{fig:nonuni_eq_KO}).
+As the depressing transitions have lower probability for the strongest states (see \autoref{fig:nonuni_model}), this, in addition to the enhanced depression of the mutant, will result in the mutant learning faster than wild-type, as seen experimentally.
+
+When comparing learning with and without gain-decrease pretraining, there are two effects to consider.
+First, the pushing synapses towards the states of stronger synaptic weight, resulting in more synapses being available for depression, speeding up learning.
+Second, it will also place synapses in states where transitions have lower probability, slowing down learning.
+The second effect will be weaker in the mutant than in wild-type, as they started further to the weaker side as explained above.
+This means that, for appropriate parameter choices, gain-decrease pretraining can speed up learning in the mutant, but also slow down learning in wild-type, as seen experimentally.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-07-22 MHCpaper/MHCsupp.tex
typos in mhc supp

@@
   with states of weak synaptic strength to the left and strong states to the right.
-  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre}\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre}\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
+  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
   with states of weak synaptic strength at the top and strong states underneath. }
@@
 
-\resultsfig{serial}{weak training}{($\Delta f=-0.1$)}{multistate_weak}
+\resultsfig{serial}{weak training}{($\Delta f=0.1$)}{multistate_weak}
 
-\resultsfig{serial}{moderate training}{($\Delta f=-0.3$)}{multistate_med}
+\resultsfig{serial}{moderate training}{($\Delta f=0.3$)}{multistate_med}
 
-\resultsfig{serial}{strong training}{($\Delta f=-0.45$)}{multistate_strong}
+\resultsfig{serial}{strong training}{($\Delta f=0.45$)}{multistate_strong}
 

### 2013-08-02 NIPS/rebuttal.txt
working on NIPS rebuttal

Reviewer 1:
Regarding the discrepancy between the envelope and the numerical results: the envelope is just an upper bound and we 
do not claim that it is a tight bound. The reason is that equation (18) is not aa complete set of constraints, as 
discussed below it. In the paragraph from line 377 to 400, we discuss this point further. Note that the numerical 
results are not always instructive as the numerical procedures can be prevented from reaching the global maximum by 
local maxima. This is shown by the fact that our hand designed models can outperform the numerical methods at late 
times. In fact, the apparent drop-off of the solid red curve at very late times comes from not allowing small enough 
epsilon. This should be fixed in fig5.

Reviewer 2:
We'll add a derivation of equation 3 to the supplement. We could add a description in the main text along
the lines of: "The factor of p^infinity describes the synapses being in the steady-state distribution before the
memory is encoded. The factor of (M^pot-M^dep) comes from the encoding of the memory at t=0, with w_ideal being +/-1 
in synapses that are potentiated/depotentiated. The factor of exp(rt W^F) describes the subsequent evolution of the
probability distribution, averaged over all sequences of plasticity events and the factor of w indicates the readout 
via the synaptic weight."
I think we referred to all of the original parts of the supplement in the main paper. I'm not sure how we could
incorporate it into the main paper without exceeding the page limit.

Reviewer 3:
line 195: in the introduction, we cited [17], which describes diversity in synaptic structure across the vertebrate 
brain. This could be related to optimisation for different timescales, but anything more than speculation would 
require a better understanding of the relation between structure and function (which is what we intend to begin with 
this work).
line 220:
line 294: agreed.
line 299: The fact that eta_i depends on M^pot/dep is taken into account. In equation (57) of the supplement, the 
term involving c_g comes from this dependence. If the reviewer is concerned that tho order of the eta_i could change 
during the maximisation procedure: note that necessary conditions for a maximum only require that there is no 
infinitesimal perturbation that increases the area. Therefor we need only consider an infinitesimal neighbourhood of 
the model, in which the order will not change.
line 304: Yes. We'll add this phrase to the text.
line 326: In fig2(a,b), we need not have an equal number of w=+/-1 states (we could change the figure to reflect 
this), but those states are not functionally relevant, as shown by their equivalence to fig2c. The model in fig2c, of 
course, has no room for such asymmetry. The models in figs4,5b do need to have equal numbers of +/- states, asymmetry 
would make them worse. For fig4, the effect of the asymmetry would be reduced in the limit as epsilon -> 0, as all
states other than the end states would have very small p^infinity.
line 415: It is only for t_0 > sqrt(N)M that the models that nearly touch the envelope are linear chains. The model 
in fig5b is not a linear chain, as it has shortcut transitions, but they are only the best models (that we have 
found) for times t_0 < sqrt(N)M. Would it help if we repeated fig4 as fig5c?

### 2013-08-02 NIPS/Figs/envPlot.m

@@ -1,0 +1,21 @@
+h=plot(t,10*env,'g',tn,10*envnum,'r--',t,10*envshort,'r','LineWidth',3);
+set(gca,'XScale','log','YScale','log');
+ylim(yl);
+xlim([t(1) t(end)]);
+xlabel('Time');
+ylabel('SNR');
+line([1;1]*bnds',yl'*ones(1,length(bnds)),'Color','k','LineWidth',1.5,'LineStyle','--');
+embiggen(gca,20);
+legend({'envelope','numerical search','hand designed'},'Location','Best');
+
+txFontSize=14;
+
+    [x,y]=dsxy2figxy([t(1) bnds], (yl(1)^0.96*yl(2)^0.04)*[1 1]);
+    annotation('doublearrow',x,y);
+    pos=dsxy2figxy([t(1)^0.9*bnds^0.1 (yl(1)^0.95*yl(2)^0.05) t(1)^0.1*bnds^0.9 (yl(1)^0.9*yl(2)^0.1)]);
+    annotation('textbox',pos,'String','Initial SNR bound active','VerticalAlignment','top','LineStyle','none','HorizontalAlignment','center','FontSize',txFontSize);
+    
+    [x,y]=dsxy2figxy([t(1) t(end)], (yl(1)^0.89*yl(2)^0.11)*[1 1]);
+    annotation('doublearrow',x,y);
+    pos=dsxy2figxy([t(1)^0.9*bnds^0.1 (yl(1)^0.88*yl(2)^0.12) t(1)^0.1*bnds^0.9 (yl(1)^0.83*yl(2)^0.17)]);
+    annotation('textbox',pos,'String','Area bound active','VerticalAlignment','bottom','LineStyle','none','HorizontalAlignment','center','FontSize',txFontSize);

### 2013-08-02 NIPS/rebuttal.txt
addressing reviewer 1

@@
-Reviewer 1:
-Regarding the discrepancy between the envelope and the numerical results: the envelope is just an upper bound and we 
-do not claim that it is a tight bound. The reason is that equation (18) is not aa complete set of constraints, as 
-discussed below it. In the paragraph from line 377 to 400, we discuss this point further. Note that the numerical 
-results are not always instructive as the numerical procedures can be prevented from reaching the global maximum by 
-local maxima. This is shown by the fact that our hand designed models can outperform the numerical methods at late 
-times. In fact, the apparent drop-off of the solid red curve at very late times comes from not allowing small enough 
-epsilon. This should be fixed in fig5.
+Reviewer 1: We briefly discussed a suggestion for biological
+experiments to test the theory in lines 428. We could expand on
+this. If we measured pre and post-synaptic spike trains, and also
+recorded changes in post-synaptic potentials to measure changes in
+synaptic weight, we could use hidden Markov model techniques to find the best-fit Markov
 
-Reviewer 2:
-We'll add a derivation of equation 3 to the supplement. We could add a description in the main text along
-the lines of: "The factor of p^infinity describes the synapses being in the steady-state distribution before the
-memory is encoded. The factor of (M^pot-M^dep) comes from the encoding of the memory at t=0, with w_ideal being +/-1 
-in synapses that are potentiated/depotentiated. The factor of exp(rt W^F) describes the subsequent evolution of the
-probability distribution, averaged over all sequences of plasticity events and the factor of w indicates the readout 
-via the synaptic weight."
-I think we referred to all of the original parts of the supplement in the main paper. I'm not sure how we could
-incorporate it into the main paper without exceeding the page limit.
+Regarding the discrepancy between the envelope and the
+numerical results: the envelope is just an upper bound and we do
+not claim that it is a tight bound. The reason is that equation
+(18) is not aa complete set of constraints, as discussed below it.
+In the paragraph from line 377 to 400, we discuss this point
+further. Note that the numerical results are not always instructive
+as the numerical procedures can be prevented from reaching the
+global maximum by local maxima. This is shown by the fact that our
+hand designed models can outperform the numerical methods at late
+times. In fact, the apparent drop-off of the solid red curve at
+very late times comes from not allowing small enough epsilon. This
+should be fixed in fig5.
+
+Reviewer 2: We'll add a derivation of equation 3 to the supplement.
+We could add a description in the main text along the lines of:
+"The factor of p^infinity describes the synapses being in the
+steady-state distribution before the memory is encoded. The factor
+of (M^pot-M^dep) comes from the encoding of the memory at t=0, with
+w_ideal being +/-1 in synapses that are potentiated/depotentiated.
+The factor of exp(rt W^F) describes the subsequent evolution of the
+probability distribution, averaged over all sequences of plasticity
+events and the factor of w indicates the readout via the synaptic
+weight."
+I think we referred to all of the original parts of the supplement
+in the main paper. I'm not sure how we could incorporate more into
+the main paper without exceeding the page limit.
 
 Reviewer 3:
-line 195: in the introduction, we cited [17], which describes diversity in synaptic structure across the vertebrate 
-brain. This could be related to optimisation for different timescales, but anything more than speculation would 
-require a better understanding of the relation between structure and function (which is what we intend to begin with 
-this work).
+line 195: in the introduction, we cited [17], which describes
+diversity in synaptic structure across the vertebrate brain. This
+could be related to optimisation for different timescales, but
+anything more than speculation would require a better understanding
+of the relation between structure and function (which is what we
+intend to begin with this work).
 line 220:
 line 294: agreed.
-line 299: The fact that eta_i depends on M^pot/dep is taken into account. In equation (57) of the supplement, the 
-term involving c_g comes from this dependence. If the reviewer is concerned that tho order of the eta_i could change 
-during the maximisation procedure: note that necessary conditions for a maximum only require that there is no 
-infinitesimal perturbation that increases the area. Therefor we need only consider an infinitesimal neighbourhood of 
+line 299: The fact that eta_i depends on M^pot/dep is taken into
+account. In equation (57) of the supplement, the term involving c_g
+comes from this dependence. If the reviewer is concerned that tho
+order of the eta_i could change during the maximisation procedure:
+note that necessary conditions for a maximum only require that
+there is no infinitesimal perturbation that increases the area.
+Therefor we need only consider an infinitesimal neighbourhood of
 the model, in which the order will not change.
 line 304: Yes. We'll add this phrase to the text.
-line 326: In fig2(a,b), we need not have an equal number of w=+/-1 states (we could change the figure to reflect 
-this), but those states are not functionally relevant, as shown by their equivalence to fig2c. The model in fig2c, of 
-course, has no room for such asymmetry. The models in figs4,5b do need to have equal numbers of +/- states, asymmetry 
-would make them worse. For fig4, the effect of the asymmetry would be reduced in the limit as epsilon -> 0, as all
-states other than the end states would have very small p^infinity.
-line 415: It is only for t_0 > sqrt(N)M that the models that nearly touch the envelope are linear chains. The model 
-in fig5b is not a linear chain, as it has shortcut transitions, but they are only the best models (that we have 
-found) for times t_0 < sqrt(N)M. Would it help if we repeated fig4 as fig5c?
+line 326: In fig2(a,b), we need not have an equal number of w=+/-1
+states (we could change the figure to reflect this), but those
+states are not functionally relevant, as shown by their equivalence
+to fig2c. The model in fig2c, of course, has no room for such
+asymmetry. The models in figs4,5b do need to have equal numbers of
++/- states, asymmetry would make them worse. For fig4, the effect
+of the asymmetry would be reduced in the limit as epsilon -> 0, as
+all states other than the end states would have very small
+p^infinity.
+line 415: It is only for t_0 > sqrt(N)M that the models that nearly
+touch the envelope are linear chains. The model in fig5b is not a
+linear chain, as it has shortcut transitions, but they are only the
+best models (that we have found) for times t_0 < sqrt(N)M. Would it
+help if we repeated fig4 as fig5c?

### 2013-08-02 NIPS/rebuttal.txt
rebuttal comments done

@@
 Reviewer 1: We briefly discussed a suggestion for biological
-experiments to test the theory in lines 428. We could expand on
+experiments to test the theory in line 428. We could expand on
 this. If we measured pre and post-synaptic spike trains, and also
 recorded changes in post-synaptic potentials to measure changes in
-synaptic weight, we could use hidden Markov model techniques to find the best-fit Markov
-
+synaptic weight, we could use hidden Markov model techniques to
+find the best-fit synaptic model. However, as shown in this paper,
+which model is best depends on the range of time-scales of
+interest, so any predictions would also require understanding of
+what timescales the particular synapse needs to operate on.
 Regarding the discrepancy between the envelope and the
@@
 should be fixed in fig5.
+Yes, the level of noise tolerance is an important consideration.
+Note that the stochasticity of the models is an expression of
+biological noise. In fact, the type of noise described by the
+reviewer could be included by adding extra states to the model with
+the different transition rates. However, the synapses should not be
+allowed to optimise all of this noise away. It would be interesting
+to see how our results change if we put lower limits on the noise
+levels. I think this would be beyond the scope of this work, but
+should be added to the discussion section.
 
@@
 intend to begin with this work).
-line 220:
+line 220: Yes, fig2a,b is an example of why it is difficult to map
+molecular states to functional states as, despite appearances,
+these models actually only have two functional states due to their
+equivalence to fig 2c. One experimental investigation of this could
+be along the lines of the experiment described in line 428 (see
+first paragraph of our reply to reviewer 1 above for elaboration).
+Presumably we will find fewer functional states than molecular
+states. Making the link between molecular and functional states
+will, unfortunately, be much harder
 line 294: agreed.

### 2013-08-05 NIPS/rebuttal.txt
my last rebuttal update

@@
-Reviewer 1: We briefly discussed a suggestion for biological
-experiments to test the theory in line 428. We could expand on
-this. If we measured pre and post-synaptic spike trains, and also
-recorded changes in post-synaptic potentials to measure changes in
-synaptic weight, we could use hidden Markov model techniques to
-find the best-fit synaptic model. However, as shown in this paper,
-which model is best depends on the range of time-scales of
-interest, so any predictions would also require understanding of
-what timescales the particular synapse needs to operate on.
+Reviewer 1:
+We would like to link our theory to experiment, but we don't think
+that the necessary experiments have been performed yet. We briefly
+discussed a suggestion for biological experiments to test the
+theory in line 428. We could expand on this. If we measured pre and
+post-synaptic spike trains, and also recorded changes in
+post-synaptic potentials/currents from a single synapse to measure
+changes in synaptic weight, we could use hidden Markov model
+techniques to find the best-fit synaptic model. However, as shown
+in this paper, which model is best depends on the range of
+time-scales of interest, so any predictions would also require
+understanding of what timescales the particular synapse needs to
+operate on. We'd also need to cut something else to stay within the
+page limit.
 Regarding the discrepancy between the envelope and the
@@
 very late times comes from not allowing small enough epsilon. This
-should be fixed in fig5.
-Yes, the level of noise tolerance is an important consideration.
-Note that the stochasticity of the models is an expression of
-biological noise. In fact, the type of noise described by the
-reviewer could be included by adding extra states to the model with
-the different transition rates. However, the synapses should not be
-allowed to optimise all of this noise away. It would be interesting
-to see how our results change if we put lower limits on the noise
-levels. I think this would be beyond the scope of this work, but
-should be added to the discussion section.
+should be fixed in fig5, showing the solid red curve hugging the
+green one at late times.
+We agree that the level of noise tolerance is an important
+consideration. Note that the stochasticity of the models is an
+expression of biological noise. In fact, the type of noise
+described by the reviewer could be included by adding extra states
+to the model with the different transition rates. However, the
+synapses should not be allowed to optimise this noise away.
+It would be interesting to see how our results change if we put
+lower limits on the noise levels, or incorporated noise in the
+manner suggested by the reviewer. I think this would be beyond the
+scope of this work, but could be mentioned in the discussion section.
 
-Reviewer 2: We'll add a derivation of equation 3 to the supplement.
+Reviewer 2:
+We'll add a derivation of equation 3 to the supplement.
 We could add a description in the main text along the lines of:
@@
 in the main paper. I'm not sure how we could incorporate more into
-the main paper without exceeding the page limit.
+the main paper without exceeding the page limit. Section 1 of the
+supplement consists of simple generalisations of discrete time
+results to continuous time, so it doesn't really merit inclusion in
+the main text. We put references at the start of the section. Would
+it make it clearer if we added these references throughout the section?
 
@@
 states. Making the link between molecular and functional states
-will, unfortunately, be much harder
+will, unfortunately, be much harder.
 line 294: agreed.
@@
 account. In equation (57) of the supplement, the term involving c_g
-comes from this dependence. If the reviewer is concerned that tho
+comes from this dependence. If the reviewer is concerned that the
 order of the eta_i could change during the maximisation procedure:
@@
 Therefor we need only consider an infinitesimal neighbourhood of
-the model, in which the order will not change.
+the model, in which the order will not change. Should this be
+mentioned in section 3.2?
 line 304: Yes. We'll add this phrase to the text.
@@
 +/- states, asymmetry would make them worse. For fig4, the effect
-of the asymmetry would be reduced in the limit as epsilon -> 0, as
+of the asymmetry would go away in the limit as epsilon -> 0, as
 all states other than the end states would have very small
-p^infinity.
+p^infinity. Should this be mentioned in section 4?
 line 415: It is only for t_0 > sqrt(N)M that the models that nearly
-touch the envelope are linear chains. The model in fig5b is not a
-linear chain, as it has shortcut transitions, but they are only the
+touch the envelope are linear chains. The models in fig5b is not a
+linear chain, as it has shortcut transitions, but these are only the
 best models (that we have found) for times t_0 < sqrt(N)M. Would it

### 2013-08-16 NIPS/SynapseNIPSsupp.tex
fixed bookmark in supp
working on teo-time env
updated NIPS supp

@@
 \begin{defn}[Fundamental matrix]
+  For discrete time, the generalized fundamental matrix was defined in \cite{Kemeny1981const}.
+  For continuous time, we define:
   %
@@
 %
-or in component form of \eqref{eq:fptfund}:
+or in component form:
 %
@@
 
-\begin{thm}\label{th:kemenyconst}
+\begin{thm} \mbox{}\label{th:kemenyconst}
   The quantity
@@
 \begin{proof}
-  We use \eqref{eq:fptfund}, \eqref{eq:fundrowsum} and the transpose of \eqref{eq:pdotD}:
+  For discrete time, a proof can be found in \cite[Th.4.4.10]{kemeny1960finite}.
+  For continuous time, we use \eqref{eq:fptfund}, \eqref{eq:fundrowsum} and the transpose of \eqref{eq:pdotD}:
   %
@@
 
-Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version \cite{hunter2006mixing} where this would only shift $\eta$ by 1.
+Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version, where this would only shift $\eta$ by 1.
 
@@
 %
+\begin{equation}\label{eq:SNRdef}
+  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}}
+     {\sqrt{\var(\wv_\text{ideal}\cdt\wv(\infty))}}.
+\end{equation}
+%
+First, let's look at the denominator, remembering that the states and plasticity events of each synapse are independent and identically distributed:
+%
+\begin{equation}\label{eq:noise}
+\begin{aligned}
+  \var(\wv_\text{ideal}\cdt\wv(\infty))
+    &= \sum_{\alpha\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty) \wv^\beta_\text{ideal}\wv^\beta(\infty)}
+    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2 \\
+    &= \sum_{\alpha} \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}
+    + \sum_{\alpha\neq\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}\av{\smash{\wv^\beta}_\text{ideal}\smash{\wv^\beta}(\infty)}
+    \\&\phantom{= \sum_{\alpha}  \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}}
+    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2  \\
+    &= N\av{1}
+    + N(N-1)\av{\wv^1_\text{ideal}\wv^1(\infty)}^2
+    - N^2\av{\wv^1_\text{ideal}\wv^1(\infty)}^2 \\
+    &= N(1-\av{\wv^1_\text{ideal}\wv^1(\infty)}^2),
+\end{aligned}
+\end{equation}
+%
+where we used $\wv^\alpha=\pm1$.
+
+For the numerator, we can write
+%
+\begin{equation}\label{eq:overlapex}
+  \av{\wv_\text{ideal}\cdt\wv(t)} = \sum_\alpha\av{\wv^\alpha_\text{ideal}\wv^\alpha(t)}
+   = N \av{\wv^1_\text{ideal}\wv^1(t)},
+\end{equation}
+%
+Noting that $\wvi=\pm1$ with probability $f\potdep$,
+%
+\begin{equation}\label{eq:overlapstate}
+\begin{aligned}
+  \av{\wv^1_\text{ideal}\wv^1(t)} &= f\pot \av{\wv^1(t)}_{\text{pot},t=0} - f\dep \av{\wv^1(t)}_{\text{dep},t=0} \\
+   &= f\pot \sum_i P(\text{state}=i,t \mid \text{pot},0) \w_i - f\dep \sum_i P(\text{state}=i,t \mid \text{dep},0) \w_i.
+\end{aligned}
+\end{equation}
+%
+From the previous section,
+%
+\begin{equation}\label{eq:stateprob}
+  P(\text{state}=i,t \mid \text{pot/dep},0) = \brk{ \eq \M\potdep\, \e^{rt\frg} }_i,
+\end{equation}
+%
+which describes the synapses starting in the equilibrium distribution, changing state due to the plasticity event at $t=0$ and subsequent evolution according to \eqref{eq:rowsum} due to plasticity events uncorrelated with the special event at $t=0$.%
+\footnote{Note that expanding the exponential gives
+%
+\begin{equation*}
+  \e^{rt\frg} =  \sum_{n=0}^\infty \frac{(rt)^n\,\e^{-rt}}{n!} \sum_{m=0}^n (f\pot)^m (f\dep)^{n-m}
+  \brk{ \M\pot \M\dep \M\pot \M\pot \ldots + \text{permutations}}.
+\end{equation*}
+%
+Thus, evolving according to \eqref{eq:rowsum} results in averaging over all sequences of plasticity events,
+as we only need linear expectations of $\wv(t)$ in the end.}
+This results in
+%
+\begin{equation}\label{eq:overlapwt}
+\begin{aligned}
+  \av{\wv^1_\text{ideal}\wv^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
+  \av{\wv^1_\text{ideal}\wv^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
+         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
+         &=  (f\pot-f\dep)\, \eq \w \\
+         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
+\end{aligned}
+\end{equation}
+%
+Combining these allows us to write the numerator as
+%
+\begin{equation}\label{eq:signal}
+\begin{aligned}
+  \av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}
+    &= N \eq (f\pot(\M\pot-\I)-f\dep(\M\dep-\I))\, \e^{rt\frg} \w \\
+    &= N \eq (f\pot(\W\pot-\frg)-f\dep(\W\dep-\frg))\, \e^{rt\frg} \w \\
+%    &= N \eq (f\pot(f\dep\W\pot-f\dep\W\dep)-f\dep(f\pot\W\dep-f\pot\W\pot))\,\e^{rt\frg}\w \\
+    &= N (2f\pot f\dep)\eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
+\end{aligned}
+\end{equation}
+%
+where we used $\eq\frg=0$ in going from the first to second lines.
+Combining with \eqref{eq:noise} gives
+%
+\begin{equation}\label{eq:SNRcurveExact}
+  \SNR(t) = \frac{\sqrt{N} (2f\pot f\dep ) \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w}
+                 {\sqrt{1-(f\pot-f\dep)^2(\eq_+-\eq_-)^2}}.
+\end{equation}
+%
+The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
+We can set the denominator to 1 without changing any of our results.
+
+Which results in our final formula:
+%
 \begin{equation}\label{eq:SNRcurve}
-  \SNR(t) = \sqrt{N} (2f\pot f\dep ) \eq (\W\pot  - \W\dep )\e^{rt\frg}\w.
+  \SNR(t) = \sqrt{N} (2f\pot f\dep )\, \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
 \end{equation}
@@
 
-\subsection{Lumpability}\label{sec:SNRlump}
+\subsection{Lumpability and the SNR curve}\label{sec:SNRlump}
 
@@
 
-Using $\eq\frg=0$, we can write the initial SNR as
+Using $\eq\frg=0$ and the first line of \eqref{eq:signal}, we can write the initial SNR as
 %
@@
 
-\subsection{Derivatives \wrt \texorpdfstring{$\W\potdep $}{W+/-}}\label{sec:deriv}
+\subsection{Derivatives \wrt \texorpdfstring{$\W\potdep $}{W(pot/dep)}}\label{sec:deriv}
 

### 2013-08-26 NIPS/SynapseNIPS.tex
working on NIPS rebuttal
updated NIPS supp
added reviewer changes to NIPS paper, too long

@@
 
-%\nipsfinalcopy % Uncomment for camera-ready version
+\nipsfinalcopy % Uncomment for camera-ready version
 
@@
 Thus the $i$'th component of $\pr(t)$ denotes the fraction of the synaptic population in state $i$.
-Furthermore, each state $i$ has its own synaptic weight, which we take, in the worst case scenario, to be restricted to two values.
+Furthermore, each state $i$ has its own synaptic weight, $\w_i$, which we take, in the worst case scenario, to be restricted to two values.
 After shifting and scaling these two values,  we can assume they are $\pm1$, without loss of generality.
@@
 \begin{equation}\label{eq:SNRdef}
-  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t) - \wv_\text{ideal}\cdt\wv(\infty)}}
+  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}}
      {\sqrt{\var(\wv_\text{ideal}\cdt\wv(\infty))}}.
@@
 \begin{equation}\label{eq:SNRcalc}
-  \snr(t) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \e^{rt\frg} \w,
+  \snr(t) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \e^{rt\frg} \w.
 \end{equation}
 %
-where the weight of the synapse when it is in its $i$'th state is given by the corresponding element of the column vector $\w$.
+The factor of $\eq$ describes the synapses being in the steady-state distribution before the memory is encoded.
+The factor of $(\M\pot-\M\dep)$ comes from the encoding of the memory at $t=0$, with $\wvi$ being $\pm1$ in synapses that are potentiated/depotentiated.
+The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
+A detailed derivation can be found in the supplementary material.
 We will frequently refer to this function as the memory curve.
@@
 %
-We can then ask the question: what are the constraints on these quantities, namely initial SNR's $\CI_a$ and time constants $\tau_a$ for the modes,  implied by the constraints in \eqref{eq:constr}?
+We can then ask the question: what are the constraints on these quantities, namely initial SNR's, $\CI_a$, and time constants, $\tau_a$, for the modes,  implied by the constraints in \eqref{eq:constr}?
 We will derive some of these constraints in the next section.
@@
 %
-The initial SNR is closely related to the flux from the states with $\w_i=-1$ to those with $\w_j=+1$:
+The initial SNR is closely related to the flux from the states with $\w_i=-1$ to those with $\w_j=+1$ (see supplementary material):
 %
@@
 In essence, in this order, a state is considered to be more potentiated if the average time it takes to reach all the strong efficacy states is shorter.
-We will see that synaptic models that optimize various measures of memory have an exceedingly simple structure when, and only when, their states are arranged in this order.
+We will see that synaptic models that optimize various measures of memory have an exceedingly simple structure when, and only when, their states are arranged in this order.%
+\footnote{Note that we do not need to worry about the order of the $\eta^\pm_i$ changing during the optimization: necessary conditions for a maximum only require that there is no infinitesimal perturbation that increases the area.
+Therefore we need only consider an infinitesimal neighborhood of the model, in which the order will not change.}
 
@@
 
-First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \eqref{eq:kemenypm}:
+First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \eqref{eq:kemenypm} (see supplementary material):
 %
@@
   (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$.
-It can no longer be used when $\M\dep$ is lower triangular.
+It can no longer be used when $\M\dep$ is lower triangular, \ie depression must move synapses to ``more depressed'' states.
   (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$.
-It can no longer be used when $\M\pot$ is upper triangular.
+It can no longer be used when $\M\pot$ is upper triangular, \ie potentiation must move synapses to ``more potentiated'' states.
   (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions.
@@
 As these perturbations do not change the equilibrium distribution, this means that the area of \emph{any} model is bounded by that of a linear chain with the same equilibrium distribution.
-The area of a linear chain model can be expressed directly in terms of its equilibrium state distribution, $\eq$, yielding the following upper bound on the area of any model with the same $\eq$:
+The area of a linear chain model can be expressed directly in terms of its equilibrium state distribution, $\eq$, yielding the following upper bound on the area of any model with the same $\eq$ (see supplementary material):
 %
@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.38\linewidth]{env.eps}}\label{fig:envplot}\hp
-  \item\aligntop{\includegraphics[width=0.3\linewidth]{diffjump.svg}}\label{fig:diffjump}
+ \parbox[t]{0.45\linewidth}{
+  \item\aligntop{\includegraphics[width=0.95\linewidth]{env.svg}}\label{fig:envplot}
+  }\hp
+  \parbox[t]{0.45\linewidth}{
+  \item\aligntop{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}\\[0.1\linewidth]
+  \item\aligntop{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
+  }
  \end{myenuma}
@@
   At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states, and all transition probabilities equal to $1$.
-  At late times they are of the form shown in \autoref{fig:max_area} with different values of $\varepsilon$.
+  At late times they are of the form shown in (\ref{fig:sticky}) with different values of $\varepsilon$.
   }\label{fig:env}
@@
 It would be interesting to systematically analyze the space of models that achieve good memory at multiple times, and understand their structural organization, and how they give rise to multiple exponentials, leading to power law memory decays.
+
+Another important consideration is the level of noise tolerance of the synapse. 
+Note that the stochasticity of the models is an expression of biological noise. 
+However, the synapses should not be allowed to optimize all of this noise away. 
+It would be interesting to consider such limits on the noise levels.
+
 Finally, it would be interesting to design physiological experiments in order to perform optimal systems identification of potential Markovian dynamical systems hiding within biological synapses.
+If we measured pre and post-synaptic spike trains, and also recorded changes in post-synaptic potentials to measure changes in synaptic weight, we could use hidden Markov model techniques to find the best-fit synaptic model.
+Then given our theory, we could match this measured synaptic model to optimal models to infer which timescales the synapse operates on.
 
@@
 
-%\subsubsection*{Acknowledgements}
-%We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
-%We thank larry Abbott, Marcus Benna, Stefano Fusi, Jascha Sohl-Dickstein and David Sussillo for useful discussions.
+\subsubsection*{Acknowledgements}
+We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
+We thank Larry Abbott, Marcus Benna, Stefano Fusi, Jascha Sohl-Dickstein and David Sussillo for useful discussions.
 

### 2013-08-26 NIPS/SynapseNIPSsupp.tex
added reviewer changes to NIPS paper, too long

@@
 %
-which describes the synapses starting in the equilibrium distribution, changing state due to the plasticity event at $t=0$ and subsequent evolution according to \eqref{eq:rowsum} due to plasticity events uncorrelated with the special event at $t=0$.%
+which describes the synapses starting in the equilibrium distribution, changing state due to the plasticity event at $t=0$ and subsequent evolution according to \eqref{eq:rowsum} due to plasticity events uncorrelated with $\wvi$.%
 \footnote{Note that expanding the exponential gives

### 2013-08-27 NIPS/SynapseNIPS.tex
typo in poster, removed repeated fig in NIPS paper

@@
 
-The extent to which the memory has been stored is described by a signal-to-noise ratio (SNR) \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}:
+The extent to which the memory has been stored is described by a signal-to-noise ratio (SNR) \cite{Fusi2005cascade,Fusi2007multistate}:
 %
@@
 With the states in the order described above, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
-Some of these perturbations are shown in \autoref{fig:perts}.
-For example, in \autoref{fig:perts}\ref{fig:tri_right}, for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
+Some of these perturbations are shown in \autoref{fig:perts}, see supplementary material for details.
+For example, in \autoref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
 $\eta^+_i > \eta^+_j$), we have proven that increasing $\M\pot_{ij}$ and decreasing $\M\dep_{ij}$ leads to an increase in area.
@@
 This determines the topology (non-zero transition probabilities) of the model with maximal area.
-It is of the form shown in \autoref{fig:max_area}, with potentiation moving one step to the right and depression moving one step to the left.
+It is of the form shown in \autoref{fig:env}(\ref{fig:sticky}),%\autoref{fig:max_area}, 
+with potentiation moving one step to the right and depression moving one step to the left.
 Any other topology would allow some perturbation to further increase the area.
@@
 We can then maximize this by pushing all of the equilibrium distribution symmetrically to the two end states.
-This can be done by reducing the transition probabilities out of these states, as in \autoref{fig:max_area}.
+This can be done by reducing the transition probabilities out of these states, as in \autoref{fig:env}(\ref{fig:sticky}).%\autoref{fig:max_area}.
 This makes it very difficult to get out of the end states once they have been entered.
@@
 The ``sticky'' end states result in very slow decay of memory, but they also make it difficult to encode the memory in the first place, since a small fraction of synapses are able to change synaptic efficacy during the storage of a new memory.
-Thus models that optimize area optimize memory at late times, at the expense of early times.
+Thus models that maximize area optimize memory at late times, at the expense of early times.
 
-\begin{figure}[tbp]
- \begin{center}
-  \includegraphics[width=0.35\linewidth]{multistate_sticky.svg}
- \end{center}
-  \caption{Model that maximizes the area.
-  Unlabeled transitions have probability 1.
-  Labeled transitions have probability $\varepsilon\to0$.
-  }\label{fig:max_area}
-\end{figure}
+%\begin{figure}[tbp]
+% \begin{center}
+%  \includegraphics[width=0.35\linewidth]{multistate_sticky.svg}
+% \end{center}
+%  \caption{Model that maximizes the area.
+%  Unlabeled transitions have probability 1.
+%  Labeled transitions have probability $\varepsilon\to0$.
+%  }\label{fig:max_area}
+%\end{figure}
 
@@
 However, we suspect that the area constraint is not the bottleneck for optimizing memory at times less than  $\CO(\frac{M}{r})$.
-We believe there is some other constraint that prevents models from approaching the envelope, and currently are exploring several mathematical conjectures for the precise form of this constraint, in order to obtain a potentially tighter envelope.
-Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \autoref{fig:max_area} can come close to the envelope.
+We believe there is some other constraint that prevents models from approaching the envelope, and currently are exploring several mathematical conjectures for the precise form of this constraint in order to obtain a potentially tighter envelope.
+Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \autoref{fig:env}(\ref{fig:sticky})%\autoref{fig:max_area}
+ can come close to the envelope.
 
@@
   At late times they are of the form shown in (\ref{fig:sticky}) with different values of $\varepsilon$.
+  The model shown in (\ref{fig:sticky}) also saturates the area bound \eqref{eq:max_area} in the limit $\varepsilon\to0$.
   }\label{fig:env}

### 2013-09-21 MHCpaper/MHCsupp.tex
pooled model for dep only

@@
 \newcommand{\wt}{_{\text{WT}}}
-\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
-\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
+%\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
+%\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
+\newcommand{\ko}{_{\text{MHC}}}
+\newcommand{\KO}{MHC}
 \newcommand{\tpre}{t_{\text{pre}}}
@@
 We will use larger values for depression in the \KO\ mutant.
+In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
+This is done by setting $q\pot\lmax=q\pot\lmin$.
 The values of these parameters are listed in \autoref{tab:params}.
@@
     % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
+    Model & \# states & pot & WT dep & \KO\ dep & $\Delta f$ & $r\tpre$ \\
     \cline{1-7}
-    Serial        & 10 & $q=0.3$  & $q=0.4$  & 0.1  & 5 & 20  &\label{tr:multistate_weak} \\
-    Serial        & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 20  &\label{tr:multistate_med} \\
-    Serial        & 10 & $q=0.3$  & $q=0.4$  & 0.45 & 5 & 30  &\label{tr:multistate_strong} \\
-    Two-state     & 2  & $q=0.1$  & $q=0.2$  & 0.1  & 5 & 5   &\label{tr:binary} \\
-    Multistate    & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 5   &\label{tr:multistate_lin} \\
-    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-                                          & 0.1 & 5 & 20 &\label{tr:pooled_plenty}\\
-    Pooled        & 10 & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
-                                          & 0.1 & 5 & 20 &\label{tr:pooled_scarce}\\
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & 0.3  & 5 & 20  &\label{tr:cascade_short} \\
-    Cascade       & 10 & $x=0.25$ & $x=0.33$ & 0.3  & 5 & 100 &\label{tr:cascade_long} \\
-    Non-uniform MS       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:nonuni} \\
+    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.1  & 20  &\label{tr:multistate_weak} \\
+    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 20  &\label{tr:multistate_med} \\
+    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.45 & 30  &\label{tr:multistate_strong} \\
+    Two-state     & 2  & $q=0.1$  & $q=0.1$  & $q=0.2$  & 0.1  & 5   &\label{tr:binary} \\
+    Multistate    & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 5   &\label{tr:multistate_lin} \\
+    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
+                                          & 0.1 & 20 &\label{tr:pooled_plenty}\\
+    Pooled        & 10 & $q\in[0.05,0.4]$ & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
+                                          & 0.1 & 20 &\label{tr:pooled_scarce}\\
+    Pooled        & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
+                                          & 0.4 & 20 &\label{tr:pooled_deponly}\\
+    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 20  &\label{tr:cascade_short} \\
+    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 100 &\label{tr:cascade_long} \\
+    Non-uniform MS& 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 150 &\label{tr:nonuni} \\
     \cline{1-7}
@@
 
+This impaired learning after pre-training in the wild-type is caused by the excessive potentiation pushing the synapses away from the central transition that generates the learning signal.
+Essentially, this model has a form of metaplasticity where repeated potentiation makes subsequent depression harder.
+This allows excessive pre-training to impair learning, despite increasing the number of potentiated synapses.
+
 Now, consider the mutant, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
@@
 The first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ and will be more significant for smaller $\beta$ or in a longer chain.
+This exponential decay in the initial distribution is what allows the saturation bias to overwhelm the increases intrinsic plasticity rate and lead to impaired learning with enhanced plasticity.
+It is important that a learning signal is only generated by the central transition so that the exponential decay can take effect.
 
@@
 
+Intuitively, this model is missing two features of the \hyperref[sec:multistate]{multistate model}.
+First, it does not have the exponential amplification of the effect of initial saturation bias as this model does not have a chain of states for the distribution to decay across.
+Second, it does not have the metaplastic effect where repeated potentiation makes future depression harder, as potentiation will merely increase the number of potentiated synapses.
+
 \subsection{Multistate model}\label{sec:multistate_lin}
@@
 
+Like the \hyperref[sec:binary]{two state model}, this model is missing two features of the \hyperref[sec:multistate]{multistate model}.
+First, it does not have the amplification of the effect of initial saturation bias as every transition contributes to the learning signal, so the exponentially decaying distribution has no effect.
+Second, it does not have the metaplastic effect where repeated potentiation makes future depression harder, as potentiation will merely increase the number of potentiated synapses without pushing them away from any boundary between strong and weak states.
+
 \subsection{Pooled resource model}\label{sec:pooled}
@@
 \resultsfig{pooled resource}{heavy depletion}{($q\lmin/q\lmax=0.125$)}{pooled_scarce}
+
+\resultsfig{pooled resource}{unaffected potentiation}{($q\pot\lmax=q\pot\lmin$)}{pooled_deponly}
 
@@
 This differs from what is seen experimentally, where the pre-trained wild-type learns slower than the untrained one.
+
+The situation discussed in the main text, with resource depletion for depression only and potentiating transitions  unaffected, is shown in \autoref{fig:pooled_deponly}.
+It shows the same learning patterns as the more symmetric versions.
+
+This model shares one feature of the \hyperref[sec:multistate]{multistate model}.
+The effect of initial saturation bias is amplified by resource depletion, allowing for impaired learning with enhanced plasticity.
+However, it does not have the metaplastic effect where repeated potentiation makes future depression harder, in fact it shows the opposite effect due to resource replenishment, so it can never show impaired learning after pre-training.
 
@@
 This effect is weaker for the mutant, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
+
+This model shares both features of the \hyperref[sec:multistate]{multistate model}.
+First, the effect of initial saturation bias is amplified by the exponential decay of transition probabilities away from the middle.
+Second, it has the metaplastic effect where repeated potentiation makes future depression harder, also due to he exponential decay of transition probabilities.
 
@@
 
+Like the \hyperref[sec:cascade]{cascade model}, this model also shares both features of the \hyperref[sec:multistate]{multistate model}.
+First, the effect of initial saturation bias is amplified by the exponential decay of transition probabilities away from the middle.
+Second, it has the metaplastic effect where repeated potentiation makes future depression harder, also due to he exponential decay of transition probabilities.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-09-21 Code/VOR/MakeVORcomparison.m

@@ -36,8 +36,10 @@ elseif strcmpi(modelname,'nonuni')
     [~,WmKn]=CascadeOriginal(paramKn,paramKn,n,0);
 elseif strcmpi(modelname,'pooled')
     w=(-1:2/(n-1):1)';
-    q=(paramWT(1):diff(paramWT)/(n-2):paramWT(2)).*(n-1:-1:1)/(n-1);
-    [Wp,Wm]=MakeSMS(q);
+    q=(paramWT(3))*(n-1:-1:1)/(n-1);
+    [Wp,~]=MakeSMS(q);
+    q=(paramWT(1):diff(paramWT(1:2))/(n-2):paramWT(2)).*(n-1:-1:1)/(n-1);
+    [~,Wm]=MakeSMS(q);
     q=(paramKn(1):diff(paramKn)/(n-2):paramKn(2)).*(n-1:-1:1)/(n-1);
     [~,WmKn]=MakeSMS(q);
 else

### 2013-09-21 Code/VOR/MakeVORfigs.m

@@ -45,7 +45,7 @@ elseif strncmpi(prefix,'binary',3)
     n=2;
 elseif strncmpi(prefix,'pooled',3)
     modelname='pooled';
-    n=10;
+    n=7;
 end
 
 

### 2013-09-21 MHCpaper/Figures/MakeFigs.m

@@ -8,3 +8,4 @@ MakeVORfigs('pooled_scarce',[0.05 0.4],[0.1 0.8],-0.1,5,20,'print','Pooled',true
 MakeVORfigs('cascade_short',0.25,0.33,-0.3,5,20,'print')
 MakeVORfigs('cascade_long',0.25,0.33,-0.3,5,100,'print')
 MakeVORfigs('nonuni',0.25,0.33,-0.3,5,100,'print','LinWeights',true)
+MakeVORfigs('pooled_deponly',[0.6 0.0006 0.008],[1 0.001],-0.4,5,20,'print','Pooled',true)

### 2013-10-01 MHCpaper/MHCsupp.tex
typos, ref and name of MHC mutant

@@
 %\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
-%\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
+\newcommand{\KO}{D$^\mathrm{b}$K$^{\mathrm{b}-/-}$}
 \newcommand{\ko}{_{\text{MHC}}}
-\newcommand{\KO}{MHC}
+%\newcommand{\KO}{MHC}
 \newcommand{\tpre}{t_{\text{pre}}}
@@
 
-The evolution of this probability is described by a forgetting matrix, $\frg$:
+The evolution of this probability is described by a continuous time Markov matrix matrix, $\frg$:
 %
@@
 
-The MHC mutant has a lower threshold for depression.
+The \KO\ mice have a lower threshold for depression.
 We can model this by changing $\M\dep\wt$ to $\M\dep\ko$, which should have larger off-diagonal matrix elements.
@@
 
-For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
-We will use a larger value for $q$ for depression in the \KO\ mutant.
+For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
+We will use a larger value for $q$ for depression in the \KO\ mice.
 
-For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
-We will use a larger value for $x$ for depression in the \KO\ mutant.
+For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
+We will use a larger value for $x$ for depression in the \KO\ mice.
 
@@
 This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
-We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
-We will use larger values for depression in the \KO\ mutant.
+We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
+We will use larger values for depression in the \KO\ mice.
 In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
@@
 This model is parameterised by the values of $x$ chosen for potentiation and depression.
-We will use a larger value for depression in the \KO\ mutant.
+We will use a larger value for depression in the \KO\ mice.
 The values used are listed in \autoref{tab:params}.
@@
 We will model this by changing $f\potdep$, leaving $\M\potdep$ unchanged.
+This approach was used to model motor learning in \cite{Smith2006savings}.
 
@@
 Therefore the rates of potentiation and depression should not change very much either.
-This implies that the mean synaptic weight in equilibrium differs between the mutant and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
+This implies that the mean synaptic weight in equilibrium differs between the \KO\ mice and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
 As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
@@
 For the most part, we set $f\dep\norm=\half$, $f\dep\inc=f\dep\norm+\Delta f$ and $f\dep\dec=f\dep\norm-\Delta f$, with $\Delta f>0$.
-We use the same values for wild-type and mutant for the reasons discussed above.
+We use the same values for wild-type and \KO\ mice for the reasons discussed above.
 We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
@@
 
-We are also assuming that the relation between VOR gain and mean synaptic weight is the same for mutant ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
+We are also assuming that the relation between VOR gain and mean synaptic weight is the same for \KO\ mice ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
 This ensures that the qualitative questions mentioned above (enhancement or impairment of learning) will not be affected.
@@
 \begin{enumerate}
-  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the mutant.
+  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the \KO\ mice.
   \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
-  \item For the mutant, gain-increase learning is significantly faster with pre-training than without it.
-  \item After pre-training, gain-increase learning is significantly faster in the mutant than in the wild-type.
+  \item For the \KO\ mice, gain-increase learning is significantly faster with pre-training than without it.
+  \item After pre-training, gain-increase learning is significantly faster in the \KO\ mice than in the wild-type.
   \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
@@
 We will not worry too much about the curvature of the learning curves, as this can be changed by the nonlinear relation between synaptic weight and VOR gain.
-However, both the wild-type and mutant should be affected in the same way, so the difference in the curvature of the mutant and wild-type curves after pre-training would be nice to capture.
+However, both the wild-type and \KO\ mice should be affected in the same way, so the difference in the curvature of the \KO\ mice and wild-type curves after pre-training would be nice to capture.
 
@@
 This should be multiplied by the difference in synaptic weights.
-But, as this quantity is constant within a model and doesn't change when comparing wild-type and mutant or with/without pre-training, it can be ignored.
+But, as this quantity is constant within a model and doesn't change when comparing wild-type and \KO\ mice or with/without pre-training, it can be ignored.
 
@@
     Multistate    & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 5   &\label{tr:multistate_lin} \\
-    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-                                          & 0.1 & 20 &\label{tr:pooled_plenty}\\
-    Pooled        & 10 & $q\in[0.05,0.4]$ & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
-                                          & 0.1 & 20 &\label{tr:pooled_scarce}\\
-    Pooled        & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
+%    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
+%                                          & 0.1 & 20 &\label{tr:pooled_plenty}\\
+%    Pooled        & 10 & $q\in[0.05,0.4]$ & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
+%                                          & 0.1 & 20 &\label{tr:pooled_scarce}\\
+    Pooled res. & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
                                           & 0.4 & 20 &\label{tr:pooled_deponly}\\
@@
     Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 100 &\label{tr:cascade_long} \\
-    Non-uniform MS& 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 150 &\label{tr:nonuni} \\
+    Non-uni. MS& 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 150 &\label{tr:nonuni} \\
     \cline{1-7}
@@
   Time is measured in units of $1/r$, the mean time between plasticity events.
-  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
+  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mice with and without pre-training.
   (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant,
+  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mice,
   with states of weak synaptic strength to the left and strong states to the right.
-  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
+  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ mice without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
   with states of weak synaptic strength at the top and strong states underneath. }
@@
 
-Now, consider the mutant, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
+Now, consider the \KO\ mice, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
 Without pre-training:
@@
 Values of $\Delta f$ that are larger than this (stronger training) will correspond to slower learning after pre-training.
-As we would like pre-training to slow down learning in the wild-type but speed it up in the mutant, it would seem that we require $\Delta f^*(1,M) < \Delta f < \Delta f^*(\beta,M)$ (remembering that the wild-type corresponds to $\beta=1$).
+As we would like pre-training to slow down learning in the wild-type but speed it up in the \KO\ mice, it would seem that we require $\Delta f^*(1,M) < \Delta f < \Delta f^*(\beta,M)$ (remembering that the wild-type corresponds to $\beta=1$).
 As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, so this range always exists.
@@
 However, this is only the initial, instantaneous rate of change.
-Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and mutant.
+Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and \KO\ mice.
 We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
-This happens sooner for the mutant than the wild-type, due to the stronger depressing transitions.
-This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the mutant, even outside the desired range of values for $\Delta f$.
+This happens sooner for the \KO\ mice than the wild-type, due to the stronger depressing transitions.
+This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the \KO\ mice, even outside the desired range of values for $\Delta f$.
 As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} (black curves), even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.
@@
  \end{center}
-  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
+  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the \KO\ mice have impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
 \end{figure}
 
-In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the mutant learns slower than wild-type without pre-training, and that pre-training speeds up learning in the mutant but slows it down in the wild-type.
-However, in this case pre-training proceeds much faster in the mutant than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.
+In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the \KO\ mice learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ mice but slows it down in the wild-type.
+However, in this case pre-training proceeds much faster in the \KO\ mice than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.
 
@@
 
-We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the mutant, the important thing is to have \emph{strong pre-training}, not strong training.
-Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the mutant than wild-type.
+We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the \KO\ mice, the important thing is to have \emph{strong pre-training}, not strong training.
+Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the \KO\ mice than wild-type.
 
-Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the mutant (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
+Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the \KO\ mice (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
 This is due to the mass of probability at the right end of the chain (see \autoref{fig:multistate_med}\ref{fig:multistate_med_eq} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) that is unable to contribute to change in synaptic weight until it has drifted to the centre.
-This will last for a shorter time in the mutant, due to the enhanced transitions.
+This will last for a shorter time in the \KO\ mice, due to the enhanced transitions.
 This will not be a feature of any of the other models that we study here.
@@
 Here $M=2$, so the leading part of the two effects will cancel.
-The subleading effects (the normalisation of the probabilities) are responsible for the faster learning in the mutant.
+The subleading effects (the normalisation of the probabilities) are responsible for the faster learning in the \KO\ mice.
 
@@
 First, we assumed that the wild type had symmetric potentiation and depression.
-Relaxing this assumption would correspond modelling the wild-type like the mutant, but with a different value for $\beta$.
-We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the mutant.
+Relaxing this assumption would correspond modelling the wild-type like the \KO\ mice, but with a different value for $\beta$.
+We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the \KO\ mice.
 
@@
 
-The only thing we have to worry about is when the wild-type catches up with the mutant.
-As the total change will be larger for the wild-type than mutant, it must overtake eventually.
-If this happens early enough, the initial period, where the mutant learns faster than wild-type, would not be seen in the experiment.
+The only thing we have to worry about is when the wild-type catches up with the \KO\ mice.
+As the total change will be larger for the wild-type than \KO\ mice, it must overtake eventually.
+If this happens early enough, the initial period, where the \KO\ mice learns faster than wild-type, would not be seen in the experiment.
 However, the timescale for this crossover would be similar to the timescale of the exponentials.
@@
 
-Intuitively, this model is missing two features of the \hyperref[sec:multistate]{multistate model}.
-First, it does not have the exponential amplification of the effect of initial saturation bias as this model does not have a chain of states for the distribution to decay across.
+Intuitively, this model is missing two features of the \hyperref[sec:multistate]{serial model}.
+First, it does not have the exponential amplification of the effect of initial saturation bias because this model does not have a chain of states for the distribution to decay across.
 Second, it does not have the metaplastic effect where repeated potentiation makes future depression harder, as potentiation will merely increase the number of potentiated synapses.
@@
 This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
-This means that the mutant will initially learn faster than the wild-type.
+This means that the \KO\ mice will initially learn faster than the wild-type.
 However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS} (solid curves).
@@
 which is always greater than 1.
-Therefore, pre-training will always enhance learning, for both mutant and wild-type, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
+Therefore, pre-training will always enhance learning, for both \KO\ mice and wild-type, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
 
-Like the \hyperref[sec:binary]{two state model}, this model is missing two features of the \hyperref[sec:multistate]{multistate model}.
+Like the \hyperref[sec:binary]{two state model}, this model is missing two features of the \hyperref[sec:multistate]{serial model}.
 First, it does not have the amplification of the effect of initial saturation bias as every transition contributes to the learning signal, so the exponentially decaying distribution has no effect.
@@
 \subsection{Pooled resource model}\label{sec:pooled}
+%
+%
+%\resultsfig{pooled resource}{light depletion}{($q\lmin/q\lmax=0.75$)}{pooled_plenty}
+%
+%\resultsfig{pooled resource}{heavy depletion}{($q\lmin/q\lmax=0.125$)}{pooled_scarce}
 
-\resultsfig{pooled resource}{light depletion}{($q\lmin/q\lmax=0.75$)}{pooled_plenty}
+\resultsfign{pooled resource}{pooled_deponly}
 
-\resultsfig{pooled resource}{heavy depletion}{($q\lmin/q\lmax=0.125$)}{pooled_scarce}
-
-\resultsfig{pooled resource}{unaffected potentiation}{($q\pot\lmax=q\pot\lmin$)}{pooled_deponly}
-
-The results of simulations of the pooled resource model can be seen in \autoref{fig:pooled_plenty} and \autoref{fig:pooled_scarce}.
+The results of simulations of the pooled resource model can be seen in \autoref{fig:pooled_deponly}.
 It is difficult to study this model analytically.
@@
 
-If we compare gain-increase learning in the mutant to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_eq}: \ref{fig:pooled_plenty_eq_WT} and \ref{fig:pooled_plenty_eq_KO} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_eq}: \ref{fig:pooled_scarce_eq_WT} and \ref{fig:pooled_scarce_eq_KO}, blue traces), where there are fewer synapses available for depression and resources are depleted.
-When resource depletion is less severe, the first effect dominates and the mutant learns faster than wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learn},\ref{fig:pooled_plenty_learnS}, solid curves).
-This is reasonable, as when resource depletion is removed it will reduce to the two-state model, which showed this behaviour.
-However, when resource depletion is more severe, the second effect dominates and the mutant learns slower than wild-type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learn},\ref{fig:pooled_scarce_learnS}, solid curves), which matches what is seen in the experiment.
+If we compare gain-increase learning in the \KO\ mice to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}: \ref{fig:pooled_deponly_eq_WT} and \ref{fig:pooled_deponly_eq_KO}, blue traces), where there are fewer synapses available for depression and resources are depleted.
+When resource depletion is sufficiently severe, the second effect dominates and the \KO\ mice learn slower than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}, solid curves), which matches what is seen in the experiment.
 
-Gain-decrease pre-training lessens the second effect, and results in the mutant learning faster than wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learn},\ref{fig:pooled_plenty_learnS} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learn},\ref{fig:pooled_scarce_learnS}), as seen in the experiment.
+Gain-decrease pre-training lessens the second effect, and results in the \KO\ mice learning faster than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}), as seen in the experiment.
 
-Gain-decrease pre-training will shift the distribution to the potentiated side, where there are more synapses available for depression and resources are more plentiful, for both the mutant and wild type (see \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_eq}, red traces).
-This means that the pre-trained animal will learn faster then the untrained one for both mutant and wild-type (see \autoref{fig:pooled_plenty}\ref{fig:pooled_plenty_learnS} or \autoref{fig:pooled_scarce}\ref{fig:pooled_scarce_learnS}).
+Gain-decrease pre-training will shift the distribution to the potentiated side, where there are more synapses available for depression and resources are more plentiful, for both the \KO\ mice and wild type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}, red traces).
+This means that the pre-trained animals will learn faster then the untrained one for both \KO\ mice and wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learnS}).
 This differs from what is seen experimentally, where the pre-trained wild-type learns slower than the untrained one.
 
-The situation discussed in the main text, with resource depletion for depression only and potentiating transitions  unaffected, is shown in \autoref{fig:pooled_deponly}.
-It shows the same learning patterns as the more symmetric versions.
-
-This model shares one feature of the \hyperref[sec:multistate]{multistate model}.
+This model shares one feature of the \hyperref[sec:multistate]{serial model}.
 The effect of initial saturation bias is amplified by resource depletion, allowing for impaired learning with enhanced plasticity.
@@
 
-In both of these, we see that the mutant is slower than wild-type without pre-training but faster with it.
+In both of these, we see that the \KO\ mice are slower than wild-type without pre-training but faster with it.
 This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
@@
 
-With shorter pre-training, we see that it speeds up learning in both wild-type and mutant (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the mutant.
+With shorter pre-training, we see that it speeds up learning in both wild-type and \KO\ mice (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the \KO\ mice.
 This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning (see \autoref{fig:cascade_short}\ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_WT_pre}).
@@
 We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_pr}\ref{fig:cascade_long_pr_WT_pre} that longer pre-training pushes the synapses further down the cascade, slowing down learning for the wild-type.
-This effect is weaker for the mutant, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
+This effect is weaker for the \KO\ mice, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
 
-This model shares both features of the \hyperref[sec:multistate]{multistate model}.
+This model shares both features of the \hyperref[sec:multistate]{serial model}.
 First, the effect of initial saturation bias is amplified by the exponential decay of transition probabilities away from the middle.
@@
 
-First, we see that the mutant is slower than wild-type without pre-training, as seen experimentally.
+First, we see that the \KO\ mice are slower than wild-type without pre-training, as seen experimentally.
 This seems to be because the enhanced depression results in a greater fraction of synapses begin in states of weaker synaptic weight (see \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_KO}) where the depressing transitions have lower probability (see \autoref{fig:nonuni_model}).
@@
 However, gain-decrease pre-training will push synapses towards the states of stronger synaptic weight.
-The wild-type will be pushed further that the mutant, as the latter started further to the weaker side as explained above (compare \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_WT} and \ref{fig:nonuni_eq_KO}).
-As the depressing transitions have lower probability for the strongest states (see \autoref{fig:nonuni_model}), this, in addition to the enhanced depression of the mutant, will result in the mutant learning faster than wild-type, as seen experimentally.
+The wild-type will be pushed further that the \KO\ mice, as the latter started further to the weaker side as explained above (compare \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_WT} and \ref{fig:nonuni_eq_KO}).
+As the depressing transitions have lower probability for the strongest states (see \autoref{fig:nonuni_model}), this, in addition to the enhanced depression of the \KO\ mice, will result in the \KO\ mice learning faster than wild-type, as seen experimentally.
 
@@
 Second, it will also place synapses in states where transitions have lower probability, slowing down learning.
-The second effect will be weaker in the mutant than in wild-type, as they started further to the weaker side as explained above.
-This means that, for appropriate parameter choices, gain-decrease pretraining can speed up learning in the mutant, but also slow down learning in wild-type, as seen experimentally.
+The second effect will be weaker in the \KO\ mice than in wild-type, as they started further to the weaker side as explained above.
+This means that, for appropriate parameter choices, gain-decrease pretraining can speed up learning in the \KO\ mice, but also slow down learning in wild-type, as seen experimentally.
 
-Like the \hyperref[sec:cascade]{cascade model}, this model also shares both features of the \hyperref[sec:multistate]{multistate model}.
+Like the \hyperref[sec:cascade]{cascade model}, this model also shares both features of the \hyperref[sec:multistate]{serial model}.
 First, the effect of initial saturation bias is amplified by the exponential decay of transition probabilities away from the middle.

### 2013-10-01 NIPS/SynapseNIPS.tex
brought NIPS paper down to page limit

@@
 %
-The factor of $\eq$ describes the synapses being in the steady-state distribution before the memory is encoded.
+%The factor of $\eq$ describes the synapses being in the steady-state distribution before the memory is encoded.
 The factor of $(\M\pot-\M\dep)$ comes from the encoding of the memory at $t=0$, with $\wvi$ being $\pm1$ in synapses that are potentiated/depotentiated.
 The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
-A detailed derivation can be found in the supplementary material.
+A detailed derivation of this formula can be found in the supplementary material.
 We will frequently refer to this function as the memory curve.
@@
   \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
-  \item\aligntop{\includegraphics[width=0.35\linewidth]{cascms.eps}}\label{fig:cascms}
+  \item\aligntop{\includegraphics[width=0.3\linewidth]{cascms.eps}}\label{fig:cascms}
  \end{myenuma}
@@
   \snr(t) &= \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a},
-  &\quad \text{where}&\;&
-  \CI_a &= \prn{2 f\pot f\dep} \eq (\M\pot-\M\dep) \uv^a  \vv^a \w,\\&
-  & \text{and}&&
-  \tau_a &= \frac{1}{q_a}.
+%  &\quad \text{where}&\;&
+%  \CI_a &= \prn{2 f\pot f\dep} \eq (\M\pot-\M\dep) \uv^a  \vv^a \w,\\&
+%  & \text{and}&&
+%  \tau_a &= \frac{1}{q_a}.
 \end{aligned}
@@
 %
-We can then ask the question: what are the constraints on these quantities, namely initial SNR's, $\CI_a$, and time constants, $\tau_a$, for the modes,  implied by the constraints in \eqref{eq:constr}?
+where $\;\CI_a = (2 f\pot f\dep) \eq (\M\pot-\M\dep) \uv^a  \vv^a \w\;$ and $\;\tau_a = {1}/{q_a}$.
+We can then ask the question: what are the constraints on these quantities, namely eigenmode initial SNR's, $\CI_a$, and time constants, $\tau_a$, implied by the constraints in \eqref{eq:constr}?
 We will derive some of these constraints in the next section.
@@
  \begin{myenuma}
-  \item\hp\aligntop{\includegraphics[width=0.15\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
-  \item\aligntop{\includegraphics[width=0.15\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
+  \item\hp\aligntop{\includegraphics[width=0.12\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
+  \item\aligntop{\includegraphics[width=0.12\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
   \item\hp\aligntop{\includegraphics[width=0.05\linewidth]{binary_det.svg}}\label{fig:binary_det}
@@
 To systematically analyze this entire space, we develop an important organizing principle using the theory of first passage times in the stochastic process of forgetting, described by $\frg$.
-The mean first passage time matrix, $\fptb_{ij}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $i$.
+The mean first passage time matrix, $\smash{\fptb_{ij}}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $i$.
 The diagonal elements are defined to be zero.
@@
 These can be thought of as the average time it takes to reach the strong/weak states respectively.
-Using these definitions, we can then impose an order on the states by arranging them in order of decreasing $\eta^+_i$ or increasing $\eta^-_i$.
-Because $\eta^+_i+\eta^-_i=\eta$ is independent of $i$, the two orderings are the same.
+Using these definitions, we can then impose an order on the states by arranging them in order of decreasing $\smash{\eta^+_i}$ or increasing $\smash{\eta^-_i}$.
+Because $\smash{\eta^+_i+\eta^-_i=\eta}$ is independent of $i$, the two orderings are the same.
 In this order, which depends sensitively on the structure of $\M\potdep$, states later (to the right in figures below) can be considered to be more potentiated than states earlier (to the left in figures below), despite the fact that they have the same synaptic efficacy.
@@
 For example, in \autoref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
-$\eta^+_i > \eta^+_j$), we have proven that increasing $\M\pot_{ij}$ and decreasing $\M\dep_{ij}$ leads to an increase in area.
+$\smash{\eta^+_i > \eta^+_j}$), we have proven that increasing $\smash{\M\pot_{ij}}$ and decreasing $\smash{\M\dep_{ij}}$ leads to an increase in area.
 The only thing that can prevent these perturbations from increasing the area is when they require the decrease of a matrix element that has already been set to 0.
@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
-  \item\aligntop{\includegraphics[width=0.1\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
-  \item\aligntop{\includegraphics[width=0.2\linewidth]{shortcut.svg}}\label{fig:shortcut}
+  \item\aligntop{\includegraphics[width=0.07\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
+  \item\aligntop{\includegraphics[width=0.07\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
+  \item\aligntop{\includegraphics[width=0.14\linewidth]{shortcut.svg}}\label{fig:shortcut}
  \end{myenuma}
@@
 %
-where we chose $\w_k=\sgn\brk{k - \sum_j j\eq_j}$.
+where we chose $\w_k=\sgn[{k - \sum_j j\eq_j}]$.
 We can then maximize this by pushing all of the equilibrium distribution symmetrically to the two end states.
@@
 
-Now we can pick some fixed time, $t_0$, and maximize the SNR at that time \wrt the parameters $\brc{\CI_a,\tau_a}$, subject to the constraints above.
-This always results in a single nonzero $\CI_a$; in essence, optimizing memory at a single time requires a single exponential.
+Now we can pick some fixed time, $\smash{t_0}$, and maximize the SNR at that time \wrt the parameters $\smash{\brc{\CI_a,\tau_a}}$, subject to the constraints above.
+This always results in a single nonzero $\smash{\CI_a}$; in essence, optimizing memory at a single time requires a single exponential.
 The resulting optimal memory curve, along with the achieved memory at the chosen time, depends on $t_0$ as follows:
@@
 At late times, only the area bound is saturated.
-The function $\snr(t_0)$, the green curve in \autoref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim t_0\inv$.
+The function $\snr(t_0)$, the green curve in \autoref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim \smash{t_0\inv}$.
 No synaptic model can have an SNR that is greater than this at any time.
@@
  \begin{myenuma}
- \parbox[t]{0.45\linewidth}{
+ \parbox[t]{0.44\linewidth}{
   \item\aligntop{\includegraphics[width=0.95\linewidth]{env.svg}}\label{fig:envplot}
@@
 In doing so, we have obtained several new mathematical results delineating the functional limits of memory achievable by synaptic complexity, and the structural characterization of synaptic dynamical systems that achieve these limits.
-In particular, operating within the ideal observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}, we have shown that for a population of $N$ synapses with $M$ internal states, (a) the initial SNR of any synaptic model cannot exceed $\sqrt{N}$, and any model that achieves this bound is equivalent to a binary synapse, (b) the area under the memory curve of any model cannot exceed that of a linear chain model with the same equilibrium distribution, (c) both the area and memory lifetime of any model cannot exceed $\CO(\sqrt{N}M)$, and the model that achieves this limit has a linear chain topology, (d) we have derived an envelope memory curve in the SNR-time plane that cannot be exceeded by the memory curve of any model, and models that approach this envelope for times greater than $\CO(\frac{M}{r})$ are linear chain models, and (e) this late-time envelope is a power-law proportional to $O({\sqrt{N} M}/{rt})$, indicating that synaptic complexity can strongly enhance the limits of achievable memory.
+In particular, operating within the ideal observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}, we have shown that for a population of $N$ synapses with $M$ internal states, (a) the initial SNR of any synaptic model cannot exceed $\sqrt{N}$, and any model that achieves this bound is equivalent to a binary synapse, (b) the area under the memory curve of any model cannot exceed that of a linear chain model with the same equilibrium distribution, (c) both the area and memory lifetime of any model cannot exceed $\CO(\sqrt{N}M)$, and the model that achieves this limit has a linear chain topology with only nearest neighbor transitions, (d) we have derived an envelope memory curve in the SNR-time plane that cannot be exceeded by the memory curve of any model, and models that approach this envelope for times greater than $\CO(\frac{M}{r})$ are linear chain models, and (e) this late-time envelope is a power-law proportional to $\smash{\CO({\sqrt{N} M}/{rt})}$, indicating that synaptic complexity can strongly enhance the limits of achievable memory.
 
@@
 
-Another important consideration is the level of noise tolerance of the synapse.
-Note that the stochasticity of the models is an expression of biological noise.
-However, the synapses should not be allowed to optimize all of this noise away.
-It would be interesting to consider such limits on the noise levels.
-
-Finally, it would be interesting to design physiological experiments in order to perform optimal systems identification of potential Markovian dynamical systems hiding within biological synapses.
-If we measured pre and post-synaptic spike trains, and also recorded changes in post-synaptic potentials to measure changes in synaptic weight, we could use hidden Markov model techniques to find the best-fit synaptic model.
+%Another important consideration is the tolerance of the synapses to biological noise, which could be investigated by imposing lower limits on the level of stochasticity of the models.
+Finally, it would be interesting to design physiological experiments in order to perform optimal systems identification of potential Markovian dynamical systems hiding within biological synapses,
+given measurements of pre and post-synaptic spike trains along with changes in post-synaptic potentials.
 Then given our theory, we could match this measured synaptic model to optimal models to infer which timescales the synapse operates on.

### 2013-10-01 NIPS/SynapseNIPSsupp.tex
brought NIPS paper down to page limit

@@
 %
-In Surya's notes, $\MM=r\frg$.
 
@@
 %
+The factor of $\eq$ describes the synapses being in the steady-state distribution before the memory is encoded.
+The factor of $(\M\pot-\M\dep)$ comes from the encoding of the memory at $t=0$, with $\wvi$ being $\pm1$ in synapses that are potentiated/depotentiated.
+The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
 

### 2013-10-17 NIPS/SynapseNIPSsupp.tex
typos

@@
 \author{Subhaneil Lahiri and Surya Ganguli\\
-Applied Physics Department, Stanford University, Stanford CA\\
+Department of Applied Physics, Stanford University, Stanford CA\\
 \emaillink{sulahiri@stanford.edu}, \emaillink{sulahiri@stanford.edu}
@@
 %
-where $\pib$ is a row vector, $\CM$ is a matrix and $\mathbf{x}$ is a column vector.
+where $\pib$ is a row vector, $\M$ is a matrix and $\mathbf{x}$ is a column vector.
 

### 2013-11-06 NIPS/SynapseNIPS.tex
typos
final NIPS paper

@@
 \author{Subhaneil Lahiri and Surya Ganguli\\
-Applied Physics Department, Stanford University, Stanford CA\\
+Department of Applied Physics, Stanford University, Stanford CA\\
 \emaillink{sulahiri@stanford.edu}, \emaillink{sganguli@stanford.edu}
@@
 
-Theoretical studies of complex, metaplastic synapses, have focused on analyzing the memory performance of a limited number of very specific, molecular dynamical systems, characterized by a number of internal states in which potentiation and depression each induce a specific set of allowable transitions between states (\eg see \autoref{fig:models} below).
+Theoretical studies of complex, metaplastic synapses have focused on analyzing the memory performance of a limited number of very specific molecular dynamical systems, characterized by a number of internal states in which potentiation and depression each induce a specific set of allowable transitions between states (\eg see \autoref{fig:models} below).
 While these models can vastly outperform simple binary synaptic switches, these analyses leave open several deep and important questions.
@@
 
-A simple mean memory curve can be derived as follows.
+A simple average memory curve can be derived as follows.
 All of the preceding plasticity events, prior to $t=0$, will put the population of synapses in its steady-state distribution, $\eq$.
-The memory we are tracking at $t=0$ will change the internal state distribution to $\eq\M\pot$ in those synapses that are potentiated and $\eq\M\dep$ in those synapses that are depressed.
+The memory we are tracking at $t=0$ will change the internal state distribution to $\eq\M\pot$ (or $\eq\M\dep$) in those synapses that experience a candidate potentiation (or depression) event.
 As the potentiating/depressing nature of the subsequent memories is independent of $\wvi$, we can average over all sequences, resulting in the evolution of the probability distribution:
@@
 The serial model performs relatively well at one particular timescale, but it performs poorly at other times.
-The cascade model does not perform quite as well at that time, but it maintains it performance over a wider range of timescales.
+The cascade model does not perform quite as well at that time, but it maintains its performance over a wider range of timescales.
 
@@
 %
-over \emph{all} possible models and also find the model that saturates this bound.
+over \emph{all} possible models and also find the class of models that saturate this bound.
 A useful quantity is the equilibrium probability flux between two disjoint sets of states, $\CA$ and $\CB$:
@@
 %
-We will find an upper bound on this quantity as well as the model that saturates it.
+We will find an upper bound on this quantity as well as the model that saturates this bound.
 
@@
 with potentiation moving one step to the right and depression moving one step to the left.
-Any other topology would allow some perturbation to further increase the area.
+Any other topology would allow some class of perturbations (\eg\ in \autoref{fig:perts}) to further increase the area.
 
@@
 We can then maximize this by pushing all of the equilibrium distribution symmetrically to the two end states.
-This can be done by reducing the transition probabilities out of these states, as in \autoref{fig:env}(\ref{fig:sticky}).%\autoref{fig:max_area}.
-This makes it very difficult to get out of the end states once they have been entered.
+This can be done by reducing the transition probabilities out of these states, as in \autoref{fig:env}(\ref{fig:sticky}). %\autoref{fig:max_area}.
+This makes it very difficult to exit these states once they have been entered.
 The resulting area is
@@
 This leaves the question of whether this bound is achievable.
-At any time, can we find a model whose memory curve touched the envelope?
+At any time, can we find a model whose memory curve touches the envelope?
 The red curves in \autoref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
@@
 given measurements of pre and post-synaptic spike trains along with changes in post-synaptic potentials.
-Then given our theory, we could match this measured synaptic model to optimal models to infer which timescales the synapse operates on.
+Then given our theory, we could match this measured synaptic model to optimal models to understand for which timescales of memory, if any, biological synaptic dynamics may be tuned.
 

### 2013-11-06 NIPS/SynapseNIPSsupp.tex
typo in nips supp
final NIPS paper

@@
 
-Although the theory and derivations outlined in the main text are relatively self-contained, here we provide more details for the aid of the reviewer.
+Here we provide more details underlying the derivations of results in the main paper.
 
@@
 
-In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time to define notation and .
-It is a straightforward generalisation of material that can be found in \cite{kemeny1960finite} with some ideas from \cite{hunter2000survey}.
+In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time to define notation.
+It is a generalization of material that can be found in \cite{kemeny1960finite} with some ideas from \cite{hunter2000survey}.
 
@@
 
-A Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$.
+A continuous time Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$ with row sums equal to zero ($\MM\onev=0$).
 The probabilities of being in each state at time $t$, the row-vector $\mathbf{p}(t)$, evolve according to
@@
   \diff{\mathbf{p}(t)}{t} = \mathbf{p}(t) \MM,
-  \qquad
-  \mathbf{p}(t)\onev=1,
-  \qquad
-  \MM\onev=0.
-\end{equation}
-%
+\end{equation}
+%
+where $\mathbf{p}(t)\onev=1$.
 
@@
 \subsection{Fundamental matrix}\label{sec:fund}
+
+For our results below regarding the integral of the memory curve, it can be useful to invert the stochastic transition matrix, $\MM$.
+However, since $\MM$ has a zero eigenvalue, it cannot be inverted.
+For this reason, the fundamental matrix arises as a useful surrogate for the inverse of $\MM$.
+It is related to the first passage times, as we will see in the \hyperref[sec:fpt]{next subsection}.
+Here we define the fundamental matrix and review its properties.
 
@@
     &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} (\fund_{jk} - \fund_{ik}) \\
-    &= \sum_{i\neq j} \diff{\MM_{ij}}{\alpha} \eq_i \eq_k (\fptb_{ik} - \fptb_{jk}),
-\end{aligned}
-\end{equation}
-%
-which is the result of \cite{cho2000markov} that we need.
+    &= \sum_{i\neq j} \diff{\MM_{ij}}{\alpha} \eq_i \eq_k (\fptb_{ik} - \fptb_{jk}).
+\end{aligned}
+\end{equation}
+%
+This is a generalization of a result of \cite{cho2000markov} from discrete to continuous time that we will need below.
 Note that the summand vanishes for $i=j$, so we can drop the restriction $i\neq j$ from the range of the sum.
@@
 %
+
 \section{Signal-to-Noise ratio (SNR)}\label{sec:SNR}
@@
 In this section we will look at the signal-to-noise curve, and put an upper bound on its initial value.
-We will only consider ergodic Markov chains.
+We need only consider ergodic Markov chains.
 Transient states would be unoccupied in equilibrium and would not be accessed by the signal creation process, therefore they could be removed from the analysis.
@@
 %
-We only require that this Markov chain is ergodic. The Markov chains described by $\M\potdep $ need not be.
+We only require that this Markov chain is ergodic.
+The Markov chains described by $\M\potdep $ need not be.
 
@@
 
-Which results in our final formula:
+This results in our final formula:
 %
@@
 As in \sref{sec:SNR}, we will only consider ergodic Markov chains.
-We will see in \sref{sec:multistate} that the optimal chain is absorbing, so it lies on the boundary of the (open) set of ergodic chains, but it still puts an upper bound on the area.
+We will see in \sref{sec:multistate} that the optimal chain is absorbing, so it lies on the boundary of the (open) set of ergodic chains, but it still puts an upper bound on the area for any chain in the interior.
 
@@
       = \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
-      = - \prn{\eq q \fptb}_k
-      = \frac{(\eq q \fund)_k}{\eq_k}, \\
+      = - \prn{\eq \enc \fptb}_k
+      = \frac{(\eq \enc \fund)_k}{\eq_k}, \\
     a_i &= \sum_j \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
@@
 \subsection{Derivatives \wrt \texorpdfstring{$\W\potdep $}{W(pot/dep)}}\label{sec:deriv}
+
+In the following, we will mathematically define the classes of perturbations pictorially described in Figure 3 of the main paper.
+In order to do so, we will need to consider expressions for derivatives of various quantities with respect to $\W\potdep_{ij}$.
 
@@
 
+Here we describe the perturbations corresponding to Figure 3(a,b) of the main paper.
+
 Consider
@@
 
-\subsubsection{Increasing \texorpdfstring{$c_k$}{c(k)}}\label{sec:areacoeffincr}
-
-Consider the following combinations of derivatives:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:areacoeffincrderiv}
-    \Delta_{gh} &\equiv
-      \frac{1}{\eq_{g}} \pdiffc{}{\frg_{gh}}
-      + \frac{1}{\eq_h} \pdiffc{}{\frg_{hg}}, \\
-%  \end{aligned}
-\end{align}
-%
-Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).
-
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators using \eqref{eq:diffpT}:
-%
-\begin{equation}\label{eq:sareacoeffincrprob}
-  \Delta_{gh} \eq_i = 0,
-\end{equation}
-%
-which makes it possible to integrate the perturbation:
-%
-\begin{equation}\label{eq:areacoeffincrfinite}
-  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon},
-  \qquad
-  \begin{aligned}
-  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
-  \\
-  \boldsymbol{\epsilon} \onev &= 0.
-  \end{aligned}
-\end{equation}
-%
-But more interestingly:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:areacoeffincr}
-    \Delta_{gh}\CL &=
-      \frac{\mu\pot _{gh}+\mu\dep _{gh}}{\eq_g} + \frac{\mu\pot _{hg}+\mu\dep _{hg}}{\eq_h}
-      + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
-%  \end{aligned}
-\end{align}
-%
-where $c_k$ were defined in \eqref{eq:areacoeffs}.
-
-Using the non-negativity of the Kuhn-Tucker multipliers, $\mu\potdep _{ij}$, \eqref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
-
-Note that, according to \sref{sec:triangular}, either $\W\pot _{gh}$ or $\W\dep _{gh}$ will be zero at the maximum, therefore we can expect one of $\mu\pot _{gh}+\mu\dep _{gh}$ to be non-zero.
-This would rule out degeneracy of the $c_k$ or $\eta^+_k$.
-Looking at \eqref{eq:shiftqderiv} closely, the only way $\mu\pot _{gh}+\mu\dep _{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
-
 \subsubsection{Shortcuts}\label{sec:shortcuts}
+
+In this subsection we will define perturbations corresponding to Figure 3(c) of the main text.
 
@@
 
-If we put the states in order of decreasing $\eta^+_k$, the results of the \sref{sec:areacoeffincr} tell us that the $c_k$ are non-decreasing.
+In the \hyperref[sec:areacoeffincr]{section below}, we will show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
 This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
@@
 This would tell us that the process with the maximal area has to have a multi-state topology.
+
+\subsubsection{Increasing \texorpdfstring{$c_k$}{c(k)}}\label{sec:areacoeffincr}
+
+In the \hyperref[sec:shortcuts]{previous subsection} we defined perturbations corresponding to Figure 3(c) of the main text.
+In order to show that those perturbations increase the area, we must now show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
+
+Consider the following combinations of derivatives:
+%
+\begin{align}
+%  \begin{aligned}
+\label{eq:areacoeffincrderiv}
+    \Delta_{gh} &\equiv
+      \frac{1}{\eq_{g}} \pdiffc{}{\frg_{gh}}
+      + \frac{1}{\eq_h} \pdiffc{}{\frg_{hg}}, \\
+%  \end{aligned}
+\end{align}
+%
+Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).
+
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators using \eqref{eq:diffpT}:
+%
+\begin{equation}\label{eq:sareacoeffincrprob}
+  \Delta_{gh} \eq_i = 0,
+\end{equation}
+%
+which makes it possible to integrate the perturbation:
+%
+\begin{equation}\label{eq:areacoeffincrfinite}
+  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon},
+  \qquad
+  \begin{aligned}
+  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
+  \\
+  \boldsymbol{\epsilon} \onev &= 0.
+  \end{aligned}
+\end{equation}
+%
+But more interestingly:
+%
+\begin{align}
+%  \begin{aligned}
+\label{eq:areacoeffincr}
+    \Delta_{gh}\CL &=
+      \frac{\mu\pot _{gh}+\mu\dep _{gh}}{\eq_g} + \frac{\mu\pot _{hg}+\mu\dep _{hg}}{\eq_h}
+      + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
+%  \end{aligned}
+\end{align}
+%
+where $c_k$ were defined in \eqref{eq:areacoeffs}.
+
+Using the non-negativity of the Kuhn-Tucker multipliers, $\mu\potdep _{ij}$, \eqref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
+
+Note that, according to \sref{sec:triangular}, either $\W\pot _{gh}$ or $\W\dep _{gh}$ will be zero at the maximum, therefore we can expect one of $\mu\pot _{gh}+\mu\dep _{gh}$ to be non-zero.
+This would rule out degeneracy of the $c_k$ or $\eta^+_k$.
+Looking at \eqref{eq:shiftqderiv} closely, the only way $\mu\pot _{gh}+\mu\dep _{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
 
@@
 
+The previous results indicate that the area under the memory curve of any model is bounded by the area under the memory curve of a model with the serial/multistate topology having the same equilibrium distribution.
+Here we compute this area, which we will see depends only on this equilibrium distribution.
+
 The multi-state/serial topology is defined by (see \cite{amit1994learning,Fusi2007multistate,Leibold2008serial}):
@@
 %
-It saturates various inequalities:
+Because it has no shortcuts, it saturates various inequalities:
 %
@@
 where we used $\w_k=\sgn(c_k)$, as discussed after \eqref{eq:areacoeffs}.
+This reproduces equation (15) of the main paper.
+
 %It will also help to define $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
@@
 
+In order to obtain an upper bound on the area under the memory curve of any model, we now maximise the area of the serial model with respect to its equilibrium distribution.
 First let us maximise \eqref{eq:multistatearea} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
@@
 %
+This yields the area bound of equation (16) of the main text.
+
 Note that the chain that achieves this is not ergodic, the two states at each end are absorbing.

### 2013-11-07 NIPS/SynapseNIPS.tex
final NIPS upload

@@
 
-In this section, we will describe the class of models of synaptic plasticity that we are studying and how we quantify their memory performance.
+In this section, we describe the class of models of synaptic plasticity that we are studying and how we quantify their memory performance.
 In the subsequent sections, we will find upper bounds on this performance.
@@
 
-For any single memory, stored at time $t=0$, we assume there will be an ideal pattern of synaptic weights across a population of $N$ synapses, the $N$-element vector $\wvi$, that is $+1$ at all synapses that experience a candidate potentiation event, and $-1$ at all synapses that experienced a candidate depression event at the time of memory storage.
-The actual pattern of synaptic weights at some later time, $t$, will change to $\wv(t)$ due to further modifications from subsequent memories.
+For any single memory, stored at time $t=0$, we assume there will be an ideal pattern of synaptic weights across a population of $N$ synapses, the $N$-element vector $\wvi$, that is $+1$ at all synapses that experience a candidate potentiation event, and $-1$ at all synapses that experience a candidate depression event at the time of memory storage.
+We assume that any pattern of synaptic weights close to $\wvi$ is sufficient to recall the memory.
+However, the actual pattern of synaptic weights at some later time, $t$, will change to $\wv(t)$ due to further modifications from the storage of subsequent memories.
 We can use the overlap between these, $\wvi\cdt\wv(t)$, as a measure of the quality of the memory.
@@
 \begin{equation}\label{eq:kemenyconst}
-  \eta \equiv \sum_j \fptb_{ij}\eq_j
+  \eta \equiv \sum_j \fptb_{ij}\eq_j,
 \end{equation}
 %
-is independent of the starting state $i$.
-It is known as Kemeny's constant (see \cite[\S4.4]{kemeny1960finite}).
+known as Kemeny's constant (see \cite[\S4.4]{kemeny1960finite}), is independent of the starting state $i$.
 Intuitively, \eqref{eq:kemenyconst} states that the average time it takes to reach any state, weighted by its equilibrium probability, is independent of the starting state, implying a hidden constancy inherent in any stochastic process.

### 2013-11-07 Notes/ContTime.tex
typos
Laplace transform of SNR

@@
 
-We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f^+f^-)\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
+We can rewrite this using \eqref{eq:diffpT}, with $rA=\sqrt{N}(2f^+f^-)\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
 %
@@
 We're better than that.
-We can choose $q_1=0$, $\uv^1=\onev$ and $\vv=\eq$.
+We can choose $q_1=0$, $\uv^1=\onev$ and $\vv^1=\eq$.
 
@@
 
+\section{Laplace transform}\label{sec:laplace}
+
+\subsection{Fundamental matrix \etc}\label{sec:lfund}
+
+In analogy to \eqref{eq:funddef}, define
+%
+\begin{equation}\label{eq:lfund}
+  \fund(s) = (s\I +\onev\pib -\MM)\inv.
+\end{equation}
+%
+This reduces to the fundamental matrix at $s=0$.
+The analogous result to \eqref{eq:fundrowsum} is
+%
+\begin{equation}\label{eq:lfundrowsum}
+  \fund(s)\onev = \frac{\tau}{1+s\tau}\onev.
+\end{equation}
+%
+The analogues of \eqref{eq:fundprob}, \eqref{eq:fundQZ} and \eqref{eq:fundZQ} are much uglier, but luckily we won't need them.
+Supose we have some row vector $\mathbf{a}$ such that $\mathbf{a}\onev=0$.
+Then $\mathbf{a}\fund(s)$ is independent of $\pib$:
+%
+\begin{equation}\label{eq:alfund}
+  \pdiff{(\mathbf{a}\fund(s))_i}{\pib_k} = (\mathbf{a}\fund(s)\onev) \fund_{ki}(s) = \frac{\tau}{1+s\tau} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
+\end{equation}
+%
+
+Then we also define
+%
+\begin{equation}\label{eq:lfptb}
+  \fptb(s) = (\onem\fund\dg(s)-\fund(s))\D,
+  \qquad
+  \fptb_{ij}(s) = \frac{\fund_{jj}(s)-\fund_{ij}(s)}{\eq_j}.
+\end{equation}
+%
+This reduces to the first passage time matrix at $s=0$, but at other $s$ there is (probably) no relation.
+
+Using the exact same proof as \autoref{th:kemenyconst}, we see that the quantity
+%
+\begin{equation}\label{eq:lkemeny}
+  \eta(s) = \sum_j \fptb_{ij}(s)\eq_j
+\end{equation}
+%
+is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau}{1+s\tau}$.
+We can the define 
+%
+\begin{equation}\label{eq:lpmixtime}
+  \eta^{\pm}_i(s) = \sum_j \fptb_{ij}(s)\eq_j \prn{\frac{1\pm\w_j}{2}} =  \sum_{j\in\pm} \fptb_{ij}(s)\eq_j.
+\end{equation}
+%
+We can arrange the states in order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
+
+\subsection{Laplace transform of SNR curve}\label{sec:laplaceSNR}
+
+Consider the Laplace transform of the evolution operator:
+%
+\begin{equation}\label{eq:levol}
+  \mathbf{G}(s) = \int_0^{\infty}\dr t\, \e^{(\MM-s) t}.
+\end{equation}
+%
+For $\Re s>0$, we have 
+%
+\begin{equation}\label{eq:levolcalc}
+  (s-\MM)\mathbf{G}(s) = \int_0^{\infty}\dr t\, (s-\MM)\e^{(\MM-s) t} = \brk{-\e^{(\MM-s) t}}_0^{\infty} = \I.
+\end{equation}
+%
+As $(s-\MM)$ is invertible for $\Re s>0$, because all eigenvalues of $\MM$ are nonpositive, we have 
+%
+\begin{equation}\label{eq:levolres}
+  \mathbf{G}(s) = (s-\MM)\inv.
+\end{equation}
+%
+For $s=0$, we can avoid our problems by replacing $\mathbf{G}(s) \to \fund(s)$.
+
+Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurve}
+%
+\begin{equation}\label{eq:lsnr}
+\begin{aligned}
+  A(s) &= \int_0^{\infty}\dr t \e^{-st}\SNR(t) \\
+   &= \int_0^{\infty}\dr t\, \sqrt{N}(2f^+f^-) \eq \enc \e^{(r\frg-s)t} \w \\
+   &= \frac{\sqrt{N}(2f^+f^-)}{r} \eq \enc \prn{s - r\frg}\inv \w.
+\end{aligned}
+\end{equation}
+%
+Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.
+
+Thanks to \eqref{eq:alfund}, we can replace $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
+%
+\begin{equation}\label{eq:larea}
+\begin{aligned}
+  \hat{A}(s) = \frac{rA(s)}{\sqrt{N}(2f^+f^-)} &=  \eq \enc \fund(s) \w \\
+    &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
+    &= 2\sum_{ij}  \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)).
+\end{aligned}
+\end{equation}
+%
+In analogy with \eqref{eq:areacoeffs}, we define
+%
+\begin{equation}\label{eq:lareacoeffs}
+  \begin{aligned}
+    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
+    a_i(s) &= \sum_j \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)),\\
+    \implies
+    \hat{A}(s) &= \sum_k c_k(s) \eq_k \w_k
+      = 2\sum_i a_i(s).
+  \end{aligned}
+\end{equation}
+%
+
+\subsection{Perturbation analysis}\label{sec:lpert}
+
+Using the same approach as the area bound, we find that the derivatives are
+%
+\begin{equation}\label{eq:lAdiff}
+  \pdiff{\hat{A}(s)}{\W^\pm_{gh}} = 2rf^\pm \eq_g \brk{ \sum_i a_i(s) (\fptb_{gi}(0)-\fptb_{hi}(0))
+     +c_g(s) (\eta^+_g(s) - \eta^+_h(s)) } \pm 2\eq_g (\eta^+_g(s) - \eta^+_h(s)).
+\end{equation}
+%
+This is exactly the same as \eqref{eq:derivA}.
+It follows that, using exactly the same perturbations, that the Laplace transform of the SNR curve for \emph{any} model is lower than that of some model with the serial topology having the same equilibrium distribution.
+
 \appendix
@@
 %
-\begin{equation}\label{derivTarea}
+\begin{equation}\label{eq:derivTarea}
   \pdiff{\hat{A}}{\fptb_{gh}} = -\eq_h\brk{

### 2013-11-08 Notes/ContTime.tex
laplacian envelope written up

@@
 \DeclareMathOperator{\SNR}{SNR}
+\DeclareMathOperator{\snr}{SNR}
+\newcommand{\CI}{\mathcal{I}}
 \newcommand{\CS}{\mathcal{S}}
@@
 %
-Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.
-
-Thanks to \eqref{eq:alfund}, we can replace $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
+This expression is ill-behaved at $s=0$.
+Thanks to \eqref{eq:alfund}, we can solve this by the replacement $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
 %
@@
 %
+Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.
+
 In analogy with \eqref{eq:areacoeffs}, we define
@@
 \begin{equation}\label{eq:lAdiff}
-  \pdiff{\hat{A}(s)}{\W^\pm_{gh}} = 2rf^\pm \eq_g \brk{ \sum_i a_i(s) (\fptb_{gi}(0)-\fptb_{hi}(0))
-     +c_g(s) (\eta^+_g(s) - \eta^+_h(s)) } \pm 2\eq_g (\eta^+_g(s) - \eta^+_h(s)).
+\begin{aligned}
+  \pdiff{\hat{A}(s)}{\W^\pm_{gh}} 
+     =&\, 2rf^\pm \eq_g \brk{ \sum_i a_i(s) (\fptb_{gi}(0)-\fptb_{hi}(0))
+     +c_g(s) (\eta^+_g(s) - \eta^+_h(s)) } \\
+     &\pm 2\eq_g (\eta^+_g(s) - \eta^+_h(s)).
+\end{aligned}
 \end{equation}
@@
 This is exactly the same as \eqref{eq:derivA}.
-It follows that, using exactly the same perturbations, that the Laplace transform of the SNR curve for \emph{any} model is lower than that of some model with the serial topology having the same equilibrium distribution.
+It follows that, using exactly the same perturbations, that the Laplace transform of the SNR curve for \emph{any} model is lower than that of a certain model with the serial topology having the same equilibrium distribution.
+
+\subsection{Envelope}\label{sec:env}
+
+Hopefully we can prove an upper bound on the Laplace transform of the memory curve:
+%
+\begin{equation}\label{eq:lbound}
+  A(s) \leq \sqrt{N}B(s) = \frac{\sqrt{N}}{C(s)}.
+\end{equation}
+%
+We will assume that $C(s)$ is a concave function.
+We already know that $C(0)=\frac{1}{M-1}$ and $C(s) \to s$ as $s\to\infty$.
+
+In the eigenmode decomposition, this reads
+%
+\begin{equation}\label{eq:lboundeig}
+  \sum_a \frac{\CI_a}{q_a+s} \leq B(s).
+\end{equation}
+%
+Now let's maximise $\snr(t_0)$ with respect to $\brc{\CI_a,q_a}$.
+Consider the Lagrangian
+%
+\begin{equation}\label{eq:envlagrange}
+  \CL = \sum_a \CI_a \e^{-q_a t_0} +\intd[_0^\infty]{s} \mu(s)\brk{B(s)-\sum_a\frac{\CI_a}{q_a+s}}.
+\end{equation}
+%
+Necessary conditions for a maximum are
+%
+\begin{alignat}{3}
+  \pdiff{\CL}{\CI_a} &= \e^{-q_a t_0} - \intd{s} \frac{\mu(s)}{q_a+s} = 0, \label{eq:envI}\\
+  \pdiff{\CL}{q_a} &= \CI_a\brk{-t_0\e^{-q_a t_0} + \intd{s} \frac{\mu(s)}{(q_a+s)^2}} = 0, \label{eq:envq}\\
+  \mu(s)\fdf{\CL}{\mu(s)} &= \mu(s)\brk{B(s)-\sum_a\frac{\CI_a}{q_a+s}} = 0, \label{eq:envKT}\\
+  \mu(s) &\geq 0, \label{eq:envKTm}\\
+  \fdf{\CL}{\mu(s)} &= B(s)-\sum_a\frac{\CI_a}{q_a+s} \geq 0 \label{eq:envKTbnd}.
+\end{alignat}
+%
+If we assume $\CI_a \neq0$ for some $a$, then \eqref{eq:envI} and \eqref{eq:envq} read
+%
+\begin{equation}\label{eq:envnumexp}
+\begin{aligned}
+  \e^{-q_a t_0} &= \intd{s} \frac{\mu(s)}{q_a+s}, \\
+  t_0\e^{-q_a t_0} &= \intd{s} \frac{\mu(s)}{(q_a+s)^2}.
+\end{aligned}
+\end{equation}
+%
+The second of these is the derivative of the first with respect to $q$, therefore the two curves in the first must touch each other tangentially at $q_a$.
+As both curves are decreasing convex functions, it seems unlikely that this could happen at multiple values of $q$.
+We will assume that there is a single non-zero $\CI_a=\CI$, with the corresponding $q_a=q$, in the following.
+
+For \eqref{eq:envI} to have a solution, $\mu(s)$ must be nonzero for at least one value of $s$.
+Then, by \eqref{eq:envKT}, we must saturate \eqref{eq:envKTbnd} at those values of $s$.
+If there were more than two such values, $\CI$ and $q$ would be overconstrained.
+Assume there are two such values, $s=s_0$ and $s=s_1$.
+%
+\begin{equation}\label{eq:sattwice}
+\begin{aligned}
+  \CI &= B(s_0)(q+s_0) = B(s_1)(q+s_1) \qquad
+  \implies q &= \frac{s_0 B(s_0) - s_1 B(s_1)}{B(s_1)-B(s_0)}.
+\end{aligned}
+\end{equation}
+%
+Now the bound \eqref{eq:envKTbnd} reads
+%
+\begin{equation}\label{eq:satwicebnd}
+\begin{aligned}
+  B(s) \geq \frac{B(s_0)(q+s_0)}{q+s} &= \frac{s_0-s_1}{(s-s_1)C(s_0)+(s_0-s)C(s_1)}, \\
+  xC(s_0)+(1-x)C(s_1) &\geq C(xs_0+(1-x)s_1),
+\end{aligned}
+\end{equation}
+%
+where we substituted $s=xs_0+(1-x)s_1$.
+As we assumed that $C(s)$ is concave, this is impossible.
+
+Therefore the bound must be saturated exactly once and $\mu(s)=\mu_0\delta(s-s_0)$.
+Now \eqref{eq:envI}, \eqref{eq:envq} and \eqref{eq:envKT} read
+%
+\begin{equation}\label{eq:satonce}
+  \mu_0 = (q+s_0)\e^{-qt_0},
+  \qquad
+  (q+s_0)t_0 = 1,
+  \qquad
+  \CI = \frac{B(s_0)}{t_0}.
+\end{equation}
+%
+This satisfies \eqref{eq:envKTm}.
+Now \eqref{eq:envKTbnd} tells us
+%
+\begin{equation}\label{eq:satoncebnd}
+\begin{aligned}
+  B(s) &\geq \frac{B(s_0)}{(q+s)t_0}, 
+  \qquad\text{or}\qquad
+  1 + (s-s_0)t_0 &\geq \frac{C(s)}{C(s_0)}.
+\end{aligned}
+\end{equation}
+%
+As the two sides are equal at $s=s_0$, it is necessary that they touch tangentially for this bound to be true elsewhere.
+The linearity of the LHS and concavity of the RHS ensure that this is sufficient for the bound to be satisfied for all $s$.
+This condition is
+%
+\begin{equation}\label{eq:envs0t0}
+  t_0 = \frac{C'(s_0)}{C(s_0)}.
+\end{equation}
+%
+
+The resulting SNR curve and envelope are
+%
+\begin{equation}\label{eq:env}
+  \snr(t) = \frac{\e^{(s_0-1/t_0)t}}{C'(s_0)},
+  \qquad
+  \snr(t_0) = \frac{\e^{s_0t_0-1}}{C'(s_0)},
+\end{equation}
+%
+where $s_0$ is determined as a function of $t_0$ by \eqref{eq:envs0t0}.
 

### 2013-11-09 Notes/ContTime.tex
cleanup of memory curve notes

@@
 \begin{aligned}
-  \hat{A} &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}) &
-    &= -2\sum_{i,j} \enc_{ij} \eq_i \eta^+_{j} \\
-    &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^-_{j} - \eta^-_{i}) &
-    &= 2\sum_{i,j} \enc_{ij} \eq_i \eta^-_{j}.
+  \hat{A} &= 2\sum_{i,j} \eq_i \enc_{ij} (\eta^+_{i} - \eta^+_{j}) &
+    &= -2\sum_{i,j} \eq_i \enc_{ij} \eta^+_{j} \\
+    &= 2\sum_{i,j} \eq_i \enc_{ij} (\eta^-_{j} - \eta^-_{i}) &
+    &= 2\sum_{i,j} \eq_i \enc_{ij} \eta^-_{j}.
 \end{aligned}
@@
 \begin{equation}\label{eq:areaZ}
-  \hat{A} = \sum_{i,j,k,l} \enc_{ij} \pib_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
-    = \pib \fund q \fund \w.
+  \hat{A} = \sum_{i,j,k,l} \pib_{l} \fund_{li} \enc_{ij} (\fund_{jk}-\fund_{ik}) \w_k
+    = \pib \fund \enc \fund \w.
 \end{equation}
@@
       = \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
-      = - \prn{\eq q \fptb}_k
-      = \frac{(\eq q \fund)_k}{\eq_k}, \\
-    a_i &= \sum_j \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
+      = - \prn{\eq \enc \fptb}_k
+      = \frac{(\eq \enc \fund)_k}{\eq_k}, \\
+    a_i &= \sum_j \eq_i \enc_{ij} (\eta^+_{i} - \eta^+_{j}),\\
     \implies
@@
 \begin{equation}\label{eq:finlagrangian}
-  \CL = \eq \enc\e^{rt\frg}\w + \sum_\pm \sum_{ij} \mu^\pm_{ij} \M^\pm_{ij}.
+  \CL = \eq \enc\, \e^{rt\frg}\w + \sum_\pm \sum_{ij} \mu^\pm_{ij} \M^\pm_{ij}.
 \end{equation}
@@
     &\pm \eq_g \prn{\brk{\e^{rt\frg} \w}_h - \brk{\e^{rt\frg} \w}_g} \\
-    &+ \sum_{ab} F_{ab} (\eq\enc\uv^a) \vv^a_g (\uv^b_h-\uv^b_g) (\vv^a\w) \\
+    &+ f^\pm \sum_{ab} F_{ab} (\eq\enc\uv^a) \vv^a_g (\uv^b_h-\uv^b_g) (\vv^a\w) \\
     &+ \mu^\pm_{gh} - \mu^\pm_{gg}.
@@
 \begin{equation}\label{eq:levol}
-  \mathbf{G}(s) = \int_0^{\infty}\dr t\, \e^{(\MM-s) t}.
+  \mathbf{G}(s) = \intd[_0^{\infty}]{t} \e^{(\MM-s) t}.
 \end{equation}
@@
 \begin{equation}\label{eq:levolcalc}
-  (s-\MM)\mathbf{G}(s) = \int_0^{\infty}\dr t\, (s-\MM)\e^{(\MM-s) t} = \brk{-\e^{(\MM-s) t}}_0^{\infty} = \I.
-\end{equation}
-%
-As $(s-\MM)$ is invertible for $\Re s>0$, because all eigenvalues of $\MM$ are nonpositive, we have 
+  (s-\MM)\mathbf{G}(s) = \intd[_0^{\infty}]{t} (s-\MM)\e^{(\MM-s) t} = \brk{-\e^{(\MM-s) t}}_0^{\infty} = \I.
+\end{equation}
+%
+As $(s-\MM)$ is invertible for $\Re s>0$, because the real part of all eigenvalues of $\MM$ are nonpositive, we have
 %
@@
 %
-For $s=0$, we can avoid our problems by replacing $\mathbf{G}(s) \to \fund(s)$.
+For $s=0$, we can avoid problems by replacing $\mathbf{G}(s) \to \fund(s)$.
 
@@
 \begin{aligned}
-  A(s) &= \int_0^{\infty}\dr t \e^{-st}\SNR(t) \\
-   &= \int_0^{\infty}\dr t\, \sqrt{N}(2f^+f^-) \eq \enc \e^{(r\frg-s)t} \w \\
+  A(s) &= \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) \\
+   &= \intd[_0^{\infty}]{t} \sqrt{N}(2f^+f^-) \eq \enc \e^{(r\frg-s)t} \w \\
    &= \frac{\sqrt{N}(2f^+f^-)}{r} \eq \enc \prn{s - r\frg}\inv \w.
@@
 We will assume that $C(s)$ is a concave function.
-We already know that $C(0)=\frac{1}{M-1}$ and $C(s) \to s$ as $s\to\infty$.
-
-In the eigenmode decomposition, this reads
+We already know that $C(0)=\frac{r}{M-1}$ and $C(s) \to s$ as $s\to\infty$.
+
+In the eigenmode decomposition \eqref{eq:eigendecomp}:
+%
+\begin{equation}\label{eq:eigenmode}
+  \snr(t) = \sqrt{N}\sum_a \CI_a\, \e^{-q_a t},
+  \qquad
+  \CI_a = (2f^+f^-) (\eq \enc \uv^a) (\vv^a \w)
+\end{equation}
+%
+the upper bound reads
 %
@@
 \begin{equation}\label{eq:envlagrange}
-  \CL = \sum_a \CI_a \e^{-q_a t_0} +\intd[_0^\infty]{s} \mu(s)\brk{B(s)-\sum_a\frac{\CI_a}{q_a+s}}.
+  \CL = \sum_a \CI_a\, \e^{-q_a t_0} +\intd[_0^\infty]{s} \mu(s)\brk{B(s)-\sum_a\frac{\CI_a}{q_a+s}}.
 \end{equation}
@@
   \pdiff{\CL}{\CI_a} &= \e^{-q_a t_0} - \intd{s} \frac{\mu(s)}{q_a+s} = 0, \label{eq:envI}\\
-  \pdiff{\CL}{q_a} &= \CI_a\brk{-t_0\e^{-q_a t_0} + \intd{s} \frac{\mu(s)}{(q_a+s)^2}} = 0, \label{eq:envq}\\
+  \pdiff{\CL}{q_a} &= \CI_a\brk{-t_0\,\e^{-q_a t_0} + \intd{s} \frac{\mu(s)}{(q_a+s)^2}} = 0, \label{eq:envq}\\
   \mu(s)\fdf{\CL}{\mu(s)} &= \mu(s)\brk{B(s)-\sum_a\frac{\CI_a}{q_a+s}} = 0, \label{eq:envKT}\\
@@
   \e^{-q_a t_0} &= \intd{s} \frac{\mu(s)}{q_a+s}, \\
-  t_0\e^{-q_a t_0} &= \intd{s} \frac{\mu(s)}{(q_a+s)^2}.
+  t_0\,\e^{-q_a t_0} &= \intd{s} \frac{\mu(s)}{(q_a+s)^2}.
 \end{aligned}
@@
 \begin{aligned}
-  \CI &= B(s_0)(q+s_0) = B(s_1)(q+s_1) \qquad
-  \implies q &= \frac{s_0 B(s_0) - s_1 B(s_1)}{B(s_1)-B(s_0)}.
+  \CI = B(s_0)(q+s_0) = B(s_1)(q+s_1) 
+  \qquad
+  \implies q = \frac{s_0 B(s_0) - s_1 B(s_1)}{B(s_1)-B(s_0)}.
 \end{aligned}
@@
 
-Therefore the bound must be saturated exactly once and $\mu(s)=\mu_0\delta(s-s_0)$.
+Therefore the bound must be saturated exactly once and $\mu(s)=\mu_0\,\delta(s-s_0)$.
 Now \eqref{eq:envI}, \eqref{eq:envq} and \eqref{eq:envKT} read
@@
 \begin{equation}\label{eq:satonce}
-  \mu_0 = (q+s_0)\e^{-qt_0},
+  \mu_0 = (q+s_0)\,\e^{-qt_0},
   \qquad
@@
 \begin{aligned}
-  B(s) &\geq \frac{B(s_0)}{(q+s)t_0}, 
+  B(s) \geq \frac{B(s_0)}{(q+s)t_0}, 
   \qquad\text{or}\qquad
-  1 + (s-s_0)t_0 &\geq \frac{C(s)}{C(s_0)}.
+  1 + (s-s_0)t_0 \geq \frac{C(s)}{C(s_0)}.
 \end{aligned}
@@
 \begin{equation}\label{eq:env}
-  \snr(t) = \frac{\e^{(s_0-1/t_0)t}}{C'(s_0)},
-  \qquad
-  \snr(t_0) = \frac{\e^{s_0t_0-1}}{C'(s_0)},
-\end{equation}
-%
-where $s_0$ is determined as a function of $t_0$ by \eqref{eq:envs0t0}.
+  \snr(t) = \frac{\sqrt{N}\,\e^{(s_0-1/t_0)t}}{C'(s_0)},
+  \qquad
+  \snr(t_0) = \frac{\sqrt{N}\,\e^{s_0t_0-1}}{C'(s_0)},
+\end{equation}
+%
+where $s_0$ is a function of $t_0$ determined by \eqref{eq:envs0t0}.
 

### 2013-11-12 Notes/ContTime.tex
bootstrap envelope

@@
 \DeclareMathOperator{\snr}{SNR}
+\DeclareMathOperator{\env}{Env}
+\newcommand{\rh}{\hat{r}}
 \newcommand{\CI}{\mathcal{I}}
@@
 \begin{equation}\label{eq:envs0t0}
-  t_0 = \frac{C'(s_0)}{C(s_0)}.
+  t_0 = \frac{C'(s_0)}{C(s_0)} = -\frac{B'(s_0)}{B(s_0)}.
 \end{equation}
@@
   \qquad
-  \snr(t_0) = \frac{\sqrt{N}\,\e^{s_0t_0-1}}{C'(s_0)},
+  \snr(t_0) = \sqrt{N}\env(t_0) = \frac{\sqrt{N}\,\e^{s_0t_0-1}}{C'(s_0)}
+   = \frac{\sqrt{N}\,B(s)\e^{s_0t_0-1}}{t_0},
 \end{equation}
@@
 where $s_0$ is a function of $t_0$ determined by \eqref{eq:envs0t0}.
+
+\subsection{Bootstrap}\label{sec:boot}
+
+Using the upper bounds on initial SNR and area, we can derive the following envelope:
+%
+\begin{equation}\label{sec:envinitarea}
+  \env_0(t_0) =
+    \begin{cases}
+    \e^{-rt_0/(M-1)},    & rt_0 \leq M-1,\\
+    \frac{M-1}{\e rt_0}, & rt_0 \geq M-1.
+    \end{cases}
+    =
+    \begin{cases}
+    \e^{-\rh t_0},        & \rh t_0 \leq 1,\\
+    \frac{1}{\e \rh t_0}, & \rh t_0 \geq 1.
+    \end{cases}
+\end{equation}
+%
+We will use the notation $\rh=r/(M-1)$ in this subsection.
+As no memory curve can cross this, the Laplace transform of any model's memory curve must be smaller than
+%
+\begin{equation}\label{eq:envlt}
+\begin{aligned}
+  \frac{A(s)}{\sqrt{N}} \leq B(s) < B_0(s) &= \intd[_0^\infty]{t_0} \e^{-st_0}\env(t_0) \\
+   &= \intd[_0^{1/\rh}]{t_0} \e^{-(s+\rh)t_0}
+   + \intd[_{1/\rh}^\infty]{t_0} \frac{\e^{-st_0-1}}{\rh t_0} \\
+   &= \frac{1-\e^{-s/\rh-1}}{s+\rh} + \ldots
+\end{aligned}
+\end{equation}
+%
+This loose bound can be used to derive a new envelope
+%
+\begin{equation}\label{eq:envb0}
+  \env_1(t_0) = \frac{B_0(s_0) \e^{s_0t_0-1}}{t_0} = \ldots
+  \qquad t_0 = -\frac{B_0'(s_0)}{B_0(s_0)}. 
+\end{equation}
+%
+As there are additional constraints, $\env_1(t_0) \leq \env_0(t_0)$.
+This can be Laplace-transformed to get a better bound, $B_1(s) \leq B_0(s)$, which can then be used to get an even better envelope.
+
+This procedure can be iterated an arbitrary number of times
+%
+\begin{equation}\label{eq:bootiter}
+  B_k(s) = \intd[_0^\infty]{t_0} \e^{-st_0}\env_k(t_0),
+  \qquad
+  \env_{k+1}(t_0)=\frac{B_k(s_0) \e^{s_0t_0-1}}{t_0},
+  \quad t_0 = -\frac{B_k'(s_0)}{B_k(s_0)}.
+\end{equation}
+%
+This will end at the fixed point
+%
+\begin{equation}\label{eq:bootfixedpt}
+  B_\infty(s) = \intd[_0^\infty]{t_0} \e^{-st_0}\env_\infty(t_0),
+  \qquad
+  \env_\infty(t_0)=\frac{B_\infty(s_0) \e^{s_0t_0-1}}{t_0},
+  \quad t_0 = -\frac{B_\infty'(s_0)}{B_\infty(s_0)}.
+\end{equation}
+%
+We can eliminate $\env_\infty$ and $t_0$ (assuming that $s_0\to \infty$ as $t_0 \to 0$ and vice versa) to get the following integro-differential equation:
+%
+\begin{equation}\label{eq:bootbnd}
+  B_\infty(s) = \intd[_0^\infty]{s_0} \e^{(s-s_0)B'_\infty(s_0)/B_\infty(s_0)} 
+      \brk{B'_\infty(s_0)-\frac{B_\infty(s_0)B''_\infty(s_0)}{B'_\infty(s_0)}}.
+\end{equation}
+%
+
+Because actual memory curves can, at best, touch the envelope at one time, there will always be a significant gap between the laplace transforms of the memory curves and of the envelope.
+Therefore both the bound $B_\infty(s)$ and the envelope $\env_\infty(t_0)$ will be loose.
 

### 2013-11-22 Notes/ContTime.tex
minor update to notes
scaling mode and laplace bound

@@
 This is exactly the same as \eqref{eq:derivA}.
-It follows that, using exactly the same perturbations, that the Laplace transform of the SNR curve for \emph{any} model is lower than that of a certain model with the serial topology having the same equilibrium distribution.
+Everything about the perturbation analysis carries over from the proof of the area bound, except for the \hyperref[sec:rescale]{scaling mode} that allowed us to ignore constraints \eqref{eq:Wmax}.
+If we could find a way around this, It would follow, using exactly the same perturbations, that the Laplace transform of the SNR curve for \emph{any} model is lower than that of a certain model with the serial topology having the same equilibrium distribution.
+
+In this case,
+%
+\begin{equation}\label{eq:scaleLaplace}
+  \sum_\pm\sum_{ij} \W^\pm_{ij} \, \pdiff{A(s)}{\W^\pm_{ij}} = -s A'(s).
+\end{equation}
+%
+The scaling mode essentially scales $s$ in the opposite direction.
 
@@
 \begin{equation}\label{eq:bootbnd}
-  B_\infty(s) = \intd[_0^\infty]{s_0} \e^{(s-s_0)B'_\infty(s_0)/B_\infty(s_0)} 
+  B_\infty(s) = \intd[_0^\infty]{s_0} \e^{(s-s_0)B'_\infty(s_0)/B_\infty(s_0)-1}
       \brk{B'_\infty(s_0)-\frac{B_\infty(s_0)B''_\infty(s_0)}{B'_\infty(s_0)}}.

### 2014-01-09 MHCpaper/MHCsupp.tex
minor edits
VOR parameter fitting

@@
 %}
+\date{}
 
@@
 
-For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
-We will use a larger value for $q$ for depression in the \KO\ mice.
+For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
+We will use a larger value for $q$ for depression in the \KO\ models.
 
-For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
-We will use a larger value for $x$ for depression in the \KO\ mice.
+For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
+We will use a larger value for $x$ for depression in the \KO\ models.
 
@@
 This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
-We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
-We will use larger values for depression in the \KO\ mice.
+We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
+We will use larger values for depression in the \KO\ models.
 In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
@@
 This model is parameterised by the values of $x$ chosen for potentiation and depression.
-We will use a larger value for depression in the \KO\ mice.
+We will use a larger value for depression in the \KO\ models.
 The values used are listed in \autoref{tab:params}.
@@
 Therefore the rates of potentiation and depression should not change very much either.
-This implies that the mean synaptic weight in equilibrium differs between the \KO\ mice and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
+This implies that the mean synaptic weight in equilibrium differs between the \KO\ models and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
 As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
@@
 For the most part, we set $f\dep\norm=\half$, $f\dep\inc=f\dep\norm+\Delta f$ and $f\dep\dec=f\dep\norm-\Delta f$, with $\Delta f>0$.
-We use the same values for wild-type and \KO\ mice for the reasons discussed above.
+We use the same values for wild-type and \KO\ models for the reasons discussed above.
 We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
@@
 
-We are also assuming that the relation between VOR gain and mean synaptic weight is the same for \KO\ mice ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
+We are also assuming that the relation between VOR gain and mean synaptic weight is the same for \KO\ mice and wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
 This ensures that the qualitative questions mentioned above (enhancement or impairment of learning) will not be affected.
@@
 This should be multiplied by the difference in synaptic weights.
-But, as this quantity is constant within a model and doesn't change when comparing wild-type and \KO\ mice or with/without pre-training, it can be ignored.
+But, as this quantity is constant within a model and doesn't change when comparing wild-type and \KO\ models or with/without pre-training, it can be ignored.
 
@@
   Time is measured in units of $1/r$, the mean time between plasticity events.
-  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mice with and without pre-training.
+  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ models with and without pre-training.
   (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mice,
+  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ models,
   with states of weak synaptic strength to the left and strong states to the right.
-  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ mice without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
+  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ models without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
   with states of weak synaptic strength at the top and strong states underneath. }
@@
 
-Now, consider the \KO\ mice, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
+Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
 Without pre-training:
@@
 Values of $\Delta f$ that are larger than this (stronger training) will correspond to slower learning after pre-training.
-As we would like pre-training to slow down learning in the wild-type but speed it up in the \KO\ mice, it would seem that we require $\Delta f^*(1,M) < \Delta f < \Delta f^*(\beta,M)$ (remembering that the wild-type corresponds to $\beta=1$).
+As we would like pre-training to slow down learning in the wild-type but speed it up in the \KO\ models, it would seem that we require $\Delta f^*(1,M) < \Delta f < \Delta f^*(\beta,M)$ (remembering that the wild-type corresponds to $\beta=1$).
 As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, so this range always exists.
@@
 However, this is only the initial, instantaneous rate of change.
-Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and \KO\ mice.
+Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and \KO\ models.
 We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
-This happens sooner for the \KO\ mice than the wild-type, due to the stronger depressing transitions.
-This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the \KO\ mice, even outside the desired range of values for $\Delta f$.
+This happens sooner for the \KO\ models than the wild-type, due to the stronger depressing transitions.
+This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the \KO\ models, even outside the desired range of values for $\Delta f$.
 As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} (black curves), even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.
@@
  \end{center}
-  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the \KO\ mice have impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
+  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the \KO\ models have impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
 \end{figure}
 
-In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the \KO\ mice learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ mice but slows it down in the wild-type.
-However, in this case pre-training proceeds much faster in the \KO\ mice than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.
+In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the \KO\ models learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ models but slows it down in the wild-type.
+However, in this case pre-training proceeds much faster in the \KO\ models than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.
 
@@
 
-We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the \KO\ mice, the important thing is to have \emph{strong pre-training}, not strong training.
-Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the \KO\ mice than wild-type.
+We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the \KO\ models, the important thing is to have \emph{strong pre-training}, not strong training.
+Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the \KO\ models than wild-type.
 
-Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the \KO\ mice (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
+Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the \KO\ models (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
 This is due to the mass of probability at the right end of the chain (see \autoref{fig:multistate_med}\ref{fig:multistate_med_eq} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) that is unable to contribute to change in synaptic weight until it has drifted to the centre.
-This will last for a shorter time in the \KO\ mice, due to the enhanced transitions.
+This will last for a shorter time in the \KO\ models, due to the enhanced transitions.
 This will not be a feature of any of the other models that we study here.
@@
 Here $M=2$, so the leading part of the two effects will cancel.
-The subleading effects (the normalisation of the probabilities) are responsible for the faster learning in the \KO\ mice.
+The subleading effects (the normalisation of the probabilities) are responsible for the faster learning in the \KO\ models.
 
@@
 Relaxing this assumption would correspond modelling the wild-type like the \KO\ mice, but with a different value for $\beta$.
-We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the \KO\ mice.
+We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the \KO\ models.
 
@@
 
-The only thing we have to worry about is when the wild-type catches up with the \KO\ mice.
-As the total change will be larger for the wild-type than \KO\ mice, it must overtake eventually.
-If this happens early enough, the initial period, where the \KO\ mice learns faster than wild-type, would not be seen in the experiment.
+The only thing we have to worry about is when the wild-type catches up with the \KO\ models.
+As the total change will be larger for the wild-type than \KO\ models, it must overtake eventually.
+If this happens early enough, the initial period, where the \KO\ models learns faster than wild-type, would not be seen in the experiment.
 However, the timescale for this crossover would be similar to the timescale of the exponentials.
@@
 This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
-This means that the \KO\ mice will initially learn faster than the wild-type.
+This means that the \KO\ models will initially learn faster than the wild-type.
 However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS} (solid curves).
@@
 which is always greater than 1.
-Therefore, pre-training will always enhance learning, for both \KO\ mice and wild-type, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
+Therefore, pre-training will always enhance learning, for both \KO\ models and wild-type, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
 
@@
 
-If we compare gain-increase learning in the \KO\ mice to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}: \ref{fig:pooled_deponly_eq_WT} and \ref{fig:pooled_deponly_eq_KO}, blue traces), where there are fewer synapses available for depression and resources are depleted.
-When resource depletion is sufficiently severe, the second effect dominates and the \KO\ mice learn slower than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}, solid curves), which matches what is seen in the experiment.
+If we compare gain-increase learning in the \KO\ models to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}: \ref{fig:pooled_deponly_eq_WT} and \ref{fig:pooled_deponly_eq_KO}, blue traces), where there are fewer synapses available for depression and resources are depleted.
+When resource depletion is sufficiently severe, the second effect dominates and the \KO\ models learn slower than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}, solid curves), which matches what is seen in the experiment.
 
-Gain-decrease pre-training lessens the second effect, and results in the \KO\ mice learning faster than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}), as seen in the experiment.
+Gain-decrease pre-training lessens the second effect, and results in the \KO\ models learning faster than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}), as seen in the experiment.
 
-Gain-decrease pre-training will shift the distribution to the potentiated side, where there are more synapses available for depression and resources are more plentiful, for both the \KO\ mice and wild type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}, red traces).
-This means that the pre-trained animals will learn faster then the untrained one for both \KO\ mice and wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learnS}).
+Gain-decrease pre-training will shift the distribution to the potentiated side, where there are more synapses available for depression and resources are more plentiful, for both the \KO\ models and wild type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}, red traces).
+This means that the pre-trained animals will learn faster then the untrained one for both \KO\ models and wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learnS}).
 This differs from what is seen experimentally, where the pre-trained wild-type learns slower than the untrained one.
@@
 
-In both of these, we see that the \KO\ mice are slower than wild-type without pre-training but faster with it.
+In both of these, we see that the \KO\ models are slower than wild-type without pre-training but faster with it.
 This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
@@
 
-With shorter pre-training, we see that it speeds up learning in both wild-type and \KO\ mice (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the \KO\ mice.
+With shorter pre-training, we see that it speeds up learning in both wild-type and \KO\ models (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the \KO\ models.
 This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning (see \autoref{fig:cascade_short}\ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_WT_pre}).
@@
 We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_pr}\ref{fig:cascade_long_pr_WT_pre} that longer pre-training pushes the synapses further down the cascade, slowing down learning for the wild-type.
-This effect is weaker for the \KO\ mice, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
+This effect is weaker for the \KO\ models, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
 
@@
 
-First, we see that the \KO\ mice are slower than wild-type without pre-training, as seen experimentally.
+First, we see that the \KO\ models are slower than wild-type without pre-training, as seen experimentally.
 This seems to be because the enhanced depression results in a greater fraction of synapses begin in states of weaker synaptic weight (see \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_KO}) where the depressing transitions have lower probability (see \autoref{fig:nonuni_model}).
@@
 However, gain-decrease pre-training will push synapses towards the states of stronger synaptic weight.
-The wild-type will be pushed further that the \KO\ mice, as the latter started further to the weaker side as explained above (compare \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_WT} and \ref{fig:nonuni_eq_KO}).
-As the depressing transitions have lower probability for the strongest states (see \autoref{fig:nonuni_model}), this, in addition to the enhanced depression of the \KO\ mice, will result in the \KO\ mice learning faster than wild-type, as seen experimentally.
+The wild-type will be pushed further that the \KO\ models, as the latter started further to the weaker side as explained above (compare \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_WT} and \ref{fig:nonuni_eq_KO}).
+As the depressing transitions have lower probability for the strongest states (see \autoref{fig:nonuni_model}), this, in addition to the enhanced depression of the \KO\ models, will result in the \KO\ models learning faster than wild-type, as seen experimentally.
 
@@
 Second, it will also place synapses in states where transitions have lower probability, slowing down learning.
-The second effect will be weaker in the \KO\ mice than in wild-type, as they started further to the weaker side as explained above.
-This means that, for appropriate parameter choices, gain-decrease pretraining can speed up learning in the \KO\ mice, but also slow down learning in wild-type, as seen experimentally.
+The second effect will be weaker in the \KO\ models than in wild-type, as they started further to the weaker side as explained above.
+This means that, for appropriate parameter choices, gain-decrease pretraining can speed up learning in the \KO\ models, but also slow down learning in wild-type, as seen experimentally.
 
@@
 Second, it has the metaplastic effect where repeated potentiation makes future depression harder, also due to he exponential decay of transition probabilities.
+
+\subsection{Parameter fitting}\label{sec:fit}
+
+So far, we have only considered the comparisons listed at the start of \hyperref[sec:results]{this section}, as they raised important conceptual questions.
+We have not considered the other two comparisons, the lack of difference between wild-type without pre-training and \KO\ with pretraining and between wild-type with pre-training and \KO\ without pretraining (the ``diagonal'' comparisons in the language of the main text), as they depend more on the values used for the parameters than on any conceptual issues.
+
+However, for purely aesthetic purposes, we found parameter values performed reasonably in these other two comparisons two for the three models that could reproduce the correct qualitative results of the four important comparisons: the \hyperref[sec:multistate]{serial model}, the \hyperref[sec:cascade]{cascade model} and the \hyperref[sec:nonuni]{non-uniform multistate model}.
+The parameters are listed in \autoref{tab:fit} and the learning curves are shown in \autoref{fig:fit}.
+
+\begin{table}
+ \begin{center}
+  \begin{tabularn}{|l|c|c|c|c|c|c|c|c|c|}
+    \cline{1-10}
+    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+    Model & \# states & \multicolumn{3}{c|}{parameter} & \multicolumn{3}{c|}{$f\dep$} & \multicolumn{2}{c|}{$rt$} \\
+    \cline{3-10}
+    & & pot & WT dep & \KO\ dep & base & inc & dec & train & pre \\
+
+    \cline{1-10}
+    Serial   & 38 & $0.308$  & $0.21$  & $0.566$  & 0.488 & 0.775 & 0.211 & 1  & 4  &\label{tr:serial_fit} \\
+    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002 & 1.5  & 200  &\label{tr:cascade_fit} \\
+    Non-uni. & 12 & $0.4$    & $0.4$    & $0.53$   & 0.5 & 0.7 & 0.1 & 5  & 500  &\label{tr:nonuni_fit} \\
+    \cline{1-10}
+  \end{tabularn}
+ \end{center}
+  \caption{Parameters used to reproduce detailed experimental results.
+For the sewrial model, the parameter listed is the transition probability between adjacent states.
+For the cascade and non-uniform multistate models, the parameter is the ratio of adjacent transition probabilities.} \label{tab:fit}
+\end{table}
+
+\begin{figure}
+ \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=4.5cm]{serial_fit_learnS.eps}}\label{fig:serial_fit}
+  \item\aligntop{\includegraphics[width=4.5cm]{cascade_fit_learnS.eps}}\label{fig:cascade_fit}
+  \item\aligntop{\includegraphics[width=4.5cm]{nonuni_fit_learnS.eps}}\label{fig:nonuni_fit}
+ \end{myenuma}
+ \end{center}
+  \caption{Learning curves restricted to gain-increase training with parameters listed in \autoref{tab:fit}, for (\ref{fig:serial_fit}) the serial model, (\ref{fig:cascade_fit}) the cascade model, and (\ref{fig:nonuni_fit}) the non-uniform multistate model.}\label{fig:fit}
+\end{figure}
 

### 2014-01-09 Code/MakeMat/CascadeOriginal.m

@@ -8,9 +8,9 @@ function [Wp,Wm,w]= CascadeOriginal( xp,xm,n,lambda )
 %   w  = Weights of states (+/-1)
 
 error(CheckSize(xp,@isscalar));
-error(CheckValue(xp,@(x) inrange(x,0,0.5)));
+% error(CheckValue(xp,@(x) inrange(x,0,0.5)));
 error(CheckSize(xm,@isscalar));
-error(CheckValue(xm,@(x) inrange(x,0,0.5)));
+% error(CheckValue(xm,@(x) inrange(x,0,0.5)));
 error(CheckSize(n,@isscalar));
 error(CheckValue(n,@isint));
 existsAndDefault('lambda',1);

### 2014-01-09 Code/VOR/@VORexperiment/VORexperiment.m

@@ -40,6 +40,7 @@ classdef VORexperiment
         PlotLearnS( obj,varargin )
         ViewFigs( obj )
         PrintFigs( obj,prefix )
+        St=LearnSdata(obj,varargin);
     end
     
     methods (Access=private)%for constructiuon

### 2014-01-09 Code/VOR/VORparamGUI.m

@@ -2,6 +2,7 @@ function VORparamGUI(varargin)
 
 %initial params
 builder_h=@SerialBuilder;
+FontSize=16;
 varargin=assignApplicable(varargin);
 % 
 %check validity of params
@@ -75,10 +76,17 @@ MakePlot;
         sl_callback(slh(which),0,which,varname);
     end
 
+    %callback for edit-boxes in slider panels
+    function edsl_callback(source,~,which,varname)
+        val=str2double(get(source,'String'));
+        set(slh(which),'Value',val);
+        sl_callback(slh(which),0,which,varname);
+    end
+
     %callback for edit-boxes
     function edb_callback(source,~,varname)
-        S.(varname)=floor(str2double(get(source,'String')));
-%         S.(varname)=str2double(get(source,'String'));
+%         S.(varname)=floor(str2double(get(source,'String')));
+        S.(varname)=str2double(get(source,'String'));
         MakePlot;
     end
 
@@ -107,11 +115,11 @@ MakePlot;
         %make panel
         phs(which)=uipanel(figure1,'Units','normalized',...
                 'Position',CalcPosVert(which,numsl,slpos{:}));%left bottom width height
-        set(phs(which),'Title',varname);
+        set(phs(which),'Title',varname,'FontSize',FontSize);
         %slider
         slh(which)=uicontrol(phs(which),'Style','slider',...
                         'Max',maxval,'Min',minval,'Value',S.(varname),...
-                        'SliderStep',[0.02 0.2],...
+                        'SliderStep',[0.002 0.1],...
                         'Units','normalized',...
                         'Position',[0.1 0.25 0.8 0.7],...
                         'Callback',{@sl_callback,which,varname});
@@ -119,37 +127,44 @@ MakePlot;
         uicontrol(phs(which),'Style','text',...
                         'String',num2str(minval),...
                         'Units','normalized',...
+                        'FontSize',FontSize,...
                         'Position',[0.1 0.05 0.08 0.2]);
         uicontrol(phs(which),'Style','text',...
                         'String',num2str(midval),...
                         'Units','normalized',...
+                        'FontSize',FontSize,...
                         'Position',[0.1+0.75*(midval-minval)/(maxval-minval) 0.05 0.05 0.2]);
         uicontrol(phs(which),'Style','text',...
                         'String',num2str(maxval),...
                         'Units','normalized',...
+                        'FontSize',FontSize,...
                         'Position',[0.85 0.05 0.05 0.2]);
         %button to set slider to middle value
[…]

### 2014-01-09 MHCpaper/MHCsupp.tex
VOR parameter fitting for serial model

@@
     \cline{1-10}
-    Serial   & 38 & $0.308$  & $0.21$  & $0.566$  & 0.488 & 0.775 & 0.211 & 1  & 4  &\label{tr:serial_fit} \\
+    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.11 & 0.89 & 5  & 100  &\label{tr:serial_fit} \\
     Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002 & 1.5  & 200  &\label{tr:cascade_fit} \\

### 2014-02-10 MHCpaper/MHCsupp.tex
Intuitive explanation of binary model failure

@@
 
+Because for this simple binary synapse model, the ratio of the numbers of potentiated and depressed synapses is equal to the ratio of the potentiating and depressing transition rates. 
+This means that the leading effects of enhanced intrinsic plasticity rates and saturation bias cancel each other. 
+This is eve true as the number of synapses available for further plasticity approaches zero, which requires that the depressing transition be infinitely stronger than the potentiating transition, compensating for the saturation.
+There is a requirement that the total number of synapses, potentiated and depressed, is fixed. 
+This normalization effect always dampens the effect of saturation bias, so that the effect of enhanced intrinsic plasticity rates always dominates. 
+Thus, we always see enhanced learning with enhanced plasticity with the binary synapse model.
+
 The only thing we have to worry about is when the wild-type catches up with the \KO\ models.

### 2014-02-14 MHCpaper/MHCsupp.tex
page numbering
updated MHC supp and slides

@@
 \documentclass[12pt]{article}
+%\usepackage[printfigures]{figcaps}
+%\usepackage[nolists,nomarkers]{endfloat}
 %Preamble:
@@
 \usepackage{array}
-\newcounter{tablerow}
+\newcounter{tablerow}[table]
 \newcommand{\tablerowautorefname}{row}
-\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
+\newenvironment{tabularn}[1]{\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
 %\usepackage{etoolbox}
 %\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
-%
+\renewcommand{\figureautorefname}{Supplementary Figure}
 % ----------------------------------------------------------------
@@
 % Title info:
-\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
+\title{Understanding both enhanced and impaired learning with enhanced plasticity: a saturation hypothesis (supplementary material)}
 %
@@
 \date{}
+\DeclareCaptionLabelFormat{supp}{\textbf{#1 S#2}}
+\captionsetup{labelformat=supp}
 
@@
 
+\setcounter{page}{6}
+\setcounter{figure}{5}
 \maketitle
@@
 
-\tableofcontents
-\listoffigures
+%\tableofcontents
+%\listoffigures
 
@@
 
-The values used for all these parameters in simulations are listed in \autoref{tab:params}.
+The values used for all these parameters in simulations are listed in \autoref{fig:sim_results}\ref{tab:params}.
 
-\begin{figure}
+\begin{figure}[p]
  \begin{center}
@@
 
-\begin{figure}
+\begin{figure}[p]
  \begin{center}
@@
   \item After pre-training, gain-increase learning is significantly faster in the \KO\ mice than in the wild-type.
-  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
+%  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
 \end{enumerate}
 These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
-We will not worry too much about \autoref{it:gaindown}, as gain-decrease learning uses a different mechanism to gain-increase.
+We will not study gain-decrease learning, as it uses a different mechanism to gain-increase.
 We are only modelling the effect of gain-decrease training on \emph{these} synapses.
-
-We will not worry too much about the curvature of the learning curves, as this can be changed by the nonlinear relation between synaptic weight and VOR gain.
-However, both the wild-type and \KO\ mice should be affected in the same way, so the difference in the curvature of the \KO\ mice and wild-type curves after pre-training would be nice to capture.
+We will also not discuss the curvature of the learning curves, as this can be changed by the nonlinear relation between synaptic weight and VOR gain.
+%However, both the wild-type and \KO\ mice should be affected in the same way, so the difference in the curvature of the \KO\ mice and wild-type curves after pre-training would be nice to capture.
 
@@
 
-The values for all parameters we will use can be found in \autoref{tab:params}.
+The learning curves resulting from these simulations can be found in \autoref{fig:sim_results}.
+The values for all parameters we will use can be found in \autoref{fig:sim_results}\ref{tab:params}.
 
-\begin{table}
- \begin{center}
+\begin{figure}[p]
+\begin{myenuma}
+  \begin{tabular}{l@{\hspace{0.05\linewidth}}l@{\hspace{0.05\linewidth}}l}
+      \item\label{fig:seial_sim}
+      \aligntop{\includegraphics[width=0.25\linewidth]{serial.svg}}
+      &
+      \item\label{fig:binary_sim}
+      \hspace{0.05\linewidth}\aligntop{\includegraphics[width=0.125\linewidth]{binary.svg}}
+      &
+      \item\label{fig:multistate_sim}
+      \aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}
+      \\[1.5cm]
+      \aligntop{\includegraphics[width=0.3\linewidth]{multistate_med_learnS.eps}}
+      &
+      \aligntop{\includegraphics[width=0.3\linewidth]{binary_learnS.eps}}
+      &
+      \aligntop{\includegraphics[width=0.3\linewidth]{multistate_lin_learnS.eps}}
+      \\[5.5cm]
+      \item\label{fig:pooled_sim}
+      \aligntop{\includegraphics[width=0.3\linewidth]{pooled_deponly.svg}}
+      &
+      \item\label{fig:cascade_sim}
+      \aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}
+      &
+      \item\label{fig:nonuni_sim}
+      \aligntop{\includegraphics[width=0.25\linewidth]{multistate_nonuni.svg}}
+      \\[2cm]
+      \aligntop{\includegraphics[width=0.3\linewidth]{pooled_deponly_learnS.eps}}
+      &
+      \aligntop{\includegraphics[width=0.3\linewidth]{cascade_long_learnS.eps}}
+      &
+      \aligntop{\includegraphics[width=0.3\linewidth]{nonuni_learnS.eps}}
+  \end{tabular}
+
+  \vspace{1cm}\item\label{tab:params}\aligntop{
   \begin{tabularn}{|l|c|c|c|c|c|c|}
     \cline{1-7}
-    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
     Model & \# states & pot & WT dep & \KO\ dep & $\Delta f$ & $r\tpre$ \\
     \cline{1-7}
-    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.1  & 20  &\label{tr:multistate_weak} \\
     Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 20  &\label{tr:multistate_med} \\
-    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.45 & 30  &\label{tr:multistate_strong} \\
     Two-state     & 2  & $q=0.1$  & $q=0.1$  & $q=0.2$  & 0.1  & 5   &\label{tr:binary} \\
     Multistate    & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 5   &\label{tr:multistate_lin} \\
-%    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-%                                          & 0.1 & 20 &\label{tr:pooled_plenty}\\
-%    Pooled        & 10 & $q\in[0.05,0.4]$ & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
-%                                          & 0.1 & 20 &\label{tr:pooled_scarce}\\
     Pooled res. & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
                                           & 0.4 & 20 &\label{tr:pooled_deponly}\\
-    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 20  &\label{tr:cascade_short} \\
     Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 100 &\label{tr:cascade_long} \\
@@
     \cline{1-7}
-%   \hline
+  \end{tabularn}}
+  \end{myenuma}
+  \caption[Simulation results for various models]{(\ref{fig:seial_sim}-\ref{fig:nonuni_sim}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (solid) and after gain-decrease pre-training (dashed) for wild-type (black) and \KO\ (red) models.
+(\ref{tab:params}) Parameters used for simulations.}\label{fig:sim_results}
+\end{figure}
+
+%\begin{table}
+% \begin{center}
+%  \begin{tabularn}{|l|c|c|c|c|c|c|}
+%    \cline{1-7}
 %    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
+%    Model & \# states & pot & WT dep & \KO\ dep & $\Delta f$ & $r\tpre$ \\
 %    \cline{1-7}
-%    \label{tr:cascade_short}&
-%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
-%    \label{tr:cascade_long}&
-%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
-%    \label{tr:multistate_weak}&
-%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
-%    \label{tr:multistate_strong}&
-%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
-%    \label{tr:binary}&
-%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
-%    \label{tr:pooled_plenty}&
-%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
-%                                          & -0.1 & 70 & 70 \\
-%    \label{tr:pooled_scarce}&
-%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
-%                                          & -0.1 & 70 & 70 \\
+%%    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.1  & 20  &\label{tr:multistate_weak} \\
+%    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 20  &\label{tr:multistate_med} \\
+%%    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.45 & 30  &\label{tr:multistate_strong} \\
+%    Two-state     & 2  & $q=0.1$  & $q=0.1$  & $q=0.2$  & 0.1  & 5   &\label{tr:binary} \\
+%    Multistate    & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 5   &\label{tr:multistate_lin} \\
+%%    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
+%%                                          & 0.1 & 20 &\label{tr:pooled_plenty}\\
+%%    Pooled        & 10 & $q\in[0.05,0.4]$ & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
+%%                                          & 0.1 & 20 &\label{tr:pooled_scarce}\\
+%    Pooled res. & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
+%                                          & 0.4 & 20 &\label{tr:pooled_deponly}\\
+%%    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 20  &\label{tr:cascade_short} \\
+%    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 100 &\label{tr:cascade_long} \\
+%    Non-uni. MS& 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 150 &\label{tr:nonuni} \\
 %    \cline{1-7}
-  \end{tabularn}
- \end{center}
-  \caption{Parameters used in simulations.} \label{tab:params}
-\end{table}
+%%   \hline
+%%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+%%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
+%%    \cline{1-7}
+%%    \label{tr:cascade_short}&
+%%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
+%%    \label{tr:cascade_long}&
+%%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
+%%    \label{tr:multistate_weak}&
+%%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
+%%    \label{tr:multistate_strong}&
+%%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
+%%    \label{tr:binary}&
+%%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
+%%    \label{tr:pooled_plenty}&
+%%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
+%%                                          & -0.1 & 70 & 70 \\
+%%    \label{tr:pooled_scarce}&
+%%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
+%%                                          & -0.1 & 70 & 70 \\
+%%    \cline{1-7}
+%  \end{tabularn}
+% \end{center}
+%  \caption{Parameters used for simulations in \autoref{fig:sim_results}.} \label{tab:params}
+%\end{table}
 
-\newcommand{\arrangefig}[1]{\begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{#1_learn.eps}}\label{fig:#1_learn}
-  \item\aligntop{\includegraphics[width=7cm]{#1_learnS.eps}}\label{fig:#1_learnS}
-  \\ \vp\vp
-  \item\label{fig:#1_eq}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_WT.eps}}\label{fig:#1_eq_WT}
-                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_KO.eps}}\label{fig:#1_eq_KO}
-                  \end{myenumi}
-  \\ \vp \vp
-  \item\label{fig:#1_pr}\begin{myenumi}
-                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_nopre.eps}}\label{fig:#1_pr_WT_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_pre.eps}}\label{fig:#1_pr_WT_pre}
-                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_nopre.eps}}\label{fig:#1_pr_KO_nopre}
-                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_pre.eps}}\label{fig:#1_pr_KO_pre}
-                  \end{myenumi}
- \end{myenuma}
- \end{center}}
-
-\newcommand{\captiontext}[1]{ can be found in \autoref{tr:#1} of \autoref{tab:params}.
-  Time is measured in units of $1/r$, the mean time between plasticity events.
-  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ models with and without pre-training.
-  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
-  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ models,
-  with states of weak synaptic strength to the left and strong states to the right.
-  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ models without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
-  with states of weak synaptic strength at the top and strong states underneath. }
-
-\newcommand{\captionstart}[2]{Simulation results for the \textbf{#1} model with #2}
-\newcommand{\captionstarts}[2]{Simulation results for the #1 model with #2}
-\newcommand{\captionstartn}[1]{Simulation results for the \textbf{#1} model}
-\newcommand{\captionstartsn}[1]{Simulation results for the #1 model}
-
-\newcommand{\resultsfig}[4]{\begin{figure}
-  \arrangefig{#4}
-  \caption[\captionstarts{#1}{#2}]{\captionstart{#1}{#2 #3}.
-  Other parameters \captiontext{#4} } \label{fig:#4}
-\end{figure}}
-
-\newcommand{\resultsfign}[2]{\begin{figure}
-  \arrangefig{#2}
-  \caption[\captionstartsn{#1}]{\captionstartn{#1}.
-  Parameters \captiontext{#2} } \label{fig:#2}
-\end{figure}}
+%\newcommand{\arrangefig}[1]{\begin{center}
+% \begin{myenuma}
+%  \item\aligntop{\includegraphics[width=7cm]{#1_learn.eps}}\label{fig:#1_learn}
+%  \item\aligntop{\includegraphics[width=7cm]{#1_learnS.eps}}\label{fig:#1_learnS}
+%  \\ \vp\vp
+%  \item\label{fig:#1_eq}\begin{myenumi}
+%                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_WT.eps}}\label{fig:#1_eq_WT}
+%                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_KO.eps}}\label{fig:#1_eq_KO}
+%                  \end{myenumi}
+%  \\ \vp \vp
+%  \item\label{fig:#1_pr}\begin{myenumi}
+%                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_nopre.eps}}\label{fig:#1_pr_WT_nopre}
+%                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_pre.eps}}\label{fig:#1_pr_WT_pre}
+%                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_nopre.eps}}\label{fig:#1_pr_KO_nopre}
+%                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_pre.eps}}\label{fig:#1_pr_KO_pre}
+%                  \end{myenumi}
+% \end{myenuma}
+% \end{center}}
+%
+%\newcommand{\captiontext}[1]{ can be found in \autoref{tr:#1} of \autoref{tab:params}.
+%  Time is measured in units of $1/r$, the mean time between plasticity events.
+%  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ models with and without pre-training.
+%  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
+%  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ models,
+%  with states of weak synaptic strength to the left and strong states to the right.
+%  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ models without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
+%  with states of weak synaptic strength at the top and strong states underneath. }
+%
+%\newcommand{\captionstart}[2]{Simulation results for the \textbf{#1} model with #2}
+%\newcommand{\captionstarts}[2]{Simulation results for the #1 model with #2}
+%\newcommand{\captionstartn}[1]{Simulation results for the \textbf{#1} model}
+%\newcommand{\captionstartsn}[1]{Simulation results for the #1 model}
+%
+%\newcommand{\resultsfig}[4]{\begin{figure}
+%  \arrangefig{#4}
+%  \caption[\captionstarts{#1}{#2}]{\captionstart{#1}{#2 #3}.
+%  Other parameters \captiontext{#4} } \label{fig:#4}
+%\end{figure}}
+%
+%\newcommand{\resultsfign}[2]{\begin{figure}
+%  \arrangefig{#2}
+%  \caption[\captionstartsn{#1}]{\captionstartn{#1}.
+%  Parameters \captiontext{#2} } \label{fig:#2}
+%\end{figure}}
 
@@
 
-\resultsfig{serial}{weak training}{($\Delta f=0.1$)}{multistate_weak}
+%\resultsfig{serial}{weak training}{($\Delta f=0.1$)}{multistate_weak}
+%
+%\resultsfig{serial}{moderate training}{($\Delta f=0.3$)}{multistate_med}
+%
+%\resultsfig{serial}{strong training}{($\Delta f=0.45$)}{multistate_strong}
 
-\resultsfig{serial}{moderate training}{($\Delta f=0.3$)}{multistate_med}
-
-\resultsfig{serial}{strong training}{($\Delta f=0.45$)}{multistate_strong}
-
-The results of simulations of the serial model can be seen in \autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
+The results of simulations of the serial model can be seen in 
+\autoref{fig:sim_results}\ref{fig:seial_sim}.
+%\autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
 However, we can get some insight into this model analytically.
@@
 %
-So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS} (black curves).
-On the other hand, if $\Delta f$ is close to $\half$, pre-training will initially slow down learning a lot due to the factor of $(1-2\Delta f)^{M/2-1}$, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.
+So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value).
+%, as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS} (black curves).
+On the other hand, if $\Delta f$ is close to $\half$, pre-training will initially slow down learning a lot due to the factor of $(1-2\Delta f)^{M/2-1}$.
+%, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.
 
 Intuitively, the flux depends on the slope of the distribution at the centre of the chain (with an offset due to the difference between $f\pot$ and $f\dep$).
-Pre-training has two effects: it produces a slope in the right direction (see \autoref{fig:multistate_weak}\ref{fig:multistate_weak_eq}, red trace), but it also reduces the distribution at the centre (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace).
+Pre-training has two effects: it produces a slope in the right direction 
+%(see \autoref{fig:multistate_weak}\ref{fig:multistate_weak_eq}, red trace)
+, but it also reduces the distribution at the centre. 
+%(see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace).
 For small $\Delta f$, the first effect is stronger and learning speeds up.
@@
 There are two effects here as well.
-Smaller $\beta$ will increase the probability of crossing the centre of the chain, speeding up learning, but it will also concentrate probability at the ends of the chain, depleting the centre and slowing down learning (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}\ref{fig:multistate_strong_eq_KO}, blue trace).
+Smaller $\beta$ will increase the probability of crossing the centre of the chain, speeding up learning, but it will also concentrate probability at the ends of the chain, depleting the centre and slowing down learning.
+ %(see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}\ref{fig:multistate_strong_eq_KO}, blue trace).
 The first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ and will be more significant for smaller $\beta$ or in a longer chain.
@@
 As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, so this range always exists.
-We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.
+%We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.
 
-However, this is only the initial, instantaneous rate of change.
-Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and \KO\ models.
-We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
-This happens sooner for the \KO\ models than the wild-type, due to the stronger depressing transitions.
-This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the \KO\ models, even outside the desired range of values for $\Delta f$.
-As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} (black curves), even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.
+%However, this is only the initial, instantaneous rate of change.
+%Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and \KO\ models.
+%We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
+%This happens sooner for the \KO\ models than the wild-type, due to the stronger depressing transitions.
+%This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the \KO\ models, even outside the desired range of values for $\Delta f$.
+%As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} (black curves), even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.
 
-\begin{figure}
+\begin{figure}[p]
  \begin{center}
@@
 
-In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the \KO\ models learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ models but slows it down in the wild-type.
-However, in this case pre-training proceeds much faster in the \KO\ models than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.
+In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f \in [\Delta f^*(1,M),\Delta f^*(\beta,M)]$ and look at intermediate time-scales, we will see that the \KO\ models learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ models but slows it down in the wild-type.
+%However, in this case pre-training proceeds much faster in the \KO\ models than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.
 
-Now consider what would happen if we did not treat gain-increase and gain-decrease symmetrically.
-Let us set $f\pot\dec = f\pot\norm - \overline{\Delta f}$.
-Then, \eqref{eq:multiKOpre} becomes
+%Now consider what would happen if we did not treat gain-increase and gain-decrease symmetrically.
+%Let us set $f\pot\dec = f\pot\norm - \overline{\Delta f}$.
+%Then, \eqref{eq:multiKOpre} becomes
+%%
+%\begin{equation}\label{eq:multiKOdiffpre}
+%\begin{aligned}
+%  \Phi &= 2\prn{\Delta f + \overline{\Delta f}} q \, \frac{(1-2\overline{\Delta f}) - \beta(1+2\overline{\Delta f})}
+%          {(1-2\overline{\Delta f})^M - \beta^M(1+2\overline{\Delta f})^M}   \,
+%          \brk{\beta (1-2\overline{\Delta f}) (1+2\overline{\Delta f})}^{M/2-1} \\
+%       &= 2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
+%\end{aligned}
+%\end{equation}
+%%
+%We can then look at the wild-type by setting $\beta=1$:
+%%
+%\begin{equation}\label{eq:multiWTdiffpre}
+%\begin{aligned}
+%  \Phi &= 8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \,
+%          \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
+%          {{(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}}   \\
+%       &= \frac{2\prn{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
+%\end{aligned}
+%\end{equation}
+%%
 %
-\begin{equation}\label{eq:multiKOdiffpre}
-\begin{aligned}
-  \Phi &= 2\prn{\Delta f + \overline{\Delta f}} q \, \frac{(1-2\overline{\Delta f}) - \beta(1+2\overline{\Delta f})}
-          {(1-2\overline{\Delta f})^M - \beta^M(1+2\overline{\Delta f})^M}   \,
-          \brk{\beta (1-2\overline{\Delta f}) (1+2\overline{\Delta f})}^{M/2-1} \\
-       &= 2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
-\end{aligned}
-\end{equation}
-%
-We can then look at the wild-type by setting $\beta=1$:
-%
-\begin{equation}\label{eq:multiWTdiffpre}
-\begin{aligned}
-  \Phi &= 8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \,
-          \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
-          {{(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}}   \\
-       &= \frac{2\prn{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
-\end{aligned}
-\end{equation}
-%
+%We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the \KO\ models, the important thing is to have \emph{strong pre-training}, not strong training.
+%Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the \KO\ models than wild-type.
 
-We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the \KO\ models, the important thing is to have \emph{strong pre-training}, not strong training.
-Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the \KO\ models than wild-type.
-
-Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the \KO\ models (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
-This is due to the mass of probability at the right end of the chain (see \autoref{fig:multistate_med}\ref{fig:multistate_med_eq} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) that is unable to contribute to change in synaptic weight until it has drifted to the centre.
-This will last for a shorter time in the \KO\ models, due to the enhanced transitions.
-This will not be a feature of any of the other models that we study here.
-Now, the curvature can be changed by the nonlinear relation between synaptic weight and VOR gain, but both curves would be affected in the same way.
-The difference in these two curvatures is therefore a good feature of this model.
+%Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the \KO\ models (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
+%This is due to the mass of probability at the right end of the chain
+%%(see \autoref{fig:multistate_med}\ref{fig:multistate_med_eq} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) 
+%that is unable to contribute to change in synaptic weight until it has drifted to the centre.
+%This will last for a shorter time in the \KO\ models, due to the enhanced transitions.
+%This will not be a feature of any of the other models that we study here.
+%Now, the curvature can be changed by the nonlinear relation between synaptic weight and VOR gain, but both curves would be affected in the same way.
+%The difference in these two curvatures is therefore a good feature of this model.
 
@@
 
-\resultsfign{two-state model}{binary}
+%\resultsfign{two-state model}{binary}
 
-The results of simulations of the two-state model can be seen in \autoref{fig:binary}.
+The results of simulations of the two-state model can be seen in 
+\autoref{fig:sim_results}\ref{fig:binary_sim}.
+%\autoref{fig:binary}.
 However, this model can be solved exactly:
@@
 So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
-This can be seen in \autoref{fig:binary}\ref{fig:binary_learnS} (black curves).
+This can be seen in 
+\autoref{fig:sim_results}\ref{fig:binary_sim} %\autoref{fig:binary}\ref{fig:binary_learnS} 
+(black curves).
 
@@
 which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
-This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS} (solid curves).
+This is also seen in 
+\autoref{fig:sim_results}\ref{fig:binary_sim} %\autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS} 
+(solid curves).
 As discussed above (below \eqref{eq:multiKOnopre}), the larger value of $q\dep=q/\beta$ has two effects.
@@
 %
-As we've already ruled out this model, we won't waste any time analysing this formula.
+As we've already ruled out this model, we will not analyse this formula any further.
 
@@
 
-As we didn't need the pre-training paradigm to rule out this model, we don't need to look at removing the symmetry between gain-increase and decrease.
-
 Because for this simple binary synapse model, the ratio of the numbers of potentiated and depressed synapses is equal to the ratio of the potentiating and depressing transition rates.
@@
 Thus, we always see enhanced learning with enhanced plasticity with the binary synapse model.
+
+As we did not need the pre-training paradigm to rule out this model, we will not look at removing the symmetry between gain-increase and decrease in detail.
+However, one can see that pre-training will always increase the fraction of synapses available for depression. 
+This model has no mechanism that could ever result in this causing impaired learning.
 
@@
 
-\resultsfign{multistate model}{multistate_lin}
+%\resultsfign{multistate model}{multistate_lin}
 
@@
 %
-The numerical results can be seen in \autoref{fig:multistate_lin}, but we can get some analytic insight into this model as well.
+The numerical results can be seen in 
+\autoref{fig:sim_results}\ref{fig:multistate_sim}, %\autoref{fig:multistate_lin}, 
+but we can get some analytic insight into this model as well.
 In essence, this model is like a series of two-state models attached to each other, in contrast to the serial model for which the synaptic weight only changes between one of the pairs of states.
@@
 This means that the \KO\ models will initially learn faster than the wild-type.
-However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS} (solid curves).
+However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in 
+\autoref{fig:sim_results}\ref{fig:multistate_sim} %\autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}
+(solid curves).
 
@@
 which is always greater than 1.
-Therefore, pre-training will always enhance learning, for both \KO\ models and wild-type, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
+Therefore, pre-training will always enhance learning, for both \KO\ models and wild-type, as seen in 
+\autoref{fig:sim_results}\ref{fig:multistate_sim}. %\autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
 
@@
 %\resultsfig{pooled resource}{heavy depletion}{($q\lmin/q\lmax=0.125$)}{pooled_scarce}
+%
+%\resultsfign{pooled resource}{pooled_deponly}
 
-\resultsfign{pooled resource}{pooled_deponly}
-
-The results of simulations of the pooled resource model can be seen in \autoref{fig:pooled_deponly}.
+The results of simulations of the pooled resource model can be seen in 
+\autoref{fig:sim_results}\ref{fig:pooled_sim}. %\autoref{fig:pooled_deponly}.
 It is difficult to study this model analytically.
@@
 
-If we compare gain-increase learning in the \KO\ models to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}: \ref{fig:pooled_deponly_eq_WT} and \ref{fig:pooled_deponly_eq_KO}, blue traces), where there are fewer synapses available for depression and resources are depleted.
-When resource depletion is sufficiently severe, the second effect dominates and the \KO\ models learn slower than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}, solid curves), which matches what is seen in the experiment.
+If we compare gain-increase learning in the \KO\ models to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side,
+% (compare \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}: \ref{fig:pooled_deponly_eq_WT} and \ref{fig:pooled_deponly_eq_KO}, blue traces),
+where there are fewer synapses available for depression and resources are depleted.
+When resource depletion is sufficiently severe, the second effect dominates and the \KO\ models learn slower than wild-type (see \autoref{fig:sim_results}\ref{fig:pooled_sim}, %\autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS},
+solid curves), which matches what is seen in the experiment.
 
-Gain-decrease pre-training lessens the second effect, and results in the \KO\ models learning faster than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}), as seen in the experiment.
+Gain-decrease pre-training lessens the second effect, and results in the \KO\ models learning faster than wild-type (see \autoref{fig:sim_results}\ref{fig:pooled_sim}, %\autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}
+red curves),
+as seen in the experiment.
 
-Gain-decrease pre-training will shift the distribution to the potentiated side, where there are more synapses available for depression and resources are more plentiful, for both the \KO\ models and wild type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}, red traces).
-This means that the pre-trained animals will learn faster then the untrained one for both \KO\ models and wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learnS}).
+Gain-decrease pre-training will shift the distribution to the potentiated side, where there are more synapses available for depression and resources are more plentiful, for both the \KO\ models and wild type.
+This means that the pre-trained animals will learn faster then the untrained one for both \KO\ models and wild-type (see \autoref{fig:sim_results}\ref{fig:pooled_sim}).
+%\autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learnS}).
 This differs from what is seen experimentally, where the pre-trained wild-type learns slower than the untrained one.
@@
 
-\resultsfig{cascade}{short pre-training}{($rt_\text{pre}=20$)}{cascade_short}
+%\resultsfig{cascade}{short pre-training}{($rt_\text{pre}=20$)}{cascade_short}
+%
+%\resultsfig{cascade}{long pre-training}{($rt_\text{pre}=100$)}{cascade_long}
 
-\resultsfig{cascade}{long pre-training}{($rt_\text{pre}=100$)}{cascade_long}
-
-The results of simulations of the cascade model can be seen in \autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
+The results of simulations of the cascade model can be seen in 
+\autoref{fig:sim_results}\ref{fig:cascade_sim}. %\autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
 It is difficult to study this model analytically.
@@
 
-In both of these, we see that the \KO\ models are slower than wild-type without pre-training but faster with it.
-This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
-With pre-training, some of them will now be potentiated (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_pre}), and the enhanced depression can speed up learning.
+We see that the \KO\ models are slower than wild-type without pre-training but faster with it.
+This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed. 
+%(\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
+With pre-training, some of them will now be potentiated, 
+%(\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_pre}), 
+and the enhanced depression can speed up learning.
 
-With shorter pre-training, we see that it speeds up learning in both wild-type and \KO\ models (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the \KO\ models.
-This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning (see \autoref{fig:cascade_short}\ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_WT_pre}).
+%With shorter pre-training, we see that it speeds up learning in both wild-type and \KO\ models (see \autoref{fig:cascade_short}\ref{fig:cascade_short_learnS}), whereas experimentally this only happens for the \KO\ models.
+%This is due to the fact that pre-training results in more synapses being potentiated, and thus ready for depression, but does not push them far enough down the cascade for the lower transition probabilities to slow down learning (see \autoref{fig:cascade_short}\ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_WT_pre}).
+%
+%We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_pr}\ref{fig:cascade_long_pr_WT_pre} that longer
 
-We can see from \autoref{fig:cascade_long}\ref{fig:cascade_long_pr}\ref{fig:cascade_long_pr_WT_pre} that longer pre-training pushes the synapses further down the cascade, slowing down learning for the wild-type.
-This effect is weaker for the \KO\ models, as the equilibrium distribution is not as heavily concentrated at the end (see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
+Sufficiently long pre-training pushes the synapses far down the cascade, slowing down learning for the wild-type (see \autoref{fig:sim_results}\ref{fig:cascade_sim}, black curves).
+This effect is weaker for the \KO\ models, as the equilibrium distribution is not as heavily concentrated at the end. 
+%(see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
 
@@
 
-\resultsfign{non-uniform multistate}{nonuni}
+%\resultsfign{non-uniform multistate}{nonuni}
 
-The results of simulations of the non-uniform multistate model can be seen in \autoref{fig:nonuni}.
+The results of simulations of the non-uniform multistate model can be seen in 
+\autoref{fig:sim_results}\ref{fig:nonuni_sim}. %\autoref{fig:nonuni}.
 It is difficult to study this model analytically.
@@
 First, we see that the \KO\ models are slower than wild-type without pre-training, as seen experimentally.
-This seems to be because the enhanced depression results in a greater fraction of synapses begin in states of weaker synaptic weight (see \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_KO}) where the depressing transitions have lower probability (see \autoref{fig:nonuni_model}).
+This seems to be because the enhanced depression results in a greater fraction of synapses begin in states of weaker synaptic weight %(see \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_KO})
+where the depressing transitions have lower probability (see \autoref{fig:nonuni_model}).
 
 However, gain-decrease pre-training will push synapses towards the states of stronger synaptic weight.
-The wild-type will be pushed further that the \KO\ models, as the latter started further to the weaker side as explained above (compare \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_WT} and \ref{fig:nonuni_eq_KO}).
+The wild-type will be pushed further that the \KO\ models, as the latter started further to the weaker side as explained above.
+%(compare \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_WT} and \ref{fig:nonuni_eq_KO}).
 As the depressing transitions have lower probability for the strongest states (see \autoref{fig:nonuni_model}), this, in addition to the enhanced depression of the \KO\ models, will result in the \KO\ models learning faster than wild-type, as seen experimentally.
 
-When comparing learning with and without gain-decrease pretraining, there are two effects to consider.
+When comparing learning with and without gain-decrease pre-training, there are two effects to consider.
 First, the pushing synapses towards the states of stronger synaptic weight, resulting in more synapses being available for depression, speeding up learning.
@@
 The second effect will be weaker in the \KO\ models than in wild-type, as they started further to the weaker side as explained above.
-This means that, for appropriate parameter choices, gain-decrease pretraining can speed up learning in the \KO\ models, but also slow down learning in wild-type, as seen experimentally.
+This means that, for appropriate parameter choices, gain-decrease pre-training can speed up learning in the \KO\ models, but also slow down learning in wild-type, as seen experimentally.
 
@@
 So far, we have only considered the comparisons listed at the start of \hyperref[sec:results]{this section}, as they raised important conceptual questions.
-We have not considered the other two comparisons, the lack of difference between wild-type without pre-training and \KO\ with pretraining and between wild-type with pre-training and \KO\ without pretraining (the ``diagonal'' comparisons in the language of the main text), as they depend more on the values used for the parameters than on any conceptual issues.
+We have not considered the other two comparisons, the lack of difference between wild-type without pre-training and \KO\ with pre-training and between wild-type with pre-training and \KO\ without pre-training (the ``diagonal'' comparisons in the language of the main text), as they depend more on the values used for the parameters than on any conceptual issues.
 
-However, for purely aesthetic purposes, we found parameter values performed reasonably in these other two comparisons two for the three models that could reproduce the correct qualitative results of the four important comparisons: the \hyperref[sec:multistate]{serial model}, the \hyperref[sec:cascade]{cascade model} and the \hyperref[sec:nonuni]{non-uniform multistate model}.
-The parameters are listed in \autoref{tab:fit} and the learning curves are shown in \autoref{fig:fit}.
+However, to avoid confusion, we found parameter values performed reasonably in these other two comparisons two for the three models that could reproduce the correct qualitative results of the four important comparisons: the \hyperref[sec:multistate]{serial model}, the \hyperref[sec:cascade]{cascade model} and the \hyperref[sec:nonuni]{non-uniform multistate model}.
+The learning curves are shown in \autoref{fig:fit} and the parameters are listed in \autoref{fig:fit}\ref{tab:fit}.
 
-\begin{table}
+%\begin{table}
+% \begin{center}
+%  \begin{tabularn}{|l|c|c|c|c|c|c|c|c|c|}
+%    \cline{1-10}
+%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+%    Model & \# states & \multicolumn{3}{c|}{parameter} & \multicolumn{3}{c|}{$f\dep$} & \multicolumn{2}{c|}{$rt$} \\
+%    \cline{3-10}
+%    & & pot & WT dep & \KO\ dep & base & inc & dec & train & pre \\
+%
+%    \cline{1-10}
+%    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.11 & 0.89 & 5  & 100  &\label{tr:serial_fit} \\
+%    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002 & 1.5  & 200  &\label{tr:cascade_fit} \\
+%    Non-uni. & 12 & $0.4$    & $0.4$    & $0.53$   & 0.5 & 0.7 & 0.1 & 5  & 500  &\label{tr:nonuni_fit} \\
+%    \cline{1-10}
+%  \end{tabularn}
+% \end{center}
+%  \caption{Parameters used to reproduce detailed experimental results.
+For the sewrial model, the parameter listed is the transition probability between adjacent states.
+For the cascade and non-uniform multistate models, the parameter is the ratio of adjacent transition probabilities.} \label{tab:fit}
+%\end{table}
+
+\begin{figure}[p]
  \begin{center}
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=4.5cm]{serial_fit_learnS.eps}}\label{fig:serial_fit}
+  \item\aligntop{\includegraphics[width=4.5cm]{cascade_fit_learnS.eps}}\label{fig:cascade_fit}
+  \item\aligntop{\includegraphics[width=4.5cm]{nonuni_fit_learnS.eps}}\label{fig:nonuni_fit}
+
+  \vspace{1cm}\item\label{tab:fit}\aligntop{
   \begin{tabularn}{|l|c|c|c|c|c|c|c|c|c|}
@@
     % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-    Model & \# states & \multicolumn{3}{c|}{parameter} & \multicolumn{3}{c|}{$f\dep$} & \multicolumn{2}{c|}{$rt$} \\
+    Model & \#  & \multicolumn{3}{c|}{parameter} & \multicolumn{3}{c|}{$f\dep$} & \multicolumn{2}{c|}{$rt$} \\
     \cline{3-10}
-    & & pot & WT dep & \KO\ dep & base & inc & dec & train & pre \\
+    & states & pot & WT dep & {\footnotesize\KO} dep & base & inc & dec & train & pre \\
 
@@
     \cline{1-10}
-  \end{tabularn}
- \end{center}
-  \caption{Parameters used to reproduce detailed experimental results.
-For the sewrial model, the parameter listed is the transition probability between adjacent states.
-For the cascade and non-uniform multistate models, the parameter is the ratio of adjacent transition probabilities.} \label{tab:fit}
-\end{table}
-
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=4.5cm]{serial_fit_learnS.eps}}\label{fig:serial_fit}
-  \item\aligntop{\includegraphics[width=4.5cm]{cascade_fit_learnS.eps}}\label{fig:cascade_fit}
-  \item\aligntop{\includegraphics[width=4.5cm]{nonuni_fit_learnS.eps}}\label{fig:nonuni_fit}
+  \end{tabularn}}
  \end{myenuma}
  \end{center}
-  \caption{Learning curves restricted to gain-increase training with parameters listed in \autoref{tab:fit}, for (\ref{fig:serial_fit}) the serial model, (\ref{fig:cascade_fit}) the cascade model, and (\ref{fig:nonuni_fit}) the non-uniform multistate model.}\label{fig:fit}
+  \caption[Learning curves for models after parameter fitting]{Learning curves restricted to gain-increase training, for (\ref{fig:serial_fit}) the serial model, (\ref{fig:cascade_fit}) the cascade model, and (\ref{fig:nonuni_fit}) the non-uniform multistate model.
+  (\ref{tab:fit}) Parameters used to reproduce detailed experimental results.
+For the serial model, the parameter listed is the transition probability between adjacent states.
+For the cascade and non-uniform multistate models, the parameter is the ratio of adjacent transition probabilities.}\label{fig:fit}
 \end{figure}

### 2014-02-18 MHCpaper/MHCsupp.tex
improved plasticity video

@@
 % ----------------------------------------------------------------
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
 %\usepackage[printfigures]{figcaps}
@@
 
-\setcounter{page}{6}
-\setcounter{figure}{5}
+\setcounter{page}{5}
+\setcounter{figure}{4}
 \maketitle
@@
 
-\begin{figure}[p]
+\begin{figure}
  \begin{center}
@@
   (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.
-  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised X
-  by these two values.
+  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised  by these two values.
   The synaptic weight varies linearly in the interval $[-1,1]$.
@@
   \caption[The non-uniform multistate model]{The non-uniform multistate model.
-  Similar to the multistate model, \autoref{fig:models}\ref{fig:multistate_model}), the synaptic weight varies linearly in the interval $[-1,1]$, but the transition probabilities between adjacent states decays exponentially away from the central transition for both potentiation and depression.} \label{fig:nonuni_model}
+  Similar to the multistate model, (\autoref{fig:models}\ref{fig:multistate_model}), the synaptic weight varies linearly in the interval $[-1,1]$, but the transition probabilities between adjacent states decays exponentially away from the central transition for both potentiation and depression.} \label{fig:nonuni_model}
 \end{figure}
@@
 We could adjust $r$ to keep $r\pot$ unchanged, if so desired, but this would only change the overall timescale and would not affect any of the qualitative comparisons that we are concerned with here.
-The values of these parameters are listed in \autoref{tab:params}.
+The values of these parameters are listed in \autoref{fig:sim_results}\ref{tab:params}.
 
@@
 
-\begin{figure}[p]
+\begin{figure}
  \begin{center}
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
-  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
+  \item\aligntop{\includegraphics[width=0.45\linewidth]{multistate_betastar.eps}}\label{fig:multistate_betastar}
+  \item\aligntop{\includegraphics[width=0.45\linewidth]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
  \end{myenuma}
@@
 
-However, to avoid confusion, we found parameter values performed reasonably in these other two comparisons two for the three models that could reproduce the correct qualitative results of the four important comparisons: the \hyperref[sec:multistate]{serial model}, the \hyperref[sec:cascade]{cascade model} and the \hyperref[sec:nonuni]{non-uniform multistate model}.
+However, to avoid confusion, we found parameter values performed reasonably in these other two comparisons for the three models that could reproduce the correct qualitative results of the four important comparisons: the \hyperref[sec:multistate]{serial model}, the \hyperref[sec:cascade]{cascade model} and the \hyperref[sec:nonuni]{non-uniform multistate model}.
 The learning curves are shown in \autoref{fig:fit} and the parameters are listed in \autoref{fig:fit}\ref{tab:fit}.
@@
 
-\begin{figure}[p]
+\begin{figure}
  \begin{center}
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=4.5cm]{serial_fit_learnS.eps}}\label{fig:serial_fit}
-  \item\aligntop{\includegraphics[width=4.5cm]{cascade_fit_learnS.eps}}\label{fig:cascade_fit}
-  \item\aligntop{\includegraphics[width=4.5cm]{nonuni_fit_learnS.eps}}\label{fig:nonuni_fit}
+  \item\aligntop{\includegraphics[width=0.28\linewidth]{serial_fit_learnS.eps}}\label{fig:serial_fit}
+  \item\aligntop{\includegraphics[width=0.28\linewidth]{cascade_fit_learnS.eps}}\label{fig:cascade_fit}
+  \item\aligntop{\includegraphics[width=0.28\linewidth]{nonuni_fit_learnS.eps}}\label{fig:nonuni_fit}
 

### 2014-03-06 MHCpaper/MHCsupp.tex
changed ko labels

@@
 \newcommand{\wt}{_{\text{WT}}}
-%\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
-\newcommand{\KO}{D$^\mathrm{b}$K$^{\mathrm{b}-/-}$}
-\newcommand{\ko}{_{\text{MHC}}}
+\newcommand{\ko}{_{\text{K$^\mathrm{b}$D$^\mathrm{b}$-/-}}}
+\newcommand{\KO}{K$^\mathrm{b}$D$^{\mathrm{b}-/-}$}
+%\newcommand{\ko}{_{\text{MHC}}}
 %\newcommand{\KO}{MHC}
@@
       \\[1.5cm]
-      \aligntop{\includegraphics[width=0.3\linewidth]{multistate_med_learnS.eps}}
+      \aligntop{\includegraphics[width=0.3\linewidth]{serial_med_learnS.eps}}
       &
@@
     \cline{1-10}
-    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.11 & 0.89 & 5  & 100  &\label{tr:serial_fit} \\
+    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.89 & 0.11 & 5  & 100  &\label{tr:serial_fit} \\
     Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002 & 1.5  & 200  &\label{tr:cascade_fit} \\

### 2014-03-06 Code/Builders/MultistateBuilder.m

@@ -1,0 +1,13 @@
+function [ Wp,Wm,w ] = MultistateBuilder( n,q )
+%[Wp,Wm,w]=SERIALBUILDER(n,q) build serial model
+%   q  = transition rates
+%   n  = number of states (even)
+%   WP = potentiation transition rates
+%   WM = depression transition rates
+%   w  = Weights of states (+/-1)
+
+
+[Wp,Wm]=MakeSMS(q*ones(1,n-1));
+w=LinearWeights(n);
+end
+

### 2014-03-06 Code/VOR/@VORexperiment/VORexperiment.m

@@ -16,7 +16,7 @@ classdef VORexperiment
     
     properties %labels
         WTlabel='WT';
-        KOlabel='D^bK^{b-/-}';
+        KOlabel='K^bD^{b-/-}';
         noprelabel='No Pre-training';
         withprelabel='w/ Pre-training';
         WTcolor='k';

### 2014-03-06 MHCpaper/Figures/MakeFigs.m

@@ -1,11 +1,11 @@
-MakeVORfigs('multistate_med',0.3,0.4,-0.3,5,20,'print')
-MakeVORfigs('multistate_strong',0.3,0.4,-0.45,5,30,'print')
-MakeVORfigs('multistate_weak',0.3,0.4,-0.1,5,20,'print')
-MakeVORfigs('binary',0.1,0.2,-0.1,5,5,'print')
-MakeVORfigs('multistate_lin',0.3,0.4,-0.3,5,5,'print','LinWeights',true)
+printVORfigs('serial_med',0.3,0.4,-0.3,5,20,'print')
+% MakeVORfigs('multistate_strong',0.3,0.4,-0.45,5,30,'print')
+% MakeVORfigs('multistate_weak',0.3,0.4,-0.1,5,20,'print')
+printVORfigs('binary',0.1,0.2,-0.1,5,5,'print')
+printVORfigs('multistate_lin',0.3,0.4,-0.3,5,5,'print','LinWeights',true)
 % MakeVORfigs('pooled_plenty',[0.3 0.4],[0.6 0.8],-0.1,5,20,'print','Pooled',true)
 % MakeVORfigs('pooled_scarce',[0.05 0.4],[0.1 0.8],-0.1,5,20,'print','Pooled',true)
-MakeVORfigs('cascade_short',0.25,0.33,-0.3,5,20,'print')
-MakeVORfigs('cascade_long',0.25,0.33,-0.3,5,100,'print')
-MakeVORfigs('nonuni',0.25,0.33,-0.3,5,100,'print','LinWeights',true)
-MakeVORfigs('pooled_deponly',[0.6 0.0006 0.008],[1 0.001],-0.4,5,20,'print','Pooled',true)
+% MakeVORfigs('cascade_short',0.25,0.33,-0.3,5,20,'print')
+printVORfigs('cascade_long',0.25,0.33,-0.3,5,100,'print')
+printVORfigs('nonuni',0.25,0.33,-0.3,5,100,'print','LinWeights',true)
+printVORfigs('pooled_deponly',[0.6 0.0006 0.008],[1 0.001],-0.4,5,20,'print','Pooled',true)

### 2014-07-03 Notes/Identification.tex
synapse fitting notes

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Pib}{\boldsymbol{\Pi}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\pr}{\mathbb{P}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:

\title{Complex Synapse Identification}%
\author{Subhaneil Lahiri}%

% ----------------------------------------------------------------
\begin{document}
% ----------------------------------------------------------------
\maketitle
% ----------------------------------------------------------------
\begin{abstract}
  Notes on using HMM techniques to fit models of complex synapses
\end{abstract}
% ----------------------------------------------------------------

\section{Introduction \label{sec:intro}}

\section{Notation \label{sec:notation}}

Let $t$ be the time of a plasticity event, assumed to be an integer. 
We denote a sequence of event times by $[0,T]$.

We denote states of a synapse by subscripts $i,j$. 
The state at a particular time during a sequence is $i(t)$.
An entire trajectory (sequence of states) is denoted by $i[0,T]$.

The synaptic weight is denoted by $\w$.
Its value when the synapse is in state $i$ is $\w_i$.
Its observed value at time $t$ during a sequence is $\w(t)$.
An entire sequence of observations is denoted by $\w[0,T]$.

The different types of plasticity (\eg potentiation and depression) are indicated by superscripts $\mu,\nu$.
The type of plasticity used at time $t$ is $\mu(t)$.
An entire sequence of plasticity types is denoted by $\mu[0,T]$.

When we are combining data from several sequences of plasticity events, the individual sequences will be denoted by superscripts $a,b$.
The set of all sequences will be denoted by the superscript $A$.

The Markov matrices describing plasticity are $\M^\mu_{ij}$.
The probabilities of the initial states are $\pib_i$.
The set of parameters defining a model (\ie $\brc{\M^\mu_{ij}}$ and $\pib_i$) are denoted by $\M$ for convenience.

\section{Expectation-Maximisation (EM) \label{sec:em}}

When we don't have any prior information about the model $\M$, we fit the model by maximising the likelihood function:
%
\begin{equation}\label{eq:like}
  \pr\cond{\w^A[0,T]}{\M} = \prod_a \pr\cond{\w^a[0,T]}{\M}.
\end{equation}
%
To do this, we define an axillary function
%
\begin{equation}\label{eq:emq}
  Q(\M,\M') = \sum_{i^A[0,T]} \pr\cond{i^A[0,T],\w^A[0,T]}{\M} \log\!\! \brk{
      \frac{ \pr\cond{i^A[0,T],\w^A[0,T]}{\M'} }{ \pr\cond{i^A[0,T],\w^A[0,T]}{\M} } }.
\end{equation}
%
It can be shown that, if $\M'$ is chosen to maximise this quantity (which is easier than maximising the likelihood), the it will have a higher likelihood than $\M'$.
In effect, $\M$ is used to estimate the hidden state trajectories, which are then used to re-estimate the model.
This can be done iteratively to find a local maximum of the likelihood.
If this is repeated several times with random initialisation, we can hope to find the global maximum.

When we have several sequences, we can write
%
\begin{multline}\label{eq:emqseveral}
%\begin{aligned}
  Q(\M,\M') = \sum_{i^A[0,T]} \brk{ \prod_a \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } \brk{\sum_a  \log\!\!  \brk{
      \frac{ \pr\cond{i^A[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^A[0,T],\w^a[0,T]}{\M} } } }\\
   = \sum_a \prod_{b \neq a} \pr\cond{\w^b[0,T]}{\M} \sum_{i^a[0,T]} \pr\cond{i^a[0,T],\w^a[0,T]}{\M} \log\!\!  \brk{
      \frac{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } } \\
   = \pr\cond{\w^A[0,T]}{\M} \sum_a \frac{Q^a(\M,\M')}{\pr\cond{\w^a[0,T]}{\M}}.
%\end{aligned}
\end{multline}
%

\subsection{Single sequence (Baum-Welch) \label{sec:bw}}

Suppose we have a single sequence of plasticity events.
Then the joint likelihood of a trajectory and sequence observed synaptic weights is given by
%
\begin{equation}\label{eq:trajlike}
\begin{aligned}
  \pr\cond{t[0,T], \w[0,T]}{\M} &= \pib_{i(0)} \, \Pib(0) 
        \prod_{t=0}^{T-1} \M^{\mu(t)}_{i(t)i(t+1)} \, \Pib(t+1),\\
  \text{where}\quad
  \Pib(t) &= \delta_{\w_{i(t)}\w(t)}
\end{aligned}
\end{equation}
%
We wish to maximise \eqref{eq:emq} subject to normalisation constraints.
Therefore we maximise the Lagrangian:
%
\begin{equation}\label{eq:lagrangesingle}
  \CL = Q(\M,\M') + \lambda\prn{1-\sum_i\pib'_i} + \sum_{\mu i} \lambda^\mu_i\prn{1-\sum_j\M'^\mu_{ij}}. 
\end{equation}
%
We find
%
\begin{equation}\label{eq:singlediffs}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \frac{N^\mu_{ij}}{\M'^\mu_{ij}} - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \frac{\alpha_i(0) \, \beta_i(0)}{\pib'_{i}} - \lambda,
\end{aligned}
\end{equation}
%
where the numerator $N^\mu_{ij}$ and the forward/backward variables $\alpha$ and $\beta$ are defined as:
%
\begin{equation}\label{eq:alphabeta}
\begin{aligned}
  N^\nu_{i(t)i(t+1)} &= \sum_{t=0}^{T-1} \pr\cond{i[t,t+1]),\w[0,T]}{\M} \\
     &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, \alpha_i(t) \, \beta_j(t+1) \, \M^{\nu}_{i(t)i(t+1)} \, \Pib(t+1) ,\\
  \alpha_{i(t)}(t) &= \pr\cond{i(t),\w[0,t]}{\M} \\
     &= \sum_{i[0,t-1]} \pib_{i(0)} \, \delta_{\w_{i(0)}\w(0)} \prod_{s=0}^{t-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib(s+1), \\
  \beta_{i(t)}(t) &= \pr\cond{i(t),\w[t+1,T]}{\M} \\
     &= \sum_{i[t+1,T]} \prod_{s=t}^{T-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib(s+1).
\end{aligned}
\end{equation}
%
Note that, whilst disjoint sequences $\w[s,s']$ and $\w[t,t']$ are not independent, they are conditionally independent given the state at any time between the two intervals.
Therefore:
%
\begin{equation}\label{eq:stateprob}
  \pr\cond{i(t),\w[0,T]}{\M} = \alpha_{i(t)}(t) \, \beta_{i(t)}(t).
\end{equation}
%

These quantities usually suffer from underflow.
This can be avoided by using normalised forwar/backward variables:
%
\begin{equation}\label{eq:normalphabeta}
\begin{aligned}
  \tilde{\alpha}_i(t) &= \alpha_i(t) \prod_{s=0}^{t} \eta(s), \\
  \tilde{\beta}_i(t) &= \beta_i(t) \prod_{s=t+1}^{T} \eta(s),
\end{aligned}
\end{equation}
%
where $\eta[0,T]$ is chosen so that all of the $\tilde{\alpha}[0,T]$ are normalised: $\sum_i \tilde{\alpha}_i(t) = 1$.
These quantities have the interpretations:
%
\begin{equation}\label{eq:normintepr}
\begin{aligned}
  \tilde{\alpha}_{i(t)}(t) &= \pr\cond{i(t)}{\w[0,t],\M}, \\
  \eta(t) &= \pr\cond{\w(t)}{\w[0,t-1],\M}, \\
  \prod_{s=0}^t \eta(t) &= \pr\cond{\w[0,t]}{\M}.
\end{aligned}
\end{equation}
%
Note that the new backward variables $\tilde{\beta}$ are not normalised and have no particularly interesting interpretation.
However:
%
\begin{equation}\label{eq:normstateprob}
\begin{aligned}
  \widetilde{N}^\nu_{i(t)i(t+1)} &= \sum_{t=0}^{T-1} \pr\cond{i[t,t+1])}{\w[0,T],\M} \\
     &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, \tilde{\alpha}_i(t) \, \tilde{\beta}_j(t+1) \, \eta(t+1) \, \M^{\nu}_{i(t)i(t+1)} \, \Pib(t+1) ,\\
  \pr\cond{i(t)}{\w[0,T],\M} &= \alpha_{i(t)}(t) \, \beta_{i(t)}(t).
\end{aligned}
\end{equation}
%
Then we can write
%
\begin{equation}\label{eq:singlediffsnorm}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr\cond{\w[0,T]}{\M} \, \frac{ \widetilde{N}^\mu_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \pr\cond{\w[0,T]}{\M} \, \frac{ \tilde{\alpha}_i(0) \, \tilde{\beta}_i(0) }{ \pib'_{i} } - \lambda,
\end{aligned}
\end{equation}
%
The factors of $\pr\cond{\w[0,T]}{\M}$ can be absorbed by a redefinition of the Lagrange multipliers.

Demanding a maximum and setting the derivatives to zero leaves us with the update rule:
%
\begin{equation}\label{eq:BWupdate}
\begin{aligned}
  \M'^\mu_{ij} &= \frac{ \widetilde{N}^\mu_{ij} }{ \sum_k \widetilde{N}^\mu_{ik} },\\
  \pib'_{i} &= \frac{ \tilde{\alpha}_i(0) \, \tilde{\beta}_i(0) }{ \sum_j \tilde{\alpha}_j(0) \, \tilde{\beta}_j(0) }.
\end{aligned}
\end{equation}
%
This can be iterated until a maximum is found.

\subsection{Multiple sequences (Rabiner-Juang) \label{sec:rj}}

When we have multiple sequences of plasticity events, we will still want to maximise the Lagrangian \eqref{eq:lagrangesingle}, but now we will use the auxiliary function \eqref{eq:emqseveral}.
This results in
%
\begin{equation}\label{eq:multidiffs}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr\cond{\w^A[0,T]}{\M} \,\sum_a \frac{1}{\pr\cond{\w^a[0,T]}{\M}} \,
    \frac{ N^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
    &= \pr\cond{\w^A[0,T]}{\M} \, \frac{ \sum_a \widetilde{N}^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \pr\cond{\w^A[0,T]}{\M} \,\sum_a \frac{1}{\pr\cond{\w^a[0,T]}{\M}} \,
    \frac{ \alpha^a_i(0) \, \beta^a_i(0) }{ \pib'_{i} } - \lambda,\\
    &= \pr\cond{\w^A[0,T]}{\M} \,
    \frac{ \sum_a \tilde{\alpha}^a_i(0) \, \tilde{\beta}^a_i(0) }{ \pib'_{i} } - \lambda.
\end{aligned}
\end{equation}
%
Once again, we can absorb the factors of $\pr\cond{\w[0,T]}{\M}$ by a redefinition of the Lagrange multipliers.

Defining $\widetilde{N}^{A\mu}_{ij} = \sum_a \widetilde{N}^{a\mu}_{ij}$ and setting the derivatives to zero leaves us with the update rule:
%
\begin{equation}\label{eq:RJupdate}
\begin{aligned}
  \M'^\mu_{ij} &= \frac{ \widetilde{N}^{A\mu}_{ij} }{ \sum_k \widetilde{N}^{A\mu}_{ik} },\\
  \pib'_{i} &= \frac{ \sum_a \tilde{\alpha}^a_i(0) \, \tilde{\beta}^a_i(0) }{ \sum_{bj} \tilde{\alpha}^b_j(0) \, \tilde{\beta}^b_j(0) }.
\end{aligned}
\end{equation}
%
This can be iterated until a maximum is found.

\section{Sparseness promoting priors \label{sec:priors}}

When we have a prior distribution for the model, $\pr(\M)$, instead of maximising the likelihood \eqref{eq:like}, we can maximise the posterior:
%
\begin{equation}\label{eq:posterior}
  \pr\cond{\M}{\w[0,T]} = \frac{ \pr(\w[0,T],\M) }{ \pr(\w[0,T) } = \frac{ \pr\cond{\w[0,T]}{\M} \pr(\M) }{ \pr(\w[0,T) }.
\end{equation}
%
As the denominator is independent of the model, this is equivalent to maximising the numerator, \ie the joint distribution.

Using the same trick as before \eqref{eq:emq}, we can define an auxilliary function
%
\begin{equation}\label{eq:emqprior}
\begin{aligned}
  \widetilde{Q}(\M,\M') &= \sum_{i^A[0,T]} \pr\!\prn{i^A[0,T],\w^A[0,T],\M} \log\!\! \brk{
      \frac{ \pr\!\prn{i^A[0,T],\w^A[0,T],\M'} }{ \pr\!\prn{i^A[0,T],\w^A[0,T],\M} } }\\
      &= \pr\!\prn{\M} Q(\M,\M') + \pr\!\prn{\w^A[0,T],\M} \log\!\! \brk{
      \frac{ \pr\!\prn{\M'} }{ \pr\!\prn{\M} } }.
\end{aligned}
\end{equation}
%

We will consider priors of the form
%
\begin{equation}\label{eq:priorsum}
  \pr\!\prn{\M} = \frac{\exp\prn{-\beta\sum_{\mu ij}E^\mu_{ij}(\M^\mu_{ij})}}{Z(\beta)}.
\end{equation}
%
We will not put any prior on $\pib$, so its update rule will be unaffected.
The normalisation constant $Z(\beta)$ will not play any role either.
Now the derivatives read:
%
\begin{equation}\label{eq:multidiffs}
  \pdiff{\CL}{\M'^\mu_{ij}} 
    = \pr\!\prn{\w[0,T],\M} \prn{ \frac{ \widetilde{N}^{A\mu}_{ij} }{ \M'^\mu_{ij} }  - \pdiff{E^\mu_{ij}}{\M'^\mu_{ij}} }- \lambda^\mu_{i}.
\end{equation}
%
Once again, we can absorb the factors of $\pr\!\prn{\w[0,T],\M}$ by a redefinition of the Lagrange multipliers.

%\subsection*{Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection*{Appendices}
%\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{utcaps_sl}
\bibliography{maths,neuro,markov}

\end{document}
% ----------------------------------------------------------------

### 2014-07-03 Notes/Identification.tex
notes on sparse priors

@@
 %
-To do this, we define an axillary function
+To do this, we define an axillary function \cite{Baum1970baumwelch,Dempster2007EM}:
 %
@@
 These quantities usually suffer from underflow.
-This can be avoided by using normalised forwar/backward variables:
+This can be avoided by using normalised forwar/backward variables \cite{Zhai2003HMMnorm}:
 %
@@
 
-Demanding a maximum and setting the derivatives to zero leaves us with the update rule:
+Demanding a maximum and setting the derivatives to zero leaves us with the update rule \cite{Baum1970baumwelch}:
 %
@@
 
-Defining $\widetilde{N}^{A\mu}_{ij} = \sum_a \widetilde{N}^{a\mu}_{ij}$ and setting the derivatives to zero leaves us with the update rule:
+Defining $\widetilde{N}^{A\mu}_{ij} = \sum_a \widetilde{N}^{a\mu}_{ij}$ and setting the derivatives to zero leaves us with the update rule \cite{rabiner1993speechrec}:
 %
@@
 %
-\begin{equation}\label{eq:multidiffs}
+\begin{equation}\label{eq:priordiffs}
   \pdiff{\CL}{\M'^\mu_{ij}} 
-    = \pr\!\prn{\w[0,T],\M} \prn{ \frac{ \widetilde{N}^{A\mu}_{ij} }{ \M'^\mu_{ij} }  - \pdiff{E^\mu_{ij}}{\M'^\mu_{ij}} }- \lambda^\mu_{i}.
+    = \pr\!\prn{\w[0,T],\M} \prn{ \frac{ \widetilde{N}^{A\mu}_{ij} }{ \M'^\mu_{ij} }  - \beta \pdiff{E^\mu_{ij}}{\M'^\mu_{ij}} }- \lambda^\mu_{i}.
 \end{equation}
@@
 Once again, we can absorb the factors of $\pr\!\prn{\w[0,T],\M}$ by a redefinition of the Lagrange multipliers.
+
+\subsection{Off-diagonal $\CL^1$ penalty \label{sec:offdiagL1}}
+
+Here we use the off-diagonal $\CL^1$ norm as a penalty:
+%
+\begin{equation}\label{eq:L1penalty}
+  E^\mu_{ij} = (1-\delta_{ij}) \M^\mu_{ij}.
+\end{equation}
+%
+Then the condition for a maximum \eqref{eq:priordiffs} reads
+%
+\begin{equation}\label{eq:L1priormax}
+\begin{aligned}
+  \frac{ \widetilde{N}^{A\mu}_{ij} }{ \M'^\mu_{ij} }  - (1-\delta_{ij})\beta - \lambda^\mu_{i} &= 0, \\
+   \M'^\mu_{ij} &= \frac{\widetilde{N}^{A\mu}_{ij}}{\lambda^\mu_{i}+(1-\delta_{ij})\beta}.
+\end{aligned}
+\end{equation}
+%
+Demanding normalisation gives
+%
+\begin{multline}\label{eq:L1norm}
+  \sum_{j \neq i}\frac{\widetilde{N}^{A\mu}_{ij}}{\lambda^\mu_{i}+\beta} +\frac{\widetilde{N}^{A\mu}_{ii}}{\lambda^\mu_{i}} =0, \\
+  \implies \quad
+  \lambda^\mu_{i} = \half \brk{ \prn{\sum_j \widetilde{N}^{A\mu}_{ij} -\beta} + 
+     \sqrt{\prn{\sum_j \widetilde{N}^{A\mu}_{ij} -\beta}^2 +4\beta\widetilde{N}^{A\mu}_{ii}}}
+\end{multline}
+%
+Which produces the update rule
+%
+\begin{equation}\label{eq:L1update}
+\begin{aligned}
+  \M'^\mu_{ij} &= \frac{ 2\widetilde{N}^{A\mu}_{ij} }{ \prn{\sum_k \widetilde{N}^{A\mu}_{ik} +\beta} +
+     \sqrt{\prn{\sum_k \widetilde{N}^{A\mu}_{ik} -\beta}^2 +4\beta\widetilde{N}^{A\mu}_{ii}} } 
+     \qquad \text{for } i\neq j, \\
+  \M'^\mu_{ii} &= \frac{ 2\widetilde{N}^{A\mu}_{ii} }{ \prn{\sum_k \widetilde{N}^{A\mu}_{ik} -\beta} +
+     \sqrt{\prn{\sum_k \widetilde{N}^{A\mu}_{ik} -\beta}^2 +4\beta\widetilde{N}^{A\mu}_{ii}} }.
+\end{aligned}
+\end{equation}
+%
+
+\subsection{$\CL^\half$ penalty \label{sec:Lhalf}}
+
+Here we use the  $\CL^\half$ norm as a penalty:
+%
+\begin{equation}\label{eq:Lhalfpenalty}
+  E^\mu_{ij} = 2\sqrt{\M^\mu_{ij}}.
+\end{equation}
+%
+Then the condition for a maximum \eqref{eq:priordiffs} reads
+%
+\begin{equation}\label{eq:Lhalfpriormax}
+  \frac{ \widetilde{N}^{A\mu}_{ij} }{ \M'^\mu_{ij} }  - \frac{\beta}{\sqrt{\M'^\mu_{ij}}} - \lambda^\mu_{i} = 0.
+\end{equation}
+%
+It is helpful to define
+%
+\begin{equation}\label{eq:Lhalfvars}
+\begin{aligned}
+  \widetilde{N}^{A\mu}_{ij} &= \beta^2 L^\mu_{ij}, \\
+  \lambda^\mu_i &= \prn{\gamma^\mu_i}^{-2}.
+\end{aligned}
+\end{equation}
+%
+Then we have the update rule
+%
+\begin{equation}\label{eq:Lhalfupdate}
+  \M'^\mu_{ij} = \frac{ \prn{\gamma^\mu_i}^2 \beta^2 }{ 4 } \prn{\sqrt{ \prn{\gamma^\mu_i}^2 + 4\prn{L^\mu_{ij}}^2 } - \gamma^\mu_i }^2, 
+\end{equation}
+%
+with $\gamma^\mu_i$ determined by:
+%
+\begin{equation}\label{eq:Lhalfnorm}
+    \frac{ \prn{\gamma^\mu_i}^2 \beta^2 }{ 4 } \sum_j \prn{\sqrt{\prn{\gamma^\mu_i}^2 + 4\prn{L^\mu_{ij}}^2} - \gamma^\mu_i}^2 = 1.
+\end{equation}
+%
 

### 2014-07-04 Notes/Identification.tex
notation changes

@@
 We denote a sequence of event times by $[0,T]$.
+Any sequence whose upper bound is smaller that its lower bound can be considered empty, and that variable is omitted.
+The same applies to sums and products.
 
-We denote states of a synapse by subscripts $i,j$. 
+We denote states of a synapse by subscripts $i,j,k$. 
 The state at a particular time during a sequence is $i(t)$.
@@
 The different types of plasticity (\eg potentiation and depression) are indicated by superscripts $\mu,\nu$.
-The type of plasticity used at time $t$ is $\mu(t)$.
-An entire sequence of plasticity types is denoted by $\mu[0,T]$.
+The type of plasticity used between times $(t-1,t)$ is $\mu(t)$.
+An entire sequence of plasticity types is denoted by $\mu[1,T]$.
 
@@
 %
-\begin{multline}\label{eq:emqseveral}
-%\begin{aligned}
-  Q(\M,\M') = \sum_{i^A[0,T]} \brk{ \prod_a \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } \brk{\sum_a  \log\!\!  \brk{
-      \frac{ \pr\cond{i^A[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^A[0,T],\w^a[0,T]}{\M} } } }\\
-   = \sum_a \prod_{b \neq a} \pr\cond{\w^b[0,T]}{\M} \sum_{i^a[0,T]} \pr\cond{i^a[0,T],\w^a[0,T]}{\M} \log\!\!  \brk{
+\begin{equation}\label{eq:emqseveral}
+\begin{gathered}
+  Q(\M,\M') = \sum_{i^A[0,T]}  \prod_b \pr\cond{i^b[0,T],\w^b[0,T]}{\M} \sum_a  \log\!\!  \brk{
+      \frac{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } } \\
+   = \sum_a \prod_{b \neq a} \pr\cond{\w^b[0,T]}{\M}\! \sum_{i^a[0,T]}\!\! \pr\cond{i^a[0,T],\w^a[0,T]}{\M} \log\!\!  \brk{
       \frac{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } } \\
    = \pr\cond{\w^A[0,T]}{\M} \sum_a \frac{Q^a(\M,\M')}{\pr\cond{\w^a[0,T]}{\M}}.
-%\end{aligned}
-\end{multline}
+\end{gathered}
+\end{equation}
 %
@@
 \begin{aligned}
-  \pr\cond{t[0,T], \w[0,T]}{\M} &= \pib_{i(0)} \, \Pib(0) 
-        \prod_{t=0}^{T-1} \M^{\mu(t)}_{i(t)i(t+1)} \, \Pib(t+1),\\
+  \pr\cond{t[0,T], \w[0,T]}{\M} &= \pib_{i(0)} \, \Pib_{i(0)}(0) 
+        \prod_{t=1}^{T} \M^{\mu(t)}_{i(t-1)i(t)} \, \Pib_{i(t)}(t),\\
   \text{where}\quad
-  \Pib(t) &= \delta_{\w_{i(t)}\w(t)}
+  \Pib_i(t) &= \delta_{\w_{i}\w(t)}
 \end{aligned}
@@
 \begin{aligned}
-  N^\nu_{i(t)i(t+1)} &= \sum_{t=0}^{T-1} \pr\cond{i[t,t+1]),\w[0,T]}{\M} \\
-     &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, \alpha_i(t) \, \beta_j(t+1) \, \M^{\nu}_{i(t)i(t+1)} \, \Pib(t+1) ,\\
-  \alpha_{i(t)}(t) &= \pr\cond{i(t),\w[0,t]}{\M} \\
-     &= \sum_{i[0,t-1]} \pib_{i(0)} \, \delta_{\w_{i(0)}\w(0)} \prod_{s=0}^{t-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib(s+1), \\
-  \beta_{i(t)}(t) &= \pr\cond{i(t),\w[t+1,T]}{\M} \\
-     &= \sum_{i[t+1,T]} \prod_{s=t}^{T-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib(s+1).
+  N^\nu_{jk} &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \pr\cond{i[t-1,t]=(jk),\w[0,T]}{\M} \\
+     &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \alpha_j(t-1) \, \beta_k(t) \, \M^{\nu}_{jk} \, \Pib_k(t) ,\\
+  \alpha_j(t) &= \pr\cond{i(t)=j,\w[0,t]}{\M} \\
+     &= \sum_{i[0,t]} \pib_{i(0)}(0) \, \Pib_{i(0)}(0) \prod_{s=1}^{t} \M^{\mu(s)}_{i(s-1)i(s)} \, \Pib_{i(s)}(s) \,\delta_{i(t)j}, \\
+  \beta_j(t) &= \pr\cond{i(t)=j,\w[t+1,T]}{\M} \\
+     &= \sum_{i[t,T]} \delta_{i(t)j} \prod_{s=t+1}^{T} \M^{\mu(s)}_{i(s-1)i(s)} \, \Pib_{i(s)}(s).
 \end{aligned}
@@
 \begin{equation}\label{eq:stateprob}
-  \pr\cond{i(t),\w[0,T]}{\M} = \alpha_{i(t)}(t) \, \beta_{i(t)}(t).
+  \alpha_j(t) \, \beta_j(t) = \pr\cond{i(t)=j,\w[0,T]}{\M}.
 \end{equation}
@@
 These quantities usually suffer from underflow.
-This can be avoided by using normalised forwar/backward variables \cite{Zhai2003HMMnorm}:
+This can be avoided by using normalised forward/backward variables \cite{Zhai2003HMMnorm}:
 %
@@
 \begin{aligned}
-  \tilde{\alpha}_{i(t)}(t) &= \pr\cond{i(t)}{\w[0,t],\M}, \\
+  \tilde{\alpha}_j(t) &= \pr\cond{i(t)=j}{\w[0,t],\M}, \\
   \eta(t) &= \pr\cond{\w(t)}{\w[0,t-1],\M}, \\
@@
 \begin{aligned}
-  \widetilde{N}^\nu_{i(t)i(t+1)} &= \sum_{t=0}^{T-1} \pr\cond{i[t,t+1])}{\w[0,T],\M} \\
-     &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, \tilde{\alpha}_i(t) \, \tilde{\beta}_j(t+1) \, \eta(t+1) \, \M^{\nu}_{i(t)i(t+1)} \, \Pib(t+1) ,\\
-  \pr\cond{i(t)}{\w[0,T],\M} &= \alpha_{i(t)}(t) \, \beta_{i(t)}(t).
+  \widetilde{N}^\nu_{jk} &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \pr\cond{i[t-1,t]=(jk)}{\w[0,T],\M} \\
+     &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \tilde{\alpha}_i(t-1) \, \tilde{\beta}_j(t) \, \eta(t) \, \M^{\nu}_{i(t-1)i(t)} \, \Pib_k(t) ,\\
+  \tilde{\alpha}_j(t) \, \tilde{\beta}_j(t) &= \pr\cond{i(t)=j}{\w[0,T],\M}.
 \end{aligned}
@@
 
-\subsection{Off-diagonal $\CL^1$ penalty \label{sec:offdiagL1}}
+\subsection{Off-diagonal \texorpdfstring{$\CL^1$}{L\textasciicircum1} penalty \label{sec:offdiagL1}}
 
@@
 
-\subsection{$\CL^\half$ penalty \label{sec:Lhalf}}
+\subsection{\texorpdfstring{$\CL^\half$}{L\textasciicircum1/2} penalty \label{sec:Lhalf}}
 
@@
 \begin{equation}\label{eq:Lhalfupdate}
-  \M'^\mu_{ij} = \frac{ \prn{\gamma^\mu_i}^2 \beta^2 }{ 4 } \prn{\sqrt{ \prn{\gamma^\mu_i}^2 + 4\prn{L^\mu_{ij}}^2 } - \gamma^\mu_i }^2, 
+  \M'^\mu_{ij} = \frac{ \prn{\gamma^\mu_i}^2 \beta^2 }{ 4 } \prn{\sqrt{ \prn{\gamma^\mu_i}^2 + 4L^\mu_{ij} } - \gamma^\mu_i }^2,
 \end{equation}
@@
 \begin{equation}\label{eq:Lhalfnorm}
-    \frac{ \prn{\gamma^\mu_i}^2 \beta^2 }{ 4 } \sum_j \prn{\sqrt{\prn{\gamma^\mu_i}^2 + 4\prn{L^\mu_{ij}}^2} - \gamma^\mu_i}^2 = 1.
+    \frac{ \prn{\gamma^\mu_i}^2 \beta^2 }{ 4 } \sum_j \prn{\sqrt{\prn{\gamma^\mu_i}^2 + 4L^\mu_{ij}} - \gamma^\mu_i}^2 = 1.
 \end{equation}

### 2014-08-22 Code/Identification/MultiExp/MultiExpGenFun.m
updated MHC supp pagination

@@ -1,0 +1,11 @@
+function [ genfun ] = MultiExpGenFun( s,data )
+%genfun=MULTIEXPGENFUN(s,data) estimate of moment generating function
+%   genfun = <exp(-st)>
+%   s      = column of s values
+%   data   = row of ssamples of t values
+
+genfun = sum(exp(-s*data),2);
+
+
+end
+

### 2014-08-22 Code/Identification/MultiExp/MultiExpGenFunAC.m

@@ -1,0 +1,32 @@
+function [ genfun ] = MultiExpGenFunAC( s,data,varargin )
+%genfun=MULTIEXPGENFUNAC(s,data) analytic continuation of estimate of
+%moment generating function
+%   genfun = <exp(-st)> for Re(s)>0
+%   s      = column of s values
+%   data   = row of ssamples of t values
+
+persistent p
+if isempty(p)
+    p=inputParser;
+    p.FunctionName='MultiExpGenFunAC';
+    p.StructExpand=true;
+    p.KeepUnmatched=false;
+    p.addParameter('R',100);
+    p.addParameter('ds',0.1);
+    p.addParameter('eps',1);
+end
+p.parse(varargin{:});
+r=p.Results;
+
+
+s_cont = (-r.R:r.ds:r.R)*1i+r.eps;
+
+genfun_cont = MultiExpGenFun(s_cont',data);
+
+kernel = 1./(bsxfun(@minus,s_cont,s));
+
+genfun = kernel*genfun_cont*r.ds/(2*pi);
+
+
+end
+

### 2015-02-03 Code/Identification/Fitters/FitSynapseSizeLump.m
attempted spectra linitialisation

@@ -25,18 +25,20 @@ w=zeros(numWvals*options.MaxStates,1);
 w(1:options.MaxStates:end)=1;
 w=cumsum(w);
 
-fitmodel=SynapseIdModel.Rand(w,'NumPlastTypes',NumPlastTypes,p.Unmatched);
-testloglike=HMMloglike(fitmodel,simobjs)+SynapsePrior(fitmodel,options);
-for i=1:options.NumReps
-    DispReps(i);
-    guessmodel=SynapseIdModel.Rand(w,'NumPlastTypes',NumPlastTypes,p.Unmatched);
-    [guessmodel,guessloglike]=FitSynapse(simobjs,guessmodel,options);
-    if guessloglike > testloglike
-        fitmodel=guessmodel;
-        testloglike=guessloglike;
-    end%if guessloglike
-end%for i
-DispReps(i+1);
+% fitmodel=SynapseIdModel.Rand(w,'NumPlastTypes',NumPlastTypes,p.Unmatched);
+% testloglike=HMMloglike(fitmodel,simobjs)+SynapsePrior(fitmodel,options);
+% for i=1:options.NumReps
+%     DispReps(i);
+%     guessmodel=SynapseIdModel.Rand(w,'NumPlastTypes',NumPlastTypes,p.Unmatched);
+%     [guessmodel,guessloglike]=FitSynapse(simobjs,guessmodel,options);
+    guessmodel=SpectralInit(simobjs,w,'NumPlastTypes',NumPlastTypes,p.Unmatched);
+    [fitmodel]=FitSynapse(simobjs,guessmodel,options);
+%     if guessloglike > testloglike
+%         fitmodel=guessmodel;
+%         testloglike=guessloglike;
+%     end%if guessloglike
+% end%for i
+% DispReps(i+1);
 
 partitions=fitmodel.FindLumps;
 if fitmodel.TestLump(partitions);

### 2015-02-03 Code/Identification/Spectral/CalcObsProbs.m

@@ -1,0 +1,56 @@
+function [ P1,P2,P3 ] = CalcObsProbs( simobj )
+%CALCOBSPROBS(simobj) Summary of this function goes here
+%   Detailed explanation goes here
+
+
+P1=hist([simobj.readouts], 1:simobj.NumWvals);
+P1=P1/sum(P1);
+
+emptymat=zeros(simobj.NumWvals);
+emptyarr=zeros(simobj.NumWvals*[1 1 1]);
+
+P2=cell(1,simobj.NumPlast);
+P3=cell(simobj.NumPlast,simobj.NumPlast);
+
+for a=1:length(P2)
+    P2{a}=emptymat;
+    for b=1:length(P3)
+        P3{a,b}=emptyarr;
+    end
+end
+
+steps=[];
+steptypes=[];
+
+dsteps=[];
+dsteptypes=[];
+
+for i=1:numel(simobj)
+
+    stepstart=simobj(i).readouts(1:end-1);
+    stepend=simobj(i).readouts(2:end);
+    steps=[steps sub2ind(size(emptymat),stepstart,stepend)];
+    steptypes=[steptypes simobj(i).potdep(1:end-1)];
+
+    stepstart=stepstart(1:end-1);
+    stepmid=stepend(1:end-1);
+    stepend=stepend(2:end);
+    dsteps=[dsteps sub2ind(size(emptyarr),stepstart,stepmid,stepend)];
+    dsteptypes=[dsteptypes [simobj(i).potdep(1:end-2); simobj(i).potdep(2:end-1)] ];
+
+end
+
+for a=1:length(P2)
+    P2{a}(1:numel(emptymat))=hist(steps(steptypes==a),1:numel(emptymat));
+    P2{a}=P2{a}/sum(P2{a}(:));
+    for b=1:length(P3)
+        P3{a,b}(1:numel(emptyarr))=hist(dsteps(dsteptypes(1,:)==a & dsteptypes(2,:)==b),1:numel(emptyarr));
+        P3{a,b}=P3{a,b}/sum(P3{a,b}(:));
+    end
+end
+
+
+
+
+end
+

### 2015-02-03 Code/Identification/Spectral/MakeObs.m

@@ -1,0 +1,13 @@
+function [ Obs ] = MakeObs( modelobj )
+%Obs=MAKEOBS(modelobj) construct observation matrix
+%   Obs(i,x) = probability of observing x modelobj.when in state i
+
+Obs=zeros(modelobj.NumStates,modelobj.NumWvals);
+
+inds=sub2ind(size(Obs),1:modelobj.NumStates,modelobj.w');
+
+Obs(inds)=1;
+
+
+end
+

### 2015-02-03 Code/Identification/Weighters/Sequentialweight.m

@@ -1,0 +1,48 @@
+function [ newmodelobj,loglike,pstate ] = Sequentialweight( modelobj,simobj,varargin )
+%[newmodelobj,loglike,pstate]=SEQUENTIAlWEIGHT(modelobj,simobj) Sequential update of estimated HMM
+%   newmodelobj = updated SynapseIdModel
+%   loglike     = log likelihood of readouts under old model (prod over simobj)
+%   pstate      = posterior prob of HMM being in each state at each time (cell, one element for each simobj)
+%   modelobj = SynapseIdModel
+%   simobj   = vector of SynapsePlastSeq
+
+% lloffset=0;%avoid overflow by making this more positive
+persistent p
+if isempty(p)
+    p=inputParser;
+    p.FunctionName='Mackayweight';
+    p.StructExpand=true;
+    p.KeepUnmatched=true;
+    p.addParameter('Algorithm','BW');
+    p.addParameter('RecurrentFrac',0.1);
+    p.addParameter('Normalise',true);
+%     p.addParameter('Normalise',true,@(x) validateattributes(x,{'logical'},{'scalar'}));
+end
+p.parse(varargin{:});
+
+updater=str2func([p.Results.Algorithm 'update']);
+
+
+
+% error(CheckSize(simobj,@isvector));
+
+pstate=cell(1,length(simobj));
+newmodelobj=modelobj;
+loglike=0;
+
+for i=1:length(simobj)
+    [chunkModel,chunkll,chunkPs]=updater(newmodelobj,simobj(i),'Normalise',false,p.Unmatched);
+    newmodelobj=newmodelobj+chunkModel +...
+        max(cellfun(@(x)max(max(x)),chunkModel.M))*p.Results.RecurrentFrac*SynapseIdModel.Rand(newmodelobj.w);
+    newmodelobj=newmodelobj.Normalise;
+    pstate{i}=chunkPs*diag(1./sum(chunkPs,1));
+    loglike=loglike+chunkll;
+end
+    
+if p.Results.Normalise
+    newmodelobj=newmodelobj.Normalise;
+    assert(newmodelobj.isvalid,'newmodelobj is invalid');
+end
+
+end
+

### 2015-05-27 Notes/laplace.tex
laplace envelope notes

--- 2013-11-09 Notes/ContTime.tex
@@
 %    Comment out showkeys
-%    Replace \input{mydefs.tex} with its contents
-%       or include mydefs.tex in zip/tar file
-%    Replace \input{newsymb.tex} with its contents
-%       or include newsymb.tex in zip/tar file
+%    Replace \input{?.tex} with its contents
+%       or include ?.tex in zip/tar file
 %    Put this file, the .bbl file, any picture or
@@
 \newcommand{\F}{\boldsymbol{\Phi}}
+\newcommand{\syn}{\vec{w}}
+\newcommand{\synid}{\syn_\text{ideal}}
 \DeclareMathOperator{\SNR}{SNR}
 \DeclareMathOperator{\snr}{SNR}
+\DeclareMathOperator{\env}{Env}
+\newcommand{\rh}{\hat{r}}
 \newcommand{\CI}{\mathcal{I}}
@@
 % Title info:
-\title{Area maximisation in continuous time}
+\title{Laplacian envelope}
 %
@@
 \begin{abstract}
-  We try to find the continuous time Markov process that has the maximal area under the signal-to-noise curve.
+  We try to find the continuous time Markov process that has the maximal Laplace tranformed signal-to-noise curve.
 \end{abstract}
@@
 
-\section{Continuous time Markov processes}\label{sec:ContMarkov}
+\section{Recognition memory}\label{sec:recog}
 
-In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time.
-It is a straightforward generalisation of material that can be found in \cite{kemeny1960finite} with some ideas from \cite{hunter2000survey}.
+We will be trying to store patterns in a set od $N$ synaptic weights, $\syn$.
+Every time we try to store a pattern, these synapses are subjected to a plasticity event where each synapse is either potentiated or depressed, depending on the pattern.
+we will assume that these patterns are spatially and temporally independent.
 
-\subsection{Notation}\label{sec:not}
+At some time, suppose we wish to determine if a given pattern is one of those that we previously attempted to store.
+We wish to answer this question by looking at the synaptic weights directly (ideal observer).
+For that given pattern there will be an ideal set of synaptic weights, $\synid$, where those synapses that were supposed to be potentiated are maximised and those that were supposed to be depressed are minimised.
+Suppose that the given pattern was actually seen at time 0 and we are observing the synapses at time $t$.
+The actual set of synaptic weights we see, $\syn(t)$, will be a vector of random variables that differs from $\synid$ due to the stochasticity of the pattern encoding and all of the other (uncorrelated) pattern that are stored after it.
+As $t\to\infty$, the synaptic weights will become independent of the patter stored at $t=0$.
+Thus, the vector of random variables $\syn(\infty)$ also describes the synaptic weights under the null hypothesis -- if that given pattern had never been stored.
 
-For any matrix $\mathbf{A}$, we define matrices $\mathbf{A}\dg$ and $\overline{\mathbf{A}}$ as
+We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold.
+For large $N$, this quantity will have a Gausssian distribution.
+There will be a ROC curve as a function of this threshold:
 %
-\begin{equation}\label{eq:dgdef}
-  \mathbf{A}\dg_{ij} \equiv \delta_{ij}\mathbf{A}_{ij},
-  \qquad
-  \overline{\mathbf{A}} \equiv \mathbf{A}-\mathbf{A}\dg.
-\end{equation}
-%
-We let $\onev$ denote a column-vector of ones and $\onem=\onev\onev\trans$ denote a matrix of ones.
-
-A Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$.
-The probabilities of being in each state at time $t$, the row-vector $\mathbf{p}(t)$, evolve according to
-%
-\begin{equation}\label{eq:rowsum}
-  \diff{\mathbf{p}(t)}{t} = \mathbf{p}(t) \MM,
-  \qquad
-  \mathbf{p}(t)\onev=1,
-  \qquad
-  \MM\onev=0.
-\end{equation}
-%
-In Surya's notes, $\MM=r\frg$.
-
-The equilibrium probabilities, $\eq$, satisfy
-%
-\begin{equation}\label{eq:equilibrium}
-  \eq\MM=0,
-  \qquad
-  \eq\onev=1.
-\end{equation}
-%
-As we assume an ergodic process, this eigenvalue is non-degenerate.
-If all other eigenvalues have strictly negative real parts, the process is regular (aperiodic).
-
-We define additional matrices
-%
-\begin{equation}\label{eq:defDLP}
-  \Lb \equiv (-\MM\dg)\inv,
-  \qquad
-  \Pb \equiv \I + \Lb\MM.
-\end{equation}
-%
-It can be shown that $\Lb_{ii}$ is the mean time it takes to leave state $i$ and $\Pb_{ij}$ is the probability the the next transition from state $i$ goes to state $j$:
-%
-\begin{equation}\label{eq:LamdaPcmpt}
-  \Lb_{ii} = \frac{1}{\sum_{j \neq j} \MM_{ij}},
-  \qquad
-  \Pb_{ij} = %\left\{
-%  \begin{aligned}
-%     &0                                         && \text{if }i=j, \\
-%     &\frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &&\text{otherwise}.
-%  \end{aligned}
-%  \right.
-  \begin{cases}
-     0                                         &\text{if }i=j, \\
-     \frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &\text{otherwise}.
-  \end{cases}
-\end{equation}
-%
-Furthermore, we also define
-%
-\begin{equation}\label{eq:pdotD}
-  \D \equiv \operatorname{diag}(\eq)\inv,
-  \qquad\implies\qquad
-  \eq\D=\onev\trans.
-\end{equation}
-%
-
-\subsection{Fundamental matrix}\label{sec:fund}
-
-\begin{defn}[Fundamental matrix]
-  %
-  \begin{equation}\label{eq:funddef}
-    \fund \equiv (-\MM + \onev\pib)\inv,
-  \end{equation}
-  %
-  where $\pib$ is any row-vector with $\pib\onev=1/\tau\neq0$.
-\end{defn}
-Note that the canonical choice for the discrete time version, $\pib=\eq$, is not available here due to problems with units.
-It will be helpful to choose $\pib$ to be independent of $\MM$, \eg $\pib=\onev\trans/(n\tau)$.
-All quantities that we calculate using $\fund$ below will be independent of this choice.
-
-\begin{thm}
-  The definition of $\fund$ is valid, \ie $(-\MM + \onev\pib)$ is invertible.
-\end{thm}
-\begin{proof}
-  Assume there exists an $\mathbf{x}$ such that
-  %
-  \begin{equation}\label{eq:fundinvkern}
-    (-\MM + \onev\pib)\mathbf{x}=0.
-  \end{equation}
-  %
-  Multiplying from the left with $\eq$ gives
-  %
-  \begin{equation}\label{eq:fundinvpix}
-    \pib \mathbf{x} = 0.
-  \end{equation}
-  %
-  Substituting back into \eqref{eq:fundinvkern} gives
-  %
-  \begin{equation*}
-    \MM\mathbf{x}=0.
-  \end{equation*}
-  %
-  As we assume an ergodic process, the zero eigenvalue is non-degenerate.
-  Therefore, $\mathbf{x}=\lambda\onev$.
-  Substituting this into \eqref{eq:fundinvpix} gives
-  %
-  \begin{equation*}
-    \lambda\pib\onev = \frac{\lambda}{\tau} = 0.
-  \end{equation*}
-  %
-  As we defined $\pib$ such that $1/\tau\neq0$, this means $\lambda=0 \implies \mathbf{x}=0$.
-\end{proof}
-
-\begin{cor}
-  %
-  \begin{align}
-    \pib\fund &= \eq, \label{eq:fundprob}\\
-    \fund\onev &= \tau\onev,\label{eq:fundrowsum}\\
-    \I+\MM\fund &= \onev\eq, \label{eq:fundQZ}\\
-    \I+\fund\MM &= \tau\onev\pib. \label{eq:fundZQ}
-  \end{align}
-  %
-\end{cor}
-\begin{proof}
-  We can deduce \eqref{eq:fundprob} and \eqref{eq:fundrowsum} be pre/post-multiplying the following equations by $\fund$:
-  %
-  \begin{equation*}
+\begin{equation}\label{eq:ROC}
   \begin{aligned}
-      \eq(-\MM + \onev\pib) &= \pib, \\
-      (-\MM + \onev\pib)\onev &= \frac{\onev}{\tau}.
-    \end{aligned}
-  \end{equation*}
-  %
-  We can then deduce \eqref{eq:fundQZ} and \eqref{eq:fundZQ} by substituting these into
-  %
-  \begin{equation*}
-    (-\MM + \onev\pib)\fund = \fund(-\MM + \onev\pib) = \I.
-  \end{equation*}
-  %
-\end{proof}
-
-%We will not need to use the concept of a g-inverse anywhere, so you can skip ahead to \sref{sec:fpt} now. For the sake of completeness, we can define $\MM^\#$ as
-%%
-%\begin{equation}\label{eq:ginvdef}
-%  \MM^\# \equiv \tau\onev\eq - \fund
-%\end{equation}
-%%
-%\begin{thm}
-%  The matrix $\MM^\#$ satisfies the following defining properties of a g-inverse of $\MM$:
-%  %
-%  \begin{align}
-%    \MM^\# \MM \MM^\# &= \MM^\#,     \label{eq:ginvpropo} \\
-%    \MM \MM^\# \MM    &= \MM,        \label{eq:ginvpropi} \\
-%    \MM \MM^\#        &= \MM^\# \MM, \label{eq:ginvcomm}
-%  \end{align}
-%  %
-%  with \eqref{eq:ginvcomm} holding iff $\eq$ and $\pib$ are proportional to each other.
-%\end{thm}
-%\begin{proof}
-%  Using \eqref{eq:rowsum}, \eqref{eq:equilibrium}, \eqref{eq:fundprob} and \eqref{eq:fundrowsum}, one can verify
-%  %
-%  \begin{alignat*}{2}
-%    \MM \MM^\# &= -\MM \fund,
-%    &\qquad
-%    \pib \MM^\# &= 0,
-%    \\
-%    \MM^\# \MM &= -\fund \MM,
-%    &
-%    \MM^\# \onev &= 0.
-%  \end{alignat*}
-%  %
-%  Substituting into \eqref{eq:fundQZ} and \eqref{eq:fundZQ} leads to
-%  %
-%  \begin{equation*}
-%    \I - \MM \MM^\# = \onev \eq,
-%    \qquad
-%    \I - \MM^\# \MM = \tau \onev \pib.
-%  \end{equation*}
-%  %
-%  Pre/post-multiplying these by $\MM$ or $\MM^\#$ produces \eqref{eq:ginvpropo} and \eqref{eq:ginvpropi}. Subtracting produces
-%  %
-%  \begin{equation*}
-%    [\MM,\MM^\#] = \onev(\tau\pib-\eq),
-%  \end{equation*}
-%  %
-%  which vanishes iff $\eq$ and $\pib$ are proportional to each other (due to $\pib\onev=1/\tau$ and $\eq\onev=1$, the constant of proportionality could only be $\tau$).
-%\end{proof}
-
-\subsection{First passage times}\label{sec:fpt}
-
-\begin{defn}[First passage time matrix]
- We define $\fptb_{ij}$ as the mean time it takes the process to reach state $j$ for the first time, starting from state $i$.
- We also define $\fpt\dg_{ii}$ as the mean time it takes the process to return to state $i$.
- As usual, $\fpt=\fptb+\fpt\dg$.
-\end{defn}
-
-Consider the first time the process leaves state $i$.
-On average, this will take time $\Lb_{ii}$.
-With probability $\Pb_{ij}$, it will go directly to $j$, so the conditional mean time would be $\Lb_{ii}$.
-On the other hand, if it goes to some other state, $k$, with probability $\Pb_{ik}$, the conditional mean time would be $\fptb_{kj}+\Lb_{ii}$.
-Combining these, we get the recursion relation
-%
-\begin{equation}\label{eq:fptrecursion}
-  \begin{aligned}
-  \fpt_{ij} &= \sum_{k\neq j} \Pb_{ik}(\fptb_{kj}+\Lb_{ii}) + \Pb_{ij}\Lb_{ii} \\
-    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \sum_{k} \Pb_{ik}\Lb_{ii}\\
-    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \Lb_{ii},\\
-   \fpt &= \Pb \fptb + \Lb\onem,
+  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \operatorname{NNR} } },
+  \quad\text{where } &
+    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
+    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
+    \operatorname{NNR} &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
   \end{aligned}
@@
 %
-or, using \eqref{eq:defDLP},
+and TPR/FPR are the true/false positive rates.
+When the signal-to-noise ratio, $\snr$ is larger than $\Phi_c\inv(\operatorname{FPR})$, it is beneficial to decrease the noise-to-noise ratio, NNR.
+In the other case, it is beneficial to increase it.
+The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events and the probability distribution of the sequence of plasticity events.
+
+The formulae above assumed that we know the time between storage and recognition.
+We should also compute the expectations and variances over probability distribution of the recall time, $t$, as well.
+If we only know average time, $\tau$, a natural choice for this distribution is
 %
-\begin{equation}\label{eq:fptrecdg}
-  \fpt\dg = \Lb\MM\fptb+\Lb\onem.
+\begin{equation}\label{eq:recogtime}
+  P(t\vert\tau) = \frac{\e^{-t/\tau}}{\tau}.
 \end{equation}
 %
-
-\begin{thm}
-  The recurrence times are given by
-  %
-  \begin{equation}\label{eq:recurtime}
-    \fpt\dg = \Lb\D.
-  \end{equation}
-  %
-\end{thm}
-\begin{proof}
-  Pre-multiply \eqref{eq:fptrecdg} by $\eq\Lb\inv$:
-  %
-  \begin{equation*}
-    \eq\Lb\inv\fpt\dg = \onev\trans,
-  \end{equation*}
-  %
-  or in component form
-  %
-  \begin{equation*}
-    \eq_i \Lb_{ii}\inv \fpt\dg_{ii} = 1.
-  \end{equation*}
-  %
-  This is the same as \eqref{eq:recurtime} in component form: $\fpt\dg_{ii} = \Lb_{ii}/\eq_i$.
-\end{proof}
-I think the extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.
-
-We can substitute this into \eqref{eq:fptrecdg} to get a recursion relation for the off diagonal part:
-%
-\begin{equation}\label{eq:fptbrec}
-  \MM\fptb = \D-\onem.
-\end{equation}
-%
-\begin{thm}
-  The off-diagonal solution of \eqref{eq:fptbrec} is unique.
-\end{thm}
-\begin{proof}
-  Suppose we have two solutions for the off-diagonal part, $\fptb$ and $\fptb'$.
-  Subtracting \eqref{eq:fptbrec} for these two solutions gives
-  %
-  \begin{equation*}
-    \MM(\fptb-\fptb')=0.
-  \end{equation*}
-  %
-  As we are assuming an ergodic process, the zero eigenvalue is non-degenerate.
-  Therefore
-  %
-  \begin{equation*}
-    \fptb-\fptb' = \onev\mathbf{y},
-  \end{equation*}
-  %
-  for some row-vector $\mathbf{y}$. Looking at the diagonal components
-  %
-  \begin{equation*}
-    \mathbf{y}_{i} = \fptb_{ii}-\fptb'_{ii} = 0,
-  \end{equation*}
-  %
-  and therefore $\fptb-\fptb'=0$.
-\end{proof}
-
-\begin{thm}
-  The off-diagonal mean first passage times are given by
-  %
-  \begin{equation}\label{eq:fptfund}
-    \fptb = (\onem\fund\dg - \fund)\D.
-  \end{equation}
-  %
-\end{thm}
-\begin{proof}
-  We just need to show that this satisfies \eqref{eq:fptbrec} and that the diagonal elements vanish.
-  Premultiply \eqref{eq:fptfund} by $\MM$:
-  %
-  \begin{equation*}
-  \begin{aligned}
-    \MM\fptb &= -\MM\fund\D\\
-     &= (\I-\onev\eq)\D \\
-     &= \D-\onem,
-  \end{aligned}
-  \end{equation*}
-  %
-  where we used \eqref{eq:fundQZ} and \eqref{eq:pdotD}.
-  The vanishing of the diagonal elements follows trivially from the component form of \eqref{eq:fptfund}:
-%
-\begin{equation}\label{eq:fptfundcmpt}
-  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
-\end{equation}
-%
-
-\end{proof}
-
-\subsection{Mixing time (Kemeny's constant)}\label{sec:mixtime}
-
-\begin{thm}\label{th:kemenyconst}
-  The quantity
-  %
-  \begin{equation}\label{eq:mixdef}
-    \eta \equiv \sum_j \fptb_{ij}\eq_j
-  \end{equation}
-  %
-  is independent of $i$.
-\end{thm}
-\begin{proof}
-  We use \eqref{eq:fptfund}, \eqref{eq:fundrowsum} and the transpose of \eqref{eq:pdotD}:
-  %
-  \begin{equation*}
-    \begin{aligned}
-      \fptb (\eq)\trans &= (\onem\fund\dg - \fund) \D (\eq)\trans \\
-        &= (\onev\onev\trans\fund\dg - \fund) \onev \\
-        &= (\onev\trans\fund\dg\onev)\onev - \fund\onev \\
-        &= (\tr\fund - \tau)\onev.
-    \end{aligned}
-  \end{equation*}
-  %
-  which proves \eqref{eq:mixdef} with $\eta = \tr\fund - \tau$.
-\end{proof}
-
-Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version \cite{hunter2006mixing} where this would only shift $\eta$ by 1.
-
-\subsection{Sensitivity of equilibrium distribution}\label{sec:sensitivity}
-
-Suppose that the Markov process, defined by $\MM$, depends on some parameter $\alpha$.
-Differentiating \eqref{eq:funddef} gives
-%
-\begin{equation}\label{eq:diffZ}
-  \diff{\fund}{\alpha} = \fund \diff{\MM}{\alpha} \fund.
-\end{equation}
-%
-We can substitute this into the derivative of \eqref{eq:fundprob}:
-%
-\begin{equation}\label{eq:diffp}
-  \diff{\eq}{\alpha} = \pib \fund \diff{\MM}{\alpha} \fund = \eq \diff{\MM}{\alpha} \fund.
-\end{equation}
-%
-We can rewrite this in component form and use the fact that $\MM_{ii} = - \sum_{i\neq j} \MM_{ij}$:
-%
-\begin{equation}\label{eq:diffpT}
-\begin{aligned}
-  \diff{\eq_k}{\alpha} &= \sum_{i,j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} \\
-    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} + \sum_i \eq_i \diff{\MM_{ii}}{\alpha} \fund_{ik} \\
-    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} (\fund_{jk} - \fund_{ik}) \\
-    &= \sum_{i\neq j} \diff{\MM_{ij}}{\alpha} \eq_i \eq_k (\fptb_{ik} - \fptb_{jk}),
-\end{aligned}
-\end{equation}
-%
-which is the result of \cite{cho2000markov} that we need.
-Note that the summand vanishes for $i=j$, so we can drop the restriction $i\neq j$ from the range of the sum.
-
-\subsection{Subsets and flux}\label{sec:subsets}
-
-Let us denote the set of states by $\CS$.
-Consider a subset $\CA\subset\CS$.
-We can define a projection operator onto this subset:
-%
-\begin{equation}\label{eq:proj}
-  \prn{\I^\CA}_{ij} =
-    \begin{cases}
-      1 &\text{if $i=j\in\CA$,}\\
-      0 &\text{otherwise.}
-    \end{cases}
-\end{equation}
-%
-We will use superscripts/subscripts to denote projection onto/summation over a subset:
-%
-\begin{equation}\label{eq:projsum}
-  \begin{aligned}
-    \pib^\CA &=\pib\I^\CA, &
-    \M^{\cdot\CA}&=\M\I^\CA, &
-    \M^{\CA\cdot}&=\I^\CA\M, &
-    \mathbf{x}^\CA &= \I^\CA\mathbf{x},
-    \\
-    \pib_\CA &=\pib\onev^\CA, &
-    \M_{\cdot\CA}&=\M\onev^\CA, &
-    \M_{\CA\cdot}&=\prn{\onev^\CA}\trans\M, &
-    \mathbf{x}_\CA &= \prn{\onev^\CA}\trans\mathbf{x},
-  \end{aligned}
-\end{equation}
-%
-where $\pib$ is a row vector, $\CM$ is a matrix and $\mathbf{x}$ is a column vector.
-
-We can define a flux matrix, a.k.a.\ ergodic flow:
-%
-\begin{equation}\label{eq:flux}
-  \F = \D\inv\MM,
-  \qquad
-  \F_{ij} = \eq_i \MM_{ij}.
-\end{equation}
-%
-This measures the flow of probability between states in the equilibrium distribution.
-Detailed balance, a.k.a.\ reversibility, is equivalent to $\F=\F\trans$.
-
-The flux between two subsets is a particularly useful quantity:
-%
-\begin{equation}\label{eq:subflux}
-  \F_{\CA\CB} = {\eq}^\CA\MM\onev^\CB.
-\end{equation}
-%
-One can show that
-%
-\begin{equation}\label{eq:compflux}
-  \F_{\CA\CA\comp} = \F_{\CA\comp\CA} = -\F_{\CA\CA} = -\F_{\CA\comp\CA\comp}
-\end{equation}
-%
-using $\prn{{\eq}^\CA + {\eq}^{\CA\comp}}\MM=0$ and $\MM\prn{{\onev}^\CA + {\onev}^{\CA\comp}}=0$.
-
-\subsection{Lumpability}\label{sec:lump}
-
-Suppose we have partitioned the states into disjoint subsets, $\brc{\CA_\alpha}$:
-%
-\begin{equation}\label{eq:partition}
-  \bigcup_\alpha \CA_\alpha = \CS,
-  \qquad
-  \CA_\alpha \cap \CA_\beta = \delta_{\alpha\beta}\CA_\alpha.
-\end{equation}
-%
-We will use $\alpha$ instead of $\CA_\alpha$ in superscripts and subscripts in the following.
-The fact that these subsets are disjoint and exhaustive allows us to define the function
-%
-\begin{equation}\label{eq:whichsey}
-  \sigma(i)=\alpha
-  \qquad\means\qquad
-  i\in\CA_\alpha.
-\end{equation}
-%
-
-We can use this partition to define a new stochastic process associated with the original Markov chain.
-At time $t$, if the state of the original process is $i$, the state of the new process is $\sigma(i)$.
-
-One may ask if this new process is a Markov chain.
-The answer is yes, if the original Markov chain has a property called lumpability \wrt the partition
-(see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time):
-%
-\begin{equation}\label{eq:lump}
-  \sigma(i)=\sigma(j)
-  \quad\implies \quad
-  \MM_{i\alpha}=\MM_{j\alpha} \equiv \sum_{k\in\CA_\alpha} \MM_{jk},
-\end{equation}
-%
-\ie the total transition rate from some state to some subset is the same for all starting states within the same subset.
-This common value is the transition rate for the new lumped Markov chain.
-
-This can be rewritten with the aid of two matrices
-%
-\begin{equation}\label{eq:lumpmats}
-  U_{\alpha i} = \frac{\delta_{\alpha\sigma(i)}}{\abs{\CA_\alpha}},
-  \qquad
-  V_{i\alpha} = \delta_{\sigma(i)\alpha}.
-\end{equation}
-%
-Left multiplication by $U$ averages over subsets, right multiplication by $V$ sums over subsets.
-For $U$, we chose the uniform measure in each subset. Any measure would work equally well, \eg one proportional to the equilibrium distribution:
-%
-\begin{equation}\label{eq:altlumpmats}
-  U_{\alpha i} = \frac{{\eq_i}^\alpha}{\eq_\alpha}.
-\end{equation}
-%
-
-One can show that $(UV)_{\alpha\beta}=\delta_{\alpha\beta}$.
-The matrix $VU$ is also interesting.
-It has a block diagonal structure, with each block corresponding to a subset.
-Each block is a discrete-time ergodic Markov matrix (it is an independent trials process with probabilities given by the measure chosen for $U$).
-This means that the right eigenvectors with eigenvalue 1 will be those that are constant in each subset:
-%
-\begin{equation}\label{eq:setconst}
-  VU\mathbf{x}=\mathbf{x}
-  \qquad\means\qquad
-  \mathbf{x} = \sum_\alpha x_\alpha \onev^\alpha.
-\end{equation}
-%
-
-This allows us to write the lumpability condition \eqref{eq:lump}, and the transition matrix for the lumped process compactly:
-%
-\begin{equation}\label{eq:lumpcompact}
-  VU\MM V = \MM V,
-  \qquad
-  \widehat{\MM} = U\MM V.
-\end{equation}
-%
-By induction, one can show that similar relations hold for all powers:
-%
-\begin{equation}\label{eq:lumppower}
-  VU\MM^nV = \MM^nV,
-  \qquad
-  \widehat{\MM}^n = U\MM^n V,
-\end{equation}
-%
-and, via the Taylor series, for the exponential as well:
-%
-\begin{equation}\label{eq:lumpexp}
-  VU \e^{t\MM} V = \e^{t\MM}V,
-  \qquad
-  \e^{t\widehat{\MM}} = U\e^{t\MM} V.
-\end{equation}
-%
-The equilibrium distribution of the lumped process is given by
-%
-\begin{equation}\label{eq:lumpeq}
-  \widehat{\mathbf{p}}^\infty = \eq V.
-\end{equation}
-%
-
-\section{Signal-to-Noise ratio (SNR)}\label{sec:SNR}
-
-In this section we will look at the signal-to-noise curve, and put an upper bound on its initial value.
-We will only consider ergodic Markov chains.
-Transient states would be unoccupied in equilibrium and would not be accessed by the signal creation process, therefore they could be removed from the analysis.
-Absorbing chains are degenerate cases: they have zero initial signal but infinite decay times, so they can only be approached as the limit of a sequence of ergodic chains.
-
-\subsection{Framework}\label{sec:framework}
-
-The individual potentiation/depression events will be described by \emph{discrete}-time Markov chains:
-%
-\begin{equation}\label{eq:MWdef}
-  \M^\pm \equiv \I + \W^\pm,
-  \qquad
-  \M^\pm\onev = \onev,
-  \qquad
-  \M^\pm_{ij} \in [0,1].
-\end{equation}
-%
-The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf^\pm$, where $f^++f^-=1$ are the fraction of plasticity events that are potentiating/depressing respectively.
-This means that the ``forgetting'' process will be described by the \emph{continuous}-time Markov chain:
-%
-\begin{equation}\label{eq:forgetting}
-  \MM = r \frg \equiv r\prn{f^+\W^+ + f^-\W^-}.
-\end{equation}
-%
-We only require that this Markov chain is ergodic. The Markov chains described by $\M^\pm$ need not be.
-
-We assume that the probability distribution starts in the equilibrium distribution \eqref{eq:equilibrium}.
-During the initial signal creation, a fraction $f^+$ will change to $\eq\M^+$ and a fraction $f^-$ will change to $\eq\M^-$.
-After this, probabilities will evolve according to \eqref{eq:rowsum}.
-
-\subsection{SNR curve}\label{sec:SNRcurve}
-
-As we ignore correlations between different synapses, the only quantity of interest will be the number of synapses with each strength.
-The signal can be defined as the increase in synaptic strength due to potentiation plus the decrease in synaptic strength due to depression.
-If there are $N$ synapses and the strength of the synapse in each state is given by the column vector, $\w$, this is given by
-%
-\begin{equation}\label{eq:signal}
-  S(t) = N f^+ [\mathbf{p}(t|+)-\eq]\w - N f^+ [\mathbf{p}(t|-)-\eq]\w,
-\end{equation}
-%
-where $\mathbf{p}(t|\pm)$ is the probability of being in each state conditional on potentiation/depression at $t=0$:
-%
-\begin{equation}\label{eq:initprob}
-  \mathbf{p}(t|\pm) = \eq\M^\pm \e^{rt\frg}.
-\end{equation}
-%
-We can rewrite the signal as
-%
-\begin{equation}\label{eq:signalr}
-\begin{aligned}
-  S(t) &= N \eq (f^+\W^+ - f^-\W^-)\e^{rt\frg}\w\\
-    & N (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\frg}\w.
-\end{aligned}
-\end{equation}
-%
-For the noise, we can use the standard deviation of the synaptic strength:
-%
-\begin{equation}\label{eq:noise}
-  \text{Noise} = \sqrt{N \prn{\sum_i \eq_i\w_i^2 - (\eq\w)^2}}.
-\end{equation}
-%
-From now on we shall assume that there are only two possible synaptic strengths, strong and weak, which we will describe by $\w_i=\pm1$. This splits the set of states into two subsets:
-%
-\begin{equation}\label{eq:plusminusstates}
-  \boldsymbol{+} = \set{k}{\w_k=+1},
-  \qquad
-  \boldsymbol{-} = \set{k}{\w_k=-1}.
-\end{equation}
-%
-This simplifies the noise:
-%
-\begin{equation}\label{eq:noisepm}
-  \text{Noise} = \sqrt{N \prn{4\eq_+\eq_-}}.
-\end{equation}
-%
-The factors of $\eq_+\eq_-$ will have little effect on anything that follows, so we will drop them.\footnote{They will not affect the optimal transition rates for most quantities, and the optima will correspond to the symmetric case when $\eq_+=\eq_-=\half$.} This leaves the signal-to-noise ratio as
-%
-\begin{equation}\label{eq:SNRcurve}
-  \SNR(t) = \sqrt{N} (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\frg}\w.
-\end{equation}
-%
-
-We can express this in terms of the one parameter family of transition matrices:
-%
-\begin{equation}\label{eq:Walpha}
-  \begin{aligned}
-  \W(\alpha) &= \alpha\W^++(1-\alpha)\W^-,
-  &\quad&\implies&
-    \frg&=\W(f^+),\\&&&&
-    \W^+ - \W^- &= \diff{\W}{\alpha},\\&&&&
-    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \frg.
-  \end{aligned}
-\end{equation}
-%
-Then \eqref{eq:SNRcurve} becomes
-%
-\begin{equation}\label{eq:SNRalpha}
-  \SNR(t) = \sqrt{N} (2f^+f^-) \diff{\eq}{\alpha}(-\frg)\,\e^{rt\frg}\w.
-\end{equation}
-%
-
-\subsection{Degrees of freedom and constraints}\label{sec:constraints}
-
-We would like to use the components of $\W^\pm$ as the independent degrees of freedom when we try to maximise quantities associated with the SNR curve \eqref{eq:SNRcurve}.
-However, \eqref{eq:MWdef} imposes some restrictions on the space of allowed values.
-We can solve some of these by setting
-%
-\begin{equation}\label{eq:Wdiag}
-  \W^\pm_{ii} = -\sum_{j\neq i} \W^\pm_{ij}.
-\end{equation}
-%
-Then we treat the off-diagonal components as the independent degrees of freedom.
-These must satisfy the following inequalities:
-%
-\begin{align}\label{eq:Wmin}
-  \W^\pm_{ij} &\geq 0 \qquad \text{for $i \neq j$,}\\
-\label{eq:Wmax}
-  \sum_{j \neq i} \W^\pm_{ij} &\leq 1.
-\end{align}
-%
-The inequalities $\W^\pm_{ij} \leq 1$, $-1\leq\W^\pm_{ii}\leq0$ follow automatically from these.
-
-In \sref{sec:rescale}, we will see that the area under the SNR curve is invariant under the scaling $\W^\pm\to\lambda\W^\pm$.
-This means that we can ignore the constraint \eqref{eq:Wmax} when maximising the area:
-after the area has been maximised, we can use this degree of freedom to enforce \eqref{eq:Wmax} without changing the area.
-However, this scaling does change the SNR curve, in particular its initial value, so we can't ignore \eqref{eq:Wmax} when maximising that.
-
-\subsection{Lumpability}\label{sec:SNRlump}
-
-Suppose that we have a partition such that $\W^+$ and $\W^-$ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \sref{sec:lump}):
-%
-\begin{equation}\label{eq:lumpablesynapse}
-  VU\W^\pm V = \W^\pm V,
-  \qquad
-  VU\w=\w.
-\end{equation}
-%
-We can define a new synapse with
-%
-\begin{equation}\label{eq:lumpedsynapse}
-  \widehat{\W}^\pm = U\W^\pm V,
-  \qquad
-  \widehat{\w} = U \w,
-  \qquad
-  \widehat{\mathbf{p}}^\infty = \eq V.
-\end{equation}
-%
-This synapse has an SNR curve:
-%
-\begin{equation}\label{eq:lumpedSNR}
-  \begin{aligned}
-    \frac{\SNR(t)}{\sqrt{N} (2f^+f^-)} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}^+ - \widehat{\W}^-)\e^{rt\widehat{\W}^F}\widehat{\w}. \\
-      &= \eq VU (\W^+ - \W^-) VU \e^{rt\frg} VU \w. \\
-      &= \eq (\W^+ - \W^-) VU \e^{rt\frg} VU \w. \\
-      &= \eq (\W^+ - \W^-)\e^{rt\frg} VU \w. \\
-      &= \eq (\W^+ - \W^-)\e^{rt\frg}\w. \\
-  \end{aligned}
-\end{equation}
-%
-\ie the lumped process has exactly the same SNR as the original one.
-
-\subsection{Initial SNR and flux}\label{sec:initflux}
-
-Using the first line of \eqref{eq:signalr}, we can write the initial SNR as
-%
-\begin{equation}\label{eq:init}
-  \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f^+\W^+-f^-\W^-)(\onev^+-\onev^-).
-\end{equation}
-%
-Using $\W^\pm(\onev^++\onev^-)=0$ and \eqref{eq:compflux}:
-%
-\begin{equation*}
-  r{\eq}^-(f^+\W^++f^-\W^-)\onev^+ = \F_{-+} = \F_{+-} = r{\eq}^+(f^+\W^++f^-\W^-)\onev^-,
-\end{equation*}
-%
-we can rewrite \eqref{eq:init} as
-%
-\begin{equation}\label{eq:initflux}
-  I = \frac{4\F_{-+}}{r} - 4{\eq}^+f^+\W^+\onev^- - 4{\eq}^-f^-\W^-\onev^+.
-\end{equation}
-%
-The last two terms are guaranteed to be negative, as the diagonal parts of $\W^\pm$ cannot contribute.
-Therefore
-%
-\begin{equation}\label{eq:initfluxineq}
-  \SNR(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}.
-\end{equation}
-%
-This inequality is saturated if potentiation never takes it from a $+$ state to a $-$ state and depression never takes it from a $-$ state to a $+$ state.
-
-\subsection{Maximising initial SNR}\label{sec:initmax}
-
-The flux will be maximised if potentiation is guaranteed to take it to a $+$ state and depression is guaranteed to take it to a $-$ state:
-%
-\begin{equation}\label{eq:guarantee}
-  \sum_{j\in+} \W^+_{ij} = 1 \quad \forall\, i\in-,
-  \qquad
-  \sum_{j\in-} \W^-_{ij} = 1 \quad \forall\, i\in+.
-\end{equation}
-%
-This implies that the process is lumpable to a two state system (see \sref{sec:lump} and \sref{sec:SNRlump}):
-%
-\begin{multline}\label{eq:binarylump}
-%\begin{aligned}
-  \W^+ = \begin{pmatrix}
-           -1 & 1 \\
-           0 & 0 \\
-         \end{pmatrix}
-  ,\quad
-  \W^- = \begin{pmatrix}
-           0 & 0 \\
-           1 & -1 \\
-         \end{pmatrix}
-  ,\quad
-  \w = \begin{pmatrix}
-           -1 \\
-           1 \\
-         \end{pmatrix}
-  ,\\ \implies\quad
-  \frg = \begin{pmatrix}
-           -f^+ & f^+ \\
-           f^- & -f^- \\
-         \end{pmatrix}
-  ,\quad
-  \eq = \begin{pmatrix}
-           f^- & f^+ \\
-         \end{pmatrix}
-  ,%&
-%  \e^{rt\frg} &= \I+(1-\e^{-rt})\frg
-%         \begin{pmatrix}
-%           f^-+f^+\e^{-rt} & f^+(1-\e^{-rt}) \\
-%           f^-(1-\e^{-rt}) & f^++f^-\e^{-rt} \\
-%         \end{pmatrix}
-%  ,
-%  \end{aligned}
-\end{multline}
-%
-which leads to the initial SNR %curve
-\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to
-%$\SNR(t) = \sqrt{N (4f^+f^-)} \,\e^{-rt}$,
-$\SNR(0) \leq \sqrt{N (4f^+f^-)}$,
-which would have no effect on what follows.}
-%
-\begin{equation}\label{eq:binarySNR}
- % \SNR(t) = \sqrt{N} (4f^+f^-) \,\e^{-rt}.
-  \SNR(0) \leq \sqrt{N} (4f^+f^-) .
-\end{equation}
-%
-%The initial SNR is maximised at $f^+=f^-=\half$:
-This is maximised at $f^+=f^-=\half$:
-%
-\begin{equation}\label{eq:initmax}
-  \SNR(0) \leq \sqrt{N}.
-\end{equation}
-%
-
-\section{Area maximisation}\label{sec:areamax}
-
-In this section we will find an upper bound on the area under the signal-to-noise curve.
-As in \sref{sec:SNR}, we will only consider ergodic Markov chains.
-We will see in \sref{sec:multistate} that the optimal chain is absorbing, so it lies on the boundary of the (open) set of ergodic chains, but it still puts an upper bound on the area.
-
-\subsection{Area under signal-to-noise curve}\label{sec:area}
-
-The signal-to-noise curve is given by \eqref{eq:SNRalpha}.
-%%
-%\begin{equation}\label{eq:SNR}
-%  \SNR(t) = \sqrt{N}(2f^+f^-) \diff{\eq}{\alpha} (-\frg) \e^{rt\frg} \w.
-%\end{equation}
-%%
-The area is computed by integrating this
-%
-\begin{equation}\label{eq:area}
-\begin{aligned}
-  A &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\frg}}_0^\infty \w \\
-    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
-    &= \frac{\sqrt{N}(2f^+f^-)}{r} \diff{\eq}{\alpha} \w.
-\end{aligned}
-\end{equation}
-%
-
-We can rewrite this using \eqref{eq:diffpT}, with $rA=\sqrt{N}(2f^+f^-)\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
-%
-\begin{equation}\label{eq:areaT}
-  \hat{A} = \sum_{i,j,k} \eq_i \enc_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
-\end{equation}
-%
-
-\begin{defn}[Partial mixing times]
-We define the $\pm$ mixing times as
-%
-\begin{equation}\label{eq:mixingpm}
-\begin{aligned}
-  \eta^\pm_i &\equiv \sum_k \fptb_{ik} \eq_k \prn{\frac{1 \pm \w_k}{2}}
-%    &&= \sum_{\set{k}{\w_k=\pm1}} \fptb_{ik} \eq_k \\
-    &&= \sum_{k\in\pm} \fptb_{ik} \eq_k \\
-    &= \sum_k \prn{\fund_{kk}-\fund_{ik}} \prn{\frac{1 \pm \w_k}{2}}
-    &&= \sum_{k\in\pm} \prn{\fund_{kk}-\fund_{ik}} .
-\end{aligned}
-\end{equation}
-%
-We can think of $\eta^+_i$ as a measure of the ``distance'' to the $\w_k=+1$ states and $\eta^-_i$ as the ``distance'' to the $\w_k=-1$ states.
-\end{defn}
-Using \eqref{eq:mixdef}, we can write:
-%
-\begin{equation}\label{eq:mixingrels}
-\begin{aligned}
-  \eta^+_i + \eta^-_i &= \eta,\\
-  2(\eta^+_i - \eta^+_j) &= \sum_k (\fptb_{ik}-\fptb_{jk}) \eq_k \w_k
-    = \sum_k (\fund_{jk}-\fund_{ik}) \w_k.
-\end{aligned}
-\end{equation}
-%
-We could arrange the states in order of decreasing $\eta^+$, which is the same as the order of increasing $\eta^-$.
-
-We can rewrite \eqref{eq:areaT} as
-%
-\begin{equation}\label{eq:areaEta}
-\begin{aligned}
-  \hat{A} &= 2\sum_{i,j} \eq_i \enc_{ij} (\eta^+_{i} - \eta^+_{j}) &
-    &= -2\sum_{i,j} \eq_i \enc_{ij} \eta^+_{j} \\
-    &= 2\sum_{i,j} \eq_i \enc_{ij} (\eta^-_{j} - \eta^-_{i}) &
-    &= 2\sum_{i,j} \eq_i \enc_{ij} \eta^-_{j}.
-\end{aligned}
-\end{equation}
-%
-We can also express it in terms of the fundamental matrix \eqref{eq:funddef} as
-%
-\begin{equation}\label{eq:areaZ}
-  \hat{A} = \sum_{i,j,k,l} \pib_{l} \fund_{li} \enc_{ij} (\fund_{jk}-\fund_{ik}) \w_k
-    = \pib \fund \enc \fund \w.
-\end{equation}
-%
-
-It is also helpful to define the following quantities:
-%
-\begin{equation}\label{eq:areacoeffs}
-  \begin{aligned}
-    c_k &= \diff{\ln\eq_k}{\alpha}
-      = \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
-      = - \prn{\eq \enc \fptb}_k
-      = \frac{(\eq \enc \fund)_k}{\eq_k}, \\
-    a_i &= \sum_j \eq_i \enc_{ij} (\eta^+_{i} - \eta^+_{j}),\\
-    \implies
-    \hat{A} &= \sum_k c_k \eq_k \w_k
-      = 2\sum_i a_i.
-  \end{aligned}
-\end{equation}
-%
-Note that the optimal choice of $\w$ is $\w_k = \sgn(c_k)$.
-
-\subsection{Derivatives \wrt \texorpdfstring{$\W^\pm$}{W+/-}}\label{sec:deriv}
-
-As discussed in \sref{sec:constraints},
-we will regard the off-diagonal elements of $\W^\pm_{ij}$ to be the independent variables,
-with $\W^\pm_{ii}=-\sum_{j \neq i} \W^\pm_{ij}$ imposed by hand.
-Thus,
-%
-\begin{equation}\label{eq:basicderivs}
-  \pdiff{\frg_{ij}}{\W^\pm_{gh}} = f^\pm \delta_{gi}(\delta_{hj}-\delta_{ij}),
-  \qquad
-  \pdiff{\enc_{ij}}{\W^\pm_{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
-\end{equation}
-%
-The implicit $g \neq h$ that comes with all derivatives is unnecessary, as the derivatives above vanish when $g=h$.
-
-In particular, differentiating \eqref{eq:funddef},
-%
-\begin{equation}\label{eq:derivZ}
-  \pdiff{\fund_{ij}}{\W^\pm_{gh}} = rf^\pm \fund_{ig} (\fund_{hj}-\fund_{gj}).
-\end{equation}
-%
-We can then differentiate expression \eqref{eq:areaZ} to get
-%
-\begin{equation}\label{eq:derivA}
-  \begin{aligned}
-    \pdiff{\hat{A}}{\W^\pm_{gh}} =&\,
-      rf^\pm \sum_{ijkl} \enc_{ij} \w_k \pib_l \brk{
-         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
-    &\pm \sum_{kl} \w_k \pib_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
-    =&\,
-      2rf^\pm \sum_{ij} \enc_{ij} \eq_i \eq_g \brk{
-         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
-         (\fptb_{ig}-\fptb_{jg})(\eta^+_{g}-\eta^+_{h}) } \\
-      &\pm 2 \eq_g (\eta^+_g - \eta^+_h) \\
-    =&\,
-      2rf^\pm \sum_{ij} \enc_{ij} \eq_i \eq_g
-         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
-      \pm 2 \eq_g (\eta^+_g - \eta^+_h)\brk{1+\pdiff{\ln(\eq_g)}{\,\ln(f^\pm)}} \\
-    =&\,
-      2rf^\pm \eq_g \brk{\sum_i a_i (\fptb_{gi}-\fptb_{hi}) + c_g (\eta^+_{g}-\eta^+_{h})}
-      \pm 2 \eq_g (\eta^+_g - \eta^+_h). \\
-  \end{aligned}
-\end{equation}
-%
-where $a_i$ and $c_k$ were defined in \eqref{eq:areacoeffs}.
-
-It is sometimes useful to consider the following derivatives:
-%
-\begin{equation}\label{eq:derivqw}
-  \begin{aligned}
-    \pdiff{}{\frg_{gh}} &\equiv     \pdiff{}{\W^+_{gh}} +     \pdiff{}{\W^-_{gh}} ,&\qquad
-    \pdiff{}{\enc_{gh}}    &\equiv f^- \pdiff{}{\W^+_{gh}} - f^+ \pdiff{}{\W^-_{gh}} .
-  \end{aligned}
-\end{equation}
-%
-Each of these derivatives behaves as their names suggest:
-%
-\begin{equation}\label{eq:derivqweff}
-  \pdiff{\frg_{ij}}{\frg_{gh}} = \pdiff{\enc_{ij}}{\enc_{gh}}
-  =  \delta_{gi}(\delta_{hj}-\delta_{ij}),
-  \qquad
-  \pdiff{\enc_{ij}}{\frg_{gh}} = \pdiff{\frg_{ij}}{q_{gh}}  = 0.
-\end{equation}
-%
-This is because we could treat $\frg$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W^\pm$.
-
-\subsubsection{Scaling mode}\label{sec:rescale}
-
-Consider the following differential operator:
-%
-\begin{equation}\label{eq:scaleop}
-  \Delta \equiv \sum_{g,h} \W^+_{gh}\pdiff{}{\W^+_{gh}} + \W^-_{gh}\pdiff{}{\W^-_{gh}}.
-\end{equation}
-%
-This corresponds to the scaling, $\W^\pm \to (1+\epsilon)\W^\pm$.
-Intuitively, this has two effects: it scales up the initial potentiation/depression and it scales down all timescales.
-This intuition is confirmed by the following results:
-%
-\begin{equation}\label{eq:scaleeffects}
-  \begin{aligned}
-    \Delta \fund &= \tau\onev\eq - \fund ,&
-    \Delta \eq  &= 0 ,&
-    \Delta \fpt  &= -\fpt ,
-    \\
-    \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
-    \Delta \enc_{ij} &= \enc_{ij} ,&
-    \Delta \hat{A}  &= 0 ,&
-  \end{aligned}
-\end{equation}
-%
-The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W^\pm$.
-
-As the area is invariant under this scaling, we can consider the $\W^\pm$ to be projective coordinates.
-Therefore we don't need to enforce the constraint \eqref{eq:Wmax} while looking for the maximum area, as we can use this null-mode to enforce it afterwards without changing the area.
-We also don't have to worry about the boundary at infinity,
-as sending some of them to infinity is equivalent to sending the rest to zero.
-
-\subsection{Kuhn-Tucker conditions}\label{sec:kuhntucker}
-
-Consider the Lagrangian
-%
-\begin{equation}\label{eq:lagrangian}
-  \CL = \hat{A} + \sum_{\pm}\sum_{i\neq j} \mu^\pm_{ij} \W^\pm_{ij} +\lambda\eq\w.
-\end{equation}
-%
-The last term ensures that we hold $\eq_\pm$ fixed during this extremisation, so that we can safely ignore the factors that we dropped from the noise in \eqref{eq:noisepm}.
-Necessary conditions for an extremum are
-%
-\begin{equation}\label{eq:extremum}
-  \pdiff{\CL}{\W^\pm_{gh}} = 0,
-  \qquad
-    \mu^\pm_{gh} \geq 0,\quad
-    \W^\pm_{gh} \geq 0,\quad
-    \mu^\pm_{gh}\W^\pm_{gh} = 0.
-\end{equation}
-%
-with $g \neq h$.
-This enforces the constraints \eqref{eq:Wmin}, but not \eqref{eq:Wmax}.
-As discussed in \sref{sec:constraints} and \sref{sec:rescale}, that can be enforced after finding the maximum with the null scaling degree of freedom.
-
-\subsubsection{Triangularity}\label{sec:triangular}
-
-Consider
-%
-\begin{equation}\label{eq:shiftqderiv}
-  \pdiff{\CL}{\enc_{gh}} =
-  f^- \pdiff{\CL}{\W^+_{gh}} - f^+ \pdiff{\CL}{\W^-_{gh}}
-   = (f^- \mu^+_{gh} - f^+ \mu^-_{gh}) + 2\eq_g (\eta^+_g - \eta^+_h)
-   = 0.
-\end{equation}
-%
-This corresponds to the shift
-%
-\begin{equation}\label{eq:shiftq}
-  \W^+_{ij} \to \W^+_{ij} + f^-\epsilon_{ij},
-  \qquad
-  \W^-_{ij} \to \W^-_{ij} - f^+\epsilon_{ij},
-  \qquad
-  \sum_j \epsilon_{ij} = 0,
-\end{equation}
-%
-which leaves $\frg$ unchanged, and therefore $\eq$, $\fpt$ and $\eta^\pm$ as well.
-
-Assume $\eta^+_g > \eta^+_h$. Then
-%
-\begin{equation}\label{eq:lowertriangular}
- f^- \mu^+_{gh} - f^+ \mu^-_{gh} <0
- \qquad\implies\qquad
- \mu^-_{gh} >0
- \qquad\implies\qquad
- \W^-_{gh}=0.
-\end{equation}
-%
-Similarly, if $\eta^+_g < \eta^+_h$, then
-%
-\begin{equation}\label{eq:uppertriangular}
- f^- \mu^+_{gh} - f^+ \mu^-_{gh} > 0
- \qquad\implies\qquad
- \mu^+_{gh} >0
- \qquad\implies\qquad
- \W^+_{gh}=0.
-\end{equation}
-%
-Thus, if we arrange the states in order of decreasing $\eta^+$, $\W^+$ is upper-triangular and $\W^-$ is lower triangular.
-
-We have ignored the possibility that $\eq_g=0$, as this would imply that $\fpt_{ig}=\infty$, which would in turn imply that the Markov process is not ergodic.
-
-%Note that if we multiply \eqref{eq:shiftqderiv} by $\enc_{gh}$ and sum over $(g,h)$, we get
-%%
-%\begin{equation}\label{eq:stupid}
-%  \hat{A} = \sum_\pm\sum_{g\neq h} \mu^\pm_{gh} \frg_{gh},
-%\end{equation}
-%%
-%which is probably useless
-
-\subsubsection{Increasing \texorpdfstring{$c_k$}{c(k)}}\label{sec:areacoeffincr}
-
-Consider the following combinations of derivatives:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:areacoeffincrderiv}
-    \Delta_{gh} &\equiv
-      \frac{1}{\eq_{g}} \pdiffc{}{\frg_{gh}}
-      + \frac{1}{\eq_h} \pdiffc{}{\frg_{hg}}, \\
-%  \end{aligned}
-\end{align}
-%
-Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).
-
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
-%
-\begin{equation}\label{eq:sareacoeffincrprob}
-  \Delta_{gh} \eq_i = 0,
-\end{equation}
-%
-which makes it possible to integrate the perturbation:
-%
-\begin{equation}\label{eq:areacoeffincrfinite}
-  \W^\pm \to \W^\pm + \D\boldsymbol{\epsilon},
-  \qquad
-  \begin{aligned}
-  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
-  \\
-  \boldsymbol{\epsilon} \onev &= 0.
-  \end{aligned}
-\end{equation}
-%
-But more interestingly:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:areacoeffincr}
-    \Delta_{gh}\CL &=
-      \frac{\mu^+_{gh}+\mu^-_{gh}}{\eq_g} + \frac{\mu^+_{hg}+\mu^-_{hg}}{\eq_h}
-      + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
-%  \end{aligned}
-\end{align}
-%
-where $c_k$ were defined in \eqref{eq:areacoeffs}.
-
-Using the non-negativity of the Kuhn-Tucker multipliers, $\mu^\pm_{ij}$, \eqref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
-
-Note that, according to \sref{sec:triangular}, either $\W^+_{gh}$ or $\W^-_{gh}$ will be zero at the maximum, therefore we can expect one of $\mu^+_{gh}+\mu^-_{gh}$ to be non-zero.
-This would rule out degeneracy of the $c_k$ or $\eta^+_k$.
-Looking at \eqref{eq:shiftqderiv} closely, the only way $\mu^+_{gh}+\mu^-_{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
-
-\subsubsection{Shortcuts}\label{sec:shortcuts}
-
-Now consider the following combinations of derivatives for $m>1$:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:shortcutderiv}
-    \widetilde{\Delta}^\pm_{g,m} &\equiv
-      \brk{ \sum_{k=0}^{m-1} \frac{1}{\eq_{g\pm k}} \pdiffc{}{\W^\pm_{g\pm k,g\pm(k+1)}} }
-      - \frac{1}{\eq_g} \pdiffc{}{\W^\pm_{g,g\pm m}}.
-%  \end{aligned}
-\end{align}
-%
-Once again, they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).
-
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
-%
-\begin{equation}\label{eq:shortcutprob}
-  \widetilde{\Delta}^\pm_{g,m} \eq_i = 0,
-\end{equation}
-%
-which makes it possible to integrate the perturbation:
-%
-\begin{equation}\label{eq:shortcutfinite}
-  \W^\pm \to \W^\pm + \D\boldsymbol{\epsilon}^{\pm(g,m)},
-  \qquad
-  \begin{aligned}
-    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g,g\pm m}
-      &\!\!\!\!\!\!&= -\epsilon,\\
-    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm(k+1)}
-      &\!\!\!\!\!\!&= \epsilon
-        &\;
-        &\forall\, k \in [0,m-1],\\
-    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm k}
-      &\!\!\!\!\!\!&= -\epsilon
-        &
-        &\forall\, k \in [1,m-1].
-  \end{aligned}
-\end{equation}
-%
-But more interestingly for our purposes:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:shortcutarea}
-    \widetilde{\Delta}^\pm_{g,m}\CL &=
-      \brk{ \sum_{k=0}^{m-1} \frac{\mu^\pm_{g\pm k,g\pm(k+1)}}{\eq_{g\pm k}}
-      - \frac{\mu^\pm_{g,g\pm m}}{\eq_g}}
-      + 2rf^\pm \sum_{k=0}^{m-1} \prn{\eta^+_{g\pm k} - \eta^+_{g\pm(k+1)}} \prn{c_{g\pm k} - c_{g}},
-%  \end{aligned}
-\end{align}
-%
-
-If we put the states in order of decreasing $\eta^+_k$, the results of the \sref{sec:areacoeffincr} tell us that the $c_k$ are non-decreasing.
-This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
-If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu^\pm_{g,g\pm m}>0$, which in turn implies that $\W^\pm_{g,g\pm m}=0$.
-This would tell us that the process with the maximal area has to have a multi-state topology.
-
-%From the expression for the area in \eqref{eq:areacoeffs}, it will clearly also be optimal to have the $\w_k$ non-decreasing, \ie all the positive $\w_k$ should lie ahead of all the negative $\w_k$.
-
-\subsubsection{Summary}\label{sec:KTsummary}
-
-Using the Kuhn-Tucker formalism, we have shown that,
-with the states arranged in order of non-increasing $\eta^+_i$:
-%
-\begin{itemize}
-  \item There can be no ergodic maximum for which $\W^+$ contains backwards transitions or $\W^-$ contains forwards transitions.
-  \item There can be no ergodic maximum with the $c_k$ decreasing.
-  \item The $c_k$ may only be degenerate at an ergodic maximum if the corresponding $\eta^+_k$ are also degenerate.
-  \item If the $c_k$ increase and the $\eta^+_i$ decrease, there can be no ergodic maximum with shortcuts.
-\end{itemize}
-%
-These were shown by finding allowed perturbations that increase the area.
-
-%Suppose we start with some arbitrary ergodic process.
-If $\W^\pm$ contain backwards/forwards transitions or the $c_k$ decrease, then we can increase the area by applying the perturbations \eqref{eq:shiftqderiv} and \eqref{eq:areacoeffincrderiv}.
-%The second perturbation will generate the unwanted backwards/forwards transitions, requiring subsequent application of first perturbation.
-%The first perturbation will change the $c_k$, possibly changing their order, requiring subsequent application of second perturbation.
-%
-%This procedure can end in several ways:
-%%
-%\begin{itemize}
-%  \item Applying \eqref{eq:areacoeffincrderiv} an infinite amount sends two transition rates to infinity, which due to the scaling mode (\sref{sec:rescale}), is equivalent to sending all other transitions to zero.
-%      Subsequent application of \eqref{eq:shiftqderiv} just removes the backwards transition from $\W^+$ and the forward transition from $\W^-$.
-%  \item Applying \eqref{eq:areacoeffincrderiv} ends in a process with degenerate $c_k$ or $\eta^+_k$, and the subsequent application of \eqref{eq:shiftqderiv} doesn't change this degeneracy.
-%  \item As above, but the subsequent application of \eqref{eq:shiftqderiv} lifts the degeneracy, resulting in a chain with increasing $c_k$.
-%\end{itemize}
-%%
-%In the third case, we can then apply the perturbation \eqref{eq:shortcutderiv} to remove all the shortcuts.
-%Either this, or the first case, results in a multi-state topology, possibly disconnected or without the full set of states.
-
-This leaves two possibilities for the maximum area Markov chain.
-Either there is no degeneracy and no shortcuts, which implies the Multi-state topology that we'll discuss in \sref{sec:multistate}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \eqref{eq:shortcutarea}).
-
-Degeneracy tends to be very delicate.
-It is usually hard to arrange without some symmetry relating degenerate states.
-Such a symmetry would imply lumpability (see \sref{sec:lump}).
-The lumped chain would not have any shortcuts, as an entire degenerate set cannot be bypassed.
-As this lumped chain has the same area (see \sref{sec:SNRlump}), we would need only consider the multi-state topology.
-
-\subsection{Multi-state topology}\label{sec:multistate}
-
-The multi-state topology is defined by (see \cite{amit1994learning,Fusi2007multistate}):
-%
-\begin{equation}\label{eq:multistatedef}
-  \W^+_{ij} = q^+_i \delta_{i+1,j},
-  \qquad
-  \W^-_{ij} = q^-_j \delta_{i,j+1}.
-\end{equation}
-%
-It saturates various inequalities:
-%
-\begin{equation}\label{eq:multistateineq}
-  \begin{aligned}
-    \fptb_{ik} - \fptb_{jk} &=
-      \begin{cases}
-        \fptb_{ij},  &\text{if}\quad i \leq j \leq k \quad\text{or}\quad i \geq j \geq k,\\
-        -\fptb_{ji}, &\text{if}\quad j \leq i \leq k \quad\text{or}\quad j \geq i \geq k,\\
-      \end{cases} \\
-    r\eq_i \frg_{ij} \prn{\fptb_{ij}+\fptb_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
-  \end{aligned}
-\end{equation}
-%
-and it satisfies detailed balance (a.k.a.\ reversibility a.k.a. $\CL^2_{\eq}$ self-adjointness):
-%
-\begin{equation}\label{eq:multistateprob}
-  f^+ q^+_i \eq_i = f^- q^-_i \eq_{i+1},
-\end{equation}
-%
-which means we can always choose the transition rates, $q^\pm_i$, to give any desired equilibrium probabilities, $\eq_i$.
-
-This allows us to calculate the $c_k$'s:
-%
-\begin{equation}\label{eq:areacoeffchain}
-\begin{aligned}
-  c_k =&\, \sum_{i<k} \fpt_{i,i+1} \prn{\eq_{i}q^+_i+\eq_{i+1}q^-_i}
-    %\\&
-    - \sum_{i\geq k} \fpt_{i+1,i} \prn{\eq_{i}q^+_i+\eq_{i+1}q^-_i}
-  ,\\
-  c_{k+1} - c_k =&\, \prn{\fpt_{k,k+1}+\fpt_{k+1,k}} \prn{\frac{\eq_{k}\frg_{k,k+1}}{f^+}+\frac{\eq_{k+1}\frg_{k+1,k}}{f^-}}
-    =%\\=&\,
-    \frac{1}{rf^+f^-},\\
-  \sum_k c_k \eq_k =&\, \sum_{ij} \eq_i \enc_{ij} (\eta-\eta) = 0,\\
-%  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f^+f^-},
-  \implies c_k =&\, \frac{k - \sum_j j\eq_j}{rf^+f^-},
-\end{aligned}
-\end{equation}
-%
-where we used \eqref{eq:multistateineq} to derive the first two equations respectively and Th.\ref{th:kemenyconst} to derive the third. This allows us to write the area as
-%
-\begin{equation}\label{eq:multistatearea}
- % A = 2\sqrt{N} \sum_k \brk{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \eq_k \w_k.
-  A = \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k
-    = \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
-\end{equation}
-%
-where we used $\w_k=\sgn(c_k)$, as discussed after \eqref{eq:areacoeffs}.
-%It will also help to define $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
-%Note that
-%%
-%\begin{equation}\label{eq:allpplusorminus}
-%  \begin{aligned}
-%    \eq_+ &=1
-%     \quad&\implies\quad
-%     \eq_k \w_k &= \eq_k
-%     \quad&\implies\quad
-%     A=0,\\
-%    \eq_- &=1
-%     \quad&\implies\quad
-%     \eq_k \w_k &= -\eq_k
-%     \quad&\implies\quad
-%     A=0,
-%  \end{aligned}
-%\end{equation}
-%%
-%neither of which are optimal.
-%
-%Consider the Lagrangian
-%%
-%\begin{equation}\label{eq:multistatelagrangian}
-%  \CL = \frac{A}{2\sqrt{N}} + \lambda\prn{1-\sum_i \eq_i} + \sum_i \mu_i \eq_i,
-%\end{equation}
-%%
-%where $\lambda$ is a Lagrange multiplier and $\mu_i$ are Kuhn-Tucker multipliers (see \eqref{eq:extremum}). Extremising \wrt $\eq_i$:
-%%
-%\begin{equation}\label{eq:multistateext}
-%%  \pdiff{\CL}{\eq_i} = \brk{\prn{i-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \w_i
-%%    - \prn{i-\frac{n+1}{2}}\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
-%  \pdiff{\CL}{\eq_i} = \brk{i - \sum_j j\eq_j} \w_i
-%    - i\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
-%\end{equation}
-%%
-%Consider an $i$ such that $\w_i=\w_{i+1}$:
-%%
-%\begin{equation}\label{eq:multistatemudiff}
-%  \mu_{i+1} - \mu_i = \prn{\eq_+-\eq_-} - \w_i.
-%\end{equation}
-%%
-%As $\abs{\eq_+-\eq_-}<1$ (unless we want $A=0$, \eqref{eq:allpplusorminus}), this means the $\mu_i$ increase amongst the negative $\w_i$ and decrease amongst the positive $\w_i$.
-%Therefore the only non-zero equilibrium probabilities at optimum are $\eq_1$ and $\eq_n$.
-
-First let us maximise \eqref{eq:multistatearea} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
-Clearly this will happen when we put all of the probability at the ends: $\eq_1=\eq_-$ and $\eq_n=\eq_+$ are the only non-zero $\eq_k$.
-This gives an area of
-\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_+\eq_-)}(n-1)/r$, which would have no effect on what follows.}
-
-%
-\begin{equation}\label{eq:multistateextarea}
-  A \leq \frac{\sqrt{N}}{r}(n-1)\prn{4\eq_+\eq_-}.
-\end{equation}
-%
-This is maximised at $\eq_+=\eq_-=\half$:
-%
-\begin{equation}\label{eq:maxarea}
-  A \leq \frac{\sqrt{N}}{r}(n-1).
-\end{equation}
-%
-Note that the chain that achieves this is not ergodic, the two states at each end are absorbing.
-This is similar to the results found numerically in \cite{Barrett2008discrete} in a slightly different situation.
-
-\section{Finite time SNR}\label{sec:finitetime}
-
-The SNR at time $t$ was given in \eqref{eq:SNRcurve} as
-%
-\begin{equation*}
-  \SNR(t) = \sqrt{N} (2f^+f^-) \eq \enc \, \e^{rt\frg}\w.
-\end{equation*}
-%
-Let us try to maximise this at time $t$ only.
-We will consider the independent parameters to be the off-diagonal elements of $\M^\pm$, with the diagonal elements determined via $\M^\pm_{ii} = \sum_{j\neq i} \M^\pm_{ij}$.
-Consider the Lagrangian
-%
-\begin{equation}\label{eq:finlagrangian}
-  \CL = \eq \enc\, \e^{rt\frg}\w + \sum_\pm \sum_{ij} \mu^\pm_{ij} \M^\pm_{ij}.
-\end{equation}
-%
-Necessary conditions for an extremum are
-%
-\begin{equation*}
-  \pdiff{\CL}{\M^\pm_{gh}} = 0,
-  \qquad
-    \mu^\pm_{gh} \geq 0,\quad
-    \M^\pm_{gh} \geq 0,\quad
-    \mu^\pm_{gh}\M^\pm_{gh} = 0.
-\end{equation*}
-%
-We can use the results of \cite{Ortiz2001diffmatexp} to write down the shift of this Lagrangian under an infinitesimal perturbation
-%
-\begin{equation}\label{eq:finpert}
-\begin{aligned}
-  \delta\CL =&\, \eq \delta\frg \fund\, \enc \, \e^{rt\frg} \w
-    + \eq \delta\enc \, \e^{rt\frg} \w %\\ &
-   + r\int_{0}^{t}\!\!\dr\tau\, \eq \enc\, \e^{r\tau\frg} \delta\frg \e^{r(t-\tau)\frg} \w\\
-   &+ \sum_\pm \sum_{ij} \mu^\pm_{ij} \delta\M^\pm_{ij}
-\end{aligned}
-\end{equation}
-%
-This can be rewritten using the eigenvector decomposition of $\frg$.
-%
-\begin{equation}\label{eq:eigendecomp}
-  \frg = \sum_a -q_a \uv^a \vv^a,
-  \quad
-  \vv^a \uv^b = \delta_{ab},
-  \quad
-  \frg \uv^a = -q_a \uv^a,
-  \quad
-  \vv^a \frg = -q_a \vv^a.
-\end{equation}
-%
-People often write $\vv$ as $\vv\trans$, presumably because they find the concept of row vectors too difficult to handle.
-We're better than that.
-We can choose $q_1=0$, $\uv^1=\onev$ and $\vv^1=\eq$.
-
-Then, defining
-%
-\begin{equation}\label{eq:eigdiff}
-  F_{ab} = \left\{
-           \begin{aligned}
-             &rt\,\e^{-q_art}, &\quad&\text{if } q_a=q_b,\\
-             &\frac{\e^{-q_art}-\e^{-q_brt}}{q_b-q_a}, &&\text{otherwise,}
-           \end{aligned}
-           \right.
-\end{equation}
-%
-we get
-%
-\begin{equation}\label{findiff}
-\begin{aligned}
-  \pdiff{\CL}{\M^\pm_{gh}} =&\,
-    f^\pm \eq_g \prn{\brk{\fund \, \enc \, \e^{rt\frg} \w}_h - \brk{\fund \, \enc \, \e^{rt\frg} \w}_g} \\
-    &\pm \eq_g \prn{\brk{\e^{rt\frg} \w}_h - \brk{\e^{rt\frg} \w}_g} \\
-    &+ f^\pm \sum_{ab} F_{ab} (\eq\enc\uv^a) \vv^a_g (\uv^b_h-\uv^b_g) (\vv^a\w) \\
-    &+ \mu^\pm_{gh} - \mu^\pm_{gg}.
-\end{aligned}
-\end{equation}
-%
-
-Consider
-%
-\begin{equation}\label{eq:finshiftq}
-\begin{aligned}
-  f^- \pdiff{\CL}{\M^+_{gh}} - f^+ \pdiff{\CL}{\M^-_{gh}}
-   =&\, \eq_g \prn{\brk{\e^{rt\frg} \w}_h - \brk{\e^{rt\frg} \w}_g}\\
-    &+ f^- \mu^+_{gh} + f^+\mu^-_{gg} - f^+ \mu^-_{gh} - f^- \mu^+_{gg}\\
-   =&\, 0,
-\end{aligned}
-\end{equation}
-%
-which corresponds to the shift considered in \S\ref{sec:triangular}. We can see that
-%
-\begin{equation}\label{eq:etalambdaineq}
-  \begin{aligned}
-    \brk{\e^{rt\frg} \w}_g &< \brk{\e^{rt\frg} \w}_h
-%    &\quad & \implies \quad&
-%    f^+ \mu^-_{gh} + f^- \mu^+_{gg} &> 0
-    &\quad & \implies \quad&
-    \mu^-_{gh} &> 0
-    &\quad & \text{or} \quad&
-    \mu^+_{gg} &> 0
-    \\ &
-    &\quad & \implies \quad&
-    \M^-_{gh} &= 0
-    &\quad & \text{or} \quad&
-    \M^+_{gg} &= 0,
-    \\
-    \brk{\e^{rt\frg} \w}_g &> \brk{\e^{rt\frg} \w}_h
-%    &\quad & \implies \quad&
-%    f^- \mu^+_{gh} + f^+\mu^-_{gg} &> 0
-    &\quad & \implies \quad&
-    \mu^+_{gh} &> 0
-    &\quad & \text{or} \quad&
-    \mu^-_{gg} &> 0
-    \\ &
-    &\quad & \implies \quad&
-    \M^+_{gh} &= 0
-    &\quad & \text{or} \quad&
-    \M^-_{gg} &= 0.
-  \end{aligned}
-\end{equation}
-%
-
-\section{Laplace transform}\label{sec:laplace}
-
-\subsection{Fundamental matrix \etc}\label{sec:lfund}
-
-In analogy to \eqref{eq:funddef}, define
-%
-\begin{equation}\label{eq:lfund}
-  \fund(s) = (s\I +\onev\pib -\MM)\inv.
-\end{equation}
-%
-This reduces to the fundamental matrix at $s=0$.
-The analogous result to \eqref{eq:fundrowsum} is
-%
-\begin{equation}\label{eq:lfundrowsum}
-  \fund(s)\onev = \frac{\tau}{1+s\tau}\onev.
-\end{equation}
-%
-The analogues of \eqref{eq:fundprob}, \eqref{eq:fundQZ} and \eqref{eq:fundZQ} are much uglier, but luckily we won't need them.
-Supose we have some row vector $\mathbf{a}$ such that $\mathbf{a}\onev=0$.
-Then $\mathbf{a}\fund(s)$ is independent of $\pib$:
-%
-\begin{equation}\label{eq:alfund}
-  \pdiff{(\mathbf{a}\fund(s))_i}{\pib_k} = (\mathbf{a}\fund(s)\onev) \fund_{ki}(s) = \frac{\tau}{1+s\tau} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
-\end{equation}
-%
-
-Then we also define
-%
-\begin{equation}\label{eq:lfptb}
-  \fptb(s) = (\onem\fund\dg(s)-\fund(s))\D,
-  \qquad
-  \fptb_{ij}(s) = \frac{\fund_{jj}(s)-\fund_{ij}(s)}{\eq_j}.
-\end{equation}
-%
-This reduces to the first passage time matrix at $s=0$, but at other $s$ there is (probably) no relation.
-
-Using the exact same proof as \autoref{th:kemenyconst}, we see that the quantity
-%
-\begin{equation}\label{eq:lkemeny}
-  \eta(s) = \sum_j \fptb_{ij}(s)\eq_j
-\end{equation}
-%
-is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau}{1+s\tau}$.
-We can the define 
-%
-\begin{equation}\label{eq:lpmixtime}
-  \eta^{\pm}_i(s) = \sum_j \fptb_{ij}(s)\eq_j \prn{\frac{1\pm\w_j}{2}} =  \sum_{j\in\pm} \fptb_{ij}(s)\eq_j.
-\end{equation}
-%
-We can arrange the states in order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
-
-\subsection{Laplace transform of SNR curve}\label{sec:laplaceSNR}
-
-Consider the Laplace transform of the evolution operator:
-%
-\begin{equation}\label{eq:levol}
-  \mathbf{G}(s) = \intd[_0^{\infty}]{t} \e^{(\MM-s) t}.
-\end{equation}
-%
-For $\Re s>0$, we have 
-%
-\begin{equation}\label{eq:levolcalc}
-  (s-\MM)\mathbf{G}(s) = \intd[_0^{\infty}]{t} (s-\MM)\e^{(\MM-s) t} = \brk{-\e^{(\MM-s) t}}_0^{\infty} = \I.
-\end{equation}
-%
-As $(s-\MM)$ is invertible for $\Re s>0$, because the real part of all eigenvalues of $\MM$ are nonpositive, we have
-%
-\begin{equation}\label{eq:levolres}
-  \mathbf{G}(s) = (s-\MM)\inv.
-\end{equation}
-%
-For $s=0$, we can avoid problems by replacing $\mathbf{G}(s) \to \fund(s)$.
-
-Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurve}
-%
-\begin{equation}\label{eq:lsnr}
-\begin{aligned}
-  A(s) &= \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) \\
-   &= \intd[_0^{\infty}]{t} \sqrt{N}(2f^+f^-) \eq \enc \e^{(r\frg-s)t} \w \\
-   &= \frac{\sqrt{N}(2f^+f^-)}{r} \eq \enc \prn{s - r\frg}\inv \w.
-\end{aligned}
-\end{equation}
-%
-This expression is ill-behaved at $s=0$.
-Thanks to \eqref{eq:alfund}, we can solve this by the replacement $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
-%
-\begin{equation}\label{eq:larea}
-\begin{aligned}
-  \hat{A}(s) = \frac{rA(s)}{\sqrt{N}(2f^+f^-)} &=  \eq \enc \fund(s) \w \\
-    &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
-    &= 2\sum_{ij}  \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)).
-\end{aligned}
-\end{equation}
-%
-Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.
-
-In analogy with \eqref{eq:areacoeffs}, we define
-%
-\begin{equation}\label{eq:lareacoeffs}
-  \begin{aligned}
-    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
-    a_i(s) &= \sum_j \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)),\\
-    \implies
-    \hat{A}(s) &= \sum_k c_k(s) \eq_k \w_k
-      = 2\sum_i a_i(s).
-  \end{aligned}
-\end{equation}
-%
-
-\subsection{Perturbation analysis}\label{sec:lpert}
-
-Using the same approach as the area bound, we find that the derivatives are
-%
-\begin{equation}\label{eq:lAdiff}
-\begin{aligned}
-  \pdiff{\hat{A}(s)}{\W^\pm_{gh}} 
-     =&\, 2rf^\pm \eq_g \brk{ \sum_i a_i(s) (\fptb_{gi}(0)-\fptb_{hi}(0))
-     +c_g(s) (\eta^+_g(s) - \eta^+_h(s)) } \\
-     &\pm 2\eq_g (\eta^+_g(s) - \eta^+_h(s)).
-\end{aligned}
-\end{equation}
-%
-This is exactly the same as \eqref{eq:derivA}.
-It follows that, using exactly the same perturbations, that the Laplace transform of the SNR curve for \emph{any} model is lower than that of a certain model with the serial topology having the same equilibrium distribution.
-
-\subsection{Envelope}\label{sec:env}
-
-Hopefully we can prove an upper bound on the Laplace transform of the memory curve:
-%
-\begin{equation}\label{eq:lbound}
-  A(s) \leq \sqrt{N}B(s) = \frac{\sqrt{N}}{C(s)}.
-\end{equation}
-%
-We will assume that $C(s)$ is a concave function.
-We already know that $C(0)=\frac{r}{M-1}$ and $C(s) \to s$ as $s\to\infty$.
-
-In the eigenmode decomposition \eqref{eq:eigendecomp}:
-%
-\begin{equation}\label{eq:eigenmode}
-  \snr(t) = \sqrt{N}\sum_a \CI_a\, \e^{-q_a t},
-  \qquad
-  \CI_a = (2f^+f^-) (\eq \enc \uv^a) (\vv^a \w)
-\end{equation}
-%
-the upper bound reads
-%
-\begin{equation}\label{eq:lboundeig}
-  \sum_a \frac{\CI_a}{q_a+s} \leq B(s).
-\end{equation}
-%
-Now let's maximise $\snr(t_0)$ with respect to $\brc{\CI_a,q_a}$.
-Consider the Lagrangian
-%
-\begin{equation}\label{eq:envlagrange}
-  \CL = \sum_a \CI_a\, \e^{-q_a t_0} +\intd[_0^\infty]{s} \mu(s)\brk{B(s)-\sum_a\frac{\CI_a}{q_a+s}}.
-\end{equation}
-%
-Necessary conditions for a maximum are
-%
-\begin{alignat}{3}
-  \pdiff{\CL}{\CI_a} &= \e^{-q_a t_0} - \intd{s} \frac{\mu(s)}{q_a+s} = 0, \label{eq:envI}\\
-  \pdiff{\CL}{q_a} &= \CI_a\brk{-t_0\,\e^{-q_a t_0} + \intd{s} \frac{\mu(s)}{(q_a+s)^2}} = 0, \label{eq:envq}\\
-  \mu(s)\fdf{\CL}{\mu(s)} &= \mu(s)\brk{B(s)-\sum_a\frac{\CI_a}{q_a+s}} = 0, \label{eq:envKT}\\
-  \mu(s) &\geq 0, \label{eq:envKTm}\\
-  \fdf{\CL}{\mu(s)} &= B(s)-\sum_a\frac{\CI_a}{q_a+s} \geq 0 \label{eq:envKTbnd}.
-\end{alignat}
-%
-If we assume $\CI_a \neq0$ for some $a$, then \eqref{eq:envI} and \eqref{eq:envq} read
-%
-\begin{equation}\label{eq:envnumexp}
-\begin{aligned}
-  \e^{-q_a t_0} &= \intd{s} \frac{\mu(s)}{q_a+s}, \\
-  t_0\,\e^{-q_a t_0} &= \intd{s} \frac{\mu(s)}{(q_a+s)^2}.
-\end{aligned}
-\end{equation}
-%
-The second of these is the derivative of the first with respect to $q$, therefore the two curves in the first must touch each other tangentially at $q_a$.
-As both curves are decreasing convex functions, it seems unlikely that this could happen at multiple values of $q$.
-We will assume that there is a single non-zero $\CI_a=\CI$, with the corresponding $q_a=q$, in the following.
-
-For \eqref{eq:envI} to have a solution, $\mu(s)$ must be nonzero for at least one value of $s$.
-Then, by \eqref{eq:envKT}, we must saturate \eqref{eq:envKTbnd} at those values of $s$.
-If there were more than two such values, $\CI$ and $q$ would be overconstrained.
-Assume there are two such values, $s=s_0$ and $s=s_1$.
-%
-\begin{equation}\label{eq:sattwice}
-\begin{aligned}
-  \CI = B(s_0)(q+s_0) = B(s_1)(q+s_1) 
-  \qquad
-  \implies q = \frac{s_0 B(s_0) - s_1 B(s_1)}{B(s_1)-B(s_0)}.
-\end{aligned}
-\end{equation}
-%
-Now the bound \eqref{eq:envKTbnd} reads
-%
-\begin{equation}\label{eq:satwicebnd}
-\begin{aligned}
-  B(s) \geq \frac{B(s_0)(q+s_0)}{q+s} &= \frac{s_0-s_1}{(s-s_1)C(s_0)+(s_0-s)C(s_1)}, \\
-  xC(s_0)+(1-x)C(s_1) &\geq C(xs_0+(1-x)s_1),
-\end{aligned}
-\end{equation}
-%
-where we substituted $s=xs_0+(1-x)s_1$.
-As we assumed that $C(s)$ is concave, this is impossible.
-
-Therefore the bound must be saturated exactly once and $\mu(s)=\mu_0\,\delta(s-s_0)$.
-Now \eqref{eq:envI}, \eqref{eq:envq} and \eqref{eq:envKT} read
-%
-\begin{equation}\label{eq:satonce}
-  \mu_0 = (q+s_0)\,\e^{-qt_0},
-  \qquad
-  (q+s_0)t_0 = 1,
-  \qquad
-  \CI = \frac{B(s_0)}{t_0}.
-\end{equation}
-%
-This satisfies \eqref{eq:envKTm}.
-Now \eqref{eq:envKTbnd} tells us
-%
-\begin{equation}\label{eq:satoncebnd}
-\begin{aligned}
-  B(s) \geq \frac{B(s_0)}{(q+s)t_0}, 
-  \qquad\text{or}\qquad
-  1 + (s-s_0)t_0 \geq \frac{C(s)}{C(s_0)}.
-\end{aligned}
-\end{equation}
-%
-As the two sides are equal at $s=s_0$, it is necessary that they touch tangentially for this bound to be true elsewhere.
-The linearity of the LHS and concavity of the RHS ensure that this is sufficient for the bound to be satisfied for all $s$.
-This condition is
-%
-\begin{equation}\label{eq:envs0t0}
-  t_0 = \frac{C'(s_0)}{C(s_0)}.
-\end{equation}
-%
-
-The resulting SNR curve and envelope are
-%
-\begin{equation}\label{eq:env}
-  \snr(t) = \frac{\sqrt{N}\,\e^{(s_0-1/t_0)t}}{C'(s_0)},
-  \qquad
-  \snr(t_0) = \frac{\sqrt{N}\,\e^{s_0t_0-1}}{C'(s_0)},
-\end{equation}
-%
-where $s_0$ is a function of $t_0$ determined by \eqref{eq:envs0t0}.
-
-\appendix
-\section{Other quantities}
-
-Here we collect useless facts.
-
-\subsection{Differences in \texorpdfstring{$c_k$}{c(k)}'s}\label{sec:areacoeffdiff}
-
-In the general case, making full use of \cite[Cor.6.2.7]{kemeny1960finite}, we can show that:
-%
-\begin{equation}\label{eq:areacoeffdiff}
-  \begin{aligned}
-    c_{k+1} - c_k &= \frac{N^{k+1}_{kk}}{\eq_k} \brk{
-      \sum_{i<j} \eq_i \frac{\frg_{ij}}{f^+} \prn{H^{k+1}_{i,k} + H^{k}_{j,k+1} - 1}
-\right.\\&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\left.
-      +\sum_{i>j} \eq_i \frac{\frg_{ij}}{f^+} \prn{H^{k+1}_{j,k} + H^{k}_{i,k+1} - 1}
-    } \\
-    &= \frac{N^{k+1}_{kk}}{\eq_k} \brk{
-      \sum_{i<j} \eq_i \frac{\frg_{ij}}{f^+} \prn{1 - H^{k+1}_{j,k} + H^{k}_{i,k+1}}
-\right.\\&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\left.
-      +\sum_{i>j} \eq_i \frac{\frg_{ij}}{f^+} \prn{1 - H^{k+1}_{i,k} + H^{k}_{j,k+1}}
-    }, \\
-    \eta^+_{k} - \eta^+_{k+1} &=
-      \fpt_{k,k+1} \eq_+ - \sum_{\set{i}{\w_i>0}} N^{k+1}_{ii} \prn{1 - H^{i}_{k,k+1}} \\&=
-      \sum_{\set{i}{\w_i>0}} N^{k}_{ii} \prn{1 - H^{i}_{k+1,k}} - \fpt_{k+1,k} \eq_+ \\
-    = \eta^-_{k+1} - \eta^-_{k} &=
-      \fpt_{k+1,k} \eq_- - \sum_{\set{i}{\w_i<0}} N^{k}_{ii} \prn{1 - H^{i}_{k+1,k}} \\&=
-      \sum_{\set{i}{\w_i<0}} N^{k+1}_{ii} \prn{1 - H^{i}_{k,k+1}} - \fpt_{k,k+1} \eq_- .
-  \end{aligned}
-\end{equation}
-%
-We need to show that the positivity of one ($\forall k$) implies that the positivity of the other.
-
-\subsection{Derivatives \wrt \texorpdfstring{$\fptb$}{T}}\label{sec:derivT}
-
-The Markov chain, $\frg$, is completely determined by the off-diagonal mean first passage times, $\fptb$, \cite[Th.4.4.12]{kemeny1960finite}:
-%
-\begin{equation}\label{eq:fptdet}
-  \begin{aligned}
-    \eq &= \eta \prn{\fptb\inv\onev}\trans, \\
-    r\frg= \MM &= (\D-\onem)\fptb\inv.
-  \end{aligned}
-\end{equation}
-%
-where $\eta$ can be determined by normalising $\eq$ and $\D$ is determined by $\eq$, \eqref{eq:pdotD}. Clearly $q$ will be undetermined.
-
-We can then define derivatives \wrt $\fptb$:
-%
-\begin{equation}\label{eq:derivTdef}
-  \pdiff{}{\fptb_{gh}} = \sum_{ij} \pdiffc{\frg_{ij}}{\fptb_{gh}} \pdiff{}{\frg_{ij}},
-\end{equation}
-%
-where
-%
-\begin{equation}\label{eq:derivTWp}
-  \pdiff{\frg_{ij}}{\fptb_{gh}} = r \eq_h \frg_{ig} \prn{\frg_{ij}-\frg_{hj}} ,
-  \qquad
-  \pdiff{\eq_{i}}{\fptb_{gh}} = -r \eq_h \eq_i \frg_{ig} .
-\end{equation}
-%
-These derivatives of $q$ will vanish.
-
-This means that the derivatives of the area are:
-%
-\begin{equation}\label{eq:derivTarea}
-  \pdiff{\hat{A}}{\fptb_{gh}} = -\eq_h\brk{
-    \sum_i \eq_i \enc_{ig} \w_h
-    + r\sum_{ijk} \prn{\frg_{ig}+\frg_{kg}} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}} \eq_k \w_k
-    }.
-\end{equation}
-%
-
-\subsection{Initial SNR}\label{sec:initSNR}
-
-Let us write $\SNR(0) = \sqrt{N} (2f^+f^-) \hat{I}$. Using \eqref{eq:SNRcurve}, we can write
-%
-\begin{equation}\label{eq:initSNR}
-  \begin{aligned}
-    \hat{I} &= \diff{\eq}{\alpha}(-\frg)\w &
-      &= - \pib \fund \diff{\MM}{\alpha} \fund \frg \w &
-      &= - \pib \fund \diff{\frg}{\alpha} \fund \MM \w \\
-      &= \pib \fund \diff{\frg}{\alpha} (\I - \tau\onev\pib) \w &
-      &= \pib \fund \diff{\frg}{\alpha} \w &
-      &= \eq \diff{\frg}{\alpha} \w .
-  \end{aligned}
-\end{equation}
-%
-We can also write this in component form
-%
-\begin{equation}\label{eq:initSNRcmpt}
-  \begin{aligned}
-     \hat{I} &= \sum_{i,j} \eq_i \enc_{ij} \w_j \\
-      &= \sum_{i \neq j} \eq_i \enc_{ij} \w_j + \sum_i \eq_i \enc_{ii} \w_i \\
-      &= \sum_{i \neq j} \enc_{ij} \eq_i (\w_j - \w_i).
-  \end{aligned}
-\end{equation}
-%
-We can differentiate the expression in terms of $\fund$ to get
-%
-\begin{equation}\label{eq:initSNRdiff}
- \begin{aligned}
-  \pdiff{\hat{I}}{\W^\pm_{gh}}
-    &= rf^\pm \sum_{ijk} \pib_k \fund_{kg} (\fund_{hi}-\fund_{gi}) \enc_{ij} \w_j
-    \pm \sum_k \pib_k \fund_{kg} (\w_h - \w_g) \\
-    &= rf^\pm \sum_{ij} \eq_g \enc_{ij} (\fptb_{gi}-\fptb_{hi}) \eq_i \w_j
-    \pm \eq_g (\w_h - \w_g).
- \end{aligned}
-\end{equation}
-%
-If we wished to maximise the area with fixed initial SNR, $\hat{I}=I_0$, we would add $\beta(\hat{I}-I_0)$ to the Lagrangian \eqref{eq:lagrangian} and extremise wrt $\beta$.
-
-\subsection{Fixing mixing time (Kemeny's  constant)}\label{sec:fixmix}
-
-The mixing time, $\eta$ \eqref{eq:mixdef}, can be thought of as the mean time to pass between two states drawn from the equilibrium distribution \cite{hunter2006mixing,levene2002kemeny}.
-It should be a measure of the time it takes to reach equilibrium, and therefore related to the time it takes the SNR to decay.
-It could be interesting to hold it fixed while maximising the area, as this could help prevent degenerate solutions with very low initial SNR but long decay time.
-This can be done by adding $\gamma(\eta-\eta_0)$ to the Lagrangian \eqref{eq:lagrangian} and extremising wrt $\gamma$.
-
-It would help to calculate its derivatives.
-%
-\begin{equation}\label{eq:mixingdiff}
-  \begin{aligned}
-    \pdiff{\eta}{\W^\pm_{gh}} &= \pdiff{}{\W^\pm_{gh}}\brk{\sum_i \fund_{ii} - \tau} \\
-      &= rf^\pm \sum_i \fund_{ig}(\fund_{hi}-\fund_{gi}) \\
-      &= rf^\pm (\fund^2_{hg} - \fund^2_{gg}).
-  \end{aligned}
-\end{equation}
-%
-
-\subsection{Scaling mode}\label{sec:rescaleconstr}
-
-Consider the following differential operator:
-%
-\begin{equation}\label{eq:scaleopapp}
-  \Delta \equiv \sum_{g,h} \W^+_{gh}\pdiff{}{\W^+_{gh}} + \W^-_{gh}\pdiff{}{\W^-_{gh}}.
-\end{equation}
-%
-This corresponds to the scaling, $\W^\pm \to (1+\epsilon)\W^\pm$.
-Intuitively, this has two effects: it scales up the initial potentiation/depression and it scales down all timescales.
-This intuition is confirmed by the following results:
-%
-\begin{equation}\label{eq:scaleeffectsapp}
-  \begin{aligned}
-    \Delta \fund &= \tau\onev\eq - \fund ,&
-    \Delta \eq  &= 0 ,&
-    \Delta \fpt  &= -\fpt ,&
-    \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
-    \Delta \enc_{ij} &= \enc_{ij} ,
-    \\&&
-    \Delta \hat{A}  &= 0 ,&
-    \Delta \hat{I} &= \hat{I} ,&
-    \Delta \eta  &= -\eta .
-  \end{aligned}
-\end{equation}
-%
-The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W^\pm$.
-
-Suppose we don't hold anything fixed when maximising the area. Then we'd have
-%
-\begin{equation*}
-  \Delta \CL = 0,
-\end{equation*}
-%
-i.e. this scaling is a null direction. Area maximisation can only determine the ratios of transition rates, not their absolute values.
-
-Now suppose we only fix $\hat{I}$ when maximising the area. Then we'd have
-%
-\begin{equation*}
-  \Delta \CL = \beta \hat{I} = \beta I_0 = 0.
-\end{equation*}
-%
-As the Lagrange multiplier is zero, the constraint doesn't change the derivatives of the Lagrangian wrt $\W^\pm$.
-The sole effect of the constraint would be to fix the scaling mode.
-It wouldn't change the ratios of transition rates, topology, etc.
-
-Exactly the same conclusions would follow if we only fixed $\eta$ when maximising the area.
-This would also only fix the scaling mode.
-However, if we held them both fixed, it would have more of an effect:
-%
-\begin{equation}\label{eq:scaleconstr}
-  \Delta \CL = \beta \hat{I} - \gamma \eta = \beta I_0 - \gamma \eta_0 = 0.
-\end{equation}
-%
-The Lagrange multipliers are non-zero, so this could affect the topology of the optimal Markov chain.
+Different parts of the brain, that store memories for different timescales, could be characterised by different values of $\tau$.
 
@@
 \bibliographystyle{utcaps_sl}
-\bibliography{maths,neuro}
+\bibliography{maths,neuro,markov}
 

### 2015-05-29 Notes/laplace.tex
defn of snr and dof

@@
 \newcommand{\MM}{\mathbf{Q}}
+%prob distribution
+\newcommand{\pr}{\mathbf{p}}
 %equilibrium distribution
-\newcommand{\eq}{\mathbf{p}^\infty}
+\newcommand{\eq}{\pr^\infty}
 %first passage times
@@
 \newcommand{\M}{\mathbf{M}}
-\newcommand{\enc}{\mathbf{q}}
+\newcommand{\enc}{\mathbf{K}}
 \newcommand{\frg}{\W^{\mathrm{F}}}
+\renewcommand{\hom}{\mathbf{H}}
+\newcommand{\Mh}{\widetilde{\M}}
+\newcommand{\frgh}{\widetilde{\W}^{\mathrm{F}}}
+\newcommand{\ench}{\widetilde{\enc}}
 \newcommand{\F}{\boldsymbol{\Phi}}
@@
 \DeclareMathOperator{\snr}{SNR}
+\newcommand{\snrb}{\overline{\snr}}
 \DeclareMathOperator{\env}{Env}
@@
 \newcommand{\comp}{^\mathrm{c}}
+%superscripts
+\newcommand{\pot}{^{\text{pot}}}
+\newcommand{\dep}{^{\text{dep}}}
+\newcommand{\potdep}{^{\text{pot/dep}}}
 % ----------------------------------------------------------------
@@
 
-\section{Recognition memory}\label{sec:recog}
+\section{Framework}\label{sec:framework}
+
+\subsection{Recognition memory}\label{sec:recog}
 
@@
 In the other case, it is beneficial to increase it.
-The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events and the probability distribution of the sequence of plasticity events.
+The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.
 
@@
 Different parts of the brain, that store memories for different timescales, could be characterised by different values of $\tau$.
+
+\subsection{Signal-to-noise ratio}\label{sec:snr}
+
+We will model the synapses as having two possible synaptic weights.
+As we are looking at the information contained in the synaptic weights, rather than modelling the readout via electrical activity of the neurons, what values the synaptic weights actually take is irrelevant.
+It will be convenient to call these two values $\pm1$.
+
+As discussed in \autoref{sec:recog}, the signal-to-noise ratio is given by
+%
+\begin{equation}\label{eq:SNRdef}
+  \snr(t) = \frac{\av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}}
+     {\sqrt{\var(\syn_\text{ideal}\cdt\syn(\infty))}},
+\end{equation}
+%
+Where the averages and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.
+If we also average over the time between storage and recognition, we have
+%
+\begin{equation}\label{eq:snrb}
+  \snrb(\tau) = \intd[_0^\infty]{t} P(t\vert\tau) \snr(t) 
+   = \frac{1}{\tau} \intd[_0^\infty]{t} \e^{-t/\tau} \snr(t),  
+\end{equation}
+%
+This is similar to the average SNR up to time $\tau$, but with the hard cut-off replaced with a smooth exponential.
+This can be expressed in terms of he Laplace transform of the SNR curve in \eqref{eq:SNRdef}:
+%
+\begin{equation}\label{eq:laplacedef}
+  A(s) = \intd[_0^\infty]{t} \e^{-st} \snr(t).
+  \qquad
+  \snrb(\tau) = \frac{A(1/\tau)}{\tau},
+\end{equation}
+%
+
+First, let's look at the variances, remembering that the states and plasticity events of each synapse are independent and identically distributed:
+%
+\begin{equation}\label{eq:noise}
+\begin{aligned}
+  \var(\syn_\text{ideal}\cdt\syn(t))
+    &= \sum_{\alpha\beta} \av{\syn^\alpha_\text{ideal}\syn^\alpha(t) \syn^\beta_\text{ideal}\syn^\beta(t)}
+    - \prn{\sum_\alpha \av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}}^2 \\
+    &= \sum_{\alpha} \av{(\syn^\alpha_\text{ideal})^2(\syn^\alpha(t))^2}
+    + \sum_{\alpha\neq\beta} \av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}\av{\smash{\syn^\beta}_\text{ideal}\smash{\syn^\beta}(t)}
+    \\&\phantom{= \sum_{\alpha}  \av{(\syn^\alpha_\text{ideal})^2(\syn^\alpha(t))^2}}
+    - \prn{\sum_\alpha \av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}}^2  \\
+    &= N\av{1}
+    + N(N-1)\av{\syn^1_\text{ideal}\syn^1(t)}^2
+    - N^2\av{\syn^1_\text{ideal}\syn^1(t)}^2 \\
+    &= N(1-\av{\syn^1_\text{ideal}\syn^1(t)}^2),
+\end{aligned}
+\end{equation}
+%
+where we used $\syn^\alpha=\pm1$.
+We can compute $\var(\syn_\text{ideal}\cdt\syn(\infty))$ by taking $t\to\infty$.
+
+For the numerator, we can write
+%
+\begin{equation}\label{eq:overlapex}
+  \av{\syn_\text{ideal}\cdt\syn(t)} = \sum_\alpha\av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}
+   = N \av{\syn^1_\text{ideal}\syn^1(t)},
+\end{equation}
+%
+If the elements of $\synid$ take values $\pm1$ with probability $f\potdep$,
+%
+\begin{equation}\label{eq:overlapone}
+  \av{\syn^1_\text{ideal}\syn^1(t)} = f\pot \av{\syn^1(t)}_{\text{pot},t=0} - f\dep \av{\syn^1(t)}_{\text{dep},t=0} 
+\end{equation}
+%
+To compute this quantity, we will need to discuss how we model individual synapses.
+
+\subsection{Markov models of synapses}\label{sec:markovsynapse}
+
+We model a synapse as having $M$ internal states.
+The synaptic weight depends deterministically on the state, given by the $M$-element column vector $\w$.
+We denote the probability distribution across these states by the row vector $\pr(t)$.
+We denote a column vector of ones by $\onev$, so the normalisation condition is $\pr(t)\onev=1$.
+
+When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M\potdep$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$.
+In between these plasticity events the synapse will undergo stochastic transitions described by a matrix $\hom$ whose $i,j$ element is the transition rate from state $i$ to state $j$ for $i \neq j$ and $\hom_{ii}=-\sum_{j \neq i} \hom_{ij}$.
+This allows us to describe homeostatic plasticity as well as plasticity processes that occur over a period of time -- $\M\potdep$ puts the synapse into some state where the process $\hom$ takes over and does the rest.
+The matrices $\M\potdep$ are transition matrices of discrete time Markov processes, whereas $\hom$ is the transition matrix of a continuous time Markov process.
+Their elements must satisfy the following constraints
+%
+\begin{equation}\label{eq:constr}
+\begin{aligned}
+  \M\potdep_{ij} &\geq 0, &\qquad
+  \sum_j \M\potdep_{ij} &= 1, \\
+  \hom_{i \neq j} &\geq 0, &\qquad
+  \sum_j \hom_{ij} &= 0. 
+\end{aligned}
+\end{equation}
+%
+The upper bounds $\M\potdep_{ij} \leq 1$ and $\hom_{ii} \leq 0$ follow automatically from these.
+There is no upper bound on $\hom_{i \neq j}$ (or lower bound on $\hom_{ii}$) as these are rates rather than probabilities.
+
+Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states is
+%
+\begin{equation}\label{eq:plastseq}
+  \pr(t_n) = \pr(t_0) \e^{\hom(t_1-t_0)} \M\pot \e^{\hom(t_2-t_1)} \M\dep \e^{\hom(t_3-t_4)} \M\dep \ldots \M\pot \e^{\hom(t_n-t_{n-1})}.
+\end{equation}
+%
+In computing the expectation in \eqref{eq:overlapone} the plasticity event that occurs at $t=0$ is fixed, but all of the previous and subsequent events are marginalised over.
+If the plasticity events occur at Poisson rate $r$, with potentiation and depression occurring independently with probability $f\potdep$, averaging over all previous and subsequent events gives
+%
+\begin{equation}\label{eq:plastseqav}
+  P(\text{state}=i,t \mid \text{pot/dep},0) = \brk{ \eq \M\potdep\, \e^{r(f\pot\M\pot+f\dep\M\dep-\I)t +\hom t} }_i,
+\end{equation}
+%
+where $\eq$ is the steady state distribution of the continuous time transition matrix that appears in the exponential:
+%
+\begin{equation}\label{eq:eq}
+  \eq(r\frg +\hom) = 0,
+  \qquad\text{where}\quad
+  \frg = f\pot\M\pot + f\dep\M\dep - \I.
+\end{equation}
+%
+
+This results in
+%
+\begin{equation}\label{eq:overlapwt}
+\begin{aligned}
+  \av{\syn^1_\text{ideal}\syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{(r\frg+\hom)t} \w, \\
+  \av{\syn^1_\text{ideal}\syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
+         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
+         &=  (f\pot-f\dep)\, \eq \w \\
+         &=  (f\pot-f\dep)\, \eq \e^{(r\frg+\hom)t} \w .
+\end{aligned}
+\end{equation}
+%
+Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
+%
+\begin{equation}\label{eq:signal}
+\begin{aligned}
+  \av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}
+    &= N \eq \enc \, \e^{(r\frg+\hom)t} \w , \\
+    \text{where} \qquad
+  \enc &= f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
+\end{aligned}
+\end{equation}
+%
+Combining with \eqref{eq:noise} gives
+%
+\begin{equation}\label{eq:SNRcurveExact}
+  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{(r\frg+\hom)t}\w}
+                 {\sqrt{1-(f\pot-f\dep)^2(\eq_+-\eq_-)^2}}.
+\end{equation}
+%
+We can absorb $\hom$ into $\MM\potdep$ by defining
+%
+\begin{equation}\label{eq:absorbhom}
+\begin{aligned}
+  \Mh\potdep &= \M\potdep + \frac{\hom}{2rf\potdep},
+  &\qquad
+  \frgh &= f\pot\Mh\pot + f\dep\Mh\dep - \I,
+  \\&&
+  \ench &= f\pot(\Mh\pot-\I)-f\dep(\Mh\dep-\I).
+\end{aligned}
+\end{equation}
+%
+Then the SNR curve is
+%
+\begin{equation}\label{eq:SNRcurvehom}
+  \SNR(t) = \frac{\sqrt{N}  \eq \ench \,\e^{rt\frgh}\w}
+                 {\sqrt{1-(f\pot-f\dep)^2(\eq_+-\eq_-)^2}}.
+\end{equation}
+%
+The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
+We can perform the maximisation in two steps.
+First maximise the numerator at fixed $\eq\w$.
+Then maximise the ratio \wrt $\eq\w$.
+This will allow us to ignore the denominator until the very end.
+
+The constraints on $\Mh\potdep$ are less stringent than the constraints on $\M\potdep$, \eqref{eq:constr}.
+Clearly
+%
+\begin{equation}\label{eq:constrhoffdiag}
+  \Mh\potdep_{i \neq j} \geq 0, \qquad
+  \sum_j \Mh\potdep_{ij} = 1.
+\end{equation}
+%
+The ``off-diagonal'' constraints are more complicated.
+There can be many $\M\potdep,\hom$ that result in the same $\Mh\potdep$.
+However, the fact that $\M\potdep_{ij} \in [0,1]$ and $\hom_{i\neq j} \geq 0$ imply that $\forall (i \neq j)$:
+%
+\begin{equation}\label{eq:homrange}
+  2r\max\brc{f\pot(\Mh\pot_{ij}-1),f\dep(\Mh\dep_{ij}-1)} \leq \hom_{ij} \leq 2r\min\brc{f\pot\Mh\pot_{ij},f\dep\Mh\dep_{ij}}
+\end{equation}
+%
+These two bounds are consistent when
+%
+\begin{equation}\label{eq:homconsistent}
+  -f\dep \leq f\pot\Mh\pot_{ij} - f\dep\Mh\dep_{ij} \leq f\pot
+  \qquad \forall i\neq j.
+\end{equation}
+%
+To ensure that $\M\potdep_{ii} \geq 0$, we also need to ensure that $\sum_{j\neq i}\M\potdep_{ij} \leq 1$.
+%
+\begin{equation}\label{eq:homdiag}
+\begin{aligned}
+  \sum_{j\neq i} \M\pot_{ij} 
+    &= \sum_{j\neq i} \prn{ \Mh\pot_{ij} - \frac{\hom_{ij}}{2rf\pot} }\\
+    &\geq \sum_{j\neq i} \prn{ \Mh\pot_{ij} - \frac{1}{f\pot}  \min\brc{ f\pot\Mh\pot_{ij},f\dep\Mh\dep_{ij} } }\\
+    &= \frac{1}{f\pot} \sum_{j\neq i} \brk{ f\pot\Mh\pot_{ij}-f\dep\Mh\dep_{ij} }_+ , \\
+  \sum _{j\neq i} \M\dep_{ij} 
+    &= \frac{1}{f\dep} \sum_{j\neq i} \brk{ f\dep\Mh\dep_{ij}-f\pot\Mh\pot_{ij} }_+ ,
+\end{aligned}
+\end{equation}
+%
+where rectification function is $[x]_\pm = x \Theta(\pm x)$, in terms of the Heaviside step function.
+Demanding that these quantities are at most 1 is sufficient to ensure that \eqref{eq:homconsistent} is satisfied.
+
+Therefore we find the set of constraints
+%
+\begin{equation}\label{eq:constrh}
+\begin{aligned}
+  \Mh\potdep_{i \neq j} &\geq 0, &\qquad
+  \sum_{j\neq i} \brk{ f\pot\Mh\pot_{ij}-f\dep\Mh\dep_{ij} }_+ &\leq f\pot, \\
+  \sum_j \Mh\potdep_{ij} &=1, &
+  \sum_{j\neq i} \brk{ f\dep\Mh\dep_{ij}-f\pot\Mh\pot_{ij} }_+  &\leq f\dep.
+\end{aligned}
+\end{equation}
+%
+We showed that these are necessary conditions for the existence of $\M\potdep,\hom$ satisfying \eqref{eq:constr}.
+We can see that they are sufficient by choosing $\hom_{ij} = 2r\min\brc{f\pot\Mh\pot_{ij},f\dep\Mh\dep_{ij}}$ for $i \neq j$ and $\hom_{ii}=-\sum_{j \neq i} \hom_{ij}$.
+
+We can then treat the off-diagonal elements of $\Mh\potdep$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
 

### 2015-06-01 Notes/laplace.tex
comment
envelope section

@@
 \newcommand{\Lb}{\boldsymbol{\Lambda}}
+%synapse models
 \newcommand{\w}{\mathbf{w}}
-\newcommand{\uv}{\mathbf{u}}
-\newcommand{\vv}{\mathbf{v}}
 \newcommand{\W}{\mathbf{W}}
@@
 \newcommand{\frg}{\W^{\mathrm{F}}}
+\newcommand{\F}{\boldsymbol{\Phi}}
+%superscripts
+\newcommand{\pot}{^{\text{pot}}}
+\newcommand{\dep}{^{\text{dep}}}
+\newcommand{\potdep}{^{\text{pot/dep}}}
+%with activity indep processes
 \renewcommand{\hom}{\mathbf{H}}
@@
 \newcommand{\ench}{\widetilde{\enc}}
-\newcommand{\F}{\boldsymbol{\Phi}}
+%eigenvectors
+\newcommand{\evr}{\mathbf{u}}
+\newcommand{\evl}{\boldsymbol{\eta}}
+%snr curves etc
 \newcommand{\syn}{\vec{w}}
@@
 \newcommand{\comp}{^\mathrm{c}}
-%superscripts
-\newcommand{\pot}{^{\text{pot}}}
-\newcommand{\dep}{^{\text{dep}}}
-\newcommand{\potdep}{^{\text{pot/dep}}}
 % ----------------------------------------------------------------
@@
 
+If we treat the off diagonal elements as the independent degrees of freedom, with the off diagonal elements determined by the row sum constraints (second column of \eqref{eq:constr}), the constraints are
+%
+\begin{equation}\label{eq:constri}
+  \M\potdep_{i \neq j} \geq 0, \qquad
+  \sum_{j\neq i} \M\potdep_{ij} \leq 1, \qquad
+  \hom_{i \neq j} \geq 0.
+\end{equation}
+%
+The second of these will be referred to as the ``diagonal'' constraint.
+
 Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states is
@@
 
+This formula takes the same form as the one in our NIPS paper \cite{Lahiri2013synapse}, where the process $\hom$ was not included.
+However, the process $\hom$ still has an impact due to the constraints.
+
 The constraints on $\Mh\potdep$ are less stringent than the constraints on $\M\potdep$, \eqref{eq:constr}.
@@
 %
-The ``off-diagonal'' constraints are more complicated.
+The ``diagonal'' constraints are more complicated.
 There can be many $\M\potdep,\hom$ that result in the same $\Mh\potdep$.
@@
 %
+The second column of this will be referred to as the ``diagonal'' constraints.
+
 We showed that these are necessary conditions for the existence of $\M\potdep,\hom$ satisfying \eqref{eq:constr}.
@@
 We can then treat the off-diagonal elements of $\Mh\potdep$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
+The diagonal elements are determined by setting the row sums to 1.
+We then have
+%
+\begin{equation}\label{eq:derivpd}
+  \pdiff{\Mh\pot_{ij}}{\Mh\pot_{mn}{}} = \pdiff{\Mh\dep_{ij}}{\Mh\dep_{mn}{}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
+  \qquad
+  \pdiff{\Mh\pot_{ij}}{\Mh\dep_{mn}{}} = \pdiff{\Mh\dep_{ij}}{\Mh\pot_{mn}{}} = 0.
+\end{equation}
+%
+It will also be convenient to define the differential operators
+%
+\begin{equation}\label{eq:pertfe}
+  \pdiff{}{\frgh_{ij}{}} = \frac{1}{2f\pot} \pdiff{}{\Mh\pot_{ij}{}} + \frac{1}{2f\dep} \pdiff{}{\Mh\dep_{ij}{}},
+  \qquad
+  \pdiff{}{\ench_{ij}{}} = \frac{1}{2f\pot} \pdiff{}{\Mh\pot_{ij}{}} - \frac{1}{2f\dep} \pdiff{}{\Mh\dep_{ij}{}},
+\end{equation}
+%
+these satisfy
+%
+\begin{equation}\label{eq:derivfe}
+  \pdiff{\frgh_{ij}}{\frgh_{mn}{}} = \pdiff{\ench_{ij}}{\ench_{mn}{}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
+  \qquad
+  \pdiff{\frgh_{ij}}{\ench_{mn}{}} = \pdiff{\ench_{ij}}{\frgh_{mn}{}} = 0.
+\end{equation}
+%
+
+\section{Laplace transform}\label{sec:laplace}
+
+\subsection{Fundamental matrix \etc}\label{sec:lfund}
+
+In analogy to the generalised fundamental matrix of a Markov chain \cite{Kemeny1981fund}, define
+%
+\begin{equation}\label{eq:lfund}
+  \fund(s) = (s\I +\onev\pib -\MM)\inv,
+\end{equation}
+%
+where $\pib$ is an arbitrary row vector satisfying $\pib\onev = \frac{1}{\tau_0} \neq 0$.
+This reduces to the fundamental matrix at $s=0$.
+We can see that $\onev$ is an eigenvector:
+%
+\begin{equation}\label{eq:lfundrowsum}
+  \fund(s)\onev = \frac{\tau_0}{1+s\tau_0}\onev.
+\end{equation}
+%
+Suppose we have some row vector $\mathbf{a}$ such that $\mathbf{a}\onev=0$.
+Then $\mathbf{a}\fund(s)$ is independent of $\pib$:
+%
+\begin{equation}\label{eq:alfund}
+  \pdiff{(\mathbf{a}\fund(s))_i}{\pib_k} = (\mathbf{a}\fund(s)\onev) \fund_{ki}(s) 
+      = \frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
+\end{equation}
+%
+
+Then we also define
+%
+\begin{equation}\label{eq:lfptb}
+  \fptb(s) = (\onem\fund\dg(s)-\fund(s))\D,
+  \qquad
+  \fptb_{ij}(s) = \frac{\fund_{jj}(s)-\fund_{ij}(s)}{\eq_j}.
+\end{equation}
+%
+This reduces to the first passage time matrix at $s=0$, but at other $s$ there is (probably) no relation.
+
+Using the exact same proof as in the case $s=0$, we see that the quantity
+%
+\begin{equation}\label{eq:lkemeny}
+  \eta(s) = \sum_j \fptb_{ij}(s)\eq_j
+\end{equation}
+%
+is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau}{1+s\tau}$.
+We can the define
+%
+\begin{equation}\label{eq:lpmixtime}
+  \eta^{\pm}_i(s) = \sum_j \fptb_{ij}(s)\eq_j \prn{\frac{1\pm\w_j}{2}} =  \sum_{j\in\pm} \fptb_{ij}(s)\eq_j.
+\end{equation}
+%
+We can arrange the states in order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
+It will be more convenient to use the quantities
+%
+\begin{equation}\label{eq:lwpmixtime}
+  \eta^w_i(s) = \sum_j \fptb_{ij}(s)\eq_j \w 
+    = \eta^+_i(s) - \eta^-_i(s) =2\eta^+_i(s) - \eta(s) =\eta(s) - 2\eta^-_i(s) .
+\end{equation}
+%
+Arranging the states in order of decreasing $\eta^w_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
+These quantities can still be used when $\w$ takes more than two values.
+
+\subsection{Laplace transform of SNR curve}\label{sec:laplaceSNR}
+
+Consider the Laplace transform of the evolution operator:
+%
+\begin{equation}\label{eq:levol}
+  \mathbf{G}(s) = \intd[_0^{\infty}]{t} \e^{(\MM-s) t}.
+\end{equation}
+%
+For $\Re s>0$, we have
+%
+\begin{equation}\label{eq:levolcalc}
+  (s-\MM)\mathbf{G}(s) = \intd[_0^{\infty}]{t} (s-\MM)\e^{(\MM-s) t} = \brk{-\e^{(\MM-s) t}}_0^{\infty} = \I.
+\end{equation}
+%
+As $(s-\MM)$ is invertible for $\Re s>0$, because the real part of all eigenvalues of $\MM$ are nonpositive, we have
+%
+\begin{equation}\label{eq:levolres}
+  \mathbf{G}(s) = (s-\MM)\inv.
+\end{equation}
+%
+For $s=0$, we can avoid problems by replacing $\mathbf{G}(s) \to \fund(s)$.
+
+Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurvehom} (ignoring the denominator)
+%
+\begin{equation}\label{eq:lsnr}
+\begin{aligned}
+  A(s) &= \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) \\
+   &= \intd[_0^{\infty}]{t} \sqrt{N} \eq \ench \e^{(r\frgh-s)t} \w \\
+   &= \sqrt{N} \eq \ench \prn{s - r\frgh}\inv \w.
+\end{aligned}
+\end{equation}
+%
+This expression is ill-behaved at $s=0$.
+Thanks to \eqref{eq:alfund}, we can solve this by the replacement $\mathbf{G}(s) \to \fund(s)$, as $\ench\onev=0$.
+%
+\begin{equation}\label{eq:larea}
+\begin{aligned}
+  \hat{A}(s) = \frac{A(s)}{\sqrt{N}} &=  \eq \ench \fund(s) \w \\
+    &= \sum_{ijk} \eq_i \ench_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
+    &= \sum_{ij}  \eq_i \ench_{ij} (\eta^w_{i}(s) - \eta^w_{j}(s)).
+\end{aligned}
+\end{equation}
+%
+Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.
+
+In analogy with the case $s=0$, we define
+%
+\begin{equation}\label{eq:lareacoeffs}
+  \begin{aligned}
+    c_k(s) &= \sum_{ij} \eq_i \ench_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
+    a_i(s) &= \sum_j \eq_i \ench_{ij} (\eta^w_{i}(s) - \eta^w_{j}(s)),\\
+    \implies
+    \hat{A}(s) &= \sum_k c_k(s) \eq_k \w_k
+      = \sum_i a_i(s).
+  \end{aligned}
+\end{equation}
+%
+
+\subsection{Derivatives}\label{sec:lderiv}
+
+Using the same approach as the area bound in \cite{Lahiri2013synapse}, we find that the derivatives are
+%
+\begin{equation}\label{eq:lAdiff}
+\begin{aligned}
+  \pdiff{\hat{A}(s)}{\Mh\potdep_{mn}{}}
+     =&\, rf\potdep \eq_m \brk{ \sum_i (\fptb_{mi}(0)-\fptb_{ni}(0)) a_i(s)
+     +c_m(s) (\eta^w_m(s) - \eta^w_n(s)) } \\
+     &\pm f\potdep \eq_m (\eta^w_m(s) - \eta^w_n(s)).
+\end{aligned}
+\end{equation}
+%
+
+\section{Upper bounds}\label{sec:upperbnds}
+
+\subsection{Area bound}\label{sec:area}
+
+%
+\begin{equation}\label{eq:areadef}
+  \int[_0^\infty]{dt} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) = A(0).
+\end{equation}
+%
+
+The only effect of $\hom$ is in the diagonal constraints (second column of \eqref{eq:constrh} vs.\ second inequality in \eqref{eq:constri}).
+However, the area is invariant under the scaling mode:
+%
+\begin{equation}\label{eq:scale}
+  \Mh\potdep - \I \to \lambda (\Mh\potdep - \I).
+\end{equation}
+%
+This means we can always take a set of matrices that violate either version of the diagonal constraints and use \eqref{eq:scale} to create a set of matrices that satisfy them without changing the area.
+Therefore, including $\hom$ has no effect on the area bound and the result from \cite{Lahiri2013synapse} applies here as well.
+
+This means that 
+%
+\begin{equation}\label{eq:areabnd}
+  A(0) \leq \frac{\sqrt{N}(M-1)}{r}.
+\end{equation}
+%
+The model that saturates this bound is one of serial topology with $\eq$ concentrated symmtrically at the two end states.
+
+\subsection{Initial SNR}\label{sec:initial}
+
+%
+\begin{equation}\label{eq:initdef}
+  \snr(0) = \snrb(0) = \lim_{s\to\infty} sA(s).
+\end{equation}
+%
+
+%
+\begin{equation}\label{eq:initbnd}
+  \snr(0) \leq \sqrt{N},
+\end{equation}
+%
+at least when $f\pot=f\dep=\frac{1}{2}$.
+
+\subsection{Envelope}\label{sec:envelope}
+
+Now we introduce an eigenvector decomposition for $-\frgh$:
+%
+\begin{equation}\label{eq:eigendecomp}
+  \frgh = - \sum_a q_a \evr^a \evl^a,
+  \quad
+  \evl^a \evr^b = \delta_{ab},
+  \quad
+  \frgh \evr^a = -q_a \evr^a,
+  \quad
+  \evl^a \frgh = -q_a \evl^a.
+\end{equation}
+%
+Then we can write
+%
+\begin{equation}\label{eqsnreig}
+  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a},
+  \qquad
+  \snrb(\tau) = \sqrt{N}\sum_a \frac{\CI_a}{1+r\tau/\tau_a},
+\end{equation}
+%
+where
+%
+\begin{equation}\label{eq:snrcoeffs}
+  \CI_a = (\eq \ench \evr^a) (\evl^a \w),
+  \qquad
+  \tau_a = \frac{1}{q_a}.
+\end{equation}
+%
+From \eqref{eq:areabnd} and \eqref{eq:initbnd}, we see that they are subject to the constraints
+%
+\begin{equation}\label{eq:coeffconstr}
+  \sum_a \CI_a \tau_a \leq M-1,
+  \qquad
+  \sum_a \CI_a \leq 1.
+\end{equation}
+%
+No doubt there are many other constraints that these quantities must satisfy, but we can see what these ones imply for the $\snrb(\tau)$ by maximising it subject to these constraints.
+
+Consider the Lagrangian
+%
+\begin{equation}\label{eq:envlagrangian}
+  \CL = \sum_a \frac{\CI_a}{1+r\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} + \mu_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
+\end{equation}
+%
+The Karush-Kuhn-Tucker necessary conditions for a maximum are
+%
+\begin{equation}\label{eq:envKTcond}
+  \pdiff{\CL}{\CI_a} = 0, \qquad
+  \pdiff{\CL}{\tau_a} = 0, \qquad
+  \mu_\alpha \geq 0, \qquad
+  \pdiff{\CL}{\mu_\alpha} \geq 0, \qquad
+  \mu_\alpha\pdiff{\CL}{\mu_\alpha} = 0. 
+\end{equation}
+%
+These are solved by
+%
+\begin{equation}\label{eq:envsol}
+\begin{gathered}
+  \CI_1 = 1, \qquad
+  \CI_{a>1} = 0, \qquad
+  \tau_1 = M-1, \\
+  \mu_\CI = \frac{(M-1)^2}{(r\tau + (M-1))^2}, \qquad
+  \mu_\CA = \frac{r\tau}{(r\tau+(M-1))^2}. \qquad
+\end{gathered}
+\end{equation}
+%
+This leads to the envelope
+%
+\begin{equation}\label{eq:env}
+  \snrb(\tau) \leq \frac{\sqrt{N}(M-1)}{r\tau + (M-1)}.
+\end{equation}
+%
+This is the envelope (``memory frontier'') that we can prove.
+Finding more constraints would lower it.
+Clearly there have to be more constraints, as \eqref{eq:envsol} indicates that this would be achieved by the same model at all times.
+This model would have to saturate both the initial SNR bound \eqref{eq:initbnd} and the area bound \eqref{eq:areabnd}, whereas we saw that they were achieved by different models, which perform well at short/long timescales respectively.
 

### 2015-06-02 Code/Optimum/ModelOpts/FindOptChainHomL.m
optimisers with fp~=0.5

@@ -1,10 +1,10 @@
 function [ qv,A ] = FindOptChainHomL( s,n,varargin )
-%[Wp,Wm]=FINDOPTCHAINHOML(t,n,reps) Find synapse model that maximises SNR(t)
-%   t    = time value
+%[qv,A]=FINDOPTCHAINHOML(s,n,reps) Find synapse model that maximises SNR(t)
+%   s    = inverse time value
 %   n    = #states
 %   reps = number of attempts we max over
-%   Wp = potentiation transition rates
-%   Wm = depression transition rates
+%   qv = nearest neighbour transition rates
+%   A  = Laplace Transf value
 
 persistent p
 if isempty(p)
@@ -37,7 +37,7 @@ if r.reps==1
     %     [Wp,Wm] = ModelOpt( Wp,Wm,t,varargin{:});
         [ qv,A ]=ModelOptChainHomL( qv,s,p.Unmatched);
     catch ME
-        A=-OptFunChainHomL(qv,s);
+        A=-OptFunChainHomL(qv,s,0.5);
         disp(ME.message);
         disp(['In function: ' ME.stack(1).name ' at line ' int2str(ME.stack(1).line) ' of ' ME.stack(1).file]);
         return;

### 2015-06-02 Code/Optimum/ModelOpts/FindOptChainHomLC.m

@@ -1,6 +1,6 @@
 function [ qv,A ] = FindOptChainHomLC( s,n,varargin )
-%[Wp,Wm]=FINDOPTCHAINHOMLC(t,n,reps) Find synapse model that maximises SNR(t)
-%   t    = time value
+%[qv,A]=FINDOPTCHAINHOMLC(s,n,reps) Find synapse model that maximises SNR(t)
+%   s    = inverse time value
 %   n    = #states
 %   reps = number of attempts we max over
 %   Wp = potentiation transition rates
@@ -37,7 +37,7 @@ if r.reps==1
     %     [Wp,Wm] = ModelOpt( Wp,Wm,t,varargin{:});
         [ qv,A ]=ModelOptChainHomLC( qv,s,p.Unmatched);
     catch ME
-        A=-OptFunChainHomLC(qv,s);
+        A=-OptFunChainHomLC(qv,s,0.5);
         disp(ME.message);
         disp(['In function: ' ME.stack(1).name ' at line ' int2str(ME.stack(1).line) ' of ' ME.stack(1).file]);
         return;

### 2015-06-02 Code/Optimum/ModelOpts/FindOptChainSL.m

@@ -1,10 +1,10 @@
 function [ qv,A ] = FindOptChainSL( s,n,varargin )
-%[Wp,Wm]=FINDOPT(t,n,reps) Find synapse model that maximises SNR(t)
-%   t    = time value
+%[qv,A]=FINDOPTCHAINSL(s,n,reps) Find synapse model that maximises SNR(t)
+%   s    = inverse time value
 %   n    = #states
 %   reps = number of attempts we max over
-%   Wp = potentiation transition rates
-%   Wm = depression transition rates
+%   qv = nearest neighbour transition rates
+%   A  = Laplace Transf value
 
 persistent p
 if isempty(p)
@@ -32,7 +32,7 @@ if r.reps==1
     %     [Wp,Wm] = ModelOpt( Wp,Wm,t,varargin{:});
         [ qv,A ]=ModelOptChainSL( qv,s,p.Unmatched);
     catch ME
-        A=-OptFunChainSL(qv,s);
+        A=-OptFunChainSL(qv,s,0.5);
         disp(ME.message);
         disp(['In function: ' ME.stack(1).name ' at line ' int2str(ME.stack(1).line) ' of ' ME.stack(1).file]);
         return;

### 2015-06-02 Code/Optimum/ModelOpts/FindOptHomL.m

@@ -1,6 +1,6 @@
 function [ Wp,Wm,Q,A ] = FindOptHomL( sm,n,varargin )
-%[Wp,Wm]=FINDOPTHOML(sm,n,reps) Find synapse model that maximises A(sm)
-%   t    = time value
+%[Wp,Wm,Q,A]=FINDOPTHOML(sm,n,reps) Find synapse model that maximises A(sm)
+%   sm   = inverse time value
 %   n    = #states
 %   reps = number of attempts we max over
 %   Wp = potentiation transition rates

### 2015-06-02 Code/Optimum/ModelOpts/ModelOptChainHomLA.m

@@ -20,6 +20,7 @@ if isempty(p)
     p.addParameter('MaxIter',1000,@(x) validateattributes(x,{'numeric'},{'scalar','integer'},'ModelOptChainHomLA','TolFun'));
     p.addParameter('Algorithm','interior-point',@(x) validatestring(x,{'trust-region-reflective','active-set','interior-point','sqp'},'ModelOptChainHomLA','TolFun'));
     p.addParameter('Display','off',@(x) validatestring(x,{'off','iter','iter-detailed','notify','notify-detailed','final','final-detailed'},'ModelOptChainHomLA','TolFun'));
+    p.addParameter('fp',0.5,@(x) validateattributes(x,{'numeric'},{'scalar','nonnegative','<=',1},'ModelOptChainHomLA','fp'));
 end
 p.parse(varargin{:});
 r=p.Results;
@@ -38,11 +39,11 @@ options = optimset(p.Unmatched,'Algorithm',r.Algorithm,'Display',r.Display,...
 
 if r.UseDerivs
     options = optimset(options,'GradObj','on');
-    [newqv,A,ef] = fmincon(@(y)OptFunGradChainHomLA(y,sm),qv,...
+    [newqv,A,ef] = fmincon(@(y)OptFunGradChainHomLA(y,sm,r.fp),qv,...
         [],[],[],[],qmin,qmax,[],... 
         options);
 else
-    [newqv,A,ef] = fmincon(@(y)OptFunChainHomLA(y,sm),qv,...
+    [newqv,A,ef] = fmincon(@(y)OptFunChainHomLA(y,sm,r.fp),qv,...
         [],[],[],[],qmin,qmax,[],... 
         options);
 end

### 2015-06-02 Code/Optimum/OptFuns/OptFunChainHomL.m

@@ -1,4 +1,4 @@
-function [ A ] = OptFunChainHomL( qv,s )
+function [ A ] = OptFunChainHomL( qv,s,fp )
 %OPTFUNGRADCHAINHOML Laplace transform of SNR curve for asymmetric serial
 %model with homeostatic plasticity for fmincon
 %   Detailed explanation goes here
@@ -15,18 +15,21 @@ qhm=qv(3*n+1:end);
 [Hp,Hm]=MakeMultistate(qhp,qhm);
 H=Hp+Hm;
 
-p=(qp+qhp)./(qm+qhm);
+p=(fp*qp+qhp)./((1-fp)*qm+qhm);
 p=[1 cumprod(p)];
 p=p/sum(p);
 
+Wp=Wp+0.5/fp*H;
+Wm=Wm+0.5/(1-fp)*H;
+
 q=Wp-Wm;
 
-Zinv=ones(length(Wp))-Wm-0.5*q-H;
+Zinv=ones(length(Wp))-fp*Wp-(1-fp)*Wm;
 Zinvs=s*eye(length(Wp))+Zinv;
 
 a=q*(Zinvs\w);
 
-A=0.5*p*a;
+A=2*fp*(1-fp)*p*a;
 
 A=-A;
 

### 2015-06-02 Code/Optimum/OptFuns/OptFunGradChainHomL.m

@@ -1,4 +1,4 @@
-function [ A,gr ] = OptFunGradChainHomL( qv,s )
+function [ A,gr ] = OptFunGradChainHomL( qv,s,fp )
 %OPTFUNGRADCHAINHOML Laplace transform of SNR curve and gradient for
 %asymmetric serial model with homeostatic plasticity for fmincon
 %   Detailed explanation goes here
@@ -15,13 +15,16 @@ qhm=qv(3*n+1:end);
 [Hp,Hm]=MakeMultistate(qhp,qhm);
 H=Hp+Hm;
 
-p=(0.5*qp+qhp)./(0.5*qm+qhm);
+p=(fp*qp+qhp)./((1-fp)*qm+qhm);
 p=[1 cumprod(p)];
 p=p/sum(p);
 
+Wp=Wp+0.5/fp*H;
+Wm=Wm+0.5/(1-fp)*H;
+
 q=Wp-Wm;
 
-Zinv=ones(length(Wp))-Wm-0.5*q-H;
+Zinv=ones(length(Wp))-fp*Wp-(1-fp)*Wm;
 Zinvs=s*eye(length(Wp))+Zinv;
 
 a=q*(Zinvs\w);
@@ -30,7 +33,7 @@ c=(p*q)/Zinvs;
 Za=diff(Zinv\a)';
 Zw=diff(Zinvs\w)';
 
-A=0.5*p*a;
+A=2*fp*(1-fp)*p*a;
 
 % %dA(s)/dq_(i,i+1)
 % dAdq = p(1:end-1).*(diff(Zinvs\w))';
@@ -39,17 +42,22 @@ A=0.5*p*a;
 % %dA/dWp_(i,i+1)+dA/dWm_(n-i+1,n-i)
 % gr = 2*dAdq+dAdW;
 
-%dA/dWp_(i,i+1)+dA/dWm_(i,i+1)
-grhp= 0.5*p(1:end-1).*Za + 0.5*c(1:end-1).*Zw;
-%dA/dWp_(i+1,i)+dA/dWm_(i+1,i)
-grhm= -0.5*p(2:end).*Za -0.5*c(2:end).*Zw;
-%dA/dWp_(i,i+1)
-grp= grhp + p(1:end-1).*Zw;
-%dA/dWm_(i+1,i)
-grm= grhm + p(2:end).*Zw;
[…]

### 2015-06-02 Code/Optimum/OptFuns/OptFunGradChainL.m

@@ -1,19 +1,20 @@
-function [ A,gr ] = OptFunGradChainL( qv,s )
+function [ A,gr ] = OptFunGradChainL( qv,s,fp )
 %OPTFUNGRADCHAINL Laplace transform of SNR curve for asymmetric serial model
 %and gradient for fmincon
 %   Detailed explanation goes here
+
 
 qp=qv(1:length(qv)/2);
 qm=qv(length(qv)/2+1:end);
 [Wp,Wm,w]=MakeMultistate(qp,qm);
 
-p=qp./qm;
+p=fp/(1-fp)*qp./qm;
 p=[1 cumprod(p)];
 p=p/sum(p);
 
 q=Wp-Wm;
 
-Zinv=ones(length(Wp))-Wm-0.5*q;
+Zinv=ones(length(Wp))-fp*Wp-(1-fp)*Wm;
 Zinvs=s*eye(length(Wp))+Zinv;
 
 a=q*(Zinvs\w);
@@ -22,7 +23,7 @@ c=(p*q)/Zinvs;
 Za=diff(Zinv\a)';
 Zw=diff(Zinvs\w)';
 
-A=0.5*p*a;
+A=2*fp*(1-fp)*p*a;
 
 % %dA(s)/dq_(i,i+1)
 % dAdq = p(1:end-1).*(diff(Zinvs\w))';
@@ -32,11 +33,11 @@ A=0.5*p*a;
 % gr = 2*dAdq+dAdW;
 
 %dA/dWp_(i,i+1)+dA/dWm_(n-i+1,n-i)
-grp= 0.5*p(1:end-1).*Za + (p(1:end-1)+0.5*c(1:end-1)).*Zw;
-grm= -0.5*p(2:end).*Za + (p(2:end)-0.5*c(2:end)).*Zw;
+grp= fp*p(1:end-1).*Za + (p(1:end-1)+fp*c(1:end-1)).*Zw;
+grm= -(1-fp)*p(2:end).*Za + (p(2:end)-(1-fp)*c(2:end)).*Zw;
 
 A=-A;
-gr=-0.5*[grp grm];
+gr=-2*fp*(1-fp)*[grp grm];
 
 end
 

### 2015-06-02 Code/Optimum/OptFuns/OptFunGradChainSL.m

@@ -1,23 +1,23 @@
-function [ A,gr ] = OptFunGradChainSL( qv,s )
+function [ A,gr ] = OptFunGradChainSL( qv,s,fp )
 %OPTFUNGRADCHAINL Laplace transform of SNR curve for symmetric serial model
 %and gradient for fmincon
 %   Detailed explanation goes here
 
 [Wp,Wm,w]=MakeSMS(qv);
 
-p=qv./wrev(qv);
+p=fp*qv./wrev((1-fp)*qv);
 p=[1 cumprod(p)];
 p=p/sum(p);
 
 q=Wp-Wm;
 
-Zinv=ones(length(Wp))-Wm-0.5*q;
+Zinv=ones(length(Wp))-fp*Wp-(1-fp)*Wm;
 Zinvs=s*eye(length(Wp))+Zinv;
 
 a=q*(Zinvs\w);
 c=(p*q)/Zinvs;
 
-A=0.5*p*a;
+A=2*fp*(1-fp)*p*a;
 
 % %dA(s)/dq_(i,i+1)
 % dAdq = p(1:end-1).*(diff(Zinvs\w))';
@@ -30,7 +30,7 @@ A=0.5*p*a;
 gr= p(1:end-1).*diff(Zinv\a)' + (2*p(1:end-1)+c(1:end-1)).*diff(Zinvs\w)';
 
 A=-A;
-gr=-gr;
+gr=-2*fp*(1-fp)*gr;
 
 end
 

### 2015-06-02 Notes/laplace.tex
optimisers with fp~=0.5
perturbations

@@
 \newcommand{\comp}{^\mathrm{c}}
+%avoid double superscripts
+\renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
+
 % ----------------------------------------------------------------
@@
   \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{(r\frg+\hom)t}\w}
-                 {\sqrt{1-(f\pot-f\dep)^2(\eq_+-\eq_-)^2}}.
+                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
 \end{equation}
@@
   \SNR(t) = \frac{\sqrt{N}  \eq \ench \,\e^{rt\frgh}\w}
-                 {\sqrt{1-(f\pot-f\dep)^2(\eq_+-\eq_-)^2}}.
+                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
 \end{equation}
@@
 \begin{equation}\label{eq:derivpd}
-  \pdiff{\Mh\pot_{ij}}{\Mh\pot_{mn}{}} = \pdiff{\Mh\dep_{ij}}{\Mh\dep_{mn}{}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
+  \pdiff{\Mh\pot_{ij}}{\Mh\pot_{mn}} = \pdiff{\Mh\dep_{ij}}{\Mh\dep_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
   \qquad
-  \pdiff{\Mh\pot_{ij}}{\Mh\dep_{mn}{}} = \pdiff{\Mh\dep_{ij}}{\Mh\pot_{mn}{}} = 0.
+  \pdiff{\Mh\pot_{ij}}{\Mh\dep_{mn}} = \pdiff{\Mh\dep_{ij}}{\Mh\pot_{mn}} = 0.
 \end{equation}
@@
 \begin{equation}\label{eq:pertfe}
-  \pdiff{}{\frgh_{ij}{}} = \frac{1}{2f\pot} \pdiff{}{\Mh\pot_{ij}{}} + \frac{1}{2f\dep} \pdiff{}{\Mh\dep_{ij}{}},
+  \pdiff{}{\frgh_{ij}} = \frac{1}{2f\pot} \pdiff{}{\Mh\pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\Mh\dep_{ij}},
   \qquad
-  \pdiff{}{\ench_{ij}{}} = \frac{1}{2f\pot} \pdiff{}{\Mh\pot_{ij}{}} - \frac{1}{2f\dep} \pdiff{}{\Mh\dep_{ij}{}},
+  \pdiff{}{\ench_{ij}} = \frac{1}{2f\pot} \pdiff{}{\Mh\pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\Mh\dep_{ij}},
 \end{equation}
@@
 \begin{equation}\label{eq:derivfe}
-  \pdiff{\frgh_{ij}}{\frgh_{mn}{}} = \pdiff{\ench_{ij}}{\ench_{mn}{}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
+  \pdiff{\frgh_{ij}}{\frgh_{mn}} = \pdiff{\ench_{ij}}{\ench_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
   \qquad
-  \pdiff{\frgh_{ij}}{\ench_{mn}{}} = \pdiff{\ench_{ij}}{\frgh_{mn}{}} = 0.
+  \pdiff{\frgh_{ij}}{\ench_{mn}} = \pdiff{\ench_{ij}}{\frgh_{mn}} = 0.
+\end{equation}
+%
+
+If we wish to use the off-diagonal elements of $\frgh$ and $\ench$ as the degrees of freedom, the constraints are
+%
+\begin{equation}\label{eq:constrfe}
+\begin{aligned}
+  \frgh_{i \neq j}   &\geq \abs{ \ench_{i \neq j} }, &\qquad
+  \sum_{j\neq i} \brk{ \ench_{ij} }_+ &\leq f\pot, \\
+  \sum_j \frgh_{ij} &=0, &
+  \sum_{j\neq i} \brk{ \ench_{ij} }_- &\geq -f\dep, \\
+  \sum_j \ench_{ij} &=0.
+\end{aligned}
 \end{equation}
@@
 \begin{equation}\label{eq:alfund}
-  \pdiff{(\mathbf{a}\fund(s))_i}{\pib_k} = (\mathbf{a}\fund(s)\onev) \fund_{ki}(s) 
-      = \frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
+  \pdiff{(\mathbf{a}\fund(s))_i}{\pib_k} = -(\mathbf{a}\fund(s)\onev) \fund_{ki}(s)
+      = -\frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
 \end{equation}
@@
   A(s) &= \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) \\
-   &= \intd[_0^{\infty}]{t} \sqrt{N} \eq \ench \e^{(r\frgh-s)t} \w \\
-   &= \sqrt{N} \eq \ench \prn{s - r\frgh}\inv \w.
+   &= \intd[_0^{\infty}]{t} \frac{ \sqrt{N} \eq \ench \e^{(r\frgh-s)t} \w }{ \sqrt{1-(f\pot-f\dep)^2(\eq\w)^2} } \\
+   &= \frac{\sqrt{N} \eq \ench \prn{s - r\frgh}\inv \w }{\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
 \end{aligned}
@@
 \begin{aligned}
-  \hat{A}(s) = \frac{A(s)}{\sqrt{N}} &=  \eq \ench \fund(s) \w \\
+  \hat{A}(s) = A(s) \sqrt{\frac{{{1-(f\pot-f\dep)^2(\eq\w)^2}}}{N}}  &=  \eq \ench \fund(s) \w \\
     &= \sum_{ijk} \eq_i \ench_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
@@
 \begin{equation}\label{eq:lAdiff}
-\begin{aligned}
-  \pdiff{\hat{A}(s)}{\Mh\potdep_{mn}{}}
-     =&\, rf\potdep \eq_m \brk{ \sum_i (\fptb_{mi}(0)-\fptb_{ni}(0)) a_i(s)
-     +c_m(s) (\eta^w_m(s) - \eta^w_n(s)) } \\
-     &\pm f\potdep \eq_m (\eta^w_m(s) - \eta^w_n(s)).
-\end{aligned}
+  \pdiff{\hat{A}(s)}{\Mh\potdep_{mn}}
+     = f\potdep \eq_m \brk{ \sum_i r[\fptb_{mi}(0)-\fptb_{ni}(0)]\, a_i(s)
+     +[rc_m(s) \pm 1] [\eta^w_m(s) - \eta^w_n(s)] } 
 \end{equation}
@@
 
+The area under the SNR curve is:
 %
 \begin{equation}\label{eq:areadef}
-  \int[_0^\infty]{dt} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) = A(0).
+  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) = A(0).
 \end{equation}
@@
 \begin{equation}\label{eq:scale}
-  \Mh\potdep - \I \to \lambda (\Mh\potdep - \I).
+  \Mh\potdep  \to (1-\lambda)\I + \lambda \Mh\potdep .
 \end{equation}
@@
 %
-The model that saturates this bound is one of serial topology with $\eq$ concentrated symmtrically at the two end states.
+The model that saturates this bound is one of serial topology with $\eq$ concentrated symmetrically at the two end states.
 
@@
 
+The initial SNR is:
 %
@@
 
+This satisfies the inequality
 %
@@
 
+\section{Finite time}\label{sec:finite}
+
+Now we'll maximise $\snrb(\tau)$ for some fixed $\tau$, subject to the constraints \eqref{eq:constrh}.
+First, we will maximise the numerator of \eqref{eq:lsnr} holding $\eq\w$ fixed. 
+We can maximise the ration \wrt $\eq\w$ later.
+Consider the Lagrangian
+%
+\begin{equation}\label{eq:finitelagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eq_i \ench_{ij} (\eta^w_i(s) - \eta^w_j(s))
+        +\sum_{i\neq j} \prn{ f\pot\mu\pot_{ij} \Mh\pot_{ij} + f\dep\mu\dep_{ij} \Mh\dep_{ij} }\\
+        &+\sum_i \prn{ \mu\pot_{ii} \brk{ f\pot - \sum_{j\neq i} \theta_{ij} \ench_{ij} } }
+        +\sum_i \prn{ \mu\dep_{ii} \brk{ f\dep + \sum_{j\neq i} \bar{\theta}_{ij} \ench_{ij} } } \\
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i\w_i },
+\end{aligned}
+\end{equation}
+%
+where $\theta_{ij} = 1-\bar{\theta}_{ij} = \Theta(\ench_{ij})$ and $\Delta\pr$ is the constant value we are holding $\eq\w$ at.
+After maximising
+The Karush-Kuhn-Tucker necessary conditions for a maximum are
+%
+\begin{equation}\label{eq:KTfinite}
+  \pdiff{\CL}{\Mh\potdep_{ij}} = 0, \quad
+  \mu\potdep_{ij} \geq 0, \quad
+  \pdiff{\CL}{\mu\potdep_{ij}} \geq 0, \quad
+  \mu\potdep_{ij}\pdiff{\CL}{\mu\potdep_{ij}} = 0, \quad
+  \pdiff{\CL}{\lambda} = 0.
+\end{equation}
+%
+The derivatives, for $m \neq n$ are
+%
+\begin{equation}\label{eq:finitederivs}
+\begin{aligned}
+  \frac{1}{f\potdep} \pdiff{\CL}{\Mh\potdep_{mn}} =& \,  
+     \eq_m \brk{ \sum_i r[\fptb_{mi}(0)-\fptb_{ni}(0)]\, a_i(s)
+     +[rc_m(s) \pm 1] [\eta^w_m(s) - \eta^w_n(s)] } \\
+     &+ \mu\potdep_{mn} \mp  \theta_{mn} \mu\pot_{mm} \pm  \bar{\theta}_{mn} \mu\dep_{mm} \\
+     &+ \lambda \eq_m \sum_i (\fptb_{mi}(0)-\fptb_{ni}(0)) \eq_i\w_i.
+\end{aligned}
+\end{equation}
+%
+Below, we will consider various combinations of these derivatives that preserve $\eq$.
+
+\subsection{Increasing \texorpdfstring{$c_k$}{c(k)}}\label{sec:areacoeffincr}
+
+Consider the following combinations of derivatives:
+%
+\begin{equation}\label{eq:areacoeffincrderiv}
+    \Delta_{mn} \equiv
+      \frac{1}{\eq_{m}} \pdiffc{}{\frgh_{mn}}
+      + \frac{1}{\eq_n} \pdiffc{}{\frgh_{nm}}, 
+\end{equation}
+%
+
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
+%
+\begin{equation}\label{eq:sareacoeffincrprob}
+  \Delta_{mn} \eq_i = 0,
+\end{equation}
+%
+which makes it possible to integrate the perturbation:
+%
+\begin{equation}\label{eq:areacoeffincrfinite}
+  \Mh\potdep \to \Mh\potdep + \frac{\D\boldsymbol{\epsilon}}{2f\potdep},
+  \qquad
+  \begin{aligned}
+  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
+  \\
+  \boldsymbol{\epsilon} \onev &= 0.
+  \end{aligned}
+\end{equation}
+%
+But more interestingly:
+%
+\begin{align}
+%  \begin{aligned}
+\label{eq:areacoeffincr}
+    \Delta_{mn}\CL &=
+      \frac{\mu\pot_{mn} + \mu\dep_{mn}}{2\eq_m} + \frac{\mu\pot_{nm} + \mu\dep_{nm}}{2\eq_n}
+      + r\prn{c_m(s)-c_n(s)}\prn{\eta^w_{m}(s) - \eta^w_{n}(s)},\\
+%  \end{aligned}
+\end{align}
+%
+where $c_k$ were defined in \eqref{eq:lareacoeffs}.
+As this derivative is zero at a maximum, and the $\mu\potdep_{ij}$ are non negative, the last term must be nonpositive.
+Therefore, if we arrange the states in order of decreasing $\eta^w_i(s)$, the $c_k(s)$ must be nondecreasing.
+
+\subsubsection{Shortcuts}\label{sec:shortcuts}
+
+Now consider the following combinations of derivatives for $n>1$:
+%
+\begin{align}
+%  \begin{aligned}
+\label{eq:shortcutderiv}
+    \widetilde{\Delta}_{m,n} &\equiv
+      \brk{ \sum_{k=0}^{n-1} \frac{1}{\eq_{m+k}} \pdiffc{}{\frgh_{m+k,m+k+1}} }
+      - \frac{1}{\eq_m} \pdiffc{}{\frgh_{m,m+ n}}.
+%  \end{aligned}
+\end{align}
+%
+
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
+%
+\begin{equation}\label{eq:shortcutprob}
+  \widetilde{\Delta}_{m,n} \eq_i = 0,
+\end{equation}
+%
+which makes it possible to integrate the perturbation:
+%
+\begin{equation}\label{eq:shortcutfinite}
+  \Mh\potdep \to \Mh\potdep + \frac{\D\boldsymbol{\epsilon}^{(m,n)}}{2f\potdep},
+  \qquad
+  \begin{aligned}
+    &\prn{\boldsymbol{\epsilon}^{(m,n)}}_{m,m+n}
+      &\!\!\!\!\!\!&= -\epsilon,\\
+    &\prn{\boldsymbol{\epsilon}^{(m,n)}}_{m+k,m+k+1}
+      &\!\!\!\!\!\!&= \epsilon
+        &\;
+        &\forall\, k \in [0,n-1],\\
+    &\prn{\boldsymbol{\epsilon}^{(m,n)}}_{m+k,m+k}
+      &\!\!\!\!\!\!&= -\epsilon
+        &
+        &\forall\, k \in [1,n-1].
+  \end{aligned}
+\end{equation}
+%
+But more interestingly for our purposes:
+%
+\begin{equation}\label{eq:shortcutarea}
+\begin{aligned}
+    \widetilde{\Delta}_{m,n}\CL =& \,
+      \brk{ \sum_{k=0}^{n-1} \frac{\mu\pot_{m+k,m+k+1}+\mu\dep_{m+k,m+k+1}}{\eq_{m+k}}
+      - \frac{\mu\pot_{m,m+n}+\mu\dep_{m,m+n}}{\eq_m}} \\
+      &+ r \sum_{k=0}^{n-1} \prn{\eta^w_{m+k}(s) - \eta^w_{m+k+1}(s)} \prn{c_{m+k}(s) - c_{m}(s)},
+\end{aligned}
+\end{equation}
+%
+
+If we put the states in order of decreasing $\eta^w_k(s)$, the results of the \sref{sec:areacoeffincr} tell us that the $c_k(s)$ are non-decreasing.
+This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
+If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu\potdep_{m,m+n}>0$, which in turn implies that $\Mh\potdep_{m,m+n}=0$ for all $n>1$.
+NO!
+ONLY ONE OF THEM.
+
+A similar technique also works for $n<1$.
+This implies that the only nonzero off-diagonal transitions are nearest neighbour.
+
 %\subsection*{Acknowledgements}

### 2015-06-15 Code/Optimum/FixNumLaplaceBnd.m
laplace envelope w/o topology constraints & homeostatic

@@ -1,11 +1,10 @@
-function [ mats ] = FixNumLaplaceBnd( oldmats,srange,nstates,trange,inds,varargin )
-%chains=FIXNUMLAPLACEBND(srange,nstates,trange,sym) numeric laplace bound
+function [ mats ] = FixNumLaplaceBnd( oldmats,srange,nstates,trange,mode,inds,varargin )
+%chains=FIXNUMLAPLACEBND(srange,nstates,trange,mode,inds) numeric laplace bound
 %   chains  = struct array (size=[1 length(srange)])
 %   srange  = values of Laplace parameter at which we maximise
 %   nstates = number of states in chain
 %   trange  = values of time for snr curve
-%   sym     = search for symmetric chains?
-%   hom     = include homeostatic plasticity?
+%   mode    = search for symmetric chains? include homeostatic plasticity?
 %   chains.s   = value of Laplace parameter at which we optimised
 %   chains.qv  = nearest neighbour transitions of optimal model
 %   chains.A   = value of Laplace transform at s for optimal model
@@ -20,7 +19,12 @@ for j=1:length(inds)
     DispCounter(j,length(inds),'s val: ');
     mats(i).s=srange(i);
     
-    [mats(i).Wp,mats(i).Wm,mats(i).Q,mats(i).A]=FindOptHomL(srange(i),nstates,reps,varargin{:});
+    switch mode
+        case 'hom'
+            [mats(i).Wp,mats(i).Wm,mats(i).Q,mats(i).A]=FindOptHomL(srange(i),nstates,reps,varargin{:});
+        otherwise
+            [mats(i).Wp,mats(i).Wm,mats(i).A]=FindOptL(srange(i),nstates,reps,varargin{:});
+    end
     
     w=BinaryWeights(nstates);
     modelobj=SynapseMemoryModel('Wp',mats(i).Wp+mats(i).Q,'Wm',mats(i).Wm+mats(i).Q,'w',w,'fp',0.5);

### 2015-06-15 Code/Optimum/ModelOpts/FindOpt.m

@@ -55,16 +55,19 @@ else
     Wm=[];
     snr=0;
     for i=1:r.reps
+        if r.DispReps
+            DispCounter(i,r.reps,'rep: ');
+        end
         [ Wpt,Wmt,St ] = FindOpt( t,n,1, p.Unmatched,'InitRand',r.InitRand,'Triangular',r.Triangular );
         if St>snr
             Wp=Wpt;
             Wm=Wmt;
             snr=St;
         end
-        if r.DispReps
-            disp([int2str(i) '/' int2str(r.reps)]);
-        end
     end
+        if r.DispReps
+            DispCounter(r.reps+1,r.reps,'rep: ');
+        end
         
 end
 

### 2015-06-15 Code/Optimum/ModelOpts/FindOptChainHomL.m

@@ -60,7 +60,9 @@ else
 %             disp([int2str(i) '/' int2str(r.reps)]);
         end
     end
+        if r.DispReps
             DispCounter(r.reps+1,r.reps,'rep: ');
+        end
     DispCounter(2,2,'init: ');
     for i=1:r.reps
         [ qvt,At ] = FindOptChainHomL( s,n,1, p.Unmatched,'InitRand',r.InitRand,'InitHomZero',true );
@@ -73,7 +75,9 @@ else
 %             disp([int2str(i) '/' int2str(r.reps)]);
         end
     end
+        if r.DispReps
             DispCounter(r.reps+1,r.reps,'rep: ');
+        end
     DispCounter(3,2,'init: ');
     warning('on','MATLAB:nearlySingularMatrix');
 end

### 2015-06-15 Code/Optimum/ModelOpts/FindOptChainHomLA.m

@@ -60,7 +60,9 @@ else
 %             disp([int2str(i) '/' int2str(r.reps)]);
         end
     end
+        if r.DispReps
             DispCounter(r.reps+1,r.reps,'rep: ');
+        end
     DispCounter(2,2,'init: ');
     for i=1:r.reps
         [ qvt,At ] = FindOptChainHomLA( s,n,1, p.Unmatched,'InitRand',r.InitRand,'InitHomZero',true );
@@ -73,7 +75,9 @@ else
 %             disp([int2str(i) '/' int2str(r.reps)]);
         end
     end
+        if r.DispReps
             DispCounter(r.reps+1,r.reps,'rep: ');
+        end
     DispCounter(3,2,'init: ');
     warning('on','MATLAB:nearlySingularMatrix');
 end

### 2015-06-15 Code/Optimum/ModelOpts/FindOptChainHomLC.m

@@ -60,7 +60,9 @@ else
 %             disp([int2str(i) '/' int2str(r.reps)]);
         end
     end
+        if r.DispReps
             DispCounter(r.reps+1,r.reps,'rep: ');
+        end
     DispCounter(2,2,'init: ');
     for i=1:r.reps
         [ qvt,At ] = FindOptChainHomLC( s,n,1, p.Unmatched,'InitRand',r.InitRand,'InitHomZero',true );
@@ -73,7 +75,9 @@ else
 %             disp([int2str(i) '/' int2str(r.reps)]);
         end
     end
+        if r.DispReps
             DispCounter(r.reps+1,r.reps,'rep: ');
+        end
     DispCounter(3,2,'init: ');
     warning('on','MATLAB:nearlySingularMatrix');
 end

### 2015-06-15 Code/Optimum/ModelOpts/FindOptHomL.m

@@ -80,7 +80,9 @@ else
             A=At;
         end
     end
+        if r.DispReps
             DispCounter(r.reps+1,r.reps,'rep: ');
+        end
     DispCounter(2,2,'init: ');
     for i=1:r.reps
         if r.DispReps
@@ -95,7 +97,9 @@ else
             A=At;
         end
     end
+        if r.DispReps
             DispCounter(r.reps+1,r.reps,'rep: ');
+        end
     DispCounter(3,2,'init: ');
     warning('on','MATLAB:nearlySingularMatrix');
     

### 2015-06-16 Code/Optimum/FixNumLaplaceBnd.m
notes on laplace env w/o hom

@@ -22,12 +22,16 @@ for j=1:length(inds)
     switch mode
         case 'hom'
             [mats(i).Wp,mats(i).Wm,mats(i).Q,mats(i).A]=FindOptHomL(srange(i),nstates,reps,varargin{:});
+            Wp=mats(i).Wp+mats(i).Q;
+            Wm=mats(i).Wm+mats(i).Q;
         otherwise
             [mats(i).Wp,mats(i).Wm,mats(i).A]=FindOptL(srange(i),nstates,reps,varargin{:});
+            Wp=mats(i).Wp;
+            Wm=mats(i).Wm;
     end
     
     w=BinaryWeights(nstates);
-    modelobj=SynapseMemoryModel('Wp',mats(i).Wp+mats(i).Q,'Wm',mats(i).Wm+mats(i).Q,'w',w,'fp',0.5);
+    modelobj=SynapseMemoryModel('Wp',Wp,'Wm',Wm,'w',w,'fp',0.5);
     
     mats(i).snr=modelobj.SNRcurve(trange);
     

### 2015-06-16 Notes/laplace.tex
writing up numerical envelope

--- 2015-06-01 Notes/laplace.tex
@@
 \usepackage[notref,notcite]{showkeys}
+%
+\input{sl_graphics_preamble.tex}
+\graphicspath{{Figures/}}
+%
 % ----------------------------------------------------------------
@@
 When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M\potdep$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$.
-In between these plasticity events the synapse will undergo stochastic transitions described by a matrix $\hom$ whose $i,j$ element is the transition rate from state $i$ to state $j$ for $i \neq j$ and $\hom_{ii}=-\sum_{j \neq i} \hom_{ij}$.
-This allows us to describe homeostatic plasticity as well as plasticity processes that occur over a period of time -- $\M\potdep$ puts the synapse into some state where the process $\hom$ takes over and does the rest.
-The matrices $\M\potdep$ are transition matrices of discrete time Markov processes, whereas $\hom$ is the transition matrix of a continuous time Markov process.
+The matrices $\M\potdep$ are transition matrices of discrete time Markov processes.
 Their elements must satisfy the following constraints
@@
   \M\potdep_{ij} &\geq 0, &\qquad
-  \sum_j \M\potdep_{ij} &= 1, \\
-  \hom_{i \neq j} &\geq 0, &\qquad
-  \sum_j \hom_{ij} &= 0.
+  \sum_j \M\potdep_{ij} &= 1.
 \end{aligned}
@@
 %
-The upper bounds $\M\potdep_{ij} \leq 1$ and $\hom_{ii} \leq 0$ follow automatically from these.
-There is no upper bound on $\hom_{i \neq j}$ (or lower bound on $\hom_{ii}$) as these are rates rather than probabilities.
+The upper bounds $\M\potdep_{ij} \leq 1$ follow automatically from these.
 
@@
   \M\potdep_{i \neq j} \geq 0, \qquad
-  \sum_{j\neq i} \M\potdep_{ij} \leq 1, \qquad
-  \hom_{i \neq j} \geq 0.
+  \sum_{j\neq i} \M\potdep_{ij} \leq 1.
 \end{equation}
@@
 
-Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states is
+Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$ is
 %
 \begin{equation}\label{eq:plastseq}
-  \pr(t_n) = \pr(t_0) \e^{\hom(t_1-t_0)} \M\pot \e^{\hom(t_2-t_1)} \M\dep \e^{\hom(t_3-t_4)} \M\dep \ldots \M\pot \e^{\hom(t_n-t_{n-1})}.
+  \pr(t_n) = \pr(t_0)  \M\pot \M\dep \M\dep \M\pot \M\dep \M\pot \M\pot \M\dep \ldots \M\pot .
 \end{equation}
 %
+If the probability of a given plasticity event being poetitiating/depressing os $f\potdep$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
+
 In computing the expectation in \eqref{eq:overlapone} the plasticity event that occurs at $t=0$ is fixed, but all of the previous and subsequent events are marginalised over.
-If the plasticity events occur at Poisson rate $r$, with potentiation and depression occurring independently with probability $f\potdep$, averaging over all previous and subsequent events gives
+If we marginalise over the types (potentiating/depressing) of these other events (conditioned on the number of events) we find
+%
+\begin{multline}\label{eq:plastseqtypes}
+  \pr(t_n) = \pr(t_0) (f\pot\M\pot+f\dep\M\dep)  \ldots (f\pot\M\pot+f\dep\M\dep) \\
+    \M\potdep (f\pot\M\pot+f\dep\M\dep) \ldots (f\pot\M\pot+f\dep\M\dep),
+\end{multline}
+%
+where the factor of $\M\potdep$ is the event at $t=0$.
+If we expand out all of the parenthetical factors, we will get a sum of terms of the form \eqref{eq:plastseq} weighted by their probabilities.
+
+If $t_0 \ll 0$, the effect of the large number of earlier events will be to put the synapse in the steady state distribution, $\eq$, of the stochastic process $(f\pot\M\pot+f\dep\M\dep)$:
+%
+\begin{equation}\label{eq:plastseqeq}
+  \pr(t_n) = \eq \M\potdep (f\pot\M\pot+f\dep\M\dep)^n,
+\end{equation}
+%
+where $n$ is the number of plasticity events that took place.
+
+If the plasticity events occur at Poisson rate $r$, the probability of $n$ events occurring in time $t$ is given by a Poisson distribution with mean $rt$.
+If we marginalise over $n$as well
 %
 \begin{equation}\label{eq:plastseqav}
-  P(\text{state}=i,t \mid \text{pot/dep},0) = \brk{ \eq \M\potdep\, \e^{r(f\pot\M\pot+f\dep\M\dep-\I)t +\hom t} }_i,
+\begin{aligned}
+  P(\text{state}=i,t \mid \text{pot/dep},0) 
+    &= \sum_n \frac{(rt)^n}{n!} \e^{-rt} \eq \M\potdep (f\pot\M\pot+f\dep\M\dep)^n \\
+    &= \brk{ \eq \M\potdep\, \e^{rt(f\pot\M\pot+f\dep\M\dep-\I) } }_i,
+\end{aligned}
 \end{equation}
@@
 \begin{equation}\label{eq:eq}
-  \eq(r\frg +\hom) = 0,
+  \eq\frg = 0,
   \qquad\text{where}\quad
@@
 \begin{aligned}
-  \av{\syn^1_\text{ideal}\syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{(r\frg+\hom)t} \w, \\
+  \av{\syn^1_\text{ideal}\syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
   \av{\syn^1_\text{ideal}\syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
@@
          &=  (f\pot-f\dep)\, \eq \w \\
-         &=  (f\pot-f\dep)\, \eq \e^{(r\frg+\hom)t} \w .
+         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
 \end{aligned}
@@
 
-This formula takes the same form as the one in our NIPS paper \cite{Lahiri2013synapse}, where the process $\hom$ was not included.
-However, the process $\hom$ still has an impact due to the constraints.
-
-The constraints on $\Mh\potdep$ are less stringent than the constraints on $\M\potdep$, \eqref{eq:constr}.
-Clearly
-%
-\begin{equation}\label{eq:constrhoffdiag}
-  \Mh\potdep_{i \neq j} \geq 0, \qquad
-  \sum_j \Mh\potdep_{ij} = 1.
-\end{equation}
-%
-The ``diagonal'' constraints are more complicated.
-There can be many $\M\potdep,\hom$ that result in the same $\Mh\potdep$.
-However, the fact that $\M\potdep_{ij} \in [0,1]$ and $\hom_{i\neq j} \geq 0$ imply that $\forall (i \neq j)$:
-%
-\begin{equation}\label{eq:homrange}
-  2r\max\brc{f\pot(\Mh\pot_{ij}-1),f\dep(\Mh\dep_{ij}-1)} \leq \hom_{ij} \leq 2r\min\brc{f\pot\Mh\pot_{ij},f\dep\Mh\dep_{ij}}
-\end{equation}
-%
-These two bounds are consistent when
-%
-\begin{equation}\label{eq:homconsistent}
-  -f\dep \leq f\pot\Mh\pot_{ij} - f\dep\Mh\dep_{ij} \leq f\pot
-  \qquad \forall i\neq j.
-\end{equation}
-%
-To ensure that $\M\potdep_{ii} \geq 0$, we also need to ensure that $\sum_{j\neq i}\M\potdep_{ij} \leq 1$.
-%
-\begin{equation}\label{eq:homdiag}
-\begin{aligned}
-  \sum_{j\neq i} \M\pot_{ij}
-    &= \sum_{j\neq i} \prn{ \Mh\pot_{ij} - \frac{\hom_{ij}}{2rf\pot} }\\
-    &\geq \sum_{j\neq i} \prn{ \Mh\pot_{ij} - \frac{1}{f\pot}  \min\brc{ f\pot\Mh\pot_{ij},f\dep\Mh\dep_{ij} } }\\
-    &= \frac{1}{f\pot} \sum_{j\neq i} \brk{ f\pot\Mh\pot_{ij}-f\dep\Mh\dep_{ij} }_+ , \\
-  \sum _{j\neq i} \M\dep_{ij}
-    &= \frac{1}{f\dep} \sum_{j\neq i} \brk{ f\dep\Mh\dep_{ij}-f\pot\Mh\pot_{ij} }_+ ,
-\end{aligned}
-\end{equation}
-%
-where rectification function is $[x]_\pm = x \Theta(\pm x)$, in terms of the Heaviside step function.
-Demanding that these quantities are at most 1 is sufficient to ensure that \eqref{eq:homconsistent} is satisfied.
-
-Therefore we find the set of constraints
-%
-\begin{equation}\label{eq:constrh}
-\begin{aligned}
-  \Mh\potdep_{i \neq j} &\geq 0, &\qquad
-  \sum_{j\neq i} \brk{ f\pot\Mh\pot_{ij}-f\dep\Mh\dep_{ij} }_+ &\leq f\pot, \\
-  \sum_j \Mh\potdep_{ij} &=1, &
-  \sum_{j\neq i} \brk{ f\dep\Mh\dep_{ij}-f\pot\Mh\pot_{ij} }_+  &\leq f\dep.
-\end{aligned}
-\end{equation}
-%
-The second column of this will be referred to as the ``diagonal'' constraints.
-
-We showed that these are necessary conditions for the existence of $\M\potdep,\hom$ satisfying \eqref{eq:constr}.
-We can see that they are sufficient by choosing $\hom_{ij} = 2r\min\brc{f\pot\Mh\pot_{ij},f\dep\Mh\dep_{ij}}$ for $i \neq j$ and $\hom_{ii}=-\sum_{j \neq i} \hom_{ij}$.
-
 We can then treat the off-diagonal elements of $\Mh\potdep$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
@@
 \begin{equation}\label{eq:derivpd}
-  \pdiff{\Mh\pot_{ij}}{\Mh\pot_{mn}{}} = \pdiff{\Mh\dep_{ij}}{\Mh\dep_{mn}{}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
+  \pdiff{\M\pot_{ij}}{\M\pot_{mn}} = \pdiff{\M\dep_{ij}}{\M\dep_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
   \qquad
-  \pdiff{\Mh\pot_{ij}}{\Mh\dep_{mn}{}} = \pdiff{\Mh\dep_{ij}}{\Mh\pot_{mn}{}} = 0.
+  \pdiff{\M\pot_{ij}}{\M\dep_{mn}} = \pdiff{\M\dep_{ij}}{\M\pot_{mn}} = 0.
 \end{equation}
@@
 \begin{equation}\label{eq:pertfe}
-  \pdiff{}{\frgh_{ij}{}} = \frac{1}{2f\pot} \pdiff{}{\Mh\pot_{ij}{}} + \frac{1}{2f\dep} \pdiff{}{\Mh\dep_{ij}{}},
+  \pdiff{}{\frg_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
   \qquad
-  \pdiff{}{\ench_{ij}{}} = \frac{1}{2f\pot} \pdiff{}{\Mh\pot_{ij}{}} - \frac{1}{2f\dep} \pdiff{}{\Mh\dep_{ij}{}},
+  \pdiff{}{\enc_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
 \end{equation}
@@
 \begin{equation}\label{eq:derivfe}
-  \pdiff{\frgh_{ij}}{\frgh_{mn}{}} = \pdiff{\ench_{ij}}{\ench_{mn}{}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
+  \pdiff{\frg_{ij}}{\frg_{mn}} = \pdiff{\enc_{ij}}{\enc_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
   \qquad
-  \pdiff{\frgh_{ij}}{\ench_{mn}{}} = \pdiff{\ench_{ij}}{\frgh_{mn}{}} = 0.
+  \pdiff{\frg_{ij}}{\enc_{mn}} = \pdiff{\enc_{ij}}{\frg_{mn}} = 0.
+\end{equation}
+%
+
+If we wish to use the off-diagonal elements of $\frg$ and $\enc$ as the degrees of freedom, the constraints are
+%
+\begin{equation}\label{eq:constrfe}
+\begin{aligned}
+  \frg_{i \neq j}   &\geq \abs{ \enc_{i \neq j} }, &\qquad
+  \sum_{j\neq i}\frg_{ij} \pm \abs{ \enc_{ij} } &\leq f\potdep, \\
+  \sum_j \frg_{ij} &=0, &
+  \sum_j \enc_{ij} &=0.
+\end{aligned}
 \end{equation}
@@
   A(s) &= \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) \\
-   &= \intd[_0^{\infty}]{t} \sqrt{N} \eq \ench \e^{(r\frgh-s)t} \w \\
-   &= \sqrt{N} \eq \ench \prn{s - r\frgh}\inv \w.
+   &= \intd[_0^{\infty}]{t} \frac{ \sqrt{N} \eq \enc \e^{(r\frg-s)t} \w }{ \sqrt{1-(f\pot-f\dep)^2(\eq\w)^2} } \\
+   &= \frac{\sqrt{N} \eq \enc \prn{s - r\frg}\inv \w }{\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
 \end{aligned}
@@
 This expression is ill-behaved at $s=0$.
-Thanks to \eqref{eq:alfund}, we can solve this by the replacement $\mathbf{G}(s) \to \fund(s)$, as $\ench\onev=0$.
+Thanks to \eqref{eq:alfund}, we can solve this by the replacement $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
 %
@@
 \begin{aligned}
-  \hat{A}(s) = \frac{A(s)}{\sqrt{N}} &=  \eq \ench \fund(s) \w \\
-    &= \sum_{ijk} \eq_i \ench_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
-    &= \sum_{ij}  \eq_i \ench_{ij} (\eta^w_{i}(s) - \eta^w_{j}(s)).
+  \hat{A}(s) = A(s) \sqrt{\frac{{{1-(f\pot-f\dep)^2(\eq\w)^2}}}{N}}  &=  \eq \enc \fund(s) \w \\
+    &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
+    &= \sum_{ij}  \eq_i \enc_{ij} (\eta^w_{i}(s) - \eta^w_{j}(s)).
 \end{aligned}
@@
   \begin{aligned}
-    c_k(s) &= \sum_{ij} \eq_i \ench_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
-    a_i(s) &= \sum_j \eq_i \ench_{ij} (\eta^w_{i}(s) - \eta^w_{j}(s)),\\
+    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
+    a_i(s) &= \sum_j \eq_i \enc_{ij} (\eta^w_{i}(s) - \eta^w_{j}(s)),\\
     \implies
@@
 
-Using the same approach as the area bound in \cite{Lahiri2013synapse}, we find that the derivatives are
+Using the same approach as the area bound in \cite{Lahiri2013synapse}, and defining $\Theta_i(s) = \sum_j \fptb_{ij}(0) a_j(s)$,  we find that the derivatives are
 %
 \begin{equation}\label{eq:lAdiff}
-\begin{aligned}
-  \pdiff{\hat{A}(s)}{\Mh\potdep_{mn}{}}
-     =&\, rf\potdep \eq_m \brk{ \sum_i (\fptb_{mi}(0)-\fptb_{ni}(0)) a_i(s)
-     +c_m(s) (\eta^w_m(s) - \eta^w_n(s)) } \\
-     &\pm f\potdep \eq_m (\eta^w_m(s) - \eta^w_n(s)).
-\end{aligned}
+  \pdiff{\hat{A}(s)}{\Mh\potdep_{mn}}
+     = f\potdep \eq_m \prn{ r \brk{\Theta_m(s) - \Theta_n(s)}
+     +\prn{rc_m(s) \pm 1} \brk{\eta^w_m(s) - \eta^w_n(s)} },
 \end{equation}
 %
+with the upper/lower sign in $\pm$ corresponding to derivatives \wrt $\M\potdep$.
+The term with $\Theta(s)$ comes from differentiating $\eq$, the term with $c(s)$ comes from differentiating $\eta^w(s)$ and the term with $\pm1$ comes from differentiating $\enc$ in \eqref{eq:larea}.
 
@@
 
-\subsection{Area bound}\label{sec:area}
-
-%
-\begin{equation}\label{eq:areadef}
-  \int[_0^\infty]{dt} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) = A(0).
-\end{equation}
-%
-
-The only effect of $\hom$ is in the diagonal constraints (second column of \eqref{eq:constrh} vs.\ second inequality in \eqref{eq:constri}).
-However, the area is invariant under the scaling mode:
-%
-\begin{equation}\label{eq:scale}
-  \Mh\potdep - \I \to \lambda (\Mh\potdep - \I).
-\end{equation}
-%
-This means we can always take a set of matrices that violate either version of the diagonal constraints and use \eqref{eq:scale} to create a set of matrices that satisfy them without changing the area.
-Therefore, including $\hom$ has no effect on the area bound and the result from \cite{Lahiri2013synapse} applies here as well.
-
-This means that 
-%
-\begin{equation}\label{eq:areabnd}
-  A(0) \leq \frac{\sqrt{N}(M-1)}{r}.
-\end{equation}
-%
-The model that saturates this bound is one of serial topology with $\eq$ concentrated symmtrically at the two end states.
-
 \subsection{Initial SNR}\label{sec:initial}
 
+The initial SNR is:
 %
@@
 
+As shown in \cite{Lahiri2013synapse}, this satisfies the inequality
 %
@@
 %
-at least when $f\pot=f\dep=\frac{1}{2}$.
+The model that saturates this bound is (equivalent to) a two-state model with deterministic transitions and $f\potdep=\frac{1}{2}$.
+
+\subsection{Area bound}\label{sec:area}
+
+The area under the SNR curve is:
+%
+\begin{equation}\label{eq:areadef}
+  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) = A(0).
+\end{equation}
+%
+
+As shown in \cite{Lahiri2013synapse}, this satisfies the inequality
+%
+\begin{equation}\label{eq:areabnd}
+  A(0) \leq \frac{\sqrt{N}(M-1)}{r}.
+\end{equation}
+%
+The model that saturates this bound is one of serial topology with $\eq$ concentrated symmetrically at the two end states.
 
@@
 
-Now we introduce an eigenvector decomposition for $-\frgh$:
+Now we introduce an eigenvector decomposition for $-\frg$:
 %
 \begin{equation}\label{eq:eigendecomp}
-  \frgh = - \sum_a q_a \evr^a \evl^a,
+  \frg = - \sum_a q_a \evr^a \evl^a,
   \quad
@@
   \quad
-  \frgh \evr^a = -q_a \evr^a,
+  \frg \evr^a = -q_a \evr^a,
   \quad
-  \evl^a \frgh = -q_a \evl^a.
+  \evl^a \frg = -q_a \evl^a.
 \end{equation}
@@
 
+This envelope is shown in \autoref{fig:envproven}.
+
+\begin{figure}[tb]
+  \centering
+  % Requires \usepackage{graphicx}
+  \includegraphics[width=0.8\linewidth]{LenvProven.eps}
+  \caption{The memory curve envelope that we can prove, \eqref{eq:env}.
+  The number of synapses $N=10^4$ and the number of states $M=12$.
+  The dashed lines indicate the initial SNR bound, \eqref{eq:initbnd}, and the area bound, \eqref{eq:areabnd}, from \cite{Lahiri2013synapse}.}\label{fig:envproven}
+\end{figure}
+
+\section{Finite time}\label{sec:finite}
+
+Now we'll maximise $\snrb(\tau)$ for some fixed $\tau$, subject to the constraints \eqref{eq:constr}.
+First, we will maximise the numerator of \eqref{eq:lsnr} holding $\eq\w$ fixed.
+We can maximise the ration \wrt $\eq\w$ later.
+Consider the Lagrangian
+%
+\begin{equation}\label{eq:finitelagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\eta^w_i(s) - \eta^w_j(s))
+        +\sum_{i j} \prn{ f\pot\mu\pot_{ij} \M\pot_{ij} + f\dep\mu\dep_{ij} \M\dep_{ij} }\\
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i\w_i },
+\end{aligned}
+\end{equation}
+%
+where $\Delta\pr$ is the constant value we are holding $\eq\w$ at.
+After maximising
+The Karush-Kuhn-Tucker necessary conditions for a maximum are
+%
+\begin{equation}\label{eq:KTfinite}
+  \pdiff{\CL}{\M\potdep_{ij}} = 0, \quad
+  \mu\potdep_{ij} \geq 0, \quad
+  \pdiff{\CL}{\mu\potdep_{ij}} \geq 0, \quad
+  \mu\potdep_{ij}\pdiff{\CL}{\mu\potdep_{ij}} = 0, \quad
+  \pdiff{\CL}{\lambda} = 0.
+\end{equation}
+%
+The derivatives, for $m \neq n$ are
+%
+\begin{equation}\label{eq:finitederivs}
+\begin{aligned}
+  \frac{1}{f\potdep} \pdiff{\CL}{\Mh\potdep_{mn}} =& \,
+      \eq_m \prn{ r \brk{\Theta_m(s) - \Theta_n(s)}
+     +\prn{rc_m(s) \pm 1} \brk{\eta^w_m(s) - \eta^w_n(s)} } \\
+     &+ \mu\potdep_{mn} -  \mu\potdep_{mm}  
+     + \lambda \eq_m \sum_i (\fptb_{mi}(0)-\fptb_{ni}(0)) \eq_i\w_i.
+\end{aligned}
+\end{equation}
+%
+These have the same form as for $s=0$, which allowed us to argue that the model with maximal area must have the serial topology, \cite{Leibold2008serial,Ben-DayanRubin2007sparse}.
+
+The one missing ingredient is the scale invariance.
+Define the scale transformation:
+%
+\begin{equation}\label{eq:scale}
+  \lambda * \M = (1-\lambda)\I + \lambda \M. 
+\end{equation}
+%
+The Laplace transform has the following scaling property
+%
+\begin{equation}\label{eqlaplacescale}
+  A(\lambda s;\lambda * \M\potdep) = A(s;\M\potdep).
+\end{equation}
+%
+At $s=0$, it is scale invariant.
+This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
+This allowed us to ignore the diagonal constraints.
+We can't do that for $s>0$.
+
+However, we do have numerical evidence that the model that maximises the Laplace transform does have the serial topology.
+In \autoref{fig:envnum} we see the result of numerical maximisation of \eqref{eq:finitelagrangian}.
+We see that restricting to the serial topology does not make the performance any worse.
+In fact, it slightly improves the performance, but this can be ascribed to the numerical optimisation problem being easier.
+Assuming that the optimiser did find the global maximum, this provides evidence for our conjecture  that the model that maximises the Laplace transform must have the serial topology.
+
+\begin{figure}[tb]
+  \centering
+  % Requires \usepackage{graphicx}
+  \includegraphics[width=0.8\linewidth]{LenvNum.eps}\\
+  \caption{Numerical envelope for the memory curve, \eqref{eq:snrb}, when $f\potdep=\frac{1}{2}$, the number of synapses $N=10^4$ and the number of states $M=12$.
+  The dashed red line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
+  The solid red line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.}\label{fig:envnum}
+\end{figure}
+
+\section{Nearly uniform serial models}\label{sec:serial}
+
+\begin{figure}[tb]
+  \centering
+  % Requires \usepackage{graphicx}
+  \begin{myenuma}
+    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_n.svg}}\hspace{0.02\linewidth}\label{fig:serial_uniform}
+    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_sticky.svg}}\hspace{0.02\linewidth}\label{fig:serial_sticky}
+    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_shorten.svg}}\label{fig:serial_shorten}
+  \end{myenuma}
+  \caption{Heuristic models that maximise memory curve.
+  At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with $M$ states.
+  At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model. 
+  At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states. 
+  Once this reaches zero, we discard the old end state and work on the new end state.} \label{fig:heuristicmodels}
+\end{figure}
+
+Looking at them models that were found by numerical maximisation in \autoref{fig:envnum} (see the video for the models), suggests we look at the particular set of models, shown in \autoref{fig:heuristicmodels}.
+
+At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with $M$ states.
+At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model as $\tau\to\infty$.
+At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
+Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state.
+Once the length has been reduced to $M=2$, the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.
+
+These models can be solved analytically, as we will see below, leading to a heuristic envelope.
+
 %\subsection*{Acknowledgements}

### 2015-06-16 Code/Figs/RunAveFig.m

@@ -1,25 +1,21 @@
-load('laplacebndNoHom.mat','AenvNoHom','srange');
-load('laplacebndHom.mat','AenvHom');
-% load('laplacebndHomC.mat','AenvHomC');
-load('laplacebndHomA.mat','AenvHomA');
-load('laplacebndHomB.mat','AenvHomB');
+
+load('laplacebnd12.mat','AenvAll','srange');
+load('laplacebndChainA12.mat','AenvChains');
+
 AenvProven=11./(1+11*srange);
 
 
 loglog( 1./srange,100*AenvProven.*srange,'g',...
-    1./srange,100*AenvHomB.*srange,'r:',...
-    1./srange,100*AenvHom.*srange,'r--',...
-    1./srange,100*AenvHomA.*srange,'r-.',...
-    1./srange,100*AenvNoHom.*srange,'r-',...
+    1./srange,100*AenvAll.*srange,'r:',...
+    1./srange,100*AenvChains.*srange,'r-',...
     'LineWidth',2);
 %     1./srange,100*AenvHomC.*srange,'r--',...
-ylim([1e-2 500]);
+ylim([1e-1 500]);
 xlabel('Mean recall time, $\tau$','Interpreter','latex','FontSize',16);
 ylabel('Recognition performance, $\overline{\mathrm{SNR}}(\tau)$','Interpreter','latex','FontSize',16);
 title('Numerical envelopes','Interpreter','latex','FontSize',20);
 legend({'Proven envelope',...
-    'Basic constraints',...
-    'Topology constraints',...
-    'All constraints',...
-    '$\mathbf{Q}=0$'},...
+    'Numeric: all topologies',...
+    'Numeric: serial topology',...
+    },...
     'Location','southwest','Interpreter','latex','FontSize',16);

### 2015-06-16 Code/Figs/RunAveFigHom.m

@@ -1,0 +1,25 @@
+load('laplacebndNoHom.mat','AenvNoHom','srange');
+load('laplacebndHom.mat','AenvHom');
+% load('laplacebndHomC.mat','AenvHomC');
+load('laplacebndHomA.mat','AenvHomA');
+load('laplacebndHomB.mat','AenvHomB');
+AenvProven=11./(1+11*srange);
+
+
+loglog( 1./srange,100*AenvProven.*srange,'g',...
+    1./srange,100*AenvHomB.*srange,'r:',...
+    1./srange,100*AenvHom.*srange,'r--',...
+    1./srange,100*AenvHomA.*srange,'r-.',...
+    1./srange,100*AenvNoHom.*srange,'r-',...
+    'LineWidth',2);
+%     1./srange,100*AenvHomC.*srange,'r--',...
+ylim([1e-2 500]);
+xlabel('Mean recall time, $\tau$','Interpreter','latex','FontSize',16);
+ylabel('Recognition performance, $\overline{\mathrm{SNR}}(\tau)$','Interpreter','latex','FontSize',16);
+title('Numerical envelopes','Interpreter','latex','FontSize',20);
+legend({'Proven envelope',...
+    'Basic constraints',...
+    'Topology constraints',...
+    'All constraints',...
+    '$\mathbf{Q}=0$'},...
+    'Location','southwest','Interpreter','latex','FontSize',16);

### 2015-06-16 Code/GUIs/LaplaceEnvGUI.m

@@ -12,6 +12,7 @@ EnvLinewidth=3;
 ModelLineWidth=2;
 TimeLineWidth=2;
 EqLineWidth=2;
+Interpreter='latex';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %Data
@@ -155,16 +156,21 @@ changeFrameNumber(frNumber);
         xlim(snr_ax,snr_xlim);
         ylim(snr_ax,snr_ylim);
         
-        set(snr_ax,'FontSize',AxFontSize)
+        set(snr_ax,'TickLabelInterpreter',Interpreter,'FontSize',AxFontSize)
         
-%         xlabel(snr_ax,'\textsf{Mean recall time}, $\mathsf{\tau}$','Interpreter','latex','FontSize',AxFontSize);
-        xlabel(snr_ax,'Mean recall time, \tau','FontSize',AxFontSize);
-        ylabel(snr_ax,'\textsf{Recognition performance}, $\overline{\mathsf{SNR}}\mathsf{(\tau)}$','Interpreter','latex','FontSize',AxFontSize);
-        title(snr_ax,'Proven and numerical envelopes','FontSize',AxFontSize);
+        switch Interpreter
+            case 'tex'
+                xlabel(snr_ax,'Mean recall time, \tau','FontSize',AxFontSize);
+                ylabel(snr_ax,'\textsf{Recognition performance}, $\overline{\mathsf{SNR}}\mathsf{(\tau)}$','Interpreter','latex','FontSize',AxFontSize);
+            case 'latex'
+                xlabel(snr_ax,'Mean recall time, $\mathsf{\tau}$','Interpreter','latex','FontSize',AxFontSize);
+                ylabel(snr_ax,'Recognition performance, $\overline{\mathrm{SNR}}(\tau)$','Interpreter','latex','FontSize',AxFontSize);
+        end
+        title(snr_ax,'Proven and numerical envelopes','Interpreter',Interpreter,'FontSize',AxFontSize);
         legend(snr_ax,{'Proven envelope',...
             'Numeric envelope',...
             'Current model'},...
-            'Location','northeast','FontSize',AxFontSize);
+            'Location','northeast','Interpreter',Interpreter,'FontSize',AxFontSize);
     end
 
    function PlotModel(frameNumber)
[…]

### 2015-06-18 Notes/laplace.tex
writing up uniform and sticky heuristics

@@
 
-Using the same approach as the area bound in \cite{Lahiri2013synapse}, and defining $\Theta_i(s) = \sum_j \fptb_{ij}(0) a_j(s)$,  we find that the derivatives are
+Using the same approach as the area bound in \cite{Lahiri2013synapse}, and defining $\theta_i(s) = \sum_j \fptb_{ij}(0) a_j(s)$,  we find that the derivatives are
 %
@@
   \pdiff{\hat{A}(s)}{\Mh\potdep_{mn}}
-     = f\potdep \eq_m \prn{ r \brk{\Theta_m(s) - \Theta_n(s)}
+     = f\potdep \eq_m \prn{ r \brk{\theta_m(s) - \theta_n(s)}
      +\prn{rc_m(s) \pm 1} \brk{\eta^w_m(s) - \eta^w_n(s)} },
@@
   \frac{1}{f\potdep} \pdiff{\CL}{\Mh\potdep_{mn}} =& \,
-      \eq_m \prn{ r \brk{\Theta_m(s) - \Theta_n(s)}
+      \eq_m \prn{ r \brk{\theta_m(s) - \theta_n(s)}
      +\prn{rc_m(s) \pm 1} \brk{\eta^w_m(s) - \eta^w_n(s)} } \\
@@
   \end{myenuma}
-  \caption{Heuristic models that maximise memory curve.
+  \caption{Heuristics for models that maximise memory curve.
+  All unlabelled arrowshave transition probabilities equal to 1.
   At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with $M$ states.
@@
 At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
-Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state.
+Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from the video, this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
 Once the length has been reduced to $M=2$, the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.
 
-These models can be solved analytically, as we will see below, leading to a heuristic envelope.
+These models' Laplace transforms can be computed analytically, as we will see below, leading to a heuristic envelope.
+All of these calculations will be performed at $f\potdep=\frac{1}{2}$, so the denominator will play no role.
+We will work in units such that $r=1$ temporarily, only restoring $r$ at the end.
+
+\subsection{Uniform serial model}\label{sec:serial_uniform}
+
+Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_uniform}.
+Looking at \eqref{eq:larea}, the only thing we need is differences in $\eta^w_i(s)$.
+In fact, given that the only nonzero elements of $\enc$ are
+%
+\begin{equation}\label{eq:unienc}
+  \enc_{ii+1}=\frac{1}{2} \quad\text{for}\quad i<M, 
+  \qquad
+  \enc_{ii-1}=-\frac{1}{2} \quad\text{for}\quad i>1,
+\end{equation}
+%
+and $\eq_i = \frac{1}{M}$, we have
+%
+\begin{equation}\label{eq:uniareaeta}
+  \hat{A}(s) = \frac{\eta^w_1(s) - \eta^w_M(s)}{M}.
+\end{equation}
+%
+
+Note that
+%
+\begin{equation}\label{eq:etafund}
+  \eta^w(s) = \text{const.} - \fund(s)\w.
+\end{equation}
+%
+As we are only interested in differences of the $\eta^w_i(s)$, we can drop the constant.
+Then
+%
+\begin{equation}\label{eq:etarecur}
+  -(s\I + \onev\pib - \frg)\eta^w(s) = \w.
+\end{equation}
+%
+We can then write
+%
+\begin{equation}\label{eq:detarecur}
+  \eta^w(s) = -\frac{(\pib\eta^w(s))\onev+\w}{s} + \delta\eta(s),
+  \qquad
+  (\frg-s\I)\delta\eta(s) = \frac{\frg\w}{s}.
+\end{equation}
+%
+As we are only interested in differences, we can drop any terms proportional to $\onev$ in $\eta^w(s)$.
+
+The $i$'th row of this equation reads
+%
+\begin{equation}\label{eq:detarow}
+  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i = 0,
+\end{equation}
+%
+with the following exceptions
+%
+\begin{equation}\label{eq:detabnduni}
+\begin{aligned}
+  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
+  \frac{\delta\eta_{M/2+1}+\delta\eta_{M/2-1}}{2} - (s+1)\delta\eta_{M/2} &= \frac{1}{s},\\
+  \frac{\delta\eta_{M/2+2}+\delta\eta_{M/2}}{2} - (s+1)\delta\eta_{M/2+1} &= -\frac{1}{s},\\
+  \frac{\delta\eta_{M-1}}{2} - \prn{s+\frac{1}{2}}\delta\eta_M &= 0.
+\end{aligned}
+\end{equation}
+%
+These equations are invariant under $\delta\eta_{M+1-i} \to -\delta\eta_i$, so these two must be equal.
+Then we have
+%
+\begin{equation}\label{eq:detahalf}
+\begin{aligned}
+  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i &= 0,
+  \quad\text{for}\quad 1<i<\frac{M}{2}\\
+  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
+  \frac{\delta\eta_{M/2-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{M/2} &= \frac{1}{s}.
+\end{aligned}
+\end{equation}
+%
+
+The first of these equations is solved by
+%
+\begin{equation}\label{eq:detagensol}
+  \delta\eta_i = B\alpha^{i-1} + C\alpha^{1-i},
+  \qquad\text{where}\qquad
+  \alpha^2 - 2(s+1)\alpha + 1 = 0,
+\end{equation}
+%
+with $B$ and $C$ determined by the boundary conditions provided by the last two equations of \eqref{eq:detahalf}:
+%
+\begin{equation}\label{eq:unicoeffs}
+  B = -\frac{2\alpha}{(1+\alpha)(\alpha^{M/2}+\alpha^{-M/2})s},
+  \qquad
+  C = -\frac{2}{(1+\alpha)(\alpha^{M/2}+\alpha^{-M/2})s}.
+\end{equation}
+%
+This can be used to compute the Laplace transform
+%
+\begin{equation}\label{eq:unilaplace}
+  A(s) = \frac{2\eta^w_1(s)}{M} 
+   = \frac{2}{M}\prn{\frac{1}{s}+B+C} 
+   = \frac{2(\alpha^{M/2}-1)^2}{Ms(\alpha^M+1)}.
+\end{equation}
+%
+
+In reality, $M$ can only take even integer values.
+However, one can find a heuristic approximation to the envelope by maximising \wrt $M$.
+We find that
+%
+\begin{multline}\label{eq:unienv}
+  s = 2\sinh^2\frac{y_*}{M_*},
+  \qquad
+  \hat{A}_*(s) = \prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\log\alpha(s)}{s},
+  \\ \text{where}\quad 
+  y_* = \tanh \frac{y_*}{2} \cosh y_* \approx 1.50553.
+  \qquad
+\end{multline}
+%
+This is only valid for
+%
+\begin{equation}\label{eq:univalid}
+  2 \leq M_* \leq M,
+  \qquad\implies\qquad
+  2\sinh^2 \frac{y_*}{M} \leq s \leq 2\sinh^2 \frac{y_*}{2}.
+\end{equation}
+%
+
+When $M \gg 2$, the validity region is approximately
+%
+\begin{equation}\label{eq:univalidapprox}
+  \frac{4.53327}{M^2} \lesssim s \lesssim 1.36423.
+\end{equation}
+%
+When $s \sim \CO(M^{-2}) \ll 1$, we have $\alpha \approx 1+\sqrt{2s}$ and the heuristic envelope is approximately
+%
+\begin{equation}\label{eq:unienvapprox}
+  \hat{A}_*(s) \approx \frac{ 1.08406 }{ \sqrt{s} }.
+\end{equation}
+%
+
+For $s\gtrsim 1.36423$, we set $M=2$ to find
+%
+\begin{equation}\label{eq:binaryenv}
+  \hat{A}(s) = \frac{1}{1+s}.
+\end{equation}
+%
+This does not match up with \eqref{eq:unienvapprox} at the boundary of the two regions, as this is outside the regime of validity of that approximation.
+However, it does match up with the full formula, \eqref{eq:unienv}.
+
+\subsection{Sticky serial model}\label{sec:serial_sticky}
+
+Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_sticky}.
+We proceed in the same manner as in \autoref{sec:serial_uniform}, with the following differences.
+
+The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
+%
+\begin{equation}\label{eq:stickyenc}
+  \begin{aligned}
+  \enc_{12} = -\enc_{MM-1} &= \frac{\varepsilon}{2},
+  &\qquad
+  \eq_1 = \eq_M &= \frac{1}{2+(M-2)\varepsilon},
+  \\ &&
+  \eq_i &=  \frac{\varepsilon}{2+(M-2)\varepsilon}
+  \quad\text{for}\quad 1<i<M.
+  \end{aligned}
+\end{equation}
+%
+This means that the Laplace transform is given by
+%
+\begin{equation}\label{eq:stickyareaeta}
+  \hat{A}(s) = \frac{\varepsilon(\eta^w_1(s)-\eta^w_M(s))}{2+(M-2)\varepsilon} 
+       = \frac{2\varepsilon\, \eta^w_1(s)}{2+(M-2)\varepsilon}.
+\end{equation}
+%
+
+The $\eta^w_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
+%
+\begin{equation}\label{eq:detabndsticky}
+\begin{aligned}
+  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{\varepsilon}{2}}\delta\eta_1 &= 0, \\
+  \frac{\delta\eta_{M/2-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{M/2} &= \frac{1}{s}.
+\end{aligned}
+\end{equation}
+%
+These are solved by
+%
+\begin{equation}\label{eq:stickycoeffs}
+\begin{aligned}
+  B &= -\frac{ 2 \alpha^{M/2} (\alpha-(1-\varepsilon)) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s}, \\
+  C &= -\frac{ 2 \alpha^{M/2} (1-(1-\varepsilon)\alpha) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s},
+\end{aligned}
+\end{equation}
+%
+which leads to the Laplace transform
+%
+\begin{equation}\label{eq:stickylaplace}
+  \hat{A}(s) = \frac{2\varepsilon}{2+(M-2)\varepsilon} \, 
+      \frac{ (\alpha^{M/2}-1)^2 - (1-\varepsilon)\alpha(\alpha^{M/2-1}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }.
+\end{equation}
+%
+
+We can maximise this \wrt $\varepsilon$.
+This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
+For small $s \ll M^{-2}$, it is given by
+%
+\begin{equation}\label{eq:stickyeps}
+  \varepsilon = \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},  
+\end{equation}
+%
+and the Laplace transform is 
+%
+\begin{equation}\label{stickyenv}
+  \hat{A}(s) = (M-1) \prn{ 1 - \sqrt{M(M-2)s} + \CO(M^2s)}.
+\end{equation}
+%
+
+\section{Shortened serial model}\label{sec:serial_shorten}
+
+Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_shorten}.
+We proceed in the same manner as in \autoref{sec:serial_uniform}, with the following differences.
+
+The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
+The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
+%
+\begin{equation}\label{eq:shortenenc}
+  \begin{aligned}
+  \enc_{M-1M} = -\enc_{21} &= \frac{\varepsilon}{2},
+  &\qquad
+  \eq_1 = \eq_M &= \frac{\varepsilon}{M-2(1-\varepsilon)},
+  \\ &&
+  \eq_i &=  \frac{1}{M-2(1-\varepsilon)}
+  \quad\text{for}\quad 1<i<M.
+  \end{aligned}
+\end{equation}
+%
+This means that the Laplace transform is given by
+%
+\begin{equation}\label{eq:shortenareaeta}
+  \hat{A}(s) = \frac{\varepsilon(\eta^w_1(s)-\eta^w_M(s)) + (1-\varepsilon)(\eta^w_2(s)-\eta^w_{M-2}(s))}{M-2(1-\varepsilon)}
+       = \frac{2\varepsilon\, \eta^w_1(s) + 2(1-\varepsilon)\eta^w_2(s)}{M-2(1-\varepsilon)}.
+\end{equation}
+%
+
+The $\eta^w_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
+%
+\begin{equation}\label{eq:detabndshorten}
+\begin{aligned}
+  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
+  \frac{\delta\eta_3 + \varepsilon\,\delta\eta_1}{2} - \prn{s+\frac{1+\varepsilon}{2}}\delta\eta_2 &= 0, \\
+  \frac{\delta\eta_{M/2-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{M/2} &= \frac{1}{s}.
+\end{aligned}
+\end{equation}
+%
+These are solved by
+%
 

### 2015-06-18 Notes/laplace.tex
written up shotened heuristics

@@
   \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{M/2+1}+\delta\eta_{M/2-1}}{2} - (s+1)\delta\eta_{M/2} &= \frac{1}{s},\\
-  \frac{\delta\eta_{M/2+2}+\delta\eta_{M/2}}{2} - (s+1)\delta\eta_{M/2+1} &= -\frac{1}{s},\\
+  \frac{\delta\eta_{\frac{M}{2}+1}+\delta\eta_{\frac{M}{2}-1}}{2} - (s+1)\delta\eta_{\frac{M}{2}} &= \frac{1}{s},\\
+  \frac{\delta\eta_{\frac{M}{2}+2}+\delta\eta_{\frac{M}{2}}}{2} - (s+1)\delta\eta_{\frac{M}{2}+1} &= -\frac{1}{s},\\
   \frac{\delta\eta_{M-1}}{2} - \prn{s+\frac{1}{2}}\delta\eta_M &= 0.
@@
   \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{M/2-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{M/2} &= \frac{1}{s}.
+  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
 \end{aligned}
@@
 \begin{equation}\label{eq:unicoeffs}
-  B = -\frac{2\alpha}{(1+\alpha)(\alpha^{M/2}+\alpha^{-M/2})s},
-  \qquad
-  C = -\frac{2}{(1+\alpha)(\alpha^{M/2}+\alpha^{-M/2})s}.
+  B = -\frac{2\alpha}{(1+\alpha)(\alpha^{\frac{M}{2}}+\alpha^{-\frac{M}{2}})s},
+  \qquad
+  C = -\frac{2}{(1+\alpha)(\alpha^{\frac{M}{2}}+\alpha^{-\frac{M}{2}})s}.
 \end{equation}
@@
    = \frac{2}{M}\prn{\frac{1}{s}+B+C} 
-   = \frac{2(\alpha^{M/2}-1)^2}{Ms(\alpha^M+1)}.
+   = \frac{2(\alpha^{\frac{M}{2}}-1)^2}{Ms(\alpha^M+1)}.
 \end{equation}
@@
   \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{\varepsilon}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{M/2-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{M/2} &= \frac{1}{s}.
+  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
 \end{aligned}
@@
 \begin{aligned}
-  B &= -\frac{ 2 \alpha^{M/2} (\alpha-(1-\varepsilon)) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s}, \\
-  C &= -\frac{ 2 \alpha^{M/2} (1-(1-\varepsilon)\alpha) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s},
+  B &= -\frac{ 2 \alpha^{\frac{M}{2}} (\alpha-(1-\varepsilon)) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s}, \\
+  C &= -\frac{ 2 \alpha^{\frac{M}{2}} (1-(1-\varepsilon)\alpha) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s},
 \end{aligned}
@@
 \begin{equation}\label{eq:stickylaplace}
-  \hat{A}(s) = \frac{2\varepsilon}{2+(M-2)\varepsilon} \, 
-      \frac{ (\alpha^{M/2}-1)^2 - (1-\varepsilon)\alpha(\alpha^{M/2-1}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }.
+  \hat{A}(s) = \frac{2\varepsilon}{[2+(M-2)\varepsilon]s} \, 
+      \frac{ (\alpha^{\frac{M}{2}}-1)^2 - (1-\varepsilon)\alpha(\alpha^{\frac{M-2}{2}}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }.
 \end{equation}
@@
 This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
+We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=1$:
+%
+\begin{equation}\label{eq:stickyvalid}
+\begin{aligned}
+  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
+    M\alpha[(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^M+1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M-2}+1)]
+     \\ &+ 2(\alpha^{\frac{M}{2}}-1)^2(\alpha^M+1) = 0,\\
+    s = &\, \frac{(\alpha-1)^2}{2\alpha}.
+\end{aligned}
+\end{equation}
+%
+The ``sticky'' models are optimal at values of $s$ smaller than this.
+This is not quite the same as the lower boundary of the region \eqref{eq:univalid}, so there is a gap between them where the uniform model with $M$ states is optimal.
+The width of this gap is insignificant for larger $M$.
+
 For small $s \ll M^{-2}$, it is given by
@@
 %
-\begin{equation}\label{stickyenv}
+\begin{equation}\label{eq:stickyenv}
   \hat{A}(s) = (M-1) \prn{ 1 - \sqrt{M(M-2)s} + \CO(M^2s)}.
@@
 
-\section{Shortened serial model}\label{sec:serial_shorten}
+\subsection{Shortened serial model}\label{sec:serial_shorten}
 
@@
   \frac{\delta\eta_3 + \varepsilon\,\delta\eta_1}{2} - \prn{s+\frac{1+\varepsilon}{2}}\delta\eta_2 &= 0, \\
-  \frac{\delta\eta_{M/2-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{M/2} &= \frac{1}{s}.
+  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
 \end{aligned}
@@
 %
+%
+\begin{equation}\label{eq:shortensol}
+\begin{aligned}
+  \delta\eta_i &= B\alpha^{i-2} + C\alpha^{2-i} \qquad
+        \text{for} \quad 2 \leq i \leq \frac{M}{2}, \\
+  \delta\eta_1 &= \frac{2\alpha^{\frac{M}{2}}}
+         {[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
+  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + \varepsilon(\alpha-1))}
+         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
+  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - \varepsilon\alpha(\alpha-1))}
+         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
+\end{aligned}
+\end{equation}
+%
+Which gives the following Laplace transform
+%
+\begin{equation}\label{eq:shortenlaplace}
+  \hat{A}(s) = \frac{2 [\varepsilon(\alpha^{\frac{M}{2}}-1)^2
+          + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
+          { [M-2(1-\varepsilon)]
+            [\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
+\end{equation}
+%
+This interpolates between \eqref{eq:unilaplace} at $\varepsilon=1$ and the same expression at $M-2$ at $\varepsilon=0$.
+
+We can maximise this \wrt $\varepsilon$.
+This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
+We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0\text{ or }1$:
+%
+\begin{equation}\label{eq:shortenvalid}
+\begin{aligned}
+  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
+    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
+      = 0,\\
+  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
+    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
+      = 0,\\
+    s = &\, \frac{(\alpha-1)^2}{2\alpha},
+\end{aligned}
+\end{equation}
+%
+with the $\varepsilon=1$ case providing the lower boundary and $\varepsilon=0$ providing the upper boundary.
+If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
+In this case, however, there is an overlap.
+This is related to the fact that the video shows that the optimal models do not6 shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
+
+\subsection{Heuristic envelope}\label{sec:heuristicenv}
+
+We can obtain a good approximation to the envelope by combining the results \autoref{sec:serial_uniform} and \autoref{sec:serial_sticky}, ignoring the complications of \autoref{sec:serial_shorten}.
+We will also ignore the small gap between the regions of validity, \eqref{eq:univalid} and \eqref{eq:stickyvalid}.
+
+We will phrase it in terms of $\snrb(\tau)$, reintroducing $r$:
+%
+\begin{equation}\label{eq:envunits}
+  \snrb(\tau) = \frac{ \sqrt{N} \hat{A}\prn{\frac{1}{r\tau}} }{ r\tau },
+  \qquad
+  \alpha(\tau) = \frac{ 1 + r\tau +\sqrt{1+2r\tau} }{ r\tau }.
+\end{equation}
+%
+Then, combining \eqref{eq:binaryenv}, \eqref{eq:unienv} and \eqref{eq:stickylaplace}, we find
+%
+\begin{equation}\label{eq:henv}
+  \snrb(\tau) = \left\{
+\begin{aligned}
+  & \frac{\sqrt{N}}{1+r\tau}, & 
+  & r\tau \leq \frac{1}{2\sinh^2\frac{y_*}{2}} , \\
+  & \sqrt{N}\prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \log\alpha, & 
+  \frac{1}{2\sinh^2 \frac{y_*}{2}} \leq & r\tau \leq \frac{1}{2\sinh^2 \frac{y_*}{M}} , \\
+  & \max_\varepsilon \brc{\frac{2\sqrt{N}\varepsilon}{2+(M-2)\varepsilon} \,
+      \frac{ (\alpha^{\frac{M}{2}}-1)^2 - (1-\varepsilon)\alpha(\alpha^{\frac{M-2}{2}}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }}, &
+  \frac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau .
+\end{aligned}
+\right.
+\end{equation}
+%
+For $M \gg 1$, and $r\tau$ much greater than the lower bound of each region, we can approximate this as 
+%
+\begin{equation}\label{eq:henvapproc}
+  \snrb(\tau) \approx \left\{
+\begin{aligned}
+  & \frac{\sqrt{N}}{1+r\tau}, &
+  & r\tau \leq 0.73 , \\
+  & \frac{ 1.1\sqrt{N}}{ \sqrt{r\tau} } , &
+  0.73 \leq & r\tau \leq 0.22 M^2 , \\
+  & \frac{ \sqrt{N} (M-1) }{ r\tau }, &\quad
+  0.22 M^2 \leq &r\tau .
+\end{aligned}
+\right.
+\end{equation}
+%
+This is plotted in \autoref{fig:heuristicenv}.
+
+\begin{figure}[tb]
+  \centering
+  % Requires \usepackage{graphicx}
+  \includegraphics[width=0.8\linewidth]{LenvHeuristic.svg}\\
+  \caption{Heuristic envelope.
+  Analytic expressions are only valid to the right of each region.}\label{fig:heuristicenv}
+\end{figure}
 

### 2015-06-18 Code/Figs/RunAveFigHeuristic.m

@@ -1,0 +1,64 @@
+
+load('laplacebndChainA12.mat','AenvChains','srange');
+
+[ AenvHeuristic,sbnd] = SerialLaplaceEnv( srange,12 );
+
+AenvChains=100*AenvChains.*srange;
+AenvHeuristic=100*AenvHeuristic.*srange;
+
+loglog( 1./srange,AenvChains,'r-',...
+    1./srange,AenvHeuristic,'g--',...
+    'LineWidth',2);
+xl=[0.1 1e3];
+yl=[0.5 500];
+ylim(yl);
+xlim(xl);
+xlabel('Mean recall time, $\tau$','Interpreter','latex','FontSize',16);
+ylabel('Recognition performance, $\overline{\mathrm{SNR}}(\tau)$','Interpreter','latex','FontSize',16);
+title('Heuristic envelope','Interpreter','latex','FontSize',20);
+legend({'Numeric envelope',...
+    'Heuristic envelope',...
+    },...
+    'Location','northeast','Interpreter','latex','FontSize',16);
+
+taubnd=1./sbnd(2:3);
+Abnd=SerialLaplaceEnv(sbnd,12);
+Abnd(1)=[];
+Abnd=100*Abnd./taubnd;
+
+line([1 1]*taubnd(1),yl,'Color','k','LineStyle','--','LineWidth',2);
+line([1 1]*taubnd(2),yl,'Color','k','LineStyle','--','LineWidth',2);
+
+
+annotation('textbox',...
+    dsxy2figxy(gca,[taubnd(2)/2 yl(1) taubnd(2)*2 yl(1)*2]),...
+    'String','$r\tau=0.73$',...
+    'Interpreter','latex','FontSize',16,...
+    'LineStyle','none','VerticalAlignment','bottom','HorizontalAlignment','center');
+
+annotation('textbox',...
+    dsxy2figxy(gca,[taubnd(1)/2 yl(1) taubnd(1)*2 yl(1)*2]),...
+    'String','$r\tau=0.22M^2$',...
+    'Interpreter','latex','FontSize',16,...
+    'LineStyle','none','VerticalAlignment','bottom','HorizontalAlignment','center');
+
+annotation('textbox',...
+    dsxy2figxy(gca,[xl(1) AenvHeuristic(end) taubnd(2) 2*AenvHeuristic(end)]),...
+    'String','$\frac{\sqrt{N}}{1+r\tau}$',...
[…]

### 2015-06-18 Notes/laplace.tex
figure corresponding to video

@@
   The dashed red line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
-  The solid red line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.}\label{fig:envnum}
+  The solid red line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
+  The solid grren line indicates the proven envelope, \eqref{eq:env}.}\label{fig:envnum}
+\end{figure}
+
+\begin{figure}[tbp]
+  \centering
+  % Requires \usepackage{graphicx}
+  \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv07.eps}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv13.eps}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv21.eps}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv27.eps}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv36.eps}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv51.eps}}
+  \end{myenuma}
+  \caption{The models that constitute the numeric envelope in \autoref{fig:envnum}.
+  The plot to the left of each frame indicates the memory curve of the model shown to the right in blue, with the numeric envelope for models with the serial topology in red and the proven envelope, \eqref{eq:env}, in green.
+  The bar graph in the upper right of each frame indicates the transition probabilities from the state to the left of each bar to the state on its right, under potentiation.
+  The bar graph in the lower right of each frame indicates the transition probabilities from the state to the right of each bar to the state on its left, under depression.
+  The heat map in the centre right of each frame indicates the equilibrium probability distribution for each state.
+  See the video for more details.}\label{fig:envvid}
 \end{figure}

### 2015-06-18 Notes/laplace.tex
conclusion

@@
 \begin{abstract}
-  We try to find the continuous time Markov process that has the maximal Laplace tranformed signal-to-noise curve.
+  We try to find the continuous time Markov process that has the maximal Laplace transformed signal-to-noise curve.
 \end{abstract}
@@
 \tableofcontents
+
+\listoffigures
 
@@
   \includegraphics[width=0.8\linewidth]{LenvProven.eps}
-  \caption{The memory curve envelope that we can prove, \eqref{eq:env}.
+  \caption[Proven envelope]{The memory curve envelope that we can prove, \eqref{eq:env}.
   The number of synapses $N=10^4$ and the number of states $M=12$.
@@
   \includegraphics[width=0.8\linewidth]{LenvNum.eps}\\
-  \caption{Numerical envelope for the memory curve, \eqref{eq:snrb}, when $f\potdep=\frac{1}{2}$, the number of synapses $N=10^4$ and the number of states $M=12$.
+  \caption[Numerical envelope]{Numerical envelope for the memory curve, \eqref{eq:snrb}, when $f\potdep=\frac{1}{2}$, the number of synapses $N=10^4$ and the number of states $M=12$.
   The dashed red line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
@@
   \end{myenuma}
-  \caption{The models that constitute the numeric envelope in \autoref{fig:envnum}.
+  \caption[Optimal models]{The models that constitute the numeric envelope in \autoref{fig:envnum}.
   The plot to the left of each frame indicates the memory curve of the model shown to the right in blue, with the numeric envelope for models with the serial topology in red and the proven envelope, \eqref{eq:env}, in green.
@@
   \end{myenuma}
-  \caption{Heuristics for models that maximise memory curve.
+  \caption[Heuristic optimal models]{Heuristics for models that maximise memory curve.
   All unlabelled arrowshave transition probabilities equal to 1.
@@
   \includegraphics[width=0.8\linewidth]{LenvHeuristic.svg}\\
-  \caption{Heuristic envelope.
+  \caption[Heuristic envelope]{Heuristic envelope.
   Analytic expressions are only valid to the right of each region.}\label{fig:heuristicenv}
 \end{figure}
+
+\section{Conclusions}\label{sec:conclusions}
+
+We do not expect to find precisely these models inside real synapses.
+First, evolution was faced with stronger constraints than us when designing these synapses as it had to build them out of real molecules.
+Second, the set of priorities faced by evolution would have been longer than just performance in recogniton memory at a single timescale.
+
+Nevertheless, the qualitative message of our findings should be more robust to reality:
+%
+\begin{itemize}
+  \item \emph{Short} timescale $\to$ \emph{intermediate} timescales: transition topology goes \\ short and wide $\to$ long and thin.
+  \item \emph{Intermediate} timescale $\to$ \emph{long} timescales: transition probabilities go \\ strong and deterministic $\to$ weak and stochastic.
+\end{itemize}
+%
+This is what we expect to find in the different brain regions that store memories for different timescales.
 

### 2015-06-22 Notes/laplace.tex
reorganised

@@
   The heat map in the centre right of each frame indicates the equilibrium probability distribution for each state.
-  See the video for more details.}\label{fig:envvid}
+  See the video for the full set of models.}\label{fig:envvid}
 \end{figure}
@@
 
-Looking at them models that were found by numerical maximisation in \autoref{fig:envnum} (see the video for the models), suggests we look at the particular set of models, shown in \autoref{fig:heuristicmodels}.
+Looking at them models that were found by numerical maximisation in \autoref{fig:envvid} (see the video for the full set of models), suggests we look at the particular set of models, shown in \autoref{fig:heuristicmodels}.
 
@@
 At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
-Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from the video, this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
+Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from \autoref{fig:envvid} and the video, this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
 Once the length has been reduced to $M=2$, the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.
@@
 In this case, however, there is an overlap.
-This is related to the fact that the video shows that the optimal models do not6 shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
+This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not6 shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
 

### 2015-07-14 Notes/laplace.tex
hessian

@@
 \newcommand{\Lb}{\boldsymbol{\Lambda}}
+\newcommand{\etw}{\boldsymbol{\eta}^w}
+\newcommand{\thb}{\boldsymbol{\theta}}
 %synapse models
@@
 \begin{equation}\label{eq:lwpmixtime}
-  \eta^w_i(s) = \sum_j \fptb_{ij}(s)\eq_j \w
+  \etw_i(s) = \sum_j \fptb_{ij}(s)\eq_j \w_j
     = \eta^+_i(s) - \eta^-_i(s) =2\eta^+_i(s) - \eta(s) =\eta(s) - 2\eta^-_i(s) .
@@
 %
-Arranging the states in order of decreasing $\eta^w_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
+Arranging the states in order of decreasing $\etw_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
 These quantities can still be used when $\w$ takes more than two values.
@@
     &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
-    &= \sum_{ij}  \eq_i \enc_{ij} (\eta^w_{i}(s) - \eta^w_{j}(s)).
+    &= \sum_{ij}  \eq_i \enc_{ij} (\etw_{i}(s) - \etw_{j}(s)).
 \end{aligned}
@@
     c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
-    a_i(s) &= \sum_j \eq_i \enc_{ij} (\eta^w_{i}(s) - \eta^w_{j}(s)),\\
+    a_i(s) &= \sum_j \eq_i \enc_{ij} (\etw_{i}(s) - \etw_{j}(s)),\\
     \implies
@@
 
-Using the same approach as the area bound in \cite{Lahiri2013synapse}, and defining $\theta_i(s) = \sum_j \fptb_{ij}(0) a_j(s)$,  we find that the derivatives are
+Using the same approach as the area bound in \cite{Lahiri2013synapse}, and defining $\thb_i(s) = \sum_j \fptb_{ij}(0) a_j(s)$,  we find that the derivatives are
 %
 \begin{equation}\label{eq:lAdiff}
-  \pdiff{\hat{A}(s)}{\Mh\potdep_{mn}}
-     = f\potdep \eq_m \prn{ r \brk{\theta_m(s) - \theta_n(s)}
-     +\prn{rc_m(s) \pm 1} \brk{\eta^w_m(s) - \eta^w_n(s)} },
+  \pdiff{\hat{A}(s)}{\M\potdep_{mn}}
+     = f\potdep \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
+     +\prn{rc_m(s) \pm 1} \brk{\etw_m(s) - \etw_n(s)} },
 \end{equation}
@@
 with the upper/lower sign in $\pm$ corresponding to derivatives \wrt $\M\potdep$.
-The term with $\Theta(s)$ comes from differentiating $\eq$, the term with $c(s)$ comes from differentiating $\eta^w(s)$ and the term with $\pm1$ comes from differentiating $\enc$ in \eqref{eq:larea}.
+The term with $\Theta(s)$ comes from differentiating $\eq$, the term with $c(s)$ comes from differentiating $\etw(s)$ and the term with $\pm1$ comes from differentiating $\enc$ in \eqref{eq:larea}.
+
+To write down the Hessian, we will first define some notation.
+Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$
+Then we denote
+%
+\begin{equation}\label{eq:hessparts}
+\begin{aligned}
+  R_{ijkl} &= 
+    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
+    &&&+ \eq_i \brc{[\fund(0)\enc\fund(s)]_{ik} - [\fund(0)\enc\fund(s)]_{jk}} \brc{\etw_k(s) - \etw_l(s)} \\
+    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)},  \\
+  S_{ijkl} &= 
+    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
+    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
+\end{aligned}
+\end{equation}
+%
+Then we can write the hessian as
+%
+\begin{equation}\label{eq:hessian}
+\begin{aligned}
+  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
+     (f\pot)^2 (R_{ijkl} + R\trans_{ijkl})
+     + f\pot(S_{ijkl} + S\trans_{ijkl}), \\
+  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
+     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
+     - f\pot S_{ijkl} + f\dep S\trans_{ijkl}, \\
+  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
+     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
+     - f\pot S\trans_{ijkl} + f\dep S_{ijkl}, \\
+  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
+     (f\dep)^2 (R_{ijkl} + R\trans_{ijkl})
+     - f\dep(S_{ijkl} + S\trans_{ijkl}), \\
+\end{aligned}
+\end{equation}
+%
+This is useful in numerical maximisation.
 
@@
 \begin{aligned}
-  \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\eta^w_i(s) - \eta^w_j(s))
+  \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\etw_i(s) - \etw_j(s))
         +\sum_{i j} \prn{ f\pot\mu\pot_{ij} \M\pot_{ij} + f\dep\mu\dep_{ij} \M\dep_{ij} }\\
@@
   \frac{1}{f\potdep} \pdiff{\CL}{\Mh\potdep_{mn}} =& \,
-      \eq_m \prn{ r \brk{\theta_m(s) - \theta_n(s)}
-     +\prn{rc_m(s) \pm 1} \brk{\eta^w_m(s) - \eta^w_n(s)} } \\
+      \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
+     +\prn{rc_m(s) \pm 1} \brk{\etw_m(s) - \etw_n(s)} } \\
      &+ \mu\potdep_{mn} -  \mu\potdep_{mm}
@@
 Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_uniform}.
-Looking at \eqref{eq:larea}, the only thing we need is differences in $\eta^w_i(s)$.
+Looking at \eqref{eq:larea}, the only thing we need is differences in $\etw_i(s)$.
 In fact, given that the only nonzero elements of $\enc$ are
@@
 \begin{equation}\label{eq:uniareaeta}
-  \hat{A}(s) = \frac{\eta^w_1(s) - \eta^w_M(s)}{M}.
+  \hat{A}(s) = \frac{\etw_1(s) - \etw_M(s)}{M}.
 \end{equation}
@@
 \begin{equation}\label{eq:etafund}
-  \eta^w(s) = \text{const.} - \fund(s)\w.
-\end{equation}
-%
-As we are only interested in differences of the $\eta^w_i(s)$, we can drop the constant.
+  \etw(s) = \text{const.} - \fund(s)\w.
+\end{equation}
+%
+As we are only interested in differences of the $\etw_i(s)$, we can drop the constant.
 Then
@@
 \begin{equation}\label{eq:etarecur}
-  -(s\I + \onev\pib - \frg)\eta^w(s) = \w.
+  -(s\I + \onev\pib - \frg)\etw(s) = \w.
 \end{equation}
@@
 \begin{equation}\label{eq:detarecur}
-  \eta^w(s) = -\frac{(\pib\eta^w(s))\onev+\w}{s} + \delta\eta(s),
+  \etw(s) = -\frac{(\pib\etw(s))\onev+\w}{s} + \delta\eta(s),
   \qquad
@@
 %
-As we are only interested in differences, we can drop any terms proportional to $\onev$ in $\eta^w(s)$.
+As we are only interested in differences, we can drop any terms proportional to $\onev$ in $\etw(s)$.
 
@@
 \begin{equation}\label{eq:unilaplace}
-  A(s) = \frac{2\eta^w_1(s)}{M}
+  A(s) = \frac{2\etw_1(s)}{M}
    = \frac{2}{M}\prn{\frac{1}{s}+B+C}
@@
 \begin{equation}\label{eq:stickyareaeta}
-  \hat{A}(s) = \frac{\varepsilon(\eta^w_1(s)-\eta^w_M(s))}{2+(M-2)\varepsilon}
-       = \frac{2\varepsilon\, \eta^w_1(s)}{2+(M-2)\varepsilon}.
-\end{equation}
-%
-
-The $\eta^w_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
+  \hat{A}(s) = \frac{\varepsilon(\etw_1(s)-\etw_M(s))}{2+(M-2)\varepsilon}
+       = \frac{2\varepsilon\, \etw_1(s)}{2+(M-2)\varepsilon}.
+\end{equation}
+%
+
+The $\etw_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
 %
@@
 \begin{equation}\label{eq:shortenareaeta}
-  \hat{A}(s) = \frac{\varepsilon(\eta^w_1(s)-\eta^w_M(s)) + (1-\varepsilon)(\eta^w_2(s)-\eta^w_{M-2}(s))}{M-2(1-\varepsilon)}
-       = \frac{2\varepsilon\, \eta^w_1(s) + 2(1-\varepsilon)\eta^w_2(s)}{M-2(1-\varepsilon)}.
-\end{equation}
-%
-
-The $\eta^w_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
+  \hat{A}(s) = \frac{\varepsilon(\etw_1(s)-\etw_M(s)) + (1-\varepsilon)(\etw_2(s)-\etw_{M-2}(s))}{M-2(1-\varepsilon)}
+       = \frac{2\varepsilon\, \etw_1(s) + 2(1-\varepsilon)\etw_2(s)}{M-2(1-\varepsilon)}.
+\end{equation}
+%
+
+The $\etw_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
 %

### 2015-09-05 Notes/laplace.tex
hzgdxccsk s

@@
 This does not match up with \eqref{eq:unienvapprox} at the boundary of the two regions, as this is outside the regime of validity of that approximation.
-However, it does match up with the full formula, \eqref{eq:unienv}.
+However, it does match up with the full formula, \eqref{eq:unienv} at the boundary.
 
@@
 %
-This interpolates between \eqref{eq:unilaplace} at $\varepsilon=1$ and the same expression at $M-2$ at $\varepsilon=0$.
+This interpolates between \eqref{eq:unilaplace} at $\varepsilon=1$ and the same expression for $M-2$ at $\varepsilon=0$.
 
@@
 This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
-We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0\text{ or }1$:
+We can find the boundary of the region of applicability by asking at what value of $s$ is the optimal value of $\varepsilon=0$ or $1$?
 %
@@
 In this case, however, there is an overlap.
-This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not6 shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
+This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
 
@@
 \begin{itemize}
-  \item \emph{Short} timescale $\to$ \emph{intermediate} timescales: transition topology goes \\ short and wide $\to$ long and thin.
-  \item \emph{Intermediate} timescale $\to$ \emph{long} timescales: transition probabilities go \\ strong and deterministic $\to$ weak and stochastic.
+  \item \emph{Short} timescale $\to$ \emph{intermediate} timescales: transition topology goes from \\ short and wide $\to$ long and thin.
+  \item \emph{Intermediate} timescale $\to$ \emph{long} timescales: transition probabilities go from \\ strong and deterministic $\to$ weak and stochastic.
 \end{itemize}

### 2015-09-24 Code/Figs/CascMSFig.m
redesigned poster

@@ -1,20 +1,26 @@
-t=10.^(-1:0.2:3);
+t=10.^(-1:0.2:4);
 x=0.5;
-q=(x.^abs((1:19)-10))/(1-x);
-[Wp,Wm,w]=CascadeMSinterp(q,q,0.5,1);
-casc=SNRcurve(t,Wp,Wm,0.5,w);
+% q=(x.^abs((1:19)-10))/(1-x);
+% [Wp,Wm,w]=CascadeMSinterp(q,q,0.5,1);
+% casc=SNRcurve(t,Wp,Wm,0.5,w);
+casc_model=SynapseMemoryModel.Build(@CascadeBuilder,0.5,{20,x});
+casc=casc_model.SNRrunAve(t);
 %%
-qm=ones(1,19);
-[Wp,Wm,w]=MakeSMS(qm);
-m1=SNRcurve(t,Wp,Wm,0.5,w);
+% qm=ones(1,19);
+% [Wp,Wm,w]=MakeSMS(qm);
+% m1=SNRcurve(t,Wp,Wm,0.5,w);
+serial_model=SynapseMemoryModel.Build(@SerialBuilder,0.5,{20,0.5});
+serial=serial_model.SNRrunAve(t);
 %%
-plot(t,casc,'b',t,m1,'g','LineWidth',2)
+plot(t,casc,'b',t,serial,'g','LineWidth',2)
 set(gca,'XScale','log','YScale','log','FontSize',20)
 xlim([t(1) t(end)])
 ylim([casc(end) casc(1)])
-xlabel('Time','FontSize',20)
-ylabel('SNR','FontSize',20)
+% xlabel('Time','FontSize',20)
+% ylabel('SNR','FontSize',20)
+xlabel('Mean recall time, \tau','FontSize',20);
+ylabel('$\overline{\mathsf{SNR}}\mathsf{(\tau)}$','Interpreter','latex','FontSize',20);
 legend({'Cascade','Serial'},'Location','Best')
 set(gcf, 'PaperPositionMode', 'auto');
 %%
-clear x q Wp Wm w t m1 casc
+clear x q Wp Wm w t serial casc casc_model serial_model

### 2015-09-24 Code/Figs/RunAveProven.m

@@ -2,19 +2,29 @@ function [ fig ] = RunAveProven( srange,M,sqrtNum )
 %RUNAVEPROVEN Summary of this function goes here
 %   Detailed explanation goes here
 
-fig=figure;
+Interpreter='tex';
+
+fig=figure('PaperPositionMode','auto');
 ax=axes('Parent',fig);
 
-tau=wrev(1./srange);
+tau=(1./srange);
 
 AenvProven=(sqrtNum)*(M-1)./(tau+(M-1));
 
 loglog(ax, tau,AenvProven,'g',...
     'LineWidth',2);
-ylim([AenvProven(1)*srange(1) 5*(sqrtNum)]);
-xlabel('Mean recall time, $\tau$','Interpreter','latex','FontSize',16);
-ylabel('Recognition performance, $\overline{\mathrm{SNR}}(\tau)$','Interpreter','latex','FontSize',16);
-title('Proven envelope','Interpreter','latex','FontSize',20);
+ylim([AenvProven(end) 2*(sqrtNum)]);
+switch Interpreter
+    case 'tex'
+        xlabel(ax,'Mean recall time, \tau','FontSize',16);
+        ylabel(ax,'\textsf{Recognition performance}, $\overline{\mathsf{SNR}}\mathsf{(\tau)}$','Interpreter','latex','FontSize',16);
+    case 'latex'
+        xlabel(ax,'Mean recall time, $\mathsf{\tau}$','Interpreter','latex','FontSize',16);
+        ylabel(ax,'Recognition performance, $\overline{\mathrm{SNR}}(\tau)$','Interpreter','latex','FontSize',16);
+end
+% xlabel('Mean recall time, $\tau$','Interpreter','latex','FontSize',16);
+% ylabel('Recognition performance, $\overline{\mathrm{SNR}}(\tau)$','Interpreter','latex','FontSize',16);
+title('Proven envelope','Interpreter',Interpreter,'FontSize',20);
 
 area = (sqrtNum)*(M-1)./tau;
 initial = (sqrtNum)*ones(size(tau));
@@ -28,24 +38,44 @@ loglog(ax, tau(1:fromstart), initial(1:fromstart), 'k--',...
     'LineWidth',1);
 
 
-annotation('textbox',...
-    dsxy2figxy(ax,[tau(1) sqrtNum tau(fromstart) 4*sqrtNum]),...
-    'String','$\sqrt{N}$',...
-    'Interpreter','latex','FontSize',16,...
[…]

### 2015-11-13 Cosyne/Timecales.tex
cosyne abstract

\documentclass[12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[a4paper,margin=0.5in]{geometry}

\usepackage{titling}
\pretitle{\begin{center}\bfseries}
\posttitle{\end{center}}
\preauthor{\begin{center}}
\postauthor{\end{center}}
\predate{}
\postdate{}

\usepackage{titlesec}
\titleformat*{\section}{\bfseries}

\usepackage{caption}
\captionsetup{font=small,labelfont=bf}
\usepackage{floatrow}

\usepackage{graphicx}
\usepackage{epstopdf}
\epstopdfsetup{update,suffix=-generated}

\usepackage{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Tuning complex synaptic plasticity for different timescales of memory}
\author{Subhaneil Lahiri and Surya Ganguli}
\date{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\pagestyle{empty}
\maketitle
\thispagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Systems consolidation suggests that different brain regions are specialized to store memories over different timescales. Similarly, synapses mediating memory have highly complex, diverse molecular signaling pathways, varying across brain regions. This suggests the possibility that synaptic diversity across brain regions may be adapted for different timescales of memory storage. We are left with the fundamental question: what type of molecular synaptic dynamics is suitable for storing memory over any given timescale?

To address this, we systematically analyze an extremely broad class of models where synaptic plasticity is implemented by stochastic transitions between internal functional states of a sub-synaptic molecular network. Such models are essential in navigating stringent tradeoffs between learning and remembering, known as the stability-plasticity dilemma. Previous work (e.g. [Fusi, Drew, Abbott, 2005]) analyzed this tradeoff in models with only one topological structure of transitions between sub-synaptic states.  This leaves open the nature of this tradeoff over all possible sub-synaptic networks.  Rather than analyze one model, we analyze the space of all possible models and elucidate principles that determine how sub-synaptic network structure can be ideally adapted to the time-scale over which memories are stored.  We find that as this timescale increases, initially synapses are forced to grow a chain of internal states with rapid transitions, while at even longer timescales, synapses are further forced to exhibit slow stochastic transitions.

We also discuss the design of synaptic physiology experiments to test our theoretical predictions.  We find conventional methods for probing synaptic plasticity cannot discern the relevant synaptic dynamics.  Instead we propose new classes of experiments and data-analysis procedures in which more subtle protocols for probing plasticity can yield systems identification of the synaptic dynamics so crucial for storing memories at a particular time-scale.

\section*{Additional Detail}

We are extremely excited about this work as our theory allows us to relate two widely disparate observations: the systems consolidation of memories from one brain region to another and the diversity of sub-synaptic molecular networks across brain regions [Emes and Grant, 2012]. This provides us with a means to interpret, as well as predict, the nature of these differences in molecular architecture.

We model the substructure of a synapse with a set of internal functional states that determine the synaptic weight, with plasticity inducing stochastic transitions between these states. We then look at the space of all possible transition topologies and transition probabilities. We combine rigorously proven theorems and numerical optimization to maximize a signal-to-noise ratio for memories as a function of the typical recall time for different brain regions. Finding the models that maximize this measure of memory at different timescales leads to an understanding of which structures are best suited for different memory timescales. The results can be seen in \autoref{fig:env}. We find that the optimal topology at each timescale has the states arranged in a chain. There are two qualitatively different mechanisms for preserving memory traces for longer times: (1) forcing the synapse to diffuse through a longer chain of states to change its strength and (2) lowering the transition rate between these states. We see that the first mechanism is used to go from short to intermediate timescales and the second mechanism is used to go from intermediate to long timescales.

%\begin{SCfigure}
%\caption{The memory curve envelope. This is an upper bound on the SNR as a function of the mean recall time for any model at each individual timescale. Thus no model, no matter what its transition topology between functional states, can have a memory curve that exceeds this envelope at any given time-scale. The models that achieve this upper bound are indicated at various timescales. The number of synapses is N and the number of internal functional states of each synapse is M. }\label{fig:env}
%\includegraphics[width=0.6\linewidth]{LenvHeuristic.eps}
%\end{SCfigure}
\begin{figure}[t]
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=0.4\linewidth}}]{figure}{
\caption{\small{\textbf{The memory curve envelope.} This is an upper bound on the SNR as a function of the mean recall time for \textbf{\textit{any}} model at each individual timescale. Thus \textbf{\textit{no}} model, \emph{no matter what its transition topology between functional states}, can have a memory curve that exceeds this envelope at any given time-scale. The models that achieve this upper bound are indicated at various timescales. The number of synapses is $N$ and the number of internal functional states of each synapse is $M$. }}\label{fig:env}}{
\includegraphics[width=\linewidth]{LenvHeuristic.eps}}
\end{figure}

We believe that our approach of studying the entire space of models in this class is a vitally important one that should be much more widespread than it is. While studying individual models does have its own value, especially in revealing new ideas, it can be difficult to interpret without knowing how many other models can reproduce the same phenomena. Looking at all such models can uncover their essential features and lead to richer experimental predictions.

The traditional style of synaptic plasticity experiments are designed to answer questions such as ``is this molecule/gene necessary/sufficient for LTP?'' Accordingly, the synapses are subjected to a protocol involving a large barrage of spikes that would ordinarily be guaranteed to induce plasticity. To test our predictions, instead one needs to take a Systems Identification approach, where the synapses are subjected to a long sequence of smaller stimuli that do not overwhelm the synapse’s internal dynamics. Then the correlations in the synaptic weight over longer timescales can be used to infer the structure of the latent states using Hidden Markov Model techniques, as seen in \autoref{fig:fit}. Overwhelming the system with large stimuli would hide these correlations, making it difficult to fit models. Given the radically different nature of these proposed experiments, we feel our work could have a substantial impact through interactions with, and feedback from, the Cosyne community.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.99\linewidth]{FitVid939.eps}
  \caption{\small{\textbf{Simulated experiment for fitting synaptic models.} Left: the matrices of transition probabilities for the true model. Right: the result of the model fitting. Top middle: the sequence of potentiation/depotentiation the synapse is subjected to and the sequence of observed synaptic weights. Middle: the true sequence of synaptic states (green line) and the model’s estimated probability distribution (heat map).}}\label{fig:fit}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2015-11-13 Cosyne/Timecales.tex
fig captions

@@
 \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=0.4\linewidth}}]{figure}{
-\caption{\small{\textbf{The memory curve envelope.} This is an upper bound on the SNR as a function of the mean recall time for \textbf{\textit{any}} model at each individual timescale.
+\caption{\textbf{The memory curve envelope.} This is an upper bound on the SNR as a function of the mean recall time for \textbf{\textit{any}} model at each individual timescale.
 Thus \textbf{\textit{no}} model, \emph{no matter what its transition topology between functional states}, can have a memory curve that exceeds this envelope at any given time-scale.
@@
 The number of synapses is $N$ and the number of internal functional states of each synapse is $M$.
-}}\label{fig:env}}{
+}\label{fig:env}}{
 \includegraphics[width=\linewidth]{LenvHeuristic.eps}}
@@
   \includegraphics[width=0.99\linewidth]{FitVid939.eps}
-  \caption{\small{\textbf{Simulated experiment for fitting synaptic models.} Left: the matrices of transition probabilities for the true model.
+  \caption{\textbf{Simulated experiment for fitting synaptic models.} Left: the matrices of transition probabilities for the true model.
 Right: the result of the model fitting.
 Top middle: the sequence of potentiation/depotentiation the synapse is subjected to and the sequence of observed synaptic weights.
-Middle: the true sequence of synaptic states (green line) and the model’s estimated probability distribution (heat map).}}\label{fig:fit}
+Middle: the true sequence of synaptic states (green line) and the model’s estimated probability distribution (heat map).}\label{fig:fit}
 \end{figure}

### 2015-11-14 Cosyne/Timecales.tex
finished cosyne abstract

@@
 
-\title{Tuning complex synaptic plasticity for different timescales of memory}
+\title{Optimal synaptic strategies for different timescales of memory}
 \author{Subhaneil Lahiri and Surya Ganguli}
@@
 Such models are essential in navigating stringent tradeoffs between learning and remembering, known as the stability-plasticity dilemma.
-Previous work (e.g.
-[Fusi, Drew, Abbott, 2005]) analyzed this tradeoff in models with only one topological structure of transitions between sub-synaptic states.
+Previous work (e.g.\ [Fusi, Drew, Abbott, 2005]) analyzed this tradeoff in models with only one topological structure of transitions between sub-synaptic states.
 This leaves open the nature of this tradeoff over all possible sub-synaptic networks.

### 2016-01-20 Cosyne/Timecales_short.tex
short cosyne abstract

\documentclass[12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[a4paper,margin=0.5in]{geometry}

\usepackage{titling}
\pretitle{\begin{center}\bfseries}
\posttitle{\end{center}}
\preauthor{\begin{center}}
\postauthor{\end{center}}
\predate{}
\postdate{}

\usepackage{titlesec}
\titleformat*{\section}{\bfseries}

\usepackage{caption}
\captionsetup{font=small,labelfont=bf}
\usepackage{floatrow}

\usepackage{graphicx}
\usepackage{epstopdf}
\epstopdfsetup{update,suffix=-generated}

\usepackage{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Optimal synaptic strategies for different timescales of memory}
\author{Subhaneil Lahiri and Surya Ganguli}
\date{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\pagestyle{empty}
\maketitle
\thispagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Systems consolidation suggests that different brain regions are specialized to store memories over different timescales. Similarly, synapses mediating memory have highly complex, diverse molecular signaling pathways, varying across brain regions. This suggests the possibility that synaptic diversity across brain regions may be adapted for different timescales of memory storage. We are left with the fundamental question: what type of molecular synaptic dynamics is suitable for storing memory over any given timescale?

To address this, we systematically analyze an extremely broad class of models where synaptic plasticity is implemented by stochastic transitions between internal functional states of a sub-synaptic molecular network. Such models are essential in navigating stringent tradeoffs between learning and remembering, known as the stability-plasticity dilemma. Previous work (e.g.\ [Fusi, Drew, Abbott, 2005]) analyzed this tradeoff in models with only one topological structure of transitions between sub-synaptic states.  This leaves open the nature of this tradeoff over all possible sub-synaptic networks.  Rather than analyze one model, we analyze the space of all possible models and elucidate principles that determine how sub-synaptic network structure can be ideally adapted to the time-scale over which memories are stored.  We find that as this timescale increases, initially synapses are forced to grow a chain of internal states with rapid transitions, while at even longer timescales, synapses are further forced to exhibit slow stochastic transitions.

We also discuss the design of synaptic physiology experiments to test our theoretical predictions.  We find conventional methods for probing synaptic plasticity cannot discern the relevant synaptic dynamics.  Instead we propose new classes of experiments and data-analysis procedures in which more subtle protocols for probing plasticity can yield systems identification of the synaptic dynamics so crucial for storing memories at a particular time-scale.

%\section*{Additional Detail}
%
%We are extremely excited about this work as our theory allows us to relate two widely disparate observations: the systems consolidation of memories from one brain region to another and the diversity of sub-synaptic molecular networks across brain regions [Emes and Grant, 2012]. This provides us with a means to interpret, as well as predict, the nature of these differences in molecular architecture.
%
%We model the substructure of a synapse with a set of internal functional states that determine the synaptic weight, with plasticity inducing stochastic transitions between these states. We then look at the space of all possible transition topologies and transition probabilities. We combine rigorously proven theorems and numerical optimization to maximize a signal-to-noise ratio for memories as a function of the typical recall time for different brain regions. Finding the models that maximize this measure of memory at different timescales leads to an understanding of which structures are best suited for different memory timescales. The results can be seen in \autoref{fig:env}. We find that the optimal topology at each timescale has the states arranged in a chain. There are two qualitatively different mechanisms for preserving memory traces for longer times: (1) forcing the synapse to diffuse through a longer chain of states to change its strength and (2) lowering the transition rate between these states. We see that the first mechanism is used to go from short to intermediate timescales and the second mechanism is used to go from intermediate to long timescales.
%
%%\begin{SCfigure}
%%\caption{The memory curve envelope. This is an upper bound on the SNR as a function of the mean recall time for any model at each individual timescale. Thus no model, no matter what its transition topology between functional states, can have a memory curve that exceeds this envelope at any given time-scale. The models that achieve this upper bound are indicated at various timescales. The number of synapses is N and the number of internal functional states of each synapse is M. }\label{fig:env}
%%\includegraphics[width=0.6\linewidth]{LenvHeuristic.eps}
%%\end{SCfigure}
%\begin{figure}[t]
%\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=0.4\linewidth}}]{figure}{
%\caption{\textbf{The memory curve envelope.} This is an upper bound on the SNR as a function of the mean recall time for \textbf{\textit{any}} model at each individual timescale. Thus \textbf{\textit{no}} model, \emph{no matter what its transition topology between functional states}, can have a memory curve that exceeds this envelope at any given time-scale. The models that achieve this upper bound are indicated at various timescales. The number of synapses is $N$ and the number of internal functional states of each synapse is $M$. }\label{fig:env}}{
%\includegraphics[width=\linewidth]{LenvHeuristic.eps}}
%\end{figure}
%
%We believe that our approach of studying the entire space of models in this class is a vitally important one that should be much more widespread than it is. While studying individual models does have its own value, especially in revealing new ideas, it can be difficult to interpret without knowing how many other models can reproduce the same phenomena. Looking at all such models can uncover their essential features and lead to richer experimental predictions.
%
%The traditional style of synaptic plasticity experiments are designed to answer questions such as ``is this molecule/gene necessary/sufficient for LTP?'' Accordingly, the synapses are subjected to a protocol involving a large barrage of spikes that would ordinarily be guaranteed to induce plasticity. To test our predictions, instead one needs to take a Systems Identification approach, where the synapses are subjected to a long sequence of smaller stimuli that do not overwhelm the synapse’s internal dynamics. Then the correlations in the synaptic weight over longer timescales can be used to infer the structure of the latent states using Hidden Markov Model techniques, as seen in \autoref{fig:fit}. Overwhelming the system with large stimuli would hide these correlations, making it difficult to fit models. Given the radically different nature of these proposed experiments, we feel our work could have a substantial impact through interactions with, and feedback from, the Cosyne community.
%
%\begin{figure}[h]
%  \centering
%  \includegraphics[width=0.99\linewidth]{FitVid939.eps}
%  \caption{\textbf{Simulated experiment for fitting synaptic models.} Left: the matrices of transition probabilities for the true model. Right: the result of the model fitting. Top middle: the sequence of potentiation/depotentiation the synapse is subjected to and the sequence of observed synaptic weights. Middle: the true sequence of synaptic states (green line) and the model’s estimated probability distribution (heat map).}\label{fig:fit}
%\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\end{document}

### 2016-02-18 Code/Figs/RunAveModels.m
single models slide

@@ -1,0 +1,102 @@
+function [ fig ] = RunAveModels( srange,M,sqrtNum,chains,inds )
+%RUNAVEPROVENNUM Summary of this function goes here
+%   Detailed explanation goes here
+
+Interpreter='tex';
+
+fig=figure('PaperPositionMode','auto');
+ax=axes('Parent',fig);
+
+tau=(1./srange);
+
+AenvNum=(sqrtNum)*[chains.A]./tau;
+
+AenvProven=(sqrtNum)*(M-1)./(tau+(M-1));
+
+loglog(ax, tau,AenvProven,'g',...
+    tau,AenvNum,'r',...
+    'LineWidth',2);
+ylim([AenvProven(end) 2*(sqrtNum)]);
+switch Interpreter
+    case 'tex'
+        xlabel(ax,'Mean recall time, \tau','FontSize',16);
+        ylabel(ax,'\textsf{Recognition performance}, $\overline{\mathsf{SNR}}\mathsf{(\tau)}$','Interpreter','latex','FontSize',16);
+    case 'latex'
+        xlabel(ax,'Mean recall time, $\mathsf{\tau}$','Interpreter','latex','FontSize',16);
+        ylabel(ax,'Recognition performance, $\overline{\mathrm{SNR}}(\tau)$','Interpreter','latex','FontSize',16);
+end
+% xlabel('Mean recall time, $\tau$','Interpreter','latex','FontSize',16);
+% ylabel('Recognition performance, $\overline{\mathrm{SNR}}(\tau)$','Interpreter','latex','FontSize',16);
+title('Proven envelope','Interpreter',Interpreter,'FontSize',20);
+
+hold(ax, 'on');
+
+for i=inds
+    loglog(ax,tau,(sqrtNum)*chains(i).snrb,'b','LineWidth',2);
+end
+
+legend({'Proven envelope',...
+    'Numeric envelope',...
+    'Single model',...
+    },...
+    'Location','southwest','Interpreter',Interpreter,'FontSize',16);
+
+
+% area = (sqrtNum)*(M-1)./tau;
+% initial = (sqrtNum)*ones(size(tau));
+% 
+% toend=find(tau > (M-1)^0.9 * tau(end)^0.1, 1,'first');
+% fromstart=find(tau < (M-1)^0.5 * tau(1)^0.5, 1,'last');
+% 
+% hold(ax, 'on');
+% loglog(ax, tau(1:fromstart), initial(1:fromstart), 'k--',...
[…]

### 2016-03-09 Paper/SynapsePaper.tex
started pnas paper

--- 2013-11-07 NIPS/SynapseNIPS.tex
@@
-% -*- TeX -*- -*- US -*-
-\documentclass{article} % For LaTeX2e
-\usepackage{nips13submit_e,times}
-\input{sl_preamble.tex}
+%% PNAStwoS.tex
+%% For two column PNAS articles
+%% Version: Apr 15, 2008
+
+%% BASIC CLASS FILE
+\documentclass{pnastwo}
+
+%% ADDITIONAL OPTIONAL STYLE FILES
+%\usepackage{graphicx}
 \input{sl_graphics_preamble.tex}
 \graphicspath{{"Figs/"}}
-% >> Only for drafts! <<
-%\usepackage[notref,notcite]{showkeys}
-% ----------------------------------------------------------------
-%\numberwithin{equation}{section}
-%\renewcommand{\baselinestretch}{1.5}
-% ----------------------------------------------------------------
-% New commands etc.
+%\usepackage{pnastwoF}
+\usepackage{amssymb,amsfonts}
+%\usepackage[centertags]{amsmath}
+%\usepackage{hyperref}
+
+%% OPTIONAL MACRO DEFINITIONS
 \input{sl_definitions.tex}
-\input{sl_symbols.tex}
+%\input{sl_symbols.tex}
 %
 %additional symbols:
+\newcommand{\half}{\frac{1}{2}}
+\newcommand{\CO}{\mathcal{O}}
+\newcommand{\dt}{\dr t}
 %
@@
 %prob distributions
-\newcommand{\pr}{\mathbf{p}}
-\newcommand{\eq}{\pr^\infty}
+\newcommand{\prob}{\mathbf{p}}
+\newcommand{\eq}{\prob^\infty}
 %first passage times
@@
 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Title info:
-\title{A memory frontier for complex synapses}
-%
-% Author List:
-%
-\author{Subhaneil Lahiri and Surya Ganguli\\
-Department of Applied Physics, Stanford University, Stanford CA\\
-\emaillink{sulahiri@stanford.edu}, \emaillink{sganguli@stanford.edu}
-%
-}
+%\renewcommand\Fref[1]{}
+\providecommand\hyperref[2][1]{#2}
 
-\nipsfinalcopy % Uncomment for camera-ready version
+%%%%%%%%%%%%
+%% For PNAS Only:
+%\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
+\renewcommand{\theurl}{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
+\copyrightyear{2016}
+\issuedate{Issue Date}
+\volume{Volume}
+\issuenumber{Issue Number}
+%\setcounter{page}{2687} %Set page number here if desired
+%%%%%%%%%%%%
 
@@
 
+\title{Optimal synaptic strategies for different timescales of memory}
+
+\author{Subhaneil Lahiri\affil{1}{Department of Applied Physics, Stanford University, Stanford CA},
+\and
+Surya Ganguli\affil{1}{}}
+
+\contributor{Submitted to Proceedings of the National Academy of Sciences
+of the United States of America}
+
 \maketitle
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\begin{article}
 \begin{abstract}
@@
 
-%Overall, we uncover general design principles governing the functional organization of complex molecular networks, and suggest new experimental observables in synaptic physiology, based on first %passage time theory, that connect molecular complexity to memory.
+  Overall, we uncover general design principles governing the functional organization of complex molecular networks, and suggest new experimental observables in synaptic physiology, based on first %passage time theory, that connect molecular complexity to memory.
 \end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\keywords{synaptic plasticity | learning | memory | neuroscience}
 
-% references to add:
-%
-% Hopfield: original paper
-% Somp: memory capacity of Hopfield
-% Gardner: memory capacity of perceptron
-% Madison/Montgomery:  State-dependent heterogeneity in synaptic depression between pyramidal cell pairs. Neuron.
-% Abraham/Bear: Metaplasticity: the plasticity of synaptic plasticity.
-% Grant:  Evolution of synapse complexity and diversity (Ann Rev)
+\abbreviations{LTP, long term potentiation; 
+LTD, long term depression}
 
-\section{Introduction}\label{sec:intro}
-
-It is widely thought that our very ability to remember the past over long time scales depends crucially on our ability to modify synapses in our brain in an experience dependent manner.
+\dropcap{I}t is widely thought that our very ability to remember the past over long time scales depends crucially on our ability to modify synapses in our brain in an experience dependent manner.
 Classical models of synaptic plasticity model synaptic efficacy as an analog scalar value, denoting the size of a post-synaptic potential injected into one neuron from another.
@@
 However, recent experimental work has shown that many synapses are more digital than analog; they cannot robustly assume an infinite continuum of analog values, but rather can only take on a finite number of distinguishable strengths, a number than can be as small as two \cite{Bliss1993LTP,Petersen1998allornone,O'Connor2005switch} (though see \cite{Enoki2009graded}).
-This one simple modification leads to a catastrophe in memory capacity:  classical models with digital synapses, when operating in a palimpset mode in which the ongoing storage of new memories can overwrite previous memories, have a memory capacity proportional to the logarithm of the number of synapses \cite{amit1992constraints,amit1994learning}.
+This one simple modification leads to a catastrophe in memory capacity:  classical models with digital synapses, when operating in a palimpsest mode in which the ongoing storage of new memories can overwrite previous memories, have a memory capacity proportional to the logarithm of the number of synapses \cite{amit1992constraints,amit1994learning}.
 Intuitively, when synapses are digital, the storage of a new memory can flip a population of synaptic switches, thereby rapidly erasing previous memories stored in the same synaptic population.
@@
 
-Theoretical studies of complex, metaplastic synapses have focused on analyzing the memory performance of a limited number of very specific molecular dynamical systems, characterized by a number of internal states in which potentiation and depression each induce a specific set of allowable transitions between states (\eg see \autoref{fig:models} below).
+Theoretical studies of complex, metaplastic synapses have focused on analyzing the memory performance of a limited number of very specific molecular dynamical systems, characterized by a number of internal states in which potentiation and depression each induce a specific set of allowable transitions between states (\eg see \Fref{fig:models} below).
 While these models can vastly outperform simple binary synaptic switches, these analyses leave open several deep and important questions.
@@
 Much of the previous work on these types of complex synaptic models has focused on understanding the memory curves of specific models, or choices of $\M\potdep$.
-Two examples of these models are shown in \autoref{fig:models}. We see that they have different memory properties.
+Two examples of these models are shown in \Fref{fig:models}. We see that they have different memory properties.
 The serial model performs relatively well at one particular timescale, but it performs poorly at other times.
@@
   (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
-  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined in \autoref{sec:setup}) as subsequent memories are stored.
+  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined \hyperref[sec:setup]{below}) as subsequent memories are stored.
   }\label{fig:models}
@@
 Instead there are tradeoffs between storing proximal and distal memories;  often attempts to increase memory at late (early) times by changing $\M\potdep$, incurs a performance loss in memory at early (late) times in specific models considered so far \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial}.
-Thus our end goal, achieved in \autoref{sec:env}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the {\it entire} memory curve of {\it any} model.
+Thus our end goal, achieved in \sref{sec:env}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the {\it entire} memory curve of {\it any} model.
 In order to achieve this goal, in this section, we must first derive upper bounds, over the space of all possible synaptic models, on two different scalar functions of the memory curve: its initial SNR, and the area under the memory curve.
@@
 To maximize this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
-An example of such a model is shown in \autoref{fig:max_init}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
+An example of such a model is shown in \Fref{fig:max_init}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
 These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \autoref{fig:max_init}(\ref{fig:binary_det}).
+They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \Fref{fig:max_init}(\ref{fig:binary_det}).
 
@@
 With the states in the order described above, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
-Some of these perturbations are shown in \autoref{fig:perts}, see supplementary material for details.
-For example, in \autoref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
+Some of these perturbations are shown in \Fref{fig:perts}, see supplementary material for details.
+For example, in \Fref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
 $\smash{\eta^+_i > \eta^+_j}$), we have proven that increasing $\smash{\M\pot_{ij}}$ and decreasing $\smash{\M\dep_{ij}}$ leads to an increase in area.
@@
 This determines the topology (non-zero transition probabilities) of the model with maximal area.
-It is of the form shown in \autoref{fig:env}(\ref{fig:sticky}),%\autoref{fig:max_area},
+It is of the form shown in \Fref{fig:env}(\ref{fig:sticky}),%\Fref{fig:max_area},
 with potentiation moving one step to the right and depression moving one step to the left.
-Any other topology would allow some class of perturbations (\eg\ in \autoref{fig:perts}) to further increase the area.
+Any other topology would allow some class of perturbations (\eg\ in \Fref{fig:perts}) to further increase the area.
 
@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.07\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
-  \item\aligntop{\includegraphics[width=0.07\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
-  \item\aligntop{\includegraphics[width=0.14\linewidth]{shortcut.svg}}\label{fig:shortcut}
+  \item\aligntop{\includegraphics[width=0.14\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
+  \item\aligntop{\includegraphics[width=0.14\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
+  \item\aligntop{\includegraphics[width=0.28\linewidth]{shortcut.svg}}\label{fig:shortcut}
  \end{myenuma}
@@
 We can then maximize this by pushing all of the equilibrium distribution symmetrically to the two end states.
-This can be done by reducing the transition probabilities out of these states, as in \autoref{fig:env}(\ref{fig:sticky}). %\autoref{fig:max_area}.
+This can be done by reducing the transition probabilities out of these states, as in \Fref{fig:env}(\ref{fig:sticky}). %\Fref{fig:max_area}.
 This makes it very difficult to exit these states once they have been entered.
@@
 At late times, only the area bound is saturated.
-The function $\snr(t_0)$, the green curve in \autoref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim \smash{t_0\inv}$.
+The function $\snr(t_0)$, the green curve in \Fref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim \smash{t_0\inv}$.
 No synaptic model can have an SNR that is greater than this at any time.
@@
 At any time, can we find a model whose memory curve touches the envelope?
-The red curves in \autoref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
+The red curves in \Fref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
 
@@
 We believe there is some other constraint that prevents models from approaching the envelope, and currently are exploring several mathematical conjectures for the precise form of this constraint in order to obtain a potentially tighter envelope.
-Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \autoref{fig:env}(\ref{fig:sticky})%\autoref{fig:max_area}
+Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \Fref{fig:env}(\ref{fig:sticky})%\Fref{fig:max_area}
  can come close to the envelope.
@@
  \begin{center}
+ \parbox[t]{0.44\linewidth}{
  \begin{myenuma}
- \parbox[t]{0.44\linewidth}{
   \item\aligntop{\includegraphics[width=0.95\linewidth]{env.svg}}\label{fig:envplot}
+ \end{myenuma}
   }\hp
   \parbox[t]{0.45\linewidth}{
+ \begin{myenuma}
+ \addtocounter{myenumai}{1}
   \item\aligntop{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}\\[0.1\linewidth]
   \item\aligntop{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
+ \end{myenuma}
   }
- \end{myenuma}
  \end{center}
@@
 
-\subsubsection*{Acknowledgements}
+\begin{acknowledgments}
 We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
 We thank Larry Abbott, Marcus Benna, Stefano Fusi, Jascha Sohl-Dickstein and David Sussillo for useful discussions.
+\end{acknowledgments}
 
@@
 
-\bibliographystyle{utcaps_sl}
+\bibliographystyle{pnas2011}
 \bibliography{maths,neuro}
 
+\end{article}
+
 \end{document}

### 2016-03-10 Paper/SynapsePaper.tex
some ref fixing

@@
 %\usepackage{graphicx}
-\input{sl_graphics_preamble.tex}
+\usepackage{ifpdf}
+%\usepackage[hypcap]{caption}%links to figures
+\usepackage{graphicx}
+%\usepackage{graphics} for finding documentation only
+%
+\usepackage{epstopdf}
+\epstopdfsetup{update,suffix=-generated}
+%
+%\newcommand{\inkscapedir}{C:/Program Files/Inkscape/}
+%\IfFileExists{"C:/Program Files (x86)/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Program Files (x86)/Inkscape/}}{}
+\IfFileExists{"C:/Program Files/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Program Files/Inkscape/}}{}
+\IfFileExists{"/Applications/Inkscape.app/Contents/Resources/bin/inkscape"}{\newcommand{\inkscapedir}{/Applications/Inkscape.app/Contents/Resources/bin/}}{}
+%
+\ifpdf
+ \epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{%
+ "\inkscapedir inkscape" -f #1 -C -A \OutputFile
+ }
+\else
+ \DeclareGraphicsRule{.png}{eps}{.bb}{}
+\fi
+%
+\usepackage{adjustbox}
+\newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
+\newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
+%
+%\usepackage{paralist}
+%\newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
+%\newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
+%\newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
+%\newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
+%\newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
+%%for nested lists, \ref only produces innermost item:  (hacky)
+%\makeatletter % for internal macros with @
+%\renewcommand\p@enumii{}
+%\makeatother
+%
+\usepackage[inline]{enumitem}
+\newlist{myenuma}{enumerate*}{4}
+\setlist[myenuma]{label=(\alph*),ref=\alph*}
+%\makeatletter
+%\DeclareRobustCommand{\ilabel}[1]{%
+%   \edef\@currentlabel{\themyenumai}%
+%   \label{#1}}
+%\makeatother
 \graphicspath{{"Figs/"}}
 %\usepackage{pnastwoF}
-\usepackage{amssymb,amsfonts}
+%\usepackage{amssymb,amsfonts}
 %\usepackage[centertags]{amsmath}
@@
 %% OPTIONAL MACRO DEFINITIONS
-\input{sl_definitions.tex}
+\newcommand{\ra}{\rightarrow}
+\newcommand{\lr}{\leftrightarrow}
+\newcommand{\cdt}{\!\cdot\!}
+\newcommand{\vp}{\vspace{0.5cm}}
+\newcommand{\hp}{\hspace{0.5cm}}
+%
+%e.g., i.e. with normal spaces
+\newcommand{\eg}{e.g.\ }
+\newcommand{\ie}{i.e.\ }
+\newcommand{\cf}{cf.\ }
+\newcommand{\etc}{etc.\ }
+\newcommand{\wrt}{wrt.\ }
+\newcommand{\Eg}{E.g.\ }
+\newcommand{\Ie}{I.e.\ }
+\newcommand{\Cf}{Cf.\ }
+\newcommand{\Etc}{Etc.\ }
+%
+% indices
+%
+% brackets etc.
+\newcommand{\prn}[1]{\left ( #1 \right )}
+\newcommand{\brc}[1]{\left\{ #1 \right\}}
+\newcommand{\brk}[1]{\left [ #1 \right ]}
+\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}
+\newcommand{\nrm}[1]{\left\lVert #1 \right\rVert}
+\newcommand{\av}[1]{\left\langle #1 \right\rangle}
+\newcommand{\avc}[2]{\left\langle #1 \,\middle | \, #2 \right\rangle}
+%\newenvironment{cases}{\left\{\begin{aligned}}{\end{aligned}\right.}
+%
+% Sets
+\newcommand{\set}[2]{\left\{ #1 \,\middle |\, #2 \right\}}
+\newcommand{\cond}[2]{\!\left( #1 \,\middle | \, #2 \right)}
+\newcommand{\condb}[2]{\!\left[ #1 \,\middle | \, #2 \right]}
+%
+% Derivatives, etc. First argument is optional.
+\newcommand{\diff}[3][\rule{0mm}{0mm}]{\frac{\mathrm{d}^{#1} #2}{\mathrm{d}{#3}^{#1}}}
+\newcommand{\pdiff}[3][\rule{0mm}{0mm}]{\frac{\partial^{#1} #2}{\partial {#3}^{#1}}}
+\newcommand{\pdiffc}[3][\rule{0mm}{0mm}]{\left (\frac{\partial #2}{\partial {#3}}\right )_{\!\!#1}}
+\newcommand{\intd}[2][\rule{0mm}{0mm}]{\int #1\!\!\dr #2\,}
+%
+% Un-italicised letters
+\newcommand{\dr}{\mathrm{d}}
+\newcommand{\e}{\mathrm{e}}
+\newcommand{\ir}{\mathrm{i}}
+\DeclareMathOperator{\tr}{tr}
+\DeclareMathOperator{\Tr}{Tr}
+\DeclareMathOperator{\Det}{Det}
+\DeclareMathOperator{\cov}{Cov}
+\DeclareMathOperator{\var}{Var}
+\DeclareMathOperator{\bias}{Bias}
+\DeclareMathOperator{\rank}{rank}
+\DeclareMathOperator{\sgn}{sgn}
+\DeclareMathOperator{\sech}{sech}
+\DeclareMathOperator{\diag}{diag}
+%
+% The default \Im and \Re look crap
+\renewcommand{\Im}{\operatorname{\mathfrak{Im}}}
+\renewcommand{\Re}{\operatorname{\mathfrak{Re}}}
+%
+% logic
+\newcommand{\means}{\Longleftrightarrow}
+\newcommand{\requires}{\Longleftarrow}
+%
+% Referencing sections, figures, etc
+\newcommand{\sref}[1]{\S\ref{#1}}
+\newcommand{\fref}[1]{fig.\ref{#1}}
+\newcommand{\Fref}[1]{Fig.~\ref{#1}}
+\newcommand{\tref}[1]{tab.\ref{#1}}
+\newcommand{\Tref}[1]{Tab.\ref{#1}}
+%
 %\input{sl_symbols.tex}
@@
 
-\author{Subhaneil Lahiri\affil{1}{Department of Applied Physics, Stanford University, Stanford CA},
+\author{Subhaneil Lahiri\affil{1}{Department of Applied Physics, Stanford University, Stanford CA}
 \and
@@
 
-\abbreviations{LTP, long term potentiation; 
+\abbreviations{SNR, signal-to-noise ratio; LTP, long term potentiation; 
 LTD, long term depression}
@@
 %
-\begin{equation}\label{eq:evol}
-  \diff{\pr(t)}{t} = r\pr(t)\frg,
-  \qquad \text{where} \quad
+\begin{multline}\label{eq:evol}
+  \diff{\prob(t)}{t} = r\prob(t)\frg,
+  \\ \text{where} \quad
   \frg = f\pot\M\pot + f\dep\M\dep - \I.
-\end{equation}
+\end{multline}
 %
@@
 \begin{equation}\label{eq:SNRcalc}
-  \snr(t) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \e^{rt\frg} \w.
+  \snr(t) = \sqrt{N}(2 f\pot f\dep) \eq \brk{\M\pot-\M\dep} \e^{rt\frg} \w.
 \end{equation}
@@
   (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined \hyperref[sec:setup]{below}) as subsequent memories are stored.
-  }\label{fig:models}
+  \label{fig:models}}
 \end{figure}
@@
 \begin{equation}\label{eq:constr}
+\begin{gathered}
 \begin{aligned}
-  \M\potdep_{ij} &\in [0,1], &\quad
-  f\potdep &\in [0,1], &\quad
-  \eq\frg &= 0, &\quad
-  \w_i &= \pm 1, \\
-  \sum_j \M\potdep_{ij} &= 1, &
-  f\pot + f\dep &= 1, &
-  \sum_i \eq_i &= 1.
+  \M\potdep_{ij} &\in [0,1], &\qquad
+  \sum_j \M\potdep_{ij} &= 1, \\
+  f\potdep &\in [0,1],&
+  f\pot + f\dep &= 1, 
 \end{aligned}
+\\
+  \eq\frg = 0, \qquad
+  \sum_i \eq_i = 1, \qquad
+  \w_i = \pm 1.
+\end{gathered}
 \end{equation}
@@
 \begin{equation}\label{eq:eigendecomp}
-  \frg = \sum_a -q_a \uv^a \vv^a,
-  \quad
-  \vv^a \uv^b = \delta_{ab},
-  \quad
-  \frg \uv^a = -q_a \uv^a,
-  \quad
-  \vv^a \frg = -q_a \vv^a.
+\begin{aligned}
+  \frg &= \sum_a -q_a \uv^a \vv^a,
+  &\quad
+  \vv^a \uv^b &= \delta_{ab},
+  \\%\quad
+  \frg \uv^a &= -q_a \uv^a,
+  &\quad
+  \vv^a \frg &= -q_a \vv^a.
+\end{aligned}
 \end{equation}
@@
 Instead there are tradeoffs between storing proximal and distal memories;  often attempts to increase memory at late (early) times by changing $\M\potdep$, incurs a performance loss in memory at early (late) times in specific models considered so far \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial}.
-Thus our end goal, achieved in \sref{sec:env}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the {\it entire} memory curve of {\it any} model.
+Thus our end goal, achieved \hyperref[sec:env]{below}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the {\it entire} memory curve of {\it any} model.
 In order to achieve this goal, in this section, we must first derive upper bounds, over the space of all possible synaptic models, on two different scalar functions of the memory curve: its initial SNR, and the area under the memory curve.
@@
  \begin{myenuma}
-  \item\hp\aligntop{\includegraphics[width=0.12\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
-  \item\aligntop{\includegraphics[width=0.12\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
-  \item\hp\aligntop{\includegraphics[width=0.05\linewidth]{binary_det.svg}}\label{fig:binary_det}
+  \item\hp\aligntop{\includegraphics[width=0.24\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
+  \item\aligntop{\includegraphics[width=0.24\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
+  \item\hp\aligntop{\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
   \end{myenuma}
@@
   (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
-  }\label{fig:max_init}
+  \label{fig:max_init}}
 \end{figure}
@@
 \begin{aligned}
-  A &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \prn{\M\pot_{ij}-\M\dep_{ij}} \prn{\eta^+_i-\eta^+_j}\\
-    &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \prn{\M\pot_{ij}-\M\dep_{ij}} \prn{\eta^-_j-\eta^-_i}.
+  A &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \!\brk{\M\pot_{ij}-\M\dep_{ij}}\! (\eta^+_i-\eta^+_j)\\
+    &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \!\brk{\M\pot_{ij}-\M\dep_{ij}}\! (\eta^-_j-\eta^-_i).
 \end{aligned}
@@
 It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
-  }\label{fig:perts}
+  \label{fig:perts}}
 \end{figure}
@@
 \begin{equation}\label{eq:multistatearea}
-  A \leq \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k
-    = \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
+\begin{aligned}
+  A &\leq \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k \\
+    &= \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
+\end{aligned}
 \end{equation}
@@
   t_0 &\leq \frac{M-1}{r} &
-   \implies
-  \snr(t) &= \sqrt{N}\e^{-rt/(M-1)} &
-   \implies
+   & \implies &
+  \snr(t) &= \sqrt{N}\e^{-rt/(M-1)} \\&&
+   & \implies &
   \snr(t_0) &= \sqrt{N}\e^{-rt_0/(M-1)},\\
   t_0 &\geq \frac{M-1}{r} &
-   \implies
-  \snr(t) &= \frac{\sqrt{N}(M-1)\e^{-t/t_0}}{rt_0} &
-   \implies
+   & \implies &
+  \snr(t) &= \frac{\sqrt{N}(M-1)\e^{-t/t_0}}{rt_0} \\&&
+   & \implies &
   \snr(t_0) &= \frac{\sqrt{N}(M-1)}{\e rt_0}.
@@
  \begin{center}
+ \parbox[t]{0.43\linewidth}{
+ \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.85\linewidth]{env.svg}}\label{fig:envplot}
+ \end{myenuma}
+  }\hp
   \parbox[t]{0.44\linewidth}{
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.95\linewidth]{env.svg}}\label{fig:envplot}
- \end{myenuma}
-  }\hp
-  \parbox[t]{0.45\linewidth}{
- \begin{myenuma}
- \addtocounter{myenumai}{1}
+ \stepcounter{myenumai}
+% \addtocounter{enumi}{1}
   \item\aligntop{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}\\[0.1\linewidth]
@@
   The model shown in (\ref{fig:sticky}) also saturates the area bound \eqref{eq:max_area} in the limit $\varepsilon\to0$.
-  }\label{fig:env}
+  \label{fig:env}}
 \end{figure}
@@
 In doing so, we have obtained several new mathematical results delineating the functional limits of memory achievable by synaptic complexity, and the structural characterization of synaptic dynamical systems that achieve these limits.
-In particular, operating within the ideal observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}, we have shown that for a population of $N$ synapses with $M$ internal states, (a) the initial SNR of any synaptic model cannot exceed $\sqrt{N}$, and any model that achieves this bound is equivalent to a binary synapse, (b) the area under the memory curve of any model cannot exceed that of a linear chain model with the same equilibrium distribution, (c) both the area and memory lifetime of any model cannot exceed $\CO(\sqrt{N}M)$, and the model that achieves this limit has a linear chain topology with only nearest neighbor transitions, (d) we have derived an envelope memory curve in the SNR-time plane that cannot be exceeded by the memory curve of any model, and models that approach this envelope for times greater than $\CO(\frac{M}{r})$ are linear chain models, and (e) this late-time envelope is a power-law proportional to $\smash{\CO({\sqrt{N} M}/{rt})}$, indicating that synaptic complexity can strongly enhance the limits of achievable memory.
+In particular, operating within the ideal observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}, we have shown that for a population of $N$ synapses with $M$ internal states, \begin{myenuma}
+\item the initial SNR of any synaptic model cannot exceed $\sqrt{N}$, and any model that achieves this bound is equivalent to a binary synapse,
+\item the area under the memory curve of any model cannot exceed that of a linear chain model with the same equilibrium distribution,
+\item both the area and memory lifetime of any model cannot exceed $\CO(\sqrt{N}M)$, and the model that achieves this limit has a linear chain topology with only nearest neighbor transitions,
+\item we have derived an envelope memory curve in the SNR-time plane that cannot be exceeded by the memory curve of any model, and models that approach this envelope for times greater than $\CO(\frac{M}{r})$ are linear chain models, and
+\item this late-time envelope is a power-law proportional to $\smash{\CO({\sqrt{N} M}/{rt})}$, indicating that synaptic complexity can strongly enhance the limits of achievable memory.
+\end{myenuma}
 
@@
 \bibliographystyle{pnas2011}
-\bibliography{maths,neuro}
+\bibliography{maths,neuro,markov}
 

### 2016-03-11 Paper/SynapsePaper.tex
thrown in supp

@@
 \newcommand{\CO}{\mathcal{O}}
+\newcommand{\CL}{\mathcal{L}}
 \newcommand{\dt}{\dr t}
@@
 
+% ----------------------------------------------------------------
+\input{sl_theorems_preamble.tex}
+% ----------------------------------------------------------------
 %\renewcommand\Fref[1]{}
 \providecommand\hyperref[2][1]{#2}
+\providecommand\texorpdfstring[2]{#1}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\begin{center}
+  \textbf{\Large Supplementary material}
+\end{center}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Here we provide more details underlying the derivations of results in the main paper.
+
+\section{Continuous time Markov processes}\label{sec:ContMarkov}
+
+In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time to define notation.
+It is a generalization of material that can be found in \cite{kemeny1960finite} with some ideas from \cite{hunter2000survey}.
+
+\subsection{Notation}\label{sec:not}
+
+For any matrix $\mathbf{A}$, we define matrices $\mathbf{A}\dg$ and $\overline{\mathbf{A}}$ as
+%
+\begin{equation}\label{eq:dgdef}
+  \mathbf{A}\dg_{ij} \equiv \delta_{ij}\mathbf{A}_{ij},
+  \qquad
+  \overline{\mathbf{A}} \equiv \mathbf{A}-\mathbf{A}\dg.
+\end{equation}
+%
+We let $\onev$ denote a column-vector of ones and $\onem=\onev\onev\trans$ denote a matrix of ones.
+
+A continuous time Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$ with row sums equal to zero ($\MM\onev=0$).
+The probabilities of being in each state at time $t$, the row-vector $\mathbf{p}(t)$, evolve according to
+%
+\begin{equation}\label{eq:rowsum}
+  \diff{\mathbf{p}(t)}{t} = \mathbf{p}(t) \MM,
+\end{equation}
+%
+where $\mathbf{p}(t)\onev=1$.
+
+The equilibrium probabilities, $\eq$, satisfy
+%
+\begin{equation}\label{eq:equilibrium}
+  \eq\MM=0,
+  \qquad
+  \eq\onev=1.
+\end{equation}
+%
+As we assume an ergodic process, this eigenvalue is non-degenerate.
+If all other eigenvalues have strictly negative real parts, the process is regular (aperiodic).
+
+We define additional matrices
+%
+\begin{equation}\label{eq:defDLP}
+  \Lb \equiv (-\MM\dg)\inv,
+  \qquad
+  \Pb \equiv \I + \Lb\MM.
+\end{equation}
+%
+It can be shown that $\Lb_{ii}$ is the mean time it takes to leave state $i$ and $\Pb_{ij}$ is the probability the the next transition from state $i$ goes to state $j$:
+%
+\begin{equation}\label{eq:LamdaPcmpt}
+  \Lb_{ii} = \frac{1}{\sum_{j \neq j} \MM_{ij}},
+  \qquad
+  \Pb_{ij} = %\left\{
+%  \begin{aligned}
+%     &0                                         && \text{if }i=j, \\
+%     &\frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &&\text{otherwise}.
+%  \end{aligned}
+%  \right.
+  \begin{cases}
+     0                                         &\text{if }i=j, \\
+     \frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &\text{otherwise}.
+  \end{cases}
+\end{equation}
+%
+Furthermore, we also define
+%
+\begin{equation}\label{eq:pdotD}
+  \D \equiv \operatorname{diag}(\eq)\inv,
+  \qquad\implies\qquad
+  \eq\D=\onev\trans.
+\end{equation}
+%
+
+\subsection{Fundamental matrix}\label{sec:fund}
+
+For our results below regarding the integral of the memory curve, it can be useful to invert the stochastic transition matrix, $\MM$.
+However, since $\MM$ has a zero eigenvalue, it cannot be inverted.
+For this reason, the fundamental matrix arises as a useful surrogate for the inverse of $\MM$.
+It is related to the first passage times, as we will see in the \hyperref[sec:fpt]{next subsection}.
+Here we define the fundamental matrix and review its properties.
+
+\begin{defn}[Fundamental matrix]
+  For discrete time, the generalized fundamental matrix was defined in \cite{Kemeny1981fund}.
+  For continuous time, we define:
+  %
+  \begin{equation}\label{eq:funddef}
+    \fund \equiv (-\MM + \onev\pib)\inv,
+  \end{equation}
+  %
+  where $\pib$ is any row-vector with $\pib\onev=1/\tau\neq0$.
+\end{defn}
+Note that the canonical choice for the discrete time version, $\pib=\eq$, is not available here due to problems with units.
+It will be helpful to choose $\pib$ to be independent of $\MM$, \eg $\pib=\onev\trans/(n\tau)$.
+All quantities that we calculate using $\fund$ below will be independent of this choice.
+
+\begin{thm}
+  The definition of $\fund$ is valid, \ie $(-\MM + \onev\pib)$ is invertible.
+\end{thm}
+\begin{proof}
+  Assume there exists an $\mathbf{x}$ such that
+  %
+  \begin{equation}\label{eq:fundinvkern}
+    (-\MM + \onev\pib)\mathbf{x}=0.
+  \end{equation}
+  %
+  Multiplying from the left with $\eq$ gives
+  %
+  \begin{equation}\label{eq:fundinvpix}
+    \pib \mathbf{x} = 0.
+  \end{equation}
+  %
+  Substituting back into \eqref{eq:fundinvkern} gives
+  %
+  \begin{equation*}
+    \MM\mathbf{x}=0.
+  \end{equation*}
+  %
+  As we assume an ergodic process, the zero eigenvalue is non-degenerate.
+  Therefore, $\mathbf{x}=\lambda\onev$.
+  Substituting this into \eqref{eq:fundinvpix} gives
+  %
+  \begin{equation*}
+    \lambda\pib\onev = \frac{\lambda}{\tau} = 0.
+  \end{equation*}
+  %
+  As we defined $\pib$ such that $1/\tau\neq0$, this means $\lambda=0 \implies \mathbf{x}=0$.
+\end{proof}
+
+\begin{cor}
+  %
+  \begin{align}
+    \pib\fund &= \eq, \label{eq:fundprob}\\
+    \fund\onev &= \tau\onev,\label{eq:fundrowsum}\\
+    \I+\MM\fund &= \onev\eq, \label{eq:fundQZ}\\
+    \I+\fund\MM &= \tau\onev\pib. \label{eq:fundZQ}
+  \end{align}
+  %
+\end{cor}
+\begin{proof}
+  We can deduce \eqref{eq:fundprob} and \eqref{eq:fundrowsum} be pre/post-multiplying the following equations by $\fund$:
+  %
+  \begin{equation*}
+    \begin{aligned}
+      \eq(-\MM + \onev\pib) &= \pib, \\
+      (-\MM + \onev\pib)\onev &= \frac{\onev}{\tau}.
+    \end{aligned}
+  \end{equation*}
+  %
+  We can then deduce \eqref{eq:fundQZ} and \eqref{eq:fundZQ} by substituting these into
+  %
+  \begin{equation*}
+    (-\MM + \onev\pib)\fund = \fund(-\MM + \onev\pib) = \I.
+  \end{equation*}
+  %
+\end{proof}
+
+%We will not need to use the concept of a g-inverse anywhere, so you can skip ahead to \sref{sec:fpt} now. For the sake of completeness, we can define $\MM^\#$ as
+%%
+%\begin{equation}\label{eq:ginvdef}
+%  \MM^\# \equiv \tau\onev\eq - \fund
+%\end{equation}
+%%
+%\begin{thm}
+%  The matrix $\MM^\#$ satisfies the following defining properties of a g-inverse of $\MM$:
+%  %
+%  \begin{align}
+%    \MM^\# \MM \MM^\# &= \MM^\#,     \label{eq:ginvpropo} \\
+%    \MM \MM^\# \MM    &= \MM,        \label{eq:ginvpropi} \\
+%    \MM \MM^\#        &= \MM^\# \MM, \label{eq:ginvcomm}
+%  \end{align}
+%  %
+%  with \eqref{eq:ginvcomm} holding iff $\eq$ and $\pib$ are proportional to each other.
+%\end{thm}
+%\begin{proof}
+%  Using \eqref{eq:rowsum}, \eqref{eq:equilibrium}, \eqref{eq:fundprob} and \eqref{eq:fundrowsum}, one can verify
+%  %
+%  \begin{alignat*}{2}
+%    \MM \MM^\# &= -\MM \fund,
+%    &\qquad
+%    \pib \MM^\# &= 0,
+%    \\
+%    \MM^\# \MM &= -\fund \MM,
+%    &
+%    \MM^\# \onev &= 0.
+%  \end{alignat*}
+%  %
+%  Substituting into \eqref{eq:fundQZ} and \eqref{eq:fundZQ} leads to
+%  %
+%  \begin{equation*}
+%    \I - \MM \MM^\# = \onev \eq,
+%    \qquad
+%    \I - \MM^\# \MM = \tau \onev \pib.
+%  \end{equation*}
+%  %
+%  Pre/post-multiplying these by $\MM$ or $\MM^\#$ produces \eqref{eq:ginvpropo} and \eqref{eq:ginvpropi}. Subtracting produces
+%  %
+%  \begin{equation*}
+%    [\MM,\MM^\#] = \onev(\tau\pib-\eq),
+%  \end{equation*}
+%  %
+%  which vanishes iff $\eq$ and $\pib$ are proportional to each other (due to $\pib\onev=1/\tau$ and $\eq\onev=1$, the constant of proportionality could only be $\tau$).
+%\end{proof}
+
+\subsection{First passage times}\label{sec:fpt}
+
+\begin{defn}[First passage time matrix]
+ We define $\fptb_{ij}$ as the mean time it takes the process to reach state $j$ for the first time, starting from state $i$.
+ We also define $\fpt\dg_{ii}$ as the mean time it takes the process to return to state $i$.
+ As usual, $\fpt=\fptb+\fpt\dg$.
+\end{defn}
+
+This matrix is given by
+%
+\begin{equation}\label{eq:fpt}
+  \fpt=(\onem\fund\dg - \fund + \Lb)\D,
+\end{equation}
+%
+see \cite{Yao1985fpt} for a proof.
+We can separate this into its diagonal and off-diagonal pieces.
+
+The recurrence times are given by
+%
+\begin{equation}\label{eq:recurtime}
+  \fpt\dg = \Lb\D.
+\end{equation}
+%
+or in component form
+%
+\begin{equation*}
+  \eq_i \Lb_{ii}\inv \fpt\dg_{ii} = 1.
+\end{equation*}
+%
+The extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.
+
+The off-diagonal mean first passage times are given by
+%
+\begin{equation}\label{eq:fptfund}
+  \fptb = (\onem\fund\dg - \fund)\D.
+\end{equation}
+%
+or in component form:
+%
+\begin{equation}\label{eq:fptfundcmpt}
+  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
+\end{equation}
+%
+
+\subsection{Mixing time (Kemeny's constant)}\label{sec:mixtime}
+
+\begin{thm} \mbox{}\label{th:kemenyconst}
+  The quantity
+  %
+  \begin{equation}\label{eq:mixdef}
+    \eta \equiv \sum_j \fptb_{ij}\eq_j
+  \end{equation}
+  %
+  is independent of $i$.
+\end{thm}
+\begin{proof}
+  For discrete time, a proof can be found in \cite[Th.4.4.10]{kemeny1960finite}.
+  For continuous time, we use \eqref{eq:fptfund}, \eqref{eq:fundrowsum} and the transpose of \eqref{eq:pdotD}:
+  %
+  \begin{equation*}
+    \begin{aligned}
+      \fptb (\eq)\trans &= (\onem\fund\dg - \fund) \D (\eq)\trans \\
+        &= (\onev\onev\trans\fund\dg - \fund) \onev \\
+        &= (\onev\trans\fund\dg\onev)\onev - \fund\onev \\
+        &= (\tr\fund - \tau)\onev.
+    \end{aligned}
+  \end{equation*}
+  %
+  which proves \eqref{eq:mixdef} with $\eta = \tr\fund - \tau$.
+\end{proof}
+
+Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version, where this would only shift $\eta$ by 1.
+
+\subsection{Sensitivity of equilibrium distribution}\label{sec:sensitivity}
+
+Suppose that the Markov process, defined by $\MM$, depends on some parameter $\alpha$.
+Differentiating \eqref{eq:funddef} gives
+%
+\begin{equation}\label{eq:diffZ}
+  \diff{\fund}{\alpha} = \fund \diff{\MM}{\alpha} \fund.
+\end{equation}
+%
+We can substitute this into the derivative of \eqref{eq:fundprob}:
+%
+\begin{equation}\label{eq:diffp}
+  \diff{\eq}{\alpha} = \pib \fund \diff{\MM}{\alpha} \fund = \eq \diff{\MM}{\alpha} \fund.
+\end{equation}
+%
+We can rewrite this in component form and use the fact that $\MM_{ii} = - \sum_{i\neq j} \MM_{ij}$:
+%
+\begin{equation}\label{eq:diffpT}
+\begin{aligned}
+  \diff{\eq_k}{\alpha} &= \sum_{i,j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} \\
+    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} + \sum_i \eq_i \diff{\MM_{ii}}{\alpha} \fund_{ik} \\
+    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} (\fund_{jk} - \fund_{ik}) \\
+    &= \sum_{i\neq j} \diff{\MM_{ij}}{\alpha} \eq_i \eq_k (\fptb_{ik} - \fptb_{jk}).
+\end{aligned}
+\end{equation}
+%
+This is a generalization of a result of \cite{cho2000markov} from discrete to continuous time that we will need below.
+Note that the summand vanishes for $i=j$, so we can drop the restriction $i\neq j$ from the range of the sum.
+
+\subsection{Subsets and flux}\label{sec:subsets}
+
+Let us denote the set of states by $\CS$.
+Consider a subset $\CA\subset\CS$.
+We can define a projection operator onto this subset:
+%
+\begin{equation}\label{eq:proj}
+  \prn{\I^\CA}_{ij} =
+    \begin{cases}
+      1 &\text{if $i=j\in\CA$,}\\
+      0 &\text{otherwise.}
+    \end{cases}
+\end{equation}
+%
+We will use superscripts/subscripts to denote projection onto/summation over a subset:
+%
+\begin{equation}\label{eq:projsum}
+  \begin{aligned}
+    \pib^\CA &=\pib\I^\CA, &
+    \M^{\cdot\CA}&=\M\I^\CA, &
+    \M^{\CA\cdot}&=\I^\CA\M, &
+    \mathbf{x}^\CA &= \I^\CA\mathbf{x},
+    \\
+    \pib_\CA &=\pib\onev^\CA, &
+    \M_{\cdot\CA}&=\M\onev^\CA, &
+    \M_{\CA\cdot}&=\prn{\onev^\CA}\trans\M, &
+    \mathbf{x}_\CA &= \prn{\onev^\CA}\trans\mathbf{x},
+  \end{aligned}
+\end{equation}
+%
+where $\pib$ is a row vector, $\M$ is a matrix and $\mathbf{x}$ is a column vector.
+
+We can define a flux matrix, a.k.a.\ ergodic flow:
+%
+\begin{equation}\label{eq:flux}
+  \F = \D\inv\MM,
+  \qquad
+  \F_{ij} = \eq_i \MM_{ij}.
+\end{equation}
+%
+This measures the flow of probability between states in the equilibrium distribution.
+Detailed balance, a.k.a.\ reversibility, is equivalent to $\F=\F\trans$.
+
+The flux between two subsets is a particularly useful quantity:
+%
+\begin{equation}\label{eq:subflux}
+  \F_{\CA\CB} = {\eq}^\CA\MM\onev^\CB.
+\end{equation}
+%
+One can show that
+%
+\begin{equation}\label{eq:compflux}
+  \F_{\CA\CA\comp} = \F_{\CA\comp\CA} = -\F_{\CA\CA} = -\F_{\CA\comp\CA\comp}
+\end{equation}
+%
+using $\prn{{\eq}^\CA + {\eq}^{\CA\comp}}\MM=0$ and $\MM\prn{{\onev}^\CA + {\onev}^{\CA\comp}}=0$.
+
+\subsection{Lumpability}\label{sec:lump}
+
+Suppose we have partitioned the states into disjoint subsets, $\brc{\CA_\alpha}$:
+%
+\begin{equation}\label{eq:partition}
+  \bigcup_\alpha \CA_\alpha = \CS,
+  \qquad
+  \CA_\alpha \cap \CA_\beta = \delta_{\alpha\beta}\CA_\alpha.
+\end{equation}
+%
+We will use $\alpha$ instead of $\CA_\alpha$ in superscripts and subscripts in the following.
+The fact that these subsets are disjoint and exhaustive allows us to define the function
+%
+\begin{equation}\label{eq:whichsey}
+  \sigma(i)=\alpha
+  \qquad\means\qquad
+  i\in\CA_\alpha.
+\end{equation}
+%
+
+We can use this partition to define a new stochastic process associated with the original Markov chain.
+At time $t$, if the state of the original process is $i$, the state of the new process is $\sigma(i)$.
+
+One may ask if this new process is a Markov chain.
+The answer is yes, if the original Markov chain has a property called lumpability \wrt the partition
+(see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time):
+%
+\begin{equation}\label{eq:lump}
+  \sigma(i)=\sigma(j)
+  \quad\implies \quad
+  \MM_{i\alpha}=\MM_{j\alpha} \equiv \sum_{k\in\CA_\alpha} \MM_{jk},
+\end{equation}
+%
+\ie the total transition rate from some state to some subset is the same for all starting states within the same subset.
+This common value is the transition rate for the new lumped Markov chain.
+
+This can be rewritten with the aid of two matrices
+%
+\begin{equation}\label{eq:lumpmats}
+  U_{\alpha i} = \frac{\delta_{\alpha\sigma(i)}}{\abs{\CA_\alpha}},
+  \qquad
+  V_{i\alpha} = \delta_{\sigma(i)\alpha}.
+\end{equation}
+%
+Left multiplication by $U$ averages over subsets, right multiplication by $V$ sums over subsets.
+For $U$, we chose the uniform measure in each subset. Any measure would work equally well, \eg one proportional to the equilibrium distribution:
+%
+\begin{equation}\label{eq:altlumpmats}
+  U_{\alpha i} = \frac{{\eq_i}^\alpha}{\eq_\alpha}.
+\end{equation}
+%
+
+One can show that $(UV)_{\alpha\beta}=\delta_{\alpha\beta}$.
+The matrix $VU$ is also interesting.
+It has a block diagonal structure, with each block corresponding to a subset.
+Each block is a discrete-time ergodic Markov matrix (it is an independent trials process with probabilities given by the measure chosen for $U$).
+This means that the right eigenvectors with eigenvalue 1 will be those that are constant in each subset:
+%
+\begin{equation}\label{eq:setconst}
+  VU\mathbf{x}=\mathbf{x}
+  \qquad\means\qquad
+  \mathbf{x} = \sum_\alpha x_\alpha \onev^\alpha.
+\end{equation}
+%
+
+This allows us to write the lumpability condition \eqref{eq:lump}, and the transition matrix for the lumped process compactly:
+%
+\begin{equation}\label{eq:lumpcompact}
+  VU\MM V = \MM V,
+  \qquad
+  \widehat{\MM} = U\MM V.
+\end{equation}
+%
+By induction, one can show that similar relations hold for all powers:
+%
+\begin{equation}\label{eq:lumppower}
+  VU\MM^nV = \MM^nV,
+  \qquad
+  \widehat{\MM}^n = U\MM^n V,
+\end{equation}
+%
+and, via the Taylor series, for the exponential as well:
+%
+\begin{equation}\label{eq:lumpexp}
+  VU \e^{t\MM} V = \e^{t\MM}V,
+  \qquad
+  \e^{t\widehat{\MM}} = U\e^{t\MM} V.
+\end{equation}
+%
+The equilibrium distribution of the lumped process is given by
+%
+\begin{equation}\label{eq:lumpeq}
+  \widehat{\mathbf{p}}^\infty = \eq V.
+\end{equation}
+%
+
+\section{Signal-to-Noise ratio (SNR)}\label{sec:SNR}
+
+In this section we will look at the signal-to-noise curve, and put an upper bound on its initial value.
+We need only consider ergodic Markov chains.
+Transient states would be unoccupied in equilibrium and would not be accessed by the signal creation process, therefore they could be removed from the analysis.
+Absorbing chains are degenerate cases: they have zero initial signal but infinite decay times, so they can only be approached as the limit of a sequence of ergodic chains.
+
+\subsection{Framework}\label{sec:framework}
+
+The individual potentiation/depression events will be described by \emph{discrete}-time Markov chains:
+%
+\begin{equation}\label{eq:MWdef}
+  \M\potdep \equiv \I + \W\potdep,
+  \qquad
+  \M\potdep\onev = \onev,
+  \qquad
+  \M\potdep_{ij} \in [0,1].
+\end{equation}
+%
+The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf\potdep$, where $f\pot +f\dep =1$ are the fraction of plasticity events that are potentiating/depressing respectively.
+This means that the ``forgetting'' process will be described by the \emph{continuous}-time Markov chain:
+%
+\begin{equation}\label{eq:forgetting}
+  \MM = r \frg \equiv r\prn{f\pot\W\pot + f\dep\W\dep}.
+\end{equation}
+%
+We only require that this Markov chain is ergodic.
+The Markov chains described by $\M\potdep $ need not be.
+
+We assume that the probability distribution starts in the equilibrium distribution \eqref{eq:equilibrium}.
+During the initial signal creation, a fraction $f\pot$ will change to $\eq\M\pot$ and a fraction $f\dep$ will change to $\eq\M\dep$.
+After this, probabilities will evolve according to \eqref{eq:rowsum}.
+
+\subsection{SNR curve}\label{sec:SNRcurve}
+
+As discussed in the main text, the signal-to-noise ratio is given by
+%
+\begin{equation}\label{eq:SNRdef}
+  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}}
+     {\sqrt{\var(\wv_\text{ideal}\cdt\wv(\infty))}}.
+\end{equation}
+%
+First, let's look at the denominator, remembering that the states and plasticity events of each synapse are independent and identically distributed:
+%
+\begin{equation}\label{eq:noise}
+\begin{aligned}
+  \var(\wv_\text{ideal}\cdt\wv(\infty))
+    &= \sum_{\alpha\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty) \wv^\beta_\text{ideal}\wv^\beta(\infty)}
+    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2 \\
+    &= \sum_{\alpha} \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}
+    + \sum_{\alpha\neq\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}\av{\smash{\wv^\beta}_\text{ideal}\smash{\wv^\beta}(\infty)}
+    \\&\phantom{= \sum_{\alpha}  \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}}
+    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2  \\
+    &= N\av{1}
+    + N(N-1)\av{\wv^1_\text{ideal}\wv^1(\infty)}^2
+    - N^2\av{\wv^1_\text{ideal}\wv^1(\infty)}^2 \\
+    &= N(1-\av{\wv^1_\text{ideal}\wv^1(\infty)}^2),
+\end{aligned}
+\end{equation}
+%
+where we used $\wv^\alpha=\pm1$.
+
+For the numerator, we can write
+%
+\begin{equation}\label{eq:overlapex}
+  \av{\wv_\text{ideal}\cdt\wv(t)} = \sum_\alpha\av{\wv^\alpha_\text{ideal}\wv^\alpha(t)}
+   = N \av{\wv^1_\text{ideal}\wv^1(t)},
+\end{equation}
+%
+Noting that $\wvi=\pm1$ with probability $f\potdep$,
+%
+\begin{equation}\label{eq:overlapstate}
+\begin{aligned}
+  \av{\wv^1_\text{ideal}\wv^1(t)} &= f\pot \av{\wv^1(t)}_{\text{pot},t=0} - f\dep \av{\wv^1(t)}_{\text{dep},t=0} \\
+   &= f\pot \sum_i P(\text{state}=i,t \mid \text{pot},0) \w_i - f\dep \sum_i P(\text{state}=i,t \mid \text{dep},0) \w_i.
+\end{aligned}
+\end{equation}
+%
+From the previous section,
+%
+\begin{equation}\label{eq:stateprob}
+  P(\text{state}=i,t \mid \text{pot/dep},0) = \brk{ \eq \M\potdep\, \e^{rt\frg} }_i,
+\end{equation}
+%
+which describes the synapses starting in the equilibrium distribution, changing state due to the plasticity event at $t=0$ and subsequent evolution according to \eqref{eq:rowsum} due to plasticity events uncorrelated with $\wvi$.%
+\footnote{Note that expanding the exponential gives
+%
+\begin{equation*}
+  \e^{rt\frg} =  \sum_{n=0}^\infty \frac{(rt)^n\,\e^{-rt}}{n!} \sum_{m=0}^n (f\pot)^m (f\dep)^{n-m}
+  \brk{ \M\pot \M\dep \M\pot \M\pot \ldots + \text{permutations}}.
+\end{equation*}
+%
+Thus, evolving according to \eqref{eq:rowsum} results in averaging over all sequences of plasticity events,
+as we only need linear expectations of $\wv(t)$ in the end.}
+This results in
+%
+\begin{equation}\label{eq:overlapwt}
+\begin{aligned}
+  \av{\wv^1_\text{ideal}\wv^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
+  \av{\wv^1_\text{ideal}\wv^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
+         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
+         &=  (f\pot-f\dep)\, \eq \w \\
+         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
+\end{aligned}
+\end{equation}
+%
+Combining these allows us to write the numerator as
+%
+\begin{equation}\label{eq:signal}
+\begin{aligned}
+  \av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}
+    &= N \eq (f\pot(\M\pot-\I)-f\dep(\M\dep-\I))\, \e^{rt\frg} \w \\
+    &= N \eq (f\pot(\W\pot-\frg)-f\dep(\W\dep-\frg))\, \e^{rt\frg} \w \\
+%    &= N \eq (f\pot(f\dep\W\pot-f\dep\W\dep)-f\dep(f\pot\W\dep-f\pot\W\pot))\,\e^{rt\frg}\w \\
+    &= N (2f\pot f\dep)\eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
+\end{aligned}
+\end{equation}
+%
+where we used $\eq\frg=0$ in going from the first to second lines.
+Combining with \eqref{eq:noise} gives
+%
+\begin{equation}\label{eq:SNRcurveExact}
+  \SNR(t) = \frac{\sqrt{N} (2f\pot f\dep ) \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w}
+                 {\sqrt{1-(f\pot-f\dep)^2(\eq_+-\eq_-)^2}}.
+\end{equation}
+%
+The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
+We can set the denominator to 1 without changing any of our results.
+
+This results in our final formula:
+%
+\begin{equation}\label{eq:SNRcurve}
+  \SNR(t) = \sqrt{N} (2f\pot f\dep )\, \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
+\end{equation}
+%
+The factor of $\eq$ describes the synapses being in the steady-state distribution before the memory is encoded.
+The factor of $(\M\pot-\M\dep)$ comes from the encoding of the memory at $t=0$, with $\wvi$ being $\pm1$ in synapses that are potentiated/depotentiated.
+The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
+
+We can express this in terms of the one parameter family of transition matrices:
+%
+\begin{equation}\label{eq:Walpha}
+  \begin{aligned}
+  \W(\alpha) &= \alpha\W\pot +(1-\alpha)\W\dep ,
+  &\quad&\implies&
+    \frg&=\W(f\pot ),\\&&&&
+    \W\pot  - \W\dep  &= \diff{\W}{\alpha},\\&&&&
+    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \frg.
+  \end{aligned}
+\end{equation}
+%
+Then \eqref{eq:SNRcurve} becomes
+%
+\begin{equation}\label{eq:SNRalpha}
+  \SNR(t) = \sqrt{N} (2f\pot f\dep ) \diff{\eq}{\alpha}(-\frg)\,\e^{rt\frg}\w.
+\end{equation}
+%
+
+\subsection{Lumpability and the SNR curve}\label{sec:SNRlump}
+
+Suppose that we have a partition such that $\W\pot $ and $\W\dep $ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \sref{sec:lump}):
+%
+\begin{equation}\label{eq:lumpablesynapse}
+  VU\W\potdep  V = \W\potdep  V,
+  \qquad
+  VU\w=\w.
+\end{equation}
+%
+We can define a new synapse with
+%
+\begin{equation}\label{eq:lumpedsynapse}
+  \widehat{\W}\potdep  = U\W\potdep  V,
+  \qquad
+  \widehat{\w} = U \w,
+  \qquad
+  \widehat{\mathbf{p}}^\infty = \eq V.
+\end{equation}
+%
+This synapse has an SNR curve:
+%
+\begin{equation}\label{eq:lumpedSNR}
+  \begin{aligned}
+    \frac{\SNR(t)}{\sqrt{N} (2f\pot f\dep )} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}\pot  - \widehat{\W}\dep )\e^{rt\widehat{\W}^F}\widehat{\w}. \\
+      &= \eq VU (\W\pot  - \W\dep ) VU \e^{rt\frg} VU \w. \\
+      &= \eq (\W\pot  - \W\dep ) VU \e^{rt\frg} VU \w. \\
+      &= \eq (\W\pot  - \W\dep )\e^{rt\frg} VU \w. \\
+      &= \eq (\W\pot  - \W\dep )\e^{rt\frg}\w. \\
+  \end{aligned}
+\end{equation}
+%
+\ie the lumped process has exactly the same SNR as the original one.
+
+\subsection{Initial SNR and flux}\label{sec:initflux}
+
+Using $\eq\frg=0$ and the first line of \eqref{eq:signal}, we can write the initial SNR as
+%
+\begin{equation}\label{eq:init}
+  \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f\pot \W\pot -f\dep \W\dep )(\onev^+-\onev^-).
+\end{equation}
+%
+Using $\W\potdep (\onev^++\onev^-)=0$ and \eqref{eq:compflux}:
+%
+\begin{equation*}
+  r{\eq}^-(f\pot \W\pot +f\dep \W\dep )\onev^+ = \F_{-+} = \F_{+-} = r{\eq}^+(f\pot \W\pot +f\dep \W\dep )\onev^-,
+\end{equation*}
+%
+we can rewrite \eqref{eq:init} as
+%
+\begin{equation}\label{eq:initflux}
+  I = \frac{4\F_{-+}}{r} - 4{\eq}^+f\pot \W\pot \onev^- - 4{\eq}^-f\dep \W\dep \onev^+.
+\end{equation}
+%
+The last two terms are guaranteed to be negative, as the diagonal parts of $\W\potdep $ cannot contribute.
+Therefore
+%
+\begin{equation}\label{eq:initfluxineq}
+  \SNR(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}.
+\end{equation}
+%
+This inequality is saturated if potentiation never takes it from a $+$ state to a $-$ state and depression never takes it from a $-$ state to a $+$ state.
+
+\section{Area maximisation}\label{sec:areamax}
+
+In this section we will find an upper bound on the area under the signal-to-noise curve.
+As in \sref{sec:SNR}, we will only consider ergodic Markov chains.
+We will see in \sref{sec:multistate} that the optimal chain is absorbing, so it lies on the boundary of the (open) set of ergodic chains, but it still puts an upper bound on the area for any chain in the interior.
+
+\subsection{Area under signal-to-noise curve}\label{sec:area}
+
+The signal-to-noise curve is given by \eqref{eq:SNRalpha}.
+%%
+%\begin{equation}\label{eq:SNR}
+%  \SNR(t) = \sqrt{N}(2f^+f\dep ) \diff{\eq}{\alpha} (-\frg) \e^{rt\frg} \w.
+%\end{equation}
+%%
+The area is computed by integrating this
+%
+\begin{equation}\label{eq:area}
+\begin{aligned}
+  A &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\frg}}_0^\infty \w \\
+    &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
+    &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} \w.
+\end{aligned}
+\end{equation}
+%
+
+We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f\pot f\dep )\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W\pot _{ij}-\W\dep _{ij}$
+%
+\begin{equation}\label{eq:areaT}
+  \hat{A} = \sum_{i,j,k} \eq_i \enc_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
+\end{equation}
+%
+
+\begin{defn}[Partial mixing times]
+We define the $\pm$ mixing times as
+%
+\begin{equation}\label{eq:mixingpm}
+\begin{aligned}
+  \eta^\pm_i &\equiv \sum_k \fptb_{ik} \eq_k \prn{\frac{1 \pm \w_k}{2}}
+%    &&= \sum_{\set{k}{\w_k=\pm1}} \fptb_{ik} \eq_k \\
+    &&= \sum_{k\in\pm} \fptb_{ik} \eq_k \\
+    &= \sum_k \prn{\fund_{kk}-\fund_{ik}} \prn{\frac{1 \pm \w_k}{2}}
+    &&= \sum_{k\in\pm} \prn{\fund_{kk}-\fund_{ik}} .
+\end{aligned}
+\end{equation}
+%
+We can think of $\eta^+_i$ as a measure of the ``distance'' to the $\w_k=+1$ states and $\eta^-_i$ as the ``distance'' to the $\w_k=-1$ states.
+\end{defn}
+Using \eqref{eq:mixdef}, we can write:
+%
+\begin{equation}\label{eq:mixingrels}
+\begin{aligned}
+  \eta^+_i + \eta^-_i &= \eta,\\
+  2(\eta^+_i - \eta^+_j) &= \sum_k (\fptb_{ik}-\fptb_{jk}) \eq_k \w_k
+    = \sum_k (\fund_{jk}-\fund_{ik}) \w_k.
+\end{aligned}
+\end{equation}
+%
+We could arrange the states in order of decreasing $\eta^+$, which is the same as the order of increasing $\eta^-$.
+
+We can rewrite \eqref{eq:areaT} as
+%
+\begin{equation}\label{eq:areaEta}
+\begin{aligned}
+  \hat{A} &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}) &
+    &= -2\sum_{i,j} \enc_{ij} \eq_i \eta^+_{j} \\
+    &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^-_{j} - \eta^-_{i}) &
+    &= 2\sum_{i,j} \enc_{ij} \eq_i \eta^-_{j}.
+\end{aligned}
+\end{equation}
+%
+We can also express it in terms of the fundamental matrix \eqref{eq:funddef} as
+%
+\begin{equation}\label{eq:areaZ}
+  \hat{A} = \sum_{i,j,k,l} \enc_{ij} \pib_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
+    = \pib \fund q \fund \w.
+\end{equation}
+%
+
+It is also helpful to define the following quantities:
+%
+\begin{equation}\label{eq:areacoeffs}
+  \begin{aligned}
+    c_k &= \diff{\ln\eq_k}{\alpha}
+      = \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
+      = - \prn{\eq \enc \fptb}_k
+      = \frac{(\eq \enc \fund)_k}{\eq_k}, \\
+    a_i &= \sum_j \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
+    \implies
+    \hat{A} &= \sum_k c_k \eq_k \w_k
+      = 2\sum_i a_i.
+  \end{aligned}
+\end{equation}
+%
+Note that the optimal choice of $\w$ is $\w_k = \sgn(c_k)$.
+
+\subsection{Derivatives \wrt \texorpdfstring{$\W\potdep $}{W(pot/dep)}}\label{sec:deriv}
+
+In the following, we will mathematically define the classes of perturbations pictorially described in Figure 3 of the main paper.
+In order to do so, we will need to consider expressions for derivatives of various quantities with respect to $\W\potdep_{ij}$.
+
+As discussed in the main text,
+we will regard the off-diagonal elements of $\W\potdep _{ij}$ to be the independent variables,
+with $\W\potdep _{ii}=-\sum_{j \neq i} \W\potdep _{ij}$ imposed by hand.
+Thus,
+%
+\begin{equation}\label{eq:basicderivs}
+  \pdiff{\frg_{ij}}{\W\potdep _{gh}} = f\potdep  \delta_{gi}(\delta_{hj}-\delta_{ij}),
+  \qquad
+  \pdiff{\enc_{ij}}{\W\potdep _{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
+\end{equation}
+%
+The implicit $g \neq h$ that comes with all derivatives is unnecessary, as the derivatives above vanish when $g=h$.
+
+In particular, differentiating \eqref{eq:funddef},
+%
+\begin{equation}\label{eq:derivZ}
+  \pdiff{\fund_{ij}}{\W\potdep _{gh}} = rf\potdep  \fund_{ig} (\fund_{hj}-\fund_{gj}).
+\end{equation}
+%
+We can then differentiate expression \eqref{eq:areaZ} to get
+%
+\begin{equation}\label{eq:derivA}
+  \begin{aligned}
+    \pdiff{\hat{A}}{\W\potdep _{gh}} =&\,
+%      rf\potdep  \sum_{ijkl} \enc_{ij} \w_k \pib_l \brk{
+%         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
+%    &\pm \sum_{kl} \w_k \pib_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
+%    =&\,
+%      2rf\potdep  \sum_{ij} \enc_{ij} \eq_i \eq_g \brk{
+%         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
+%         (\fptb_{ig}-\fptb_{jg})(\eta^+_{g}-\eta^+_{h}) } \\
+%      &\pm 2 \eq_g (\eta^+_g - \eta^+_h) \\
+%    =&\,
+%      2rf\potdep  \sum_{ij} \enc_{ij} \eq_i \eq_g
+%         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
+%      \pm 2 \eq_g (\eta^+_g - \eta^+_h)\brk{1+\pdiff{\ln(\eq_g)}{\,\ln(f\potdep )}} \\
+%    =&\,
+      2rf\potdep  \eq_g \brk{\sum_i a_i (\fptb_{gi}-\fptb_{hi}) + c_g (\eta^+_{g}-\eta^+_{h})}
+      \pm 2 \eq_g (\eta^+_g - \eta^+_h). \\
+  \end{aligned}
+\end{equation}
+%
+where $a_i$ and $c_k$ were defined in \eqref{eq:areacoeffs}.
+
+It is sometimes useful to consider the following derivatives:
+%
+\begin{equation}\label{eq:derivqw}
+  \begin{aligned}
+    \pdiff{}{\frg_{gh}} &\equiv     \pdiff{}{\W\pot _{gh}} +     \pdiff{}{\W\dep _{gh}} ,&\qquad
+    \pdiff{}{\enc_{gh}}    &\equiv f\dep  \pdiff{}{\W\pot _{gh}} - f\pot  \pdiff{}{\W\dep _{gh}} .
+  \end{aligned}
+\end{equation}
+%
+Each of these derivatives behaves as their names suggest:
+%
+\begin{equation}\label{eq:derivqweff}
+  \pdiff{\frg_{ij}}{\frg_{gh}} = \pdiff{\enc_{ij}}{\enc_{gh}}
+  =  \delta_{gi}(\delta_{hj}-\delta_{ij}),
+  \qquad
+  \pdiff{\enc_{ij}}{\frg_{gh}} = \pdiff{\frg_{ij}}{q_{gh}}  = 0.
+\end{equation}
+%
+This is because we could treat $\frg$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W\potdep $.
+
+\subsubsection{Scaling mode}\label{sec:rescale}
+
+Consider the following differential operator:
+%
+\begin{equation}\label{eq:scaleop}
+  \Delta \equiv \sum_{g,h} \W\pot _{gh}\pdiff{}{\W\pot _{gh}} + \W\dep _{gh}\pdiff{}{\W\dep _{gh}}.
+\end{equation}
+%
+This corresponds to the scaling, $\W\potdep  \to (1+\epsilon)\W\potdep $.
+Intuitively, this has two effects: it scales up the initial potentiation/depression and it scales down all timescales.
+This intuition is confirmed by the following results:
+%
+\begin{equation}\label{eq:scaleeffects}
+  \begin{aligned}
+    \Delta \fund &= \tau\onev\eq - \fund ,&
+    \Delta \eq  &= 0 ,&
+    \Delta \fpt  &= -\fpt ,
+    \\
+    \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
+    \Delta \enc_{ij} &= \enc_{ij} ,&
+    \Delta \hat{A}  &= 0 ,&
+  \end{aligned}
+\end{equation}
+%
+The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W\potdep $.
+
+As the area is invariant under this scaling, we can consider the $\W\potdep $ to be projective coordinates.
+Therefore we don't need to enforce the lower bound on the diagonal matrix elements while looking for the maximum area, as we can use this null-mode to enforce it afterwards without changing the area.
+%We also don't have to worry about the boundary at infinity, as sending some of them to infinity is equivalent to sending the rest to zero.
+
+\subsection{Kuhn-Tucker conditions}\label{sec:kuhntucker}
+
+Consider the Lagrangian
+%
+\begin{equation}\label{eq:lagrangian}
+  \CL = \hat{A} + \sum_{\text{pot/dep}}\sum_{i\neq j} \mu\potdep _{ij} \W\potdep _{ij} .%  +\lambda\eq\w.
+\end{equation}
+%
+%The last term ensures that we hold $\eq_\pm$ fixed during this extremisation, so that we can safely ignore the factors that we dropped from the noise in \eqref{eq:noisepm}.
+Necessary conditions for an extremum are
+%
+\begin{equation}\label{eq:extremum}
+  \pdiff{\CL}{\W\potdep _{gh}} = 0,
+  \qquad
+    \mu\potdep _{gh} \geq 0,\quad
+    \W\potdep _{gh} \geq 0,\quad
+    \mu\potdep _{gh}\W\potdep _{gh} = 0.
+\end{equation}
+%
+with $g \neq h$.
+This enforces the positivity constraints on the off-diagonal elements, but not the diagonals.
+As discussed in \sref{sec:rescale}, that can be enforced after finding the maximum using the null scaling degree of freedom.
+
+\subsubsection{Triangularity}\label{sec:triangular}
+
+Here we describe the perturbations corresponding to Figure 3(a,b) of the main paper.
+
+Consider
+%
+\begin{equation}\label{eq:shiftqderiv}
+  \pdiff{\CL}{\enc_{gh}} =
+  f\dep  \pdiff{\CL}{\W\pot _{gh}} - f\pot  \pdiff{\CL}{\W\dep _{gh}}
+   = (f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh}) + 2\eq_g (\eta^+_g - \eta^+_h)
+   = 0.
+\end{equation}
+%
+This corresponds to the shift
+%
+\begin{equation}\label{eq:shiftq}
+  \W\pot _{ij} \to \W\pot _{ij} + f\dep \epsilon_{ij},
+  \qquad
+  \W\dep _{ij} \to \W\dep _{ij} - f\pot \epsilon_{ij},
+  \qquad
+  \sum_j \epsilon_{ij} = 0,
+\end{equation}
+%
+which leaves $\frg$ unchanged, and therefore $\eq$, $\fpt$ and $\eta^\pm$ as well.
+
+Assume $\eta^+_g > \eta^+_h$. Then
+%
+\begin{equation}\label{eq:lowertriangular}
+ f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh} <0
+ \qquad\implies\qquad
+ \mu\dep _{gh} >0
+ \qquad\implies\qquad
+ \W\dep _{gh}=0.
+\end{equation}
+%
+Similarly, if $\eta^+_g < \eta^+_h$, then
+%
+\begin{equation}\label{eq:uppertriangular}
+ f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh} > 0
+ \qquad\implies\qquad
+ \mu\pot _{gh} >0
+ \qquad\implies\qquad
+ \W\pot _{gh}=0.
+\end{equation}
+%
+Thus, if we arrange the states in order of decreasing $\eta^+$, $\W\pot $ is upper-triangular and $\W\dep $ is lower triangular.
+
+We have ignored the possibility that $\eq_g=0$, as this would imply that $\fpt_{ig}=\infty$, which would in turn imply that the Markov process is not ergodic.
+
+%Note that if we multiply \eqref{eq:shiftqderiv} by $\enc_{gh}$ and sum over $(g,h)$, we get
+%%
+%\begin{equation}\label{eq:stupid}
+%  \hat{A} = \sum_\pm\sum_{g\neq h} \mu\potdep _{gh} \frg_{gh},
+%\end{equation}
+%%
+%which is probably useless
+
+\subsubsection{Shortcuts}\label{sec:shortcuts}
+
+In this subsection we will define perturbations corresponding to Figure 3(c) of the main text.
+
+Now consider the following combinations of derivatives for $m>1$:
+%
+\begin{align}
+%  \begin{aligned}
+\label{eq:shortcutderiv}
+    \widetilde{\Delta}\potdep _{g,m} &\equiv
+      \brk{ \sum_{k=0}^{m-1} \frac{1}{\eq_{g\pm k}} \pdiffc{}{\W\potdep _{g\pm k,g\pm(k+1)}} }
+      - \frac{1}{\eq_g} \pdiffc{}{\W\potdep _{g,g\pm m}}.
+%  \end{aligned}
+\end{align}
+%
+Once again, they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).
+
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators \eqref{eq:diffpT}:
+%
+\begin{equation}\label{eq:shortcutprob}
+  \widetilde{\Delta}\potdep _{g,m} \eq_i = 0,
+\end{equation}
+%
+which makes it possible to integrate the perturbation:
+%
+\begin{equation}\label{eq:shortcutfinite}
+  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon}^{\pm(g,m)},
+  \qquad
+  \begin{aligned}
+    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g,g\pm m}
+      &\!\!\!\!\!\!&= -\epsilon,\\
+    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm(k+1)}
+      &\!\!\!\!\!\!&= \epsilon
+        &\;
+        &\forall\, k \in [0,m-1],\\
+    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm k}
+      &\!\!\!\!\!\!&= -\epsilon
+        &
+        &\forall\, k \in [1,m-1].
+  \end{aligned}
+\end{equation}
+%
+But more interestingly for our purposes:
+%
+\begin{align}
+%  \begin{aligned}
+\label{eq:shortcutarea}
+    \widetilde{\Delta}\potdep _{g,m}\CL &=
+      \brk{ \sum_{k=0}^{m-1} \frac{\mu\potdep _{g\pm k,g\pm(k+1)}}{\eq_{g\pm k}}
+      - \frac{\mu\potdep _{g,g\pm m}}{\eq_g}}
+      + 2rf\potdep  \sum_{k=0}^{m-1} \prn{\eta^+_{g\pm k} - \eta^+_{g\pm(k+1)}} \prn{c_{g\pm k} - c_{g}},
+%  \end{aligned}
+\end{align}
+%
+
+In the \hyperref[sec:areacoeffincr]{section below}, we will show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
+This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
+If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu\potdep _{g,g\pm m}>0$, which in turn implies that $\W\potdep _{g,g\pm m}=0$.
+This would tell us that the process with the maximal area has to have a multi-state topology.
+
+\subsubsection{Increasing \texorpdfstring{$c_k$}{c(k)}}\label{sec:areacoeffincr}
+
+In the \hyperref[sec:shortcuts]{previous subsection} we defined perturbations corresponding to Figure 3(c) of the main text.
+In order to show that those perturbations increase the area, we must now show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
+
+Consider the following combinations of derivatives:
+%
+\begin{align}
+%  \begin{aligned}
+\label{eq:areacoeffincrderiv}
+    \Delta_{gh} &\equiv
+      \frac{1}{\eq_{g}} \pdiffc{}{\frg_{gh}}
+      + \frac{1}{\eq_h} \pdiffc{}{\frg_{hg}}, \\
+%  \end{aligned}
+\end{align}
+%
+Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).
+
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators using \eqref{eq:diffpT}:
+%
+\begin{equation}\label{eq:sareacoeffincrprob}
+  \Delta_{gh} \eq_i = 0,
+\end{equation}
+%
+which makes it possible to integrate the perturbation:
+%
+\begin{equation}\label{eq:areacoeffincrfinite}
+  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon},
+  \qquad
+  \begin{aligned}
+  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
+  \\
+  \boldsymbol{\epsilon} \onev &= 0.
+  \end{aligned}
+\end{equation}
+%
+But more interestingly:
+%
+\begin{align}
+%  \begin{aligned}
+\label{eq:areacoeffincr}
+    \Delta_{gh}\CL &=
+      \frac{\mu\pot _{gh}+\mu\dep _{gh}}{\eq_g} + \frac{\mu\pot _{hg}+\mu\dep _{hg}}{\eq_h}
+      + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
+%  \end{aligned}
+\end{align}
+%
+where $c_k$ were defined in \eqref{eq:areacoeffs}.
+
+Using the non-negativity of the Kuhn-Tucker multipliers, $\mu\potdep _{ij}$, \eqref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
+
+Note that, according to \sref{sec:triangular}, either $\W\pot _{gh}$ or $\W\dep _{gh}$ will be zero at the maximum, therefore we can expect one of $\mu\pot _{gh}+\mu\dep _{gh}$ to be non-zero.
+This would rule out degeneracy of the $c_k$ or $\eta^+_k$.
+Looking at \eqref{eq:shiftqderiv} closely, the only way $\mu\pot _{gh}+\mu\dep _{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
+
+%From the expression for the area in \eqref{eq:areacoeffs}, it will clearly also be optimal to have the $\w_k$ non-decreasing, \ie all the positive $\w_k$ should lie ahead of all the negative $\w_k$.
+
+\subsubsection{Summary}\label{sec:KTsummary}
+
+Using the Kuhn-Tucker formalism, we have shown that,
+with the states arranged in order of non-increasing $\eta^+_i$:
+%
+\begin{itemize}
+  \item There can be no ergodic maximum for which $\W\pot $ contains backwards transitions or $\W\dep $ contains forwards transitions.
+  \item There can be no ergodic maximum with the $c_k$ decreasing.
+  \item The $c_k$ may only be degenerate at an ergodic maximum if the corresponding $\eta^+_k$ are also degenerate.
+  \item If the $c_k$ increase and the $\eta^+_i$ decrease, there can be no ergodic maximum with shortcuts.
+\end{itemize}
+%
+These were shown by finding allowed perturbations that increase the area.
+
+%Suppose we start with some arbitrary ergodic process.
+If $\W\potdep $ contain backwards/forwards transitions or the $c_k$ decrease, then we can increase the area by applying the perturbations \eqref{eq:shiftqderiv} and \eqref{eq:areacoeffincrderiv}.
+%The second perturbation will generate the unwanted backwards/forwards transitions, requiring subsequent application of first perturbation.
+%The first perturbation will change the $c_k$, possibly changing their order, requiring subsequent application of second perturbation.
+%
+%This procedure can end in several ways:
+%%
+%\begin{itemize}
+%  \item Applying \eqref{eq:areacoeffincrderiv} an infinite amount sends two transition rates to infinity, which due to the scaling mode (\sref{sec:rescale}), is equivalent to sending all other transitions to zero.
+%      Subsequent application of \eqref{eq:shiftqderiv} just removes the backwards transition from $\W\pot $ and the forward transition from $\W\dep $.
+%  \item Applying \eqref{eq:areacoeffincrderiv} ends in a process with degenerate $c_k$ or $\eta^+_k$, and the subsequent application of \eqref{eq:shiftqderiv} doesn't change this degeneracy.
+%  \item As above, but the subsequent application of \eqref{eq:shiftqderiv} lifts the degeneracy, resulting in a chain with increasing $c_k$.
+%\end{itemize}
+%%
+%In the third case, we can then apply the perturbation \eqref{eq:shortcutderiv} to remove all the shortcuts.
+%Either this, or the first case, results in a multi-state topology, possibly disconnected or without the full set of states.
+
+This leaves two possibilities for the maximum area Markov chain.
+Either there is no degeneracy and no shortcuts, which implies the Multi-state/serial topology that we'll discuss in \sref{sec:multistate}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \eqref{eq:shortcutarea}).
+
+Degeneracy tends to be very delicate.
+It is usually hard to arrange without some symmetry relating degenerate states.
+Such a symmetry would imply lumpability (see \sref{sec:lump}).
+The lumped chain would not have any shortcuts, as an entire degenerate set cannot be bypassed.
+As this lumped chain has the same area (see \sref{sec:SNRlump}), we would need only consider the multi-state topology.
+
+\subsection{Multi-state/Serial topology}\label{sec:multistate}
+
+The previous results indicate that the area under the memory curve of any model is bounded by the area under the memory curve of a model with the serial/multistate topology having the same equilibrium distribution.
+Here we compute this area, which we will see depends only on this equilibrium distribution.
+
+The multi-state/serial topology is defined by (see \cite{amit1994learning,Fusi2007multistate,Leibold2008serial}):
+%
+\begin{equation}\label{eq:multistatedef}
+  \W\pot _{ij} = q\pot _i \delta_{i+1,j},
+  \qquad
+  \W\dep _{ij} = q\dep _j \delta_{i,j+1}.
+\end{equation}
+%
+Because it has no shortcuts, it saturates various inequalities:
+%
+\begin{equation}\label{eq:multistateineq}
+  \begin{aligned}
+    \fptb_{ik} - \fptb_{jk} &=
+      \begin{cases}
+        \fptb_{ij},  &\text{if}\quad i \leq j \leq k \quad\text{or}\quad i \geq j \geq k,\\
+        -\fptb_{ji}, &\text{if}\quad j \leq i \leq k \quad\text{or}\quad j \geq i \geq k,\\
+      \end{cases} \\
+    r\eq_i \frg_{ij} \prn{\fptb_{ij}+\fptb_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
+  \end{aligned}
+\end{equation}
+%
+and it satisfies detailed balance (a.k.a.\ reversibility a.k.a. $\CL^2_{\eq}$ self-adjointness):
+%
+\begin{equation}\label{eq:multistateprob}
+  f\pot  q\pot _i \eq_i = f\dep  q\dep _i \eq_{i+1},
+\end{equation}
+%
+which means we can always choose the transition rates, $q\potdep _i$, to give any desired equilibrium probabilities, $\eq_i$.
+
+This allows us to calculate the $c_k$'s:
+%
+\begin{equation}\label{eq:areacoeffchain}
+\begin{aligned}
+  c_k =&\, \sum_{i<k} \fpt_{i,i+1} \prn{\eq_{i}q\pot _i+\eq_{i+1}q\dep _i}
+    %\\&
+    - \sum_{i\geq k} \fpt_{i+1,i} \prn{\eq_{i}q\pot _i+\eq_{i+1}q\dep _i}
+  ,\\
+  c_{k+1} - c_k =&\, \prn{\fpt_{k,k+1}+\fpt_{k+1,k}} \prn{\frac{\eq_{k}\frg_{k,k+1}}{f\pot }+\frac{\eq_{k+1}\frg_{k+1,k}}{f\dep }}
+    =%\\=&\,
+    \frac{1}{rf\pot f\dep },\\
+  \sum_k c_k \eq_k =&\, \sum_{ij} \eq_i \enc_{ij} (\eta-\eta) = 0,\\
+%  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f\pot f\dep },
+  \implies c_k =&\, \frac{k - \sum_j j\eq_j}{rf\pot f\dep },
+\end{aligned}
+\end{equation}
+%
+where we used \eqref{eq:multistateineq} to derive the first two equations respectively and Th.\ref{th:kemenyconst} to derive the third. This allows us to write the area as
+%
+\begin{equation}\label{eq:multistatearea}
+ % A = 2\sqrt{N} \sum_k \brk{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \eq_k \w_k.
+  A = \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k
+    = \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
+\end{equation}
+%
+where we used $\w_k=\sgn(c_k)$, as discussed after \eqref{eq:areacoeffs}.
+This reproduces equation (15) of the main paper.
+
+%It will also help to define $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
+%Note that
+%%
+%\begin{equation}\label{eq:allpplusorminus}
+%  \begin{aligned}
+%    \eq_+ &=1
+%     \quad&\implies\quad
+%     \eq_k \w_k &= \eq_k
+%     \quad&\implies\quad
+%     A=0,\\
+%    \eq_- &=1
+%     \quad&\implies\quad
+%     \eq_k \w_k &= -\eq_k
+%     \quad&\implies\quad
+%     A=0,
+%  \end{aligned}
+%\end{equation}
+%%
+%neither of which are optimal.
+%
+%Consider the Lagrangian
+%%
+%\begin{equation}\label{eq:multistatelagrangian}
+%  \CL = \frac{A}{2\sqrt{N}} + \lambda\prn{1-\sum_i \eq_i} + \sum_i \mu_i \eq_i,
+%\end{equation}
+%%
+%where $\lambda$ is a Lagrange multiplier and $\mu_i$ are Kuhn-Tucker multipliers (see \eqref{eq:extremum}). Extremising \wrt $\eq_i$:
+%%
+%\begin{equation}\label{eq:multistateext}
+%%  \pdiff{\CL}{\eq_i} = \brk{\prn{i-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \w_i
+%%    - \prn{i-\frac{n+1}{2}}\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
+%  \pdiff{\CL}{\eq_i} = \brk{i - \sum_j j\eq_j} \w_i
+%    - i\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
+%\end{equation}
+%%
+%Consider an $i$ such that $\w_i=\w_{i+1}$:
+%%
+%\begin{equation}\label{eq:multistatemudiff}
+%  \mu_{i+1} - \mu_i = \prn{\eq_+-\eq_-} - \w_i.
+%\end{equation}
+%%
+%As $\abs{\eq_+-\eq_-}<1$ (unless we want $A=0$, \eqref{eq:allpplusorminus}), this means the $\mu_i$ increase amongst the negative $\w_i$ and decrease amongst the positive $\w_i$.
+%Therefore the only non-zero equilibrium probabilities at optimum are $\eq_1$ and $\eq_n$.
+
+In order to obtain an upper bound on the area under the memory curve of any model, we now maximise the area of the serial model with respect to its equilibrium distribution.
+First let us maximise \eqref{eq:multistatearea} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
+Clearly this will happen when we put all of the probability at the ends: $\eq_1=\eq_-$ and $\eq_n=\eq_+$ are the only non-zero $\eq_k$.
+This gives an area of
+%\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_+\eq_-)}(n-1)/r$, which would have no effect on what follows.}
+
+%
+\begin{equation}\label{eq:multistateextarea}
+  A \leq \frac{\sqrt{N}}{r}(M-1)\prn{4\eq_+\eq_-}.
+\end{equation}
+%
+This is maximised at $\eq_+=\eq_-=\half$:
+%
+\begin{equation}\label{eq:maxarea}
+  A \leq \frac{\sqrt{N}}{r}(M-1).
+\end{equation}
+%
+This yields the area bound of equation (16) of the main text.
+
+Note that the chain that achieves this is not ergodic, the two states at each end are absorbing.
+This is similar to the results found numerically in \cite{Barrett2008discrete} in a slightly different situation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\bibliographystyle{utcaps_sl}
+%\bibliography{maths,neuro}
+
 \end{document}

### 2016-03-21 Notes/laplace.tex
intuition for topology vs. stochasticity

@@
 If the plasticity events occur at Poisson rate $r$, the probability of $n$ events occurring in time $t$ is given by a Poisson distribution with mean $rt$.
-If we marginalise over $n$as well
+If we marginalise over $n$ as well
 %
@@
 First, evolution was faced with stronger constraints than us when designing these synapses as it had to build them out of real molecules.
-Second, the set of priorities faced by evolution would have been longer than just performance in recogniton memory at a single timescale.
+Second, the set of priorities faced by evolution would have been longer than just performance in recognition memory at a single timescale.
 
@@
 
+The transition between the two regimes is set by the maximum number of states, $M$ (we can't set the number of states to be exactly $M$, as bigger models can always mimic smaller models).
+So, if we want to lengthen the timescale at which a synapse is optimal, we start by adding extra states, using the topology mechanism.
+When we run out of states, we switch to the stochasticity mechanism and make the end states sticky.
+The stochasticity mechanism is only used when we no longer have the option of using the topology mechanism.
+It is always better to add more states, if allowed, than to make transitions stochastic.
+
+We can get some insight into why this is by considering a serial model with $M$ states and all transition topologies equal to $q$.
+%
+\begin{equation*}
+\begin{aligned}
+  \snr(0) &\propto q,
+  &\qquad
+  \max_a \tau_a &\propto \frac{1}{q}, \\
+  \snr(0) &\propto \frac{1}{M},
+  &\qquad
+  \max_a \tau_a &\propto M^2. \\
+\end{aligned}
+\end{equation*}
+%
+The initial SNR scales as described because it is related to the equilibrium probability flux between the weak and strong states, and the equilibrium distribution is uniform over $M$ states.
+The lifetime scales as described because $rq$ sets the timescale for transitions, and for diffusion distance ($M$) scales as the square root of time.
+
+If we want to lengthen the lifetime with a fixed hit on initial SNR, we get a larger increase by increasing $M$ than by decreasing $q$.
+If we want to lengthen the lifetime by a fixed amount, we take a smaller hit on initial SNR by increasing $M$ than by decreasing $q$.
+Thus, we always get more bang for our buck using the topology mechanism than the stochasticity mechanism.
+
 %\subsection*{Acknowledgements}

### 2016-06-09 MHCpaper/MHCsupp.tex
updated MHC supp pagination
attempted spectra linitialisation
typos in mhc supp
rejigging MHC supp

@@
 \newcommand{\ttrain}{t_{\text{train}}}
-\newcommand{\lmax}{_{\text{max}}}
-\newcommand{\lmin}{_{\text{min}}}
+%\newcommand{\lmax}{_{\text{max}}}
+%\newcommand{\lmin}{_{\text{min}}}
+\newcommand{\modelfig}[1][a]{Figure 4#1 of the main text}
 % ----------------------------------------------------------------
@@
 
-\setcounter{page}{5}
-\setcounter{figure}{4}
+\setcounter{page}{8}
+\setcounter{figure}{7}
 \maketitle
@@
 
-The independence and identicalness of synapses means that the state of the system can be completely described by the probability distribution of the internal states, the row vector $\pr(t)$.
-
-The evolution of this probability is described by a continuous time Markov matrix matrix, $\frg$:
+The independence and identicalness of synapses means that the state of the system can be completely described by the probability distribution over the internal states, the \emph{row} vector $\pr(t)$ which evolves as
 %
@@
 %
-where $\I$ is the identity matrix.
+where $\frg$ is a continuous time Markov matrix matrix and $\I$ is the identity matrix.
 Eventually, this will settle into the equilibrium distribution:
@@
 \begin{equation}\label{eq:eqprob}
-  \eq\frg=0.
+  \eq\frg = \mathbf{0}.
 \end{equation}
@@
 We will look at several different models,
-the serial model (see \cite{Leibold2008serial} and \autoref{fig:models}\ref{fig:serial_model}) which has only two values for the synaptic weight,
+the serial model (see \cite{Leibold2008serial,Ben-DayanRubin2007sparse} and \autoref{fig:models}\ref{fig:serial_model}) which has only two values for the synaptic weight,
 the two-state model (which can be thought of as a special case of the serial model, see \autoref{fig:models}\ref{fig:binary_model}),
@@
 This is done by setting $q\pot\lmax=q\pot\lmin$.
-The values of these parameters are listed in \autoref{tab:params}.
+The values of these parameters are listed in \autoref{fig:sim_results}\ref{tab:params}.
 
@@
 We will use a larger value for depression in the \KO\ models.
-The values used are listed in \autoref{tab:params}.
+The values used are listed in \autoref{fig:sim_results}\ref{tab:params}.
 
@@
 %
-If we imagine something like an LN model for the Purkinje cell, the output would depend on some linear combination of the synaptic weights (weighted by the activities of the corresponding parallel fibres).
-If we're not keeping track of synaptic identity, the most natural linear combination to use would be an equal sum of them all.
+One can approximate the input to a Purkinje cell as some linear combination of the synaptic weights (weighted by the activities of the corresponding parallel fibres).
+If we are not keeping track of synaptic identity, the most natural linear combination to use would be an equal sum of them all.
 The behavioural output (VOR gain) will be some unknown, non-linear function of the synaptic weights, so the best we can hope for is to reproduce qualitative features of the experiment, such as whether learning is enhanced or impaired by the mutation or pre-training.
 
+As the knockout produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
+We will model this compensation as a simple linear offset, as could be produced by another population of neurons/synapses whose effect cancels with these neurons/synapses.
+
 We will assume $f\dep\wt=f\dep\ko$.
-This is because the effects of the mutation are well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
+This is because the relevant effects of the knockout are well localised to the Purkinje cells, as shown by the rescue data, so the activity of the parallel and climbing fibres should not change very much.
 Therefore the rates of potentiation and depression should not change very much either.
-This implies that the mean synaptic weight in equilibrium differs between the \KO\ models and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
-As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
-We will model this compensation as a simple linear offset, as could be produced by another population of neurons/synapses whose effect cancels with these neurons/synapses.
+%This implies that the mean synaptic weight in equilibrium differs between the \KO\ models and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
 
@@
 
-The features of the actual experiments that we'd like to capture are:
+The features of the experiments shown in \modelfig\ that we'd like to capture are:
+%
 \begin{enumerate}
@@
 \end{enumerate}
+%
 These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
@@
 We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
-This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measured using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
-This should be multiplied by the difference in synaptic weights.
-But, as this quantity is constant within a model and doesn't change when comparing wild-type and \KO\ models or with/without pre-training, it can be ignored.
+This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measured using the transition rates for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
+%This should be multiplied by the difference in synaptic weights.
+%But, as this quantity is constant within a model and doesn't change when comparing wild-type and \KO\ models or with/without pre-training, it can be ignored.
 
@@
 
-Consider the general uniform multistate model.
+Consider the general uniform serial model.
 Then the equilibrium distribution is given by
@@
 \begin{equation}\label{eq:multiflux}
-  \Phi = \eq_{M/2+1}f'{}\dep q\dep - \eq_{M/2}f'{}\pot q\pot = \frac{1-\alpha}{1-\alpha^M}\,\alpha^{M/2-1}\prn{\alpha-\alpha'}f'{}\dep q\dep,
-\end{equation}
-%
-where primed values correspond to the new value of $f\pot$.
+  \Phi = \eq_{M/2+1}f'{}\dep q\dep - \eq_{M/2}f'{}\pot q\pot = \frac{1-\alpha}{1-\alpha^M} \, \alpha^{M/2-1} \prn{\frac{\alpha}{\alpha\inc} - 1} f\inc\pot  q\pot,
+\end{equation}
+%
+where $\alpha\inc$ is $\alpha$, but with $f\potdep$ replaced by $f\potdep\inc$.
+
+To see the effects of the knockout, we can see if this increases or decreases with $q\dep$:
+%
+\begin{equation}\label{eq:serialderq}
+  q\dep \pdiff{\Phi}{q\dep} = 
+    \frac{\alpha^{\frac{M}{2} - 1}}{(1 - \alpha^M)^2} 
+    \brk{(1 - \alpha^M) - \frac{M}{2} (1 - \alpha) (1 + \alpha^M) } 
+    \prn{\frac{\alpha}{\alpha\inc} - 1} f\inc\pot  q\pot.
+\end{equation}
+%
+This quantity can have either sign, depending on $\alpha$ and $M$, so $\Phi$ is not a monotonic function of $q\dep$.
+Therefore, enhancing plasticity in this way can either enhance or impair the initial learning rate, depending on the parameters.
+
+The effect of pre-training is to replace $f\potdep$ with $f\potdep\dec$ without changing $f\potdep\inc$, which increases $f\pot$ and decreases $f\dep$.
+Therefore, we can see the effects of pre-training on the initial learning rate by seeing if it increases or decreases with $f\pot$ (with $f\dep = 1 - f\pot$):
+%
+\begin{equation}\label{eq:serialderf}
+  f\pot f\dep \pdiff{\Phi}{f\pot} =  
+    \frac{\alpha^{\frac{M}{2} - 1}}{(1 - \alpha^M)^2} 
+      \brk{ \prn{\frac{\alpha}{\alpha\inc} - 1} \frac{M}{2} (1 - \alpha) (1 + \alpha^M) 
+      - \prn{\frac{\alpha^2}{\alpha\inc} - 1} (1 - \alpha^M) }  
+      f\inc\pot  q\pot.
+\end{equation}
+%
+This quantity can have either sign, depending on $\alpha$ and $M$, so $\Phi$ is not a monotonic function of $f\pot$.
+Therefore, pre-training can either enhance or impair the initial learning rate, depending on the parameters.
+
+The lack of monotonicity shown above allows this model to reproduce the key qualitative features of the experiments outlined in \autoref{sec:results}.
+This can be illustrated by looking at a special case, where $q\pot = q\dep = q$ for the wild type, $f\dep_0 = \frac{1}{2}$, $f\dep\dec = \frac{1}{2} - \Delta f$ and $f\dep\inc = \frac{1}{2} - \Delta f$.
 
@@
 %
-So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value).
+So, we see that pre-training will speed up learning when $\Delta f$ is small.
 %, as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS} (black curves).
@@
 The first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ and will be more significant for smaller $\beta$ or in a longer chain.
-This exponential decay in the initial distribution is what allows the saturation bias to overwhelm the increases intrinsic plasticity rate and lead to impaired learning with enhanced plasticity.
+This exponential decay in the initial distribution is what allows the saturation bias to overwhelm the increased intrinsic plasticity rate and lead to impaired learning with enhanced plasticity.
 It is important that a learning signal is only generated by the central transition so that the exponential decay can take effect.
@@
   \qquad
-  \pr(t) = \eq + (\pr(t)-\eq)\e^{-\lambda rt},\\
+  \pr(t) = \eq + (\pr(0)-\eq)\e^{-\lambda rt},\\
   \qquad \text{where} \quad
@@
 But it is easier to just substitute $M=2$ into the formulae in \autoref{sec:multistate}.
-In this case, the initial rate of change and the total change encapsulate the whole solution, as there is only a single exponential decay.
+%In this case, the initial rate of change and the total change encapsulate the whole solution, as there is only a single exponential decay.
+
+We find that
+%
+\begin{equation}\label{eq:binaryflux}
+\begin{aligned}
+  \Phi &= \frac{q\pot q\dep \Delta f}{f\pot q\pot + f\dep q\dep}, &\qquad
+  \pdiff{\Phi}{q\dep} &= \frac{f\pot (q\pot)^2 \Delta f}{(f\pot q\pot + f\dep q\dep)^2}, \\ &&
+  \pdiff{\Phi}{f\pot} &= \frac{(f\pot\inc q\pot + f\dep\inc q\dep) q\pot q\dep}{(f\pot q\pot + f\dep q\dep)^2}.
+\end{aligned}
+\end{equation}
+%
+The flux is a monotonically increasing function of $q\dep$ and $f\pot$.
+Therefore, enhancing plasticity in this way or pre-training can only enhance the initial learning rate, unlike the experimental results shown in \modelfig.
+
+The only thing we have to worry about is when the wild-type catches up with the \KO\ models.
+As the total change will be larger for the wild-type than \KO\ models, it must overtake eventually.
+If this happens early enough, the initial period, where the \KO\ models learns faster than wild-type, would not be seen in the experiment.
+However, the timescale for this crossover would be similar to the timescale of the exponentials.
+In the data, it looks like that timescale is longer than the gaps between successive measurements, so we needn't worry about this.
+
+The monotonicity shown above means that this model cannot reproduce the key qualitative features of the experiments outlined in \autoref{sec:results}, at least as for the initial learning rate.
+This can be illustrated by looking at a special case, where $q\pot = q\dep = q$ for the wild type, $f\dep_0 = \frac{1}{2}$, $f\dep\dec = \frac{1}{2} - \Delta f$ and $f\dep\inc = \frac{1}{2} - \Delta f$.
 
@@
 %
-So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
+So, we see that pre-training will speed up learning for all parameter values, unlike what is seen in the experiment.
 This can be seen in
@@
 %
-which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
+which is always larger than \eqref{eq:binWTnopre}, unlike what is seen in the experiments (see \modelfig).
 This is also seen in
@@
 
-We made a number of assumptions in our analysis so far.
-Before we declare this model to be falsified, we should think about what would happen if we relaxed these assumptions and explored a larger parameter space.
-
-First, we assumed that the wild type had symmetric potentiation and depression.
-Relaxing this assumption would correspond modelling the wild-type like the \KO\ mice, but with a different value for $\beta$.
-We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the \KO\ models.
-
-Secondly, we assumed that there would be equal rates of potentiation and depression in the untrained situation.
-However, relaxing this assumption merely changes \eqref{eq:binKOnopre} to
-%
-\begin{equation}\label{eq:binKOnoprediff}
-  \Phi = \frac{{\Delta f}\, q}{f\dep+\beta f\pot} = \frac{{\Delta f}\, q\pot q\dep}{f\dep q\dep+ f\pot q\pot}.
-\end{equation}
-%
-This will not change our conclusions either.
-
-Because for this simple binary synapse model, the ratio of the numbers of potentiated and depressed synapses is equal to the ratio of the potentiating and depressing transition rates.
+For this simple binary synapse model, the ratio of the numbers of potentiated and depressed synapses is equal to the ratio of the potentiating and depressing transition rates.
 This means that the leading effects of enhanced intrinsic plasticity rates and saturation bias cancel each other.
-This is eve true as the number of synapses available for further plasticity approaches zero, which requires that the depressing transition be infinitely stronger than the potentiating transition, compensating for the saturation.
+This is even true as the number of synapses available for further plasticity approaches zero, which requires that the depressing transition be infinitely stronger than the potentiating transition, compensating for the saturation.
 There is a requirement that the total number of synapses, potentiated and depressed, is fixed.
@@
 
-As we did not need the pre-training paradigm to rule out this model, we will not look at removing the symmetry between gain-increase and decrease in detail.
-However, one can see that pre-training will always increase the fraction of synapses available for depression.
+One can see that pre-training will always increase the fraction of synapses available for depression.
 This model has no mechanism that could ever result in this causing impaired learning.
-
-The only thing we have to worry about is when the wild-type catches up with the \KO\ models.
-As the total change will be larger for the wild-type than \KO\ models, it must overtake eventually.
-If this happens early enough, the initial period, where the \KO\ models learns faster than wild-type, would not be seen in the experiment.
-However, the timescale for this crossover would be similar to the timescale of the exponentials.
-In the data, it looks like that timescale is longer than the gaps between successive measurements, so we needn't worry about this.
 
@@
     \Phi &= \sum_{i=1}^{M-1} \eq_{i+1} f'{}\dep q\dep - \eq_i f'{}\pot q\pot
-         &= \sum_{i=1}^{M-1} \eq_i (\alpha-\alpha') f'{}\dep q\dep \\
-         &= \frac{1-\alpha^{M-1}}{1-\alpha^M} \, (\alpha-\alpha') f'{}\dep q\dep.
+         %&= \sum_{i=1}^{M-1} \eq_i (\alpha-\alpha') f'{}\dep q\dep \\
+         &= \frac{1-\alpha^{M-1}}{1-\alpha^M} \, \prn{\frac{\alpha}{\alpha\inc}} f\inc\pot q\pot.
   \end{aligned}
@@
 %
+
+To see the effects of the knockout, we can see if this increases or decreases with $q\dep$:
+%
+\begin{equation}\label{eq:multiderq}
+  q\dep \pdiff{\Phi}{q\dep} = 
+    \frac{\alpha^{M - 1}}{(1 - \alpha^M)^2}
+    \brk{1 - \alpha}
+    \brk{M - \sum_{i=1}^M \alpha^{i-1}} 
+    \prn{\frac{\alpha}{\alpha\inc} - 1} f\inc\pot  q\pot.
+\end{equation}
+%
+If $\alpha < 1$, both quantities in square brackets are positive, as the sum contains $M$ terms that are all at most $1$.
+If $\alpha > 1$, both of them are negative.
+Therefore this quantity is positive, so $\Phi$ is a monotonically increasing function of $q\dep$.
+Therefore, enhancing plasticity in this way can only enhance the initial learning rate, for all values the parameters.
+
+The effect of pre-training is to replace $f\potdep$ with $f\potdep\dec$ without changing $f\potdep\inc$, which increases $f\pot$ and decreases $f\dep$.
+Therefore, we can see the effects of pre-training on the initial learning rate by seeing if it increases or decreases with $f\pot$ (with $f\dep = 1 - f\pot$):
+%
+\begin{equation}\label{eq:multiderf}
+  f\pot f\dep \pdiff{\Phi}{f\pot} =  
+    \frac{\alpha^{M - 2}}{(1 - \alpha^M)^2}
+      \brk{1 - \alpha} 
+      \prn{\brk{M - \sum_{i=1}^M \alpha^{i-1}} 
+      + \frac{\alpha}{\alpha\inc} \brk{\sum_{i=1}^M \alpha^{1-i} - M}}
+      f\inc\pot  q\pot.
+\end{equation}
+%
+This quantity is positive, for similar reasons to \eqref{eq:multiderq}, so $\Phi$ is a monotonically increasing function of $f\pot$.
+Therefore, pre-training can can only enhance the initial learning rate, for all values the parameters.
+
+The monotonicity shown above means that this model cannot reproduce the key qualitative features of the experiments outlined in \autoref{sec:results}, at least as for the initial learning rate.
+This can be illustrated by looking at a special case, where $q\pot = q\dep = q$ for the wild type, $f\dep_0 = \frac{1}{2}$, $f\dep\dec = \frac{1}{2} - \Delta f$ and $f\dep\inc = \frac{1}{2} - \Delta f$.
 
@@
 \begin{equation}\label{eq:multiLinWTnopre}
-  \Phi = 2{\Delta f}\,q\,\frac{M-1}{M},
-\end{equation}
-%
-where, once more, we should note that $\Delta f<0$.
+  \Phi = 2{\Delta f}\,q\,\frac{M-1}{M}.
+\end{equation}
+%
 With pre-training:
@@
 This means that the \KO\ models will initially learn faster than the wild-type.
-However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in
+However the wild-type will eventually catch up, and this can happen very quickly, as seen in
 \autoref{fig:sim_results}\ref{fig:multistate_sim} %\autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}
@@
 which is always greater than 1.
-Therefore, pre-training will always enhance learning, for both \KO\ models and wild-type, as seen in
+Therefore, pre-training will enhance the initial learning rate for all parameter values, for both \KO\ models and wild-type, as seen in
 \autoref{fig:sim_results}\ref{fig:multistate_sim}. %\autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
@@
 Sufficiently long pre-training pushes the synapses far down the cascade, slowing down learning for the wild-type (see \autoref{fig:sim_results}\ref{fig:cascade_sim}, black curves).
-This effect is weaker for the \KO\ models, as the equilibrium distribution is not as heavily concentrated at the end.
+This effect is weaker for the \KO\ models, as the equilibrium distribution is not as heavily concentrated at the end, as the distribution before pre-training is shifted in the other direction.
 %(see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
@@
 So far, we have only considered the comparisons listed at the start of \hyperref[sec:results]{this section}, as they raised important conceptual questions.
-We have not considered the other two comparisons, the lack of difference between wild-type without pre-training and \KO\ with pre-training and between wild-type with pre-training and \KO\ without pre-training (the ``diagonal'' comparisons in the language of the main text), as they depend more on the values used for the parameters than on any conceptual issues.
+We have not considered the other two comparisons, the lack of difference between wild-type without pre-training and \KO\ with pre-training and between wild-type with pre-training and \KO\ without pre-training (the ``diagonal'' comparisons in the language of \modelfig), as they depend more on the values used for the parameters than on any conceptual issues.
 

### 2016-06-10 MHCpaper/MHCsupp.tex
removed unnecessary stuff from supp

@@
 \usepackage{array}
-\newcounter{tablerow}[table]
-\newcommand{\tablerowautorefname}{row}
-\newenvironment{tabularn}[1]{\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
+%\newcounter{tablerow}[table]
+%\newcommand{\tablerowautorefname}{row}
+%\newenvironment{tabularn}[1]{\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
 %\usepackage{etoolbox}
@@
 \newcommand{\modelfig}[1][a]{Figure 4#1 of the main text}
+\newcommand{\datafig}[1][b]{Figure 3#1 of the main text}
 % ----------------------------------------------------------------
@@
 \begin{itemize}
-  \item There are $N$ identical synapses with $M$ internal functional states.
-  \item The states of different synapses are independent of each other.
-  \item The synapses that are eligible for plasticity are chosen randomly.
+  \item There synapses are identical and have $M$ internal functional states, where the states of different synapses are independent of each other.
+%  \item The synapses that are eligible for plasticity are chosen randomly.
   \item Candidate potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $r\potdep$.
-  \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
+  \item Candidate potentiation and depression events are described by Markov processes with transition probabilities $\M\potdep$.
   \item The synaptic weights of the internal states are given by the column vector $\w$.
@@
 
-In practice, the overall rate of candidate plasticity events will only set the timescale, and only the relative rates will matter.
+The overall rate of candidate plasticity events will only set the timescale, and only the relative rates will matter.
 With this in mind, we define
@@
   \qquad
-  f\dep = \frac{r\dep}{r}.
-\end{equation}
-%
-The quantities $f\potdep$ measure the fraction of candidate plasticity events that are potentiating/depressing.
+  f\dep = \frac{r\dep}{r},
+\end{equation}
+%
+where $f\potdep$ is the fraction of candidate plasticity events that are potentiating/depressing.
 
@@
 
-For models with only two possible synaptic weights, the distribution of synaptic weights is completely described by the mean, $\pr(t)\w$.
+%For models with only two possible synaptic weights, the distribution of synaptic weights is completely described by the mean, $\pr(t)\w$.
 
 The \KO\ mice have a lower threshold for depression.
-We can model this by changing $\M\dep\wt$ to $\M\dep\ko$, which should have larger off-diagonal matrix elements.
+We can model this by changing $\M\dep\wt$ to $\M\dep\ko$, which has larger off-diagonal matrix elements.
 
@@
 
-For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
-We will use a larger value for $q$ for depression in the \KO\ models.
-
-For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
-We will use a larger value for $x$ for depression in the \KO\ models.
+For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation in the wild-type and \KO\ models.
+We will use a larger value for $q$ for depression in the \KO\ models than in the wild-type.
+
+For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation in the wild-type and \KO\ models.
+We will use a larger value for $x$ for depression in the \KO\ models than in the wild-type.
 
@@
  \begin{center}
- \parbox{0.8\linewidth}{%
    \begin{myenuma}
-  \item\aligntop{\includegraphics[height=1.7cm]{serial.svg}}\label{fig:serial_model}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\hspace{0.5cm}\\[1cm]
-  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
-  \item\aligntop{\includegraphics[height=2.5cm]{cascade.svg}}\label{fig:cascade_model}
+  \begin{tabular}{l@{\hspace{0.05\linewidth}}l}
+    \item\aligntop{\includegraphics[height=1.7cm]{serial.svg}}\label{fig:serial_model}&
+    \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\\[1.7cm]
+    \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}&
+    \item\aligntop{\includegraphics[height=2.5cm]{cascade.svg}}\label{fig:cascade_model}\\[2.5cm]
+    \item\aligntop{\includegraphics[height=2.3cm]{pooled.svg}}\label{fig:pooled_model}&
+    \item\aligntop{\includegraphics[height=2.3cm]{multistate_nonuni.svg}}\label{fig:nonuni_model}
+  \end{tabular}
    \end{myenuma}
- }
  \end{center}
@@
   (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}) and synaptic weight takes only two values.
+  (\ref{fig:pooled_model}) In the pooled resource model,
+  Several two-state synapses share resources that are required for potentiation and depression.
+  These resources are depleted as more synapses are potentiated or depressed.
+  This pool of synapses can be modelled as one compound synapse.
+  (\ref{fig:nonuni_model}) In the non-uniform multistate model
+  the synaptic weight varies linearly in the interval $[-1,1]$, similar to the multistate model, but the transition probabilities between adjacent states decays exponentially away from the central transition for both potentiation and depression.
   } \label{fig:models}
@@
 
-Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed and replenished when this is reversed.
+Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed and replenished when this is reversed, as shown in \autoref{fig:models}\ref{fig:pooled_model}.
 We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single compound synapse.
@@
 
-The Markov process is lumpable \wrt this partition of states (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$.
-The result must be the same for all states in lump $i$.
+The Markov process is lumpable \wrt this partition of states (see \cite{kemeny1960finite,burke1958markovian,Ball1993Lumpability}).
+The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$,
+which must be the same for all states in lump $i$.
 
@@
 This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
-We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
-We will use larger values for depression in the \KO\ models.
-In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
+We will use the same values for potentiation in the wild-type asnd \KO\ models.
+We will use larger values for depression in the \KO\ models than in the wild-type.
+we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
 This is done by setting $q\pot\lmax=q\pot\lmin$.
@@
 
-\begin{figure}
- \begin{center}
-  \aligntop{\includegraphics[height=4cm]{pooled.svg}}
- \end{center}
-  \caption[The pooled resource model]{The pooled resource model.
-  Several two-state synapses share resources that are required for potentiation and depression.
-  These resources are depleted as more synapses are potentiated or depressed.
-  This pool of synapses can be modelled as one compound synapse.} \label{fig:pooled_model}
-\end{figure}
+%\begin{figure}
+% \begin{center}
+%  \aligntop{\includegraphics[height=4cm]{pooled.svg}}
+% \end{center}
+%  \caption[The pooled resource model]{The pooled resource model.
+%  Several two-state synapses share resources that are required for potentiation and depression.
+%  These resources are depleted as more synapses are potentiated or depressed.
+%  This pool of synapses can be modelled as one compound synapse.} \label{fig:pooled_model}
+%\end{figure}
 
@@
 
-This model is similar to the multistate model (see \autoref{fig:models}\ref{fig:multistate_model})), as it only has transitions between adjacent states and a linearly varying synaptic weight.
+This model, shown in \autoref{fig:models}\ref{fig:nonuni_model}, is similar to the multistate model (see \autoref{fig:models}\ref{fig:multistate_model})), as it only has transitions between adjacent states and a linearly varying synaptic weight.
 However, like the cascade model (see \cite{Fusi2005cascade} and \autoref{fig:models}\ref{fig:cascade_model}), the transition probabilities decay exponentially away from the central transition.
@@
 
-\begin{figure}
- \begin{center}
-  \aligntop{\includegraphics[height=3cm]{multistate_nonuni.svg}}
- \end{center}
-  \caption[The non-uniform multistate model]{The non-uniform multistate model.
-  Similar to the multistate model, (\autoref{fig:models}\ref{fig:multistate_model}), the synaptic weight varies linearly in the interval $[-1,1]$, but the transition probabilities between adjacent states decays exponentially away from the central transition for both potentiation and depression.} \label{fig:nonuni_model}
-\end{figure}
+%\begin{figure}
+% \begin{center}
+%  \aligntop{\includegraphics[height=3cm]{multistate_nonuni.svg}}
+% \end{center}
+%  \caption[The non-uniform multistate model]{The non-uniform multistate model.
+%  Similar to the multistate model, (\autoref{fig:models}\ref{fig:multistate_model}), the synaptic weight varies linearly in the interval $[-1,1]$, but the transition probabilities between adjacent states decays exponentially away from the central transition for both potentiation and depression.} \label{fig:nonuni_model}
+%\end{figure}
 
@@
 It will change the environment, which will lead to a change in how often potentiation and depression occur.
-It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
+%It could be manifested in a change in which synapses are potentiated/depressed, but this could not be captured in this type of model.
 We will model this by changing $f\potdep$, leaving $\M\potdep$ unchanged.
@@
 We use the same values for wild-type and \KO\ models for the reasons discussed above.
-We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
+%We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
 We could adjust $r$ to keep $r\pot$ unchanged, if so desired, but this would only change the overall timescale and would not affect any of the qualitative comparisons that we are concerned with here.
@@
 \begin{myenuma}
-  \begin{tabular}{l@{\hspace{0.05\linewidth}}l@{\hspace{0.05\linewidth}}l}
+  \begin{tabular}{lll}
       \item\label{fig:seial_sim}
-      \aligntop{\includegraphics[width=0.25\linewidth]{serial.svg}}
+      \aligntop{\includegraphics[width=0.24\linewidth]{serial.svg}}
       &
       \item\label{fig:binary_sim}
-      \hspace{0.05\linewidth}\aligntop{\includegraphics[width=0.125\linewidth]{binary.svg}}
+      \hspace{0.05\linewidth}\aligntop{\includegraphics[width=0.12\linewidth]{binary.svg}}
       &
       \item\label{fig:multistate_sim}
-      \aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}
-      \\[1.5cm]
-      \aligntop{\includegraphics[width=0.3\linewidth]{serial_med_learnS.eps}}
+      \aligntop{\includegraphics[width=0.24\linewidth]{multistate.svg}}
+      \\[1.3cm]
+      \aligntop{\includegraphics[width=0.29\linewidth]{serial_fit_learnS.eps}}
       &
-      \aligntop{\includegraphics[width=0.3\linewidth]{binary_learnS.eps}}
+      \aligntop{\includegraphics[width=0.29\linewidth]{binary_learnS.eps}}
       &
-      \aligntop{\includegraphics[width=0.3\linewidth]{multistate_lin_learnS.eps}}
-      \\[5.5cm]
+      \aligntop{\includegraphics[width=0.29\linewidth]{multistate_lin_learnS.eps}}
+      \\[3.5cm]
       \item\label{fig:pooled_sim}
-      \aligntop{\includegraphics[width=0.3\linewidth]{pooled_deponly.svg}}
+      \aligntop{\includegraphics[width=0.29\linewidth]{pooled_deponly.svg}}
       &
       \item\label{fig:cascade_sim}
-      \aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}
+      \aligntop{\includegraphics[width=0.24\linewidth]{cascade.svg}}
       &
       \item\label{fig:nonuni_sim}
-      \aligntop{\includegraphics[width=0.25\linewidth]{multistate_nonuni.svg}}
-      \\[2cm]
-      \aligntop{\includegraphics[width=0.3\linewidth]{pooled_deponly_learnS.eps}}
+      \aligntop{\includegraphics[width=0.24\linewidth]{multistate_nonuni.svg}}
+      \\[1.8cm]
+      \aligntop{\includegraphics[width=0.29\linewidth]{pooled_deponly_learnS.eps}}
       &
-      \aligntop{\includegraphics[width=0.3\linewidth]{cascade_long_learnS.eps}}
+      \aligntop{\includegraphics[width=0.29\linewidth]{cascade_fit_learnS.eps}}
       &
-      \aligntop{\includegraphics[width=0.3\linewidth]{nonuni_learnS.eps}}
+      \aligntop{\includegraphics[width=0.29\linewidth]{nonuni_fit_learnS.eps}}
   \end{tabular}
 
-  \vspace{1cm}\item\label{tab:params}\aligntop{
-  \begin{tabularn}{|l|c|c|c|c|c|c|}
-    \cline{1-7}
-    Model & \# states & pot & WT dep & \KO\ dep & $\Delta f$ & $r\tpre$ \\
-    \cline{1-7}
-    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 20  &\label{tr:multistate_med} \\
-    Two-state     & 2  & $q=0.1$  & $q=0.1$  & $q=0.2$  & 0.1  & 5   &\label{tr:binary} \\
-    Multistate    & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 5   &\label{tr:multistate_lin} \\
-    Pooled res. & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
-                                          & 0.4 & 20 &\label{tr:pooled_deponly}\\
-    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 100 &\label{tr:cascade_long} \\
-    Non-uni. MS& 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 150 &\label{tr:nonuni} \\
-    \cline{1-7}
-  \end{tabularn}}
+  \vspace{0.5cm}
+  \item\label{tab:params}\aligntop{
+  \begin{tabular}{|l|c|c|c|c|c|c|c|c|}
+    \cline{1-9}
+    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+    Model & \#  & \multicolumn{3}{c|}{Plasticity parameter} & \multicolumn{3}{c|}{$f\dep$} & $r\tpre$ \\
+    \cline{3-8}
+    & states & pot & WT dep & {\footnotesize\KO} dep & base & inc & dec & {} \\
+
+    \cline{1-9}
+    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.89 & 0.11  & 100  \\%&\label{tr:serial_fit} \\
+    Two-state     & 2  & $0.1$  & $0.1$  & $0.2$  & 0.5 & 0.6 & 0.4  & 5   \\%&\label{tr:binary} \\
+    Multistate    & 10 & $0.3$  & $0.3$  & $0.4$  & 0.5 & 0.8 & 0.2  & 5   \\%&\label{tr:multistate_lin} \\
+    Pooled res.\ & 7  & $0.008$        & $[0.0006,0.6]$  & $[0.001,1]$
+                                          & 0.5 & 0.9 & 0.1 & 20 \\%&\label{tr:pooled_deponly}\\
+    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002  & 200  \\%&\label{tr:cascade_fit} \\
+    Non-uni.\ & 12 & $0.4$    & $0.4$    & $0.53$   & 0.5 & 0.7 & 0.1 & 500  \\%&\label{tr:nonuni_fit} \\
+    \cline{1-9}
+  \end{tabular}}
   \end{myenuma}
   \caption[Simulation results for various models]{(\ref{fig:seial_sim}-\ref{fig:nonuni_sim}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (solid) and after gain-decrease pre-training (dashed) for wild-type (black) and \KO\ (red) models.
-(\ref{tab:params}) Parameters used for simulations.}\label{fig:sim_results}
+(\ref{tab:params}) Parameters used for simulations.
+For the serial, two state and multistate models, the plasticity parameter listed is the transition probability between adjacent states.
+For the cascade and non-uniform multistate models, the plasticity parameter is the ratio of adjacent transition probabilities.}\label{fig:sim_results}
 \end{figure}
@@
 
-First, consider the wild-type, for which $q\pot=q\dep=q$.
+First, consider the wild-type.
 Without pre-training:
@@
 
-Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
+Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and $q\pot=q$.
 Without pre-training:
@@
 
-In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f \in [\Delta f^*(1,M),\Delta f^*(\beta,M)]$ and look at intermediate time-scales, we will see that the \KO\ models learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ models but slows it down in the wild-type.
+In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f \in [\Delta f^*(1,M),\Delta f^*(\beta,M)]$, we will see that the \KO\ models learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ models but slows it down in the wild-type.
 %However, in this case pre-training proceeds much faster in the \KO\ models than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.
@@
 \begin{equation}\label{eq:binaryflux}
-\begin{aligned}
-  \Phi &= \frac{q\pot q\dep \Delta f}{f\pot q\pot + f\dep q\dep}, &\qquad
-  \pdiff{\Phi}{q\dep} &= \frac{f\pot (q\pot)^2 \Delta f}{(f\pot q\pot + f\dep q\dep)^2}, \\ &&
-  \pdiff{\Phi}{f\pot} &= \frac{(f\pot\inc q\pot + f\dep\inc q\dep) q\pot q\dep}{(f\pot q\pot + f\dep q\dep)^2}.
-\end{aligned}
+  \Phi = \frac{q\pot q\dep \Delta f}{\lambda}, \qquad
+  \pdiff{\Phi}{q\dep} = \frac{f\pot (q\pot)^2 \Delta f}{\lambda^2}, \qquad
+  \pdiff{\Phi}{f\pot} = \frac{\lambda\inc q\pot q\dep}{\lambda^2}.
 \end{equation}
@@
 
-The only thing we have to worry about is when the wild-type catches up with the \KO\ models.
+We have only looked at the initial learning rate here, but for the experiments we looked at the learning after a finite period of time.
 As the total change will be larger for the wild-type than \KO\ models, it must overtake eventually.
-If this happens early enough, the initial period, where the \KO\ models learns faster than wild-type, would not be seen in the experiment.
-However, the timescale for this crossover would be similar to the timescale of the exponentials.
-In the data, it looks like that timescale is longer than the gaps between successive measurements, so we needn't worry about this.
+If this happens early enough (\ie between the first and second data point in \datafig), the initial period where the \KO\ models learns faster than wild-type would not be seen in the experiment.
+However, the timescale for this crossover would be similar to the timescale of the exponentials, which is at least as long as the timescale for saturation of learning.
+In \datafig\ we see that this timescale is longer than the gaps between successive measurements, so we would not miss any initial period of enhanced learning in the \KO\ mice if this were the correct model.
+In addition, the models without pre-training can never catch up with those with pre-training, and this is sufficient to rule out this model.
 
@@
 
-With pre-training:
-%
-\begin{equation}\label{eq:binKOpre}
-\begin{aligned}
-  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
-          {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
-\end{aligned}
-\end{equation}
-%
-As we've already ruled out this model, we will not analyse this formula any further.
+%With pre-training:
+%%
+%\begin{equation}\label{eq:binKOpre}
+%\begin{aligned}
+%  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
+%          {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
+%\end{aligned}
+%\end{equation}
+%%
+%As we've already ruled out this model, we will not analyse this formula any further.
+As this model is already ruled out, we will not analyse pre-training in the \KO\ models.
 
@@
 
-In this section, we will consider the multistate model as originally defined in \cite{amit1994learning}, \ie with linearly varying synaptic weight:
-%
-\begin{equation}\label{eq:multistateLinWeight}
-  \w_i = \frac{2i-M-1}{M-1}.
-\end{equation}
-%
+In this section, we will consider the multistate model as defined in \cite{amit1994learning}, \ie with linearly varying synaptic weight.
 The numerical results can be seen in
@@
 
-The effect of pre-training is to replace $f\potdep$ with $f\potdep\dec$ without changing $f\potdep\inc$, which increases $f\pot$ and decreases $f\dep$.
-Therefore, we can see the effects of pre-training on the initial learning rate by seeing if it increases or decreases with $f\pot$ (with $f\dep = 1 - f\pot$):
+We can also see the effects of pre-training on the initial learning rate by seeing if it increases or decreases with $f\pot$ (with $f\dep = 1 - f\pot$):
 %
@@
 The monotonicity shown above means that this model cannot reproduce the key qualitative features of the experiments outlined in \autoref{sec:results}, at least as for the initial learning rate.
-This can be illustrated by looking at a special case, where $q\pot = q\dep = q$ for the wild type, $f\dep_0 = \frac{1}{2}$, $f\dep\dec = \frac{1}{2} - \Delta f$ and $f\dep\inc = \frac{1}{2} - \Delta f$.
-
-First, consider the wild-type, for which $q\pot=q\dep=q$.
-Without pre-training:
-%
-\begin{equation}\label{eq:multiLinWTnopre}
-  \Phi = 2{\Delta f}\,q\,\frac{M-1}{M}.
-\end{equation}
-%
-With pre-training:
-%
-\begin{equation}\label{eq:multiLinWTpre}
-\begin{aligned}
-  \Phi &= 4{\Delta f}\, q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
-          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
-       &= {4{\Delta f}\, q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
-\end{aligned}
-\end{equation}
-%
-
-Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
-Without pre-training:
-%
-\begin{equation}\label{eq:multiLinKOnopre}
-  \Phi = 2{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
-\end{equation}
-%
-This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
-This means that the \KO\ models will initially learn faster than the wild-type.
-However the wild-type will eventually catch up, and this can happen very quickly, as seen in
-\autoref{fig:sim_results}\ref{fig:multistate_sim} %\autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}
-(solid curves).
-
-With pre-training:
-%
-\begin{equation}\label{eq:multiLinKOpre}
-\begin{aligned}
-  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f)^{M-1} - \beta^{M-1}(1+2\Delta f)^{M-1}}
-          {(1-2\Delta f)^M - \beta^M(1+2\Delta f)^M} \\
-       &= 4{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
-\end{aligned}
-\end{equation}
-%
-The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the upper end of the interval $\Delta f \in \brk{0,\half}$, where it takes the value
-%
-\begin{equation}\label{eq:multiLinprenopre}
-  \frac{\Phi_{\text{w/ pre}}}{\Phi_{\text{w/o pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
-   \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,
-\end{equation}
-%
-which is always greater than 1.
-Therefore, pre-training will enhance the initial learning rate for all parameter values, for both \KO\ models and wild-type, as seen in
-\autoref{fig:sim_results}\ref{fig:multistate_sim}. %\autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
+%This can be illustrated by looking at a special case, where $q\pot = q\dep = q$ for the wild type, $f\dep_0 = \frac{1}{2}$, $f\dep\dec = \frac{1}{2} - \Delta f$ and $f\dep\inc = \frac{1}{2} - \Delta f$.
+%
+%First, consider the wild-type, for which $q\pot=q\dep=q$.
+%Without pre-training:
+%%
+%\begin{equation}\label{eq:multiLinWTnopre}
+%  \Phi = 2{\Delta f}\,q\,\frac{M-1}{M}.
+%\end{equation}
+%%
+%With pre-training:
+%%
+%\begin{equation}\label{eq:multiLinWTpre}
+%\begin{aligned}
+%  \Phi &= 4{\Delta f}\, q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
+%          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
+%       &= {4{\Delta f}\, q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
+%\end{aligned}
+%\end{equation}
+%%
+%
+%Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
+%Without pre-training:
+%%
+%\begin{equation}\label{eq:multiLinKOnopre}
+%  \Phi = 2{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
+%\end{equation}
+%%
+%This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
+%This means that the \KO\ models will initially learn faster than the wild-type.
+%However the wild-type will eventually catch up, and this can happen very quickly, as seen in
+%\autoref{fig:sim_results}\ref{fig:multistate_sim} %\autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}
+%(solid curves).
+%
+%With pre-training:
+%%
+%\begin{equation}\label{eq:multiLinKOpre}
+%\begin{aligned}
+%  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f)^{M-1} - \beta^{M-1}(1+2\Delta f)^{M-1}}
+%          {(1-2\Delta f)^M - \beta^M(1+2\Delta f)^M} \\
+%       &= 4{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
+%\end{aligned}
+%\end{equation}
+%%
+%The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the upper end of the interval $\Delta f \in \brk{0,\half}$, where it takes the value
+%%
+%\begin{equation}\label{eq:multiLinprenopre}
+%  \frac{\Phi_{\text{w/ pre}}}{\Phi_{\text{w/o pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
+%   \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,
+%\end{equation}
+%%
+%which is always greater than 1.
+%Therefore, pre-training will enhance the initial learning rate for all parameter values, for both \KO\ models and wild-type, as seen in
+%\autoref{fig:sim_results}\ref{fig:multistate_sim}. %\autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.
+
+We have only looked at the initial learning rate here, but for the experiments we looked at the learning after a finite period of time.
+Just as the case of the  \hyperref[sec:binary]{two state model}, the wild-type will eventually catch up with \KO\ model.
+This can happen very quickly, as seen in \autoref{fig:sim_results}\ref{fig:multistate_sim} (solid curves).
+This means that the initial period, where the \KO\ models learns faster than wild-type, might not be seen in the experiment.
+However, the models without pre-training can never catch up with those with pre-training, and this is sufficient to rule out this model.
 
@@
 \autoref{fig:sim_results}\ref{fig:pooled_sim}. %\autoref{fig:pooled_deponly}.
-It is difficult to study this model analytically.
-However, the numerical results can helps us understand it qualitatively.
+%It is difficult to study this model analytically.
+%However, the numerical results can help us understand it qualitatively.
+The numerical results can help us understand it qualitatively.
 
@@
 
-\subsection{Cascade model}\label{sec:cascade}
+\subsection{Cascade and non-uniform multistate models}\label{sec:cascade}
 
@@
 
-The results of simulations of the cascade model can be seen in
-\autoref{fig:sim_results}\ref{fig:cascade_sim}. %\autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
-It is difficult to study this model analytically.
-However, the numerical results can helps us understand it qualitatively.
-
-We see that the \KO\ models are slower than wild-type without pre-training but faster with it.
+The results of simulations of the cascade and non-uniform multistate models can be seen in
+\autoref{fig:sim_results}\ref{fig:cascade_sim},\ref{fig:nonuni_sim}. %\autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
+%It is difficult to study these models analytically.
+%However, the numerical results can help us understand them qualitatively.
+The numerical results can help us understand them qualitatively.
+
+For the cascade model, we see that the \KO\ models are slower than wild-type without pre-training but faster with it.
 This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed.
@@
 Sufficiently long pre-training pushes the synapses far down the cascade, slowing down learning for the wild-type (see \autoref{fig:sim_results}\ref{fig:cascade_sim}, black curves).
-This effect is weaker for the \KO\ models, as the equilibrium distribution is not as heavily concentrated at the end, as the distribution before pre-training is shifted in the other direction.
+This effect is weaker for the \KO\ models, as the equilibrium distribution is not as heavily concentrated at the end, as the distribution before pre-training is shifted in the opposite direction.
 %(see \autoref{fig:cascade_long}\ref{fig:cascade_long_eq}\ref{fig:cascade_long_eq_KO}).
 
-This model shares both features of the \hyperref[sec:multistate]{serial model}.
-First, the effect of initial saturation bias is amplified by the exponential decay of transition probabilities away from the middle.
-Second, it has the metaplastic effect where repeated potentiation makes future depression harder, also due to he exponential decay of transition probabilities.
-
-\subsection{Non-uniform multistate model}\label{sec:nonuni}
-
-%\resultsfign{non-uniform multistate}{nonuni}
-
-The results of simulations of the non-uniform multistate model can be seen in
-\autoref{fig:sim_results}\ref{fig:nonuni_sim}. %\autoref{fig:nonuni}.
-It is difficult to study this model analytically.
-However, the numerical results can helps us understand it qualitatively.
-
-First, we see that the \KO\ models are slower than wild-type without pre-training, as seen experimentally.
+For the non-uniform multistate model, we see that the \KO\ models are slower than wild-type without pre-training, as seen experimentally.
 This seems to be because the enhanced depression results in a greater fraction of synapses begin in states of weaker synaptic weight %(see \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_KO})
-where the depressing transitions have lower probability (see \autoref{fig:nonuni_model}).
+where the depressing transitions have lower probability (see \autoref{fig:models}\ref{fig:nonuni_model}).
 
@@
 %(compare \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_WT} and \ref{fig:nonuni_eq_KO}).
-As the depressing transitions have lower probability for the strongest states (see \autoref{fig:nonuni_model}), this, in addition to the enhanced depression of the \KO\ models, will result in the \KO\ models learning faster than wild-type, as seen experimentally.
+As the depressing transitions have lower probability for the strongest states (see \autoref{fig:models}\ref{fig:nonuni_model}), this, in addition to the enhanced depression of the \KO\ models, will result in the \KO\ models learning faster than wild-type, as seen experimentally.
 
 When comparing learning with and without gain-decrease pre-training, there are two effects to consider.
-First, the pushing synapses towards the states of stronger synaptic weight, resulting in more synapses being available for depression, speeding up learning.
+First, the synapses are pushing towards the states of stronger synaptic weight, resulting in more synapses being available for depression, speeding up learning.
 Second, it will also place synapses in states where transitions have lower probability, slowing down learning.
@@
 
-Like the \hyperref[sec:cascade]{cascade model}, this model also shares both features of the \hyperref[sec:multistate]{serial model}.
+These models share both key features of the \hyperref[sec:multistate]{serial model}.
 First, the effect of initial saturation bias is amplified by the exponential decay of transition probabilities away from the middle.
-Second, it has the metaplastic effect where repeated potentiation makes future depression harder, also due to he exponential decay of transition probabilities.
-
-\subsection{Parameter fitting}\label{sec:fit}
-
-So far, we have only considered the comparisons listed at the start of \hyperref[sec:results]{this section}, as they raised important conceptual questions.
-We have not considered the other two comparisons, the lack of difference between wild-type without pre-training and \KO\ with pre-training and between wild-type with pre-training and \KO\ without pre-training (the ``diagonal'' comparisons in the language of \modelfig), as they depend more on the values used for the parameters than on any conceptual issues.
-
-However, to avoid confusion, we found parameter values performed reasonably in these other two comparisons for the three models that could reproduce the correct qualitative results of the four important comparisons: the \hyperref[sec:multistate]{serial model}, the \hyperref[sec:cascade]{cascade model} and the \hyperref[sec:nonuni]{non-uniform multistate model}.
+Second, they both have the metaplastic effect where repeated potentiation makes future depression harder, also due to he exponential decay of transition probabilities.
+
+%\subsection{Non-uniform multistate model}\label{sec:nonuni}
+%
+%%\resultsfign{non-uniform multistate}{nonuni}
+%
+%
+%
+%The results of simulations of the non-uniform multistate model can be seen in
+%\autoref{fig:sim_results}\ref{fig:nonuni_sim}. %\autoref{fig:nonuni}.
+%It is difficult to study this model analytically.
+%However, the numerical results can helps us understand it qualitatively.
+%
+%First, we see that the \KO\ models are slower than wild-type without pre-training, as seen experimentally.
+%This seems to be because the enhanced depression results in a greater fraction of synapses begin in states of weaker synaptic weight %(see \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_KO})
+%where the depressing transitions have lower probability (see \autoref{fig:nonuni_model}).
+%
+%However, gain-decrease pre-training will push synapses towards the states of stronger synaptic weight.
+%The wild-type will be pushed further that the \KO\ models, as the latter started further to the weaker side as explained above.
+%(compare \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_WT} and \ref{fig:nonuni_eq_KO}).
+%As the depressing transitions have lower probability for the strongest states (see \autoref{fig:nonuni_model}), this, in addition to the enhanced depression of the \KO\ models, will result in the \KO\ models learning faster than wild-type, as seen experimentally.
+%
+%When comparing learning with and without gain-decrease pre-training, there are two effects to consider.
+%First, the pushing synapses towards the states of stronger synaptic weight, resulting in more synapses being available for depression, speeding up learning.
+%Second, it will also place synapses in states where transitions have lower probability, slowing down learning.
+%The second effect will be weaker in the \KO\ models than in wild-type, as they started further to the weaker side as explained above.
+%This means that, for appropriate parameter choices, gain-decrease pre-training can speed up learning in the \KO\ models, but also slow down learning in wild-type, as seen experimentally.
+%
+%Like the \hyperref[sec:cascade]{cascade model}, this model also shares both features of the \hyperref[sec:multistate]{serial model}.
+%First, the effect of initial saturation bias is amplified by the exponential decay of transition probabilities away from the middle.
+%Second, it has the metaplastic effect where repeated potentiation makes future depression harder, also due to he exponential decay of transition probabilities.
+
+%\subsection{Parameter fitting}\label{sec:fit}
+%
+%So far, we have only considered the comparisons listed at the start of \hyperref[sec:results]{this section}, as they raised important conceptual questions.
+%We have not considered the other two comparisons, the lack of difference between wild-type without pre-training and \KO\ with pre-training and between wild-type with pre-training and \KO\ without pre-training (the ``diagonal'' comparisons in the language of \modelfig), as they depend more on the values used for the parameters than on any conceptual issues.
+%
+%However, to avoid confusion, we found parameter values performed reasonably in these other two comparisons for the three models that could reproduce the correct qualitative results of the four important comparisons: the \hyperref[sec:multistate]{serial model}, the \hyperref[sec:cascade]{cascade model} and the \hyperref[sec:nonuni]{non-uniform multistate model}.
 The learning curves are shown in \autoref{fig:fit} and the parameters are listed in \autoref{fig:fit}\ref{tab:fit}.
-
-%\begin{table}
-% \begin{center}
-%  \begin{tabularn}{|l|c|c|c|c|c|c|c|c|c|}
-%    \cline{1-10}
-%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-%    Model & \# states & \multicolumn{3}{c|}{parameter} & \multicolumn{3}{c|}{$f\dep$} & \multicolumn{2}{c|}{$rt$} \\
-%    \cline{3-10}
-%    & & pot & WT dep & \KO\ dep & base & inc & dec & train & pre \\
-%
-%    \cline{1-10}
-%    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.11 & 0.89 & 5  & 100  &\label{tr:serial_fit} \\
-%    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002 & 1.5  & 200  &\label{tr:cascade_fit} \\
-%    Non-uni. & 12 & $0.4$    & $0.4$    & $0.53$   & 0.5 & 0.7 & 0.1 & 5  & 500  &\label{tr:nonuni_fit} \\
-%    \cline{1-10}
-%  \end{tabularn}
-% \end{center}
-%  \caption{Parameters used to reproduce detailed experimental results.
+%
+%
+%%\begin{table}
+%% \begin{center}
+%%  \begin{tabularn}{|l|c|c|c|c|c|c|c|c|c|}
+%%    \cline{1-10}
+%%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+%%    Model & \# states & \multicolumn{3}{c|}{parameter} & \multicolumn{3}{c|}{$f\dep$} & \multicolumn{2}{c|}{$rt$} \\
+%%    \cline{3-10}
+%%    & & pot & WT dep & \KO\ dep & base & inc & dec & train & pre \\
+%%
+%%    \cline{1-10}
+%%    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.11 & 0.89 & 5  & 100  &\label{tr:serial_fit} \\
+%%    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002 & 1.5  & 200  &\label{tr:cascade_fit} \\
+%%    Non-uni. & 12 & $0.4$    & $0.4$    & $0.53$   & 0.5 & 0.7 & 0.1 & 5  & 500  &\label{tr:nonuni_fit} \\
+%%    \cline{1-10}
+%%  \end{tabularn}
+%% \end{center}
+%%  \caption{Parameters used to reproduce detailed experimental results.
 For the sewrial model, the parameter listed is the transition probability between adjacent states.
 For the cascade and non-uniform multistate models, the parameter is the ratio of adjacent transition probabilities.} \label{tab:fit}
-%\end{table}
-
-\begin{figure}
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.28\linewidth]{serial_fit_learnS.eps}}\label{fig:serial_fit}
-  \item\aligntop{\includegraphics[width=0.28\linewidth]{cascade_fit_learnS.eps}}\label{fig:cascade_fit}
-  \item\aligntop{\includegraphics[width=0.28\linewidth]{nonuni_fit_learnS.eps}}\label{fig:nonuni_fit}
-
-  \vspace{1cm}\item\label{tab:fit}\aligntop{
-  \begin{tabularn}{|l|c|c|c|c|c|c|c|c|c|}
-    \cline{1-10}
-    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
-    Model & \#  & \multicolumn{3}{c|}{parameter} & \multicolumn{3}{c|}{$f\dep$} & \multicolumn{2}{c|}{$rt$} \\
-    \cline{3-10}
-    & states & pot & WT dep & {\footnotesize\KO} dep & base & inc & dec & train & pre \\
-
-    \cline{1-10}
-    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.89 & 0.11 & 5  & 100  &\label{tr:serial_fit} \\
-    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002 & 1.5  & 200  &\label{tr:cascade_fit} \\
-    Non-uni. & 12 & $0.4$    & $0.4$    & $0.53$   & 0.5 & 0.7 & 0.1 & 5  & 500  &\label{tr:nonuni_fit} \\
-    \cline{1-10}
-  \end{tabularn}}
- \end{myenuma}
- \end{center}
-  \caption[Learning curves for models after parameter fitting]{Learning curves restricted to gain-increase training, for (\ref{fig:serial_fit}) the serial model, (\ref{fig:cascade_fit}) the cascade model, and (\ref{fig:nonuni_fit}) the non-uniform multistate model.
-  (\ref{tab:fit}) Parameters used to reproduce detailed experimental results.
+%%\end{table}
+%
+%\begin{figure}
+% \begin{center}
+% \begin{myenuma}
+%  \item\aligntop{\includegraphics[width=0.28\linewidth]{serial_fit_learnS.eps}}\label{fig:serial_fit}
+%  \item\aligntop{\includegraphics[width=0.28\linewidth]{cascade_fit_learnS.eps}}\label{fig:cascade_fit}
+%  \item\aligntop{\includegraphics[width=0.28\linewidth]{nonuni_fit_learnS.eps}}\label{fig:nonuni_fit}
+%
+%  \vspace{1cm}\item\label{tab:fit}\aligntop{
+%  \begin{tabularn}{|l|c|c|c|c|c|c|c|c|}
+%    \cline{1-9}
+%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+%    Model & \#  & \multicolumn{3}{c|}{parameter} & \multicolumn{3}{c|}{$f\dep$} & $r\tpre$ \\
+%    \cline{3-8}
+%    & states & pot & WT dep & {\footnotesize\KO} dep & base & inc & dec &  \\
+%
+%    \cline{1-9}
+%    Serial   & 10 & $0.12$  & $0.14$  & $0.2$  & 0.5 & 0.89 & 0.11  & 100  &\label{tr:serial_fit} \\
+%    Two-state     & 2  & $0.1$  & $0.1$  & $0.2$  & 0.5 & 0.6 & 0.4  & 5   &\label{tr:binary} \\
+%    Multistate    & 10 & $0.3$  & $0.3$  & $0.4$  & 0.5 & 0.8 & 0.2  & 5   &\label{tr:multistate_lin} \\
+%    Pooled res. & 7  & $0.008$        & $[0.0006,0.6]$  & $[0.001,1]$
+%                                          & 0.5 & 0.9 & 0.1 & 20 &\label{tr:pooled_deponly}\\
+%    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002  & 200  &\label{tr:cascade_fit} \\
+%    Non-uni. & 12 & $0.4$    & $0.4$    & $0.53$   & 0.5 & 0.7 & 0.1 & 500  &\label{tr:nonuni_fit} \\
+%    \cline{1-9}
+%  \end{tabularn}}
+% \end{myenuma}
+% \end{center}
+%  \caption[Learning curves for models after parameter fitting]{Learning curves restricted to gain-increase training, for (\ref{fig:serial_fit}) the serial model, (\ref{fig:cascade_fit}) the cascade model, and (\ref{fig:nonuni_fit}) the non-uniform multistate model.
+%  (\ref{tab:fit}) Parameters used to reproduce detailed experimental results.
 For the serial model, the parameter listed is the transition probability between adjacent states.
 For the cascade and non-uniform multistate models, the parameter is the ratio of adjacent transition probabilities.}\label{fig:fit}
-\end{figure}
+%\end{figure}
 

### 2016-06-11 MHCpaper/MHCsupp.tex
wrote about popn of different binary synapses

@@
 %\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
-\renewcommand{\figureautorefname}{Supplementary Figure}
+\renewcommand{\figureautorefname}{Supp.~Fig.}
 % ----------------------------------------------------------------
@@
 These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
-We will not study gain-decrease learning, as it uses a different mechanism to gain-increase.
+\emph{We will not study gain-decrease learning}, as its mechanisms are not fully understood, and known to be different to to gain-increase.
 We are only modelling the effect of gain-decrease training on \emph{these} synapses.
@@
   For the serial, two state and multistate models, the plasticity parameter listed is the transition probability between adjacent states.
+  For the pooled resource model, the plasticity parameters are the minimum and maximum transition probability for the constituent two-state synapses.
   For the cascade and non-uniform multistate models, the plasticity parameter is the ratio of adjacent transition probabilities.}\label{fig:sim_results}
@@
 %
-But it is easier to just substitute $M=2$ into the formulae in \autoref{sec:multistate}.
+In practice, it is easier to just substitute $M=2$ into the formulae in \autoref{sec:multistate}.
 %In this case, the initial rate of change and the total change encapsulate the whole solution, as there is only a single exponential decay.
@@
   \pdiff{\Phi}{q\dep} = \frac{f\pot (q\pot)^2 \Delta f}{\lambda^2}, \qquad
-  \pdiff{\Phi}{f\pot} = \frac{\lambda\inc q\pot q\dep}{\lambda^2}.
+  \pdiff{\Phi}{f\pot} = \frac{q\pot q\dep \lambda\inc}{\lambda^2}.
 \end{equation}
@@
 The flux is a monotonically increasing function of $q\dep$ and $f\pot$.
-Therefore, enhancing plasticity in this way or pre-training can only enhance the initial learning rate, unlike the experimental results shown in \modelfig.
+Therefore, enhancing plasticity in this way or pre-training can only enhance the initial learning rate for all values of the parameters, unlike the experimental results shown in \modelfig.
 
@@
 
+This assumed that all synapses have the same parameters.
+What would happen if there were a population of synapses with different parameters?
+In this case, the learning rate would be determined by the average flux over this distribution:
+%
+\begin{equation}\label{eq:meanflux}
+  \overline{\Phi} = \int \Phi(q\dep, \ldots) \,
+                            p(q\dep, \ldots) \,
+                          \dr q\dep \ldots,
+\end{equation}
+%
+where the dependence on all other parameters is omitted.
+
+The \KO\ mice will have a different distribution, $p\ko(q\dep, \ldots)$.
+It is helpful to introduce a function $P(q\dep, \ldots)$ that is a cumulative distribution \wrt $q\dep$, but a probability density \wrt all other parameters:
+%
+\begin{equation}\label{eq:cumul}
+  P(q\dep, \ldots) = \int_0^{q\pot} \!\!\! p(q, \ldots) \, \dr q.
+\end{equation}
+%
+Then, integration by parts leads to
+%
+\begin{equation}\label{eq:fluxcumul}
+  \overline{\Phi} = \av{\Phi(1, \ldots)} - 
+             \int \pdiff{\Phi}{q\pot} \,
+                            P(q\dep, \ldots) \,
+                          \dr q\dep \ldots,
+\end{equation}
+%
+where the first term is averaged over all the other omitted parameters and the partial derivative is positive for this model \eqref{eq:binaryflux}.
+
+The statement that the \KO\ models have larger $q\dep$ than wild-type is now a statement about the distributions, with the \KO\ models having more density at larger values.
+This is not a precise concept, but it will tend to lead to $P(q\dep, \ldots)$ being smaller for the \KO\ models than wild type, which makes the second term less negative and the average flux larger, in contradiction with the experimental results in \modelfig.
+
+If $P\ko \leq P\wt$ at all parameter values and the marginal distribution for all other parameters is the same for both models, then $\overline{\Phi}\ko \geq \overline{\Phi}\wt$.
+This is guaranteed if, for example, there is a one-to-one map between $q\dep$ for the two models, so that the cumulative distributions are the same at $q\dep\wt$ and $q\dep\ko = g(q\dep\wt)$ for some function $g(x) \geq x$.
+
+If the dependence of the distribution on $q\dep$ is mixed with the other parameters in some complicated way, or if the edges of the distributions are very unusual, it could be possible for the enhanced plasticity models to have an impaired initial learning rate, but this seems contrived and unlikely.
+A similar argument applies to the effect of pre-training and the distributions over $f\pot$.
+
 The monotonicity shown above means that this model cannot reproduce the key qualitative features of the experiments outlined in \autoref{sec:results}, at least as for the initial learning rate.
@@
 %
-which is always larger than \eqref{eq:binWTnopre}, unlike what is seen in the experiments (see \modelfig).
+which is larger than \eqref{eq:binWTnopre} for all $\beta$, unlike what is seen in the experiments (see \modelfig).
 This is also seen in
@@
 There is a requirement that the total number of synapses, potentiated and depressed, is fixed.
-This normalization effect always dampens the effect of saturation bias, so that the effect of enhanced intrinsic plasticity rates always dominates.
-Thus, we always see enhanced learning with enhanced plasticity with the binary synapse model.
-
-One can see that pre-training will always increase the fraction of synapses available for depression.
+This normalization effect dampens the effect of saturation bias, so that the effect of enhanced intrinsic plasticity rates dominates for all parameter values.
+Thus, we must see enhanced initial learning with enhanced plasticity in the binary synapse model.
+
+One can see that pre-training can only increase the fraction of synapses available for depression.
 This model has no mechanism that could ever result in this causing impaired learning.
@@
 Therefore this quantity is positive, so $\Phi$ is a monotonically increasing function of $q\dep$.
-Therefore, enhancing plasticity in this way can only enhance the initial learning rate, for all values the parameters.
+Therefore, enhancing plasticity in this way can only enhance the initial learning rate, for all values of the parameters.
 
@@
 This quantity is positive, for similar reasons to \eqref{eq:multiderq}, so $\Phi$ is a monotonically increasing function of $f\pot$.
-Therefore, pre-training can can only enhance the initial learning rate, for all values the parameters.
+Therefore, pre-training can can only enhance the initial learning rate, for all values of the parameters.
 

### 2016-10-18 MHCpaper/MHCsupp.tex
format supp for elife

@@
 %\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
-\renewcommand{\figureautorefname}{Supp.~Fig.}
+\renewcommand{\figurename}{Appendix-figure}
+\renewcommand{\figureautorefname}{Appendix-figure}
+\DeclareCaptionLabelFormat{supp}{\textbf{#1 #2}}
+\captionsetup{labelformat=supp}
 % ----------------------------------------------------------------
@@
 %\newcommand{\lmin}{_{\text{min}}}
-\newcommand{\modelfig}[1][a]{Figure 4#1 of the main text}
-\newcommand{\datafig}[1][b]{Figure 3#1 of the main text}
+\newcommand{\modelfig}[1][a]{Figure 3#1 of the main text}
+\newcommand{\datafig}[1][b]{Figure 2#1 of the main text}
 % ----------------------------------------------------------------
@@
 % Title info:
-\title{Understanding both enhanced and impaired learning with enhanced plasticity: a saturation hypothesis (supplementary material)}
+\title{Appendix to ``Understanding both enhanced and impaired learning with enhanced plasticity: a saturation hypothesis''}
 %
@@
 \date{}
-\DeclareCaptionLabelFormat{supp}{\textbf{#1 S#2}}
-\captionsetup{labelformat=supp}
 
@@
 
-\setcounter{page}{8}
-\setcounter{figure}{7}
+%\setcounter{page}{8}
+%\setcounter{figure}{7}
 \maketitle
@@
  \end{center}
-  \caption[Transition probabilities for different models]{Transition probabilities for different models.
+  \caption[Transition probabilities for different models]{{Transition probabilities for different models.}
   Potentiation induces transitions indicated by orange arrows, depression indicated by blue arrows.

### 2016-10-24 MHCpaper/MHCsupp.tex
scan & chr2 figs

@@
 
+\hypertarget{par:pooled_scan}{To} verify that this always happens, we scanned over acceptable values of the six parameters relevant to gain-decrease pre-training in wild-type models: $q\pot$, $q\dep\lmin < q\dep\lmax$, and $f\dep\dec < f\dep\norm < f\dep\inc$.
+Each parameter was scanned over 10 values from 0.05 to 0.95, rejecting values that do not respect the inequalities above.
+For each parameter set, we computed the initial learning rate, $\dot{L}(0)$ from \eqref{eq:learning}, with and without pre-training.
+To qualitatively reproduce the experimental results, we would need $\dot{L}_\text{no pre} - \dot{L}_\text{pre} > 0$ at $t = 0$.
+The range of values obtained is plotted in \autoref{fig:pooled_scan}, where we see that this difference is always negative, ruling out this model.
+
+\begin{figure}
+  \centering
+  % Requires \usepackage{graphicx}
+  \includegraphics[width=0.4\linewidth]{pooled_deponly_scan.eps}
+  \caption{Difference in initial learning rate in wild-type models, without and with gain-decrease pre-training.
+  Learning, $L(t)$ is defined in \eqref{eq:learning}.
+  Here, we show the maximum and minimum values of this quantity when scanning all 6 relevant parameters (see \protect\hyperlink{par:pooled_scan}{text}).
+  We see that this quantity is always negative, whereas it was positive in the experiments.
+  }\label{fig:pooled_scan}
+\end{figure}
+
 This model shares one feature of the \hyperref[sec:multistate]{serial model}.

### 2016-10-24 Code/VOR/Plots/printVORfigsChR2.m

@@ -48,13 +48,13 @@ end
 vexpt=VORbuilderChR2(builder_h, n, paramPot, paramDep, df{:}, T_train, T_pre, pooled);
 % vexpt=VORbuilderChR2eq(builder_h, n, paramPot, paramDep, df{:}, T_train, pooled);
 
-fig=figure('WindowStyle','docked','PaperPositionMode','auto');
+fig=figure('PaperPositionMode','auto','Position',[60 60 1000 1000]);
 Parent=axes('Parent',fig);
 vexpt.PlotLearn('LineWidth',2,'Parent',Parent);
 print(fig,[prefix '_ChR2_learn.eps'],'-depsc');
 close(fig);
 
-figs=figure('WindowStyle','docked','PaperPositionMode','auto');
+figs=figure('PaperPositionMode','auto','Position',[60 60 1000 1000]);
 Parent=axes('Parent',figs);
 vexpt.PlotLearnS('LineWidth',2,'Parent',Parent);
 print(figs,[prefix '_ChR2_learnS.eps'],'-depsc');

### 2016-11-01 MHCpaper/MHCsupp.tex
added chr2 fig

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Model of VOR learning experiment}\label{sec:learning}
@@
 This ensures that the qualitative questions mentioned above (enhancement or impairment of learning) will not be affected.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Serial model}\label{sec:multistate}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Two-state model}\label{sec:binary}
@@
 The flux is a monotonically increasing function of $q\dep$ and $f\pot$.
-Therefore, enhancing plasticity in this way or pre-training can only enhance the initial learning rate for all values of the parameters, unlike the experimental results shown in \modelfig.
+Therefore, enhancing plasticity in this way, or pre-training, can only enhance the initial learning rate for all values of the parameters, unlike the experimental results shown in \modelfig.
 
@@
 %
-So, we see that pre-training will speed up learning for all parameter values, unlike what is seen in the experiment.
+So, we see that pre-training will speed up learning, unlike what is seen in the experiment.
 This can be seen in
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Multistate model}\label{sec:multistate_lin}
@@
 Second, it does not have the metaplastic effect where repeated potentiation makes future depression harder, as potentiation will merely increase the number of potentiated synapses without pushing them away from any boundary between strong and weak states.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 However, it does not have the metaplastic effect where repeated potentiation makes future depression harder, in fact it shows the opposite effect due to resource replenishment, so it can never show impaired learning after pre-training.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+\section{Model of non-specific CF stimulation}\label{sec:chr2}
+
+\begin{figure}[p]
+\begin{myenuma}
+  \begin{tabular}{lll}
+      \item\label{fig:seial_ChR2}
+      \aligntop{\includegraphics[width=0.24\linewidth]{serial.svg}}
+      &
+      \item\label{fig:binary_ChR2}
+      \hspace{0.05\linewidth}\aligntop{\includegraphics[width=0.12\linewidth]{binary.svg}}
+      &
+      \item\label{fig:multistate_ChR2}
+      \aligntop{\includegraphics[width=0.24\linewidth]{multistate.svg}}
+      \\[1.3cm]
+      \aligntop{\includegraphics[width=0.29\linewidth]{serial_fit_ChR2_learnS.eps}}
+      &
+      \aligntop{\includegraphics[width=0.29\linewidth]{binary_ChR2_learnS.eps}}
+      &
+      \aligntop{\includegraphics[width=0.29\linewidth]{multistate_lin_ChR2_learnS.eps}}
+      \\[4.5cm]
+      \item\label{fig:pooled_ChR2}
+      \aligntop{\includegraphics[width=0.29\linewidth]{pooled_deponly.svg}}
+      &
+      \item\label{fig:cascade_ChR2}
+      \aligntop{\includegraphics[width=0.24\linewidth]{cascade.svg}}
+      &
+      \item\label{fig:nonuni_ChR2}
+      \aligntop{\includegraphics[width=0.24\linewidth]{multistate_nonuni.svg}}
+      \\[1.8cm]
+      \aligntop{\includegraphics[width=0.29\linewidth]{pooled_deponly_ChR2_learnS.eps}}
+      &
+      \aligntop{\includegraphics[width=0.29\linewidth]{cascade_fit_ChR2_learnS.eps}}
+      &
+      \aligntop{\includegraphics[width=0.29\linewidth]{nonuni_fit_ChR2_learnS.eps}}
+  \end{tabular}
+
+  \vspace{0.5cm}
+  \item\label{tab:params_ChR2}\aligntop{
+  \begin{tabular}{|l|c|c|c|c|c|c|c|}
+    \cline{1-8}
+    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
+    Model & \#  & \multicolumn{2}{c|}{Plasticity param.} & \multicolumn{3}{c|}{$f\dep$} & $r\tpre$ \\
+    \cline{3-7}
+    & states & pot & dep & base & inc & CF Stim & {} \\
+
+    \cline{1-8}
+    Serial       & 10 & $0.12$  & $0.14$        & 0.5   & 0.89 & 0.9879 & 100 \\
+    Two-state    & 2  & $0.1$   & $0.4$         & 0.5   & 0.7  & 0.91   & 5   \\
+    Multistate   & 10 & $0.3$   & $0.3$         & 0.5   & 0.8  & 0.96   & 5   \\
+    Pooled res.\ & 7  & $0.08$  & $[0.006,0.6]$ & 0.5   & 0.9  & 0.99   & 20  \\
+    Cascade      & 14 & $0.386$ & $0.398$       & 0.522 & 0.63 & 0.99   & 200  \\ 
+    Non-uni.\    & 12 & $0.4$   & $0.4$         & 0.5   & 0.7  & 0.99   & 500  \\
+    \cline{1-8}
+  \end{tabular}}
+  \end{myenuma}
+  \caption[Simulation results for various models]{(\ref{fig:seial_sim}-\ref{fig:nonuni_sim}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (black) and after non-specific CF stimulation (cyan) for wild-type models.
+  (\ref{tab:params}) Parameters used for simulations.
+  For the serial, two state and multistate models, the plasticity parameter listed is the transition probability between adjacent states.
+  For the pooled resource model, the plasticity parameters are the minimum and maximum transition probability for the constituent two-state synapses.
+  For the cascade and non-uniform multistate models, the plasticity parameter is the ratio of adjacent transition probabilities.}\label{fig:sim_ChR2}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \bibliographystyle{utcaps_sl}
-\bibliography{maths,neuro,markov}
+\bibliography{neuro,markov}
 

### 2016-11-03 MHCpaper/MHCsupp.tex
talking about chr2

@@
 \newcommand{\datafig}[1][b]{Figure 2#1 of the main text}
+\newcommand{\chrfig}[1][b]{Figure 4#1 of the main text}
+\newcommand{\compfig}[1][a]{Figure 4 -- figure supplement 2#1}
 % ----------------------------------------------------------------
@@
 
+Here, we model the effects of non-specific CF stimulation, as in \chrfig. 
+It is necessary to also model the ``other'' population of synapses that compensate for the change in baseline synaptic weight of the synapses considered up to now, either due to enhanced plasticity or non-specific CF stimulation.
+
+To be concrete, we model these other synapses as described in \compfig.
+We set the relative rate of depression events for these synapses during training to $f\dep = f\dep\norm$, \ie the baseline value, so that they are unaffected by training.
+For both sets of synapses during non-specific CF stimulation, we set $f\dep = > f\dep\norm$.
+
 \begin{figure}[p]

### 2016-11-04 MHCpaper/MHCsupp.tex
describe chr2 results

@@
 We set the relative rate of depression events for these synapses during training to $f\dep = f\dep\norm$, \ie the baseline value, so that they are unaffected by training.
-For both sets of synapses during non-specific CF stimulation, we set $f\dep = > f\dep\norm$.
+Here we set the rate of plasticity events causing forgetting of the CF stimulation in these synapses is the same as the rate of plasticity events in the original synapses during training, although similar results could be acheived with other choices.
+For both sets of synapses during non-specific CF stimulation, we set $f\dep > f\dep\norm$ to model the increased rate of depression.
+The values of all parameters used can be seen in \autoref{fig:sim_ChR2}\ref{tab:params_ChR2}.
+
+In \autoref{fig:sim_ChR2}\ref{fig:seial_ChR2}--\ref{fig:nonuni_ChR2} we see examples of simulations of all of the models we have considered.
+We see that every single one of them is capable of reproducing the experimental results of \chrfig, for suitable parameter choices.
+This is because this intervention has the effect of depleting the population of synapses available for depression \emph{without} the competing effect of enhanced plasticity.
+Therefore the depletion effect is stronger (than 0) and learning is impaired.
+This means that these experiments cannot rule out any of the models.
+However these experiments do agree with the idea of impaired learning due to depletion of the population of labile synapses.
 

### 2016-11-04 Code/VOR/Plots/PooledScan.m

@@ -8,7 +8,7 @@ ncomps=ScanPooledNum(paramvals,4:10);
 txopts={'Interpreter','latex','FontSize',20};
 ax=gca;
 bh=bar(ax, ns-1, ncomps','stacked');
-xlabel('Pool size', txopts{:});
+xlabel('$\#$ of synapses in pool', txopts{:});
 ylabel('max/min $\{\dot{L}(0) - \dot{L}_\mathrm{pre}(0)\}$', txopts{:});
 title('Pooled resource model, WT', txopts{:});
 bh(1).FaceColor='none';

### 2016-11-18 MHCpaper/MHCsupp.tex
working scans?

@@
 %\newcommand{\lmin}{_{\text{min}}}
-\newcommand{\modelfig}[1][a]{Figure 3#1 of the main text}
-\newcommand{\datafig}[1][b]{Figure 2#1 of the main text}
-\newcommand{\chrfig}[1][b]{Figure 4#1 of the main text}
-\newcommand{\compfig}[1][a]{Figure 4 -- figure supplement 2#1}
+\newcommand{\modelfig}[1][A]{Figure 3#1 of the main text}
+\newcommand{\datafig}[1][B]{Figure 2#1 of the main text}
+\newcommand{\chrfig}[1][B]{Figure 4#1 of the main text}
+\newcommand{\compfig}[1][A]{Figure 4 -- figure supplement 2#1}
+\newcommand{\basefig}{Figure 1 -- figure supplement 2}
 % ----------------------------------------------------------------
@@
 
+\section{Balanced potentiation and depression}\label{sec:bal}
+
+As the baseline oculomotor performance was normal in the \KO\ mice (see \basefig), one might ask if it is possible that potentiation is also enhanced by the knockout, in such a way that the baseline mean synaptic weight is unchanged.
+
+Intuitively, this seems impossible.
+The impaired learning observed in some of the models considered depended on the competition between the enhanced intrinsic plasticity rate and the depletion of the population of synapses available for depression.
+If the second effect is removed, the first must dominate and learning will be enhanced.
+
+However, enhancing potentiation will increase the number of synapses making transitions in the ``wrong'' direction, impairing learning.
+Furthermore, the distribution of synapses across labile and stubborn states may be altered, also affecting the learning rate.
+
+For the \hyperref[sec:multistate]{serial}, \hyperref[sec:binary]{two-state} and \hyperref[sec:multistate_lin]{multistate} models, the validity of the intuition can be seen analytically.
+From \eqref{eq:mutltieq}, we can see that the baseline mean synaptic weight depends monotonically on the quantity
+%
+\begin{equation*}
+  \alpha=\frac{f\pot q\pot}{f\dep q\dep}.
+\end{equation*}
+%
+Maintaining the same baseline would then require increasing $q\pot$ in proportion to $q\dep$.
+As one can see from \eqref{eq:multiflux}, \eqref{eq:binaryflux} and \eqref{eq:multiLinFlux}, this will always lead to an enhanced initial learning rate.
+
+For the other models, a numerical analysis is required, as shown in \autoref{fig:bal}.
+\hypertarget{par:bal_scan}{To} see if learning is impaired with enhanced plasticity each relevant parameter was scanned over 10 values from 0.05 to 0.95, rejecting values that do not respect the inequalities:
+$0 \leq f\dep\norm < f\dep\inc \leq 1$, for the pooled resource model $0 \leq q\dep_\text{min,WT} < q\dep_\text{min,\KO}, q\dep_\text{max,WT} < q\dep_\text{max,\KO}$ and $0 \leq q\pot\wt$, for the cascade model $0 \leq x\dep\wt < x\dep\ko \leq \frac{1}{2}$ and $0 \leq x\pot\wt \leq \frac{1}{2}$, and for the nonuniform multistate model $0 \leq x\dep\wt < x\dep\ko \leq 1$ and $0 \leq x\pot\wt \leq 1$.
+In each case, the \KO\ potentiation parameter, $q\pot\ko$ for the pooled resource model and $x\pot\ko$ for the others, was adjusted to match the baseline mean synaptic weight.
+$\pr(0)\w$, with wild-type.
+
+\begin{figure}
+  \centering
+  \begin{myenuma}
+    \item\aligntop{\includegraphics[width=0.29\linewidth]{pooled_deponly_scan.eps}}\label{fig:pooled_bal}
+    \item\aligntop{\includegraphics[width=0.29\linewidth]{cascade_scan.eps}}\label{fig:cascade_bal}
+    \item\aligntop{\includegraphics[width=0.29\linewidth]{nonuni_scan.eps}}\label{fig:nonuni_bal}
+  \end{myenuma}
+  \caption{Difference in initial learning rate in wild-type and \KO\ models, without gain-decrease pre-training.
+  Learning, $L(t)$ is defined in \eqref{eq:learning}.
+  Here, we show the maximum and minimum values of this quantity when scanning all 7/5 relevant parameters (see \protect\hyperlink{par:bal_scan}{text}).
+  We see that this quantity is usually negative, whereas it was positive in the experiments.
+  }\label{fig:bal}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \bibliographystyle{utcaps_sl}

### 2016-11-18 Code/VOR/Plots/PooledScanTop.m

@@ -1,0 +1,21 @@
+%parameter ranges
+paramvals = (0.05:0.1:0.95);
+ns = 4:10;
+reps=2;
+%%
+%calculations
+ncomps=ScanPooledNum(paramvals,ns,'Top',reps);
+%%
+%plot
+txopts={'Interpreter','latex','FontSize',20};
+ax=gca;
+bh=bar(ax, ns-1, ncomps','stacked');
+xlabel('$\#$ of synapses in pool', txopts{:});
+ylabel('max/min $\{\dot{L}_{\mathrm{WT}}(0) - \dot{L}_{\mathrm{K}^b\mathrm{D}^{b-/-}}(0)\}$', txopts{:});
+title('Pooled resource model, WT', txopts{:});
+bh(1).FaceColor='none';
+bh(1).EdgeColor='none';
+bh(2).FaceColor=[0 0.4470 0.7410];
+ax.YAxis.Scale = 'log';
+%%
+print('pooled_scan.eps','-depsc');

### 2016-11-18 Code/VOR/Scan/ScanPooledTop.m

@@ -1,0 +1,85 @@
+function [ comps,wh ] = ScanPooledtop( prm, n, reps  )
+%comps=SCANCNTOP(ranges,n,reps) parameter scan for pooled
+%resource model (dep only)
+%   comps  = learning rate differences: WT_nopre - KO_nopre
+%   ranges = range for each parameter scan
+%   n      = number of states
+%   parametrs: pot_wt, dep_wt, dep_ko, fp_norm, fp_inc
+
+builder_h = @PooledBuilder;
+minv = 1e-4;
+maxv = 1 - minv;
+
+vexpt=VORbuilderKO(builder_h, n, prm(1), prm(1), prm(2), prm(2), prm(1), 0.5, 1,1, false);
+m = 1:length(prm);
+M = m(end);
+wh=[];
+
+comps = -Inf(length(prm) * ones(1,5));
+
+for i1 = m
+    DispCounter(i1,M,'i1:');
+    
+    Wp = builder_h(n, prm(i1));
+    vexpt.WT = vexpt.WT.setWp(Wp);
+    
+    for i2 = m(1:end-2)
+        DispCounter(i2,M-2,'i2:');
+        DispCounter(1,M-1,'i2b:');
+        for i2b = m(m > i2 & m < M)
+            DispCounter(i2b,M-1,'i2b:');
+
+            [~,Wm] = builder_h(n, [prm(i2b) prm(i2)]);
+            vexpt.WT = vexpt.WT.setWm(Wm);
+
+            DispCounter(1,M-1,'i3:');
+            for i3 = m(m > i2 & m < M)
+                DispCounter(i3,M-1,'i3:');
+                DispCounter(1,M,'i3b:');
+                for i3b = m(m > i2b & m > i3)
+                    DispCounter(i3b,M,'i3b:');
+
+                    [~,Wm] = builder_h(n, [prm(i3b) prm(i3)]);
+                    vexpt.KO = vexpt.KO.setWm(Wm);
+
+                    DispCounter(1,M,'i4:');
+                    for i4 = m(2:end)
+                        DispCounter(i4,M,'i4:');
+
+                        vexpt.nopre = vexpt.nopre.setFp(prm(i4),1);
+
+                        x = Find_pot_KO(builder_h,n,prm(i1),[prm(i2b) prm(i2)],[prm(i3b) prm(i3)],prm(i4),reps,minv,maxv);
+                        if isnan(x)
+                            continue;
+                        end
[…]

### 2016-11-18 MHCpaper/MHCsupp.tex
rewording for referees

@@
 
-The overall rate of candidate plasticity events will only set the timescale, and only the relative rates will matter.
+The overall rate of candidate plasticity events will only affect the units we measure time in.
+For the questions we are investigating, qualitative understanding of when learning is enhanced/impaired rather than detailed numerical matching of learning curves, this is completely irrelevant.
+Only the relative rates will matter.
 With this in mind, we define
@@
 %
-where $\frg$ is a continuous time Markov matrix matrix and $\I$ is the identity matrix.
+where $\frg$ is a continuous time Markov transition matrix and $\I$ is the identity matrix.
 Eventually, this will settle into the equilibrium distribution:
@@
 
-The \KO\ mice have a lower threshold for depression.
-We can model this by changing $\M\dep\wt$ to $\M\dep\ko$, which has larger off-diagonal matrix elements.
+We model the \KO\ mice by changing $\M\dep\wt$ to $\M\dep\ko$, which has larger off-diagonal matrix elements.
+This reduces the typical number of candidate depression events required to induce a change in synaptic weight, modelling the observed lower threshold for LTD \cite{McConnell2009MHCcerebellum}.
 
@@
 The monotonicity shown above means that this model cannot reproduce the key qualitative features of the experiments outlined in \autoref{sec:results}, at least as for the initial learning rate.
-This can be illustrated by looking at a special case, where $q\pot = q\dep = q$ for the wild type, $f\dep_0 = \frac{1}{2}$, $f\dep\dec = \frac{1}{2} - \Delta f$ and $f\dep\inc = \frac{1}{2} - \Delta f$.
-
-First, consider the wild-type.
-Without pre-training:
-%
-\begin{equation}\label{eq:binWTnopre}
-  \Phi = {\Delta f}\, q.
-\end{equation}
-%
-With pre-training:
-%
-\begin{equation}\label{eq:binWTpre}
-\begin{aligned}
-  \Phi &= 2{\Delta f}\, q.
-\end{aligned}
-\end{equation}
-%
-So, we see that pre-training will speed up learning, unlike what is seen in the experiment.
-This can be seen in
-\autoref{fig:sim_results}\ref{fig:binary_sim} %\autoref{fig:binary}\ref{fig:binary_learnS}
-(black curves).
-
-Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and $q\pot=q$.
-Without pre-training:
-%
-\begin{equation}\label{eq:binKOnopre}
-  \Phi = \frac{2{\Delta f}\, q}{1+\beta},
-\end{equation}
-%
-which is larger than \eqref{eq:binWTnopre} for all $\beta$, unlike what is seen in the experiments (see \modelfig).
-This is also seen in
-\autoref{fig:sim_results}\ref{fig:binary_sim} %\autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS}
-(solid curves).
-As discussed above (below \eqref{eq:multiKOnopre}), the larger value of $q\dep=q/\beta$ has two effects.
-Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
-For the serial model, we argued that the first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ to leading order.
-Here $M=2$, so the leading part of the two effects will cancel.
-The subleading effects (the normalisation of the probabilities) are responsible for the faster learning in the \KO\ models.
-
+%This can be illustrated by looking at a special case, where $q\pot = q\dep = q$ for the wild type, $f\dep_0 = \frac{1}{2}$, $f\dep\dec = \frac{1}{2} - \Delta f$ and $f\dep\inc = \frac{1}{2} - \Delta f$.
+%
+%First, consider the wild-type.
+%Without pre-training:
+%%
+%\begin{equation}\label{eq:binWTnopre}
+%  \Phi = {\Delta f}\, q.
+%\end{equation}
+%%
 %With pre-training:
 %%
-%\begin{equation}\label{eq:binKOpre}
+%\begin{equation}\label{eq:binWTpre}
 %\begin{aligned}
-%  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
-%          {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
+%  \Phi &= 2{\Delta f}\, q.
 %\end{aligned}
@@
 %%
-%As we've already ruled out this model, we will not analyse this formula any further.
-As this model is already ruled out, we will not analyse pre-training in the \KO\ models.
+%So, we see that pre-training will speed up learning, unlike what is seen in the experiment.
+%This can be seen in
+%\autoref{fig:sim_results}\ref{fig:binary_sim} %\autoref{fig:binary}\ref{fig:binary_learnS}
+%(black curves).
+%
+%Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and $q\pot=q$.
+%Without pre-training:
+%%
+%\begin{equation}\label{eq:binKOnopre}
+%  \Phi = \frac{2{\Delta f}\, q}{1+\beta},
+%\end{equation}
+%%
+%which is larger than \eqref{eq:binWTnopre} for all $\beta$, unlike what is seen in the experiments (see \modelfig).
+%This is also seen in
+%\autoref{fig:sim_results}\ref{fig:binary_sim} %\autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS}
+%(solid curves).
+%As discussed above (below \eqref{eq:multiKOnopre}), the larger value of $q\dep=q/\beta$ has two effects.
+%Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
+%For the serial model, we argued that the first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ to leading order.
+%Here $M=2$, so the leading part of the two effects will cancel.
+%The subleading effects (the normalisation of the probabilities) are responsible for the faster learning in the \KO\ models.
+%
+%%With pre-training:
+%%%
+%%\begin{equation}\label{eq:binKOpre}
+%%\begin{aligned}
+%%  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
+%%          {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
+%%\end{aligned}
+%%\end{equation}
+%%%
+%%As we've already ruled out this model, we will not analyse this formula any further.
+%As this model is already ruled out, we will not analyse pre-training in the \KO\ models.
 
@@
 
+We see that, for most parameter values, such models do not reproduce the observed learning impairment with enhanced plasticity.
+In some cases a very small learning impairment was seen, but these occurred at extreme parameter values where floating point precision could become an issue.
+
 \begin{figure}
@@
   We see that this quantity is usually negative, whereas it was positive in the experiments.
+  The exceptions are very small, and tend to occur at extreme parameter values where rounding errors may occur.
   }\label{fig:bal}

### 2016-11-19 MHCpaper/MHCsupp.tex
changing precision of xpot fitting

@@
 We model the \KO\ mice by changing $\M\dep\wt$ to $\M\dep\ko$, which has larger off-diagonal matrix elements.
-This reduces the typical number of candidate depression events required to induce a change in synaptic weight, modelling the observed lower threshold for LTD \cite{McConnell2009MHCcerebellum}.
+This makes it easier to induce a decrease in synaptic weight, modelling the observed lower threshold for LTD \cite{McConnell2009MHCcerebellum}.
 

### 2016-11-22 MHCpaper/MHCsupp.tex
scan figs

@@
 We see that, for most parameter values, such models do not reproduce the observed learning impairment with enhanced plasticity.
-In some cases a very small learning impairment was seen, but these occurred at extreme parameter values where floating point precision could become an issue.
+In some cases a very small learning impairment was seen ($\CO(10^{-5})$), but these occurred at extreme parameter values where the tolerances used in fitting the potentiation parameter could have a noticeable impact.
 
@@
   \begin{myenuma}
-    \item\aligntop{\includegraphics[width=0.29\linewidth]{pooled_deponly_scan.eps}}\label{fig:pooled_bal}
+    \item\aligntop{\includegraphics[width=0.29\linewidth]{pooled_scan.eps}}\label{fig:pooled_bal}
     \item\aligntop{\includegraphics[width=0.29\linewidth]{cascade_scan.eps}}\label{fig:cascade_bal}
@@
   We see that this quantity is usually negative, whereas it was positive in the experiments.
-  The exceptions are very small, and tend to occur at extreme parameter values where rounding errors may occur.
+  The exceptions are very small, $\CO(10^{-5})$, and tend to occur at extreme parameter values where the accuracy of fitting the potentiation parameter could become an issue.
   }\label{fig:bal}

### 2016-11-22 Code/VOR/Plots/PooledScanTop.m

@@ -10,12 +10,15 @@ ncomps=ScanPooledNum(paramvals,ns,'Top',reps);
 txopts={'Interpreter','latex','FontSize',20};
 ax=gca;
 bh=bar(ax, ns-1, ncomps','stacked');
+ax.FontSize=16;
 xlabel('$\#$ of synapses in pool', txopts{:});
 ylabel('max/min $\{\dot{L}_{\mathrm{WT}}(0) - \dot{L}_{\mathrm{K}^b\mathrm{D}^{b-/-}}(0)\}$', txopts{:});
 title('Pooled resource model, no pre', txopts{:});
 bh(1).FaceColor='none';
 bh(1).EdgeColor='none';
 bh(2).FaceColor=[0 0.4470 0.7410];
-ax.YAxis.Scale = 'log';
+xlim([ns(1)-1.5 ns(end)-0.5]);
+ylim([1.1*min(ncomps(:)) 0.02]);
+% ax.YAxis.Scale = 'log';
 %%
 print('pooled_scan.eps','-depsc');

### 2016-11-28 MHCpaper/MHCsupp.tex
Talked about longer training after pre

@@
 %}
+\author{}
 \date{}
@@
   \begin{tabular}{lll}
-      \item\label{fig:seial_sim}
+      \item\label{fig:serial_sim}
       \aligntop{\includegraphics[width=0.24\linewidth]{serial.svg}}
@@
   \end{myenuma}
-  \caption[Simulation results for various models]{(\ref{fig:seial_sim}-\ref{fig:nonuni_sim}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (solid) and after gain-decrease pre-training (dashed) for wild-type (black) and \KO\ (red) models.
+  \caption[Simulation results for various models]{(\ref{fig:serial_sim}-\ref{fig:nonuni_sim}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (solid) and after gain-decrease pre-training (dashed) for wild-type (black) and \KO\ (red) models.
   (\ref{tab:params}) Parameters used for simulations.
@@
 The results of simulations of the serial model can be seen in
-\autoref{fig:sim_results}\ref{fig:seial_sim}.
+\autoref{fig:sim_results}\ref{fig:serial_sim}.
 %\autoref{fig:multistate_weak}, \autoref{fig:multistate_med} and \autoref{fig:multistate_strong}.
@@
 
+\section{Long term effects of pre-training}\label{sec:longterm}
+
+The fact that after pre-training, gain-increase learning is significantly faster in the \KO\ models than in the wild-type, as discussed in the \hyperref[sec:results]{previous section}, is true of the initial learning rate.
+With a long enough duration of training, memory of the initial conditions should be erased and the final learning outcome should be the same with or without pre-training.
+In \autoref{fig:longterm} we show the result of longer training durations on the three models that were not ruled out on the basis of the initial learning rate, the serial, cascade and nonuniform multistate models.
+For all of these models, we see that the wild-type models eventually catch up and overtake the \KO\.
+The initial learning enhancement is a transient effect, and at later times it turns into a learning impairment.
+
+\begin{figure}
+  \centering
+  \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.29\linewidth]{serial_longterm.eps}}\label{fig:serial_longterm}
+  \item\aligntop{\includegraphics[width=0.29\linewidth]{cascade_longterm.eps}}\label{fig:cascade_longterm}
+  \item\aligntop{\includegraphics[width=0.29\linewidth]{nonuni_longterm.eps}}\label{fig:nonuni_longterm}
+  \end{myenuma}
+  \caption{ Simulation of training for long times after pre-training shows that it erases the effect of pre-training.
+  These simulations were performed with the same parameters as in \autoref{fig:sim_results}\ref{tab:params}, but with much longer training times, for the (\ref{fig:serial_longterm}) serial, (\ref{fig:cascade_longterm} cascade and (\ref{fig:nonuni_longterm}) non-uniform multistate models.
+  Compare with \autoref{fig:sim_results}\ref{fig:serial_sim},\ref{fig:cascade_sim},\ref{fig:nonuni_sim}.
+  }\label{fig:longterm}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Model of non-specific CF stimulation}\label{sec:chr2}
-
-Here, we model the effects of non-specific CF stimulation, as in \chrfig.
-It is necessary to also model the ``other'' population of synapses that compensate for the change in baseline synaptic weight of the synapses considered up to now, either due to enhanced plasticity or non-specific CF stimulation.
-
-To be concrete, we model these other synapses as described in \compfig.
-We set the relative rate of depression events for these synapses during training to $f\dep = f\dep\norm$, \ie the baseline value, so that they are unaffected by training.
-Here we set the rate of plasticity events causing forgetting of the CF stimulation in these synapses is the same as the rate of plasticity events in the original synapses during training, although similar results could be acheived with other choices.
-For both sets of synapses during non-specific CF stimulation, we set $f\dep > f\dep\norm$ to model the increased rate of depression.
-The values of all parameters used can be seen in \autoref{fig:sim_ChR2}\ref{tab:params_ChR2}.
-
-In \autoref{fig:sim_ChR2}\ref{fig:seial_ChR2}--\ref{fig:nonuni_ChR2} we see examples of simulations of all of the models we have considered.
-We see that every single one of them is capable of reproducing the experimental results of \chrfig, for suitable parameter choices.
-This is because this intervention has the effect of depleting the population of synapses available for depression \emph{without} the competing effect of enhanced plasticity.
-Therefore the depletion effect is stronger (than 0) and learning is impaired.
-This means that these experiments cannot rule out any of the models.
-However these experiments do agree with the idea of impaired learning due to depletion of the population of labile synapses.
 
@@
   \begin{tabular}{lll}
-      \item\label{fig:seial_ChR2}
+      \item\label{fig:serial_ChR2}
       \aligntop{\includegraphics[width=0.24\linewidth]{serial.svg}}
@@
   \end{myenuma}
-  \caption[Simulation results for various models]{(\ref{fig:seial_sim}-\ref{fig:nonuni_sim}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (black) and after non-specific CF stimulation (cyan) for wild-type models.
+  \caption[Simulation results for various models]{(\ref{fig:serial_ChR2}-\ref{fig:nonuni_ChR2}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (black) and after non-specific CF stimulation (cyan) for wild-type models.
   (\ref{tab:params}) Parameters used for simulations.
@@
 \end{figure}
+
+Here, we model the effects of non-specific climbing fibre stimulation, as in \chrfig.
+It is necessary to also model the ``other'' population of synapses that compensate for the change in baseline synaptic weight of the synapses considered up to now, either due to enhanced plasticity or non-specific CF stimulation.
+
+To be concrete, we model these other synapses as described in \compfig.
+We set the relative rate of depression events for these synapses during training to $f\dep = f\dep\norm$, \ie the baseline value, so that they are unaffected by training.
+Here we set the rate of plasticity events causing forgetting of the CF stimulation in these synapses is the same as the rate of plasticity events in the original synapses during training, although similar results could be acheived with other choices.
+For both sets of synapses during non-specific CF stimulation, we set $f\dep > f\dep\norm$ to model the increased rate of depression.
+The values of all parameters used can be seen in \autoref{fig:sim_ChR2}\ref{tab:params_ChR2}.
+
+In \autoref{fig:sim_ChR2}\ref{fig:serial_ChR2}--\ref{fig:nonuni_ChR2} we see examples of simulations of all of the models we have considered.
+We see that every single one of them is capable of reproducing the experimental results of \chrfig, for suitable parameter choices.
+This is because this intervention has the effect of depleting the population of synapses available for depression \emph{without} the competing effect of enhanced plasticity.
+Therefore the depletion effect is stronger (than 0) and learning is impaired.
+This means that these experiments cannot rule out any of the models.
+However these experiments do agree with the idea of impaired learning due to depletion of the population of labile synapses.
 

### 2016-11-28 MHCpaper/MHCsupp.tex
Paragraph about balanced pot ready
temp balanced cascade fig

@@
 
-We see that, for most parameter values, such models do not reproduce the observed learning impairment with enhanced plasticity.
-In some cases a very small learning impairment was seen ($\CO(10^{-5})$), but these occurred at extreme parameter values where the tolerances used in fitting the potentiation parameter could have a noticeable impact.
+%We see that, for most parameter values, such models do not reproduce the observed learning impairment with enhanced plasticity.
+%In some cases a very small learning impairment was seen ($\CO(10^{-5})$), but these occurred at extreme parameter values where the tolerances used in fitting the potentiation parameter could have a noticeable impact.
+
+We see that, for all parameter values, such models do not reproduce the observed learning impairment with enhanced plasticity.
+The normal baseline oculomotor performance must have some other explanation, such as those outlined in \compfig.
 

### 2016-11-30 Code/VOR/Plots/CascNuniScan.m
sentence about long term

@@ -8,7 +8,7 @@ paramvals = (0.05:0.1:0.95);
 ns = 4:10;
 reps=20;
 if cn
-    ns = 14:2:16;
+    ns = 4:2:16;
     paramvals = paramvals/2;
 end
 %%
@@ -18,7 +18,8 @@ ncomps2=ScanCNNum(paramvals,ns,reps,cn,'StepTolerance',1e-12,'ConstraintToleranc
 %plot
 txopts={'Interpreter','latex','FontSize',20};
 ax=gca;
-bh=bar(ax, ns, ncomps(:,1:6)','stacked');
+bh=bar(ax, ns, ncomps','stacked');
+% bh=bar(ax, ns, ncomps(:,1:6)','stacked');
 ax.FontSize=16;
 bh(1).FaceColor='none';
 bh(1).EdgeColor='none';

### 2016-12-03 MHCpaper/MHCsupp.tex
sentence about long term
main figure references
surya's edits

@@
 %\newcommand{\lmin}{_{\text{min}}}
-\newcommand{\modelfig}[1][A]{Figure 3#1 of the main text}
-\newcommand{\datafig}[1][B]{Figure 2#1 of the main text}
-\newcommand{\chrfig}[1][B]{Figure 4#1 of the main text}
-\newcommand{\compfig}[1][A]{Figure 4 -- figure supplement 2#1}
+\newcommand{\modelfig}[1][A]{Figure 4#1 of the main text}
+\newcommand{\datafig}[1][B]{Figure 3#1 of the main text}
+\newcommand{\chrfig}[1][B]{Figure 2#1 of the main text}
+\newcommand{\compfig}[1][A]{Figure 2 -- figure supplement 2#1}
 \newcommand{\basefig}{Figure 1 -- figure supplement 2}
@@
 \begin{itemize}
-  \item There synapses are identical and have $M$ internal functional states, where the states of different synapses are independent of each other.
+  \item The synapses are identical and have $M$ internal functional states, where the states of different synapses are independent of each other.
 %  \item The synapses that are eligible for plasticity are chosen randomly.
@@
 
-For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation in the wild-type and \KO\ models.
+For the serial, multistate and two state models, we will use the same value for $q$, the transition rate between adjacent states, for potentiation in the wild-type and \KO\ models.
 We will use a larger value for $q$ for depression in the \KO\ models than in the wild-type.
 
-For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation in the wild-type and \KO\ models.
+For the cascade model, we will use the same value for the parameter $x$ (which controls the ratio of adjacent transition rates, see \cite{Fusi2005cascade}) for potentiation in the wild-type and \KO\ models.
 We will use a larger value for $x$ for depression in the \KO\ models than in the wild-type.
@@
 Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed and replenished when this is reversed, as shown in \autoref{fig:models}\ref{fig:pooled_model}.
-We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single compound synapse.
+We can avoid going beyond the independent synapse model by modelling the pool of synapses that share the resource as a single compound synapse.
 
@@
 However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
+We can group those states of the pool of synapses that have the same total number of potentiated synapses into a single state of the equivalent compound synapse.
 The transition network will then have a multistate topology (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
@@
 
-The Markov process is lumpable \wrt this partition of states (see \cite{kemeny1960finite,burke1958markovian,Ball1993Lumpability}).
+The Markov process is lumpable when partitioning states by total number of potentiated synapses (see \cite{kemeny1960finite,burke1958markovian,Ball1993Lumpability}).
 The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$,
@@
 These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
-\emph{We will not study gain-decrease learning}, as its mechanisms are not fully understood, and known to be different to to gain-increase.
+\emph{We will not study gain-decrease learning}, as its mechanisms are not fully understood, and known to be different to gain-increase.
 We are only modelling the effect of gain-decrease training on \emph{these} synapses.
@@
 The initial learning enhancement is a transient effect, and at later times it turns into a learning impairment.
+This is consistent with an extrapolation of the trend in \datafig.
 
@@
 We see that, for all parameter values, such models do not reproduce the observed learning impairment with enhanced plasticity.
-The normal baseline oculomotor performance must have some other explanation, such as those outlined in \compfig.
+The normal baseline oculomotor performance must have some other explanation, such as those outlined in \compfig[].
 

### 2016-12-03 Code/VOR/Scan/ScanCNNum.m

@@ -13,7 +13,7 @@ for n = ns
     pcomps = ScanCNtop(ranges, n, reps, useCnotN, varargin{:});
 %     pcomps = ScanCNtop2(ranges, n, useCnotN, varargin{:});
     comps(1,ns==n) = max(pcomps(~isnan(pcomps)));
-    comps(2,ns==n) = min(pcomps(~isnan(pcomps)));
+    comps(2,ns==n) = min(pcomps(~(isnan(pcomps) | isinf(pcomps))));
 end
 DispCounter(ns(end)+1,ns(end),'n:');
 end

### 2016-12-03 MHCpaper/MHCsupp.tex
surya's edits

@@
 For the non-uniform multistate model, we see that the \KO\ models are slower than wild-type without pre-training, as seen experimentally.
-This seems to be because the enhanced depression results in a greater fraction of synapses begin in states of weaker synaptic weight %(see \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_KO})
+This seems to be because the enhanced depression results in a greater fraction of synapses starting in states of weaker synaptic weight %(see \autoref{fig:nonuni}\ref{fig:nonuni_eq}\ref{fig:nonuni_eq_KO})
 where the depressing transitions have lower probability (see \autoref{fig:models}\ref{fig:nonuni_model}).

### 2016-12-03 MHCpaper/MHCsupp.tex
tweaked order of lumping

@@
 This compound synapse would seem to have $2^P$ internal states.
-However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
-We can group those states of the pool of synapses that have the same total number of potentiated synapses into a single state of the equivalent compound synapse.
+However, we need only keep track of the number of potentiated synapses, not their identity.
+We can group those states of the pool of synapses that have the same total number of potentiated synapses into a single state of the equivalent compound synapse, leaving $M=P+1$ states.
 The transition network will then have a multistate topology (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:

### 2016-12-16 MHCpaper/MHCsupp.tex
pooled fig fixed

@@
     \item\aligntop{\includegraphics[height=2.5cm]{cascade.svg}}\label{fig:cascade_model}\\[2.5cm]
-    \item\aligntop{\includegraphics[height=2.3cm]{pooled.svg}}\label{fig:pooled_model}&
+    \item\aligntop{\includegraphics[height=2.3cm]{pooled_deponly.svg}}\label{fig:pooled_model}&
     \item\aligntop{\includegraphics[height=2.3cm]{multistate_nonuni.svg}}\label{fig:nonuni_model}
@@
 We will use larger values for depression in the \KO\ models than in the wild-type.
-we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
+We consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
 This is done by setting $q\pot\lmax=q\pot\lmin$.
@@
 For the other models, a numerical analysis is required, as shown in \autoref{fig:bal}.
-\hypertarget{par:bal_scan}{To} see if learning is impaired with enhanced plasticity each relevant parameter was scanned over 10 values from 0.05 to 0.95, rejecting values that do not respect the inequalities:
-$0 \leq f\dep\norm < f\dep\inc \leq 1$, for the pooled resource model $0 \leq q\dep_\text{min,WT} < q\dep_\text{min,\KO}, q\dep_\text{max,WT} < q\dep_\text{max,\KO}$ and $0 \leq q\pot\wt$, for the cascade model $0 \leq x\dep\wt < x\dep\ko \leq \frac{1}{2}$ and $0 \leq x\pot\wt \leq \frac{1}{2}$, and for the nonuniform multistate model $0 \leq x\dep\wt < x\dep\ko \leq 1$ and $0 \leq x\pot\wt \leq 1$.
+\hypertarget{par:bal_scan}{To} see if learning is impaired with enhanced plasticity each relevant parameter was scanned over 10 values from 0.05 to 0.95 (or from 0.05 to 0.5 for the $x$ parameter of the cascade model), rejecting values that do not respect the inequalities:
+$0 \leq f\dep\norm < f\dep\inc \leq 1$, for the pooled resource model $0 \leq q\dep_\text{min,WT} < q\dep_\text{min,\KO}, q\dep_\text{max,WT} < q\dep_\text{max,\KO} \leq 1$ and $0 \leq q\pot\wt \leq 1$, for the cascade model $0 \leq x\dep\wt < x\dep\ko \leq \frac{1}{2}$ and $0 \leq x\pot\wt \leq \frac{1}{2}$, and for the nonuniform multistate model $0 \leq x\dep\wt < x\dep\ko \leq 1$ and $0 \leq x\pot\wt \leq 1$.
 In each case, the \KO\ potentiation parameter, $q\pot\ko$ for the pooled resource model and $x\pot\ko$ for the others, was adjusted to match the baseline mean synaptic weight.

### 2016-12-16 MHCpaper/MHCsupp.tex
long term figs for wrong models in supp

@@
 With a long enough duration of training, memory of the initial conditions should be erased and the final learning outcome should be the same with or without pre-training.
-In \autoref{fig:longterm} we show the result of longer training durations on the three models that were not ruled out on the basis of the initial learning rate, the serial, cascade and nonuniform multistate models.
+In \autoref{fig:longterm} we show the result of longer training durations on all models considered in \autoref{sec:results}.
 For all of these models, we see that the wild-type models eventually catch up and overtake the \KO\.
@@
   \item\aligntop{\includegraphics[width=0.29\linewidth]{serial_longterm.eps}}\label{fig:serial_longterm}
+  \item\aligntop{\includegraphics[width=0.29\linewidth]{binary_longterm.eps}}\label{fig:binary_longterm}
+  \item\aligntop{\includegraphics[width=0.29\linewidth]{multistate_longterm.eps}}\label{fig:multistate_longterm} \\
+  \item\aligntop{\includegraphics[width=0.29\linewidth]{pooled_longterm.eps}}\label{fig:pooled_longterm}
   \item\aligntop{\includegraphics[width=0.29\linewidth]{cascade_longterm.eps}}\label{fig:cascade_longterm}
@@
   \caption{ Simulation of training for long times after pre-training shows that it erases the effect of pre-training.
-  These simulations were performed with the same parameters as in \autoref{fig:sim_results}\ref{tab:params}, but with much longer training times, for the (\ref{fig:serial_longterm}) serial, (\ref{fig:cascade_longterm} cascade and (\ref{fig:nonuni_longterm}) non-uniform multistate models.
-  Compare with \autoref{fig:sim_results}\ref{fig:serial_sim},\ref{fig:cascade_sim},\ref{fig:nonuni_sim}.
+  These simulations were performed with the same parameters as in \autoref{fig:sim_results}\ref{tab:params}, but with much longer training times, for the (\ref{fig:serial_longterm}) serial, (\ref{fig:binary_longterm}) two-state, (\ref{fig:multistate_longterm}) multistate, (\ref{fig:pooled_longterm}) pooled resource, (\ref{fig:cascade_longterm} cascade and (\ref{fig:nonuni_longterm}) non-uniform multistate models.
+  Compare with \autoref{fig:sim_results}.
   }\label{fig:longterm}

### 2016-12-23 MHCpaper/MHCsupp.tex
change of format

@@
 %Preamble:
+\usepackage[inactive,tightpage]{preview}
 \usepackage{a4wide}
@@
 \captionsetup{labelformat=supp}
-% ----------------------------------------------------------------
+\renewenvironment{myenumA}{\begin{inparaenum}[\bfseries A]}{\end{inparaenum}}
+% ---------------------------------------------------------------
 % New commands etc.
@@
 \begin{figure}
+\begin{preview}
  \begin{center}
-   \begin{myenuma}
+   \begin{myenumA}
   \begin{tabular}{l@{\hspace{0.05\linewidth}}l}
@@
   \end{tabular}
-   \end{myenuma}
+   \end{myenumA}
  \end{center}
+\end{preview}
   \caption[Transition probabilities for different models]{{Transition probabilities for different models.}
@@
 \begin{figure}[p]
-\begin{myenuma}
+\begin{preview}
+\begin{myenumA}
   \begin{tabular}{lll}
@@
   \end{tabular}}
-  \end{myenuma}
+  \end{myenumA}
+\end{preview}
   \caption[Simulation results for various models]{(\ref{fig:serial_sim}-\ref{fig:nonuni_sim}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (solid) and after gain-decrease pre-training (dashed) for wild-type (black) and \KO\ (red) models.
@@
 %\newcommand{\arrangefig}[1]{\begin{center}
-% \begin{myenuma}
+% \begin{myenumA}
 %  \item\aligntop{\includegraphics[width=7cm]{#1_learn.eps}}\label{fig:#1_learn}
@@
 %                  \end{myenumi}
-% \end{myenuma}
+% \end{myenumA}
 % \end{center}}
@@
 The lack of monotonicity shown above allows this model to reproduce the key qualitative features of the experiments outlined in \autoref{sec:results}.
-This can be illustrated by looking at a special case, where $q\pot = q\dep = q$ for the wild type, $f\dep_0 = \frac{1}{2}$, $f\dep\dec = \frac{1}{2} - \Delta f$ and $f\dep\inc = \frac{1}{2} - \Delta f$.
+This can be illustrated by looking at a special case, where $q\pot = q\dep = q$ for the wild type, $f\dep_0 = \frac{1}{2}$, $f\dep\dec = \frac{1}{2} - \Delta f$ and $f\dep\inc = \frac{1}{2} + \Delta f$.
 
@@
 %
-Once again, we see that pre-training will speed up learning when $\Delta f$ is small, whereas, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning.
+Once again, we see that pre-training will speed up learning when $\Delta f$ is small, whereas, if $\Delta f$ is close to $\half$, pre-training will initially slow down learning.
 
@@
 \begin{figure}
+\begin{preview}
  \begin{center}
- \begin{myenuma}
+ \begin{myenumA}
   \item\aligntop{\includegraphics[width=0.45\linewidth]{multistate_betastar.eps}}\label{fig:multistate_betastar}
   \item\aligntop{\includegraphics[width=0.45\linewidth]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
- \end{myenuma}
+ \end{myenumA}
  \end{center}
-  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the \KO\ models have impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
+\end{preview}
+  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions
+  (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the \KO\ models have impaired learning,
+  and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
 \end{figure}
@@
 \begin{figure}
+\begin{preview}
   \centering
@@
   \includegraphics[width=0.4\linewidth]{pooled_deponly_scan.eps}
+\end{preview}
   \caption{Difference in initial learning rate in wild-type models, without and with gain-decrease pre-training.
@@
 % \begin{center}
-% \begin{myenuma}
+% \begin{myenumA}
 %  \item\aligntop{\includegraphics[width=0.28\linewidth]{serial_fit_learnS.eps}}\label{fig:serial_fit}
@@
 %  \end{tabularn}}
-% \end{myenuma}
+% \end{myenumA}
 % \end{center}
@@
 \begin{figure}
+\begin{preview}
   \centering
-  \begin{myenuma}
+  \begin{myenumA}
   \item\aligntop{\includegraphics[width=0.29\linewidth]{serial_longterm.eps}}\label{fig:serial_longterm}
@@
   \item\aligntop{\includegraphics[width=0.29\linewidth]{nonuni_longterm.eps}}\label{fig:nonuni_longterm}
-  \end{myenuma}
+  \end{myenumA}
+\end{preview}
   \caption{ Simulation of training for long times after pre-training shows that it erases the effect of pre-training.
@@
 
-\begin{figure}[p]
-\begin{myenuma}
+\begin{figure}
+\begin{preview}
+\begin{myenumA}
   \begin{tabular}{lll}
@@
   \end{tabular}}
-  \end{myenuma}
+  \end{myenumA}
+\end{preview}
   \caption[Simulation results for various models]{(\ref{fig:serial_ChR2}-\ref{fig:nonuni_ChR2}) Simulation results for various models, showing decrease in mean synaptic weight over time during gain increase training from the normal state (black) and after non-specific CF stimulation (cyan) for wild-type models.
-  (\ref{tab:params}) Parameters used for simulations.
+  (\ref{tab:params_ChR2}) Parameters used for simulations.
   For the serial, two state and multistate models, the plasticity parameter listed is the transition probability between adjacent states.
@@
 \begin{figure}
+\begin{preview}
   \centering
-  \begin{myenuma}
+  \begin{myenumA}
     \item\aligntop{\includegraphics[width=0.29\linewidth]{pooled_scan.eps}}\label{fig:pooled_bal}
@@
     \item\aligntop{\includegraphics[width=0.29\linewidth]{nonuni_scan.eps}}\label{fig:nonuni_bal}
-  \end{myenuma}
+  \end{myenumA}
+\end{preview}
   \caption{Difference in initial learning rate in wild-type and \KO\ models, without gain-decrease pre-training.
@@
   Here, we show the maximum and minimum values of this quantity when scanning all 7/5 relevant parameters (see \protect\hyperlink{par:bal_scan}{text}).
-  We see that this quantity is usually negative, whereas it was positive in the experiments.
-  The exceptions are very small, $\CO(10^{-5})$, and tend to occur at extreme parameter values where the accuracy of fitting the potentiation parameter could become an issue.
+  We see that this quantity is always negative, whereas it was positive in the experiments.
   }\label{fig:bal}

### 2017-02-05 MHCpaper/MHCsupp.tex
ko name
typos

@@
 \newcommand{\wt}{_{\text{WT}}}
-\newcommand{\ko}{_{\text{K$^\mathrm{b}$D$^\mathrm{b}$-/-}}}
-\newcommand{\KO}{K$^\mathrm{b}$D$^{\mathrm{b}-/-}$}
+\newcommand{\ko}{_{\text{\KO}}}
+\newcommand{\KO}{DKO}
 %\newcommand{\ko}{_{\text{MHC}}}
@@
                                           & 0.5 & 0.9 & 0.1 & 20 \\%&\label{tr:pooled_deponly}\\
-    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.522 & 0.63 & 0.002  & 200  \\%&\label{tr:cascade_fit} \\
+    Cascade  & 14 & $0.386$  & $0.398$  & $0.466$  & 0.478 & 0.63 & 0.002  & 200  \\%&\label{tr:cascade_fit} \\
     Non-uni.\ & 12 & $0.4$    & $0.4$    & $0.53$   & 0.5 & 0.7 & 0.1 & 500  \\%&\label{tr:nonuni_fit} \\
@@
 \begin{preview}
-  \centering
+  \begin{center}
   % Requires \usepackage{graphicx}
   \includegraphics[width=0.4\linewidth]{pooled_deponly_scan.eps}
+  \end{center}
 \end{preview}
@@
     Pooled res.\ & 7  & $0.08$  & $[0.006,0.6]$ & 0.5   & 0.9  & 0.99   & 20  \\
-    Cascade      & 14 & $0.386$ & $0.398$       & 0.522 & 0.63 & 0.99   & 200  \\
+    Cascade      & 14 & $0.386$ & $0.398$       & 0.478 & 0.63 & 0.99   & 200  \\
     Non-uni.\    & 12 & $0.4$   & $0.4$         & 0.5   & 0.7  & 0.99   & 500  \\

### 2018-01-20 Notes/laplace.tex
chernoff etc

@@
 %matrices
-\newcommand{\inv}{^{-1}}
-\newcommand{\dg}{^\mathrm{dg}}
-\newcommand{\trans}{^\mathrm{T}}
-\newcommand{\I}{\mathbf{I}}
 %vec of ones
@@
 \newcommand{\syn}{\vec{w}}
-\newcommand{\synid}{\syn_\text{ideal}}
+\newcommand{\synid}{\syn_\text{id}}
 \DeclareMathOperator{\SNR}{SNR}
 \DeclareMathOperator{\snr}{SNR}
+\DeclareMathOperator{\NNR}{NNR}
+\DeclareMathOperator{\nnr}{NNR}
 \newcommand{\snrb}{\overline{\snr}}
@@
 \newcommand{\rh}{\hat{r}}
-\newcommand{\CI}{\mathcal{I}}
-\newcommand{\CS}{\mathcal{S}}
 \newcommand{\CA}{\mathcal{A}}
 \newcommand{\CB}{\mathcal{B}}
-\newcommand{\comp}{^\mathrm{c}}
 %avoid double superscripts
@@
   \begin{aligned}
-  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \operatorname{NNR} } },
+  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \NNR } },
   \quad\text{where } &
@@
     \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
-    \operatorname{NNR} &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
+    \NNR &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
   \end{aligned}
@@
 
+There are other measures of memory one could consider, \eg asymptotic error exponents.
+For the Neyman-Pearson approach to hypothesis testing, the error exponent is the KL divergence:
+%
+\begin{equation*}
+  D_\text{KL}\prn{P_0 \middle\Vert P_1} = \ln\nnr + \frac{1 + \snr^2}{2\nnr^2} - \frac{1}{2}.
+\end{equation*}
+%
+For Bayesian hypothesis testing, the error exponent is the Chernoff distance:
+%
+\begin{equation*}
+  D^*\prn{P_0 \middle\Vert P_1} = \max_\alpha \brc{
+    \frac{1}{2} \ln\brk{\frac{1 + \alpha(\nnr^2-1)}{\nnr^{2\alpha}}} + \frac{\alpha(1-\alpha)\snr^2}{2[1 + \alpha(\nnr^2-1)]}
+  }.
+\end{equation*}
+%
+
 The formulae above assumed that we know the time between storage and recognition.
@@
 \begin{equation}\label{eq:laplacedef}
-  A(s) = \intd[_0^\infty]{t} \e^{-st} \snr(t).
-  \qquad
-  \snrb(\tau) = \frac{A(1/\tau)}{\tau},
+  A(s) = \intd[_0^\infty]{t} \e^{-st} \snr(t),
+  \qquad
+  \snrb(\tau) = \frac{A(1/\tau)}{\tau}.
 \end{equation}
@@
 \begin{aligned}
-  \var(\syn_\text{ideal}\cdt\syn(t))
-    &= \sum_{\alpha\beta} \av{\syn^\alpha_\text{ideal}\syn^\alpha(t) \syn^\beta_\text{ideal}\syn^\beta(t)}
-    - \prn{\sum_\alpha \av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}}^2 \\
-    &= \sum_{\alpha} \av{(\syn^\alpha_\text{ideal})^2(\syn^\alpha(t))^2}
-    + \sum_{\alpha\neq\beta} \av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}\av{\smash{\syn^\beta}_\text{ideal}\smash{\syn^\beta}(t)}
-    \\&\phantom{= \sum_{\alpha}  \av{(\syn^\alpha_\text{ideal})^2(\syn^\alpha(t))^2}}
-    - \prn{\sum_\alpha \av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}}^2  \\
+  \var(\synid\cdt\syn(t))
+    &= \sum_{\alpha\beta} \av{\synid^\alpha \syn^\alpha(t) \synid^\beta \syn^\beta(t)}
+    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2 \\
+    &= \sum_{\alpha} \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}
+    + \sum_{\alpha\neq\beta} \av{\synid^\alpha\syn^\alpha(t)}\av{\smash{\synid^\beta}\smash{\syn^\beta}(t)}
+    %\\&\phantom{= \sum_{\alpha}  \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}}
+    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2  \\
     &= N\av{1}
-    + N(N-1)\av{\syn^1_\text{ideal}\syn^1(t)}^2
-    - N^2\av{\syn^1_\text{ideal}\syn^1(t)}^2 \\
-    &= N(1-\av{\syn^1_\text{ideal}\syn^1(t)}^2),
+    + N(N-1)\av{\synid^1 \syn^1(t)}^2
+    - N^2\av{\synid^1 \syn^1(t)}^2 \\
+    &= N(1-\av{\synid^1 \syn^1(t)}^2),
 \end{aligned}
@@
      - f\pot S\trans_{ijkl} + f\dep S_{ijkl}, \\
-  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
+  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\dep_{kl}} &=
      (f\dep)^2 (R_{ijkl} + R\trans_{ijkl})

### 2018-03-12 Notes/ContTime.tex
typo
markov bib file in notes
laplace envelope notes
laplace envelope w/o topology constraints & homeostatic
chernoff etc

@@
 \newcommand{\rh}{\hat{r}}
-\newcommand{\CI}{\mathcal{I}}
-\newcommand{\CS}{\mathcal{S}}
 \newcommand{\CA}{\mathcal{A}}
 \newcommand{\CB}{\mathcal{B}}
-\newcommand{\comp}{^\mathrm{c}}
 % ----------------------------------------------------------------
@@
     &\pm \eq_g \prn{\brk{\e^{rt\frg} \w}_h - \brk{\e^{rt\frg} \w}_g} \\
-    &+ f^\pm \sum_{ab} F_{ab} (\eq\enc\uv^a) \vv^a_g (\uv^b_h-\uv^b_g) (\vv^a\w) \\
+    &+ f^\pm \sum_{ab} F_{ab} (\eq\enc\uv^a) \vv^a_g (\uv^b_h-\uv^b_g) (\vv^b\w) \\
     &+ \mu^\pm_{gh} - \mu^\pm_{gg}.
@@
 \begin{equation}\label{eq:alfund}
-  \pdiff{(\mathbf{a}\fund(s))_i}{\pib_k} = (\mathbf{a}\fund(s)\onev) \fund_{ki}(s) = \frac{\tau}{1+s\tau} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
+  \pdiff{(\mathbf{a}\fund(s))_i}{\pib_k} = -(\mathbf{a}\fund(s)\onev) \fund_{ki}(s) = -\frac{\tau}{1+s\tau} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
 \end{equation}
@@
   \snr(t_0) = \sqrt{N}\env(t_0) = \frac{\sqrt{N}\,\e^{s_0t_0-1}}{C'(s_0)}
-   = \frac{\sqrt{N}\,B(s)\e^{s_0t_0-1}}{t_0},
+   = \frac{\sqrt{N}\,B(s_0)\e^{s_0t_0-1}}{t_0},
 \end{equation}

### 2018-03-16 Paper/SynapsePaper.tex
new pnas doc class

@@
-%% PNAStwoS.tex
-%% For two column PNAS articles
-%% Version: Apr 15, 2008
-
-%% BASIC CLASS FILE
-\documentclass{pnastwo}
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
+
+\templatetype{pnasresearcharticle} % Choose template
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} = Template for a one-column mathematics article
+% {pnasinvited} = Template for a PNAS invited submission
 
 %% ADDITIONAL OPTIONAL STYLE FILES
-%\usepackage{graphicx}
-\usepackage{ifpdf}
 %\usepackage[hypcap]{caption}%links to figures
@@
 %\usepackage{graphics} for finding documentation only
-%
+\pdfsuppresswarningpagegroup=1
 \usepackage{epstopdf}
@@
 %
-%\newcommand{\inkscapedir}{C:/Program Files/Inkscape/}
-%\IfFileExists{"C:/Program Files (x86)/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Program Files (x86)/Inkscape/}}{}
-\IfFileExists{"C:/Program Files/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Program Files/Inkscape/}}{}
+\IfFileExists{"C:/Progra\string~1/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Progra\string~1/Inkscape/}}{}
 \IfFileExists{"/Applications/Inkscape.app/Contents/Resources/bin/inkscape"}{\newcommand{\inkscapedir}{/Applications/Inkscape.app/Contents/Resources/bin/}}{}
-%
-\ifpdf
- \epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{%
- "\inkscapedir inkscape" -f #1 -C -A \OutputFile
- }
-\else
- \DeclareGraphicsRule{.png}{eps}{.bb}{}
-\fi
+\providecommand{\inkscapedir}{}
+%
+\epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{"\inkscapedir inkscape" -f #1 -C -A \OutputFile}
 %
@@
 \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
-\newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
 %
@@
 %\newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
-%\newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
 %\newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
-%\newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
-%\newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
-%%for nested lists, \ref only produces innermost item:  (hacky)
-%\makeatletter % for internal macros with @
-%\renewcommand\p@enumii{}
-%\makeatother
-%
-\usepackage[inline]{enumitem}
+%
+%\usepackage[inline]{enumitem}
 \newlist{myenuma}{enumerate*}{4}
 \setlist[myenuma]{label=(\alph*),ref=\alph*}
-%\makeatletter
-%\DeclareRobustCommand{\ilabel}[1]{%
-%   \edef\@currentlabel{\themyenumai}%
-%   \label{#1}}
-%\makeatother
+\newcommand{\itemp}[1]{\item\aligntop{#1}}
+
 \graphicspath{{"Figs/"}}
-%\usepackage{pnastwoF}
-%\usepackage{amssymb,amsfonts}
+\usepackage{amssymb,amsfonts}
 %\usepackage[centertags]{amsmath}
 %\usepackage{hyperref}
+\usepackage[capitalise,nameinlink]{cleveref}
+\creflabelformat{equation}{[#2#1#3]}
+\crefrangelabelformat{equation}{[#3#1#4--#5#2#6]}
+%\crefformat{equation}{#2Eq.~[#1]#3}
+%\crefrangeformat{equation}{Eqs.~[#3#1#4--#5#2#6]}
+%\crefmultiformat{equation}{Eqs.~[#2#1#3}{,2#1#3]}{,#2#1#3}{,#2#1#3]}
+%\crefrangemultiformat{equation}{Eqs.~[#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}{,#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}
 
@@
 %
-% indices
-%
+\usepackage{delim}
 % brackets etc.
-\newcommand{\prn}[1]{\left ( #1 \right )}
-\newcommand{\brc}[1]{\left\{ #1 \right\}}
-\newcommand{\brk}[1]{\left [ #1 \right ]}
-\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}
-\newcommand{\nrm}[1]{\left\lVert #1 \right\rVert}
-\newcommand{\av}[1]{\left\langle #1 \right\rangle}
-\newcommand{\avc}[2]{\left\langle #1 \,\middle | \, #2 \right\rangle}
-%\newenvironment{cases}{\left\{\begin{aligned}}{\end{aligned}\right.}
+\delimdef\prn#1{\dleft( #1 \dright)}
+\delimdef\brc#1{\dleft\{ #1 \dright\}}
+\delimdef\brk#1{\dleft[ #1 \dright]}
+\delimdef\abs#1{\dleft\lvert #1 \dright\rvert}
+\delimdef\nrm#1{\dleft\lVert #1 \dright\rVert}
+\delimdef\av#1{\dleft\langle #1 \dright\rangle}
+\delimdef\avc#1#2{\dleft\langle #1 \dmiddle\vert #2 \dright\rangle}
+%\newenvironment{cases}{\dleft\{\begin{aligned}}{\end{aligned}\dright.}
 %
 % Sets
-\newcommand{\set}[2]{\left\{ #1 \,\middle |\, #2 \right\}}
-\newcommand{\cond}[2]{\!\left( #1 \,\middle | \, #2 \right)}
-\newcommand{\condb}[2]{\!\left[ #1 \,\middle | \, #2 \right]}
+\delimdef\set#1#2{\dleft\{ #1 \dmiddle\vert #2 \dright\}}
+\delimdef\cond#1#2{\!\dleft( #1 \dmiddle\vert #2 \dright)}
+\delimdef\condb#1#2{\!\dleft[ #1 \dmiddle\vert #2 \dright]}
 %
 % Derivatives, etc. First argument is optional.
-\newcommand{\diff}[3][\rule{0mm}{0mm}]{\frac{\mathrm{d}^{#1} #2}{\mathrm{d}{#3}^{#1}}}
-\newcommand{\pdiff}[3][\rule{0mm}{0mm}]{\frac{\partial^{#1} #2}{\partial {#3}^{#1}}}
-\newcommand{\pdiffc}[3][\rule{0mm}{0mm}]{\left (\frac{\partial #2}{\partial {#3}}\right )_{\!\!#1}}
-\newcommand{\intd}[2][\rule{0mm}{0mm}]{\int #1\!\!\dr #2\,}
+\newcommand{\diff}[3][]{\frac{\mathrm{d}^{#1} #2}{\mathrm{d}{#3}^{#1}}}
+\newcommand{\pdiff}[3][]{\frac{\partial^{#1} #2}{\partial {#3}^{#1}}}
+\newcommand{\pdiffc}[3][]{\left(\frac{\partial #2}{\partial {#3}}\right)_{\!\!#1}}
+\newcommand{\intd}[2][]{\int #1\!\!\dr #2\,}
 %
@@
 \newcommand{\requires}{\Longleftarrow}
-%
-% Referencing sections, figures, etc
-\newcommand{\sref}[1]{\S\ref{#1}}
-\newcommand{\fref}[1]{fig.\ref{#1}}
-\newcommand{\Fref}[1]{Fig.~\ref{#1}}
-\newcommand{\tref}[1]{tab.\ref{#1}}
-\newcommand{\Tref}[1]{Tab.\ref{#1}}
 %
@@
 % ----------------------------------------------------------------
-%\renewcommand\Fref[1]{}
 \providecommand\hyperref[2][1]{#2}
@@
 
-%%%%%%%%%%%%
-%% For PNAS Only:
-%\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-\renewcommand{\theurl}{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-\copyrightyear{2016}
-\issuedate{Issue Date}
-\volume{Volume}
-\issuenumber{Issue Number}
-%\setcounter{page}{2687} %Set page number here if desired
-%%%%%%%%%%%%
-
-\begin{document}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \title{Optimal synaptic strategies for different timescales of memory}
-
-\author{Subhaneil Lahiri\affil{1}{Department of Applied Physics, Stanford University, Stanford CA}
-\and
-Surya Ganguli\affil{1}{}}
-
-\contributor{Submitted to Proceedings of the National Academy of Sciences
-of the United States of America}
-
-\maketitle
-
-\begin{article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1,2]{Subhaneil Lahiri}
+\author[b,1]{Surya Ganguli}
+
+\affil[a]{Department of Applied Physics, Stanford University, Stanford CA}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Lahiri}
+
+% Please add here a significance statement to explain the relevance of your work
+\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.}
+
+% Please include corresponding author, author contribution and author declaration information
+\authorcontributions{Please provide details of author contributions here.}
+\authordeclaration{Please declare any conflict of interest here.}
+%\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: sulahiri@stanford.edu}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+\keywords{synaptic plasticity | learning | memory | neuroscience}
+
+%\abbreviations{SNR, signal-to-noise ratio; LTP, long term potentiation;LTD, long term depression}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{article}
 \begin{abstract}
@@
 
-\keywords{synaptic plasticity | learning | memory | neuroscience}
-
-\abbreviations{SNR, signal-to-noise ratio; LTP, long term potentiation;
-LTD, long term depression}
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+
+\begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{article}
+
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalised the article contents.
+\verticaladjustment{-2pt}
+
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
@@
 
-Theoretical studies of complex, metaplastic synapses have focused on analyzing the memory performance of a limited number of very specific molecular dynamical systems, characterized by a number of internal states in which potentiation and depression each induce a specific set of allowable transitions between states (\eg see \Fref{fig:models} below).
+Theoretical studies of complex, metaplastic synapses have focused on analyzing the memory performance of a limited number of very specific molecular dynamical systems, characterized by a number of internal states in which potentiation and depression each induce a specific set of allowable transitions between states (\eg see \cref{fig:models} below).
 While these models can vastly outperform simple binary synaptic switches, these analyses leave open several deep and important questions.
@@
 Furthermore, we assume our synaptic model has $M$ internal molecular functional states, and that a candidate potentiating (depotentiating) event induces a stochastic transition in the internal state described by an $M \times M$ discrete time Markov transition matrix  $\M\pot$ ($\M\dep$).
-In this framework, the states of different synapses will be independent, and the entire synaptic population can be fully described by the probability distribution across these states, which we will indicate with the row-vector $\pr(t)$.
-Thus the $i$'th component of $\pr(t)$ denotes the fraction of the synaptic population in state $i$.
+In this framework, the states of different synapses will be independent, and the entire synaptic population can be fully described by the probability distribution across these states, which we will indicate with the row-vector $\prob(t)$.
+Thus the $i$'th component of $\prob(t)$ denotes the fraction of the synaptic population in state $i$.
 Furthermore, each state $i$ has its own synaptic weight, $\w_i$, which we take, in the worst case scenario, to be restricted to two values.
@@
 Much of the previous work on these types of complex synaptic models has focused on understanding the memory curves of specific models, or choices of $\M\potdep$.
-Two examples of these models are shown in \Fref{fig:models}. We see that they have different memory properties.
+Two examples of these models are shown in \cref{fig:models}. We see that they have different memory properties.
 The serial model performs relatively well at one particular timescale, but it performs poorly at other times.
@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
+  \protect\item\aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
   \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
@@
 
-The critical question is: what do these constraints imply about the space of achievable memory curves in \eqref{eq:SNRcalc}?
+The critical question is: what do these constraints imply about the space of achievable memory curves in \cref{eq:SNRcalc}?
 To answer this question, especially for limits on achievable memory at finite times, it will be useful to employ the eigenmode decomposition:
@@
 where $\;\CI_a = (2 f\pot f\dep) \eq (\M\pot-\M\dep) \uv^a  \vv^a \w\;$ and $\;\tau_a = {1}/{q_a}$.
-We can then ask the question: what are the constraints on these quantities, namely eigenmode initial SNR's, $\CI_a$, and time constants, $\tau_a$, implied by the constraints in \eqref{eq:constr}?
+We can then ask the question: what are the constraints on these quantities, namely eigenmode initial SNR's, $\CI_a$, and time constants, $\tau_a$, implied by the constraints in \cref{eq:constr}?
 We will derive some of these constraints in the next section.
@@
 
-In the previous section, in \eqref{eq:SNRcalc} we have described an analytic expression for a memory curve as a function of the structure of a synaptic dynamical system, described by the pair of stochastic transition matrices  $\M\potdep$.
+In the previous section, in \cref{eq:SNRcalc} we have described an analytic expression for a memory curve as a function of the structure of a synaptic dynamical system, described by the pair of stochastic transition matrices  $\M\potdep$.
 Since the performance measure for memory is an entire memory curve, and not just a single number, there is no universal scalar notion of optimal memory in the space of synaptic dynamical systems.
@@
 To maximize this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
-An example of such a model is shown in \Fref{fig:max_init}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
+An example of such a model is shown in \cref{fig:max_init}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
 These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \Fref{fig:max_init}(\ref{fig:binary_det}).
+They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \cref{fig:max_init}(\ref{fig:binary_det}).
 
@@
 known as Kemeny's constant (see \cite[\S4.4]{kemeny1960finite}), is independent of the starting state $i$.
-Intuitively, \eqref{eq:kemenyconst} states that the average time it takes to reach any state, weighted by its equilibrium probability, is independent of the starting state, implying a hidden constancy inherent in any stochastic process.
+Intuitively, \cref{eq:kemenyconst} states that the average time it takes to reach any state, weighted by its equilibrium probability, is independent of the starting state, implying a hidden constancy inherent in any stochastic process.
 
@@
 
-First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \eqref{eq:kemenypm} (see supplementary material):
+First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \cref{eq:kemenypm} (see supplementary material):
 %
@@
 With the states in the order described above, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
-Some of these perturbations are shown in \Fref{fig:perts}, see supplementary material for details.
-For example, in \Fref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
+Some of these perturbations are shown in \cref{fig:perts}, see supplementary material for details.
+For example, in \cref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
 $\smash{\eta^+_i > \eta^+_j}$), we have proven that increasing $\smash{\M\pot_{ij}}$ and decreasing $\smash{\M\dep_{ij}}$ leads to an increase in area.
@@
 This determines the topology (non-zero transition probabilities) of the model with maximal area.
-It is of the form shown in \Fref{fig:env}(\ref{fig:sticky}),%\Fref{fig:max_area},
+It is of the form shown in \cref{fig:env}(\ref{fig:sticky}),%\cref{fig:max_area},
 with potentiation moving one step to the right and depression moving one step to the left.
-Any other topology would allow some class of perturbations (\eg\ in \Fref{fig:perts}) to further increase the area.
+Any other topology would allow some class of perturbations (\eg\ in \cref{fig:perts}) to further increase the area.
 
@@
 We can then maximize this by pushing all of the equilibrium distribution symmetrically to the two end states.
-This can be done by reducing the transition probabilities out of these states, as in \Fref{fig:env}(\ref{fig:sticky}). %\Fref{fig:max_area}.
+This can be done by reducing the transition probabilities out of these states, as in \cref{fig:env}(\ref{fig:sticky}). %\cref{fig:max_area}.
 This makes it very difficult to exit these states once they have been entered.
@@
 Now we will look at the implications of the upper bounds found in the \hyperref[sec:bounds]{previous section} for the SNR at finite times.
-As argued in \eqref{eq:SNReigen}, the memory curve can be written in the form
+As argued in \cref{eq:SNReigen}, the memory curve can be written in the form
 %
@@
 %
-The upper bounds on the initial SNR, \eqref{eq:max_init}, and the area, \eqref{eq:max_area}, imply the following constraints on the parameters $\brc{\CI_a,\tau_a}$:
+The upper bounds on the initial SNR, \cref{eq:max_init}, and the area, \cref{eq:max_area}, imply the following constraints on the parameters $\brc{\CI_a,\tau_a}$:
 %
@@
 At late times, only the area bound is saturated.
-The function $\snr(t_0)$, the green curve in \Fref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim \smash{t_0\inv}$.
+The function $\snr(t_0)$, the green curve in \cref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim \smash{t_0\inv}$.
 No synaptic model can have an SNR that is greater than this at any time.
@@
 At any time, can we find a model whose memory curve touches the envelope?
-The red curves in \Fref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
+The red curves in \cref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
 
@@
 We believe there is some other constraint that prevents models from approaching the envelope, and currently are exploring several mathematical conjectures for the precise form of this constraint in order to obtain a potentially tighter envelope.
-Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \Fref{fig:env}(\ref{fig:sticky})%\Fref{fig:max_area}
+Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \cref{fig:env}(\ref{fig:sticky})%\cref{fig:max_area}
  can come close to the envelope.
@@
  \begin{center}
- \parbox[t]{0.43\linewidth}{
+% \parbox[t]{0.43\linewidth}{%
  \begin{myenuma}
    \item\aligntop{\includegraphics[width=0.85\linewidth]{env.svg}}\label{fig:envplot}
- \end{myenuma}
-  }\hp
-  \parbox[t]{0.44\linewidth}{
- \begin{myenuma}
- \stepcounter{myenumai}
-% \addtocounter{enumi}{1}
-  \item\aligntop{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}\\[0.1\linewidth]
+% \end{myenuma}}\hp
+% \parbox[t]{0.44\linewidth}{%
+% \begin{myenuma}
+%  \stepcounter{myenumai}
+%% \addtocounter{enumi}{1}
+   \item\aligntop{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}%\\[0.1\linewidth]
    \item\aligntop{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
- \end{myenuma}
-  }
+ \end{myenuma}%}
  \end{center}
   \caption{The memory curve envelope for $N=100$, $M=12$.
-  (\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
+  (\protect\protect\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
   The red dashed curve shows the result of numerical optimization of synaptic models with random initialization.
@@
   At late times they are of the form shown in (\ref{fig:sticky}) with different values of $\varepsilon$.
-  The model shown in (\ref{fig:sticky}) also saturates the area bound \eqref{eq:max_area} in the limit $\varepsilon\to0$.
+  The model shown in (\ref{fig:sticky}) also saturates the area bound \cref{eq:max_area} in the limit $\varepsilon\to0$.
   \label{fig:env}}
@@
 
-\begin{acknowledgments}
-We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
-We thank Larry Abbott, Marcus Benna, Stefano Fusi, Jascha Sohl-Dickstein and David Sussillo for useful discussions.
-\end{acknowledgments}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\bibliographystyle{pnas2011}
+\acknow{We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
+We thank Larry Abbott, Marcus Benna, Stefano Fusi, Jascha Sohl-Dickstein and David Sussillo for useful discussions.}
+\showacknow % Display the acknowledgments section
+
+% \pnasbreak splits and balances the columns before the references.
+% If you see unexpected formatting errors, try commenting out this line
+% as it can run into problems with floats and footnotes on the final page.
+%\pnasbreak
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\bibliographystyle{pnas-new}
 \bibliography{maths,neuro,markov}
 
-\end{article}
-
+%\end{article}
+\clearpage
+
+\begin{strip}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
+\appendix
 \begin{center}
@@
   %
-  Substituting back into \eqref{eq:fundinvkern} gives
+  Substituting back into \cref{eq:fundinvkern} gives
   %
@@
   Therefore, $\mathbf{x}=\lambda\onev$.
-  Substituting this into \eqref{eq:fundinvpix} gives
+  Substituting this into \cref{eq:fundinvpix} gives
   %
@@
 \begin{proof}
-  We can deduce \eqref{eq:fundprob} and \eqref{eq:fundrowsum} be pre/post-multiplying the following equations by $\fund$:
+  We can deduce \cref{eq:fundprob,eq:fundrowsum} be pre/post-multiplying the following equations by $\fund$:
   %
@@
   %
-  We can then deduce \eqref{eq:fundQZ} and \eqref{eq:fundZQ} by substituting these into
+  We can then deduce \cref{eq:fundQZ,eq:fundZQ} by substituting these into
   %
@@
 
-%We will not need to use the concept of a g-inverse anywhere, so you can skip ahead to \sref{sec:fpt} now. For the sake of completeness, we can define $\MM^\#$ as
+%We will not need to use the concept of a g-inverse anywhere, so you can skip ahead to \cref{sec:fpt} now. For the sake of completeness, we can define $\MM^\#$ as
 %%
@@
 %  %
-%  with \eqref{eq:ginvcomm} holding iff $\eq$ and $\pib$ are proportional to each other.
+%  with \cref{eq:ginvcomm} holding iff $\eq$ and $\pib$ are proportional to each other.
 %\end{thm}
 %\begin{proof}
-%  Using \eqref{eq:rowsum}, \eqref{eq:equilibrium}, \eqref{eq:fundprob} and \eqref{eq:fundrowsum}, one can verify
+%  Using \cref{eq:rowsum}, \cref{eq:equilibrium}, \cref{eq:fundprob} and \cref{eq:fundrowsum}, one can verify
 %  %
@@
 %  %
-%  Substituting into \eqref{eq:fundQZ} and \eqref{eq:fundZQ} leads to
+%  Substituting into \cref{eq:fundQZ} and \cref{eq:fundZQ} leads to
 %  %
@@
 %  %
-%  Pre/post-multiplying these by $\MM$ or $\MM^\#$ produces \eqref{eq:ginvpropo} and \eqref{eq:ginvpropi}. Subtracting produces
+%  Pre/post-multiplying these by $\MM$ or $\MM^\#$ produces \cref{eq:ginvpropo} and \cref{eq:ginvpropi}. Subtracting produces
 %  %
@@
   For discrete time, a proof can be found in \cite[Th.4.4.10]{kemeny1960finite}.
-  For continuous time, we use \eqref{eq:fptfund}, \eqref{eq:fundrowsum} and the transpose of \eqref{eq:pdotD}:
+  For continuous time, we use \cref{eq:fptfund}, \cref{eq:fundrowsum} and the transpose of \cref{eq:pdotD}:
   %
@@
   %
-  which proves \eqref{eq:mixdef} with $\eta = \tr\fund - \tau$.
+  which proves \cref{eq:mixdef} with $\eta = \tr\fund - \tau$.
 \end{proof}
@@
 Suppose that the Markov process, defined by $\MM$, depends on some parameter $\alpha$.
-Differentiating \eqref{eq:funddef} gives
+Differentiating \cref{eq:funddef} gives
 %
@@
 %
-We can substitute this into the derivative of \eqref{eq:fundprob}:
+We can substitute this into the derivative of \cref{eq:fundprob}:
 %
@@
 %
-\begin{equation}\label{eq:flux}
+\begin{equation}\label{eq:flux:s}
   \F = \D\inv\MM,
@@
 
-This allows us to write the lumpability condition \eqref{eq:lump}, and the transition matrix for the lumped process compactly:
+This allows us to write the lumpability condition \cref{eq:lump}, and the transition matrix for the lumped process compactly:
 %
@@
 
-We assume that the probability distribution starts in the equilibrium distribution \eqref{eq:equilibrium}.
+We assume that the probability distribution starts in the equilibrium distribution \cref{eq:equilibrium}.
 During the initial signal creation, a fraction $f\pot$ will change to $\eq\M\pot$ and a fraction $f\dep$ will change to $\eq\M\dep$.
-After this, probabilities will evolve according to \eqref{eq:rowsum}.
+After this, probabilities will evolve according to \cref{eq:rowsum}.
 
@@
 %
-\begin{equation}\label{eq:SNRdef}
+\begin{equation}\label{eq:SNRdef:s}
   \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}}
@@
 %
-which describes the synapses starting in the equilibrium distribution, changing state due to the plasticity event at $t=0$ and subsequent evolution according to \eqref{eq:rowsum} due to plasticity events uncorrelated with $\wvi$.%
+which describes the synapses starting in the equilibrium distribution, changing state due to the plasticity event at $t=0$ and subsequent evolution according to \cref{eq:rowsum} due to plasticity events uncorrelated with $\wvi$.%
 \footnote{Note that expanding the exponential gives
@@
 %
-Thus, evolving according to \eqref{eq:rowsum} results in averaging over all sequences of plasticity events,
+Thus, evolving according to \cref{eq:rowsum} results in averaging over all sequences of plasticity events,
 as we only need linear expectations of $\wv(t)$ in the end.}
@@
 where we used $\eq\frg=0$ in going from the first to second lines.
-Combining with \eqref{eq:noise} gives
+Combining with \cref{eq:noise} gives
 %
@@
 %
-Then \eqref{eq:SNRcurve} becomes
+Then \cref{eq:SNRcurve} becomes
 %
@@
 
-Suppose that we have a partition such that $\W\pot $ and $\W\dep $ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \sref{sec:lump}):
+Suppose that we have a partition such that $\W\pot $ and $\W\dep $ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \cref{sec:lump}):
 %
@@
 
-Using $\eq\frg=0$ and the first line of \eqref{eq:signal}, we can write the initial SNR as
-%
-\begin{equation}\label{eq:init}
+Using $\eq\frg=0$ and the first line of \cref{eq:signal}, we can write the initial SNR as
+%
+\begin{equation}\label{eq:init:s}
   \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f\pot \W\pot -f\dep \W\dep )(\onev^+-\onev^-).
@@
 %
-Using $\W\potdep (\onev^++\onev^-)=0$ and \eqref{eq:compflux}:
+Using $\W\potdep (\onev^++\onev^-)=0$ and \cref{eq:compflux}:
 %
@@
 %
-we can rewrite \eqref{eq:init} as
-%
-\begin{equation}\label{eq:initflux}
+we can rewrite \cref{eq:init} as
+%
+\begin{equation}\label{eq:initflux:s}
   I = \frac{4\F_{-+}}{r} - 4{\eq}^+f\pot \W\pot \onev^- - 4{\eq}^-f\dep \W\dep \onev^+.
@@
 In this section we will find an upper bound on the area under the signal-to-noise curve.
-As in \sref{sec:SNR}, we will only consider ergodic Markov chains.
-We will see in \sref{sec:multistate} that the optimal chain is absorbing, so it lies on the boundary of the (open) set of ergodic chains, but it still puts an upper bound on the area for any chain in the interior.
-
-\subsection{Area under signal-to-noise curve}\label{sec:area}
-
-The signal-to-noise curve is given by \eqref{eq:SNRalpha}.
+As in \cref{sec:SNR}, we will only consider ergodic Markov chains.
+We will see in \cref{sec:multistate} that the optimal chain is absorbing, so it lies on the boundary of the (open) set of ergodic chains, but it still puts an upper bound on the area for any chain in the interior.
+
+\subsection{Area under signal-to-noise curve}\label{sec:areadef}
+
+The signal-to-noise curve is given by \cref{eq:SNRalpha}.
 %%
@@
 %
-\begin{equation}\label{eq:area}
+\begin{equation}\label{eq:area:s}
 \begin{aligned}
@@
 
-We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f\pot f\dep )\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W\pot _{ij}-\W\dep _{ij}$
+We can rewrite this using \cref{eq:diffpT}, with $A=\sqrt{N}(2f\pot f\dep )\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W\pot _{ij}-\W\dep _{ij}$
 %
@@
 \end{defn}
-Using \eqref{eq:mixdef}, we can write:
+Using \cref{eq:mixdef}, we can write:
 %
@@
 
-We can rewrite \eqref{eq:areaT} as
-%
-\begin{equation}\label{eq:areaEta}
+We can rewrite \cref{eq:areaT} as
+%
+\begin{equation}\label{eq:areaEta:s}
 \begin{aligned}
@@
 %
-We can also express it in terms of the fundamental matrix \eqref{eq:funddef} as
+We can also express it in terms of the fundamental matrix \cref{eq:funddef} as
 %
@@
 
-In particular, differentiating \eqref{eq:funddef},
+In particular, differentiating \cref{eq:funddef},
 %
@@
 %
-We can then differentiate expression \eqref{eq:areaZ} to get
+We can then differentiate expression \cref{eq:areaZ} to get
 %
@@
 %
-where $a_i$ and $c_k$ were defined in \eqref{eq:areacoeffs}.
+where $a_i$ and $c_k$ were defined in \cref{eq:areacoeffs}.
 
@@
 %
-%The last term ensures that we hold $\eq_\pm$ fixed during this extremisation, so that we can safely ignore the factors that we dropped from the noise in \eqref{eq:noisepm}.
+%The last term ensures that we hold $\eq_\pm$ fixed during this extremisation, so that we can safely ignore the factors that we dropped from the noise in \cref{eq:noisepm}.
 Necessary conditions for an extremum are
@@
 This enforces the positivity constraints on the off-diagonal elements, but not the diagonals.
-As discussed in \sref{sec:rescale}, that can be enforced after finding the maximum using the null scaling degree of freedom.
+As discussed in \cref{sec:rescale}, that can be enforced after finding the maximum using the null scaling degree of freedom.
 
@@
 
-%Note that if we multiply \eqref{eq:shiftqderiv} by $\enc_{gh}$ and sum over $(g,h)$, we get
+%Note that if we multiply \cref{eq:shiftqderiv} by $\enc_{gh}$ and sum over $(g,h)$, we get
 %%
@@
 %
-Once again, they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).
-
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators \eqref{eq:diffpT}:
+Once again, they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \cref{sec:triangular} about this being satisfied for ergodic chains).
+
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators \cref{eq:diffpT}:
 %
@@
 In the \hyperref[sec:areacoeffincr]{section below}, we will show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
-This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
+This implies that the last term of the final expression in \cref{eq:shortcutarea} is non-negative.
 If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu\potdep _{g,g\pm m}>0$, which in turn implies that $\W\potdep _{g,g\pm m}=0$.
@@
 %
-Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \sref{sec:triangular} about this being satisfied for ergodic chains).
-
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators using \eqref{eq:diffpT}:
+Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \cref{sec:triangular} about this being satisfied for ergodic chains).
+
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators using \cref{eq:diffpT}:
 %
@@
 %
-where $c_k$ were defined in \eqref{eq:areacoeffs}.
-
-Using the non-negativity of the Kuhn-Tucker multipliers, $\mu\potdep _{ij}$, \eqref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
-
-Note that, according to \sref{sec:triangular}, either $\W\pot _{gh}$ or $\W\dep _{gh}$ will be zero at the maximum, therefore we can expect one of $\mu\pot _{gh}+\mu\dep _{gh}$ to be non-zero.
+where $c_k$ were defined in \cref{eq:areacoeffs}.
+
+Using the non-negativity of the Kuhn-Tucker multipliers, $\mu\potdep _{ij}$, \cref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
+
+Note that, according to \cref{sec:triangular}, either $\W\pot _{gh}$ or $\W\dep _{gh}$ will be zero at the maximum, therefore we can expect one of $\mu\pot _{gh}+\mu\dep _{gh}$ to be non-zero.
 This would rule out degeneracy of the $c_k$ or $\eta^+_k$.
-Looking at \eqref{eq:shiftqderiv} closely, the only way $\mu\pot _{gh}+\mu\dep _{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
-
-%From the expression for the area in \eqref{eq:areacoeffs}, it will clearly also be optimal to have the $\w_k$ non-decreasing, \ie all the positive $\w_k$ should lie ahead of all the negative $\w_k$.
+Looking at \cref{eq:shiftqderiv} closely, the only way $\mu\pot _{gh}+\mu\dep _{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
+
+%From the expression for the area in \cref{eq:areacoeffs}, it will clearly also be optimal to have the $\w_k$ non-decreasing, \ie all the positive $\w_k$ should lie ahead of all the negative $\w_k$.
 
@@
 %Suppose we start with some arbitrary ergodic process.
-If $\W\potdep $ contain backwards/forwards transitions or the $c_k$ decrease, then we can increase the area by applying the perturbations \eqref{eq:shiftqderiv} and \eqref{eq:areacoeffincrderiv}.
+If $\W\potdep $ contain backwards/forwards transitions or the $c_k$ decrease, then we can increase the area by applying the perturbations \cref{eq:shiftqderiv} and \cref{eq:areacoeffincrderiv}.
 %The second perturbation will generate the unwanted backwards/forwards transitions, requiring subsequent application of first perturbation.
@@
 %\begin{itemize}
-%  \item Applying \eqref{eq:areacoeffincrderiv} an infinite amount sends two transition rates to infinity, which due to the scaling mode (\sref{sec:rescale}), is equivalent to sending all other transitions to zero.
-%      Subsequent application of \eqref{eq:shiftqderiv} just removes the backwards transition from $\W\pot $ and the forward transition from $\W\dep $.
-%  \item Applying \eqref{eq:areacoeffincrderiv} ends in a process with degenerate $c_k$ or $\eta^+_k$, and the subsequent application of \eqref{eq:shiftqderiv} doesn't change this degeneracy.
-%  \item As above, but the subsequent application of \eqref{eq:shiftqderiv} lifts the degeneracy, resulting in a chain with increasing $c_k$.
+%  \item Applying \cref{eq:areacoeffincrderiv} an infinite amount sends two transition rates to infinity, which due to the scaling mode (\cref{sec:rescale}), is equivalent to sending all other transitions to zero.
+%      Subsequent application of \cref{eq:shiftqderiv} just removes the backwards transition from $\W\pot $ and the forward transition from $\W\dep $.
+%  \item Applying \cref{eq:areacoeffincrderiv} ends in a process with degenerate $c_k$ or $\eta^+_k$, and the subsequent application of \cref{eq:shiftqderiv} doesn't change this degeneracy.
+%  \item As above, but the subsequent application of \cref{eq:shiftqderiv} lifts the degeneracy, resulting in a chain with increasing $c_k$.
 %\end{itemize}
 %%
-%In the third case, we can then apply the perturbation \eqref{eq:shortcutderiv} to remove all the shortcuts.
+%In the third case, we can then apply the perturbation \cref{eq:shortcutderiv} to remove all the shortcuts.
 %Either this, or the first case, results in a multi-state topology, possibly disconnected or without the full set of states.
@@
 This leaves two possibilities for the maximum area Markov chain.
-Either there is no degeneracy and no shortcuts, which implies the Multi-state/serial topology that we'll discuss in \sref{sec:multistate}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \eqref{eq:shortcutarea}).
+Either there is no degeneracy and no shortcuts, which implies the Multi-state/serial topology that we'll discuss in \cref{sec:multistate}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \cref{eq:shortcutarea}).
 
@@
 It is usually hard to arrange without some symmetry relating degenerate states.
-Such a symmetry would imply lumpability (see \sref{sec:lump}).
+Such a symmetry would imply lumpability (see \cref{sec:lump}).
 The lumped chain would not have any shortcuts, as an entire degenerate set cannot be bypassed.
-As this lumped chain has the same area (see \sref{sec:SNRlump}), we would need only consider the multi-state topology.
+As this lumped chain has the same area (see \cref{sec:SNRlump}), we would need only consider the multi-state topology.
 
@@
 %
-where we used \eqref{eq:multistateineq} to derive the first two equations respectively and Th.\ref{th:kemenyconst} to derive the third. This allows us to write the area as
-%
-\begin{equation}\label{eq:multistatearea}
+where we used \cref{eq:multistateineq} to derive the first two equations respectively and Th.\ref{th:kemenyconst} to derive the third. This allows us to write the area as
+%
+\begin{equation}\label{eq:multistatearea:s}
  % A = 2\sqrt{N} \sum_k \brk{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \eq_k \w_k.
@@
 %
-where we used $\w_k=\sgn(c_k)$, as discussed after \eqref{eq:areacoeffs}.
+where we used $\w_k=\sgn(c_k)$, as discussed after \cref{eq:areacoeffs}.
 This reproduces equation (15) of the main paper.
@@
 %%
-%where $\lambda$ is a Lagrange multiplier and $\mu_i$ are Kuhn-Tucker multipliers (see \eqref{eq:extremum}). Extremising \wrt $\eq_i$:
+%where $\lambda$ is a Lagrange multiplier and $\mu_i$ are Kuhn-Tucker multipliers (see \cref{eq:extremum}). Extremising \wrt $\eq_i$:
 %%
@@
 %%
-%As $\abs{\eq_+-\eq_-}<1$ (unless we want $A=0$, \eqref{eq:allpplusorminus}), this means the $\mu_i$ increase amongst the negative $\w_i$ and decrease amongst the positive $\w_i$.
+%As $\abs{\eq_+-\eq_-}<1$ (unless we want $A=0$, \cref{eq:allpplusorminus}), this means the $\mu_i$ increase amongst the negative $\w_i$ and decrease amongst the positive $\w_i$.
 %Therefore the only non-zero equilibrium probabilities at optimum are $\eq_1$ and $\eq_n$.
@@
 In order to obtain an upper bound on the area under the memory curve of any model, we now maximise the area of the serial model with respect to its equilibrium distribution.
-First let us maximise \eqref{eq:multistatearea} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
+First let us maximise \cref{eq:multistatearea} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
 Clearly this will happen when we put all of the probability at the ends: $\eq_1=\eq_-$ and $\eq_n=\eq_+$ are the only non-zero $\eq_k$.
 This gives an area of
-%\footnote{Note that including the dropped factors in \eqref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_+\eq_-)}(n-1)/r$, which would have no effect on what follows.}
+%\footnote{Note that including the dropped factors in \cref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_+\eq_-)}(n-1)/r$, which would have no effect on what follows.}
 
@@
 
+\end{strip}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \pnasbreak splits and balances the columns before the references.
+% If you see unexpected formatting errors, try commenting out this line
+% as it can run into problems with floats and footnotes on the final page.
+%\pnasbreak
 
 %\bibliographystyle{utcaps_sl}
-%\bibliography{maths,neuro}
+%\bibliography{maths,neuro,markov}
 

### 2018-03-16 Paper/SynapsePaper.tex
cleaning

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% ADDITIONAL OPTIONAL STYLE FILES
-%\usepackage[hypcap]{caption}%links to figures
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \usepackage{graphicx}
@@
 %
-%\usepackage{paralist}
-%\newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
-%\newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
-%
 %\usepackage[inline]{enumitem}
@@
 \setlist[myenuma]{label=(\alph*),ref=\alph*}
-\newcommand{\itemp}[1]{\item\aligntop{#1}}
-
+\newcommand{\panel}[1]{\item\aligntop{#1}}
 \graphicspath{{"Figs/"}}
+
 \usepackage{amssymb,amsfonts}
+\usepackage{amsthm}
 %\usepackage[centertags]{amsmath}
@@
 
+% ----------------------------------------------------------------
+\newtheoremstyle{sldefinition}%
+  {3pt}%space above
+  {3pt}%space below
+  {}%body font
+  {}%indent amount
+  {\bfseries}%theorem head font
+  {}%theorem head punctuation
+  {\newline}%space after head
+  {\thmname{#1}\thmnumber{ #2}:{\bfseries\thmnote{ #3}}}%head spec
+\newtheoremstyle{slplain}%
+  {3pt}%space above
+  {3pt}%space below
+  {}%body font
+  {}%indent amount
+  {\bfseries}%theorem head font
+  {}%theorem head punctuation
+  {\newline}%space after head
+  {\thmname{#1}\thmnumber{ #2}{\bfseries\thmnote{ (#3)}}: }%head spec
+%
+\theoremstyle{slplain}
+\newtheorem{thm}{Theorem}
+\newtheorem{cor}[thm]{Corollary}
+\newtheorem{lem}[thm]{Lemma}
+%
+\theoremstyle{sldefinition}
+\newtheorem{defn}{Definition}
+%
+\theoremstyle{remark}
+\newtheorem*{notn}{Notation}
+\newtheorem*{rem}{Remark}
+% ----------------------------------------------------------------
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% OPTIONAL MACRO DEFINITIONS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newcommand{\ra}{\rightarrow}
@@
 \newcommand{\CI}{\mathcal{I}}
-
-% ----------------------------------------------------------------
-\input{sl_theorems_preamble.tex}
-% ----------------------------------------------------------------
-\providecommand\hyperref[2][1]{#2}
-\providecommand\texorpdfstring[2]{#1}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
@@
 
+\setboolean{displaywatermark}{false}
 \begin{document}
@@
  \begin{myenuma}
-  \protect\item\aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
-  \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
-  \item\aligntop{\includegraphics[width=0.3\linewidth]{cascms.eps}}\label{fig:cascms}
+  \panel{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
+  \panel{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
+  \panel{\includegraphics[width=0.3\linewidth]{cascms.eps}}\label{fig:cascms}
  \end{myenuma}
@@
  \begin{myenuma}
-  \item\hp\aligntop{\includegraphics[width=0.24\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
-  \item\aligntop{\includegraphics[width=0.24\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
-  \item\hp\aligntop{\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
+  \panel{\hp\includegraphics[width=0.24\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
+  \panel{\includegraphics[width=0.24\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
+  \panel{\hp\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
   \end{myenuma}
@@
 \footnote{Note that we do not need to worry about the order of the $\eta^\pm_i$ changing during the optimization: necessary conditions for a maximum only require that there is no infinitesimal perturbation that increases the area.
-Therefore we need only consider an infinitesimal neighborhood of the model, in which the order will not change.}
+Therefore we need only consider an infinitesimal neighborhood of the model, in which the order will not change.\label{ft:etaorder}}
 
@@
  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.14\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
-  \item\aligntop{\includegraphics[width=0.14\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
-  \item\aligntop{\includegraphics[width=0.28\linewidth]{shortcut.svg}}\label{fig:shortcut}
+  \panel{\includegraphics[width=0.14\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
+  \panel{\includegraphics[width=0.14\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
+  \panel{\includegraphics[width=0.28\linewidth]{shortcut.svg}}\label{fig:shortcut}
  \end{myenuma}
@@
  \begin{center}
-% \parbox[t]{0.43\linewidth}{%
+ \parbox[t]{0.43\linewidth}{%
  \begin{myenuma}
-   \item\aligntop{\includegraphics[width=0.85\linewidth]{env.svg}}\label{fig:envplot}
-% \end{myenuma}}\hp
-% \parbox[t]{0.44\linewidth}{%
-% \begin{myenuma}
-%  \stepcounter{myenumai}
-%% \addtocounter{enumi}{1}
-   \item\aligntop{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}%\\[0.1\linewidth]
-   \item\aligntop{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
- \end{myenuma}%}
+   \panel{\includegraphics[width=0.85\linewidth]{env.svg}}\label{fig:envplot}
+ \end{myenuma}}\hp
+ \parbox[t]{0.44\linewidth}{%
+ \begin{myenuma}
+  \stepcounter{myenumai}
+% \addtocounter{enumi}{1}
+   \panel{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}\\[0.1\linewidth]
+   \panel{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
+ \end{myenuma}}
  \end{center}
@@
 %\bibliographystyle{pnas-new}
-\bibliography{maths,neuro,markov}
+%\bibliography{maths,neuro,markov}
 
 %\end{article}
-\clearpage
+%\clearpage
 
@@
 \begin{center}
-  \textbf{\Large Supplementary material}
+  \textbf{\Large Supporting information}
 \end{center}
@@
 Thus, evolving according to \cref{eq:rowsum} results in averaging over all sequences of plasticity events,
-as we only need linear expectations of $\wv(t)$ in the end.}
+as we only need linear expectations of $\wv(t)$ in the end.\label{ft:avexp}}
 This results in
@@
 
-In the \hyperref[sec:areacoeffincr]{section below}, we will show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
+In the \hyperref[sec:areacoeffincr]{\lcnamecref{sec:areacoeffincr} below}, we will show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
 This implies that the last term of the final expression in \cref{eq:shortcutarea} is non-negative.
@@
 
-In the \hyperref[sec:shortcuts]{previous subsection} we defined perturbations corresponding to Figure 3(c) of the main text.
+In the \hyperref[sec:shortcuts]{previous \lcnamecref{sec:shortcuts}} we defined perturbations corresponding to Figure 3(c) of the main text.
 In order to show that those perturbations increase the area, we must now show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
@@
 %\bibliographystyle{utcaps_sl}
-%\bibliography{maths,neuro,markov}
+\bibliography{maths,neuro,markov}
 

### 2018-03-17 Paper/SynapsePaper.tex
cross refs customised

@@
 \providecommand{\inkscapedir}{}
-%
-\epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{"\inkscapedir inkscape" -f #1 -C -A \OutputFile}
+\epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{%
+  "\inkscapedir inkscape" -f #1 -C -A \OutputFile}
 %
@@
 \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
-%
 %\usepackage[inline]{enumitem}
@@
 \graphicspath{{"Figs/"}}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \usepackage{amssymb,amsfonts}
 \usepackage{amsthm}
+%\usepackage[notref,notcite]{showkeys}
 %\usepackage[centertags]{amsmath}
+%\usepackage{bibunits}
 %\usepackage{hyperref}
+\hypersetup{bookmarksopenlevel=1,bookmarksnumbered}
 \usepackage[capitalise,nameinlink]{cleveref}
+%\Crefname{appendix}{App.}{Apps.}
 \creflabelformat{equation}{[#2#1#3]}
@@
 %\crefrangemultiformat{equation}{Eqs.~[#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}{,#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}
-
+\newcommand{\main}{the main paper}
+\newcommand{\supp}{supporting information}
 % ----------------------------------------------------------------
@@
 %
-%\input{sl_symbols.tex}
-%
 %additional symbols:
@@
 \newcommand{\prob}{\mathbf{p}}
-\newcommand{\eq}{\prob^\infty}
+\newcommand{\eq}{\boldsymbol{\pi}}
 %first passage times
@@
 \newcommand{\D}{\mathbf{D}}
-\newcommand{\pib}{\boldsymbol{\pi}}
+\newcommand{\arow}{\boldsymbol{\xi}}
 \newcommand{\Lb}{\boldsymbol{\Lambda}}
@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1,2]{Subhaneil Lahiri}
-\author[b,1]{Surya Ganguli}
+\author[a,1]{Subhaneil Lahiri}
+\author[a]{Surya Ganguli}
 
@@
 %\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: sulahiri@stanford.edu}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: sulahiri@stanford.edu}
 
@@
 Theoretical work has shown that such models have a reasonable, extensive memory capacity, in which the number of long term associations that can be stored by a neuron is proportional its number of afferent synapses \cite{Hopfield1982model,Amit1985hopfield,Gardner1988perceptron}.
-However, recent experimental work has shown that many synapses are more digital than analog; they cannot robustly assume an infinite continuum of analog values, but rather can only take on a finite number of distinguishable strengths, a number than can be as small as two \cite{Bliss1993LTP,Petersen1998allornone,O'Connor2005switch} (though see \cite{Enoki2009graded}).
+However, recent experimental work has shown that many synapses are more digital than analog; they cannot robustly assume an infinite continuum of analog values, but rather can only take on a finite number of distinguishable strengths, a number than can be as small as two \cite{Bliss1993LTP,Petersen1998allornone,O'Connor2005switch} \cite[though see][] {Enoki2009graded}.
 This one simple modification leads to a catastrophe in memory capacity:  classical models with digital synapses, when operating in a palimpsest mode in which the ongoing storage of new memories can overwrite previous memories, have a memory capacity proportional to the logarithm of the number of synapses \cite{amit1992constraints,amit1994learning}.
@@
 Recent work \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial} has suggested that a way out of this logarithmic catastrophe is to expand our theoretical conception of a synapse from a single scalar value to an entire stochastic dynamical system in its own right.
-This conceptual expansion is further necessitated by the experimental reality that synapses contain within them immensely complex molecular signaling pathways, with many internal molecular functional states (\eg see \cite{Bliss1993LTP,Bredt2003AMPA,Coba2009phosphorylation}).
+This conceptual expansion is further necessitated by the experimental reality that synapses contain within them immensely complex molecular signaling pathways, with many internal molecular functional states \cite[\eg see][]{Bliss1993LTP,Bredt2003AMPA,Coba2009phosphorylation}.
 While externally, synaptic efficacy could be digital, candidate patterns of electrical activity leading to potentiation or depression could yield transitions between these internal molecular states without necessarily inducing an associated change in synaptic efficacy.
@@
 The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
-A detailed derivation of this formula can be found in the supplementary material.
+A detailed derivation of this formula can be found in the \supp.
 We will frequently refer to this function as the memory curve.
@@
  \end{center}
-  \caption{Models of complex synapses.
+  \caption[Models of complex synapses]{Models of complex synapses.
   (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
@@
 %
-The initial SNR is closely related to the flux from the states with $\w_i=-1$ to those with $\w_j=+1$ (see supplementary material):
+The initial SNR is closely related to the flux from the states with $\w_i=-1$ to those with $\w_j=+1$ (see \supp):
 %
@@
 These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \cref{fig:max_init}(\ref{fig:binary_det}).
+They are completely equivalent to a two state model with transition probabilities equal to 1, as shown in \cref{fig:max_init}(\ref{fig:binary_det}).
+In particular, they have the same memory curve.
 
@@
  \end{center}
-  \caption{Synaptic models that maximize initial SNR.
+  \caption[Synaptic models that maximize initial SNR]{Synaptic models that maximize initial SNR.
   (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
@@
 
-Our goal of understanding the relationship between structure and function in the space of all possible synaptic models is complicated by the fact that this space contains many different possible network topologies, encoded in the nonzero matrix elements of $\M\potdep$.
+Our goal is to understand the relationship between structure and function in the space of all possible synaptic models.
+This goal is complicated by the fact that this space contains many different possible network topologies, encoded in the nonzero matrix elements of $\M\potdep$.
 To systematically analyze this entire space, we develop an important organizing principle using the theory of first passage times in the stochastic process of forgetting, described by $\frg$.
@@
 %
-known as Kemeny's constant (see \cite[\S4.4]{kemeny1960finite}), is independent of the starting state $i$.
+is independent of the starting state $i$.
+It is known as Kemeny's constant \cite[see][\S4.4]{kemeny1960finite}.
 Intuitively, \cref{eq:kemenyconst} states that the average time it takes to reach any state, weighted by its equilibrium probability, is independent of the starting state, implying a hidden constancy inherent in any stochastic process.
@@
 
-First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \cref{eq:kemenypm} (see supplementary material):
+First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \cref{eq:kemenypm} (see \supp):
 %
@@
 With the states in the order described above, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
-Some of these perturbations are shown in \cref{fig:perts}, see supplementary material for details.
+Some of these perturbations are shown in \cref{fig:perts}, see \supp\ for details.
 For example, in \cref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
@@
 with potentiation moving one step to the right and depression moving one step to the left.
-Any other topology would allow some class of perturbations (\eg\ in \cref{fig:perts}) to further increase the area.
+Any other topology would allow some class of perturbations (\eg in \cref{fig:perts}) to further increase the area.
 
@@
  \end{center}
-  \caption{Perturbations that increase the area.
+  \caption[Perturbations that increase the area]{Perturbations that increase the area.
   (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$.
@@
 As these perturbations do not change the equilibrium distribution, this means that the area of \emph{any} model is bounded by that of a linear chain with the same equilibrium distribution.
-The area of a linear chain model can be expressed directly in terms of its equilibrium state distribution, $\eq$, yielding the following upper bound on the area of any model with the same $\eq$ (see supplementary material):
+The area of a linear chain model can be expressed directly in terms of its equilibrium state distribution, $\eq$, yielding the following upper bound on the area of any model with the same $\eq$ (see \supp):
 %
@@
  \end{center}
-  \caption{The memory curve envelope for $N=100$, $M=12$.
+  \caption[The memory curve envelope for $N=100$, $M=12$]{The memory curve envelope for $N=100$, $M=12$.
   (\protect\protect\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
@@
 %\bibliographystyle{pnas-new}
-%\bibliography{maths,neuro,markov}
+\bibliography{maths,neuro,markov}
 
@@
 \appendix
+\pdfbookmark[0]{Supporting information}{appendix}
+\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
 \begin{center}
@@
 
-Here we provide more details underlying the derivations of results in the main paper.
+Here we provide more details underlying the derivations of results in \main.
 
@@
   \begin{equation}\label{eq:funddef}
-    \fund \equiv (-\MM + \onev\pib)\inv,
-  \end{equation}
-  %
-  where $\pib$ is any row-vector with $\pib\onev=1/\tau\neq0$.
+    \fund \equiv (-\MM + \onev\arow)\inv,
+  \end{equation}
+  %
+  where $\arow$ is any row-vector with $\arow\onev=1/\tau\neq0$.
 \end{defn}
-Note that the canonical choice for the discrete time version, $\pib=\eq$, is not available here due to problems with units.
-It will be helpful to choose $\pib$ to be independent of $\MM$, \eg $\pib=\onev\trans/(n\tau)$.
+Note that the canonical choice for the discrete time version, $\arow=\eq$, is not available here due to problems with units.
+It will be helpful to choose $\arow$ to be independent of $\MM$, \eg $\arow=\onev\trans/(n\tau)$.
 All quantities that we calculate using $\fund$ below will be independent of this choice.
@@
 \begin{thm}
-  The definition of $\fund$ is valid, \ie $(-\MM + \onev\pib)$ is invertible.
+  The definition of $\fund$ is valid, \ie $(-\MM + \onev\arow)$ is invertible.
 \end{thm}
@@
   \begin{equation}\label{eq:fundinvkern}
-    (-\MM + \onev\pib)\mathbf{x}=0.
+    (-\MM + \onev\arow)\mathbf{x}=0.
   \end{equation}
@@
   \begin{equation}\label{eq:fundinvpix}
-    \pib \mathbf{x} = 0.
+    \arow \mathbf{x} = 0.
   \end{equation}
@@
   \begin{equation*}
-    \lambda\pib\onev = \frac{\lambda}{\tau} = 0.
+    \lambda\arow\onev = \frac{\lambda}{\tau} = 0.
   \end{equation*}
   %
-  As we defined $\pib$ such that $1/\tau\neq0$, this means $\lambda=0 \implies \mathbf{x}=0$.
+  As we defined $\arow$ such that $1/\tau\neq0$, this means $\lambda=0 \implies \mathbf{x}=0$.
 \end{proof}
@@
   \begin{align}
-    \pib\fund &= \eq, \label{eq:fundprob}\\
+    \arow\fund &= \eq, \label{eq:fundprob}\\
     \fund\onev &= \tau\onev,\label{eq:fundrowsum}\\
     \I+\MM\fund &= \onev\eq, \label{eq:fundQZ}\\
-    \I+\fund\MM &= \tau\onev\pib. \label{eq:fundZQ}
+    \I+\fund\MM &= \tau\onev\arow. \label{eq:fundZQ}
   \end{align}
@@
     \begin{aligned}
-      \eq(-\MM + \onev\pib) &= \pib, \\
-      (-\MM + \onev\pib)\onev &= \frac{\onev}{\tau}.
+      \eq(-\MM + \onev\arow) &= \arow, \\
+      (-\MM + \onev\arow)\onev &= \frac{\onev}{\tau}.
     \end{aligned}
@@
   \begin{equation*}
-    (-\MM + \onev\pib)\fund = \fund(-\MM + \onev\pib) = \I.
+    (-\MM + \onev\arow)\fund = \fund(-\MM + \onev\arow) = \I.
   \end{equation*}
@@
 %  %
-%  with \cref{eq:ginvcomm} holding iff $\eq$ and $\pib$ are proportional to each other.
+%  with \cref{eq:ginvcomm} holding iff $\eq$ and $\arow$ are proportional to each other.
 %\end{thm}
@@
 %    &\qquad
-%    \pib \MM^\# &= 0,
+%    \arow \MM^\# &= 0,
 %    \\
@@
 %    \qquad
-%    \I - \MM^\# \MM = \tau \onev \pib.
+%    \I - \MM^\# \MM = \tau \onev \arow.
 %  \end{equation*}
@@
 %  \begin{equation*}
-%    [\MM,\MM^\#] = \onev(\tau\pib-\eq),
+%    [\MM,\MM^\#] = \onev(\tau\arow-\eq),
 %  \end{equation*}
 %  %
-%  which vanishes iff $\eq$ and $\pib$ are proportional to each other (due to $\pib\onev=1/\tau$ and $\eq\onev=1$, the constant of proportionality could only be $\tau$).
+%  which vanishes iff $\eq$ and $\arow$ are proportional to each other (due to $\arow\onev=1/\tau$ and $\eq\onev=1$, the constant of proportionality could only be $\tau$).
 %\end{proof}
@@
 \begin{equation}\label{eq:diffp}
-  \diff{\eq}{\alpha} = \pib \fund \diff{\MM}{\alpha} \fund = \eq \diff{\MM}{\alpha} \fund.
+  \diff{\eq}{\alpha} = \arow \fund \diff{\MM}{\alpha} \fund = \eq \diff{\MM}{\alpha} \fund.
 \end{equation}
@@
     &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} (\fund_{jk} - \fund_{ik}) \\
-    &= \sum_{i\neq j} \diff{\MM_{ij}}{\alpha} \eq_i \eq_k (\fptb_{ik} - \fptb_{jk}).
+    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} \eq_k (\fptb_{ik} - \fptb_{jk}).
 \end{aligned}
@@
   \begin{aligned}
-    \pib^\CA &=\pib\I^\CA, &
+    \arow^\CA &=\arow\I^\CA, &
     \M^{\cdot\CA}&=\M\I^\CA, &
@@
     \\
-    \pib_\CA &=\pib\onev^\CA, &
+    \arow_\CA &=\arow\onev^\CA, &
     \M_{\cdot\CA}&=\M\onev^\CA, &
@@
 %
-where $\pib$ is a row vector, $\M$ is a matrix and $\mathbf{x}$ is a column vector.
+where $\arow$ is a row vector, $\M$ is a matrix and $\mathbf{x}$ is a column vector.
 
@@
 
-As discussed in the main text, the signal-to-noise ratio is given by
+As discussed in \main, the signal-to-noise ratio is given by
 %
@@
     + \sum_{\alpha\neq\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}\av{\smash{\wv^\beta}_\text{ideal}\smash{\wv^\beta}(\infty)}
-    \\&\phantom{= \sum_{\alpha}  \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}}
+%    \\&\phantom{= \sum_{\alpha}  \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}}
     - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2  \\
@@
 \begin{equation}\label{eq:areaZ}
-  \hat{A} = \sum_{i,j,k,l} \enc_{ij} \pib_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
-    = \pib \fund q \fund \w.
+  \hat{A} = \sum_{i,j,k,l} \enc_{ij} \arow_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
+    = \arow \fund q \fund \w.
 \end{equation}
@@
 
-In the following, we will mathematically define the classes of perturbations pictorially described in Figure 3 of the main paper.
+In the following, we will mathematically define the classes of perturbations pictorially described in \cref{fig:perts} of \main.
 In order to do so, we will need to consider expressions for derivatives of various quantities with respect to $\W\potdep_{ij}$.
 
-As discussed in the main text,
+As discussed in \main,
 we will regard the off-diagonal elements of $\W\potdep _{ij}$ to be the independent variables,
@@
     \pdiff{\hat{A}}{\W\potdep _{gh}} =&\,
-%      rf\potdep  \sum_{ijkl} \enc_{ij} \w_k \pib_l \brk{
+%      rf\potdep  \sum_{ijkl} \enc_{ij} \w_k \arow_l \brk{
 %         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
-%    &\pm \sum_{kl} \w_k \pib_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
+%    &\pm \sum_{kl} \w_k \arow_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
 %    =&\,
@@
 %
-The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W\potdep $.
+The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\arow$ and $\tau$ on $\W\potdep $.
 
@@
 
-Here we describe the perturbations corresponding to Figure 3(a,b) of the main paper.
+Here we describe the perturbations corresponding to \cref{fig:perts}(\ref{fig:tri_right},\ref{fig:tri_left}) of \main.
 
@@
 
-In this subsection we will define perturbations corresponding to Figure 3(c) of the main text.
+In this subsection we will define perturbations corresponding to \cref{fig:perts}(\ref{fig:shortcut}) of \main.
 
@@
 
-In the \hyperref[sec:shortcuts]{previous \lcnamecref{sec:shortcuts}} we defined perturbations corresponding to Figure 3(c) of the main text.
+In the \hyperref[sec:shortcuts]{previous \lcnamecref{sec:shortcuts}} we defined perturbations corresponding to \cref{fig:perts}(\ref{fig:shortcut}) of \main.
 In order to show that those perturbations increase the area, we must now show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
@@
 \begin{equation}\label{eq:multistatedef}
-  \W\pot _{ij} = q\pot _i \delta_{i+1,j},
-  \qquad
-  \W\dep _{ij} = q\dep _j \delta_{i,j+1}.
+  \W\pot _{ij} = q\pot _i (\delta_{i+1,j} - \delta_{ij}),
+  \qquad
+  \W\dep _{ij} = q\dep _j (\delta_{i,j+1} - \delta_{ij}).
 \end{equation}
@@
 %
-where we used \cref{eq:multistateineq} to derive the first two equations respectively and Th.\ref{th:kemenyconst} to derive the third. This allows us to write the area as
+where we used \cref{eq:multistateineq} to derive the first two equations respectively and \cref{th:kemenyconst} to derive the third. This allows us to write the area as
 %
@@
 where we used $\w_k=\sgn(c_k)$, as discussed after \cref{eq:areacoeffs}.
-This reproduces equation (15) of the main paper.
+This reproduces \cref{eq:multistatearea} of \main.
 
@@
 In order to obtain an upper bound on the area under the memory curve of any model, we now maximise the area of the serial model with respect to its equilibrium distribution.
-First let us maximise \cref{eq:multistatearea} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
+First let us maximise \cref{eq:multistatearea:s} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
 Clearly this will happen when we put all of the probability at the ends: $\eq_1=\eq_-$ and $\eq_n=\eq_+$ are the only non-zero $\eq_k$.
@@
 %
-This yields the area bound of equation (16) of the main text.
+This yields the area bound of equation \cref{eq:max_area} of \main.
 
@@
 %\bibliographystyle{utcaps_sl}
+\begingroup
+\renewcommand{\bibitem}[1]{\item[\citenum{#1}.]}
 \bibliography{maths,neuro,markov}
+\endgroup
 

### 2018-03-19 Paper/synapse_main.tex
separated main, sup, etc

--- 2018-03-17 Paper/SynapsePaper.tex
@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
-% Note that the use of elements such as single-column equations
-% may affect the guide line number alignment.
-
-\templatetype{pnasresearcharticle} % Choose template
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ADDITIONAL OPTIONAL STYLE FILES
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{graphicx}
-%\usepackage{graphics} for finding documentation only
-\pdfsuppresswarningpagegroup=1
-\usepackage{epstopdf}
-\epstopdfsetup{update,suffix=-generated}
-%
-\IfFileExists{"C:/Progra\string~1/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Progra\string~1/Inkscape/}}{}
-\IfFileExists{"/Applications/Inkscape.app/Contents/Resources/bin/inkscape"}{\newcommand{\inkscapedir}{/Applications/Inkscape.app/Contents/Resources/bin/}}{}
-\providecommand{\inkscapedir}{}
-\epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{%
-  "\inkscapedir inkscape" -f #1 -C -A \OutputFile}
-%
-\usepackage{adjustbox}
-\newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
-%\usepackage[inline]{enumitem}
-\newlist{myenuma}{enumerate*}{4}
-\setlist[myenuma]{label=(\alph*),ref=\alph*}
-\newcommand{\panel}[1]{\item\aligntop{#1}}
-\graphicspath{{"Figs/"}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{amssymb,amsfonts}
-\usepackage{amsthm}
-%\usepackage[notref,notcite]{showkeys}
-%\usepackage[centertags]{amsmath}
-%\usepackage{bibunits}
-%\usepackage{hyperref}
-\hypersetup{bookmarksopenlevel=1,bookmarksnumbered}
-\usepackage[capitalise,nameinlink]{cleveref}
-%\Crefname{appendix}{App.}{Apps.}
-\creflabelformat{equation}{[#2#1#3]}
-\crefrangelabelformat{equation}{[#3#1#4--#5#2#6]}
-%\crefformat{equation}{#2Eq.~[#1]#3}
-%\crefrangeformat{equation}{Eqs.~[#3#1#4--#5#2#6]}
-%\crefmultiformat{equation}{Eqs.~[#2#1#3}{,2#1#3]}{,#2#1#3}{,#2#1#3]}
-%\crefrangemultiformat{equation}{Eqs.~[#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}{,#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}
-\newcommand{\main}{the main paper}
-\newcommand{\supp}{supporting information}
-% ----------------------------------------------------------------
-\newtheoremstyle{sldefinition}%
-  {3pt}%space above
-  {3pt}%space below
-  {}%body font
-  {}%indent amount
-  {\bfseries}%theorem head font
-  {}%theorem head punctuation
-  {\newline}%space after head
-  {\thmname{#1}\thmnumber{ #2}:{\bfseries\thmnote{ #3}}}%head spec
-\newtheoremstyle{slplain}%
-  {3pt}%space above
-  {3pt}%space below
-  {}%body font
-  {}%indent amount
-  {\bfseries}%theorem head font
-  {}%theorem head punctuation
-  {\newline}%space after head
-  {\thmname{#1}\thmnumber{ #2}{\bfseries\thmnote{ (#3)}}: }%head spec
-%
-\theoremstyle{slplain}
-\newtheorem{thm}{Theorem}
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{lem}[thm]{Lemma}
-%
-\theoremstyle{sldefinition}
-\newtheorem{defn}{Definition}
-%
-\theoremstyle{remark}
-\newtheorem*{notn}{Notation}
-\newtheorem*{rem}{Remark}
-% ----------------------------------------------------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL MACRO DEFINITIONS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\ra}{\rightarrow}
-\newcommand{\lr}{\leftrightarrow}
-\newcommand{\cdt}{\!\cdot\!}
-\newcommand{\vp}{\vspace{0.5cm}}
-\newcommand{\hp}{\hspace{0.5cm}}
-%
-%e.g., i.e. with normal spaces
-\newcommand{\eg}{e.g.\ }
-\newcommand{\ie}{i.e.\ }
-\newcommand{\cf}{cf.\ }
-\newcommand{\etc}{etc.\ }
-\newcommand{\wrt}{wrt.\ }
-\newcommand{\Eg}{E.g.\ }
-\newcommand{\Ie}{I.e.\ }
-\newcommand{\Cf}{Cf.\ }
-\newcommand{\Etc}{Etc.\ }
-%
-\usepackage{delim}
-% brackets etc.
-\delimdef\prn#1{\dleft( #1 \dright)}
-\delimdef\brc#1{\dleft\{ #1 \dright\}}
-\delimdef\brk#1{\dleft[ #1 \dright]}
-\delimdef\abs#1{\dleft\lvert #1 \dright\rvert}
-\delimdef\nrm#1{\dleft\lVert #1 \dright\rVert}
-\delimdef\av#1{\dleft\langle #1 \dright\rangle}
-\delimdef\avc#1#2{\dleft\langle #1 \dmiddle\vert #2 \dright\rangle}
-%\newenvironment{cases}{\dleft\{\begin{aligned}}{\end{aligned}\dright.}
-%
-% Sets
-\delimdef\set#1#2{\dleft\{ #1 \dmiddle\vert #2 \dright\}}
-\delimdef\cond#1#2{\!\dleft( #1 \dmiddle\vert #2 \dright)}
-\delimdef\condb#1#2{\!\dleft[ #1 \dmiddle\vert #2 \dright]}
-%
-% Derivatives, etc. First argument is optional.
-\newcommand{\diff}[3][]{\frac{\mathrm{d}^{#1} #2}{\mathrm{d}{#3}^{#1}}}
-\newcommand{\pdiff}[3][]{\frac{\partial^{#1} #2}{\partial {#3}^{#1}}}
-\newcommand{\pdiffc}[3][]{\left(\frac{\partial #2}{\partial {#3}}\right)_{\!\!#1}}
-\newcommand{\intd}[2][]{\int #1\!\!\dr #2\,}
-%
-% Un-italicised letters
-\newcommand{\dr}{\mathrm{d}}
-\newcommand{\e}{\mathrm{e}}
-\newcommand{\ir}{\mathrm{i}}
-\DeclareMathOperator{\tr}{tr}
-\DeclareMathOperator{\Tr}{Tr}
-\DeclareMathOperator{\Det}{Det}
-\DeclareMathOperator{\cov}{Cov}
-\DeclareMathOperator{\var}{Var}
-\DeclareMathOperator{\bias}{Bias}
-\DeclareMathOperator{\rank}{rank}
-\DeclareMathOperator{\sgn}{sgn}
-\DeclareMathOperator{\sech}{sech}
-\DeclareMathOperator{\diag}{diag}
-%
-% The default \Im and \Re look crap
-\renewcommand{\Im}{\operatorname{\mathfrak{Im}}}
-\renewcommand{\Re}{\operatorname{\mathfrak{Re}}}
-%
-% logic
-\newcommand{\means}{\Longleftrightarrow}
-\newcommand{\requires}{\Longleftarrow}
-%
-%additional symbols:
-\newcommand{\half}{\frac{1}{2}}
-\newcommand{\CO}{\mathcal{O}}
-\newcommand{\CL}{\mathcal{L}}
-\newcommand{\dt}{\dr t}
-%
-\DeclareMathOperator{\SNR}{SNR}
-\DeclareMathOperator{\snr}{SNR}
-\newcommand{\wv}{\vec{w}}
-\newcommand{\wvi}{\vec{w}_\text{ideal}}
-%matrices
-\newcommand{\inv}{^{-1}}
-\newcommand{\dg}{^\mathrm{dg}}
-\newcommand{\trans}{^\mathrm{T}}
-\newcommand{\I}{\mathbf{I}}
-%vec of ones
-\newcommand{\onev}{\mathbf{e}}
-%mat of ones
-\newcommand{\onem}{\mathbf{E}}
-%Markov matrix
-\newcommand{\MM}{\mathbf{Q}}
-%prob distributions
-\newcommand{\prob}{\mathbf{p}}
-\newcommand{\eq}{\boldsymbol{\pi}}
-%first passage times
-\newcommand{\fpt}{\mathbf{T}}
-%off-diag first passage times
-\newcommand{\fptb}{\overline{\fpt}}
-%fundamental matrix
-\newcommand{\fund}{\mathbf{Z}}
-%other symbols for matrices
-\newcommand{\Pb}{\mathbf{P}}
-\newcommand{\D}{\mathbf{D}}
-\newcommand{\arow}{\boldsymbol{\xi}}
-\newcommand{\Lb}{\boldsymbol{\Lambda}}
-\newcommand{\w}{\mathbf{w}}
-\newcommand{\W}{\mathbf{W}}
-\newcommand{\M}{\mathbf{M}}
-\newcommand{\enc}{\mathbf{q}}
-\newcommand{\frg}{\W^{\mathrm{F}}}
-\newcommand{\F}{\boldsymbol{\Phi}}
-%superscripts
-\newcommand{\pot}{^{\text{pot}}}
-\newcommand{\dep}{^{\text{dep}}}
-\newcommand{\potdep}{^{\text{pot/dep}}}
-%sets
-\newcommand{\CS}{\mathcal{S}}
-\newcommand{\CA}{\mathcal{A}}
-\newcommand{\CB}{\mathcal{B}}
-\newcommand{\comp}{^\mathrm{c}}
-%eigenmodes
-\newcommand{\uv}{\mathbf{u}}
-\newcommand{\vv}{\mathbf{v}}
-\newcommand{\CI}{\mathcal{I}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Optimal synaptic strategies for different timescales of memory}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1]{Subhaneil Lahiri}
-\author[a]{Surya Ganguli}
-
-\affil[a]{Department of Applied Physics, Stanford University, Stanford CA}
-
-% Please give the surname of the lead author for the running footer
-\leadauthor{Lahiri}
-
-% Please add here a significance statement to explain the relevance of your work
-\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.}
-
-% Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any conflict of interest here.}
-%\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: sulahiri@stanford.edu}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{synaptic plasticity | learning | memory | neuroscience}
-
-%\abbreviations{SNR, signal-to-noise ratio; LTP, long term potentiation;LTD, long term depression}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\begin{article}
-\begin{abstract}
-  An incredible gulf separates theoretical models of synapses, often described solely by a single scalar value denoting the size of a postsynaptic potential, from the immense complexity of molecular signaling pathways underlying real synapses.
-  To understand the functional contribution of such molecular complexity to learning and memory, it is essential to expand our theoretical conception of a synapse from a single scalar to an entire dynamical system with many internal molecular functional states.
-  Moreover, theoretical considerations alone demand such an expansion; network models with scalar synapses assuming finite numbers of distinguishable synaptic strengths have strikingly limited memory capacity.
-  This raises the fundamental question, how does synaptic complexity give rise to memory?
-  To address this, we develop new mathematical theorems elucidating the relationship between the structural organization and memory properties of complex synapses that are themselves molecular networks.
-  Moreover, in proving such theorems, we uncover a framework, based on first passage time theory, to impose an order on the internal states of complex synaptic models, thereby simplifying the relationship between synaptic structure and function.
-
-  Overall, we uncover general design principles governing the functional organization of complex molecular networks, and suggest new experimental observables in synaptic physiology, based on first %passage time theory, that connect molecular complexity to memory.
-\end{abstract}
-
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\setboolean{displaywatermark}{false}
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{article}
-
-% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
-% You should only change this length when you've finalised the article contents.
-\verticaladjustment{-2pt}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
 \dropcap{I}t is widely thought that our very ability to remember the past over long time scales depends crucially on our ability to modify synapses in our brain in an experience dependent manner.
@@
 Along the way we develop principles based on first passage time theory to order the structure of synaptic dynamical systems and relate this structure to memory performance.
-We summarize our main results in the \hyperref[sec:disc]{discussion section}.
+We summarize our main results in the \hyperref[sec:disc]{discussion \lcnamecref{sec:disc}}.
 
@@
 The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
-A detailed derivation of this formula can be found in the \supp.
+A detailed derivation of this formula can be found in \cref{sec:SNRcurve} of the \supp.
 We will frequently refer to this function as the memory curve.
@@
 We can then ask the question: what are the constraints on these quantities, namely eigenmode initial SNR's, $\CI_a$, and time constants, $\tau_a$, implied by the constraints in \cref{eq:constr}?
-We will derive some of these constraints in the next section.
+We will derive some of these constraints in the next \lcnamecref{sec:bounds}.
 
@@
 
-In the previous section, in \cref{eq:SNRcalc} we have described an analytic expression for a memory curve as a function of the structure of a synaptic dynamical system, described by the pair of stochastic transition matrices  $\M\potdep$.
+In the previous \lcnamecref{sec:setup}, in \cref{eq:SNRcalc} we have described an analytic expression for a memory curve as a function of the structure of a synaptic dynamical system, described by the pair of stochastic transition matrices  $\M\potdep$.
 Since the performance measure for memory is an entire memory curve, and not just a single number, there is no universal scalar notion of optimal memory in the space of synaptic dynamical systems.
 Instead there are tradeoffs between storing proximal and distal memories;  often attempts to increase memory at late (early) times by changing $\M\potdep$, incurs a performance loss in memory at early (late) times in specific models considered so far \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial}.
-Thus our end goal, achieved \hyperref[sec:env]{below}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the {\it entire} memory curve of {\it any} model.
+Thus our end goal, achieved \hyperref[eq:env]{below}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the {\it entire} memory curve of {\it any} model.
 In order to achieve this goal, in this section, we must first derive upper bounds, over the space of all possible synaptic models, on two different scalar functions of the memory curve: its initial SNR, and the area under the memory curve.
@@
 These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-They are completely equivalent to a two state model with transition probabilities equal to 1, as shown in \cref{fig:max_init}(\ref{fig:binary_det}).
-In particular, they have the same memory curve.
+In terms of their memory curves, they are completely equivalent to a two state model with transition probabilities equal to 1, as shown in \cref{fig:max_init}(\ref{fig:binary_det}).
 
@@
 \footnote{Note that we do not need to worry about the order of the $\eta^\pm_i$ changing during the optimization: necessary conditions for a maximum only require that there is no infinitesimal perturbation that increases the area.
-Therefore we need only consider an infinitesimal neighborhood of the model, in which the order will not change.\label{ft:etaorder}}
+Therefore we need only consider an infinitesimal neighborhood of the model, in which the order will not change.}%\label{ft:etaorder}}
 
@@
 
-First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \cref{eq:kemenypm} (see \supp):
+First passage time theory introduced in the \hyperref[sec:order]{previous \lcnamecref{sec:order}} becomes useful because the area has a simple expression in terms of quantities introduced in \cref{eq:kemenypm} (see \supp, \cref{sec:areadef}):
 %
@@
 With the states in the order described above, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
-Some of these perturbations are shown in \cref{fig:perts}, see \supp\ for details.
+Some of these perturbations are shown in \cref{fig:perts}, see \cref{sec:kuhntucker} of the \supp\ for details.
 For example, in \cref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
@@
 
-Now we will look at the implications of the upper bounds found in the \hyperref[sec:bounds]{previous section} for the SNR at finite times.
+Now we will look at the implications of the upper bounds found in the \hyperref[sec:bounds]{previous \lcnamecref{sec:bounds}} for the SNR at finite times.
 As argued in \cref{eq:SNReigen}, the memory curve can be written in the form
@@
 \showacknow % Display the acknowledgments section
-
-% \pnasbreak splits and balances the columns before the references.
-% If you see unexpected formatting errors, try commenting out this line
-% as it can run into problems with floats and footnotes on the final page.
-%\pnasbreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{pnas-new}
-\bibliography{maths,neuro,markov}
-
-%\end{article}
-%\clearpage
-
-\begin{strip}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\appendix
-\pdfbookmark[0]{Supporting information}{appendix}
-\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
-\begin{center}
-  \textbf{\Large Supporting information}
-\end{center}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Here we provide more details underlying the derivations of results in \main.
-
-\section{Continuous time Markov processes}\label{sec:ContMarkov}
-
-In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time to define notation.
-It is a generalization of material that can be found in \cite{kemeny1960finite} with some ideas from \cite{hunter2000survey}.
-
-\subsection{Notation}\label{sec:not}
-
-For any matrix $\mathbf{A}$, we define matrices $\mathbf{A}\dg$ and $\overline{\mathbf{A}}$ as
-%
-\begin{equation}\label{eq:dgdef}
-  \mathbf{A}\dg_{ij} \equiv \delta_{ij}\mathbf{A}_{ij},
-  \qquad
-  \overline{\mathbf{A}} \equiv \mathbf{A}-\mathbf{A}\dg.
-\end{equation}
-%
-We let $\onev$ denote a column-vector of ones and $\onem=\onev\onev\trans$ denote a matrix of ones.
-
-A continuous time Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$ with row sums equal to zero ($\MM\onev=0$).
-The probabilities of being in each state at time $t$, the row-vector $\mathbf{p}(t)$, evolve according to
-%
-\begin{equation}\label{eq:rowsum}
-  \diff{\mathbf{p}(t)}{t} = \mathbf{p}(t) \MM,
-\end{equation}
-%
-where $\mathbf{p}(t)\onev=1$.
-
-The equilibrium probabilities, $\eq$, satisfy
-%
-\begin{equation}\label{eq:equilibrium}
-  \eq\MM=0,
-  \qquad
-  \eq\onev=1.
-\end{equation}
-%
-As we assume an ergodic process, this eigenvalue is non-degenerate.
-If all other eigenvalues have strictly negative real parts, the process is regular (aperiodic).
-
-We define additional matrices
-%
-\begin{equation}\label{eq:defDLP}
-  \Lb \equiv (-\MM\dg)\inv,
-  \qquad
-  \Pb \equiv \I + \Lb\MM.
-\end{equation}
-%
-It can be shown that $\Lb_{ii}$ is the mean time it takes to leave state $i$ and $\Pb_{ij}$ is the probability the the next transition from state $i$ goes to state $j$:
-%
-\begin{equation}\label{eq:LamdaPcmpt}
-  \Lb_{ii} = \frac{1}{\sum_{j \neq j} \MM_{ij}},
-  \qquad
-  \Pb_{ij} = %\left\{
-%  \begin{aligned}
-%     &0                                         && \text{if }i=j, \\
-%     &\frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &&\text{otherwise}.
-%  \end{aligned}
-%  \right.
-  \begin{cases}
-     0                                         &\text{if }i=j, \\
-     \frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &\text{otherwise}.
-  \end{cases}
-\end{equation}
-%
-Furthermore, we also define
-%
-\begin{equation}\label{eq:pdotD}
-  \D \equiv \operatorname{diag}(\eq)\inv,
-  \qquad\implies\qquad
-  \eq\D=\onev\trans.
-\end{equation}
-%
-
-\subsection{Fundamental matrix}\label{sec:fund}
-
-For our results below regarding the integral of the memory curve, it can be useful to invert the stochastic transition matrix, $\MM$.
-However, since $\MM$ has a zero eigenvalue, it cannot be inverted.
-For this reason, the fundamental matrix arises as a useful surrogate for the inverse of $\MM$.
-It is related to the first passage times, as we will see in the \hyperref[sec:fpt]{next subsection}.
-Here we define the fundamental matrix and review its properties.
-
-\begin{defn}[Fundamental matrix]
-  For discrete time, the generalized fundamental matrix was defined in \cite{Kemeny1981fund}.
-  For continuous time, we define:
-  %
-  \begin{equation}\label{eq:funddef}
-    \fund \equiv (-\MM + \onev\arow)\inv,
-  \end{equation}
-  %
-  where $\arow$ is any row-vector with $\arow\onev=1/\tau\neq0$.
-\end{defn}
-Note that the canonical choice for the discrete time version, $\arow=\eq$, is not available here due to problems with units.
-It will be helpful to choose $\arow$ to be independent of $\MM$, \eg $\arow=\onev\trans/(n\tau)$.
-All quantities that we calculate using $\fund$ below will be independent of this choice.
-
-\begin{thm}
-  The definition of $\fund$ is valid, \ie $(-\MM + \onev\arow)$ is invertible.
-\end{thm}
-\begin{proof}
-  Assume there exists an $\mathbf{x}$ such that
-  %
-  \begin{equation}\label{eq:fundinvkern}
-    (-\MM + \onev\arow)\mathbf{x}=0.
-  \end{equation}
-  %
-  Multiplying from the left with $\eq$ gives
-  %
-  \begin{equation}\label{eq:fundinvpix}
-    \arow \mathbf{x} = 0.
-  \end{equation}
-  %
-  Substituting back into \cref{eq:fundinvkern} gives
-  %
-  \begin{equation*}
-    \MM\mathbf{x}=0.
-  \end{equation*}
-  %
-  As we assume an ergodic process, the zero eigenvalue is non-degenerate.
-  Therefore, $\mathbf{x}=\lambda\onev$.
-  Substituting this into \cref{eq:fundinvpix} gives
-  %
-  \begin{equation*}
-    \lambda\arow\onev = \frac{\lambda}{\tau} = 0.
-  \end{equation*}
-  %
-  As we defined $\arow$ such that $1/\tau\neq0$, this means $\lambda=0 \implies \mathbf{x}=0$.
-\end{proof}
-
-\begin{cor}
-  %
-  \begin{align}
-    \arow\fund &= \eq, \label{eq:fundprob}\\
-    \fund\onev &= \tau\onev,\label{eq:fundrowsum}\\
-    \I+\MM\fund &= \onev\eq, \label{eq:fundQZ}\\
-    \I+\fund\MM &= \tau\onev\arow. \label{eq:fundZQ}
-  \end{align}
-  %
-\end{cor}
-\begin{proof}
-  We can deduce \cref{eq:fundprob,eq:fundrowsum} be pre/post-multiplying the following equations by $\fund$:
-  %
-  \begin{equation*}
-    \begin{aligned}
-      \eq(-\MM + \onev\arow) &= \arow, \\
-      (-\MM + \onev\arow)\onev &= \frac{\onev}{\tau}.
-    \end{aligned}
-  \end{equation*}
-  %
-  We can then deduce \cref{eq:fundQZ,eq:fundZQ} by substituting these into
-  %
-  \begin{equation*}
-    (-\MM + \onev\arow)\fund = \fund(-\MM + \onev\arow) = \I.
-  \end{equation*}
-  %
-\end{proof}
-
-%We will not need to use the concept of a g-inverse anywhere, so you can skip ahead to \cref{sec:fpt} now. For the sake of completeness, we can define $\MM^\#$ as
-%%
-%\begin{equation}\label{eq:ginvdef}
-%  \MM^\# \equiv \tau\onev\eq - \fund
-%\end{equation}
-%%
-%\begin{thm}
-%  The matrix $\MM^\#$ satisfies the following defining properties of a g-inverse of $\MM$:
-%  %
-%  \begin{align}
-%    \MM^\# \MM \MM^\# &= \MM^\#,     \label{eq:ginvpropo} \\
-%    \MM \MM^\# \MM    &= \MM,        \label{eq:ginvpropi} \\
-%    \MM \MM^\#        &= \MM^\# \MM, \label{eq:ginvcomm}
-%  \end{align}
-%  %
-%  with \cref{eq:ginvcomm} holding iff $\eq$ and $\arow$ are proportional to each other.
-%\end{thm}
-%\begin{proof}
-%  Using \cref{eq:rowsum}, \cref{eq:equilibrium}, \cref{eq:fundprob} and \cref{eq:fundrowsum}, one can verify
-%  %
-%  \begin{alignat*}{2}
-%    \MM \MM^\# &= -\MM \fund,
-%    &\qquad
-%    \arow \MM^\# &= 0,
-%    \\
-%    \MM^\# \MM &= -\fund \MM,
-%    &
-%    \MM^\# \onev &= 0.
-%  \end{alignat*}
-%  %
-%  Substituting into \cref{eq:fundQZ} and \cref{eq:fundZQ} leads to
-%  %
-%  \begin{equation*}
-%    \I - \MM \MM^\# = \onev \eq,
-%    \qquad
-%    \I - \MM^\# \MM = \tau \onev \arow.
-%  \end{equation*}
-%  %
-%  Pre/post-multiplying these by $\MM$ or $\MM^\#$ produces \cref{eq:ginvpropo} and \cref{eq:ginvpropi}. Subtracting produces
-%  %
-%  \begin{equation*}
-%    [\MM,\MM^\#] = \onev(\tau\arow-\eq),
-%  \end{equation*}
-%  %
-%  which vanishes iff $\eq$ and $\arow$ are proportional to each other (due to $\arow\onev=1/\tau$ and $\eq\onev=1$, the constant of proportionality could only be $\tau$).
-%\end{proof}
-
-\subsection{First passage times}\label{sec:fpt}
-
-\begin{defn}[First passage time matrix]
- We define $\fptb_{ij}$ as the mean time it takes the process to reach state $j$ for the first time, starting from state $i$.
- We also define $\fpt\dg_{ii}$ as the mean time it takes the process to return to state $i$.
- As usual, $\fpt=\fptb+\fpt\dg$.
-\end{defn}
-
-This matrix is given by
-%
-\begin{equation}\label{eq:fpt}
-  \fpt=(\onem\fund\dg - \fund + \Lb)\D,
-\end{equation}
-%
-see \cite{Yao1985fpt} for a proof.
-We can separate this into its diagonal and off-diagonal pieces.
-
-The recurrence times are given by
-%
-\begin{equation}\label{eq:recurtime}
-  \fpt\dg = \Lb\D.
-\end{equation}
-%
-or in component form
-%
-\begin{equation*}
-  \eq_i \Lb_{ii}\inv \fpt\dg_{ii} = 1.
-\end{equation*}
-%
-The extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.
-
-The off-diagonal mean first passage times are given by
-%
-\begin{equation}\label{eq:fptfund}
-  \fptb = (\onem\fund\dg - \fund)\D.
-\end{equation}
-%
-or in component form:
-%
-\begin{equation}\label{eq:fptfundcmpt}
-  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
-\end{equation}
-%
-
-\subsection{Mixing time (Kemeny's constant)}\label{sec:mixtime}
-
-\begin{thm} \mbox{}\label{th:kemenyconst}
-  The quantity
-  %
-  \begin{equation}\label{eq:mixdef}
-    \eta \equiv \sum_j \fptb_{ij}\eq_j
-  \end{equation}
-  %
-  is independent of $i$.
-\end{thm}
-\begin{proof}
-  For discrete time, a proof can be found in \cite[Th.4.4.10]{kemeny1960finite}.
-  For continuous time, we use \cref{eq:fptfund}, \cref{eq:fundrowsum} and the transpose of \cref{eq:pdotD}:
-  %
-  \begin{equation*}
-    \begin{aligned}
-      \fptb (\eq)\trans &= (\onem\fund\dg - \fund) \D (\eq)\trans \\
-        &= (\onev\onev\trans\fund\dg - \fund) \onev \\
-        &= (\onev\trans\fund\dg\onev)\onev - \fund\onev \\
-        &= (\tr\fund - \tau)\onev.
-    \end{aligned}
-  \end{equation*}
-  %
-  which proves \cref{eq:mixdef} with $\eta = \tr\fund - \tau$.
-\end{proof}
-
-Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version, where this would only shift $\eta$ by 1.
-
-\subsection{Sensitivity of equilibrium distribution}\label{sec:sensitivity}
-
-Suppose that the Markov process, defined by $\MM$, depends on some parameter $\alpha$.
-Differentiating \cref{eq:funddef} gives
-%
-\begin{equation}\label{eq:diffZ}
-  \diff{\fund}{\alpha} = \fund \diff{\MM}{\alpha} \fund.
-\end{equation}
-%
-We can substitute this into the derivative of \cref{eq:fundprob}:
-%
-\begin{equation}\label{eq:diffp}
-  \diff{\eq}{\alpha} = \arow \fund \diff{\MM}{\alpha} \fund = \eq \diff{\MM}{\alpha} \fund.
-\end{equation}
-%
-We can rewrite this in component form and use the fact that $\MM_{ii} = - \sum_{i\neq j} \MM_{ij}$:
-%
-\begin{equation}\label{eq:diffpT}
-\begin{aligned}
-  \diff{\eq_k}{\alpha} &= \sum_{i,j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} \\
-    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} + \sum_i \eq_i \diff{\MM_{ii}}{\alpha} \fund_{ik} \\
-    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} (\fund_{jk} - \fund_{ik}) \\
-    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} \eq_k (\fptb_{ik} - \fptb_{jk}).
-\end{aligned}
-\end{equation}
-%
-This is a generalization of a result of \cite{cho2000markov} from discrete to continuous time that we will need below.
-Note that the summand vanishes for $i=j$, so we can drop the restriction $i\neq j$ from the range of the sum.
-
-\subsection{Subsets and flux}\label{sec:subsets}
-
-Let us denote the set of states by $\CS$.
-Consider a subset $\CA\subset\CS$.
-We can define a projection operator onto this subset:
-%
-\begin{equation}\label{eq:proj}
-  \prn{\I^\CA}_{ij} =
-    \begin{cases}
-      1 &\text{if $i=j\in\CA$,}\\
-      0 &\text{otherwise.}
-    \end{cases}
-\end{equation}
-%
-We will use superscripts/subscripts to denote projection onto/summation over a subset:
-%
-\begin{equation}\label{eq:projsum}
-  \begin{aligned}
-    \arow^\CA &=\arow\I^\CA, &
-    \M^{\cdot\CA}&=\M\I^\CA, &
-    \M^{\CA\cdot}&=\I^\CA\M, &
-    \mathbf{x}^\CA &= \I^\CA\mathbf{x},
-    \\
-    \arow_\CA &=\arow\onev^\CA, &
-    \M_{\cdot\CA}&=\M\onev^\CA, &
-    \M_{\CA\cdot}&=\prn{\onev^\CA}\trans\M, &
-    \mathbf{x}_\CA &= \prn{\onev^\CA}\trans\mathbf{x},
-  \end{aligned}
-\end{equation}
-%
-where $\arow$ is a row vector, $\M$ is a matrix and $\mathbf{x}$ is a column vector.
-
-We can define a flux matrix, a.k.a.\ ergodic flow:
-%
-\begin{equation}\label{eq:flux:s}
-  \F = \D\inv\MM,
-  \qquad
-  \F_{ij} = \eq_i \MM_{ij}.
-\end{equation}
-%
-This measures the flow of probability between states in the equilibrium distribution.
-Detailed balance, a.k.a.\ reversibility, is equivalent to $\F=\F\trans$.
-
-The flux between two subsets is a particularly useful quantity:
-%
-\begin{equation}\label{eq:subflux}
-  \F_{\CA\CB} = {\eq}^\CA\MM\onev^\CB.
-\end{equation}
-%
-One can show that
-%
-\begin{equation}\label{eq:compflux}
-  \F_{\CA\CA\comp} = \F_{\CA\comp\CA} = -\F_{\CA\CA} = -\F_{\CA\comp\CA\comp}
-\end{equation}
-%
-using $\prn{{\eq}^\CA + {\eq}^{\CA\comp}}\MM=0$ and $\MM\prn{{\onev}^\CA + {\onev}^{\CA\comp}}=0$.
-
-\subsection{Lumpability}\label{sec:lump}
-
-Suppose we have partitioned the states into disjoint subsets, $\brc{\CA_\alpha}$:
-%
-\begin{equation}\label{eq:partition}
-  \bigcup_\alpha \CA_\alpha = \CS,
-  \qquad
-  \CA_\alpha \cap \CA_\beta = \delta_{\alpha\beta}\CA_\alpha.
-\end{equation}
-%
-We will use $\alpha$ instead of $\CA_\alpha$ in superscripts and subscripts in the following.
-The fact that these subsets are disjoint and exhaustive allows us to define the function
-%
-\begin{equation}\label{eq:whichsey}
-  \sigma(i)=\alpha
-  \qquad\means\qquad
-  i\in\CA_\alpha.
-\end{equation}
-%
-
-We can use this partition to define a new stochastic process associated with the original Markov chain.
-At time $t$, if the state of the original process is $i$, the state of the new process is $\sigma(i)$.
-
-One may ask if this new process is a Markov chain.
-The answer is yes, if the original Markov chain has a property called lumpability \wrt the partition
-(see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time):
-%
-\begin{equation}\label{eq:lump}
-  \sigma(i)=\sigma(j)
-  \quad\implies \quad
-  \MM_{i\alpha}=\MM_{j\alpha} \equiv \sum_{k\in\CA_\alpha} \MM_{jk},
-\end{equation}
-%
-\ie the total transition rate from some state to some subset is the same for all starting states within the same subset.
-This common value is the transition rate for the new lumped Markov chain.
-
-This can be rewritten with the aid of two matrices
-%
-\begin{equation}\label{eq:lumpmats}
-  U_{\alpha i} = \frac{\delta_{\alpha\sigma(i)}}{\abs{\CA_\alpha}},
-  \qquad
-  V_{i\alpha} = \delta_{\sigma(i)\alpha}.
-\end{equation}
-%
-Left multiplication by $U$ averages over subsets, right multiplication by $V$ sums over subsets.
-For $U$, we chose the uniform measure in each subset. Any measure would work equally well, \eg one proportional to the equilibrium distribution:
-%
-\begin{equation}\label{eq:altlumpmats}
-  U_{\alpha i} = \frac{{\eq_i}^\alpha}{\eq_\alpha}.
-\end{equation}
-%
-
-One can show that $(UV)_{\alpha\beta}=\delta_{\alpha\beta}$.
-The matrix $VU$ is also interesting.
-It has a block diagonal structure, with each block corresponding to a subset.
-Each block is a discrete-time ergodic Markov matrix (it is an independent trials process with probabilities given by the measure chosen for $U$).
-This means that the right eigenvectors with eigenvalue 1 will be those that are constant in each subset:
-%
-\begin{equation}\label{eq:setconst}
-  VU\mathbf{x}=\mathbf{x}
-  \qquad\means\qquad
-  \mathbf{x} = \sum_\alpha x_\alpha \onev^\alpha.
-\end{equation}
-%
-
-This allows us to write the lumpability condition \cref{eq:lump}, and the transition matrix for the lumped process compactly:
-%
-\begin{equation}\label{eq:lumpcompact}
-  VU\MM V = \MM V,
-  \qquad
-  \widehat{\MM} = U\MM V.
-\end{equation}
-%
-By induction, one can show that similar relations hold for all powers:
-%
-\begin{equation}\label{eq:lumppower}
-  VU\MM^nV = \MM^nV,
-  \qquad
-  \widehat{\MM}^n = U\MM^n V,
-\end{equation}
-%
-and, via the Taylor series, for the exponential as well:
-%
-\begin{equation}\label{eq:lumpexp}
-  VU \e^{t\MM} V = \e^{t\MM}V,
-  \qquad
-  \e^{t\widehat{\MM}} = U\e^{t\MM} V.
-\end{equation}
-%
-The equilibrium distribution of the lumped process is given by
-%
-\begin{equation}\label{eq:lumpeq}
-  \widehat{\mathbf{p}}^\infty = \eq V.
-\end{equation}
-%
-
-\section{Signal-to-Noise ratio (SNR)}\label{sec:SNR}
-
-In this section we will look at the signal-to-noise curve, and put an upper bound on its initial value.
-We need only consider ergodic Markov chains.
-Transient states would be unoccupied in equilibrium and would not be accessed by the signal creation process, therefore they could be removed from the analysis.
-Absorbing chains are degenerate cases: they have zero initial signal but infinite decay times, so they can only be approached as the limit of a sequence of ergodic chains.
-
-\subsection{Framework}\label{sec:framework}
-
-The individual potentiation/depression events will be described by \emph{discrete}-time Markov chains:
-%
-\begin{equation}\label{eq:MWdef}
-  \M\potdep \equiv \I + \W\potdep,
-  \qquad
-  \M\potdep\onev = \onev,
-  \qquad
-  \M\potdep_{ij} \in [0,1].
-\end{equation}
-%
-The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf\potdep$, where $f\pot +f\dep =1$ are the fraction of plasticity events that are potentiating/depressing respectively.
-This means that the ``forgetting'' process will be described by the \emph{continuous}-time Markov chain:
-%
-\begin{equation}\label{eq:forgetting}
-  \MM = r \frg \equiv r\prn{f\pot\W\pot + f\dep\W\dep}.
-\end{equation}
-%
-We only require that this Markov chain is ergodic.
-The Markov chains described by $\M\potdep $ need not be.
-
-We assume that the probability distribution starts in the equilibrium distribution \cref{eq:equilibrium}.
-During the initial signal creation, a fraction $f\pot$ will change to $\eq\M\pot$ and a fraction $f\dep$ will change to $\eq\M\dep$.
-After this, probabilities will evolve according to \cref{eq:rowsum}.
-
-\subsection{SNR curve}\label{sec:SNRcurve}
-
-As discussed in \main, the signal-to-noise ratio is given by
-%
-\begin{equation}\label{eq:SNRdef:s}
-  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}}
-     {\sqrt{\var(\wv_\text{ideal}\cdt\wv(\infty))}}.
-\end{equation}
-%
-First, let's look at the denominator, remembering that the states and plasticity events of each synapse are independent and identically distributed:
-%
-\begin{equation}\label{eq:noise}
-\begin{aligned}
-  \var(\wv_\text{ideal}\cdt\wv(\infty))
-    &= \sum_{\alpha\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty) \wv^\beta_\text{ideal}\wv^\beta(\infty)}
-    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2 \\
-    &= \sum_{\alpha} \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}
-    + \sum_{\alpha\neq\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}\av{\smash{\wv^\beta}_\text{ideal}\smash{\wv^\beta}(\infty)}
-%    \\&\phantom{= \sum_{\alpha}  \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}}
-    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2  \\
-    &= N\av{1}
-    + N(N-1)\av{\wv^1_\text{ideal}\wv^1(\infty)}^2
-    - N^2\av{\wv^1_\text{ideal}\wv^1(\infty)}^2 \\
-    &= N(1-\av{\wv^1_\text{ideal}\wv^1(\infty)}^2),
-\end{aligned}
-\end{equation}
-%
-where we used $\wv^\alpha=\pm1$.
-
-For the numerator, we can write
-%
-\begin{equation}\label{eq:overlapex}
-  \av{\wv_\text{ideal}\cdt\wv(t)} = \sum_\alpha\av{\wv^\alpha_\text{ideal}\wv^\alpha(t)}
-   = N \av{\wv^1_\text{ideal}\wv^1(t)},
-\end{equation}
-%
-Noting that $\wvi=\pm1$ with probability $f\potdep$,
-%
-\begin{equation}\label{eq:overlapstate}
-\begin{aligned}
-  \av{\wv^1_\text{ideal}\wv^1(t)} &= f\pot \av{\wv^1(t)}_{\text{pot},t=0} - f\dep \av{\wv^1(t)}_{\text{dep},t=0} \\
-   &= f\pot \sum_i P(\text{state}=i,t \mid \text{pot},0) \w_i - f\dep \sum_i P(\text{state}=i,t \mid \text{dep},0) \w_i.
-\end{aligned}
-\end{equation}
-%
-From the previous section,
-%
-\begin{equation}\label{eq:stateprob}
-  P(\text{state}=i,t \mid \text{pot/dep},0) = \brk{ \eq \M\potdep\, \e^{rt\frg} }_i,
-\end{equation}
-%
-which describes the synapses starting in the equilibrium distribution, changing state due to the plasticity event at $t=0$ and subsequent evolution according to \cref{eq:rowsum} due to plasticity events uncorrelated with $\wvi$.%
-\footnote{Note that expanding the exponential gives
-%
-\begin{equation*}
-  \e^{rt\frg} =  \sum_{n=0}^\infty \frac{(rt)^n\,\e^{-rt}}{n!} \sum_{m=0}^n (f\pot)^m (f\dep)^{n-m}
-  \brk{ \M\pot \M\dep \M\pot \M\pot \ldots + \text{permutations}}.
-\end{equation*}
-%
-Thus, evolving according to \cref{eq:rowsum} results in averaging over all sequences of plasticity events,
-as we only need linear expectations of $\wv(t)$ in the end.\label{ft:avexp}}
-This results in
-%
-\begin{equation}\label{eq:overlapwt}
-\begin{aligned}
-  \av{\wv^1_\text{ideal}\wv^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
-  \av{\wv^1_\text{ideal}\wv^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
-         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
-         &=  (f\pot-f\dep)\, \eq \w \\
-         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
-\end{aligned}
-\end{equation}
-%
-Combining these allows us to write the numerator as
-%
-\begin{equation}\label{eq:signal}
-\begin{aligned}
-  \av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}
-    &= N \eq (f\pot(\M\pot-\I)-f\dep(\M\dep-\I))\, \e^{rt\frg} \w \\
-    &= N \eq (f\pot(\W\pot-\frg)-f\dep(\W\dep-\frg))\, \e^{rt\frg} \w \\
-%    &= N \eq (f\pot(f\dep\W\pot-f\dep\W\dep)-f\dep(f\pot\W\dep-f\pot\W\pot))\,\e^{rt\frg}\w \\
-    &= N (2f\pot f\dep)\eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
-\end{aligned}
-\end{equation}
-%
-where we used $\eq\frg=0$ in going from the first to second lines.
-Combining with \cref{eq:noise} gives
-%
-\begin{equation}\label{eq:SNRcurveExact}
-  \SNR(t) = \frac{\sqrt{N} (2f\pot f\dep ) \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w}
-                 {\sqrt{1-(f\pot-f\dep)^2(\eq_+-\eq_-)^2}}.
-\end{equation}
-%
-The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
-We can set the denominator to 1 without changing any of our results.
-
-This results in our final formula:
-%
-\begin{equation}\label{eq:SNRcurve}
-  \SNR(t) = \sqrt{N} (2f\pot f\dep )\, \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
-\end{equation}
-%
-The factor of $\eq$ describes the synapses being in the steady-state distribution before the memory is encoded.
-The factor of $(\M\pot-\M\dep)$ comes from the encoding of the memory at $t=0$, with $\wvi$ being $\pm1$ in synapses that are potentiated/depotentiated.
-The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
-
-We can express this in terms of the one parameter family of transition matrices:
-%
-\begin{equation}\label{eq:Walpha}
-  \begin{aligned}
-  \W(\alpha) &= \alpha\W\pot +(1-\alpha)\W\dep ,
-  &\quad&\implies&
-    \frg&=\W(f\pot ),\\&&&&
-    \W\pot  - \W\dep  &= \diff{\W}{\alpha},\\&&&&
-    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \frg.
-  \end{aligned}
-\end{equation}
-%
-Then \cref{eq:SNRcurve} becomes
-%
-\begin{equation}\label{eq:SNRalpha}
-  \SNR(t) = \sqrt{N} (2f\pot f\dep ) \diff{\eq}{\alpha}(-\frg)\,\e^{rt\frg}\w.
-\end{equation}
-%
-
-\subsection{Lumpability and the SNR curve}\label{sec:SNRlump}
-
-Suppose that we have a partition such that $\W\pot $ and $\W\dep $ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \cref{sec:lump}):
-%
-\begin{equation}\label{eq:lumpablesynapse}
-  VU\W\potdep  V = \W\potdep  V,
-  \qquad
-  VU\w=\w.
-\end{equation}
-%
-We can define a new synapse with
-%
-\begin{equation}\label{eq:lumpedsynapse}
-  \widehat{\W}\potdep  = U\W\potdep  V,
-  \qquad
-  \widehat{\w} = U \w,
-  \qquad
-  \widehat{\mathbf{p}}^\infty = \eq V.
-\end{equation}
-%
-This synapse has an SNR curve:
-%
-\begin{equation}\label{eq:lumpedSNR}
-  \begin{aligned}
-    \frac{\SNR(t)}{\sqrt{N} (2f\pot f\dep )} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}\pot  - \widehat{\W}\dep )\e^{rt\widehat{\W}^F}\widehat{\w}. \\
-      &= \eq VU (\W\pot  - \W\dep ) VU \e^{rt\frg} VU \w. \\
-      &= \eq (\W\pot  - \W\dep ) VU \e^{rt\frg} VU \w. \\
-      &= \eq (\W\pot  - \W\dep )\e^{rt\frg} VU \w. \\
-      &= \eq (\W\pot  - \W\dep )\e^{rt\frg}\w. \\
-  \end{aligned}
-\end{equation}
-%
-\ie the lumped process has exactly the same SNR as the original one.
-
-\subsection{Initial SNR and flux}\label{sec:initflux}
-
-Using $\eq\frg=0$ and the first line of \cref{eq:signal}, we can write the initial SNR as
-%
-\begin{equation}\label{eq:init:s}
-  \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f\pot \W\pot -f\dep \W\dep )(\onev^+-\onev^-).
-\end{equation}
-%
-Using $\W\potdep (\onev^++\onev^-)=0$ and \cref{eq:compflux}:
-%
-\begin{equation*}
-  r{\eq}^-(f\pot \W\pot +f\dep \W\dep )\onev^+ = \F_{-+} = \F_{+-} = r{\eq}^+(f\pot \W\pot +f\dep \W\dep )\onev^-,
-\end{equation*}
-%
-we can rewrite \cref{eq:init} as
-%
-\begin{equation}\label{eq:initflux:s}
-  I = \frac{4\F_{-+}}{r} - 4{\eq}^+f\pot \W\pot \onev^- - 4{\eq}^-f\dep \W\dep \onev^+.
-\end{equation}
-%
-The last two terms are guaranteed to be negative, as the diagonal parts of $\W\potdep $ cannot contribute.
-Therefore
-%
-\begin{equation}\label{eq:initfluxineq}
-  \SNR(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}.
-\end{equation}
-%
-This inequality is saturated if potentiation never takes it from a $+$ state to a $-$ state and depression never takes it from a $-$ state to a $+$ state.
-
-\section{Area maximisation}\label{sec:areamax}
-
-In this section we will find an upper bound on the area under the signal-to-noise curve.
-As in \cref{sec:SNR}, we will only consider ergodic Markov chains.
-We will see in \cref{sec:multistate} that the optimal chain is absorbing, so it lies on the boundary of the (open) set of ergodic chains, but it still puts an upper bound on the area for any chain in the interior.
-
-\subsection{Area under signal-to-noise curve}\label{sec:areadef}
-
-The signal-to-noise curve is given by \cref{eq:SNRalpha}.
-%%
-%\begin{equation}\label{eq:SNR}
-%  \SNR(t) = \sqrt{N}(2f^+f\dep ) \diff{\eq}{\alpha} (-\frg) \e^{rt\frg} \w.
-%\end{equation}
-%%
-The area is computed by integrating this
-%
-\begin{equation}\label{eq:area:s}
-\begin{aligned}
-  A &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\frg}}_0^\infty \w \\
-    &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
-    &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} \w.
-\end{aligned}
-\end{equation}
-%
-
-We can rewrite this using \cref{eq:diffpT}, with $A=\sqrt{N}(2f\pot f\dep )\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W\pot _{ij}-\W\dep _{ij}$
-%
-\begin{equation}\label{eq:areaT}
-  \hat{A} = \sum_{i,j,k} \eq_i \enc_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
-\end{equation}
-%
-
-\begin{defn}[Partial mixing times]
-We define the $\pm$ mixing times as
-%
-\begin{equation}\label{eq:mixingpm}
-\begin{aligned}
-  \eta^\pm_i &\equiv \sum_k \fptb_{ik} \eq_k \prn{\frac{1 \pm \w_k}{2}}
-%    &&= \sum_{\set{k}{\w_k=\pm1}} \fptb_{ik} \eq_k \\
-    &&= \sum_{k\in\pm} \fptb_{ik} \eq_k \\
-    &= \sum_k \prn{\fund_{kk}-\fund_{ik}} \prn{\frac{1 \pm \w_k}{2}}
-    &&= \sum_{k\in\pm} \prn{\fund_{kk}-\fund_{ik}} .
-\end{aligned}
-\end{equation}
-%
-We can think of $\eta^+_i$ as a measure of the ``distance'' to the $\w_k=+1$ states and $\eta^-_i$ as the ``distance'' to the $\w_k=-1$ states.
-\end{defn}
-Using \cref{eq:mixdef}, we can write:
-%
-\begin{equation}\label{eq:mixingrels}
-\begin{aligned}
-  \eta^+_i + \eta^-_i &= \eta,\\
-  2(\eta^+_i - \eta^+_j) &= \sum_k (\fptb_{ik}-\fptb_{jk}) \eq_k \w_k
-    = \sum_k (\fund_{jk}-\fund_{ik}) \w_k.
-\end{aligned}
-\end{equation}
-%
-We could arrange the states in order of decreasing $\eta^+$, which is the same as the order of increasing $\eta^-$.
-
-We can rewrite \cref{eq:areaT} as
-%
-\begin{equation}\label{eq:areaEta:s}
-\begin{aligned}
-  \hat{A} &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}) &
-    &= -2\sum_{i,j} \enc_{ij} \eq_i \eta^+_{j} \\
-    &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^-_{j} - \eta^-_{i}) &
-    &= 2\sum_{i,j} \enc_{ij} \eq_i \eta^-_{j}.
-\end{aligned}
-\end{equation}
-%
-We can also express it in terms of the fundamental matrix \cref{eq:funddef} as
-%
-\begin{equation}\label{eq:areaZ}
-  \hat{A} = \sum_{i,j,k,l} \enc_{ij} \arow_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
-    = \arow \fund q \fund \w.
-\end{equation}
-%
-
-It is also helpful to define the following quantities:
-%
-\begin{equation}\label{eq:areacoeffs}
-  \begin{aligned}
-    c_k &= \diff{\ln\eq_k}{\alpha}
-      = \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
-      = - \prn{\eq \enc \fptb}_k
-      = \frac{(\eq \enc \fund)_k}{\eq_k}, \\
-    a_i &= \sum_j \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
-    \implies
-    \hat{A} &= \sum_k c_k \eq_k \w_k
-      = 2\sum_i a_i.
-  \end{aligned}
-\end{equation}
-%
-Note that the optimal choice of $\w$ is $\w_k = \sgn(c_k)$.
-
-\subsection{Derivatives \wrt \texorpdfstring{$\W\potdep $}{W(pot/dep)}}\label{sec:deriv}
-
-In the following, we will mathematically define the classes of perturbations pictorially described in \cref{fig:perts} of \main.
-In order to do so, we will need to consider expressions for derivatives of various quantities with respect to $\W\potdep_{ij}$.
-
-As discussed in \main,
-we will regard the off-diagonal elements of $\W\potdep _{ij}$ to be the independent variables,
-with $\W\potdep _{ii}=-\sum_{j \neq i} \W\potdep _{ij}$ imposed by hand.
-Thus,
-%
-\begin{equation}\label{eq:basicderivs}
-  \pdiff{\frg_{ij}}{\W\potdep _{gh}} = f\potdep  \delta_{gi}(\delta_{hj}-\delta_{ij}),
-  \qquad
-  \pdiff{\enc_{ij}}{\W\potdep _{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
-\end{equation}
-%
-The implicit $g \neq h$ that comes with all derivatives is unnecessary, as the derivatives above vanish when $g=h$.
-
-In particular, differentiating \cref{eq:funddef},
-%
-\begin{equation}\label{eq:derivZ}
-  \pdiff{\fund_{ij}}{\W\potdep _{gh}} = rf\potdep  \fund_{ig} (\fund_{hj}-\fund_{gj}).
-\end{equation}
-%
-We can then differentiate expression \cref{eq:areaZ} to get
-%
-\begin{equation}\label{eq:derivA}
-  \begin{aligned}
-    \pdiff{\hat{A}}{\W\potdep _{gh}} =&\,
-%      rf\potdep  \sum_{ijkl} \enc_{ij} \w_k \arow_l \brk{
-%         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
-%    &\pm \sum_{kl} \w_k \arow_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
-%    =&\,
-%      2rf\potdep  \sum_{ij} \enc_{ij} \eq_i \eq_g \brk{
-%         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
-%         (\fptb_{ig}-\fptb_{jg})(\eta^+_{g}-\eta^+_{h}) } \\
-%      &\pm 2 \eq_g (\eta^+_g - \eta^+_h) \\
-%    =&\,
-%      2rf\potdep  \sum_{ij} \enc_{ij} \eq_i \eq_g
-%         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
-%      \pm 2 \eq_g (\eta^+_g - \eta^+_h)\brk{1+\pdiff{\ln(\eq_g)}{\,\ln(f\potdep )}} \\
-%    =&\,
-      2rf\potdep  \eq_g \brk{\sum_i a_i (\fptb_{gi}-\fptb_{hi}) + c_g (\eta^+_{g}-\eta^+_{h})}
-      \pm 2 \eq_g (\eta^+_g - \eta^+_h). \\
-  \end{aligned}
-\end{equation}
-%
-where $a_i$ and $c_k$ were defined in \cref{eq:areacoeffs}.
-
-It is sometimes useful to consider the following derivatives:
-%
-\begin{equation}\label{eq:derivqw}
-  \begin{aligned}
-    \pdiff{}{\frg_{gh}} &\equiv     \pdiff{}{\W\pot _{gh}} +     \pdiff{}{\W\dep _{gh}} ,&\qquad
-    \pdiff{}{\enc_{gh}}    &\equiv f\dep  \pdiff{}{\W\pot _{gh}} - f\pot  \pdiff{}{\W\dep _{gh}} .
-  \end{aligned}
-\end{equation}
-%
-Each of these derivatives behaves as their names suggest:
-%
-\begin{equation}\label{eq:derivqweff}
-  \pdiff{\frg_{ij}}{\frg_{gh}} = \pdiff{\enc_{ij}}{\enc_{gh}}
-  =  \delta_{gi}(\delta_{hj}-\delta_{ij}),
-  \qquad
-  \pdiff{\enc_{ij}}{\frg_{gh}} = \pdiff{\frg_{ij}}{q_{gh}}  = 0.
-\end{equation}
-%
-This is because we could treat $\frg$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W\potdep $.
-
-\subsubsection{Scaling mode}\label{sec:rescale}
-
-Consider the following differential operator:
-%
-\begin{equation}\label{eq:scaleop}
-  \Delta \equiv \sum_{g,h} \W\pot _{gh}\pdiff{}{\W\pot _{gh}} + \W\dep _{gh}\pdiff{}{\W\dep _{gh}}.
-\end{equation}
-%
-This corresponds to the scaling, $\W\potdep  \to (1+\epsilon)\W\potdep $.
-Intuitively, this has two effects: it scales up the initial potentiation/depression and it scales down all timescales.
-This intuition is confirmed by the following results:
-%
-\begin{equation}\label{eq:scaleeffects}
-  \begin{aligned}
-    \Delta \fund &= \tau\onev\eq - \fund ,&
-    \Delta \eq  &= 0 ,&
-    \Delta \fpt  &= -\fpt ,
-    \\
-    \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
-    \Delta \enc_{ij} &= \enc_{ij} ,&
-    \Delta \hat{A}  &= 0 ,&
-  \end{aligned}
-\end{equation}
-%
-The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\arow$ and $\tau$ on $\W\potdep $.
-
-As the area is invariant under this scaling, we can consider the $\W\potdep $ to be projective coordinates.
-Therefore we don't need to enforce the lower bound on the diagonal matrix elements while looking for the maximum area, as we can use this null-mode to enforce it afterwards without changing the area.
-%We also don't have to worry about the boundary at infinity, as sending some of them to infinity is equivalent to sending the rest to zero.
-
-\subsection{Kuhn-Tucker conditions}\label{sec:kuhntucker}
-
-Consider the Lagrangian
-%
-\begin{equation}\label{eq:lagrangian}
-  \CL = \hat{A} + \sum_{\text{pot/dep}}\sum_{i\neq j} \mu\potdep _{ij} \W\potdep _{ij} .%  +\lambda\eq\w.
-\end{equation}
-%
-%The last term ensures that we hold $\eq_\pm$ fixed during this extremisation, so that we can safely ignore the factors that we dropped from the noise in \cref{eq:noisepm}.
-Necessary conditions for an extremum are
-%
-\begin{equation}\label{eq:extremum}
-  \pdiff{\CL}{\W\potdep _{gh}} = 0,
-  \qquad
-    \mu\potdep _{gh} \geq 0,\quad
-    \W\potdep _{gh} \geq 0,\quad
-    \mu\potdep _{gh}\W\potdep _{gh} = 0.
-\end{equation}
-%
-with $g \neq h$.
-This enforces the positivity constraints on the off-diagonal elements, but not the diagonals.
-As discussed in \cref{sec:rescale}, that can be enforced after finding the maximum using the null scaling degree of freedom.
-
-\subsubsection{Triangularity}\label{sec:triangular}
-
-Here we describe the perturbations corresponding to \cref{fig:perts}(\ref{fig:tri_right},\ref{fig:tri_left}) of \main.
-
-Consider
-%
-\begin{equation}\label{eq:shiftqderiv}
-  \pdiff{\CL}{\enc_{gh}} =
-  f\dep  \pdiff{\CL}{\W\pot _{gh}} - f\pot  \pdiff{\CL}{\W\dep _{gh}}
-   = (f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh}) + 2\eq_g (\eta^+_g - \eta^+_h)
-   = 0.
-\end{equation}
-%
-This corresponds to the shift
-%
-\begin{equation}\label{eq:shiftq}
-  \W\pot _{ij} \to \W\pot _{ij} + f\dep \epsilon_{ij},
-  \qquad
-  \W\dep _{ij} \to \W\dep _{ij} - f\pot \epsilon_{ij},
-  \qquad
-  \sum_j \epsilon_{ij} = 0,
-\end{equation}
-%
-which leaves $\frg$ unchanged, and therefore $\eq$, $\fpt$ and $\eta^\pm$ as well.
-
-Assume $\eta^+_g > \eta^+_h$. Then
-%
-\begin{equation}\label{eq:lowertriangular}
- f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh} <0
- \qquad\implies\qquad
- \mu\dep _{gh} >0
- \qquad\implies\qquad
- \W\dep _{gh}=0.
-\end{equation}
-%
-Similarly, if $\eta^+_g < \eta^+_h$, then
-%
-\begin{equation}\label{eq:uppertriangular}
- f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh} > 0
- \qquad\implies\qquad
- \mu\pot _{gh} >0
- \qquad\implies\qquad
- \W\pot _{gh}=0.
-\end{equation}
-%
-Thus, if we arrange the states in order of decreasing $\eta^+$, $\W\pot $ is upper-triangular and $\W\dep $ is lower triangular.
-
-We have ignored the possibility that $\eq_g=0$, as this would imply that $\fpt_{ig}=\infty$, which would in turn imply that the Markov process is not ergodic.
-
-%Note that if we multiply \cref{eq:shiftqderiv} by $\enc_{gh}$ and sum over $(g,h)$, we get
-%%
-%\begin{equation}\label{eq:stupid}
-%  \hat{A} = \sum_\pm\sum_{g\neq h} \mu\potdep _{gh} \frg_{gh},
-%\end{equation}
-%%
-%which is probably useless
-
-\subsubsection{Shortcuts}\label{sec:shortcuts}
-
-In this subsection we will define perturbations corresponding to \cref{fig:perts}(\ref{fig:shortcut}) of \main.
-
-Now consider the following combinations of derivatives for $m>1$:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:shortcutderiv}
-    \widetilde{\Delta}\potdep _{g,m} &\equiv
-      \brk{ \sum_{k=0}^{m-1} \frac{1}{\eq_{g\pm k}} \pdiffc{}{\W\potdep _{g\pm k,g\pm(k+1)}} }
-      - \frac{1}{\eq_g} \pdiffc{}{\W\potdep _{g,g\pm m}}.
-%  \end{aligned}
-\end{align}
-%
-Once again, they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \cref{sec:triangular} about this being satisfied for ergodic chains).
-
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators \cref{eq:diffpT}:
-%
-\begin{equation}\label{eq:shortcutprob}
-  \widetilde{\Delta}\potdep _{g,m} \eq_i = 0,
-\end{equation}
-%
-which makes it possible to integrate the perturbation:
-%
-\begin{equation}\label{eq:shortcutfinite}
-  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon}^{\pm(g,m)},
-  \qquad
-  \begin{aligned}
-    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g,g\pm m}
-      &\!\!\!\!\!\!&= -\epsilon,\\
-    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm(k+1)}
-      &\!\!\!\!\!\!&= \epsilon
-        &\;
-        &\forall\, k \in [0,m-1],\\
-    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm k}
-      &\!\!\!\!\!\!&= -\epsilon
-        &
-        &\forall\, k \in [1,m-1].
-  \end{aligned}
-\end{equation}
-%
-But more interestingly for our purposes:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:shortcutarea}
-    \widetilde{\Delta}\potdep _{g,m}\CL &=
-      \brk{ \sum_{k=0}^{m-1} \frac{\mu\potdep _{g\pm k,g\pm(k+1)}}{\eq_{g\pm k}}
-      - \frac{\mu\potdep _{g,g\pm m}}{\eq_g}}
-      + 2rf\potdep  \sum_{k=0}^{m-1} \prn{\eta^+_{g\pm k} - \eta^+_{g\pm(k+1)}} \prn{c_{g\pm k} - c_{g}},
-%  \end{aligned}
-\end{align}
-%
-
-In the \hyperref[sec:areacoeffincr]{\lcnamecref{sec:areacoeffincr} below}, we will show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
-This implies that the last term of the final expression in \cref{eq:shortcutarea} is non-negative.
-If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu\potdep _{g,g\pm m}>0$, which in turn implies that $\W\potdep _{g,g\pm m}=0$.
-This would tell us that the process with the maximal area has to have a multi-state topology.
-
-\subsubsection{Increasing \texorpdfstring{$c_k$}{c(k)}}\label{sec:areacoeffincr}
-
-In the \hyperref[sec:shortcuts]{previous \lcnamecref{sec:shortcuts}} we defined perturbations corresponding to \cref{fig:perts}(\ref{fig:shortcut}) of \main.
-In order to show that those perturbations increase the area, we must now show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
-
-Consider the following combinations of derivatives:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:areacoeffincrderiv}
-    \Delta_{gh} &\equiv
-      \frac{1}{\eq_{g}} \pdiffc{}{\frg_{gh}}
-      + \frac{1}{\eq_h} \pdiffc{}{\frg_{hg}}, \\
-%  \end{aligned}
-\end{align}
-%
-Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \cref{sec:triangular} about this being satisfied for ergodic chains).
-
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators using \cref{eq:diffpT}:
-%
-\begin{equation}\label{eq:sareacoeffincrprob}
-  \Delta_{gh} \eq_i = 0,
-\end{equation}
-%
-which makes it possible to integrate the perturbation:
-%
-\begin{equation}\label{eq:areacoeffincrfinite}
-  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon},
-  \qquad
-  \begin{aligned}
-  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
-  \\
-  \boldsymbol{\epsilon} \onev &= 0.
-  \end{aligned}
-\end{equation}
-%
-But more interestingly:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:areacoeffincr}
-    \Delta_{gh}\CL &=
-      \frac{\mu\pot _{gh}+\mu\dep _{gh}}{\eq_g} + \frac{\mu\pot _{hg}+\mu\dep _{hg}}{\eq_h}
-      + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
-%  \end{aligned}
-\end{align}
-%
-where $c_k$ were defined in \cref{eq:areacoeffs}.
-
-Using the non-negativity of the Kuhn-Tucker multipliers, $\mu\potdep _{ij}$, \cref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
-
-Note that, according to \cref{sec:triangular}, either $\W\pot _{gh}$ or $\W\dep _{gh}$ will be zero at the maximum, therefore we can expect one of $\mu\pot _{gh}+\mu\dep _{gh}$ to be non-zero.
-This would rule out degeneracy of the $c_k$ or $\eta^+_k$.
-Looking at \cref{eq:shiftqderiv} closely, the only way $\mu\pot _{gh}+\mu\dep _{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
-
-%From the expression for the area in \cref{eq:areacoeffs}, it will clearly also be optimal to have the $\w_k$ non-decreasing, \ie all the positive $\w_k$ should lie ahead of all the negative $\w_k$.
-
-\subsubsection{Summary}\label{sec:KTsummary}
-
-Using the Kuhn-Tucker formalism, we have shown that,
-with the states arranged in order of non-increasing $\eta^+_i$:
-%
-\begin{itemize}
-  \item There can be no ergodic maximum for which $\W\pot $ contains backwards transitions or $\W\dep $ contains forwards transitions.
-  \item There can be no ergodic maximum with the $c_k$ decreasing.
-  \item The $c_k$ may only be degenerate at an ergodic maximum if the corresponding $\eta^+_k$ are also degenerate.
-  \item If the $c_k$ increase and the $\eta^+_i$ decrease, there can be no ergodic maximum with shortcuts.
-\end{itemize}
-%
-These were shown by finding allowed perturbations that increase the area.
-
-%Suppose we start with some arbitrary ergodic process.
-If $\W\potdep $ contain backwards/forwards transitions or the $c_k$ decrease, then we can increase the area by applying the perturbations \cref{eq:shiftqderiv} and \cref{eq:areacoeffincrderiv}.
-%The second perturbation will generate the unwanted backwards/forwards transitions, requiring subsequent application of first perturbation.
-%The first perturbation will change the $c_k$, possibly changing their order, requiring subsequent application of second perturbation.
-%
-%This procedure can end in several ways:
-%%
-%\begin{itemize}
-%  \item Applying \cref{eq:areacoeffincrderiv} an infinite amount sends two transition rates to infinity, which due to the scaling mode (\cref{sec:rescale}), is equivalent to sending all other transitions to zero.
-%      Subsequent application of \cref{eq:shiftqderiv} just removes the backwards transition from $\W\pot $ and the forward transition from $\W\dep $.
-%  \item Applying \cref{eq:areacoeffincrderiv} ends in a process with degenerate $c_k$ or $\eta^+_k$, and the subsequent application of \cref{eq:shiftqderiv} doesn't change this degeneracy.
-%  \item As above, but the subsequent application of \cref{eq:shiftqderiv} lifts the degeneracy, resulting in a chain with increasing $c_k$.
-%\end{itemize}
-%%
-%In the third case, we can then apply the perturbation \cref{eq:shortcutderiv} to remove all the shortcuts.
-%Either this, or the first case, results in a multi-state topology, possibly disconnected or without the full set of states.
-
-This leaves two possibilities for the maximum area Markov chain.
-Either there is no degeneracy and no shortcuts, which implies the Multi-state/serial topology that we'll discuss in \cref{sec:multistate}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \cref{eq:shortcutarea}).
-
-Degeneracy tends to be very delicate.
-It is usually hard to arrange without some symmetry relating degenerate states.
-Such a symmetry would imply lumpability (see \cref{sec:lump}).
-The lumped chain would not have any shortcuts, as an entire degenerate set cannot be bypassed.
-As this lumped chain has the same area (see \cref{sec:SNRlump}), we would need only consider the multi-state topology.
-
-\subsection{Multi-state/Serial topology}\label{sec:multistate}
-
-The previous results indicate that the area under the memory curve of any model is bounded by the area under the memory curve of a model with the serial/multistate topology having the same equilibrium distribution.
-Here we compute this area, which we will see depends only on this equilibrium distribution.
-
-The multi-state/serial topology is defined by (see \cite{amit1994learning,Fusi2007multistate,Leibold2008serial}):
-%
-\begin{equation}\label{eq:multistatedef}
-  \W\pot _{ij} = q\pot _i (\delta_{i+1,j} - \delta_{ij}),
-  \qquad
-  \W\dep _{ij} = q\dep _j (\delta_{i,j+1} - \delta_{ij}).
-\end{equation}
-%
-Because it has no shortcuts, it saturates various inequalities:
-%
-\begin{equation}\label{eq:multistateineq}
-  \begin{aligned}
-    \fptb_{ik} - \fptb_{jk} &=
-      \begin{cases}
-        \fptb_{ij},  &\text{if}\quad i \leq j \leq k \quad\text{or}\quad i \geq j \geq k,\\
-        -\fptb_{ji}, &\text{if}\quad j \leq i \leq k \quad\text{or}\quad j \geq i \geq k,\\
-      \end{cases} \\
-    r\eq_i \frg_{ij} \prn{\fptb_{ij}+\fptb_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
-  \end{aligned}
-\end{equation}
-%
-and it satisfies detailed balance (a.k.a.\ reversibility a.k.a. $\CL^2_{\eq}$ self-adjointness):
-%
-\begin{equation}\label{eq:multistateprob}
-  f\pot  q\pot _i \eq_i = f\dep  q\dep _i \eq_{i+1},
-\end{equation}
-%
-which means we can always choose the transition rates, $q\potdep _i$, to give any desired equilibrium probabilities, $\eq_i$.
-
-This allows us to calculate the $c_k$'s:
-%
-\begin{equation}\label{eq:areacoeffchain}
-\begin{aligned}
-  c_k =&\, \sum_{i<k} \fpt_{i,i+1} \prn{\eq_{i}q\pot _i+\eq_{i+1}q\dep _i}
-    %\\&
-    - \sum_{i\geq k} \fpt_{i+1,i} \prn{\eq_{i}q\pot _i+\eq_{i+1}q\dep _i}
-  ,\\
-  c_{k+1} - c_k =&\, \prn{\fpt_{k,k+1}+\fpt_{k+1,k}} \prn{\frac{\eq_{k}\frg_{k,k+1}}{f\pot }+\frac{\eq_{k+1}\frg_{k+1,k}}{f\dep }}
-    =%\\=&\,
-    \frac{1}{rf\pot f\dep },\\
-  \sum_k c_k \eq_k =&\, \sum_{ij} \eq_i \enc_{ij} (\eta-\eta) = 0,\\
-%  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f\pot f\dep },
-  \implies c_k =&\, \frac{k - \sum_j j\eq_j}{rf\pot f\dep },
-\end{aligned}
-\end{equation}
-%
-where we used \cref{eq:multistateineq} to derive the first two equations respectively and \cref{th:kemenyconst} to derive the third. This allows us to write the area as
-%
-\begin{equation}\label{eq:multistatearea:s}
- % A = 2\sqrt{N} \sum_k \brk{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \eq_k \w_k.
-  A = \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k
-    = \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
-\end{equation}
-%
-where we used $\w_k=\sgn(c_k)$, as discussed after \cref{eq:areacoeffs}.
-This reproduces \cref{eq:multistatearea} of \main.
-
-%It will also help to define $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
-%Note that
-%%
-%\begin{equation}\label{eq:allpplusorminus}
-%  \begin{aligned}
-%    \eq_+ &=1
-%     \quad&\implies\quad
-%     \eq_k \w_k &= \eq_k
-%     \quad&\implies\quad
-%     A=0,\\
-%    \eq_- &=1
-%     \quad&\implies\quad
-%     \eq_k \w_k &= -\eq_k
-%     \quad&\implies\quad
-%     A=0,
-%  \end{aligned}
-%\end{equation}
-%%
-%neither of which are optimal.
-%
-%Consider the Lagrangian
-%%
-%\begin{equation}\label{eq:multistatelagrangian}
-%  \CL = \frac{A}{2\sqrt{N}} + \lambda\prn{1-\sum_i \eq_i} + \sum_i \mu_i \eq_i,
-%\end{equation}
-%%
-%where $\lambda$ is a Lagrange multiplier and $\mu_i$ are Kuhn-Tucker multipliers (see \cref{eq:extremum}). Extremising \wrt $\eq_i$:
-%%
-%\begin{equation}\label{eq:multistateext}
-%%  \pdiff{\CL}{\eq_i} = \brk{\prn{i-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \w_i
-%%    - \prn{i-\frac{n+1}{2}}\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
-%  \pdiff{\CL}{\eq_i} = \brk{i - \sum_j j\eq_j} \w_i
-%    - i\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
-%\end{equation}
-%%
-%Consider an $i$ such that $\w_i=\w_{i+1}$:
-%%
-%\begin{equation}\label{eq:multistatemudiff}
-%  \mu_{i+1} - \mu_i = \prn{\eq_+-\eq_-} - \w_i.
-%\end{equation}
-%%
-%As $\abs{\eq_+-\eq_-}<1$ (unless we want $A=0$, \cref{eq:allpplusorminus}), this means the $\mu_i$ increase amongst the negative $\w_i$ and decrease amongst the positive $\w_i$.
-%Therefore the only non-zero equilibrium probabilities at optimum are $\eq_1$ and $\eq_n$.
-
-In order to obtain an upper bound on the area under the memory curve of any model, we now maximise the area of the serial model with respect to its equilibrium distribution.
-First let us maximise \cref{eq:multistatearea:s} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
-Clearly this will happen when we put all of the probability at the ends: $\eq_1=\eq_-$ and $\eq_n=\eq_+$ are the only non-zero $\eq_k$.
-This gives an area of
-%\footnote{Note that including the dropped factors in \cref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_+\eq_-)}(n-1)/r$, which would have no effect on what follows.}
-
-%
-\begin{equation}\label{eq:multistateextarea}
-  A \leq \frac{\sqrt{N}}{r}(M-1)\prn{4\eq_+\eq_-}.
-\end{equation}
-%
-This is maximised at $\eq_+=\eq_-=\half$:
-%
-\begin{equation}\label{eq:maxarea}
-  A \leq \frac{\sqrt{N}}{r}(M-1).
-\end{equation}
-%
-This yields the area bound of equation \cref{eq:max_area} of \main.
-
-Note that the chain that achieves this is not ergodic, the two states at each end are absorbing.
-This is similar to the results found numerically in \cite{Barrett2008discrete} in a slightly different situation.
-
-\end{strip}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \pnasbreak splits and balances the columns before the references.
-% If you see unexpected formatting errors, try commenting out this line
-% as it can run into problems with floats and footnotes on the final page.
-%\pnasbreak
-
-%\bibliographystyle{utcaps_sl}
-\begingroup
-\renewcommand{\bibitem}[1]{\item[\citenum{#1}.]}
-\bibliography{maths,neuro,markov}
-\endgroup
-
-\end{document}

### 2018-03-19 Paper/synapse_supp.tex
separated main, sup, etc

--- 2018-03-17 Paper/SynapsePaper.tex
@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
-% Note that the use of elements such as single-column equations
-% may affect the guide line number alignment.
-
-\templatetype{pnasresearcharticle} % Choose template
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ADDITIONAL OPTIONAL STYLE FILES
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{graphicx}
-%\usepackage{graphics} for finding documentation only
-\pdfsuppresswarningpagegroup=1
-\usepackage{epstopdf}
-\epstopdfsetup{update,suffix=-generated}
-%
-\IfFileExists{"C:/Progra\string~1/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Progra\string~1/Inkscape/}}{}
-\IfFileExists{"/Applications/Inkscape.app/Contents/Resources/bin/inkscape"}{\newcommand{\inkscapedir}{/Applications/Inkscape.app/Contents/Resources/bin/}}{}
-\providecommand{\inkscapedir}{}
-\epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{%
-  "\inkscapedir inkscape" -f #1 -C -A \OutputFile}
-%
-\usepackage{adjustbox}
-\newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
-%\usepackage[inline]{enumitem}
-\newlist{myenuma}{enumerate*}{4}
-\setlist[myenuma]{label=(\alph*),ref=\alph*}
-\newcommand{\panel}[1]{\item\aligntop{#1}}
-\graphicspath{{"Figs/"}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{amssymb,amsfonts}
-\usepackage{amsthm}
-%\usepackage[notref,notcite]{showkeys}
-%\usepackage[centertags]{amsmath}
-%\usepackage{bibunits}
-%\usepackage{hyperref}
-\hypersetup{bookmarksopenlevel=1,bookmarksnumbered}
-\usepackage[capitalise,nameinlink]{cleveref}
-%\Crefname{appendix}{App.}{Apps.}
-\creflabelformat{equation}{[#2#1#3]}
-\crefrangelabelformat{equation}{[#3#1#4--#5#2#6]}
-%\crefformat{equation}{#2Eq.~[#1]#3}
-%\crefrangeformat{equation}{Eqs.~[#3#1#4--#5#2#6]}
-%\crefmultiformat{equation}{Eqs.~[#2#1#3}{,2#1#3]}{,#2#1#3}{,#2#1#3]}
-%\crefrangemultiformat{equation}{Eqs.~[#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}{,#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}
-\newcommand{\main}{the main paper}
-\newcommand{\supp}{supporting information}
-% ----------------------------------------------------------------
-\newtheoremstyle{sldefinition}%
-  {3pt}%space above
-  {3pt}%space below
-  {}%body font
-  {}%indent amount
-  {\bfseries}%theorem head font
-  {}%theorem head punctuation
-  {\newline}%space after head
-  {\thmname{#1}\thmnumber{ #2}:{\bfseries\thmnote{ #3}}}%head spec
-\newtheoremstyle{slplain}%
-  {3pt}%space above
-  {3pt}%space below
-  {}%body font
-  {}%indent amount
-  {\bfseries}%theorem head font
-  {}%theorem head punctuation
-  {\newline}%space after head
-  {\thmname{#1}\thmnumber{ #2}{\bfseries\thmnote{ (#3)}}: }%head spec
-%
-\theoremstyle{slplain}
-\newtheorem{thm}{Theorem}
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{lem}[thm]{Lemma}
-%
-\theoremstyle{sldefinition}
-\newtheorem{defn}{Definition}
-%
-\theoremstyle{remark}
-\newtheorem*{notn}{Notation}
-\newtheorem*{rem}{Remark}
-% ----------------------------------------------------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL MACRO DEFINITIONS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\ra}{\rightarrow}
-\newcommand{\lr}{\leftrightarrow}
-\newcommand{\cdt}{\!\cdot\!}
-\newcommand{\vp}{\vspace{0.5cm}}
-\newcommand{\hp}{\hspace{0.5cm}}
-%
-%e.g., i.e. with normal spaces
-\newcommand{\eg}{e.g.\ }
-\newcommand{\ie}{i.e.\ }
-\newcommand{\cf}{cf.\ }
-\newcommand{\etc}{etc.\ }
-\newcommand{\wrt}{wrt.\ }
-\newcommand{\Eg}{E.g.\ }
-\newcommand{\Ie}{I.e.\ }
-\newcommand{\Cf}{Cf.\ }
-\newcommand{\Etc}{Etc.\ }
-%
-\usepackage{delim}
-% brackets etc.
-\delimdef\prn#1{\dleft( #1 \dright)}
-\delimdef\brc#1{\dleft\{ #1 \dright\}}
-\delimdef\brk#1{\dleft[ #1 \dright]}
-\delimdef\abs#1{\dleft\lvert #1 \dright\rvert}
-\delimdef\nrm#1{\dleft\lVert #1 \dright\rVert}
-\delimdef\av#1{\dleft\langle #1 \dright\rangle}
-\delimdef\avc#1#2{\dleft\langle #1 \dmiddle\vert #2 \dright\rangle}
-%\newenvironment{cases}{\dleft\{\begin{aligned}}{\end{aligned}\dright.}
-%
-% Sets
-\delimdef\set#1#2{\dleft\{ #1 \dmiddle\vert #2 \dright\}}
-\delimdef\cond#1#2{\!\dleft( #1 \dmiddle\vert #2 \dright)}
-\delimdef\condb#1#2{\!\dleft[ #1 \dmiddle\vert #2 \dright]}
-%
-% Derivatives, etc. First argument is optional.
-\newcommand{\diff}[3][]{\frac{\mathrm{d}^{#1} #2}{\mathrm{d}{#3}^{#1}}}
-\newcommand{\pdiff}[3][]{\frac{\partial^{#1} #2}{\partial {#3}^{#1}}}
-\newcommand{\pdiffc}[3][]{\left(\frac{\partial #2}{\partial {#3}}\right)_{\!\!#1}}
-\newcommand{\intd}[2][]{\int #1\!\!\dr #2\,}
-%
-% Un-italicised letters
-\newcommand{\dr}{\mathrm{d}}
-\newcommand{\e}{\mathrm{e}}
-\newcommand{\ir}{\mathrm{i}}
-\DeclareMathOperator{\tr}{tr}
-\DeclareMathOperator{\Tr}{Tr}
-\DeclareMathOperator{\Det}{Det}
-\DeclareMathOperator{\cov}{Cov}
-\DeclareMathOperator{\var}{Var}
-\DeclareMathOperator{\bias}{Bias}
-\DeclareMathOperator{\rank}{rank}
-\DeclareMathOperator{\sgn}{sgn}
-\DeclareMathOperator{\sech}{sech}
-\DeclareMathOperator{\diag}{diag}
-%
-% The default \Im and \Re look crap
-\renewcommand{\Im}{\operatorname{\mathfrak{Im}}}
-\renewcommand{\Re}{\operatorname{\mathfrak{Re}}}
-%
-% logic
-\newcommand{\means}{\Longleftrightarrow}
-\newcommand{\requires}{\Longleftarrow}
-%
-%additional symbols:
-\newcommand{\half}{\frac{1}{2}}
-\newcommand{\CO}{\mathcal{O}}
-\newcommand{\CL}{\mathcal{L}}
-\newcommand{\dt}{\dr t}
-%
-\DeclareMathOperator{\SNR}{SNR}
-\DeclareMathOperator{\snr}{SNR}
-\newcommand{\wv}{\vec{w}}
-\newcommand{\wvi}{\vec{w}_\text{ideal}}
-%matrices
-\newcommand{\inv}{^{-1}}
-\newcommand{\dg}{^\mathrm{dg}}
-\newcommand{\trans}{^\mathrm{T}}
-\newcommand{\I}{\mathbf{I}}
-%vec of ones
-\newcommand{\onev}{\mathbf{e}}
-%mat of ones
-\newcommand{\onem}{\mathbf{E}}
-%Markov matrix
-\newcommand{\MM}{\mathbf{Q}}
-%prob distributions
-\newcommand{\prob}{\mathbf{p}}
-\newcommand{\eq}{\boldsymbol{\pi}}
-%first passage times
-\newcommand{\fpt}{\mathbf{T}}
-%off-diag first passage times
-\newcommand{\fptb}{\overline{\fpt}}
-%fundamental matrix
-\newcommand{\fund}{\mathbf{Z}}
-%other symbols for matrices
-\newcommand{\Pb}{\mathbf{P}}
-\newcommand{\D}{\mathbf{D}}
-\newcommand{\arow}{\boldsymbol{\xi}}
-\newcommand{\Lb}{\boldsymbol{\Lambda}}
-\newcommand{\w}{\mathbf{w}}
-\newcommand{\W}{\mathbf{W}}
-\newcommand{\M}{\mathbf{M}}
-\newcommand{\enc}{\mathbf{q}}
-\newcommand{\frg}{\W^{\mathrm{F}}}
-\newcommand{\F}{\boldsymbol{\Phi}}
-%superscripts
-\newcommand{\pot}{^{\text{pot}}}
-\newcommand{\dep}{^{\text{dep}}}
-\newcommand{\potdep}{^{\text{pot/dep}}}
-%sets
-\newcommand{\CS}{\mathcal{S}}
-\newcommand{\CA}{\mathcal{A}}
-\newcommand{\CB}{\mathcal{B}}
-\newcommand{\comp}{^\mathrm{c}}
-%eigenmodes
-\newcommand{\uv}{\mathbf{u}}
-\newcommand{\vv}{\mathbf{v}}
-\newcommand{\CI}{\mathcal{I}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Optimal synaptic strategies for different timescales of memory}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1]{Subhaneil Lahiri}
-\author[a]{Surya Ganguli}
-
-\affil[a]{Department of Applied Physics, Stanford University, Stanford CA}
-
-% Please give the surname of the lead author for the running footer
-\leadauthor{Lahiri}
-
-% Please add here a significance statement to explain the relevance of your work
-\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.}
-
-% Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any conflict of interest here.}
-%\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: sulahiri@stanford.edu}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{synaptic plasticity | learning | memory | neuroscience}
-
-%\abbreviations{SNR, signal-to-noise ratio; LTP, long term potentiation;LTD, long term depression}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\begin{article}
-\begin{abstract}
-  An incredible gulf separates theoretical models of synapses, often described solely by a single scalar value denoting the size of a postsynaptic potential, from the immense complexity of molecular signaling pathways underlying real synapses.
-  To understand the functional contribution of such molecular complexity to learning and memory, it is essential to expand our theoretical conception of a synapse from a single scalar to an entire dynamical system with many internal molecular functional states.
-  Moreover, theoretical considerations alone demand such an expansion; network models with scalar synapses assuming finite numbers of distinguishable synaptic strengths have strikingly limited memory capacity.
-  This raises the fundamental question, how does synaptic complexity give rise to memory?
-  To address this, we develop new mathematical theorems elucidating the relationship between the structural organization and memory properties of complex synapses that are themselves molecular networks.
-  Moreover, in proving such theorems, we uncover a framework, based on first passage time theory, to impose an order on the internal states of complex synaptic models, thereby simplifying the relationship between synaptic structure and function.
-
-  Overall, we uncover general design principles governing the functional organization of complex molecular networks, and suggest new experimental observables in synaptic physiology, based on first %passage time theory, that connect molecular complexity to memory.
-\end{abstract}
-
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\setboolean{displaywatermark}{false}
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{article}
-
-% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
-% You should only change this length when you've finalised the article contents.
-\verticaladjustment{-2pt}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-\dropcap{I}t is widely thought that our very ability to remember the past over long time scales depends crucially on our ability to modify synapses in our brain in an experience dependent manner.
-Classical models of synaptic plasticity model synaptic efficacy as an analog scalar value, denoting the size of a post-synaptic potential injected into one neuron from another.
-Theoretical work has shown that such models have a reasonable, extensive memory capacity, in which the number of long term associations that can be stored by a neuron is proportional its number of afferent synapses \cite{Hopfield1982model,Amit1985hopfield,Gardner1988perceptron}.
-However, recent experimental work has shown that many synapses are more digital than analog; they cannot robustly assume an infinite continuum of analog values, but rather can only take on a finite number of distinguishable strengths, a number than can be as small as two \cite{Bliss1993LTP,Petersen1998allornone,O'Connor2005switch} \cite[though see][] {Enoki2009graded}.
-This one simple modification leads to a catastrophe in memory capacity:  classical models with digital synapses, when operating in a palimpsest mode in which the ongoing storage of new memories can overwrite previous memories, have a memory capacity proportional to the logarithm of the number of synapses \cite{amit1992constraints,amit1994learning}.
-Intuitively, when synapses are digital, the storage of a new memory can flip a population of synaptic switches, thereby rapidly erasing previous memories stored in the same synaptic population.
-This result indicates that the dominant theoretical basis for the storage of long term memories in modifiable synaptic switches is flawed.
-
-Recent work \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial} has suggested that a way out of this logarithmic catastrophe is to expand our theoretical conception of a synapse from a single scalar value to an entire stochastic dynamical system in its own right.
-This conceptual expansion is further necessitated by the experimental reality that synapses contain within them immensely complex molecular signaling pathways, with many internal molecular functional states \cite[\eg see][]{Bliss1993LTP,Bredt2003AMPA,Coba2009phosphorylation}.
-While externally, synaptic efficacy could be digital, candidate patterns of electrical activity leading to potentiation or depression could yield transitions between these internal molecular states without necessarily inducing an associated change in synaptic efficacy.
-This form of synaptic change, known as metaplasticity \cite{Abraham1996metaplasticity,Montgomery2002765}, can allow the probability of synaptic potentiation or depression to acquire a rich dependence on the history of prior changes in efficacy, thereby potentially improving memory capacity.
-
-Theoretical studies of complex, metaplastic synapses have focused on analyzing the memory performance of a limited number of very specific molecular dynamical systems, characterized by a number of internal states in which potentiation and depression each induce a specific set of allowable transitions between states (\eg see \cref{fig:models} below).
-While these models can vastly outperform simple binary synaptic switches, these analyses leave open several deep and important questions.
-For example, how does the structure of a synaptic dynamical system determine its memory performance?
-What are the fundamental limits of memory performance over the space of all possible synaptic dynamical systems?
-What is the structural organization of synaptic dynamical systems that achieve these limits?
-Moreover, from an experimental perspective, it is unlikely that all synapses can be described by a single canonical synaptic model;  just like the case of neurons, there is an incredible diversity of molecular networks underlying synapses both across species and across brain regions within a single organism \cite{Emes2012synapserev}.
-In order to elucidate the functional contribution of this diverse molecular complexity to learning and memory, it is essential to move beyond the analysis of specific models and instead develop a general theory of learning and memory for complex synapses.
-Moreover, such a general theory of complex synapses could aid in development of novel artificial memory storage devices.  % Sutskever, memory for learning?
-
-Here we initiate such a general theory by proving upper bounds on the memory curve associated with any synaptic dynamical system, within the well established ideal
-observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}.
-Along the way we develop principles based on first passage time theory to order the structure of synaptic dynamical systems and relate this structure to memory performance.
-We summarize our main results in the \hyperref[sec:disc]{discussion section}.
-
-\section{Overall framework: synaptic models and their memory curves}\label{sec:setup}
-
-In this section, we describe the class of models of synaptic plasticity that we are studying and how we quantify their memory performance.
-In the subsequent sections, we will find upper bounds on this performance.
-
-We use a well established formalism for the study of learning and memory with complex synapses (see \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}).
-In this approach, electrical patterns of activity corresponding to candidate potentiating and depressing plasticity events occur randomly and independently at all synapses at a Poisson rate $r$.
-These events reflect possible synaptic changes due to either spontaneous network activity, or the storage of new memories.
-We let  $f\pot$ and  $f\dep$ denote the fraction of these events that are candidate potentiating or depressing events respectively.
-Furthermore, we assume our synaptic model has $M$ internal molecular functional states, and that a candidate potentiating (depotentiating) event induces a stochastic transition in the internal state described by an $M \times M$ discrete time Markov transition matrix  $\M\pot$ ($\M\dep$).
-In this framework, the states of different synapses will be independent, and the entire synaptic population can be fully described by the probability distribution across these states, which we will indicate with the row-vector $\prob(t)$.
-Thus the $i$'th component of $\prob(t)$ denotes the fraction of the synaptic population in state $i$.
-Furthermore, each state $i$ has its own synaptic weight, $\w_i$, which we take, in the worst case scenario, to be restricted to two values.
-After shifting and scaling these two values,  we can assume they are $\pm1$, without loss of generality.
-
-We also employ an ``ideal observer'' approach to the memory readout, where the synaptic weights are read directly.
-This provides an upper bound on the quality of any readout using neural activity.
-
-For any single memory, stored at time $t=0$, we assume there will be an ideal pattern of synaptic weights across a population of $N$ synapses, the $N$-element vector $\wvi$, that is $+1$ at all synapses that experience a candidate potentiation event, and $-1$ at all synapses that experience a candidate depression event at the time of memory storage.
-We assume that any pattern of synaptic weights close to $\wvi$ is sufficient to recall the memory.
-However, the actual pattern of synaptic weights at some later time, $t$, will change to $\wv(t)$ due to further modifications from the storage of subsequent memories.
-We can use the overlap between these, $\wvi\cdt\wv(t)$, as a measure of the quality of the memory.
-As $t\to\infty$, the system will return to its steady state distribution which will be uncorrelated with the memory stored at $t=0$.
-The probability distribution of the quantity $\wvi\cdt\wv(\infty)$ can be used as a ``null model'' for comparison.
-
-The extent to which the memory has been stored is described by a signal-to-noise ratio (SNR) \cite{Fusi2005cascade,Fusi2007multistate}:
-%
-\begin{equation}\label{eq:SNRdef}
-  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}}
-     {\sqrt{\var(\wv_\text{ideal}\cdt\wv(\infty))}}.
-\end{equation}
-%
-The noise in the denominator is essentially $\sqrt{N}$.
-There is a correction when potentiation and depression are imbalanced, but this will not affect the upper bounds that we will discuss below and will be ignored in the subsequent formulae.
-
-A simple average memory curve can be derived as follows.
-All of the preceding plasticity events, prior to $t=0$, will put the population of synapses in its steady-state distribution, $\eq$.
-The memory we are tracking at $t=0$ will change the internal state distribution to $\eq\M\pot$ (or $\eq\M\dep$) in those synapses that experience a candidate potentiation (or depression) event.
-As the potentiating/depressing nature of the subsequent memories is independent of $\wvi$, we can average over all sequences, resulting in the evolution of the probability distribution:
-%
-\begin{multline}\label{eq:evol}
-  \diff{\prob(t)}{t} = r\prob(t)\frg,
-  \\ \text{where} \quad
-  \frg = f\pot\M\pot + f\dep\M\dep - \I.
-\end{multline}
-%
-Here $\frg$ is a continuous time transition matrix that models the process of forgetting the memory stored at time $t=0$ due to random candidate potentiation/depression events occurring at each synapse due to the storage of subsequent memories.
-Its stationary distribution is $\eq$.
-
-This results in the following SNR
-%
-\begin{equation}\label{eq:SNRcalc}
-  \snr(t) = \sqrt{N}(2 f\pot f\dep) \eq \brk{\M\pot-\M\dep} \e^{rt\frg} \w.
-\end{equation}
-%
-%The factor of $\eq$ describes the synapses being in the steady-state distribution before the memory is encoded.
-The factor of $(\M\pot-\M\dep)$ comes from the encoding of the memory at $t=0$, with $\wvi$ being $\pm1$ in synapses that are potentiated/depotentiated.
-The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
-A detailed derivation of this formula can be found in the \supp.
-We will frequently refer to this function as the memory curve.
-It can be thought of as the excess fraction of synapses (relative to equilibrium) that maintain their ideal synaptic strength at time $t$, as dictated by the stored memory at time $t=0$.
-
-Much of the previous work on these types of complex synaptic models has focused on understanding the memory curves of specific models, or choices of $\M\potdep$.
-Two examples of these models are shown in \cref{fig:models}. We see that they have different memory properties.
-The serial model performs relatively well at one particular timescale, but it performs poorly at other times.
-The cascade model does not perform quite as well at that time, but it maintains its performance over a wider range of timescales.
-
-\begin{figure}[tbp]
- \begin{center}
- \begin{myenuma}
-  \panel{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
-  \panel{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
-  \panel{\includegraphics[width=0.3\linewidth]{cascms.eps}}\label{fig:cascms}
- \end{myenuma}
- \end{center}
-  \caption[Models of complex synapses]{Models of complex synapses.
-  (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
-  (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
-  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined \hyperref[sec:setup]{below}) as subsequent memories are stored.
-  \label{fig:models}}
-\end{figure}
-
-In this work, rather than analyzing specific models,  we take a different approach, in order to obtain a more general theory.
-We consider the \emph{entire} space of these models and find upper bounds on the memory capacity of \emph{any} of them.
-The space of models with a fixed number of internal states $M$ is parameterized by the pair of $M \times M$ discrete time stochastic
-transition matrices $\M\pot$ and $\M\dep$, in addition to $f\potdep$.  The parameters must satisfy the following constraints:
-%
-\begin{equation}\label{eq:constr}
-\begin{gathered}
-\begin{aligned}
-  \M\potdep_{ij} &\in [0,1], &\qquad
-  \sum_j \M\potdep_{ij} &= 1, \\
-  f\potdep &\in [0,1],&
-  f\pot + f\dep &= 1,
-\end{aligned}
-\\
-  \eq\frg = 0, \qquad
-  \sum_i \eq_i = 1, \qquad
-  \w_i = \pm 1.
-\end{gathered}
-\end{equation}
-%
-The upper bounds on $\M\potdep_{ij}$ and $f\potdep$ follow automatically from the other constraints.
-
-The critical question is: what do these constraints imply about the space of achievable memory curves in \cref{eq:SNRcalc}?
-To answer this question, especially for limits on achievable memory at finite times, it will be useful to employ the eigenmode decomposition:
-%
-\begin{equation}\label{eq:eigendecomp}
-\begin{aligned}
-  \frg &= \sum_a -q_a \uv^a \vv^a,
-  &\quad
-  \vv^a \uv^b &= \delta_{ab},
-  \\%\quad
-  \frg \uv^a &= -q_a \uv^a,
-  &\quad
-  \vv^a \frg &= -q_a \vv^a.
-\end{aligned}
-\end{equation}
-%
-Here $q_a$ are the negative of the eigenvalues of the forgetting process $\frg$, $\uv^a$ are the right (column) eigenvectors and $\vv^a$ are the left (row) eigenvectors.
-This decomposition allows us to write the memory curve as a sum of exponentials,
-%
-\begin{equation}\label{eq:SNReigen}
-\begin{aligned}
-  \snr(t) &= \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a},
-%  &\quad \text{where}&\;&
-%  \CI_a &= \prn{2 f\pot f\dep} \eq (\M\pot-\M\dep) \uv^a  \vv^a \w,\\&
-%  & \text{and}&&
-%  \tau_a &= \frac{1}{q_a}.
-\end{aligned}
-\end{equation}
-%
-where $\;\CI_a = (2 f\pot f\dep) \eq (\M\pot-\M\dep) \uv^a  \vv^a \w\;$ and $\;\tau_a = {1}/{q_a}$.
-We can then ask the question: what are the constraints on these quantities, namely eigenmode initial SNR's, $\CI_a$, and time constants, $\tau_a$, implied by the constraints in \cref{eq:constr}?
-We will derive some of these constraints in the next section.
-
-\section{Upper bounds on achievable memory capacity}\label{sec:bounds}
-
-In the previous section, in \cref{eq:SNRcalc} we have described an analytic expression for a memory curve as a function of the structure of a synaptic dynamical system, described by the pair of stochastic transition matrices  $\M\potdep$.
-Since the performance measure for memory is an entire memory curve, and not just a single number, there is no universal scalar notion of optimal memory in the space of synaptic dynamical systems.
-Instead there are tradeoffs between storing proximal and distal memories;  often attempts to increase memory at late (early) times by changing $\M\potdep$, incurs a performance loss in memory at early (late) times in specific models considered so far \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial}.
-Thus our end goal, achieved \hyperref[sec:env]{below}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the {\it entire} memory curve of {\it any} model.
-In order to achieve this goal, in this section, we must first derive upper bounds, over the space of all possible synaptic models, on two different scalar functions of the memory curve: its initial SNR, and the area under the memory curve.
-In the process of upper-bounding the area, we will develop an essential framework to organize the structure of synaptic dynamical systems based on first passage time theory.
-
-\subsection{Bounding initial SNR}\label{sec:initial}
-
-We now give an upper bound on the initial SNR,
-%
-\begin{equation}\label{eq:init}
-  \snr(0) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \w,
-\end{equation}
-%
-over \emph{all} possible models and also find the class of models that saturate this bound.
-A useful quantity is the equilibrium probability flux between two disjoint sets of states, $\CA$ and $\CB$:
-%
-\begin{equation}\label{eq:flux}
-  \F_{\CA\CB} = \sum_{i\in\CA} \sum_{j\in\CB} r \eq_i \frg_{ij}.
-\end{equation}
-%
-The initial SNR is closely related to the flux from the states with $\w_i=-1$ to those with $\w_j=+1$ (see \supp):
-%
-\begin{equation}\label{eq:initflux}
-  \snr(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}\,.
-\end{equation}
-%
-This inequality becomes an equality if potentiation never decreases the synaptic weight and depression never increases it, which should be a property of any sensible model.
-
-To maximize this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
-An example of such a model is shown in \cref{fig:max_init}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
-These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-They are completely equivalent to a two state model with transition probabilities equal to 1, as shown in \cref{fig:max_init}(\ref{fig:binary_det}).
-In particular, they have the same memory curve.
-
-This two state model has the equilibrium distribution $\eq=(f\dep,f\pot)$ and its flux is given by $\F_{-+} = rf\pot f\dep$.
-This is maximized when $f\pot=f\dep=\half$, leading to the upper bound:
-%
-\begin{equation}\label{eq:max_init}
-  \snr(0) \leq \sqrt{N}.
-\end{equation}
-%
-We note that while this model has high initial SNR, it also has very fast memory decay -- with a timescale $\tau\sim\frac{1}{r}$.
-As the synapse is very plastic, the initial memory is encoded very easily, but the subsequent memories also overwrite it rapidly.
-This is one example of the tradeoff between optimizing memory at early versus late times.
-\begin{figure}[tbp]
- \begin{center}
- \begin{myenuma}
-  \panel{\hp\includegraphics[width=0.24\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
-  \panel{\includegraphics[width=0.24\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
-  \panel{\hp\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
-  \end{myenuma}
- \end{center}
-  \caption[Synaptic models that maximize initial SNR]{Synaptic models that maximize initial SNR.
-  (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
-  (\ref{fig:max_init_pot}) Depression is similar to potentiation, but with strong and weak interchanged.
-  (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
-  \label{fig:max_init}}
-\end{figure}
-
-\subsection{Imposing order on internal states through first passage times}\label{sec:order}
-
-Our goal is to understand the relationship between structure and function in the space of all possible synaptic models.
-This goal is complicated by the fact that this space contains many different possible network topologies, encoded in the nonzero matrix elements of $\M\potdep$.
-To systematically analyze this entire space, we develop an important organizing principle using the theory of first passage times in the stochastic process of forgetting, described by $\frg$.
-The mean first passage time matrix, $\smash{\fptb_{ij}}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $i$.
-The diagonal elements are defined to be zero.
-
-A remarkable theorem we will exploit is that the quantity
-%
-\begin{equation}\label{eq:kemenyconst}
-  \eta \equiv \sum_j \fptb_{ij}\eq_j,
-\end{equation}
-%
-is independent of the starting state $i$.
-It is known as Kemeny's constant \cite[see][\S4.4]{kemeny1960finite}.
-Intuitively, \cref{eq:kemenyconst} states that the average time it takes to reach any state, weighted by its equilibrium probability, is independent of the starting state, implying a hidden constancy inherent in any stochastic process.
-
-In the context of complex synapses, we can define the partial sums
-%
-\begin{equation}\label{eq:kemenypm}
-   \eta^+_i = \sum_{j\in+} \fptb_{ij} \eq_j,
-   \qquad
-   \eta^-_i = \sum_{j\in-} \fptb_{ij} \eq_j.
-\end{equation}
-%
-These can be thought of as the average time it takes to reach the strong/weak states respectively.
-Using these definitions, we can then impose an order on the states by arranging them in order of decreasing $\smash{\eta^+_i}$ or increasing $\smash{\eta^-_i}$.
-Because $\smash{\eta^+_i+\eta^-_i=\eta}$ is independent of $i$, the two orderings are the same.
-In this order, which depends sensitively on the structure of $\M\potdep$, states later (to the right in figures below) can be considered to be more potentiated than states earlier (to the left in figures below), despite the fact that they have the same synaptic efficacy.
-In essence, in this order, a state is considered to be more potentiated if the average time it takes to reach all the strong efficacy states is shorter.
-We will see that synaptic models that optimize various measures of memory have an exceedingly simple structure when, and only when, their states are arranged in this order.%
-\footnote{Note that we do not need to worry about the order of the $\eta^\pm_i$ changing during the optimization: necessary conditions for a maximum only require that there is no infinitesimal perturbation that increases the area.
-Therefore we need only consider an infinitesimal neighborhood of the model, in which the order will not change.\label{ft:etaorder}}
-
-\subsection{Bounding area}\label{sec:area}
-
-Now consider the area under the memory curve:
-%
-\begin{equation}\label{eq:area}
-  A = \int_0^\infty\!\!\dr t \, \snr(t).
-\end{equation}
-%
-We will find an upper bound on this quantity as well as the model that saturates this bound.
-
-First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \cref{eq:kemenypm} (see \supp):
-%
-\begin{equation}\label{eq:areaEta}
-\begin{aligned}
-  A &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \!\brk{\M\pot_{ij}-\M\dep_{ij}}\! (\eta^+_i-\eta^+_j)\\
-    &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \!\brk{\M\pot_{ij}-\M\dep_{ij}}\! (\eta^-_j-\eta^-_i).
-\end{aligned}
-\end{equation}
-%
-
-With the states in the order described above, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
-Some of these perturbations are shown in \cref{fig:perts}, see \supp\ for details.
-For example, in \cref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
-$\smash{\eta^+_i > \eta^+_j}$), we have proven that increasing $\smash{\M\pot_{ij}}$ and decreasing $\smash{\M\dep_{ij}}$ leads to an increase in area.
-The only thing that can prevent these perturbations from increasing the area is when they require the decrease of a matrix element that has already been set to 0.
-This determines the topology (non-zero transition probabilities) of the model with maximal area.
-It is of the form shown in \cref{fig:env}(\ref{fig:sticky}),%\cref{fig:max_area},
-with potentiation moving one step to the right and depression moving one step to the left.
-Any other topology would allow some class of perturbations (\eg in \cref{fig:perts}) to further increase the area.
-
-\begin{figure}[tbp]
- \begin{center}
- \begin{myenuma}
-  \panel{\includegraphics[width=0.14\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
-  \panel{\includegraphics[width=0.14\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
-  \panel{\includegraphics[width=0.28\linewidth]{shortcut.svg}}\label{fig:shortcut}
- \end{myenuma}
- \end{center}
-  \caption[Perturbations that increase the area]{Perturbations that increase the area.
-  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$.
-It can no longer be used when $\M\dep$ is lower triangular, \ie depression must move synapses to ``more depressed'' states.
-  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$.
-It can no longer be used when $\M\pot$ is upper triangular, \ie potentiation must move synapses to ``more potentiated'' states.
-  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions.
-It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
-  \label{fig:perts}}
-\end{figure}
-
-As these perturbations do not change the equilibrium distribution, this means that the area of \emph{any} model is bounded by that of a linear chain with the same equilibrium distribution.
-The area of a linear chain model can be expressed directly in terms of its equilibrium state distribution, $\eq$, yielding the following upper bound on the area of any model with the same $\eq$ (see \supp):
-%
-\begin{equation}\label{eq:multistatearea}
-\begin{aligned}
-  A &\leq \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k \\
-    &= \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
-\end{aligned}
-\end{equation}
-%
-where we chose $\w_k=\sgn[{k - \sum_j j\eq_j}]$.
-We can then maximize this by pushing all of the equilibrium distribution symmetrically to the two end states.
-This can be done by reducing the transition probabilities out of these states, as in \cref{fig:env}(\ref{fig:sticky}). %\cref{fig:max_area}.
-This makes it very difficult to exit these states once they have been entered.
-The resulting area is
-%
-\begin{equation}\label{eq:max_area}
-  A \leq \frac{\sqrt{N}(M-1)}{r}.
-\end{equation}
-%
-This analytical result is similar to a numerical result found in \cite{Barrett2008discrete} under a slightly different information theoretic measure of memory performance.
-
-The ``sticky'' end states result in very slow decay of memory, but they also make it difficult to encode the memory in the first place, since a small fraction of synapses are able to change synaptic efficacy during the storage of a new memory.
-Thus models that maximize area optimize memory at late times, at the expense of early times.
-
-%\begin{figure}[tbp]
-% \begin{center}
-%  \includegraphics[width=0.35\linewidth]{multistate_sticky.svg}
-% \end{center}
-%  \caption{Model that maximizes the area.
-%  Unlabeled transitions have probability 1.
-%  Labeled transitions have probability $\varepsilon\to0$.
-%  }\label{fig:max_area}
-%\end{figure}
-
-\section{Memory curve envelope}\label{sec:env}
-
-Now we will look at the implications of the upper bounds found in the \hyperref[sec:bounds]{previous section} for the SNR at finite times.
-As argued in \cref{eq:SNReigen}, the memory curve can be written in the form
-%
-\begin{equation}\label{eq:SNRmodes}
-  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a}.
-\end{equation}
-%
-The upper bounds on the initial SNR, \cref{eq:max_init}, and the area, \cref{eq:max_area}, imply the following constraints on the parameters $\brc{\CI_a,\tau_a}$:
-%
-\begin{equation}\label{eq:modeconstr}
-  \sum_a \CI_a \leq 1,
-  \qquad\qquad
-  \sum_a \CI_a \tau_a \leq M-1.
-\end{equation}
-%
-We are not claiming that these are a complete set of constraints: not every set $\brc{\CI_a,\tau_a}$ that satisfies these inequalities will actually be achievable by a synaptic model.
-However, any set that violates either inequality will definitely not be achievable.
-
-Now we can pick some fixed time, $\smash{t_0}$, and maximize the SNR at that time \wrt the parameters $\smash{\brc{\CI_a,\tau_a}}$, subject to the constraints above.
-This always results in a single nonzero $\smash{\CI_a}$; in essence, optimizing memory at a single time requires a single exponential.
-The resulting optimal memory curve, along with the achieved memory at the chosen time, depends on $t_0$ as follows:
-%
-\begin{equation}\label{eq:env}
-\begin{aligned}
-  t_0 &\leq \frac{M-1}{r} &
-   & \implies &
-  \snr(t) &= \sqrt{N}\e^{-rt/(M-1)} \\&&
-   & \implies &
-  \snr(t_0) &= \sqrt{N}\e^{-rt_0/(M-1)},\\
-  t_0 &\geq \frac{M-1}{r} &
-   & \implies &
-  \snr(t) &= \frac{\sqrt{N}(M-1)\e^{-t/t_0}}{rt_0} \\&&
-   & \implies &
-  \snr(t_0) &= \frac{\sqrt{N}(M-1)}{\e rt_0}.
-\end{aligned}
-\end{equation}
-%
-Both the initial SNR bound and the area bound are saturated at early times.
-At late times, only the area bound is saturated.
-The function $\snr(t_0)$, the green curve in \cref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim \smash{t_0\inv}$.
-No synaptic model can have an SNR that is greater than this at any time.
-We can use this to find an upper bound on the memory lifetime, $\tau(\epsilon)$, by finding the point at which the envelope crosses $\epsilon$:
-%
-\begin{equation}\label{eq:lifebound}
-  \tau(\epsilon) \leq \frac{\sqrt{N}(M-1)}{\epsilon\e r},
-\end{equation}
-%
-where we assume $N>(\epsilon\e)^2$.
-Intriguingly, both the lifetime and memory envelope expand linearly with the number of internal states $M$, and increase as the square root of the number of synapses $N$.
-
-This leaves the question of whether this bound is achievable.
-At any time, can we find a model whose memory curve touches the envelope?
-The red curves in \cref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
-
-We see that at early, but not late times, there is a gap between the upper bound that we can prove and what we can achieve with actual models.
-There may be other models we haven't found that could beat the ones we have, and come closer to our proven envelope.
-However, we suspect that the area constraint is not the bottleneck for optimizing memory at times less than  $\CO(\frac{M}{r})$.
-We believe there is some other constraint that prevents models from approaching the envelope, and currently are exploring several mathematical conjectures for the precise form of this constraint in order to obtain a potentially tighter envelope.
-Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \cref{fig:env}(\ref{fig:sticky})%\cref{fig:max_area}
- can come close to the envelope.
-
-\begin{figure}
- \begin{center}
- \parbox[t]{0.43\linewidth}{%
- \begin{myenuma}
-   \panel{\includegraphics[width=0.85\linewidth]{env.svg}}\label{fig:envplot}
- \end{myenuma}}\hp
- \parbox[t]{0.44\linewidth}{%
- \begin{myenuma}
-  \stepcounter{myenumai}
-% \addtocounter{enumi}{1}
-   \panel{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}\\[0.1\linewidth]
-   \panel{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
- \end{myenuma}}
- \end{center}
-  \caption[The memory curve envelope for $N=100$, $M=12$]{The memory curve envelope for $N=100$, $M=12$.
-  (\protect\protect\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
-  The red dashed curve shows the result of numerical optimization of synaptic models with random initialization.
-  The solid red curve shows the highest SNR we have found with hand designed models.
-  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states, and all transition probabilities equal to $1$.
-  At late times they are of the form shown in (\ref{fig:sticky}) with different values of $\varepsilon$.
-  The model shown in (\ref{fig:sticky}) also saturates the area bound \cref{eq:max_area} in the limit $\varepsilon\to0$.
-  \label{fig:env}}
-\end{figure}
-
-\section{Discussion}\label{sec:disc}
-
-We have initiated the development of  a general theory of learning and memory with complex synapses, allowing for an exploration of the entire space of complex synaptic models, rather than analyzing individual models one at a time.
-In doing so, we have obtained several new mathematical results delineating the functional limits of memory achievable by synaptic complexity, and the structural characterization of synaptic dynamical systems that achieve these limits.
-In particular, operating within the ideal observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}, we have shown that for a population of $N$ synapses with $M$ internal states, \begin{myenuma}
-\item the initial SNR of any synaptic model cannot exceed $\sqrt{N}$, and any model that achieves this bound is equivalent to a binary synapse,
-\item the area under the memory curve of any model cannot exceed that of a linear chain model with the same equilibrium distribution,
-\item both the area and memory lifetime of any model cannot exceed $\CO(\sqrt{N}M)$, and the model that achieves this limit has a linear chain topology with only nearest neighbor transitions,
-\item we have derived an envelope memory curve in the SNR-time plane that cannot be exceeded by the memory curve of any model, and models that approach this envelope for times greater than $\CO(\frac{M}{r})$ are linear chain models, and
-\item this late-time envelope is a power-law proportional to $\smash{\CO({\sqrt{N} M}/{rt})}$, indicating that synaptic complexity can strongly enhance the limits of achievable memory.
-\end{myenuma}
-
-This theoretical study opens up several avenues for further inquiry.
-In particular, the tightness of our envelope for early times, less than $\CO(\frac{M}{r})$, remains an open question, and we are currently pursuing several conjectures.
-We have also derived memory constrained envelopes, by asking in the space of models that achieve a given SNR at a given time, what is the maximal SNR achievable at other times.
-If these two times are beyond a threshold separation, optimal constrained models require two exponentials.
-It would be interesting to systematically analyze the space of models that achieve good memory at multiple times, and understand their structural organization, and how they give rise to multiple exponentials, leading to power law memory decays.
-
-%Another important consideration is the tolerance of the synapses to biological noise, which could be investigated by imposing lower limits on the level of stochasticity of the models.
-Finally, it would be interesting to design physiological experiments in order to perform optimal systems identification of potential Markovian dynamical systems hiding within biological synapses,
-given measurements of pre and post-synaptic spike trains along with changes in post-synaptic potentials.
-Then given our theory, we could match this measured synaptic model to optimal models to understand for which timescales of memory, if any, biological synaptic dynamics may be tuned.
-
-In summary, we hope that a deeper theoretical understanding of the functional role of synaptic complexity, initiated here, will help advance our understanding of the neurobiology of learning and memory, aid in the design of engineered memory circuits, and lead to new mathematical theorems about stochastic processes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\acknow{We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
-We thank Larry Abbott, Marcus Benna, Stefano Fusi, Jascha Sohl-Dickstein and David Sussillo for useful discussions.}
-\showacknow % Display the acknowledgments section
-
-% \pnasbreak splits and balances the columns before the references.
-% If you see unexpected formatting errors, try commenting out this line
-% as it can run into problems with floats and footnotes on the final page.
-%\pnasbreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{pnas-new}
-\bibliography{maths,neuro,markov}
-
-%\end{article}
-%\clearpage
-
-\begin{strip}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\appendix
-\pdfbookmark[0]{Supporting information}{appendix}
-\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
-\begin{center}
-  \textbf{\Large Supporting information}
-\end{center}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 Here we provide more details underlying the derivations of results in \main.
@@
 \begin{equation}\label{eq:lumpeq}
-  \widehat{\mathbf{p}}^\infty = \eq V.
+  \widehat{\eq} = \eq V.
 \end{equation}
@@
 Thus, evolving according to \cref{eq:rowsum} results in averaging over all sequences of plasticity events,
-as we only need linear expectations of $\wv(t)$ in the end.\label{ft:avexp}}
+as we only need linear expectations of $\wv(t)$ in the end.}%\label{ft:avexp}}
 This results in
@@
 This is similar to the results found numerically in \cite{Barrett2008discrete} in a slightly different situation.
-
-\end{strip}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \pnasbreak splits and balances the columns before the references.
-% If you see unexpected formatting errors, try commenting out this line
-% as it can run into problems with floats and footnotes on the final page.
-%\pnasbreak
-
-%\bibliographystyle{utcaps_sl}
-\begingroup
-\renewcommand{\bibitem}[1]{\item[\citenum{#1}.]}
-\bibliography{maths,neuro,markov}
-\endgroup
-
-\end{document}

### 2018-03-19 Paper/synapse_title.tex
separated main, sup, etc

--- 2018-03-16 Paper/SynapsePaper.tex
@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
-% Note that the use of elements such as single-column equations
-% may affect the guide line number alignment.
-
-\templatetype{pnasresearcharticle} % Choose template
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
-
-%% ADDITIONAL OPTIONAL STYLE FILES
-%\usepackage[hypcap]{caption}%links to figures
-\usepackage{graphicx}
-%\usepackage{graphics} for finding documentation only
-\pdfsuppresswarningpagegroup=1
-\usepackage{epstopdf}
-\epstopdfsetup{update,suffix=-generated}
-%
-\IfFileExists{"C:/Progra\string~1/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Progra\string~1/Inkscape/}}{}
-\IfFileExists{"/Applications/Inkscape.app/Contents/Resources/bin/inkscape"}{\newcommand{\inkscapedir}{/Applications/Inkscape.app/Contents/Resources/bin/}}{}
-\providecommand{\inkscapedir}{}
-%
-\epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{"\inkscapedir inkscape" -f #1 -C -A \OutputFile}
-%
-\usepackage{adjustbox}
-\newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
-%
-%\usepackage{paralist}
-%\newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
-%\newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
-%
-%\usepackage[inline]{enumitem}
-\newlist{myenuma}{enumerate*}{4}
-\setlist[myenuma]{label=(\alph*),ref=\alph*}
-\newcommand{\itemp}[1]{\item\aligntop{#1}}
-
-\graphicspath{{"Figs/"}}
-\usepackage{amssymb,amsfonts}
-%\usepackage[centertags]{amsmath}
-%\usepackage{hyperref}
-\usepackage[capitalise,nameinlink]{cleveref}
-\creflabelformat{equation}{[#2#1#3]}
-\crefrangelabelformat{equation}{[#3#1#4--#5#2#6]}
-%\crefformat{equation}{#2Eq.~[#1]#3}
-%\crefrangeformat{equation}{Eqs.~[#3#1#4--#5#2#6]}
-%\crefmultiformat{equation}{Eqs.~[#2#1#3}{,2#1#3]}{,#2#1#3}{,#2#1#3]}
-%\crefrangemultiformat{equation}{Eqs.~[#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}{,#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}
-
-%% OPTIONAL MACRO DEFINITIONS
-\newcommand{\ra}{\rightarrow}
-\newcommand{\lr}{\leftrightarrow}
-\newcommand{\cdt}{\!\cdot\!}
-\newcommand{\vp}{\vspace{0.5cm}}
-\newcommand{\hp}{\hspace{0.5cm}}
-%
-%e.g., i.e. with normal spaces
-\newcommand{\eg}{e.g.\ }
-\newcommand{\ie}{i.e.\ }
-\newcommand{\cf}{cf.\ }
-\newcommand{\etc}{etc.\ }
-\newcommand{\wrt}{wrt.\ }
-\newcommand{\Eg}{E.g.\ }
-\newcommand{\Ie}{I.e.\ }
-\newcommand{\Cf}{Cf.\ }
-\newcommand{\Etc}{Etc.\ }
-%
-\usepackage{delim}
-% brackets etc.
-\delimdef\prn#1{\dleft( #1 \dright)}
-\delimdef\brc#1{\dleft\{ #1 \dright\}}
-\delimdef\brk#1{\dleft[ #1 \dright]}
-\delimdef\abs#1{\dleft\lvert #1 \dright\rvert}
-\delimdef\nrm#1{\dleft\lVert #1 \dright\rVert}
-\delimdef\av#1{\dleft\langle #1 \dright\rangle}
-\delimdef\avc#1#2{\dleft\langle #1 \dmiddle\vert #2 \dright\rangle}
-%\newenvironment{cases}{\dleft\{\begin{aligned}}{\end{aligned}\dright.}
-%
-% Sets
-\delimdef\set#1#2{\dleft\{ #1 \dmiddle\vert #2 \dright\}}
-\delimdef\cond#1#2{\!\dleft( #1 \dmiddle\vert #2 \dright)}
-\delimdef\condb#1#2{\!\dleft[ #1 \dmiddle\vert #2 \dright]}
-%
-% Derivatives, etc. First argument is optional.
-\newcommand{\diff}[3][]{\frac{\mathrm{d}^{#1} #2}{\mathrm{d}{#3}^{#1}}}
-\newcommand{\pdiff}[3][]{\frac{\partial^{#1} #2}{\partial {#3}^{#1}}}
-\newcommand{\pdiffc}[3][]{\left(\frac{\partial #2}{\partial {#3}}\right)_{\!\!#1}}
-\newcommand{\intd}[2][]{\int #1\!\!\dr #2\,}
-%
-% Un-italicised letters
-\newcommand{\dr}{\mathrm{d}}
-\newcommand{\e}{\mathrm{e}}
-\newcommand{\ir}{\mathrm{i}}
-\DeclareMathOperator{\tr}{tr}
-\DeclareMathOperator{\Tr}{Tr}
-\DeclareMathOperator{\Det}{Det}
-\DeclareMathOperator{\cov}{Cov}
-\DeclareMathOperator{\var}{Var}
-\DeclareMathOperator{\bias}{Bias}
-\DeclareMathOperator{\rank}{rank}
-\DeclareMathOperator{\sgn}{sgn}
-\DeclareMathOperator{\sech}{sech}
-\DeclareMathOperator{\diag}{diag}
-%
-% The default \Im and \Re look crap
-\renewcommand{\Im}{\operatorname{\mathfrak{Im}}}
-\renewcommand{\Re}{\operatorname{\mathfrak{Re}}}
-%
-% logic
-\newcommand{\means}{\Longleftrightarrow}
-\newcommand{\requires}{\Longleftarrow}
-%
-%\input{sl_symbols.tex}
-%
-%additional symbols:
-\newcommand{\half}{\frac{1}{2}}
-\newcommand{\CO}{\mathcal{O}}
-\newcommand{\CL}{\mathcal{L}}
-\newcommand{\dt}{\dr t}
-%
-\DeclareMathOperator{\SNR}{SNR}
-\DeclareMathOperator{\snr}{SNR}
-\newcommand{\wv}{\vec{w}}
-\newcommand{\wvi}{\vec{w}_\text{ideal}}
-%matrices
-\newcommand{\inv}{^{-1}}
-\newcommand{\dg}{^\mathrm{dg}}
-\newcommand{\trans}{^\mathrm{T}}
-\newcommand{\I}{\mathbf{I}}
-%vec of ones
-\newcommand{\onev}{\mathbf{e}}
-%mat of ones
-\newcommand{\onem}{\mathbf{E}}
-%Markov matrix
-\newcommand{\MM}{\mathbf{Q}}
-%prob distributions
-\newcommand{\prob}{\mathbf{p}}
-\newcommand{\eq}{\prob^\infty}
-%first passage times
-\newcommand{\fpt}{\mathbf{T}}
-%off-diag first passage times
-\newcommand{\fptb}{\overline{\fpt}}
-%fundamental matrix
-\newcommand{\fund}{\mathbf{Z}}
-%other symbols for matrices
-\newcommand{\Pb}{\mathbf{P}}
-\newcommand{\D}{\mathbf{D}}
-\newcommand{\pib}{\boldsymbol{\pi}}
-\newcommand{\Lb}{\boldsymbol{\Lambda}}
-\newcommand{\w}{\mathbf{w}}
-\newcommand{\W}{\mathbf{W}}
-\newcommand{\M}{\mathbf{M}}
-\newcommand{\enc}{\mathbf{q}}
-\newcommand{\frg}{\W^{\mathrm{F}}}
-\newcommand{\F}{\boldsymbol{\Phi}}
-%superscripts
-\newcommand{\pot}{^{\text{pot}}}
-\newcommand{\dep}{^{\text{dep}}}
-\newcommand{\potdep}{^{\text{pot/dep}}}
-%sets
-\newcommand{\CS}{\mathcal{S}}
-\newcommand{\CA}{\mathcal{A}}
-\newcommand{\CB}{\mathcal{B}}
-\newcommand{\comp}{^\mathrm{c}}
-%eigenmodes
-\newcommand{\uv}{\mathbf{u}}
-\newcommand{\vv}{\mathbf{v}}
-\newcommand{\CI}{\mathcal{I}}
-
-% ----------------------------------------------------------------
-\input{sl_theorems_preamble.tex}
-% ----------------------------------------------------------------
-\providecommand\hyperref[2][1]{#2}
-\providecommand\texorpdfstring[2]{#1}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1,2]{Subhaneil Lahiri}
-\author[b,1]{Surya Ganguli}
-
-\affil[a]{Department of Applied Physics, Stanford University, Stanford CA}
+\iftoggle{main} {\author[1]{Subhaneil Lahiri}
+\author{Surya Ganguli}}
+{\author{Lahiri \textit{et al.}}}
 
@@
 
+\iftoggle{main}{%
+\affil{Department of Applied Physics, Stanford University, Stanford CA}
 % Please add here a significance statement to explain the relevance of your work
 \significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.}
-
 % Please include corresponding author, author contribution and author declaration information
@@
 %\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: sulahiri@stanford.edu}
-
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: sulahiri@stanford.edu}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{synaptic plasticity | learning | memory | neuroscience}
-
 %\abbreviations{SNR, signal-to-noise ratio; LTP, long term potentiation;LTD, long term depression}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\begin{article}
 \begin{abstract}
@@
 \end{abstract}
+}{}
 
@@
 \doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{article}
-
-% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
-% You should only change this length when you've finalised the article contents.
-\verticaladjustment{-2pt}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-\dropcap{I}t is widely thought that our very ability to remember the past over long time scales depends crucially on our ability to modify synapses in our brain in an experience dependent manner.
-Classical models of synaptic plasticity model synaptic efficacy as an analog scalar value, denoting the size of a post-synaptic potential injected into one neuron from another.
-Theoretical work has shown that such models have a reasonable, extensive memory capacity, in which the number of long term associations that can be stored by a neuron is proportional its number of afferent synapses \cite{Hopfield1982model,Amit1985hopfield,Gardner1988perceptron}.
-However, recent experimental work has shown that many synapses are more digital than analog; they cannot robustly assume an infinite continuum of analog values, but rather can only take on a finite number of distinguishable strengths, a number than can be as small as two \cite{Bliss1993LTP,Petersen1998allornone,O'Connor2005switch} (though see \cite{Enoki2009graded}).
-This one simple modification leads to a catastrophe in memory capacity:  classical models with digital synapses, when operating in a palimpsest mode in which the ongoing storage of new memories can overwrite previous memories, have a memory capacity proportional to the logarithm of the number of synapses \cite{amit1992constraints,amit1994learning}.
-Intuitively, when synapses are digital, the storage of a new memory can flip a population of synaptic switches, thereby rapidly erasing previous memories stored in the same synaptic population.
-This result indicates that the dominant theoretical basis for the storage of long term memories in modifiable synaptic switches is flawed.
-
-Recent work \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial} has suggested that a way out of this logarithmic catastrophe is to expand our theoretical conception of a synapse from a single scalar value to an entire stochastic dynamical system in its own right.
-This conceptual expansion is further necessitated by the experimental reality that synapses contain within them immensely complex molecular signaling pathways, with many internal molecular functional states (\eg see \cite{Bliss1993LTP,Bredt2003AMPA,Coba2009phosphorylation}).
-While externally, synaptic efficacy could be digital, candidate patterns of electrical activity leading to potentiation or depression could yield transitions between these internal molecular states without necessarily inducing an associated change in synaptic efficacy.
-This form of synaptic change, known as metaplasticity \cite{Abraham1996metaplasticity,Montgomery2002765}, can allow the probability of synaptic potentiation or depression to acquire a rich dependence on the history of prior changes in efficacy, thereby potentially improving memory capacity.
-
-Theoretical studies of complex, metaplastic synapses have focused on analyzing the memory performance of a limited number of very specific molecular dynamical systems, characterized by a number of internal states in which potentiation and depression each induce a specific set of allowable transitions between states (\eg see \cref{fig:models} below).
-While these models can vastly outperform simple binary synaptic switches, these analyses leave open several deep and important questions.
-For example, how does the structure of a synaptic dynamical system determine its memory performance?
-What are the fundamental limits of memory performance over the space of all possible synaptic dynamical systems?
-What is the structural organization of synaptic dynamical systems that achieve these limits?
-Moreover, from an experimental perspective, it is unlikely that all synapses can be described by a single canonical synaptic model;  just like the case of neurons, there is an incredible diversity of molecular networks underlying synapses both across species and across brain regions within a single organism \cite{Emes2012synapserev}.
-In order to elucidate the functional contribution of this diverse molecular complexity to learning and memory, it is essential to move beyond the analysis of specific models and instead develop a general theory of learning and memory for complex synapses.
-Moreover, such a general theory of complex synapses could aid in development of novel artificial memory storage devices.  % Sutskever, memory for learning?
-
-Here we initiate such a general theory by proving upper bounds on the memory curve associated with any synaptic dynamical system, within the well established ideal
-observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}.
-Along the way we develop principles based on first passage time theory to order the structure of synaptic dynamical systems and relate this structure to memory performance.
-We summarize our main results in the \hyperref[sec:disc]{discussion section}.
-
-\section{Overall framework: synaptic models and their memory curves}\label{sec:setup}
-
-In this section, we describe the class of models of synaptic plasticity that we are studying and how we quantify their memory performance.
-In the subsequent sections, we will find upper bounds on this performance.
-
-We use a well established formalism for the study of learning and memory with complex synapses (see \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}).
-In this approach, electrical patterns of activity corresponding to candidate potentiating and depressing plasticity events occur randomly and independently at all synapses at a Poisson rate $r$.
-These events reflect possible synaptic changes due to either spontaneous network activity, or the storage of new memories.
-We let  $f\pot$ and  $f\dep$ denote the fraction of these events that are candidate potentiating or depressing events respectively.
-Furthermore, we assume our synaptic model has $M$ internal molecular functional states, and that a candidate potentiating (depotentiating) event induces a stochastic transition in the internal state described by an $M \times M$ discrete time Markov transition matrix  $\M\pot$ ($\M\dep$).
-In this framework, the states of different synapses will be independent, and the entire synaptic population can be fully described by the probability distribution across these states, which we will indicate with the row-vector $\prob(t)$.
-Thus the $i$'th component of $\prob(t)$ denotes the fraction of the synaptic population in state $i$.
-Furthermore, each state $i$ has its own synaptic weight, $\w_i$, which we take, in the worst case scenario, to be restricted to two values.
-After shifting and scaling these two values,  we can assume they are $\pm1$, without loss of generality.
-
-We also employ an ``ideal observer'' approach to the memory readout, where the synaptic weights are read directly.
-This provides an upper bound on the quality of any readout using neural activity.
-
-For any single memory, stored at time $t=0$, we assume there will be an ideal pattern of synaptic weights across a population of $N$ synapses, the $N$-element vector $\wvi$, that is $+1$ at all synapses that experience a candidate potentiation event, and $-1$ at all synapses that experience a candidate depression event at the time of memory storage.
-We assume that any pattern of synaptic weights close to $\wvi$ is sufficient to recall the memory.
-However, the actual pattern of synaptic weights at some later time, $t$, will change to $\wv(t)$ due to further modifications from the storage of subsequent memories.
-We can use the overlap between these, $\wvi\cdt\wv(t)$, as a measure of the quality of the memory.
-As $t\to\infty$, the system will return to its steady state distribution which will be uncorrelated with the memory stored at $t=0$.
-The probability distribution of the quantity $\wvi\cdt\wv(\infty)$ can be used as a ``null model'' for comparison.
-
-The extent to which the memory has been stored is described by a signal-to-noise ratio (SNR) \cite{Fusi2005cascade,Fusi2007multistate}:
-%
-\begin{equation}\label{eq:SNRdef}
-  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}}
-     {\sqrt{\var(\wv_\text{ideal}\cdt\wv(\infty))}}.
-\end{equation}
-%
-The noise in the denominator is essentially $\sqrt{N}$.
-There is a correction when potentiation and depression are imbalanced, but this will not affect the upper bounds that we will discuss below and will be ignored in the subsequent formulae.
-
-A simple average memory curve can be derived as follows.
-All of the preceding plasticity events, prior to $t=0$, will put the population of synapses in its steady-state distribution, $\eq$.
-The memory we are tracking at $t=0$ will change the internal state distribution to $\eq\M\pot$ (or $\eq\M\dep$) in those synapses that experience a candidate potentiation (or depression) event.
-As the potentiating/depressing nature of the subsequent memories is independent of $\wvi$, we can average over all sequences, resulting in the evolution of the probability distribution:
-%
-\begin{multline}\label{eq:evol}
-  \diff{\prob(t)}{t} = r\prob(t)\frg,
-  \\ \text{where} \quad
-  \frg = f\pot\M\pot + f\dep\M\dep - \I.
-\end{multline}
-%
-Here $\frg$ is a continuous time transition matrix that models the process of forgetting the memory stored at time $t=0$ due to random candidate potentiation/depression events occurring at each synapse due to the storage of subsequent memories.
-Its stationary distribution is $\eq$.
-
-This results in the following SNR
-%
-\begin{equation}\label{eq:SNRcalc}
-  \snr(t) = \sqrt{N}(2 f\pot f\dep) \eq \brk{\M\pot-\M\dep} \e^{rt\frg} \w.
-\end{equation}
-%
-%The factor of $\eq$ describes the synapses being in the steady-state distribution before the memory is encoded.
-The factor of $(\M\pot-\M\dep)$ comes from the encoding of the memory at $t=0$, with $\wvi$ being $\pm1$ in synapses that are potentiated/depotentiated.
-The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
-A detailed derivation of this formula can be found in the supplementary material.
-We will frequently refer to this function as the memory curve.
-It can be thought of as the excess fraction of synapses (relative to equilibrium) that maintain their ideal synaptic strength at time $t$, as dictated by the stored memory at time $t=0$.
-
-Much of the previous work on these types of complex synaptic models has focused on understanding the memory curves of specific models, or choices of $\M\potdep$.
-Two examples of these models are shown in \cref{fig:models}. We see that they have different memory properties.
-The serial model performs relatively well at one particular timescale, but it performs poorly at other times.
-The cascade model does not perform quite as well at that time, but it maintains its performance over a wider range of timescales.
-
-\begin{figure}[tbp]
- \begin{center}
- \begin{myenuma}
-  \protect\item\aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
-  \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
-  \item\aligntop{\includegraphics[width=0.3\linewidth]{cascms.eps}}\label{fig:cascms}
- \end{myenuma}
- \end{center}
-  \caption{Models of complex synapses.
-  (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
-  (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
-  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined \hyperref[sec:setup]{below}) as subsequent memories are stored.
-  \label{fig:models}}
-\end{figure}
-
-In this work, rather than analyzing specific models,  we take a different approach, in order to obtain a more general theory.
-We consider the \emph{entire} space of these models and find upper bounds on the memory capacity of \emph{any} of them.
-The space of models with a fixed number of internal states $M$ is parameterized by the pair of $M \times M$ discrete time stochastic
-transition matrices $\M\pot$ and $\M\dep$, in addition to $f\potdep$.  The parameters must satisfy the following constraints:
-%
-\begin{equation}\label{eq:constr}
-\begin{gathered}
-\begin{aligned}
-  \M\potdep_{ij} &\in [0,1], &\qquad
-  \sum_j \M\potdep_{ij} &= 1, \\
-  f\potdep &\in [0,1],&
-  f\pot + f\dep &= 1,
-\end{aligned}
-\\
-  \eq\frg = 0, \qquad
-  \sum_i \eq_i = 1, \qquad
-  \w_i = \pm 1.
-\end{gathered}
-\end{equation}
-%
-The upper bounds on $\M\potdep_{ij}$ and $f\potdep$ follow automatically from the other constraints.
-
-The critical question is: what do these constraints imply about the space of achievable memory curves in \cref{eq:SNRcalc}?
-To answer this question, especially for limits on achievable memory at finite times, it will be useful to employ the eigenmode decomposition:
-%
-\begin{equation}\label{eq:eigendecomp}
-\begin{aligned}
-  \frg &= \sum_a -q_a \uv^a \vv^a,
-  &\quad
-  \vv^a \uv^b &= \delta_{ab},
-  \\%\quad
-  \frg \uv^a &= -q_a \uv^a,
-  &\quad
-  \vv^a \frg &= -q_a \vv^a.
-\end{aligned}
-\end{equation}
-%
-Here $q_a$ are the negative of the eigenvalues of the forgetting process $\frg$, $\uv^a$ are the right (column) eigenvectors and $\vv^a$ are the left (row) eigenvectors.
-This decomposition allows us to write the memory curve as a sum of exponentials,
-%
-\begin{equation}\label{eq:SNReigen}
-\begin{aligned}
-  \snr(t) &= \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a},
-%  &\quad \text{where}&\;&
-%  \CI_a &= \prn{2 f\pot f\dep} \eq (\M\pot-\M\dep) \uv^a  \vv^a \w,\\&
-%  & \text{and}&&
-%  \tau_a &= \frac{1}{q_a}.
-\end{aligned}
-\end{equation}
-%
-where $\;\CI_a = (2 f\pot f\dep) \eq (\M\pot-\M\dep) \uv^a  \vv^a \w\;$ and $\;\tau_a = {1}/{q_a}$.
-We can then ask the question: what are the constraints on these quantities, namely eigenmode initial SNR's, $\CI_a$, and time constants, $\tau_a$, implied by the constraints in \cref{eq:constr}?
-We will derive some of these constraints in the next section.
-
-\section{Upper bounds on achievable memory capacity}\label{sec:bounds}
-
-In the previous section, in \cref{eq:SNRcalc} we have described an analytic expression for a memory curve as a function of the structure of a synaptic dynamical system, described by the pair of stochastic transition matrices  $\M\potdep$.
-Since the performance measure for memory is an entire memory curve, and not just a single number, there is no universal scalar notion of optimal memory in the space of synaptic dynamical systems.
-Instead there are tradeoffs between storing proximal and distal memories;  often attempts to increase memory at late (early) times by changing $\M\potdep$, incurs a performance loss in memory at early (late) times in specific models considered so far \cite{Fusi2005cascade,Fusi2007multistate,Leibold2008serial}.
-Thus our end goal, achieved \hyperref[sec:env]{below}, is to derive an envelope memory curve in the SNR-time plane, or a curve that forms an upper-bound on the {\it entire} memory curve of {\it any} model.
-In order to achieve this goal, in this section, we must first derive upper bounds, over the space of all possible synaptic models, on two different scalar functions of the memory curve: its initial SNR, and the area under the memory curve.
-In the process of upper-bounding the area, we will develop an essential framework to organize the structure of synaptic dynamical systems based on first passage time theory.
-
-\subsection{Bounding initial SNR}\label{sec:initial}
-
-We now give an upper bound on the initial SNR,
-%
-\begin{equation}\label{eq:init}
-  \snr(0) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \w,
-\end{equation}
-%
-over \emph{all} possible models and also find the class of models that saturate this bound.
-A useful quantity is the equilibrium probability flux between two disjoint sets of states, $\CA$ and $\CB$:
-%
-\begin{equation}\label{eq:flux}
-  \F_{\CA\CB} = \sum_{i\in\CA} \sum_{j\in\CB} r \eq_i \frg_{ij}.
-\end{equation}
-%
-The initial SNR is closely related to the flux from the states with $\w_i=-1$ to those with $\w_j=+1$ (see supplementary material):
-%
-\begin{equation}\label{eq:initflux}
-  \snr(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}\,.
-\end{equation}
-%
-This inequality becomes an equality if potentiation never decreases the synaptic weight and depression never increases it, which should be a property of any sensible model.
-
-To maximize this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
-An example of such a model is shown in \cref{fig:max_init}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
-These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \cref{fig:max_init}(\ref{fig:binary_det}).
-
-This two state model has the equilibrium distribution $\eq=(f\dep,f\pot)$ and its flux is given by $\F_{-+} = rf\pot f\dep$.
-This is maximized when $f\pot=f\dep=\half$, leading to the upper bound:
-%
-\begin{equation}\label{eq:max_init}
-  \snr(0) \leq \sqrt{N}.
-\end{equation}
-%
-We note that while this model has high initial SNR, it also has very fast memory decay -- with a timescale $\tau\sim\frac{1}{r}$.
-As the synapse is very plastic, the initial memory is encoded very easily, but the subsequent memories also overwrite it rapidly.
-This is one example of the tradeoff between optimizing memory at early versus late times.
-\begin{figure}[tbp]
- \begin{center}
- \begin{myenuma}
-  \item\hp\aligntop{\includegraphics[width=0.24\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
-  \item\aligntop{\includegraphics[width=0.24\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
-  \item\hp\aligntop{\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
-  \end{myenuma}
- \end{center}
-  \caption{Synaptic models that maximize initial SNR.
-  (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
-  (\ref{fig:max_init_pot}) Depression is similar to potentiation, but with strong and weak interchanged.
-  (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
-  \label{fig:max_init}}
-\end{figure}
-
-\subsection{Imposing order on internal states through first passage times}\label{sec:order}
-
-Our goal of understanding the relationship between structure and function in the space of all possible synaptic models is complicated by the fact that this space contains many different possible network topologies, encoded in the nonzero matrix elements of $\M\potdep$.
-To systematically analyze this entire space, we develop an important organizing principle using the theory of first passage times in the stochastic process of forgetting, described by $\frg$.
-The mean first passage time matrix, $\smash{\fptb_{ij}}$, is defined as the average time it takes to reach state $j$ for the first time, starting from state $i$.
-The diagonal elements are defined to be zero.
-
-A remarkable theorem we will exploit is that the quantity
-%
-\begin{equation}\label{eq:kemenyconst}
-  \eta \equiv \sum_j \fptb_{ij}\eq_j,
-\end{equation}
-%
-known as Kemeny's constant (see \cite[\S4.4]{kemeny1960finite}), is independent of the starting state $i$.
-Intuitively, \cref{eq:kemenyconst} states that the average time it takes to reach any state, weighted by its equilibrium probability, is independent of the starting state, implying a hidden constancy inherent in any stochastic process.
-
-In the context of complex synapses, we can define the partial sums
-%
-\begin{equation}\label{eq:kemenypm}
-   \eta^+_i = \sum_{j\in+} \fptb_{ij} \eq_j,
-   \qquad
-   \eta^-_i = \sum_{j\in-} \fptb_{ij} \eq_j.
-\end{equation}
-%
-These can be thought of as the average time it takes to reach the strong/weak states respectively.
-Using these definitions, we can then impose an order on the states by arranging them in order of decreasing $\smash{\eta^+_i}$ or increasing $\smash{\eta^-_i}$.
-Because $\smash{\eta^+_i+\eta^-_i=\eta}$ is independent of $i$, the two orderings are the same.
-In this order, which depends sensitively on the structure of $\M\potdep$, states later (to the right in figures below) can be considered to be more potentiated than states earlier (to the left in figures below), despite the fact that they have the same synaptic efficacy.
-In essence, in this order, a state is considered to be more potentiated if the average time it takes to reach all the strong efficacy states is shorter.
-We will see that synaptic models that optimize various measures of memory have an exceedingly simple structure when, and only when, their states are arranged in this order.%
-\footnote{Note that we do not need to worry about the order of the $\eta^\pm_i$ changing during the optimization: necessary conditions for a maximum only require that there is no infinitesimal perturbation that increases the area.
-Therefore we need only consider an infinitesimal neighborhood of the model, in which the order will not change.}
-
-\subsection{Bounding area}\label{sec:area}
-
-Now consider the area under the memory curve:
-%
-\begin{equation}\label{eq:area}
-  A = \int_0^\infty\!\!\dr t \, \snr(t).
-\end{equation}
-%
-We will find an upper bound on this quantity as well as the model that saturates this bound.
-
-First passage time theory introduced in the \hyperref[sec:order]{previous section} becomes useful because the area has a simple expression in terms of quantities introduced in \cref{eq:kemenypm} (see supplementary material):
-%
-\begin{equation}\label{eq:areaEta}
-\begin{aligned}
-  A &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \!\brk{\M\pot_{ij}-\M\dep_{ij}}\! (\eta^+_i-\eta^+_j)\\
-    &= \sqrt{N}(4f\pot f\dep) \sum_{ij} \eq_i \!\brk{\M\pot_{ij}-\M\dep_{ij}}\! (\eta^-_j-\eta^-_i).
-\end{aligned}
-\end{equation}
-%
-
-With the states in the order described above, we can find perturbations of $\M\potdep$ that will always increase the area, whilst leaving the equilibrium distribution, $\eq$, unchanged.
-Some of these perturbations are shown in \cref{fig:perts}, see supplementary material for details.
-For example, in \cref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
-$\smash{\eta^+_i > \eta^+_j}$), we have proven that increasing $\smash{\M\pot_{ij}}$ and decreasing $\smash{\M\dep_{ij}}$ leads to an increase in area.
-The only thing that can prevent these perturbations from increasing the area is when they require the decrease of a matrix element that has already been set to 0.
-This determines the topology (non-zero transition probabilities) of the model with maximal area.
-It is of the form shown in \cref{fig:env}(\ref{fig:sticky}),%\cref{fig:max_area},
-with potentiation moving one step to the right and depression moving one step to the left.
-Any other topology would allow some class of perturbations (\eg\ in \cref{fig:perts}) to further increase the area.
-
-\begin{figure}[tbp]
- \begin{center}
- \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.14\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
-  \item\aligntop{\includegraphics[width=0.14\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
-  \item\aligntop{\includegraphics[width=0.28\linewidth]{shortcut.svg}}\label{fig:shortcut}
- \end{myenuma}
- \end{center}
-  \caption{Perturbations that increase the area.
-  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$.
-It can no longer be used when $\M\dep$ is lower triangular, \ie depression must move synapses to ``more depressed'' states.
-  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$.
-It can no longer be used when $\M\pot$ is upper triangular, \ie potentiation must move synapses to ``more potentiated'' states.
-  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions.
-It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
-  \label{fig:perts}}
-\end{figure}
-
-As these perturbations do not change the equilibrium distribution, this means that the area of \emph{any} model is bounded by that of a linear chain with the same equilibrium distribution.
-The area of a linear chain model can be expressed directly in terms of its equilibrium state distribution, $\eq$, yielding the following upper bound on the area of any model with the same $\eq$ (see supplementary material):
-%
-\begin{equation}\label{eq:multistatearea}
-\begin{aligned}
-  A &\leq \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k \\
-    &= \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
-\end{aligned}
-\end{equation}
-%
-where we chose $\w_k=\sgn[{k - \sum_j j\eq_j}]$.
-We can then maximize this by pushing all of the equilibrium distribution symmetrically to the two end states.
-This can be done by reducing the transition probabilities out of these states, as in \cref{fig:env}(\ref{fig:sticky}). %\cref{fig:max_area}.
-This makes it very difficult to exit these states once they have been entered.
-The resulting area is
-%
-\begin{equation}\label{eq:max_area}
-  A \leq \frac{\sqrt{N}(M-1)}{r}.
-\end{equation}
-%
-This analytical result is similar to a numerical result found in \cite{Barrett2008discrete} under a slightly different information theoretic measure of memory performance.
-
-The ``sticky'' end states result in very slow decay of memory, but they also make it difficult to encode the memory in the first place, since a small fraction of synapses are able to change synaptic efficacy during the storage of a new memory.
-Thus models that maximize area optimize memory at late times, at the expense of early times.
-
-%\begin{figure}[tbp]
-% \begin{center}
-%  \includegraphics[width=0.35\linewidth]{multistate_sticky.svg}
-% \end{center}
-%  \caption{Model that maximizes the area.
-%  Unlabeled transitions have probability 1.
-%  Labeled transitions have probability $\varepsilon\to0$.
-%  }\label{fig:max_area}
-%\end{figure}
-
-\section{Memory curve envelope}\label{sec:env}
-
-Now we will look at the implications of the upper bounds found in the \hyperref[sec:bounds]{previous section} for the SNR at finite times.
-As argued in \cref{eq:SNReigen}, the memory curve can be written in the form
-%
-\begin{equation}\label{eq:SNRmodes}
-  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a}.
-\end{equation}
-%
-The upper bounds on the initial SNR, \cref{eq:max_init}, and the area, \cref{eq:max_area}, imply the following constraints on the parameters $\brc{\CI_a,\tau_a}$:
-%
-\begin{equation}\label{eq:modeconstr}
-  \sum_a \CI_a \leq 1,
-  \qquad\qquad
-  \sum_a \CI_a \tau_a \leq M-1.
-\end{equation}
-%
-We are not claiming that these are a complete set of constraints: not every set $\brc{\CI_a,\tau_a}$ that satisfies these inequalities will actually be achievable by a synaptic model.
-However, any set that violates either inequality will definitely not be achievable.
-
-Now we can pick some fixed time, $\smash{t_0}$, and maximize the SNR at that time \wrt the parameters $\smash{\brc{\CI_a,\tau_a}}$, subject to the constraints above.
-This always results in a single nonzero $\smash{\CI_a}$; in essence, optimizing memory at a single time requires a single exponential.
-The resulting optimal memory curve, along with the achieved memory at the chosen time, depends on $t_0$ as follows:
-%
-\begin{equation}\label{eq:env}
-\begin{aligned}
-  t_0 &\leq \frac{M-1}{r} &
-   & \implies &
-  \snr(t) &= \sqrt{N}\e^{-rt/(M-1)} \\&&
-   & \implies &
-  \snr(t_0) &= \sqrt{N}\e^{-rt_0/(M-1)},\\
-  t_0 &\geq \frac{M-1}{r} &
-   & \implies &
-  \snr(t) &= \frac{\sqrt{N}(M-1)\e^{-t/t_0}}{rt_0} \\&&
-   & \implies &
-  \snr(t_0) &= \frac{\sqrt{N}(M-1)}{\e rt_0}.
-\end{aligned}
-\end{equation}
-%
-Both the initial SNR bound and the area bound are saturated at early times.
-At late times, only the area bound is saturated.
-The function $\snr(t_0)$, the green curve in \cref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim \smash{t_0\inv}$.
-No synaptic model can have an SNR that is greater than this at any time.
-We can use this to find an upper bound on the memory lifetime, $\tau(\epsilon)$, by finding the point at which the envelope crosses $\epsilon$:
-%
-\begin{equation}\label{eq:lifebound}
-  \tau(\epsilon) \leq \frac{\sqrt{N}(M-1)}{\epsilon\e r},
-\end{equation}
-%
-where we assume $N>(\epsilon\e)^2$.
-Intriguingly, both the lifetime and memory envelope expand linearly with the number of internal states $M$, and increase as the square root of the number of synapses $N$.
-
-This leaves the question of whether this bound is achievable.
-At any time, can we find a model whose memory curve touches the envelope?
-The red curves in \cref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
-
-We see that at early, but not late times, there is a gap between the upper bound that we can prove and what we can achieve with actual models.
-There may be other models we haven't found that could beat the ones we have, and come closer to our proven envelope.
-However, we suspect that the area constraint is not the bottleneck for optimizing memory at times less than  $\CO(\frac{M}{r})$.
-We believe there is some other constraint that prevents models from approaching the envelope, and currently are exploring several mathematical conjectures for the precise form of this constraint in order to obtain a potentially tighter envelope.
-Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \cref{fig:env}(\ref{fig:sticky})%\cref{fig:max_area}
- can come close to the envelope.
-
-\begin{figure}
- \begin{center}
-% \parbox[t]{0.43\linewidth}{%
- \begin{myenuma}
-   \item\aligntop{\includegraphics[width=0.85\linewidth]{env.svg}}\label{fig:envplot}
-% \end{myenuma}}\hp
-% \parbox[t]{0.44\linewidth}{%
-% \begin{myenuma}
-%  \stepcounter{myenumai}
-%% \addtocounter{enumi}{1}
-   \item\aligntop{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}%\\[0.1\linewidth]
-   \item\aligntop{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
- \end{myenuma}%}
- \end{center}
-  \caption{The memory curve envelope for $N=100$, $M=12$.
-  (\protect\protect\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
-  The red dashed curve shows the result of numerical optimization of synaptic models with random initialization.
-  The solid red curve shows the highest SNR we have found with hand designed models.
-  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states, and all transition probabilities equal to $1$.
-  At late times they are of the form shown in (\ref{fig:sticky}) with different values of $\varepsilon$.
-  The model shown in (\ref{fig:sticky}) also saturates the area bound \cref{eq:max_area} in the limit $\varepsilon\to0$.
-  \label{fig:env}}
-\end{figure}
-
-\section{Discussion}\label{sec:disc}
-
-We have initiated the development of  a general theory of learning and memory with complex synapses, allowing for an exploration of the entire space of complex synaptic models, rather than analyzing individual models one at a time.
-In doing so, we have obtained several new mathematical results delineating the functional limits of memory achievable by synaptic complexity, and the structural characterization of synaptic dynamical systems that achieve these limits.
-In particular, operating within the ideal observer framework of \cite{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}, we have shown that for a population of $N$ synapses with $M$ internal states, \begin{myenuma}
-\item the initial SNR of any synaptic model cannot exceed $\sqrt{N}$, and any model that achieves this bound is equivalent to a binary synapse,
-\item the area under the memory curve of any model cannot exceed that of a linear chain model with the same equilibrium distribution,
-\item both the area and memory lifetime of any model cannot exceed $\CO(\sqrt{N}M)$, and the model that achieves this limit has a linear chain topology with only nearest neighbor transitions,
-\item we have derived an envelope memory curve in the SNR-time plane that cannot be exceeded by the memory curve of any model, and models that approach this envelope for times greater than $\CO(\frac{M}{r})$ are linear chain models, and
-\item this late-time envelope is a power-law proportional to $\smash{\CO({\sqrt{N} M}/{rt})}$, indicating that synaptic complexity can strongly enhance the limits of achievable memory.
-\end{myenuma}
-
-This theoretical study opens up several avenues for further inquiry.
-In particular, the tightness of our envelope for early times, less than $\CO(\frac{M}{r})$, remains an open question, and we are currently pursuing several conjectures.
-We have also derived memory constrained envelopes, by asking in the space of models that achieve a given SNR at a given time, what is the maximal SNR achievable at other times.
-If these two times are beyond a threshold separation, optimal constrained models require two exponentials.
-It would be interesting to systematically analyze the space of models that achieve good memory at multiple times, and understand their structural organization, and how they give rise to multiple exponentials, leading to power law memory decays.
-
-%Another important consideration is the tolerance of the synapses to biological noise, which could be investigated by imposing lower limits on the level of stochasticity of the models.
-Finally, it would be interesting to design physiological experiments in order to perform optimal systems identification of potential Markovian dynamical systems hiding within biological synapses,
-given measurements of pre and post-synaptic spike trains along with changes in post-synaptic potentials.
-Then given our theory, we could match this measured synaptic model to optimal models to understand for which timescales of memory, if any, biological synaptic dynamics may be tuned.
-
-In summary, we hope that a deeper theoretical understanding of the functional role of synaptic complexity, initiated here, will help advance our understanding of the neurobiology of learning and memory, aid in the design of engineered memory circuits, and lead to new mathematical theorems about stochastic processes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\acknow{We thank Sloan, Genenetech, Burroughs-Wellcome, and Swartz foundations for support.
-We thank Larry Abbott, Marcus Benna, Stefano Fusi, Jascha Sohl-Dickstein and David Sussillo for useful discussions.}
-\showacknow % Display the acknowledgments section
-
-% \pnasbreak splits and balances the columns before the references.
-% If you see unexpected formatting errors, try commenting out this line
-% as it can run into problems with floats and footnotes on the final page.
-%\pnasbreak
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{pnas-new}
-\bibliography{maths,neuro,markov}
-
-%\end{article}
-\clearpage
-
-\begin{strip}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\appendix
-\begin{center}
-  \textbf{\Large Supplementary material}
-\end{center}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Here we provide more details underlying the derivations of results in the main paper.
-
-\section{Continuous time Markov processes}\label{sec:ContMarkov}
-
-In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time to define notation.
-It is a generalization of material that can be found in \cite{kemeny1960finite} with some ideas from \cite{hunter2000survey}.
-
-\subsection{Notation}\label{sec:not}
-
-For any matrix $\mathbf{A}$, we define matrices $\mathbf{A}\dg$ and $\overline{\mathbf{A}}$ as
-%
-\begin{equation}\label{eq:dgdef}
-  \mathbf{A}\dg_{ij} \equiv \delta_{ij}\mathbf{A}_{ij},
-  \qquad
-  \overline{\mathbf{A}} \equiv \mathbf{A}-\mathbf{A}\dg.
-\end{equation}
-%
-We let $\onev$ denote a column-vector of ones and $\onem=\onev\onev\trans$ denote a matrix of ones.
-
-A continuous time Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$ with row sums equal to zero ($\MM\onev=0$).
-The probabilities of being in each state at time $t$, the row-vector $\mathbf{p}(t)$, evolve according to
-%
-\begin{equation}\label{eq:rowsum}
-  \diff{\mathbf{p}(t)}{t} = \mathbf{p}(t) \MM,
-\end{equation}
-%
-where $\mathbf{p}(t)\onev=1$.
-
-The equilibrium probabilities, $\eq$, satisfy
-%
-\begin{equation}\label{eq:equilibrium}
-  \eq\MM=0,
-  \qquad
-  \eq\onev=1.
-\end{equation}
-%
-As we assume an ergodic process, this eigenvalue is non-degenerate.
-If all other eigenvalues have strictly negative real parts, the process is regular (aperiodic).
-
-We define additional matrices
-%
-\begin{equation}\label{eq:defDLP}
-  \Lb \equiv (-\MM\dg)\inv,
-  \qquad
-  \Pb \equiv \I + \Lb\MM.
-\end{equation}
-%
-It can be shown that $\Lb_{ii}$ is the mean time it takes to leave state $i$ and $\Pb_{ij}$ is the probability the the next transition from state $i$ goes to state $j$:
-%
-\begin{equation}\label{eq:LamdaPcmpt}
-  \Lb_{ii} = \frac{1}{\sum_{j \neq j} \MM_{ij}},
-  \qquad
-  \Pb_{ij} = %\left\{
-%  \begin{aligned}
-%     &0                                         && \text{if }i=j, \\
-%     &\frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &&\text{otherwise}.
-%  \end{aligned}
-%  \right.
-  \begin{cases}
-     0                                         &\text{if }i=j, \\
-     \frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &\text{otherwise}.
-  \end{cases}
-\end{equation}
-%
-Furthermore, we also define
-%
-\begin{equation}\label{eq:pdotD}
-  \D \equiv \operatorname{diag}(\eq)\inv,
-  \qquad\implies\qquad
-  \eq\D=\onev\trans.
-\end{equation}
-%
-
-\subsection{Fundamental matrix}\label{sec:fund}
-
-For our results below regarding the integral of the memory curve, it can be useful to invert the stochastic transition matrix, $\MM$.
-However, since $\MM$ has a zero eigenvalue, it cannot be inverted.
-For this reason, the fundamental matrix arises as a useful surrogate for the inverse of $\MM$.
-It is related to the first passage times, as we will see in the \hyperref[sec:fpt]{next subsection}.
-Here we define the fundamental matrix and review its properties.
-
-\begin{defn}[Fundamental matrix]
-  For discrete time, the generalized fundamental matrix was defined in \cite{Kemeny1981fund}.
-  For continuous time, we define:
-  %
-  \begin{equation}\label{eq:funddef}
-    \fund \equiv (-\MM + \onev\pib)\inv,
-  \end{equation}
-  %
-  where $\pib$ is any row-vector with $\pib\onev=1/\tau\neq0$.
-\end{defn}
-Note that the canonical choice for the discrete time version, $\pib=\eq$, is not available here due to problems with units.
-It will be helpful to choose $\pib$ to be independent of $\MM$, \eg $\pib=\onev\trans/(n\tau)$.
-All quantities that we calculate using $\fund$ below will be independent of this choice.
-
-\begin{thm}
-  The definition of $\fund$ is valid, \ie $(-\MM + \onev\pib)$ is invertible.
-\end{thm}
-\begin{proof}
-  Assume there exists an $\mathbf{x}$ such that
-  %
-  \begin{equation}\label{eq:fundinvkern}
-    (-\MM + \onev\pib)\mathbf{x}=0.
-  \end{equation}
-  %
-  Multiplying from the left with $\eq$ gives
-  %
-  \begin{equation}\label{eq:fundinvpix}
-    \pib \mathbf{x} = 0.
-  \end{equation}
-  %
-  Substituting back into \cref{eq:fundinvkern} gives
-  %
-  \begin{equation*}
-    \MM\mathbf{x}=0.
-  \end{equation*}
-  %
-  As we assume an ergodic process, the zero eigenvalue is non-degenerate.
-  Therefore, $\mathbf{x}=\lambda\onev$.
-  Substituting this into \cref{eq:fundinvpix} gives
-  %
-  \begin{equation*}
-    \lambda\pib\onev = \frac{\lambda}{\tau} = 0.
-  \end{equation*}
-  %
-  As we defined $\pib$ such that $1/\tau\neq0$, this means $\lambda=0 \implies \mathbf{x}=0$.
-\end{proof}
-
-\begin{cor}
-  %
-  \begin{align}
-    \pib\fund &= \eq, \label{eq:fundprob}\\
-    \fund\onev &= \tau\onev,\label{eq:fundrowsum}\\
-    \I+\MM\fund &= \onev\eq, \label{eq:fundQZ}\\
-    \I+\fund\MM &= \tau\onev\pib. \label{eq:fundZQ}
-  \end{align}
-  %
-\end{cor}
-\begin{proof}
-  We can deduce \cref{eq:fundprob,eq:fundrowsum} be pre/post-multiplying the following equations by $\fund$:
-  %
-  \begin{equation*}
-    \begin{aligned}
-      \eq(-\MM + \onev\pib) &= \pib, \\
-      (-\MM + \onev\pib)\onev &= \frac{\onev}{\tau}.
-    \end{aligned}
-  \end{equation*}
-  %
-  We can then deduce \cref{eq:fundQZ,eq:fundZQ} by substituting these into
-  %
-  \begin{equation*}
-    (-\MM + \onev\pib)\fund = \fund(-\MM + \onev\pib) = \I.
-  \end{equation*}
-  %
-\end{proof}
-
-%We will not need to use the concept of a g-inverse anywhere, so you can skip ahead to \cref{sec:fpt} now. For the sake of completeness, we can define $\MM^\#$ as
-%%
-%\begin{equation}\label{eq:ginvdef}
-%  \MM^\# \equiv \tau\onev\eq - \fund
-%\end{equation}
-%%
-%\begin{thm}
-%  The matrix $\MM^\#$ satisfies the following defining properties of a g-inverse of $\MM$:
-%  %
-%  \begin{align}
-%    \MM^\# \MM \MM^\# &= \MM^\#,     \label{eq:ginvpropo} \\
-%    \MM \MM^\# \MM    &= \MM,        \label{eq:ginvpropi} \\
-%    \MM \MM^\#        &= \MM^\# \MM, \label{eq:ginvcomm}
-%  \end{align}
-%  %
-%  with \cref{eq:ginvcomm} holding iff $\eq$ and $\pib$ are proportional to each other.
-%\end{thm}
-%\begin{proof}
-%  Using \cref{eq:rowsum}, \cref{eq:equilibrium}, \cref{eq:fundprob} and \cref{eq:fundrowsum}, one can verify
-%  %
-%  \begin{alignat*}{2}
-%    \MM \MM^\# &= -\MM \fund,
-%    &\qquad
-%    \pib \MM^\# &= 0,
-%    \\
-%    \MM^\# \MM &= -\fund \MM,
-%    &
-%    \MM^\# \onev &= 0.
-%  \end{alignat*}
-%  %
-%  Substituting into \cref{eq:fundQZ} and \cref{eq:fundZQ} leads to
-%  %
-%  \begin{equation*}
-%    \I - \MM \MM^\# = \onev \eq,
-%    \qquad
-%    \I - \MM^\# \MM = \tau \onev \pib.
-%  \end{equation*}
-%  %
-%  Pre/post-multiplying these by $\MM$ or $\MM^\#$ produces \cref{eq:ginvpropo} and \cref{eq:ginvpropi}. Subtracting produces
-%  %
-%  \begin{equation*}
-%    [\MM,\MM^\#] = \onev(\tau\pib-\eq),
-%  \end{equation*}
-%  %
-%  which vanishes iff $\eq$ and $\pib$ are proportional to each other (due to $\pib\onev=1/\tau$ and $\eq\onev=1$, the constant of proportionality could only be $\tau$).
-%\end{proof}
-
-\subsection{First passage times}\label{sec:fpt}
-
-\begin{defn}[First passage time matrix]
- We define $\fptb_{ij}$ as the mean time it takes the process to reach state $j$ for the first time, starting from state $i$.
- We also define $\fpt\dg_{ii}$ as the mean time it takes the process to return to state $i$.
- As usual, $\fpt=\fptb+\fpt\dg$.
-\end{defn}
-
-This matrix is given by
-%
-\begin{equation}\label{eq:fpt}
-  \fpt=(\onem\fund\dg - \fund + \Lb)\D,
-\end{equation}
-%
-see \cite{Yao1985fpt} for a proof.
-We can separate this into its diagonal and off-diagonal pieces.
-
-The recurrence times are given by
-%
-\begin{equation}\label{eq:recurtime}
-  \fpt\dg = \Lb\D.
-\end{equation}
-%
-or in component form
-%
-\begin{equation*}
-  \eq_i \Lb_{ii}\inv \fpt\dg_{ii} = 1.
-\end{equation*}
-%
-The extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.
-
-The off-diagonal mean first passage times are given by
-%
-\begin{equation}\label{eq:fptfund}
-  \fptb = (\onem\fund\dg - \fund)\D.
-\end{equation}
-%
-or in component form:
-%
-\begin{equation}\label{eq:fptfundcmpt}
-  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
-\end{equation}
-%
-
-\subsection{Mixing time (Kemeny's constant)}\label{sec:mixtime}
-
-\begin{thm} \mbox{}\label{th:kemenyconst}
-  The quantity
-  %
-  \begin{equation}\label{eq:mixdef}
-    \eta \equiv \sum_j \fptb_{ij}\eq_j
-  \end{equation}
-  %
-  is independent of $i$.
-\end{thm}
-\begin{proof}
-  For discrete time, a proof can be found in \cite[Th.4.4.10]{kemeny1960finite}.
-  For continuous time, we use \cref{eq:fptfund}, \cref{eq:fundrowsum} and the transpose of \cref{eq:pdotD}:
-  %
-  \begin{equation*}
-    \begin{aligned}
-      \fptb (\eq)\trans &= (\onem\fund\dg - \fund) \D (\eq)\trans \\
-        &= (\onev\onev\trans\fund\dg - \fund) \onev \\
-        &= (\onev\trans\fund\dg\onev)\onev - \fund\onev \\
-        &= (\tr\fund - \tau)\onev.
-    \end{aligned}
-  \end{equation*}
-  %
-  which proves \cref{eq:mixdef} with $\eta = \tr\fund - \tau$.
-\end{proof}
-
-Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version, where this would only shift $\eta$ by 1.
-
-\subsection{Sensitivity of equilibrium distribution}\label{sec:sensitivity}
-
-Suppose that the Markov process, defined by $\MM$, depends on some parameter $\alpha$.
-Differentiating \cref{eq:funddef} gives
-%
-\begin{equation}\label{eq:diffZ}
-  \diff{\fund}{\alpha} = \fund \diff{\MM}{\alpha} \fund.
-\end{equation}
-%
-We can substitute this into the derivative of \cref{eq:fundprob}:
-%
-\begin{equation}\label{eq:diffp}
-  \diff{\eq}{\alpha} = \pib \fund \diff{\MM}{\alpha} \fund = \eq \diff{\MM}{\alpha} \fund.
-\end{equation}
-%
-We can rewrite this in component form and use the fact that $\MM_{ii} = - \sum_{i\neq j} \MM_{ij}$:
-%
-\begin{equation}\label{eq:diffpT}
-\begin{aligned}
-  \diff{\eq_k}{\alpha} &= \sum_{i,j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} \\
-    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} + \sum_i \eq_i \diff{\MM_{ii}}{\alpha} \fund_{ik} \\
-    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} (\fund_{jk} - \fund_{ik}) \\
-    &= \sum_{i\neq j} \diff{\MM_{ij}}{\alpha} \eq_i \eq_k (\fptb_{ik} - \fptb_{jk}).
-\end{aligned}
-\end{equation}
-%
-This is a generalization of a result of \cite{cho2000markov} from discrete to continuous time that we will need below.
-Note that the summand vanishes for $i=j$, so we can drop the restriction $i\neq j$ from the range of the sum.
-
-\subsection{Subsets and flux}\label{sec:subsets}
-
-Let us denote the set of states by $\CS$.
-Consider a subset $\CA\subset\CS$.
-We can define a projection operator onto this subset:
-%
-\begin{equation}\label{eq:proj}
-  \prn{\I^\CA}_{ij} =
-    \begin{cases}
-      1 &\text{if $i=j\in\CA$,}\\
-      0 &\text{otherwise.}
-    \end{cases}
-\end{equation}
-%
-We will use superscripts/subscripts to denote projection onto/summation over a subset:
-%
-\begin{equation}\label{eq:projsum}
-  \begin{aligned}
-    \pib^\CA &=\pib\I^\CA, &
-    \M^{\cdot\CA}&=\M\I^\CA, &
-    \M^{\CA\cdot}&=\I^\CA\M, &
-    \mathbf{x}^\CA &= \I^\CA\mathbf{x},
-    \\
-    \pib_\CA &=\pib\onev^\CA, &
-    \M_{\cdot\CA}&=\M\onev^\CA, &
-    \M_{\CA\cdot}&=\prn{\onev^\CA}\trans\M, &
-    \mathbf{x}_\CA &= \prn{\onev^\CA}\trans\mathbf{x},
-  \end{aligned}
-\end{equation}
-%
-where $\pib$ is a row vector, $\M$ is a matrix and $\mathbf{x}$ is a column vector.
-
-We can define a flux matrix, a.k.a.\ ergodic flow:
-%
-\begin{equation}\label{eq:flux:s}
-  \F = \D\inv\MM,
-  \qquad
-  \F_{ij} = \eq_i \MM_{ij}.
-\end{equation}
-%
-This measures the flow of probability between states in the equilibrium distribution.
-Detailed balance, a.k.a.\ reversibility, is equivalent to $\F=\F\trans$.
-
-The flux between two subsets is a particularly useful quantity:
-%
-\begin{equation}\label{eq:subflux}
-  \F_{\CA\CB} = {\eq}^\CA\MM\onev^\CB.
-\end{equation}
-%
-One can show that
-%
-\begin{equation}\label{eq:compflux}
-  \F_{\CA\CA\comp} = \F_{\CA\comp\CA} = -\F_{\CA\CA} = -\F_{\CA\comp\CA\comp}
-\end{equation}
-%
-using $\prn{{\eq}^\CA + {\eq}^{\CA\comp}}\MM=0$ and $\MM\prn{{\onev}^\CA + {\onev}^{\CA\comp}}=0$.
-
-\subsection{Lumpability}\label{sec:lump}
-
-Suppose we have partitioned the states into disjoint subsets, $\brc{\CA_\alpha}$:
-%
-\begin{equation}\label{eq:partition}
-  \bigcup_\alpha \CA_\alpha = \CS,
-  \qquad
-  \CA_\alpha \cap \CA_\beta = \delta_{\alpha\beta}\CA_\alpha.
-\end{equation}
-%
-We will use $\alpha$ instead of $\CA_\alpha$ in superscripts and subscripts in the following.
-The fact that these subsets are disjoint and exhaustive allows us to define the function
-%
-\begin{equation}\label{eq:whichsey}
-  \sigma(i)=\alpha
-  \qquad\means\qquad
-  i\in\CA_\alpha.
-\end{equation}
-%
-
-We can use this partition to define a new stochastic process associated with the original Markov chain.
-At time $t$, if the state of the original process is $i$, the state of the new process is $\sigma(i)$.
-
-One may ask if this new process is a Markov chain.
-The answer is yes, if the original Markov chain has a property called lumpability \wrt the partition
-(see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time):
-%
-\begin{equation}\label{eq:lump}
-  \sigma(i)=\sigma(j)
-  \quad\implies \quad
-  \MM_{i\alpha}=\MM_{j\alpha} \equiv \sum_{k\in\CA_\alpha} \MM_{jk},
-\end{equation}
-%
-\ie the total transition rate from some state to some subset is the same for all starting states within the same subset.
-This common value is the transition rate for the new lumped Markov chain.
-
-This can be rewritten with the aid of two matrices
-%
-\begin{equation}\label{eq:lumpmats}
-  U_{\alpha i} = \frac{\delta_{\alpha\sigma(i)}}{\abs{\CA_\alpha}},
-  \qquad
-  V_{i\alpha} = \delta_{\sigma(i)\alpha}.
-\end{equation}
-%
-Left multiplication by $U$ averages over subsets, right multiplication by $V$ sums over subsets.
-For $U$, we chose the uniform measure in each subset. Any measure would work equally well, \eg one proportional to the equilibrium distribution:
-%
-\begin{equation}\label{eq:altlumpmats}
-  U_{\alpha i} = \frac{{\eq_i}^\alpha}{\eq_\alpha}.
-\end{equation}
-%
-
-One can show that $(UV)_{\alpha\beta}=\delta_{\alpha\beta}$.
-The matrix $VU$ is also interesting.
-It has a block diagonal structure, with each block corresponding to a subset.
-Each block is a discrete-time ergodic Markov matrix (it is an independent trials process with probabilities given by the measure chosen for $U$).
-This means that the right eigenvectors with eigenvalue 1 will be those that are constant in each subset:
-%
-\begin{equation}\label{eq:setconst}
-  VU\mathbf{x}=\mathbf{x}
-  \qquad\means\qquad
-  \mathbf{x} = \sum_\alpha x_\alpha \onev^\alpha.
-\end{equation}
-%
-
-This allows us to write the lumpability condition \cref{eq:lump}, and the transition matrix for the lumped process compactly:
-%
-\begin{equation}\label{eq:lumpcompact}
-  VU\MM V = \MM V,
-  \qquad
-  \widehat{\MM} = U\MM V.
-\end{equation}
-%
-By induction, one can show that similar relations hold for all powers:
-%
-\begin{equation}\label{eq:lumppower}
-  VU\MM^nV = \MM^nV,
-  \qquad
-  \widehat{\MM}^n = U\MM^n V,
-\end{equation}
-%
-and, via the Taylor series, for the exponential as well:
-%
-\begin{equation}\label{eq:lumpexp}
-  VU \e^{t\MM} V = \e^{t\MM}V,
-  \qquad
-  \e^{t\widehat{\MM}} = U\e^{t\MM} V.
-\end{equation}
-%
-The equilibrium distribution of the lumped process is given by
-%
-\begin{equation}\label{eq:lumpeq}
-  \widehat{\mathbf{p}}^\infty = \eq V.
-\end{equation}
-%
-
-\section{Signal-to-Noise ratio (SNR)}\label{sec:SNR}
-
-In this section we will look at the signal-to-noise curve, and put an upper bound on its initial value.
-We need only consider ergodic Markov chains.
-Transient states would be unoccupied in equilibrium and would not be accessed by the signal creation process, therefore they could be removed from the analysis.
-Absorbing chains are degenerate cases: they have zero initial signal but infinite decay times, so they can only be approached as the limit of a sequence of ergodic chains.
-
-\subsection{Framework}\label{sec:framework}
-
-The individual potentiation/depression events will be described by \emph{discrete}-time Markov chains:
-%
-\begin{equation}\label{eq:MWdef}
-  \M\potdep \equiv \I + \W\potdep,
-  \qquad
-  \M\potdep\onev = \onev,
-  \qquad
-  \M\potdep_{ij} \in [0,1].
-\end{equation}
-%
-The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf\potdep$, where $f\pot +f\dep =1$ are the fraction of plasticity events that are potentiating/depressing respectively.
-This means that the ``forgetting'' process will be described by the \emph{continuous}-time Markov chain:
-%
-\begin{equation}\label{eq:forgetting}
-  \MM = r \frg \equiv r\prn{f\pot\W\pot + f\dep\W\dep}.
-\end{equation}
-%
-We only require that this Markov chain is ergodic.
-The Markov chains described by $\M\potdep $ need not be.
-
-We assume that the probability distribution starts in the equilibrium distribution \cref{eq:equilibrium}.
-During the initial signal creation, a fraction $f\pot$ will change to $\eq\M\pot$ and a fraction $f\dep$ will change to $\eq\M\dep$.
-After this, probabilities will evolve according to \cref{eq:rowsum}.
-
-\subsection{SNR curve}\label{sec:SNRcurve}
-
-As discussed in the main text, the signal-to-noise ratio is given by
-%
-\begin{equation}\label{eq:SNRdef:s}
-  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}}
-     {\sqrt{\var(\wv_\text{ideal}\cdt\wv(\infty))}}.
-\end{equation}
-%
-First, let's look at the denominator, remembering that the states and plasticity events of each synapse are independent and identically distributed:
-%
-\begin{equation}\label{eq:noise}
-\begin{aligned}
-  \var(\wv_\text{ideal}\cdt\wv(\infty))
-    &= \sum_{\alpha\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty) \wv^\beta_\text{ideal}\wv^\beta(\infty)}
-    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2 \\
-    &= \sum_{\alpha} \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}
-    + \sum_{\alpha\neq\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}\av{\smash{\wv^\beta}_\text{ideal}\smash{\wv^\beta}(\infty)}
-    \\&\phantom{= \sum_{\alpha}  \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}}
-    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2  \\
-    &= N\av{1}
-    + N(N-1)\av{\wv^1_\text{ideal}\wv^1(\infty)}^2
-    - N^2\av{\wv^1_\text{ideal}\wv^1(\infty)}^2 \\
-    &= N(1-\av{\wv^1_\text{ideal}\wv^1(\infty)}^2),
-\end{aligned}
-\end{equation}
-%
-where we used $\wv^\alpha=\pm1$.
-
-For the numerator, we can write
-%
-\begin{equation}\label{eq:overlapex}
-  \av{\wv_\text{ideal}\cdt\wv(t)} = \sum_\alpha\av{\wv^\alpha_\text{ideal}\wv^\alpha(t)}
-   = N \av{\wv^1_\text{ideal}\wv^1(t)},
-\end{equation}
-%
-Noting that $\wvi=\pm1$ with probability $f\potdep$,
-%
-\begin{equation}\label{eq:overlapstate}
-\begin{aligned}
-  \av{\wv^1_\text{ideal}\wv^1(t)} &= f\pot \av{\wv^1(t)}_{\text{pot},t=0} - f\dep \av{\wv^1(t)}_{\text{dep},t=0} \\
-   &= f\pot \sum_i P(\text{state}=i,t \mid \text{pot},0) \w_i - f\dep \sum_i P(\text{state}=i,t \mid \text{dep},0) \w_i.
-\end{aligned}
-\end{equation}
-%
-From the previous section,
-%
-\begin{equation}\label{eq:stateprob}
-  P(\text{state}=i,t \mid \text{pot/dep},0) = \brk{ \eq \M\potdep\, \e^{rt\frg} }_i,
-\end{equation}
-%
-which describes the synapses starting in the equilibrium distribution, changing state due to the plasticity event at $t=0$ and subsequent evolution according to \cref{eq:rowsum} due to plasticity events uncorrelated with $\wvi$.%
-\footnote{Note that expanding the exponential gives
-%
-\begin{equation*}
-  \e^{rt\frg} =  \sum_{n=0}^\infty \frac{(rt)^n\,\e^{-rt}}{n!} \sum_{m=0}^n (f\pot)^m (f\dep)^{n-m}
-  \brk{ \M\pot \M\dep \M\pot \M\pot \ldots + \text{permutations}}.
-\end{equation*}
-%
-Thus, evolving according to \cref{eq:rowsum} results in averaging over all sequences of plasticity events,
-as we only need linear expectations of $\wv(t)$ in the end.}
-This results in
-%
-\begin{equation}\label{eq:overlapwt}
-\begin{aligned}
-  \av{\wv^1_\text{ideal}\wv^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
-  \av{\wv^1_\text{ideal}\wv^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
-         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
-         &=  (f\pot-f\dep)\, \eq \w \\
-         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
-\end{aligned}
-\end{equation}
-%
-Combining these allows us to write the numerator as
-%
-\begin{equation}\label{eq:signal}
-\begin{aligned}
-  \av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}
-    &= N \eq (f\pot(\M\pot-\I)-f\dep(\M\dep-\I))\, \e^{rt\frg} \w \\
-    &= N \eq (f\pot(\W\pot-\frg)-f\dep(\W\dep-\frg))\, \e^{rt\frg} \w \\
-%    &= N \eq (f\pot(f\dep\W\pot-f\dep\W\dep)-f\dep(f\pot\W\dep-f\pot\W\pot))\,\e^{rt\frg}\w \\
-    &= N (2f\pot f\dep)\eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
-\end{aligned}
-\end{equation}
-%
-where we used $\eq\frg=0$ in going from the first to second lines.
-Combining with \cref{eq:noise} gives
-%
-\begin{equation}\label{eq:SNRcurveExact}
-  \SNR(t) = \frac{\sqrt{N} (2f\pot f\dep ) \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w}
-                 {\sqrt{1-(f\pot-f\dep)^2(\eq_+-\eq_-)^2}}.
-\end{equation}
-%
-The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
-We can set the denominator to 1 without changing any of our results.
-
-This results in our final formula:
-%
-\begin{equation}\label{eq:SNRcurve}
-  \SNR(t) = \sqrt{N} (2f\pot f\dep )\, \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
-\end{equation}
-%
-The factor of $\eq$ describes the synapses being in the steady-state distribution before the memory is encoded.
-The factor of $(\M\pot-\M\dep)$ comes from the encoding of the memory at $t=0$, with $\wvi$ being $\pm1$ in synapses that are potentiated/depotentiated.
-The factor of $\e^{rt\frg}$ describes the subsequent evolution of the probability distribution, averaged over all sequences of plasticity events, and the factor of $\w$ indicates the readout via the synaptic weight.
-
-We can express this in terms of the one parameter family of transition matrices:
-%
-\begin{equation}\label{eq:Walpha}
-  \begin{aligned}
-  \W(\alpha) &= \alpha\W\pot +(1-\alpha)\W\dep ,
-  &\quad&\implies&
-    \frg&=\W(f\pot ),\\&&&&
-    \W\pot  - \W\dep  &= \diff{\W}{\alpha},\\&&&&
-    \eq \diff{\W}{\alpha} &= -\diff{\eq}{\alpha} \frg.
-  \end{aligned}
-\end{equation}
-%
-Then \cref{eq:SNRcurve} becomes
-%
-\begin{equation}\label{eq:SNRalpha}
-  \SNR(t) = \sqrt{N} (2f\pot f\dep ) \diff{\eq}{\alpha}(-\frg)\,\e^{rt\frg}\w.
-\end{equation}
-%
-
-\subsection{Lumpability and the SNR curve}\label{sec:SNRlump}
-
-Suppose that we have a partition such that $\W\pot $ and $\W\dep $ are simultaneously lumpable, and that all the states in each subset have the same synaptic strength (see \cref{sec:lump}):
-%
-\begin{equation}\label{eq:lumpablesynapse}
-  VU\W\potdep  V = \W\potdep  V,
-  \qquad
-  VU\w=\w.
-\end{equation}
-%
-We can define a new synapse with
-%
-\begin{equation}\label{eq:lumpedsynapse}
-  \widehat{\W}\potdep  = U\W\potdep  V,
-  \qquad
-  \widehat{\w} = U \w,
-  \qquad
-  \widehat{\mathbf{p}}^\infty = \eq V.
-\end{equation}
-%
-This synapse has an SNR curve:
-%
-\begin{equation}\label{eq:lumpedSNR}
-  \begin{aligned}
-    \frac{\SNR(t)}{\sqrt{N} (2f\pot f\dep )} &=  \widehat{\mathbf{p}}^\infty (\widehat{\W}\pot  - \widehat{\W}\dep )\e^{rt\widehat{\W}^F}\widehat{\w}. \\
-      &= \eq VU (\W\pot  - \W\dep ) VU \e^{rt\frg} VU \w. \\
-      &= \eq (\W\pot  - \W\dep ) VU \e^{rt\frg} VU \w. \\
-      &= \eq (\W\pot  - \W\dep )\e^{rt\frg} VU \w. \\
-      &= \eq (\W\pot  - \W\dep )\e^{rt\frg}\w. \\
-  \end{aligned}
-\end{equation}
-%
-\ie the lumped process has exactly the same SNR as the original one.
-
-\subsection{Initial SNR and flux}\label{sec:initflux}
-
-Using $\eq\frg=0$ and the first line of \cref{eq:signal}, we can write the initial SNR as
-%
-\begin{equation}\label{eq:init:s}
-  \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f\pot \W\pot -f\dep \W\dep )(\onev^+-\onev^-).
-\end{equation}
-%
-Using $\W\potdep (\onev^++\onev^-)=0$ and \cref{eq:compflux}:
-%
-\begin{equation*}
-  r{\eq}^-(f\pot \W\pot +f\dep \W\dep )\onev^+ = \F_{-+} = \F_{+-} = r{\eq}^+(f\pot \W\pot +f\dep \W\dep )\onev^-,
-\end{equation*}
-%
-we can rewrite \cref{eq:init} as
-%
-\begin{equation}\label{eq:initflux:s}
-  I = \frac{4\F_{-+}}{r} - 4{\eq}^+f\pot \W\pot \onev^- - 4{\eq}^-f\dep \W\dep \onev^+.
-\end{equation}
-%
-The last two terms are guaranteed to be negative, as the diagonal parts of $\W\potdep $ cannot contribute.
-Therefore
-%
-\begin{equation}\label{eq:initfluxineq}
-  \SNR(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}.
-\end{equation}
-%
-This inequality is saturated if potentiation never takes it from a $+$ state to a $-$ state and depression never takes it from a $-$ state to a $+$ state.
-
-\section{Area maximisation}\label{sec:areamax}
-
-In this section we will find an upper bound on the area under the signal-to-noise curve.
-As in \cref{sec:SNR}, we will only consider ergodic Markov chains.
-We will see in \cref{sec:multistate} that the optimal chain is absorbing, so it lies on the boundary of the (open) set of ergodic chains, but it still puts an upper bound on the area for any chain in the interior.
-
-\subsection{Area under signal-to-noise curve}\label{sec:areadef}
-
-The signal-to-noise curve is given by \cref{eq:SNRalpha}.
-%%
-%\begin{equation}\label{eq:SNR}
-%  \SNR(t) = \sqrt{N}(2f^+f\dep ) \diff{\eq}{\alpha} (-\frg) \e^{rt\frg} \w.
-%\end{equation}
-%%
-The area is computed by integrating this
-%
-\begin{equation}\label{eq:area:s}
-\begin{aligned}
-  A &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} \brk{-\e^{rt\frg}}_0^\infty \w \\
-    &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} (\I-\onev\eq) \w \\
-    &= \frac{\sqrt{N}(2f\pot f\dep )}{r} \diff{\eq}{\alpha} \w.
-\end{aligned}
-\end{equation}
-%
-
-We can rewrite this using \cref{eq:diffpT}, with $A=\sqrt{N}(2f\pot f\dep )\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W\pot _{ij}-\W\dep _{ij}$
-%
-\begin{equation}\label{eq:areaT}
-  \hat{A} = \sum_{i,j,k} \eq_i \enc_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
-\end{equation}
-%
-
-\begin{defn}[Partial mixing times]
-We define the $\pm$ mixing times as
-%
-\begin{equation}\label{eq:mixingpm}
-\begin{aligned}
-  \eta^\pm_i &\equiv \sum_k \fptb_{ik} \eq_k \prn{\frac{1 \pm \w_k}{2}}
-%    &&= \sum_{\set{k}{\w_k=\pm1}} \fptb_{ik} \eq_k \\
-    &&= \sum_{k\in\pm} \fptb_{ik} \eq_k \\
-    &= \sum_k \prn{\fund_{kk}-\fund_{ik}} \prn{\frac{1 \pm \w_k}{2}}
-    &&= \sum_{k\in\pm} \prn{\fund_{kk}-\fund_{ik}} .
-\end{aligned}
-\end{equation}
-%
-We can think of $\eta^+_i$ as a measure of the ``distance'' to the $\w_k=+1$ states and $\eta^-_i$ as the ``distance'' to the $\w_k=-1$ states.
-\end{defn}
-Using \cref{eq:mixdef}, we can write:
-%
-\begin{equation}\label{eq:mixingrels}
-\begin{aligned}
-  \eta^+_i + \eta^-_i &= \eta,\\
-  2(\eta^+_i - \eta^+_j) &= \sum_k (\fptb_{ik}-\fptb_{jk}) \eq_k \w_k
-    = \sum_k (\fund_{jk}-\fund_{ik}) \w_k.
-\end{aligned}
-\end{equation}
-%
-We could arrange the states in order of decreasing $\eta^+$, which is the same as the order of increasing $\eta^-$.
-
-We can rewrite \cref{eq:areaT} as
-%
-\begin{equation}\label{eq:areaEta:s}
-\begin{aligned}
-  \hat{A} &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}) &
-    &= -2\sum_{i,j} \enc_{ij} \eq_i \eta^+_{j} \\
-    &= 2\sum_{i,j} \enc_{ij} \eq_i (\eta^-_{j} - \eta^-_{i}) &
-    &= 2\sum_{i,j} \enc_{ij} \eq_i \eta^-_{j}.
-\end{aligned}
-\end{equation}
-%
-We can also express it in terms of the fundamental matrix \cref{eq:funddef} as
-%
-\begin{equation}\label{eq:areaZ}
-  \hat{A} = \sum_{i,j,k,l} \enc_{ij} \pib_{l} \fund_{li} (\fund_{jk}-\fund_{ik}) \w_k
-    = \pib \fund q \fund \w.
-\end{equation}
-%
-
-It is also helpful to define the following quantities:
-%
-\begin{equation}\label{eq:areacoeffs}
-  \begin{aligned}
-    c_k &= \diff{\ln\eq_k}{\alpha}
-      = \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}}
-      = - \prn{\eq \enc \fptb}_k
-      = \frac{(\eq \enc \fund)_k}{\eq_k}, \\
-    a_i &= \sum_j \enc_{ij} \eq_i (\eta^+_{i} - \eta^+_{j}),\\
-    \implies
-    \hat{A} &= \sum_k c_k \eq_k \w_k
-      = 2\sum_i a_i.
-  \end{aligned}
-\end{equation}
-%
-Note that the optimal choice of $\w$ is $\w_k = \sgn(c_k)$.
-
-\subsection{Derivatives \wrt \texorpdfstring{$\W\potdep $}{W(pot/dep)}}\label{sec:deriv}
-
-In the following, we will mathematically define the classes of perturbations pictorially described in Figure 3 of the main paper.
-In order to do so, we will need to consider expressions for derivatives of various quantities with respect to $\W\potdep_{ij}$.
-
-As discussed in the main text,
-we will regard the off-diagonal elements of $\W\potdep _{ij}$ to be the independent variables,
-with $\W\potdep _{ii}=-\sum_{j \neq i} \W\potdep _{ij}$ imposed by hand.
-Thus,
-%
-\begin{equation}\label{eq:basicderivs}
-  \pdiff{\frg_{ij}}{\W\potdep _{gh}} = f\potdep  \delta_{gi}(\delta_{hj}-\delta_{ij}),
-  \qquad
-  \pdiff{\enc_{ij}}{\W\potdep _{gh}} = \pm\delta_{gi}(\delta_{hj}-\delta_{ij}).
-\end{equation}
-%
-The implicit $g \neq h$ that comes with all derivatives is unnecessary, as the derivatives above vanish when $g=h$.
-
-In particular, differentiating \cref{eq:funddef},
-%
-\begin{equation}\label{eq:derivZ}
-  \pdiff{\fund_{ij}}{\W\potdep _{gh}} = rf\potdep  \fund_{ig} (\fund_{hj}-\fund_{gj}).
-\end{equation}
-%
-We can then differentiate expression \cref{eq:areaZ} to get
-%
-\begin{equation}\label{eq:derivA}
-  \begin{aligned}
-    \pdiff{\hat{A}}{\W\potdep _{gh}} =&\,
-%      rf\potdep  \sum_{ijkl} \enc_{ij} \w_k \pib_l \brk{
-%         \fund_{lg}(\fund_{hi}-\fund_{gi})(\fund_{jk}-\fund_{ik}) + \fund_{li}(\fund_{jg}-\fund_{ig})(\fund_{hk}-\fund_{gk}) } \\
-%    &\pm \sum_{kl} \w_k \pib_l \fund_{lg} (\fund_{hk} - \fund_{gk}) \\
-%    =&\,
-%      2rf\potdep  \sum_{ij} \enc_{ij} \eq_i \eq_g \brk{
-%         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j}) +
-%         (\fptb_{ig}-\fptb_{jg})(\eta^+_{g}-\eta^+_{h}) } \\
-%      &\pm 2 \eq_g (\eta^+_g - \eta^+_h) \\
-%    =&\,
-%      2rf\potdep  \sum_{ij} \enc_{ij} \eq_i \eq_g
-%         (\fptb_{gi}-\fptb_{hi})(\eta^+_{i}-\eta^+_{j})
-%      \pm 2 \eq_g (\eta^+_g - \eta^+_h)\brk{1+\pdiff{\ln(\eq_g)}{\,\ln(f\potdep )}} \\
-%    =&\,
-      2rf\potdep  \eq_g \brk{\sum_i a_i (\fptb_{gi}-\fptb_{hi}) + c_g (\eta^+_{g}-\eta^+_{h})}
-      \pm 2 \eq_g (\eta^+_g - \eta^+_h). \\
-  \end{aligned}
-\end{equation}
-%
-where $a_i$ and $c_k$ were defined in \cref{eq:areacoeffs}.
-
-It is sometimes useful to consider the following derivatives:
-%
-\begin{equation}\label{eq:derivqw}
-  \begin{aligned}
-    \pdiff{}{\frg_{gh}} &\equiv     \pdiff{}{\W\pot _{gh}} +     \pdiff{}{\W\dep _{gh}} ,&\qquad
-    \pdiff{}{\enc_{gh}}    &\equiv f\dep  \pdiff{}{\W\pot _{gh}} - f\pot  \pdiff{}{\W\dep _{gh}} .
-  \end{aligned}
-\end{equation}
-%
-Each of these derivatives behaves as their names suggest:
-%
-\begin{equation}\label{eq:derivqweff}
-  \pdiff{\frg_{ij}}{\frg_{gh}} = \pdiff{\enc_{ij}}{\enc_{gh}}
-  =  \delta_{gi}(\delta_{hj}-\delta_{ij}),
-  \qquad
-  \pdiff{\enc_{ij}}{\frg_{gh}} = \pdiff{\frg_{ij}}{q_{gh}}  = 0.
-\end{equation}
-%
-This is because we could treat $\frg$ and $q$ as the independent variables. However, the boundaries of the allowed region are more easily expressed in terms of $\W\potdep $.
-
-\subsubsection{Scaling mode}\label{sec:rescale}
-
-Consider the following differential operator:
-%
-\begin{equation}\label{eq:scaleop}
-  \Delta \equiv \sum_{g,h} \W\pot _{gh}\pdiff{}{\W\pot _{gh}} + \W\dep _{gh}\pdiff{}{\W\dep _{gh}}.
-\end{equation}
-%
-This corresponds to the scaling, $\W\potdep  \to (1+\epsilon)\W\potdep $.
-Intuitively, this has two effects: it scales up the initial potentiation/depression and it scales down all timescales.
-This intuition is confirmed by the following results:
-%
-\begin{equation}\label{eq:scaleeffects}
-  \begin{aligned}
-    \Delta \fund &= \tau\onev\eq - \fund ,&
-    \Delta \eq  &= 0 ,&
-    \Delta \fpt  &= -\fpt ,
-    \\
-    \Delta \eta^\pm_i  &= - \eta^\pm_i ,&
-    \Delta \enc_{ij} &= \enc_{ij} ,&
-    \Delta \hat{A}  &= 0 ,&
-  \end{aligned}
-\end{equation}
-%
-The anomalous bit in the scaling of $\fund$ is due to the lack of dependence of $\pib$ and $\tau$ on $\W\potdep $.
-
-As the area is invariant under this scaling, we can consider the $\W\potdep $ to be projective coordinates.
-Therefore we don't need to enforce the lower bound on the diagonal matrix elements while looking for the maximum area, as we can use this null-mode to enforce it afterwards without changing the area.
-%We also don't have to worry about the boundary at infinity, as sending some of them to infinity is equivalent to sending the rest to zero.
-
-\subsection{Kuhn-Tucker conditions}\label{sec:kuhntucker}
-
-Consider the Lagrangian
-%
-\begin{equation}\label{eq:lagrangian}
-  \CL = \hat{A} + \sum_{\text{pot/dep}}\sum_{i\neq j} \mu\potdep _{ij} \W\potdep _{ij} .%  +\lambda\eq\w.
-\end{equation}
-%
-%The last term ensures that we hold $\eq_\pm$ fixed during this extremisation, so that we can safely ignore the factors that we dropped from the noise in \cref{eq:noisepm}.
-Necessary conditions for an extremum are
-%
-\begin{equation}\label{eq:extremum}
-  \pdiff{\CL}{\W\potdep _{gh}} = 0,
-  \qquad
-    \mu\potdep _{gh} \geq 0,\quad
-    \W\potdep _{gh} \geq 0,\quad
-    \mu\potdep _{gh}\W\potdep _{gh} = 0.
-\end{equation}
-%
-with $g \neq h$.
-This enforces the positivity constraints on the off-diagonal elements, but not the diagonals.
-As discussed in \cref{sec:rescale}, that can be enforced after finding the maximum using the null scaling degree of freedom.
-
-\subsubsection{Triangularity}\label{sec:triangular}
-
-Here we describe the perturbations corresponding to Figure 3(a,b) of the main paper.
-
-Consider
-%
-\begin{equation}\label{eq:shiftqderiv}
-  \pdiff{\CL}{\enc_{gh}} =
-  f\dep  \pdiff{\CL}{\W\pot _{gh}} - f\pot  \pdiff{\CL}{\W\dep _{gh}}
-   = (f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh}) + 2\eq_g (\eta^+_g - \eta^+_h)
-   = 0.
-\end{equation}
-%
-This corresponds to the shift
-%
-\begin{equation}\label{eq:shiftq}
-  \W\pot _{ij} \to \W\pot _{ij} + f\dep \epsilon_{ij},
-  \qquad
-  \W\dep _{ij} \to \W\dep _{ij} - f\pot \epsilon_{ij},
-  \qquad
-  \sum_j \epsilon_{ij} = 0,
-\end{equation}
-%
-which leaves $\frg$ unchanged, and therefore $\eq$, $\fpt$ and $\eta^\pm$ as well.
-
-Assume $\eta^+_g > \eta^+_h$. Then
-%
-\begin{equation}\label{eq:lowertriangular}
- f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh} <0
- \qquad\implies\qquad
- \mu\dep _{gh} >0
- \qquad\implies\qquad
- \W\dep _{gh}=0.
-\end{equation}
-%
-Similarly, if $\eta^+_g < \eta^+_h$, then
-%
-\begin{equation}\label{eq:uppertriangular}
- f\dep  \mu\pot _{gh} - f\pot  \mu\dep _{gh} > 0
- \qquad\implies\qquad
- \mu\pot _{gh} >0
- \qquad\implies\qquad
- \W\pot _{gh}=0.
-\end{equation}
-%
-Thus, if we arrange the states in order of decreasing $\eta^+$, $\W\pot $ is upper-triangular and $\W\dep $ is lower triangular.
-
-We have ignored the possibility that $\eq_g=0$, as this would imply that $\fpt_{ig}=\infty$, which would in turn imply that the Markov process is not ergodic.
-
-%Note that if we multiply \cref{eq:shiftqderiv} by $\enc_{gh}$ and sum over $(g,h)$, we get
-%%
-%\begin{equation}\label{eq:stupid}
-%  \hat{A} = \sum_\pm\sum_{g\neq h} \mu\potdep _{gh} \frg_{gh},
-%\end{equation}
-%%
-%which is probably useless
-
-\subsubsection{Shortcuts}\label{sec:shortcuts}
-
-In this subsection we will define perturbations corresponding to Figure 3(c) of the main text.
-
-Now consider the following combinations of derivatives for $m>1$:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:shortcutderiv}
-    \widetilde{\Delta}\potdep _{g,m} &\equiv
-      \brk{ \sum_{k=0}^{m-1} \frac{1}{\eq_{g\pm k}} \pdiffc{}{\W\potdep _{g\pm k,g\pm(k+1)}} }
-      - \frac{1}{\eq_g} \pdiffc{}{\W\potdep _{g,g\pm m}}.
-%  \end{aligned}
-\end{align}
-%
-Once again, they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \cref{sec:triangular} about this being satisfied for ergodic chains).
-
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators \cref{eq:diffpT}:
-%
-\begin{equation}\label{eq:shortcutprob}
-  \widetilde{\Delta}\potdep _{g,m} \eq_i = 0,
-\end{equation}
-%
-which makes it possible to integrate the perturbation:
-%
-\begin{equation}\label{eq:shortcutfinite}
-  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon}^{\pm(g,m)},
-  \qquad
-  \begin{aligned}
-    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g,g\pm m}
-      &\!\!\!\!\!\!&= -\epsilon,\\
-    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm(k+1)}
-      &\!\!\!\!\!\!&= \epsilon
-        &\;
-        &\forall\, k \in [0,m-1],\\
-    &\prn{\boldsymbol{\epsilon}^{\pm(g,m)}}_{g\pm k,g\pm k}
-      &\!\!\!\!\!\!&= -\epsilon
-        &
-        &\forall\, k \in [1,m-1].
-  \end{aligned}
-\end{equation}
-%
-But more interestingly for our purposes:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:shortcutarea}
-    \widetilde{\Delta}\potdep _{g,m}\CL &=
-      \brk{ \sum_{k=0}^{m-1} \frac{\mu\potdep _{g\pm k,g\pm(k+1)}}{\eq_{g\pm k}}
-      - \frac{\mu\potdep _{g,g\pm m}}{\eq_g}}
-      + 2rf\potdep  \sum_{k=0}^{m-1} \prn{\eta^+_{g\pm k} - \eta^+_{g\pm(k+1)}} \prn{c_{g\pm k} - c_{g}},
-%  \end{aligned}
-\end{align}
-%
-
-In the \hyperref[sec:areacoeffincr]{section below}, we will show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
-This implies that the last term of the final expression in \cref{eq:shortcutarea} is non-negative.
-If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu\potdep _{g,g\pm m}>0$, which in turn implies that $\W\potdep _{g,g\pm m}=0$.
-This would tell us that the process with the maximal area has to have a multi-state topology.
-
-\subsubsection{Increasing \texorpdfstring{$c_k$}{c(k)}}\label{sec:areacoeffincr}
-
-In the \hyperref[sec:shortcuts]{previous subsection} we defined perturbations corresponding to Figure 3(c) of the main text.
-In order to show that those perturbations increase the area, we must now show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
-
-Consider the following combinations of derivatives:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:areacoeffincrderiv}
-    \Delta_{gh} &\equiv
-      \frac{1}{\eq_{g}} \pdiffc{}{\frg_{gh}}
-      + \frac{1}{\eq_h} \pdiffc{}{\frg_{hg}}, \\
-%  \end{aligned}
-\end{align}
-%
-Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \cref{sec:triangular} about this being satisfied for ergodic chains).
-
-One can show that the equilibrium probabilities, $\eq$, are invariant under these operators using \cref{eq:diffpT}:
-%
-\begin{equation}\label{eq:sareacoeffincrprob}
-  \Delta_{gh} \eq_i = 0,
-\end{equation}
-%
-which makes it possible to integrate the perturbation:
-%
-\begin{equation}\label{eq:areacoeffincrfinite}
-  \W\potdep  \to \W\potdep  + \D\boldsymbol{\epsilon},
-  \qquad
-  \begin{aligned}
-  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
-  \\
-  \boldsymbol{\epsilon} \onev &= 0.
-  \end{aligned}
-\end{equation}
-%
-But more interestingly:
-%
-\begin{align}
-%  \begin{aligned}
-\label{eq:areacoeffincr}
-    \Delta_{gh}\CL &=
-      \frac{\mu\pot _{gh}+\mu\dep _{gh}}{\eq_g} + \frac{\mu\pot _{hg}+\mu\dep _{hg}}{\eq_h}
-      + 2r\prn{c_g-c_h}\prn{\eta^+_{g} - \eta^+_{h}},\\
-%  \end{aligned}
-\end{align}
-%
-where $c_k$ were defined in \cref{eq:areacoeffs}.
-
-Using the non-negativity of the Kuhn-Tucker multipliers, $\mu\potdep _{ij}$, \cref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\eta^+_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\eta^+_k$ are degenerate, we can choose their order to ensure this).
-
-Note that, according to \cref{sec:triangular}, either $\W\pot _{gh}$ or $\W\dep _{gh}$ will be zero at the maximum, therefore we can expect one of $\mu\pot _{gh}+\mu\dep _{gh}$ to be non-zero.
-This would rule out degeneracy of the $c_k$ or $\eta^+_k$.
-Looking at \cref{eq:shiftqderiv} closely, the only way $\mu\pot _{gh}+\mu\dep _{gh}$ could be zero is if $\eta^+_g=\eta^+_h$ or $\eq_g=0$.
-
-%From the expression for the area in \cref{eq:areacoeffs}, it will clearly also be optimal to have the $\w_k$ non-decreasing, \ie all the positive $\w_k$ should lie ahead of all the negative $\w_k$.
-
-\subsubsection{Summary}\label{sec:KTsummary}
-
-Using the Kuhn-Tucker formalism, we have shown that,
-with the states arranged in order of non-increasing $\eta^+_i$:
-%
-\begin{itemize}
-  \item There can be no ergodic maximum for which $\W\pot $ contains backwards transitions or $\W\dep $ contains forwards transitions.
-  \item There can be no ergodic maximum with the $c_k$ decreasing.
-  \item The $c_k$ may only be degenerate at an ergodic maximum if the corresponding $\eta^+_k$ are also degenerate.
-  \item If the $c_k$ increase and the $\eta^+_i$ decrease, there can be no ergodic maximum with shortcuts.
-\end{itemize}
-%
-These were shown by finding allowed perturbations that increase the area.
-
-%Suppose we start with some arbitrary ergodic process.
-If $\W\potdep $ contain backwards/forwards transitions or the $c_k$ decrease, then we can increase the area by applying the perturbations \cref{eq:shiftqderiv} and \cref{eq:areacoeffincrderiv}.
-%The second perturbation will generate the unwanted backwards/forwards transitions, requiring subsequent application of first perturbation.
-%The first perturbation will change the $c_k$, possibly changing their order, requiring subsequent application of second perturbation.
-%
-%This procedure can end in several ways:
-%%
-%\begin{itemize}
-%  \item Applying \cref{eq:areacoeffincrderiv} an infinite amount sends two transition rates to infinity, which due to the scaling mode (\cref{sec:rescale}), is equivalent to sending all other transitions to zero.
-%      Subsequent application of \cref{eq:shiftqderiv} just removes the backwards transition from $\W\pot $ and the forward transition from $\W\dep $.
-%  \item Applying \cref{eq:areacoeffincrderiv} ends in a process with degenerate $c_k$ or $\eta^+_k$, and the subsequent application of \cref{eq:shiftqderiv} doesn't change this degeneracy.
-%  \item As above, but the subsequent application of \cref{eq:shiftqderiv} lifts the degeneracy, resulting in a chain with increasing $c_k$.
-%\end{itemize}
-%%
-%In the third case, we can then apply the perturbation \cref{eq:shortcutderiv} to remove all the shortcuts.
-%Either this, or the first case, results in a multi-state topology, possibly disconnected or without the full set of states.
-
-This leaves two possibilities for the maximum area Markov chain.
-Either there is no degeneracy and no shortcuts, which implies the Multi-state/serial topology that we'll discuss in \cref{sec:multistate}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \cref{eq:shortcutarea}).
-
-Degeneracy tends to be very delicate.
-It is usually hard to arrange without some symmetry relating degenerate states.
-Such a symmetry would imply lumpability (see \cref{sec:lump}).
-The lumped chain would not have any shortcuts, as an entire degenerate set cannot be bypassed.
-As this lumped chain has the same area (see \cref{sec:SNRlump}), we would need only consider the multi-state topology.
-
-\subsection{Multi-state/Serial topology}\label{sec:multistate}
-
-The previous results indicate that the area under the memory curve of any model is bounded by the area under the memory curve of a model with the serial/multistate topology having the same equilibrium distribution.
-Here we compute this area, which we will see depends only on this equilibrium distribution.
-
-The multi-state/serial topology is defined by (see \cite{amit1994learning,Fusi2007multistate,Leibold2008serial}):
-%
-\begin{equation}\label{eq:multistatedef}
-  \W\pot _{ij} = q\pot _i \delta_{i+1,j},
-  \qquad
-  \W\dep _{ij} = q\dep _j \delta_{i,j+1}.
-\end{equation}
-%
-Because it has no shortcuts, it saturates various inequalities:
-%
-\begin{equation}\label{eq:multistateineq}
-  \begin{aligned}
-    \fptb_{ik} - \fptb_{jk} &=
-      \begin{cases}
-        \fptb_{ij},  &\text{if}\quad i \leq j \leq k \quad\text{or}\quad i \geq j \geq k,\\
-        -\fptb_{ji}, &\text{if}\quad j \leq i \leq k \quad\text{or}\quad j \geq i \geq k,\\
-      \end{cases} \\
-    r\eq_i \frg_{ij} \prn{\fptb_{ij}+\fptb_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
-  \end{aligned}
-\end{equation}
-%
-and it satisfies detailed balance (a.k.a.\ reversibility a.k.a. $\CL^2_{\eq}$ self-adjointness):
-%
-\begin{equation}\label{eq:multistateprob}
-  f\pot  q\pot _i \eq_i = f\dep  q\dep _i \eq_{i+1},
-\end{equation}
-%
-which means we can always choose the transition rates, $q\potdep _i$, to give any desired equilibrium probabilities, $\eq_i$.
-
-This allows us to calculate the $c_k$'s:
-%
-\begin{equation}\label{eq:areacoeffchain}
-\begin{aligned}
-  c_k =&\, \sum_{i<k} \fpt_{i,i+1} \prn{\eq_{i}q\pot _i+\eq_{i+1}q\dep _i}
-    %\\&
-    - \sum_{i\geq k} \fpt_{i+1,i} \prn{\eq_{i}q\pot _i+\eq_{i+1}q\dep _i}
-  ,\\
-  c_{k+1} - c_k =&\, \prn{\fpt_{k,k+1}+\fpt_{k+1,k}} \prn{\frac{\eq_{k}\frg_{k,k+1}}{f\pot }+\frac{\eq_{k+1}\frg_{k+1,k}}{f\dep }}
-    =%\\=&\,
-    \frac{1}{rf\pot f\dep },\\
-  \sum_k c_k \eq_k =&\, \sum_{ij} \eq_i \enc_{ij} (\eta-\eta) = 0,\\
-%  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f\pot f\dep },
-  \implies c_k =&\, \frac{k - \sum_j j\eq_j}{rf\pot f\dep },
-\end{aligned}
-\end{equation}
-%
-where we used \cref{eq:multistateineq} to derive the first two equations respectively and Th.\ref{th:kemenyconst} to derive the third. This allows us to write the area as
-%
-\begin{equation}\label{eq:multistatearea:s}
- % A = 2\sqrt{N} \sum_k \brk{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \eq_k \w_k.
-  A = \frac{2\sqrt{N}}{r} \sum_k \brk{k - \sum_j j\eq_j} \eq_k \w_k
-    = \frac{2\sqrt{N}}{r} \sum_k \abs{k - \sum_j j\eq_j} \eq_k ,
-\end{equation}
-%
-where we used $\w_k=\sgn(c_k)$, as discussed after \cref{eq:areacoeffs}.
-This reproduces equation (15) of the main paper.
-
-%It will also help to define $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
-%Note that
-%%
-%\begin{equation}\label{eq:allpplusorminus}
-%  \begin{aligned}
-%    \eq_+ &=1
-%     \quad&\implies\quad
-%     \eq_k \w_k &= \eq_k
-%     \quad&\implies\quad
-%     A=0,\\
-%    \eq_- &=1
-%     \quad&\implies\quad
-%     \eq_k \w_k &= -\eq_k
-%     \quad&\implies\quad
-%     A=0,
-%  \end{aligned}
-%\end{equation}
-%%
-%neither of which are optimal.
-%
-%Consider the Lagrangian
-%%
-%\begin{equation}\label{eq:multistatelagrangian}
-%  \CL = \frac{A}{2\sqrt{N}} + \lambda\prn{1-\sum_i \eq_i} + \sum_i \mu_i \eq_i,
-%\end{equation}
-%%
-%where $\lambda$ is a Lagrange multiplier and $\mu_i$ are Kuhn-Tucker multipliers (see \cref{eq:extremum}). Extremising \wrt $\eq_i$:
-%%
-%\begin{equation}\label{eq:multistateext}
-%%  \pdiff{\CL}{\eq_i} = \brk{\prn{i-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \w_i
-%%    - \prn{i-\frac{n+1}{2}}\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
-%  \pdiff{\CL}{\eq_i} = \brk{i - \sum_j j\eq_j} \w_i
-%    - i\prn{\eq_+-\eq_-} - \lambda + \mu_i = 0.
-%\end{equation}
-%%
-%Consider an $i$ such that $\w_i=\w_{i+1}$:
-%%
-%\begin{equation}\label{eq:multistatemudiff}
-%  \mu_{i+1} - \mu_i = \prn{\eq_+-\eq_-} - \w_i.
-%\end{equation}
-%%
-%As $\abs{\eq_+-\eq_-}<1$ (unless we want $A=0$, \cref{eq:allpplusorminus}), this means the $\mu_i$ increase amongst the negative $\w_i$ and decrease amongst the positive $\w_i$.
-%Therefore the only non-zero equilibrium probabilities at optimum are $\eq_1$ and $\eq_n$.
-
-In order to obtain an upper bound on the area under the memory curve of any model, we now maximise the area of the serial model with respect to its equilibrium distribution.
-First let us maximise \cref{eq:multistatearea} at fixed $\eq_\pm = \sum_k \eq_k \prn{\frac{1\pm\w_k}{2}}$.
-Clearly this will happen when we put all of the probability at the ends: $\eq_1=\eq_-$ and $\eq_n=\eq_+$ are the only non-zero $\eq_k$.
-This gives an area of
-%\footnote{Note that including the dropped factors in \cref{eq:noisepm} would only change this to $A \leq \sqrt{N(4\eq_+\eq_-)}(n-1)/r$, which would have no effect on what follows.}
-
-%
-\begin{equation}\label{eq:multistateextarea}
-  A \leq \frac{\sqrt{N}}{r}(M-1)\prn{4\eq_+\eq_-}.
-\end{equation}
-%
-This is maximised at $\eq_+=\eq_-=\half$:
-%
-\begin{equation}\label{eq:maxarea}
-  A \leq \frac{\sqrt{N}}{r}(M-1).
-\end{equation}
-%
-This yields the area bound of equation (16) of the main text.
-
-Note that the chain that achieves this is not ergodic, the two states at each end are absorbing.
-This is similar to the results found numerically in \cite{Barrett2008discrete} in a slightly different situation.
-
-\end{strip}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \pnasbreak splits and balances the columns before the references.
-% If you see unexpected formatting errors, try commenting out this line
-% as it can run into problems with floats and footnotes on the final page.
-%\pnasbreak
-
-%\bibliographystyle{utcaps_sl}
-%\bibliography{maths,neuro,markov}
-
-\end{document}

### 2018-04-01 Paper/synapse_title.tex
reference formatting
removed spurious %

@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\iftoggle{main} {\author[1]{Subhaneil Lahiri}
-\author{Surya Ganguli}}
+\iftoggle{main} {\author[a,1]{Subhaneil Lahiri}
+\author[a]{Surya Ganguli}}
 {\author{Lahiri \textit{et al.}}}
@@
 \iftoggle{main}{%
-\affil{Department of Applied Physics, Stanford University, Stanford CA}
+\affil[a]{Department of Applied Physics, Stanford University, Stanford CA}
 % Please add here a significance statement to explain the relevance of your work
@@
 
-  Overall, we uncover general design principles governing the functional organization of complex molecular networks, and suggest new experimental observables in synaptic physiology, based on first %passage time theory, that connect molecular complexity to memory.
+  Overall, we uncover general design principles governing the functional organization of complex molecular networks, and suggest new experimental observables in synaptic physiology, based on first passage time theory, that connect molecular complexity to memory.
 \end{abstract}

### 2018-04-24 Paper/synapse_main.tex
simpler preamble

@@
  \begin{center}
- \begin{myenuma}
-  \panel{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
-  \panel{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
-  \panel{\includegraphics[width=0.3\linewidth]{cascms.eps}}\label{fig:cascms}
- \end{myenuma}
+ %\begin{myenuma}
+  \panel{\includegraphics[width=0.25\linewidth]{cascade.svg}\label{fig:cascade}}
+  \panel{\includegraphics[width=0.25\linewidth]{multistate.svg}\label{fig:multistate}}
+  \panel{\includegraphics[width=0.3\linewidth]{cascms.eps}\label{fig:cascms}}
+ %\end{myenuma}
  \end{center}
   \caption[Models of complex synapses]{Models of complex synapses.
-  (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
-  (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
-  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined \hyperref[sec:setup]{below}) as subsequent memories are stored.
+  \panelcap{fig:cascade} The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
+  \panelcap{fig:multistate} The serial model of \cite{Leibold2008serial}.
+  \panelcap{fig:cascms} The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined \hyperref[sec:setup]{below}) as subsequent memories are stored.
   \label{fig:models}}
@@
 To maximize this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
-An example of such a model is shown in \cref{fig:max_init}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
+An example of such a model is shown in \cref{fig:max_init_pot,fig:max_init_dep}.
 These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
-In terms of their memory curves, they are completely equivalent to a two state model with transition probabilities equal to 1, as shown in \cref{fig:max_init}(\ref{fig:binary_det}).
+In terms of their memory curves, they are completely equivalent to a two state model with transition probabilities equal to 1, as shown in \cref{fig:binary_det}.
 
@@
  \begin{center}
- \begin{myenuma}
-  \panel{\hp\includegraphics[width=0.24\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
-  \panel{\includegraphics[width=0.24\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
-  \panel{\hp\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
-  \end{myenuma}
+ %%\begin{myenuma}
+  \panel{\hp\includegraphics[width=0.24\linewidth]{max_init_pot.svg}\label{fig:max_init_pot}}\hp
+  \panel{\includegraphics[width=0.24\linewidth]{max_init_dep.svg}\label{fig:max_init_dep}}\hp
+  \panel{\hp\includegraphics[width=0.1\linewidth]{binary_det.svg}\label{fig:binary_det}}
+  %\end{myenuma}
  \end{center}
   \caption[Synaptic models that maximize initial SNR]{Synaptic models that maximize initial SNR.
-  (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
-  (\ref{fig:max_init_pot}) Depression is similar to potentiation, but with strong and weak interchanged.
-  (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
+  \panelcap{fig:max_init_pot} For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
+  \panelcap{fig:max_init_dep} Depression is similar to potentiation, but with strong and weak interchanged.
+  \panelcap{fig:binary_det} The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
   \label{fig:max_init}}
@@
 Some of these perturbations are shown in \cref{fig:perts}, see \cref{sec:kuhntucker} of the \supp\ for details.
-For example, in \cref{fig:perts}(\ref{fig:tri_right}), for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
+For example, in \cref{fig:tri_right}, for two states $i$ on the left and $j$ on the right, with $j$ being more ``potentiated" than $i$ (i.e.
 $\smash{\eta^+_i > \eta^+_j}$), we have proven that increasing $\smash{\M\pot_{ij}}$ and decreasing $\smash{\M\dep_{ij}}$ leads to an increase in area.
@@
 This determines the topology (non-zero transition probabilities) of the model with maximal area.
-It is of the form shown in \cref{fig:env}(\ref{fig:sticky}),%\cref{fig:max_area},
+It is of the form shown in \cref{fig:sticky}, %\cref{fig:max_area},
 with potentiation moving one step to the right and depression moving one step to the left.
@@
  \begin{center}
- \begin{myenuma}
-  \panel{\includegraphics[width=0.14\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
-  \panel{\includegraphics[width=0.14\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
-  \panel{\includegraphics[width=0.28\linewidth]{shortcut.svg}}\label{fig:shortcut}
- \end{myenuma}
+ %%\begin{myenuma}
+  \panel{\includegraphics[width=0.14\linewidth]{triangular_right.svg}\label{fig:tri_right}}\hp
+  \panel{\includegraphics[width=0.14\linewidth]{triangular_left.svg}\label{fig:tri_left}}\hp
+  \panel{\includegraphics[width=0.28\linewidth]{shortcut.svg}\label{fig:shortcut}}
+ %\end{myenuma}
  \end{center}
   \caption[Perturbations that increase the area]{Perturbations that increase the area.
-  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$.
+  \panelcap{fig:tri_right} Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$.
 It can no longer be used when $\M\dep$ is lower triangular, \ie depression must move synapses to ``more depressed'' states.
-  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$.
+  \panelcap{fig:tri_left} Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$.
 It can no longer be used when $\M\pot$ is upper triangular, \ie potentiation must move synapses to ``more potentiated'' states.
-  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions.
+  \panelcap{fig:shortcut} Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions.
 It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
@@
 We can then maximize this by pushing all of the equilibrium distribution symmetrically to the two end states.
-This can be done by reducing the transition probabilities out of these states, as in \cref{fig:env}(\ref{fig:sticky}). %\cref{fig:max_area}.
+This can be done by reducing the transition probabilities out of these states, as in \cref{fig:sticky}. %\cref{fig:max_area}.
 This makes it very difficult to exit these states once they have been entered.
@@
 At late times, only the area bound is saturated.
-The function $\snr(t_0)$, the green curve in \cref{fig:env}(\ref{fig:envplot}), above forms a memory curve envelope with late-time power-law decay $\sim \smash{t_0\inv}$.
+The function $\snr(t_0)$, the green curve in \cref{fig:envplot}, above forms a memory curve envelope with late-time power-law decay $\sim \smash{t_0\inv}$.
 No synaptic model can have an SNR that is greater than this at any time.
@@
 At any time, can we find a model whose memory curve touches the envelope?
-The red curves in \cref{fig:env}(\ref{fig:envplot}) show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
+The red curves in \cref{fig:envplot} show the closest we have come to the envelope with actual models, by repeated numerical optimization of $\SNR(t_0)$ over $\M\potdep$ with random initialization and by hand designed models.
 
@@
 We believe there is some other constraint that prevents models from approaching the envelope, and currently are exploring several mathematical conjectures for the precise form of this constraint in order to obtain a potentially tighter envelope.
-Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \cref{fig:env}(\ref{fig:sticky})%\cref{fig:max_area}
+Nevertheless, we have proven rigorously that no model's memory curve can ever exceed this envelope, and that it is at least tight for late times, longer than $\CO(\frac{M}{r})$, where models of the form in \cref{fig:sticky} %\cref{fig:max_area}
  can come close to the envelope.
@@
  \parbox[t]{0.43\linewidth}{%
- \begin{myenuma}
-   \panel{\includegraphics[width=0.85\linewidth]{env.svg}}\label{fig:envplot}
- \end{myenuma}}\hp
+ %%\begin{myenuma}
+   \panel{\includegraphics[width=0.85\linewidth]{env.svg}\label{fig:envplot}}
+ %\end{myenuma}
+ }\hp
  \parbox[t]{0.44\linewidth}{%
- \begin{myenuma}
+ %%\begin{myenuma}
   \stepcounter{myenumai}
 % \addtocounter{enumi}{1}
-   \panel{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}\\[0.1\linewidth]
-   \panel{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
- \end{myenuma}}
+   \panel{\includegraphics[width=0.8\linewidth]{diffjump.svg}\label{fig:diffjump}}\\[0.1\linewidth]
+   \panel{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}\label{fig:sticky}}
+ %\end{myenuma}
+ }
  \end{center}
@@
   The solid red curve shows the highest SNR we have found with hand designed models.
-  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states, and all transition probabilities equal to $1$.
-  At late times they are of the form shown in (\ref{fig:sticky}) with different values of $\varepsilon$.
-  The model shown in (\ref{fig:sticky}) also saturates the area bound \cref{eq:max_area} in the limit $\varepsilon\to0$.
+  At early times these models are of the form shown in \panelcap{fig:diffjump} with different numbers of states, and all transition probabilities equal to $1$.
+  At late times they are of the form shown in \panelcap{fig:sticky} with different values of $\varepsilon$.
+  The model shown in \panelcap{fig:sticky} also saturates the area bound \cref{eq:max_area} in the limit $\varepsilon\to0$.
   \label{fig:env}}

### 2019-03-30 Notes/laplace.tex
typo
removed some slides

@@
 %
-If the probability of a given plasticity event being poetitiating/depressing os $f\potdep$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
+If the probability of a given plasticity event being potentiating/depressing is $f\potdep$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
 
@@
 \begin{equation}\label{eq:SNRcurveExact}
-  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{(r\frg+\hom)t}\w}
+  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt\frg}\w}
                  {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
@@
 %
-We can absorb $\hom$ into $\MM\potdep$ by defining
-%
-\begin{equation}\label{eq:absorbhom}
-\begin{aligned}
-  \Mh\potdep &= \M\potdep + \frac{\hom}{2rf\potdep},
-  &\qquad
-  \frgh &= f\pot\Mh\pot + f\dep\Mh\dep - \I,
-  \\&&
-  \ench &= f\pot(\Mh\pot-\I)-f\dep(\Mh\dep-\I).
-\end{aligned}
-\end{equation}
-%
-Then the SNR curve is
-%
-\begin{equation}\label{eq:SNRcurvehom}
-  \SNR(t) = \frac{\sqrt{N}  \eq \ench \,\e^{rt\frgh}\w}
-                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
-\end{equation}
-%
+%We can absorb $\hom$ into $\M\potdep$ by defining
+%%
+%\begin{equation}\label{eq:absorbhom}
+%\begin{aligned}
+%  \Mh\potdep &= \M\potdep + \frac{\hom}{2rf\potdep},
+%  &\qquad
+%  \frgh &= f\pot\Mh\pot + f\dep\Mh\dep - \I,
+%  \\&&
+%  \ench &= f\pot(\Mh\pot-\I)-f\dep(\Mh\dep-\I).
+%\end{aligned}
+%\end{equation}
+%%
+%Then the SNR curve is
+%%
+%\begin{equation}\label{eq:SNRcurvehom}
+%  \SNR(t) = \frac{\sqrt{N}  \eq \enc \,\e^{rt\frg}\w}
+%                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
+%\end{equation}
+%%
 The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
@@
 
-We can then treat the off-diagonal elements of $\Mh\potdep$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
+We can then treat the off-diagonal elements of $\M\potdep$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
 The diagonal elements are determined by setting the row sums to 1.
@@
 
-Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurvehom} (ignoring the denominator)
+Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurveExact} (ignoring the denominator)
 %
@@
   \begin{aligned}
-    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
-    a_i(s) &= \sum_j \eq_i \enc_{ij} (\etw_{i}(s) - \etw_{j}(s)),\\
+    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)}, \\
+    a_j(s) &= \sum_{jk} \eq_j \enc_{jk} (\etw_{j}(s) - \etw_{k}(s)), \\
+    \thb_i(s) &= \sum_j \fptb_{ij}(0)\ a_j(s), \\
+  \end{aligned}
     \implies
-    \hat{A}(s) &= \sum_k c_k(s) \eq_k \w_k
+    \hat{A}(s) = \sum_k c_k(s) \eq_k \w_k
       = \sum_i a_i(s).
-  \end{aligned}
 \end{equation}
@@
 
-Using the same approach as the area bound in \cite{Lahiri2013synapse}, and defining $\thb_i(s) = \sum_j \fptb_{ij}(0) a_j(s)$,  we find that the derivatives are
+Using the same approach as the area bound in \cite{Lahiri2013synapse} we find that the derivatives are
 %
@@
 To write down the Hessian, we will first define some notation.
-Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$
+We define the matrix $\mathbf{X}(s) = \fund(0) \enc \fund(s)$.
+Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$.
 Then we denote
@@
     &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
-    &&&+ \eq_i \brc{[\fund(0)\enc\fund(s)]_{ik} - [\fund(0)\enc\fund(s)]_{jk}} \brc{\etw_k(s) - \etw_l(s)} \\
+    &&&+ \eq_i \brc{\mathbf{X}_{ik}(s) - \mathbf{X}_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)} \\
     &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)},  \\
@@
 %
-Then we can write the hessian as
+Then we can write the Hessian as
 %
@@
   \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
-     (f\pot)^2 (R_{ijkl} + R\trans_{ijkl})
-     + f\pot(S_{ijkl} + S\trans_{ijkl}), \\
+     (f\pot)^2   [R + R\trans + S + S\trans]_{ijkl}, \\
   \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
-     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
-     - f\pot S_{ijkl} + f\dep S\trans_{ijkl}, \\
+     f\pot f\dep [R + R\trans - S + S\trans]_{ijkl}, \\
   \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
-     f\pot f\dep (R_{ijkl} + R\trans_{ijkl})
-     - f\pot S\trans_{ijkl} + f\dep S_{ijkl}, \\
+     f\pot f\dep [R + R\trans + S - S\trans]_{ijkl}, \\
   \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\dep_{kl}} &=
-     (f\dep)^2 (R_{ijkl} + R\trans_{ijkl})
-     - f\dep(S_{ijkl} + S\trans_{ijkl}), \\
+     (f\dep)^2   [R + R\trans - S - S\trans]_{ijkl}, \\
 \end{aligned}

### 2020-06-18 Notes/laplace.tex
details re shortening chain

@@
     c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)}, \\
-    a_j(s) &= \sum_{jk} \eq_j \enc_{jk} (\etw_{j}(s) - \etw_{k}(s)), \\
-    \thb_i(s) &= \sum_j \fptb_{ij}(0)\ a_j(s), \\
+    a_j(s) &= \sum_{k} \enc_{jk} (\etw_{j}(s) - \etw_{k}(s)), \\
+    \thb_i(s) &= \sum_j \fptb_{ij}(0)\, \eq_j a_j(s), \\
   \end{aligned}
@@
     \hat{A}(s) = \sum_k c_k(s) \eq_k \w_k
-      = \sum_i a_i(s).
+      = \sum_i  \eq_i a_i(s).
 \end{equation}
@@
 \begin{equation}\label{eq:hessparts}
-\begin{aligned}
-  R_{ijkl} &=
+\begin{alignedat}{3}
+    R_{ijkl} &= \
     &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
-    &&&+ \eq_i \brc{\mathbf{X}_{ik}(s) - \mathbf{X}_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)} \\
-    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} \brc{\etw_k(s) - \etw_l(s)},  \\
+    &&&+ \eq_i \brc{\mathbf{X}_{ik}(s) - \mathbf{X}_{jk}(s)} 
+          \brc{\etw_k(s) - \etw_l(s)} \\
+    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} 
+          \brc{\etw_k(s) - \etw_l(s)},  \\
   S_{ijkl} &=
@@
     &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
-\end{aligned}
+\end{alignedat}
 \end{equation}
@@
 \begin{multline}\label{eq:unienv}
+  \alpha = \exp\prn{\pm\frac{2y_*}{M_*}},
+  \quad
   s = 2\sinh^2\frac{y_*}{M_*},
   \qquad
-  \hat{A}_*(s) = \prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\log\alpha(s)}{s},
+  \hat{A}_*(s) = \prn{\frac{ 2\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\abs{\log\alpha(s)}}{s},
   \\ \text{where}\quad
@@
 %
-When $s \sim \CO(M^{-2}) \ll 1$, we have $\alpha \approx 1+\sqrt{2s}$ and the heuristic envelope is approximately
+When $s \sim \CO(M^{-2}) \ll 1$, we have $\alpha \approx 1 \pm \sqrt{2s}$ and the heuristic envelope is approximately
 %
 \begin{equation}\label{eq:unienvapprox}
-  \hat{A}_*(s) \approx \frac{ 1.08406 }{ \sqrt{s} }.
+  \hat{A}_*(s) \approx \frac{ 0.54203 }{ \sqrt{s} }.
 \end{equation}
@@
 
-The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
-The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
+The following elements of $\enc$ and $\eq$ differ from \eqref{eq:unienc}:
 %
@@
   \begin{aligned}
-  \enc_{M-1M} = -\enc_{21} &= \frac{\varepsilon}{2},
+  \enc_{M-1M} = -\enc_{21} &= \frac{(1-\varepsilon)}{2},
   &\qquad
-  \eq_1 = \eq_M &= \frac{\varepsilon}{M-2(1-\varepsilon)},
+  \eq_1 = \eq_M &= \frac{(1-\varepsilon)}{M-2\varepsilon},
   \\ &&
-  \eq_i &=  \frac{1}{M-2(1-\varepsilon)}
+  \eq_i &=  \frac{1}{M-2\varepsilon}
   \quad\text{for}\quad 1<i<M.
@@
 \begin{equation}\label{eq:shortenareaeta}
-  \hat{A}(s) = \frac{\varepsilon(\etw_1(s)-\etw_M(s)) + (1-\varepsilon)(\etw_2(s)-\etw_{M-2}(s))}{M-2(1-\varepsilon)}
-       = \frac{2\varepsilon\, \etw_1(s) + 2(1-\varepsilon)\etw_2(s)}{M-2(1-\varepsilon)}.
+  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_M(s)) + \varepsilon(\etw_2(s)-\etw_{M-1}(s))}{M-2\varepsilon}
+       = \frac{2(1-\varepsilon) \etw_1(s) + 2\varepsilon\,\etw_2(s)}{M-2\varepsilon}.
 \end{equation}
@@
   \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_3 + \varepsilon\,\delta\eta_1}{2} - \prn{s+\frac{1+\varepsilon}{2}}\delta\eta_2 &= 0, \\
+  \frac{\delta\eta_3 + (1-\varepsilon)\delta\eta_1}{2} - \prn{s+\frac{2-\varepsilon}{2}}\delta\eta_2 &= 0, \\
   \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
@@
   \delta\eta_1 &= \frac{2\alpha^{\frac{M}{2}}}
-         {[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
-  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + \varepsilon(\alpha-1))}
-         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
-  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - \varepsilon\alpha(\alpha-1))}
-         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
+         {[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
+  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + (1-\varepsilon)(\alpha-1))}
+         {(1+\alpha)[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
+  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - (1-\varepsilon)\alpha(\alpha-1))}
+         {(1+\alpha)[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
 \end{aligned}
@@
 \begin{equation}\label{eq:shortenlaplace}
-  \hat{A}(s) = \frac{2 [\varepsilon(\alpha^{\frac{M}{2}}-1)^2
-          + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
+  \hat{A}(s) = \frac{2 [(1-\varepsilon)(\alpha^{\frac{M}{2}}-1)^2
+          + \varepsilon(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
           { [M-2(1-\varepsilon)]
-            [\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
-\end{equation}
-%
-This interpolates between \eqref{eq:unilaplace} at $\varepsilon=1$ and the same expression for $M-2$ at $\varepsilon=0$.
+            [(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
+\end{equation}
+%
+This interpolates between \eqref{eq:unilaplace} at $\varepsilon=0$ and the same expression for $M-2$ at $\varepsilon=1$.
+This is easier to see if we define
+%
+\begin{equation}\label{eq:shortdefs}
+\begin{aligned}
+  F_M(\alpha) &= (\alpha^{\frac{M}{2}} - 1)^2, &
+  \Delta F_M(\alpha) &= F_M(\alpha) - \zeta F_{M-2}(\alpha), &
+  f_m(\alpha) = \frac{F_M(\alpha)}{\Delta F_M(\alpha)}, \\
+  G_M(\alpha) &= \alpha^{M} + 1, &
+  \Delta G_M(\alpha) &= G_M(\alpha) - \zeta G_{M-2}(\alpha), &
+  g_m(\alpha) = \frac{G_M(\alpha)}{\Delta G_M(\alpha)}, \\
+  \text{where}\quad \zeta &= \alpha^2 - \alpha + 1, &
+  m &= \frac{M}{2}. \\
+\end{aligned}
+\end{equation}
+%
+Then the uniform model \eqref{eq:unilaplace} has \( \hat{A}(s) = \frac{F_m(\alpha)}{s m G_m(\alpha)} \).
+Then \cref{eq:shortenlaplace} can be be written as
+%
+\begin{equation}\label{eq:shortsimple}
+  \hat{A}(s) =
+    \frac{(1-\varepsilon)F_m(\alpha) + \varepsilon \zeta F_{m-1}(\alpha)}
+         {s (m - \varepsilon) ((1-\varepsilon)G_m(\alpha) + \varepsilon \zeta G_{m-1}(\alpha))}
+    % = \frac{F_m - \varepsilon \Delta F_{m}}
+    %      {s (m - \varepsilon) (G_m - \varepsilon \Delta G_{m})}
+    = \brk{\frac{\Delta F_{m}}{s \Delta G_{m}}}
+    \frac{f_m - \varepsilon}{(m - \varepsilon) (g_m - \varepsilon)}.
+\end{equation}
+%
+
+\begin{figure}[ht]
+\begin{center}
+\begin{myenuma}
+  \item\label{fig:short_opt}\aligntop{%
+    \includegraphics[width=0.45\linewidth]{shorten_opt.pdf}}
+  \item\label{fig:short_ratio}\aligntop{%
+    \includegraphics[width=0.45\linewidth]{shorten_ratio.pdf}}
+\end{myenuma}
+\caption{Signal-to-noise ratio for optimal shortened models, \cref{eq:shortenenc} optimised over \(\varepsilon\).
+(\ref{fig:short_opt}): Solid grey line is \cref{eq:unienv}, the uniform model with \(M\) treated as a continuous parameter and optimised.
+Dotted grey line is the approximate expression from \cref{eq:unienvapprox}.
+The other curves are from \cref{eq:shortopt}, optimised shortened models for various values of \(M\).
+(\ref{fig:short_ratio}): The ratios of the quantities plotted in (\ref{fig:short_opt}) to \cref{eq:unienv}.
+The dotted orange line is the extension of the \(M=4\) line with \(\varepsilon\) held at 0, as discussed below \cref{eq:shortenvalid}.
+\label{fig:shorten}}
+\end{center}
+\end{figure}
 
 We can maximise this \wrt $\varepsilon$.
-This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
+%
+\begin{equation}\label{eq:shortopt}
+  (\varepsilon - f_m)^2 = (m - f_m)(g_m - f_m),
+  \qquad
+  \hat{A}(s) = \brk{\frac{\Delta F_{m}}{s \Delta G_{m}}}
+    \frac{1}{\brk{\sqrt{\abs{m - f_m}} - \sqrt{\abs{g_m - f_m}}}^2}.
+\end{equation}
+%
 We can find the boundary of the region of applicability by asking at what value of $s$ is the optimal value of $\varepsilon=0$ or $1$?
@@
 \begin{aligned}
+  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
+    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^{M-1}-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
+      = 0,\\
   \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
-    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
+    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^{M-1}-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
       = 0,\\
-  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
-    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
-      = 0,\\
+    \text{where}\quad
     s = &\, \frac{(\alpha-1)^2}{2\alpha},
@@
 %
-with the $\varepsilon=1$ case providing the lower boundary and $\varepsilon=0$ providing the upper boundary.
+with the $\varepsilon=0$ case providing the lower boundary and $\varepsilon=1$ providing the upper boundary.
+For lower \(s\), \(\varepsilon = 0\) is the optimum. 
+For higher \(s\), \(\varepsilon = 1\) is the optimum, which is eqivalent to \(\varepsilon = 0\) with \(M-2\) states.
+This is depicted by the dotted orange line in \cref{fig:shorten}\ref{fig:short_ratio}.
+
+\(M=2\) is a special case:
+%
+\begin{equation*}
+  \hat{A}(s) = \frac{(\alpha-1)^2}
+                    {[2(\alpha^2-\alpha+1) - (1-\varepsilon) (\alpha-1)^2]s},
+\end{equation*}
+%
+so \(\varepsilon=0\) is always the maximum.
+
 If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
-In this case, however, there is an overlap.
+In this case, however, there is an overlap as seen in \cref{fig:shorten}\ref{fig:short_ratio}.
+The \(M=2\) region of validity overlaps everything else.
+The other overlaps are less extreme, but they do exist.
 This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
@@
   & r\tau \leq 0.73 , \\
-  & \frac{ 1.1\sqrt{N}}{ \sqrt{r\tau} } , &
+  & \frac{ 0.54\sqrt{N}}{ \sqrt{r\tau} } , &
   0.73 \leq & r\tau \leq 0.22 M^2 , \\

### 2020-06-18 Code/Python/complex_synapse/synapse_base.py

@@ -10,7 +10,7 @@ from numbers import Number
 import numpy as np
 import numpy_linalg as la
 from . import builders as bld
-cvl = la.convert_loop
+cvl = la.convert
 
 # types that can multiply with/add to a matrix
 ArrayLike = Union[Number, Sequence[Number], np.ndarray]
@@ -71,17 +71,16 @@ class SynapseBase(np.lib.mixins.NDArrayOperatorsMixin):
         self.fix()
 
     def __array_ufunc__(self, ufunc, method, *inputs, **kwargs):
-        """Handling ufunce with SynapseBases
+        """Handling ufuncs with SynapseBases
         """
-        args, _ = cvl.conv_loop_in_attr('plast', SynapseBase, inputs)
+        args, _ = cvl.conv_in_attr('plast', SynapseBase, inputs)
 
         conv = [True] + [False] * (ufunc.nout-1)
-        outputs, conv = cvl.conv_loop_in_attr(
-            'plast', SynapseBase, kwargs, conv)
+        outputs, conv = cvl.conv_in_attr('plast', SynapseBase, kwargs, conv)
 
         results = self.plast.__array_ufunc__(ufunc, method, *args, **kwargs)
 
-        return cvl.conv_loop_out_attr(self, 'plast', results, outputs, conv)
+        return cvl.conv_out_attr(self, 'plast', results, outputs, conv)
 
     # -------------------------------------------------------------------------
     # %%* Housekeeping

### 2020-06-18 Code/Python/complex_synapse/synapse_memory_model.py

@@ -13,6 +13,8 @@ from sl_py_tools.numpy_tricks import markov as ma
 from .synapse_base import SynapseBase, ArrayLike
 from .builders import scalarise
 
+wrap = la.wrappers.Wrappers(la.lnarray)
+
 
 class SynapseMemoryModel(SynapseBase):
     """Class for memory curves of complex synapse models.
@@ -358,7 +360,7 @@ class SynapseMemoryModel(SynapseBase):
         inita : la.lnarray
             Initial SNR from eigenmodes.
         """
-        eig = la.wrappers.wrap_several(np.linalg.eig)
+        eig = wrap.several(np.linalg.eig)
         qas, evs = eig(-self.markov())
         mask = la.ones_like(qas, dtype=bool).put(qas.real.argmin(), False)
         taua = 1. / qas[mask]

### 2020-06-18 Code/Python/complex_synapse/synapse_opt.py

@@ -14,7 +14,8 @@ from .builders import scalarise
 from .synapse_memory_model import SynapseMemoryModel as _SynapseMemoryModel
 from .synapse_base import SynapseBase as _SynapseBase
 
-eig = la.wrappers.wrap_several(np.linalg.eig)
+wrap = la.wrappers.Wrappers(la.lnarray)
+eig = wrap.several(np.linalg.eig)
 
 
 class SynapseOptModel(_SynapseMemoryModel):
@@ -111,7 +112,7 @@ class SynapseOptModel(_SynapseMemoryModel):
 
         fundi = self.zinv()
         fundis = self.zinv(rate)
- 
+
         peq = self.peq()
         c_s = peq @ self.enc() @ fundis.inv
 
@@ -195,7 +196,7 @@ class SynapseOptModel(_SynapseMemoryModel):
         """
         # (p,c), (eta,theta)
         rows, cols, mats = self._derivs(rate, inv)
-        rcond = 1. / np.linalg.cond(mats[:2]).max()       
+        rcond = 1. / np.linalg.cond(mats[:2]).max()
         if rcond < self.RCondThresh:
             func = np.array(np.inf)
         else:
@@ -441,6 +442,7 @@ def _dbl_diagsub(tens: la.lnarray) -> la.lnarray:
     """
     tens_trn = _diagsub(_trnsp4(_diagsub(tens)))
     return la.stack((_trnsp4(tens_trn), tens_trn))
+
 
 if __name__ == "__main__":
     pass

### 2020-06-20 Notes/laplace.tex
changed notation sor uniform serial models

@@
   \begin{myenuma}
-    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_n.svg}}\hspace{0.02\linewidth}\label{fig:serial_uniform}
+    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_uni.svg}}\hspace{0.02\linewidth}\label{fig:serial_uniform}
     \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_sticky.svg}}\hspace{0.02\linewidth}\label{fig:serial_sticky}
@@
 We will work in units such that $r=1$ temporarily, only restoring $r$ at the end.
+We study models with \(m\) strong and \(m\) weak states for a total of  \(M = 2 m\).
 
@@
 \begin{equation}\label{eq:unienc}
-  \enc_{ii+1}=\frac{1}{2} \quad\text{for}\quad i<M,
+  \enc_{ii+1}=\frac{1}{2} \quad\text{for}\quad i<2m,
   \qquad
@@
 %
-and $\eq_i = \frac{1}{M}$, we have
+and $\eq_i = \frac{1}{2m}$, we have
 %
 \begin{equation}\label{eq:uniareaeta}
-  \hat{A}(s) = \frac{\etw_1(s) - \etw_M(s)}{M}.
+  \hat{A}(s) = \frac{\etw_1(s) - \etw_{2m}(s)}{2m}.
 \end{equation}
@@
   \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{\frac{M}{2}+1}+\delta\eta_{\frac{M}{2}-1}}{2} - (s+1)\delta\eta_{\frac{M}{2}} &= \frac{1}{s},\\
-  \frac{\delta\eta_{\frac{M}{2}+2}+\delta\eta_{\frac{M}{2}}}{2} - (s+1)\delta\eta_{\frac{M}{2}+1} &= -\frac{1}{s},\\
-  \frac{\delta\eta_{M-1}}{2} - \prn{s+\frac{1}{2}}\delta\eta_M &= 0.
-\end{aligned}
-\end{equation}
-%
-These equations are invariant under $\delta\eta_{M+1-i} \to -\delta\eta_i$, so these two must be equal.
+  \frac{\delta\eta_{m+1}+\delta\eta_{m-1}}{2} - (s+1)\delta\eta_{m} &= \frac{1}{s},\\
+  \frac{\delta\eta_{m+2}+\delta\eta_{m}}{2} - (s+1)\delta\eta_{m+1} &= -\frac{1}{s},\\
+  \frac{\delta\eta_{2m-1}}{2} - \prn{s+\frac{1}{2}}\delta\eta_{2m} &= 0.
+\end{aligned}
+\end{equation}
+%
+These equations are invariant under $\delta\eta_{2m+1-i} \to -\delta\eta_i$, so these two must be equal.
 Then we have
@@
   \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i &= 0,
-  \quad\text{for}\quad 1<i<\frac{M}{2}\\
+  \quad\text{for}\quad 1<i<m, \\
   \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
+  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
 \end{aligned}
@@
 \begin{equation}\label{eq:detagensol}
-  \delta\eta_i = B\alpha^{i-1} + C\alpha^{1-i},
+  \delta\eta_i = B\e^{\beta(i-1)} + C\e^{\beta(1-i)},
   \qquad\text{where}\qquad
-  \alpha^2 - 2(s+1)\alpha + 1 = 0,
+  s = S(\beta) \equiv 2 \sinh^2 \prn{\frac{\beta}{2}},
 \end{equation}
@@
 \begin{equation}\label{eq:unicoeffs}
-  B = -\frac{2\alpha}{(1+\alpha)(\alpha^{\frac{M}{2}}+\alpha^{-\frac{M}{2}})s},
-  \qquad
-  C = -\frac{2}{(1+\alpha)(\alpha^{\frac{M}{2}}+\alpha^{-\frac{M}{2}})s}.
+  B = -\frac{1}{s(\e^{-\beta} + 1) \cosh \prn{m \beta}},
+  \qquad
+  C = -\frac{1}{s(\e^{\beta} + 1) \cosh \prn{m \beta}}.
 \end{equation}
@@
 \begin{equation}\label{eq:unilaplace}
-  A(s) = \frac{2\etw_1(s)}{M}
-   = \frac{2}{M}\prn{\frac{1}{s}+B+C}
-   = \frac{2(\alpha^{\frac{M}{2}}-1)^2}{Ms(\alpha^M+1)}.
+  A(s) = \frac{2\etw_1(s)}{2m}
+   = \frac{1}{m}\prn{\frac{1}{s}+B+C}
+   = \frac{2\sinh^2\prn{\frac{m \beta}{2}}}{m s \cosh\prn{m \beta}}
+   = \frac{2 S\prn{m \beta}}{M s \prn{S\prn{m \beta} + 1}}.
 \end{equation}
@@
 \begin{multline}\label{eq:unienv}
-  \alpha = \exp\prn{\pm\frac{2y_*}{M_*}},
+  M_* = \pm\frac{2y_*}{\beta},
   \quad
@@
   \qquad
-  \hat{A}_*(s) = \prn{\frac{ 2\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\abs{\log\alpha(s)}}{s},
+  \hat{A}_*(s) = \prn{\frac{ 2\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\abs{\beta}}{s},
   \\ \text{where}\quad
@@
 %
-When $s \sim \CO(M^{-2}) \ll 1$, we have $\alpha \approx 1 \pm \sqrt{2s}$ and the heuristic envelope is approximately
+When $s \sim \CO(M^{-2}) \ll 1$, we have \(\beta \approx \pm \sqrt{2s}\) and the heuristic envelope is approximately
 %
@@
   \begin{aligned}
-  \enc_{12} = -\enc_{MM-1} &= \frac{\varepsilon}{2},
-  &\qquad
-  \eq_1 = \eq_M &= \frac{1}{2+(M-2)\varepsilon},
+  \enc_{1,2} = -\enc_{2m,2m-1} &= \frac{1-\varepsilon}{2},
+  &\quad
+  \eq_1 = \eq_{2m} &= \frac{1}{2(m-(m-1)\varepsilon)},
   \\ &&
-  \eq_i &=  \frac{\varepsilon}{2+(M-2)\varepsilon}
-  \quad\text{for}\quad 1<i<M.
+  \eq_i &=  \frac{1-\varepsilon}{2(m-(m-1)\varepsilon)}
+  \quad\text{for}\quad 1<i<2m.
   \end{aligned}
@@
 \begin{equation}\label{eq:stickyareaeta}
-  \hat{A}(s) = \frac{\varepsilon(\etw_1(s)-\etw_M(s))}{2+(M-2)\varepsilon}
-       = \frac{2\varepsilon\, \etw_1(s)}{2+(M-2)\varepsilon}.
+  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_M(s))}
+                    {2(m-(m-1)\varepsilon)}
+       = \frac{2(1-\varepsilon)\, \etw_1(s)}{2(m-(m-1)\varepsilon)}.
 \end{equation}
@@
 \begin{aligned}
-  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{\varepsilon}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
+  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{1-\varepsilon}{2}}\delta\eta_1 &= 0, \\
+  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
 \end{aligned}
@@
 \begin{aligned}
-  B &= -\frac{ 2 \alpha^{\frac{M}{2}} (\alpha-(1-\varepsilon)) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s}, \\
-  C &= -\frac{ 2 \alpha^{\frac{M}{2}} (1-(1-\varepsilon)\alpha) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s},
+  B &= -\frac{\e^{\beta} - \varepsilon}
+      {s [\e^{\beta} + 1] [\cosh(m\beta) - \varepsilon \cosh((m-1)\beta)]}, \\
+  C &= -\frac{1 - \varepsilon\e^{\beta}}
+      {s [\e^{\beta} + 1] [\cosh(m\beta) - \varepsilon \cosh((m-1)\beta)]},
 \end{aligned}
@@
 \begin{equation}\label{eq:stickylaplace}
-  \hat{A}(s) = \frac{2\varepsilon}{[2+(M-2)\varepsilon]s} \,
-      \frac{ (\alpha^{\frac{M}{2}}-1)^2 - (1-\varepsilon)\alpha(\alpha^{\frac{M-2}{2}}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }.
+  \hat{A}(s) = \frac{(1-\varepsilon)}{[m-(m-1)\varepsilon]s} \,
+      \brk{\frac{S(m\beta) - \varepsilon S((m-1)\beta)}
+      {S(m\beta) - \varepsilon S((m-1)\beta) + 1 - \varepsilon}}.
 \end{equation}
@@
 This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
-We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=1$:
+We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0$:
 %
@@
 \begin{aligned}
-  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
-    M\alpha[(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^M+1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M-2}+1)]
-     \\ &+ 2(\alpha^{\frac{M}{2}}-1)^2(\alpha^M+1) = 0,\\
-    s = &\, \frac{(\alpha-1)^2}{2\alpha}.
+  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} &\propto
+    m \prn{S(m\beta) - S((m-1)\beta)} + (m-1) S(m\beta) \prn{S(m\beta) + 1} = 0,\\
+   &\text{where} \quad s =  S(\beta) = 2 \sinh^2(\beta/2).
 \end{aligned}
@@
 \begin{equation}\label{eq:stickyeps}
-  \varepsilon = \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},
+  \varepsilon = 1 - \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},
 \end{equation}
@@
   \begin{aligned}
-  \enc_{M-1M} = -\enc_{21} &= \frac{(1-\varepsilon)}{2},
+  \enc_{2m-1,2m} = -\enc_{2,1} &= \frac{(1-\varepsilon)}{2},
   &\qquad
-  \eq_1 = \eq_M &= \frac{(1-\varepsilon)}{M-2\varepsilon},
+  \eq_1 = \eq_M &= \frac{(1-\varepsilon)}{2(m-\varepsilon)},
   \\ &&
-  \eq_i &=  \frac{1}{M-2\varepsilon}
-  \quad\text{for}\quad 1<i<M.
+  \eq_i &=  \frac{1}{2(m-\varepsilon)}
+  \quad\text{for}\quad 1<i<M2m.
   \end{aligned}
@@
 \begin{equation}\label{eq:shortenareaeta}
-  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_M(s)) + \varepsilon(\etw_2(s)-\etw_{M-1}(s))}{M-2\varepsilon}
-       = \frac{2(1-\varepsilon) \etw_1(s) + 2\varepsilon\,\etw_2(s)}{M-2\varepsilon}.
+  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_{2m}(s)) 
+                      + \varepsilon(\etw_2(s)-\etw_{2m-1}(s))}
+                    {2(m-\varepsilon)}
+       = \frac{2(1-\varepsilon) \etw_1(s) + 2\varepsilon\,\etw_2(s)}
+              {2(m-\varepsilon)}.
 \end{equation}
@@
   \frac{\delta\eta_3 + (1-\varepsilon)\delta\eta_1}{2} - \prn{s+\frac{2-\varepsilon}{2}}\delta\eta_2 &= 0, \\
-  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
+  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
 \end{aligned}
@@
 \begin{aligned}
-  \delta\eta_i &= B\alpha^{i-2} + C\alpha^{2-i} \qquad
-        \text{for} \quad 2 \leq i \leq \frac{M}{2}, \\
-  \delta\eta_1 &= \frac{2\alpha^{\frac{M}{2}}}
-         {[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
-  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + (1-\varepsilon)(\alpha-1))}
-         {(1+\alpha)[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
-  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - (1-\varepsilon)\alpha(\alpha-1))}
-         {(1+\alpha)[(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
+  \delta\eta_i &= B\e^{(i-2)\beta} + C\e^{(2-i)\beta} \qquad
+        \text{for} \quad 2 \leq i \leq m, \\
+  \delta\eta_1 &= \frac{1}
+         {[(1-\varepsilon)\cosh(m\beta) + \varepsilon(2s+1)\cosh((m-1)\beta)]s}, \\
+  B &= - \frac{(2s+1) \e^{\beta/2} + (1-\varepsilon) \tanh(\beta/2)}
+         {s [(1-\varepsilon)\cosh(m\beta) + \varepsilon(2s+1)\cosh((m-1)\beta)]}, \\
+  C &= - \frac{((2s+1) \e^{-\beta/2} - (1-\varepsilon) \tanh(\beta/2)}
+         {s [(1-\varepsilon)\cosh(m\beta) + \varepsilon(2s+1)\cosh((m-1)\beta)]}, \\
 \end{aligned}
@@
 \begin{equation}\label{eq:shortenlaplace}
-  \hat{A}(s) = \frac{2 [(1-\varepsilon)(\alpha^{\frac{M}{2}}-1)^2
-          + \varepsilon(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
-          { [M-2(1-\varepsilon)]
-            [(1-\varepsilon)(\alpha^{M}+1) + \varepsilon(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
-\end{equation}
-%
-This interpolates between \eqref{eq:unilaplace} at $\varepsilon=0$ and the same expression for $M-2$ at $\varepsilon=1$.
-This is easier to see if we define
+  \hat{A}(s) = \frac{(1-\varepsilon) S(m\beta) + \varepsilon(2s+1) S((m-1)\beta)}
+          { s [m-\varepsilon]
+            [(1-\varepsilon)[S(m\beta) + 1] + \varepsilon(2s+1)[S((m-1)\beta) + 1]] }.
+\end{equation}
+%
+This interpolates between \eqref{eq:unilaplace} at $\varepsilon=0$ and the same expression for $m-1$ at $\varepsilon=1$.
+
+To find the optimal \(\varepsilon\) at any given \(s\), it helps if we define
 %
@@
 \begin{aligned}
-  F_M(\alpha) &= (\alpha^{\frac{M}{2}} - 1)^2, &
-  \Delta F_M(\alpha) &= F_M(\alpha) - \zeta F_{M-2}(\alpha), &
-  f_m(\alpha) = \frac{F_M(\alpha)}{\Delta F_M(\alpha)}, \\
-  G_M(\alpha) &= \alpha^{M} + 1, &
-  \Delta G_M(\alpha) &= G_M(\alpha) - \zeta G_{M-2}(\alpha), &
-  g_m(\alpha) = \frac{G_M(\alpha)}{\Delta G_M(\alpha)}, \\
-  \text{where}\quad \zeta &= \alpha^2 - \alpha + 1, &
-  m &= \frac{M}{2}. \\
-\end{aligned}
-\end{equation}
-%
-Then the uniform model \eqref{eq:unilaplace} has \( \hat{A}(s) = \frac{F_m(\alpha)}{s m G_m(\alpha)} \).
+  \Delta f_m(\beta) &= S(m\beta) - \zeta S((m-1)\beta), &
+  f_m(\beta) &= \frac{S(m\beta)}{\Delta f_m(\beta)}, \\
+  \Delta g_m(\beta) &= \Delta f_m(\beta) + 1 - \zeta , &
+  g_m(\beta) &= \frac{S(m\beta) + 1}{\Delta g_m(\beta)}, \\
+  \text{where}\quad \zeta &= 2s + 1. \\
+\end{aligned}
+\end{equation}
+%
 Then \cref{eq:shortenlaplace} can be be written as
@@
   \hat{A}(s) =
-    \frac{(1-\varepsilon)F_m(\alpha) + \varepsilon \zeta F_{m-1}(\alpha)}
-         {s (m - \varepsilon) ((1-\varepsilon)G_m(\alpha) + \varepsilon \zeta G_{m-1}(\alpha))}
-    % = \frac{F_m - \varepsilon \Delta F_{m}}
-    %      {s (m - \varepsilon) (G_m - \varepsilon \Delta G_{m})}
-    = \brk{\frac{\Delta F_{m}}{s \Delta G_{m}}}
+    = \brk{\frac{\Delta f_{m}}{s \Delta g_{m}}}
     \frac{f_m - \varepsilon}{(m - \varepsilon) (g_m - \varepsilon)}.
@@
 (\ref{fig:short_ratio}): The ratios of the quantities plotted in (\ref{fig:short_opt}) to \cref{eq:unienv}.
-The dotted orange line is the extension of the \(M=4\) line with \(\varepsilon\) held at 0, as discussed below \cref{eq:shortenvalid}.
+The dotted blue and orange lines are the extensions of the \(M=2,4\) lines with \(\varepsilon\) held at 0, as discussed below \cref{eq:shortenvalid}.
+Noting the scale of the y-axis, we see that \cref{eq:unienv,eq:unienvapprox} do provide good approximations, especially for larger \(M\).
 \label{fig:shorten}}
@@
   \qquad
-  \hat{A}(s) = \brk{\frac{\Delta F_{m}}{s \Delta G_{m}}}
+  \hat{A}(s) = \brk{\frac{\Delta f_{m}}{s \Delta g_{m}}}
     \frac{1}{\brk{\sqrt{\abs{m - f_m}} - \sqrt{\abs{g_m - f_m}}}^2}.
@@
   \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
-    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^{M-1}-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
+    m(2s+1) \sqrt{2s} \sinh\brk{{\prn{m-\tfrac{1}{2}}\beta}} 
+    - S(m\beta) [S(m\beta) + 1]
       = 0,\\
   \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
-    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^{M-1}-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
+    (m-1) \sqrt{2s} \sinh\brk{{\prn{m-\tfrac{1}{2}}\beta}} 
+    - (2s+1) S((m-1)\beta) [S((m-1)\beta) + 1]
       = 0,\\
-    \text{where}\quad
-    s = &\, \frac{(\alpha-1)^2}{2\alpha},
+    &\text{where}\quad
+    s = S(\beta) = 2 \sinh^2(\beta/2),
 \end{aligned}
@@
 For higher \(s\), \(\varepsilon = 1\) is the optimum, which is eqivalent to \(\varepsilon = 0\) with \(M-2\) states.
-This is depicted by the dotted orange line in \cref{fig:shorten}\ref{fig:short_ratio}.
+This is depicted by the dotted lines in \cref{fig:shorten}\ref{fig:short_ratio}.
 
@@
 If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
-In this case, however, there is an overlap as seen in \cref{fig:shorten}\ref{fig:short_ratio}.
-The \(M=2\) region of validity overlaps everything else.
-The other overlaps are less extreme, but they do exist.
+In fact there are overlaps, as seen in \cref{fig:shorten}\ref{fig:short_ratio} for \(M=6\) and higher.
 This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
@@
   & \frac{\sqrt{N}}{1+r\tau}, &
-  & r\tau \leq \frac{1}{2\sinh^2\frac{y_*}{2}} , \\
-  & \sqrt{N}\prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \log\alpha, &
-  \frac{1}{2\sinh^2 \frac{y_*}{2}} \leq & r\tau \leq \frac{1}{2\sinh^2 \frac{y_*}{M}} , \\
-  & \max_\varepsilon \brc{\frac{2\sqrt{N}\varepsilon}{2+(M-2)\varepsilon} \,
-      \frac{ (\alpha^{\frac{M}{2}}-1)^2 - (1-\varepsilon)\alpha(\alpha^{\frac{M-2}{2}}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }}, &
-  \frac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau .
+  & r\tau \leq \tfrac{1}{2\sinh^2\frac{y_*}{2}} , \\
+  & \sqrt{N}\prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } }\sinh\inv \!\!\sqrt{1/2r\tau}, &
+  \tfrac{1}{2\sinh^2 \frac{y_*}{2}} \leq & r\tau \leq \tfrac{1}{2\sinh^2 \frac{y_*}{M}} , \\
+  & \max_{\varepsilon \in [0,1]} \brc{
+      \frac{2\sqrt{N}(1-\varepsilon) \Delta_{M,\varepsilon} \brk{S\prn{\tfrac{M\beta}{2}}}}
+      {\Delta_{M,\varepsilon}[M]\, \Delta_{M,\varepsilon}\! \brk{S\prn{\tfrac{M\beta}{2}} + 1} } }, &
+  \tfrac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau ,
 \end{aligned}
@@
 %
+where  we use the notation \( \Delta_{M,\varepsilon}[T(M, \beta)] = T(M, \beta) - \varepsilon T(M-2, \beta) \).
 For $M \gg 1$, and $r\tau$ much greater than the lower bound of each region, we can approximate this as

### 2020-06-20 Code/Python/complex_synapse/shorten.py

@@ -129,7 +129,7 @@ def limits(num: int, debug: bool = False) -> float:
     Note: higher t -> lower s -> higher M
     """
     if num == 2:
-        return (0.99, 0.01)
+        return (0.01, np.exp(-Y_STAR))
     x_0, x_1 = alpha_star(num - 2)[0], alpha_star(num)[0]
     lo_sol = sco.root_scalar(lower, args=(num,), x0=x_0, x1=x_1)
     if debug:

### 2020-06-21 Code/Python/complex_synapse/sticky.py
graph of optimised sticky models

@@ -1,0 +1,136 @@
+# To add a new cell, type '# %%'
+# To add a new markdown cell, type '# %% [markdown]'
+# -------------------------------------
+"""Properties of optimal shortened models
+"""
+from typing import Tuple
+import numpy as np
+import scipy.optimize as sco
+from .shorten import Y_STAR
+
+Arrays = Tuple[np.ndarray, ...]
+
+
+# -------------------------------------
+def beta_to_s(beta: float) -> float:
+    """Convert alpha to s"""
+    return np.cosh(beta) - 1
+
+
+def s_to_beta(sval: float) -> (float, float):
+    """Convert s to beta. two possibilities: negatives"""
+    return np.arccosh(sval + 1)
+
+
+# -------------------------------------
+def beta_star(num: int) -> float:
+    """Value of beta at which num is optimal"""
+    return 2 * Y_STAR / num
+
+
+# =============================================================================
+
+
+def components(beta: float, num: int) -> Tuple[float, ...]:
+    """Components of numerator and denominator
+
+    numerator = numer - eps * dnumer
+    denominator = denom - eps * ddenom
+    """
+    denom = np.cosh(num * beta / 2)
+    numer = denom - 1
+    ddenom = np.cosh((num - 2) * beta / 2)
+    dnumer = ddenom - 1
+    return numer, denom, dnumer, ddenom
+
+
+# -------------------------------------
+def sticky(eps: float, beta: float, num: int) -> float:
+    """Laplace-SNR for shortened serial model"""
+    sval = beta_to_s(beta)
+    numer, denom, dnumer, ddenom = components(beta, num)
+    fnumer = numer - eps * dnumer
+    fdenom = denom - eps * ddenom
+    fmu = num - eps * (num - 2)
+    return 2 * (1 - eps) * fnumer / (fmu * sval * fdenom)
+
+
+def neg_sticky(eps: float, alpha: float, num: int) -> float:
+    """negative Laplace-SNR for shortened serial model"""
+    return - sticky(eps, np.log(alpha), num)
+
+
+# -------------------------------------
+def eps_stars(beta: float, num: int) -> float:
+    """Optimal epsilon for sticky serial model"""
+    if num == 2:
+        return 0, 0
+    numer, denom, dnumer, ddenom = components(beta, num)
[…]

### 2020-06-22 Notes/laplace.tex
graph of optimised sticky models
started alterative optimisation approach

@@
 \input{sl_preamble.tex}
+\usepackage{mathtools}
 %
@@
 %
-
 When $M \gg 2$, the validity region is approximately
@@
 %
+We will lok at a more accurate interpolation between different \(M\) in the \hyperref[sec:serial_shorten]{next section}.
 
@@
 However, it does match up with the full formula, \eqref{eq:unienv} at the boundary.
-
-\subsection{Sticky serial model}\label{sec:serial_sticky}
-
-Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_sticky}.
-We proceed in the same manner as in \autoref{sec:serial_uniform}, with the following differences.
-
-The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
-%
-\begin{equation}\label{eq:stickyenc}
-  \begin{aligned}
-  \enc_{1,2} = -\enc_{2m,2m-1} &= \frac{1-\varepsilon}{2},
-  &\quad
-  \eq_1 = \eq_{2m} &= \frac{1}{2(m-(m-1)\varepsilon)},
-  \\ &&
-  \eq_i &=  \frac{1-\varepsilon}{2(m-(m-1)\varepsilon)}
-  \quad\text{for}\quad 1<i<2m.
-  \end{aligned}
-\end{equation}
-%
-This means that the Laplace transform is given by
-%
-\begin{equation}\label{eq:stickyareaeta}
-  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_M(s))}
-                    {2(m-(m-1)\varepsilon)}
-       = \frac{2(1-\varepsilon)\, \etw_1(s)}{2(m-(m-1)\varepsilon)}.
-\end{equation}
-%
-
-The $\etw_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
-%
-\begin{equation}\label{eq:detabndsticky}
-\begin{aligned}
-  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{1-\varepsilon}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
-\end{aligned}
-\end{equation}
-%
-These are solved by
-%
-\begin{equation}\label{eq:stickycoeffs}
-\begin{aligned}
-  B &= -\frac{\e^{\beta} - \varepsilon}
-      {s [\e^{\beta} + 1] [\cosh(m\beta) - \varepsilon \cosh((m-1)\beta)]}, \\
-  C &= -\frac{1 - \varepsilon\e^{\beta}}
-      {s [\e^{\beta} + 1] [\cosh(m\beta) - \varepsilon \cosh((m-1)\beta)]},
-\end{aligned}
-\end{equation}
-%
-which leads to the Laplace transform
-%
-\begin{equation}\label{eq:stickylaplace}
-  \hat{A}(s) = \frac{(1-\varepsilon)}{[m-(m-1)\varepsilon]s} \,
-      \brk{\frac{S(m\beta) - \varepsilon S((m-1)\beta)}
-      {S(m\beta) - \varepsilon S((m-1)\beta) + 1 - \varepsilon}}.
-\end{equation}
-%
-
-We can maximise this \wrt $\varepsilon$.
-This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
-We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0$:
-%
-\begin{equation}\label{eq:stickyvalid}
-\begin{aligned}
-  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} &\propto
-    m \prn{S(m\beta) - S((m-1)\beta)} + (m-1) S(m\beta) \prn{S(m\beta) + 1} = 0,\\
-   &\text{where} \quad s =  S(\beta) = 2 \sinh^2(\beta/2).
-\end{aligned}
-\end{equation}
-%
-The ``sticky'' models are optimal at values of $s$ smaller than this.
-This is not quite the same as the lower boundary of the region \eqref{eq:univalid}, so there is a gap between them where the uniform model with $M$ states is optimal.
-The width of this gap is insignificant for larger $M$.
-
-For small $s \ll M^{-2}$, it is given by
-%
-\begin{equation}\label{eq:stickyeps}
-  \varepsilon = 1 - \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},
-\end{equation}
-%
-and the Laplace transform is
-%
-\begin{equation}\label{eq:stickyenv}
-  \hat{A}(s) = (M-1) \prn{ 1 - \sqrt{M(M-2)s} + \CO(M^2s)}.
-\end{equation}
-%
 
@@
 %
-\begin{equation}\label{eq:shortdefs}
+\begin{equation*}
 \begin{aligned}
@@
   g_m(\beta) &= \frac{S(m\beta) + 1}{\Delta g_m(\beta)}, \\
-  \text{where}\quad \zeta &= 2s + 1. \\
-\end{aligned}
-\end{equation}
-%
-Then \cref{eq:shortenlaplace} can be be written as
-%
-\begin{equation}\label{eq:shortsimple}
-  \hat{A}(s) =
-    = \brk{\frac{\Delta f_{m}}{s \Delta g_{m}}}
-    \frac{f_m - \varepsilon}{(m - \varepsilon) (g_m - \varepsilon)}.
-\end{equation}
-%
+  \text{where}\quad \zeta &= 2s + 1, 
+  \qquad\qquad \implies & 
+  \hat{A}(s) &= \frac{\Delta f_{m} (f_m - \varepsilon)}
+           {s \Delta g_{m} (m - \varepsilon) (g_m - \varepsilon)}.
+  \\
+\end{aligned}
+\end{equation*}
+%
+The results are plotted in \cref{fig:shorten}.
 
@@
   \qquad
-  \hat{A}(s) = \brk{\frac{\Delta f_{m}}{s \Delta g_{m}}}
-    \frac{1}{\brk{\sqrt{\abs{m - f_m}} - \sqrt{\abs{g_m - f_m}}}^2}.
+  \hat{A}(s) = \frac{\Delta f_{m}}
+      {s \Delta g_{m} \brk{\sqrt{\abs{m - f_m}} - \sqrt{\abs{g_m - f_m}}}^2}.
 \end{equation}
@@
 \begin{aligned}
-  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
-    m(2s+1) \sqrt{2s} \sinh\brk{{\prn{m-\tfrac{1}{2}}\beta}} 
-    - S(m\beta) [S(m\beta) + 1]
+  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\mathrlap{\varepsilon=0}} \propto &\,
+    m(2s+1) \brk{S(m\beta) - S((m-1)\beta)} - S(m\beta) [S(m\beta) + 1]
       = 0,\\
-  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
-    (m-1) \sqrt{2s} \sinh\brk{{\prn{m-\tfrac{1}{2}}\beta}} 
-    - (2s+1) S((m-1)\beta) [S((m-1)\beta) + 1]
+  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\mathrlap{\varepsilon=1}} \propto &\,
+    (m-1)\brk{S(m\beta) - S((m-1)\beta)} - (2s+1) S((m-1)\beta) [S((m-1)\beta) + 1]
       = 0,\\
@@
 
+\subsection{Sticky serial model}\label{sec:serial_sticky}
+
+Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_sticky}.
+We proceed in the same manner as in \autoref{sec:serial_uniform}, with the following differences.
+
+The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
+%
+\begin{equation}\label{eq:stickyenc}
+  \begin{aligned}
+  \enc_{1,2} = -\enc_{2m,2m-1} &= \frac{1-\varepsilon}{2},
+  &\quad
+  \eq_1 = \eq_{2m} &= \frac{1}{2(m-(m-1)\varepsilon)},
+  \\ &&
+  \eq_i &=  \frac{1-\varepsilon}{2(m-(m-1)\varepsilon)}
+  \quad\text{for}\quad 1<i<2m.
+  \end{aligned}
+\end{equation}
+%
+This means that the Laplace transform is given by
+%
+\begin{equation}\label{eq:stickyareaeta}
+  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_M(s))}
+                    {2(m-(m-1)\varepsilon)}
+       = \frac{2(1-\varepsilon)\, \etw_1(s)}{2(m-(m-1)\varepsilon)}.
+\end{equation}
+%
+
+The $\etw_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
+%
+\begin{equation}\label{eq:detabndsticky}
+\begin{aligned}
+  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{1-\varepsilon}{2}}\delta\eta_1 &= 0, \\
+  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
+\end{aligned}
+\end{equation}
+%
+These are solved by
+%
+\begin{equation}\label{eq:stickycoeffs}
+\begin{aligned}
+  B &= -\frac{\e^{\beta} - \varepsilon}
+      {s [\e^{\beta} + 1] [\cosh(m\beta) - \varepsilon \cosh((m-1)\beta)]}, \\
+  C &= -\frac{1 - \varepsilon\e^{\beta}}
+      {s [\e^{\beta} + 1] [\cosh(m\beta) - \varepsilon \cosh((m-1)\beta)]},
+\end{aligned}
+\end{equation}
+%
+which leads to the Laplace transform
+%
+\begin{equation}\label{eq:stickylaplace}
+  \hat{A}(s) = \frac{(1-\varepsilon)}{[m-(m-1)\varepsilon]s} \,
+      \brk{\frac{S(m\beta) - \varepsilon S((m-1)\beta)}
+      {S(m\beta) - \varepsilon S((m-1)\beta) + 1 - \varepsilon}}.
+\end{equation}
+%
+
+\begin{figure}[ht]
+\begin{center}
+  \includegraphics[width=0.4\linewidth]{sticky_opt.pdf}
+\caption{Signal-to-noise ratio for optimal sticky models, \cref{eq:stickyenc} optimised over \(\varepsilon\) for \(M=10\), in blue.
+  The orange line is the SNR for the optimised shortened models, \cref{eq:shortenenc} with \(M=10\).
+  The grey line is our theoretical envelope from \cref{eq:env}.
+  \label{fig:sticky}}
+\end{center}
+\end{figure}
+
+We can maximise this \wrt $\varepsilon$.
+This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
+We define
+%
+\begin{equation*}
+\begin{aligned}
+  \Delta m &= (m-1), &
+  \mu &= \frac{m}{\Delta m}, \\
+  \Delta f_m(\beta) &= S((m-1)\beta), &
+  f_m(\beta) &= \frac{S(m\beta)}{\Delta f_m(\beta)}, \\
+  \Delta g_m(\beta) &= \Delta f_m(\beta) + 1 , &
+  g_m(\beta) &= \frac{S(m\beta) + 1}{\Delta g_m(\beta)}, \\
+\end{aligned}
+ \; \implies \quad
+ \hat{A}(s) = \frac{\Delta f_m (1 - \varepsilon) (f_m - \varepsilon)}
+                    {s \Delta m \Delta g_m (\mu - \varepsilon) (g_m - \varepsilon)}.
+\end{equation*}
+%
+The minimum is at the solution of
+%
+\begin{equation*}
+  \brk{(1 + f_m -  g_m - \mu)\varepsilon - f_m + \mu g_m}^2 
+    = (\mu - 1) (g_m - 1) (\mu - f_m) (g_m - f_m).
+\end{equation*}
+%
+The result is plotted in \cref{fig:sticky}.
+
+We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0$:
+%
+\begin{equation}\label{eq:stickyvalid}
+\begin{aligned}
+  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\mathrlap{\varepsilon=0}} &\propto
+    m \brk{S(m\beta) - S((m-1)\beta)} - S(m\beta) \brk{S(m\beta) + 1} = 0,\\
+   &\text{where} \quad s =  S(\beta) = 2 \sinh^2(\beta/2).
+\end{aligned}
+\end{equation}
+%
+The ``sticky'' models are optimal at values of $s$ smaller than this.
+This is not quite the same as the lower boundary of the region \cref{eq:shortenvalid}, or its approximate version \cref{eq:univalid}, so there is a gap between them where the uniform model with $M$ states is optimal.
+The width of this gap is insignificant for larger $M$, as seen in \cref{fig:sticky}.
+
+For small $s \ll M^{-2}$, it is given by
+%
+\begin{equation}\label{eq:stickyeps}
+  \varepsilon = 1 - \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},
+\end{equation}
+%
+and the Laplace transform is
+%
+\begin{equation}\label{eq:stickyenv}
+  \hat{A}(s) = (M-1) \prn{ 1 - \sqrt{M(M-2)s} + \CO(M^2s)}.
+\end{equation}
+%
+
 \subsection{Heuristic envelope}\label{sec:heuristicenv}
@@
   \qquad
-  \alpha(\tau) = \frac{ 1 + r\tau +\sqrt{1+2r\tau} }{ r\tau }.
-\end{equation}
-%
-Then, combining \eqref{eq:binaryenv}, \eqref{eq:unienv} and \eqref{eq:stickylaplace}, we find
+  \beta(\tau) = \sinh\inv \!\!\sqrt{1/2r\tau}.
+\end{equation}
+%
+Then, combining \eqref{eq:binaryenv}, \eqref{eq:unienv} and \eqref{eq:stickylaplace}, using the notation \( \Delta_{M,\varepsilon}[T(M, \beta)] = T(M, \beta) - \varepsilon T(M-2, \beta) \)., we find
 %
@@
   & r\tau \leq \tfrac{1}{2\sinh^2\frac{y_*}{2}} , \\
-  & \sqrt{N}\prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } }\sinh\inv \!\!\sqrt{1/2r\tau}, &
+  & \sqrt{N}\prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \beta(\tau), &
   \tfrac{1}{2\sinh^2 \frac{y_*}{2}} \leq & r\tau \leq \tfrac{1}{2\sinh^2 \frac{y_*}{M}} , \\
   & \max_{\varepsilon \in [0,1]} \brc{
-      \frac{2\sqrt{N}(1-\varepsilon) \Delta_{M,\varepsilon} \brk{S\prn{\tfrac{M\beta}{2}}}}
+      \frac{2\sqrt{N}(1-\varepsilon)\, \Delta_{M,\varepsilon}\! \brk{S\prn{\tfrac{M\beta}{2}}}}
       {\Delta_{M,\varepsilon}[M]\, \Delta_{M,\varepsilon}\! \brk{S\prn{\tfrac{M\beta}{2}} + 1} } }, &
-  \tfrac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau ,
+  \tfrac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau .
 \end{aligned}
@@
 %
-where  we use the notation \( \Delta_{M,\varepsilon}[T(M, \beta)] = T(M, \beta) - \varepsilon T(M-2, \beta) \).
 For $M \gg 1$, and $r\tau$ much greater than the lower bound of each region, we can approximate this as

### 2020-06-22 Code/Python/complex_synapse/soptimise.py

@@ -1,0 +1,107 @@
+# -*- coding: utf-8 -*-
+"""Optimising synapse models
+"""
+from numbers import Number
+import numpy as np
+import scipy.optimize as sco
+from sl_py_tools.iter_tricks import dcount, denumerate, delay_warnings, dzip
+from sl_py_tools.numpy_tricks import markov_param as mp
+import numpy_linalg as la
+from .synapse_opt import SynapseOptModel
+from . import optimise as opt
+
+# =============================================================================
+# Optimisation helper functions
+# =============================================================================
+
+
+def make_laplace_problem(rate: Number, nst: int, **kwds) -> (dict, dict):
+    """Make an optimize problem.
+    """
+    modelopts = opt.get_model_opts(kwds)
+    opt.get_param_opts(kwds)
+    keep_feasible = kwds.pop('keep_feasible', False)
+    method = kwds.get('method', 'SLSQP')
+
+    model = SynapseOptModel.rand(nst, **modelopts)
+    fun = opt.make_loss_function(model, model.laplace_grad, None)
+    x_init = model.get_params()
+    hess = opt.make_loss_function(model, model.laplace_hess, None)
+
+    con_coeff = opt.constraint_coeff(nst, modelopts['npl'])
+    bounds = sco.Bounds(0, 1 + rate, keep_feasible)
+    ofun = opt.make_loss_function(model, model.peq_min_fun, rate)
+    ojac = opt.make_loss_function(model, model.peq_min_grad, rate)
+    if method not in {'SLSQP', 'trust-constr'}:
+        raise ValueError('method must be one of SLSQP, trust-constr')
+    if method == 'trust-constr':
+        ohess = opt.make_loss_function(model, model.peq_min_hess, rate)
+        dconstraint = sco.LinearConstraint(con_coeff, rate, 1 + rate,
+                                           keep_feasible)
+        oconstraint = sco.NonlinearConstraint(ofun, 0, 1, ojac, ohess,
+                                              keep_feasible)
+    else:
+        hess = None
+        dconstraint = {'type': 'ineq', 'args': (),
+                       'fun': lambda x: 1 + rate - con_coeff @ x,
+                       'jac': lambda x: -con_coeff}
+        oconstraint = {'type': 'ineq', 'args': (), 'fun': ofun, 'jac': ojac}
+
+    problem = {'fun': fun, 'x0': x_init, 'jac': True, 'hess': hess,
[…]

### 2020-06-22 Code/Python/complex_synapse/synapse_opt.py

@@ -10,7 +10,7 @@ from typing import Tuple, Optional, Union, Dict, ClassVar
 import numpy as np
 import numpy_linalg as la
 from sl_py_tools.numpy_tricks import markov_param as mp
-from .builders import scalarise
+from . import builders as bld
 from .synapse_memory_model import SynapseMemoryModel as _SynapseMemoryModel
 from .synapse_base import SynapseBase as _SynapseBase
 
@@ -35,6 +35,12 @@ class SynapseOptModel(_SynapseMemoryModel):
         super().__init__(*args, **kwds)
         self._saved = (0, (), (), ())
         self._unchanged = False
+
+    @property
+    def nparam(self) -> int:
+        """number of plasticity types."""
+        npr = mp.num_param(self.nstates, drn=self.directions[0], **self.type)
+        return self.nplast * npr
 
     def get_params(self) -> la.lnarray:
         """Independent parameters of transition matrices.
@@ -104,16 +110,29 @@ class SynapseOptModel(_SynapseMemoryModel):
             if inv: (Z,Zs,ZQZs)
             if not inv: (Z^{-1},Zs^{-1})
         """
-        if all((self._unchanged,
-                len(self._saved[3]) == 2 + inv,
-                np.isclose(rate, self._saved[0]))):
+        if (all((self._unchanged, rate is None, self._saved[0] is None)) or
+            rate is not None and self._saved[0] is not None and
+            all((self._unchanged,
+                 len(self._saved[3]) == 2 + inv,
+                 np.isclose(rate, self._saved[0])))):
             return self._saved[1:]
         self._unchanged = True
 
         fundi = self.zinv()
-        fundis = self.zinv(rate)
-
         peq = self.peq()
+
+        if rate is None:
+            c_s = peq @ self.enc() @ fundi.inv
+            eta = fundi.inv @ self.weight
+            theta = fundi.inv @ self.enc() @ eta
+            if inv:
+                mats = (fundi.inv(),)
+            else:
+                mats = (fundi,)
+            self._saved = rate, (peq, c_s), (eta, theta), mats
+            return self._saved[1:]
+
+        fundis = self.zinv(rate)
         c_s = peq @ self.enc() @ fundis.inv
 
         eta = fundis.inv @ self.weight
@@ -176,7 +195,7 @@ class SynapseOptModel(_SynapseMemoryModel):
                 for m, d in zip([dsdwp, dsdwm], self.directions)]
[…]

### 2020-07-05 Notes/laplace.tex
pot/dep as index

@@
 
-We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold.
+We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold~\cite{Sommer1998retrieval}.
 For large $N$, this quantity will have a Gausssian distribution.
@@
 \begin{equation}\label{eq:SNRdef}
-  \snr(t) = \frac{\av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}}
-     {\sqrt{\var(\syn_\text{ideal}\cdt\syn(\infty))}},
+  \snr(t) = \frac{\av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)}}
+     {\sqrt{\var(\synid\cdt\syn(\infty))}},
 \end{equation}
@@
 where we used $\syn^\alpha=\pm1$.
-We can compute $\var(\syn_\text{ideal}\cdt\syn(\infty))$ by taking $t\to\infty$.
+We can compute $\var(\synid\cdt\syn(\infty))$ by taking $t\to\infty$.
 
@@
 \begin{equation}\label{eq:overlapex}
-  \av{\syn_\text{ideal}\cdt\syn(t)} = \sum_\alpha\av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}
-   = N \av{\syn^1_\text{ideal}\syn^1(t)},
+  \av{\synid\cdt\syn(t)} = \sum_\alpha\av{\synid^\alpha\syn^\alpha(t)}
+   = N \av{\synid^1\syn^1(t)},
 \end{equation}
@@
 \begin{equation}\label{eq:overlapone}
-  \av{\syn^1_\text{ideal}\syn^1(t)} = f\pot \av{\syn^1(t)}_{\text{pot},t=0} - f\dep \av{\syn^1(t)}_{\text{dep},t=0}
+  \av{\synid^1\syn^1(t)} = f\pot \av{\syn^1(t)}_{\text{pot},t=0} - f\dep \av{\syn^1(t)}_{\text{dep},t=0}
 \end{equation}
@@
 
-When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M\potdep$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$.
-The matrices $\M\potdep$ are transition matrices of discrete time Markov processes.
+When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M^\mu$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$,
+and \(\mu\) labels the diffrerent types of plasticity event.
+Potentiating events have \(\mu=\)pot, depressing events have \(\mu=\)dep.
+We also use the notation
+%
+\begin{equation*}
+  (-1)^\mu = 
+    \begin{cases}
+       1, & \mu = \text{pot}, \\
+      -1, & \mu = \text{dep}.
+    \end{cases}
+\end{equation*}
+%
+The matrices $\M^\mu$ are transition matrices of discrete time Markov processes.
 Their elements must satisfy the following constraints
@@
 \begin{equation}\label{eq:constr}
-\begin{aligned}
-  \M\potdep_{ij} &\geq 0, &\qquad
-  \sum_j \M\potdep_{ij} &= 1.
-\end{aligned}
-\end{equation}
-%
-The upper bounds $\M\potdep_{ij} \leq 1$ follow automatically from these.
+  \M^\mu_{ij} \geq 0 
+    \quad \forall \; \mu, i, j, \qquad
+  \sum_j \M^\mu_{ij} = 1
+    \quad \forall \; \mu, i.
+\end{equation}
+%
+The upper bounds $\M^\mu_{ij} \leq 1$ follow automatically from these.
 
@@
 \begin{equation}\label{eq:constri}
-  \M\potdep_{i \neq j} \geq 0, \qquad
-  \sum_{j\neq i} \M\potdep_{ij} \leq 1.
+  \M^\mu_{i j} \geq 0
+    \quad \forall \; \mu, i, j : i \neq j, \qquad
+  \sum_{j : j\neq i} \M^\mu_{ij} \leq 1
+    \quad \forall \; \mu, i.
 \end{equation}
@@
 %
-If the probability of a given plasticity event being potentiating/depressing is $f\potdep$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
+If the probability of a given plasticity event being potentiating/depressing is $f^\mu$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
 
@@
   \qquad\text{where}\quad
-  \frg = f\pot\M\pot + f\dep\M\dep - \I.
+  \frg = \sum_\alpha f^\mu \M^\mu - \I
+    = f\pot\M\pot + f\dep\M\dep - \I.
 \end{equation}
@@
 \begin{aligned}
-  \av{\syn^1_\text{ideal}\syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
-  \av{\syn^1_\text{ideal}\syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
+  \av{\synid^1 \syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
+  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
          &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
@@
 \begin{aligned}
-  \av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}
-    &= N \eq \enc \, \e^{(r\frg+\hom)t} \w , \\
+  \av{\synid \cdt \syn(t)} - \av{\synid \cdt \syn(\infty)}
+    &= N \eq \enc \, \e^{rt \frg} \w , \\
     \text{where} \qquad
-  \enc &= f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
+  \enc &= \sum_\mu (-1)^\mu f^\mu (\M^\mu - \I)
+    = f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
 \end{aligned}
@@
 %
-%We can absorb $\hom$ into $\M\potdep$ by defining
-%%
-%\begin{equation}\label{eq:absorbhom}
-%\begin{aligned}
-%  \Mh\potdep &= \M\potdep + \frac{\hom}{2rf\potdep},
-%  &\qquad
-%  \frgh &= f\pot\Mh\pot + f\dep\Mh\dep - \I,
-%  \\&&
-%  \ench &= f\pot(\Mh\pot-\I)-f\dep(\Mh\dep-\I).
-%\end{aligned}
-%\end{equation}
-%%
-%Then the SNR curve is
-%%
-%\begin{equation}\label{eq:SNRcurvehom}
-%  \SNR(t) = \frac{\sqrt{N}  \eq \enc \,\e^{rt\frg}\w}
-%                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
-%\end{equation}
-%%
 The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
@@
 
-We can then treat the off-diagonal elements of $\M\potdep$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
+We can then treat the off-diagonal elements of $\M^\nu$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
 The diagonal elements are determined by setting the row sums to 1.
@@
 \begin{equation}\label{eq:derivpd}
-  \pdiff{\M\pot_{ij}}{\M\pot_{mn}} = \pdiff{\M\dep_{ij}}{\M\dep_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
-  \qquad
-  \pdiff{\M\pot_{ij}}{\M\dep_{mn}} = \pdiff{\M\dep_{ij}}{\M\pot_{mn}} = 0.
+  \pdiff{\M^\mu_{ij}}{\M^\nu_{mn}} 
+      = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}),
 \end{equation}
@@
   \frg_{i \neq j}   &\geq \abs{ \enc_{i \neq j} }, &\qquad
-  \sum_{j\neq i}\frg_{ij} \pm \abs{ \enc_{ij} } &\leq f\potdep, \\
+  \sum_{j\neq i}\frg_{ij} + (-1)^\mu \enc_{ij}  &\leq f^\mu, \\
   \sum_j \frg_{ij} &=0, &
@@
 \begin{equation}\label{eq:lAdiff}
-  \pdiff{\hat{A}(s)}{\M\potdep_{mn}}
-     = f\potdep \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
-     +\prn{rc_m(s) \pm 1} \brk{\etw_m(s) - \etw_n(s)} },
-\end{equation}
-%
-with the upper/lower sign in $\pm$ corresponding to derivatives \wrt $\M\potdep$.
+  \pdiff{\hat{A}(s)}{\M^\mu_{mn}}
+     = f^\mu \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
+     +\prn{rc_m(s) + (-1)^\mu} \brk{\etw_m(s) - \etw_n(s)} }.
+\end{equation}
+%
 The term with $\Theta(s)$ comes from differentiating $\eq$, the term with $c(s)$ comes from differentiating $\etw(s)$ and the term with $\pm1$ comes from differentiating $\enc$ in \eqref{eq:larea}.
@@
 \begin{equation}\label{eq:hessian}
-\begin{aligned}
-  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
-     (f\pot)^2   [R + R\trans + S + S\trans]_{ijkl}, \\
-  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
-     f\pot f\dep [R + R\trans - S + S\trans]_{ijkl}, \\
-  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
-     f\pot f\dep [R + R\trans + S - S\trans]_{ijkl}, \\
-  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\dep_{kl}} &=
-     (f\dep)^2   [R + R\trans - S - S\trans]_{ijkl}, \\
-\end{aligned}
+  \pdiff{^2\hat{A}(s)}{\M^\mu_{ij} \partial\M^\nu_{kl}} =
+     f^\mu f^\nu [R + R\trans + (-1)^\nu S + (-1)^\mu S\trans]_{ijkl}.
 \end{equation}
@@
   \includegraphics[width=0.8\linewidth]{LenvProven.eps}
-  \caption[Proven envelope]{The memory curve envelope that we can prove, \eqref{eq:env}.
+  \caption[Proven envelope for the signal-to-noise ratio]
+  {The memory curve envelope that we can prove, \eqref{eq:env}.
   The number of synapses $N=10^4$ and the number of states $M=12$.
@@
   \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\etw_i(s) - \etw_j(s))
-        +\sum_{i j} \prn{ f\pot\mu\pot_{ij} \M\pot_{ij} + f\dep\mu\dep_{ij} \M\dep_{ij} }\\
+        +\sum_{\nu i j} \prn{ f^\nu \mu^\nu_{ij} \M\pot_{ij} }
         &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i\w_i },
@@
 \begin{equation}\label{eq:KTfinite}
-  \pdiff{\CL}{\M\potdep_{ij}} = 0, \quad
-  \mu\potdep_{ij} \geq 0, \quad
-  \pdiff{\CL}{\mu\potdep_{ij}} \geq 0, \quad
-  \mu\potdep_{ij}\pdiff{\CL}{\mu\potdep_{ij}} = 0, \quad
+  \pdiff{\CL}{\M^\nu_{ij}} = 0, \quad
+  \mu^\nu_{ij} \geq 0, \quad
+  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
+  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
   \pdiff{\CL}{\lambda} = 0.
@@
 \begin{aligned}
-  \frac{1}{f\potdep} \pdiff{\CL}{\Mh\potdep_{mn}} =& \,
+  \frac{1}{f^\nu} \pdiff{\CL}{\Mh^\nu_{mn}} = & \,
       \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
-     +\prn{rc_m(s) \pm 1} \brk{\etw_m(s) - \etw_n(s)} } \\
-     &+ \mu\potdep_{mn} -  \mu\potdep_{mm}
+     +\prn{rc_m(s) + (-1)^\nu} \brk{\etw_m(s) - \etw_n(s)} } \\
+     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
      + \lambda \eq_m \sum_i (\fptb_{mi}(0)-\fptb_{ni}(0)) \eq_i\w_i.
@@
 \begin{equation}\label{eqlaplacescale}
-  A(\lambda s;\lambda * \M\potdep) = A(s;\M\potdep).
+  A(\lambda s;\lambda * \M^\nu) = A(s;\M^\nu).
 \end{equation}
@@
   % Requires \usepackage{graphicx}
-  \includegraphics[width=0.8\linewidth]{LenvNum.eps}\\
-  \caption[Numerical envelope]{Numerical envelope for the memory curve, \eqref{eq:snrb}, when $f\potdep=\frac{1}{2}$, the number of synapses $N=10^4$ and the number of states $M=12$.
+  \includegraphics[width=0.8\linewidth]{LenvNum.svg}\\
+  \caption[Numerical envelope for the signal-to-noise ratio]
+  {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
+  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numericallly with $f\potdep=\frac{1}{2}$, the number of synapses $N=10^4$ and the number of states $M=12$.
   The dashed red line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
@@
   \qquad\text{where}\qquad
-  s = S(\beta) \equiv 2 \sinh^2 \prn{\frac{\beta}{2}},
+  s = S(\beta) \equiv 2 \sinh^2 \prn{\frac{\beta}{2}}
+    = \cosh (m\beta) - 1,
 \end{equation}
@@
 \begin{equation}\label{eq:unicoeffs}
-  B = -\frac{1}{s(\e^{-\beta} + 1) \cosh \prn{m \beta}},
+  B = -\frac{1}{s(1 + \e^{-\beta}) \cosh \prn{m \beta}},
   \qquad
@@
 \begin{equation}\label{eq:unilaplace}
-  A(s) = \frac{2\etw_1(s)}{2m}
+  A(s) = \frac{\etw_1(s)}{m}
    = \frac{1}{m}\prn{\frac{1}{s}+B+C}
@@
 \end{myenuma}
-\caption{Signal-to-noise ratio for optimal shortened models, \cref{eq:shortenenc} optimised over \(\varepsilon\).
+\caption[Signal-to-noise ratio for optimal shortened models]
+{Signal-to-noise ratio for optimal shortened models, \cref{eq:shortenenc} optimised over \(\varepsilon\).
 (\ref{fig:short_opt}): Solid grey line is \cref{eq:unienv}, the uniform model with \(M\) treated as a continuous parameter and optimised.
@@
   \includegraphics[width=0.4\linewidth]{sticky_opt.pdf}
-\caption{Signal-to-noise ratio for optimal sticky models, \cref{eq:stickyenc} optimised over \(\varepsilon\) for \(M=10\), in blue.
+\caption[Signal-to-noise ratio for optimal sticky models]
+{Signal-to-noise ratio for optimal sticky models, \cref{eq:stickyenc} optimised over \(\varepsilon\) for \(M=10\), in blue.
   The orange line is the SNR for the optimised shortened models, \cref{eq:shortenenc} with \(M=10\).
@@
 
-We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0$:
+We can find the boundary of the region of applicability by asking: at what value of $s$ is the optimal value of $\varepsilon=0$?
 %
@@
 %
-\begin{equation}\label{eq:henvapproc}
+\begin{equation}\label{eq:henvapprox}
   \snrb(\tau) \approx \left\{
@@
   \includegraphics[width=0.8\linewidth]{LenvHeuristic.svg}\\
-  \caption[Heuristic envelope]{Heuristic envelope.
-  Analytic expressions are only valid to the right of each region.}\label{fig:heuristicenv}
+  \caption[Heuristic envelope for the signal-to-noise ratio]
+  {Heuristic envelope  for the signal-to-noise ratio, \cref{eq:henv}.
+  Analytic expressions from \cref{eq:henvapprox} are only valid to the right of each region.}\label{fig:heuristicenv}
 \end{figure}

### 2020-07-07 Notes/laplace.tex
moved model creation

@@
 \newcommand{\pib}{\boldsymbol{\pi}}
+\newcommand{\arow}{\boldsymbol{\xi}}
 \newcommand{\Lb}{\boldsymbol{\Lambda}}
@@
 We denote the probability distribution across these states by the row vector $\pr(t)$.
-We denote a column vector of ones by $\onev$, so the normalisation condition is $\pr(t)\onev=1$.
-
-When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M^\mu$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$,
-and \(\mu\) labels the diffrerent types of plasticity event.
-Potentiating events have \(\mu=\)pot, depressing events have \(\mu=\)dep.
+We denote a column vector of ones by $\onev$, so the normalisation condition is $\pr(t) \, \onev=1$.
+
+When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M^xn  $ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$,
+and \(\nu \) labels the diffrerent types of plasticity event.
+Potentiating events have \(\nu=\)pot, depressing events have \(\nu=\)dep.
 We also use the notation
@@
 \begin{equation*}
-  (-1)^\mu = 
+  (-1)^\nu = 
     \begin{cases}
-       1, & \mu = \text{pot}, \\
-      -1, & \mu = \text{dep}.
+       1, & \nu = \text{pot}, \\
+      -1, & \nu = \text{dep}.
     \end{cases}
@@
 %
-The matrices $\M^\mu$ are transition matrices of discrete time Markov processes.
+The matrices $\M^\nu$ are transition matrices of discrete time Markov processes.
 Their elements must satisfy the following constraints
@@
 \begin{equation}\label{eq:constr}
-  \M^\mu_{ij} \geq 0 
-    \quad \forall \; \mu, i, j, \qquad
-  \sum_j \M^\mu_{ij} = 1
-    \quad \forall \; \mu, i.
-\end{equation}
-%
-The upper bounds $\M^\mu_{ij} \leq 1$ follow automatically from these.
+  \M^\nu_{ij} \geq 0 
+    \quad \forall \; \nu, i, j, \qquad
+  \sum_j \M^\nu_{ij} = 1
+    \quad \forall \; \nu, i.
+\end{equation}
+%
+The upper bounds $\M^\nu_{ij} \leq 1$ follow automatically from these.
 
@@
 \begin{equation}\label{eq:constri}
-  \M^\mu_{i j} \geq 0
-    \quad \forall \; \mu, i, j : i \neq j, \qquad
-  \sum_{j : j\neq i} \M^\mu_{ij} \leq 1
-    \quad \forall \; \mu, i.
+  \M^\nu_{i j} \geq 0
+    \quad \forall \; \nu, i, j : i \neq j, \qquad
+  \sum_{j : j\neq i} \M^\nu_{ij} \leq 1
+    \quad \forall \; \nu, i.
 \end{equation}
@@
 
-Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$ is
+Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$, 
+mariginalising over states at earlier times, is
 %
@@
 %
-If the probability of a given plasticity event being potentiating/depressing is $f^\mu$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
+If the probability of a given plasticity event being potentiating/depressing is $f\potdep$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
 
@@
   \qquad\text{where}\quad
-  \frg = \sum_\alpha f^\mu \M^\mu - \I
+  \frg = \sum_\nu f^\nu \M^\nu - \I
     = f\pot\M\pot + f\dep\M\dep - \I.
@@
     \text{where} \qquad
-  \enc &= \sum_\mu (-1)^\mu f^\mu (\M^\mu - \I)
+  \enc &= \sum_\nu (-1)^\nu f^\nu (\M^\nu - \I)
     = f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
@@
 \begin{equation}\label{eq:lfund}
-  \fund(s) = (s\I +\onev\pib -\MM)\inv,
-\end{equation}
-%
-where $\pib$ is an arbitrary row vector satisfying $\pib\onev = \frac{1}{\tau_0} \neq 0$.
+  \fund(s) = (s \I + \onev \arow -\MM)\inv,
+\end{equation}
+%
+where $\arow$ is an arbitrary row vector satisfying $\arow \onev = \frac{1}{\tau_0} \neq 0$.
 This reduces to the fundamental matrix at $s=0$.
@@
 Suppose we have some row vector $\mathbf{a}$ such that $\mathbf{a}\onev=0$.
-Then $\mathbf{a}\fund(s)$ is independent of $\pib$:
+Then $\mathbf{a}\fund(s)$ is independent of $\arow$:
 %
 \begin{equation}\label{eq:alfund}
-  \pdiff{(\mathbf{a}\fund(s))_i}{\pib_k} = -(\mathbf{a}\fund(s)\onev) \fund_{ki}(s)
+  \pdiff{(\mathbf{a}\fund(s))_i}{\arow_k} = -(\mathbf{a}\fund(s)\onev) \fund_{ki}(s)
       = -\frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
@@
 \begin{equation}\label{eq:lfptb}
-  \fptb(s) = (\onem\fund\dg(s)-\fund(s))\D,
+  \fptb(s) = (\onem \fund\dg(s)-\fund(s))\D,
   \qquad
@@
 is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau}{1+s\tau}$.
-We can the define
+We can then define
 %
 \begin{equation}\label{eq:lpmixtime}
-  \eta^{\pm}_i(s) = \sum_j \fptb_{ij}(s)\eq_j \prn{\frac{1\pm\w_j}{2}} =  \sum_{j\in\pm} \fptb_{ij}(s)\eq_j.
+  \eta^{\pm}_i(s) = \sum_j \fptb_{ij}(s) \eq_j \prn{\frac{1 \pm \w_j}{2}} 
+      =  \sum_{j\in \pm} \fptb_{ij}(s) \eq_j.
 \end{equation}
@@
   \etw_i(s) = \sum_j \fptb_{ij}(s)\eq_j \w_j
-    = \eta^+_i(s) - \eta^-_i(s) =2\eta^+_i(s) - \eta(s) =\eta(s) - 2\eta^-_i(s) .
+    = \eta^+_i(s) - \eta^-_i(s) 
+    = 2\eta^+_i(s) - \eta(s) 
+    = \eta(s) - 2\eta^-_i(s) .
 \end{equation}
@@
 Arranging the states in order of decreasing $\etw_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
-These quantities can still be used when $\w$ takes more than two values.
+The $\etw_i(s)$ can still be used when $\w$ takes more than two values.
 
@@
 
-Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurveExact} (ignoring the denominator)
+Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurveExact}
 %
@@
 \begin{aligned}
-  A(s) &= \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) \\
-   &= \intd[_0^{\infty}]{t} \frac{ \sqrt{N} \eq \enc \e^{(r\frg-s)t} \w }{ \sqrt{1-(f\pot-f\dep)^2(\eq\w)^2} } \\
+  A(s) = \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) 
+   &= \intd[_0^{\infty}]{t} \frac{ \sqrt{N} \eq \enc \e^{(r\frg-s)t} \w }
+                                { \sqrt{1-(f\pot-f\dep)^2(\eq\w)^2} } \\
    &= \frac{\sqrt{N} \eq \enc \prn{s - r\frg}\inv \w }{\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.

### 2020-07-07 Code/Python/complex_synapse/__init__.py

@@ -8,6 +8,4 @@ from . import builders
 from .synapse_memory_model import SynapseMemoryModel
 from .synapse_opt import SynapseOptModel
 from . import optimise
-from .optimise import make_loss_function, make_laplace_problem
-assert any((builders, SynapseMemoryModel, SynapseOptModel))
-assert any((make_loss_function, make_laplace_problem, optimise))
+assert any((True, builders, SynapseMemoryModel, SynapseOptModel, optimise))

### 2020-07-07 Code/Python/complex_synapse/synapse_base.py

@@ -83,7 +83,7 @@ class SynapseBase(np.lib.mixins.NDArrayOperatorsMixin):
         return cvl.conv_out_attr(self, 'plast', results, outputs, conv)
 
     # -------------------------------------------------------------------------
-    # %%* Housekeeping
+    # Housekeeping
     # -------------------------------------------------------------------------
 
     def fix(self):
@@ -105,7 +105,7 @@ class SynapseBase(np.lib.mixins.NDArrayOperatorsMixin):
         return f"{type(self).__name__} with M={self.nstates}, fpm={self.frac}"
 
     # -------------------------------------------------------------------------
-    # %%* Utility methods
+    # Utility methods
     # -------------------------------------------------------------------------
 
     def view(self, typ):
@@ -141,7 +141,7 @@ class SynapseBase(np.lib.mixins.NDArrayOperatorsMixin):
         return self.plast.shape[-1]
 
     # -------------------------------------------------------------------------
-    # %%* Factory methods
+    # Factory methods
     # -------------------------------------------------------------------------
 
     @classmethod

### 2020-07-07 Code/Python/complex_synapse/synapse_opt.py

@@ -18,9 +18,7 @@ from . import builders as bld
 from .synapse_base import SynapseBase as _SynapseBase
 from .synapse_memory_model import SynapseMemoryModel as _SynapseMemoryModel
 from .synapse_memory_model import well_behaved
-
-wrap = la.wrappers.Wrappers(la.lnarray)
-eig = wrap.several(np.linalg.eig)
+# =============================================================================
 
 
 class SynapseOptModel(_SynapseMemoryModel):
@@ -91,6 +89,56 @@ class SynapseOptModel(_SynapseMemoryModel):
             mp.mat_update_params(self.plast[i], plast[i], drn=drn, **self.type)
         self._unchanged = False
 
+    def _calc_fab(self, time, qas, evs, peq_enc):
+        """Calculate eigenvalue differences
+        """
+        expqt = np.exp(qas * time)
+        fab = qas.c - qas.r
+        degenerate = np.fabs(fab) < self.DegThresh
+        fab[degenerate] = 1.
+        fab = (expqt.c - expqt.r) / fab
+        fab[degenerate] = expqt[degenerate.nonzero()[0]] * time
+        fab *= evs.inv @ self.weight
+        fab *= (peq_enc @ evs).c
+        expwftw = (evs * expqt) @ (evs.inv @ self.weight)
+        return (evs.inv @ fab @ evs), expwftw
+
+    def snr_grad(self, time: Number) -> (float, la.lnarray):
+        """Gradient of instantaneous SNR memory curve.
+
+        Parameters
+        ----------
+        time : float
+            Recall time.
+
+        Returns
+        -------
+        func : float
+            Value of ``SNR`` at ``time``.
+        grad : la.lnarray (2n(n-1),)
+            Gradient of ``SNR`` at ``time`` with respect to parameters.
+        """
+        peq = self.peq()
+        peq_enc = peq @ self.enc()
+
+        evd = np.linalg.eig(self.markov())
+        fab, expqtw = self._calc_fab(time, *evd, peq_enc)
+
+        func = - peq_enc @ expqtw
+
+        dsdq = - _diagsub(peq.c * expqtw)
+        dsdw = - _diagsub(peq.c * (self.zinv().inv @ (self.enc() @ expqtw)))
+        dsdw -= _diagsub(fab)
+
+        dsdwp = self.frac[0] * (dsdw + dsdq)
+        dsdwm = self.frac[1] * (dsdw - dsdq)
+
+        grad = [mp.mat_to_params(m, drn=d, **self.type)
[…]

### 2020-07-15 Notes/laplace.tex
interpretation of laplace FPT

@@
 %equilibrium distribution
-\newcommand{\eq}{\pr^\infty}
+\newcommand{\eq}{\pib}
+%equilibrium metric
+\newcommand{\Eq}{\boldsymbol{\Pi}}
 %first passage times
@@
 \newcommand{\thb}{\boldsymbol{\theta}}
+\newcommand{\omb}{\boldsymbol{\omega}}
+\newcommand{\alb}{\boldsymbol{\alpha}}
 %synapse models
@@
 \renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
+% shifted process
+\newcommand{\shift}[1]{\widehat{#1}}
 
@@
 %
-This reduces to the first passage time matrix at $s=0$, but at other $s$ there is (probably) no relation.
-
-Using the exact same proof as in the case $s=0$, we see that the quantity
+This reduces to the mean first-passage time matrix at $s=0$.
+
+At other values of \(s\) can they be interpreted as the mean first-passage times of a different process?
+In order to have \cref{eq:lfund} as its fundamental matrix, the extra bit must draw probability from each state at a rate \(s\).
+In other words, it should have the form \(s(\onev \boldsymbol{\zeta} - \I)\) for some \(\boldsymbol{\zeta} \) 
+(which can be absorbed into \(\arow\) in \cref{eq:lfund}).
+This other process must have the same steady-state distribution, so that its mean first-passage times are given by \cref{eq:lfptb}, so the probability drawn from each state must be redistributed proportional to \(\eq\).
+This other stochastic process is
+%
+\begin{equation}\label{eq:lshift}
+  \shift{\MM}(s) = \MM + s (\onev \eq - \I).
+\end{equation}
+%
+This is a valid continuous-time Markov process.
+Its fundamental matrix is given by \cref{eq:lfund} with \( \arow_0 = \arow + s \eq \).
+
+This means that the \( \fptb(s) \) have all the same properties as in the case $s=0$.
+I particular, we see that the quantity
 %
@@
 %
-is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau}{1+s\tau}$.
+is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau_0}{1 + s \tau_0}$.
 We can then define
@@
                                 { \sqrt{1-(f\pot-f\dep)^2(\eq\w)^2} } \\
-   &= \frac{\sqrt{N} \eq \enc \prn{s - r\frg}\inv \w }{\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
+   &= \frac{\sqrt{N} \eq \enc \prn{s \I - r\frg}\inv \w }{\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
 \end{aligned}
@@
 
+We can also express this in terms of the shifted process in \cref{eq:larea}.
+We can choose how to distribute the shift across \(\M\pot\) and \(\M\dep\).
+Two choices from a continuum%
+\footnote{We could choose \( \shift{\M}^\nu(s) = \M^\nu + \frac{s g^\nu}{r f^\nu} (\onev \eq - \I) \), for any \(g^\nu \) with \( \sum_\nu g^\nu = 1 \).
+The two possibilities in \cref{eq:lshiftpd} come from \(g^\nu = \frac{1}{2}\) and \(g^\nu = f^\nu\) respectively.}
+are:
+%
+\begin{equation}\label{eq:lshiftpd}
+  \shift{\M}^\nu(s) = \M^\nu + \frac{s}{2 r f^\nu} (\onev \eq - \I)
+  \qquad \text{or} \qquad
+  \shift{\M}^\nu(s) = \M^\nu + \frac{s}{r} (\onev \eq - \I).
+\end{equation}
+%
+The first choice leaves \( \shift{\enc} = \enc \), so it only affects the \(\etw\) piece of \cref{eq:larea}.
+The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + (f\pot - d\dep) (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
+Either way:
+%
+\begin{equation}\label{eq:lareashift}
+  \hat{A}\prn{s; \M^\nu} = \hat{A}\prn{0; \shift{\M}^\nu(s)}.
+\end{equation}
+%
+In other words, the laplace transform of the SNR is the area under the SNR for the shifted process.
+
+Note that while \cref{eq:lshift} produces a valid continuous-time Markov process, \cref{eq:lshiftpd} does not always produce valid discrete-time Markov processes, as the diagonal elements can become negative.
+However, \(r (\shift{\M}^\nu - \I)\) is a valid continuous-time Markov process.
+
 In analogy with the case $s=0$, we define
@@
   \begin{aligned}
-    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)}, \\
-    a_j(s) &= \sum_{k} \enc_{jk} (\etw_{j}(s) - \etw_{k}(s)), \\
-    \thb_i(s) &= \sum_j \fptb_{ij}(0)\, \eq_j a_j(s), \\
+    \mathbf{c}_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)}, \\
+    \mathbf{a}_j(s) &= \sum_{k} \enc_{jk} (\etw_{j}(s) - \etw_{k}(s)), \\
+    \thb_i(s) &= \sum_j \fptb_{ij}(0)\, \eq_j \mathbf{a}_j(s), \\
   \end{aligned}
     \implies
-    \hat{A}(s) = \sum_k c_k(s) \eq_k \w_k
-      = \sum_i  \eq_i a_i(s).
+    \hat{A}(s) 
+      = \sum_k \eq_k \mathbf{c}_k(s) \w_k
+      % = \mathbf{c}^\dag(s) \w
+      = \eq \mathbf{a}(s).
 \end{equation}
@@
      = f^\mu \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
-     +\prn{rc_m(s) + (-1)^\mu} \brk{\etw_m(s) - \etw_n(s)} }.
+     +\prn{r \mathbf{c}_m(s) + (-1)^\mu} \brk{\etw_m(s) - \etw_n(s)} }.
 \end{equation}
@@
 \begin{alignedat}{3}
-    R_{ijkl} &= \
+    R_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frg_{ij} \circ \partial \frg_{kl}}} 
+    &= \
     &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
@@
           \brc{\etw_k(s) - \etw_l(s)} \\
-    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} 
+    &&&+ \eq_i \mathbf{c}_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} 
           \brc{\etw_k(s) - \etw_l(s)},  \\
-  S_{ijkl} &=
+  S_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frg_{ij} \partial \enc_{kl}} }
+  &= \
     &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
-    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
+    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)},
 \end{alignedat}
@@
 %
+where \(R\) contains only those terms where \( \frg_{ij} \) precedes \( \frg_{kl} \) in the matrix products from the expression for \( \hat{A} \) in \cref{eq:larea}.
+The other terms sum to \(R\trans\).
+
 Then we can write the Hessian as
@@
   \pdiff{^2\hat{A}(s)}{\M^\mu_{ij} \partial\M^\nu_{kl}} =
-     f^\mu f^\nu [R + R\trans + (-1)^\nu S + (-1)^\mu S\trans]_{ijkl}.
+     f^\mu f^\nu [R + R\trans + (-1)^\mu S\trans + (-1)^\nu S]_{ijkl}.
 \end{equation}
@@
   \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\etw_i(s) - \etw_j(s))
-        +\sum_{\nu i j} \prn{ f^\nu \mu^\nu_{ij} \M\pot_{ij} }
+        +\sum_{\nu i j} \prn{ f^\nu \mu^\nu_{ij} \M^\nu_{ij} }
         &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i\w_i },
@@
 where $\Delta\pr$ is the constant value we are holding $\eq\w$ at.
-After maximising
 The Karush-Kuhn-Tucker necessary conditions for a maximum are
@@
 %
+We have scaled the KKT multipliers by \(f^\nu\) for later convenience.
+
 The derivatives, for $m \neq n$ are
@@
     \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
-     +\prn{rc_m(s) + (-1)^\nu} \brk{\etw_m(s) - \etw_n(s)} } \\
+     + \prn{r \mathbf{c}_m(s) + (-1)^\nu} \brk{\etw_m(s) - \etw_n(s)} } \\
      &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     + \lambda \eq_m \sum_i (\fptb_{mi}(0)-\fptb_{ni}(0)) \eq_i\w_i.
-\end{aligned}
-\end{equation}
-%
-These have the same form as for $s=0$, which allowed us to argue that the model with maximal area must have the serial topology, \cite{Leibold2008serial,Ben-DayanRubin2007sparse}.
+     + r \lambda \eq_m (\etw_{m}(0)-\etw_{n}(0)) 
+    = 0.
+\end{aligned}
+\end{equation}
+%
+We can solve for the KKT multipliers by noting that
+%
+\begin{equation*}
+  \mu^\nu_{ij} \M^\nu_{ij} = 0,
+  \qquad \text{and} \qquad
+  \sum_{j: j \neq i} \mu^\nu_{ii} \M^\nu_{ij} 
+      = \mu^\nu_{ii} (1 - \M^\nu_{ii}) = \mu^\nu_{ii}.
+\end{equation*}
+%
+Then, with some new vectors defined:
+%
+\begin{equation}\label{eq:kktsol}
+\begin{aligned}
+  \mu^\nu_{mm} &= - \eq_m \prn{ r \alb^\nu_m(s) 
+    + r \lambda \omb^\nu_i(0) 
+    + \prn{r \mathbf{c}_m(s) + (-1)^\nu} \omb^\nu_m(s)
+    }, \\
+  \mu^\nu_{mn} &= \eq_m \big( 
+      r \brk{ \thb_n(s) - \thb_m(s) - \alb^\nu_m(s) }
+    + r \lambda \brk{ \etw_n(0) - \etw_m(0) - \omb^\nu_m(0) } 
+  \\ &\phantom{= - \eq_m \big( }
+    + \brk{rc_m(s) + (-1)^\nu} 
+        \brk{ \etw_n(s) - \etw_m(s) - \omb^\nu_m(s) }
+     \big), \\
+  &\text{where} \quad 
+  \boldsymbol{\alpha}^\nu(s) = (\M^\nu - \I) \thb(s), \qquad
+  \boldsymbol{\omega}^\nu(s) = (\M^\nu - \I) \etw(s).
+\end{aligned}
+\end{equation}
+%
+
+The expressions in \cref{eq:finitederivs} have the same form as for $s=0$, which allowed us to argue that the model with maximal area must have the serial topology, \cite{Lahiri2013synapse}.
 
@@
 \begin{equation}\label{eq:scale}
-  \lambda * \M = (1-\lambda)\I + \lambda \M.
+  \lambda * \M^\nu \equiv (1-\lambda)\I + \lambda \M^\nu.
 \end{equation}
@@
 %
-\begin{equation}\label{eqlaplacescale}
-  A(\lambda s;\lambda * \M^\nu) = A(s;\M^\nu).
+\begin{equation}\label{eq:laplacescale}
+  \hat{A} \prn{\lambda s;\lambda * \M^\nu} = \hat{A} \prn{s; \M^\nu}.
 \end{equation}
@@
 We can't do that for $s>0$.
+
+The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
+%
+\begin{equation}\label{eq:shiftscale}
+  \lambda {\circledast}_{s} \M \equiv (1 - \lambda) \I + \lambda \M^\nu
+                + \frac{(\lambda - 1) s}{r} (\onev \eq - \I).
+\end{equation}
+%
+Then we have
+%
+\begin{equation}\label{eq:lshiftscale}
+  \hat{A} \prn{s; \M^\nu} 
+    = \hat{A} \prn{0, \shift{\M}^\nu(s)}
+    = \hat{A} \prn{0, \lambda * \shift{\M}^\nu(s)}
+    = \hat{A} \prn{s, \lambda {\circledast}_{s} \M^\nu}.
+\end{equation}
+%
+The problem here is that the lower bound, \(\M^\nu_{ij} \geq 0\), is not invariant under this transformation, unlike the \(s = 0\) case.
+Using this transformation to fix the upper bound can break the lower bound.
 
@@
 \end{figure}
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Shifted problem}\label{sec:lshift}
+
+The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
+treat the off-diagonal components of \(\shift{\M}^\nu \) as the degrees of freedom and maximise the area instead.
+The constraints on \(\shift{\M}^\nu_{ij}\) are different from \cref{eq:constri}:
+%
+\begin{equation}\label{eq:lshiftconstr}
+  \shift{\M}^\nu_{ij} \geq \frac{s \eq_j}{r},
+    \quad \forall \; \nu, i, j : i \neq j, \qquad
+  \sum_{j: j \neq i} \shift{\M}^\nu_{ij} \leq 1 + \frac{s}{r} (1 - \eq_i).
+    \quad \forall \; \nu, i.
+\end{equation}
+%
+The diagonal elements are still given by \( \shift{\M}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\M}^\nu_{ij} \), but now they are allowed to be negative.
+
+The Lagrangian we should consider is
+%
+\begin{equation}\label{eq:lshiftlagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\etw_i - \etw_j)
+        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\M^\nu_{ij} -  \frac{s}{r} (\eq_j - \delta_{ij}) }
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i\w_i },
+\end{aligned}
+\end{equation}
+%
+where all of the first-passage times are evaluated at \(s = 0\).
+
+The derivatives, for $m \neq n$ are
+%
+\begin{equation}\label{eq:lshiftderivs}
+\begin{aligned}
+  \frac{1}{f^\nu} \pdiff{\CL}{\Mh^\nu_{mn}} = & \,
+    \eq_m \prn{ r \brk{\thb_m - \thb_n}
+     + \prn{r \mathbf{c}_m + (-1)^\nu + r \lambda} \brk{\etw_m - \etw_n} } \\
+     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
+     - \sum_{ij} \mu^\nu_{ij} s f^\nu \eq_m (\fptb_{mj} - \fptb_{nj}) \eq_j
+    = 0.
+\end{aligned}
+\end{equation}
+%
+We can absorb the new terms by defining
+%
+\begin{equation*}
+\begin{aligned}
+  \shift{\thb}^\nu(s) = \thb(0) - \frac{s f^\nu}{r} \fptb \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
+  \qquad
+  \shift{\alb}^\nu(s) &= (\M^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
+  \\
+  \shift{\omb}^\nu(s) &= (\M^\nu + (s - 1) \I - s \onev \eq) \etw(0).
+\end{aligned}
+\end{equation*}
+%
+We note that
+%
+\begin{equation*}
+  \mu^\nu_{ij} (\M^\nu_{ij} - s \eq_j) = 0,
+  \qquad \text{and} \qquad
+  \sum_{j: j \neq i} \mu^\nu_{ii} (\M^\nu_{ij} - s \eq_j)
+      = \mu^\nu_{ii} (1 - \M^\nu_{ii} - s (1 - \eq_i)) = \mu^\nu_{ii},
+\end{equation*}
+%
+then we can ``solve'' for the KKT multipliers:
+%
+\begin{equation}\label{eq:lshiftsol}
+\begin{aligned}
+  \mu^\nu_{mm} &= - \eq_m \prn{ r \shift{\alb}^\nu_m(s) 
+    + \brk{ r \mathbf{c}_m + (-1)^\nu + r \lambda } \shift{\omb}^\nu_m(s)
+    }, \\
+  \mu^\nu_{mn} &= \eq_m \prn{ 
+      r \brk{ \shift{\thb}_n(s) - \shift{\thb}_m(s) - \shift{\alb}^\nu_m(s) }
+  % \\ & \phantom{= - \eq_m \big( } 
+    + \brk{ r \mathbf{c}_m + (-1)^\nu + r \lambda } 
+        \brk{ \etw_n + \etw_m - \shift{\omb}^\nu_m(s) }
+     }, \\
+\end{aligned}
+\end{equation}
+%
+although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\).
+
+While this problem involves the scale invariant quantity \(\hat{A}(0)\), neither the upper nor lower bounds in \cref{eq:lshiftconstr} are invariant.
+We then run into the same problems in proving our serial-topology hypothesis.
 

### 2020-07-15 Code/Python/complex_synapse/synapse_opt.py

@@ -43,8 +43,7 @@ class SynapseOptModel(_SynapseMemoryModel):
     @property
     def nparam(self) -> int:
         """number of plasticity types."""
-        npr = mp.num_param(self.nstates, **self.directed(0))
-        return self.nplast * npr
+        return self.nplast * mp.num_param(self.nstates, **self.directed(0))
 
     def get_params(self) -> la.lnarray:
         """Independent parameters of transition matrices.

### 2020-07-22 Notes/laplace.tex
numeric optimised shifted problem

@@
   % Requires \usepackage{graphicx}
-  \includegraphics[width=0.8\linewidth]{LenvProven.eps}
+  \includegraphics[width=0.8\linewidth]{LenvProven.pdf}
   \caption[Proven envelope for the signal-to-noise ratio]
@@
 We can maximise the ration \wrt $\eq\w$ later.
+From this point on, we work in units where \(r = 1\).
+
 Consider the Lagrangian
@@
         +\sum_{\nu i j} \prn{ f^\nu \mu^\nu_{ij} \M^\nu_{ij} }
-        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i\w_i },
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i \w_i },
 \end{aligned}
@@
   \frac{1}{f^\nu} \pdiff{\CL}{\Mh^\nu_{mn}} = & \,
-    \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
-     + \prn{r \mathbf{c}_m(s) + (-1)^\nu} \brk{\etw_m(s) - \etw_n(s)} } \\
+    \eq_m \prn{ \brk{\thb_m(s) - \thb_n(s)}
+     + \prn{\mathbf{c}_m(s) + (-1)^\nu} \brk{\etw_m(s) - \etw_n(s)} } \\
      &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     + r \lambda \eq_m (\etw_{m}(0)-\etw_{n}(0)) 
+     + \lambda \eq_m (\etw_{m}(0)-\etw_{n}(0)) 
     = 0.
@@
 \begin{aligned}
-  \mu^\nu_{mm} &= - \eq_m \prn{ r \alb^\nu_m(s) 
-    + r \lambda \omb^\nu_i(0) 
-    + \prn{r \mathbf{c}_m(s) + (-1)^\nu} \omb^\nu_m(s)
+  \mu^\nu_{mm} &= - \eq_m \prn{ \alb^\nu_m(s) 
+    + \lambda \omb^\nu_i(0) 
+    + \prn{\mathbf{c}_m(s) + (-1)^\nu} \omb^\nu_m(s)
     }, \\
   \mu^\nu_{mn} &= \eq_m \big( 
-      r \brk{ \thb_n(s) - \thb_m(s) - \alb^\nu_m(s) }
-    + r \lambda \brk{ \etw_n(0) - \etw_m(0) - \omb^\nu_m(0) } 
-  \\ &\phantom{= - \eq_m \big( }
-    + \brk{rc_m(s) + (-1)^\nu} 
+      \brk{ \thb_n(s) - \thb_m(s) - \alb^\nu_m(s) }
+    + \lambda \brk{ \etw_n(0) - \etw_m(0) - \omb^\nu_m(0) } 
+  \\ &\phantom{= \eq_m \big( }
+    + \brk{\mathbf{c}_m(s) + (-1)^\nu} 
         \brk{ \etw_n(s) - \etw_m(s) - \omb^\nu_m(s) }
@@
 \begin{equation}\label{eq:shiftscale}
-  \lambda {\circledast}_{s} \M \equiv (1 - \lambda) \I + \lambda \M^\nu
-                + \frac{(\lambda - 1) s}{r} (\onev \eq - \I).
+  \lambda {\circledast}_{s} \M^\nu 
+      \equiv (1 - \lambda) \brk{(1 + s)\I - s \onev \eq} 
+      + \lambda \M^\nu.
 \end{equation}
@@
 However, we do have numerical evidence that the model that maximises the Laplace transform does have the serial topology.
-In \autoref{fig:envnum} we see the result of numerical maximisation of \eqref{eq:finitelagrangian}.
+In \cref{fig:envnum} we see the result of numerical maximisation of \cref{eq:finitelagrangian}.
 We see that restricting to the serial topology does not make the performance any worse.
@@
   % Requires \usepackage{graphicx}
-  \includegraphics[width=0.8\linewidth]{LenvNum.svg}\\
+  \includegraphics[width=0.8\linewidth]{LenvNum.pdf}\\
   \caption[Numerical envelope for the signal-to-noise ratio]
   {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
-  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numericallly with $f\potdep=\frac{1}{2}$, the number of synapses $N=10^4$ and the number of states $M=12$.
-  The dashed red line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
-  The solid red line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
-  The solid grren line indicates the proven envelope, \eqref{eq:env}.}\label{fig:envnum}
+  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numericallly with $f\potdep=\frac{1}{2}$, the number of synapses $N=1$ and the number of states $M=10$.
+  The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
+  The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
+  The blue line indicates the proven envelope, \eqref{eq:env}.}\label{fig:envnum}
 \end{figure}
@@
 \begin{equation}\label{eq:lshiftconstr}
-  \shift{\M}^\nu_{ij} \geq \frac{s \eq_j}{r},
+  \shift{\M}^\nu_{ij} \geq s \eq_j,
     \quad \forall \; \nu, i, j : i \neq j, \qquad
-  \sum_{j: j \neq i} \shift{\M}^\nu_{ij} \leq 1 + \frac{s}{r} (1 - \eq_i).
+  \sum_{j: j \neq i} \shift{\M}^\nu_{ij} \leq 1 + s (1 - \eq_i).
     \quad \forall \; \nu, i.
@@
   \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\etw_i - \etw_j)
-        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\M^\nu_{ij} -  \frac{s}{r} (\eq_j - \delta_{ij}) }
-        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i\w_i },
+        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\M^\nu_{ij} -  s (\eq_j - \delta_{ij}) }
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i \w_i },
 \end{aligned}
@@
 
+In \cref{fig:shifted} we show the result of solving this problem with numerical optimisation.
+We see essentially the same results as the original problem, \cref{eq:finitelagrangian,fig:envnum}, as expected.
+
 The derivatives, for $m \neq n$ are
@@
   \frac{1}{f^\nu} \pdiff{\CL}{\Mh^\nu_{mn}} = & \,
-    \eq_m \prn{ r \brk{\thb_m - \thb_n}
-     + \prn{r \mathbf{c}_m + (-1)^\nu + r \lambda} \brk{\etw_m - \etw_n} } \\
+    \eq_m \prn{ \brk{\thb_m - \thb_n}
+     + \prn{\mathbf{c}_m + (-1)^\nu + \lambda} \brk{\etw_m - \etw_n} } \\
      &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
@@
 \begin{aligned}
-  \shift{\thb}^\nu(s) = \thb(0) - \frac{s f^\nu}{r} \fptb \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
+  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
   \qquad
@@
 \begin{aligned}
-  \mu^\nu_{mm} &= - \eq_m \prn{ r \shift{\alb}^\nu_m(s) 
-    + \brk{ r \mathbf{c}_m + (-1)^\nu + r \lambda } \shift{\omb}^\nu_m(s)
+  \mu^\nu_{mm} &= - \eq_m \prn{ \shift{\alb}^\nu_m(s) 
+    + \brk{ \mathbf{c}_m + (-1)^\nu + \lambda } \shift{\omb}^\nu_m(s)
     }, \\
   \mu^\nu_{mn} &= \eq_m \prn{ 
-      r \brk{ \shift{\thb}_n(s) - \shift{\thb}_m(s) - \shift{\alb}^\nu_m(s) }
+      \brk{ \shift{\thb}^\nu_n(s) - \shift{\thb}^\nu_m(s) - \shift{\alb}^\nu_m(s) }
   % \\ & \phantom{= - \eq_m \big( } 
-    + \brk{ r \mathbf{c}_m + (-1)^\nu + r \lambda } 
+    + \brk{ \mathbf{c}_m + (-1)^\nu + \lambda } 
         \brk{ \etw_n + \etw_m - \shift{\omb}^\nu_m(s) }
@@
 We then run into the same problems in proving our serial-topology hypothesis.
+
+\begin{figure}[ht]
+\begin{center}
+\begin{myenuma}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{LenvShift.pdf}}
+  \label{fig:envshift}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{shift_cond.pdf}}
+  \label{fig:shiftcond}
+\end{myenuma}
+\caption[Shifted problem vs.\ original]{Shifted problem vs.\ original.
+(\ref*{fig:envshift}) Comparison of numerical optimisation of the shifted problem in \cref{eq:lshiftlagrangian,fig:envnum} with the original problemin \cref{eq:finitelagrangian}.
+We see that the two approaches produce very similar numerical envelopes.
+The shifted problem involves non-linear constraints and is thus harder to solve.
+(\ref{fig:shiftcond}) Plot of the discrepancy in (\ref*{fig:envshift}) against the condition number of the fundamental matrix, \(\fund \).
+We see that most of the large discrepancies occur where the problem is numerically ill-conditioned.
+\label{fig:shifted}}
+\end{center}
+\end{figure}
 
@@
   % Requires \usepackage{graphicx}
-  \includegraphics[width=0.8\linewidth]{LenvHeuristic.svg}\\
+  \includegraphics[width=0.8\linewidth]{LenvHeuristic.pdf}\\
   \caption[Heuristic envelope for the signal-to-noise ratio]

### 2020-07-22 Code/Python/complex_synapse/__init__.py

@@ -4,8 +4,7 @@ Spyder Editor
 
 This is a temporary script file.
 """
-from . import builders
+from . import builders, optimise, sticky, shorten
 from .synapse_memory_model import SynapseMemoryModel
 from .synapse_opt import SynapseOptModel
-from . import optimise
 assert any((True, builders, SynapseMemoryModel, SynapseOptModel, optimise))

### 2020-07-22 Code/Python/complex_synapse/optimise.py

@@ -1,6 +1,8 @@
 # -*- coding: utf-8 -*-
 """Optimising synapse modelopts
 """
+from __future__ import annotations
+
 from functools import wraps
 from numbers import Number
 from typing import Callable, List, Optional, Tuple, TypeVar, Union
@@ -18,12 +20,6 @@ from sl_py_tools.numpy_tricks.markov import params as mp
 from . import builders as bld
 from .synapse_opt import SynapseOptModel
 
-Data = TypeVar('Data', Number, np.ndarray)
-Func = Callable[[np.ndarray], Data]
-Constraint = Union[sco.LinearConstraint, sco.NonlinearConstraint]
-Problem = Tuple[Func[Number], Func[np.ndarray], Func[np.ndarray],
-                sco.Bounds, List[Constraint]]
-Maker = Callable[[SynapseOptModel, Number, bool, bool], Problem]
 # =============================================================================
 # Problem creation helper functions
 # =============================================================================
@@ -210,6 +206,8 @@ def normal_problem(model: SynapseOptModel, rate: Number, inv: bool = False,
                    keep_feasible: bool = False, **kwds) -> Problem:
     """Make an optimize problem.
     """
+    svd = kwds.pop('svd', False)
+    kwds['cond'] = not svd
     fun = make_fn(model, model.laplace_fun, rate, inv=inv, **kwds)
     jac = make_fn(model, model.laplace_grad, rate, inv=inv, **kwds)
     hess = make_fn(model, model.laplace_hess, rate, **kwds) if inv else None
@@ -217,7 +215,7 @@ def normal_problem(model: SynapseOptModel, rate: Number, inv: bool = False,
     con_coeff = constraint_coeff(model)
     bounds = sco.Bounds(0, 1, keep_feasible)
     diag = [sco.LinearConstraint(con_coeff, -np.inf, 1, keep_feasible)]
-    if kwds.get('svd', False):
+    if svd:
         diag.append(cond_limit(model, rate, keep_feasible, inv=inv))
 
     return fun, jac, hess, bounds, diag
@@ -235,6 +233,8 @@ def shifted_problem(model: SynapseOptModel, rate: Number, inv: bool = False,
     if any(SynapseOptModel.type.values()):
         raise ValueError('Shifted problem cannot have special topology')
 
+    svd = kwds.pop('svd', False)
+    kwds['cond'] = not svd
     fun = make_fn(model, model.area_fun, inv=True, **kwds)
[…]

### 2020-07-22 Code/Python/complex_synapse/synapse_opt.py

@@ -228,7 +228,7 @@ class SynapseOptModel(_SynapseMemoryModel):
         func : float
             Value of ``snr_laplace`` at ``s``.
         """
-        if not well_behaved(self, rate):
+        if not well_behaved(self, rate, kwds.pop('cond', True)):
             return np.array(1.e10)
         # (p,c), (eta,theta)
         rows = self._derivs(rate, **kwds)[0]
@@ -253,7 +253,7 @@ class SynapseOptModel(_SynapseMemoryModel):
         grad : la.lnarray (2n(n-1),)
             Gradient of ``snr_laplace`` at ``s`` with respect to parameters.
         """
-        if not well_behaved(self, rate):
+        if not well_behaved(self, rate, kwds.pop('cond', True)):
             return bld.RNG.random(self.nparam)
         # (p,c), (eta,theta)
         rows, cols, _ = self._derivs(rate, **kwds)
@@ -281,6 +281,8 @@ class SynapseOptModel(_SynapseMemoryModel):
         hess : la.lnarray (2n(n-1),2n(n-1))
             Hessian of ``snr_laplace`` at ``s`` with respect to parameters.
         """
+        if not well_behaved(self, rate, kwds.pop('cond', True)):
+            return bld.RNG.random((self.nparam,) * 2)
         # (p,c), (eta,theta), (Z,Zs,ZQZs)
         kwds['inv'] = True
         rows, cols, mats = self._derivs(rate, **kwds)
@@ -323,6 +325,8 @@ class SynapseOptModel(_SynapseMemoryModel):
             Change in parameters matrix-multiplied by hessian of
             ``snr_laplace`` at ``s`` with respect to parameters.
         """
+        if not well_behaved(self, rate, kwds.pop('cond', True)):
+            return bld.RNG.random(self.nparam)
         # (p,c), (eta,theta), (Zi,Zis)
         kwds.setdefault('inv', False)
         rows, cols, mats = self._derivs(rate, **kwds)
@@ -402,7 +406,7 @@ class SynapseOptModel(_SynapseMemoryModel):
         func : la.lnarray (2n**2,)
             Value of ``W - s * e @ peq``.
         """
-        if not well_behaved(self, rate):
+        if not well_behaved(self, rate, kwds.pop('cond', True)):
             return -np.ones(self.nplast*self.nstates**2)
         rate = rate / (2 * self.frac.s)
         # (p,c), (eta,theta), (Z,Zs,ZQZs)
[…]

### 2020-07-25 Notes/laplace.tex
heuristic envelope functions

@@
 \newcommand{\Lb}{\boldsymbol{\Lambda}}
+\newcommand{\lb}{\boldsymbol{\lambda}}
 \newcommand{\etw}{\boldsymbol{\eta}^w}
@@
 The expressions in \cref{eq:finitederivs} have the same form as for $s=0$, which allowed us to argue that the model with maximal area must have the serial topology, \cite{Lahiri2013synapse}.
-
 The one missing ingredient is the scale invariance.
-Define the scale transformation:
-%
-\begin{equation}\label{eq:scale}
-  \lambda * \M^\nu \equiv (1-\lambda)\I + \lambda \M^\nu.
-\end{equation}
-%
-The Laplace transform has the following scaling property
-%
-\begin{equation}\label{eq:laplacescale}
-  \hat{A} \prn{\lambda s;\lambda * \M^\nu} = \hat{A} \prn{s; \M^\nu}.
-\end{equation}
-%
-At $s=0$, it is scale invariant.
-This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
-This allowed us to ignore the diagonal constraints.
-We can't do that for $s>0$.
-
-The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
-%
-\begin{equation}\label{eq:shiftscale}
-  \lambda {\circledast}_{s} \M^\nu 
-      \equiv (1 - \lambda) \brk{(1 + s)\I - s \onev \eq} 
-      + \lambda \M^\nu.
-\end{equation}
-%
-Then we have
-%
-\begin{equation}\label{eq:lshiftscale}
-  \hat{A} \prn{s; \M^\nu} 
-    = \hat{A} \prn{0, \shift{\M}^\nu(s)}
-    = \hat{A} \prn{0, \lambda * \shift{\M}^\nu(s)}
-    = \hat{A} \prn{s, \lambda {\circledast}_{s} \M^\nu}.
-\end{equation}
-%
-The problem here is that the lower bound, \(\M^\nu_{ij} \geq 0\), is not invariant under this transformation, unlike the \(s = 0\) case.
-Using this transformation to fix the upper bound can break the lower bound.
 
@@
 
+\subsection{Scale invariance}\label{sec:scale}
+
+Define the scale transformation:
+%
+\begin{equation}\label{eq:scale}
+  \zeta * \M \equiv \zeta \M + (1-\zeta)\I .
+\end{equation}
+%
+The Laplace transform has the following scaling property
+%
+\begin{equation}\label{eq:laplacescale}
+  \hat{A} \prn{\zeta s;\zeta * \M^\nu} = \hat{A} \prn{s; \M^\nu}.
+\end{equation}
+%
+At $s = 0$, it is scale invariant.
+This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
+This allowed us to ignore the diagonal constraints.
+We can't do that for $s > 0$.
+
+The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
+%
+\begin{equation}\label{eq:shiftscale}
+  \zeta {\circledast}_{s} \M 
+      \equiv \zeta \M
+      + (1 - \zeta) \brk{(1 + s)\I - s \onev \eq} .
+\end{equation}
+%
+Then we have
+%
+\begin{equation}\label{eq:lshiftscale}
+  \hat{A} \prn{s; \M^\nu} 
+    = \hat{A} \prn{0, \shift{\M}^\nu(s)}
+    = \hat{A} \prn{0, \zeta * \shift{\M}^\nu(s)}
+    = \hat{A} \prn{s, \zeta \circledast_s \M^\nu}.
+\end{equation}
+%
+The problem here is that the lower bound, \(\M^\nu_{ij} \geq 0\), 
+is not invariant under this transformation, unlike the \(s = 0\) case.
+Using this transformation to fix the upper bound can break the lower bound.
+
+We might ask what conditions are required so that a model is fixable?
+This means that the \(\zeta \circledast_s \) orbit of the model contains some valid models.
+The conditions are
+%
+\begin{equation*}
+\begin{alignedat}{5}
+  [\zeta \circledast \M^\nu]_{ij}  
+    &= \zeta \M^\nu_{ij} - (1-\zeta) s \eq_j 
+    &&\geq 0 &
+  \quad &\implies \quad &
+  \zeta &\geq \frac{s \eq_j}{\M^\nu_{ij} + s \eq_j} &
+    &\forall \nu, i, j : i \neq j , \\
+  \sum_{j : i \neq j} [\zeta \circledast \M^\nu]_{ij} 
+    &= \zeta \lb^\nu_i - (1-\zeta) s (1-\eq_i) 
+    &&\leq 1 &
+  &\implies &
+  \zeta &\leq \frac{1 + s (1-\eq_i)}{\lb^\nu_i + s (1-\eq_i)} &
+    \quad &\forall \nu, i ,
+\end{alignedat}
+\end{equation*}
+%
+where \(\lb^\nu_i \equiv \sum_{j : i \neq j} \M^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
+The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
+%
+\begin{equation}\label{eq:fixable}
+  \frac{s \eq_j}{\M^\nu_{ij} + s \eq_j} \leq \frac{1 + s (1-\eq_k)}{\lb^\mu_k + s (1-\eq_k)}
+  \quad \implies \quad
+  \frac{\lb^\mu_k - 1}{1 + s (1-\eq_k)} \leq \frac{\M^\nu_{ij}}{s \eq_j} 
+  \quad \forall \mu, \nu, i, j, k : i \neq j.
+\end{equation}
+%
+Note that when \(\M^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lb^\mu_k \leq 1\), the original diagonal constraint.
+Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
+
+%-------Section------------------------------------------------------------------
+
 \subsection{Shifted problem}\label{sec:lshift}
@@
 %
-The diagonal elements are still given by \( \shift{\M}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\M}^\nu_{ij} \), but now they are allowed to be negative.
+The diagonal elements are still given by \( \shift{\M}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\M}^\nu_{ij} \), but now they are allowed to be negative so long as \(\M^\nu_{ii} \geq - s (1 - \eq_i)\).
 
@@
     + \brk{ \mathbf{c}_m + (-1)^\nu + \lambda } 
-        \brk{ \etw_n + \etw_m - \shift{\omb}^\nu_m(s) }
+        \brk{ \etw_n - \etw_m - \shift{\omb}^\nu_m(s) }
      }, \\
@@
   s = S(\beta) \equiv 2 \sinh^2 \prn{\frac{\beta}{2}}
-    = \cosh (m\beta) - 1,
+    = \cosh (\beta) - 1,
 \end{equation}
@@
 This interpolates between \eqref{eq:unilaplace} at $\varepsilon=0$ and the same expression for $m-1$ at $\varepsilon=1$.
+
+This expression breaks down for \(M = 2\), as there are no \( 2 \leq i \leq m \) when \(m = 1\).
+In that case, 
+%
+\begin{equation}\label{eq:binaryshorten}
+  \hat{A}(s) = \frac{1-\varepsilon}{1-\varepsilon + s}.
+\end{equation}
+%
+We can see that the optimal \(\varepsilon = 0\), reducing \cref{eq:binaryshorten} to \cref{eq:binaryenv}.
 

### 2020-07-25 Code/Python/complex_synapse/sticky.py

@@ -5,18 +5,16 @@ from typing import Tuple
 import numpy as np
 import scipy.optimize as sco
 
-from .shorten import Y_STAR
+from .shorten import Y_STAR, Values
+# =============================================================================
 
-Arrays = Tuple[np.ndarray, ...]
 
-
-# -------------------------------------
-def beta_to_s(beta: float) -> float:
+def beta_to_s(beta: Values) -> Values:
     """Convert alpha to s"""
     return np.cosh(beta) - 1
 
 
-def s_to_beta(sval: float) -> float:
+def s_to_beta(sval: Values) -> Values:
     """Convert s to beta. positive one of two possibilities"""
     return np.arccosh(sval + 1)
 
@@ -30,7 +28,7 @@ def beta_star(num: int) -> float:
 # =============================================================================
 
 
-def components(beta: float, num: int) -> Tuple[float, ...]:
+def _components(beta: Values, num: int) -> Tuple[Values, ...]:
     """Components of numerator and denominator
 
     numerator = numer - eps * dnumer
@@ -44,27 +42,22 @@ def components(beta: float, num: int) -> Tuple[float, ...]:
 
 
 # -------------------------------------
-def sticky(eps: float, beta: float, num: int) -> float:
+def sticky(eps: Values, beta: Values, num: int) -> Values:
     """Laplace-SNR for sticky serial model"""
     sval = beta_to_s(beta)
-    numer, denom, dnumer, ddenom = components(beta, num)
+    numer, denom, dnumer, ddenom = _components(beta, num)
     fnumer = numer - eps * dnumer
     fdenom = denom - eps * ddenom
     fmu = num - eps * (num - 2)
     return 2 * (1 - eps) * fnumer / (fmu * sval * fdenom)
 
 
-def neg_sticky(eps: float, alpha: float, num: int) -> float:
-    """negative Laplace-SNR for sticky serial model"""
-    return - sticky(eps, np.log(alpha), num)
-
-
 # -------------------------------------
-def eps_stars(beta: float, num: int) -> float:
+def eps_stars(beta: Values, num: int) -> Tuple[Values, Values]:
     """Optimal epsilon for sticky serial model, raw"""
     if num == 2:
         return 0, 0
-    numer, denom, dnumer, ddenom = components(beta, num)
+    numer, denom, dnumer, ddenom = _components(beta, num)
     fnumer = numer / dnumer
[…]

### 2020-08-05 Notes/Identification.tex
basics for identification in place

@@
 The different types of plasticity (\eg potentiation and depression) are indicated by superscripts $\mu,\nu$.
-The type of plasticity used between times $(t-1,t)$ is $\mu(t)$.
-An entire sequence of plasticity types is denoted by $\mu[1,T]$.
+The type of plasticity used between times $(t,t+1)$ is $\mu(t)$.
+An entire sequence of plasticity types is denoted by $\mu[0,T-1]$.
 
@@
 \begin{equation}\label{eq:like}
-  \pr\cond{\w^A[0,T]}{\M} = \prod_a \pr\cond{\w^a[0,T]}{\M}.
+\begin{aligned}
+  \pr^A_{[0,T]}\cond{\w}{\mu,\M}
+    &\equiv \pr\cond{\w^A[0,T]}{\mu^A[0,T-1],\M}  \\
+    &= \prod_a \pr\cond{\w^a[0,T]}{\mu^a[0,T-1],\M}
+      \equiv \prod_a \pr^a_{[0,T]}\cond{\w}{\mu,\M}.
+\end{aligned}
 \end{equation}
@@
 \begin{equation}\label{eq:emq}
-  Q(\M,\M') = \sum_{i^A[0,T]} \pr\cond{i^A[0,T],\w^A[0,T]}{\M} \log\!\! \brk{
-      \frac{ \pr\cond{i^A[0,T],\w^A[0,T]}{\M'} }{ \pr\cond{i^A[0,T],\w^A[0,T]}{\M} } }.
+  Q(\M,\M') = \sum_{{i^A[0,T]}} 
+      \pr^A_{[0,T]}\cond{i,\w}{\mu,\M} \log\!\! \brk{
+      \frac{ \pr^A_{[0,T]}\cond{i,\w}{\mu,\M'} }
+           { \pr^A_{[0,T]}\cond{i,\w}{\mu,\M} } }.
 \end{equation}
@@
 \begin{equation}\label{eq:emqseveral}
-\begin{gathered}
-  Q(\M,\M') = \sum_{i^A[0,T]}  \prod_b \pr\cond{i^b[0,T],\w^b[0,T]}{\M} \sum_a  \log\!\!  \brk{
-      \frac{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } } \\
-   = \sum_a \prod_{b \neq a} \pr\cond{\w^b[0,T]}{\M}\! \sum_{i^a[0,T]}\!\! \pr\cond{i^a[0,T],\w^a[0,T]}{\M} \log\!\!  \brk{
-      \frac{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } } \\
-   = \pr\cond{\w^A[0,T]}{\M} \sum_a \frac{Q^a(\M,\M')}{\pr\cond{\w^a[0,T]}{\M}}.
-\end{gathered}
+\begin{aligned}
+  Q(\M,\M') &= \sum_{i^A[0,T]} \prod_b 
+    \pr^b_{[0,T]}\cond{i,\w}{\mu,\M} \sum_a  \log\!\!  \brk{
+      \frac{ \pr^a_{[0,T]}\cond{i,\w}{\mu,\M'} }
+           { \pr^a_{[0,T]}\cond{i,\w}{\mu,\M} } } \\
+   &= \sum_a \prod_{b \neq a} 
+   \pr^b_{[0,T]}\cond{\w}{\mu,\M}\! 
+   \sum_{i^a[0,T]}\!\! \pr^a_{[0,T]}\cond{i,\w}{\mu,\M} \log\!\!  \brk{
+      \frac{ \pr^a_{[0,T]}\cond{i,\w}{\mu,\M'} }
+           { \pr^a_{[0,T]}\cond{i,\w}{\mu,\M} } } \\
+   &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \sum_a 
+      \frac{Q^a(\M,\M')}{\pr^a_{[0,T]}\cond{\w}{\mu,\M}}.
+\end{aligned}
 \end{equation}
@@
 \begin{aligned}
-  \pr\cond{t[0,T], \w[0,T]}{\M} &= \pib_{i(0)} \, \Pib_{i(0)}(0) 
-        \prod_{t=1}^{T} \M^{\mu(t)}_{i(t-1)i(t)} \, \Pib_{i(t)}(t),\\
+  \pr\cond{i[0,T], \w[0,T]}{\mu[0,T-1],\M} &= \pib_{i(0)} \, \Pib_{i(0)}(0) 
+        \prod_{t=0}^{T-1} \M^{\mu(t)}_{i(t)i(t+1)} \, \Pib_{i(t+1)}(t+1),\\
   \text{where}\quad
@@
 \begin{aligned}
-  N^\nu_{jk} &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \pr\cond{i[t-1,t]=(jk),\w[0,T]}{\M} \\
-     &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \alpha_j(t-1) \, \beta_k(t) \, \M^{\nu}_{jk} \, \Pib_k(t) ,\\
-  \alpha_j(t) &= \pr\cond{i(t)=j,\w[0,t]}{\M} \\
-     &= \sum_{i[0,t]} \pib_{i(0)}(0) \, \Pib_{i(0)}(0) \prod_{s=1}^{t} \M^{\mu(s)}_{i(s-1)i(s)} \, \Pib_{i(s)}(s) \,\delta_{i(t)j}, \\
-  \beta_j(t) &= \pr\cond{i(t)=j,\w[t+1,T]}{\M} \\
-     &= \sum_{i[t,T]} \delta_{i(t)j} \prod_{s=t+1}^{T} \M^{\mu(s)}_{i(s-1)i(s)} \, \Pib_{i(s)}(s).
+  N^\nu_{jk} &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, 
+        \pr\cond{i[t,t+1]=(j,k),\w[0,T]}{\mu[0,T-1],\M} \\
+     &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, 
+        \alpha_j(t) \, \beta_k(t+1) \, \M^{\nu}_{jk} \, \Pib_k(t+1) ,\\
+  \alpha_j(t) &= \pr\cond{i(t)=j,\w[0,t]}{\mu[0,t-1],\M} \\
+     &= \sum_{i[0,t]} \pib_{i(0)}(0) \, \Pib_{i(0)}(0) \prod_{s=0}^{t-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib_{i(s+1)}(s+1) \,\delta_{i(t)j}, \\
+  \beta_j(t) &= \pr\cond{\w[t+1,T]}{i(t)=j,\mu[t,T-1],\M} \\
+     &= \sum_{i[t,T]} \delta_{i(t)j} \prod_{s=t}^{T-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib_{i(s+1)}(s+1).
 \end{aligned}
@@
 \begin{aligned}
-  \tilde{\alpha}_j(t) &= \pr\cond{i(t)=j}{\w[0,t],\M}, \\
-  \eta(t) &= \pr\cond{\w(t)}{\w[0,t-1],\M}, \\
-  \prod_{s=0}^t \eta(t) &= \pr\cond{\w[0,t]}{\M}.
+  \tilde{\alpha}_j(t) &= \pr\cond{i(t)=j}{\w[0,t],\mu[0,t-1],\M}, \\
+  \eta(t)\inv &= \pr\cond{\w(t)}{\w[0,t-1],\mu[0,t-1],\M}, \\
+  \brk{\prod_{s=0}^t \eta(s)}\inv &= \pr\cond{\w[0,t]}{\mu[0,t-1],\M}, \\
+  \brk{\prod_{s=t_0}^{t_1} \eta(s)}\inv &= 
+          \pr\cond{\w[t_0,t_1]}{\w[0,t_0-1],\mu[0,t_1-1],\M}.
 \end{aligned}
@@
 %
-Note that the new backward variables $\tilde{\beta}$ are not normalised and have no particularly interesting interpretation.
-However:
+Note that the new backward variables $\tilde{\beta}$ are not normalised and have no particularly interesting interpretation, beyond the following:
 %
@@
 \begin{aligned}
-  \widetilde{N}^\nu_{jk} &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \pr\cond{i[t-1,t]=(jk)}{\w[0,T],\M} \\
-     &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \tilde{\alpha}_i(t-1) \, \tilde{\beta}_j(t) \, \eta(t) \, \M^{\nu}_{i(t-1)i(t)} \, \Pib_k(t) ,\\
-  \tilde{\alpha}_j(t) \, \tilde{\beta}_j(t) &= \pr\cond{i(t)=j}{\w[0,T],\M}.
+  \tilde{\alpha}_j(t) \, \tilde{\beta}_j(t) &= \pr\cond{i(t)=j}{\w[0,T],\mu[0,T-1],\M}.
 \end{aligned}
@@
 \begin{aligned}
-  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr\cond{\w[0,T]}{\M} \, \frac{ \widetilde{N}^\mu_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
-  \pdiff{\CL}{\pib'_{i}} &= \pr\cond{\w[0,T]}{\M} \, \frac{ \tilde{\alpha}_i(0) \, \tilde{\beta}_i(0) }{ \pib'_{i} } - \lambda,
+  \widetilde{N}^\nu_{jk} &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, 
+      \pr\cond{i[t,t+1]=(jk)}{\w[0,T],\mu[0,T-1],\M} \\
+     &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, 
+      \tilde{\alpha}_j(t) \, \tilde{\beta}_k(t+1) \, \eta(t+1) \, 
+      \M^{\nu}_{jk} \, \Pib_k(t+1) ,\\
+  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr_{[0,T]}\cond{\w}{\mu,\M} \, 
+      \frac{ \widetilde{N}^\mu_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
+  \pdiff{\CL}{\pib'_{i}} &= \pr_{[0,T]}\cond{\w}{\mu,\M} \, 
+      \frac{ \tilde{\alpha}_i(0) \, \tilde{\beta}_i(0) }{ \pib'_{i} } - \lambda,
 \end{aligned}
@@
 %
-The factors of $\pr\cond{\w[0,T]}{\M}$ can be absorbed by a redefinition of the Lagrange multipliers.
+The factors of $\pr_{[0,T]}\cond{\w}{\mu,\M}$ can be absorbed by a redefinition of the Lagrange multipliers.
 
@@
 \begin{aligned}
-  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr\cond{\w^A[0,T]}{\M} \,\sum_a \frac{1}{\pr\cond{\w^a[0,T]}{\M}} \,
+  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \,\sum_a \frac{1}{\pr^a_{[0,T]}\cond{\w}{\mu,\M}} \,
     \frac{ N^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
-    &= \pr\cond{\w^A[0,T]}{\M} \, \frac{ \sum_a \widetilde{N}^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
-  \pdiff{\CL}{\pib'_{i}} &= \pr\cond{\w^A[0,T]}{\M} \,\sum_a \frac{1}{\pr\cond{\w^a[0,T]}{\M}} \,
+    &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \, \frac{ \sum_a \widetilde{N}^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
+  \pdiff{\CL}{\pib'_{i}} &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \,\sum_a \frac{1}{\pr^a_{[0,T]}\cond{\w}{\mu,\M}} \,
     \frac{ \alpha^a_i(0) \, \beta^a_i(0) }{ \pib'_{i} } - \lambda,\\
-    &= \pr\cond{\w^A[0,T]}{\M} \,
+    &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \,
     \frac{ \sum_a \tilde{\alpha}^a_i(0) \, \tilde{\beta}^a_i(0) }{ \pib'_{i} } - \lambda.
@@
 %
-Once again, we can absorb the factors of $\pr\cond{\w[0,T]}{\M}$ by a redefinition of the Lagrange multipliers.
+Once again, we can absorb the factors of $\pr^A_{[0,T]}\cond{\w}{\mu,\M}$ by a redefinition of the Lagrange multipliers.
 
@@
 \begin{equation}\label{eq:posterior}
-  \pr\cond{\M}{\w[0,T]} = \frac{ \pr(\w[0,T],\M) }{ \pr(\w[0,T) } = \frac{ \pr\cond{\w[0,T]}{\M} \pr(\M) }{ \pr(\w[0,T) }.
+  \pr\cond{\M}{\w[0,T]} = \frac{ \pr(\w[0,T],\M) }{ \pr(\w[0,T]) } = \frac{ \pr\cond{\w[0,T]}{\M} \pr(\M) }{ \pr(\w[0,T]) }.
 \end{equation}
@@
 
-Using the same trick as before \eqref{eq:emq}, we can define an auxilliary function
+Using the same trick that was used for \eqref{eq:emq}, we can define an auxilliary function
 %
@@
 \begin{multline}\label{eq:L1norm}
-  \sum_{j \neq i}\frac{\widetilde{N}^{A\mu}_{ij}}{\lambda^\mu_{i}+\beta} +\frac{\widetilde{N}^{A\mu}_{ii}}{\lambda^\mu_{i}} =0, \\
+  \sum_{j \neq i}\frac{\widetilde{N}^{A\mu}_{ij}}{\lambda^\mu_{i}+\beta} +\frac{\widetilde{N}^{A\mu}_{ii}}{\lambda^\mu_{i}} = 1, \\
   \implies \quad

### 2020-08-05 Code/Python/complex_synapse/identify/plast_seq.py

@@ -2,9 +2,10 @@
 """
 from __future__ import annotations
 
-from typing import List, Tuple, Union
+from typing import List, Tuple, Union, Sequence
 
 import numpy as np
+import matplotlib as mpl
 
 import numpy_linalg as la
 
@@ -68,7 +69,7 @@ class PlasticitySequence:
         attrs.update(kwds)
         return type(self)(**attrs)
 
-    def copy(self, order='C', **kwargs) -> PlasticitySequence:
+    def copy(self, order: str = 'C', **kwargs) -> PlasticitySequence:
         """Copy of object, with copies of array attributes
 
         Requires `__init__` parameter names to be the same as attribute names.
@@ -118,6 +119,14 @@ class PlasticitySequence:
             `dwells[i]` is an array of dwell times in readout group `i`.
         """
         return _dwells(self.readouts, self.nreadout)
+
+    def plot(self, axs: Sequence[Union[Image, Axes]], **kwds) -> List[Image]:
+        """Plot heatmaps for plast and initial"""
+        kwds['norm'] = mpl.colors.Normalize(0., 1. * self.nplast)
+        imh = [set_plot(axs[0], self.plast_type, **kwds)]
+        kwds['norm'] = mpl.colors.Normalize(0., 1. * self.nreadout)
+        imh.append(set_plot(axs[1], self.readouts, **kwds))
+        return imh
 
 
 class SimPlasticitySequence(PlasticitySequence):
@@ -186,6 +195,14 @@ class SimPlasticitySequence(PlasticitySequence):
         """
         return _dwells(self.states, self.nstate)
 
+    def plot(self, axs: Sequence[Union[Image, Line, Axes]], **kwds
+             ) -> List[Union[Image, Line]]:
+        """Plot heatmaps for plast and initial"""
+        imh = super().plot(axs, **kwds)
+        kwds['norm'] = mpl.colors.Normalize(0., 1. * self.nstate)
+        imh.append(set_plot(axs[2], self.states, **kwds))
+        return imh
+
 
 def _dwells(values: np.ndarray, num: int) -> List[la.lnarray]:
     """Set of dwell times
@@ -213,4 +230,24 @@ def _dwells(values: np.ndarray, num: int) -> List[la.lnarray]:
         dwells.append(stays[before == i])
     return dwells
 
-# TODO: plot
+
+def set_plot(handle: Union[Axes, Image, Line], data: np.ndarray, **kwds
+             ) -> Union[Image, Line]:
+    """Make a plot"""
[…]

### 2020-08-05 Code/Python/complex_synapse/identify/synapse_id.py

@@ -1,20 +1,22 @@
 """Model that is being fit to data
 """
 from __future__ import annotations
+from complex_synapse.synapse_mem import normalise
 
-from typing import Optional, Sequence, Union
+from typing import List, Optional, Sequence, Union
 
 import numpy as np
+import matplotlib as mpl
 
 import numpy_linalg as la
 from numpy_linalg import convert as cvl
 from sl_py_tools.arg_tricks import default_non_eval as default
 from sl_py_tools.containers import tuplify
-from sl_py_tools.numpy_tricks.markov import isstochastic_c
+from sl_py_tools.numpy_tricks.markov import isstochastic_c, stochastify_d
 
 from ..builders import RNG
 from ..synapse_base import ArrayLike, SynapseBase
-from .plast_seq import SimPlasticitySequence
+from .plast_seq import SimPlasticitySequence, Axes, Image, set_plot
 
 
 class SynapseIdModel(SynapseBase):
@@ -78,6 +80,12 @@ class SynapseIdModel(SynapseBase):
         results = self.initial.__array_ufunc__(ufunc, method, *args, **kwargs)
         return cvl.conv_out_attr(base_result, 'initial', results, outs, conv)
 
+    @property
+    def nreadout(self) -> int:
+        """Number of readout values, R
+        """
+        return self.readout.max() + 1
+
     def moveaxis(self, source: Union[int, Sequence[int]],
                  destination: Union[int, Sequence[int]]) -> SynapseIdModel:
         """Change order of axes in self.plast and self.ini6tial.
@@ -98,27 +106,13 @@ class SynapseIdModel(SynapseBase):
         newobj.plast = np.moveaxis(self.plast, source, destination)
         newobj.initial = np.moveaxis(self.initial, source, destination)
 
-    @property
-    def nreadout(self) -> int:
-        """Number of readout values, R
-        """
-        return self.readout.max() + 1
-
-    def readout_indicator(self) -> la.lnarray:
-        """1 in sets with each readout value, 0 elsewhere (R,M)
-        """
-        indicators = la.zeros((self.nreadout, self.nstate))
-        for i in range(self.nreadout):
-            indicators[i, self.readout == i] = 1
-        return indicators
-
-    def readout_proj(self) -> la.lnarray:
-        """Projection onto sets with each readout value, (R,M,M)
-        """
-        return self.readout_indicator().c * la.identity(self.nstate)
+    def normalise(self) -> None:
[…]

### 2020-10-13 Code/Python/src/baum_welch.c
memory leak somewhere in m_init

@@ -76,6 +76,28 @@ PyDoc_STRVAR(alpha_beta__doc__,
 "    Norm of Baum-Welch forward variable.\n"
 );
 
+/* update_m_init_signature = "(r,p,m,m),(tm),(t),(t,m),(t,m),(t)->(p,m,m),(m)"; */
+
+PyDoc_STRVAR(m_init__doc__,
+/* "m_init(M, P, R, A, B, E: ndarray) -> (UM, US: ndarray)\n\n" */
+"One Baum-Welch/Rabiner-Juang update of the model\n"
+"\nParameters\n-----------\n"
+"updaters : la.lnarray, (R,P,M,M) float[0:1]\n"
+"    Plasticity matrices multiplied by readout probability given 'to' state.\n"
+"plast_type : la.lnarray, (T-1,E), int[0:P]\n"
+"    id of plasticity type after each time-step\n"
+"alpha : la.lnarray, (T,M) float[0:1]\n"
+"    Normalised Baum-Welch forward variable\n"
+"beta : la.lnarray, (T,M) float\n"
+"    Scaled Baum-Welch backward variable\n"
+"eta : la.lnarray, (T,) float[1:]\n"
+"    Norm of Baum-Welch forward variable\n"
+"\nReturns\n-------\n"
+"plast : array_like, (P,M,M), float[0:1]\n"
+"    new estimate of transition probability matrix.\n"
+"initial : array_like, (M,) float[0:1]\n"
+"    new estimate of distribution of initial state.\n"
+);
 
 /*
 *****************************************************************************
@@ -85,9 +107,7 @@ PyDoc_STRVAR(alpha_beta__doc__,
 
 /* copy vector x into y */
 extern void
-FNAME(dcopy)(int *n,
-            double *sx, int *incx,
-            double *sy, int *incy);
+FNAME(dcopy)(int *n, double *x, int *incx, double *y, int *incy);
 
 /* set vector x to constant */
 extern void
@@ -95,22 +115,43 @@ FNAME(dlaset)(char trans[], int *m, int *n,
             double *alpha, double *beta,
             double a[], int *lda);
 
+/* y -> a x + y */
+extern void
+FNAME(daxpy)(int *n, double *a, double x[], int *incx,
+            double y[], int *incy);
+
 /* x -> x / a */
 extern void
 FNAME(drscl)(int *n, double *alpha, double x[], int *incx);
 
[…]

### 2020-10-13 Code/Python/src/gufunc_common.h

@@ -60,60 +60,86 @@ Copyright/licence info for that file:
 Macros for looping over non-core dimensions provided by numpy
 */
 
-/* Declare and initialise loop variables:
-    dN = length of non-core dimension
-    N_ = loop variable
-    s0 = stride along non-core dimension of first argument
-When number of ufunc inputs + outputs == 1 */
+/* Declare and initialise loop variables
+when number of ufunc inputs + outputs == 1:
+    dN = length of non-core dimension,
+    N_ = loop variable,
+    s0 = stride along non-core dimension of first argument.
+*/
 #define INIT_OUTER_LOOP_1         \
     npy_intp dN = *dimensions++;  \
     npy_intp N_;                  \
     npy_intp s0 = *steps++;
 
-/* Declare and initialise loop variables:
-    s1 = stride along non-core dimension of second argument
-When number of ufunc inputs + outputs == 2 */
+/* Declare and initialise loop variables
+when number of ufunc inputs + outputs == 2:
+    s1 = stride along non-core dimension of second argument.
+*/
 #define INIT_OUTER_LOOP_2   \
     INIT_OUTER_LOOP_1       \
     npy_intp s1 = *steps++;
 
-/* Declare and initialise loop variables:
-    s2 = stride along non-core dimension of third argument
-When number of ufunc inputs + outputs == 3 */
+/* Declare and initialise loop variables
+when number of ufunc inputs + outputs == 3:
+    s2 = stride along non-core dimension of third argument.
+*/
 #define INIT_OUTER_LOOP_3   \
     INIT_OUTER_LOOP_2       \
     npy_intp s2 = *steps++;
 
-/* Declare and initialise loop variables:
-    s3 = stride along non-core dimension of fourth argument
-When number of ufunc inputs + outputs == 4 */
+/* Declare and initialise loop variables
+when number of ufunc inputs + outputs == 4:
+    s3 = stride along non-core dimension of fourth argument.
+*/
 #define INIT_OUTER_LOOP_4   \
     INIT_OUTER_LOOP_3       \
     npy_intp s3 = *steps++;
 
-/* Declare and initialise loop variables:
-    s4 = stride along non-core dimension of fifth argument
[…]

### 2020-10-17 Notes/laplace.tex
notes on sqrt bound

@@
 %Preamble:
+\usepackage[T1]{fontenc}
 \usepackage{a4wide}
-\input{sl_preamble.tex}
+\usepackage{stix}
+% Preamble:
+%\usepackage{a4wide}
+%\usepackage{ams} for finding documentation only
+\usepackage[centertags]{amsmath}
+% \usepackage{amssymb}
+\usepackage{xcolor}
+\usepackage[sort&compress,numbers]{natbib}
+\usepackage{ifpdf}
+%
+\ifpdf
+ \usepackage[pdftex,bookmarks,bookmarksopen,pdfstartview=FitH,bookmarksnumbered]{hyperref}
+\else
+ \usepackage[hypertex]{hyperref}
+\fi
+% ----------------------------------------------------------------
+\vfuzz2pt % Don't report over-full v-boxes if over-edge is small
+\hfuzz2pt % Don't report over-full h-boxes if over-edge is small
+% ----------------------------------------------------------------
+\def\Snospace~{\S{}}
+\renewcommand{\sectionautorefname}{\Snospace}
+\renewcommand{\subsectionautorefname}{\Snospace}
+\renewcommand{\subsubsectionautorefname}{\Snospace}
+\usepackage{cleveref}
 \usepackage{mathtools}
@@
 
+%--------------------------------------------------------------------------------
+
 \subsection{Recognition memory}\label{sec:recog}
@@
 Different parts of the brain, that store memories for different timescales, could be characterised by different values of $\tau$.
+
+%--------------------------------------------------------------------------------
 
@@
 
+%--------------------------------------------------------------------------------
+
 \subsection{Markov models of synapses}\label{sec:markovsynapse}
@@
 %
-If the probability of a given plasticity event being potentiating/depressing is $f\potdep$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
+If the probability of a given plasticity event being potentiating or depressing is $f\pot$ or $f\dep$ respectively, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
 
@@
 \begin{multline}\label{eq:plastseqtypes}
-  \pr(t_n) = \pr(t_0) (f\pot\M\pot+f\dep\M\dep)  \ldots (f\pot\M\pot+f\dep\M\dep) \\
-    \M\potdep (f\pot\M\pot+f\dep\M\dep) \ldots (f\pot\M\pot+f\dep\M\dep),
+  \pr(t_n) = \pr(t_0) (f\pot\M\pot+f\dep\M\dep)  \ldots \\
+    \ldots (f\pot\M\pot+f\dep\M\dep) 
+    \M\potdep (f\pot\M\pot+f\dep\M\dep) \ldots \\
+    \ldots (f\pot\M\pot+f\dep\M\dep),
 \end{multline}
@@
 
+%================================================================================
+
 \section{Laplace transform}\label{sec:laplace}
+
+%--------------------------------------------------------------------------------
 
@@
 
+%--------------------------------------------------------------------------------
+
 \subsection{Laplace transform of SNR curve}\label{sec:laplaceSNR}
@@
 The first choice leaves \( \shift{\enc} = \enc \), so it only affects the \(\etw\) piece of \cref{eq:larea}.
-The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + (f\pot - d\dep) (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
+The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + (f\pot - f\dep) (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
 Either way:
@@
 %
+
+%--------------------------------------------------------------------------------
 
@@
 
+%================================================================================
+
 \section{Upper bounds}\label{sec:upperbnds}
 
-\subsection{Initial SNR}\label{sec:initial}
-
-The initial SNR is:
-%
-\begin{equation}\label{eq:initdef}
-  \snr(0) = \snrb(0) = \lim_{s\to\infty} sA(s).
-\end{equation}
-%
-
-As shown in \cite{Lahiri2013synapse}, this satisfies the inequality
-%
-\begin{equation}\label{eq:initbnd}
-  \snr(0) \leq \sqrt{N},
-\end{equation}
-%
-The model that saturates this bound is (equivalent to) a two-state model with deterministic transitions and $f\potdep=\frac{1}{2}$.
-
-\subsection{Area bound}\label{sec:area}
-
-The area under the SNR curve is:
-%
-\begin{equation}\label{eq:areadef}
-  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) = A(0).
-\end{equation}
-%
-
-As shown in \cite{Lahiri2013synapse}, this satisfies the inequality
-%
-\begin{equation}\label{eq:areabnd}
-  A(0) \leq \frac{\sqrt{N}(M-1)}{r}.
-\end{equation}
-%
-The model that saturates this bound is one of serial topology with $\eq$ concentrated symmetrically at the two end states.
-
-\subsection{Envelope}\label{sec:envelope}
-
-Now we introduce an eigenvector decomposition for $-\frg$:
+Now to study the SNR as a function of  time, it helps to consider some variables that are more directly related to the time dependance than the matrix elements. 
+So we introduce an eigenvector decomposition for $-\frg$:
 %
@@
 %
-\begin{equation}\label{eqsnreig}
+\begin{equation}\label{eq:snreig}
   \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a},
@@
 \begin{equation}\label{eq:snrcoeffs}
+  \CI_a = (\eq \enc \evr^a) (\evl^a \w),
+  \qquad
+  \tau_a = \frac{1}{q_a}.
+\end{equation}
+%
+Now the question is what are the allowed values of \( (\CI_a, \tau_a) \)?
+We have to translate the constraints of \cref{eq:constri} into constraints on the eigenmode variables.
+Each constraint we find will be a necessary condition for the existence of a set of Markov processes \( \M_{ij}^\nu \) that result in \( (\CI_a, \tau_a) \).
+We can imagine the notion of a complete set of constraints: constraints that, when combined, are also sufficient for the existence of a \( \M_{ij}^\nu \).
+But finding such a set seems intractable for now.
+Most likely there are a large number of such constraints, but if we can find a few of them they will provide us with some limits on the space of possible memory curves.
+These limits may not be tight - there may not be models that can reach these limits, but they will imply a forbidden region in the space of memory curves.
+
+%--------------------------------------------------------------------------------
+
+\subsection{Initial SNR}\label{sec:initial}
+
+The initial SNR is:
+%
+\begin{equation}\label{eq:initdef}
+  \snr(0) = \snrb(0) = \lim_{s\to\infty} sA(s).
+\end{equation}
+%
+
+As shown in \cite{Lahiri2013synapse}, this satisfies the inequality
+%
+\begin{equation}\label{eq:initbnd}
+  \snr(0) \leq \sqrt{N},
+\end{equation}
+%
+The model that saturates this bound is (equivalent to) a two-state model with deterministic transitions and $f\potdep=\frac{1}{2}$.
+
+%--------------------------------------------------------------------------------
+
+\subsection{Area bound}\label{sec:area}
+
+The area under the SNR curve is:
+%
+\begin{equation}\label{eq:areadef}
+  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) = A(0).
+\end{equation}
+%
+
+As shown in \cite{Lahiri2013synapse}, this satisfies the inequality
+%
+\begin{equation}\label{eq:areabnd}
+  A(0) \leq \frac{\sqrt{N}(M-1)}{r}.
+\end{equation}
+%
+The model that saturates this bound is one of serial topology with $\eq$ concentrated symmetrically at the two end states.
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Square-root bound}\label{sec:sqrt}
+
+\newcommand{\aav}[1]{\mathopen{}\left\lAngle #1 \right\rAngle\mathclose{}}
+\newcommand{\nnrm}[1]{\mathopen{}\left\Vvert #1 \right\Vvert\mathclose{}}
+
+This proof follows that of \cite{Lawler1988cheeger}.
+We can define two inner-products that will be of use.
+The first is the usual inner product of two vectors (or functions on state space in the continuum case) using the steady state distribution as the measure:
+%
+\begin{equation}\label{eq:innereq}
+  \av{\mathbf{u}, \mathbf{v}} = \sum_i \eq_i u_i^\ast v_i ,
+  \qquad
+  \nrm{\mathbf{v}} = \sqrt{\av{\mathbf{v}, \mathbf{v}}} .
+\end{equation}
+%
+The second is an inner product of rank 2 tensors (functions of pairs of points on state space) using the steady-state flux as the measure:
+%
+\begin{equation}\label{eq:innerflux}
+  \aav{\mathbf{F}, \mathbf{G}} = \sum_{i \neq j} \F_{ij} F_{ij}^\ast G_{ij} ,
+  \qquad
+  \nnrm{\mathbf{F}} = \sqrt{\aav{\mathbf{F}, \mathbf{F}}}.
+\end{equation}
+%
+The omission of the diagonal terms leaves only non-negative terms.
+
+When the process satisfies detailed balance, \ie \(\F_{ij}\) is symmetric, we can relate these inner products to quantities of interest.
+
+We can write
+%
+\begin{equation*}
+  \av{\mathbf{u}, -\MM\mathbf{v}} = \sum_{i \neq j} \eq_i u_i^\ast \MM_{ij} (v_i - v_j),
+\end{equation*}
+%
+under detailed balance, we can symmetrize
+%
+\begin{equation*}
+  \av{\mathbf{u}, -\MM \mathbf{v}} 
+      = \frac{1}{2}\sum_{i \neq j} \F_{ij} (u_i^\ast - u_j^\ast) (v_i - v_j) 
+      = \frac{1}{2} \aav{\boldsymbol{\Delta}\mathbf{u}, \boldsymbol{\Delta}\mathbf{v}}.
+\end{equation*}
+%
+
+The eigenvalue spectrum of \(\MM\) is closely related to its numerical range:
+%
+\begin{equation}\label{eq:numrange}
+  \mathcal{W}(-\MM; \mathcal{V}) 
+    = \set{\frac{\av{\mathbf{u}, -\MM \mathbf{u}}}{\av{\mathbf{u}, \mathbf{u}}}}
+          {\mathbf{u} \in \mathcal{V}} 
+    = \set{\frac{\nnrm{\boldsymbol{\Delta}\mathbf{u}}^2}{2\nrm{\mathbf{u}}^2}}
+          {\mathbf{u} \in \mathcal{V}} .
+\end{equation}
+%
+When \(\MM\) satisfies detailed balance (more generally when it commutes with its time reverse), the numerical range is the convex hull of the eigenvalues.
+
+The other quantity of interest in \cref{eq:snrcoeffs} is the amplitude of the eigenmodes.
+In analogy, we define the following functional (for real \(\mathbf{u}\))
+%
+\begin{equation*}
+  \CI(\mathbf{u}) = \frac{(\eq \enc \mathbf{u}) (\mathbf{u}^\dag \w)}{\mathbf{u}^\dag \mathbf{u}}
+      = \frac{\sum_{i \neq j} \eq_i \enc_{ij} (u_j - u_i) \sum_k \eq_k u_k \w_k} 
+             {\sum_{l} \eq_l u_l^2} .
+\end{equation*}
+%
+Following \cite{Lawler1988cheeger}, we can write
+%
+\begin{equation*}
+  u_j - u_i = \intd[_{u_i}^{u_j}]{\alpha} 
+      = \intd[_{-\infty}^{\infty}]{\alpha} 
+        \brk{\vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
+           - \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)} ,
+\end{equation*}
+%
+where \(\vartheta\) is the Heaviside step-function.
+Then 
+%
+\begin{equation*}
+  \begin{aligned}[t]
+  r \eq \enc \mathbf{u} &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
+        \F\pot_{\CA_\alpha \CA_\alpha^c} - \F\dep_{\CA_\alpha \CA_\alpha^c} 
+      - \F\pot_{\CA_\alpha^c \CA_\alpha} + \F\dep_{\CA_\alpha^c \CA_\alpha} }
+    \\
+    &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
+        \F_{\CA_\alpha \CA_\alpha^c} + \F_{\CA_\alpha^c \CA_\alpha} 
+        - 2\F\dep_{\CA_\alpha \CA_\alpha^c} - 2\F\pot_{\CA_\alpha^c \CA_\alpha} }
+  \end{aligned}
+  %
+  \quad \text{where} \quad
+  %
+  \begin{aligned}[t]
+    \CA_\alpha &= \set{i}{u_i < \alpha} 
+    , \\
+    \F^\nu_{\CA \CB} &= r f^\nu \sum_{\substack{i \in \CA \\ j \in \CB}} 
+                        \prn{\M^\nu_{ij} - \delta_{ij}}
+    .
+  \end{aligned}
+\end{equation*}
+%
+In exactly the same way as the proof of \cref{eq:initbnd}, we can argue that the two terms with the negative signs are indeed non-positive as they do not involve the negative diagonal elements of~\( \F_{ij}^\nu \).
+Therefore the other terms provide an upper bound.
+%
+\begin{equation*}
+\begin{aligned}
+  r \nrm{\mathbf{u}}^2 \CI(\mathbf{u}) 
+      &\leq \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
+        \F_{\CA_\alpha \CA_\alpha^c} + \F_{\CA_\alpha^c \CA_\alpha} } 
+        \av{\mathbf{u}, \w}
+  \\  &= \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \F_{ij} \brk{
+            \vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
+          + \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)
+        } \av{\mathbf{u}, \w}
+  \\  &= \sum_{i \neq j} \F_{ij} \abs{u_j - u_i} \av{\mathbf{u}, \w}
+  \\  &= \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem} \av{\mathbf{u}, \w} .
+\end{aligned}
+\end{equation*}
+%
+Now we can apply the Cauchy-Schwartz inequality to both factors to find
+%
+\begin{equation*}
+  r \CI(\mathbf{u}) \leq \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem} 
+                          \nrm{\mathbf{u}} \nrm{\w}.
+\end{equation*}
+%
+We also have \( \nrm{\w}^2 = \sum_i \eq_i \w_i^2 = 1 \).
+Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eq_i \MM_{ij} \leq r \sum_i \eq_i = r\).
+Then, remembering \cref{eq:numrange},
+%
+\begin{equation}\label{eq:sqrtnumrange}
+  \sqrt{r} \CI(\mathbf{u}) \leq \sqrt{2\mathcal{W}(-\MM, \mathbf{u})}.
+\end{equation}
+%
+Now, simply choosing \(\mathbf{u} = \evr_a \), one of the orthonormal eigenvectors, we find a bound on the eigenmodes:
+%
+\begin{equation}\label{eq:sqrtbound}
+  \text{assuming} \quad 
+  \eq_i \MM_{ij} = \eq_j \MM_{ji},
+  \qquad \text{we have} \quad
+  \CI_a \sqrt{\tau_a} \leq \sqrt{2}.
+\end{equation}
+%
+
+%--------------------------------------------------------------------------------
+
+\subsection{Envelope}\label{sec:envelope}
+
+Recall the eigenmode description of the memory curve from \cref{eq:snreig}
+%
+\begin{equation*}
+  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a},
+  \qquad
+  \snrb(\tau) = \sqrt{N}\sum_a \frac{\CI_a}{1+r\tau/\tau_a},
+\end{equation*}
+%
+where
+%
+\begin{equation*}
   \CI_a = (\eq \ench \evr^a) (\evl^a \w),
@@
   \tau_a = \frac{1}{q_a}.
-\end{equation}
+\end{equation*}
 %
@@
 \end{figure}
+
+%================================================================================
 
@@
 
+%================================================================================
+
 \section{Nearly uniform serial models}\label{sec:serial}
@@
 
+%--------------------------------------------------------------------------------
+
 \subsection{Uniform serial model}\label{sec:serial_uniform}
@@
 However, it does match up with the full formula, \eqref{eq:unienv} at the boundary.
+
+%--------------------------------------------------------------------------------
 
@@
 
+%--------------------------------------------------------------------------------
+
 \subsection{Sticky serial model}\label{sec:serial_sticky}
@@
 
+%--------------------------------------------------------------------------------
+
 \subsection{Heuristic envelope}\label{sec:heuristicenv}
@@
 
+%================================================================================
+
 \section{Conclusions}\label{sec:conclusions}

### 2020-10-22 Notes/laplace.tex
numerics on sqrt bound

@@
 \usepackage{a4wide}
-\usepackage{stix}
-% Preamble:
-%\usepackage{a4wide}
-%\usepackage{ams} for finding documentation only
-\usepackage[centertags]{amsmath}
-% \usepackage{amssymb}
-\usepackage{xcolor}
-\usepackage[sort&compress,numbers]{natbib}
-\usepackage{ifpdf}
-%
-\ifpdf
- \usepackage[pdftex,bookmarks,bookmarksopen,pdfstartview=FitH,bookmarksnumbered]{hyperref}
-\else
- \usepackage[hypertex]{hyperref}
-\fi
-% ----------------------------------------------------------------
-\vfuzz2pt % Don't report over-full v-boxes if over-edge is small
-\hfuzz2pt % Don't report over-full h-boxes if over-edge is small
-% ----------------------------------------------------------------
+\usepackage[tables,graphics,draft]{sl-preamble}
+% \usepackage{stix}
+%\usepackage{amsldoc} for finding documentation only
 \def\Snospace~{\S{}}
@@
 \renewcommand{\subsubsectionautorefname}{\Snospace}
-\usepackage{cleveref}
-\usepackage{mathtools}
-%
-% >> Only for drafts! <<
-\usepackage[notref,notcite]{showkeys}
-%
-\input{sl_graphics_preamble.tex}
+%
 \graphicspath{{Figures/}}
@@
 %\numberwithin{equation}{section}
-%\renewcommand{\baselinestretch}{1.5}
-% ----------------------------------------------------------------
-% New commands etc.
-\input{sl_definitions.tex}
-\input{sl_symbols.tex}
+%================================================================================
+\makeatletter
+\DeclareFontEncoding{LS2}{}{\noaccents@}
+\makeatother
+
+\DeclareFontSubstitution{LS2}{stix}{m}{n}
+\DeclareSymbolFont{xlargesymbols}{LS2}{stixex}{m}{n}
+\DeclareMathDelimiter{\lAngle}{\mathopen}{xlargesymbols}{"EC}{xlargesymbols}{"12}
+\DeclareMathDelimiter{\rAngle}{\mathclose}{xlargesymbols}{"ED}{largesymbols}{"13}
+\DeclareMathDelimiter{\lVvert}{\mathopen}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
+\DeclareMathDelimiter{\rVvert}{\mathclose}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
+
+\newcommand{\aav}[1]{\mathopen{}\left\lAngle #1 \right\rAngle\mathclose{}}
+\newcommand{\nnrm}[1]{\mathopen{}\left\lVvert #1 \right\rVvert\mathclose{}}
+%================================================================================
+%
+\usepackage{adjustbox}
+\newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
+\newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
+%
+\usepackage{paralist}
+\newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
+\newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
+\newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
+\newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
+\newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
+%for nested lists, \ref only produces innermost item:  (hacky)
+\makeatletter % for internal macros with @
+\renewcommand\p@enumii{}
+\makeatother
+%
+%================================================================================
 %matrices
@@
 %Markov matrix
-\newcommand{\MM}{\mathbf{Q}}
+\newcommand{\MMm}{Q}
+\newcommand{\MM}{\mathbf{\MMm}}
 %prob distribution
-\newcommand{\pr}{\mathbf{p}}
+\newcommand{\prm}{p}
+\newcommand{\pr}{\mathbf{\prm}}
 %equilibrium distribution
-\newcommand{\eq}{\pib}
+\newcommand{\eqm}{\pi}
+\newcommand{\eq}{\boldsymbol{\eqm}}
 %equilibrium metric
-\newcommand{\Eq}{\boldsymbol{\Pi}}
+\newcommand{\Eqm}{\Pi}
+\newcommand{\Eq}{\boldsymbol{\Eqm}}
 %first passage times
-\newcommand{\fpt}{\mathbf{T}}
+\newcommand{\fptm}{T}
+\newcommand{\fpt}{\mathbf{\fptm}}
 %off-diag first passage times
+\newcommand{\fptbm}{\overline{\fptm}}
 \newcommand{\fptb}{\overline{\fpt}}
 %fundamental matrix
-\newcommand{\fund}{\mathbf{Z}}
+\newcommand{\fundm}{Z}
+\newcommand{\fund}{\mathbf{\fundm}}
 %other symbols
 \newcommand{\Pb}{\mathbf{P}}
-\newcommand{\D}{\mathbf{D}}
 \newcommand{\pib}{\boldsymbol{\pi}}
-\newcommand{\arow}{\boldsymbol{\xi}}
-\newcommand{\Lb}{\boldsymbol{\Lambda}}
-\newcommand{\lb}{\boldsymbol{\lambda}}
+\newcommand{\arowm}{\xi}
+\newcommand{\arow}{\boldsymbol{\arowm}}
+\newcommand{\Lbm}{\Lambda}
+\newcommand{\Lb}{\boldsymbol{\Lbm}}
+\newcommand{\lbm}{\lambda}
+\newcommand{\lb}{\boldsymbol{\lbm}}
+\newcommand{\etwm}{\eta^w}
 \newcommand{\etw}{\boldsymbol{\eta}^w}
-\newcommand{\thb}{\boldsymbol{\theta}}
-\newcommand{\omb}{\boldsymbol{\omega}}
-\newcommand{\alb}{\boldsymbol{\alpha}}
+\newcommand{\thbm}{\theta}
+\newcommand{\thb}{\boldsymbol{\thbm}}
+\newcommand{\ombm}{\omega}
+\newcommand{\omb}{\boldsymbol{\ombm}}
+\newcommand{\albm}{\alpha}
+\newcommand{\alb}{\boldsymbol{\albm}}
 %synapse models
-\newcommand{\w}{\mathbf{w}}
-\newcommand{\W}{\mathbf{W}}
-\newcommand{\M}{\mathbf{M}}
-\newcommand{\enc}{\mathbf{K}}
+\newcommand{\wm}{w}
+\newcommand{\w}{\mathbf{\wm}}
+\newcommand{\Wm}{W}
+\newcommand{\W}{\mathbf{\Wm}}
+\newcommand{\MMdm}{M}
+\newcommand{\MMd}{\mathbf{\MMdm}}
+\newcommand{\encm}{K}
+\newcommand{\enc}{\mathbf{\encm}}
+\newcommand{\frgm}{\Wm^{\mathrm{F}}}
 \newcommand{\frg}{\W^{\mathrm{F}}}
-\newcommand{\F}{\boldsymbol{\Phi}}
+\newcommand{\Fm}{\Phi}
+\newcommand{\F}{\boldsymbol{\Fm}}
+%eigenvectors
+\newcommand{\evrm}{u}
+\newcommand{\evr}{\mathbf{\evrm}}
+\newcommand{\evlm}{\eta}
+\newcommand{\evl}{\boldsymbol{\evlm}}
 %superscripts
@@
 \newcommand{\potdep}{^{\text{pot/dep}}}
-%with activity indep processes
-\renewcommand{\hom}{\mathbf{H}}
-\newcommand{\Mh}{\widetilde{\M}}
-\newcommand{\frgh}{\widetilde{\W}^{\mathrm{F}}}
-\newcommand{\ench}{\widetilde{\enc}}
-%eigenvectors
-\newcommand{\evr}{\mathbf{u}}
-\newcommand{\evl}{\boldsymbol{\eta}}
 %snr curves etc
@@
 \DeclareMathOperator{\env}{Env}
-\newcommand{\rh}{\hat{r}}
-\newcommand{\CA}{\mathcal{A}}
-\newcommand{\CB}{\mathcal{B}}
+% \newcommand{\CA}{\mathcal{A}}
+% \newcommand{\CB}{\mathcal{B}}
 %avoid double superscripts
@@
 % ----------------------------------------------------------------
-\input{sl_theorems_preamble.tex}
+% \input{sl_theorems_preamble.tex}
 % ----------------------------------------------------------------
@@
 
-When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M^xn  $ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$,
+When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\MMd^\nu$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$,
 and \(\nu \) labels the diffrerent types of plasticity event.
-Potentiating events have \(\nu=\)pot, depressing events have \(\nu=\)dep.
+Potentiating events have \(\nu=\) pot, depressing events have \(\nu=\) dep.
 We also use the notation
@@
 %
-The matrices $\M^\nu$ are transition matrices of discrete time Markov processes.
+The matrices $\MMd^\nu$ are transition matrices of discrete time Markov processes.
 Their elements must satisfy the following constraints
@@
 \begin{equation}\label{eq:constr}
-  \M^\nu_{ij} \geq 0 
+  \MMdm^\nu_{ij} \geq 0 
     \quad \forall \; \nu, i, j, \qquad
-  \sum_j \M^\nu_{ij} = 1
+  \sum_j \MMdm^\nu_{ij} = 1
     \quad \forall \; \nu, i.
@@
 %
-The upper bounds $\M^\nu_{ij} \leq 1$ follow automatically from these.
+The upper bounds $\MMdm^\nu_{ij} \leq 1$ follow automatically from these.
 
@@
 \begin{equation}\label{eq:constri}
-  \M^\nu_{i j} \geq 0
+  \MMdm^\nu_{i j} \geq 0
     \quad \forall \; \nu, i, j : i \neq j, \qquad
-  \sum_{j : j\neq i} \M^\nu_{ij} \leq 1
+  \sum_{j : j\neq i} \MMdm^\nu_{ij} \leq 1
     \quad \forall \; \nu, i.
@@
 \begin{equation}\label{eq:plastseq}
-  \pr(t_n) = \pr(t_0)  \M\pot \M\dep \M\dep \M\pot \M\dep \M\pot \M\pot \M\dep \ldots \M\pot .
+  \pr(t_n) = \pr(t_0)  \MMd \pot \MMd \dep \MMd \dep \MMd \pot \MMd \dep \MMd \pot \MMd \pot \MMd \dep \ldots \MMd \pot .
 \end{equation}
@@
 \begin{multline}\label{eq:plastseqtypes}
-  \pr(t_n) = \pr(t_0) (f\pot\M\pot+f\dep\M\dep)  \ldots \\
-    \ldots (f\pot\M\pot+f\dep\M\dep) 
-    \M\potdep (f\pot\M\pot+f\dep\M\dep) \ldots \\
-    \ldots (f\pot\M\pot+f\dep\M\dep),
+  \pr(t_n) = \pr(t_0) (f\pot \MMd \pot + f\dep \MMd \dep)  \ldots \\
+    \ldots (f\pot \MMd \pot + f\dep \MMd \dep) 
+     \MMd \potdep (f\pot \MMd \pot + f\dep \MMd \dep) \ldots \\
+    \ldots (f\pot \MMd \pot + f\dep \MMd \dep),
 \end{multline}
 % 
-where the factor of $\M\potdep$ is the event at $t=0$.
+where the factor of $ \MMd \potdep$ is the event at $t=0$.
 If we expand out all of the parenthetical factors, we will get a sum of terms of the form \eqref{eq:plastseq} weighted by their probabilities.
 
-If $t_0 \ll 0$, the effect of the large number of earlier events will be to put the synapse in the steady state distribution, $\eq$, of the stochastic process $(f\pot\M\pot+f\dep\M\dep)$:
+If $t_0 \ll 0$, the effect of the large number of earlier events will be to put the synapse in the steady state distribution, $\eq$, of the stochastic process $(f\pot \MMd \pot+f\dep \MMd \dep)$:
 %
 \begin{equation}\label{eq:plastseqeq}
-  \pr(t_n) = \eq \M\potdep (f\pot\M\pot+f\dep\M\dep)^n,
+  \pr(t_n) = \eq  \MMd \potdep (f\pot \MMd \pot+f\dep \MMd \dep)^n,
 \end{equation}
@@
   P(\text{state}=i,t \mid \text{pot/dep},0)
-    &= \sum_n \frac{(rt)^n}{n!} \e^{-rt} \eq \M\potdep (f\pot\M\pot+f\dep\M\dep)^n \\
-    &= \brk{ \eq \M\potdep\, \e^{rt(f\pot\M\pot+f\dep\M\dep-\I) } }_i,
+    &= \sum_n \frac{(rt)^n}{n!} \e^{-rt} \eq  \MMd \potdep (f\pot \MMd \pot+f\dep \MMd \dep)^n \\
+    &= \brk{ \eq  \MMd \potdep\, \e^{rt(f\pot \MMd \pot+f\dep \MMd \dep-\I) } }_i,
 \end{aligned}
@@
   \qquad\text{where}\quad
-  \frg = \sum_\nu f^\nu \M^\nu - \I
-    = f\pot\M\pot + f\dep\M\dep - \I.
+  \frg = \sum_\nu f^\nu \MMd^\nu - \I
+    = f\pot \MMd \pot + f\dep \MMd \dep - \I.
 \end{equation}
@@
 \begin{aligned}
-  \av{\synid^1 \syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
-  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
+  \av{\synid^1 \syn^1(t)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \e^{rt\frg} \w, \\
+  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \onev\eq \w \\
          &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
@@
     \text{where} \qquad
-  \enc &= \sum_\nu (-1)^\nu f^\nu (\M^\nu - \I)
-    = f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
+  \enc &= \sum_\nu (-1)^\nu f^\nu (\MMd^\nu - \I)
+    = f\pot(\MMd \pot-\I)-f\dep(\MMd \dep-\I)
 \end{aligned}
@@
 \begin{equation}\label{eq:SNRcurveExact}
-  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt\frg}\w}
-                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
+  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt \frg}\w}
+                 {\sqrt{1-(f\pot-f\dep)^2(\eq \w)^2}}.
 \end{equation}
@@
 We can perform the maximisation in two steps.
-First maximise the numerator at fixed $\eq\w$.
+First maximise the numerator at fixed $\eq \w$.
 Then maximise the ratio \wrt $\eq\w$.
@@
 
-We can then treat the off-diagonal elements of $\M^\nu$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
+We can then treat the off-diagonal elements of $\MMd^\nu$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
 The diagonal elements are determined by setting the row sums to 1.
@@
 \begin{equation}\label{eq:derivpd}
-  \pdiff{\M^\mu_{ij}}{\M^\nu_{mn}} 
+  \pdiff{\MMdm^\mu_{ij}}{\MMdm^\nu_{mn}} 
       = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}),
@@
 \begin{equation}\label{eq:pertfe}
-  \pdiff{}{\frg_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
+  \pdiff{}{\frgm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\MMd \dep_{ij}},
   \qquad
-  \pdiff{}{\enc_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
+  \pdiff{}{\encm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
 \end{equation}
@@
 \begin{equation}\label{eq:derivfe}
-  \pdiff{\frg_{ij}}{\frg_{mn}} = \pdiff{\enc_{ij}}{\enc_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
+  \pdiff{\frgm_{ij}}{\frgm_{mn}} = \pdiff{\encm_{ij}}{\encm_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
   \qquad
-  \pdiff{\frg_{ij}}{\enc_{mn}} = \pdiff{\enc_{ij}}{\frg_{mn}} = 0.
+  \pdiff{\frgm_{ij}}{\encm_{mn}} = \pdiff{\encm_{ij}}{\frgm_{mn}} = 0.
 \end{equation}
@@
 \begin{aligned}
-  \frg_{i \neq j}   &\geq \abs{ \enc_{i \neq j} }, &\qquad
-  \sum_{j\neq i}\frg_{ij} + (-1)^\mu \enc_{ij}  &\leq f^\mu, \\
-  \sum_j \frg_{ij} &=0, &
-  \sum_j \enc_{ij} &=0.
+  \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
+  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq f^\mu, \\
+  \sum_j \frgm_{ij} &=0, &
+  \sum_j \encm_{ij} &=0.
 \end{aligned}
@@
 \begin{equation}\label{eq:alfund}
-  \pdiff{(\mathbf{a}\fund(s))_i}{\arow_k} = -(\mathbf{a}\fund(s)\onev) \fund_{ki}(s)
-      = -\frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
+  \pdiff{(\mathbf{a}\fund(s))_i}{\arowm_k} = -(\mathbf{a}\fund(s)\onev) \fundm_{ki}(s)
+      = -\frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fundm_{ki}(s) = 0.
 \end{equation}
@@
 \begin{equation}\label{eq:lfptb}
-  \fptb(s) = (\onem \fund\dg(s)-\fund(s))\D,
+  \fptb(s) = (\onem \fund\dg(s)-\fund(s))\Eq\inv,
   \qquad
-  \fptb_{ij}(s) = \frac{\fund_{jj}(s)-\fund_{ij}(s)}{\eq_j}.
+  \fptbm_{ij}(s) = \frac{\fundm_{jj}(s)-\fundm_{ij}(s)}{\eqm_j}.
 \end{equation}
@@
 This means that the \( \fptb(s) \) have all the same properties as in the case $s=0$.
-I particular, we see that the quantity
+In particular, we see that the quantity
 %
 \begin{equation}\label{eq:lkemeny}
-  \eta(s) = \sum_j \fptb_{ij}(s)\eq_j
+  \eta(s) = \sum_j \fptbm_{ij}(s) \eqm_j
 \end{equation}
@@
 \begin{equation}\label{eq:lpmixtime}
-  \eta^{\pm}_i(s) = \sum_j \fptb_{ij}(s) \eq_j \prn{\frac{1 \pm \w_j}{2}} 
-      =  \sum_{j\in \pm} \fptb_{ij}(s) \eq_j.
+  \eta^{\pm}_i(s) = \sum_j \fptbm_{ij}(s) \eqm_j \prn{\frac{1 \pm \wm_j}{2}} 
+      =  \sum_{j\in \pm} \fptbm_{ij}(s) \eqm_j.
 \end{equation}
@@
 \begin{equation}\label{eq:lwpmixtime}
-  \etw_i(s) = \sum_j \fptb_{ij}(s)\eq_j \w_j
+  \etwm_i(s) = \sum_j \fptbm_{ij}(s)\eqm_j \wm_j
     = \eta^+_i(s) - \eta^-_i(s) 
@@
 %
-Arranging the states in order of decreasing $\etw_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
-The $\etw_i(s)$ can still be used when $\w$ takes more than two values.
+Arranging the states in order of decreasing $\etwm_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
+The $\etwm_i(s)$ can still be used when $\w$ takes more than two values.
 
@@
   \hat{A}(s) = A(s) \sqrt{\frac{{{1-(f\pot-f\dep)^2(\eq\w)^2}}}{N}}  &=  \eq \enc \fund(s) \w \\
-    &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
-    &= \sum_{ij}  \eq_i \enc_{ij} (\etw_{i}(s) - \etw_{j}(s)).
+    &= \sum_{ijk} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)} \eqm_k \wm_k \\
+    &= \sum_{ij}  \eqm_i \encm_{ij} (\etwm_{i}(s) - \etwm_{j}(s)).
 \end{aligned}
@@
 We can also express this in terms of the shifted process in \cref{eq:larea}.
-We can choose how to distribute the shift across \(\M\pot\) and \(\M\dep\).
+We can choose how to distribute the shift across \(\MMd \pot\) and \(\MMd \dep\).
 Two choices from a continuum%
-\footnote{We could choose \( \shift{\M}^\nu(s) = \M^\nu + \frac{s g^\nu}{r f^\nu} (\onev \eq - \I) \), for any \(g^\nu \) with \( \sum_\nu g^\nu = 1 \).
+\footnote{We could choose \( \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s g^\nu}{r f^\nu} (\onev \eq - \I) \), for any \(g^\nu \) with \( \sum_\nu g^\nu = 1 \).
 The two possibilities in \cref{eq:lshiftpd} come from \(g^\nu = \frac{1}{2}\) and \(g^\nu = f^\nu\) respectively.}
@@
 \begin{equation}\label{eq:lshiftpd}
-  \shift{\M}^\nu(s) = \M^\nu + \frac{s}{2 r f^\nu} (\onev \eq - \I)
+  \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s}{2 r f^\nu} (\onev \eq - \I)
   \qquad \text{or} \qquad
-  \shift{\M}^\nu(s) = \M^\nu + \frac{s}{r} (\onev \eq - \I).
+  \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s}{r} (\onev \eq - \I).
 \end{equation}
@@
 The first choice leaves \( \shift{\enc} = \enc \), so it only affects the \(\etw\) piece of \cref{eq:larea}.
-The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + (f\pot - f\dep) (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
+The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + \frac{s}{r} (f\pot - f\dep) (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
 Either way:
@@
 \begin{equation}\label{eq:lareashift}
-  \hat{A}\prn{s; \M^\nu} = \hat{A}\prn{0; \shift{\M}^\nu(s)}.
+  \hat{A}\prn{s; \MMd^\nu} = \hat{A}\prn{0; \shift{\MMd}^\nu(s)}.
 \end{equation}
@@
 Note that while \cref{eq:lshift} produces a valid continuous-time Markov process, \cref{eq:lshiftpd} does not always produce valid discrete-time Markov processes, as the diagonal elements can become negative.
-However, \(r (\shift{\M}^\nu - \I)\) is a valid continuous-time Markov process.
+However, \(r (\shift{\MMd}^\nu - \I)\) is a valid continuous-time Markov process.
 
@@
   \begin{aligned}
-    \mathbf{c}_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)}, \\
-    \mathbf{a}_j(s) &= \sum_{k} \enc_{jk} (\etw_{j}(s) - \etw_{k}(s)), \\
-    \thb_i(s) &= \sum_j \fptb_{ij}(0)\, \eq_j \mathbf{a}_j(s), \\
+    {c}_k(s) &= \sum_{ij} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)}, \\
+    {a}_j(s) &= \sum_{k} \encm_{jk} (\etwm_{j}(s) - \etwm_{k}(s)), \\
+    \thbm_i(s) &= \sum_j \fptbm_{ij}(0)\, \eqm_j {a}_j(s), \\
   \end{aligned}
@@
     \hat{A}(s) 
-      = \sum_k \eq_k \mathbf{c}_k(s) \w_k
+      = \sum_k \eqm_k {c}_k(s) \wm_k
       % = \mathbf{c}^\dag(s) \w
@@
 \begin{equation}\label{eq:lAdiff}
-  \pdiff{\hat{A}(s)}{\M^\mu_{mn}}
-     = f^\mu \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
-     +\prn{r \mathbf{c}_m(s) + (-1)^\mu} \brk{\etw_m(s) - \etw_n(s)} }.
+  \pdiff{\hat{A}(s)}{\MMdm^\mu_{mn}}
+     = f^\mu \eqm_m \prn{ r \brk{\thbm_m(s) - \thbm_n(s)}
+     +\prn{r {c}_m(s) + (-1)^\mu} \brk{\etwm_m(s) - \etwm_n(s)} }.
 \end{equation}
@@
 \begin{alignedat}{3}
-    R_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frg_{ij} \circ \partial \frg_{kl}}} 
+    R_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frgm_{ij} \circ \partial \frgm_{kl}}} 
     &= \
-    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
-    &&&+ \eq_i \brc{\mathbf{X}_{ik}(s) - \mathbf{X}_{jk}(s)} 
-          \brc{\etw_k(s) - \etw_l(s)} \\
-    &&&+ \eq_i \mathbf{c}_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} 
-          \brc{\etw_k(s) - \etw_l(s)},  \\
-  S_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frg_{ij} \partial \enc_{kl}} }
+    &&\eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\thbm_k(s) - \thbm_l(s)} \\
+    &&&+ \eqm_i \brc{{X}_{ik}(s) - {X}_{jk}(s)} 
+          \brc{\etwm_k(s) - \etwm_l(s)} \\
+    &&&+ \eqm_i {c}_i(s) \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} 
+          \brc{\etwm_k(s) - \etwm_l(s)},  \\
+  S_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frgm_{ij} \partial \encm_{kl}} }
   &= \
-    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
-    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)},
+    &&\eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\etwm_k(s) - \etwm_l(s)} \\
+    &&&+ \eqm_k \brc{\fundm_{ki}(s) - \fundm_{li}(s)} \brc{\etwm_i(s) - \etwm_j(s)},
 \end{alignedat}
@@
 %
-where \(R\) contains only those terms where \( \frg_{ij} \) precedes \( \frg_{kl} \) in the matrix products from the expression for \( \hat{A} \) in \cref{eq:larea}.
+where \(R\) contains only those terms where \( \frgm_{ij} \) precedes \( \frgm_{kl} \) in the matrix products from the expression for \( \hat{A} \) in \cref{eq:larea}.
 The other terms sum to \(R\trans\).
@@
 \begin{equation}\label{eq:hessian}
-  \pdiff{^2\hat{A}(s)}{\M^\mu_{ij} \partial\M^\nu_{kl}} =
+  \pdiff{^2\hat{A}(s)}{\MMdm^\mu_{ij} \partial \MMdm^\nu_{kl}} =
      f^\mu f^\nu [R + R\trans + (-1)^\mu S\trans + (-1)^\nu S]_{ijkl}.
@@
 We have to translate the constraints of \cref{eq:constri} into constraints on the eigenmode variables.
-Each constraint we find will be a necessary condition for the existence of a set of Markov processes \( \M_{ij}^\nu \) that result in \( (\CI_a, \tau_a) \).
-We can imagine the notion of a complete set of constraints: constraints that, when combined, are also sufficient for the existence of a \( \M_{ij}^\nu \).
+Each constraint we find will be a necessary condition for the existence of a set of Markov processes \( \MMdm_{ij}^\nu \) that result in \( (\CI_a, \tau_a) \).
+We can imagine the notion of a complete set of constraints: constraints that, when combined, are also sufficient for the existence of a \( \MMdm_{ij}^\nu \).
 But finding such a set seems intractable for now.
@@
 \begin{equation}\label{eq:initdef}
-  \snr(0) = \snrb(0) = \lim_{s\to\infty} sA(s).
+  \snr(0) = \snrb(0) = \lim_{s\to\infty} sA(s)
+      = \sqrt{N} \sum_a \CI_a.
 \end{equation}
@@
 \begin{equation}\label{eq:areadef}
-  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) = A(0).
+  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) 
+      = A(0) = \sqrt{N} \sum_a \CI_a \tau_a.
 \end{equation}
@@
 
-\newcommand{\aav}[1]{\mathopen{}\left\lAngle #1 \right\rAngle\mathclose{}}
-\newcommand{\nnrm}[1]{\mathopen{}\left\Vvert #1 \right\Vvert\mathclose{}}
+% \newcommand{\aav}[1]{\mathopen{}\left\lAngle #1 \right\rAngle\mathclose{}}
+% \newcommand{\nnrm}[1]{\mathopen{}\left\Vvert #1 \right\Vvert\mathclose{}}
 
@@
 \begin{equation}\label{eq:innereq}
-  \av{\mathbf{u}, \mathbf{v}} = \sum_i \eq_i u_i^\ast v_i ,
+  \av{\mathbf{u}, \mathbf{v}} = \sum_i \eqm_i u_i^\ast v_i ,
   \qquad
@@
 \begin{equation}\label{eq:innerflux}
-  \aav{\mathbf{F}, \mathbf{G}} = \sum_{i \neq j} \F_{ij} F_{ij}^\ast G_{ij} ,
+  \aav{\mathbf{F}, \mathbf{G}} = \sum_{i \neq j} \Fm_{ij} F_{ij}^\ast G_{ij} ,
   \qquad
@@
 
-When the process satisfies detailed balance, \ie \(\F_{ij}\) is symmetric, we can relate these inner products to quantities of interest.
+When the process satisfies detailed balance, \ie \(\Fm_{ij}\) is symmetric, we can relate these inner products to quantities of interest.
 
@@
 \begin{equation*}
-  \av{\mathbf{u}, -\MM\mathbf{v}} = \sum_{i \neq j} \eq_i u_i^\ast \MM_{ij} (v_i - v_j),
+  \av{\mathbf{u}, -\MM\mathbf{v}} = \sum_{i \neq j} \eqm_i u_i^\ast \MMm_{ij} (v_i - v_j),
 \end{equation*}
@@
   \av{\mathbf{u}, -\MM \mathbf{v}} 
-      = \frac{1}{2}\sum_{i \neq j} \F_{ij} (u_i^\ast - u_j^\ast) (v_i - v_j) 
+      = \frac{1}{2}\sum_{i \neq j} \Fm_{ij} (u_i^\ast - u_j^\ast) (v_i - v_j) 
       = \frac{1}{2} \aav{\boldsymbol{\Delta}\mathbf{u}, \boldsymbol{\Delta}\mathbf{v}}.
@@
   \CI(\mathbf{u}) = \frac{(\eq \enc \mathbf{u}) (\mathbf{u}^\dag \w)}{\mathbf{u}^\dag \mathbf{u}}
-      = \frac{\sum_{i \neq j} \eq_i \enc_{ij} (u_j - u_i) \sum_k \eq_k u_k \w_k} 
-             {\sum_{l} \eq_l u_l^2} .
+      = \frac{\sum_{i \neq j} \eqm_i \encm_{ij} (u_j - u_i) \sum_k \eqm_k u_k \wm_k} 
+             {\sum_{l} \eqm_l u_l^2} .
 \end{equation*}
@@
   r \eq \enc \mathbf{u} &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
-        \F\pot_{\CA_\alpha \CA_\alpha^c} - \F\dep_{\CA_\alpha \CA_\alpha^c} 
-      - \F\pot_{\CA_\alpha^c \CA_\alpha} + \F\dep_{\CA_\alpha^c \CA_\alpha} }
+        \Fm\pot_{\CA_\alpha \CA_\alpha^c} - \Fm\dep_{\CA_\alpha \CA_\alpha^c} 
+      - \Fm\pot_{\CA_\alpha^c \CA_\alpha} + \Fm\dep_{\CA_\alpha^c \CA_\alpha} }
+  %
+  \; \text{where} 
     \\
     &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
-        \F_{\CA_\alpha \CA_\alpha^c} + \F_{\CA_\alpha^c \CA_\alpha} 
-        - 2\F\dep_{\CA_\alpha \CA_\alpha^c} - 2\F\pot_{\CA_\alpha^c \CA_\alpha} }
-  \end{aligned}
-  %
-  \quad \text{where} \quad
+        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} 
+        - 2\Fm\dep_{\CA_\alpha \CA_\alpha^c} - 2\Fm\pot_{\CA_\alpha^c \CA_\alpha} } ,
+  \end{aligned}
   %
   \begin{aligned}[t]
-    \CA_\alpha &= \set{i}{u_i < \alpha} 
-    , \\
+    \CA_\alpha &= \set{i}{u_i < \alpha} , 
+    \\
     \F^\nu_{\CA \CB} &= r f^\nu \sum_{\substack{i \in \CA \\ j \in \CB}} 
-                        \prn{\M^\nu_{ij} - \delta_{ij}}
-    .
-  \end{aligned}
-\end{equation*}
-%
-In exactly the same way as the proof of \cref{eq:initbnd}, we can argue that the two terms with the negative signs are indeed non-positive as they do not involve the negative diagonal elements of~\( \F_{ij}^\nu \).
-Therefore the other terms provide an upper bound.
+                        \eqm_i \prn{\MMdm^\nu_{ij} - \delta_{ij}} .
+  \end{aligned}
+\end{equation*}
+%
+In exactly the same way as the proof of \cref{eq:initbnd}, we can argue that the two terms with the negative signs are indeed non-positive as they do not involve the negative diagonal elements of~\( \Fm_{ij}^\nu \).
+Therefore the other terms provide an upper bound:
 %
@@
   r \nrm{\mathbf{u}}^2 \CI(\mathbf{u}) 
-      &\leq \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
-        \F_{\CA_\alpha \CA_\alpha^c} + \F_{\CA_\alpha^c \CA_\alpha} } 
-        \av{\mathbf{u}, \w}
-  \\  &= \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \F_{ij} \brk{
+      &\leq \av{\mathbf{u}, \w} \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
+        \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} } 
+  \\  &= \av{\mathbf{u}, \w} \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \Fm_{ij} \brk{
             \vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
           + \vartheta(u_i - \alpha) \vartheta(\alpha - u_j)
-        } \av{\mathbf{u}, \w}
-  \\  &= \sum_{i \neq j} \F_{ij} \abs{u_j - u_i} \av{\mathbf{u}, \w}
-  \\  &= \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem} \av{\mathbf{u}, \w} .
+        }
+  \\  &= \av{\mathbf{u}, \w} \sum_{i \neq j} \Fm_{ij} \abs{u_j - u_i}
+  \\  &= \av{\mathbf{u}, \w} \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem}.
 \end{aligned}
@@
 \begin{equation*}
-  r \CI(\mathbf{u}) \leq \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem} 
-                          \nrm{\mathbf{u}} \nrm{\w}.
-\end{equation*}
-%
-We also have \( \nrm{\w}^2 = \sum_i \eq_i \w_i^2 = 1 \).
-Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eq_i \MM_{ij} \leq r \sum_i \eq_i = r\).
+  r \nrm{\mathbf{u}}^2 \CI(\mathbf{u}) \leq \nrm{\mathbf{u}} \nrm{\w} 
+      \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem} .
+\end{equation*}
+%
+We also have \( \nrm{\w}^2 = \sum_i \eqm_i \wm_i^2 = 1 \).
+Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \MMm_{ij} \leq r \sum_i \eqm_i = r\).
 Then, remembering \cref{eq:numrange},
@@
 \begin{equation}\label{eq:sqrtnumrange}
-  \sqrt{r} \CI(\mathbf{u}) \leq \sqrt{2\mathcal{W}(-\MM, \mathbf{u})}.
+  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\MM \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
+      = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})}.
 \end{equation}
@@
   \text{assuming} \quad 
-  \eq_i \MM_{ij} = \eq_j \MM_{ji},
+  \eqm_i \frgm_{ij} = \eqm_j \frgm_{ji},
   \qquad \text{we have} \quad
@@
 %
+Numerical maximisation of the quantity above produces the same result as in \cref{sec:initial}:
+the two-state model with deterministic transitions.
+This model has \( \CI_1 = \tau_1 = 1 \), suggestint that there's a factor of 2 looseness in this bound.
 
@@
 \begin{equation*}
-  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a},
+  \snr(t) = \sqrt{N} \sum_a \CI_a \e^{-rt/\tau_a},
   \qquad
-  \snrb(\tau) = \sqrt{N}\sum_a \frac{\CI_a}{1+r\tau/\tau_a},
+  \snrb(\tau) = \sqrt{N} \sum_a \frac{\CI_a}{1+r\tau/\tau_a},
 \end{equation*}
@@
 \begin{equation*}
-  \CI_a = (\eq \ench \evr^a) (\evl^a \w),
+  \CI_a = (\eq \enc \evr^a) (\evl^a \w),
   \qquad
@@
 
+    \newcolumntype{B}{R@{\,}C@{\,}L}
+% \(
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{BACCCCCC}
+\toprule
+  \multicolumn{5}{c}{Conditions} & \CI_a & \tau_a 
+    & \mu_\CI & \mu_a & \mu_\CA 
+    & \snrb(\tau) \\
+\midrule
+  0 &\leq r \tau &\leq 2 & M &\geq 3 & 1 & 2 
+    & \frac{4 - 2r \tau}{[2 + r \tau]^2} & \frac{2\sqrt{2}r\tau}{[2 + r \tau]^2} & 0 
+    & \frac{2 \sqrt{N}}{2 + r \tau} 
+  \\
+  2 &< r \tau &< \frac{(M-1)^2}{2} & M &> 3 & \sqrt{\frac{2}{\tau}} & \tau 
+    & 0 & \frac{1}{2\sqrt{r \tau}} & 0 
+    & \sqrt{\frac{N}{2 r \tau}} 
+  \\ \addlinespace[0.2ex]
+  \frac{(M-1)^2}{2} &\leq r \tau & & M &\geq 3 & \frac{2}{M-1} & \frac{(M-1)^2}{2} 
+    & 0 & \frac{\sqrt{2}(M-1)^3}{[(M-1)^2 + 2r \tau]^2} 
+    & \frac{4r \tau - 2(M-1)^2}{[(M-1)^2 + 2r \tau]^2} 
+    & \frac{2 \sqrt{N} (M-1)}{(M-1)^2 + 2r \tau} 
+  \\
+  &-& & M &\leq 3 & 1 & M - 1 
+    & \frac{(M-1)^2}{(M-1) + r \tau} %- \frac{\mu_a \sqrt{M-1}}{2} 
+    & 0 & \frac{r \tau}{[(M-1) + r \tau]^2} %- \frac{\mu_a}{2 \sqrt{M-1}} 
+    & \frac{\sqrt{N} (M-1)}{(M-1) + r \tau} 
+  \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{The envelope under different conditions. The range of \(\tau\) for the second row does not exist for \(M = 3\), in which case any of the other rows can be used.}
+\label{tab:env}
+\end{table}
+% \)
+
 %================================================================================
@@
 \begin{aligned}
-  \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\etw_i(s) - \etw_j(s))
-        +\sum_{\nu i j} \prn{ f^\nu \mu^\nu_{ij} \M^\nu_{ij} }
-        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i \w_i },
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
+        +\sum_{\nu i j} \prn{ f^\nu \mu^\nu_{ij} \MMdm^\nu_{ij} }
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
 \end{aligned}
@@
 \begin{equation}\label{eq:KTfinite}
-  \pdiff{\CL}{\M^\nu_{ij}} = 0, \quad
+  \pdiff{\CL}{\MMdm^\nu_{ij}} = 0, \quad
   \mu^\nu_{ij} \geq 0, \quad
@@
 \begin{aligned}
-  \frac{1}{f^\nu} \pdiff{\CL}{\Mh^\nu_{mn}} = & \,
-    \eq_m \prn{ \brk{\thb_m(s) - \thb_n(s)}
-     + \prn{\mathbf{c}_m(s) + (-1)^\nu} \brk{\etw_m(s) - \etw_n(s)} } \\
+  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+    \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
+     + \prn{\mathbf{c}_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
      &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     + \lambda \eq_m (\etw_{m}(0)-\etw_{n}(0)) 
+     + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
     = 0.
@@
 \begin{equation*}
-  \mu^\nu_{ij} \M^\nu_{ij} = 0,
+  \mu^\nu_{ij} \MMdm^\nu_{ij} = 0,
   \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} \M^\nu_{ij} 
-      = \mu^\nu_{ii} (1 - \M^\nu_{ii}) = \mu^\nu_{ii}.
+  \sum_{j: j \neq i} \mu^\nu_{ii} \MMdm^\nu_{ij} 
+      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii}) = \mu^\nu_{ii}.
 \end{equation*}
@@
 \begin{aligned}
-  \mu^\nu_{mm} &= - \eq_m \prn{ \alb^\nu_m(s) 
-    + \lambda \omb^\nu_i(0) 
-    + \prn{\mathbf{c}_m(s) + (-1)^\nu} \omb^\nu_m(s)
+  \mu^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
+    + \lambda \ombm^\nu_i(0) 
+    + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
     }, \\
-  \mu^\nu_{mn} &= \eq_m \big( 
-      \brk{ \thb_n(s) - \thb_m(s) - \alb^\nu_m(s) }
-    + \lambda \brk{ \etw_n(0) - \etw_m(0) - \omb^\nu_m(0) } 
-  \\ &\phantom{= \eq_m \big( }
-    + \brk{\mathbf{c}_m(s) + (-1)^\nu} 
-        \brk{ \etw_n(s) - \etw_m(s) - \omb^\nu_m(s) }
+  \mu^\nu_{mn} &= \eqm_m \big( 
+      \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
+    + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
+  \\ &\phantom{= \eqm_m \big( }
+    + \brk{{c}_m(s) + (-1)^\nu} 
+        \brk{ \etwm_n(s) - \etwm_m(s) - \ombm^\nu_m(s) }
      \big), \\
   &\text{where} \quad 
-  \boldsymbol{\alpha}^\nu(s) = (\M^\nu - \I) \thb(s), \qquad
-  \boldsymbol{\omega}^\nu(s) = (\M^\nu - \I) \etw(s).
+  \alb^\nu(s) = (\MMd^\nu - \I) \thb(s), \qquad
+  \omb^\nu(s) = (\MMd^\nu - \I) \etw(s).
 \end{aligned}
@@
 \begin{equation}\label{eq:scale}
-  \zeta * \M \equiv \zeta \M + (1-\zeta)\I .
+  \zeta * \MMd \equiv \zeta \MMd + (1-\zeta)\I .
 \end{equation}
@@
 \begin{equation}\label{eq:laplacescale}
-  \hat{A} \prn{\zeta s;\zeta * \M^\nu} = \hat{A} \prn{s; \M^\nu}.
+  \hat{A} \prn{\zeta s;\zeta * \MMd^\nu} = \hat{A} \prn{s; \MMd^\nu}.
 \end{equation}
@@
 \begin{equation}\label{eq:shiftscale}
-  \zeta {\circledast}_{s} \M 
-      \equiv \zeta \M
+  \zeta {\circledast}_{s} \MMd 
+      \equiv \zeta \MMd
       + (1 - \zeta) \brk{(1 + s)\I - s \onev \eq} .
@@
 \begin{equation}\label{eq:lshiftscale}
-  \hat{A} \prn{s; \M^\nu} 
-    = \hat{A} \prn{0, \shift{\M}^\nu(s)}
-    = \hat{A} \prn{0, \zeta * \shift{\M}^\nu(s)}
-    = \hat{A} \prn{s, \zeta \circledast_s \M^\nu}.
-\end{equation}
-%
-The problem here is that the lower bound, \(\M^\nu_{ij} \geq 0\), 
+  \hat{A} \prn{s; \MMd^\nu} 
+    = \hat{A} \prn{0, \shift{\MMd}^\nu(s)}
+    = \hat{A} \prn{0, \zeta * \shift{\MMd}^\nu(s)}
+    = \hat{A} \prn{s, \zeta \circledast_s \MMd^\nu}.
+\end{equation}
+%
+The problem here is that the lower bound, \(\MMdm^\nu_{ij} \geq 0\), 
 is not invariant under this transformation, unlike the \(s = 0\) case.
@@
 \begin{alignedat}{5}
-  [\zeta \circledast \M^\nu]_{ij}  
-    &= \zeta \M^\nu_{ij} - (1-\zeta) s \eq_j 
+  [\zeta \circledast \MMd^\nu]_{ij}  
+    &= \zeta \MMdm^\nu_{ij} - (1-\zeta) s \eqm_j 
     &&\geq 0 &
   \quad &\implies \quad &
-  \zeta &\geq \frac{s \eq_j}{\M^\nu_{ij} + s \eq_j} &
+  \zeta &\geq \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} &
     &\forall \nu, i, j : i \neq j , \\
-  \sum_{j : i \neq j} [\zeta \circledast \M^\nu]_{ij} 
-    &= \zeta \lb^\nu_i - (1-\zeta) s (1-\eq_i) 
+  \sum_{j : i \neq j} [\zeta \circledast \MMd^\nu]_{ij} 
+    &= \zeta \lbm^\nu_i - (1-\zeta) s (1-\eqm_i) 
     &&\leq 1 &
   &\implies &
-  \zeta &\leq \frac{1 + s (1-\eq_i)}{\lb^\nu_i + s (1-\eq_i)} &
+  \zeta &\leq \frac{1 + s (1-\eqm_i)}{\lbm^\nu_i + s (1-\eqm_i)} &
     \quad &\forall \nu, i ,
@@
 %
-where \(\lb^\nu_i \equiv \sum_{j : i \neq j} \M^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
+where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \MMdm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
 The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
@@
 \begin{equation}\label{eq:fixable}
-  \frac{s \eq_j}{\M^\nu_{ij} + s \eq_j} \leq \frac{1 + s (1-\eq_k)}{\lb^\mu_k + s (1-\eq_k)}
+  \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} \leq \frac{1 + s (1-\eqm_k)}{\lbm^\mu_k + s (1-\eqm_k)}
   \quad \implies \quad
-  \frac{\lb^\mu_k - 1}{1 + s (1-\eq_k)} \leq \frac{\M^\nu_{ij}}{s \eq_j} 
+  \frac{\lbm^\mu_k - 1}{1 + s (1-\eqm_k)} \leq \frac{\MMdm^\nu_{ij}}{s \eqm_j} 
   \quad \forall \mu, \nu, i, j, k : i \neq j.
@@
 %
-Note that when \(\M^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lb^\mu_k \leq 1\), the original diagonal constraint.
+Note that when \(\MMdm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq 1\), the original diagonal constraint.
 Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
@@
 The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
-treat the off-diagonal components of \(\shift{\M}^\nu \) as the degrees of freedom and maximise the area instead.
-The constraints on \(\shift{\M}^\nu_{ij}\) are different from \cref{eq:constri}:
+treat the off-diagonal components of \(\shift{\MMd}^\nu \) as the degrees of freedom and maximise the area instead.
+The constraints on \(\shift{\MMdm}^\nu_{ij}\) are different from \cref{eq:constri}:
 %
 \begin{equation}\label{eq:lshiftconstr}
-  \shift{\M}^\nu_{ij} \geq s \eq_j,
+  \shift{\MMdm}^\nu_{ij} \geq s \eqm_j,
     \quad \forall \; \nu, i, j : i \neq j, \qquad
-  \sum_{j: j \neq i} \shift{\M}^\nu_{ij} \leq 1 + s (1 - \eq_i).
+  \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \leq 1 + s (1 - \eqm_i).
     \quad \forall \; \nu, i.
@@
 %
-The diagonal elements are still given by \( \shift{\M}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\M}^\nu_{ij} \), but now they are allowed to be negative so long as \(\M^\nu_{ii} \geq - s (1 - \eq_i)\).
+The diagonal elements are still given by \( \shift{\MMdm}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \), but now they are allowed to be negative so long as \(\shift{\MMdm}^\nu_{ii} \geq - s (1 - \eqm_i)\).
 
@@
 \begin{aligned}
-  \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\etw_i - \etw_j)
-        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\M^\nu_{ij} -  s (\eq_j - \delta_{ij}) }
-        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i \w_i },
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
+        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\MMdm^\nu_{ij} -  s (\eqm_j - \delta_{ij}) }
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
 \end{aligned}
@@
 \begin{aligned}
-  \frac{1}{f^\nu} \pdiff{\CL}{\Mh^\nu_{mn}} = & \,
-    \eq_m \prn{ \brk{\thb_m - \thb_n}
-     + \prn{\mathbf{c}_m + (-1)^\nu + \lambda} \brk{\etw_m - \etw_n} } \\
+  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+    \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
+     + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
      &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     - \sum_{ij} \mu^\nu_{ij} s f^\nu \eq_m (\fptb_{mj} - \fptb_{nj}) \eq_j
+     - \sum_{ij} \mu^\nu_{ij} s f^\nu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
     = 0.
@@
   \qquad
-  \shift{\alb}^\nu(s) &= (\M^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
+  \shift{\alb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
   \\
-  \shift{\omb}^\nu(s) &= (\M^\nu + (s - 1) \I - s \onev \eq) \etw(0).
+  \shift{\omb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \etw(0).
 \end{aligned}
@@
 \begin{equation*}
-  \mu^\nu_{ij} (\M^\nu_{ij} - s \eq_j) = 0,
+  \mu^\nu_{ij} (\MMdm^\nu_{ij} - s \eqm_j) = 0,
   \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} (\M^\nu_{ij} - s \eq_j)
-      = \mu^\nu_{ii} (1 - \M^\nu_{ii} - s (1 - \eq_i)) = \mu^\nu_{ii},
+  \sum_{j: j \neq i} \mu^\nu_{ii} (\MMdm^\nu_{ij} - s \eqm_j)
+      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
 \end{equation*}
@@
 \begin{aligned}
-  \mu^\nu_{mm} &= - \eq_m \prn{ \shift{\alb}^\nu_m(s) 
-    + \brk{ \mathbf{c}_m + (-1)^\nu + \lambda } \shift{\omb}^\nu_m(s)
+  \mu^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
     }, \\
-  \mu^\nu_{mn} &= \eq_m \prn{ 
-      \brk{ \shift{\thb}^\nu_n(s) - \shift{\thb}^\nu_m(s) - \shift{\alb}^\nu_m(s) }
-  % \\ & \phantom{= - \eq_m \big( } 
-    + \brk{ \mathbf{c}_m + (-1)^\nu + \lambda } 
-        \brk{ \etw_n - \etw_m - \shift{\omb}^\nu_m(s) }
+  \mu^\nu_{mn} &= \eqm_m \prn{ 
+      \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
+  % \\ & \phantom{= - \eqm_m \big( } 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } 
+        \brk{ \etwm_n - \etwm_m - \shift{\ombm}^\nu_m(s) }
      }, \\
@@
 Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_uniform}.
-Looking at \eqref{eq:larea}, the only thing we need is differences in $\etw_i(s)$.
+Looking at \eqref{eq:larea}, the only thing we need is differences in $\etwm_i(s)$.
 In fact, given that the only nonzero elements of $\enc$ are
@@
 \begin{equation}\label{eq:unienc}
-  \enc_{ii+1}=\frac{1}{2} \quad\text{for}\quad i<2m,
+  \encm_{ii+1}=\frac{1}{2} \quad\text{for}\quad i<2m,
   \qquad
-  \enc_{ii-1}=-\frac{1}{2} \quad\text{for}\quad i>1,
-\end{equation}
-%
-and $\eq_i = \frac{1}{2m}$, we have
+  \encm_{ii-1}=-\frac{1}{2} \quad\text{for}\quad i>1,
+\end{equation}
+%
+and $\eqm_i = \frac{1}{2m}$, we have
 %
 \begin{equation}\label{eq:uniareaeta}
-  \hat{A}(s) = \frac{\etw_1(s) - \etw_{2m}(s)}{2m}.
+  \hat{A}(s) = \frac{\etwm_1(s) - \etwm_{2m}(s)}{2m}.
 \end{equation}
@@
 %
-As we are only interested in differences of the $\etw_i(s)$, we can drop the constant.
+As we are only interested in differences of the $\etwm_i(s)$, we can drop the constant.
 Then
@@
 \begin{equation}\label{eq:detarecur}
-  \etw(s) = -\frac{(\pib\etw(s))\onev+\w}{s} + \delta\eta(s),
+  \etw(s) = -\frac{(\pib\etw(s))\onev+\w}{s} + \delta \boldsymbol{\eta}(s),
   \qquad
-  (\frg-s\I)\delta\eta(s) = \frac{\frg\w}{s}.
+  (\frg-s\I)\delta \boldsymbol{\eta}(s) = \frac{\frg\w}{s}.
 \end{equation}
@@
 \begin{equation}\label{eq:unilaplace}
-  A(s) = \frac{\etw_1(s)}{m}
+  A(s) = \frac{\etwm_1(s)}{m}
    = \frac{1}{m}\prn{\frac{1}{s}+B+C}
@@
   \begin{aligned}
-  \enc_{2m-1,2m} = -\enc_{2,1} &= \frac{(1-\varepsilon)}{2},
+  \encm_{2m-1,2m} = -\encm_{2,1} &= \frac{(1-\varepsilon)}{2},
   &\qquad
-  \eq_1 = \eq_M &= \frac{(1-\varepsilon)}{2(m-\varepsilon)},
+  \eqm_1 = \eqm_M &= \frac{(1-\varepsilon)}{2(m-\varepsilon)},
   \\ &&
-  \eq_i &=  \frac{1}{2(m-\varepsilon)}
-  \quad\text{for}\quad 1<i<M2m.
+  \eqm_i &=  \frac{1}{2(m-\varepsilon)}
+  \quad\text{for}\quad 1<i<2m.
   \end{aligned}
@@
 \begin{equation}\label{eq:shortenareaeta}
-  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_{2m}(s)) 
-                      + \varepsilon(\etw_2(s)-\etw_{2m-1}(s))}
+  \hat{A}(s) = \frac{(1-\varepsilon)(\etwm_1(s)-\etwm_{2m}(s)) 
+                      + \varepsilon(\etwm_2(s)-\etwm_{2m-1}(s))}
                     {2(m-\varepsilon)}
-       = \frac{2(1-\varepsilon) \etw_1(s) + 2\varepsilon\,\etw_2(s)}
+       = \frac{2(1-\varepsilon) \etwm_1(s) + 2\varepsilon\,\etwm_2(s)}
               {2(m-\varepsilon)}.
@@
 
-The $\etw_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
+The $\etwm_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
 %
@@
   \begin{aligned}
-  \enc_{1,2} = -\enc_{2m,2m-1} &= \frac{1-\varepsilon}{2},
+  \encm_{1,2} = -\encm_{2m,2m-1} &= \frac{1-\varepsilon}{2},
   &\quad
-  \eq_1 = \eq_{2m} &= \frac{1}{2(m-(m-1)\varepsilon)},
+  \eqm_1 = \eqm_{2m} &= \frac{1}{2(m-(m-1)\varepsilon)},
   \\ &&
-  \eq_i &=  \frac{1-\varepsilon}{2(m-(m-1)\varepsilon)}
+  \eqm_i &=  \frac{1-\varepsilon}{2(m-(m-1)\varepsilon)}
   \quad\text{for}\quad 1<i<2m.
@@
 \begin{equation}\label{eq:stickyareaeta}
-  \hat{A}(s) = \frac{(1-\varepsilon)(\etw_1(s)-\etw_M(s))}
+  \hat{A}(s) = \frac{(1-\varepsilon)(\etwm_1(s)-\etwm_M(s))}
                     {2(m-(m-1)\varepsilon)}
-       = \frac{2(1-\varepsilon)\, \etw_1(s)}{2(m-(m-1)\varepsilon)}.
-\end{equation}
-%
-
-The $\etw_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
+       = \frac{2(1-\varepsilon)\, \etwm_1(s)}{2(m-(m-1)\varepsilon)}.
+\end{equation}
+%
+
+The $\etwm_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
 %

### 2020-10-22 Code/Python/complex_synapse/optimise/optimise.py

@@ -488,6 +488,8 @@ class OptimProblem:
         """Verify that solution satisfies constraints"""
         # maxcv = constr_violation(prob, result)
         # return maxcv < prob.get('tol', 1e-3)
+        if not np.isfinite(result.fun):
+            return False
         itol = self.opts.extra.get('tol', 1e-3)
         maxcv = getattr(result, 'constr_violation', None)
         if maxcv is not None:
@@ -697,6 +699,8 @@ def optim_laplace(rate: float, prob: Optional[OptimProblem] = None,
     else:
         prob.update_rate(rate)
     res = first_good(prob)
+    if not prob.verify_solution(res):
+        res.fun = 0
     for _ in _it.dcount('repeats', prob.opts.repeats, disp_step=1):
         prob.update_init()
         new_res = sco.minimize(**prob.for_scipy())

### 2020-10-22 Code/Python/complex_synapse/sqrt.py

@@ -138,10 +138,10 @@ def get_valid(nst: int, frac: float = 0.5, rng: np.random.Generator = bld.RNG):
 class SynapseSqrt(sb.SynapseParam, sm.SynapseModel):
     """class for testing sqrt bound"""
     _param: la.lnarray
-    _ind: Optional[int]
     _peq: la.lnarray
     _pneq: la.lnarray
     _flux: la.lnarray
+    _ind: Optional[int]
     _changed: bool
     _saved: Tuple[la.lnarray, ...]
 
@@ -149,8 +149,12 @@ class SynapseSqrt(sb.SynapseParam, sm.SynapseModel):
     def __init__(self, *args, **kwds) -> None:
         super().__init__(*args, **kwds)
         self._param = la.zeros((self.nstate, self.nstate))
+        self._flux = la.zeros((self.nstate, self.nstate))
+        self._peq = la.zeros(self.nstate)
+        self._pneq = la.zeros(self.nstate)
         self._ind = None
         self._changed = True
+        self._saved = (la.array(0), la.array(0), la.array(0))
 
     def biggest(self) -> float:
         """max(I_a * sqrt(t_a / 2))"""
@@ -167,13 +171,13 @@ class SynapseSqrt(sb.SynapseParam, sm.SynapseModel):
         if self._ind is None:
             self.biggest()
         ind = self._ind
-        qas, init_ab, dq_ab, di_ab = self._eig()
+        qas, init_ab, dq_ab, di_aa = self._eig()
         i_by_q = init_ab / (qas.c - qas)
         i_by_q[np.diag_indices(qas.size)] = np.diagonal(init_ab) / qas
         grad = dq_ab[..., ind] @ i_by_q[ind, :]
         grad -= dq_ab[..., ind, :] @ i_by_q[:, ind]
-        grad -= i_by_q[ind, ind] * dq_ab[..., ind, ind] / 2
-        grad += di_ab[..., ind, ind]
+        grad -= dq_ab[..., ind, ind] * i_by_q[ind, ind] / 2
+        grad += di_aa[..., ind]
         grad = grad.ravel()
         return - (grad[:-1] - grad[-1]) / np.sqrt(2 * qas[ind])
 
@@ -184,17 +188,17 @@ class SynapseSqrt(sb.SynapseParam, sm.SynapseModel):
         self._changed = False
         self._ind = None
         frac = self.frac[0]
-        evals, evc = sla.eigh(la.eye(self.nstate) - self._flux,
[…]

### 2020-10-23 Notes/laplace.tex
plotting new envelope

@@
 %
+Note that the first two terms of the integrand are of the form \( \Fm_{\CA \CB} = \Fm_{\CA \CB}\pot + \Fm_{\CA \CB}\dep \).
 In exactly the same way as the proof of \cref{eq:initbnd}, we can argue that the two terms with the negative signs are indeed non-positive as they do not involve the negative diagonal elements of~\( \Fm_{ij}^\nu \).
@@
 \begin{aligned}
-  r \nrm{\mathbf{u}}^2 \CI(\mathbf{u}) 
-      &\leq \av{\mathbf{u}, \w} \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
+  \CI(\mathbf{u}) 
+      &\leq \frac{\av{\mathbf{u},\w}}{r \nrm{\mathbf{u}}^2} 
+        \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
         \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} } 
-  \\  &= \av{\mathbf{u}, \w} \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \Fm_{ij} \brk{
+  \\  &= \frac{\av{\mathbf{u},\w}}{r \nrm{\mathbf{u}}^2} 
+        \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \Fm_{ij} \brk{
             \vartheta(u_j - \alpha) \vartheta(\alpha - u_i) 
@@
         }
-  \\  &= \av{\mathbf{u}, \w} \sum_{i \neq j} \Fm_{ij} \abs{u_j - u_i}
-  \\  &= \av{\mathbf{u}, \w} \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem}.
+  \\  &= \frac{\av{\mathbf{u},\w}}{r \nrm{\mathbf{u}}^2} 
+      \sum_{i \neq j} \Fm_{ij} \abs{u_j - u_i}
+  \\  &= \frac{\av{\mathbf{u},\w} \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem}}
+              {r \nrm{\mathbf{u}}^2}.
 \end{aligned}
@@
 \begin{equation*}
-  r \nrm{\mathbf{u}}^2 \CI(\mathbf{u}) \leq \nrm{\mathbf{u}} \nrm{\w} 
-      \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem} .
+  \CI(\mathbf{u}) \leq \frac{\nrm{\w} \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem}}
+                            {r \nrm{\mathbf{u}}} .
 \end{equation*}
@@
 the two-state model with deterministic transitions.
-This model has \( \CI_1 = \tau_1 = 1 \), suggestint that there's a factor of 2 looseness in this bound.
+This model has \(\CI_1 = \tau_1 = 1\), suggesting that there's a factor of 2 looseness in this bound.
+
+We can look at the various inequalities used in this derivation to see which models could saturate them:
+%
+\begin{equation*}
+\begin{aligned}
+  \Fm_{\CA_\alpha \CA_\alpha^c}\dep + \Fm_{\CA_\alpha^c \CA_\alpha}\pot &\geq 0 , &
+  \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem} 
+                        &\leq \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem} , &
+  \av{\mathbf{u},\w} &\leq \nrm{\mathbf{u}} \nrm{\w} , &
+  \sum_{j: j \neq i} \frg_{ij} &\leq 1.
+\end{aligned}
+\end{equation*}
+%
+To saturate the first inequality it helps to arrange the states in order of increasing \(u_i\).
+For any \(\alpha\), the sets \(\CA_\alpha\) and \(\CA_\alpha^c\) will be the states to the left and right of \(u = \alpha\) respectively.
+We see that the inequality is saturated iff \(\W \pot \) contains rightwards transitions only and \(\W \dep \) only contains leftwards transitions only.
+
+The second inequality is saturated iff \(\abs{\boldsymbol{\Delta}\mathbf{u}} \propto \onem \), but this need only hold for pairs of states for which \(\Fm_{ij} \neq 0\).
+This can be achieved if \(u_i\) varies linearly with \(i\), and only nearest neighbour transitions are allowed
+(we are allowed states that are degenerate if they are not connected by any nonzero transitions).
+This almost forces the serial topology on us (if we can argue that degenerate states have the same outgoing transition rates, they would be lumpable).
+
+The third inequality is saturated iff \(\mathbf{u} \propto \w\).
+This contradicts the requirements of the previous paragraph unless \(M=2\).
+
+The fourth inequality can only be saturated if the outgoing potentitation rates and depression rates separately sum to one for each state.
+This does force degenerate states to be lumpable, and we are left with a serial topology with deterministic transitions.
+However, this cannot be done for the two end states without violating the requirements for saturating one of the other inequalities.
+
+So we see that all four inequalities cannot be simultaneously saturated, and the bound must be loose.
 
@@
 
-This envelope is shown in \autoref{fig:envproven}.
+This envelope is shown in \cref{fig:envproven}\ref{fig:provengen}.
+
+If we know that the optimal model for any timescale satisfies detailed balance, we have one more constraint, \cref{eq:sqrtbound}.
+Then the Lagrangian reads
+%
+\begin{equation}\label{eq:envDBlagrangian}
+  \CL = \sum_a \frac{\CI_a}{1+r\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} 
+      + \sum_a \mu_a \prn{\sqrt{2} - \CI_a \sqrt{\tau_a}} + \mu_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
+\end{equation}
+%
+The solution to \cref{eq:envKTcond} takes different forms for different ranges of \(\tau\) and \(M\),
+as described in \cref{tab:env}.
+The resulting envelope is (for \(M > 3\))
+%
+\begin{equation}\label{eq:envDB}
+  \snrb(\tau) \leq 
+  \left\{ 
+  \begin{alignedat}{3}
+    &\frac{2 \sqrt{N}}{2 + r \tau}
+      , &
+      0 &\leq r \tau &&\leq 2
+      , \\
+    &\sqrt{\frac{N}{2 r \tau}}
+      , &
+      2 &< r \tau &&< \frac{(M-1)^2}{2}
+      , \\
+    &\frac{2 \sqrt{N} (M-1)}{(M-1)^2 + 2r \tau}
+      , &\qquad
+      \frac{(M-1)^2}{2} &\leq r \tau 
+      .
+  \end{alignedat}
+  \right.
+\end{equation}
+%
+This envelope is shown in \cref{fig:envproven}\ref{fig:provendetbal}.
+For \(M \leq 3\), \cref{eq:env} still applies.
 
@@
   % Requires \usepackage{graphicx}
-  \includegraphics[width=0.8\linewidth]{LenvProven.pdf}
+  \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.4\linewidth]{LenvProven.pdf}}\label{fig:provengen}
+  \item\aligntop{\includegraphics[width=0.4\linewidth]{LenvConj.pdf}}\label{fig:provendetbal}
+  \end{myenuma}
   \caption[Proven envelope for the signal-to-noise ratio]
-  {The memory curve envelope that we can prove, \eqref{eq:env}.
+  {\ref{fig:provengen} The memory curve envelope that we can prove, \cref{eq:env}.
+  \ref{fig:provendetbal} Conjectured memory curve envelope when we assume the optimal models satisfy detailed balance, \cref{eq:envDB}.
   The number of synapses $N=10^4$ and the number of states $M=12$.
@@
     & \mu_\CI & \mu_a & \mu_\CA 
-    & \snrb(\tau) \\
+    % & \snrb(\tau) 
+  \\
 \midrule
@@
     & \frac{4 - 2r \tau}{[2 + r \tau]^2} & \frac{2\sqrt{2}r\tau}{[2 + r \tau]^2} & 0 
-    & \frac{2 \sqrt{N}}{2 + r \tau} 
+    % & \frac{2 \sqrt{N}}{2 + r \tau} 
   \\
@@
     & 0 & \frac{1}{2\sqrt{r \tau}} & 0 
-    & \sqrt{\frac{N}{2 r \tau}} 
+    % & \sqrt{\frac{N}{2 r \tau}} 
   \\ \addlinespace[0.2ex]
@@
     & \frac{4r \tau - 2(M-1)^2}{[(M-1)^2 + 2r \tau]^2} 
-    & \frac{2 \sqrt{N} (M-1)}{(M-1)^2 + 2r \tau} 
+    % & \frac{2 \sqrt{N} (M-1)}{(M-1)^2 + 2r \tau} 
   \\
@@
     & 0 & \frac{r \tau}{[(M-1) + r \tau]^2} %- \frac{\mu_a}{2 \sqrt{M-1}} 
-    & \frac{\sqrt{N} (M-1)}{(M-1) + r \tau} 
+    % & \frac{\sqrt{N} (M-1)}{(M-1) + r \tau} 
   \\
@@
 \end{center}
-\caption{The envelope under different conditions. The range of \(\tau\) for the second row does not exist for \(M = 3\), in which case any of the other rows can be used.}
+\caption{The solution to the maximum SNR \cref{eq:envKTcond} for the Lagrangian in \cref{eq:envDBlagrangian}, under different conditions. 
+The resulting envelope is \cref{eq:envDB}.
+The range of \(\tau\) for the second row does not exist for \(M = 3\), in which case any of the other rows can be used.}
 \label{tab:env}
@@
 
+Note that models with the serial topology must satisfy detailed balance.
+Therefore, this conjecture implies that the square-root bound of \cref{sec:sqrt} applies, implying the stricter envelope of \cref{eq:envDB}.
+In \cref{fig:envnum}, we can see that this envelope is close to tight, being off by a constant factor.
+
 \begin{figure}[tb]
@@
   The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
-  The blue line indicates the proven envelope, \eqref{eq:env}.}\label{fig:envnum}
+  The dotted blue line indicates the proven envelope, \cref{eq:env} and the solid blue line is the conjectured envelope assuming detailed balance, see \cref{eq:envDB}.}\label{fig:envnum}
 \end{figure}
@@
   \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv07.eps}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv13.eps}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv21.eps}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv27.eps}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv36.eps}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv51.eps}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=7]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=13]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=21]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=27]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=36]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=50]{envelope.pdf}}
+  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv07.eps}}
+  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv13.eps}}
+  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv21.eps}}
+  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv27.eps}}
+  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv36.eps}}
+  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv51.eps}}
   \end{myenuma}

### 2020-10-23 Code/Python/complex_synapse/optimise/__init__.py

@@ -7,7 +7,7 @@ from .optimise import (ModelOptions, OptimOptions, OptimProblem,
                        heuristic_envelope_laplace, normal_problem,
                        optim_laplace, optim_laplace_range,
                        proven_envelope_laplace, reoptim_laplace_range,
-                       shifted_problem)
+                       shifted_problem, equlibrium_envelope_laplace)
 from .plot import GraphPlots, load_data, save_data
 from .synapse_opt import SynapseOptModel
 from .video import (EnvelopeFig, ModelPlots, VideoLabels, VideoLayout,
@@ -21,4 +21,4 @@ assert any((True, ModelOptions, ProblemOptions, OptimOptions, VideoLabels,
 assert any((True, load_data, save_data, animate, normal_problem,
             shifted_problem, optim_laplace, optim_laplace_range,
             reoptim_laplace_range, check_cond_range, proven_envelope_laplace,
-            heuristic_envelope_laplace))
+            heuristic_envelope_laplace, equlibrium_envelope_laplace))

### 2020-10-23 Code/Python/complex_synapse/optimise/optimise.py

@@ -788,14 +788,14 @@ def check_cond_range(rates: np.ndarray, models: np.ndarray,
 
 
 def proven_envelope_laplace(rate: Data, nst: int) -> Data:
-    """Theoretical envelope for Laplace transform
+    """Theoretical envelope for Laplace transform.
 
     Parameters
     ----------
-    rates : Number or ndarray
-        Rate parameter of Laplace transform
+    rates : Number|ndarray
+        Rate parameter of Laplace transform.
     nst : int
-        Number of states
+        Number of states.
 
     Returns
     -------
@@ -805,15 +805,40 @@ def proven_envelope_laplace(rate: Data, nst: int) -> Data:
     return (nst - 1) / (rate * (nst - 1) + 1)
 
 
-def heuristic_envelope_laplace(rate: Data, nst: int) -> Data:
-    """Heuristic envelope for Laplace transform
+def equlibrium_envelope_laplace(rate: Data, nst: int) -> Data:
+    """Theoretical envelope for Laplace transform assuming detailed balance.
 
     Parameters
     ----------
-    rate : Number or ndarray
-        Rate parameter of Laplace transform
+    rates : Number|ndarray
+        Rate parameter of Laplace transform.
     nst : int
-        Number of states
+        Number of states.
+
+    Returns
+    -------
+    envelope : Data
+        Putative maximum A(s) for all models.
+    """
+    if nst < 3:
+        return (nst - 1) / (rate * (nst - 1) + 1)
+    s_two, s_sticky = 0.5, 2 / (nst-1)**2
+    i_a = np.sqrt(s_sticky/s_two)
+    env = 0.5 / np.sqrt(s_two * rate)
+    env = np.where(rate > s_two, 1 / (rate + s_two), env)
+    env = np.where(rate < s_sticky, i_a / (rate + s_sticky), env)
+    return env
+
+
+def heuristic_envelope_laplace(rate: Data, nst: int) -> Data:
+    """Heuristic envelope for Laplace transform.
+
+    Parameters
+    ----------
+    rate : Number|ndarray
+        Rate parameter of Laplace transform.
+    nst : int
+        Number of states.
 
     Returns
     -------

### 2020-10-23 Code/Python/complex_synapse/optimise/plot.py

@@ -2,6 +2,7 @@
 """
 from __future__ import annotations
 import typing as ty
+from typing import Sequence
 
 import matplotlib as mpl
 import matplotlib.pyplot as plt
@@ -157,7 +158,7 @@ def all_data(nst: int, sss: Array, **opts) -> ty.Tuple[Arrays, Arrays]:
 # =============================================================================
 
 
-def mem_plot(sss: Array, **mem: ty.Dict[str, Array]) -> ty.Tuple[Figure, Axes]:
+def mem_plot(sss: Array, **mem: ty.Dict[str, Array]) -> ty.Tuple[Figure, Axes, Line]:
     """Plot memory curves
 
     Parameters
@@ -175,11 +176,12 @@ def mem_plot(sss: Array, **mem: ty.Dict[str, Array]) -> ty.Tuple[Figure, Axes]:
         Axes containing plots.
     """
     fig, axs = plt.subplots()
+    lns = []
     for label, mem in mem.items():
-        axs.loglog(1/sss, mem * sss, label=label)
+        lns.extend(axs.loglog(1/sss, mem * sss, label=label))
     axs.set_xlabel(r"Time, $r\tau$")
     axs.set_ylabel("SNR")
-    return fig, axs
+    return fig, axs, lns
 
 
 
@@ -204,7 +206,7 @@ def theory_plot(nst: int, sss: Array) -> Figure:
         The figure object containing the plot.
     """
     env_th = cso.proven_envelope_laplace(sss, nst)
-    fig, axs = mem_plot(sss, theory=env_th)
+    fig, axs, _ = mem_plot(sss, theory=env_th)
     t_pts = np.array([1e-1, 0.9e1, 1.1e1, 1e4])
     axs.loglog(t_pts[:2], np.ones(2), 'k:')
     axs.loglog(t_pts[-2:], (nst-1) / t_pts[-2:], 'k:')
@@ -219,6 +221,52 @@ def theory_plot(nst: int, sss: Array) -> Figure:
     axs.text(1/sss[ind], env_th[ind] * sss[ind],
              r"$\frac{\sqrt{N}(M-1)}{r\tau + (M-1)}$", fontsize=24,
              ha='right', va='top')
+    mplt.clean_axes(axs)
+    return fig
+
+
+def equilibrium_plot(nst: int, sss: Array) -> Figure:
+    """Theoretical envelope plot
+
+    Parameters
[…]

### 2020-10-27 Notes/laplace.tex
typos & tweaks to envelope presentation

@@
 The third inequality is saturated iff \(\mathbf{u} \propto \w\).
-This contradicts the requirements of the previous paragraph unless \(M=2\).
-
-The fourth inequality can only be saturated if the outgoing potentitation rates and depression rates separately sum to one for each state.
+This contradicts the requirements of the previous paragraph unless \(M=2\), or if it is a multistate model rather than serial.
+
+The fourth inequality can only be saturated if, for each state, the outgoing potentitation rates and depression rates separately sum to one.
 This does force degenerate states to be lumpable, and we are left with a serial topology with deterministic transitions.
-However, this cannot be done for the two end states without violating the requirements for saturating one of the other inequalities.
+However, this cannot be done for the two end states without violating the requirements for saturating the first inequality.
 
@@
   \tau_1 = M-1, \\
-  \mu_\CI = \frac{(M-1)^2}{(r\tau + (M-1))^2}, \qquad
-  \mu_\CA = \frac{r\tau}{(r\tau+(M-1))^2}. \qquad
+  \mu_\CI = \frac{(M-1)^2}{[r\tau + (M-1)]^2}, \qquad
+  \mu_\CA = \frac{r\tau}{[r\tau+(M-1)]^2}. \qquad
 \end{gathered}
@@
 
-This envelope is shown in \cref{fig:envproven}\ref{fig:provengen}.
+This envelope is shown in \cref{fig:envproven}(\ref{fig:provengen}).
 
@@
 The solution to \cref{eq:envKTcond} takes different forms for different ranges of \(\tau\) and \(M\),
-as described in \cref{tab:env}.
+as described in \cref{tab:envDB}.
 The resulting envelope is (for \(M > 3\))
@@
 %
-This envelope is shown in \cref{fig:envproven}\ref{fig:provendetbal}.
+This envelope is shown in \cref{fig:envproven}(\ref{fig:provendetbal}).
 For \(M \leq 3\), \cref{eq:env} still applies.
@@
   \caption[Proven envelope for the signal-to-noise ratio]
-  {\ref{fig:provengen} The memory curve envelope that we can prove, \cref{eq:env}.
-  \ref{fig:provendetbal} Conjectured memory curve envelope when we assume the optimal models satisfy detailed balance, \cref{eq:envDB}.
-  The number of synapses $N=10^4$ and the number of states $M=12$.
+  {(\ref{fig:provengen}) The memory curve envelope that we can prove, \cref{eq:env}.
+  (\ref{fig:provendetbal}) Conjectured memory curve envelope when we assume the optimal models satisfy detailed balance, \cref{eq:envDB}.
+  The number of synapses $N=1$ and the number of states $M=10$.
   The dashed lines indicate the initial SNR bound, \eqref{eq:initbnd}, and the area bound, \eqref{eq:areabnd}, from \cite{Lahiri2013synapse}.}\label{fig:envproven}
@@
 % \(
-\begin{table}[ht]
+\begin{table}[tb]
   \begin{center}
@@
     \toprule
-  \multicolumn{5}{c}{Conditions} & \CI_a & \tau_a 
+      \multicolumn{5}{c}{Domain of validity} & \CI_a & \tau_a 
         & \mu_\CI & \mu_a & \mu_\CA 
@@
       &-& & M &\leq 3 & 1 & M - 1 
-    & \frac{(M-1)^2}{(M-1) + r \tau} %- \frac{\mu_a \sqrt{M-1}}{2} 
-    & 0 & \frac{r \tau}{[(M-1) + r \tau]^2} %- \frac{\mu_a}{2 \sqrt{M-1}} 
+        & \frac{(M-1)^2}{[(M-1) + r \tau]^2} %- \frac{\mu_a \sqrt{\tau_a}}{2} 
+        & 0 & \frac{r \tau}{[(M-1) + r \tau]^2} %- \frac{\mu_a}{2 \sqrt{\tau_a}} 
         % & \frac{\sqrt{N} (M-1)}{(M-1) + r \tau} 
@@
 The range of \(\tau\) for the second row does not exist for \(M = 3\), in which case any of the other rows can be used.}
-\label{tab:env}
+\label{tab:envDB}
 \end{table}
@@
   \mu^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
-    + \lambda \ombm^\nu_i(0) 
+    + \lambda \ombm^\nu_m(0) 
     + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
@@
 
-We can maximise this \wrt $\varepsilon$.
+We can maximise this \wrt $\varepsilon$, which yields
 %
@@
   (\varepsilon - f_m)^2 = (m - f_m)(g_m - f_m),
-  \qquad
+  \quad \implies \quad
   \hat{A}(s) = \frac{\Delta f_{m}}
@@
   \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\mathrlap{\varepsilon=0}} \propto &\,
-    m(2s+1) \brk{S(m\beta) - S((m-1)\beta)} - S(m\beta) [S(m\beta) + 1]
+    m \zeta \brk{S(m\beta) - S((m-1)\beta)} - S(m\beta) [S(m\beta) + 1]
       = 0,\\
   \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\mathrlap{\varepsilon=1}} \propto &\,
-    (m-1)\brk{S(m\beta) - S((m-1)\beta)} - (2s+1) S((m-1)\beta) [S((m-1)\beta) + 1]
+    (m-1)\brk{S(m\beta) - S((m-1)\beta)} - \zeta S((m-1)\beta) [S((m-1)\beta) + 1]
       = 0,\\
     &\text{where}\quad
+    \zeta = 2s + 1, \qquad
     s = S(\beta) = 2 \sinh^2(\beta/2),
@@
 \caption[Signal-to-noise ratio for optimal sticky models]
-{Signal-to-noise ratio for optimal sticky models, \cref{eq:stickyenc} optimised over \(\varepsilon\) for \(M=10\), in blue.
-  The orange line is the SNR for the optimised shortened models, \cref{eq:shortenenc} with \(M=10\).
+{Signal-to-noise ratio for optimal sticky models, \cref{eq:stickyopt} for \(M=10\), in blue.
+  The orange line is the SNR for the optimised shortened models, \cref{eq:shortopt} with \(M=10\).
   The grey line is our theoretical envelope from \cref{eq:env}.
@@
 We can maximise this \wrt $\varepsilon$.
-This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
 We define
@@
 %
-The minimum is at the solution of
-%
-\begin{equation*}
+The maximum is at the solution of
+%
+\begin{multline}\label{eq:stickyopt}
+% \begin{aligned}
   \brk{(1 + f_m -  g_m - \mu)\varepsilon - f_m + \mu g_m}^2 
-    = (\mu - 1) (g_m - 1) (\mu - f_m) (g_m - f_m).
-\end{equation*}
+    = (\mu - 1) (g_m - 1) (f_m - \mu) (f_m - g_m) , \\
+  \implies \qquad
+  \hat{A}(s) = \frac{\Delta f_m}{s \Delta m \Delta g_m}
+                \brk{\frac{\sqrt{(f_m - \mu)(f_m - g_m)} + \sqrt{(\mu - 1)(g_m - 1)}}
+                          {\sqrt{(f_m - g_m)(g_m - 1)} + \sqrt{(f_m - \mu)(\mu - 1)}}}^2.
+% \end{aligned}
+\end{multline}
 %

### 2020-10-28 Notes/laplace.tex
copied proof of init bound from older notes

@@
 
-We will be trying to store patterns in a set od $N$ synaptic weights, $\syn$.
+We will be trying to store patterns in a set od \(N\) synaptic weights, \(\syn\).
 Every time we try to store a pattern, these synapses are subjected to a plasticity event where each synapse is either potentiated or depressed, depending on the pattern.
@@
 We wish to answer this question by looking at the synaptic weights directly (ideal observer).
-For that given pattern there will be an ideal set of synaptic weights, $\synid$, where those synapses that were supposed to be potentiated are maximised and those that were supposed to be depressed are minimised.
-Suppose that the given pattern was actually seen at time 0 and we are observing the synapses at time $t$.
-The actual set of synaptic weights we see, $\syn(t)$, will be a vector of random variables that differs from $\synid$ due to the stochasticity of the pattern encoding and all of the other (uncorrelated) pattern that are stored after it.
-As $t\to\infty$, the synaptic weights will become independent of the patter stored at $t=0$.
-Thus, the vector of random variables $\syn(\infty)$ also describes the synaptic weights under the null hypothesis -- if that given pattern had never been stored.
-
-We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold~\cite{Sommer1998retrieval}.
-For large $N$, this quantity will have a Gausssian distribution.
+For that given pattern there will be an ideal set of synaptic weights, \(\synid\), where those synapses that were supposed to be potentiated are maximised and those that were supposed to be depressed are minimised.
+Suppose that the given pattern was actually seen at time 0 and we are observing the synapses at time \(t\).
+The actual set of synaptic weights we see, \(\syn(t)\), will be a vector of random variables that differs from \(\synid\) due to the stochasticity of the pattern encoding and all of the other (uncorrelated) pattern that are stored after it.
+As \(t\to\infty\), the synaptic weights will become independent of the patter stored at \(t=0\).
+Thus, the vector of random variables \(\syn(\infty)\) also describes the synaptic weights under the null hypothesis -- if that given pattern had never been stored.
+
+We can test if the pattern had been previously stored by computing \(\synid\cdt\syn\) and comparing it to some threshold~\cite{Sommer1998retrieval}.
+For large \(N\), this quantity will have a Gausssian distribution.
 There will be a ROC curve as a function of this threshold:
@@
 and TPR/FPR are the true/false positive rates.
-When the signal-to-noise ratio, $\snr$ is larger than $\Phi_c\inv(\operatorname{FPR})$, it is beneficial to decrease the noise-to-noise ratio, NNR.
+When the signal-to-noise ratio, \(\snr\) is larger than \(\Phi_c\inv(\operatorname{FPR})\), it is beneficial to decrease the noise-to-noise ratio, NNR.
 In the other case, it is beneficial to increase it.
-The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.
+The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, \(\synid\).
 
@@
 The formulae above assumed that we know the time between storage and recognition.
-We should also compute the expectations and variances over probability distribution of the recall time, $t$, as well.
-If we only know average time, $\tau$, a natural choice for this distribution is
+We should also compute the expectations and variances over probability distribution of the recall time, \(t\), as well.
+If we only know average time, \(\tau\), a natural choice for this distribution is
 %
@@
 %
-Different parts of the brain, that store memories for different timescales, could be characterised by different values of $\tau$.
+Different parts of the brain, that store memories for different timescales, could be characterised by different values of \(\tau\).
 
@@
 As we are looking at the information contained in the synaptic weights, rather than modelling the readout via electrical activity of the neurons, what values the synaptic weights actually take is irrelevant.
-It will be convenient to call these two values $\pm1$.
+It will be convenient to call these two values \(\pm1\).
 
@@
 %
-Where the averages and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.
+Where the averages and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, \(\synid\).
 If we also average over the time between storage and recognition, we have
@@
 %
-This is similar to the average SNR up to time $\tau$, but with the hard cut-off replaced with a smooth exponential.
+This is similar to the average SNR up to time \(\tau\), but with the hard cut-off replaced with a smooth exponential.
 This can be expressed in terms of he Laplace transform of the SNR curve in \eqref{eq:SNRdef}:
@@
 %
-where we used $\syn^\alpha=\pm1$.
-We can compute $\var(\synid\cdt\syn(\infty))$ by taking $t\to\infty$.
+where we used \(\syn^\alpha=\pm1\).
+We can compute \(\var(\synid\cdt\syn(\infty))\) by taking \(t\to\infty\).
 
@@
 %
-If the elements of $\synid$ take values $\pm1$ with probability $f\potdep$,
+If the elements of \(\synid\) take values \(\pm1\) with probability \(f\potdep\),
 %
@@
 
-We model a synapse as having $M$ internal states.
-The synaptic weight depends deterministically on the state, given by the $M$-element column vector $\w$.
-We denote the probability distribution across these states by the row vector $\pr(t)$.
-We denote a column vector of ones by $\onev$, so the normalisation condition is $\pr(t) \, \onev=1$.
-
-When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\MMd^\nu$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$,
+We model a synapse as having \(M\) internal states.
+The synaptic weight depends deterministically on the state, given by the \(M\)-element column vector \(\w\).
+We denote the probability distribution across these states by the row vector \(\pr(t)\).
+We denote a column vector of ones by \(\onev\), so the normalisation condition is \(\pr(t) \, \onev=1\).
+
+When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices \(\MMd^\nu\) whose \(i,j\) elements are the transition probabilities from state \(i\) to state \(j\),
 and \(\nu \) labels the diffrerent types of plasticity event.
@@
 %
-The matrices $\MMd^\nu$ are transition matrices of discrete time Markov processes.
+The matrices \(\MMd^\nu\) are transition matrices of discrete time Markov processes.
 Their elements must satisfy the following constraints
@@
 %
-The upper bounds $\MMdm^\nu_{ij} \leq 1$ follow automatically from these.
+The upper bounds \(\MMdm^\nu_{ij} \leq 1\) follow automatically from these.
 
@@
 
-Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$, 
+Given a sequence of plasticity events at times \(t_1,t_2,t_3,\ldots\), the probability distribution across the internal states at time \(t_n\), 
 mariginalising over states at earlier times, is
@@
 %
-If the probability of a given plasticity event being potentiating or depressing is $f\pot$ or $f\dep$ respectively, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.
-
-In computing the expectation in \eqref{eq:overlapone} the plasticity event that occurs at $t=0$ is fixed, but all of the previous and subsequent events are marginalised over.
+If the probability of a given plasticity event being potentiating or depressing is \(f\pot\) or \(f\dep\) respectively, the probability of the sequence above (conditioned on the number of events) is given by \(f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot\).
+
+In computing the expectation in \eqref{eq:overlapone} the plasticity event that occurs at \(t=0\) is fixed, but all of the previous and subsequent events are marginalised over.
 If we marginalise over the types (potentiating/depressing) of these other events (conditioned on the number of events) we find
@@
 % 
-where the factor of $ \MMd \potdep$ is the event at $t=0$.
+where the factor of \( \MMd \potdep\) is the event at \(t=0\).
 If we expand out all of the parenthetical factors, we will get a sum of terms of the form \eqref{eq:plastseq} weighted by their probabilities.
 
-If $t_0 \ll 0$, the effect of the large number of earlier events will be to put the synapse in the steady state distribution, $\eq$, of the stochastic process $(f\pot \MMd \pot+f\dep \MMd \dep)$:
+If \(t_0 \ll 0\), the effect of the large number of earlier events will be to put the synapse in the steady state distribution, \(\eq\), of the stochastic process \((f\pot \MMd \pot+f\dep \MMd \dep)\):
 %
@@
 %
-where $n$ is the number of plasticity events that took place.
-
-If the plasticity events occur at Poisson rate $r$, the probability of $n$ events occurring in time $t$ is given by a Poisson distribution with mean $rt$.
-If we marginalise over $n$ as well
+where \(n\) is the number of plasticity events that took place.
+
+If the plasticity events occur at Poisson rate \(r\), the probability of \(n\) events occurring in time \(t\) is given by a Poisson distribution with mean \(rt\).
+If we marginalise over \(n\) as well
 %
@@
 %
-where $\eq$ is the steady state distribution of the continuous time transition matrix that appears in the exponential:
+where \(\eq\) is the steady state distribution of the continuous time transition matrix that appears in the exponential:
 %
@@
 %
-The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
+The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with \(f\pot=f\dep\) or \(\eq_+=\eq_-\).
 We can perform the maximisation in two steps.
-First maximise the numerator at fixed $\eq \w$.
-Then maximise the ratio \wrt $\eq\w$.
+First maximise the numerator at fixed \(\eq \w\).
+Then maximise the ratio \wrt \(\eq\w\).
 This will allow us to ignore the denominator until the very end.
 
-We can then treat the off-diagonal elements of $\MMd^\nu$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
+We can then treat the off-diagonal elements of \(\MMd^\nu\) as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
 The diagonal elements are determined by setting the row sums to 1.
@@
 
-If we wish to use the off-diagonal elements of $\frg$ and $\enc$ as the degrees of freedom, the constraints are
+If we wish to use the off-diagonal elements of \(\frg\) and \(\enc\) as the degrees of freedom, the constraints are
 %
@@
 %
-where $\arow$ is an arbitrary row vector satisfying $\arow \onev = \frac{1}{\tau_0} \neq 0$.
-This reduces to the fundamental matrix at $s=0$.
-We can see that $\onev$ is an eigenvector:
+where \(\arow\) is an arbitrary row vector satisfying \(\arow \onev = \frac{1}{\tau_0} \neq 0\).
+This reduces to the fundamental matrix at \(s=0\).
+We can see that \(\onev\) is an eigenvector:
 %
@@
 %
-Suppose we have some row vector $\mathbf{a}$ such that $\mathbf{a}\onev=0$.
-Then $\mathbf{a}\fund(s)$ is independent of $\arow$:
+Suppose we have some row vector \(\mathbf{a}\) such that \(\mathbf{a}\onev=0\).
+Then \(\mathbf{a}\fund(s)\) is independent of \(\arow\):
 %
@@
 %
-This reduces to the mean first-passage time matrix at $s=0$.
+This reduces to the mean first-passage time matrix at \(s=0\).
 
@@
 
-This means that the \( \fptb(s) \) have all the same properties as in the case $s=0$.
+This means that the \( \fptb(s) \) have all the same properties as in the case \(s=0\).
 In particular, we see that the quantity
@@
 %
-is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau_0}{1 + s \tau_0}$.
+is independent of the starting state \(i\), but now \(\eta(s) = \tr\fund(s) - \frac{\tau_0}{1 + s \tau_0}\).
 We can then define
@@
 %
-We can arrange the states in order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
+We can arrange the states in order of decreasing \(\eta^{+}_i(s)\) or increasing \(\eta^{-}_i(s)\).
 It will be more convenient to use the quantities
@@
 %
-Arranging the states in order of decreasing $\etwm_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
-The $\etwm_i(s)$ can still be used when $\w$ takes more than two values.
+Arranging the states in order of decreasing \(\etwm_i(s)\) is the same as the order of decreasing \(\eta^{+}_i(s)\) or increasing \(\eta^{-}_i(s)\).
+The \(\etwm_i(s)\) can still be used when \(\w\) takes more than two values.
 
@@
 %
-For $\Re s>0$, we have
+For \(\Re s>0\), we have
 %
@@
 %
-As $(s-\MM)$ is invertible for $\Re s>0$, because the real part of all eigenvalues of $\MM$ are nonpositive, we have
+As \((s-\MM)\) is invertible for \(\Re s>0\), because the real part of all eigenvalues of \(\MM\) are nonpositive, we have
 %
@@
 %
-For $s=0$, we can avoid problems by replacing $\mathbf{G}(s) \to \fund(s)$.
+For \(s=0\), we can avoid problems by replacing \(\mathbf{G}(s) \to \fund(s)\).
 
@@
 %
-This expression is ill-behaved at $s=0$.
-Thanks to \eqref{eq:alfund}, we can solve this by the replacement $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
+This expression is ill-behaved at \(s=0\).
+Thanks to \eqref{eq:alfund}, we can solve this by the replacement \(\mathbf{G}(s) \to \fund(s)\), as \(\enc\onev=0\).
 %
@@
 %
-Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.
+Note that \(A(0)=A\), the area, and \(\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)\), the initial SNR.
 
@@
 
-In analogy with the case $s=0$, we define
+In analogy with the case \(s=0\), we define
 %
@@
 %
-The term with $\Theta(s)$ comes from differentiating $\eq$, the term with $c(s)$ comes from differentiating $\etw(s)$ and the term with $\pm1$ comes from differentiating $\enc$ in \eqref{eq:larea}.
+The term with \(\thb(s)\) comes from differentiating \(\eq\), the term with \(c(s)\) comes from differentiating \(\etw(s)\) and the term with \((-1)^\mu\) comes from differentiating \(\enc\) in \eqref{eq:larea}.
 
 To write down the Hessian, we will first define some notation.
-We define the matrix $\mathbf{X}(s) = \fund(0) \enc \fund(s)$.
-Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$.
+We define the matrix \(\mathbf{X}(s) = \fund(0) \enc \fund(s)\).
+Given a four-index array \(H_{ijkl}\), we define \(H\trans_{ijkl} = H_{klij}\).
 Then we denote
@@
 Now to study the SNR as a function of  time, it helps to consider some variables that are more directly related to the time dependance than the matrix elements. 
-So we introduce an eigenvector decomposition for $-\frg$:
+So we introduce an eigenvector decomposition for \(-\frg\):
 %
@@
 %
+We have ignored thedenominator of \cref{eq:SNRcurveExact} for now.
+This will not have any effect on the results: it will not change the location of the maxima, which are at \( f\pot - f\dep = \eq \w = 0 \).
+We will confine such discussion to the footnotes.
+
 Now the question is what are the allowed values of \( (\CI_a, \tau_a) \)?
@@
 %
-
-As shown in \cite{Lahiri2013synapse}, this satisfies the inequality
+We can write the initial SNR as
+%
+\begin{equation*}
+  \frac{\SNR(0)}{\sqrt{N}} = \CI = \eq \enc \w.
+\end{equation*}
+%
+First, we define two sets of states: \( + = \set{i}{\wm_i = 1} \) and \( - = \set{i}{\wm_i = - 1} \), and the fluxes \( \Fm^\nu_{\CA \CB} = r f^\nu \displaystyle\sum_{\mathclap{i \in \CA, j \in \CB}} \eqm_i \prn{\MMdm^\nu_{ij} - \delta_{ij}} \).
+Then we can write
+%
+\begin{equation*}
+\begin{aligned}
+  r\CI &= \Fm\pot_{++} + \Fm\pot_{-+} - \Fm\dep_{++} - \Fm\dep_{-+} - \Fm\pot_{+-} - \Fm\pot_{--} + \Fm\dep_{+-} + \Fm\dep_{--} \\
+      &= 2\prn{\Fm\pot_{-+} - \Fm\dep_{-+} - \Fm\pot_{+-} + \Fm\dep_{+-}} \\
+      &= 2\prn{\Fm_{-+} + \Fm_{+-}} - 4\prn{\Fm\pot_{+-} + \Fm\dep_{-+}} \\
+      &= 4\prn{\Fm_{-+} - \Fm\pot_{+-} - \Fm\dep_{-+}}, \\
+\end{aligned}
+\end{equation*}
+%
+where we used \( \Fm^\nu_{\CA+} + \Fm^\nu_{\CA-} = 0 \) to derive the second line, \( \Fm_{\CA\CB} = \Fm\pot_{\CA\CB} + \Fm\dep_{\CA\CB} \) to derive the third and additionally \( \Fm_{+\CA} + \Fm_{-\CA} = 0 \) to derive the fourth.
+
+The last two terms are guaranteed to be negative, as the diagonal parts of \(\MMd \potdep - \I\) cannot contribute.
+Therefore
+%
+\begin{equation}\label{eq:initfluxineq}
+  \SNR(0) \leq \frac{4\sqrt{N}\Fm_{-+}}{r}.
+\end{equation}
+%
+This inequality is saturated if potentiation never takes it from a \(+\) state to a \(-\) state and depression never takes it from a \(-\) state to a \(+\) state.
+
+The flux will be maximised if potentiation is guaranteed to take it to a \(+\) state and depression is guaranteed to take it to a \(-\) state:
+%
+\begin{equation*}
+  \sum_{j \in +} \MMd \pot_{ij} = 1 \quad \forall\, i \in -,
+  \qquad
+  \sum_{j \in -} \MMd \dep_{ij} = 1 \quad \forall\, i \in +.
+\end{equation*}
+%
+This implies that the process is lumpable to a two state system (see \cite{kemeny1960finite,Ball1993Lumpability,burke1958markovian}):
+%
+\begin{multline}\label{eq:binarylump}
+%\begin{aligned}
+  \MMd \pot = \begin{pmatrix}
+                0 & 1 \\
+                0 & 1 \\
+              \end{pmatrix}
+  ,\quad
+  \MMd \dep = \begin{pmatrix}
+                1 & 0 \\
+                1 & 0 \\
+              \end{pmatrix}
+  ,\quad
+  \w = \begin{pmatrix}
+           -1 \\
+           1 \\
+         \end{pmatrix}
+  ,\\ \implies\quad
+  \frg = \begin{pmatrix}
+           -f\pot & f\pot \\
+           f\dep & -f\dep \\
+         \end{pmatrix}
+  ,\quad
+  \eq = \begin{pmatrix}
+           f\dep & f\pot \\
+         \end{pmatrix}
+  ,
+\end{multline}s
+%
+which leads to the initial SNR
+\footnote{Note that including the dropped factors in \cref{eq:SNRcurveExact} would only change this to
+\(\SNR(0) \leq \frac{\sqrt{N} (4f^+f^-)}{\sqrt{1 - (f\pot - f\dep)^4}}\),
+which is also maximised at \(f\potdep = \frac{1}{2}\) and would therefore have no effect on what follows.}
+%
+\begin{equation}\label{eq:binarySNR}
+  \SNR(0) \leq \sqrt{N} (4 f\pot f\dep) .
+\end{equation}
+%
+This is maximised at \(f\pot = f\dep = \frac{1}{2}\),
+leading to the inequality firt shown in \cite{Lahiri2013synapse}:
 %
@@
 %
-The model that saturates this bound is (equivalent to) a two-state model with deterministic transitions and $f\potdep=\frac{1}{2}$.
+The model that saturates this bound is (equivalent to) a two-state model with deterministic transitions and \(f\potdep=\frac{1}{2}\).
 
@@
 %
+To maximise this quantity, consider the Lagrangian
+%
+\begin{equation}\label{eq:arealagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(0) - \etwm_j(0))
+        +\sum_{\nu i j} \prn{ \mu^\nu_{ij} \MMdm^\nu_{ij} }
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
+\end{aligned}
+\end{equation}
+%
+where \(\Delta\pr\) is the constant value we are holding \(\eq\w\) at.
+The Karush-Kuhn-Tucker necessary conditions for a maximum are
+%
+\begin{equation}\label{eq:KTarea}
+  \pdiff{\CL}{\MMdm^\nu_{ij}} = 0, \quad
+  \mu^\nu_{ij} \geq 0, \quad
+  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
+  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
+  \pdiff{\CL}{\lambda} = 0.
+\end{equation}
+%
+From \cref{eq:lAdiff}, the derivatives for \(m \neq n\) are
+%
+\begin{equation}\label{eq:areaderivs}
+\begin{aligned}
+  \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+    f^\nu \eqm_m \prn{ \brk{\thbm_m(0) - \thbm_n(0)}
+     + \prn{\mathbf{c}_m(0) + (-1)^\nu} \brk{\etwm_m(0) - \etwm_n(0)} } \\
+     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
+     + \lambda f^\nu \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
+    = 0.
+\end{aligned}
+\end{equation}
+%
+See \cref{eq:lareacoeffs} for the definitions of these quantities.
+
+The area (and \(\Delta\pr\)) is invariant under the scale transformation
+%
+\begin{equation}\label{eq:scale}
+  \zeta * \MMd \equiv \zeta \MMd + (1-\zeta)\I .
+\end{equation}
+%
+It follows that
+%
+\begin{equation*}
+  \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} = - \sum_{\nu, m} \mu^\nu_{mm} = 0 .
+\end{equation*}
+%
+As the KKT multipliers are nonnegative, this implies that all the diagonal KKT multipliers vanish and we can ignore the diagonal constraints.
+
+Another way of seeing this is to note that all but the diagonal constraints in \cref{eq:constri} are invariant under \cref{eq:scale}.
+This means that if we have a set of parameters that satisfy the off-diagonal constraints, it can always be scaled so that it meets the diagonal constraints.
+Therefore, we can ignore the diagonal constraints whilst maximising the area and fix them afterwards by  scaling.
 
@@
 %
-The model that saturates this bound is one of serial topology with $\eq$ concentrated symmetrically at the two end states.
+The model that saturates this bound is one of serial topology with \(\eq\) concentrated symmetrically at the two end states.
 
@@
     \\
-    \F^\nu_{\CA \CB} &= r f^\nu \sum_{\substack{i \in \CA \\ j \in \CB}} 
+    \Fm^\nu_{\CA \CB} &= r f^\nu \sum_{\substack{i \in \CA \\ j \in \CB}} 
                         \eqm_i \prn{\MMdm^\nu_{ij} - \delta_{ij}} .
@@
 %
-No doubt there are many other constraints that these quantities must satisfy, but we can see what these ones imply for the $\snrb(\tau)$ by maximising it subject to these constraints.
+No doubt there are many other constraints that these quantities must satisfy, but we can see what these ones imply for the \(\snrb(\tau)\) by maximising it subject to these constraints.
 
@@
   (\ref{fig:provendetbal}) Conjectured memory curve envelope when we assume the optimal models satisfy detailed balance, \cref{eq:envDB}.
-  The number of synapses $N=1$ and the number of states $M=10$.
+  The number of synapses \(N=1\) and the number of states \(M=10\).
   The dashed lines indicate the initial SNR bound, \eqref{eq:initbnd}, and the area bound, \eqref{eq:areabnd}, from \cite{Lahiri2013synapse}.}\label{fig:envproven}
@@
 
-Now we'll maximise $\snrb(\tau)$ for some fixed $\tau$, subject to the constraints \eqref{eq:constr}.
-First, we will maximise the numerator of \eqref{eq:lsnr} holding $\eq\w$ fixed.
-We can maximise the ration \wrt $\eq\w$ later.
+Now we'll maximise \(\snrb(\tau)\) for some fixed \(\tau\), subject to the constraints \eqref{eq:constr}.
+First, we will maximise the numerator of \eqref{eq:lsnr} holding \(\eq\w\) fixed.
+We can maximise the ration \wrt \(\eq\w\) later.
 From this point on, we work in units where \(r = 1\).
@@
 %
-where $\Delta\pr$ is the constant value we are holding $\eq\w$ at.
+where \(\Delta\pr\) is the constant value we are holding \(\eq\w\) at.
 The Karush-Kuhn-Tucker necessary conditions for a maximum are
@@
 
-The derivatives, for $m \neq n$ are
+The derivatives, for \(m \neq n\) are
 %
@@
 
-The expressions in \cref{eq:finitederivs} have the same form as for $s=0$, which allowed us to argue that the model with maximal area must have the serial topology, \cite{Lahiri2013synapse}.
+The expressions in \cref{eq:finitederivs} have the same form as for \(s=0\), which allowed us to argue that the model with maximal area must have the serial topology, \cite{Lahiri2013synapse}.
 The one missing ingredient is the scale invariance.
@@
   {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
-  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numericallly with $f\potdep=\frac{1}{2}$, the number of synapses $N=1$ and the number of states $M=10$.
+  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numericallly with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
   The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
@@
 
-Define the scale transformation:
-%
-\begin{equation}\label{eq:scale}
-  \zeta * \MMd \equiv \zeta \MMd + (1-\zeta)\I .
-\end{equation}
-%
-The Laplace transform has the following scaling property
+Under the transformation from \cref{eq:scale}, the Laplace transform has the following scaling property
 %
@@
 %
-At $s = 0$, it is scale invariant.
+At \(s = 0\), it is scale invariant.
 This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
 This allowed us to ignore the diagonal constraints.
-We can't do that for $s > 0$.
+We can't do that for \(s > 0\).
 
@@
   All unlabelled arrowshave transition probabilities equal to 1.
-  At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with $M$ states.
+  At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with \(M\) states.
   At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model.
@@
 
-At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with $M$ states.
-At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model as $\tau\to\infty$.
+At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with \(M\) states.
+At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model as \(\tau \to \infty\).
 At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
 Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from \autoref{fig:envvid} and the video, this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
-Once the length has been reduced to $M=2$, the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.
+Once the length has been reduced to \(M=2\), the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.
 
 These models' Laplace transforms can be computed analytically, as we will see below, leading to a heuristic envelope.
-All of these calculations will be performed at $f\potdep=\frac{1}{2}$, so the denominator will play no role.
-We will work in units such that $r=1$ temporarily, only restoring $r$ at the end.
+All of these calculations will be performed at \(f\potdep=\frac{1}{2}\), so the denominator will play no role.
+We will work in units such that \(r=1\) temporarily, only restoring \(r\) at the end.
 We study models with \(m\) strong and \(m\) weak states for a total of  \(M = 2 m\).
@@
 Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_uniform}.
-Looking at \eqref{eq:larea}, the only thing we need is differences in $\etwm_i(s)$.
-In fact, given that the only nonzero elements of $\enc$ are
+Looking at \eqref{eq:larea}, the only thing we need is differences in \(\etwm_i(s)\).
+In fact, given that the only nonzero elements of \(\enc\) are
 %
@@
 %
-and $\eqm_i = \frac{1}{2m}$, we have
+and \(\eqm_i = \frac{1}{2m}\), we have
 %
@@
 %
-As we are only interested in differences of the $\etwm_i(s)$, we can drop the constant.
+As we are only interested in differences of the \(\etwm_i(s)\), we can drop the constant.
 Then
@@
 %
-As we are only interested in differences, we can drop any terms proportional to $\onev$ in $\etw(s)$.
-
-The $i$'th row of this equation reads
+As we are only interested in differences, we can drop any terms proportional to \(\onev\) in \(\etw(s)\).
+
+The \(i\)'th row of this equation reads
 %
@@
 %
-These equations are invariant under $\delta\eta_{2m+1-i} \to -\delta\eta_i$, so these two must be equal.
+These equations are invariant under \(\delta\eta_{2m+1-i} \to -\delta\eta_i\), so these two must be equal.
 Then we have
@@
 %
-with $B$ and $C$ determined by the boundary conditions provided by the last two equations of \eqref{eq:detahalf}:
+with \(B\) and \(C\) determined by the boundary conditions provided by the last two equations of \eqref{eq:detahalf}:
 %
@@
 
-In reality, $M$ can only take even integer values.
-However, one can find a heuristic approximation to the envelope by maximising \wrt $M$.
+In reality, \(M\) can only take even integer values.
+However, one can find a heuristic approximation to the envelope by maximising \wrt \(M\).
 We find that
@@
 %
-When $M \gg 2$, the validity region is approximately
+When \(M \gg 2\), the validity region is approximately
 %
@@
 %
-When $s \sim \CO(M^{-2}) \ll 1$, we have \(\beta \approx \pm \sqrt{2s}\) and the heuristic envelope is approximately
+When \(s \sim \CO(M^{-2}) \ll 1\), we have \(\beta \approx \pm \sqrt{2s}\) and the heuristic envelope is approximately
 %
@@
 
-For $s\gtrsim 1.36423$, we set $M=2$ to find
+For \(s\gtrsim 1.36423\), we set \(M=2\) to find
 %
@@
 
-The following elements of $\enc$ and $\eq$ differ from \eqref{eq:unienc}:
+The following elements of \(\enc\) and \(\eq\) differ from \eqref{eq:unienc}:
 %
@@
 
-The $\etwm_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
+The \(\etwm_i(s)\) take the same general form \eqref{eq:detagensol} except for \(i=1\), and with different coefficients due to the different boundary conditions
 %
@@
 %
-This interpolates between \eqref{eq:unilaplace} at $\varepsilon=0$ and the same expression for $m-1$ at $\varepsilon=1$.
+This interpolates between \eqref{eq:unilaplace} at \(\varepsilon=0\) and the same expression for \(m-1\) at \(\varepsilon=1\).
 
@@
 
-We can maximise this \wrt $\varepsilon$, which yields
+We can maximise this \wrt \(\varepsilon\), which yields
 %
@@
 %
-We can find the boundary of the region of applicability by asking at what value of $s$ is the optimal value of $\varepsilon=0$ or $1$?
+We can find the boundary of the region of applicability by asking at what value of \(s\) is the optimal value of \(\varepsilon=0\) or \(1\)?
 %
@@
 %
-with the $\varepsilon=0$ case providing the lower boundary and $\varepsilon=1$ providing the upper boundary.
+with the \(\varepsilon=0\) case providing the lower boundary and \(\varepsilon=1\) providing the upper boundary.
 For lower \(s\), \(\varepsilon = 0\) is the optimum. 
@@
 
-If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
+If these were really the models that maximised \(A(s)\), the region of validity for \(M\) would not overlap with that for \(M-2\).
 In fact there are overlaps, as seen in \cref{fig:shorten}\ref{fig:short_ratio} for \(M=6\) and higher.
@@
 
-The following elements of $\enc$ and $\eq$ are differ from \eqref{eq:unienc}:
+The following elements of \(\enc\) and \(\eq\) are differ from \eqref{eq:unienc}:
 %
@@
 
-The $\etwm_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
+The \(\etwm_i(s)\) take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
 %
@@
 
-We can maximise this \wrt $\varepsilon$.
+We can maximise this \wrt \(\varepsilon\).
 We define
@@
 
-We can find the boundary of the region of applicability by asking: at what value of $s$ is the optimal value of $\varepsilon=0$?
+We can find the boundary of the region of applicability by asking: at what value of \(s\) is the optimal value of \(\varepsilon=0\)?
 %
@@
 %
-The ``sticky'' models are optimal at values of $s$ smaller than this.
-This is not quite the same as the lower boundary of the region \cref{eq:shortenvalid}, or its approximate version \cref{eq:univalid}, so there is a gap between them where the uniform model with $M$ states is optimal.
-The width of this gap is insignificant for larger $M$, as seen in \cref{fig:sticky}.
-
-For small $s \ll M^{-2}$, it is given by
+The ``sticky'' models are optimal at values of \(s\) smaller than this.
+This is not quite the same as the lower boundary of the region \cref{eq:shortenvalid}, or its approximate version \cref{eq:univalid}, so there is a gap between them where the uniform model with \(M\) states is optimal.
+The width of this gap is insignificant for larger \(M\), as seen in \cref{fig:sticky}.
+
+For small \(s \ll M^{-2}\), it is given by
 %
@@
 
-We will phrase it in terms of $\snrb(\tau)$, reintroducing $r$:
+We will phrase it in terms of \(\snrb(\tau)\), reintroducing \(r\):
 %
@@
 %
-For $M \gg 1$, and $r\tau$ much greater than the lower bound of each region, we can approximate this as
+For \(M \gg 1\), and \(r\tau\) much greater than the lower bound of each region, we can approximate this as
 %
@@
 \begin{itemize}
-  \item \emph{Short} timescale $\to$ \emph{intermediate} timescales: transition topology goes from \\ short and wide $\to$ long and thin.
-  \item \emph{Intermediate} timescale $\to$ \emph{long} timescales: transition probabilities go from \\ strong and deterministic $\to$ weak and stochastic.
+  \item \emph{Short} timescale \(\to\) \emph{intermediate} timescales: transition topology goes from \\ short and wide \(\to\) long and thin.
+  \item \emph{Intermediate} timescale \(\to\) \emph{long} timescales: transition probabilities go from \\ strong and deterministic \(\to\) weak and stochastic.
 \end{itemize}
@@
 
-The transition between the two regimes is set by the maximum number of states, $M$ (we can't set the number of states to be exactly $M$, as bigger models can always mimic smaller models).
+The transition between the two regimes is set by the maximum number of states, \(M\) (we can't set the number of states to be exactly \(M\), as bigger models can always mimic smaller models).
 So, if we want to lengthen the timescale at which a synapse is optimal, we start by adding extra states, using the topology mechanism.
@@
 
-We can get some insight into why this is by considering a serial model with $M$ states and all transition topologies equal to $q$.
+We can get some insight into why this is by considering a serial model with \(M\) states and all transition topologies equal to \(q\).
 %
@@
 %
-The initial SNR scales as described because it is related to the equilibrium probability flux between the weak and strong states, and the equilibrium distribution is uniform over $M$ states.
-The lifetime scales as described because $rq$ sets the timescale for transitions, and for diffusion distance ($M$) scales as the square root of time.
-
-If we want to lengthen the lifetime with a fixed hit on initial SNR, we get a larger increase by increasing $M$ than by decreasing $q$.
-If we want to lengthen the lifetime by a fixed amount, we take a smaller hit on initial SNR by increasing $M$ than by decreasing $q$.
+The initial SNR scales as described because it is related to the equilibrium probability flux between the weak and strong states, and the equilibrium distribution is uniform over \(M\) states.
+The lifetime scales as described because \(rq\) sets the timescale for transitions, and for diffusion distance (\(M\)) scales as the square root of time.
+
+If we want to lengthen the lifetime with a fixed hit on initial SNR, we get a larger increase by increasing \(M\) than by decreasing \(q\).
+If we want to lengthen the lifetime by a fixed amount, we take a smaller hit on initial SNR by increasing \(M\) than by decreasing \(q\).
 Thus, we always get more bang for our buck using the topology mechanism than the stochasticity mechanism.

### 2020-11-02 Notes/laplace.tex
num_synapse parameter for plots

@@
 %
-which leads to the initial SNR
+which leads to the initial SNR%
 \footnote{Note that including the dropped factors in \cref{eq:SNRcurveExact} would only change this to
@@
 This is maximised at \(f\pot = f\dep = \frac{1}{2}\),
-leading to the inequality firt shown in \cite{Lahiri2013synapse}:
+leading to the inequality (first shown in \cite{Lahiri2013synapse}):
 %

### 2020-11-02 Code/Python/complex_synapse/examples.py

@@ -17,7 +17,7 @@ np.set_printoptions(precision=3, suppress=False, linewidth=90)
 mplt.rc_fonts()
 mplt.rc_colours()
 
-def ser_casc_data(nst: int, jmp: float, time: la.lnarray
+def ser_casc_data(nst: int, jmp: float, time: la.lnarray, nsyn: float = 1
                   ) -> ty.Tuple[la.lnarray, la.lnarray, la.lnarray]:
     """Data for example plot
 
@@ -29,6 +29,8 @@ def ser_casc_data(nst: int, jmp: float, time: la.lnarray
         Model parameter
     time : la.lnarray
         Vector of times to evaluate SNR
+    nsyn : float, optional
+        Number of synapses, by default 1.
 
     Returns
     -------
@@ -42,17 +44,17 @@ def ser_casc_data(nst: int, jmp: float, time: la.lnarray
     nst = ag.default(nst, 10)
     time = ag.default(time, 1 / la.geomspace(1e-4, 10, 50))
     jmp = ag.default(jmp, 0.7)
+    prefactor = np.sqrt(nsyn)
 
     serial = cs.SynapseMemoryModel.build(cs.builders.build_serial,
                                          nst, jmp=jmp)
     cascade = cs.SynapseMemoryModel.build(cs.builders.build_cascade,
                                           nst, jmp=jmp)
 
-
     serial_snr = serial.snr_exp_ave(time)
     cascade_snr = cascade.snr_exp_ave(time)
 
-    return time, serial_snr, cascade_snr
+    return time, prefactor * serial_snr, prefactor * cascade_snr
 
 
 def ser_casc_plot(time: la.lnarray, serial_snr: la.lnarray,
@@ -80,7 +82,7 @@ def ser_casc_plot(time: la.lnarray, serial_snr: la.lnarray,
 
 
 if __name__ == "__main__":
-    j, n, t = 0.7, 10, 1 / la.geomspace(1e-4, 10, 50)
-    t, snr_serial, snr_cascade = ser_casc_data(n, j, t)
-    fig_sc = ser_casc_plot(t, snr_serial, snr_cascade)
+    jump, nstate, num, times = 0.7, 10, 1e4, 1 / la.geomspace(1e-4, 10, 50)
+    times, snr_serial, snr_cascade = ser_casc_data(nstate, jump, times, num)
+    fig_sc = ser_casc_plot(times, snr_serial, snr_cascade)
     fig_sc.savefig("../../Notes/Figures/serial_vs_cascade.pdf")

### 2020-11-02 Code/Python/complex_synapse/optimise/plot.py

@@ -19,6 +19,7 @@ import sl_py_tools.graph_tricks as gt
 import complex_synapse.builders as bld
 import complex_synapse.optimise.optimise as cso
 import complex_synapse.optimise.shorten as sh
+import complex_synapse.synapse_mem as sm
 
 np.set_printoptions(precision=3, suppress=False, linewidth=90)
 mplt.rc_fonts()
@@ -158,13 +159,16 @@ def all_data(nst: int, sss: Array, **opts) -> ty.Tuple[Arrays, Arrays]:
 # =============================================================================
 
 
-def mem_plot(sss: Array, **mem: ty.Dict[str, Array]) -> ty.Tuple[Figure, Axes, Line]:
+def mem_plot(sss: Array, nsyn: float = 1, **mem: Array
+             ) -> ty.Tuple[Figure, Axes, Line]:
     """Plot memory curves
 
     Parameters
     ----------
     sss : Array
         1/time
+    nsyn : float, optional
+        Number of synapses, by default 1.
     mem : Dict[str, Array]
         Laplace transforms of memory curves to plot, keyed by plot label
 
@@ -175,14 +179,56 @@ def mem_plot(sss: Array, **mem: ty.Dict[str, Array]) -> ty.Tuple[Figure, Axes, L
     axs : mpl.axes.Axes
         Axes containing plots.
     """
+    pref = np.sqrt(nsyn)
     fig, axs = plt.subplots()
     lns = []
     for label, mem in mem.items():
-        lns.extend(axs.loglog(1/sss, mem * sss, label=label))
+        lns.extend(axs.loglog(1/sss, mem * sss * pref, label=label))
     axs.set_xlabel(r"Time, $r\tau$")
     axs.set_ylabel("SNR")
     return fig, axs, lns
 
+
+# =============================================================================
+# Examples
+# =============================================================================
+
+
+def ser_casc_plot(nst: int, jmp: float, sss: la.lnarray, nsyn: float = 1
+                  ) -> mpl.figure.Figure:
+    """Example plot
+
+    Parameters
+    ----------
+    nst : int
+        Number of states
+    jmp : float
+        Model parameter
+    time : la.lnarray
+        Vector of times to evaluate SNR
+    nsyn : float, optional
+        Number of synapses, by default 1.
+
+    Returns
+    -------
[…]

### 2020-11-02 Code/Python/test_optim_vid.py

@@ -14,7 +14,7 @@ mpt.rc_fonts('sans-serif')
 fmt = 'pdf'
 # -----------------------------------------------------------------------------
 nst, sss, opts, envs, mods = cso.plot.load_data()
-vid = cso.video.EnvelopeFig(sss, envs[1], mods[1], vmax=0.5)
+vid = cso.video.EnvelopeFig(sss, envs[1], mods[1], nsyn=1e4, vmax=0.5)
 ani = cso.video.animate(vid)
 disp = dt.FormattedTempDisplay("frame: {:2d}/{:2d}")
 folder = Path('~/Documents/videos/envelope').expanduser()

### 2020-11-16 Notes/laplace.tex
copiedarea bound proof to laplace notess

@@
 %
+If we regard the \(\wm_k\) as optimisable parameters (limited to the range \( [-1,1] \) \wlg),s the optimal choice is clearly \( \wm_k = \sgn(c_k) \).
 
@@
   ,
-\end{multline}s
+\end{multline}
 %
@@
   \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) 
-      = A(0) = \sqrt{N} \sum_a \CI_a \tau_a.
+      = A(0) = \sqrt{N} \CA = \sqrt{N} \sum_a \CI_a \tau_a.
 \end{equation}
@@
 
-As shown in \cite{Lahiri2013synapse}, this satisfies the inequality
+\subsubsection{Perturbations}\label{sec:areapert}
+
+\paragraph{Triangularity:}\label{sec:triangular}
+
+Consider
+%
+\begin{equation}\label{eq:shiftqderiv}
+  \pdiff{\CL}{\enc_{mn}} =
+  \frac{1}{2f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} - \frac{1}{2f\pot} \pdiff{\CL}{\MMdm\dep_{mn}}
+   = \frac{1}{2}(\mu\pot_{mn} - \mu\dep_{mn}) + \eqm_m (\etwm_m - \etwm_n)
+   = 0.
+\end{equation}
+%
+This corresponds to the shift
+%
+\begin{equation}\label{eq:shiftq}
+  \MMdm\pot_{ij} \to \MMdm\pot_{ij} + \frac{\epsilon_{ij}}{2f\pot},
+  \qquad
+  \MMdm\dep_{ij} \to \MMdm\dep_{ij} - \frac{\epsilon_{ij}}{2f\dep},
+  \qquad
+  \sum_j \epsilon_{ij} = 0,
+\end{equation}
+%
+which leaves $\frg$ unchanged, and therefore $\eq$, $\fptb$ and $\etw$ as well.
+
+Assume $\etwm_m > \etwm_n$. Then
+%
+\begin{equation}\label{eq:lowertriangular}
+ \mu\pot_{mn} - \mu\dep_{mn} < 0
+ \qquad\implies\qquad
+ \mu\dep_{mn} > 0
+ \qquad\implies\qquad
+ \MMdm\dep_{mn} = 0.
+\end{equation}
+%
+Similarly, if $\etwm_m < \etwm_n$, then
+%
+\begin{equation}\label{eq:uppertriangular}
+ \mu\pot_{mn} - \mu\dep_{mn} > 0
+ \qquad\implies\qquad
+ \mu\pot_{mn} >0
+ \qquad\implies\qquad
+ \MMdm\pot_{mn}=0.
+\end{equation}
+%
+Thus, if we arrange the states in order of decreasing $\etw$, $\MMd\pot$ is upper-triangular and $\MMd\dep$ is lower triangular.
+
+We have ignored the possibility that $\eqm_m = 0$, as this would imply that $\fptbm_{im} = \infty$, which would in turn imply that the Markov process is not ergodic.
+
+\paragraph{Increasing \texorpdfstring{$c_k$}{c(k)}:}\label{sec:areacoeffincr}
+
+Consider the following combinations of derivatives:
+%
+\begin{equation}\label{eq:areacoeffincrderiv}
+    \Delta^\nu_{mn} \equiv
+      \frac{1}{\eqm_m} \pdiff{}{\MMdm^\nu_{mn}}
+      + \frac{1}{\eqm_n} \pdiff{}{\MMdm^\nu_{nm}}, 
+\end{equation}
+%
+Note that they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \cref{sec:triangular} about this being satisfied for ergodic chains).
+
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
+%
+\begin{equation}\label{eq:sareacoeffincrprob}
+  \Delta^\nu_{mn} \eqm_i = 0,
+\end{equation}
+%
+which makes it possible to integrate the perturbation:
+%
+\begin{equation}\label{eq:areacoeffincrfinite}
+  \MMd^\nu \to \MMd^\nu + \Eq\inv \boldsymbol{\epsilon},
+  \qquad
+  \begin{aligned}
+  \boldsymbol{\epsilon} &= \boldsymbol{\epsilon}\trans,
+  \\
+  \boldsymbol{\epsilon} \onev &= 0.
+  \end{aligned}
+\end{equation}
+%
+But more interestingly:
+%
+\begin{equation}\label{eq:areacoeffincr}
+    \Delta^\nu_{mn}\CL =
+      \frac{\mu^\nu_{mn}}{\eqm_m} + \frac{\mu^\nu_{nm}}{\eqm_n}
+      + r \prn{c_m - c_n}\prn{\etwm_m - \etwm_n} = 0,
+\end{equation}
+%
+where $c_k$ were defined in \cref{eq:lareacoeffs}, with \(s = 0\).
+
+Using the non-negativity of the Kuhn-Tucker multipliers, $\mu^\nu_{ij}$, \cref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\etwm_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\etwm_k$ are degenerate, we can choose their order to ensure this).
+
+Note that, according to \cref{sec:triangular}, either $\MMdm\pot_{mn}$ or $\MMdm\dep_{mn}$ will be zero at the maximum, therefore we can expect one of the $\mu^\nu_{mn}$ to be non-zero.
+This would rule out degeneracy of the $c_k$ or $\etwm_k$.
+Looking at \cref{eq:shiftqderiv} closely, the only way both $\mu^\nu_{mn}$ could be zero is if $\etwm_m = \etwm_n$ or $\eqm_m = \eqm_n = 0$.
+
+\paragraph{Shortcuts:}\label{sec:shortcuts}
+
+Now consider the following combinations of derivatives for $(n - m)(-1)^\nu > 1$:
+%
+\begin{equation}\label{eq:shortcutderiv}
+  \widetilde{\Delta}^\nu_{mn} \equiv
+    \brk{ \sum_{k=m}^{n \mp 1} \frac{1}{\eqm_{k}} \pdiff{}{\MMdm^\nu_{k,k \pm 1}} }
+    - \frac{1}{\eq_m} \pdiff{}{\MMdm^\nu_{mn}},
+  \qquad \pm = (-1)^\nu.
+\end{equation}
+%
+Once again, they are only well defined if all the states have non-zero equilibrium probabilities (see the comment in \cref{sec:triangular} about this being satisfied for ergodic chains).
+
+One can show that the equilibrium probabilities, $\eq$, are invariant under these operators:
+%
+\begin{equation}\label{eq:shortcutprob}
+  \widetilde{\Delta}^\nu_{g,m} \eqm_i = 0,
+\end{equation}
+%
+which makes it possible to integrate the perturbation:
+%
+\begin{equation}\label{eq:shortcutfinite}
+  \MMd^\nu \to \MMd^\nu + \Eq\inv \boldsymbol{\epsilon}^{\nu mn},
+  \qquad
+  \begin{alignedat}[t]{3}
+    &\prn{{\epsilon}^{\nu mn}}_{mn}
+      &&= -\epsilon,\\
+    &\prn{{\epsilon}^{\nu mn}}_{k,k\pm1}
+      &&= \epsilon
+      &\;&\forall\, k \in [m,n),\\
+    &\prn{{\epsilon}^{\nu mn)}}_{kk}
+      &&= -\epsilon
+      &&\forall\, k \in (m,n).
+  \end{alignedat}
+\end{equation}
+%
+But more interestingly for our purposes:
+%
+\begin{equation}\label{eq:shortcutarea}
+    \widetilde{\Delta}^\nu_{mn}\CL =
+      \brk{ \sum_{k=m}^{n\mp1} \frac{\mu^\nu_{k,k\pm1}}{\eqm_k}
+      - \frac{\mu^\nu_{m,n}}{\eq_m}}
+      + rf^\nu \sum_{k=m}^{n\mp1} \prn{\etwm_{k} - \etwm_{k\pm1}} \prn{c_k - c_m},
+\end{equation}
+%
+
+If we put the states in order of decreasing $\etwm_k$, the results of the \cref{sec:areacoeffincr} tell us that the $c_k$ are non-decreasing.
+This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
+If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu^\nu_{mn} > 0$, which in turn implies that $\MMdm^\nu_{mn}=0$.
+This would tell us that the process with the maximal area has to have a serial topology.
+
+\paragraph{Summary:}\label{sec:pertsummary}
+
+Using the Kuhn-Tucker formalism, we have shown that,
+with the states arranged in order of non-increasing $\etwm_i$:
+%
+\begin{itemize}
+  \item There can be no ergodic maximum for which $\MMd\pot$ contains backwards transitions or $\MMd\dep$ contains forwards transitions.
+  \item There can be no ergodic maximum with the $c_k$ decreasing.
+  \item The $c_k$ may only be degenerate at an ergodic maximum if the corresponding $\etwm_k$ are also degenerate.
+  \item If the $c_k$ increase and the $\etwm_i$ decrease, there can be no ergodic maximum with shortcuts.
+\end{itemize}
+%
+These were shown by finding allowed perturbations that increase the area.
+
+This leaves two possibilities for the maximum area Markov chain.
+Either there is no degeneracy and no shortcuts, which implies the serial topology that we'll discuss in \cref{sec:serialarea}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \cref{eq:shortcutarea}).
+
+Degeneracy tends to be very delicate. 
+It is usually hard to arrange without some symmetry relating degenerate states. 
+Such a symmetry would imply lumpability (see \cite{kemeny1960finite,Ball1993Lumpability,burke1958markovian}). 
+The lumped chain would not have any shortcuts, as an entire degenerate set cannot be bypassed. 
+As this lumped chain has the same area (see \cite{Lahiri2013synapse}), we would need only consider the serial topology.
+
+\subsubsection{Serial topology}\label{sec:serialarea}
+
+The serial topology is defined by (see \cite{Leibold2008serial,Ben-DayanRubin2007sparse}):
+%
+\begin{equation}\label{eq:serialdef}
+  \MMdm\pot_{ij} = q\pot_i \delta_{i+1,j},
+  \qquad
+  \MMdm\dep_{ij} = q\dep_j \delta_{i,j+1},
+  \qquad \text{for}\quad i \neq j
+\end{equation}
+%
+It saturates various inequalities:
+%
+\begin{equation}\label{eq:serialineq}
+  \begin{aligned}
+    \fptbm_{ik} - \fptbm_{jk} &=
+      \begin{cases}
+        \fptbm_{ij},  &\text{if}\quad i \leq j \leq k \quad\text{or}\quad i \geq j \geq k,\\
+        -\fptbm_{ji}, &\text{if}\quad j \leq i \leq k \quad\text{or}\quad j \geq i \geq k,\\
+      \end{cases} \\
+    r\eqm_i \frgm_{ij} \prn{\fptbm_{ij}+\fptbm_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
+  \end{aligned}
+\end{equation}
+%
+and it satisfies detailed balance (a.k.a.\ reversibility):
+%
+\begin{equation}\label{eq:multistateprob}
+  f\pot q\pot_i \eqm_i = f\dep q\dep_i \eqm_{i+1},
+\end{equation}
+%
+which means we can always choose the transition rates, $q^\nu_i$, to give any desired equilibrium probabilities, $\eqm_i$.
+
+This allows us to calculate the $c_k$'s:
+%
+\begin{equation}\label{eq:serialareacoeff}
+\begin{aligned}
+  c_k &= \sum_{i<k} \fptbm_{i,i+1} \prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i}
+    %\\&
+    - \sum_{i\geq k} \fptbm_{i+1,i} \prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i}
+  ,\\
+  c_{k+1} - c_k &= \prn{\fptbm_{k,k+1} + \fptbm_{k+1,k}} \prn{\eqm_{k} \frgm_{k,k+1} + \eqm_{k+1} \frgm_{k+1,k}}
+    =%\\=&\,
+    \frac{2}{r},\\
+  \sum_k c_k \eqm_k &= \sum_{ij} \eqm_i \encm_{ij} (\eta-\eta) = 0,\\
+%  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f^+f^-},
+  \implies c_k &= \frac{2}{r} \prn{k - \sum_j j\eqm_j},
+\end{aligned}
+\end{equation}
+%
+where we used \cref{eq:serialineq} to derive the first two equations respectively and \cref{eq:lkemeny} to derive the third. 
+This allows us to write the area as
+%
+\begin{equation}\label{eq:serialarea}
+ % A = 2\sqrt{N} \sum_k \brk{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \eq_k \w_k.
+  \CA = \frac{2}{r} \sum_k \brk{k - \sum_j j\eq_j} \eqm_k \wm_k
+    = \frac{2}{r} \sum_k \abs{k - \sum_j j\eq_j} \eqm_k ,
+\end{equation}
+%
+where we used $\wm_k=\sgn(c_k)$, as discussed after \cref{eq:lareacoeffs}.
+
+First let us maximise \cref{eq:serialarea} at fixed $\eqm_\pm = \sum_k \eqm_k \prn{\frac{1\pm\wm_k}{2}}$.
+Clearly this will happen when we put all of the probability at the ends: $\eqm_1=\eqm_-$ and $\eqm_n=\eqm_+$ are the only non-zero $\eqm_k$.
+This gives an area of
+\footnote{Note that including the dropped factors in \cref{eq:lsnr} would only change this to \(\CA \leq (M-1)(4\eq_+\eq_-)/r\sqrt{1 - (f\pot-f\dep)^2(\eqm_+ - \eqm_-)^2}\), whose maximum also lies at \(\eqm_+=\eqm_-= \frac{1}{2}\).}
+%
+\begin{equation}\label{eq:serialextarea}
+  \CA \leq \frac{(M-1)}{r}\prn{4\eqm_+\eqm_-}.
+\end{equation}
+%
+This is maximised at \(\eqm_+=\eqm_-= \frac{1}{2}\):
+%
+\begin{equation}\label{eq:maxarea}
+  \CA \leq \frac{(M-1)}{r}.
+\end{equation}
+%
+Note that the markov process that achieves this is not ergodic, the two states at each end are absorbing. 
+This is similar to the results found numerically in \cite{Barrett2008discrete} in a slightly different situation.
+
+This leaves us with the result, first shown in \cite{Lahiri2013synapse}, that
 %
@@
 the two-state model with deterministic transitions.
-This model has \(\CI_1 = \tau_1 = 1\), suggesting that there's a factor of 2 looseness in this bound.
+This model has \(\CI_1 = \tau_1 = 1\), suggesting that there's a factor of \(\sqrt{2}\) looseness in this bound.
 
@@
   \caption[Optimal models]{The models that constitute the numeric envelope in \autoref{fig:envnum}.
-  The plot to the left of each frame indicates the memory curve of the model shown to the right in blue, with the numeric envelope for models with the serial topology in red and the proven envelope, \eqref{eq:env}, in green.
+  The plot to the left of each frame indicates the memory curve of the model shown to the right in green, with the numeric envelope for models with the serial topology in orange and the proven envelope, \eqref{eq:env}, in blue.
+  Above the left plot we have a graphical display of the model, described with more detail to the right.
+  The area of each node is proportional to the steady state probability in that state and the thickness of the red and blue edges are proportional to the transition probabilities under potentiation and depression.
   The bar graph in the upper right of each frame indicates the transition probabilities from the state to the left of each bar to the state on its right, under potentiation.

### 2021-04-22 Code/Python/complex_synapse/optimise/plot.py
using property.setter in options

@@ -257,7 +257,7 @@ def theory_plot(nst: int, sss: Array, nsyn: float = 1) -> Figure:
     axs.loglog(t_pts[:2], np.ones(2), 'k:')
     axs.loglog(t_pts[-2:], (nst-1) / t_pts[-2:], 'k:')
     axs.set_xlim(1e-1, 1e3)
-    axs.set_ylim(1e-2, 3)
+    axs.set_ylim(1e-2 * np.sqrt(nsyn), 3 * np.sqrt(nsyn))
     ind = 30
     axs.text(t_pts[1]/2, 1.1, r"$\sqrt{N}$", fontsize=18,
              ha='right', va='bottom')
@@ -298,7 +298,7 @@ def equilibrium_plot(nst: int, sss: Array, nsyn: float = 1) -> Figure:
     # axs.loglog(t_pts[:2], np.ones(2), 'k:')
     # axs.loglog(t_pts[-2:], (nst-1) / t_pts[-2:], 'k:')
     axs.set_xlim(1e-1, 1e3)
-    ylim = axs.set_ylim(1e-2, 3)
+    ylim = axs.set_ylim(1e-2 * np.sqrt(nsyn), 3 * np.sqrt(nsyn))
     s_two, s_sticky = 0.5, 2 / (nst-1)**2
     _annotate_axis(axs, nst, ylim[0], [s_two, s_sticky],
                 ["2", r"\frac{(M-1)^2}{2}"])
@@ -447,7 +447,7 @@ def heuristic_plot(nst: int, sss: Array, env_srl: Array, models_srl: Array,
     lns[2].set_c(lns[1].get_c())
     lns[1].set_c(lns[0].get_c())
     axs.set_xlim(1/sss[-1], 1e4)
-    ylim = axs.set_ylim(2e-3, 1.1)
+    ylim = axs.set_ylim(2e-3 * np.sqrt(nsyn), 1.1 * np.sqrt(nsyn))
 
     _annotate_curves(axs, 1/sss, env_heu * sss, models_srl, **kwds)
     _annotate_axis(axs, nst, ylim[0], [sh.s_star(2), sh.s_star(nst)],

### 2021-07-27 Notes/laplace.tex
using property.setter in options
tex problems
typos

@@
 %
-\usepackage{adjustbox}
-\newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
-\newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
+% \usepackage{adjustbox}
+% \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
+% \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
 %
@@
          &=  (f\pot-f\dep)\, \eq \w \\
-         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
-\end{aligned}
-\end{equation}
-%
-Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
+         &=  \Delta f\, \eq \e^{rt\frg} \w ,
+\end{aligned}
+\end{equation}
+%
+where \(\Delta f = f\pot - f\dep \).
+Combining the expressions in \cref{eq:overlapwt} allows us to write the numerator of \cref{eq:SNRdef} as
 %
@@
   \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt \frg}\w}
-                 {\sqrt{1-(f\pot-f\dep)^2(\eq \w)^2}}.
+                 {\sqrt{1-(\Delta f)^2(\eq \w)^2}}.
 \end{equation}
@@
 This reduces to the fundamental matrix at \(s=0\).
-We can see that \(\onev\) is an eigenvector:
+We can see that it obeys similar identities, including that \(\onev\) is an eigenvector:
 %
 \begin{equation}\label{eq:lfundrowsum}
-  \fund(s)\onev = \frac{\tau_0}{1+s\tau_0}\onev.
+\begin{aligned}
+  \fund(s) \onev &= \frac{\tau_0}{1+s\tau_0}\onev, \qquad &
+  \MM \fund(s) &= (\onev \eq - \I)(\I - s \fund), \\ 
+  \arow \fund(s) &= \eq (\I - s \fund),  &
+  \fund(s) \MM &= (\I - s \fund)(\tau_0 \onev \arow - \I).
+\end{aligned}
 \end{equation}
@@
 This is a valid continuous-time Markov process.
-Its fundamental matrix is given by \cref{eq:lfund} with \( \arow_0 = \arow + s \eq \).
+Its fundamental matrix is given by \cref{eq:lfund} with \( \arow(s) = \arow + s \eq \) (although this last change is unnecessary).
 
@@
    &= \intd[_0^{\infty}]{t} \frac{ \sqrt{N} \eq \enc \e^{(r\frg-s)t} \w }
-                                { \sqrt{1-(f\pot-f\dep)^2(\eq\w)^2} } \\
-   &= \frac{\sqrt{N} \eq \enc \prn{s \I - r\frg}\inv \w }{\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
+                                { \sqrt{1-(\Delta f)^2(\eq\w)^2} } \\
+   &= \frac{\sqrt{N} \eq \enc \prn{s \I - r\frg}\inv \w }{\sqrt{1-(\Delta f)^2(\eq\w)^2}}.
 \end{aligned}
@@
 \begin{aligned}
-  \hat{A}(s) = A(s) \sqrt{\frac{{{1-(f\pot-f\dep)^2(\eq\w)^2}}}{N}}  &=  \eq \enc \fund(s) \w \\
+  \hat{A}(s) = A(s) \sqrt{\frac{{{1-(\Delta f)^2(\eq\w)^2}}}{N}}  &=  \eq \enc \fund(s) \w \\
     &= \sum_{ijk} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)} \eqm_k \wm_k \\
@@
 
-This allows us to calculate the $c_k$'s:
+This allows us to calculate the $c_k$'s from \cref{eq:lareacoeffs}:
 %
@@
 \begin{aligned}
-  c_k &= \sum_{i<k} \fptbm_{i,i+1} \prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i}
+  c_k &= \sum_{i<k} \prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i} \fptbm_{i,i+1}
     %\\&
-    - \sum_{i\geq k} \fptbm_{i+1,i} \prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i}
+    - \sum_{i\geq k} \prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i} \fptbm_{i+1,i}
   ,\\
-  c_{k+1} - c_k &= \prn{\fptbm_{k,k+1} + \fptbm_{k+1,k}} \prn{\eqm_{k} \frgm_{k,k+1} + \eqm_{k+1} \frgm_{k+1,k}}
-    =%\\=&\,
-    \frac{2}{r},\\
+  c_{k+1} - c_k &= \prn{\eqm_{k} \frgm_{k,k+1} + \eqm_{k+1} \frgm_{k+1,k}}
+                    \prn{\fptbm_{k,k+1} + \fptbm_{k+1,k}}
+    = \frac{2}{r},\\
   \sum_k c_k \eqm_k &= \sum_{ij} \eqm_i \encm_{ij} (\eta-\eta) = 0,\\
-%  \implies c_k =&\, \frac{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j}{f^+f^-},
   \implies c_k &= \frac{2}{r} \prn{k - \sum_j j\eqm_j},
@@
 This gives an area of
-\footnote{Note that including the dropped factors in \cref{eq:lsnr} would only change this to \(\CA \leq (M-1)(4\eq_+\eq_-)/r\sqrt{1 - (f\pot-f\dep)^2(\eqm_+ - \eqm_-)^2}\), whose maximum also lies at \(\eqm_+=\eqm_-= \frac{1}{2}\).}
+\footnote{Note that including the dropped factors in \cref{eq:lsnr} would only change this to \(\CA \leq (M-1)(4\eq_+\eq_-)/r\sqrt{1 - (\Delta f)^2(\eqm_+ - \eqm_-)^2}\), whose maximum also lies at \(\eqm_+=\eqm_-= \frac{1}{2}\).}
 %
@@
      &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     - \sum_{ij} \mu^\nu_{ij} s f^\nu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
+     - \sum_{\mu ij} \mu^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
     = 0.
@@
   \item\aligntop{\includegraphics[width=0.45\linewidth,page=50]{envelope.pdf}}
-  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv07.eps}}
-  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv13.eps}}
-  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv21.eps}}
-  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv27.eps}}
-  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv36.eps}}
-  % \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv51.eps}}
   \end{myenuma}

### 2021-07-29 Notes/laplace.tex
topology constraints

@@
 \usepackage{a4wide}
-\usepackage[tables,graphics,draft]{sl-preamble}
-% \usepackage{stix}
+% \usepackage{stix2}
+\usepackage[tables,graphics,draft,panels=paralist]{sl-preamble}
 %\usepackage{amsldoc} for finding documentation only
@@
 %================================================================================
-\makeatletter
-\DeclareFontEncoding{LS2}{}{\noaccents@}
-\makeatother
-
-\DeclareFontSubstitution{LS2}{stix}{m}{n}
-\DeclareSymbolFont{xlargesymbols}{LS2}{stixex}{m}{n}
-\DeclareMathDelimiter{\lAngle}{\mathopen}{xlargesymbols}{"EC}{xlargesymbols}{"12}
-\DeclareMathDelimiter{\rAngle}{\mathclose}{xlargesymbols}{"ED}{largesymbols}{"13}
-\DeclareMathDelimiter{\lVvert}{\mathopen}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
-\DeclareMathDelimiter{\rVvert}{\mathclose}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
-
-\newcommand{\aav}[1]{\mathopen{}\left\lAngle #1 \right\rAngle\mathclose{}}
-\newcommand{\nnrm}[1]{\mathopen{}\left\lVvert #1 \right\rVvert\mathclose{}}
+% \makeatletter
+% \DeclareFontEncoding{LS2}{}{\noaccents@}
+% \makeatother
+
+% \DeclareFontSubstitution{LS2}{stix}{m}{n}
+% \DeclareSymbolFont{xlargesymbols}{LS2}{stixex}{m}{n}
+% \DeclareMathDelimiter{\lAngle}{\mathopen}{xlargesymbols}{"EC}{xlargesymbols}{"12}
+% \DeclareMathDelimiter{\rAngle}{\mathclose}{xlargesymbols}{"ED}{largesymbols}{"13}
+% \DeclareMathDelimiter{\lVvert}{\mathopen}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
+% \DeclareMathDelimiter{\rVvert}{\mathclose}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
+
+\newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
+\newcommand{\nnrm}[1]{\mathopen{}\left\lWavy #1 \right\rWavy\mathclose{}}
 %================================================================================
@@
 % \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
-%
-\usepackage{paralist}
-\newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
-\newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
-\newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
-\newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
-\newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
-%for nested lists, \ref only produces innermost item:  (hacky)
-\makeatletter % for internal macros with @
-\renewcommand\p@enumii{}
-\makeatother
+% %
+% \usepackage{paralist}
+% \newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
+% \newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
+% \newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
+% \newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
+% \newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
+% %for nested lists, \ref only produces innermost item:  (hacky)
+% \makeatletter % for internal macros with @
+% \renewcommand\p@enumii{}
+% \makeatother
 %
@@
 \newcommand{\alb}{\boldsymbol{\albm}}
+\newcommand{\dgnm}{\zeta}
+\newcommand{\dgn}{\boldsymbol{\dgnm}}
 %synapse models
@@
 
-As discussed in \autoref{sec:recog}, the signal-to-noise ratio is given by
+As discussed in \cref{sec:recog}, the signal-to-noise ratio is given by
 %
@@
   \fund(s) \onev &= \frac{\tau_0}{1+s\tau_0}\onev, \qquad &
-  \MM \fund(s) &= (\onev \eq - \I)(\I - s \fund), \\ 
-  \arow \fund(s) &= \eq (\I - s \fund),  &
-  \fund(s) \MM &= (\I - s \fund)(\tau_0 \onev \arow - \I).
+  \MM \fund(s) &= (\onev \eq - \I)(\I - s \fund(s)), \\ 
+  \arow \fund(s) &= \eq (\I - s \fund(s)),  &
+  \fund(s) \MM &= (\I - s \fund(s))(\tau_0 \onev \arow - \I).
 \end{aligned}
@@
 This is a valid continuous-time Markov process.
-Its fundamental matrix is given by \cref{eq:lfund} with \( \arow(s) = \arow + s \eq \) (although this last change is unnecessary).
+Its fundamental matrix is given by \cref{eq:lfund} with \( \arow(s) = \arow - s \eq \) (although this last change is unnecessary).
+
+The usual fundamental matrix identities will hold provided we use \cref{eq:lshift} as the transition matrix and we note that the \( \arow \) we used should be interpreted as \( \arow(s)\) and the ``original'' \( \arow \) is \( \arow + s \eq \equiv \shift{\arow}\) and the ``original'' \( \frac{1}{\tau_0} \) is \( \frac{1}{\tau_0} + s \equiv \frac{1}{\shift{\tau}_0}\).
+We could rewrite \cref{eq:lfundrowsum} as
+%
+\begin{equation}\label{eq:lshiftrowsum}
+\begin{aligned}
+  \fund(s) \onev &= {\shift{\tau}_0}\onev, \qquad &
+  \shift{\MM}(s) \fund(s) &= \onev \eq - \I, \\ 
+  \shift{\arow} \fund(s) &= \eq,  &
+  \fund(s) \shift{\MM}(s) &= \shift{\tau} \onev \shift{\arow} - \I.
+\end{aligned}
+\end{equation}
+%
 
 This means that the \( \fptb(s) \) have all the same properties as in the case \(s=0\).
+This includes non-negativity and all of the triangle inequalities.
+Identities involving mean first-passage times and the transition matrix will also hold, provided that \cref{eq:lshift} is used as the transition matrix.
 In particular, we see that the quantity
@@
 
-In analogy with the case \(s=0\), we define
+%--------------------------------------------------------------------------------
+
+\subsection{Derivatives}\label{sec:lderiv}
+
+In analogy with the case \(s=0\) in \cite{Lahiri2013synapse}, we define
 %
@@
     \hat{A}(s) 
-      = \sum_k \eqm_k {c}_k(s) \wm_k
-      % = \mathbf{c}^\dag(s) \w
+      % = \sum_k \eqm_k {c}_k(s) \wm_k
+      = \mathbf{c}^\dag(s) \w
       = \eq \mathbf{a}(s).
@@
 %
-If we regard the \(\wm_k\) as optimisable parameters (limited to the range \( [-1,1] \) \wlg),s the optimal choice is clearly \( \wm_k = \sgn(c_k) \).
-
-%--------------------------------------------------------------------------------
-
-\subsection{Derivatives}\label{sec:lderiv}
+If we regard the \(\wm_k\) as optimisable parameters (limited to the range \( [-1,1] \) wlg.),s the optimal choice is clearly \( \wm_k = \sgn(c_k) \).
 
@@
 
+\section{Finite time topology}\label{sec:finite}
+
+Now we'll maximise \(\snrb(\tau)\) for some fixed \(\tau\), subject to the constraints \eqref{eq:constr}.
+First, we will maximise the numerator of \eqref{eq:lsnr} holding \(\eq\w\) fixed.
+We can maximise the ratio \wrt \(\eq\w\) later.
+From this point on, we work in units where \(r = 1\).
+
+Consider the Lagrangian
+%
+\begin{equation}\label{eq:finitelagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
+        +\sum_{\nu i j} { f^\nu \mu^\nu_{ij} \MMdm^\nu_{ij} }
+        + \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
+\end{aligned}
+\end{equation}
+%
+where \(\Delta\pr\) is the constant value we are holding \(\eq\w\) at.
+The Karush-Kuhn-Tucker necessary conditions for a maximum are
+%
+\begin{equation}\label{eq:KTfinite}
+  \pdiff{\CL}{\MMdm^\nu_{ij}} = 0, \quad
+  \mu^\nu_{ij} \geq 0, \quad
+  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
+  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
+  \pdiff{\CL}{\lambda} = 0.
+\end{equation}
+%
+We have scaled the KKT multipliers by \(f^\nu\) for later convenience.
+
+The derivatives, for \(m \neq n\) are
+%
+\begin{equation}\label{eq:finitederivs}
+\begin{aligned}
+  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+    \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
+     + \prn{\mathbf{c}_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
+     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
+     + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
+    = 0.
+\end{aligned}
+\end{equation}
+%
+We can solve for the KKT multipliers by noting that
+%
+\begin{equation*}
+  \mu^\nu_{ij} \MMdm^\nu_{ij} = 0,
+  \qquad \text{and} \qquad
+  \sum_{j: j \neq i} \mu^\nu_{ii} \MMdm^\nu_{ij} 
+      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii}) = \mu^\nu_{ii}.
+\end{equation*}
+%
+Then, with some new vectors defined:
+%
+\begin{equation}\label{eq:kktsol}
+\begin{aligned}
+  \mu^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
+    + \lambda \ombm^\nu_m(0) 
+    + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
+    }, \\
+  \mu^\nu_{mn} &= \eqm_m \big( 
+      \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
+    + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
+  \\ &\phantom{= \eqm_m \big( }
+    + \brk{{c}_m(s) + (-1)^\nu} 
+        \brk{ \etwm_n(s) - \etwm_m(s) - \ombm^\nu_m(s) }
+     \big), \\
+  &\text{where} \quad 
+  \alb^\nu(s) = (\MMd^\nu - \I) \thb(s), \qquad
+  \omb^\nu(s) = (\MMd^\nu - \I) \etw(s).
+\end{aligned}
+\end{equation}
+%
+If we note that we can write \( \MMd^\nu - \I = \prn{\frg + (-1)^\nu \enc} / 2f^\nu \), we find that
+%
+\begin{equation}\label{eq:diagKKT}
+\begin{gathered}
+  \alb^\nu(s) = \frac{\delta \mathbf{a}(s) + (-1)^\nu \enc \thb(s)}{2f^\nu},
+  \qquad 
+  \omb^\nu(s) = \frac{\delta \w + s \delta \etw(s) - (-1)^\nu \mathbf{a}(s)}{2 f^\nu}, \\
+\begin{aligned}
+  \mu^\nu_{mm} &= - \frac{\eqm_m}{2f^\nu} 
+    \left[ 
+      c_m(s) (\delta \wm_m + s \delta \etwm_m(s)) - \hat{A}(s)
+      + \lambda \delta w_m
+      \right. \\ &\phantom{= - \frac{\eqm_m}{2f^\nu} []}\left.
+      + (-1)^\nu \prn{ 
+        [\encm \thbm(s)]_m - c_m(s) a_m(s)
+        + \delta \wm_m + s \delta \etwm_m(s)
+        - \lambda a_m(0)
+      }
+    \right], \\
+\end{aligned}
+\end{gathered}
+\end{equation}
+%
+where \(\delta\) indicates mean subtraction and \(\eq \mathbf{a}(s) = \hat{A}(s)\).
+
+The expressions in \cref{eq:finitederivs} have the same form as for \(s=0\), which allowed us to argue that the model with maximal area must have the serial topology, \cite{Lahiri2013synapse}.
+The one missing ingredient is the scale invariance.
+
+However, we do have numerical evidence that the model that maximises the Laplace transform does have the serial topology.
+In \cref{fig:envnum} we see the result of numerical maximisation of \cref{eq:finitelagrangian}.
+We see that restricting to the serial topology does not make the performance any worse.
+In fact, it slightly improves the performance, but this can be ascribed to the numerical optimisation problem being easier.
+Assuming that the optimiser did find the global maximum, this provides evidence for our conjecture that the model that maximises the Laplace transform must have the serial topology.
+
+Note that models with the serial topology must satisfy detailed balance.
+Therefore, this conjecture implies that the square-root bound of \cref{sec:sqrt} applies, implying the stricter envelope of \cref{eq:envDB}.
+In \cref{fig:envnum}, we can see that this envelope is close to tight, being off by a constant factor.
+
+\begin{figure}[tb]
+  \centering
+  % Requires \usepackage{graphicx}
+  \includegraphics[width=0.8\linewidth]{LenvNum.pdf}\\
+  \caption[Numerical envelope for the signal-to-noise ratio]
+  {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
+  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numericallly with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
+  The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
+  The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
+  The dotted blue line indicates the proven envelope, \cref{eq:env} and the solid blue line is the conjectured envelope assuming detailed balance, see \cref{eq:envDB}.}\label{fig:envnum}
+\end{figure}
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Scale invariance}\label{sec:scale}
+
+Under the transformation from \cref{eq:scale}, the Laplace transform has the following scaling property
+%
+\begin{equation}\label{eq:laplacescale}
+  \hat{A} \prn{\zeta s;\zeta * \MMd^\nu} = \hat{A} \prn{s; \MMd^\nu}.
+\end{equation}
+%
+At \(s = 0\), it is scale invariant.
+This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
+This allowed us to ignore the diagonal constraints.
+We can't do that for \(s > 0\).
+
+The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
+%
+\begin{equation}\label{eq:shiftscale}
+  \zeta {\circledast}_{s} \MMd 
+      \equiv \zeta \MMd
+      + (1 - \zeta) \brk{(1 + s)\I - s \onev \eq} .
+\end{equation}
+%
+Then we have
+%
+\begin{equation}\label{eq:lshiftscale}
+  \hat{A} \prn{s; \MMd^\nu} 
+    = \hat{A} \prn{0, \shift{\MMd}^\nu(s)}
+    = \hat{A} \prn{0, \zeta * \shift{\MMd}^\nu(s)}
+    = \hat{A} \prn{s, \zeta \circledast_s \MMd^\nu}.
+\end{equation}
+%
+The problem here is that the lower bound, \(\MMdm^\nu_{ij} \geq 0\), 
+is not invariant under this transformation, unlike the \(s = 0\) case.
+Using this transformation to fix the upper bound can break the lower bound.
+
+We might ask what conditions are required so that a model is fixable?
+This means that the \(\zeta \circledast_s \) orbit of the model contains some valid models.
+The conditions are
+%
+\begin{equation*}
+\begin{alignedat}{5}
+  [\zeta \circledast \MMd^\nu]_{ij}  
+    &= \zeta \MMdm^\nu_{ij} - (1-\zeta) s \eqm_j 
+    &&\geq 0 &
+  \quad &\implies \quad &
+  \zeta &\geq \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} &
+    &\forall \nu, i, j : i \neq j , \\
+  \sum_{j : i \neq j} [\zeta \circledast \MMd^\nu]_{ij} 
+    &= \zeta \lbm^\nu_i - (1-\zeta) s (1-\eqm_i) 
+    &&\leq 1 &
+  &\implies &
+  \zeta &\leq \frac{1 + s (1-\eqm_i)}{\lbm^\nu_i + s (1-\eqm_i)} &
+    \quad &\forall \nu, i ,
+\end{alignedat}
+\end{equation*}
+%
+where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \MMdm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
+The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
+%
+\begin{equation}\label{eq:fixable}
+  \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} \leq \frac{1 + s (1-\eqm_k)}{\lbm^\mu_k + s (1-\eqm_k)}
+  \quad \implies \quad
+  \frac{\lbm^\mu_k - 1}{1 + s (1-\eqm_k)} \leq \frac{\MMdm^\nu_{ij}}{s \eqm_j} 
+  \quad \forall \mu, \nu, i, j, k : i \neq j.
+\end{equation}
+%
+Note that when \(\MMdm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq 1\), the original diagonal constraint.
+Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Shifted problem}\label{sec:lshift}
+
+The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
+treat the off-diagonal components of \(\shift{\MMd}^\nu \) as the degrees of freedom and maximise the area instead.
+The constraints on \(\shift{\MMdm}^\nu_{ij}\) are different from \cref{eq:constri}:
+%
+\begin{equation}\label{eq:lshiftconstr}
+  \shift{\MMdm}^\nu_{ij} \geq s \eqm_j,
+    \quad \forall \; \nu, i, j : i \neq j, \qquad
+  \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \leq 1 + s (1 - \eqm_i).
+    \quad \forall \; \nu, i.
+\end{equation}
+%
+The diagonal elements are still given by \( \shift{\MMdm}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \), but now they are allowed to be negative so long as \(\shift{\MMdm}^\nu_{ii} \geq - s (1 - \eqm_i)\).
+
+The Lagrangian we should consider is
+%
+\begin{equation}\label{eq:lshiftlagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
+        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\MMdm^\nu_{ij} -  s (\eqm_j - \delta_{ij}) }
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
+\end{aligned}
+\end{equation}
+%
+where all of the first-passage times are evaluated at \(s = 0\).
+
+In \cref{fig:shifted} we show the result of solving this problem with numerical optimisation.
+We see essentially the same results as the original problem, \cref{eq:finitelagrangian,fig:envnum}, as expected.
+
+The derivatives, for \(m \neq n\) are
+%
+\begin{equation}\label{eq:lshiftderivs}
+\begin{aligned}
+  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+    \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
+     + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
+     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
+     - \sum_{\mu ij} \mu^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
+    = 0.
+\end{aligned}
+\end{equation}
+%
+We can absorb the new terms by defining
+%
+\begin{equation*}
+\begin{aligned}
+  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
+  \qquad
+  \shift{\alb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
+  \\
+  \shift{\omb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \etw(0).
+\end{aligned}
+\end{equation*}
+%
+We note that
+%
+\begin{equation*}
+  \mu^\nu_{ij} (\MMdm^\nu_{ij} - s \eqm_j) = 0,
+  \qquad \text{and} \qquad
+  \sum_{j: j \neq i} \mu^\nu_{ii} (\MMdm^\nu_{ij} - s \eqm_j)
+      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
+\end{equation*}
+%
+then we can ``solve'' for the KKT multipliers:
+%
+\begin{equation}\label{eq:lshiftsol}
+\begin{aligned}
+  \mu^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
+    }, \\
+  \mu^\nu_{mn} &= \eqm_m \prn{ 
+      \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
+  % \\ & \phantom{= - \eqm_m \big( } 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } 
+        \brk{ \etwm_n - \etwm_m - \shift{\ombm}^\nu_m(s) }
+     }, \\
+\end{aligned}
+\end{equation}
+%
+although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\).
+
+While this problem involves the scale invariant quantity \(\hat{A}(0)\), neither the upper nor lower bounds in \cref{eq:lshiftconstr} are invariant.
+We then run into the same problems in proving our serial-topology hypothesis.
+
+\begin{figure}[ht]
+\begin{center}
+\begin{myenuma}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{LenvShift.pdf}}
+  \label{fig:envshift}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{shift_cond.pdf}}
+  \label{fig:shiftcond}
+\end{myenuma}
+\caption[Shifted problem vs.\ original]{Shifted problem vs.\ original.
+(\ref*{fig:envshift}) Comparison of numerical optimisation of the shifted problem in \cref{eq:lshiftlagrangian,fig:envnum} with the original problemin \cref{eq:finitelagrangian}.
+We see that the two approaches produce very similar numerical envelopes.
+The shifted problem involves non-linear constraints and is thus harder to solve.
+(\ref{fig:shiftcond}) Plot of the discrepancy in (\ref*{fig:envshift}) against the condition number of the fundamental matrix, \(\fund \).
+We see that most of the large discrepancies occur where the problem is numerically ill-conditioned.
+\label{fig:shifted}}
+\end{center}
+\end{figure}
+
+\begin{figure}[tbp]
+  \centering
+  % Requires \usepackage{graphicx}
+  \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=7]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=13]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=21]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=27]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=36]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=50]{envelope.pdf}}
+  \end{myenuma}
+  \caption[Optimal models]{The models that constitute the numeric envelope in \cref{fig:envnum}.
+  The plot to the left of each frame indicates the memory curve of the model shown to the right in green, with the numeric envelope for models with the serial topology in orange and the proven envelope, \eqref{eq:env}, in blue.
+  Above the left plot we have a graphical display of the model, described with more detail to the right.
+  The area of each node is proportional to the steady state probability in that state and the thickness of the red and blue edges are proportional to the transition probabilities under potentiation and depression.
+  The bar graph in the upper right of each frame indicates the transition probabilities from the state to the left of each bar to the state on its right, under potentiation.
+  The bar graph in the lower right of each frame indicates the transition probabilities from the state to the right of each bar to the state on its left, under depression.
+  The heat map in the centre right of each frame indicates the equilibrium probability distribution for each state.
+  See the video for the full set of models.}\label{fig:envvid}
+\end{figure}
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Topology constraining perturbations}\label{sec:topology}
+
+In this section we will see how far we can get in constraining the topology, 
+despite the lack of scale invariance.
+
+%-------Section------------------------------------------------------------------
+
+\subsubsection{Number of exits}\label{sec:exit}
+
+Let us define the following
+%
+\begin{equation}\label{eq:degenerate}
+  \dgn_{m} = \thb + c_m \etw,
+  \quad
+  \dgn^\nu_m = \dgn_m + (-1)^\nu \etw,
+  \quad \implies \;
+  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
+\end{equation}
+%
+They should be thought of as collections of vectors rather than matrices.
+
+Suppose that a state \(n\) has a nonzero transition rate from state \(m\) for one \(\nu\).
+Then \(\mu^\nu_{mn} = 0\) and:
+%
+\begin{equation}\label{eq:oneexit}
+  \MMdm^\nu_{mn} > 0
+  \quad \implies \quad
+  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n})
+  \quad \implies \quad
+  \dgnm_{m,m} - \dgnm_{m,n} \geq - (-1)^\nu (\etwm_m - \etwm_n).
+\end{equation}
+%
+If both \(\nu\) have nonzero rates from \(m\) to \(n\):
+%
+\begin{equation}\label{eq:doubleexit}
+  \MMdm\pot_{mn}, \MMdm\dep_{mn} > 0
+  \qquad \implies \qquad
+  \dgnm_{m,m} - \dgnm_{m,n} \geq\abs{\etwm_m - \etwm_n}.
+\end{equation}
+%
+
+Suppose there are two states, \(i,j\), with nonzer transitions from \(m\) for one \(\nu\).
+Both must statisfy \cref{eq:oneexit} for the same \(\mu^\nu_{mm}\), so
+%
+\begin{equation}\label{eq:twoexits}
+  \MMdm^\nu_{mi}, \MMdm^\nu_{mj} > 0
+  \qquad \implies \qquad
+  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
+\end{equation}
+%
+Furthermore, for a state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
+%
+\begin{equation}\label{eq:nonexit}
+  \MMdm^\nu_{mi} > 0, \MMdm^\nu_{mk} = 0 
+  \;\;\; \implies \;\;\;
+  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
+  \;\; \implies \;\;
+  \dgnm_{m,k} - \dgnm_{m,i} \geq - (-1)^\nu (\etwm_k - \etwm_i).
+\end{equation}
+%
+If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
+%
+\begin{equation}\label{eq:twodoubleexits}
+  \MMdm\pot_{mi}, \MMdm\dep_{mi}, \MMdm\pot_{mj}, \MMdm\dep_{mj} > 0
+  \qquad \implies \qquad
+  \dgnm_{m,i} = \dgnm_{m,j} 
+  \quad \text{and} \quad
+  \etwm_i = \etwm_j.
+\end{equation}
+%
+Both \cref{eq:twoexits,eq:twodoubleexits} require degeneracy.
+The physicist's intuition states that degeneracy is fragile and requires a symmetry to protect it.
+Such a symmetry would imply that the states are lumpable.
+This would imply that, after lumping all lumpable states together, the situation in \cref{eq:twoexits} is impossible and \cref{eq:twodoubleexits} is even more impossible.
+
+%-------Section------------------------------------------------------------------
+
+\subsubsection{Triangularity}\label{sec:parttriangular}
+
+Suppose we have two states, \(i\) and \(j\), with \( \etwm_i > \etwm_j \).
+Consider the following perturbation:
+%
+\begin{equation}\label{eq:pertKupdn}
+\begin{aligned}
+  2 \brk{\pdiff{\CL}{\encm_{mj}} - \pdiff{\CL}{\encm_{mi}}}
+  &= \frac{1}{f\pot} \brk{\pdiff{\CL}{\MMdm\pot_{mj}} - \pdiff{\CL}{\MMdm\pot_{mi}}} -
+  \frac{1}{f\dep} \brk{\pdiff{\CL}{\MMdm\dep_{mj}} - \pdiff{\CL}{\MMdm\dep_{mi}}} \\ 
+    &= 2 \eqm_m (\etwm_i - \etwm_j) 
+    + \mu\pot_{mj} - \mu\pot_{mi} - \mu\dep_{mj} + \mu\dep_{mi}.
+\end{aligned}
+\end{equation}
+%
+Because \( \etwm_i > \etwm_j \), we must have \( \mu\pot_{mi} + \mu\dep_{mj} > 0 \).
+So \( \MMdm\pot_{mi} > 0 \) would imply that \( \MMdm\dep_{mj} = 0 \) and vice versa.
+Let us arrange the states left to right in order of decreasing \(\etw\), so that \(i\) is to the left of \(j\).
+For the unconstrained transitions \( \MMdm\pot_{mj} \) and \( \MMdm\dep_{mi} \) the potentiating transition goes to the \emph{right} of the depressing transition.
+For the other transitions \( \MMdm\pot_{mi} \) and \( \MMdm\dep_{mj} \) the potentiating transition goes to the \emph{left} of the depressing transition, and they cannot both be nonzero.
+
+This logic can be applied to any pair of states with different values of \(\etw\).
+Therefore, when starting from the same state, no potentiating transition can go to the left of any depressing transition, and no depressing transition can go to the right of any potentiating transition.
+
+For any state \(m\), there is a state \(n\) such that all potentiating transitions go to equal or smaller values of \(\etw\) all depressing transitions go to equal or greater values.
+%
+\begin{equation}\label{eq:separator}
+  \forall\ m, \;\; \exists\ n : 
+  \begin{cases}
+    \etwm_i > \etwm_n &\implies \MMdm\pot_{mi} = 0, \\
+    \etwm_i < \etwm_n &\implies \MMdm\dep_{mi} = 0.
+  \end{cases}
+\end{equation}
+%
+We have not yet ruled the possibility that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero.
+
+Consider the perturbation
+%
+\begin{equation}\label{eq:pertK}
+  \frac{1}{f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} -
+  \frac{1}{f\dep} \pdiff{\CL}{\MMdm\dep_{mn}} =
+    2 \eqm_m (\etwm_m - \etwm_n) 
+    + \mu\pot_{mn} - \mu\pot_{mm} - \mu\dep_{mn} + \mu\dep_{mm}.
+\end{equation}
+%
+If we assume that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero, we find that \( \mu\pot_{mm} - \mu\dep_{mm} = 2 \eqm_m (\etwm_m - \etwm_n) \).
+Therefore \( \etwm_m > \etwm_n \implies \MMdm\pot_{mm} = 0 \) and \( \etwm_m < \etwm_n \implies \MMdm\dep_{mm} = 0 \).
+It is also possible that both of the diagonal elements are zero.
+
+%================================================================================
+
 \section{Upper bounds}\label{sec:upperbnds}
@@
 %
-We have ignored thedenominator of \cref{eq:SNRcurveExact} for now.
+We have ignored the denominator of \cref{eq:SNRcurveExact} for now.
 This will not have any effect on the results: it will not change the location of the maxima, which are at \( f\pot - f\dep = \eq \w = 0 \).
@@
 
-\section{Finite time}\label{sec:finite}
-
-Now we'll maximise \(\snrb(\tau)\) for some fixed \(\tau\), subject to the constraints \eqref{eq:constr}.
-First, we will maximise the numerator of \eqref{eq:lsnr} holding \(\eq\w\) fixed.
-We can maximise the ration \wrt \(\eq\w\) later.
-From this point on, we work in units where \(r = 1\).
-
-Consider the Lagrangian
-%
-\begin{equation}\label{eq:finitelagrangian}
-\begin{aligned}
-  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
-        +\sum_{\nu i j} \prn{ f^\nu \mu^\nu_{ij} \MMdm^\nu_{ij} }
-        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
-\end{aligned}
-\end{equation}
-%
-where \(\Delta\pr\) is the constant value we are holding \(\eq\w\) at.
-The Karush-Kuhn-Tucker necessary conditions for a maximum are
-%
-\begin{equation}\label{eq:KTfinite}
-  \pdiff{\CL}{\MMdm^\nu_{ij}} = 0, \quad
-  \mu^\nu_{ij} \geq 0, \quad
-  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
-  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
-  \pdiff{\CL}{\lambda} = 0.
-\end{equation}
-%
-We have scaled the KKT multipliers by \(f^\nu\) for later convenience.
-
-The derivatives, for \(m \neq n\) are
-%
-\begin{equation}\label{eq:finitederivs}
-\begin{aligned}
-  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
-    \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
-     + \prn{\mathbf{c}_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
-     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
-    = 0.
-\end{aligned}
-\end{equation}
-%
-We can solve for the KKT multipliers by noting that
-%
-\begin{equation*}
-  \mu^\nu_{ij} \MMdm^\nu_{ij} = 0,
-  \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} \MMdm^\nu_{ij} 
-      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii}) = \mu^\nu_{ii}.
-\end{equation*}
-%
-Then, with some new vectors defined:
-%
-\begin{equation}\label{eq:kktsol}
-\begin{aligned}
-  \mu^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
-    + \lambda \ombm^\nu_m(0) 
-    + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
-    }, \\
-  \mu^\nu_{mn} &= \eqm_m \big( 
-      \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
-    + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
-  \\ &\phantom{= \eqm_m \big( }
-    + \brk{{c}_m(s) + (-1)^\nu} 
-        \brk{ \etwm_n(s) - \etwm_m(s) - \ombm^\nu_m(s) }
-     \big), \\
-  &\text{where} \quad 
-  \alb^\nu(s) = (\MMd^\nu - \I) \thb(s), \qquad
-  \omb^\nu(s) = (\MMd^\nu - \I) \etw(s).
-\end{aligned}
-\end{equation}
-%
-
-The expressions in \cref{eq:finitederivs} have the same form as for \(s=0\), which allowed us to argue that the model with maximal area must have the serial topology, \cite{Lahiri2013synapse}.
-The one missing ingredient is the scale invariance.
-
-However, we do have numerical evidence that the model that maximises the Laplace transform does have the serial topology.
-In \cref{fig:envnum} we see the result of numerical maximisation of \cref{eq:finitelagrangian}.
-We see that restricting to the serial topology does not make the performance any worse.
-In fact, it slightly improves the performance, but this can be ascribed to the numerical optimisation problem being easier.
-Assuming that the optimiser did find the global maximum, this provides evidence for our conjecture that the model that maximises the Laplace transform must have the serial topology.
-
-Note that models with the serial topology must satisfy detailed balance.
-Therefore, this conjecture implies that the square-root bound of \cref{sec:sqrt} applies, implying the stricter envelope of \cref{eq:envDB}.
-In \cref{fig:envnum}, we can see that this envelope is close to tight, being off by a constant factor.
-
-\begin{figure}[tb]
-  \centering
-  % Requires \usepackage{graphicx}
-  \includegraphics[width=0.8\linewidth]{LenvNum.pdf}\\
-  \caption[Numerical envelope for the signal-to-noise ratio]
-  {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
-  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numericallly with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
-  The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
-  The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
-  The dotted blue line indicates the proven envelope, \cref{eq:env} and the solid blue line is the conjectured envelope assuming detailed balance, see \cref{eq:envDB}.}\label{fig:envnum}
-\end{figure}
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Scale invariance}\label{sec:scale}
-
-Under the transformation from \cref{eq:scale}, the Laplace transform has the following scaling property
-%
-\begin{equation}\label{eq:laplacescale}
-  \hat{A} \prn{\zeta s;\zeta * \MMd^\nu} = \hat{A} \prn{s; \MMd^\nu}.
-\end{equation}
-%
-At \(s = 0\), it is scale invariant.
-This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
-This allowed us to ignore the diagonal constraints.
-We can't do that for \(s > 0\).
-
-The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
-%
-\begin{equation}\label{eq:shiftscale}
-  \zeta {\circledast}_{s} \MMd 
-      \equiv \zeta \MMd
-      + (1 - \zeta) \brk{(1 + s)\I - s \onev \eq} .
-\end{equation}
-%
-Then we have
-%
-\begin{equation}\label{eq:lshiftscale}
-  \hat{A} \prn{s; \MMd^\nu} 
-    = \hat{A} \prn{0, \shift{\MMd}^\nu(s)}
-    = \hat{A} \prn{0, \zeta * \shift{\MMd}^\nu(s)}
-    = \hat{A} \prn{s, \zeta \circledast_s \MMd^\nu}.
-\end{equation}
-%
-The problem here is that the lower bound, \(\MMdm^\nu_{ij} \geq 0\), 
-is not invariant under this transformation, unlike the \(s = 0\) case.
-Using this transformation to fix the upper bound can break the lower bound.
-
-We might ask what conditions are required so that a model is fixable?
-This means that the \(\zeta \circledast_s \) orbit of the model contains some valid models.
-The conditions are
-%
-\begin{equation*}
-\begin{alignedat}{5}
-  [\zeta \circledast \MMd^\nu]_{ij}  
-    &= \zeta \MMdm^\nu_{ij} - (1-\zeta) s \eqm_j 
-    &&\geq 0 &
-  \quad &\implies \quad &
-  \zeta &\geq \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} &
-    &\forall \nu, i, j : i \neq j , \\
-  \sum_{j : i \neq j} [\zeta \circledast \MMd^\nu]_{ij} 
-    &= \zeta \lbm^\nu_i - (1-\zeta) s (1-\eqm_i) 
-    &&\leq 1 &
-  &\implies &
-  \zeta &\leq \frac{1 + s (1-\eqm_i)}{\lbm^\nu_i + s (1-\eqm_i)} &
-    \quad &\forall \nu, i ,
-\end{alignedat}
-\end{equation*}
-%
-where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \MMdm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
-The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
-%
-\begin{equation}\label{eq:fixable}
-  \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} \leq \frac{1 + s (1-\eqm_k)}{\lbm^\mu_k + s (1-\eqm_k)}
-  \quad \implies \quad
-  \frac{\lbm^\mu_k - 1}{1 + s (1-\eqm_k)} \leq \frac{\MMdm^\nu_{ij}}{s \eqm_j} 
-  \quad \forall \mu, \nu, i, j, k : i \neq j.
-\end{equation}
-%
-Note that when \(\MMdm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq 1\), the original diagonal constraint.
-Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Shifted problem}\label{sec:lshift}
-
-The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
-treat the off-diagonal components of \(\shift{\MMd}^\nu \) as the degrees of freedom and maximise the area instead.
-The constraints on \(\shift{\MMdm}^\nu_{ij}\) are different from \cref{eq:constri}:
-%
-\begin{equation}\label{eq:lshiftconstr}
-  \shift{\MMdm}^\nu_{ij} \geq s \eqm_j,
-    \quad \forall \; \nu, i, j : i \neq j, \qquad
-  \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \leq 1 + s (1 - \eqm_i).
-    \quad \forall \; \nu, i.
-\end{equation}
-%
-The diagonal elements are still given by \( \shift{\MMdm}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \), but now they are allowed to be negative so long as \(\shift{\MMdm}^\nu_{ii} \geq - s (1 - \eqm_i)\).
-
-The Lagrangian we should consider is
-%
-\begin{equation}\label{eq:lshiftlagrangian}
-\begin{aligned}
-  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
-        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\MMdm^\nu_{ij} -  s (\eqm_j - \delta_{ij}) }
-        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
-\end{aligned}
-\end{equation}
-%
-where all of the first-passage times are evaluated at \(s = 0\).
-
-In \cref{fig:shifted} we show the result of solving this problem with numerical optimisation.
-We see essentially the same results as the original problem, \cref{eq:finitelagrangian,fig:envnum}, as expected.
-
-The derivatives, for \(m \neq n\) are
-%
-\begin{equation}\label{eq:lshiftderivs}
-\begin{aligned}
-  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
-    \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
-     + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
-     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     - \sum_{\mu ij} \mu^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
-    = 0.
-\end{aligned}
-\end{equation}
-%
-We can absorb the new terms by defining
-%
-\begin{equation*}
-\begin{aligned}
-  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
-  \qquad
-  \shift{\alb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
-  \\
-  \shift{\omb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \etw(0).
-\end{aligned}
-\end{equation*}
-%
-We note that
-%
-\begin{equation*}
-  \mu^\nu_{ij} (\MMdm^\nu_{ij} - s \eqm_j) = 0,
-  \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} (\MMdm^\nu_{ij} - s \eqm_j)
-      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
-\end{equation*}
-%
-then we can ``solve'' for the KKT multipliers:
-%
-\begin{equation}\label{eq:lshiftsol}
-\begin{aligned}
-  \mu^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
-    + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
-    }, \\
-  \mu^\nu_{mn} &= \eqm_m \prn{ 
-      \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
-  % \\ & \phantom{= - \eqm_m \big( } 
-    + \brk{ {c}_m + (-1)^\nu + \lambda } 
-        \brk{ \etwm_n - \etwm_m - \shift{\ombm}^\nu_m(s) }
-     }, \\
-\end{aligned}
-\end{equation}
-%
-although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\).
-
-While this problem involves the scale invariant quantity \(\hat{A}(0)\), neither the upper nor lower bounds in \cref{eq:lshiftconstr} are invariant.
-We then run into the same problems in proving our serial-topology hypothesis.
-
-\begin{figure}[ht]
-\begin{center}
-\begin{myenuma}
-  \item \aligntop{\includegraphics[width=0.4\linewidth]{LenvShift.pdf}}
-  \label{fig:envshift}
-  \item \aligntop{\includegraphics[width=0.4\linewidth]{shift_cond.pdf}}
-  \label{fig:shiftcond}
-\end{myenuma}
-\caption[Shifted problem vs.\ original]{Shifted problem vs.\ original.
-(\ref*{fig:envshift}) Comparison of numerical optimisation of the shifted problem in \cref{eq:lshiftlagrangian,fig:envnum} with the original problemin \cref{eq:finitelagrangian}.
-We see that the two approaches produce very similar numerical envelopes.
-The shifted problem involves non-linear constraints and is thus harder to solve.
-(\ref{fig:shiftcond}) Plot of the discrepancy in (\ref*{fig:envshift}) against the condition number of the fundamental matrix, \(\fund \).
-We see that most of the large discrepancies occur where the problem is numerically ill-conditioned.
-\label{fig:shifted}}
-\end{center}
-\end{figure}
-
-\begin{figure}[tbp]
-  \centering
-  % Requires \usepackage{graphicx}
-  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=7]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=13]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=21]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=27]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=36]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=50]{envelope.pdf}}
-  \end{myenuma}
-  \caption[Optimal models]{The models that constitute the numeric envelope in \autoref{fig:envnum}.
-  The plot to the left of each frame indicates the memory curve of the model shown to the right in green, with the numeric envelope for models with the serial topology in orange and the proven envelope, \eqref{eq:env}, in blue.
-  Above the left plot we have a graphical display of the model, described with more detail to the right.
-  The area of each node is proportional to the steady state probability in that state and the thickness of the red and blue edges are proportional to the transition probabilities under potentiation and depression.
-  The bar graph in the upper right of each frame indicates the transition probabilities from the state to the left of each bar to the state on its right, under potentiation.
-  The bar graph in the lower right of each frame indicates the transition probabilities from the state to the right of each bar to the state on its left, under depression.
-  The heat map in the centre right of each frame indicates the equilibrium probability distribution for each state.
-  See the video for the full set of models.}\label{fig:envvid}
-\end{figure}
-
-%================================================================================
-
 \section{Nearly uniform serial models}\label{sec:serial}
@@
 
-Looking at them models that were found by numerical maximisation in \autoref{fig:envvid} (see the video for the full set of models), suggests we look at the particular set of models, shown in \autoref{fig:heuristicmodels}.
+Looking at them models that were found by numerical maximisation in \cref{fig:envvid} (see the video for the full set of models), suggests we look at the particular set of models, shown in \cref{fig:heuristicmodels}.
 
@@
 At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
-Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from \autoref{fig:envvid} and the video, this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
+Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from \cref{fig:envvid} and the video, this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
 Once the length has been reduced to \(M=2\), the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.
@@
 
-Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_uniform}.
+Here we will look at the model in \cref{fig:heuristicmodels}\ref{fig:serial_uniform}.
 Looking at \eqref{eq:larea}, the only thing we need is differences in \(\etwm_i(s)\).
@@
 
-Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_shorten}.
-We proceed in the same manner as in \autoref{sec:serial_uniform}, with the following differences.
+Here we will look at the model in \cref{fig:heuristicmodels}\ref{fig:serial_shorten}.
+We proceed in the same manner as in \cref{sec:serial_uniform}, with the following differences.
 
@@
 In fact there are overlaps, as seen in \cref{fig:shorten}\ref{fig:short_ratio} for \(M=6\) and higher.
-This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
+This is related to the fact that \cref{fig:envvid} and the video shows that the optimal models do not shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
 
@@
 
-Here we will look at the model in \autoref{fig:heuristicmodels}\ref{fig:serial_sticky}.
-We proceed in the same manner as in \autoref{sec:serial_uniform}, with the following differences.
+Here we will look at the model in \cref{fig:heuristicmodels}\ref{fig:serial_sticky}.
+We proceed in the same manner as in \cref{sec:serial_uniform}, with the following differences.
 
@@
 
-We can obtain a good approximation to the envelope by combining the results \autoref{sec:serial_uniform} and \autoref{sec:serial_sticky}, ignoring the complications of \autoref{sec:serial_shorten}.
+We can obtain a good approximation to the envelope by combining the results \cref{sec:serial_uniform} and \cref{sec:serial_sticky}, ignoring the complications of \cref{sec:serial_shorten}.
 We will also ignore the small gap between the regions of validity, \eqref{eq:univalid} and \eqref{eq:stickyvalid}.
@@
 %
-This is plotted in \autoref{fig:heuristicenv}.
+This is plotted in \cref{fig:heuristicenv}.
 

### 2021-07-30 Notes/laplace.tex
figure of triangular perturbations
spacing fig

@@
 
+\subsubsection{Triangularity}\label{sec:parttriangular}
+
+Suppose we have two states, \(i\) and \(j\), with \( \etwm_i > \etwm_j \).
+Consider the following perturbation (see also \cref{fig:triangle_s}(\ref{fig:triangle_pert})):
+%
+\begin{equation}\label{eq:pertKupdn}
+\begin{aligned}
+  2 \brk{\pdiff{\CL}{\encm_{mj}} - \pdiff{\CL}{\encm_{mi}}}
+  &= \frac{1}{f\pot} \brk{\pdiff{\CL}{\MMdm\pot_{mj}} - \pdiff{\CL}{\MMdm\pot_{mi}}} -
+  \frac{1}{f\dep} \brk{\pdiff{\CL}{\MMdm\dep_{mj}} - \pdiff{\CL}{\MMdm\dep_{mi}}} \\ 
+    &= 2 \eqm_m (\etwm_i - \etwm_j) 
+    + \mu\pot_{mj} - \mu\pot_{mi} - \mu\dep_{mj} + \mu\dep_{mi}.
+\end{aligned}
+\end{equation}
+%
+Because \( \etwm_i > \etwm_j \), we must have \( \mu\pot_{mi} + \mu\dep_{mj} > 0 \).
+So \( \MMdm\pot_{mi} > 0 \) would imply that \( \MMdm\dep_{mj} = 0 \) and vice versa.
+These two possibilities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_left},\ref{fig:triangle_right}).
+
+Let us arrange the states left to right in order of decreasing \(\etw\), so that \(i\) is to the left of \(j\).
+For the unconstrained transitions \( \MMdm\pot_{mj} \) and \( \MMdm\dep_{mi} \) the potentiating transition goes to the \emph{right} of the depressing transition.
+For the other transitions \( \MMdm\pot_{mi} \) and \( \MMdm\dep_{mj} \) the potentiating transition goes to the \emph{left} of the depressing transition, and they cannot both be nonzero.
+
+This logic can be applied to any pair of states with different values of \(\etw\).
+Therefore, when starting from the same state, no potentiating transition can go to the left of any depressing transition, and no depressing transition can go to the right of any potentiating transition.
+
+For any state \(m\), there is a state \(n\) such that all potentiating transitions go to equal or smaller values of \(\etw\) all depressing transitions go to equal or greater values.
+%
+\begin{equation}\label{eq:separator}
+  \forall\ m, \;\; \exists\ n : 
+  \begin{cases}
+    \etwm_i > \etwm_n &\implies \MMdm\pot_{mi} = 0, \\
+    \etwm_i < \etwm_n &\implies \MMdm\dep_{mi} = 0.
+  \end{cases}
+\end{equation}
+%
+We have not yet ruled the possibility that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero.
+The remaining allowed transitions from \(m\) are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_allowed}).
+If one of those transitions is zero, there will be two states with the property above: the targets of right-most depressing and left-most potentiating transitions.
+
+Consider the perturbation
+%
+\begin{equation}\label{eq:pertK}
+  \frac{1}{f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} -
+  \frac{1}{f\dep} \pdiff{\CL}{\MMdm\dep_{mn}} =
+    2 \eqm_m (\etwm_m - \etwm_n) 
+    + \mu\pot_{mn} - \mu\pot_{mm} - \mu\dep_{mn} + \mu\dep_{mm}.
+\end{equation}
+%
+If we assume that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero, we find that \( \mu\pot_{mm} - \mu\dep_{mm} = 2 \eqm_m (\etwm_m - \etwm_n) \).
+Therefore \( \etwm_m > \etwm_n \implies \MMdm\pot_{mm} = 0 \) and \( \etwm_m < \etwm_n \implies \MMdm\dep_{mm} = 0 \).
+It is also possible that both of the diagonal elements are zero.
+
+\begin{figure}[ht]
+\begin{center}
+\begin{myenuma}
+  \item \label{fig:triangle_pert} \aligntop{%
+  \includegraphics[width=0.28\linewidth]{pert_triangle.svg}}
+  \item \label{fig:triangle_left} \aligntop{%
+  \includegraphics[width=0.28\linewidth]{left_triangle.svg}}
+  \item \label{fig:triangle_right} \aligntop{%
+  \includegraphics[width=0.28\linewidth]{right_triangle.svg}}
+  \\[1em]
+  \item \label{fig:triangle_allowed} \aligntop{%
+  \includegraphics[width=0.4\linewidth]{allowed_triangle.svg}}
+  \item \label{fig:triangle_spacing} \aligntop{%
+  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
+\end{myenuma}
+\caption[Triangularising perturbation]{Triangularising perturbation.
+(\ref*{fig:triangle_pert}) The type of perturbation considered in \cref{eq:pertKupdn}.
+(\ref*{fig:triangle_left},\ref*{fig:triangle_right}) Possible end-points of this perturbation. 
+(\ref*{fig:triangle_allowed}) The not-as-yet ruled out transitions when states are sorted so that \(\etw\) decreases to the right, with \(\etwm_m\) unconstrained (as yet).
+(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, with \(i\) a state with a nonzero depressing transition from \(m\), \(j\) a state with a nonzero potentiating transition, \(n\) a state with both depressing and potentiating transitions and \(k\) a state with no nonzero transitions from \(m\).
+The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
+\label{fig:triangle_s}}
+\end{center}
+\end{figure}
+
+%================================================================================
+
 \subsubsection{Number of exits}\label{sec:exit}
@@
 
-Suppose there are two states, \(i,j\), with nonzer transitions from \(m\) for one \(\nu\).
+Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
 Both must statisfy \cref{eq:oneexit} for the same \(\mu^\nu_{mm}\), so
@@
 %
+The consequences of these inequalities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}).
+We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
+Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.
+
 Both \cref{eq:twoexits,eq:twodoubleexits} require degeneracy.
@@
 %-------Section------------------------------------------------------------------
-
-\subsubsection{Triangularity}\label{sec:parttriangular}
-
-Suppose we have two states, \(i\) and \(j\), with \( \etwm_i > \etwm_j \).
-Consider the following perturbation:
-%
-\begin{equation}\label{eq:pertKupdn}
-\begin{aligned}
-  2 \brk{\pdiff{\CL}{\encm_{mj}} - \pdiff{\CL}{\encm_{mi}}}
-  &= \frac{1}{f\pot} \brk{\pdiff{\CL}{\MMdm\pot_{mj}} - \pdiff{\CL}{\MMdm\pot_{mi}}} -
-  \frac{1}{f\dep} \brk{\pdiff{\CL}{\MMdm\dep_{mj}} - \pdiff{\CL}{\MMdm\dep_{mi}}} \\ 
-    &= 2 \eqm_m (\etwm_i - \etwm_j) 
-    + \mu\pot_{mj} - \mu\pot_{mi} - \mu\dep_{mj} + \mu\dep_{mi}.
-\end{aligned}
-\end{equation}
-%
-Because \( \etwm_i > \etwm_j \), we must have \( \mu\pot_{mi} + \mu\dep_{mj} > 0 \).
-So \( \MMdm\pot_{mi} > 0 \) would imply that \( \MMdm\dep_{mj} = 0 \) and vice versa.
-Let us arrange the states left to right in order of decreasing \(\etw\), so that \(i\) is to the left of \(j\).
-For the unconstrained transitions \( \MMdm\pot_{mj} \) and \( \MMdm\dep_{mi} \) the potentiating transition goes to the \emph{right} of the depressing transition.
-For the other transitions \( \MMdm\pot_{mi} \) and \( \MMdm\dep_{mj} \) the potentiating transition goes to the \emph{left} of the depressing transition, and they cannot both be nonzero.
-
-This logic can be applied to any pair of states with different values of \(\etw\).
-Therefore, when starting from the same state, no potentiating transition can go to the left of any depressing transition, and no depressing transition can go to the right of any potentiating transition.
-
-For any state \(m\), there is a state \(n\) such that all potentiating transitions go to equal or smaller values of \(\etw\) all depressing transitions go to equal or greater values.
-%
-\begin{equation}\label{eq:separator}
-  \forall\ m, \;\; \exists\ n : 
-  \begin{cases}
-    \etwm_i > \etwm_n &\implies \MMdm\pot_{mi} = 0, \\
-    \etwm_i < \etwm_n &\implies \MMdm\dep_{mi} = 0.
-  \end{cases}
-\end{equation}
-%
-We have not yet ruled the possibility that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero.
-
-Consider the perturbation
-%
-\begin{equation}\label{eq:pertK}
-  \frac{1}{f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} -
-  \frac{1}{f\dep} \pdiff{\CL}{\MMdm\dep_{mn}} =
-    2 \eqm_m (\etwm_m - \etwm_n) 
-    + \mu\pot_{mn} - \mu\pot_{mm} - \mu\dep_{mn} + \mu\dep_{mm}.
-\end{equation}
-%
-If we assume that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero, we find that \( \mu\pot_{mm} - \mu\dep_{mm} = 2 \eqm_m (\etwm_m - \etwm_n) \).
-Therefore \( \etwm_m > \etwm_n \implies \MMdm\pot_{mm} = 0 \) and \( \etwm_m < \etwm_n \implies \MMdm\dep_{mm} = 0 \).
-It is also possible that both of the diagonal elements are zero.
-
-%================================================================================
 

### 2021-07-31 Notes/laplace.tex
fixed labels in spacing fig

@@
 \begin{gathered}
-  \alb^\nu(s) = \frac{\delta \mathbf{a}(s) + (-1)^\nu \enc \thb(s)}{2f^\nu},
+  \alb^\nu(s) = \frac{\delta \mathbf{a}(s) + (-1)^\nu \enc \thb(s)}
+                     {2f^\nu},
   \qquad 
-  \omb^\nu(s) = \frac{\delta \w + s \delta \etw(s) - (-1)^\nu \mathbf{a}(s)}{2 f^\nu}, \\
-\begin{aligned}
-  \mu^\nu_{mm} &= - \frac{\eqm_m}{2f^\nu} 
+  \omb^\nu(s) = \frac{\delta \w + s \delta \etw(s) - (-1)^\nu \mathbf{a}(s)}
+                     {2 f^\nu}, \\
+\begin{aligned}
+  \mu^\nu_{mm} &= \frac{\eqm_m}{2f^\nu} 
     \left[ 
-      c_m(s) (\delta \wm_m + s \delta \etwm_m(s)) - \hat{A}(s)
-      + \lambda \delta w_m
-      \right. \\ &\phantom{= - \frac{\eqm_m}{2f^\nu} []}\left.
+      \hat{A}(s) - c_m(s) (\delta \wm_m + s \delta \etwm_m(s))
+      - \lambda \delta w_m
+    \right. \\ &\phantom{= \frac{\eqm_m}{2f^\nu} }
+    \left.
       + (-1)^\nu \prn{ 
-        [\encm \thbm(s)]_m - c_m(s) a_m(s)
-        + \delta \wm_m + s \delta \etwm_m(s)
-        - \lambda a_m(0)
+        c_m(s) a_m(s) - [\encm \thbm(s)]_m
+        - \delta \wm_m - s \delta \etwm_m(s)
+        + \lambda a_m(0)
       }
@@
   \zeta {\circledast}_{s} \MMd 
-      \equiv \zeta \MMd
-      + (1 - \zeta) \brk{(1 + s)\I - s \onev \eq} .
+      \equiv (1 - \zeta) \MMd
+      + \zeta \brk{(1 + s)\I - s \onev \eq} .
 \end{equation}
@@
 %
-We have not yet ruled the possibility that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero.
 The remaining allowed transitions from \(m\) are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_allowed}).
+Note that we have not yet ruled the possibility that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero.
 If one of those transitions is zero, there will be two states with the property above: the targets of right-most depressing and left-most potentiating transitions.
@@
 \caption[Triangularising perturbation]{Triangularising perturbation.
-(\ref*{fig:triangle_pert}) The type of perturbation considered in \cref{eq:pertKupdn}.
+(\ref*{fig:triangle_pert}) The type of perturbation considered in \cref{eq:pertKupdn}, indicating which transitions have their rates increased/decreased.
 (\ref*{fig:triangle_left},\ref*{fig:triangle_right}) Possible end-points of this perturbation. 
@@
 They should be thought of as collections of vectors rather than matrices.
+We can also write \cref{eq:diagKKT} as
+%
+\begin{equation}\label{eq:diagKKTz}
+\begin{aligned}
+  \mu^\nu_{mm} &= \frac{\eqm_m}{2f^\nu} 
+    \left[ 
+      \hat{A}(s) - c_m(s) (\delta \wm_m + s \delta \etwm_m(s))
+      - \lambda \delta w_m
+    \right. \\ &\phantom{= \frac{\eqm_m}{2f^\nu} }
+    \left.
+      - (-1)^\nu \prn{ 
+        [\encm \dgnm(s)]_m
+        + \delta \wm_m + s \delta \etwm_m(s)
+        - \lambda a_m(0)
+      }
+    \right], \\
+\end{aligned}
+\end{equation}
+%
 
@@
 
-Both \cref{eq:twoexits,eq:twodoubleexits} require degeneracy.
+Both of \cref{eq:twoexits,eq:twodoubleexits} require degeneracy.
 The physicist's intuition states that degeneracy is fragile and requires a symmetry to protect it.
@@
 This would imply that, after lumping all lumpable states together, the situation in \cref{eq:twoexits} is impossible and \cref{eq:twodoubleexits} is even more impossible.
+If this intuition holds, we expect that \cref{fig:triangle_s}(\ref{fig:triangle_spacing}) would have at most one state on each of the two diagonal lines.
+Either there would be one state like \(n\), or there would be one state like \(i\) and one state like \(j\).
 
@@
 \begin{equation}\label{eq:scale}
-  \zeta * \MMd \equiv \zeta \MMd + (1-\zeta)\I .
+  \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I .
 \end{equation}

### 2021-07-31 Notes/laplace.tex
rearranged figs

@@
     + \sum_{\alpha\neq\beta} \av{\synid^\alpha\syn^\alpha(t)}\av{\smash{\synid^\beta}\smash{\syn^\beta}(t)}
-    %\\&\phantom{= \sum_{\alpha}  \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}}
+    %\\&\hphantom{= \sum_{\alpha}  \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}}
     - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2  \\
@@
     + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
-  \\ &\phantom{= \eqm_m \big( }
+  \\ &\hphantom{= \eqm_m \big( }
     + \brk{{c}_m(s) + (-1)^\nu} 
@@
      \big), \\
-  &\text{where} \quad 
-  \alb^\nu(s) = (\MMd^\nu - \I) \thb(s), \qquad
+  \text{where} \quad 
+  \alb^\nu(s) &= (\MMd^\nu - \I) \thb(s), \qquad
   \omb^\nu(s) = (\MMd^\nu - \I) \etw(s).
@@
 \begin{equation}\label{eq:diagKKT}
-\begin{gathered}
-  \alb^\nu(s) = \frac{\delta \mathbf{a}(s) + (-1)^\nu \enc \thb(s)}
+\begin{aligned}
+  \alb^\nu(s) &= \frac{\delta\mathbf{a}(s) + (-1)^\nu \enc \thb(s)}
                      {2f^\nu},
   \qquad 
-  \omb^\nu(s) = \frac{\delta \w + s \delta \etw(s) - (-1)^\nu \mathbf{a}(s)}
-                     {2 f^\nu}, \\
-\begin{aligned}
+  \omb^\nu(s) = \frac{\delta\w + s \delta\etw(s) - (-1)^\nu \mathbf{a}(s)}
+                     {2 f^\nu}, 
+  \\
   \mu^\nu_{mm} &= \frac{\eqm_m}{2f^\nu} 
     \left[ 
-      \hat{A}(s) - c_m(s) (\delta \wm_m + s \delta \etwm_m(s))
+      \hat{A}(s) - c_m(s) (\delta\wm_m + s \delta\etwm_m(s))
       - \lambda \delta w_m
-    \right. \\ &\phantom{= \frac{\eqm_m}{2f^\nu} }
-    \left.
+    \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} \big[}
+    \left. {}
       + (-1)^\nu \prn{ 
         c_m(s) a_m(s) - [\encm \thbm(s)]_m
-        - \delta \wm_m - s \delta \etwm_m(s)
+        - \delta\wm_m - s \delta\etwm_m(s)
         + \lambda a_m(0)
       }
-    \right], \\
-\end{aligned}
-\end{gathered}
+    \right], 
+\end{aligned}
 \end{equation}
@@
 Assuming that the optimiser did find the global maximum, this provides evidence for our conjecture that the model that maximises the Laplace transform must have the serial topology.
-
-Note that models with the serial topology must satisfy detailed balance.
-Therefore, this conjecture implies that the square-root bound of \cref{sec:sqrt} applies, implying the stricter envelope of \cref{eq:envDB}.
-In \cref{fig:envnum}, we can see that this envelope is close to tight, being off by a constant factor.
+The models found numerically are described in \cref{fig:envvid}.
 
@@
   % Requires \usepackage{graphicx}
-  \includegraphics[width=0.8\linewidth]{LenvNum.pdf}\\
+  \includegraphics[width=0.8\linewidth]{LenvNumOnly.pdf}\\
   \caption[Numerical envelope for the signal-to-noise ratio]
   {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
-  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numericallly with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
+  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numerically with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
   The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
   The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
-  The dotted blue line indicates the proven envelope, \cref{eq:env} and the solid blue line is the conjectured envelope assuming detailed balance, see \cref{eq:envDB}.}\label{fig:envnum}
-\end{figure}
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Scale invariance}\label{sec:scale}
-
-Under the transformation from \cref{eq:scale}, the Laplace transform has the following scaling property
-%
-\begin{equation}\label{eq:laplacescale}
-  \hat{A} \prn{\zeta s;\zeta * \MMd^\nu} = \hat{A} \prn{s; \MMd^\nu}.
-\end{equation}
-%
-At \(s = 0\), it is scale invariant.
-This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
-This allowed us to ignore the diagonal constraints.
-We can't do that for \(s > 0\).
-
-The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
-%
-\begin{equation}\label{eq:shiftscale}
-  \zeta {\circledast}_{s} \MMd 
-      \equiv (1 - \zeta) \MMd
-      + \zeta \brk{(1 + s)\I - s \onev \eq} .
-\end{equation}
-%
-Then we have
-%
-\begin{equation}\label{eq:lshiftscale}
-  \hat{A} \prn{s; \MMd^\nu} 
-    = \hat{A} \prn{0, \shift{\MMd}^\nu(s)}
-    = \hat{A} \prn{0, \zeta * \shift{\MMd}^\nu(s)}
-    = \hat{A} \prn{s, \zeta \circledast_s \MMd^\nu}.
-\end{equation}
-%
-The problem here is that the lower bound, \(\MMdm^\nu_{ij} \geq 0\), 
-is not invariant under this transformation, unlike the \(s = 0\) case.
-Using this transformation to fix the upper bound can break the lower bound.
-
-We might ask what conditions are required so that a model is fixable?
-This means that the \(\zeta \circledast_s \) orbit of the model contains some valid models.
-The conditions are
-%
-\begin{equation*}
-\begin{alignedat}{5}
-  [\zeta \circledast \MMd^\nu]_{ij}  
-    &= \zeta \MMdm^\nu_{ij} - (1-\zeta) s \eqm_j 
-    &&\geq 0 &
-  \quad &\implies \quad &
-  \zeta &\geq \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} &
-    &\forall \nu, i, j : i \neq j , \\
-  \sum_{j : i \neq j} [\zeta \circledast \MMd^\nu]_{ij} 
-    &= \zeta \lbm^\nu_i - (1-\zeta) s (1-\eqm_i) 
-    &&\leq 1 &
-  &\implies &
-  \zeta &\leq \frac{1 + s (1-\eqm_i)}{\lbm^\nu_i + s (1-\eqm_i)} &
-    \quad &\forall \nu, i ,
-\end{alignedat}
-\end{equation*}
-%
-where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \MMdm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
-The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
-%
-\begin{equation}\label{eq:fixable}
-  \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} \leq \frac{1 + s (1-\eqm_k)}{\lbm^\mu_k + s (1-\eqm_k)}
-  \quad \implies \quad
-  \frac{\lbm^\mu_k - 1}{1 + s (1-\eqm_k)} \leq \frac{\MMdm^\nu_{ij}}{s \eqm_j} 
-  \quad \forall \mu, \nu, i, j, k : i \neq j.
-\end{equation}
-%
-Note that when \(\MMdm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq 1\), the original diagonal constraint.
-Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Shifted problem}\label{sec:lshift}
-
-The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
-treat the off-diagonal components of \(\shift{\MMd}^\nu \) as the degrees of freedom and maximise the area instead.
-The constraints on \(\shift{\MMdm}^\nu_{ij}\) are different from \cref{eq:constri}:
-%
-\begin{equation}\label{eq:lshiftconstr}
-  \shift{\MMdm}^\nu_{ij} \geq s \eqm_j,
-    \quad \forall \; \nu, i, j : i \neq j, \qquad
-  \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \leq 1 + s (1 - \eqm_i).
-    \quad \forall \; \nu, i.
-\end{equation}
-%
-The diagonal elements are still given by \( \shift{\MMdm}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \), but now they are allowed to be negative so long as \(\shift{\MMdm}^\nu_{ii} \geq - s (1 - \eqm_i)\).
-
-The Lagrangian we should consider is
-%
-\begin{equation}\label{eq:lshiftlagrangian}
-\begin{aligned}
-  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
-        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\MMdm^\nu_{ij} -  s (\eqm_j - \delta_{ij}) }
-        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
-\end{aligned}
-\end{equation}
-%
-where all of the first-passage times are evaluated at \(s = 0\).
-
-In \cref{fig:shifted} we show the result of solving this problem with numerical optimisation.
-We see essentially the same results as the original problem, \cref{eq:finitelagrangian,fig:envnum}, as expected.
-
-The derivatives, for \(m \neq n\) are
-%
-\begin{equation}\label{eq:lshiftderivs}
-\begin{aligned}
-  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
-    \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
-     + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
-     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     - \sum_{\mu ij} \mu^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
-    = 0.
-\end{aligned}
-\end{equation}
-%
-We can absorb the new terms by defining
-%
-\begin{equation*}
-\begin{aligned}
-  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
-  \qquad
-  \shift{\alb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
-  \\
-  \shift{\omb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \etw(0).
-\end{aligned}
-\end{equation*}
-%
-We note that
-%
-\begin{equation*}
-  \mu^\nu_{ij} (\MMdm^\nu_{ij} - s \eqm_j) = 0,
-  \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} (\MMdm^\nu_{ij} - s \eqm_j)
-      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
-\end{equation*}
-%
-then we can ``solve'' for the KKT multipliers:
-%
-\begin{equation}\label{eq:lshiftsol}
-\begin{aligned}
-  \mu^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
-    + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
-    }, \\
-  \mu^\nu_{mn} &= \eqm_m \prn{ 
-      \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
-  % \\ & \phantom{= - \eqm_m \big( } 
-    + \brk{ {c}_m + (-1)^\nu + \lambda } 
-        \brk{ \etwm_n - \etwm_m - \shift{\ombm}^\nu_m(s) }
-     }, \\
-\end{aligned}
-\end{equation}
-%
-although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\).
-
-While this problem involves the scale invariant quantity \(\hat{A}(0)\), neither the upper nor lower bounds in \cref{eq:lshiftconstr} are invariant.
-We then run into the same problems in proving our serial-topology hypothesis.
-
-\begin{figure}[ht]
-\begin{center}
-\begin{myenuma}
-  \item \aligntop{\includegraphics[width=0.4\linewidth]{LenvShift.pdf}}
-  \label{fig:envshift}
-  \item \aligntop{\includegraphics[width=0.4\linewidth]{shift_cond.pdf}}
-  \label{fig:shiftcond}
-\end{myenuma}
-\caption[Shifted problem vs.\ original]{Shifted problem vs.\ original.
-(\ref*{fig:envshift}) Comparison of numerical optimisation of the shifted problem in \cref{eq:lshiftlagrangian,fig:envnum} with the original problemin \cref{eq:finitelagrangian}.
-We see that the two approaches produce very similar numerical envelopes.
-The shifted problem involves non-linear constraints and is thus harder to solve.
-(\ref{fig:shiftcond}) Plot of the discrepancy in (\ref*{fig:envshift}) against the condition number of the fundamental matrix, \(\fund \).
-We see that most of the large discrepancies occur where the problem is numerically ill-conditioned.
-\label{fig:shifted}}
-\end{center}
+  }\label{fig:envnum}
 \end{figure}
@@
   See the video for the full set of models.}\label{fig:envvid}
+\end{figure}
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Scale invariance}\label{sec:scale}
+
+Under the transformation from \cref{eq:scale}, the Laplace transform has the following scaling property
+%
+\begin{equation}\label{eq:laplacescale}
+  \hat{A} \prn{\zeta s;\zeta * \MMd^\nu} = \hat{A} \prn{s; \MMd^\nu}.
+\end{equation}
+%
+At \(s = 0\), it is scale invariant.
+This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
+This allowed us to ignore the diagonal constraints.
+We can't do that for \(s > 0\).
+
+The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
+%
+\begin{equation}\label{eq:shiftscale}
+  \zeta {\circledast}_{s} \MMd 
+      \equiv (1 - \zeta) \MMd
+      + \zeta \brk{(1 + s)\I - s \onev \eq} .
+\end{equation}
+%
+Then we have
+%
+\begin{equation}\label{eq:lshiftscale}
+  \hat{A} \prn{s; \MMd^\nu} 
+    = \hat{A} \prn{0, \shift{\MMd}^\nu(s)}
+    = \hat{A} \prn{0, \zeta * \shift{\MMd}^\nu(s)}
+    = \hat{A} \prn{s, \zeta \circledast_s \MMd^\nu}.
+\end{equation}
+%
+The problem here is that the lower bound, \(\MMdm^\nu_{ij} \geq 0\), 
+is not invariant under this transformation, unlike the \(s = 0\) case.
+Using this transformation to fix the upper bound can break the lower bound.
+
+We might ask what conditions are required so that a model is fixable?
+This means that the \(\zeta \circledast_s \) orbit of the model contains some valid models.
+The conditions are
+%
+\begin{equation*}
+\begin{alignedat}{5}
+  [\zeta \circledast \MMd^\nu]_{ij}  
+    &= \zeta \MMdm^\nu_{ij} - (1-\zeta) s \eqm_j 
+    &&\geq 0 &
+  \quad &\implies \quad &
+  \zeta &\geq \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} &
+    &\forall \nu, i, j : i \neq j , \\
+  \sum_{j : i \neq j} [\zeta \circledast \MMd^\nu]_{ij} 
+    &= \zeta \lbm^\nu_i - (1-\zeta) s (1-\eqm_i) 
+    &&\leq 1 &
+  &\implies &
+  \zeta &\leq \frac{1 + s (1-\eqm_i)}{\lbm^\nu_i + s (1-\eqm_i)} &
+    \quad &\forall \nu, i ,
+\end{alignedat}
+\end{equation*}
+%
+where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \MMdm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
+The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
+%
+\begin{equation}\label{eq:fixable}
+  \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} \leq \frac{1 + s (1-\eqm_k)}{\lbm^\mu_k + s (1-\eqm_k)}
+  \quad \implies \quad
+  \frac{\lbm^\mu_k - 1}{1 + s (1-\eqm_k)} \leq \frac{\MMdm^\nu_{ij}}{s \eqm_j} 
+  \quad \forall \mu, \nu, i, j, k : i \neq j.
+\end{equation}
+%
+Note that when \(\MMdm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq 1\), the original diagonal constraint.
+Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Shifted problem}\label{sec:lshift}
+
+The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
+treat the off-diagonal components of \(\shift{\MMd}^\nu \) as the degrees of freedom and maximise the area instead.
+The constraints on \(\shift{\MMdm}^\nu_{ij}\) are different from \cref{eq:constri}:
+%
+\begin{equation}\label{eq:lshiftconstr}
+  \shift{\MMdm}^\nu_{ij} \geq s \eqm_j,
+    \quad \forall \; \nu, i, j : i \neq j, \qquad
+  \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \leq 1 + s (1 - \eqm_i).
+    \quad \forall \; \nu, i.
+\end{equation}
+%
+The diagonal elements are still given by \( \shift{\MMdm}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \), but now they are allowed to be negative so long as \(\shift{\MMdm}^\nu_{ii} \geq - s (1 - \eqm_i)\).
+
+The Lagrangian we should consider is
+%
+\begin{equation}\label{eq:lshiftlagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
+        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\MMdm^\nu_{ij} -  s (\eqm_j - \delta_{ij}) }
+        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
+\end{aligned}
+\end{equation}
+%
+where all of the first-passage times are evaluated at \(s = 0\).
+
+In \cref{fig:shifted} we show the result of solving this problem with numerical optimisation.
+We see essentially the same results as the original problem, \cref{eq:finitelagrangian,fig:envnum}, as expected.
+
+The derivatives, for \(m \neq n\) are
+%
+\begin{equation}\label{eq:lshiftderivs}
+\begin{aligned}
+  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+    \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
+     + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
+     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
+     - \sum_{\mu ij} \mu^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
+    = 0.
+\end{aligned}
+\end{equation}
+%
+We can absorb the new terms by defining
+%
+\begin{equation*}
+\begin{aligned}
+  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
+  \qquad
+  \shift{\alb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
+  \\
+  \shift{\omb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \etw(0).
+\end{aligned}
+\end{equation*}
+%
+We note that
+%
+\begin{equation*}
+  \mu^\nu_{ij} (\MMdm^\nu_{ij} - s \eqm_j) = 0,
+  \qquad \text{and} \qquad
+  \sum_{j: j \neq i} \mu^\nu_{ii} (\MMdm^\nu_{ij} - s \eqm_j)
+      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
+\end{equation*}
+%
+then we can ``solve'' for the KKT multipliers:
+%
+\begin{equation}\label{eq:lshiftsol}
+\begin{aligned}
+  \mu^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
+    }, \\
+  \mu^\nu_{mn} &= \eqm_m \prn{ 
+      \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
+  % \\ & \hphantom{= - \eqm_m \big( } 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } 
+        \brk{ \etwm_n - \etwm_m - \shift{\ombm}^\nu_m(s) }
+     }, \\
+\end{aligned}
+\end{equation}
+%
+although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\).
+
+While this problem involves the scale invariant quantity \(\hat{A}(0)\), neither the upper nor lower bounds in \cref{eq:lshiftconstr} are invariant.
+We then run into the same problems in proving our serial-topology hypothesis.
+
+\begin{figure}[ht]
+\begin{center}
+\begin{myenuma}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{LenvShift.pdf}}
+  \label{fig:envshift}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{shift_cond.pdf}}
+  \label{fig:shiftcond}
+\end{myenuma}
+\caption[Shifted problem vs.\ original]{Shifted problem vs.\ original.
+(\ref*{fig:envshift}) Comparison of numerical optimisation of the shifted problem in \cref{eq:lshiftlagrangian,fig:envnum} with the original problemin \cref{eq:finitelagrangian}.
+We see that the two approaches produce very similar numerical envelopes.
+The shifted problem involves non-linear constraints and is thus harder to solve.
+(\ref{fig:shiftcond}) Plot of the discrepancy in (\ref*{fig:envshift}) against the condition number of the fundamental matrix, \(\fund \).
+We see that most of the large discrepancies occur where the problem is numerically ill-conditioned.
+\label{fig:shifted}}
+\end{center}
 \end{figure}
@@
     \left[ 
-      \hat{A}(s) - c_m(s) (\delta \wm_m + s \delta \etwm_m(s))
+      \hat{A}
+      - (c_m + (-1)^\nu) (\delta \wm_m + s \delta \etwm_m)
       - \lambda \delta w_m
-    \right. \\ &\phantom{= \frac{\eqm_m}{2f^\nu} }
-    \left.
+    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
+    % \left.
       - (-1)^\nu \prn{ 
-        [\encm \dgnm(s)]_m
-        + \delta \wm_m + s \delta \etwm_m(s)
+        [\encm \dgnm_m]_m
         - \lambda a_m(0)
       }
-    \right], \\
-\end{aligned}
-\end{equation}
-%
+    \right],
+\end{aligned}
+\end{equation}
+%
+with all quantities evaluated at \(s\) unless labelled otherwise.
 
@@
   \end{myenuma}
-  \caption[Proven envelope for the signal-to-noise ratio]
-  {(\ref{fig:provengen}) The memory curve envelope that we can prove, \cref{eq:env}.
+  \caption[Proven and conjectured envelopes for the signal-to-noise ratio]
+  {Proven and conjectured envelopes for the signal-to-noise ratio.
+  (\ref{fig:provengen}) The memory curve envelope that we can prove, \cref{eq:env}.
   (\ref{fig:provendetbal}) Conjectured memory curve envelope when we assume the optimal models satisfy detailed balance, \cref{eq:envDB}.
@@
 
+Note that models with the serial topology must satisfy detailed balance.
+Therefore, the conjecture from \cref{sec:finite} implies that the square-root bound of \cref{sec:sqrt} applies, implying the stricter envelope of \cref{eq:envDB}.
+In \cref{fig:envnum} we compare these theoretical envelopes with the numerical maximisation from \cref{fig:envnum}.
+We can see that this conjectured envelope is close to tight, being off by a constant factor.
+
+\begin{figure}[tb]
+  \centering
+  % Requires \usepackage{graphicx}
+  \includegraphics[width=0.8\linewidth]{LenvNum.pdf}\\
+  \caption[Numerical and theoretical envelopes for the signal-to-noise ratio]
+  {Numerical  and theoretical envelopes  for the signal-to-noise ratio, \eqref{eq:snrb}. 
+  As in \cref{fig:envnum}, \cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numerically with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
+  The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
+  The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
+  The dotted blue line indicates the proven envelope, \cref{eq:env} and the solid blue line is the conjectured envelope assuming detailed balance, see \cref{eq:envDB}.}\label{fig:envnumth}
+\end{figure}
+
 %================================================================================

### 2021-08-12 Notes/laplace.tex
diagonal constraints

@@
 % \newcommand{\CB}{\mathcal{B}}
-%avoid double superscripts
-\renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
+% %avoid double superscripts
+% \renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
 % shifted process
@@
 First maximise the numerator at fixed \(\eq \w\).
-Then maximise the ratio \wrt \(\eq\w\).
+Then maximise the ratio \wrt \(\Delta \pr = \eq\w\).
 This will allow us to ignore the denominator until the very end.
@@
 \begin{equation}\label{eq:pertfe}
-  \pdiff{}{\frgm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\MMd \dep_{ij}},
+  \pdiff{}{\frgm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
   \qquad
@@
   \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
-  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq f^\mu, \\
+  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq 2f^\mu, \\
   \sum_j \frgm_{ij} &=0, &
@@
 \begin{aligned}
-  \fund(s) \onev &= {\shift{\tau}_0}\onev, \qquad &
+  \fund(s) \onev &= \shift{\tau}_0 \onev, \qquad &
   \shift{\MM}(s) \fund(s) &= \onev \eq - \I, \\ 
   \shift{\arow} \fund(s) &= \eq,  &
-  \fund(s) \shift{\MM}(s) &= \shift{\tau} \onev \shift{\arow} - \I.
+  \fund(s) \shift{\MM}(s) &= \shift{\tau}_0 \onev \shift{\arow} - \I.
 \end{aligned}
@@
 %
-is independent of the starting state \(i\), but now \(\eta(s) = \tr\fund(s) - \frac{\tau_0}{1 + s \tau_0}\).
+is independent of the starting state \(i\), but now \(\eta(s) = \tr\fund(s) - \shift{\tau}_0 \).
 We can then define
@@
 \footnote{We could choose \( \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s g^\nu}{r f^\nu} (\onev \eq - \I) \), for any \(g^\nu \) with \( \sum_\nu g^\nu = 1 \).
+It shifts \(\frg\) correctly and also shifts \( \shift{\enc}(s) = \enc + \frac{s}{r} \Delta g (\onev \eq - \I) \).
 The two possibilities in \cref{eq:lshiftpd} come from \(g^\nu = \frac{1}{2}\) and \(g^\nu = f^\nu\) respectively.}
@@
 The first choice leaves \( \shift{\enc} = \enc \), so it only affects the \(\etw\) piece of \cref{eq:larea}.
-The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + \frac{s}{r} (f\pot - f\dep) (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
+The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + \frac{s}{r} \Delta f (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
 Either way:
@@
   \begin{aligned}
-    {c}_k(s) &= \sum_{ij} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)}, \\
-    {a}_j(s) &= \sum_{k} \encm_{jk} (\etwm_{j}(s) - \etwm_{k}(s)), \\
+    {c}_k(s) &= r\sum_{ij} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)}, \\
+    {a}_j(s) &= r\sum_{k} \encm_{jk} (\etwm_{j}(s) - \etwm_{k}(s)), \\
     \thbm_i(s) &= \sum_j \fptbm_{ij}(0)\, \eqm_j {a}_j(s), \\
@@
   \pdiff{\hat{A}(s)}{\MMdm^\mu_{mn}}
-     = f^\mu \eqm_m \prn{ r \brk{\thbm_m(s) - \thbm_n(s)}
-     +\prn{r {c}_m(s) + (-1)^\mu} \brk{\etwm_m(s) - \etwm_n(s)} }.
+     = f^\mu \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
+     +\prn{{c}_m(s) + (-1)^\mu} \brk{\etwm_m(s) - \etwm_n(s)} }.
 \end{equation}
@@
 To write down the Hessian, we will first define some notation.
-We define the matrix \(\mathbf{X}(s) = \fund(0) \enc \fund(s)\).
+We define the matrix \(\mathbf{X}(s) = r\fund(0) \enc \fund(s)\).
 Given a four-index array \(H_{ijkl}\), we define \(H\trans_{ijkl} = H_{klij}\).
@@
 \begin{equation}\label{eq:hessparts}
-\begin{alignedat}{3}
-    R_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frgm_{ij} \circ \partial \frgm_{kl}}} 
-    &= \
-    &&\eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\thbm_k(s) - \thbm_l(s)} \\
-    &&&+ \eqm_i \brc{{X}_{ik}(s) - {X}_{jk}(s)} 
+\begin{alignedat}{2}
+    R_{ijkl} &= \smash{\pdiff{\hat{A}}{\frgm_{ij} \circ ; \frgm_{kl}}} &
+    &=
+    \eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\thbm_k(s) - \thbm_l(s)} \\
+    &&&\hphantom{=} {}
+    + \eqm_i \brc{{X}_{ik}(s) - {X}_{jk}(s)} 
           \brc{\etwm_k(s) - \etwm_l(s)} \\
-    &&&+ \eqm_i {c}_i(s) \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} 
+    &&& \hphantom{=} {}
+    + \eqm_i {c}_i(s) \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} 
           \brc{\etwm_k(s) - \etwm_l(s)},  \\
-  S_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frgm_{ij} \partial \encm_{kl}} }
-  &= \
-    &&\eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\etwm_k(s) - \etwm_l(s)} \\
-    &&&+ \eqm_k \brc{\fundm_{ki}(s) - \fundm_{li}(s)} \brc{\etwm_i(s) - \etwm_j(s)},
+  S_{ijkl} &= \smash{\pdiff{\hat{A}}{\encm_{ij} ; \frgm_{kl}} } &
+  &= \eqm_k \brc{\fundm_{ki}(0) - \fundm_{li}(0)} \brc{\etwm_i(s) - \etwm_j(s)} \\
+    &&& \hphantom{=} {}
+    + \eqm_i \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} \brc{\etwm_k(s) - \etwm_l(s)},
 \end{alignedat}
@@
 where \(R\) contains only those terms where \( \frgm_{ij} \) precedes \( \frgm_{kl} \) in the matrix products from the expression for \( \hat{A} \) in \cref{eq:larea}.
-The other terms sum to \(R\trans\).
+The missing terms sum to \(R\trans\).
 
@@
   \pdiff{^2\hat{A}(s)}{\MMdm^\mu_{ij} \partial \MMdm^\nu_{kl}} =
-     f^\mu f^\nu [R + R\trans + (-1)^\mu S\trans + (-1)^\nu S]_{ijkl}.
+     f^\mu f^\nu [R + R\trans + (-1)^\mu S + (-1)^\nu S\trans]_{ijkl}.
 \end{equation}
@@
     \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
-     + \prn{\mathbf{c}_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
+     + \prn{c_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
      &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
@@
 
-The expressions in \cref{eq:finitederivs} have the same form as for \(s=0\), which allowed us to argue that the model with maximal area must have the serial topology, \cite{Lahiri2013synapse}.
+The expressions in \cref{eq:finitederivs} have the same form as for \(s=0\), which allowed us to argue that the model with maximal area must have the serial topology in \cref{sec:areapert} and \cite{Lahiri2013synapse}.
 The one missing ingredient is the scale invariance.
@@
 
-\begin{figure}[tbp]
-  \centering
-  % Requires \usepackage{graphicx}
-  \begin{myenuma}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=7]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=13]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=21]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=27]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=36]{envelope.pdf}}
-  \item\aligntop{\includegraphics[width=0.45\linewidth,page=50]{envelope.pdf}}
-  \end{myenuma}
-  \caption[Optimal models]{The models that constitute the numeric envelope in \cref{fig:envnum}.
-  The plot to the left of each frame indicates the memory curve of the model shown to the right in green, with the numeric envelope for models with the serial topology in orange and the proven envelope, \eqref{eq:env}, in blue.
-  Above the left plot we have a graphical display of the model, described with more detail to the right.
-  The area of each node is proportional to the steady state probability in that state and the thickness of the red and blue edges are proportional to the transition probabilities under potentiation and depression.
-  The bar graph in the upper right of each frame indicates the transition probabilities from the state to the left of each bar to the state on its right, under potentiation.
-  The bar graph in the lower right of each frame indicates the transition probabilities from the state to the right of each bar to the state on its left, under depression.
-  The heat map in the centre right of each frame indicates the equilibrium probability distribution for each state.
-  See the video for the full set of models.}\label{fig:envvid}
-\end{figure}
-
 %-------Section------------------------------------------------------------------
@@
 
-Under the transformation from \cref{eq:scale}, the Laplace transform has the following scaling property
+Under the scale transformation
+%
+\begin{equation}\label{eq:scale}
+  \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I ,
+\end{equation}
+%
+the Laplace transform has the following scaling property
 %
@@
 At \(s = 0\), it is scale invariant.
-This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
+This means that in \cite{Lahiri2013synapse} we could take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
 This allowed us to ignore the diagonal constraints.
@@
 %
-\begin{equation}\label{eq:degenerate}
-  \dgn_{m} = \thb + c_m \etw,
+\begin{multline}\label{eq:degenerate}
+  \dgn_{m}(s) = \thb(s) + c_m \etw(s) + \lambda \etw(0),
+  \qquad
+  \dgn^\nu_m(s) = \dgn_m(s) + (-1)^\nu \etw(s), \\
+  \implies \quad
+  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
+\end{multline}
+%
+They should be thought of as collections of vectors rather than matrices.
+
+Suppose that a state \(n\) has a nonzero transition rate from state \(m\) for one \(\nu\).
+Then \(\mu^\nu_{mn} = 0\) and:
+%
+\begin{equation}\label{eq:oneexit}
+  \MMdm^\nu_{mn} > 0
+  \quad \implies \quad
+  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n})
+  \quad \implies \quad
+  \dgnm_{m,m} - \dgnm_{m,n} \geq - (-1)^\nu (\etwm_m - \etwm_n).
+\end{equation}
+%
+If both \(\nu\) have nonzero rates from \(m\) to \(n\):
+%
+\begin{equation}\label{eq:doubleexit}
+  \MMdm\pot_{mn}, \MMdm\dep_{mn} > 0
+  \qquad \implies \qquad
+  \dgnm_{m,m} - \dgnm_{m,n} \geq \abs{\etwm_m - \etwm_n}.
+\end{equation}
+%
+
+Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
+Both must statisfy \cref{eq:oneexit} for the same \(\mu^\nu_{mm}\), so
+%
+\begin{equation}\label{eq:twoexits}
+  \MMdm^\nu_{mi}, \MMdm^\nu_{mj} > 0
+  \qquad \implies \qquad
+  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
+\end{equation}
+%
+Furthermore, for a state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
+%
+\begin{equation}\label{eq:nonexit}
+  \MMdm^\nu_{mi} > 0, \MMdm^\nu_{mk} = 0 
+  \;\;\; \implies \;\;\;
+  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
+  \;\; \implies \;\;
+  \dgnm_{m,k} - \dgnm_{m,i} \geq - (-1)^\nu (\etwm_k - \etwm_i).
+\end{equation}
+%
+If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
+%
+\begin{equation}\label{eq:twodoubleexits}
+  \MMdm\pot_{mi}, \MMdm\dep_{mi}, \MMdm\pot_{mj}, \MMdm\dep_{mj} > 0
+  \qquad \implies \qquad
+  \dgnm_{m,i} = \dgnm_{m,j} 
+  \quad \text{and} \quad
+  \etwm_i = \etwm_j.
+\end{equation}
+%
+The consequences of these inequalities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}).
+We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
+Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.
+
+We use \(m*\) to label the intersection of the two diagonal lines in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}), wether or not there is a state there.
+It can be defined for each \(m\), provided that it has at least one outgoing transition for each plasticity type, via
+%
+\begin{equation*}
+\begin{lgathered}  
+  \dgnm\pot_{m*} = \dgnm\pot_{m,j} = \min_k \dgnm\pot_{m,k},
+  \qquad
+  \dgnm\dep_{m*} = \dgnm\dep_{m,i} = \min_k \dgnm\dep_{m,k},
+  \\
+  \text{and also} \quad
+  \dgnm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} + \dgnm\dep_{m*}},
   \quad
-  \dgn^\nu_m = \dgn_m + (-1)^\nu \etw,
-  \quad \implies \;
-  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
-\end{equation}
-%
-They should be thought of as collections of vectors rather than matrices.
-We can also write \cref{eq:diagKKT} as
-%
-\begin{equation}\label{eq:diagKKTz}
+  \etwm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} - \dgnm\dep_{m*}},
+  \quad
+  \thbm_{m*} = \dgnm_{m*} - c_m \etwm_{m*}.
+\end{lgathered}
+\end{equation*}
+%
+As shown in \cref{eq:twoexits}, this produces the same point regardless of which states \(i\) and \(j\) we use, 
+so long as there is a nonzero depressing/potentiating transition from \(m\) to that state.
+The KKT multipliers are related to these quantities by
+%
+\begin{equation}\label{eq:KKTz}
+  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m*}),
+  \qquad
+  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m*}).
+\end{equation}
+%
+
+Both of \cref{eq:twoexits,eq:twodoubleexits} require degeneracy -- 
+those two situations would require \(\dgn^\nu_m\) to have degenerate minima.
+The physicist's intuition states that degeneracy is fragile and requires a symmetry to protect it.
+Such a symmetry would imply that the states are lumpable.
+This would imply that, after lumping all lumpable states together, the situation in \cref{eq:twoexits} is impossible and \cref{eq:twodoubleexits} is even more impossible.
+If this intuition holds, we expect that \cref{fig:triangle_s}(\ref{fig:triangle_spacing}) would have at most one state on each of the two diagonal lines.
+Either there would be one state like \(n\), or there would be one state like \(i\) and one state like \(j\), and the first of these possibilities is rather implausible.
+
+%-------Section------------------------------------------------------------------
+
+\subsubsection{Diagonal constraints}\label{sec:diagKKT}
+
+We can write \cref{eq:diagKKT} as
+%
+\begin{equation*}
 \begin{aligned}
@@
     % \left.
-      - (-1)^\nu \prn{ 
-        [\encm \dgnm_m]_m
-        - \lambda a_m(0)
+      - (-1)^\nu [\encm \dgnm_m]_m
+    \right],
+\end{aligned}
+\end{equation*}
+%
+with all quantities evaluated at \(s\) unless labelled otherwise.
+
+From \cref{eq:oneexit,eq:twoexits} we see that
+%
+\begin{equation*}
+\begin{gathered}
+  \begin{aligned}
+    \MMdm^\nu_{mi} &> 0 &
+    \quad &\implies \quad &
+    \dgnm^\nu_{m,i} &= \dgnm^\nu_{m*}
+    \\
+    \sum_{k \neq m} \MMdm^\nu_{mk} &< 1 &
+    &\implies &
+    \dgnm^\nu_{m,m} &= \dgnm^\nu_{m*},
+    % \quad \forall \ i : \MMdm^\nu_{mi} > 0,
+  \end{aligned}
+  \\
+  \begin{aligned}
+    \implies \quad
+    [(\MMd^\nu - \I) \dgn_m]_m
+      &= \sum_k (\MMdm^\nu_{mk} - \delta_{mk}) \dgnm_{m,k}
+      = \sum_k (\MMdm^\nu_{mk} - \delta_{mk}) 
+        (\dgnm^\nu_{m,k} - (-1)^\nu \etwm_k) \\
+      &= \sum_{k \neq m} \MMdm^\nu_{mk} 
+        (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,m}) 
+        - (-1)^\nu [(\MMd^\nu - \I) \etw]_m \\
+      &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
+        \sum_{k \neq m} \MMdm^\nu_{mk}
+        - (-1)^\nu [(\MMd^\nu - \I) \etw]_m \\
+      &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
+        - (-1)^\nu [(\MMd^\nu - \I) \etw]_m \\
+      &= (\dgnm_{m*} - \dgnm_{m,m}) 
+        + (-1)^\nu \prn{
+          (\etwm_{m*} - \etwm_m)
+          - [(\MMd^\nu - \I) \etw]_m
+        },
+  \end{aligned}
+\end{gathered}
+\end{equation*}
+%
+where we used the fact that \( (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) \) is only nonzero when \( \sum_{k \neq m} \MMdm^\nu_{mk} = 1 \).
+Therefore
+%
+\begin{equation*}
+\begin{aligned}
+  [\enc \dgn_m]_m
+    &= \sum_{\nu} (-1)^\nu f^\nu 
+      (\dgnm_{m*} - \dgnm_{m,m}) 
+      + \sum_{\nu} f^\nu \prn{
+        (\etwm_{m*} - \etwm_m)
+        - [(\MMd^\nu - \I) \etw]_m
+      } \\
+    &= \Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + (\etwm_{m*} - \etwm_m)
+      - [\frg \etw]_m \\
+    &= \Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + (\etwm_{m*} - \etwm_m)
+      - (\delta w_m + s \delta \etwm_m).
+\end{aligned}
+\end{equation*}
+%
+Then we can also write \cref{eq:diagKKT} as
+%
+\begin{equation}\label{eq:diagKKTz}
+\begin{aligned}
+  \mu^\nu_{mm} &= \frac{\eqm_m}{2f^\nu} 
+    \left[ 
+      \hat{A}
+      - c_m (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m
+    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
+    % \left.
+      + (-1)^\nu \prn{ 
+        \Delta f (\dgnm_{m,m} - \dgnm_{m*}) 
+        + (\etwm_m - \etwm_{m*})
       }
@@
 %
-with all quantities evaluated at \(s\) unless labelled otherwise.
-
-Suppose that a state \(n\) has a nonzero transition rate from state \(m\) for one \(\nu\).
-Then \(\mu^\nu_{mn} = 0\) and:
-%
-\begin{equation}\label{eq:oneexit}
-  \MMdm^\nu_{mn} > 0
-  \quad \implies \quad
-  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n})
-  \quad \implies \quad
-  \dgnm_{m,m} - \dgnm_{m,n} \geq - (-1)^\nu (\etwm_m - \etwm_n).
-\end{equation}
-%
-If both \(\nu\) have nonzero rates from \(m\) to \(n\):
-%
-\begin{equation}\label{eq:doubleexit}
-  \MMdm\pot_{mn}, \MMdm\dep_{mn} > 0
-  \qquad \implies \qquad
-  \dgnm_{m,m} - \dgnm_{m,n} \geq\abs{\etwm_m - \etwm_n}.
-\end{equation}
-%
-
-Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
-Both must statisfy \cref{eq:oneexit} for the same \(\mu^\nu_{mm}\), so
-%
-\begin{equation}\label{eq:twoexits}
-  \MMdm^\nu_{mi}, \MMdm^\nu_{mj} > 0
-  \qquad \implies \qquad
-  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
-\end{equation}
-%
-Furthermore, for a state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
-%
-\begin{equation}\label{eq:nonexit}
-  \MMdm^\nu_{mi} > 0, \MMdm^\nu_{mk} = 0 
-  \;\;\; \implies \;\;\;
-  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
-  \;\; \implies \;\;
-  \dgnm_{m,k} - \dgnm_{m,i} \geq - (-1)^\nu (\etwm_k - \etwm_i).
-\end{equation}
-%
-If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
-%
-\begin{equation}\label{eq:twodoubleexits}
-  \MMdm\pot_{mi}, \MMdm\dep_{mi}, \MMdm\pot_{mj}, \MMdm\dep_{mj} > 0
-  \qquad \implies \qquad
-  \dgnm_{m,i} = \dgnm_{m,j} 
-  \quad \text{and} \quad
-  \etwm_i = \etwm_j.
-\end{equation}
-%
-The consequences of these inequalities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}).
-We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
-Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.
-
-Both of \cref{eq:twoexits,eq:twodoubleexits} require degeneracy.
-The physicist's intuition states that degeneracy is fragile and requires a symmetry to protect it.
-Such a symmetry would imply that the states are lumpable.
-This would imply that, after lumping all lumpable states together, the situation in \cref{eq:twoexits} is impossible and \cref{eq:twodoubleexits} is even more impossible.
-If this intuition holds, we expect that \cref{fig:triangle_s}(\ref{fig:triangle_spacing}) would have at most one state on each of the two diagonal lines.
-Either there would be one state like \(n\), or there would be one state like \(i\) and one state like \(j\).
+again with all quantities evaluated at \(s\) unless labelled otherwise.
+
+\Cref{eq:KKTz} provides another expression for \(\mu^\nu_{mm}\).
+That expression will agree with \cref{eq:diagKKTz} provided that
+%
+\begin{equation}\label{eq:diagAgree}
+    \hat{A} - c_m (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m 
+      = (\dgnm_{m,m} - \dgnm_{m*}) 
+        + \Delta f (\etwm_m - \etwm_{m*}).
+\end{equation}
+%
 
@@
 
-The area (and \(\Delta\pr\)) is invariant under the scale transformation
-%
-\begin{equation}\label{eq:scale}
+The area and \(\Delta\pr\) are invariant under the scale transformation \cref{eq:scale}
+%
+\begin{equation*}%\label{eq:scale}
   \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I .
-\end{equation}
+\end{equation*}
 %
@@
 \begin{equation*}
-  \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} = - \sum_{\nu, m} \mu^\nu_{mm} = 0 .
+  \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} 
+    = - \sum_{\nu, m} \mu^\nu_{mm} = 0 .
 \end{equation*}
@@
 \begin{equation}\label{eq:serialextarea}
-  \CA \leq \frac{(M-1)}{r}\prn{4\eqm_+\eqm_-}.
+  \CA \leq \frac{(M-1)}{r}\prn{4\eqm_+\eqm_-} = \frac{(M-1)}{r}\prn{1 - (\eqm_+ - \eqm_-)^2}.
 \end{equation}
@@
   {Numerical  and theoretical envelopes  for the signal-to-noise ratio, \eqref{eq:snrb}. 
-  As in \cref{fig:envnum}, \cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numerically with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
-  The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
-  The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
-  The dotted blue line indicates the proven envelope, \cref{eq:env} and the solid blue line is the conjectured envelope assuming detailed balance, see \cref{eq:envDB}.}\label{fig:envnumth}
+  As in \cref{fig:envnum}, \cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} 
+  are solved numerically with \(f\potdep=\frac{1}{2}\), 
+  the number of synapses \(N=1\) and the number of states \(M=10\).
+  The orange line indicates the maximum memory curve over the space of all matrices 
+  satisfying the constraints \eqref{eq:constr}.
+  The green line indicates maximum over models with the serial topology, 
+  \ie potentiation only moves up one state and depression only moves down one state.
+  The dotted blue line indicates the proven envelope, 
+  \cref{eq:env} and the solid blue line is the conjectured envelope assuming detailed balance, 
+  see \cref{eq:envDB}.}\label{fig:envnumth}
 \end{figure}
+
+\begin{figure}[tbp]
+  \centering
+  % Requires \usepackage{graphicx}
+  \begin{myenuma}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=7]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=13]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=21]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=27]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=36]{envelope.pdf}}
+  \item\aligntop{\includegraphics[width=0.45\linewidth,page=50]{envelope.pdf}}
+  \end{myenuma}
+  \caption[Optimal models]{The models that constitute the numeric envelope in \cref{fig:envnum}.
+  The plot to the left of each frame indicates the memory curve of the model shown to the right in green, with the numeric envelope for models with the serial topology in orange and the proven envelope, \eqref{eq:env}, in blue.
+  Above the left plot we have a graphical display of the model, described with more detail to the right.
+  The area of each node is proportional to the steady state probability in that state and the thickness of the red and blue edges are proportional to the transition probabilities under potentiation and depression.
+  The bar graph in the upper right of each frame indicates the transition probabilities from the state to the left of each bar to the state on its right, under potentiation.
+  The bar graph in the lower right of each frame indicates the transition probabilities from the state to the right of each bar to the state on its left, under depression.
+  The heat map in the centre right of each frame indicates the equilibrium probability distribution for each state.
+  See the video for the full set of models.}\label{fig:envvid}
+\end{figure}
+\
 
@@
   At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with \(M\) states.
-  At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model.
+  At later times the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model.
   At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
-  Once this reaches zero, we discard the old end state and work on the new end state.} \label{fig:heuristicmodels}
+  Once this reaches zero we discard the old end state and work on the new end state.
+  } \label{fig:heuristicmodels}
 \end{figure}
 
-Looking at them models that were found by numerical maximisation in \cref{fig:envvid} (see the video for the full set of models), suggests we look at the particular set of models, shown in \cref{fig:heuristicmodels}.
-
-At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with \(M\) states.
-At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model as \(\tau \to \infty\).
-At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
-Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from \cref{fig:envvid} and the video, this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
-Once the length has been reduced to \(M=2\), the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.
+Looking at them models that were found by numerical maximisation in \cref{fig:envvid} (see the video for the full set of models) suggests we look at the particular set of models shown in \cref{fig:heuristicmodels}.
+
+At one particular time the model is (\ref{fig:serial_uniform}) the uniform serial model with \(M\) states.
+At later times the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model as \(\tau \to \infty\).
+At earlier times the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
+Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from \cref{fig:envvid} and the video this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
+Once the length has been reduced to \(M=2\) the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.
 
@@
 \begin{equation}\label{eq:etarecur}
-  -(s\I + \onev\pib - \frg)\etw(s) = \w.
+  -(s\I + \ onev \arow - \frg)\etw(s) = \w.
 \end{equation}
@@
 \begin{equation}\label{eq:detarecur}
-  \etw(s) = -\frac{(\pib\etw(s))\onev+\w}{s} + \delta \boldsymbol{\eta}(s),
+  \etw(s) = -\frac{(\arow \etw(s)) \onev + \w}{s} + \delta \etw(s),
   \qquad
-  (\frg-s\I)\delta \boldsymbol{\eta}(s) = \frac{\frg\w}{s}.
+  (\frg - s\I) \delta \etw(s) = \frac{\frg \w}{s}.
 \end{equation}

### 2021-08-27 Notes/laplace.tex
changimg units

@@
   \qquad\text{where}\quad
-  \frg = \sum_\nu f^\nu \MMd^\nu - \I
-    = f\pot \MMd \pot + f\dep \MMd \dep - \I.
+  \frg = \sum_\nu r f^\nu (\MMd^\nu - \I)
+    = r f\pot \MMd \pot + r f\dep \MMd \dep - r \I.
 \end{equation}
@@
 \begin{aligned}
-  \av{\synid^1 \syn^1(t)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \e^{rt\frg} \w, \\
+  \av{\synid^1 \syn^1(t)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \e^{t\frg} \w, \\
   \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \onev\eq \w \\
@@
          &=  (f\pot-f\dep)\, \eq \w \\
-         &=  \Delta f\, \eq \e^{rt\frg} \w ,
+         &=  \Delta f\, \eq \e^{t\frg} \w ,
 \end{aligned}
@@
   \av{\synid \cdt \syn(t)} - \av{\synid \cdt \syn(\infty)}
-    &= N \eq \enc \, \e^{rt \frg} \w , \\
+    &= \frac{N}{r} \eq \enc \, \e^{t \frg} \w , \\
     \text{where} \qquad
-  \enc &= \sum_\nu (-1)^\nu f^\nu (\MMd^\nu - \I)
-    = f\pot(\MMd \pot-\I)-f\dep(\MMd \dep-\I)
+  \enc &= \sum_\nu (-1)^\nu r f^\nu (\MMd^\nu - \I)
+    = r f\pot \MMd\pot - r f\dep \MMd\dep - r \Delta f \I.
 \end{aligned}
@@
 \begin{equation}\label{eq:SNRcurveExact}
-  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt \frg}\w}
-                 {\sqrt{1-(\Delta f)^2(\eq \w)^2}}.
+  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{t \frg}\w}
+                 {r \sqrt{1-(\Delta f)^2(\eq \w)^2}}.
 \end{equation}
@@
 \begin{equation}\label{eq:pertfe}
-  \pdiff{}{\frgm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
+  \pdiff{}{\frgm_{ij}} = \frac{1}{2rf\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2rf\dep} \pdiff{}{\MMdm \dep_{ij}},
   \qquad
-  \pdiff{}{\encm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
+  \pdiff{}{\encm_{ij}} = \frac{1}{2rf\pot} \pdiff{}{\MMdm \pot_{ij}} - \frac{1}{2rf\dep} \pdiff{}{\MMdm \dep_{ij}},
 \end{equation}
@@
   \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
-  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq 2f^\mu, \\
+  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq 2rf^\mu, \\
   \sum_j \frgm_{ij} &=0, &
@@
   A(s) = \intd[_0^{\infty}]{t} \e^{-st}\SNR(t) 
-   &= \intd[_0^{\infty}]{t} \frac{ \sqrt{N} \eq \enc \e^{(r\frg-s)t} \w }
-                                { \sqrt{1-(\Delta f)^2(\eq\w)^2} } \\
-   &= \frac{\sqrt{N} \eq \enc \prn{s \I - r\frg}\inv \w }{\sqrt{1-(\Delta f)^2(\eq\w)^2}}.
+   &= \intd[_0^{\infty}]{t} \frac{ \sqrt{N} \eq \enc \e^{(\frg-s)t} \w }
+                                { r\sqrt{1-(\Delta f)^2(\eq\w)^2} } \\
+   &= \frac{\sqrt{N} \eq \enc \prn{s \I - \frg}\inv \w }{r\sqrt{1-(\Delta f)^2(\eq\w)^2}}.
 \end{aligned}
@@
 \begin{aligned}
-  \hat{A}(s) = A(s) \sqrt{\frac{{{1-(\Delta f)^2(\eq\w)^2}}}{N}}  &=  \eq \enc \fund(s) \w \\
+  \hat{A}(s) = r A(s) \sqrt{\frac{{{1-(\Delta f)^2(\eq\w)^2}}}{N}}  &=  \eq \enc \fund(s) \w \\
     &= \sum_{ijk} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)} \eqm_k \wm_k \\
@@
   \begin{aligned}
-    {c}_k(s) &= r\sum_{ij} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)}, \\
-    {a}_j(s) &= r\sum_{k} \encm_{jk} (\etwm_{j}(s) - \etwm_{k}(s)), \\
+    {c}_k(s) &= \sum_{ij} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)}, \\
+    {a}_j(s) &= \sum_{k} \encm_{jk} (\etwm_{j}(s) - \etwm_{k}(s)), \\
     \thbm_i(s) &= \sum_j \fptbm_{ij}(0)\, \eqm_j {a}_j(s), \\
@@
   \pdiff{\hat{A}(s)}{\MMdm^\mu_{mn}}
-     = f^\mu \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
+     = rf^\mu \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
      +\prn{{c}_m(s) + (-1)^\mu} \brk{\etwm_m(s) - \etwm_n(s)} }.
@@
 To write down the Hessian, we will first define some notation.
-We define the matrix \(\mathbf{X}(s) = r\fund(0) \enc \fund(s)\).
+We define the matrix \(\mathbf{X}(s) = \fund(0) \enc \fund(s)\).
 Given a four-index array \(H_{ijkl}\), we define \(H\trans_{ijkl} = H_{klij}\).
@@
   \pdiff{^2\hat{A}(s)}{\MMdm^\mu_{ij} \partial \MMdm^\nu_{kl}} =
-     f^\mu f^\nu [R + R\trans + (-1)^\mu S + (-1)^\nu S\trans]_{ijkl}.
+     r^2 f^\mu f^\nu [R + R\trans + (-1)^\mu S + (-1)^\nu S\trans]_{ijkl}.
 \end{equation}
@@
   \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
-        +\sum_{\nu i j} { f^\nu \mu^\nu_{ij} \MMdm^\nu_{ij} }
+        +\sum_{\nu i j} { r f^\nu \mu^\nu_{ij} \MMdm^\nu_{ij} }
         + \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
@@
 %
-We have scaled the KKT multipliers by \(f^\nu\) for later convenience.
+We have scaled the KKT multipliers by \(r f^\nu\) for later convenience.
 
@@
 \begin{aligned}
-  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+  \frac{1}{rf^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
     \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
@@
 %
-If we note that we can write \( \MMd^\nu - \I = \prn{\frg + (-1)^\nu \enc} / 2f^\nu \), we find that
+If we note that we can write \( \MMd^\nu - \I = \prn{\frg + (-1)^\nu \enc} / 2rf^\nu \), we find that
 %
@@
   \alb^\nu(s) &= \frac{\delta\mathbf{a}(s) + (-1)^\nu \enc \thb(s)}
-                     {2f^\nu},
+                     {2rf^\nu},
   \qquad 
   \omb^\nu(s) = \frac{\delta\w + s \delta\etw(s) - (-1)^\nu \mathbf{a}(s)}
-                     {2 f^\nu}, 
+                     {2rf^\nu}, 
   \\
-  \mu^\nu_{mm} &= \frac{\eqm_m}{2f^\nu} 
+  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
     \left[ 
@@
   \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
-        +\sum_{\nu i j}  f^\nu \mu^\nu_{ij}\prn{\MMdm^\nu_{ij} -  s (\eqm_j - \delta_{ij}) }
+        +\sum_{\nu i j} r f^\nu \mu^\nu_{ij}\prn{\MMdm^\nu_{ij} -  s (\eqm_j - \delta_{ij}) }
         &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
@@
 \begin{aligned}
-  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+  \frac{1}{r f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
     \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
@@
   2 \brk{\pdiff{\CL}{\encm_{mj}} - \pdiff{\CL}{\encm_{mi}}}
-  &= \frac{1}{f\pot} \brk{\pdiff{\CL}{\MMdm\pot_{mj}} - \pdiff{\CL}{\MMdm\pot_{mi}}} -
-  \frac{1}{f\dep} \brk{\pdiff{\CL}{\MMdm\dep_{mj}} - \pdiff{\CL}{\MMdm\dep_{mi}}} \\ 
+  &= \frac{1}{rf\pot} \brk{\pdiff{\CL}{\MMdm\pot_{mj}} - \pdiff{\CL}{\MMdm\pot_{mi}}} -
+  \frac{1}{rf\dep} \brk{\pdiff{\CL}{\MMdm\dep_{mj}} - \pdiff{\CL}{\MMdm\dep_{mi}}} \\ 
     &= 2 \eqm_m (\etwm_i - \etwm_j) 
@@
 If one of those transitions is zero, there will be two states with the property above: the targets of right-most depressing and left-most potentiating transitions.
+Note that we have not yet proved anything about where \(\etwm_m\) lies with respect to any of the others.
 
@@
 \begin{equation}\label{eq:pertK}
-  \frac{1}{f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} -
-  \frac{1}{f\dep} \pdiff{\CL}{\MMdm\dep_{mn}} =
+  \frac{1}{rf\pot} \pdiff{\CL}{\MMdm\pot_{mn}} -
+  \frac{1}{rf\dep} \pdiff{\CL}{\MMdm\dep_{mn}} =
     2 \eqm_m (\etwm_m - \etwm_n) 
@@
   \implies \quad
-  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
+  \frac{1}{rf^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
 \end{multline}
@@
 \begin{aligned}
-  \mu^\nu_{mm} &= \frac{\eqm_m}{2f^\nu} 
+  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
     \left[ 
@@
   [\enc \dgn_m]_m
-    &= \sum_{\nu} (-1)^\nu f^\nu 
+    &= \sum_{\nu} (-1)^\nu rf^\nu 
       (\dgnm_{m*} - \dgnm_{m,m}) 
@@
       } \\
-    &= \Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
-      + (\etwm_{m*} - \etwm_m)
-      - [\frg \etw]_m \\
-    &= \Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
-      + (\etwm_{m*} - \etwm_m)
+    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + r(\etwm_{m*} - \etwm_m)
+      - r[\frg \etw]_m \\
+    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + r(\etwm_{m*} - \etwm_m)
       - (\delta w_m + s \delta \etwm_m).
@@
 \begin{equation}\label{eq:snreig}
-  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-rt/\tau_a},
+  \snr(t) = \sqrt{N}\sum_a \CI_a \e^{-t/\tau_a},
   \qquad
-  \snrb(\tau) = \sqrt{N}\sum_a \frac{\CI_a}{1+r\tau/\tau_a},
+  \snrb(\tau) = \sqrt{N}\sum_a \frac{\CI_a}{1+\tau/\tau_a},
 \end{equation}
@@
       \end{cases} \\
-    r\eqm_i \frgm_{ij} \prn{\fptbm_{ij}+\fptbm_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
+    \eqm_i \frgm_{ij} \prn{\fptbm_{ij}+\fptbm_{ji}} &= 1 \quad\text{if}\quad i=j\pm1,
   \end{aligned}
@@
 \begin{aligned}
-  c_k &= \sum_{i<k} \prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i} \fptbm_{i,i+1}
+  c_k &= \sum_{i<k} r\prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i} \fptbm_{i,i+1}
     %\\&
-    - \sum_{i\geq k} \prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i} \fptbm_{i+1,i}
+    - \sum_{i\geq k} r\prn{\eqm_{i} f\pot q\pot_i + \eqm_{i+1} f\dep q\dep_i} \fptbm_{i+1,i}
   ,\\
-  c_{k+1} - c_k &= \prn{\eqm_{k} \frgm_{k,k+1} + \eqm_{k+1} \frgm_{k+1,k}}
+  c_{k+1} - c_k &= r\prn{\eqm_{k} \frgm_{k,k+1} + \eqm_{k+1} \frgm_{k+1,k}}
                     \prn{\fptbm_{k,k+1} + \fptbm_{k+1,k}}
-    = \frac{2}{r},\\
+    = 2,\\
   \sum_k c_k \eqm_k &= \sum_{ij} \eqm_i \encm_{ij} (\eta-\eta) = 0,\\
-  \implies c_k &= \frac{2}{r} \prn{k - \sum_j j\eqm_j},
+  \implies c_k &= 2 \prn{k - \sum_j j\eqm_j},
 \end{aligned}
@@
 \begin{equation}\label{eq:serialextarea}
-  \CA \leq \frac{(M-1)}{r}\prn{4\eqm_+\eqm_-} = \frac{(M-1)}{r}\prn{1 - (\eqm_+ - \eqm_-)^2}.
+  \CA \leq (M-1) \prn{4\eqm_+\eqm_-} = (M-1)\prn{1 - (\eqm_+ - \eqm_-)^2}.
 \end{equation}
@@
 \begin{equation}\label{eq:maxarea}
-  \CA \leq \frac{(M-1)}{r}.
+  \CA \leq (M-1).
 \end{equation}
@@
 \begin{equation*}
-  \snr(t) = \sqrt{N} \sum_a \CI_a \e^{-rt/\tau_a},
+  \snr(t) = \sqrt{N} \sum_a \CI_a \e^{-t/\tau_a},
   \qquad
-  \snrb(\tau) = \sqrt{N} \sum_a \frac{\CI_a}{1+r\tau/\tau_a},
+  \snrb(\tau) = \sqrt{N} \sum_a \frac{\CI_a}{1+\tau/\tau_a},
 \end{equation*}
@@
 \begin{aligned}
-  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{1-\varepsilon}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
+  \frac{1 - \varepsilon}{2} \delta \eta_2 
+    - \prn{s+\frac{1-\varepsilon}{2}} \delta \eta_1 
+    &= 0, \\
+  \frac{\delta \eta_{m-1}}{2} 
+    - \prn{s+\frac{3}{2}} \delta \eta_{m} 
+    &= \frac{1}{s}.
 \end{aligned}

### 2021-09-01 Notes/laplace.tex
M to W in finite time section

@@
 %
+It will be useful to introduce continuous time Markov processes corresponding to \(\MMd^\nu \):
+%
+\begin{equation*}
+  \W^\nu = r f^\nu (\MMd^\nu - \I),
+  \qquad
+  \frg = \sum_{\nu} \W^\nu.
+\end{equation*}
+%
 
@@
     \text{where} \qquad
-  \enc &= \sum_\nu (-1)^\nu r f^\nu (\MMd^\nu - \I)
+  \enc &= \sum_\nu (-1)^\nu \W^\nu
     = r f\pot \MMd\pot - r f\dep \MMd\dep - r \Delta f \I.
@@
 %
-The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with \(f\pot=f\dep\) or \(\eq_+=\eq_-\).
+The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with \(f\pot=f\dep\) or \(\eqm_+=\eqm_-\).
 We can perform the maximisation in two steps.
@@
 %
+It will sometimes be more convenient to use the off-diagonal elements of \(\W^\nu \) as the independent degrees of freedom.
+Then
+%
+\begin{equation}\label{eq:wdof}
+\begin{gathered}
+  \Wm^\nu_{i j} \geq 0
+    \quad \forall \; \nu, i, j : i \neq j, \qquad
+  \sum_{j : j\neq i} \Wm^\nu_{ij} \leq r f^\nu
+    \quad \forall \; \nu, i, \qquad
+  \Wm^\nu_{ii} = - \sum_{j : j\neq i} \Wm^\nu_{ij}, \\
+  \pdiff{}{\Wm^\nu_{ij}} = \frac{1}{rf^\nu} \pdiff{}{\MMdm^\nu_{ij}} 
+    \quad \forall \; \nu, i, j : i \neq j, \qquad
+  \pdiff{\Wm^\mu_{ij}}{\Wm^\nu_{mn}} 
+      = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}).
+\end{gathered}
+\end{equation}
+%
+Replacing \(\MMd\) with \(\W\) makes dimensional analysis easier and absorbs factors of \(f^\nu\) in several places.
+
 It will also be convenient to define the differential operators
@@
 %
-these satisfy
+which satisfy
 %
@@
 
-If we wish to use the off-diagonal elements of \(\frg\) and \(\enc\) as the degrees of freedom, the constraints are
-%
-\begin{equation}\label{eq:constrfe}
-\begin{aligned}
-  \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
-  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq 2rf^\mu, \\
-  \sum_j \frgm_{ij} &=0, &
-  \sum_j \encm_{ij} &=0.
-\end{aligned}
-\end{equation}
-%
+% If we wish to use the off-diagonal elements of \(\frg\) and \(\enc\) as the degrees of freedom, the constraints are
+% %
+% \begin{equation}\label{eq:constrfe}
+% \begin{aligned}
+%   \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
+%   \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq 2rf^\mu, \\
+%   \sum_j \frgm_{ij} &=0, &
+%   \sum_j \encm_{ij} &=0.
+% \end{aligned}
+% \end{equation}
+% %
 
@@
 We can also express this in terms of the shifted process in \cref{eq:larea}.
-We can choose how to distribute the shift across \(\MMd \pot\) and \(\MMd \dep\).
+We can choose how to distribute the shift across \(\W \pot\) and \(\W \dep\).
 Two choices from a continuum%
-\footnote{We could choose \( \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s g^\nu}{r f^\nu} (\onev \eq - \I) \), for any \(g^\nu \) with \( \sum_\nu g^\nu = 1 \).
-It shifts \(\frg\) correctly and also shifts \( \shift{\enc}(s) = \enc + \frac{s}{r} \Delta g (\onev \eq - \I) \).
+\footnote{We could choose \( \shift{\W}^\nu(s) = \W^\nu + s g^\nu (\onev \eq - \I) \), for any \(g^\nu \) with \( \sum_\nu g^\nu = 1 \).
+It shifts \(\frg\) correctly and also shifts \( \shift{\enc}(s) = \enc + s \Delta g (\onev \eq - \I) \).
 The two possibilities in \cref{eq:lshiftpd} come from \(g^\nu = \frac{1}{2}\) and \(g^\nu = f^\nu\) respectively.}
@@
 \begin{equation}\label{eq:lshiftpd}
-  \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s}{2 r f^\nu} (\onev \eq - \I)
+  \shift{\W}^\nu(s) = \W^\nu + \frac{s}{2} (\onev \eq - \I)
   \qquad \text{or} \qquad
-  \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s}{r} (\onev \eq - \I).
+  \shift{\W}^\nu(s) = \W^\nu + s f^\nu (\onev \eq - \I).
 \end{equation}
@@
 The first choice leaves \( \shift{\enc} = \enc \), so it only affects the \(\etw\) piece of \cref{eq:larea}.
-The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + \frac{s}{r} \Delta f (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
+The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + s \Delta f (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
 Either way:
@@
 \begin{equation}\label{eq:lareashift}
-  \hat{A}\prn{s; \MMd^\nu} = \hat{A}\prn{0; \shift{\MMd}^\nu(s)}.
+  \hat{A}\prn{s; \W^\nu} = \hat{A}\prn{0; \shift{\W}^\nu(s)}.
 \end{equation}
@@
 
-Note that while \cref{eq:lshift} produces a valid continuous-time Markov process, \cref{eq:lshiftpd} does not always produce valid discrete-time Markov processes, as the diagonal elements can become negative.
-However, \(r (\shift{\MMd}^\nu - \I)\) is a valid continuous-time Markov process.
+Note that while \cref{eq:lshift} produces a valid continuous-time Markov process, \cref{eq:lshiftpd} does not always produce valid discrete-time Markov processes, \( \shift{\MMd}^\nu = \frac{\shift{\W}^\nu}{rf^\nu} + \I \), as the diagonal elements can become negative.
+However, \(\shift{\W}^\nu \) is a valid continuous-time Markov process.
 
@@
 We can maximise the ratio \wrt \(\eq\w\) later.
-From this point on, we work in units where \(r = 1\).
+% From this point on, we work in units where \(r = 1\).
 
@@
   \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
-        +\sum_{\nu i j} { r f^\nu \mu^\nu_{ij} \MMdm^\nu_{ij} }
+        +\sum_{\nu i j} \mu^\nu_{ij} \prn{ \Wm^\nu_{ij} + r f^\nu \delta_{ij} }
         + \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
@@
 \begin{equation}\label{eq:KTfinite}
-  \pdiff{\CL}{\MMdm^\nu_{ij}} = 0, \quad
+  \pdiff{\CL}{\Wm^\nu_{ij}} = 0, \quad
   \mu^\nu_{ij} \geq 0, \quad
@@
 \begin{aligned}
-  \frac{1}{rf^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+  \pdiff{\CL}{\Wm^\nu_{mn}} = & \,
     \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
@@
 \begin{equation*}
-  \mu^\nu_{ij} \MMdm^\nu_{ij} = 0,
+  \mu^\nu_{ij} \Wm^\nu_{ij} = 0,
   \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} \MMdm^\nu_{ij} 
-      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii}) = \mu^\nu_{ii}.
+  \sum_{j: j \neq i} \mu^\nu_{ii} \Wm^\nu_{ij} 
+      = \mu^\nu_{ii} ( - \Wm^\nu_{ii}) = r f^\nu \mu^\nu_{ii}.
 \end{equation*}
@@
   \text{where} \quad 
-  \alb^\nu(s) &= (\MMd^\nu - \I) \thb(s), \qquad
-  \omb^\nu(s) = (\MMd^\nu - \I) \etw(s).
-\end{aligned}
-\end{equation}
-%
-If we note that we can write \( \MMd^\nu - \I = \prn{\frg + (-1)^\nu \enc} / 2rf^\nu \), we find that
+  \alb^\nu(s) &= \frac{\W^\nu \thb(s)}{rf^\nu}, \qquad
+  \omb^\nu(s) = \frac{\W^\nu \etw(s)}{rf^\nu}.
+\end{aligned}
+\end{equation}
+%
+If we note that we can write \( \W^\nu = \prn{\frg + (-1)^\nu \enc} / 2 \), we find that
 %
@@
 
-Under the scale transformation
-%
-\begin{equation}\label{eq:scale}
-  \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I ,
-\end{equation}
-%
-the Laplace transform has the following scaling property
+% Under the scale transformation
+% %
+% \begin{equation}\label{eq:scale}
+%   \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I ,
+% \end{equation}
+% %
+The Laplace transform has the following scaling property
 %
 \begin{equation}\label{eq:laplacescale}
-  \hat{A} \prn{\zeta s;\zeta * \MMd^\nu} = \hat{A} \prn{s; \MMd^\nu}.
+  \hat{A} \prn{\zeta s;\zeta \W^\nu} = \hat{A} \prn{s; \W^\nu}.
 \end{equation}
@@
 At \(s = 0\), it is scale invariant.
-This means that in \cite{Lahiri2013synapse} we could take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
+This means that in \cite{Lahiri2013synapse} we could take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \cref{eq:constri}) and use the scale transform to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
 This allowed us to ignore the diagonal constraints.
@@
 \begin{equation}\label{eq:shiftscale}
-  \zeta {\circledast}_{s} \MMd 
-      \equiv (1 - \zeta) \MMd
-      + \zeta \brk{(1 + s)\I - s \onev \eq} .
+  \zeta {\circledast}_{s} \W^\nu
+      \equiv \zeta \W^\nu
+      + (1 - \zeta) s f^\nu \brk{\I - \onev \eq} .
 \end{equation}
@@
 \begin{equation}\label{eq:lshiftscale}
-  \hat{A} \prn{s; \MMd^\nu} 
-    = \hat{A} \prn{0, \shift{\MMd}^\nu(s)}
-    = \hat{A} \prn{0, \zeta * \shift{\MMd}^\nu(s)}
-    = \hat{A} \prn{s, \zeta \circledast_s \MMd^\nu}.
-\end{equation}
-%
-The problem here is that the lower bound, \(\MMdm^\nu_{ij} \geq 0\), 
+  \hat{A} \prn{s; \W^\nu} 
+    = \hat{A} \prn{0, \shift{\W}^\nu(s)}
+    = \hat{A} \prn{0, \zeta \shift{\W}^\nu(s)}
+    = \hat{A} \prn{s, \zeta \circledast_s \W^\nu}.
+\end{equation}
+%
+The problem here is that the lower bound, \(\Wm^\nu_{ij} \geq 0\), 
 is not invariant under this transformation, unlike the \(s = 0\) case.
@@
 \begin{alignedat}{5}
-  [\zeta \circledast \MMd^\nu]_{ij}  
-    &= \zeta \MMdm^\nu_{ij} - (1-\zeta) s \eqm_j 
+  [\zeta \circledast \W^\nu]_{ij}  
+    &= \zeta \Wm^\nu_{ij} - (1-\zeta) s f^\nu \eqm_j 
     &&\geq 0 &
-  \quad &\implies \quad &
-  \zeta &\geq \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} &
+  \;\; &\implies \;\; &
+  \zeta &\geq \frac{s \eqm_j}{\Wm^\nu_{ij} /  f^\nu + s \eqm_j} &
     &\forall \nu, i, j : i \neq j , \\
-  \sum_{j : i \neq j} [\zeta \circledast \MMd^\nu]_{ij} 
-    &= \zeta \lbm^\nu_i - (1-\zeta) s (1-\eqm_i) 
-    &&\leq 1 &
+  \sum_{j : i \neq j} [\zeta \circledast \W^\nu]_{ij} 
+    &= \zeta \lbm^\nu_i - (1-\zeta) s f^\nu (1-\eqm_i) 
+    &&\leq r f^\nu &
   &\implies &
-  \zeta &\leq \frac{1 + s (1-\eqm_i)}{\lbm^\nu_i + s (1-\eqm_i)} &
+  \zeta &\leq \frac{r + s (1-\eqm_i)}{\lbm^\nu_i / f^\nu + s (1-\eqm_i)} &
     \quad &\forall \nu, i ,
@@
 %
-where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \MMdm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
+where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \Wm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
 The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
@@
 \begin{equation}\label{eq:fixable}
-  \frac{s \eqm_j}{\MMdm^\nu_{ij} + s \eqm_j} \leq \frac{1 + s (1-\eqm_k)}{\lbm^\mu_k + s (1-\eqm_k)}
+  \frac{s \eqm_j}{\Wm^\nu_{ij} / f^\nu + s \eqm_j} \leq \frac{r + s (1-\eqm_k)}{\lbm^\mu_k / f^\nu + s (1-\eqm_k)}
   \quad \implies \quad
-  \frac{\lbm^\mu_k - 1}{1 + s (1-\eqm_k)} \leq \frac{\MMdm^\nu_{ij}}{s \eqm_j} 
+  \frac{\lbm^\mu_k - r f^\nu}{r + s (1-\eqm_k)} \leq \frac{\Wm^\nu_{ij}}{s \eqm_j} 
   \quad \forall \mu, \nu, i, j, k : i \neq j.
@@
 %
-Note that when \(\MMdm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq 1\), the original diagonal constraint.
+Note that when \(\Wm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq r f^\nu\), the original diagonal constraint.
 Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
@@
 The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
-treat the off-diagonal components of \(\shift{\MMd}^\nu \) as the degrees of freedom and maximise the area instead.
-The constraints on \(\shift{\MMdm}^\nu_{ij}\) are different from \cref{eq:constri}:
+treat the off-diagonal components of \(\shift{\W}^\nu \) as the degrees of freedom and maximise the area instead.
+The constraints on \(\shift{\Wm}^\nu_{ij}\) are different from \cref{eq:constri}:
 %
 \begin{equation}\label{eq:lshiftconstr}
-  \shift{\MMdm}^\nu_{ij} \geq s \eqm_j,
+  \shift{\Wm}^\nu_{ij} \geq s f^\nu \eqm_j,
     \quad \forall \; \nu, i, j : i \neq j, \qquad
-  \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \leq 1 + s (1 - \eqm_i).
+  \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \leq r f^\nu + s f^\nu (1 - \eqm_i).
     \quad \forall \; \nu, i.
@@
 %
-The diagonal elements are still given by \( \shift{\MMdm}^\nu_{ii} = 1 - \sum_{j: j \neq i} \shift{\MMdm}^\nu_{ij} \), but now they are allowed to be negative so long as \(\shift{\MMdm}^\nu_{ii} \geq - s (1 - \eqm_i)\).
+The diagonal elements are still given by \( \shift{\Wm}^\nu_{ii} = - \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \), but now they are be below \(-r f^\nu \) so long as \(\shift{\Wm}^\nu_{ii} \geq -r f^\nu - s f^\nu (1 - \eqm_i)\).
 
@@
   \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
-        +\sum_{\nu i j} r f^\nu \mu^\nu_{ij}\prn{\MMdm^\nu_{ij} -  s (\eqm_j - \delta_{ij}) }
+        +\sum_{\nu i j} \mu^\nu_{ij}\prn{\Wm^\nu_{ij} + f^\nu (r + s) \delta_{ij} -  s f^\nu \eqm_j }
         &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
@@
 \begin{aligned}
-  \frac{1}{r f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
+  \pdiff{\CL}{\Wm^\nu_{mn}} = & \,
     \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
@@
   \qquad
-  \shift{\alb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
+  \shift{\alb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
   \\
-  \shift{\omb}^\nu(s) &= (\MMd^\nu + (s - 1) \I - s \onev \eq) \etw(0).
+  \shift{\omb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \etw(0).
 \end{aligned}
@@
 \begin{equation*}
-  \mu^\nu_{ij} (\MMdm^\nu_{ij} - s \eqm_j) = 0,
+  \mu^\nu_{ij} (\Wm^\nu_{ij} - s \eqm_j) = 0,
   \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} (\MMdm^\nu_{ij} - s \eqm_j)
-      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
+  \sum_{j: j \neq i} \mu^\nu_{ii} (\Wm^\nu_{ij} - s \eqm_j)
+      = \mu^\nu_{ii} (1 - \Wm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
 \end{equation*}
@@
 %
-although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\).
+although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\), so these expressions are not explicit.
 
@@
   2 \brk{\pdiff{\CL}{\encm_{mj}} - \pdiff{\CL}{\encm_{mi}}}
-  &= \frac{1}{rf\pot} \brk{\pdiff{\CL}{\MMdm\pot_{mj}} - \pdiff{\CL}{\MMdm\pot_{mi}}} -
-  \frac{1}{rf\dep} \brk{\pdiff{\CL}{\MMdm\dep_{mj}} - \pdiff{\CL}{\MMdm\dep_{mi}}} \\ 
+  &= \pdiff{\CL}{\Wm\pot_{mj}} - \pdiff{\CL}{\Wm\pot_{mi}} -
+  \pdiff{\CL}{\Wm\dep_{mj}} + \pdiff{\CL}{\Wm\dep_{mi}} \\ 
     &= 2 \eqm_m (\etwm_i - \etwm_j) 
@@
 Because \( \etwm_i > \etwm_j \), we must have \( \mu\pot_{mi} + \mu\dep_{mj} > 0 \).
-So \( \MMdm\pot_{mi} > 0 \) would imply that \( \MMdm\dep_{mj} = 0 \) and vice versa.
+So \( \Wm\pot_{mi} > 0 \) would imply that \( \Wm\dep_{mj} = 0 \) and vice versa.
 These two possibilities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_left},\ref{fig:triangle_right}).
@@
 Let us arrange the states left to right in order of decreasing \(\etw\), so that \(i\) is to the left of \(j\).
-For the unconstrained transitions \( \MMdm\pot_{mj} \) and \( \MMdm\dep_{mi} \) the potentiating transition goes to the \emph{right} of the depressing transition.
-For the other transitions \( \MMdm\pot_{mi} \) and \( \MMdm\dep_{mj} \) the potentiating transition goes to the \emph{left} of the depressing transition, and they cannot both be nonzero.
+For the unconstrained transitions \( \Wm\pot_{mj} \) and \( \Wm\dep_{mi} \) the potentiating transition goes to the \emph{right} of the depressing transition.
+For the other transitions \( \Wm\pot_{mi} \) and \( \Wm\dep_{mj} \) the potentiating transition goes to the \emph{left} of the depressing transition, and they cannot both be nonzero.
 
@@
   \begin{cases}
-    \etwm_i > \etwm_n &\implies \MMdm\pot_{mi} = 0, \\
-    \etwm_i < \etwm_n &\implies \MMdm\dep_{mi} = 0.
+    \etwm_i > \etwm_n &\implies \Wm\pot_{mi} = 0, \\
+    \etwm_i < \etwm_n &\implies \Wm\dep_{mi} = 0.
   \end{cases}
@@
 The remaining allowed transitions from \(m\) are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_allowed}).
-Note that we have not yet ruled the possibility that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero.
+Note that we have not yet ruled the possibility that both \( \Wm\pot_{mn} \) and \( \Wm\dep_{mn} \) are nonzero.
 If one of those transitions is zero, there will be two states with the property above: the targets of right-most depressing and left-most potentiating transitions.
@@
 \begin{equation}\label{eq:pertK}
-  \frac{1}{rf\pot} \pdiff{\CL}{\MMdm\pot_{mn}} -
-  \frac{1}{rf\dep} \pdiff{\CL}{\MMdm\dep_{mn}} =
+  \pdiff{\CL}{\Wm\pot_{mn}} -
+  \pdiff{\CL}{\Wm\dep_{mn}} =
     2 \eqm_m (\etwm_m - \etwm_n) 
@@
 %
-If we assume that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero, we find that \( \mu\pot_{mm} - \mu\dep_{mm} = 2 \eqm_m (\etwm_m - \etwm_n) \).
-Therefore \( \etwm_m > \etwm_n \implies \MMdm\pot_{mm} = 0 \) and \( \etwm_m < \etwm_n \implies \MMdm\dep_{mm} = 0 \).
-It is also possible that both of the diagonal elements are zero.
+If we assume that both \( \Wm\pot_{mn} \) and \( \Wm\dep_{mn} \) are nonzero, we find that \( \mu\pot_{mm} - \mu\dep_{mm} = 2 \eqm_m (\etwm_m - \etwm_n) \).
+Therefore \( \etwm_m > \etwm_n \implies \Wm\pot_{mm} = -rf\pot \) and \( \etwm_m < \etwm_n \implies \Wm\dep_{mm} = -rf\dep \).
+It is also possible that both of the diagonal constraints are saturated.
 
@@
 %
-\begin{multline}\label{eq:degenerate}
-  \dgn_{m}(s) = \thb(s) + c_m \etw(s) + \lambda \etw(0),
+\begin{equation}\label{eq:degenerate}
+\begin{aligned}
+    \dgn_{m}(s) &= \thb(s) + c_m \etw(s) + \lambda \etw(0), \\
     \qquad
-  \dgn^\nu_m(s) = \dgn_m(s) + (-1)^\nu \etw(s), \\
+    \dgn^\nu_m(s) &= \dgn_m(s) + (-1)^\nu \etw(s),   
+\end{aligned}
   \implies \quad
-  \frac{1}{rf^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
-\end{multline}
-%
-They should be thought of as collections of vectors rather than matrices.
+  \pdiff{\CL}{\Wm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
+\end{equation}
+%
+They should be thought of as collections of vectors rather than matrices, with the first \(m\) a label rather than an index.
 
@@
 \begin{equation}\label{eq:oneexit}
-  \MMdm^\nu_{mn} > 0
+\begin{aligned}
+  \Wm^\nu_{mn} > 0
   \quad \implies \quad
@@
   \quad \implies \quad
-  \dgnm_{m,m} - \dgnm_{m,n} \geq - (-1)^\nu (\etwm_m - \etwm_n).
+  \dgnm^\nu_{m,m} &\geq \dgnm^\nu_{m,n} \\
+  \text{or} \quad
+  \dgnm_{m,m} - \dgnm_{m,n} &\geq - (-1)^\nu (\etwm_m - \etwm_n).
+\end{aligned}
 \end{equation}
@@
 \begin{equation}\label{eq:doubleexit}
-  \MMdm\pot_{mn}, \MMdm\dep_{mn} > 0
+  \Wm\pot_{mn}, \Wm\dep_{mn} > 0
   \qquad \implies \qquad
@@
 \begin{equation}\label{eq:twoexits}
-  \MMdm^\nu_{mi}, \MMdm^\nu_{mj} > 0
+  \Wm^\nu_{mi}, \Wm^\nu_{mj} > 0
   \qquad \implies \qquad
+  \dgnm^\mu_{m,i} = \dgnm^\nu_{m,j}
+  \quad \text{or}\quad
   \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
@@
 %
-Furthermore, for a state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
+Furthermore, for another state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
 %
 \begin{equation}\label{eq:nonexit}
-  \MMdm^\nu_{mi} > 0, \MMdm^\nu_{mk} = 0 
+\begin{aligned}
+  \Wm^\nu_{mi} > 0, \Wm^\nu_{mk} = 0 
   \;\;\; \implies \;\;\;
@@
   \;\; \implies \;\;
-  \dgnm_{m,k} - \dgnm_{m,i} \geq - (-1)^\nu (\etwm_k - \etwm_i).
+  \dgnm^\nu_{m,k} &\geq \dgnm^\nu_{m,i} \\
+  \text{or}\quad
+  \dgnm_{m,k} - \dgnm_{m,i} &\geq - (-1)^\nu (\etwm_k - \etwm_i).
+\end{aligned}
 \end{equation}
@@
 \begin{equation}\label{eq:twodoubleexits}
-  \MMdm\pot_{mi}, \MMdm\dep_{mi}, \MMdm\pot_{mj}, \MMdm\dep_{mj} > 0
+  \Wm\pot_{mi}, \Wm\dep_{mi}, \Wm\pot_{mj}, \Wm\dep_{mj} > 0
   \qquad \implies \qquad
@@
   \begin{aligned}
-    \MMdm^\nu_{mi} &> 0 &
+    \W^\nu_{mi} &> 0 &
     \quad &\implies \quad &
@@
     \\
-    \sum_{k \neq m} \MMdm^\nu_{mk} &< 1 &
+    \sum_{k \neq m} \W^\nu_{mk} &< r f^\nu &
     &\implies &
     \dgnm^\nu_{m,m} &= \dgnm^\nu_{m*},
-    % \quad \forall \ i : \MMdm^\nu_{mi} > 0,
   \end{aligned}
@@
     \implies \quad
-    [(\MMd^\nu - \I) \dgn_m]_m
-      &= \sum_k (\MMdm^\nu_{mk} - \delta_{mk}) \dgnm_{m,k}
-      = \sum_k (\MMdm^\nu_{mk} - \delta_{mk}) 
+    [\W^\nu \dgn_m]_m
+      &= \sum_k \Wm^\nu_{mk}  \dgnm_{m,k}
+      = \sum_k \Wm^\nu_{mk}
         (\dgnm^\nu_{m,k} - (-1)^\nu \etwm_k) \\
-      &= \sum_{k \neq m} \MMdm^\nu_{mk} 
+      &= \sum_{k \neq m} \Wm^\nu_{mk} 
         (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,m}) 
-        - (-1)^\nu [(\MMd^\nu - \I) \etw]_m \\
+        - (-1)^\nu [\W^\nu \etw]_m \\
       &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
-        \sum_{k \neq m} \MMdm^\nu_{mk}
-        - (-1)^\nu [(\MMd^\nu - \I) \etw]_m \\
-      &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
-        - (-1)^\nu [(\MMd^\nu - \I) \etw]_m \\
-      &= (\dgnm_{m*} - \dgnm_{m,m}) 
+        \sum_{k \neq m} \Wm^\nu_{mk}
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= r f^\nu (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= r f^\nu (\dgnm_{m*} - \dgnm_{m,m}) 
         + (-1)^\nu \prn{
-          (\etwm_{m*} - \etwm_m)
-          - [(\MMd^\nu - \I) \etw]_m
+          r f^\nu (\etwm_{m*} - \etwm_m)
+          - [\W^\nu \etw]_m
         },
@@
 %
-where we used the fact that \( (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) \) is only nonzero when \( \sum_{k \neq m} \MMdm^\nu_{mk} = 1 \).
+where we used the fact that \( (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) \) is only nonzero when \( \sum_{k \neq m} \Wm^\nu_{mk} = rf^\nu \).
 Therefore
@@
       (\dgnm_{m*} - \dgnm_{m,m}) 
-      + \sum_{\nu} f^\nu \prn{
-        (\etwm_{m*} - \etwm_m)
-        - [(\MMd^\nu - \I) \etw]_m
+      + \sum_{\nu} \prn{
+        r f^\nu (\etwm_{m*} - \etwm_m)
+        -  [\W^\nu \etw]_m
       } \\
@@
       + r(\etwm_{m*} - \etwm_m)
-      - r[\frg \etw]_m \\
+      - [\frg \etw]_m \\
     &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
@@
 
+\subsection{Scaling with a floating cutoff}\label{sec:cutoff}
+
+%-------Section------------------------------------------------------------------
+
 \section{Upper bounds}\label{sec:upperbnds}
@@
 \begin{equation}\label{eq:snrcoeffs}
-  \CI_a = (\eq \enc \evr^a) (\evl^a \w),
+  \CI_a = \frac{(\eq \enc \evr^a) (\evl^a \w)}{r},
   \qquad
@@
 We have ignored the denominator of \cref{eq:SNRcurveExact} for now.
-This will not have any effect on the results: it will not change the location of the maxima, which are at \( f\pot - f\dep = \eq \w = 0 \).
+This will not have any effect on the results: it will not change the location of the maxima, which are at \( \Delta f = \eq \w = 0 \).
 We will confine such discussion to the footnotes.
+We will also work in units where \(r = 1\) until the very end.
 
@@
 \begin{equation*}
-  \frac{\SNR(0)}{\sqrt{N}} = \CI = \eq \enc \w.
-\end{equation*}
-%
-First, we define two sets of states: \( + = \set{i}{\wm_i = 1} \) and \( - = \set{i}{\wm_i = - 1} \), and the fluxes \( \Fm^\nu_{\CA \CB} = r f^\nu \displaystyle\sum_{\mathclap{i \in \CA, j \in \CB}} \eqm_i \prn{\MMdm^\nu_{ij} - \delta_{ij}} \).
+  \frac{\SNR(0)}{\sqrt{N}} = \CI = \frac{\eq \enc \w}{r}.
+\end{equation*}
+%
+First, we define two sets of states: \( + = \set{i}{\wm_i = 1} \) and \( - = \set{i}{\wm_i = - 1} \), and the fluxes \( \Fm^\nu_{\CA \CB} = \displaystyle\sum_{\mathclap{i \in \CA, j \in \CB}} \eqm_i \Wm^\nu_{ij} \).
 Then we can write
@@
 \begin{aligned}
-  r\CI &= \Fm\pot_{++} + \Fm\pot_{-+} - \Fm\dep_{++} - \Fm\dep_{-+} - \Fm\pot_{+-} - \Fm\pot_{--} + \Fm\dep_{+-} + \Fm\dep_{--} \\
+  \CI &= \Fm\pot_{++} + \Fm\pot_{-+} - \Fm\dep_{++} - \Fm\dep_{-+} - \Fm\pot_{+-} - \Fm\pot_{--} + \Fm\dep_{+-} + \Fm\dep_{--} \\
       &= 2\prn{\Fm\pot_{-+} - \Fm\dep_{-+} - \Fm\pot_{+-} + \Fm\dep_{+-}} \\
@@
 
-The last two terms are guaranteed to be negative, as the diagonal parts of \(\MMd \potdep - \I\) cannot contribute.
+The last two terms are guaranteed to be negative, as the diagonal parts of \(\W^\nu \) cannot contribute.
 Therefore
@@
   ,\\ \implies\quad
-  \frg = \begin{pmatrix}
+  \frg = r\begin{pmatrix}
             -f\pot & f\pot \\
@@
 
-The area and \(\Delta\pr\) are invariant under the scale transformation \cref{eq:scale}
-%
-\begin{equation*}%\label{eq:scale}
+The area and \(\Delta\pr\) are invariant under the scale transformation %\cref{eq:scale}
+%
+\begin{equation}\label{eq:scale}
   \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I .
-\end{equation*}
+\end{equation}
 %
@@
     \brk{ \sum_{k=m}^{n \mp 1} \frac{1}{\eqm_{k}} \pdiff{}{\MMdm^\nu_{k,k \pm 1}} }
-    - \frac{1}{\eq_m} \pdiff{}{\MMdm^\nu_{mn}},
+    - \frac{1}{\eqm_m} \pdiff{}{\MMdm^\nu_{mn}},
   \qquad \pm = (-1)^\nu.
@@
       \brk{ \sum_{k=m}^{n\mp1} \frac{\mu^\nu_{k,k\pm1}}{\eqm_k}
-      - \frac{\mu^\nu_{m,n}}{\eq_m}}
+      - \frac{\mu^\nu_{m,n}}{\eqm_m}}
       + rf^\nu \sum_{k=m}^{n\mp1} \prn{\etwm_{k} - \etwm_{k\pm1}} \prn{c_k - c_m},
@@
   ,\\
-  c_{k+1} - c_k &= r\prn{\eqm_{k} \frgm_{k,k+1} + \eqm_{k+1} \frgm_{k+1,k}}
+  c_{k+1} - c_k &= \prn{\eqm_{k} \frgm_{k,k+1} + \eqm_{k+1} \frgm_{k+1,k}}
                     \prn{\fptbm_{k,k+1} + \fptbm_{k+1,k}}
@@
 \begin{equation}\label{eq:serialarea}
- % A = 2\sqrt{N} \sum_k \brk{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eq_j} \eq_k \w_k.
-  \CA = \frac{2}{r} \sum_k \brk{k - \sum_j j\eq_j} \eqm_k \wm_k
-    = \frac{2}{r} \sum_k \abs{k - \sum_j j\eq_j} \eqm_k ,
+ % A = 2\sqrt{N} \sum_k \brk{\prn{k-\frac{n+1}{2}} - \sum_j\prn{j-\frac{n+1}{2}}\eqm_j} \eqm_k \w_k.
+  \CA = {2} \sum_k \brk{k - \sum_j j\eqm_j} \eqm_k \wm_k
+    = {2} \sum_k \abs{k - \sum_j j\eqm_j} \eqm_k ,
 \end{equation}
@@
 This gives an area of
-\footnote{Note that including the dropped factors in \cref{eq:lsnr} would only change this to \(\CA \leq (M-1)(4\eq_+\eq_-)/r\sqrt{1 - (\Delta f)^2(\eqm_+ - \eqm_-)^2}\), whose maximum also lies at \(\eqm_+=\eqm_-= \frac{1}{2}\).}
+\footnote{Note that including the dropped factors in \cref{eq:lsnr} would only change this to \(\CA \leq (M-1)(4\eqm_+\eqm_-)/r\sqrt{1 - (\Delta f)^2(\eqm_+ - \eqm_-)^2}\), whose maximum also lies at \(\eqm_+=\eqm_-= \frac{1}{2}\).}
 %
@@
   \begin{aligned}[t]
-  r \eq \enc \mathbf{u} &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
+  \eq \enc \mathbf{u} &= \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
         \Fm\pot_{\CA_\alpha \CA_\alpha^c} - \Fm\dep_{\CA_\alpha \CA_\alpha^c} 
@@
     \\
-    \Fm^\nu_{\CA \CB} &= r f^\nu \sum_{\substack{i \in \CA \\ j \in \CB}} 
-                        \eqm_i \prn{\MMdm^\nu_{ij} - \delta_{ij}} .
+    \Fm^\nu_{\CA \CB} &= \sum_{\substack{i \in \CA \\ j \in \CB}} 
+                        \eqm_i \Wm^\nu_{ij} .
   \end{aligned}
@@
   \CI(\mathbf{u}) 
-      &\leq \frac{\av{\mathbf{u},\w}}{r \nrm{\mathbf{u}}^2} 
+      &\leq \frac{\av{\mathbf{u},\w}}{\nrm{\mathbf{u}}^2} 
         \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
         \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} } 
-  \\  &= \frac{\av{\mathbf{u},\w}}{r \nrm{\mathbf{u}}^2} 
+  \\  &= \frac{\av{\mathbf{u},\w}}{\nrm{\mathbf{u}}^2} 
         \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \Fm_{ij} \brk{
@@
         }
-  \\  &= \frac{\av{\mathbf{u},\w}}{r \nrm{\mathbf{u}}^2} 
+  \\  &= \frac{\av{\mathbf{u},\w}}{\nrm{\mathbf{u}}^2} 
       \sum_{i \neq j} \Fm_{ij} \abs{u_j - u_i}
   \\  &= \frac{\av{\mathbf{u},\w} \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem}}
-              {r \nrm{\mathbf{u}}^2}.
+              {\nrm{\mathbf{u}}^2}.
 \end{aligned}
@@
   \CI(\mathbf{u}) \leq \frac{\nrm{\w} \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem}}
-                            {r \nrm{\mathbf{u}}} .
+                            {\nrm{\mathbf{u}}} .
 \end{equation*}
@@
 We also have \( \nrm{\w}^2 = \sum_i \eqm_i \wm_i^2 = 1 \).
-Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \MMm_{ij} \leq r \sum_i \eqm_i = r\).
+Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \MMm_{ij} \leq r \sum_i \eqm_i = r\) (which is 1 in our units!).
 Then, remembering \cref{eq:numrange},
@@
 \begin{equation}\label{eq:sqrtnumrange}
-  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\MM \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
+  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\MM \mathbf{u}}}{\av{\mathbf{u}, \mathbf{u}}}}
       = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})}.
@@
   \qquad \text{we have} \quad
-  \CI_a \sqrt{\tau_a} \leq \sqrt{2}.
+  \CI_a \sqrt{r\tau_a} \leq \sqrt{2}.
 \end{equation}
@@
 \begin{equation}\label{eq:envlagrangian}
-  \CL = \sum_a \frac{\CI_a}{1+r\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} + \mu_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
+  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} + \mu_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
 \end{equation}
@@
 \begin{equation}\label{eq:envDBlagrangian}
-  \CL = \sum_a \frac{\CI_a}{1+r\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} 
+  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} 
       + \sum_a \mu_a \prn{\sqrt{2} - \CI_a \sqrt{\tau_a}} + \mu_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
@@
     \midrule
-      0 &\leq r \tau &\leq 2 & M &\geq 3 & 1 & 2 
+      0 &\leq r \tau &\leq 2 & M &\geq 3 & 1 & 2/r 
         & \frac{4 - 2r \tau}{[2 + r \tau]^2} & \frac{2\sqrt{2}r\tau}{[2 + r \tau]^2} & 0 
@@
       \\
-      2 &< r \tau &< \frac{(M-1)^2}{2} & M &> 3 & \sqrt{\frac{2}{\tau}} & \tau 
+      2 &< r \tau &< \frac{(M-1)^2}{2} & M &> 3 & \sqrt{\frac{2}{r\tau}} & \tau 
         & 0 & \frac{1}{2\sqrt{r \tau}} & 0 
@@
       \\ \addlinespace[0.2ex]
-      \frac{(M-1)^2}{2} &\leq r \tau & & M &\geq 3 & \frac{2}{M-1} & \frac{(M-1)^2}{2} 
+      \frac{(M-1)^2}{2} &\leq r \tau & & M &\geq 3 & \frac{2}{M-1} & \frac{(M-1)^2}{2r} 
         & 0 & \frac{\sqrt{2}(M-1)^3}{[(M-1)^2 + 2r \tau]^2} 
@@
       \\
-      &-& & M &\leq 3 & 1 & M - 1 
+      &-& & M &\leq 3 & 1 & \frac{M - 1}{r} 
         & \frac{(M-1)^2}{[(M-1) + r \tau]^2} %- \frac{\mu_a \sqrt{\tau_a}}{2} 
@@
 \begin{equation*}
-  \hat{A}(s) = \frac{(\alpha-1)^2}
-                    {[2(\alpha^2-\alpha+1) - (1-\varepsilon) (\alpha-1)^2]s},
+  \hat{A}(s) = \frac{1}{1 + (1+\epsilon)s},
 \end{equation*}
@@
 This is plotted in \cref{fig:heuristicenv}.
+It should be compared to \cref{eq:env,eq:envDB}.
 

### 2021-09-01 Notes/laplace.tex
swap secs 3 & 4 back

@@
 
-\section{Finite time topology}\label{sec:finite}
-
-Now we'll maximise \(\snrb(\tau)\) for some fixed \(\tau\), subject to the constraints \eqref{eq:constr}.
-First, we will maximise the numerator of \eqref{eq:lsnr} holding \(\eq\w\) fixed.
-We can maximise the ratio \wrt \(\eq\w\) later.
-% From this point on, we work in units where \(r = 1\).
-
-Consider the Lagrangian
-%
-\begin{equation}\label{eq:finitelagrangian}
-\begin{aligned}
-  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
-        +\sum_{\nu i j} \mu^\nu_{ij} \prn{ \Wm^\nu_{ij} + r f^\nu \delta_{ij} }
-        + \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
-\end{aligned}
-\end{equation}
-%
-where \(\Delta\pr\) is the constant value we are holding \(\eq\w\) at.
-The Karush-Kuhn-Tucker necessary conditions for a maximum are
-%
-\begin{equation}\label{eq:KTfinite}
-  \pdiff{\CL}{\Wm^\nu_{ij}} = 0, \quad
-  \mu^\nu_{ij} \geq 0, \quad
-  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
-  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
-  \pdiff{\CL}{\lambda} = 0.
-\end{equation}
-%
-We have scaled the KKT multipliers by \(r f^\nu\) for later convenience.
-
-The derivatives, for \(m \neq n\) are
-%
-\begin{equation}\label{eq:finitederivs}
-\begin{aligned}
-  \pdiff{\CL}{\Wm^\nu_{mn}} = & \,
-    \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
-     + \prn{c_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
-     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
-    = 0.
-\end{aligned}
-\end{equation}
-%
-We can solve for the KKT multipliers by noting that
-%
-\begin{equation*}
-  \mu^\nu_{ij} \Wm^\nu_{ij} = 0,
-  \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} \Wm^\nu_{ij} 
-      = \mu^\nu_{ii} ( - \Wm^\nu_{ii}) = r f^\nu \mu^\nu_{ii}.
-\end{equation*}
-%
-Then, with some new vectors defined:
-%
-\begin{equation}\label{eq:kktsol}
-\begin{aligned}
-  \mu^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
-    + \lambda \ombm^\nu_m(0) 
-    + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
-    }, \\
-  \mu^\nu_{mn} &= \eqm_m \big( 
-      \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
-    + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
-  \\ &\hphantom{= \eqm_m \big( }
-    + \brk{{c}_m(s) + (-1)^\nu} 
-        \brk{ \etwm_n(s) - \etwm_m(s) - \ombm^\nu_m(s) }
-     \big), \\
-  \text{where} \quad 
-  \alb^\nu(s) &= \frac{\W^\nu \thb(s)}{rf^\nu}, \qquad
-  \omb^\nu(s) = \frac{\W^\nu \etw(s)}{rf^\nu}.
-\end{aligned}
-\end{equation}
-%
-If we note that we can write \( \W^\nu = \prn{\frg + (-1)^\nu \enc} / 2 \), we find that
-%
-\begin{equation}\label{eq:diagKKT}
-\begin{aligned}
-  \alb^\nu(s) &= \frac{\delta\mathbf{a}(s) + (-1)^\nu \enc \thb(s)}
-                     {2rf^\nu},
-  \qquad 
-  \omb^\nu(s) = \frac{\delta\w + s \delta\etw(s) - (-1)^\nu \mathbf{a}(s)}
-                     {2rf^\nu}, 
-  \\
-  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
-    \left[ 
-      \hat{A}(s) - c_m(s) (\delta\wm_m + s \delta\etwm_m(s))
-      - \lambda \delta w_m
-    \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} \big[}
-    \left. {}
-      + (-1)^\nu \prn{ 
-        c_m(s) a_m(s) - [\encm \thbm(s)]_m
-        - \delta\wm_m - s \delta\etwm_m(s)
-        + \lambda a_m(0)
-      }
-    \right], 
-\end{aligned}
-\end{equation}
-%
-where \(\delta\) indicates mean subtraction and \(\eq \mathbf{a}(s) = \hat{A}(s)\).
-
-The expressions in \cref{eq:finitederivs} have the same form as for \(s=0\), which allowed us to argue that the model with maximal area must have the serial topology in \cref{sec:areapert} and \cite{Lahiri2013synapse}.
-The one missing ingredient is the scale invariance.
-
-However, we do have numerical evidence that the model that maximises the Laplace transform does have the serial topology.
-In \cref{fig:envnum} we see the result of numerical maximisation of \cref{eq:finitelagrangian}.
-We see that restricting to the serial topology does not make the performance any worse.
-In fact, it slightly improves the performance, but this can be ascribed to the numerical optimisation problem being easier.
-Assuming that the optimiser did find the global maximum, this provides evidence for our conjecture that the model that maximises the Laplace transform must have the serial topology.
-The models found numerically are described in \cref{fig:envvid}.
-
-\begin{figure}[tb]
-  \centering
-  % Requires \usepackage{graphicx}
-  \includegraphics[width=0.8\linewidth]{LenvNumOnly.pdf}\\
-  \caption[Numerical envelope for the signal-to-noise ratio]
-  {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
-  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numerically with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
-  The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
-  The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
-  }\label{fig:envnum}
-\end{figure}
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Scale invariance}\label{sec:scale}
-
-% Under the scale transformation
-% %
-% \begin{equation}\label{eq:scale}
-%   \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I ,
-% \end{equation}
-% %
-The Laplace transform has the following scaling property
-%
-\begin{equation}\label{eq:laplacescale}
-  \hat{A} \prn{\zeta s;\zeta \W^\nu} = \hat{A} \prn{s; \W^\nu}.
-\end{equation}
-%
-At \(s = 0\), it is scale invariant.
-This means that in \cite{Lahiri2013synapse} we could take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \cref{eq:constri}) and use the scale transform to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
-This allowed us to ignore the diagonal constraints.
-We can't do that for \(s > 0\).
-
-The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
-%
-\begin{equation}\label{eq:shiftscale}
-  \zeta {\circledast}_{s} \W^\nu
-      \equiv \zeta \W^\nu
-      + (1 - \zeta) s f^\nu \brk{\I - \onev \eq} .
-\end{equation}
-%
-Then we have
-%
-\begin{equation}\label{eq:lshiftscale}
-  \hat{A} \prn{s; \W^\nu} 
-    = \hat{A} \prn{0, \shift{\W}^\nu(s)}
-    = \hat{A} \prn{0, \zeta \shift{\W}^\nu(s)}
-    = \hat{A} \prn{s, \zeta \circledast_s \W^\nu}.
-\end{equation}
-%
-The problem here is that the lower bound, \(\Wm^\nu_{ij} \geq 0\), 
-is not invariant under this transformation, unlike the \(s = 0\) case.
-Using this transformation to fix the upper bound can break the lower bound.
-
-We might ask what conditions are required so that a model is fixable?
-This means that the \(\zeta \circledast_s \) orbit of the model contains some valid models.
-The conditions are
-%
-\begin{equation*}
-\begin{alignedat}{5}
-  [\zeta \circledast \W^\nu]_{ij}  
-    &= \zeta \Wm^\nu_{ij} - (1-\zeta) s f^\nu \eqm_j 
-    &&\geq 0 &
-  \;\; &\implies \;\; &
-  \zeta &\geq \frac{s \eqm_j}{\Wm^\nu_{ij} /  f^\nu + s \eqm_j} &
-    &\forall \nu, i, j : i \neq j , \\
-  \sum_{j : i \neq j} [\zeta \circledast \W^\nu]_{ij} 
-    &= \zeta \lbm^\nu_i - (1-\zeta) s f^\nu (1-\eqm_i) 
-    &&\leq r f^\nu &
-  &\implies &
-  \zeta &\leq \frac{r + s (1-\eqm_i)}{\lbm^\nu_i / f^\nu + s (1-\eqm_i)} &
-    \quad &\forall \nu, i ,
-\end{alignedat}
-\end{equation*}
-%
-where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \Wm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
-The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
-%
-\begin{equation}\label{eq:fixable}
-  \frac{s \eqm_j}{\Wm^\nu_{ij} / f^\nu + s \eqm_j} \leq \frac{r + s (1-\eqm_k)}{\lbm^\mu_k / f^\nu + s (1-\eqm_k)}
-  \quad \implies \quad
-  \frac{\lbm^\mu_k - r f^\nu}{r + s (1-\eqm_k)} \leq \frac{\Wm^\nu_{ij}}{s \eqm_j} 
-  \quad \forall \mu, \nu, i, j, k : i \neq j.
-\end{equation}
-%
-Note that when \(\Wm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq r f^\nu\), the original diagonal constraint.
-Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Shifted problem}\label{sec:lshift}
-
-The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
-treat the off-diagonal components of \(\shift{\W}^\nu \) as the degrees of freedom and maximise the area instead.
-The constraints on \(\shift{\Wm}^\nu_{ij}\) are different from \cref{eq:constri}:
-%
-\begin{equation}\label{eq:lshiftconstr}
-  \shift{\Wm}^\nu_{ij} \geq s f^\nu \eqm_j,
-    \quad \forall \; \nu, i, j : i \neq j, \qquad
-  \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \leq r f^\nu + s f^\nu (1 - \eqm_i).
-    \quad \forall \; \nu, i.
-\end{equation}
-%
-The diagonal elements are still given by \( \shift{\Wm}^\nu_{ii} = - \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \), but now they are be below \(-r f^\nu \) so long as \(\shift{\Wm}^\nu_{ii} \geq -r f^\nu - s f^\nu (1 - \eqm_i)\).
-
-The Lagrangian we should consider is
-%
-\begin{equation}\label{eq:lshiftlagrangian}
-\begin{aligned}
-  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
-        +\sum_{\nu i j} \mu^\nu_{ij}\prn{\Wm^\nu_{ij} + f^\nu (r + s) \delta_{ij} -  s f^\nu \eqm_j }
-        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
-\end{aligned}
-\end{equation}
-%
-where all of the first-passage times are evaluated at \(s = 0\).
-
-In \cref{fig:shifted} we show the result of solving this problem with numerical optimisation.
-We see essentially the same results as the original problem, \cref{eq:finitelagrangian,fig:envnum}, as expected.
-
-The derivatives, for \(m \neq n\) are
-%
-\begin{equation}\label{eq:lshiftderivs}
-\begin{aligned}
-  \pdiff{\CL}{\Wm^\nu_{mn}} = & \,
-    \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
-     + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
-     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     - \sum_{\mu ij} \mu^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
-    = 0.
-\end{aligned}
-\end{equation}
-%
-We can absorb the new terms by defining
-%
-\begin{equation*}
-\begin{aligned}
-  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
-  \qquad
-  \shift{\alb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
-  \\
-  \shift{\omb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \etw(0).
-\end{aligned}
-\end{equation*}
-%
-We note that
-%
-\begin{equation*}
-  \mu^\nu_{ij} (\Wm^\nu_{ij} - s \eqm_j) = 0,
-  \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} (\Wm^\nu_{ij} - s \eqm_j)
-      = \mu^\nu_{ii} (1 - \Wm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
-\end{equation*}
-%
-then we can ``solve'' for the KKT multipliers:
-%
-\begin{equation}\label{eq:lshiftsol}
-\begin{aligned}
-  \mu^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
-    + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
-    }, \\
-  \mu^\nu_{mn} &= \eqm_m \prn{ 
-      \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
-  % \\ & \hphantom{= - \eqm_m \big( } 
-    + \brk{ {c}_m + (-1)^\nu + \lambda } 
-        \brk{ \etwm_n - \etwm_m - \shift{\ombm}^\nu_m(s) }
-     }, \\
-\end{aligned}
-\end{equation}
-%
-although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\), so these expressions are not explicit.
-
-While this problem involves the scale invariant quantity \(\hat{A}(0)\), neither the upper nor lower bounds in \cref{eq:lshiftconstr} are invariant.
-We then run into the same problems in proving our serial-topology hypothesis.
-
-\begin{figure}[ht]
-\begin{center}
-\begin{myenuma}
-  \item \aligntop{\includegraphics[width=0.4\linewidth]{LenvShift.pdf}}
-  \label{fig:envshift}
-  \item \aligntop{\includegraphics[width=0.4\linewidth]{shift_cond.pdf}}
-  \label{fig:shiftcond}
-\end{myenuma}
-\caption[Shifted problem vs.\ original]{Shifted problem vs.\ original.
-(\ref*{fig:envshift}) Comparison of numerical optimisation of the shifted problem in \cref{eq:lshiftlagrangian,fig:envnum} with the original problemin \cref{eq:finitelagrangian}.
-We see that the two approaches produce very similar numerical envelopes.
-The shifted problem involves non-linear constraints and is thus harder to solve.
-(\ref{fig:shiftcond}) Plot of the discrepancy in (\ref*{fig:envshift}) against the condition number of the fundamental matrix, \(\fund \).
-We see that most of the large discrepancies occur where the problem is numerically ill-conditioned.
-\label{fig:shifted}}
-\end{center}
-\end{figure}
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Topology constraining perturbations}\label{sec:topology}
-
-In this section we will see how far we can get in constraining the topology, 
-despite the lack of scale invariance.
-
-%-------Section------------------------------------------------------------------
-
-\subsubsection{Triangularity}\label{sec:parttriangular}
-
-Suppose we have two states, \(i\) and \(j\), with \( \etwm_i > \etwm_j \).
-Consider the following perturbation (see also \cref{fig:triangle_s}(\ref{fig:triangle_pert})):
-%
-\begin{equation}\label{eq:pertKupdn}
-\begin{aligned}
-  2 \brk{\pdiff{\CL}{\encm_{mj}} - \pdiff{\CL}{\encm_{mi}}}
-  &= \pdiff{\CL}{\Wm\pot_{mj}} - \pdiff{\CL}{\Wm\pot_{mi}} -
-  \pdiff{\CL}{\Wm\dep_{mj}} + \pdiff{\CL}{\Wm\dep_{mi}} \\ 
-    &= 2 \eqm_m (\etwm_i - \etwm_j) 
-    + \mu\pot_{mj} - \mu\pot_{mi} - \mu\dep_{mj} + \mu\dep_{mi}.
-\end{aligned}
-\end{equation}
-%
-Because \( \etwm_i > \etwm_j \), we must have \( \mu\pot_{mi} + \mu\dep_{mj} > 0 \).
-So \( \Wm\pot_{mi} > 0 \) would imply that \( \Wm\dep_{mj} = 0 \) and vice versa.
-These two possibilities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_left},\ref{fig:triangle_right}).
-
-Let us arrange the states left to right in order of decreasing \(\etw\), so that \(i\) is to the left of \(j\).
-For the unconstrained transitions \( \Wm\pot_{mj} \) and \( \Wm\dep_{mi} \) the potentiating transition goes to the \emph{right} of the depressing transition.
-For the other transitions \( \Wm\pot_{mi} \) and \( \Wm\dep_{mj} \) the potentiating transition goes to the \emph{left} of the depressing transition, and they cannot both be nonzero.
-
-This logic can be applied to any pair of states with different values of \(\etw\).
-Therefore, when starting from the same state, no potentiating transition can go to the left of any depressing transition, and no depressing transition can go to the right of any potentiating transition.
-
-For any state \(m\), there is a state \(n\) such that all potentiating transitions go to equal or smaller values of \(\etw\) all depressing transitions go to equal or greater values.
-%
-\begin{equation}\label{eq:separator}
-  \forall\ m, \;\; \exists\ n : 
-  \begin{cases}
-    \etwm_i > \etwm_n &\implies \Wm\pot_{mi} = 0, \\
-    \etwm_i < \etwm_n &\implies \Wm\dep_{mi} = 0.
-  \end{cases}
-\end{equation}
-%
-The remaining allowed transitions from \(m\) are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_allowed}).
-Note that we have not yet ruled the possibility that both \( \Wm\pot_{mn} \) and \( \Wm\dep_{mn} \) are nonzero.
-If one of those transitions is zero, there will be two states with the property above: the targets of right-most depressing and left-most potentiating transitions.
-Note that we have not yet proved anything about where \(\etwm_m\) lies with respect to any of the others.
-
-Consider the perturbation
-%
-\begin{equation}\label{eq:pertK}
-  \pdiff{\CL}{\Wm\pot_{mn}} -
-  \pdiff{\CL}{\Wm\dep_{mn}} =
-    2 \eqm_m (\etwm_m - \etwm_n) 
-    + \mu\pot_{mn} - \mu\pot_{mm} - \mu\dep_{mn} + \mu\dep_{mm}.
-\end{equation}
-%
-If we assume that both \( \Wm\pot_{mn} \) and \( \Wm\dep_{mn} \) are nonzero, we find that \( \mu\pot_{mm} - \mu\dep_{mm} = 2 \eqm_m (\etwm_m - \etwm_n) \).
-Therefore \( \etwm_m > \etwm_n \implies \Wm\pot_{mm} = -rf\pot \) and \( \etwm_m < \etwm_n \implies \Wm\dep_{mm} = -rf\dep \).
-It is also possible that both of the diagonal constraints are saturated.
-
-\begin{figure}[ht]
-\begin{center}
-\begin{myenuma}
-  \item \label{fig:triangle_pert} \aligntop{%
-  \includegraphics[width=0.28\linewidth]{pert_triangle.svg}}
-  \item \label{fig:triangle_left} \aligntop{%
-  \includegraphics[width=0.28\linewidth]{left_triangle.svg}}
-  \item \label{fig:triangle_right} \aligntop{%
-  \includegraphics[width=0.28\linewidth]{right_triangle.svg}}
-  \\[1em]
-  \item \label{fig:triangle_allowed} \aligntop{%
-  \includegraphics[width=0.4\linewidth]{allowed_triangle.svg}}
-  \item \label{fig:triangle_spacing} \aligntop{%
-  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
-\end{myenuma}
-\caption[Triangularising perturbation]{Triangularising perturbation.
-(\ref*{fig:triangle_pert}) The type of perturbation considered in \cref{eq:pertKupdn}, indicating which transitions have their rates increased/decreased.
-(\ref*{fig:triangle_left},\ref*{fig:triangle_right}) Possible end-points of this perturbation. 
-(\ref*{fig:triangle_allowed}) The not-as-yet ruled out transitions when states are sorted so that \(\etw\) decreases to the right, with \(\etwm_m\) unconstrained (as yet).
-(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, with \(i\) a state with a nonzero depressing transition from \(m\), \(j\) a state with a nonzero potentiating transition, \(n\) a state with both depressing and potentiating transitions and \(k\) a state with no nonzero transitions from \(m\).
-The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
-\label{fig:triangle_s}}
-\end{center}
-\end{figure}
-
-%================================================================================
-
-\subsubsection{Number of exits}\label{sec:exit}
-
-Let us define the following
-%
-\begin{equation}\label{eq:degenerate}
-\begin{aligned}
-    \dgn_{m}(s) &= \thb(s) + c_m \etw(s) + \lambda \etw(0), \\
-    \qquad
-    \dgn^\nu_m(s) &= \dgn_m(s) + (-1)^\nu \etw(s),   
-\end{aligned}
-  \implies \quad
-  \pdiff{\CL}{\Wm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
-\end{equation}
-%
-They should be thought of as collections of vectors rather than matrices, with the first \(m\) a label rather than an index.
-
-Suppose that a state \(n\) has a nonzero transition rate from state \(m\) for one \(\nu\).
-Then \(\mu^\nu_{mn} = 0\) and:
-%
-\begin{equation}\label{eq:oneexit}
-\begin{aligned}
-  \Wm^\nu_{mn} > 0
-  \quad \implies \quad
-  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n})
-  \quad \implies \quad
-  \dgnm^\nu_{m,m} &\geq \dgnm^\nu_{m,n} \\
-  \text{or} \quad
-  \dgnm_{m,m} - \dgnm_{m,n} &\geq - (-1)^\nu (\etwm_m - \etwm_n).
-\end{aligned}
-\end{equation}
-%
-If both \(\nu\) have nonzero rates from \(m\) to \(n\):
-%
-\begin{equation}\label{eq:doubleexit}
-  \Wm\pot_{mn}, \Wm\dep_{mn} > 0
-  \qquad \implies \qquad
-  \dgnm_{m,m} - \dgnm_{m,n} \geq \abs{\etwm_m - \etwm_n}.
-\end{equation}
-%
-
-Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
-Both must statisfy \cref{eq:oneexit} for the same \(\mu^\nu_{mm}\), so
-%
-\begin{equation}\label{eq:twoexits}
-  \Wm^\nu_{mi}, \Wm^\nu_{mj} > 0
-  \qquad \implies \qquad
-  \dgnm^\mu_{m,i} = \dgnm^\nu_{m,j}
-  \quad \text{or}\quad
-  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
-\end{equation}
-%
-Furthermore, for another state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
-%
-\begin{equation}\label{eq:nonexit}
-\begin{aligned}
-  \Wm^\nu_{mi} > 0, \Wm^\nu_{mk} = 0 
-  \;\;\; \implies \;\;\;
-  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
-  \;\; \implies \;\;
-  \dgnm^\nu_{m,k} &\geq \dgnm^\nu_{m,i} \\
-  \text{or}\quad
-  \dgnm_{m,k} - \dgnm_{m,i} &\geq - (-1)^\nu (\etwm_k - \etwm_i).
-\end{aligned}
-\end{equation}
-%
-If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
-%
-\begin{equation}\label{eq:twodoubleexits}
-  \Wm\pot_{mi}, \Wm\dep_{mi}, \Wm\pot_{mj}, \Wm\dep_{mj} > 0
-  \qquad \implies \qquad
-  \dgnm_{m,i} = \dgnm_{m,j} 
-  \quad \text{and} \quad
-  \etwm_i = \etwm_j.
-\end{equation}
-%
-The consequences of these inequalities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}).
-We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
-Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.
-
-We use \(m*\) to label the intersection of the two diagonal lines in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}), wether or not there is a state there.
-It can be defined for each \(m\), provided that it has at least one outgoing transition for each plasticity type, via
-%
-\begin{equation*}
-\begin{lgathered}  
-  \dgnm\pot_{m*} = \dgnm\pot_{m,j} = \min_k \dgnm\pot_{m,k},
-  \qquad
-  \dgnm\dep_{m*} = \dgnm\dep_{m,i} = \min_k \dgnm\dep_{m,k},
-  \\
-  \text{and also} \quad
-  \dgnm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} + \dgnm\dep_{m*}},
-  \quad
-  \etwm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} - \dgnm\dep_{m*}},
-  \quad
-  \thbm_{m*} = \dgnm_{m*} - c_m \etwm_{m*}.
-\end{lgathered}
-\end{equation*}
-%
-As shown in \cref{eq:twoexits}, this produces the same point regardless of which states \(i\) and \(j\) we use, 
-so long as there is a nonzero depressing/potentiating transition from \(m\) to that state.
-The KKT multipliers are related to these quantities by
-%
-\begin{equation}\label{eq:KKTz}
-  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m*}),
-  \qquad
-  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m*}).
-\end{equation}
-%
-
-Both of \cref{eq:twoexits,eq:twodoubleexits} require degeneracy -- 
-those two situations would require \(\dgn^\nu_m\) to have degenerate minima.
-The physicist's intuition states that degeneracy is fragile and requires a symmetry to protect it.
-Such a symmetry would imply that the states are lumpable.
-This would imply that, after lumping all lumpable states together, the situation in \cref{eq:twoexits} is impossible and \cref{eq:twodoubleexits} is even more impossible.
-If this intuition holds, we expect that \cref{fig:triangle_s}(\ref{fig:triangle_spacing}) would have at most one state on each of the two diagonal lines.
-Either there would be one state like \(n\), or there would be one state like \(i\) and one state like \(j\), and the first of these possibilities is rather implausible.
-
-%-------Section------------------------------------------------------------------
-
-\subsubsection{Diagonal constraints}\label{sec:diagKKT}
-
-We can write \cref{eq:diagKKT} as
-%
-\begin{equation*}
-\begin{aligned}
-  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
-    \left[ 
-      \hat{A}
-      - (c_m + (-1)^\nu) (\delta \wm_m + s \delta \etwm_m)
-      - \lambda \delta w_m
-    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
-    % \left.
-      - (-1)^\nu [\encm \dgnm_m]_m
-    \right],
-\end{aligned}
-\end{equation*}
-%
-with all quantities evaluated at \(s\) unless labelled otherwise.
-
-From \cref{eq:oneexit,eq:twoexits} we see that
-%
-\begin{equation*}
-\begin{gathered}
-  \begin{aligned}
-    \W^\nu_{mi} &> 0 &
-    \quad &\implies \quad &
-    \dgnm^\nu_{m,i} &= \dgnm^\nu_{m*}
-    \\
-    \sum_{k \neq m} \W^\nu_{mk} &< r f^\nu &
-    &\implies &
-    \dgnm^\nu_{m,m} &= \dgnm^\nu_{m*},
-  \end{aligned}
-  \\
-  \begin{aligned}
-    \implies \quad
-    [\W^\nu \dgn_m]_m
-      &= \sum_k \Wm^\nu_{mk}  \dgnm_{m,k}
-      = \sum_k \Wm^\nu_{mk}
-        (\dgnm^\nu_{m,k} - (-1)^\nu \etwm_k) \\
-      &= \sum_{k \neq m} \Wm^\nu_{mk} 
-        (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,m}) 
-        - (-1)^\nu [\W^\nu \etw]_m \\
-      &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
-        \sum_{k \neq m} \Wm^\nu_{mk}
-        - (-1)^\nu [\W^\nu \etw]_m \\
-      &= r f^\nu (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
-        - (-1)^\nu [\W^\nu \etw]_m \\
-      &= r f^\nu (\dgnm_{m*} - \dgnm_{m,m}) 
-        + (-1)^\nu \prn{
-          r f^\nu (\etwm_{m*} - \etwm_m)
-          - [\W^\nu \etw]_m
-        },
-  \end{aligned}
-\end{gathered}
-\end{equation*}
-%
-where we used the fact that \( (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) \) is only nonzero when \( \sum_{k \neq m} \Wm^\nu_{mk} = rf^\nu \).
-Therefore
-%
-\begin{equation*}
-\begin{aligned}
-  [\enc \dgn_m]_m
-    &= \sum_{\nu} (-1)^\nu rf^\nu 
-      (\dgnm_{m*} - \dgnm_{m,m}) 
-      + \sum_{\nu} \prn{
-        r f^\nu (\etwm_{m*} - \etwm_m)
-        -  [\W^\nu \etw]_m
-      } \\
-    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
-      + r(\etwm_{m*} - \etwm_m)
-      - [\frg \etw]_m \\
-    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
-      + r(\etwm_{m*} - \etwm_m)
-      - (\delta w_m + s \delta \etwm_m).
-\end{aligned}
-\end{equation*}
-%
-Then we can also write \cref{eq:diagKKT} as
-%
-\begin{equation}\label{eq:diagKKTz}
-\begin{aligned}
-  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
-    \left[ 
-      \hat{A}
-      - c_m (\delta \wm_m + s \delta \etwm_m)
-      - \lambda \delta w_m
-    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
-    % \left.
-      + (-1)^\nu \prn{ 
-        \Delta f (\dgnm_{m,m} - \dgnm_{m*}) 
-        + (\etwm_m - \etwm_{m*})
-      }
-    \right],
-\end{aligned}
-\end{equation}
-%
-again with all quantities evaluated at \(s\) unless labelled otherwise.
-
-\Cref{eq:KKTz} provides another expression for \(\mu^\nu_{mm}\).
-That expression will agree with \cref{eq:diagKKTz} provided that
-%
-\begin{equation}\label{eq:diagAgree}
-    \hat{A} - c_m (\delta \wm_m + s \delta \etwm_m)
-      - \lambda \delta w_m 
-      = (\dgnm_{m,m} - \dgnm_{m*}) 
-        + \Delta f (\etwm_m - \etwm_{m*}).
-\end{equation}
-%
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Scaling with a floating cutoff}\label{sec:cutoff}
-
-%-------Section------------------------------------------------------------------
-
 \section{Upper bounds}\label{sec:upperbnds}
@@
 We will confine such discussion to the footnotes.
-We will also work in units where \(r = 1\) until the very end.
 
@@
   \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) 
-      = A(0) = \sqrt{N} \CA = \sqrt{N} \sum_a \CI_a \tau_a.
+      = A(0) = \frac{\sqrt{N}}{r} \CA = \sqrt{N} \sum_a \CI_a \tau_a.
 \end{equation}
@@
         +\sum_{\nu i j} \prn{ \mu^\nu_{ij} \MMdm^\nu_{ij} }
-        &+ \lambda \prn{ \Delta\pr - \sum_i \eqm_i \wm_i },
-\end{aligned}
-\end{equation}
-%
-where \(\Delta\pr\) is the constant value we are holding \(\eq\w\) at.
+        &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
+\end{aligned}
+\end{equation}
+%
+where \(\Delta p\) is the constant value we are holding \(\eq\w\) at.
 The Karush-Kuhn-Tucker necessary conditions for a maximum are
@@
   \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
-    f^\nu \eqm_m \prn{ \brk{\thbm_m(0) - \thbm_n(0)}
-     + \prn{\mathbf{c}_m(0) + (-1)^\nu} \brk{\etwm_m(0) - \etwm_n(0)} } \\
+    r f^\nu \eqm_m \prn{ \brk{\thbm_m(0) - \thbm_n(0)}
+     + \prn{c_m(0) + (-1)^\nu} \brk{\etwm_m(0) - \etwm_n(0)} } \\
      &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     + \lambda f^\nu \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
+     + \lambda r f^\nu \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
     = 0.
@@
 
-The area and \(\Delta\pr\) are invariant under the scale transformation %\cref{eq:scale}
+The area and \(\Delta p\) are invariant under the scale transformation %\cref{eq:scale}
 %
@@
 \begin{equation}\label{eq:shiftqderiv}
-  \pdiff{\CL}{\enc_{mn}} =
-  \frac{1}{2f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} - \frac{1}{2f\pot} \pdiff{\CL}{\MMdm\dep_{mn}}
-   = \frac{1}{2}(\mu\pot_{mn} - \mu\dep_{mn}) + \eqm_m (\etwm_m - \etwm_n)
+  2r\pdiff{\CL}{\encm_{mn}} =
+  \frac{1}{f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} - \frac{1}{f\pot} \pdiff{\CL}{\MMdm\dep_{mn}}
+   = \mu\pot_{mn} - \mu\dep_{mn} + 2\eqm_m (\etwm_m - \etwm_n)
    = 0.
@@
 \begin{equation}\label{eq:shiftq}
-  \MMdm\pot_{ij} \to \MMdm\pot_{ij} + \frac{\epsilon_{ij}}{2f\pot},
+  \MMdm\pot_{ij} \to \MMdm\pot_{ij} + \frac{\epsilon_{ij}}{f\pot},
   \qquad
-  \MMdm\dep_{ij} \to \MMdm\dep_{ij} - \frac{\epsilon_{ij}}{2f\dep},
+  \MMdm\dep_{ij} \to \MMdm\dep_{ij} - \frac{\epsilon_{ij}}{f\dep},
   \qquad
@@
 
-Now consider the following combinations of derivatives for $(n - m)(-1)^\nu > 1$:
+Now consider the following combinations of derivatives for $(-1)^\nu (n - m) > 1$:
 %
@@
 \begin{equation*}
-  \CI(\mathbf{u}) = \frac{(\eq \enc \mathbf{u}) (\mathbf{u}^\dag \w)}{\mathbf{u}^\dag \mathbf{u}}
+  \CI(\mathbf{u}) = \frac{(\eq \enc \mathbf{u}) (\mathbf{u}^\dag \w)}{r\mathbf{u}^\dag \mathbf{u}}
       = \frac{\sum_{i \neq j} \eqm_i \encm_{ij} (u_j - u_i) \sum_k \eqm_k u_k \wm_k} 
-             {\sum_{l} \eqm_l u_l^2} .
+             {r\sum_{l} \eqm_l u_l^2} .
 \end{equation*}
@@
   \CI(\mathbf{u}) 
-      &\leq \frac{\av{\mathbf{u},\w}}{\nrm{\mathbf{u}}^2} 
+      &\leq \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
         \intd[_{-\infty}^{\infty}]{\alpha}  \brk{
         \Fm_{\CA_\alpha \CA_\alpha^c} + \Fm_{\CA_\alpha^c \CA_\alpha} } 
-  \\  &= \frac{\av{\mathbf{u},\w}}{\nrm{\mathbf{u}}^2} 
+  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
         \intd[_{-\infty}^{\infty}]{\alpha} \sum_{i \neq j} \Fm_{ij} \brk{
@@
         }
-  \\  &= \frac{\av{\mathbf{u},\w}}{\nrm{\mathbf{u}}^2} 
+  \\  &= \frac{\av{\mathbf{u},\w}}{r\nrm{\mathbf{u}}^2} 
       \sum_{i \neq j} \Fm_{ij} \abs{u_j - u_i}
   \\  &= \frac{\av{\mathbf{u},\w} \aav{\abs{\boldsymbol{\Delta}\mathbf{u}}, \onem}}
-              {\nrm{\mathbf{u}}^2}.
+              {r\nrm{\mathbf{u}}^2}.
 \end{aligned}
@@
   \CI(\mathbf{u}) \leq \frac{\nrm{\w} \nnrm{\boldsymbol{\Delta}\mathbf{u}} \nnrm{\onem}}
-                            {\nrm{\mathbf{u}}} .
+                            {r\nrm{\mathbf{u}}} .
 \end{equation*}
@@
 We also have \( \nrm{\w}^2 = \sum_i \eqm_i \wm_i^2 = 1 \).
-Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \MMm_{ij} \leq r \sum_i \eqm_i = r\) (which is 1 in our units!).
+Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \MMm_{ij} \leq r \sum_i \eqm_i = r\).
 Then, remembering \cref{eq:numrange},
@@
 \begin{equation}\label{eq:sqrtnumrange}
-  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\MM \mathbf{u}}}{\av{\mathbf{u}, \mathbf{u}}}}
-      = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})}.
+  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\MM \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
+      = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})/r}.
 \end{equation}
@@
 %
-Numerical maximisation of the quantity above produces the same result as in \cref{sec:initial}:
+bNumerical maximisation of the quantity above produces the same result as in \cref{sec:initial}:
 the two-state model with deterministic transitions.
@@
   \av{\mathbf{u},\w} &\leq \nrm{\mathbf{u}} \nrm{\w} , &
-  \sum_{j: j \neq i} \frg_{ij} &\leq 1.
+  \sum_{j: j \neq i} \frg_{ij} &\leq r.
 \end{aligned}
@@
 \begin{equation*}
-  \CI_a = (\eq \enc \evr^a) (\evl^a \w),
+  \CI_a = \frac{(\eq \enc \evr^a) (\evl^a \w)}{r},
   \qquad
@@
 %
+In this section we will work in units where \(r = 1\) until the very end.
+
 From \eqref{eq:areabnd} and \eqref{eq:initbnd}, we see that they are subject to the constraints
@@
 
+\section{Finite time topology}\label{sec:finite}
+
+Now we'll maximise \(\snrb(\tau)\) for some fixed \(\tau\), subject to the constraints \eqref{eq:constr}.
+First, we will maximise the numerator of \eqref{eq:lsnr} holding \(\eq\w\) fixed.
+We can maximise the ratio \wrt \(\eq\w\) later.
+% From this point on, we work in units where \(r = 1\).
+
+For this section we will treatthe off-diagonal components of \(\W^\nu\) as the independent degrees of freedom, as in \cref{eq:wdof}, rather than the \(\MMd^\nu\).
+Consider the Lagrangian
+%
+\begin{equation}\label{eq:finitelagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
+        +\sum_{\nu i j} \mu^\nu_{ij} \prn{ \Wm^\nu_{ij} + r f^\nu \delta_{ij} }
+        + \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
+\end{aligned}
+\end{equation}
+%
+where \(\Delta p\) is the constant value we are holding \(\eq\w\) at.
+The Karush-Kuhn-Tucker necessary conditions for a maximum are
+%
+\begin{equation}\label{eq:KTfinite}
+  \pdiff{\CL}{\Wm^\nu_{ij}} = 0, \quad
+  \mu^\nu_{ij} \geq 0, \quad
+  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
+  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
+  \pdiff{\CL}{\lambda} = 0.
+\end{equation}
+%
+% We have scaled the KKT multipliers by \(r f^\nu\) for later convenience.
+
+The derivatives, for \(m \neq n\) are
+%
+\begin{equation}\label{eq:finitederivs}
+\begin{aligned}
+  \pdiff{\CL}{\Wm^\nu_{mn}} = & \,
+    \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
+     + \prn{c_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
+     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
+     + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
+    = 0.
+\end{aligned}
+\end{equation}
+%
+We can solve for the KKT multipliers by noting that
+%
+\begin{equation*}
+  \mu^\nu_{ij} \Wm^\nu_{ij} = 0,
+  \qquad \text{and} \qquad
+  \sum_{j: j \neq i} \mu^\nu_{ii} \Wm^\nu_{ij} 
+      = \mu^\nu_{ii} ( - \Wm^\nu_{ii}) = r f^\nu \mu^\nu_{ii}.
+\end{equation*}
+%
+Then, with some new vectors defined:
+%
+\begin{equation}\label{eq:kktsol}
+\begin{aligned}
+  \mu^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
+    + \lambda \ombm^\nu_m(0) 
+    + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
+    }, \\
+  \mu^\nu_{mn} &= \eqm_m \big( 
+      \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
+    + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
+  \\ &\hphantom{= \eqm_m \big( }
+    + \brk{{c}_m(s) + (-1)^\nu} 
+        \brk{ \etwm_n(s) - \etwm_m(s) - \ombm^\nu_m(s) }
+     \big), \\
+  \text{where} \quad 
+  \alb^\nu(s) &= \frac{\W^\nu \thb(s)}{rf^\nu}, \qquad
+  \omb^\nu(s) = \frac{\W^\nu \etw(s)}{rf^\nu}.
+\end{aligned}
+\end{equation}
+%
+If we note that we can write \( \W^\nu = \prn{\frg + (-1)^\nu \enc} / 2 \), we find that
+%
+\begin{equation}\label{eq:diagKKT}
+\begin{aligned}
+  \alb^\nu(s) &= \frac{\delta\mathbf{a}(s) + (-1)^\nu \enc \thb(s)}
+                     {2rf^\nu},
+  \qquad 
+  \omb^\nu(s) = \frac{\delta\w + s \delta\etw(s) - (-1)^\nu \mathbf{a}(s)}
+                     {2rf^\nu}, 
+  \\
+  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+    \left[ 
+      \hat{A}(s) - c_m(s) (\delta\wm_m + s \delta\etwm_m(s))
+      - \lambda \delta w_m
+    \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} \big[}
+    \left. {}
+      + (-1)^\nu \prn{ 
+        c_m(s) a_m(s) - [\encm \thbm(s)]_m
+        - \delta\wm_m - s \delta\etwm_m(s)
+        + \lambda a_m(0)
+      }
+    \right], 
+\end{aligned}
+\end{equation}
+%
+where \(\delta\) indicates mean subtraction and \(\eq \mathbf{a}(s) = \hat{A}(s)\).
+
+The expressions in \cref{eq:finitederivs} have the same form as for \(s=0\), which allowed us to argue that the model with maximal area must have the serial topology in \cref{sec:areapert} and \cite{Lahiri2013synapse}.
+The one missing ingredient is the scale invariance.
+
+However, we do have numerical evidence that the model that maximises the Laplace transform does have the serial topology.
+In \cref{fig:envnum} we see the result of numerical maximisation of \cref{eq:finitelagrangian}.
+We see that restricting to the serial topology does not make the performance any worse.
+In fact, it slightly improves the performance, but this can be ascribed to the numerical optimisation problem being easier.
+Assuming that the optimiser did find the global maximum, this provides evidence for our conjecture that the model that maximises the Laplace transform must have the serial topology.
+The models found numerically are described in \cref{fig:envvid}.
+
+\begin{figure}[tb]
+  \centering
+  % Requires \usepackage{graphicx}
+  \includegraphics[width=0.8\linewidth]{LenvNumOnly.pdf}\\
+  \caption[Numerical envelope for the signal-to-noise ratio]
+  {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
+  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numerically with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
+  The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
+  The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
+  }\label{fig:envnum}
+\end{figure}
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Scale invariance}\label{sec:scale}
+
+% Under the scale transformation
+% %
+% \begin{equation}\label{eq:scale}
+%   \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I ,
+% \end{equation}
+% %
+The Laplace transform has the following scaling property
+%
+\begin{equation}\label{eq:laplacescale}
+  \hat{A} \prn{\zeta s;\zeta \W^\nu} = \hat{A} \prn{s; \W^\nu}.
+\end{equation}
+%
+At \(s = 0\), it is scale invariant.
+This means that in \cite{Lahiri2013synapse} we could take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \cref{eq:constri}) and use the scale transform to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
+This allowed us to ignore the diagonal constraints.
+We can't do that for \(s > 0\).
+
+The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
+%
+\begin{equation}\label{eq:shiftscale}
+  \zeta {\circledast}_{s} \W^\nu
+      \equiv \zeta \W^\nu
+      + (1 - \zeta) s f^\nu \brk{\I - \onev \eq} .
+\end{equation}
+%
+Then we have
+%
+\begin{equation}\label{eq:lshiftscale}
+  \hat{A} \prn{s; \W^\nu} 
+    = \hat{A} \prn{0, \shift{\W}^\nu(s)}
+    = \hat{A} \prn{0, \zeta \shift{\W}^\nu(s)}
+    = \hat{A} \prn{s, \zeta \circledast_s \W^\nu}.
+\end{equation}
+%
+The problem here is that the lower bound, \(\Wm^\nu_{ij} \geq 0\), 
+is not invariant under this transformation, unlike the \(s = 0\) case.
+Using this transformation to fix the upper bound can break the lower bound.
+
+We might ask what conditions are required so that a model is fixable?
+This means that the \(\zeta \circledast_s \) orbit of the model contains some valid models.
+The conditions are
+%
+\begin{equation*}
+\begin{alignedat}{5}
+  [\zeta \circledast \W^\nu]_{ij}  
+    &= \zeta \Wm^\nu_{ij} - (1-\zeta) s f^\nu \eqm_j 
+    &&\geq 0 &
+  \;\; &\implies \;\; &
+  \zeta &\geq \frac{s \eqm_j}{\Wm^\nu_{ij} /  f^\nu + s \eqm_j} &
+    &\forall \nu, i, j : i \neq j , \\
+  \sum_{j : i \neq j} [\zeta \circledast \W^\nu]_{ij} 
+    &= \zeta \lbm^\nu_i - (1-\zeta) s f^\nu (1-\eqm_i) 
+    &&\leq r f^\nu &
+  &\implies &
+  \zeta &\leq \frac{r + s (1-\eqm_i)}{\lbm^\nu_i / f^\nu + s (1-\eqm_i)} &
+    \quad &\forall \nu, i ,
+\end{alignedat}
+\end{equation*}
+%
+where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \Wm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
+The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
+%
+\begin{equation}\label{eq:fixable}
+  \frac{s \eqm_j}{\Wm^\nu_{ij} / f^\nu + s \eqm_j} \leq \frac{r + s (1-\eqm_k)}{\lbm^\mu_k / f^\nu + s (1-\eqm_k)}
+  \quad \implies \quad
+  \frac{\lbm^\mu_k - r f^\nu}{r + s (1-\eqm_k)} \leq \frac{\Wm^\nu_{ij}}{s \eqm_j} 
+  \quad \forall \mu, \nu, i, j, k : i \neq j.
+\end{equation}
+%
+Note that when \(\Wm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq r f^\nu\), the original diagonal constraint.
+Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Shifted problem}\label{sec:lshift}
+
+The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
+treat the off-diagonal components of \(\shift{\W}^\nu \) as the degrees of freedom and maximise the area instead.
+The constraints on \(\shift{\Wm}^\nu_{ij}\) are different from \cref{eq:constri}:
+%
+\begin{equation}\label{eq:lshiftconstr}
+  \shift{\Wm}^\nu_{ij} \geq s f^\nu \eqm_j,
+    \quad \forall \; \nu, i, j : i \neq j, \qquad
+  \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \leq r f^\nu + s f^\nu (1 - \eqm_i).
+    \quad \forall \; \nu, i.
+\end{equation}
+%
+The diagonal elements are still given by \( \shift{\Wm}^\nu_{ii} = - \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \), but now they can be below \(-r f^\nu \) so long as \(\shift{\Wm}^\nu_{ii} \geq -r f^\nu - s f^\nu (1 - \eqm_i)\).
+
+The Lagrangian we should consider is
+%
+\begin{equation}\label{eq:lshiftlagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
+        +\sum_{\nu i j} \mu^\nu_{ij}\prn{\Wm^\nu_{ij} + f^\nu (r + s) \delta_{ij} -  s f^\nu \eqm_j }
+        &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
+\end{aligned}
+\end{equation}
+%
+where all of the first-passage times are evaluated at \(s = 0\).
+
+In \cref{fig:shifted} we show the result of solving this problem with numerical optimisation.
+We see essentially the same results as the original problem, \cref{eq:finitelagrangian,fig:envnum}, as expected.
+
+The derivatives, for \(m \neq n\) are
+%
+\begin{equation}\label{eq:lshiftderivs}
+\begin{aligned}
+  \pdiff{\CL}{\Wm^\nu_{mn}} = & \,
+    \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
+     + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
+     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
+     - \sum_{\mu ij} \mu^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
+    = 0.
+\end{aligned}
+\end{equation}
+%
+We can absorb the new terms by defining
+%
+\begin{equation*}
+\begin{aligned}
+  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
+  \qquad
+  \shift{\alb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
+  \\
+  \shift{\omb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \etw(0).
+\end{aligned}
+\end{equation*}
+%
+We note that
+%
+\begin{equation*}
+  \mu^\nu_{ij} (\Wm^\nu_{ij} - s \eqm_j) = 0,
+  \qquad \text{and} \qquad
+  \sum_{j: j \neq i} \mu^\nu_{ii} (\Wm^\nu_{ij} - s \eqm_j)
+      = \mu^\nu_{ii} (1 - \Wm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
+\end{equation*}
+%
+then we can ``solve'' for the KKT multipliers:
+%
+\begin{equation}\label{eq:lshiftsol}
+\begin{aligned}
+  \mu^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
+    }, \\
+  \mu^\nu_{mn} &= \eqm_m \prn{ 
+      \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
+  % \\ & \hphantom{= - \eqm_m \big( } 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } 
+        \brk{ \etwm_n - \etwm_m - \shift{\ombm}^\nu_m(s) }
+     }, \\
+\end{aligned}
+\end{equation}
+%
+although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\), so these expressions are not explicit.
+
+While this problem involves the scale invariant quantity \(\hat{A}(0)\), neither the upper nor lower bounds in \cref{eq:lshiftconstr} are invariant.
+We then run into the same problems in proving our serial-topology hypothesis.
+
+\begin{figure}[ht]
+\begin{center}
+\begin{myenuma}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{LenvShift.pdf}}
+  \label{fig:envshift}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{shift_cond.pdf}}
+  \label{fig:shiftcond}
+\end{myenuma}
+\caption[Shifted problem vs.\ original]{Shifted problem vs.\ original.
+(\ref*{fig:envshift}) Comparison of numerical optimisation of the shifted problem in \cref{eq:lshiftlagrangian,fig:envnum} with the original problemin \cref{eq:finitelagrangian}.
+We see that the two approaches produce very similar numerical envelopes.
+The shifted problem involves non-linear constraints and is thus harder to solve.
+(\ref{fig:shiftcond}) Plot of the discrepancy in (\ref*{fig:envshift}) against the condition number of the fundamental matrix, \(\fund \).
+We see that most of the large discrepancies occur where the problem is numerically ill-conditioned.
+\label{fig:shifted}}
+\end{center}
+\end{figure}
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Topology constraining perturbations}\label{sec:topology}
+
+In this section we will see how far we can get in constraining the topology, 
+despite the lack of scale invariance.
+
+%-------Section------------------------------------------------------------------
+
+\subsubsection{Triangularity}\label{sec:parttriangular}
+
+Suppose we have two states, \(i\) and \(j\), with \( \etwm_i > \etwm_j \).
+Consider the following perturbation (see also \cref{fig:triangle_s}(\ref{fig:triangle_pert})):
+%
+\begin{equation}\label{eq:pertKupdn}
+\begin{aligned}
+  2 \brk{\pdiff{\CL}{\encm_{mj}} - \pdiff{\CL}{\encm_{mi}}}
+  &= \pdiff{\CL}{\Wm\pot_{mj}} - \pdiff{\CL}{\Wm\pot_{mi}} -
+  \pdiff{\CL}{\Wm\dep_{mj}} + \pdiff{\CL}{\Wm\dep_{mi}} \\ 
+    &= 2 \eqm_m (\etwm_i - \etwm_j) 
+    + \mu\pot_{mj} - \mu\pot_{mi} - \mu\dep_{mj} + \mu\dep_{mi}.
+\end{aligned}
+\end{equation}
+%
+Because \( \etwm_i > \etwm_j \), we must have \( \mu\pot_{mi} + \mu\dep_{mj} > 0 \).
+So \( \Wm\pot_{mi} > 0 \) would imply that \( \Wm\dep_{mj} = 0 \) and vice versa.
+These two possibilities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_left},\ref{fig:triangle_right}).
+
+Let us arrange the states left to right in order of decreasing \(\etw\), so that \(i\) is to the left of \(j\).
+For the unconstrained transitions \( \Wm\pot_{mj} \) and \( \Wm\dep_{mi} \) the potentiating transition goes to the \emph{right} of the depressing transition.
+For the other transitions \( \Wm\pot_{mi} \) and \( \Wm\dep_{mj} \) the potentiating transition goes to the \emph{left} of the depressing transition, and they cannot both be nonzero.
+
+This logic can be applied to any pair of states with different values of \(\etw\).
+Therefore, when starting from the same state, no potentiating transition can go to the left of any depressing transition, and no depressing transition can go to the right of any potentiating transition.
+
+For any state \(m\), there is a state \(n\) such that all potentiating transitions go to equal or smaller values of \(\etw\) all depressing transitions go to equal or greater values.
+%
+\begin{equation}\label{eq:separator}
+  \forall\ m, \;\; \exists\ n : 
+  \begin{cases}
+    \etwm_i > \etwm_n &\implies \Wm\pot_{mi} = 0, \\
+    \etwm_i < \etwm_n &\implies \Wm\dep_{mi} = 0.
+  \end{cases}
+\end{equation}
+%
+The remaining allowed transitions from \(m\) are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_allowed}).
+Note that we have not yet ruled the possibility that both \( \Wm\pot_{mn} \) and \( \Wm\dep_{mn} \) are nonzero.
+If one of those transitions is zero, there will be two states with the property above: the targets of right-most depressing and left-most potentiating transitions.
+Note that we have not yet proved anything about where \(\etwm_m\) lies with respect to any of the others.
+
+Consider the perturbation
+%
+\begin{equation}\label{eq:pertK}
+  \pdiff{\CL}{\Wm\pot_{mn}} -
+  \pdiff{\CL}{\Wm\dep_{mn}} =
+    2 \eqm_m (\etwm_m - \etwm_n) 
+    + \mu\pot_{mn} - \mu\pot_{mm} - \mu\dep_{mn} + \mu\dep_{mm}.
+\end{equation}
+%
+If we assume that both \( \Wm\pot_{mn} \) and \( \Wm\dep_{mn} \) are nonzero, we find that \( \mu\pot_{mm} - \mu\dep_{mm} = 2 \eqm_m (\etwm_m - \etwm_n) \).
+Therefore \( \etwm_m > \etwm_n \implies \Wm\pot_{mm} = -rf\pot \) and \( \etwm_m < \etwm_n \implies \Wm\dep_{mm} = -rf\dep \).
+It is also possible that both of the diagonal constraints are saturated.
+
+\begin{figure}[ht]
+\begin{center}
+\begin{myenuma}
+  \item \label{fig:triangle_pert} \aligntop{%
+  \includegraphics[width=0.28\linewidth]{pert_triangle.svg}}
+  \item \label{fig:triangle_left} \aligntop{%
+  \includegraphics[width=0.28\linewidth]{left_triangle.svg}}
+  \item \label{fig:triangle_right} \aligntop{%
+  \includegraphics[width=0.28\linewidth]{right_triangle.svg}}
+  \\[1em]
+  \item \label{fig:triangle_allowed} \aligntop{%
+  \includegraphics[width=0.4\linewidth]{allowed_triangle.svg}}
+  \item \label{fig:triangle_spacing} \aligntop{%
+  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
+\end{myenuma}
+\caption[Triangularising perturbation]{Triangularising perturbation.
+(\ref*{fig:triangle_pert}) The type of perturbation considered in \cref{eq:pertKupdn}, indicating which transitions have their rates increased/decreased.
+(\ref*{fig:triangle_left},\ref*{fig:triangle_right}) Possible end-points of this perturbation. 
+(\ref*{fig:triangle_allowed}) The not-as-yet ruled out transitions when states are sorted so that \(\etw\) decreases to the right, with \(\etwm_m\) unconstrained (as yet).
+(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, with \(i\) a state with a nonzero depressing transition from \(m\), \(j\) a state with a nonzero potentiating transition, \(n\) a state with both depressing and potentiating transitions and \(k\) a state with no nonzero transitions from \(m\).
+The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
+\label{fig:triangle_s}}
+\end{center}
+\end{figure}
+
+%================================================================================
+
+\subsubsection{Number of exits}\label{sec:exit}
+
+Let us define the following
+%
+\begin{equation}\label{eq:degenerate}
+\begin{aligned}
+    \dgn_{m}(s) &= \thb(s) + c_m \etw(s) + \lambda \etw(0), \\
+    \qquad
+    \dgn^\nu_m(s) &= \dgn_m(s) + (-1)^\nu \etw(s),   
+\end{aligned}
+  \implies \quad
+  \pdiff{\CL}{\Wm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
+\end{equation}
+%
+They should be thought of as collections of vectors rather than matrices, with the first \(m\) a label rather than an index.
+
+Suppose that a state \(n\) has a nonzero transition rate from state \(m\) for one \(\nu\).
+Then \(\mu^\nu_{mn} = 0\) and:
+%
+\begin{equation}\label{eq:oneexit}
+\begin{aligned}
+  \Wm^\nu_{mn} > 0
+  \quad \implies \quad
+  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n})
+  \quad \implies \quad
+  \dgnm^\nu_{m,m} &\geq \dgnm^\nu_{m,n} \\
+  \text{or} \quad
+  \dgnm_{m,m} - \dgnm_{m,n} &\geq - (-1)^\nu (\etwm_m - \etwm_n).
+\end{aligned}
+\end{equation}
+%
+If both \(\nu\) have nonzero rates from \(m\) to \(n\):
+%
+\begin{equation}\label{eq:doubleexit}
+  \Wm\pot_{mn}, \Wm\dep_{mn} > 0
+  \qquad \implies \qquad
+  \dgnm_{m,m} - \dgnm_{m,n} \geq \abs{\etwm_m - \etwm_n}.
+\end{equation}
+%
+
+Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
+Both must statisfy \cref{eq:oneexit} for the same \(\mu^\nu_{mm}\), so
+%
+\begin{equation}\label{eq:twoexits}
+  \Wm^\nu_{mi}, \Wm^\nu_{mj} > 0
+  \qquad \implies \qquad
+  \dgnm^\mu_{m,i} = \dgnm^\nu_{m,j}
+  \quad \text{or}\quad
+  \dgnm_{m,i} - \dgnm_{m,j} = - (-1)^\nu (\etwm_i - \etwm_j).
+\end{equation}
+%
+Furthermore, for another state \(k\) with zero transition rate from \(m\) for the same \(\nu\):
+%
+\begin{equation}\label{eq:nonexit}
+\begin{aligned}
+  \Wm^\nu_{mi} > 0, \Wm^\nu_{mk} = 0 
+  \;\;\; \implies \;\;\;
+  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
+  \;\; \implies \;\;
+  \dgnm^\nu_{m,k} &\geq \dgnm^\nu_{m,i} \\
+  \text{or}\quad
+  \dgnm_{m,k} - \dgnm_{m,i} &\geq - (-1)^\nu (\etwm_k - \etwm_i).
+\end{aligned}
+\end{equation}
+%
+If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
+%
+\begin{equation}\label{eq:twodoubleexits}
+  \Wm\pot_{mi}, \Wm\dep_{mi}, \Wm\pot_{mj}, \Wm\dep_{mj} > 0
+  \qquad \implies \qquad
+  \dgnm_{m,i} = \dgnm_{m,j} 
+  \quad \text{and} \quad
+  \etwm_i = \etwm_j.
+\end{equation}
+%
+The consequences of these inequalities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}).
+We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
+Even if there is no state \(n\) that is the target of both potentiating and depressing transitions, there will be two lines at \(45^\circ \) passing through any targets of potentiating/depressing transitions that demarcate the forbidden zone.
+
+We use \(m*\) to label the intersection of the two diagonal lines in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}), wether or not there is a state there.
+It can be defined for each \(m\), provided that it has at least one outgoing transition for each plasticity type, via
+%
+\begin{equation*}
+\begin{lgathered}  
+  \dgnm\pot_{m*} = \dgnm\pot_{m,j} = \min_k \dgnm\pot_{m,k},
+  \qquad
+  \dgnm\dep_{m*} = \dgnm\dep_{m,i} = \min_k \dgnm\dep_{m,k},
+  \\
+  \text{and also} \quad
+  \dgnm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} + \dgnm\dep_{m*}},
+  \quad
+  \etwm_{m*} = \frac{1}{2} \prn{\dgnm\pot_{m*} - \dgnm\dep_{m*}},
+  \quad
+  \thbm_{m*} = \dgnm_{m*} - c_m \etwm_{m*}.
+\end{lgathered}
+\end{equation*}
+%
+As shown in \cref{eq:twoexits}, this produces the same point regardless of which states \(i\) and \(j\) we use, 
+so long as there is a nonzero depressing/potentiating transition from \(m\) to that state.
+The KKT multipliers are related to these quantities by
+%
+\begin{equation}\label{eq:KKTz}
+  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m*}),
+  \qquad
+  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m*}).
+\end{equation}
+%
+
+Both of \cref{eq:twoexits,eq:twodoubleexits} require degeneracy -- 
+those two situations would require \(\dgn^\nu_m\) to have degenerate minima.
+The physicist's intuition states that degeneracy is fragile and requires a symmetry to protect it.
+Such a symmetry would imply that the states are lumpable.
+This would imply that, after lumping all lumpable states together, the situation in \cref{eq:twoexits} is impossible and \cref{eq:twodoubleexits} is even more impossible.
+If this intuition holds, we expect that \cref{fig:triangle_s}(\ref{fig:triangle_spacing}) would have at most one state on each of the two diagonal lines.
+Either there would be one state like \(n\), or there would be one state like \(i\) and one state like \(j\), and the first of these possibilities is rather implausible.
+
+%-------Section------------------------------------------------------------------
+
+\subsubsection{Diagonal constraints}\label{sec:diagKKT}
+
+We can write \cref{eq:diagKKT} as
+%
+\begin{equation*}
+\begin{aligned}
+  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+    \left[ 
+      \hat{A}
+      - (c_m + (-1)^\nu) (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m
+    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
+    % \left.
+      - (-1)^\nu [\encm \dgnm_m]_m
+    \right],
+\end{aligned}
+\end{equation*}
+%
+with all quantities evaluated at \(s\) unless labelled otherwise.
+
+From \cref{eq:oneexit,eq:twoexits} we see that
+%
+\begin{equation*}
+\begin{gathered}
+  \begin{aligned}
+    \W^\nu_{mi} &> 0 &
+    \quad &\implies \quad &
+    \dgnm^\nu_{m,i} &= \dgnm^\nu_{m*}
+    \\
+    \sum_{k \neq m} \W^\nu_{mk} &< r f^\nu &
+    &\implies &
+    \dgnm^\nu_{m,m} &= \dgnm^\nu_{m*},
+  \end{aligned}
+  \\
+  \begin{aligned}
+    \implies \quad
+    [\W^\nu \dgn_m]_m
+      &= \sum_k \Wm^\nu_{mk}  \dgnm_{m,k}
+      = \sum_k \Wm^\nu_{mk}
+        (\dgnm^\nu_{m,k} - (-1)^\nu \etwm_k) \\
+      &= \sum_{k \neq m} \Wm^\nu_{mk} 
+        (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,m}) 
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
+        \sum_{k \neq m} \Wm^\nu_{mk}
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= r f^\nu (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= r f^\nu (\dgnm_{m*} - \dgnm_{m,m}) 
+        + (-1)^\nu \prn{
+          r f^\nu (\etwm_{m*} - \etwm_m)
+          - [\W^\nu \etw]_m
+        },
+  \end{aligned}
+\end{gathered}
+\end{equation*}
+%
+where we used the fact that \( (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) \) is only nonzero when \( \sum_{k \neq m} \Wm^\nu_{mk} = rf^\nu \).
+Therefore
+%
+\begin{equation*}
+\begin{aligned}
+  [\enc \dgn_m]_m
+    &= \sum_{\nu} (-1)^\nu rf^\nu 
+      (\dgnm_{m*} - \dgnm_{m,m}) 
+      + \sum_{\nu} \prn{
+        r f^\nu (\etwm_{m*} - \etwm_m)
+        -  [\W^\nu \etw]_m
+      } \\
+    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + r(\etwm_{m*} - \etwm_m)
+      - [\frg \etw]_m \\
+    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + r(\etwm_{m*} - \etwm_m)
+      - (\delta w_m + s \delta \etwm_m).
+\end{aligned}
+\end{equation*}
+%
+Then we can also write \cref{eq:diagKKT} as
+%
+\begin{equation}\label{eq:diagKKTz}
+\begin{aligned}
+  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+    \left[ 
+      \hat{A}
+      - c_m (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m
+    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
+    % \left.
+      + (-1)^\nu \prn{ 
+        \Delta f (\dgnm_{m,m} - \dgnm_{m*}) 
+        + (\etwm_m - \etwm_{m*})
+      }
+    \right],
+\end{aligned}
+\end{equation}
+%
+again with all quantities evaluated at \(s\) unless labelled otherwise.
+
+\Cref{eq:KKTz} provides another expression for \(\mu^\nu_{mm}\).
+That expression will agree with \cref{eq:diagKKTz} provided that
+%
+\begin{equation}\label{eq:diagAgree}
+    \hat{A} - c_m (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m 
+      = (\dgnm_{m,m} - \dgnm_{m*}) 
+        + \Delta f (\etwm_m - \etwm_{m*}).
+\end{equation}
+%
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Scaling with a floating cutoff}\label{sec:cutoff}
+
+%-------Section------------------------------------------------------------------
+
 \section{Nearly uniform serial models}\label{sec:serial}
@@
   {Heuristic envelope  for the signal-to-noise ratio, \cref{eq:henv}.
-  Analytic expressions from \cref{eq:henvapprox} are only valid to the right of each region.}\label{fig:heuristicenv}
+  Analytic expressions from \cref{eq:henvapprox} describe the yellow curve and are only valid to the right of each region.}\label{fig:heuristicenv}
 \end{figure}

### 2021-09-06 Notes/laplace.tex
sign of hessian

@@
   &=
-    \eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\thbm_k(s) - \thbm_l(s)} \\
+    - \eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\thbm_k(s) - \thbm_l(s)} \\
   &&&\hphantom{=} {}
-    + \eqm_i \brc{{X}_{ik}(s) - {X}_{jk}(s)} 
+    - \eqm_i \brc{{X}_{ik}(s) - {X}_{jk}(s)} 
           \brc{\etwm_k(s) - \etwm_l(s)} \\
   &&& \hphantom{=} {}
-    + \eqm_i {c}_i(s) \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} 
+    - \eqm_i {c}_i(s) \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} 
           \brc{\etwm_k(s) - \etwm_l(s)},  \\
   S_{ijkl} &= \smash{\pdiff{\hat{A}}{\encm_{ij} ; \frgm_{kl}} } &
-  &= \eqm_k \brc{\fundm_{ki}(0) - \fundm_{li}(0)} \brc{\etwm_i(s) - \etwm_j(s)} \\
+  &= 
+    - \eqm_k \brc{\fundm_{ki}(0) - \fundm_{li}(0)} \brc{\etwm_i(s) - \etwm_j(s)} \\
   &&& \hphantom{=} {}
-    + \eqm_i \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} \brc{\etwm_k(s) - \etwm_l(s)},
+    - \eqm_i \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} \brc{\etwm_k(s) - \etwm_l(s)},
 \end{alignedat}
@@
 \begin{equation}\label{eq:etarecur}
-  -(s\I + \ onev \arow - \frg)\etw(s) = \w.
+  -(s\I + \onev \arow - \frg)\etw(s) = \w.
 \end{equation}

### 2021-09-06 Code/Python/complex_synapse/optimise/synapse_opt.py

@@ -261,12 +261,12 @@ class SynapseOptModel(_sm.SynapseModel, _sb.SynapseParam):
         rows, cols, mats = self._derivs(rate, **kwds)
 
         # (...,n,n,n,n)
-        hessww = (_outer3(rows[0], mats[0], cols[1])
-                  + _outer3(rows[0], mats[2], cols[0])
-                  + _outer3(rows[1], mats[1], cols[0])).sum(0)
+        hessww = -(_outer3(rows[0], mats[0], cols[1])
+                   + _outer3(rows[0], mats[2], cols[0])
+                   + _outer3(rows[1], mats[1], cols[0])).sum(0)
         # (2,...,n,n,n,n) - axis 0: h_wq, h_qw
-        hesswq = (_outer3(rows[0], mats[0], cols[0])
-                  + _trnsp4(_outer3(rows[0], mats[1], cols[0])))
+        hesswq = -(_outer3(rows[0], mats[0], cols[0])
+                   + _trnsp4(_outer3(rows[0], mats[1], cols[0])))
         # (...,1,n,n,1,n,n), (2,...,1,n,n,1,n,n)
         nax = (-6, -3)
         hessww, hesswq = hessww.expand_dims(nax), hesswq.expand_dims(nax)
@@ -314,13 +314,13 @@ class SynapseOptModel(_sm.SynapseModel, _sb.SynapseParam):
         # p Z theta V  -> (2,2,n,n) -> (n,n)
         # c Zs eta V   -> (2,2,n,n) -> (n,n)
         # p ZQZs eta V -> (2,2,n,n) -> (n,n)
-        h_ww = (_outer3p(rows[0], mats[0].inv, cols[1], vecm)
-                + _outer3p(rows[1], mats[1].inv, cols[0], vecm)
-                + _outer3p(rows[0], mats[2], cols[0], vecm)).sum((0, -3))
+        h_ww = -(_outer3p(rows[0], mats[0].inv, cols[1], vecm)
+                 + _outer3p(rows[1], mats[1].inv, cols[0], vecm)
+                 + _outer3p(rows[0], mats[2], cols[0], vecm)).sum((0, -3))
         # p Z eta V  -> (2,2,n,n) - h_wq, h_qw
         # p Zs eta V -> (2,2,n,n) - h_wq, h_qw
-        h_wqqw = (_outer3p(rows[0], mats[0].inv, cols[0], vecm)
[…]

### 2021-09-29 Notes/laplace.tex
started scaling argument

@@
 \newcommand{\evl}{\boldsymbol{\evlm}}
+%KKT multipliers
+\newcommand{\kktm}{\mu}
+\newcommand{\kkt}{\boldsymbol{\kktm}}
 %superscripts
@@
 \DeclareMathOperator{\env}{Env}
+\newcommand{\cnst}{\gamma}
 % \newcommand{\CA}{\mathcal{A}}
@@
   \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(0) - \etwm_j(0))
-        +\sum_{\nu i j} \prn{ \mu^\nu_{ij} \MMdm^\nu_{ij} }
+        +\sum_{\nu i j} \prn{ \kktm^\nu_{ij} \MMdm^\nu_{ij} }
         &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
@@
   \pdiff{\CL}{\MMdm^\nu_{ij}} = 0, \quad
-  \mu^\nu_{ij} \geq 0, \quad
-  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
-  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
+  \kktm^\nu_{ij} \geq 0, \quad
+  \pdiff{\CL}{\kktm^\nu_{ij}} \geq 0, \quad
+  \kktm^\nu_{ij}\pdiff{\CL}{\kktm^\nu_{ij}} = 0, \quad
   \pdiff{\CL}{\lambda} = 0.
@@
      + \prn{c_m(0) + (-1)^\nu} \brk{\etwm_m(0) - \etwm_n(0)} } \\
-     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
+     &+ \kktm^\nu_{mn} -  \kktm^\nu_{mm}
      + \lambda r f^\nu \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
@@
   \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} 
-    = - \sum_{\nu, m} \mu^\nu_{mm} = 0 .
+    = - \sum_{\nu, m} \kktm^\nu_{mm} = 0 .
 \end{equation*}
@@
   \frac{1}{f\pot} \pdiff{\CL}{\MMdm\pot_{mn}} - \frac{1}{f\pot} \pdiff{\CL}{\MMdm\dep_{mn}}
-   = \mu\pot_{mn} - \mu\dep_{mn} + 2\eqm_m (\etwm_m - \etwm_n)
+   = \kktm\pot_{mn} - \kktm\dep_{mn} + 2\eqm_m (\etwm_m - \etwm_n)
    = 0.
@@
 \begin{equation}\label{eq:lowertriangular}
- \mu\pot_{mn} - \mu\dep_{mn} < 0
+ \kktm\pot_{mn} - \kktm\dep_{mn} < 0
  \qquad\implies\qquad
- \mu\dep_{mn} > 0
+ \kktm\dep_{mn} > 0
  \qquad\implies\qquad
@@
 \begin{equation}\label{eq:uppertriangular}
- \mu\pot_{mn} - \mu\dep_{mn} > 0
+ \kktm\pot_{mn} - \kktm\dep_{mn} > 0
  \qquad\implies\qquad
- \mu\pot_{mn} >0
+ \kktm\pot_{mn} >0
  \qquad\implies\qquad
@@
     \Delta^\nu_{mn}\CL =
-      \frac{\mu^\nu_{mn}}{\eqm_m} + \frac{\mu^\nu_{nm}}{\eqm_n}
+      \frac{\kktm^\nu_{mn}}{\eqm_m} + \frac{\kktm^\nu_{nm}}{\eqm_n}
       + r \prn{c_m - c_n}\prn{\etwm_m - \etwm_n} = 0,
@@
 
-Using the non-negativity of the Kuhn-Tucker multipliers, $\mu^\nu_{ij}$, \cref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\etwm_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\etwm_k$ are degenerate, we can choose their order to ensure this).
-
-Note that, according to \cref{sec:triangular}, either $\MMdm\pot_{mn}$ or $\MMdm\dep_{mn}$ will be zero at the maximum, therefore we can expect one of the $\mu^\nu_{mn}$ to be non-zero.
+Using the non-negativity of the Kuhn-Tucker multipliers, $\kktm^\nu_{ij}$, \cref{eq:areacoeffincr} tells us that if we arrange the states in order of decreasing $\etwm_i$, the optimal process will have non-decreasing $c_k$ (if any of the $\etwm_k$ are degenerate, we can choose their order to ensure this).
+
+Note that, according to \cref{sec:triangular}, either $\MMdm\pot_{mn}$ or $\MMdm\dep_{mn}$ will be zero at the maximum, 
+therefore we can expect one of the $\kktm^\nu_{mn}$ to be non-zero.
 This would rule out degeneracy of the $c_k$ or $\etwm_k$.
-Looking at \cref{eq:shiftqderiv} closely, the only way both $\mu^\nu_{mn}$ could be zero is if $\etwm_m = \etwm_n$ or $\eqm_m = \eqm_n = 0$.
+Looking at \cref{eq:shiftqderiv} closely, the only way both $\kktm^\nu_{mn}$ could be zero is if $\etwm_m = \etwm_n$ or $\eqm_m = \eqm_n = 0$.
 
@@
     \widetilde{\Delta}^\nu_{mn}\CL =
-      \brk{ \sum_{k=m}^{n\mp1} \frac{\mu^\nu_{k,k\pm1}}{\eqm_k}
-      - \frac{\mu^\nu_{m,n}}{\eqm_m}}
+      \brk{ \sum_{k=m}^{n\mp1} \frac{\kktm^\nu_{k,k\pm1}}{\eqm_k}
+      - \frac{\kktm^\nu_{m,n}}{\eqm_m}}
       + rf^\nu \sum_{k=m}^{n\mp1} \prn{\etwm_{k} - \etwm_{k\pm1}} \prn{c_k - c_m},
@@
 This implies that the last term of the final expression in \eqref{eq:shortcutarea} is non-negative.
-If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu^\nu_{mn} > 0$, which in turn implies that $\MMdm^\nu_{mn}=0$.
+If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\kktm^\nu_{mn} > 0$, which in turn implies that $\MMdm^\nu_{mn}=0$.
 This would tell us that the process with the maximal area has to have a serial topology.
@@
 The model that saturates this bound is one of serial topology with \(\eq\) concentrated symmetrically at the two end states.
+
+Note that in the case of a model that is not of the serial topology, but is lumpable to one, 
+it will satisfy \cref{eq:areabnd} with a lower value of \(M\).
+Therefore the maximal area will come from a model with the full serial topology.
 
@@
 %
-bNumerical maximisation of the quantity above produces the same result as in \cref{sec:initial}:
+Numerical maximisation of the quantity above produces the same result as in \cref{sec:initial}:
 the two-state model with deterministic transitions.
@@
 So we see that all four inequalities cannot be simultaneously saturated, and the bound must be loose.
+Because we think that \cref{eq:sqrtbound} can be improved, we will rewrite it as
+%
+\begin{equation}\label{eq:gammabnd}
+  \CI_a \sqrt{r\tau_a} \leq \sqrt{\cnst}.
+\end{equation}
+% 
+We have shown that the smallest \(\cnst\) for which \cref{eq:gammabnd} is true is less than 2.
+Considering the two-state model with deterministic transitions, we see that \(\cnst\) is at least 1.
+Later, looking at uniform serial models in \cref{sec:heuristicenv} we will see near \cref{eq:henvapprox} that \(\gamma\) is at least 1.18.
 
@@
 \begin{equation}\label{eq:envlagrangian}
-  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} + \mu_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
+  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \kktm_\CI \prn{1 - \sum_a \CI_a} + \kktm_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
 \end{equation}
@@
   \pdiff{\CL}{\tau_a} = 0, \qquad
-  \mu_\alpha \geq 0, \qquad
-  \pdiff{\CL}{\mu_\alpha} \geq 0, \qquad
-  \mu_\alpha\pdiff{\CL}{\mu_\alpha} = 0.
+  \kktm_\alpha \geq 0, \qquad
+  \pdiff{\CL}{\kktm_\alpha} \geq 0, \qquad
+  \kktm_\alpha\pdiff{\CL}{\kktm_\alpha} = 0.
 \end{equation}
@@
   \tau_1 = M-1, \\
-  \mu_\CI = \frac{(M-1)^2}{[r\tau + (M-1)]^2}, \qquad
-  \mu_\CA = \frac{r\tau}{[r\tau+(M-1)]^2}. \qquad
+  \kktm_\CI = \frac{(M-1)^2}{[r\tau + (M-1)]^2}, \qquad
+  \kktm_\CA = \frac{r\tau}{[r\tau+(M-1)]^2}. \qquad
 \end{gathered}
@@
 \begin{equation}\label{eq:envDBlagrangian}
-  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \mu_\CI \prn{1 - \sum_a \CI_a} 
-      + \sum_a \mu_a \prn{\sqrt{2} - \CI_a \sqrt{\tau_a}} + \mu_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
+  \CL = \sum_a \frac{\CI_a}{1+\tau/\tau_a} + \kktm_\CI \prn{1 - \sum_a \CI_a} 
+      + \sum_a \kktm_a \prn{\sqrt{\cnst} - \CI_a \sqrt{\tau_a}} + \kktm_\CA \prn{(M-1) - \sum_a \CI_a \tau_a}.
 \end{equation}
@@
   \begin{alignedat}{3}
-    &\frac{2 \sqrt{N}}{2 + r \tau}
+    &\frac{\sqrt{N} \cnst}{\cnst + r \tau}
       , &
-      0 &\leq r \tau &&\leq 2
+      0 &\leq r \tau &&\leq \cnst
       , \\
-    &\sqrt{\frac{N}{2 r \tau}}
+    &\sqrt{\frac{N\cnst}{4 r \tau}}
       , &
-      2 &< r \tau &&< \frac{(M-1)^2}{2}
+      \cnst &< r \tau &&< \frac{(M-1)^2}{\cnst}
       , \\
-    &\frac{2 \sqrt{N} (M-1)}{(M-1)^2 + 2r \tau}
+    &\frac{\sqrt{N} \cnst (M-1)}{(M-1)^2 + \cnst r \tau}
       , &\qquad
-      \frac{(M-1)^2}{2} &\leq r \tau 
+      \frac{(M-1)^2}{\cnst} &\leq r \tau 
       .
@@
       \multicolumn{5}{c}{Domain of validity} & \CI_a & \tau_a 
-        & \mu_\CI & \mu_a & \mu_\CA 
+        & \kktm_\CI & \kktm_a & \kktm_\CA 
         % & \snrb(\tau) 
@@
     \midrule
-      0 &\leq r \tau &\leq 2 & M &\geq 3 & 1 & 2/r 
-        & \frac{4 - 2r \tau}{[2 + r \tau]^2} & \frac{2\sqrt{2}r\tau}{[2 + r \tau]^2} & 0 
+      0 &\leq r \tau &\leq \cnst & M &\geq 3 & 1 & \cnst/r 
+        & \frac{\cnst(\cnst - r \tau)}{[\cnst + r \tau]^2} & \frac{2\sqrt{\cnst}r\tau}{[\cnst + r \tau]^2} & 0 
         % & \frac{2 \sqrt{N}}{2 + r \tau} 
       \\
-      2 &< r \tau &< \frac{(M-1)^2}{2} & M &> 3 & \sqrt{\frac{2}{r\tau}} & \tau 
+      \cnst &< r \tau &< \frac{(M-1)^2}{\cnst} & M &> 3 & \sqrt{\frac{\cnst}{r\tau}} & \tau 
         & 0 & \frac{1}{2\sqrt{r \tau}} & 0 
@@
       \\ \addlinespace[0.2ex]
-      \frac{(M-1)^2}{2} &\leq r \tau & & M &\geq 3 & \frac{2}{M-1} & \frac{(M-1)^2}{2r} 
-        & 0 & \frac{\sqrt{2}(M-1)^3}{[(M-1)^2 + 2r \tau]^2} 
-        & \frac{4r \tau - 2(M-1)^2}{[(M-1)^2 + 2r \tau]^2} 
+      \frac{(M-1)^2}{\cnst} &\leq r \tau & & M &\geq 3 & \frac{\cnst}{M-1} & \frac{(M-1)^2}{\cnst r} 
+        & 0 & \frac{2 \sqrt{\cnst} (M-1)^3}{[(M-1)^2 + \cnst r \tau]^2} 
+        & \frac{\cnst (\cnst r \tau - (M-1)^2)}{[(M-1)^2 + 2r \tau]^2} 
         % & \frac{2 \sqrt{N} (M-1)}{(M-1)^2 + 2r \tau} 
@@
       &-& & M &\leq 3 & 1 & \frac{M - 1}{r} 
-        & \frac{(M-1)^2}{[(M-1) + r \tau]^2} %- \frac{\mu_a \sqrt{\tau_a}}{2} 
-        & 0 & \frac{r \tau}{[(M-1) + r \tau]^2} %- \frac{\mu_a}{2 \sqrt{\tau_a}} 
+        & \frac{(M-1)^2}{[(M-1) + r \tau]^2} %- \frac{\kktm_a \sqrt{\tau_a}}{2} 
+        & 0 & \frac{r \tau}{[(M-1) + r \tau]^2} %- \frac{\kktm_a}{2 \sqrt{\tau_a}} 
         % & \frac{\sqrt{N} (M-1)}{(M-1) + r \tau} 
@@
 
-For this section we will treatthe off-diagonal components of \(\W^\nu\) as the independent degrees of freedom, as in \cref{eq:wdof}, rather than the \(\MMd^\nu\).
+For this section we will treat the off-diagonal components of \(\W^\nu\) as the independent degrees of freedom, as in \cref{eq:wdof}, rather than the \(\MMd^\nu\).
 Consider the Lagrangian
@@
   \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
-        +\sum_{\nu i j} \mu^\nu_{ij} \prn{ \Wm^\nu_{ij} + r f^\nu \delta_{ij} }
+        +\sum_{\nu i j} \kktm^\nu_{ij} \prn{ \Wm^\nu_{ij} + r f^\nu \delta_{ij} }
         + \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
@@
   \pdiff{\CL}{\Wm^\nu_{ij}} = 0, \quad
-  \mu^\nu_{ij} \geq 0, \quad
-  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
-  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
+  \kktm^\nu_{ij} \geq 0, \quad
+  \pdiff{\CL}{\kktm^\nu_{ij}} \geq 0, \quad
+  \kktm^\nu_{ij}\pdiff{\CL}{\kktm^\nu_{ij}} = 0, \quad
   \pdiff{\CL}{\lambda} = 0.
@@
      + \prn{c_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
-     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
+     &+ \kktm^\nu_{mn} -  \kktm^\nu_{mm}
      + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
@@
 \begin{equation*}
-  \mu^\nu_{ij} \Wm^\nu_{ij} = 0,
+  \kktm^\nu_{ij} \Wm^\nu_{ij} = 0,
   \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} \Wm^\nu_{ij} 
-      = \mu^\nu_{ii} ( - \Wm^\nu_{ii}) = r f^\nu \mu^\nu_{ii}.
+  \sum_{j: j \neq i} \kktm^\nu_{ii} \Wm^\nu_{ij} 
+      = \kktm^\nu_{ii} ( - \Wm^\nu_{ii}) = r f^\nu \kktm^\nu_{ii}.
 \end{equation*}
@@
 \begin{aligned}
-  \mu^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
+  \kktm^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
     + \lambda \ombm^\nu_m(0) 
@@
     }, \\
-  \mu^\nu_{mn} &= \eqm_m \big( 
+  \kktm^\nu_{mn} &= \eqm_m \big( 
       \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
@@
   \\
-  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
     \left[ 
@@
   \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
-        +\sum_{\nu i j} \mu^\nu_{ij}\prn{\Wm^\nu_{ij} + f^\nu (r + s) \delta_{ij} -  s f^\nu \eqm_j }
+        +\sum_{\nu i j} \kktm^\nu_{ij}\prn{\Wm^\nu_{ij} + f^\nu (r + s) \delta_{ij} -  s f^\nu \eqm_j }
         &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
@@
      + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
-     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
-     - \sum_{\mu ij} \mu^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
+     &+ \kktm^\nu_{mn} -  \kktm^\nu_{mm}
+     - \sum_{\mu ij} \kktm^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
     = 0.
@@
 \begin{aligned}
-  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\boldsymbol{\mu}^\nu)\trans \onev,
+  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\kkt^\nu)\trans \onev,
   \qquad
@@
 \begin{equation*}
-  \mu^\nu_{ij} (\Wm^\nu_{ij} - s \eqm_j) = 0,
+  \kktm^\nu_{ij} (\Wm^\nu_{ij} - s \eqm_j) = 0,
   \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \mu^\nu_{ii} (\Wm^\nu_{ij} - s \eqm_j)
-      = \mu^\nu_{ii} (1 - \Wm^\nu_{ii} - s (1 - \eqm_i)) = \mu^\nu_{ii},
+  \sum_{j: j \neq i} \kktm^\nu_{ii} (\Wm^\nu_{ij} - s \eqm_j)
+      = \kktm^\nu_{ii} (1 - \Wm^\nu_{ii} - s (1 - \eqm_i)) = \kktm^\nu_{ii},
 \end{equation*}
@@
 \begin{aligned}
-  \mu^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
+  \kktm^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
     + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
     }, \\
-  \mu^\nu_{mn} &= \eqm_m \prn{ 
+  \kktm^\nu_{mn} &= \eqm_m \prn{ 
       \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
@@
     &= 2 \eqm_m (\etwm_i - \etwm_j) 
-    + \mu\pot_{mj} - \mu\pot_{mi} - \mu\dep_{mj} + \mu\dep_{mi}.
-\end{aligned}
-\end{equation}
-%
-Because \( \etwm_i > \etwm_j \), we must have \( \mu\pot_{mi} + \mu\dep_{mj} > 0 \).
+    + \kktm\pot_{mj} - \kktm\pot_{mi} - \kktm\dep_{mj} + \kktm\dep_{mi}.
+\end{aligned}
+\end{equation}
+%
+Because \( \etwm_i > \etwm_j \), we must have \( \kktm\pot_{mi} + \kktm\dep_{mj} > 0 \).
 So \( \Wm\pot_{mi} > 0 \) would imply that \( \Wm\dep_{mj} = 0 \) and vice versa.
@@
     2 \eqm_m (\etwm_m - \etwm_n) 
-    + \mu\pot_{mn} - \mu\pot_{mm} - \mu\dep_{mn} + \mu\dep_{mm}.
-\end{equation}
-%
-If we assume that both \( \Wm\pot_{mn} \) and \( \Wm\dep_{mn} \) are nonzero, we find that \( \mu\pot_{mm} - \mu\dep_{mm} = 2 \eqm_m (\etwm_m - \etwm_n) \).
+    + \kktm\pot_{mn} - \kktm\pot_{mm} - \kktm\dep_{mn} + \kktm\dep_{mm}.
+\end{equation}
+%
+If we assume that both \( \Wm\pot_{mn} \) and \( \Wm\dep_{mn} \) are nonzero, we find that \( \kktm\pot_{mm} - \kktm\dep_{mm} = 2 \eqm_m (\etwm_m - \etwm_n) \).
 Therefore \( \etwm_m > \etwm_n \implies \Wm\pot_{mm} = -rf\pot \) and \( \etwm_m < \etwm_n \implies \Wm\dep_{mm} = -rf\dep \).
@@
   \includegraphics[width=0.4\linewidth]{allowed_triangle.svg}}
-  \item \label{fig:triangle_spacing} \aligntop{%
-  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
 \end{myenuma}
@@
   (\ref*{fig:triangle_allowed}) The not-as-yet ruled out transitions when states are sorted so that \(\etw\) decreases to the right, with \(\etwm_m\) unconstrained (as yet).
-(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, with \(i\) a state with a nonzero depressing transition from \(m\), \(j\) a state with a nonzero potentiating transition, \(n\) a state with both depressing and potentiating transitions and \(k\) a state with no nonzero transitions from \(m\).
-The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
 \label{fig:triangle_s}}
@@
   \implies \quad
-  \pdiff{\CL}{\Wm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
+  \pdiff{\CL}{\Wm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \kktm^\nu_{mn} - \kktm^\nu_{mm}.
 \end{equation}
@@
 Suppose that a state \(n\) has a nonzero transition rate from state \(m\) for one \(\nu\).
-Then \(\mu^\nu_{mn} = 0\) and:
+Then \(\kktm^\nu_{mn} = 0\) and:
 %
@@
   \quad \implies \quad
-  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n})
+  \kktm^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n})
   \quad \implies \quad
@@
 Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
-Both must statisfy \cref{eq:oneexit} for the same \(\mu^\nu_{mm}\), so
+Both must statisfy \cref{eq:oneexit} for the same \(\kktm^\nu_{mm}\), so
 %
@@
   \;\;\; \implies \;\;\;
-  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
+  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i})
   \;\; \implies \;\;
@@
 %
-The consequences of these inequalities are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}).
+The consequences of these inequalities are depicted in \cref{fig:planes}(\ref{fig:triangle_spacing}).
 We see that \(n\) must be a global minimum of \(\dgn_m\), and if there are multiple transitions of a single type (potentiating or depressing) the targets are constrained to lie on a single line.
@@
 
-We use \(m*\) to label the intersection of the two diagonal lines in \cref{fig:triangle_s}(\ref{fig:triangle_spacing}), wether or not there is a state there.
+We use \(m*\) to label the intersection of the two diagonal lines in \cref{fig:planes}(\ref{fig:triangle_spacing}), wether or not there is a state there.
 It can be defined for each \(m\), provided that it has at least one outgoing transition for each plasticity type, via
@@
 \begin{equation}\label{eq:KKTz}
-  \mu^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m*}),
-  \qquad
-  \mu^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m*}).
+  \kktm^\nu_{mm} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m*}),
+  \qquad
+  \kktm^\nu_{mk} = \eqm_m (\dgnm^\nu_{m,k} - \dgnm^\nu_{m*}).
 \end{equation}
@@
 This would imply that, after lumping all lumpable states together, the situation in \cref{eq:twoexits} is impossible and \cref{eq:twodoubleexits} is even more impossible.
-If this intuition holds, we expect that \cref{fig:triangle_s}(\ref{fig:triangle_spacing}) would have at most one state on each of the two diagonal lines.
+If this intuition holds, we expect that \cref{fig:planes}(\ref{fig:triangle_spacing}) would have at most one state on each of the two diagonal lines.
 Either there would be one state like \(n\), or there would be one state like \(i\) and one state like \(j\), and the first of these possibilities is rather implausible.
 
+\begin{figure}[ht]
+\begin{center}
+\begin{myenuma}
+  \item \label{fig:triangle_spacing} \aligntop{ \hspace{-2em}%
+  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
+  \hspace{0.2em}
+  \item \label{fig:scaling_plane}\aligntop{ \hspace{-1em}%
+  \includegraphics[width=0.4\linewidth]{scaling_plane.svg}}
+\end{myenuma}
+\caption[Depicting constraints on topology]{Depicting constraints on topology.
+(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, 
+  with \(i\) a state with a nonzero depressing transition from \(m\), 
+  \(j\) a state with a nonzero potentiating transition, 
+  \(n\) a state with both depressing and potentiating transitions 
+  and \(k\) a state with no nonzero transitions from \(m\).
+  The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
+(\ref*{fig:scaling_plane}) Scaling of optimal models in the \((s,r)\)-plane.
+  Optimal models on grey ``rays'' are related by scaling, and therefore have the same topology.
+  The line at \(r = r_0\) is where we need the solutions.
+  The line at \(s = \epsilon\) is where we can determine the topology.
+\label{fig:planes}}
+\end{center}
+\end{figure}
+
 %-------Section------------------------------------------------------------------
@@
 \begin{aligned}
-  \mu^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
     \left[ 
@@
 
-\Cref{eq:KKTz} provides another expression for \(\mu^\nu_{mm}\).
+\Cref{eq:KKTz} provides another expression for \(\kktm^\nu_{mm}\).
 That expression will agree with \cref{eq:diagKKTz} provided that
@@
 \subsection{Scaling with a floating cutoff}\label{sec:cutoff}
+
+We have seen that invariance under the scale transformation \(\W \to \lambda\W \) is broken in two places: adding \(s\I \) in \cref{eq:lfund} and the upper bound on transition probabilities in \cref{eq:wdof}.
+It is helpful to describe the broken symmetry by assigning scaling properties to the parameters.
+To do that we need to create a parameter for the upper bound in \cref{eq:wdof}.
+
+If we look at \cref{eq:finitelagrangian}, this parameter would appear in the \(r f^\nu \delta_{ij}\) term.
+Furthermore, when we phrase the problem in terms of \(\W^\nu \) rather than \(\MMd^\nu \), this is the only place where \(r\) appears.
+Therefore we can absorb this parameter into \(r\) -- we will call the original value \(r_0\).
+
+We have the following invariances:
+%
+\begin{equation*}
+  A(\lambda s, \lambda \W) = A(s, \W),
+  \qquad
+  \CL(\lambda s, \lambda r, \lambda \W, \lambda\inv \kkt) = \CL(s, r, \W, \kkt).
+\end{equation*}
+%
+or in differential form:
+%
+\begin{equation*}
+  s \pdiff{A}{s} + \sum_{\nu i j} \Wm^\nu_{ij} \pdiff{A}{\Wm^\nu_{ij}} = 0,
+  \qquad
+  s \pdiff{\CL}{s} + r \pdiff{\CL}{r} 
+    + \sum_{\nu i j} \brk{\Wm^\nu_{ij} \pdiff{\CL}{\Wm^\nu_{ij}} 
+                      - \kktm^\nu_{ij} \pdiff{\CL}{\kktm^\nu_{ij}}} = 0.
+\end{equation*}
+%
+
+If we denote the solution to the maximisation problem by \(\W^\ast(s,r), \kkt^\ast(s,r)\), we have
+%
+\begin{equation}\label{eq:optmodelscale}
+\begin{aligned}
+  \W^\ast(\lambda s, \lambda r) &= \lambda \W^\ast(s,r), &
+  \qquad \text{or} \qquad
+  s \pdiff{\W^\ast}{s} + r \pdiff{\W^\ast}{r} &= \W^\ast, \\
+  \kkt^\ast(\lambda s, \lambda r) &= \lambda\inv \kkt^\ast(s,r), &
+  s \pdiff{\kkt^\ast}{s} + r \pdiff{\kkt^\ast}{r} &= -\kkt^\ast.
+\end{aligned}
+\end{equation}
+%
+We can also denote the envelope by
+%
+\begin{equation*}
+  A^\ast(s,r) = A(s, \W^\ast(s,r))
+  \qquad \implies \qquad
+\begin{aligned}[t]
+  A^\ast(\lambda s, \lambda r) &= A^\ast(s,r), \\
+  s \pdiff{A^\ast}{s} + r \pdiff{A^\ast}{r} &= 0. \\
+\end{aligned}
+\end{equation*}
+%
+This means that the topology of the optimal model and the envelope are fixed along any of the ``rays'' of constant \(r/s\) in the \((s,r)\)-plane, as depicted in \cref{fig:planes}(\ref{fig:scaling_plane}).
+The transition probabilities of the optimal model obey a simple scaling relation along any such ray.
+
+Note from \cref{fig:planes}(\ref{fig:scaling_plane}) that the line we are interested in is the horizontal line, \(r = r_0\).
+This line intersects every ray except \(r = 0\).
+The same is true of any horizontal line of constant \(r \neq 0\).
+Any vertical line of constant \(s \neq 0\) will also intersect every ray except \(s = 0\).
+
+From \cref{sec:serialarea}, we know the optimal models at \(s = 0\) (There is still an ambiguity concering the scale, but it makes sense to choose to saturate the diagonal constraints in \cref{eq:wdof} so that \cref{eq:optmodelscale} holds).
+At \(r = 0\) there is only one allowed model: \(\W = \mathbf{0}\).
+
+Therefore we will find the full range of optimal topologies if we look at any horizontal or vertical line.
+Given that we know the optimal models at \(s = 0\), we should be able to determine the topology of the optimal models at infinitesimal \(s = \epsilon\).
+More precisely:
+%
+\begin{equation}\label{eq:scaleopt}
+  A^\ast\prn{s, r_0} = A^\ast\prn{\epsilon, \frac{r_0 \epsilon}{s}},
+  \quad
+  \W^\ast\prn{s, r_0} = \prnfrac{s}{\epsilon} \W^\ast\prn{\epsilon, \frac{r_0 \epsilon}{s}},
+  \quad
+  \kkt^\ast\prn{s, r_0} = \prnfrac{\epsilon}{s} \kkt^\ast\prn{\epsilon, \frac{r_0 \epsilon}{s}}.
+\end{equation}
+%
 
@@
   \caption[Heuristic optimal models]{Heuristics for models that maximise memory curve.
-  All unlabelled arrowshave transition probabilities equal to 1.
+  All unlabelled arrows have transition probabilities equal to 1.
   At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with \(M\) states.
@@
   \qquad
-  \hat{A}_*(s) = \prn{\frac{ 2\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \frac{\abs{\beta}}{s},
+  \hat{A}_*(s) = \prn{\frac{ \sinh y_* }{ \cosh^2 y_* }} 
+      \frac{\abs{\beta(s)}}{s},
   \\ \text{where}\quad
-  y_* = \tanh \frac{y_*}{2} \cosh y_* \approx 1.50553.
+  y_* = \tanh \frac{y_*}{2} \cosh y_* \approx 1.50553,
   \qquad
@@
 %
+and \(\beta(s)\) is given by the second formula of \cref{eq:detagensol}.
+
 This is only valid for
@@
   \hat{A}(s) = \frac{\Delta f_m}{s \Delta m \Delta g_m}
-                \brk{\frac{\sqrt{(f_m - \mu)(f_m - g_m)} + \sqrt{(\mu - 1)(g_m - 1)}}
-                          {\sqrt{(f_m - g_m)(g_m - 1)} + \sqrt{(f_m - \mu)(\mu - 1)}}}^2.
+      \brk{\frac{\sqrt{(f_m - \mu)(f_m - g_m)} 
+                  + \sqrt{(\mu - 1)(g_m - 1)}}
+                {\sqrt{(f_m - g_m)(g_m - 1)} 
+                  + \sqrt{(f_m - \mu)(\mu - 1)}}}^2.
 % \end{aligned}
@@
 \begin{equation}\label{eq:stickyeps}
-  \varepsilon = 1 - \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},
+  \varepsilon = 1 - \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} 
+    \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},
 \end{equation}
@@
 %
-Then, combining \eqref{eq:binaryenv}, \eqref{eq:unienv} and \eqref{eq:stickylaplace}, using the notation \( \Delta_{M,\varepsilon}[T(M, \beta)] = T(M, \beta) - \varepsilon T(M-2, \beta) \)., we find
+Then, combining \cref{eq:binaryenv,eq:unienv,eq:stickylaplace}, using the notation \( \Delta_{\varepsilon}[T(M)] = T(M) - \varepsilon T(M-2) \) for any function \(T(M)\), we find
 %
@@
   & \max_{\varepsilon \in [0,1]} \brc{
-      \frac{2\sqrt{N}(1-\varepsilon)\, \Delta_{M,\varepsilon}\! \brk{S\prn{\tfrac{M\beta}{2}}}}
-      {\Delta_{M,\varepsilon}[M]\, \Delta_{M,\varepsilon}\! \brk{S\prn{\tfrac{M\beta}{2}} + 1} } }, &
-  \tfrac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau .
+      \frac{2\sqrt{N}(1-\varepsilon) \Delta_{\varepsilon}\brk{S\prn{\tfrac{M\beta}{2}}}}
+      {\Delta_{\varepsilon}\brk{M} \Delta_{\varepsilon}\brk{S\prn{\tfrac{M\beta}{2}} + 1} } 
+      }, &
+  \tfrac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau ,
 \end{aligned}
@@
 %
-For \(M \gg 1\), and \(r\tau\) much greater than the lower bound of each region, we can approximate this as
+where \(\beta(\tau)\) and \(S(\beta)\) are given by the second formula of \cref{eq:detagensol}.
+
+For \(M \gg 1\), and \(r\tau\) much greater than the lower bound of each region, we can approximate the envelope as
 %
@@
 It should be compared to \cref{eq:env,eq:envDB}.
+\Cref{eq:henvapprox} has a similar form as \cref{eq:envDB}.
+The left edge of the middle section is at later times in the theoretical version if \(\gamma \gtrsim \frac{1}{2} \csch^2 \frac{y_\ast}{2} \approx 0.73 \).
+The right edge of the middle section is at later times in the theoretical version if \(\gamma \lesssim 2 y_\ast^2 \approx 4.53 \).
+The middle section has greater \(\snrb\) in the theoretical version if \(\gamma \gtrsim \frac{8\sinh^2y_\ast}{\cosh^4y_\ast} \approx 1.18 \).
 

### 2021-09-30 Notes/laplace.tex
area maxers section

@@
   \begin{alignedat}{3}
-    &\frac{\sqrt{N} \cnst}{\cnst + r \tau}
+    &\frac{\cnst \sqrt{N}}{\cnst + r \tau}
       , &
@@
       , \\
-    &\sqrt{\frac{N\cnst}{4 r \tau}}
+    &\sqrt{\frac{\cnst N}{4 r \tau}}
       , &
@@
       , \\
-    &\frac{\sqrt{N} \cnst (M-1)}{(M-1)^2 + \cnst r \tau}
+    &\frac{\cnst \sqrt{N} (M-1)}{(M-1)^2 + \cnst r \tau}
       , &\qquad
@@
 \begin{equation*}
-  s \pdiff{A}{s} + \sum_{\nu i j} \Wm^\nu_{ij} \pdiff{A}{\Wm^\nu_{ij}} = 0,
-  \qquad
-  s \pdiff{\CL}{s} + r \pdiff{\CL}{r} 
-    + \sum_{\nu i j} \brk{\Wm^\nu_{ij} \pdiff{\CL}{\Wm^\nu_{ij}} 
-                      - \kktm^\nu_{ij} \pdiff{\CL}{\kktm^\nu_{ij}}} = 0.
-\end{equation*}
-%
-
-If we denote the solution to the maximisation problem by \(\W^\ast(s,r), \kkt^\ast(s,r)\), we have
+  s \pdiffc<\Wm>{A}{s} + \sum_{\nu i j} \Wm^\nu_{ij} \pdiffc<s>{A}{\Wm^\nu_{ij}} 
+  = 
+  s \pdiffc<r \Wm \mathrlap{\kktm}>{\CL}{s} 
+    + r \pdiffc<s \Wm \mathrlap{\kktm}>{\CL}{r} 
+    + \sum_{\nu i j} \brk{\Wm^\nu_{ij} \pdiffc<s r \mathrlap{\kktm}>{\CL}{\Wm^\nu_{ij}} 
+                      - \kktm^\nu_{ij} \pdiffc<s r {\Wm}>{\CL}{\kktm^\nu_{ij}}} = 0.
+\end{equation*}
+%
+In \cref{eq:kktsol} we have an expression for the KKT multipliers at the maximum in terms of \(s\) and \(\W \).
+We denote this expression by \( \kkt^\ddagger(s, \W) \).
+If we denote the solution to the maximisation problem by \(\W^\ast(s,r), \kkt^\ast(s,r)\), we have \( \kkt^\ast(s,r) = \kkt^\ddagger(s, \W^\ast(s,r)) \) and
 %
@@
 \begin{aligned}
-  \W^\ast(\lambda s, \lambda r) &= \lambda \W^\ast(s,r), &
+  \kkt^\ddagger(\lambda s, \lambda \W) 
+    &= \lambda\inv \kkt^\ddagger(s, \W), &
   \qquad \text{or} \qquad
-  s \pdiff{\W^\ast}{s} + r \pdiff{\W^\ast}{r} &= \W^\ast, \\
+  s \pdiffc<\Wm>{\kkt^\ddagger}{s} 
+      + \sum_{\nu i j} \Wm^\nu_{ij} \pdiffc<s>{\kkt^\ddagger}{\Wm^\nu_{ij}} 
+    &= -\kkt^\ddagger, \\
+  \W^\ast(\lambda s, \lambda r) 
+    &= \lambda \W^\ast(s,r), &
+  s \pdiffc<r>{\W^\ast}{s} + r \pdiffc<s>{\W^\ast}{r} 
+    &= \W^\ast, \\
   \kkt^\ast(\lambda s, \lambda r) &= \lambda\inv \kkt^\ast(s,r), &
-  s \pdiff{\kkt^\ast}{s} + r \pdiff{\kkt^\ast}{r} &= -\kkt^\ast.
-\end{aligned}
-\end{equation}
-%
-We can also denote the envelope by
-%
-\begin{equation*}
-  A^\ast(s,r) = A(s, \W^\ast(s,r))
-  \qquad \implies \qquad
+  s \pdiffc<r>{\kkt^\ast}{s} + r \pdiffc<s>{\kkt^\ast}{r} 
+    &= -\kkt^\ast.
+\end{aligned}
+\end{equation}
+%
+We can also denote the envelope by \( A^\ast(s,r) = A(s, \W^\ast(s,r) \)
+%
+\begin{equation*}
 \begin{aligned}[t]
-  A^\ast(\lambda s, \lambda r) &= A^\ast(s,r), \\
-  s \pdiff{A^\ast}{s} + r \pdiff{A^\ast}{r} &= 0. \\
+  A^\ast(\lambda s, \lambda r) &= A^\ast(s,r), 
+  \qquad \text{or} \qquad
+  s \pdiff{A^\ast}{s} + r \pdiff{A^\ast}{r} &= 0.
 \end{aligned}
@@
 
-Note from \cref{fig:planes}(\ref{fig:scaling_plane}) that the line we are interested in is the horizontal line, \(r = r_0\).
+Note from \cref{fig:planes}(\ref{fig:scaling_plane}) that the line of interest is the horizontal line, \(r = r_0\).
 This line intersects every ray except \(r = 0\).
@@
 
-From \cref{sec:serialarea}, we know the optimal models at \(s = 0\) (There is still an ambiguity concering the scale, but it makes sense to choose to saturate the diagonal constraints in \cref{eq:wdof} so that \cref{eq:optmodelscale} holds).
+From \cref{sec:serialarea}, we know the optimal models at \(s = 0\).
+% (There is still an ambiguity concering the scale, but it makes sense to choose to saturate the diagonal constraints in \cref{eq:wdof} so that \cref{eq:optmodelscale} holds).
 At \(r = 0\) there is only one allowed model: \(\W = \mathbf{0}\).
@@
 
+\subsection{Optimal models at \texorpdfstring{\(s=0\)}{s=0}}\label{sec:areamodels}
+
+We saw in \cref{sec:serialarea} that the optimal models at \(s = 0\) are those with the serial topology with \(\eq\) concentrated symmetrically at the two end states.
+The set of serial models with such a \(\eq\) are the limit as \(\delta \rightarrow 0\) of
+%
+\begin{equation}\label{eq:areamodels}
+\begin{aligned}
+  \W\pot_{ii+1} = \W\dep_{i+1i} = q_i, 
+  \text{ for } 2 \leq i \leq M-2,
+  \qquad
+  \W\pot_{12} &= \delta, \quad &
+  \W\pot_{M-1M} &= q_{M-1}, \\
+  \W\dep_{21} &= q_1, &
+  \W\dep_{MM-1} &= \delta.
+\end{aligned}
+\end{equation}
+%
+Where \(0 \leq q_i \leq rf^\nu\) for both \(\nu\).
+We have an \(M-1\) parameter space of maxima.
+Using results from \cref{sec:serialarea}:
+%
+\begin{equation*}
+% \begin{aligned}
+  \eqm_i = \frac{\delta}{2z} 
+  \text{ for } 2 \leq i \leq M-1,
+  \qquad
+  \eqm_1 = \eqm_M = \frac{1}{2z},
+  \text{ where } z = 1 + \frac{\delta (M-2)}{2},
+  \qquad
+  c_i = 2i - M - 1.
+% \end{aligned}
+\end{equation*}
+%
+We can find the other vectors of interest by:
+solving \(\frg \etw = (\I - \onev \eq) \w \) for \(\etw\),
+computing \(\mathbf{a} = - \enc \etw\),
+and solving \(\frg \thb = (\I - \onev \eq) \mathbf{a} \) for \(\thb\).
+First, some notation:
+%
+\begin{equation*}
+  \sum_k^{i:j} f(k) 
+    \equiv \sgn(j-i) \smashoperator{\sum_{k = \min(i,j)}^{\max(i,j)-1}} f(k)
+    = \begin{cases}
+        \sum_{k=i}^{j-1} f(k) & i < j,\\
+        - \sum_{k=j}^{i-1} f(k) & i > j.
+      \end{cases}
+\end{equation*}
+%
+Then, to leading order in small \(\delta\), we find:
+%
+\begin{equation}\label{eq:areavecs}
+\begin{aligned}
+  \etwm_i - \etwm_j &= \frac{z}{\delta} \sum_k^{i:j} \frac{1}{q_k}, & \qquad
+  \thbm_i - \thbm_j &= \frac{z}{\delta} \sum_k^{j:i} \frac{2k - M}{q_k}, \\
+  a_i &= \frac{1}{\eqm_i} 
+    \quad \text{for } 2 \leq i \leq M-1, &
+  a_1 &= a_M = 1.
+\end{aligned}
+\end{equation}
+%
+Now we can use \cref{eq:kktsol} to calculate the KKT multipliers
+%
+\begin{equation}\label{eq:areaKKT}
+\begin{aligned}
+    \kktm^\nu_{ij} = \sum_k^{i:j} \frac{k - i - \delta\dep_\nu}{q_k}
+      \quad \text{ for } 2 \leq i \leq M-1, \qquad
+    \kktm^\nu_{1j} &= \frac{1}{\delta} \sum_{k=1}^{j-1} \frac{k - 1 - \delta\dep_\nu}{q_k},
+      \\
+    \kktm^\nu_{Mj} &= \frac{1}{\delta} \sum_{k=j}^{M-1} \frac{M - k + \delta\dep_\nu}{q_k}.
+\end{aligned}
+\end{equation}
+%
+They are all nonnegative, as the numerator is only negative when \(i > j\) and the negative of the sum is used.
+There are only two ways a KKT multiplier can be zero. 
+If the sum is empty (\(j = i\)) or it contains exactly one term that is zero (\( j = i + (-1)^\nu \)).
+It is impossible for more than one term in the sum to be zero.
+Those transitions ddefine the serial topology.
+
+%-------Section------------------------------------------------------------------
+
 \section{Nearly uniform serial models}\label{sec:serial}
@@
   \begin{myenuma}
-    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_uni.svg}}\hspace{0.02\linewidth}\label{fig:serial_uniform}
-    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_sticky.svg}}\hspace{0.02\linewidth}\label{fig:serial_sticky}
-    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_shorten.svg}}\label{fig:serial_shorten}
+    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_uni.svg}}\label{fig:serial_uni}
+    \hspace{0.02\linewidth}
+    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_sticky.svg}}\label{fig:serial_stk}
+    \hspace{0.02\linewidth}
+    \item\aligntop{\includegraphics[width=0.26\linewidth]{serial_shorten.svg}}\label{fig:serial_sht}
   \end{myenuma}
@@
   All unlabelled arrows have transition probabilities equal to 1.
-  At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with \(M\) states.
-  At later times the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model.
-  At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
+  At one particular time, the model is (\ref{fig:serial_uni}) the uniform serial model with \(M\) states.
+  At later times the model has (\ref{fig:serial_stk}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model.
+  At earlier times, the model is (\ref{fig:serial_sht}) progressively shortened by reducing the transition probability to the end states.
   Once this reaches zero we discard the old end state and work on the new end state.
@@
 
-Looking at them models that were found by numerical maximisation in \cref{fig:envvid} (see the video for the full set of models) suggests we look at the particular set of models shown in \cref{fig:heuristicmodels}.
-
-At one particular time the model is (\ref{fig:serial_uniform}) the uniform serial model with \(M\) states.
-At later times the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model as \(\tau \to \infty\).
-At earlier times the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
+Looking at them models that were found by numerical maximisation in \cref{fig:envvid} 
+(see the video for the full set of models) suggests we look at the particular set of models shown in \cref{fig:heuristicmodels}.
+
+At one particular time the model is (\ref{fig:serial_uni}) the uniform serial model with \(M\) states.
+At later times the model has (\ref{fig:serial_stk}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model as \(\tau \to \infty\).
+At earlier times the model is (\ref{fig:serial_sht}) progressively shortened by reducing the transition probability to the end states.
 Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from \cref{fig:envvid} and the video this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
@@
 
-Here we will look at the model in \cref{fig:heuristicmodels}\ref{fig:serial_uniform}.
+Here we will look at the model in \cref{fig:heuristicmodels}\ref{fig:serial_uni}.
 Looking at \eqref{eq:larea}, the only thing we need is differences in \(\etwm_i(s)\).
@@
 
-Here we will look at the model in \cref{fig:heuristicmodels}\ref{fig:serial_shorten}.
+Here we will look at the model in \cref{fig:heuristicmodels}\ref{fig:serial_sht}.
 We proceed in the same manner as in \cref{sec:serial_uniform}, with the following differences.
@@
 
-Here we will look at the model in \cref{fig:heuristicmodels}\ref{fig:serial_sticky}.
+Here we will look at the model in \cref{fig:heuristicmodels}\ref{fig:serial_stk}.
 We proceed in the same manner as in \cref{sec:serial_uniform}, with the following differences.
@@
 %
-The maximum is at the solution of
-%
-\begin{multline}\label{eq:stickyopt}
-% \begin{aligned}
-  \brk{(1 + f_m -  g_m - \mu)\varepsilon - f_m + \mu g_m}^2 
-    = (\mu - 1) (g_m - 1) (f_m - \mu) (f_m - g_m) , \\
+The maximum is at
+%
+\begin{equation}\label{eq:stickyopt}
+\begin{aligned}
+  % \brk{(1 + f_m -  g_m - \mu)\varepsilon - f_m + \mu g_m}^2 
+  %   = (\mu - 1) (g_m - 1) (f_m - \mu) (f_m - g_m) , \\
+  \varepsilon &= \frac{f_m - \mu g_m 
+        \pm \sqrt{(\mu - 1) (g_m - 1) (f_m - \mu) (f_m - g_m)}}
+      {1 + f_m -  g_m - \mu} \\
   \implies \qquad
-  \hat{A}(s) = \frac{\Delta f_m}{s \Delta m \Delta g_m}
+  \hat{A}(s) &= \frac{\Delta f_m}{s \Delta m \Delta g_m}
       \brk{\frac{\sqrt{(f_m - \mu)(f_m - g_m)} 
@@
                   + \sqrt{(f_m - \mu)(\mu - 1)}}}^2.
-% \end{aligned}
-\end{multline}
+\end{aligned}
+\end{equation}
 %

### 2021-10-01 Notes/laplace.tex
written about area maxers KKT

@@
 The omission of the diagonal terms leaves only non-negative terms.
+We can think of \cref{eq:innereq} as the inner product in node-space with metric \( \diag(\eq) \), 
+and \cref{eq:innerflux} as the inner product in edge-space with metric \( \diag(\F) \).
+The edge-space metric is positive-definite if the sum in \cref{eq:innerflux} is restricted to the edges that have transitions.
 
@@
 We also have \( \nrm{\w}^2 = \sum_i \eqm_i \wm_i^2 = 1 \).
-Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \MMm_{ij} \leq r \sum_i \eqm_i = r\).
+Furthermore, using \cref{eq:constri}, \( \nnrm{\onem}^2 = \sum_{i \neq j} \eqm_i \frgm_{ij} \leq r \sum_i \eqm_i = r\).
 Then, remembering \cref{eq:numrange},
@@
 \begin{equation}\label{eq:sqrtnumrange}
-  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\MM \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
+  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\frg \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
       = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})/r}.
@@
   \qquad \implies \qquad
-  \dgnm^\mu_{m,i} = \dgnm^\nu_{m,j}
+  \dgnm^\nu_{m,i} = \dgnm^\nu_{m,j}
   \quad \text{or}\quad
@@
   The line at \(r = r_0\) is where we need the solutions.
-  The line at \(s = \epsilon\) is where we can determine the topology.
+  The line at \(s = \epsilon\) is close to where we can determine the topology.
 \label{fig:planes}}
@@
 %
-We can also denote the envelope by \( A^\ast(s,r) = A(s, \W^\ast(s,r) \)
+We can also denote the envelope by \( A^\ast(s,r) = A(s, \W^\ast(s,r)) \)
 %
@@
   \W\pot_{ii+1} = \W\dep_{i+1i} = q_i, 
-  \text{ for } 2 \leq i \leq M-2,
-  \qquad
-  \W\pot_{12} &= \delta, \quad &
-  \W\pot_{M-1M} &= q_{M-1}, \\
-  \W\dep_{21} &= q_1, &
-  \W\dep_{MM-1} &= \delta.
+  % \text{ for } 2 \leq i \leq M-2,
+  \qquad
+  \text{except} \quad
+  \W\pot_{12} &= \delta q_1, \quad &
+  % \W\pot_{M-1M} &= q_{M-1}, \\
+  % \W\dep_{21} &= q_1, &
+  \W\dep_{MM-1} &= \delta q_{M-1}.
 \end{aligned}
@@
   \etwm_i - \etwm_j &= \frac{z}{\delta} \sum_k^{i:j} \frac{1}{q_k}, & \qquad
-  \thbm_i - \thbm_j &= \frac{z}{\delta} \sum_k^{j:i} \frac{2k - M}{q_k}, \\
+  \thbm_i - \thbm_j &= \frac{z}{\delta} \sum_k^{i:j} \frac{M - 2k}{q_k}, \\
   a_i &= \frac{1}{\eqm_i} 
     \quad \text{for } 2 \leq i \leq M-1, &
-  a_1 &= a_M = 1.
+  a_1 &= a_M = 1, \\
+  \alpha^\nu_i &= - \frac{z ((-1)^\nu c_i + 1)}{r f^\nu \delta}
+    \quad \text{for } 2 \leq i \leq M-1, &
+  \alpha\pot_1 &= \frac{z(M-2)}{r f^\nu q_1}, \\
+  \alpha\pot_M &= \alpha\dep_1 = 0, &
+  \alpha\dep_M &= \frac{z(M-2)}{r f^\nu q_{M-1}}, \\
+  \omega^\nu_i &= \frac{(-1)^\nu z}{r f^\nu \delta}
+    \quad \text{for } 2 \leq i \leq M-1, &
+  \omega\pot_1 &= \frac{z}{r f^\nu q_1}, \\
+  \omega\pot_M &= \omega\dep_1 = 0, &
+  \omega\dep_M &= - \frac{z}{r f^\nu q_{M-1}}.
 \end{aligned}
@@
 \begin{aligned}
-    \kktm^\nu_{ij} = \sum_k^{i:j} \frac{k - i - \delta\dep_\nu}{q_k}
+    \kktm^\nu_{ij} = \sum_k^{i:j} \frac{k - i + \delta\dep_\nu}{q_k}
       \quad \text{ for } 2 \leq i \leq M-1, \qquad
-    \kktm^\nu_{1j} &= \frac{1}{\delta} \sum_{k=1}^{j-1} \frac{k - 1 - \delta\dep_\nu}{q_k},
+    \kktm^\nu_{1j} &= \frac{1}{\delta} \sum_{k=1}^{j-1} \frac{k - 1 + \delta\dep_\nu}{q_k},
       \\
-    \kktm^\nu_{Mj} &= \frac{1}{\delta} \sum_{k=j}^{M-1} \frac{M - k + \delta\dep_\nu}{q_k}.
+    \kktm^\nu_{Mj} &= \frac{1}{\delta} \sum_{k=j}^{M-1} \frac{M - k - \delta\dep_\nu}{q_k}.
 \end{aligned}
@@
 They are all nonnegative, as the numerator is only negative when \(i > j\) and the negative of the sum is used.
-There are only two ways a KKT multiplier can be zero. 
-If the sum is empty (\(j = i\)) or it contains exactly one term that is zero (\( j = i + (-1)^\nu \)).
-It is impossible for more than one term in the sum to be zero.
-Those transitions ddefine the serial topology.
+There are only two ways one of these KKT multipliers could be zero: 
+the sum is empty; 
+the sum contains exactly one term and that term is zero
+(It is impossible for more than one term in the sum to be zero).
+The first occurs for \(j = i\), the second for \( j = i + (-1)^\nu \).
+Those transitions define the serial topology.
 
@@
 %
-These equations are invariant under \(\delta\eta_{2m+1-i} \to -\delta\eta_i\), so these two must be equal.
+These equations are invariant under \(\delta\eta_{2m+1-i} \to -\delta\eta_i\), so these two must be equal (up to a constant that we can drop).
 Then we have
@@
 The other curves are from \cref{eq:shortopt}, optimised shortened models for various values of \(M\).
-(\ref{fig:short_ratio}): The ratios of the quantities plotted in (\ref{fig:short_opt}) to \cref{eq:unienv}.
+(\ref{fig:short_ratio}): The quantities plotted in (\ref{fig:short_opt}) divided by \cref{eq:unienv}.
 The dotted blue and orange lines are the extensions of the \(M=2,4\) lines with \(\varepsilon\) held at 0, as discussed below \cref{eq:shortenvalid}.

### 2021-10-11 Notes/laplace.tex
simpler heuristic envelope

@@
 We can maximise this \wrt \(\varepsilon\).
-We define
-%
-\begin{equation*}
-\begin{aligned}
-  \Delta m &= (m-1), &
-  \mu &= \frac{m}{\Delta m}, \\
-  \Delta f_m(\beta) &= S((m-1)\beta), &
-  f_m(\beta) &= \frac{S(m\beta)}{\Delta f_m(\beta)}, \\
-  \Delta g_m(\beta) &= \Delta f_m(\beta) + 1 , &
-  g_m(\beta) &= \frac{S(m\beta) + 1}{\Delta g_m(\beta)}, \\
-\end{aligned}
-  \; \implies \quad
-  \hat{A}(s) = \frac{\Delta f_m (1 - \varepsilon) (f_m - \varepsilon)}
-                    {s \Delta m \Delta g_m (\mu - \varepsilon) (g_m - \varepsilon)}.
-\end{equation*}
+% We define
+% %
+% \begin{equation*}
+% \begin{aligned}
+%   \Delta m &= (m-1), &
+%   \mu &= \frac{m}{\Delta m}, \\
+%   \Delta f_m(\beta) &= S((m-1)\beta), &
+%   f_m(\beta) &= \frac{S(m\beta)}{\Delta f_m(\beta)}, \\
+%   \Delta g_m(\beta) &= \Delta f_m(\beta) + 1 , &
+%   g_m(\beta) &= \frac{S(m\beta) + 1}{\Delta g_m(\beta)}, \\
+% \end{aligned}
+%   \; \implies \quad
+%   \hat{A}(s) = \frac{\Delta f_m (1 - \varepsilon) (f_m - \varepsilon)}
+%                     {s \Delta m \Delta g_m (\mu - \varepsilon) (g_m - \varepsilon)}.
+% \end{equation*}
 %
@@
   %   = (\mu - 1) (g_m - 1) (f_m - \mu) (f_m - g_m) , \\
-  \varepsilon &= \frac{f_m - \mu g_m 
-        \pm \sqrt{(\mu - 1) (g_m - 1) (f_m - \mu) (f_m - g_m)}}
-      {1 + f_m -  g_m - \mu} \\
-  \implies \qquad
-  \hat{A}(s) &= \frac{\Delta f_m}{s \Delta m \Delta g_m}
-      \brk{\frac{\sqrt{(f_m - \mu)(f_m - g_m)} 
-                  + \sqrt{(\mu - 1)(g_m - 1)}}
-                {\sqrt{(f_m - g_m)(g_m - 1)} 
-                  + \sqrt{(f_m - \mu)(\mu - 1)}}}^2.
-\end{aligned}
-\end{equation}
-%
+  \varepsilon 
+    % &= \frac{f_m - \mu g_m 
+    %     - \sqrt{(\mu - 1) (g_m - 1) (f_m - \mu) (f_m - g_m)}}
+    %   {1 + f_m -  g_m - \mu} \\
+    &= \frac{D - S(m\beta)}{D - S((m-1)\beta)} &
+  \implies \quad
+  \hat{A}(s) 
+    % &= \frac{\Delta f_m}{s \Delta m \Delta g_m}
+    %   \brk{\frac{\sqrt{(f_m - \mu)(f_m - g_m)} 
+    %               + \sqrt{(\mu - 1)(g_m - 1)}}
+    %             {\sqrt{(f_m - g_m)(g_m - 1)} 
+    %               + \sqrt{(f_m - \mu)(\mu - 1)}}}^2.
+    &= \frac{S(m\beta) - S((m-1)\beta)}{s\prn{D + 1}^2} ,
+\end{aligned}
+\end{equation}
+%
+where \( D = \sqrt{(m-1)S(m\beta) - m S((m-1)\beta)} \).
 The result is plotted in \cref{fig:sticky}.
@@
 %
-Then, combining \cref{eq:binaryenv,eq:unienv,eq:stickylaplace}, using the notation \( \Delta_{\varepsilon}[T(M)] = T(M) - \varepsilon T(M-2) \) for any function \(T(M)\), we find
+Then, combining \cref{eq:binaryenv,eq:unienv,eq:stickylaplace}, 
+% using the notation \( \Delta_{\varepsilon}[T(M)] = T(M) - \varepsilon T(M-2) \) for any function \(T(M)\), 
+we find
 %
@@
   & r\tau \leq \tfrac{1}{2\sinh^2\frac{y_*}{2}} , \\
-  & \sqrt{N}\prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \beta(\tau), &
+  & \sqrt{N} \prn{\frac{ \sinh y_* }{ \cosh^2 y_* }} \beta(\tau), &
   \tfrac{1}{2\sinh^2 \frac{y_*}{2}} \leq & r\tau \leq \tfrac{1}{2\sinh^2 \frac{y_*}{M}} , \\
-  & \max_{\varepsilon \in [0,1]} \brc{
-      \frac{2\sqrt{N}(1-\varepsilon) \Delta_{\varepsilon}\brk{S\prn{\tfrac{M\beta}{2}}}}
-      {\Delta_{\varepsilon}\brk{M} \Delta_{\varepsilon}\brk{S\prn{\tfrac{M\beta}{2}} + 1} } 
-      }, &
+  & \frac{\sqrt{N}\brk{S\prn{\tfrac{M\beta}{2}} - S\prn{\tfrac{(M-2)\beta}{2}}}}
+        {\brk{\sqrt{{\frac{M-2}{2} S\prn{\tfrac{M\beta}{2}} - \frac{M}{2} S\prn{\tfrac{(M-2)\beta}{2}}}} + 1}^2} , &
+  % & \max_{\varepsilon \in [0,1]} \brc{
+  %     \frac{2\sqrt{N}(1-\varepsilon) \Delta_{\varepsilon}\brk{S\prn{\tfrac{M\beta}{2}}}}
+  %     {\Delta_{\varepsilon}\brk{M} \Delta_{\varepsilon}\brk{S\prn{\tfrac{M\beta}{2}} + 1} } 
+  %     }, &
   \tfrac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau ,
@@
 %
-where \(\beta(\tau)\) and \(S(\beta)\) are given by the second formula of \cref{eq:detagensol}.
+where \(S(\beta) = \cosh \beta(\tau) - 1\) (see \cref{eq:detagensol}) and \(y_*\) is defined in \cref{eq:unienv}.
 
@@
 It should be compared to \cref{eq:env,eq:envDB}.
-\Cref{eq:henvapprox} has a similar form as \cref{eq:envDB}.
+\Cref{eq:henvapprox} has a similar form to the theoretical envelope in \cref{eq:envDB}.
 The left edge of the middle section is at later times in the theoretical version if \(\gamma \gtrsim \frac{1}{2} \csch^2 \frac{y_\ast}{2} \approx 0.73 \).

### 2021-10-11 Code/Python/complex_synapse/optimise/sticky.py

@@ -28,15 +28,32 @@ def beta_star(num: int) -> float:
 # =============================================================================
 
 
-def _components(beta: Values, num: int) -> Tuple[Values, ...]:
+def _components(beta: Values, numh: int) -> Tuple[Values, ...]:
     """Components of numerator and denominator
 
     numerator = numer - eps * dnumer
+
     denominator = denom - eps * ddenom
+
+    Parameters
+    ----------
+    numh
+        num/2
+
+    Returns
+    -------
+    numer
+        cosh(num*beta/2) - 1
+    denom
+        cosh(num*beta/2)
+    dnumer
+        cosh((num-2)*beta/2) - 1
+    ddenom
+        cosh((num-2)*beta/2)
     """
-    denom = np.cosh(num * beta / 2)
+    denom = np.cosh(numh * beta)
     numer = denom - 1
-    ddenom = np.cosh((num - 2) * beta / 2)
+    ddenom = np.cosh((numh - 1) * beta)
     dnumer = ddenom - 1
     return numer, denom, dnumer, ddenom
 
@@ -45,7 +62,7 @@ def _components(beta: Values, num: int) -> Tuple[Values, ...]:
 def sticky(eps: Values, beta: Values, num: int) -> Values:
     """Laplace-SNR for sticky serial model"""
     sval = beta_to_s(beta)
-    numer, denom, dnumer, ddenom = _components(beta, num)
+    numer, denom, dnumer, ddenom = _components(beta, num / 2)
     fnumer = numer - eps * dnumer
     fdenom = denom - eps * ddenom
     fmu = num - eps * (num - 2)
@@ -57,7 +74,7 @@ def eps_stars(beta: Values, num: int) -> Tuple[Values, Values]:
     """Optimal epsilon for sticky serial model, raw"""
     if num == 2:
         return 0, 0
-    numer, denom, dnumer, ddenom = _components(beta, num)
+    numer, denom, dnumer, ddenom = _components(beta, num / 2)
     fnumer = numer / dnumer
     fdenom = denom / ddenom
     fmu = num / (num - 2)
@@ -74,27 +91,54 @@ def eps_stars(beta: Values, num: int) -> Tuple[Values, Values]:
 def eps_star(beta: Values, num: int) -> Values:
     """Optimal epsilon for sticky serial model, clipped"""
     epss = eps_stars(beta, num)[0]
-    return np.minimum(np.maximum(epss, 0), 1)
+    return np.clip(epss, 0, 1)
 
 
[…]

### 2021-11-13 Notes/laplace.tex
reordered notes

@@
 Therefore, the conjecture from \cref{sec:finite} implies that the square-root bound of \cref{sec:sqrt} applies, implying the stricter envelope of \cref{eq:envDB}.
-In \cref{fig:envnum} we compare these theoretical envelopes with the numerical maximisation from \cref{fig:envnum}.
+In \cref{fig:envnumth} we compare these theoretical envelopes with the numerical maximisation from \cref{fig:envnum}.
 We can see that this conjectured envelope is close to tight, being off by a constant factor.
@@
   see \cref{eq:envDB}.}\label{fig:envnumth}
+\end{figure}
+
+%================================================================================
+
+\section{Finite time topology}\label{sec:finite}
+
+Now we'll maximise \(\snrb(\tau)\) for some fixed \(\tau\), subject to the constraints \eqref{eq:constr}.
+First, we will maximise the numerator of \eqref{eq:lsnr} holding \(\eq\w\) fixed.
+We can maximise the ratio \wrt \(\eq\w\) later.
+% From this point on, we work in units where \(r = 1\).
+
+For this section we will treat the off-diagonal components of \(\W^\nu\) as the independent degrees of freedom, as in \cref{eq:wdof}, rather than the \(\MMd^\nu\).
+Consider the Lagrangian
+%
+\begin{equation}\label{eq:finitelagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
+        +\sum_{\nu i j} \kktm^\nu_{ij} \prn{ \Wm^\nu_{ij} + r f^\nu \delta_{ij} }
+        + \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
+\end{aligned}
+\end{equation}
+%
+where \(\Delta p\) is the constant value we are holding \(\eq\w\) at.
+The Karush-Kuhn-Tucker necessary conditions for a maximum are
+%
+\begin{equation}\label{eq:KTfinite}
+  \pdiff{\CL}{\Wm^\nu_{ij}} = 0, \quad
+  \kktm^\nu_{ij} \geq 0, \quad
+  \pdiff{\CL}{\kktm^\nu_{ij}} \geq 0, \quad
+  \kktm^\nu_{ij}\pdiff{\CL}{\kktm^\nu_{ij}} = 0, \quad
+  \pdiff{\CL}{\lambda} = 0.
+\end{equation}
+%
+% We have scaled the KKT multipliers by \(r f^\nu\) for later convenience.
+
+The derivatives, for \(m \neq n\) are
+%
+\begin{equation}\label{eq:finitederivs}
+\begin{aligned}
+  \pdiff{\CL}{\Wm^\nu_{mn}} = & \,
+    \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
+     + \prn{c_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
+     &+ \kktm^\nu_{mn} -  \kktm^\nu_{mm}
+     + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
+    = 0.
+\end{aligned}
+\end{equation}
+%
+We can solve for the diagonal KKT multipliers by noting that
+%
+\begin{equation*}
+  \kktm^\nu_{ij} \Wm^\nu_{ij} = 0,
+  \qquad \text{and} \qquad
+  \sum_{j: j \neq i} \kktm^\nu_{ii} \Wm^\nu_{ij} 
+      = \kktm^\nu_{ii} ( - \Wm^\nu_{ii}) = r f^\nu \kktm^\nu_{ii}.
+\end{equation*}
+%
+Then, with some new vectors defined:
+%
+\begin{equation}\label{eq:kktsol}
+\begin{aligned}
+  \kktm^\nu_{mn} &= \eqm_m \big( 
+      \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
+    + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
+  \\ &\hphantom{= \eqm_m \big( }
+    + \brk{{c}_m(s) + (-1)^\nu} 
+        \brk{ \etwm_n(s) - \etwm_m(s) - \ombm^\nu_m(s) }
+     \big), \\
+  \text{where} \quad 
+  \alb^\nu(s) &= \frac{\W^\nu \thb(s)}{rf^\nu}, \qquad
+  \omb^\nu(s) = \frac{\W^\nu \etw(s)}{rf^\nu}.
+\end{aligned}
+\end{equation}
+%
+
+Let us define the following
+%
+\begin{equation}\label{eq:dgndef}
+\begin{aligned}
+    \dgn_{m}(s) &= \thb(s) + c_m \etw(s) + \lambda \etw(0), \\
+    \qquad
+    \dgn^\nu_m(s) &= \dgn_m(s) + (-1)^\nu \etw(s),   
+\end{aligned}
+  \implies \quad
+  \pdiff{\CL}{\Wm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \kktm^\nu_{mn} - \kktm^\nu_{mm}.
+\end{equation}
+%
+They should be thought of as collections of vectors rather than matrices, with the first \(m\) a label rather than an index.
+
+The expressions in \cref{eq:finitederivs} have the same form as for \(s=0\), which allowed us to argue that the model with maximal area must have the serial topology in \cref{sec:areapert} and \cite{Lahiri2013synapse}.
+The one missing ingredient is the scale invariance.
+
+However, we do have numerical evidence that the model that maximises the Laplace transform does have the serial topology.
+In \cref{fig:envnum} we see the result of numerical maximisation of \cref{eq:finitelagrangian}.
+We see that restricting to the serial topology does not make the performance any worse.
+In fact, it slightly improves the performance, but this can be ascribed to the numerical optimisation problem being easier.
+Assuming that the optimiser did find the global maximum, this provides evidence for our conjecture that the model that maximises the Laplace transform must have the serial topology.
+The models found numerically are described in \cref{fig:envvid}.
+
+\begin{figure}[tb]
+  \centering
+  % Requires \usepackage{graphicx}
+  \includegraphics[width=0.8\linewidth]{LenvNumOnly.pdf}\\
+  \caption[Numerical envelope for the signal-to-noise ratio]
+  {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
+  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numerically with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
+  The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
+  The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
+  }\label{fig:envnum}
 \end{figure}
@@
 \end{figure}
-\
-
-%================================================================================
-
-\section{Finite time topology}\label{sec:finite}
-
-Now we'll maximise \(\snrb(\tau)\) for some fixed \(\tau\), subject to the constraints \eqref{eq:constr}.
-First, we will maximise the numerator of \eqref{eq:lsnr} holding \(\eq\w\) fixed.
-We can maximise the ratio \wrt \(\eq\w\) later.
-% From this point on, we work in units where \(r = 1\).
-
-For this section we will treat the off-diagonal components of \(\W^\nu\) as the independent degrees of freedom, as in \cref{eq:wdof}, rather than the \(\MMd^\nu\).
-Consider the Lagrangian
-%
-\begin{equation}\label{eq:finitelagrangian}
-\begin{aligned}
-  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i(s) - \etwm_j(s))
-        +\sum_{\nu i j} \kktm^\nu_{ij} \prn{ \Wm^\nu_{ij} + r f^\nu \delta_{ij} }
-        + \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
-\end{aligned}
-\end{equation}
-%
-where \(\Delta p\) is the constant value we are holding \(\eq\w\) at.
-The Karush-Kuhn-Tucker necessary conditions for a maximum are
-%
-\begin{equation}\label{eq:KTfinite}
-  \pdiff{\CL}{\Wm^\nu_{ij}} = 0, \quad
-  \kktm^\nu_{ij} \geq 0, \quad
-  \pdiff{\CL}{\kktm^\nu_{ij}} \geq 0, \quad
-  \kktm^\nu_{ij}\pdiff{\CL}{\kktm^\nu_{ij}} = 0, \quad
-  \pdiff{\CL}{\lambda} = 0.
-\end{equation}
-%
-% We have scaled the KKT multipliers by \(r f^\nu\) for later convenience.
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Scale invariance}\label{sec:scale}
+
+% Under the scale transformation
+% %
+% \begin{equation}\label{eq:scale}
+%   \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I ,
+% \end{equation}
+% %
+The Laplace transform has the following scaling property
+%
+\begin{equation}\label{eq:laplacescale}
+  \hat{A} \prn{\zeta s;\zeta \W^\nu} = \hat{A} \prn{s; \W^\nu}.
+\end{equation}
+%
+At \(s = 0\), it is scale invariant.
+This means that in \cite{Lahiri2013synapse} we could take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \cref{eq:constri}) and use the scale transform to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
+This allowed us to ignore the diagonal constraints.
+We can't do that for \(s > 0\).
+
+The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
+%
+\begin{equation}\label{eq:shiftscale}
+  \zeta {\circledast}_{s} \W^\nu
+      \equiv \zeta \W^\nu
+      + (1 - \zeta) s f^\nu \brk{\I - \onev \eq} .
+\end{equation}
+%
+Then we have
+%
+\begin{equation}\label{eq:lshiftscale}
+  \hat{A} \prn{s; \W^\nu} 
+    = \hat{A} \prn{0, \shift{\W}^\nu(s)}
+    = \hat{A} \prn{0, \zeta \shift{\W}^\nu(s)}
+    = \hat{A} \prn{s, \zeta \circledast_s \W^\nu}.
+\end{equation}
+%
+The problem here is that the lower bound, \(\Wm^\nu_{ij} \geq 0\), 
+is not invariant under this transformation, unlike the \(s = 0\) case.
+Using this transformation to fix the upper bound can break the lower bound.
+
+We might ask what conditions are required so that a model is fixable?
+This means that the \(\zeta \circledast_s \) orbit of the model contains some valid models.
+The conditions are
+%
+\begin{equation*}
+\begin{alignedat}{5}
+  [\zeta \circledast \W^\nu]_{ij}  
+    &= \zeta \Wm^\nu_{ij} - (1-\zeta) s f^\nu \eqm_j 
+    &&\geq 0 &
+  \;\; &\implies \;\; &
+  \zeta &\geq \frac{s \eqm_j}{\Wm^\nu_{ij} /  f^\nu + s \eqm_j} &
+    &\forall \nu, i, j : i \neq j , \\
+  \sum_{j : i \neq j} [\zeta \circledast \W^\nu]_{ij} 
+    &= \zeta \lbm^\nu_i - (1-\zeta) s f^\nu (1-\eqm_i) 
+    &&\leq r f^\nu &
+  &\implies &
+  \zeta &\leq \frac{r + s (1-\eqm_i)}{\lbm^\nu_i / f^\nu + s (1-\eqm_i)} &
+    \quad &\forall \nu, i ,
+\end{alignedat}
+\end{equation*}
+%
+where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \Wm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
+The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
+%
+\begin{equation}\label{eq:fixable}
+  \frac{s \eqm_j}{\Wm^\nu_{ij} / f^\nu + s \eqm_j} \leq \frac{r + s (1-\eqm_k)}{\lbm^\mu_k / f^\nu + s (1-\eqm_k)}
+  \quad \implies \quad
+  \frac{\lbm^\mu_k - r f^\nu}{r + s (1-\eqm_k)} \leq \frac{\Wm^\nu_{ij}}{s \eqm_j} 
+  \quad \forall \mu, \nu, i, j, k : i \neq j.
+\end{equation}
+%
+Note that when \(\Wm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq r f^\nu\), the original diagonal constraint.
+Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
+
+%-------Section------------------------------------------------------------------
+
+\subsection{Shifted problem}\label{sec:lshift}
+
+The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
+treat the off-diagonal components of \(\shift{\W}^\nu \) as the degrees of freedom and maximise the area instead.
+The constraints on \(\shift{\Wm}^\nu_{ij}\) are different from \cref{eq:constri}:
+%
+\begin{equation}\label{eq:lshiftconstr}
+  \shift{\Wm}^\nu_{ij} \geq s f^\nu \eqm_j,
+    \quad \forall \; \nu, i, j : i \neq j, \qquad
+  \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \leq r f^\nu + s f^\nu (1 - \eqm_i).
+    \quad \forall \; \nu, i.
+\end{equation}
+%
+The diagonal elements are still given by \( \shift{\Wm}^\nu_{ii} = - \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \), but now they can be below \(-r f^\nu \) so long as \(\shift{\Wm}^\nu_{ii} \geq -r f^\nu - s f^\nu (1 - \eqm_i)\).
+
+The Lagrangian we should consider is
+%
+\begin{equation}\label{eq:lshiftlagrangian}
+\begin{aligned}
+  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
+        +\sum_{\nu i j} \kktm^\nu_{ij}\prn{\Wm^\nu_{ij} + f^\nu (r + s) \delta_{ij} -  s f^\nu \eqm_j }
+        &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
+\end{aligned}
+\end{equation}
+%
+where all of the first-passage times are evaluated at \(s = 0\).
+
+In \cref{fig:shifted} we show the result of solving this problem with numerical optimisation.
+We see essentially the same results as the original problem, \cref{eq:finitelagrangian,fig:envnum}, as expected.
 
@@
 %
-\begin{equation}\label{eq:finitederivs}
+\begin{equation}\label{eq:lshiftderivs}
 \begin{aligned}
   \pdiff{\CL}{\Wm^\nu_{mn}} = & \,
-    \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
-     + \prn{c_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
+    \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
+     + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
      &+ \kktm^\nu_{mn} -  \kktm^\nu_{mm}
-     + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
+     - \sum_{\mu ij} \kktm^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
     = 0.
@@
 %
-We can solve for the KKT multipliers by noting that
-%
-\begin{equation*}
-  \kktm^\nu_{ij} \Wm^\nu_{ij} = 0,
+We can absorb the new terms by defining
+%
+\begin{equation*}
+\begin{aligned}
+  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\kkt^\nu)\trans \onev,
+  \qquad
+  \shift{\alb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
+  \\
+  \shift{\omb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \etw(0).
+\end{aligned}
+\end{equation*}
+%
+We note that
+%
+\begin{equation*}
+  \kktm^\nu_{ij} (\Wm^\nu_{ij} - s \eqm_j) = 0,
   \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \kktm^\nu_{ii} \Wm^\nu_{ij} 
-      = \kktm^\nu_{ii} ( - \Wm^\nu_{ii}) = r f^\nu \kktm^\nu_{ii}.
-\end{equation*}
-%
-Then, with some new vectors defined:
-%
-\begin{equation}\label{eq:kktsol}
+  \sum_{j: j \neq i} \kktm^\nu_{ii} (\Wm^\nu_{ij} - s \eqm_j)
+      = \kktm^\nu_{ii} (1 - \Wm^\nu_{ii} - s (1 - \eqm_i)) = \kktm^\nu_{ii},
+\end{equation*}
+%
+then we can ``solve'' for the KKT multipliers:
+%
+\begin{equation}\label{eq:lshiftsol}
+\begin{aligned}
+  \kktm^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
+    }, \\
+  \kktm^\nu_{mn} &= \eqm_m \prn{ 
+      \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
+  % \\ & \hphantom{= - \eqm_m \big( } 
+    + \brk{ {c}_m + (-1)^\nu + \lambda } 
+        \brk{ \etwm_n - \etwm_m - \shift{\ombm}^\nu_m(s) }
+     }, \\
+\end{aligned}
+\end{equation}
+%
+although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\), so these expressions are not explicit.
+
+While this problem involves the scale invariant quantity \(\hat{A}(0)\), neither the upper nor lower bounds in \cref{eq:lshiftconstr} are invariant.
+We then run into the same problems in proving our serial-topology hypothesis.
+
+\begin{figure}[ht]
+\begin{center}
+\begin{myenuma}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{LenvShift.pdf}}
+  \label{fig:envshift}
+  \item \aligntop{\includegraphics[width=0.4\linewidth]{shift_cond.pdf}}
+  \label{fig:shiftcond}
+\end{myenuma}
+\caption[Shifted problem vs.\ original]{Shifted problem vs.\ original.
+(\ref*{fig:envshift}) Comparison of numerical optimisation of the shifted problem in \cref{eq:lshiftlagrangian,fig:envnum} with the original problem in \cref{eq:finitelagrangian}.
+We see that the two approaches produce very similar numerical envelopes.
+The shifted problem involves non-linear constraints and is thus harder to solve.
+We used a limit on the condition number of \(\fund\) of \(10^3\) as a constraint during optimisation.
+(\ref*{fig:shiftcond}) Plot of the discrepancy in (\ref*{fig:envshift}) against the condition number of the fundamental matrix, \(\fund \).
+We see that most of the large discrepancies occur where the problem is numerically ill-conditioned.
+\label{fig:shifted}}
+\end{center}
+\end{figure}
+
+%-------Section------------------------------------------------------------------
+
+subsection{Diagonal constraints}\label{sec:diagKKT}
+
+In \cref{eq:kktsol} we saw that
+%
+\begin{equation*}
 \begin{aligned}
@@
     }, \\
-  \kktm^\nu_{mn} &= \eqm_m \big( 
-      \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
-    + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
-  \\ &\hphantom{= \eqm_m \big( }
-    + \brk{{c}_m(s) + (-1)^\nu} 
-        \brk{ \etwm_n(s) - \etwm_m(s) - \ombm^\nu_m(s) }
-     \big), \\
   \text{where} \quad 
@@
 \end{aligned}
-\end{equation}
+\end{equation*}
 %
@@
 where \(\delta\) indicates mean subtraction and \(\eq \mathbf{a}(s) = \hat{A}(s)\).
-
-The expressions in \cref{eq:finitederivs} have the same form as for \(s=0\), which allowed us to argue that the model with maximal area must have the serial topology in \cref{sec:areapert} and \cite{Lahiri2013synapse}.
-The one missing ingredient is the scale invariance.
-
-However, we do have numerical evidence that the model that maximises the Laplace transform does have the serial topology.
-In \cref{fig:envnum} we see the result of numerical maximisation of \cref{eq:finitelagrangian}.
-We see that restricting to the serial topology does not make the performance any worse.
-In fact, it slightly improves the performance, but this can be ascribed to the numerical optimisation problem being easier.
-Assuming that the optimiser did find the global maximum, this provides evidence for our conjecture that the model that maximises the Laplace transform must have the serial topology.
-The models found numerically are described in \cref{fig:envvid}.
-
-\begin{figure}[tb]
-  \centering
-  % Requires \usepackage{graphicx}
-  \includegraphics[width=0.8\linewidth]{LenvNumOnly.pdf}\\
-  \caption[Numerical envelope for the signal-to-noise ratio]
-  {Numerical envelope for the signal-to-noise ratio, \eqref{eq:snrb}. 
-  \Cref{eq:finitelagrangian,eq:KTfinite,eq:finitederivs} are solved numerically with \(f\potdep=\frac{1}{2}\), the number of synapses \(N=1\) and the number of states \(M=10\).
-  The orange line indicates maximisation of the memory curve over the space of all matrices satisfying the constraints \eqref{eq:constr}.
-  The green line indicates maximisation over models with the serial topology, \ie potentiation only moves up one state and depression only moves down one state.
-  }\label{fig:envnum}
-\end{figure}
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Scale invariance}\label{sec:scale}
-
-% Under the scale transformation
-% %
-% \begin{equation}\label{eq:scale}
-%   \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I ,
-% \end{equation}
-% %
-The Laplace transform has the following scaling property
-%
-\begin{equation}\label{eq:laplacescale}
-  \hat{A} \prn{\zeta s;\zeta \W^\nu} = \hat{A} \prn{s; \W^\nu}.
-\end{equation}
-%
-At \(s = 0\), it is scale invariant.
-This means that in \cite{Lahiri2013synapse} we could take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \cref{eq:constri}) and use the scale transform to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
-This allowed us to ignore the diagonal constraints.
-We can't do that for \(s > 0\).
-
-The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
-%
-\begin{equation}\label{eq:shiftscale}
-  \zeta {\circledast}_{s} \W^\nu
-      \equiv \zeta \W^\nu
-      + (1 - \zeta) s f^\nu \brk{\I - \onev \eq} .
-\end{equation}
-%
-Then we have
-%
-\begin{equation}\label{eq:lshiftscale}
-  \hat{A} \prn{s; \W^\nu} 
-    = \hat{A} \prn{0, \shift{\W}^\nu(s)}
-    = \hat{A} \prn{0, \zeta \shift{\W}^\nu(s)}
-    = \hat{A} \prn{s, \zeta \circledast_s \W^\nu}.
-\end{equation}
-%
-The problem here is that the lower bound, \(\Wm^\nu_{ij} \geq 0\), 
-is not invariant under this transformation, unlike the \(s = 0\) case.
-Using this transformation to fix the upper bound can break the lower bound.
-
-We might ask what conditions are required so that a model is fixable?
-This means that the \(\zeta \circledast_s \) orbit of the model contains some valid models.
-The conditions are
-%
-\begin{equation*}
-\begin{alignedat}{5}
-  [\zeta \circledast \W^\nu]_{ij}  
-    &= \zeta \Wm^\nu_{ij} - (1-\zeta) s f^\nu \eqm_j 
-    &&\geq 0 &
-  \;\; &\implies \;\; &
-  \zeta &\geq \frac{s \eqm_j}{\Wm^\nu_{ij} /  f^\nu + s \eqm_j} &
-    &\forall \nu, i, j : i \neq j , \\
-  \sum_{j : i \neq j} [\zeta \circledast \W^\nu]_{ij} 
-    &= \zeta \lbm^\nu_i - (1-\zeta) s f^\nu (1-\eqm_i) 
-    &&\leq r f^\nu &
+We can write \cref{eq:diagKKT} as
+%
+\begin{equation*}
+\begin{aligned}
+  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+    \left[ 
+      \hat{A}
+      - (c_m + (-1)^\nu) (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m
+    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
+    % \left.
+      - (-1)^\nu [\encm \dgnm_m]_m
+    \right],
+\end{aligned}
+\end{equation*}
+%
+with all quantities evaluated at \(s\) unless labelled otherwise.
+
+From \cref{eq:oneexit,eq:twoexits} we see that
+%
+\begin{equation*}
+\begin{gathered}
+  \begin{aligned}
+    \W^\nu_{mi} &> 0 &
+    \quad &\implies \quad &
+    \dgnm^\nu_{m,i} &= \dgnm^\nu_{m*}
+    \\
+    \sum_{k \neq m} \W^\nu_{mk} &< r f^\nu &
     &\implies &
-  \zeta &\leq \frac{r + s (1-\eqm_i)}{\lbm^\nu_i / f^\nu + s (1-\eqm_i)} &
-    \quad &\forall \nu, i ,
-\end{alignedat}
-\end{equation*}
-%
-where \(\lbm^\nu_i \equiv \sum_{j : i \neq j} \Wm^\nu_{ij}\) is the inverse mean  lifetime of state \(i\).
-The model is fixable if it is possible to find a \(\zeta\) that satisfies all of the above conditions:
-%
-\begin{equation}\label{eq:fixable}
-  \frac{s \eqm_j}{\Wm^\nu_{ij} / f^\nu + s \eqm_j} \leq \frac{r + s (1-\eqm_k)}{\lbm^\mu_k / f^\nu + s (1-\eqm_k)}
-  \quad \implies \quad
-  \frac{\lbm^\mu_k - r f^\nu}{r + s (1-\eqm_k)} \leq \frac{\Wm^\nu_{ij}}{s \eqm_j} 
-  \quad \forall \mu, \nu, i, j, k : i \neq j.
-\end{equation}
-%
-Note that when \(\Wm^\nu_{ij} = 0\) for some \(i,j \), this condition reduces to \(\lbm^\mu_k \leq r f^\nu\), the original diagonal constraint.
-Thus, relaxing the constraints from ``the model is valid'' to ``the model is fixable'' doesn't help us prove anything about the topology.
-
-%-------Section------------------------------------------------------------------
-
-\subsection{Shifted problem}\label{sec:lshift}
-
-The relation of the Laplace transform to the shifted process from \cref{eq:lshiftpd,eq:lareashift} suggests an alternative approach: 
-treat the off-diagonal components of \(\shift{\W}^\nu \) as the degrees of freedom and maximise the area instead.
-The constraints on \(\shift{\Wm}^\nu_{ij}\) are different from \cref{eq:constri}:
-%
-\begin{equation}\label{eq:lshiftconstr}
-  \shift{\Wm}^\nu_{ij} \geq s f^\nu \eqm_j,
-    \quad \forall \; \nu, i, j : i \neq j, \qquad
-  \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \leq r f^\nu + s f^\nu (1 - \eqm_i).
-    \quad \forall \; \nu, i.
-\end{equation}
-%
-The diagonal elements are still given by \( \shift{\Wm}^\nu_{ii} = - \sum_{j: j \neq i} \shift{\Wm}^\nu_{ij} \), but now they can be below \(-r f^\nu \) so long as \(\shift{\Wm}^\nu_{ii} \geq -r f^\nu - s f^\nu (1 - \eqm_i)\).
-
-The Lagrangian we should consider is
-%
-\begin{equation}\label{eq:lshiftlagrangian}
-\begin{aligned}
-  \CL =&\, \sum_{ij} \eqm_i \encm_{ij} (\etwm_i - \etwm_j)
-        +\sum_{\nu i j} \kktm^\nu_{ij}\prn{\Wm^\nu_{ij} + f^\nu (r + s) \delta_{ij} -  s f^\nu \eqm_j }
-        &+ \lambda \prn{ \Delta p - \sum_i \eqm_i \wm_i },
-\end{aligned}
-\end{equation}
-%
-where all of the first-passage times are evaluated at \(s = 0\).
-
-In \cref{fig:shifted} we show the result of solving this problem with numerical optimisation.
-We see essentially the same results as the original problem, \cref{eq:finitelagrangian,fig:envnum}, as expected.
-
-The derivatives, for \(m \neq n\) are
-%
-\begin{equation}\label{eq:lshiftderivs}
-\begin{aligned}
-  \pdiff{\CL}{\Wm^\nu_{mn}} = & \,
-    \eqm_m \prn{ \brk{\thbm_m - \thbm_n}
-     + \prn{{c}_m + (-1)^\nu + \lambda} \brk{\etwm_m - \etwm_n} } \\
-     &+ \kktm^\nu_{mn} -  \kktm^\nu_{mm}
-     - \sum_{\mu ij} \kktm^\mu_{ij} s f^\mu \eqm_m (\fptbm_{mj} - \fptbm_{nj}) \eqm_j
-    = 0.
-\end{aligned}
-\end{equation}
-%
-We can absorb the new terms by defining
-%
-\begin{equation*}
-\begin{aligned}
-  \shift{\thb}^\nu(s) = \thb(0) - s f^\nu \fptb(0) \Eq (\kkt^\nu)\trans \onev,
-  \qquad
-  \shift{\alb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \shift{\thb}^\nu(s),
+    \dgnm^\nu_{m,m} &= \dgnm^\nu_{m*},
+  \end{aligned}
   \\
-  \shift{\omb}^\nu(s) &= (\W^\nu + (s - 1) \I - s \onev \eq) \etw(0).
-\end{aligned}
-\end{equation*}
-%
-We note that
-%
-\begin{equation*}
-  \kktm^\nu_{ij} (\Wm^\nu_{ij} - s \eqm_j) = 0,
-  \qquad \text{and} \qquad
-  \sum_{j: j \neq i} \kktm^\nu_{ii} (\Wm^\nu_{ij} - s \eqm_j)
-      = \kktm^\nu_{ii} (1 - \Wm^\nu_{ii} - s (1 - \eqm_i)) = \kktm^\nu_{ii},
-\end{equation*}
-%
-then we can ``solve'' for the KKT multipliers:
-%
-\begin{equation}\label{eq:lshiftsol}
-\begin{aligned}
-  \kktm^\nu_{mm} &= - \eqm_m \prn{ \shift{\albm}^\nu_m(s) 
-    + \brk{ {c}_m + (-1)^\nu + \lambda } \shift{\ombm}^\nu_m(s)
-    }, \\
-  \kktm^\nu_{mn} &= \eqm_m \prn{ 
-      \brk{ \shift{\thbm}^\nu_n(s) - \shift{\thbm}^\nu_m(s) - \shift{\albm}^\nu_m(s) }
-  % \\ & \hphantom{= - \eqm_m \big( } 
-    + \brk{ {c}_m + (-1)^\nu + \lambda } 
-        \brk{ \etwm_n - \etwm_m - \shift{\ombm}^\nu_m(s) }
-     }, \\
-\end{aligned}
-\end{equation}
-%
-although there are KKT multipliers hiding in \(\shift{\thb}^\nu(s)\) and \(\shift{\alb}^\nu(s)\), so these expressions are not explicit.
-
-While this problem involves the scale invariant quantity \(\hat{A}(0)\), neither the upper nor lower bounds in \cref{eq:lshiftconstr} are invariant.
-We then run into the same problems in proving our serial-topology hypothesis.
-
-\begin{figure}[ht]
-\begin{center}
-\begin{myenuma}
-  \item \aligntop{\includegraphics[width=0.4\linewidth]{LenvShift.pdf}}
-  \label{fig:envshift}
-  \item \aligntop{\includegraphics[width=0.4\linewidth]{shift_cond.pdf}}
-  \label{fig:shiftcond}
-\end{myenuma}
-\caption[Shifted problem vs.\ original]{Shifted problem vs.\ original.
-(\ref*{fig:envshift}) Comparison of numerical optimisation of the shifted problem in \cref{eq:lshiftlagrangian,fig:envnum} with the original problemin \cref{eq:finitelagrangian}.
-We see that the two approaches produce very similar numerical envelopes.
-The shifted problem involves non-linear constraints and is thus harder to solve.
-(\ref{fig:shiftcond}) Plot of the discrepancy in (\ref*{fig:envshift}) against the condition number of the fundamental matrix, \(\fund \).
-We see that most of the large discrepancies occur where the problem is numerically ill-conditioned.
-\label{fig:shifted}}
-\end{center}
-\end{figure}
+  \begin{aligned}
+    \implies \quad
+    [\W^\nu \dgn_m]_m
+      &= \sum_k \Wm^\nu_{mk}  \dgnm_{m,k}
+      = \sum_k \Wm^\nu_{mk}
+        (\dgnm^\nu_{m,k} - (-1)^\nu \etwm_k) \\
+      &= \sum_{k \neq m} \Wm^\nu_{mk} 
+        (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,m}) 
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
+        \sum_{k \neq m} \Wm^\nu_{mk}
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= r f^\nu (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= r f^\nu (\dgnm_{m*} - \dgnm_{m,m}) 
+        + (-1)^\nu \prn{
+          r f^\nu (\etwm_{m*} - \etwm_m)
+          - [\W^\nu \etw]_m
+        },
+  \end{aligned}
+\end{gathered}
+\end{equation*}
+%
+where we used the fact that \( (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) \) is only nonzero when \( \sum_{k \neq m} \Wm^\nu_{mk} = rf^\nu \).
+Therefore
+%
+\begin{equation*}
+\begin{aligned}
+  [\enc \dgn_m]_m
+    &= \sum_{\nu} (-1)^\nu rf^\nu 
+      (\dgnm_{m*} - \dgnm_{m,m}) 
+      + \sum_{\nu} \prn{
+        r f^\nu (\etwm_{m*} - \etwm_m)
+        -  [\W^\nu \etw]_m
+      } \\
+    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + r(\etwm_{m*} - \etwm_m)
+      - [\frg \etw]_m \\
+    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + r(\etwm_{m*} - \etwm_m)
+      - (\delta w_m + s \delta \etwm_m).
+\end{aligned}
+\end{equation*}
+%
+Then we can also write \cref{eq:diagKKT} as
+%
+\begin{equation}\label{eq:diagKKTz}
+\begin{aligned}
+  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+    \left[ 
+      \hat{A}
+      - c_m (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m
+    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
+    % \left.
+      + (-1)^\nu \prn{ 
+        \Delta f (\dgnm_{m,m} - \dgnm_{m*}) 
+        + (\etwm_m - \etwm_{m*})
+      }
+    \right],
+\end{aligned}
+\end{equation}
+%
+again with all quantities evaluated at \(s\) unless labelled otherwise.
+
+\Cref{eq:KKTz} will provide another expression for \(\kktm^\nu_{mm}\).
+That expression will agree with \cref{eq:diagKKTz} provided that
+%
+\begin{equation}\label{eq:diagAgree}
+    \hat{A} - c_m (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m 
+      = (\dgnm_{m,m} - \dgnm_{m*}) 
+        + \Delta f (\etwm_m - \etwm_{m*}).
+\end{equation}
+%
 
@@
 
-Let us define the following
-%
-\begin{equation}\label{eq:degenerate}
-\begin{aligned}
-    \dgn_{m}(s) &= \thb(s) + c_m \etw(s) + \lambda \etw(0), \\
-    \qquad
-    \dgn^\nu_m(s) &= \dgn_m(s) + (-1)^\nu \etw(s),   
-\end{aligned}
-  \implies \quad
-  \pdiff{\CL}{\Wm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \kktm^\nu_{mn} - \kktm^\nu_{mm}.
-\end{equation}
-%
-They should be thought of as collections of vectors rather than matrices, with the first \(m\) a label rather than an index.
-
+The arguments in this section will break down if \(\eqm_m = 0\).
 Suppose that a state \(n\) has a nonzero transition rate from state \(m\) for one \(\nu\).
@@
 \begin{lgathered}  
-  \dgnm\pot_{m*} = \dgnm\pot_{m,j} = \min_k \dgnm\pot_{m,k},
-  \qquad
-  \dgnm\dep_{m*} = \dgnm\dep_{m,i} = \min_k \dgnm\dep_{m,k},
+  \dgnm\pot_{m*} = \dgnm\pot_{m,j} = \min_{k\neq m} \dgnm\pot_{m,k},
+  \qquad
+  \dgnm\dep_{m*} = \dgnm\dep_{m,i} = \min_{k\neq m} \dgnm\dep_{m,k},
   \\
@@
 
-\subsubsection{Diagonal constraints}\label{sec:diagKKT}
-
-We can write \cref{eq:diagKKT} as
-%
-\begin{equation*}
-\begin{aligned}
-  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
-    \left[ 
-      \hat{A}
-      - (c_m + (-1)^\nu) (\delta \wm_m + s \delta \etwm_m)
-      - \lambda \delta w_m
-    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
-    % \left.
-      - (-1)^\nu [\encm \dgnm_m]_m
-    \right],
-\end{aligned}
-\end{equation*}
-%
-with all quantities evaluated at \(s\) unless labelled otherwise.
-
-From \cref{eq:oneexit,eq:twoexits} we see that
-%
-\begin{equation*}
-\begin{gathered}
-  \begin{aligned}
-    \W^\nu_{mi} &> 0 &
-    \quad &\implies \quad &
-    \dgnm^\nu_{m,i} &= \dgnm^\nu_{m*}
-    \\
-    \sum_{k \neq m} \W^\nu_{mk} &< r f^\nu &
-    &\implies &
-    \dgnm^\nu_{m,m} &= \dgnm^\nu_{m*},
-  \end{aligned}
-  \\
-  \begin{aligned}
-    \implies \quad
-    [\W^\nu \dgn_m]_m
-      &= \sum_k \Wm^\nu_{mk}  \dgnm_{m,k}
-      = \sum_k \Wm^\nu_{mk}
-        (\dgnm^\nu_{m,k} - (-1)^\nu \etwm_k) \\
-      &= \sum_{k \neq m} \Wm^\nu_{mk} 
-        (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,m}) 
-        - (-1)^\nu [\W^\nu \etw]_m \\
-      &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
-        \sum_{k \neq m} \Wm^\nu_{mk}
-        - (-1)^\nu [\W^\nu \etw]_m \\
-      &= r f^\nu (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
-        - (-1)^\nu [\W^\nu \etw]_m \\
-      &= r f^\nu (\dgnm_{m*} - \dgnm_{m,m}) 
-        + (-1)^\nu \prn{
-          r f^\nu (\etwm_{m*} - \etwm_m)
-          - [\W^\nu \etw]_m
-        },
-  \end{aligned}
-\end{gathered}
-\end{equation*}
-%
-where we used the fact that \( (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) \) is only nonzero when \( \sum_{k \neq m} \Wm^\nu_{mk} = rf^\nu \).
-Therefore
-%
-\begin{equation*}
-\begin{aligned}
-  [\enc \dgn_m]_m
-    &= \sum_{\nu} (-1)^\nu rf^\nu 
-      (\dgnm_{m*} - \dgnm_{m,m}) 
-      + \sum_{\nu} \prn{
-        r f^\nu (\etwm_{m*} - \etwm_m)
-        -  [\W^\nu \etw]_m
-      } \\
-    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
-      + r(\etwm_{m*} - \etwm_m)
-      - [\frg \etw]_m \\
-    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
-      + r(\etwm_{m*} - \etwm_m)
-      - (\delta w_m + s \delta \etwm_m).
-\end{aligned}
-\end{equation*}
-%
-Then we can also write \cref{eq:diagKKT} as
-%
-\begin{equation}\label{eq:diagKKTz}
-\begin{aligned}
-  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
-    \left[ 
-      \hat{A}
-      - c_m (\delta \wm_m + s \delta \etwm_m)
-      - \lambda \delta w_m
-    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
-    % \left.
-      + (-1)^\nu \prn{ 
-        \Delta f (\dgnm_{m,m} - \dgnm_{m*}) 
-        + (\etwm_m - \etwm_{m*})
-      }
-    \right],
-\end{aligned}
-\end{equation}
-%
-again with all quantities evaluated at \(s\) unless labelled otherwise.
-
-\Cref{eq:KKTz} provides another expression for \(\kktm^\nu_{mm}\).
-That expression will agree with \cref{eq:diagKKTz} provided that
-%
-\begin{equation}\label{eq:diagAgree}
-    \hat{A} - c_m (\delta \wm_m + s \delta \etwm_m)
-      - \lambda \delta w_m 
-      = (\dgnm_{m,m} - \dgnm_{m*}) 
-        + \Delta f (\etwm_m - \etwm_{m*}).
-\end{equation}
-%
-
-%-------Section------------------------------------------------------------------
-
 \subsection{Scaling with a floating cutoff}\label{sec:cutoff}
@@
 In \cref{eq:kktsol} we have an expression for the KKT multipliers at the maximum in terms of \(s\) and \(\W \).
-We denote this expression by \( \kkt^\ddagger(s, \W) \).
-If we denote the solution to the maximisation problem by \(\W^\ast(s,r), \kkt^\ast(s,r)\), we have \( \kkt^\ast(s,r) = \kkt^\ddagger(s, \W^\ast(s,r)) \) and
+We denote this expression by \( \kkt^\ddagger(s, r, \W) \)
+(note there is \(s\) dependance through \(\thb(s), \etw(s)\) \etc and \(r\) dependance through \(\alb\) and \(\omb\)).
+If we denote the solution to the maximisation problem by \(\W^\ast(s,r), \kkt^\ast(s,r)\), we have \( \kkt^\ast(s,r) = \kkt^\ddagger(s, r, \W^\ast(s,r)) \) and
 %
@@
 \begin{aligned}
-  \kkt^\ddagger(\lambda s, \lambda \W) 
-    &= \lambda\inv \kkt^\ddagger(s, \W), &
-  \qquad \text{or} \qquad
-  s \pdiffc<\Wm>{\kkt^\ddagger}{s} 
-      + \sum_{\nu i j} \Wm^\nu_{ij} \pdiffc<s>{\kkt^\ddagger}{\Wm^\nu_{ij}} 
+  \kkt^\ddagger(\lambda s, \lambda r, \lambda \W) 
+    &= \lambda\inv \kkt^\ddagger(s, r, \W), &
+  \quad \text{or} \quad
+  s \pdiffc<r \mathrlap{\Wm}>{\kkt^\ddagger}{s} +
+  r \pdiffc<s \mathrlap{\Wm}>{\kkt^\ddagger}{r} 
+      + \sum_{\nu i j} \Wm^\nu_{ij} \pdiffc<s r>{\kkt^\ddagger}{\Wm^\nu_{ij}} 
     &= -\kkt^\ddagger, \\
@@
   \sum_k^{i:j} f(k) 
-    \equiv \sgn(j-i) \smashoperator{\sum_{k = \min(i,j)}^{\max(i,j)-1}} f(k)
+    \equiv \sum_{k = 1}^{j-1} f(k) - \sum_{k = 1}^{i-1} f(k)
+    = \sum_{k = i}^{M} f(k) - \sum_{k = j}^{M} f(k)
+    % = \sgn(j-i) \smashoperator{\sum_{k = \min(i,j)}^{\max(i,j)-1}} f(k)
     = \begin{cases}
@@
 \begin{aligned}
-  \etwm_i - \etwm_j &= \frac{z}{\delta} \sum_k^{i:j} \frac{1}{q_k}, & \qquad
+  && \mathllap
+  {\etwm_i - \etwm_j} &= \frac{z}{\delta} \sum_k^{i:j} \frac{1}{q_k}, & \quad
   \thbm_i - \thbm_j &= \frac{z}{\delta} \sum_k^{i:j} \frac{M - 2k}{q_k}, \\
-  a_i &= \frac{1}{\eqm_i} 
-    \quad \text{for } 2 \leq i \leq M-1, &
+    i &\in [2, M-1]: &
+  a_i &= \frac{1}{\eqm_i}, &
   a_1 &= a_M = 1, \\
-  \alpha^\nu_i &= - \frac{z ((-1)^\nu c_i + 1)}{r f^\nu \delta}
-    \quad \text{for } 2 \leq i \leq M-1, &
-  \alpha\pot_1 &= \frac{z(M-2)}{r f^\nu q_1}, \\
-  \alpha\pot_M &= \alpha\dep_1 = 0, &
-  \alpha\dep_M &= \frac{z(M-2)}{r f^\nu q_{M-1}}, \\
-  \omega^\nu_i &= \frac{(-1)^\nu z}{r f^\nu \delta}
-    \quad \text{for } 2 \leq i \leq M-1, &
-  \omega\pot_1 &= \frac{z}{r f^\nu q_1}, \\
-  \omega\pot_M &= \omega\dep_1 = 0, &
-  \omega\dep_M &= - \frac{z}{r f^\nu q_{M-1}}.
+    i &\in [2, M-1]: &
+  \albm^\nu_i &= - \frac{z ((-1)^\nu c_i + 1)}{r f^\nu \delta}, &
+  \albm\pot_1 &= \frac{z(M-2)}{r f^\nu q_1}, \\
+  &&
+  \albm\pot_M &= \albm\dep_1 = 0, &
+  \albm\dep_M &= \frac{z(M-2)}{r f^\nu q_{M-1}}, \\
+    i &\in [2, M-1]: &
+  \ombm^\nu_i &= \frac{(-1)^\nu z}{r f^\nu \delta}, &
+  \ombm\pot_1 &= \frac{z}{r f^\nu q_1}, \\
+  &&
+  \ombm\pot_M &= \ombm\dep_1 = 0, &
+  \ombm\dep_M &= - \frac{z}{r f^\nu q_{M-1}}.
 \end{aligned}
@@
 \begin{aligned}
-    \kktm^\nu_{ij} = \sum_k^{i:j} \frac{k - i + \delta\dep_\nu}{q_k}
-      \quad \text{ for } 2 \leq i \leq M-1, \qquad
+    i &\in [2, M-1]: &
+    \kktm^\nu_{ij} &= \sum_k^{i:j} \frac{k - i + \delta\dep_\nu}{q_k}, &
     \kktm^\nu_{1j} &= \frac{1}{\delta} \sum_{k=1}^{j-1} \frac{k - 1 + \delta\dep_\nu}{q_k},
-      \\
+      \\ &&&&
     \kktm^\nu_{Mj} &= \frac{1}{\delta} \sum_{k=j}^{M-1} \frac{M - k - \delta\dep_\nu}{q_k}.
@@
 the sum contains exactly one term and that term is zero
-(It is impossible for more than one term in the sum to be zero).
+(it is impossible for more than one term in the sum to be zero).
 The first occurs for \(j = i\), the second for \( j = i + (-1)^\nu \).
 Those transitions define the serial topology.
+
+Unfortunately, arguments based on continuity are not enough to reach the \(s=\epsilon\) line in \cref{fig:planes}\ref{fig:scaling_plane}.
+The largest \(\epsilon\) we could use would be proportional to \(r\) with a coefficient \( \sim \CO(\delta) \).
+The bottom end of the line would have to curve towards the origin and fail to intersect every ray.
+In essence, continuity can get us to a nearby ray, not a nearby vertical line.
 

### 2021-11-20 Notes/laplace.tex
start degeneracy section

@@
 
-This leaves two possibilities for the maximum area Markov chain.
+%-------Section------------------------------------------------------------------
+
+\subsubsection{Degeneracy}\label{sec:degeneracy_area}
+
+\hyperref[sec:areapert]{The previous section} leaves two possibilities for the maximum area Markov chain.
 Either there is no degeneracy and no shortcuts, which implies the serial topology that we'll discuss in \cref{sec:serialarea}, or there is some degeneracy, which would allow shortcuts provided that they do not bypass an entire degenerate set (see \cref{eq:shortcutarea}).
@@
 As this lumped chain has the same area (see \cite{Lahiri2013synapse}), we would need only consider the serial topology.
+
+We can almost see that degeneracy implies lumpability as follows:
+suppose there is a nonzero transition between states \(i\) and \(j\) of type \(\nu\),
+Then \(\kktm^\nu_{ij} = 0\) and \cref{eq:areaderivs} becomes
+%
+\begin{equation}\label{eq:dgn_theta}
+\begin{gathered}
+  \pdiff{\CL}{\MMdm^\nu_{ij}} =
+    r f^\nu \eqm_i \prn{ \brk{\thbm_i - \thbm_j}
+     + \prn{c_i + (-1)^\nu + \lambda} \brk{\etwm_i - \etwm_j} } 
+    = 0, \\
+    \implies \quad
+    r f^\nu \eqm_i = 0
+    \qquad \text{or} \qquad
+    \thbm_i - \thbm_j = -  \prn{c_i + (-1)^\nu + \lambda} \brk{\etwm_i - \etwm_j}. 
+\end{gathered}
+\end{equation}
+%
+Therefore, provided that state \(i\) has nonzero equilibrium probability, degeneracy in \(\etw\) implies degeneracy in \(\thb\).
+
+Let \(C_a\) denote the \(a\)'th clump of states that are degenerate in \(\etw\) and \(\thb\), in order of decreasing \(\etw\):
 

### 2021-11-22 Notes/laplace.tex
unrearranged a bit
degeneracy section

@@
 %
-Therefore, provided that state \(i\) has nonzero equilibrium probability, degeneracy in \(\etw\) implies degeneracy in \(\thb\).
-
-Let \(C_a\) denote the \(a\)'th clump of states that are degenerate in \(\etw\) and \(\thb\), in order of decreasing \(\etw\):
+Therefore, provided that state \(i\) has nonzero equilibrium probability, degeneracy in \(\etw\) implies degeneracy in \(\thb\) for an otimal model.
+
+Let \(C_a\) denote the \(a\)'th clump of states that are degenerate in \(\etw\) and \(\thb\), in order of decreasing \(\etw\).
+The arguments from \cref{sec:areapert} tell us that the transitions in \(\W\pot\) from states in \(C_a\) can only go to states in \(C_{a+1}\).%
+\footnote{We haven't said anything about intra-clump transitions yet.
+Perturbing such transitions does not break the degeneracy, so these perturbations can be extended finitely.
+Those perturbations also leave the area invariant, so they are completely null degrees of freedom.
+We can choose to set all intra-clump transitions to zero.}
+Similarly, \(\W\dep\) can only go from states in \(C_a\) to states in \(C_{a-1}\).
+
+Suppose we have two states, \(i,j\), in \(C_a\), and two states, \(k,l\), in \(C_{a+1}\), with \(\Wm\pot_{ik}, \Wm\pot_{jl} > 0\).
+Then \cref{eq:dgn_theta} tells us that
+%
+\begin{equation}\label{eq:dgn_c}
+\begin{aligned}
+  \thbm_i - \thbm_k &= - \prn{c_i + 1 + \lambda} \brk{\etwm_i - \etwm_k} 
+  \quad \implies & 
+  \thbm_a - \thbm_{a+1} &= - \prn{c_i + 1 + \lambda} \brk{\etwm_a - \etwm_{a+1}}, \\
+  \thbm_j - \thbm_l &= - \prn{c_j + 1 + \lambda} \brk{\etwm_j - \etwm_l}
+  \quad \implies & 
+  \thbm_a - \thbm_{a+1} &= - \prn{c_j + 1 + \lambda} \brk{\etwm_a - \etwm_{a+1}}, \\
+  && \implies \quad 
+  0 &= \prn{c_i - c_j} \brk{\etwm_a - \etwm_{a+1}}. \\
+\end{aligned}
+\end{equation}
+%
+Therefore \(\mathbf{c}\) has the same degeneracy as \(\etw\) and \(\thb\).
+If we make the choice \(\wm_i = \sgn c_i\) (as discussed after \cref{eq:lareacoeffs}) we see that \(\w\) also has the same degeneracy.
+A result similar to \cref{eq:dgn_c} holds with \( \text{pot} \leftrightarrow \text{dep} \).
+
+Note that, for any state \(i \in C_a\)
+%
+\begin{equation*}
+\begin{aligned}
+  [\W\pot \etw]_i &= \sum_{j \in C_{a+1}} \Wm\pot_{ij} (\etwm_j - \etwm_i)
+    = (\etwm_{a+1} - \etwm_a) \sum_{\mathclap{j \in C_{a+1}}} \Wm\pot_{ij}
+    = (\etwm_{a+1} - \etwm_a) \lambda\pot_i, \\
+  [\W\dep \etw]_i &= (\etwm_{a-1} - \etwm_a) \lambda\dep_i,
+  \quad\text{where} \quad
+  \lambda^\nu_i =  \sum_{j \neq i} \Wm^\nu_{ij}.
+\end{aligned}
+\end{equation*}
+%
+which also applies wth \(\etw\) replaced with \(\thb\) or \(\mathbf{c}\).
+All models, not just the optimal ones, satisfy
+%
+\begin{equation*}
+\begin{aligned}
+  \frg \thb(0) &= (\I - \onev \eq) \mathbf{a} = - \enc \etw(0) - \CA,
+  \\
+  \frg \etw(0) &= (\I - \onev \eq) \w = \w - \Delta p\, \onev.
+\end{aligned}
+\end{equation*}
+%
+Using the results above
+%
+\begin{equation}\label{eq:dgn_lump}
+\begin{gathered}
+\begin{aligned}
+  (\thbm_{a+1} - \thbm_a) \lambda\pot_i + (\thbm_{a-1} - \thbm_a) \lambda\dep_i
+    &= - (\etwm_{a+1} - \etwm_a) \lambda\pot_i + (\etwm_{a-1} - \etwm_a) \lambda\dep_i - \CA,
+  \\
+  (\etwm_{a+1} - \etwm_a) \lambda\pot_i + (\etwm_{a-1} - \etwm_a) \lambda\dep_i &= \wm_a - \Delta p,
+\end{aligned}
+\\
+  \implies
+\begin{aligned}
+  \lambda\pot_i &= \frac{(\CA - \wm_a + \Delta p) (\etwm_a - \etwm_{a-1}) 
+                        + (\wm_a - \Delta p) (\thbm_a - \thbm_{a-1})}
+                      {(\etwm_{a+1} - \etwm_a) (\thbm_a - \thbm_{a-1})
+                        - (\etwm_a - \etwm_{a-1}) (\thbm_{a+1} - \thbm_a)
+                        - 2 (\etwm_{a+1} - \etwm_a) (\etwm_a - \etwm_{a-1})}, \\
+  \lambda\dep_i &= \frac{(\CA + \wm_a - \Delta p) (\etwm_{a+1} - \etwm_a) 
+                        + (\wm_a - \Delta p) (\thbm_{a+1} - \thbm_a)}
+                      {(\etwm_{a+1} - \etwm_a) (\thbm_a - \thbm_{a-1})
+                        - (\etwm_a - \etwm_{a-1}) (\thbm_{a+1} - \thbm_a)
+                        - 2 (\etwm_{a+1} - \etwm_a) (\etwm_a - \etwm_{a-1})}.
+\end{aligned}
+\end{gathered}
+\end{equation}
+%
+Because the right-hand sides above depend on \(a\) and not \(i\), \cref{eq:dgn_lump} implies that \(\boldsymbol{\lambda}^\nu\) also has the same degeneracy as \(\etw\), which is the condition for lumpability under the partitions \(\brc{C_a}\) \cite{kemeny1960finite,Ball1993Lumpability,burke1958markovian}.
+
+However, if we substitute the last of \cref{eq:dgn_theta} into \cref{eq:dgn_lump} we find that the denominators vanish and the numerators reduce to \( [\CA - (c_a + \lambda) (\wm_a - \Delta p)] (\etwm_a - \etwm_{a-1}) \), etc.
+When the numerator vanishes, the first two equations of \cref{eq:dgn_lump} are degenerate rather than contradictory, leaving one linear combination of the \(\lambda^\nu_i\) undetermined.
+This means we must have \( \CA = (c_a + \lambda) (\wm_a - \Delta p) \), which would seem to imply that there are only two allowed values of \(c_a\).
+
+This bizarre situation can be avoided if we note the other possibility in \cref{eq:dgn_theta}: we could have \(\eqm_i = 0\).
+As we will see below in \cref{eq:serialextarea}, the optimal models have \(\eqm_i = 0\) for all but two states.
+Using this alternative loses the degeneracy of \(\thb\) and \(\w\), leaving us just short of proving lumpability.
 
@@
 
-subsection{Diagonal constraints}\label{sec:diagKKT}
-
-In \cref{eq:kktsol} we saw that
-%
-\begin{equation*}
-\begin{aligned}
-  \kktm^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
-    + \lambda \ombm^\nu_m(0) 
-    + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
-    }, \\
-  \text{where} \quad 
-  \alb^\nu(s) &= \frac{\W^\nu \thb(s)}{rf^\nu}, \qquad
-  \omb^\nu(s) = \frac{\W^\nu \etw(s)}{rf^\nu}.
-\end{aligned}
-\end{equation*}
-%
-If we note that we can write \( \W^\nu = \prn{\frg + (-1)^\nu \enc} / 2 \), we find that
-%
-\begin{equation}\label{eq:diagKKT}
-\begin{aligned}
-  \alb^\nu(s) &= \frac{\delta\mathbf{a}(s) + (-1)^\nu \enc \thb(s)}
-                     {2rf^\nu},
-  \qquad 
-  \omb^\nu(s) = \frac{\delta\w + s \delta\etw(s) - (-1)^\nu \mathbf{a}(s)}
-                     {2rf^\nu}, 
-  \\
-  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
-    \left[ 
-      \hat{A}(s) - c_m(s) (\delta\wm_m + s \delta\etwm_m(s))
-      - \lambda \delta w_m
-    \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} \big[}
-    \left. {}
-      + (-1)^\nu \prn{ 
-        c_m(s) a_m(s) - [\encm \thbm(s)]_m
-        - \delta\wm_m - s \delta\etwm_m(s)
-        + \lambda a_m(0)
-      }
-    \right], 
-\end{aligned}
-\end{equation}
-%
-where \(\delta\) indicates mean subtraction and \(\eq \mathbf{a}(s) = \hat{A}(s)\).
-We can write \cref{eq:diagKKT} as
-%
-\begin{equation*}
-\begin{aligned}
-  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
-    \left[ 
-      \hat{A}
-      - (c_m + (-1)^\nu) (\delta \wm_m + s \delta \etwm_m)
-      - \lambda \delta w_m
-    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
-    % \left.
-      - (-1)^\nu [\encm \dgnm_m]_m
-    \right],
-\end{aligned}
-\end{equation*}
-%
-with all quantities evaluated at \(s\) unless labelled otherwise.
-
-From \cref{eq:oneexit,eq:twoexits} we see that
-%
-\begin{equation*}
-\begin{gathered}
-  \begin{aligned}
-    \W^\nu_{mi} &> 0 &
-    \quad &\implies \quad &
-    \dgnm^\nu_{m,i} &= \dgnm^\nu_{m*}
-    \\
-    \sum_{k \neq m} \W^\nu_{mk} &< r f^\nu &
-    &\implies &
-    \dgnm^\nu_{m,m} &= \dgnm^\nu_{m*},
-  \end{aligned}
-  \\
-  \begin{aligned}
-    \implies \quad
-    [\W^\nu \dgn_m]_m
-      &= \sum_k \Wm^\nu_{mk}  \dgnm_{m,k}
-      = \sum_k \Wm^\nu_{mk}
-        (\dgnm^\nu_{m,k} - (-1)^\nu \etwm_k) \\
-      &= \sum_{k \neq m} \Wm^\nu_{mk} 
-        (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,m}) 
-        - (-1)^\nu [\W^\nu \etw]_m \\
-      &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
-        \sum_{k \neq m} \Wm^\nu_{mk}
-        - (-1)^\nu [\W^\nu \etw]_m \\
-      &= r f^\nu (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
-        - (-1)^\nu [\W^\nu \etw]_m \\
-      &= r f^\nu (\dgnm_{m*} - \dgnm_{m,m}) 
-        + (-1)^\nu \prn{
-          r f^\nu (\etwm_{m*} - \etwm_m)
-          - [\W^\nu \etw]_m
-        },
-  \end{aligned}
-\end{gathered}
-\end{equation*}
-%
-where we used the fact that \( (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) \) is only nonzero when \( \sum_{k \neq m} \Wm^\nu_{mk} = rf^\nu \).
-Therefore
-%
-\begin{equation*}
-\begin{aligned}
-  [\enc \dgn_m]_m
-    &= \sum_{\nu} (-1)^\nu rf^\nu 
-      (\dgnm_{m*} - \dgnm_{m,m}) 
-      + \sum_{\nu} \prn{
-        r f^\nu (\etwm_{m*} - \etwm_m)
-        -  [\W^\nu \etw]_m
-      } \\
-    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
-      + r(\etwm_{m*} - \etwm_m)
-      - [\frg \etw]_m \\
-    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
-      + r(\etwm_{m*} - \etwm_m)
-      - (\delta w_m + s \delta \etwm_m).
-\end{aligned}
-\end{equation*}
-%
-Then we can also write \cref{eq:diagKKT} as
-%
-\begin{equation}\label{eq:diagKKTz}
-\begin{aligned}
-  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
-    \left[ 
-      \hat{A}
-      - c_m (\delta \wm_m + s \delta \etwm_m)
-      - \lambda \delta w_m
-    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
-    % \left.
-      + (-1)^\nu \prn{ 
-        \Delta f (\dgnm_{m,m} - \dgnm_{m*}) 
-        + (\etwm_m - \etwm_{m*})
-      }
-    \right],
-\end{aligned}
-\end{equation}
-%
-again with all quantities evaluated at \(s\) unless labelled otherwise.
-
-\Cref{eq:KKTz} will provide another expression for \(\kktm^\nu_{mm}\).
-That expression will agree with \cref{eq:diagKKTz} provided that
-%
-\begin{equation}\label{eq:diagAgree}
-    \hat{A} - c_m (\delta \wm_m + s \delta \etwm_m)
-      - \lambda \delta w_m 
-      = (\dgnm_{m,m} - \dgnm_{m*}) 
-        + \Delta f (\etwm_m - \etwm_{m*}).
-\end{equation}
-%
-
-%-------Section------------------------------------------------------------------
-
 \subsection{Topology constraining perturbations}\label{sec:topology}
@@
 
+\subsection{Diagonal constraints}\label{sec:diagKKT}
+
+In \cref{eq:kktsol} we saw that
+%
+\begin{equation*}
+\begin{aligned}
+  \kktm^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
+    + \lambda \ombm^\nu_m(0) 
+    + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
+    }, \\
+  \text{where} \quad 
+  \alb^\nu(s) &= \frac{\W^\nu \thb(s)}{rf^\nu}, \qquad
+  \omb^\nu(s) = \frac{\W^\nu \etw(s)}{rf^\nu}.
+\end{aligned}
+\end{equation*}
+%
+If we note that we can write \( \W^\nu = \prn{\frg + (-1)^\nu \enc} / 2 \), we find that
+%
+\begin{equation}\label{eq:diagKKT}
+\begin{aligned}
+  \alb^\nu(s) &= \frac{\delta\mathbf{a}(s) + (-1)^\nu \enc \thb(s)}
+                     {2rf^\nu},
+  \qquad 
+  \omb^\nu(s) = \frac{\delta\w + s \delta\etw(s) - (-1)^\nu \mathbf{a}(s)}
+                     {2rf^\nu}, 
+  \\
+  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+    \left[ 
+      \hat{A}(s) - c_m(s) (\delta\wm_m + s \delta\etwm_m(s))
+      - \lambda \delta w_m
+    \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} \big[}
+    \left. {}
+      + (-1)^\nu \prn{ 
+        c_m(s) a_m(s) - [\encm \thbm(s)]_m
+        - \delta\wm_m - s \delta\etwm_m(s)
+        + \lambda a_m(0)
+      }
+    \right], 
+\end{aligned}
+\end{equation}
+%
+where \(\delta\) indicates mean subtraction and \(\eq \mathbf{a}(s) = \hat{A}(s)\).
+We can write \cref{eq:diagKKT} as
+%
+\begin{equation*}
+\begin{aligned}
+  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+    \left[ 
+      \hat{A}
+      - (c_m + (-1)^\nu) (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m
+    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
+    % \left.
+      - (-1)^\nu [\encm \dgnm_m]_m
+    \right],
+\end{aligned}
+\end{equation*}
+%
+with all quantities evaluated at \(s\) unless labelled otherwise.
+
+From \cref{eq:oneexit,eq:twoexits} we see that
+%
+\begin{equation*}
+\begin{gathered}
+  \begin{aligned}
+    \W^\nu_{mi} &> 0 &
+    \quad &\implies \quad &
+    \dgnm^\nu_{m,i} &= \dgnm^\nu_{m*}
+    \\
+    \sum_{k \neq m} \W^\nu_{mk} &< r f^\nu &
+    &\implies &
+    \dgnm^\nu_{m,m} &= \dgnm^\nu_{m*},
+  \end{aligned}
+  \\
+  \begin{aligned}
+    \implies \quad
+    [\W^\nu \dgn_m]_m
+      &= \sum_k \Wm^\nu_{mk}  \dgnm_{m,k}
+      = \sum_k \Wm^\nu_{mk}
+        (\dgnm^\nu_{m,k} - (-1)^\nu \etwm_k) \\
+      &= \sum_{k \neq m} \Wm^\nu_{mk} 
+        (\dgnm^\nu_{m,k} - \dgnm^\nu_{m,m}) 
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
+        \sum_{k \neq m} \Wm^\nu_{mk}
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= r f^\nu (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) 
+        - (-1)^\nu [\W^\nu \etw]_m \\
+      &= r f^\nu (\dgnm_{m*} - \dgnm_{m,m}) 
+        + (-1)^\nu \prn{
+          r f^\nu (\etwm_{m*} - \etwm_m)
+          - [\W^\nu \etw]_m
+        },
+  \end{aligned}
+\end{gathered}
+\end{equation*}
+%
+where we used the fact that \( (\dgnm^\nu_{m*} - \dgnm^\nu_{m,m}) \) is only nonzero when \( \sum_{k \neq m} \Wm^\nu_{mk} = rf^\nu \).
+Therefore
+%
+\begin{equation*}
+\begin{aligned}
+  [\enc \dgn_m]_m
+    &= \sum_{\nu} (-1)^\nu rf^\nu 
+      (\dgnm_{m*} - \dgnm_{m,m}) 
+      + \sum_{\nu} \prn{
+        r f^\nu (\etwm_{m*} - \etwm_m)
+        -  [\W^\nu \etw]_m
+      } \\
+    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + r(\etwm_{m*} - \etwm_m)
+      - [\frg \etw]_m \\
+    &= r\Delta f (\dgnm_{m*} - \dgnm_{m,m}) 
+      + r(\etwm_{m*} - \etwm_m)
+      - (\delta w_m + s \delta \etwm_m).
+\end{aligned}
+\end{equation*}
+%
+Then we can also write \cref{eq:diagKKT} as
+%
+\begin{equation}\label{eq:diagKKTz}
+\begin{aligned}
+  \kktm^\nu_{mm} &= \frac{\eqm_m}{2rf^\nu} 
+    \left[ 
+      \hat{A}
+      - c_m (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m
+    % \right. \\ &\hphantom{= \frac{\eqm_m}{2f^\nu} }
+    % \left.
+      + (-1)^\nu \prn{ 
+        \Delta f (\dgnm_{m,m} - \dgnm_{m*}) 
+        + (\etwm_m - \etwm_{m*})
+      }
+    \right],
+\end{aligned}
+\end{equation}
+%
+again with all quantities evaluated at \(s\) unless labelled otherwise.
+
+\Cref{eq:KKTz} provided another expression for \(\kktm^\nu_{mm}\).
+That expression will agree with \cref{eq:diagKKTz} provided that
+%
+\begin{equation}\label{eq:diagAgree}
+    \hat{A} - c_m (\delta \wm_m + s \delta \etwm_m)
+      - \lambda \delta w_m 
+      = (\dgnm_{m,m} - \dgnm_{m*}) 
+        + \Delta f (\etwm_m - \etwm_{m*}).
+\end{equation}
+%
+
+%-------Section------------------------------------------------------------------
+
 \subsection{Scaling with a floating cutoff}\label{sec:cutoff}

### 2021-11-22 Notes/laplace.tex
use notation macros

@@
 %
-If we regard the \(\wm_k\) as optimisable parameters (limited to the range \( [-1,1] \) wlg.),s the optimal choice is clearly \( \wm_k = \sgn(c_k) \).
+If we regard the \(\wm_k\) as optimisable parameters (limited to the range \( [-1,1] \) wlg.), the optimal choice is clearly \( \wm_k = \sgn(c_k) \).
 
@@
 %
-where \(\Delta p\) is the constant value we are holding \(\eq\w\) at.
+where \(\Delta p\) is the constant value at which we are holding \(\eq\w\).
 The Karush-Kuhn-Tucker necessary conditions for a maximum are
@@
 \begin{equation}\label{eq:detarow}
-  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i = 0,
+  \frac{\delta\etwm_{i+1}+\delta\etwm_{i-1}}{2} - (s+1)\delta\etwm_i = 0,
 \end{equation}
@@
 \begin{aligned}
-  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{m+1}+\delta\eta_{m-1}}{2} - (s+1)\delta\eta_{m} &= \frac{1}{s},\\
-  \frac{\delta\eta_{m+2}+\delta\eta_{m}}{2} - (s+1)\delta\eta_{m+1} &= -\frac{1}{s},\\
-  \frac{\delta\eta_{2m-1}}{2} - \prn{s+\frac{1}{2}}\delta\eta_{2m} &= 0.
-\end{aligned}
-\end{equation}
-%
-These equations are invariant under \(\delta\eta_{2m+1-i} \to -\delta\eta_i\), so these two must be equal (up to a constant that we can drop).
+  \frac{\delta\etwm_2}{2} - \prn{s+\frac{1}{2}}\delta\etwm_1 &= 0, \\
+  \frac{\delta\etwm_{m+1}+\delta\etwm_{m-1}}{2} - (s+1)\delta\etwm_{m} &= \frac{1}{s},\\
+  \frac{\delta\etwm_{m+2}+\delta\etwm_{m}}{2} - (s+1)\delta\etwm_{m+1} &= -\frac{1}{s},\\
+  \frac{\delta\etwm_{2m-1}}{2} - \prn{s+\frac{1}{2}}\delta\etwm_{2m} &= 0.
+\end{aligned}
+\end{equation}
+%
+These equations are invariant under \(\delta\etwm_{2m+1-i} \to -\delta\etwm_i\), so these two must be equal (up to a constant that we can drop).
 Then we have
@@
 \begin{aligned}
-  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i &= 0,
+  \frac{\delta\etwm_{i+1}+\delta\etwm_{i-1}}{2} - (s+1)\delta\etwm_i &= 0,
   \quad\text{for}\quad 1<i<m, \\
-  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
+  \frac{\delta\etwm_2}{2} - \prn{s+\frac{1}{2}}\delta\etwm_1 &= 0, \\
+  \frac{\delta\etwm_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\etwm_{m} &= \frac{1}{s}.
 \end{aligned}
@@
 \begin{equation}\label{eq:detagensol}
-  \delta\eta_i = B\e^{\beta(i-1)} + C\e^{\beta(1-i)},
+  \delta\etwm_i = B\e^{\beta(i-1)} + C\e^{\beta(1-i)},
   \qquad\text{where}\qquad
@@
 \begin{aligned}
-  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
-  \frac{\delta\eta_3 + (1-\varepsilon)\delta\eta_1}{2} - \prn{s+\frac{2-\varepsilon}{2}}\delta\eta_2 &= 0, \\
-  \frac{\delta\eta_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{m} &= \frac{1}{s}.
+  \frac{\delta\etwm_2}{2} - \prn{s+\frac{1}{2}}\delta\etwm_1 &= 0, \\
+  \frac{\delta\etwm_3 + (1-\varepsilon)\delta\etwm_1}{2} - \prn{s+\frac{2-\varepsilon}{2}}\delta\etwm_2 &= 0, \\
+  \frac{\delta\etwm_{m-1}}{2} - \prn{s+\frac{3}{2}}\delta\etwm_{m} &= \frac{1}{s}.
 \end{aligned}
@@
 \begin{aligned}
-  \delta\eta_i &= B\e^{(i-2)\beta} + C\e^{(2-i)\beta} \qquad
+  \delta\etwm_i &= B\e^{(i-2)\beta} + C\e^{(2-i)\beta} \qquad
         \text{for} \quad 2 \leq i \leq m, \\
-  \delta\eta_1 &= \frac{1}
+  \delta\etwm_1 &= \frac{1}
          {[(1-\varepsilon)\cosh(m\beta) + \varepsilon(2s+1)\cosh((m-1)\beta)]s}, \\
@@
 \begin{aligned}
-  \frac{1 - \varepsilon}{2} \delta \eta_2 
-    - \prn{s+\frac{1-\varepsilon}{2}} \delta \eta_1 
+  \frac{1 - \varepsilon}{2} \delta \etwm_2 
+    - \prn{s+\frac{1-\varepsilon}{2}} \delta \etwm_1 
     &= 0, \\
-  \frac{\delta \eta_{m-1}}{2} 
-    - \prn{s+\frac{3}{2}} \delta \eta_{m} 
+  \frac{\delta \etwm_{m-1}}{2} 
+    - \prn{s+\frac{3}{2}} \delta \etwm_{m} 
     &= \frac{1}{s}.

### 2021-11-23 Notes/laplace.tex
typos

@@
 \begin{aligned}
-  [\W\pot \etw]_i &= \sum_{j \in C_{a+1}} \Wm\pot_{ij} (\etwm_j - \etwm_i)
+  [\W\pot \etw]_i &= \sum_{\mathclap{j \in C_{a+1}}} \Wm\pot_{ij} (\etwm_j - \etwm_i)
     = (\etwm_{a+1} - \etwm_a) \sum_{\mathclap{j \in C_{a+1}}} \Wm\pot_{ij}
@@
 %
-Because the right-hand sides above depend on \(a\) and not \(i\), \cref{eq:dgn_lump} implies that \(\boldsymbol{\lambda}^\nu\) also has the same degeneracy as \(\etw\), which is the condition for lumpability under the partitions \(\brc{C_a}\) \cite{kemeny1960finite,Ball1993Lumpability,burke1958markovian}.
+Because the right-hand sides above depend on \(a\) and not \(i\), \cref{eq:dgn_lump} implies that \(\boldsymbol{\lambda}^\nu\) also has the same degeneracy as \(\etw\), which is the condition for lumpability under the partitions \(\brc{C_a}\) \cite{kemeny1960finite,Ball1993Lumpability,burke1958markovian}:
+for any two states in \(C_a\), we get the same value for the sum of transition rates from that state to all states in \(C_b\).
 
@@
 First let us maximise \cref{eq:serialarea} at fixed $\eqm_\pm = \sum_k \eqm_k \prn{\frac{1\pm\wm_k}{2}}$.
-Clearly this will happen when we put all of the probability at the ends: $\eqm_1=\eqm_-$ and $\eqm_n=\eqm_+$ are the only non-zero $\eqm_k$.
+Clearly this will happen when we put all of the probability at the ends: $\eqm_1=\eqm_-$ and $\eqm_M=\eqm_+$ are the only non-zero $\eqm_k$.
 This gives an area of

### 2022-06-03 Notes/laplace.tex
typos

@@
 \newcommand{\MM}{\mathbf{\MMm}}
-%prob distribution
+%probability distribution
 \newcommand{\prm}{p}
@@
 \newcommand{\fpt}{\mathbf{\fptm}}
-%off-diag first passage times
+%off-diagonal first passage times
 \newcommand{\fptbm}{\overline{\fptm}}
@@
 We can test if the pattern had been previously stored by computing \(\synid\cdt\syn\) and comparing it to some threshold~\cite{Sommer1998retrieval}.
-For large \(N\), this quantity will have a Gausssian distribution.
+For large \(N\), this quantity will have a Gaussian distribution.
 There will be a ROC curve as a function of this threshold:
@@
 When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices \(\MMd^\nu\) whose \(i,j\) elements are the transition probabilities from state \(i\) to state \(j\),
-and \(\nu \) labels the diffrerent types of plasticity event.
+and \(\nu \) labels the different types of plasticity event.
 Potentiating events have \(\nu=\) pot, depressing events have \(\nu=\) dep.
@@
 Given a sequence of plasticity events at times \(t_1,t_2,t_3,\ldots\), the probability distribution across the internal states at time \(t_n\), 
-mariginalising over states at earlier times, is
+marginalising over states at earlier times, is
 %
@@
 
-Now to study the SNR as a function of  time, it helps to consider some variables that are more directly related to the time dependance than the matrix elements. 
+Now to study the SNR as a function of  time, it helps to consider some variables that are more directly related to the time dependence than the matrix elements. 
 So we introduce an eigenvector decomposition for \(-\frg\):
@@
 %
-As the KKT multipliers are nonnegative, this implies that all the diagonal KKT multipliers vanish and we can ignore the diagonal constraints.
+As the KKT multipliers are non-negative, this implies that all the diagonal KKT multipliers vanish and we can ignore the diagonal constraints.
 
@@
 %
-Therefore, provided that state \(i\) has nonzero equilibrium probability, degeneracy in \(\etw\) implies degeneracy in \(\thb\) for an otimal model.
+Therefore, provided that state \(i\) has nonzero equilibrium probability, degeneracy in \(\etw\) implies degeneracy in \(\thb\) for an optimal model.
 
@@
 
-The fourth inequality can only be saturated if, for each state, the outgoing potentitation rates and depression rates separately sum to one.
+The fourth inequality can only be saturated if, for each state, the outgoing potentiation rates and depression rates separately sum to one.
 This does force degenerate states to be lumpable, and we are left with a serial topology with deterministic transitions.
@@
 Suppose there are two states, \(i,j\), with nonzero transitions from \(m\) for one \(\nu\).
-Both must statisfy \cref{eq:oneexit} for the same \(\kktm^\nu_{mm}\), so
+Both must satisfy \cref{eq:oneexit} for the same \(\kktm^\nu_{mm}\), so
 %
@@
 
-We use \(m*\) to label the intersection of the two diagonal lines in \cref{fig:planes}(\ref{fig:triangle_spacing}), wether or not there is a state there.
+We use \(m*\) to label the intersection of the two diagonal lines in \cref{fig:planes}(\ref{fig:triangle_spacing}), whether or not there is a state there.
 It can be defined for each \(m\), provided that it has at least one outgoing transition for each plasticity type, via
@@
 We denote this expression by \( \kkt^\ddagger(s, r, \W) \)
-(note there is \(s\) dependance through \(\thb(s), \etw(s)\) \etc and \(r\) dependance through \(\alb\) and \(\omb\)).
+(note there is \(s\) dependence through \(\thb(s), \etw(s)\) \etc and \(r\) dependence through \(\alb\) and \(\omb\)).
 If we denote the solution to the maximisation problem by \(\W^\ast(s,r), \kkt^\ast(s,r)\), we have \( \kkt^\ast(s,r) = \kkt^\ddagger(s, r, \W^\ast(s,r)) \) and
@@
 From \cref{sec:serialarea}, we know the optimal models at \(s = 0\).
-% (There is still an ambiguity concering the scale, but it makes sense to choose to saturate the diagonal constraints in \cref{eq:wdof} so that \cref{eq:optmodelscale} holds).
+% (There is still an ambiguity concerning the scale, but it makes sense to choose to saturate the diagonal constraints in \cref{eq:wdof} so that \cref{eq:optmodelscale} holds).
 At \(r = 0\) there is only one allowed model: \(\W = \mathbf{0}\).
@@
 For lower \(s\), \(\varepsilon = 0\) is the optimum. 
-For higher \(s\), \(\varepsilon = 1\) is the optimum, which is eqivalent to \(\varepsilon = 0\) with \(M-2\) states.
+For higher \(s\), \(\varepsilon = 1\) is the optimum, which is equivalent to \(\varepsilon = 0\) with \(M-2\) states.
 This is depicted by the dotted lines in \cref{fig:shorten}\ref{fig:short_ratio}.

### 2022-09-05 Notes/laplace.tex
eigenmoodes for uni serial

@@
 % \DeclareMathDelimiter{\rVvert}{\mathclose}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
-
+\usepackage{MnSymbol}
 \newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
@@
 
+Note that
+%
+\begin{equation}\label{eq:laplaceeig}
+  \hat{A}(s) = \sum_a \frac{\CI_a}{s + q_a},
+\end{equation}
+%
+Therefore the eigenmodes with nonzero initial SNR can be found from singularity structure of \( \hat{A}(s) \) in the complex plane.
+The poles are at \(s = - \frac{1}{\tau_a}\), with residue \(\CI_a\).
+
 Now the question is what are the allowed values of \( (\CI_a, \tau_a) \)?
@@
 Considering the two-state model with deterministic transitions, we see that \(\cnst\) is at least 1.
-Later, looking at uniform serial models in \cref{sec:heuristicenv} we will see near \cref{eq:henvapprox} that \(\gamma\) is at least 1.18.
+For the \(M\)-state uniform serial model with deterministic transitions, in \cref{eq:unieig}, we find
+%
+\begin{equation*}
+  \CI_n \sqrt{r\tau_n} = 
+      \frac{2 \sqrt{2} \sin\frac{n\pi}{2} \cos\frac{n\pi}{2M}}
+           {\prn{M \sin \frac{n\pi}{2M}}^2},
+\end{equation*}
+%
+which is maximised at \(n=1, M \to \infty\).
+Thus we see that \(\cnst \geq \frac{128}{\pi^4} = 1.31 \).
 
@@
 %
+This has poles as \( \beta = \frac{(2n-1)\pi\ir}{2m} \), therefore
+%
+\begin{equation}\label{eq:unieig}
+  q_n = 1 - \cos\frac{(2n-1)\pi}{2m} = 2 \sin^2\frac{(2n-1)\pi}{4m},
+  \qquad
+  \CI_n = \frac{(-1)^{n-1} \sin\frac{(2n-1)\pi}{2m}}{m^2 q_n},
+  \qquad n = 1,\dots,m.
+\end{equation}
+%
+Note that these are only the eigenmodes with nonzero \(\CI_n\).
 

### 2022-12-08 Notes/laplace.tex
uni eigenmode and heuristic env

@@
 %
-This has poles as \( \beta = \frac{(2n-1)\pi\ir}{2m} \), therefore
+This has poles as \( \beta = \frac{(2n-1)\pi\ir}{2m} \).
+From \cref{eq:laplaceeig} we see that the poles and residues in the \(s\)-plane are \( (-q_n, \CI_n) \), therefore:
 %
 \begin{equation}\label{eq:unieig}
-  q_n = 1 - \cos\frac{(2n-1)\pi}{2m} = 2 \sin^2\frac{(2n-1)\pi}{4m},
+  q_n = 1 - \cos \brkfrac{(2n-1)\pi}{2m} 
+      % = 2 \sin^2\frac{(2n-1)\pi}{4m},
   \qquad
-  \CI_n = \frac{(-1)^{n-1} \sin\frac{(2n-1)\pi}{2m}}{m^2 q_n},
+  \CI_n = \frac{(-1)^{n-1}}{m^2} \cot \brkfrac{(2n-1)\pi}{4m},
   \qquad n = 1,\dots,m.
@@
   \hat{A}_*(s) = \prn{\frac{ \sinh y_* }{ \cosh^2 y_* }} 
-      \frac{\abs{\beta(s)}}{s},
+      \frac{\operatorname{arccosh}(1+s)}{s},
   \\ \text{where}\quad
-  y_* = \tanh \frac{y_*}{2} \cosh y_* \approx 1.50553,
-  \qquad
+  y_* = \tanh \frac{y_*}{2} \cosh y_* \approx 1.50553.
 \end{multline}
 %
-and \(\beta(s)\) is given by the second formula of \cref{eq:detagensol}.
+% and \(\beta(s)\) is given by the second formula of \cref{eq:detagensol}.
 
@@
   \tfrac{1}{2\sinh^2 \frac{y_*}{2}} \leq & r\tau \leq \tfrac{1}{2\sinh^2 \frac{y_*}{M}} , \\
-  & \frac{\sqrt{N}\brk{S\prn{\tfrac{M\beta}{2}} - S\prn{\tfrac{(M-2)\beta}{2}}}}
-        {\brk{\sqrt{{\frac{M-2}{2} S\prn{\tfrac{M\beta}{2}} - \frac{M}{2} S\prn{\tfrac{(M-2)\beta}{2}}}} + 1}^2} , &
+  & \frac{\sqrt{N}\brk{S\prn{\tfrac{M\beta(\tau)}{2}} - S\prn{\tfrac{(M-2)\beta(\tau)}{2}}}}
+        {\brk{\sqrt{{\frac{M-2}{2} S\prn{\tfrac{M\beta(\tau)}{2}} - \frac{M}{2} S\prn{\tfrac{(M-2)\beta(\tau)}{2}}}} + 1}^2} , &
   % & \max_{\varepsilon \in [0,1]} \brc{
@@
 %
-where \(S(\beta) = \cosh \beta(\tau) - 1\) (see \cref{eq:detagensol}) and \(y_*\) is defined in \cref{eq:unienv}.
+where \(S(\beta) = \cosh \beta - 1\), \( S(\beta(\tau)) = 1/\tau \) (see \cref{eq:detagensol}) and \(y_*\) is defined in \cref{eq:unienv}.
 

### 2024-05-14 Notes/laplace.tex
changing planes

--- 2022-09-05 Notes/laplace.tex
@@
 % \usepackage{stix2}
-\usepackage[tables,graphics,draft,panels=paralist]{sl-preamble}
+\usepackage[tables,graphics,draft,panels=paralist,noeqcref]{sl-preamble}
 %\usepackage{amsldoc} for finding documentation only
@@
 When \(\MM\) satisfies detailed balance (more generally when it commutes with its time reverse), the numerical range is the convex hull of the eigenvalues.
+For any vector \(\mathbf{u}\) that is perpendicular to \(\onev\), we define the following functional
+%
+\begin{equation*}
+    \tau(\mathbf{u}) = - \frac{\mathbf{u}^\dag \mathbf{u}}{\mathbf{u}^\dag \frg \mathbf{u}}
+      = \frac{1}{\mathcal{W}(-\frg, \mathbf{u})},
+\end{equation*}
+%
+with the \(\mathbf{u}\) in \( \mathcal{W}(-\frg, \mathbf{u}) \) representing the vector space spanned by \(\mathbf{u}\).
 
 The other quantity of interest in \cref{eq:snrcoeffs} is the amplitude of the eigenmodes.
-In analogy, we define the following functional (for real \(\mathbf{u}\))
+In analogy, we define the following functionals (for real \(\mathbf{u}\))
 %
@@
 \begin{equation}\label{eq:sqrtnumrange}
-  \CI(\mathbf{u}) \leq \sqrt{\frac{2\av{\mathbf{u}, -\frg \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
-      = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})/r}.
+\begin{aligned}
+  \CI(\mathbf{u}) &\leq \sqrt{\frac{2\av{\mathbf{u}, -\frg \mathbf{u}}}{r\av{\mathbf{u}, \mathbf{u}}}}
+      = \sqrt{2\mathcal{W}(-\frg;\mathbf{u})/r} , &\quad \text{or} \quad
+  \CI(\mathbf{u}) \sqrt{r \tau(\mathbf{u})} &\leq \sqrt{2}.
+\end{aligned}
 \end{equation}
@@
 the two-state model with deterministic transitions.
-This model has \(\CI_1 = \tau_1 = 1\), suggesting that there's a factor of \(\sqrt{2}\) looseness in this bound.
+This model has \(\CI_1 = \tau_1 = 1\), suggesting that there's a factor of \(\sqrt{2}\) looseness in this bound, although we will see below that there is at least one model for which this quantity is higher.
 
@@
 % 
-We have shown that the smallest \(\cnst\) for which \cref{eq:gammabnd} is true is less than 2.
-Considering the two-state model with deterministic transitions, we see that \(\cnst\) is at least 1.
+We have shown that the smallest \(\cnst\) for which \cref{eq:gammabnd} is true is at most 2.
 For the \(M\)-state uniform serial model with deterministic transitions, in \cref{eq:unieig}, we find
@@
 which is maximised at \(n=1, M \to \infty\).
-Thus we see that \(\cnst \geq \frac{128}{\pi^4} = 1.31 \).
+Thus we see that \(\cnst\) is at least \(\frac{128}{\pi^4} = 1.31 \).
 
@@
 %
+Therefore, the only nonzero transitions from \(m\) in \(\W^\nu\) must go to the global minimum of \(\dgn^\nu_m\).
+This can also be seen by considering a perturbation that increases \(\Wm^\nu_{mi}\) and decreases \(\Wm^\nu_{mk}\) by the same amount (thus evading the diagonal constraint):
+%
+\begin{equation}\label{eq:exitminimum}
+  \pdiff{\mathcal{L}}{\Wm^\nu_{mi}} - \pdiff{\mathcal{L}}{\Wm^\nu_{mk}}
+    = \eqm_m \prn{\dgnm^\nu_{m,k} - \dgnm^\nu_{m,i}} + \kktm^\nu_{mi} - \kktm^\nu_{mk}.
+\end{equation}
+%
+If \( \dgnm^\nu_{m,k} > \dgnm^\nu_{m,i} \) we must have \( \kktm^\nu_{mk} > 0 \) and therefore \( \Wm^\nu_{mk} = 0 \).
+
 If both \(\nu\) have nonzero rates from \(m\) to \(i\) and \(j\):
@@
 \caption[Depicting constraints on topology]{Depicting constraints on topology.
-(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, 
+(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (-\etwm_l, \dgnm_{m,l}) \)-plane, 
   with \(i\) a state with a nonzero depressing transition from \(m\), 
@@
 \end{figure}
+
+%-------Section------------------------------------------------------------------
+
+\subsubsection{Changing planes}\label{sec:planes}
+
+Consider the setup of the previous section.
+Suppose there was a nonzero transition in \(\W\pot\) from some state \(p\) to state \(m\).
+Now \(m\) must be the minimum of \(\dgn\pot_p\), where \(j\) was the minimum of \(\dgn\pot_m\).
+
+As we move from the \( (-\etwm_l, \dgnm_{m,l}) \)-plane in \cref{fig:triangle_spacing} to the \( (-\etwm_l, \dgnm_{p,l}) \)-plane, the different states will rise or fall depending on \(c_p - c_m\) and the value of \(\etwm_l\).
+Let's focus on the rise/fall \emph{relative} to that of \(m\).
+If \(c_p > c_m\), states to the left of \(m\) will rise and those to its right will fall relative to it.
+If \(c_p < c_m\), states to the left of \(m\) will fall and those to its right will rise relative to it.
+
+If state \(j\) is to the left of \(m\) we must have \(c_p > c_m\).
+If \(j\) is to the right, as would be implied by the triangularity of \cref{sec:triangular} if it could be proven more generally,  we must have \(c_p < c_m\).
+
+Similarly, suppose  there was a nonzero transition in \(\W\dep\) from some state \(q\) to state \(m\).
+If state \(i\) is to the left of \(m\), as would be implied by the triangularity of \cref{sec:triangular}, we must have \(c_q > c_m\).
+If \(i\) is to the right we must have \(c_q < c_m\).
+
+The state \(j\) needs to do more than rise when we change planes.
+It must rise further than the gap between the diagonals passing through \(j\) and \(m\), which is \( \frac{\kktm\pot_{mm}}{\eqm_m} \).
+Regardless of the signs of the two factors, this rise is given by
+%
+\begin{equation}\label{eq:plane_rise}
+  \prn{c_m - c_p} \prn{\etwm_m - \etwm_j} \geq \frac{\kktm\pot_{mm}}{\eqm_m}.
+\end{equation}
+%
+A similar relation holds between \(c_q\), \(\etwm_i\) and \(\kktm\dep_{mm}\).
 
@@
   \end{myenuma}
-  \caption[Heuristic optimal models]{Heuristics for models that maximise memory curve.
+  \caption[Semi-empirically optimal models]{Semi-empirical models that maximise memory curve.
   All unlabelled arrows have transition probabilities equal to 1.
@@
 
-Looking at them models that were found by numerical maximisation in \cref{fig:envvid} 
+Looking at the models that were found by numerical maximisation in \cref{fig:envvid} 
 (see the video for the full set of models) suggests we look at the particular set of models shown in \cref{fig:heuristicmodels}.
+The two families of models are not identical, but are similar enough to provide some insight.
 
@@
 
-These models' Laplace transforms can be computed analytically, as we will see below, leading to a heuristic envelope.
+These models' Laplace transforms can be computed analytically, as we will see below, leading to a semi-empirical envelope.
 All of these calculations will be performed at \(f\potdep=\frac{1}{2}\), so the denominator will play no role.
@@
 %
-When \(s \sim \CO(M^{-2}) \ll 1\), we have \(\beta \approx \pm \sqrt{2s}\) and the heuristic envelope is approximately
+When \(s \sim \CO(M^{-2}) \ll 1\), we have \(\beta \approx \pm \sqrt{2s}\) and the semi-empirical envelope is approximately
 %
@@
 
-\subsection{Heuristic envelope}\label{sec:heuristicenv}
+\subsection{Semi-empirical envelope}\label{sec:heuristicenv}
 
@@
   \includegraphics[width=0.8\linewidth]{LenvHeuristic.pdf}\\
-  \caption[Heuristic envelope for the signal-to-noise ratio]
-  {Heuristic envelope  for the signal-to-noise ratio, \cref{eq:henv}.
+  \caption[Semi-empirical envelope for the signal-to-noise ratio]
+  {Semi-empirical envelope  for the signal-to-noise ratio, \cref{eq:henv}.
   Analytic expressions from \cref{eq:henvapprox} describe the yellow curve and are only valid to the right of each region.}\label{fig:heuristicenv}
